%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\input{defs}
% \input{largeprint} \renewcommand{\gab}[1]{\marginpar[\tiny\raggedleft #1]{\tiny #1}}


\title{Portrait of the Artist as a Young Man\\
       The Christmas Dinner}

\begin{document}

\maketitle


\begin{comment}

    \N1:
    A great fire,
    banked high and red,
    flamed in the grate
    and under the ivytwined branches of the chandelier
    the Christmas table was spread.
    They had come home a little late
    and still dinner was not ready:
    but it would be ready in a jiffy,
    his mother had said.
    They were waiting for the door to open
    and for the servants to come in,
    holding the big dishes covered
    with their heavy metal covers.

    \N2:
    All were waiting:
    uncle Charles, who sat far away in the shadow of the window,
    Dante and Mr~Casey,
    who sat in the easychairs at either side of the hearth,
    Stephen, seated on a chair between them,
    his feet resting on the toasted boss.
    Mr~Dedalus looked at himself in the pierglass above the mantelpiece,
    waxed out his moustache ends
    and then,
    parting his coat tails,
    stood with his back to the glowing fire:
    and still from time to time
    he withdrew a hand from his coat tail
    to wax out one of his moustache ends.
    Mr~Casey leaned his head to one side
    and, smiling,
    tapped the gland of his neck with his fingers.
    And Stephen smiled too
    for he knew now that it was not true
    that Mr~Casey had a purse of silver in his throat.
    He smiled to think how the silvery noise which Mr~Casey used to make
    had deceived him.
    And when he had tried to open Mr~Casey’s hand
    to see if the purse of silver was hidden there
    he had seen that the fingers could not be straightened out:
    and Mr~Casey had told him
    that he had got those three cramped fingers
    making a birthday present for Queen Victoria.
    Mr~Casey tapped the gland of his neck
    and smiled at Stephen with sleepy eyes:
    and Mr~Dedalus said to him:

    \simon:
    Yes.
    Well now,
    that’s all right.
    O, we had a good walk,
    hadn’t we,
    John?
    Yes...
    I wonder if there’s any likelihood of dinner this evening.
    Yes...
    O, well now,
    we got a good breath of ozone round the Head today.
    Ay, bedad.

    \N1:
    He turned to Dante and said:

    \simon:
    You didn’t stir out at all, Mrs~Riordan?

    \N2:
    Dante frowned and said shortly:

    \dante:
    No.

    \N1:
    Mr~Dedalus dropped his coat tails
    and went over to the sideboard.
    He brought forth a great stone jar of whisky
    from the locker and filled the decanter slowly,
    bending now and then
    to see how much he had poured in.
    Then replacing the jar in the locker
    he poured a little of the whisky into two glasses,
    added a little water
    and came back with them to the fireplace.

    \simon:
    A thimbleful, John,
    \sout{he said,}
    just to whet your appetite.

    \N1:
    Mr~Casey took the glass,
    drank,
    and placed it near him on the mantelpiece.
    Then he said:

    \casey:
    Well, I can’t help thinking of our friend Christopher
    manufacturing...

    \N1:
    He broke into a fit of laughter and coughing
    and added:

    \casey:
    ...manufacturing that champagne for those fellows.

    \N1:
    Mr~Dedalus laughed loudly.

    \simon:
    Is it Christy?
    \sout{he said.}
    There’s more cunning in one of those warts on his bald head
    than in a pack of jack foxes.

    \N1:
    He inclined his head,
    closed his eyes,
    and, licking his lips profusely,
    began to speak with the voice of the hotel keeper.

    \simon:
    And he has such a soft mouth when he’s speaking to you,
    don’t you know.
    He’s very moist and watery about the dewlaps,
    God bless him.

    \N1:
    Mr~Casey was still struggling through
    his fit of coughing and laughter.
    Stephen, seeing and hearing the hotel keeper
    through his father’s face and voice,
    laughed.

    \N1:
    Mr~Dedalus put up his eyeglass
    and, staring down at him,
    said quietly and kindly:

    \simon:
    What are you laughing at,
    you little puppy,
    you?

\end{comment}


\N1:
The servants entered
and placed the dishes on the table.

\N2:
Mrs~Dedalus followed
and the places were arranged.

\marydedalus:
Sit over,

\N2:
she said.

\N1:
Mr~Dedalus went to the end of the table
and said:

\simon:
Now, Mrs~Riordan,
sit over.
John,
sit you down,
my hearty.

\N1:
He looked round to where uncle Charles sat and said:

\simon:
Now then, sir,
there’s a bird here waiting for you.

\N1:
When all had taken their seats |
he laid his hand on the cover
and then said quickly,
withdrawing it:

\simon:
Now, Stephen.

\N2:
Stephen stood up in his place
to say the grace before meals:

\Stephen:
Bless us, O Lord,
and these Thy gifts
which through Thy bounty
we are about to receive
through Christ our Lord.
Amen.

\N1:
All blessed themselves
and Mr~Dedalus with a sigh of pleasure |
lifted from the dish
the heavy cover |
pearled around the edge with glistening drops.

\N2:
Stephen looked at the plump turkey
which had lain, trussed and skewered,
on the kitchen table.


\begin{comment}

    \N2:
    He knew that his father had paid a guinea for it
    in Dunn’s of D’Olier Street
    and that the man had prodded it often at the breastbone
    to show how good it was:
    and he remembered the man’s voice when he had said:

    \man:
    Take that one, sir.
    That’s the real Ally Daly.

    \StephenInt:
    Why did Mr~Barrett in Clongowes
    call his pandybat a turkey?
    But Clongowes was far away:
    and the warm heavy smell of turkey and ham and celery
    rose from the plates and dishes
    and the great fire was banked high and red in the grate
    and the green ivy and red holly
    made you feel so happy
    and when dinner was ended
    the big plum pudding would be carried in,
    studded with peeled almonds and sprigs of holly,
    with bluish fire running around it
    and a little green flag flying from the top.

\end{comment}


\N2:
It was his first Christmas dinner
and he thought of his little brothers and sisters
who were waiting in the nursery,
as he had often waited,
till the pudding came.
\sout{The deep low collar and the Eton jacket
made him feel queer and oldish:
and that morning
when his mother had brought him down to the parlour,
dressed for mass,
his father had cried.
That was because he was thinking of his own father.
And uncle Charles had said so too.}

\N1:
Mr~Dedalus covered the dish
and began to eat hungrily.
Then he said:

\simon:
Poor old Christy,
he’s nearly lopsided now with roguery.

\marydedalus:
Simon,
\sout{said Mrs~Dedalus,}
you haven’t given Mrs~Riordan any sauce.

\N1:
Mr~Dedalus seized the sauceboat.

\simon:
Haven’t I?
\sout{he cried.}
Mrs~Riordan,
pity the poor blind.

\N2:
Dante covered her plate with her hands and said:

\dante:
No, thanks.

\N1:
Mr~Dedalus turned to uncle Charles.

\simon:
How are you off, sir?

\charles:
Right as the mail, Simon.

\simon:
You, John?

\casey:
I’m all right.
Go on yourself.

\simon:
Mary?
Here, Stephen,
here’s something to make your hair curl.

\N1:
He poured sauce freely over Stephen’s plate
and set the boat again on the table.
Then he asked uncle Charles was it tender.
Uncle Charles could not speak
because his mouth was full
but he nodded that it was.

\simon:
That was a good answer
our friend made to the canon.
What?
\sout{said Mr~Dedalus.}

\casey:
I didn’t think he had that much in him,
\sout{said Mr~Casey.}

\casey:
\stage{[as parishioner]}
I’ll pay your dues,
father,
when you cease turning the house of God into a polling-booth.

\dante:
A nice answer,
\sout{said Dante,}
for any man calling himself a catholic to give to his priest.

\simon:
They have only themselves to blame,
\sout{said Mr~Dedalus suavely.}
If they took a fool’s advice |
they would confine their attention to religion.

\dante:
It is religion,
\sout{Dante said.}
They are doing their duty in warning the people.

\casey:
We go to the house of God,
\sout{Mr~Casey said,}
in all humility to pray to our Maker |
and not to hear election addresses.

\dante:
It is religion,
\sout{Dante said again.}
They are right.
They must direct their flocks.

\simon:
And preach politics from the altar,
is it?
\sout{asked Mr~Dedalus.}

\dante:
Certainly,
\sout{said Dante.}
It is a question of public morality.
A priest would not be a priest |
if he did not tell his flock
what is right and what is wrong.

\N2:
Mrs~Dedalus laid down her knife and fork,
saying:

\marydedalus:
For pity sake and for pity sake |
let us have no political discussion
on this day |
of all days in the year.

\charles:
Quite right, ma’am,
\sout{said uncle Charles.}
Now Simon, that’s quite enough now.
Not another word now.

\simon:
Yes, yes,
\sout{said Mr~Dedalus quickly.}

\N1:
He uncovered the dish boldly
and said:

\simon:
Now then,
who’s for more turkey?

\N2:
Nobody answered.
Dante said:

\dante:
Nice language for any catholic to use!

\marydedalus:
Mrs~Riordan,
I appeal to you,
\sout{said Mrs~Dedalus,}
to let the matter drop now.

\N2:
Dante turned on her and said:

\dante:
And am I to sit here
and listen to the pastors of my church
being flouted?

\simon:
Nobody is saying a word against them,
\sout{said Mr~Dedalus,}
so long as they don’t meddle in politics.

\dante:
The bishops and priests of Ireland have spoken,
\sout{said Dante,}
and they must be obeyed.

\casey:
Let them leave politics alone,
\sout{said Mr~Casey,}
or the people may leave their church alone.

\dante:
You hear?

\N2:
said Dante,
turning to Mrs~Dedalus.

\marydedalus:
Mr~Casey!
Simon!
\sout{said Mrs~Dedalus,}
let it end now.

\charles:
Too bad!
Too bad!
\sout{said uncle Charles.}

\simon:
What?
\sout{cried Mr~Dedalus.}
Were we to desert him
at the bidding of the English people?

\dante:
He was no longer worthy to lead,
\sout{said Dante.}
He was a public sinner.

\casey:
We are all sinners and black sinners,
\sout{said Mr~Casey coldly.}

\dante:
Woe be to the man
by whom the scandal cometh!
\sout{said Mrs~Riordan.}
It would be better for him |
that a millstone were tied about his neck |
and that he were cast into the depths of the sea |
rather than that he should scandalise one of these,
my least little ones.
That is the language of the Holy Ghost.

\simon:
And very bad language
if you ask me,
\sout{said Mr~Dedalus coolly.}

\charles:
Simon!
Simon!
\sout{said uncle Charles.}
The boy.

\simon:
Yes, yes,
\sout{said Mr~Dedalus.}
I meant about the...
I was thinking about the bad language of the railway porter.
Well now, that’s all right.
Here, Stephen,
show me your plate,
old chap.
Eat away now.
Here.

\N1:
He heaped up the food on Stephen’s plate
and served uncle Charles and Mr~Casey
to large pieces of turkey and splashes of sauce.

\N2:
Mrs~Dedalus was eating little
and Dante sat with her hands in her lap.
She was red in the face.

\N1:
Mr~Dedalus rooted with the carvers at the end of the dish
and said:

\simon:
There’s a tasty bit here
we call the pope’s nose.
If any lady or gentleman...

\N1:
He held a piece of fowl up
on the prong of the carvingfork.
Nobody spoke.
He put it on his own plate,
saying:

\simon:
Well, you can’t say but you were asked.
I think I had better eat it myself
because I’m not well in my health lately.

\N1:
He winked at Stephen
and, replacing the dishcover,
began to eat again.

\N1:
There was a silence while he ate.
Then he said:

\simon:
Well now, the day kept up fine after all.
There were plenty of strangers down too.

\N1:
Nobody spoke.
He said again:

\simon:
I think there were more strangers down than last Christmas.

\N1:
He looked round at the others
whose faces were bent towards their plates |
and, receiving no reply,
waited for a moment and said bitterly:

\simon:
Well, my Christmas dinner has been spoiled anyhow.

\dante:
There could be neither luck nor grace,
\sout{Dante said,}
in a house where there is no respect for the pastors of the church.

\N1:
Mr~Dedalus threw his knife and fork noisily on his plate.

\simon:
Respect!
\sout{he said.}
Is it for Billy with the lip
or for the tub of guts up in Armagh?
Respect!

\casey:
Princes of the church,
\sout{said Mr~Casey with slow scorn.}

\simon:
Lord Leitrim’s coachman,
yes,
\sout{said Mr~Dedalus.}

\dante:
They are the Lord’s anointed,
\sout{Dante said.}
They are an honour to their country.

\simon:
Tub of guts,
\sout{said Mr~Dedalus coarsely.}
He has a handsome face,
mind you,
in repose.
You should see that fellow
lapping up his bacon and cabbage
of a cold winter’s day.
O Johnny!

\N1:
He twisted his features
into a grimace of heavy bestiality |
and made a lapping noise with his lips.

\marydedalus:
Really, Simon,
you should not speak that way before Stephen.
It’s not right.

\dante:
O, he’ll remember all this when he grows up,
\sout{said Dante hotly,}
the language he heard against God and religion and priests
in his own home.

\casey:
Let him remember too,

\N1:
cried Mr~Casey to her from across the table,

\casey:
the language with which the priests
and the priests’ pawns
broke Parnell’s heart
and hounded him into his grave.
Let him remember that too when he grows up.

\simon:
Sons of bitches!
\sout{cried Mr~Dedalus.}
When he was down
they turned on him
to betray him and rend him like rats in a sewer.
Lowlived dogs!
And they look it!
By Christ, they look it!

\dante:
They behaved rightly,
\sout{cried Dante.}
They obeyed their bishops and their priests.
Honour to them!

\marydedalus:
Well, it is perfectly dreadful
to say that not even for one day in the year,
\sout{said Mrs~Dedalus,}
can we be free from these dreadful disputes!

\N1:
Uncle Charles raised his hands mildly and said:

\charles:
Come now, come now, come now!
Can we not have our opinions
whatever they are
without this bad temper and this bad language?
It is too bad surely.

\N2:
Mrs~Dedalus spoke to Dante in a low voice
but Dante said loudly:

\dante:
I will not say nothing.
I will defend my church and my religion
when it is insulted and spit on
by renegade catholics.

\N1:
Mr~Casey pushed his plate rudely
into the middle of the table
and, resting his elbows before him,
said in a hoarse voice to his host:

\casey:
Tell me, did I tell you that story
about a very famous spit?

\simon:
You did not, John,
\sout{said Mr~Dedalus.}

\casey:
Why then,
\sout{said Mr~Casey,}
it is a most instructive story.
It happened not long ago in the county Wicklow
where we are now.

\N1:
He broke off
and, turning towards Dante,
said with quiet indignation:

\casey:
And I may tell you, ma’am,
that I, if you mean me,
am no renegade catholic.
I am a catholic as my father was
and his father before him
and his father before him again |
when we gave up our lives
rather than sell our faith.

\dante:
The more shame to you now,
\sout{Dante said,}
to speak as you do.

\simon:
The story, John,
\sout{said Mr~Dedalus smiling.}
Let us have the story anyhow.

\dante:
Catholic indeed!
\sout{repeated Dante ironically.}
The blackest protestant in the land
would not speak the language I have heard this evening.

\N1:
Mr~Dedalus began to sway his head to and fro,
crooning like a country singer.

\casey:
I am no protestant,
I tell you again,
\sout{said Mr~Casey,
flushing.}

\N1:
Mr~Dedalus,
still crooning and swaying his head,
began to sing in a grunting nasal tone:

\simon:
O, come all you Roman catholics \\
That never went to mass.

\N1:
He took up his knife and fork again
in good humour
and set to eating,
saying to Mr~Casey:

\simon:
Let us have the story, John.
It will help us to digest.


\begin{comment}

    \N2:
    Stephen looked with affection
    at Mr~Casey’s face
    which stared across the table over his joined hands.
    He liked to sit near him at the fire,
    looking up at his dark fierce face.
    But his dark eyes were never fierce
    and his slow voice was good to listen to.
    But why was he then against the priests?
    Because Dante must be right then.
    But he had heard his father say
    that she was a spoiled nun
    and that she had come out of the convent in the Alleghanies
    when her brother had got the money from the savages
    for the trinkets and the chainies.
    Perhaps that made her severe against Parnell.
    And she did not like him to play with Eileen
    because Eileen was a protestant
    and when she was young
    she knew children that used to play with protestants
    and the protestants used to make fun
    of the litany of the Blessed Virgin.
    Tower of Ivory,
    they used to say,
    House of Gold!

    \StephenInt:
    How could a woman be a tower of ivory
    or a house of gold?
    Who was right then?

    \N2:
    And he remembered the evening in the infirmary in Clongowes,
    the dark waters,
    the light at the pierhead
    and the moan of sorrow from the people when they had heard.

    \StephenInt:
    Eileen had long white hands.
    One evening
    when playing tig
    she had put her hands over his eyes:
    long and white and thin and cold and soft.
    That was ivory: a cold white thing.
    That was the meaning of Tower of Ivory.

\end{comment}


\casey:
The story is very short and sweet,
\sout{Mr~Casey said.}
It was one day down in Arklow,
a cold bitter day,
not long before the chief died.
May God have mercy on him!

\N1:
He closed his eyes wearily and paused.
Mr~Dedalus took a bone from his plate
and tore some meat from it with his teeth,
saying:

\simon:
Before he was killed,
you mean.

\N1:
Mr~Casey opened his eyes,
sighed
and went on:

\casey:
It was down in Arklow one day.
We were down there at a meeting
and after the meeting was over
we had to make our way to the railway station through the crowd.
Such booing and baaing, man, you never heard.
They called us all the names in the world.
Well there was one old lady,
and a drunken old harridan she was surely,
that paid all her attention to me.
She kept dancing along beside me in the mud
bawling and screaming into my face:
\stage{[as old harridan]}
Priesthunter!
The Paris Funds!
Mr~Fox!
Kitty O’Shea!

\simon:
And what did you do, John?
\sout{asked Mr~Dedalus.}

\casey:
I let her bawl away,
\sout{said Mr~Casey.}
It was a cold day
and to keep up my heart I had
(saving your presence, ma’am)
a quid of Tullamore in my mouth
and sure I couldn’t say a word in any case
because my mouth was full of tobacco juice.

\simon:
Well, John?

\casey:
Well.
I let her bawl away,
to her heart’s content,
Kitty O’Shea and the rest of it
till at last she called that lady
a name that I won’t sully this Christmas board
nor your ears, ma’am,
nor my own lips
by repeating.

\N1:
He paused.
Mr~Dedalus,
lifting his head from the bone,
asked:

\simon:
And what did you do, John?

\casey:
Do!
\sout{said Mr~Casey.}
She stuck her ugly old face up at me
when she said it
and I had my mouth full of tobacco juice.
I bent down to her
and Phth!
says I to her
like that.

\N1:
He turned aside
and made the act of spitting.

\casey:
Phth!
says I to her like that,
right into her eye.

\N1:
He clapped his hand to his eye
and gave a hoarse scream of pain.

\casey:
\stage{[as old harridan]}
O Jesus, Mary and Joseph!
says she.
I’m blinded! I’m blinded and drownded!

\N1:
He stopped in a fit of coughing and laughter,
repeating:

\casey:
\stage{[as old harridan]}
I’m blinded entirely.

\N1:
Mr~Dedalus laughed loudly
and lay back in his chair
while uncle Charles swayed his head to and fro.

\N2:
Dante looked terribly angry
and repeated while they laughed:

\dante:
Very nice!
Ha!
Very nice!


\begin{comment}

    \StephenInt:
    It was not nice
    about the spit in the woman’s eye.

    \StephenInt:
    But what was the name the woman had called Kitty O’Shea
    that Mr~Casey would not repeat?

    \N2:
    He thought of Mr~Casey walking through the crowds of people
    and making speeches from a wagonette.
    That was what he had been in prison for
    and he remembered that one night
    Sergeant O’Neill had come to the house
    and had stood in the hall,
    talking in a low voice with his father
    and chewing nervously at the chinstrap of his cap.
    And that night
    Mr~Casey had not gone to Dublin by train
    but a car had come to the door
    and he had heard his father say something
    about the Cabinteely road.

    \N2:
    He was for Ireland and Parnell and so was his father:
    and so was Dante too
    for one night at the band on the esplanade
    she had hit a gentleman on the head with her umbrella
    because he had taken off his hat
    when the band played God save the Queen at the end.

\end{comment}


\N1:
Mr~Dedalus gave a snort of contempt.

\simon:
Ah, John,
\sout{he said.}
It is true for them.
We are an unfortunate priest\-ridden race
and always were
and always will be
till the end of the chapter.

\N1:
Uncle Charles shook his head,
saying:

\charles:
A bad business!
A bad business!

\N1:
Mr~Dedalus repeated:

\simon:
A priest\-ridden Godforsaken race!

\N1:
He pointed to the portrait of his grandfather
on the wall to his right.

\simon:
Do you see that old chap up there, John?
\sout{he said.}
He was a good Irishman when there was no money in the job.
He was condemned to death as a whiteboy.
But he had a saying about our clerical friends,
that he would never let one of them
put his two feet under his mahogany.

\N2:
Dante broke in angrily:

\dante:
If we are a priest\-ridden race
we ought to be proud of it!
They are the apple of God’s eye.
Touch them not,
says Christ,
for they are the apple of My eye.

\casey:
And can we not love our country then?
\sout{asked Mr~Casey.}
Are we not to follow the man that was born to lead us?

\dante:
A traitor to his country!
\sout{replied Dante.}
A traitor, an adulterer!
The priests were right to abandon him.
The priests were always the true friends of Ireland.

\casey:
Were they, faith?
\sout{said Mr~Casey.}

\N1:
He threw his fist on the table
and, frowning angrily,
protruded one finger after another.

\casey:
Didn’t the bishops of Ireland
betray us in the time of the union
when Bishop Lanigan presented an address of loyalty
to the Marquess Cornwallis?
Didn’t the bishops and priests
sell the aspirations of their country in 1829 |
in return for catholic emancipation?
Didn’t they denounce the fenian movement
from the pulpit and in the confession box?
And didn’t they dishonour the ashes of Terence Bellew MacManus?

\N1:
His face was glowing with anger
and Stephen felt the glow rise to his own cheek
as the spoken words thrilled him.
Mr~Dedalus uttered a guffaw of coarse scorn.

\simon:
O, by God,
\sout{he cried,}
I forgot little old Paul Cullen!
Another apple of God’s eye!

\N2:
Dante bent across the table and cried to Mr~Casey:

\dante:
Right!
Right!
They were always right!
God and morality and religion come first.

\N2:
Mrs~Dedalus,
seeing her excitement,
said to her:

\marydedalus:
Mrs~Riordan,
don’t excite yourself answering them.

\dante:
God and religion before everything!
\sout{Dante cried.}
God and religion before the world.

\N1:
Mr~Casey raised his clenched fist
and brought it down on the table with a crash.

\casey:
Very well then,
\sout{he shouted hoarsely,}
if it comes to that,
no God for Ireland!

\simon:
John!
John!

\N1:
cried Mr~Dedalus,
seizing his guest by the coat sleeve.

\N2:
Dante stared across the table,
her cheeks shaking.

\N1:
Mr~Casey struggled up from his chair
and bent across the table towards her,
scraping the air from before his eyes with one hand
as though he were tearing aside a cobweb.

\casey:
No God for Ireland!
\sout{he cried.}
We have had too much God in Ireland.
Away with God!

\dante:
Blasphemer!
Devil!

\N2:
screamed Dante,
starting to her feet
and almost spitting in his face.

\N1:
Uncle Charles and Mr~Dedalus
pulled Mr~Casey back into his chair again,
talking to him from both sides reasonably.
He stared before him out of his dark flaming eyes,
repeating:

\casey:
Away with God,
I say!

\N2:
Dante shoved her chair violently aside
and left the table,
upsetting her napkinring
which rolled slowly along the carpet
and came to rest against the foot of an easychair.
Mrs~Dedalus rose quickly and followed her towards the door.
At the door Dante turned round violently
and shouted down the room,
her cheeks flushed and quivering with rage:

\dante:
Devil out of hell!
We won!
We crushed him to death!
Fiend!

\N2:
The door slammed behind her.

\N1:
Mr~Casey,
freeing his arms from his holders,
suddenly bowed his head on his hands with a sob of pain.

\casey:
Poor Parnell!
\sout{he cried loudly.}
My dead king!

\N1:
He sobbed loudly and bitterly.

\N2:
Stephen,
raising his terrorstricken face,
saw that his father’s eyes were full of tears.

\end{document}
