\section*{Rock I: Father Conmee}

\N1:
The superior, the very reverend John Conmee~S.J. |
reset his smooth watch in his interior pocket |
as he came down the presbytery steps.

\conmeeint:
Five to three.
Just nice time to walk to Artane.
What is that boy's name again?
Dignam.
Yes.
\latin{Vere dignum et iustum est.}
Brother Swan is the person to see.
Mr Cunningham's letter.
Yes.
Oblige him, if possible.
Good practical catholic:
useful at mission time.

\stage{Enter sailor}

\N3:
A onelegged sailor,
swinging himself onward by lazy jerks of his crutches,
growled some notes.
He jerked short before the convent of the sisters of charity |
and held out a peaked cap for alms
towards the very reverend John Conmee~S.J.
\gab{10}

\N1:
Father Conmee blessed him in the sun |
for his purse held,
he knew,
one silver crown.

\stage{Exit sailor}

\N1:
Father Conmee crossed to Mountjoy square.
He thought,
but not for long, |
of soldiers and sailors,
whose legs had been shot off by cannonballs,
ending their days in some pauper ward, |
and of cardinal Wolsey's words:

\conmeeint:
\emph{If I had served my God |
as I have served my King |
He would not have abandoned me in my old days.}

\N1:
He walked by the treeshade of sunnywinking leaves:
and towards him |
came the wife of Mr David Sheehy~M.P.

\stage{Enter Mrs~Sheehy}

\mrssheehy:
Very well, indeed, father.
And you, father?

\N1:
Father Conmee was wonderfully well indeed.
He would go to Buxton probably for the waters.
\gab{20}

\conmee:
And your boys, are they getting on well at Belvedere?

\conmee:
Is that so?

\N1:
Father Conmee was very glad indeed to hear that.

\conmee:
And Mr Sheehy himself?

\mrssheehy:
Still in London.
The house is still sitting, to be sure it is.

\mrssheehy:
Beautiful weather it is, delightful indeed.

\conmee:
Yes, it is very probable that Father Bernard Vaughan will come again to preach.

\conmee:
O, yes: a very great success.
A wonderful man really.

\N1:
Father Conmee was very glad 
to see the wife of Mr David Sheehy~M.P. looking so well |
and he begged to be remembered to Mr David Sheehy~M.P.
Yes, he would certainly call.

\conmee:
Good afternoon, Mrs~Sheehy.

\stage{Exit Mrs~Sheehy}

\N1:
Father Conmee doffed his silk hat and smiled,
\gab{30}
as he took leave,
at the jet beads of her mantilla inkshining in the sun.
And smiled yet again, in going.
He had cleaned his teeth, he knew, with arecanut paste.

\N1:
Father Conmee walked |
and, walking, smiled |
for he thought on
Father Bernard Vaughan's droll eyes and cockney voice.

\conmeeint: \stage{[Cockney accent]\/}
Pilate!
Wy don't you old back that owlin mob?

\conmeeint:
A zealous man, however.
Really he is.
And really does great good in his way.
Beyond a doubt.
He loves Ireland, he says, and he loves the Irish.
Of good family too |
would one think it?
Welsh, are they not?

\conmeeint:
O, lest I forget.
That letter to father provincial.

\N1:
Father Conmee stopped three little schoolboys
\gab{40}
at the corner of Mountjoy square.

\Boys:
Yes: we are from Belvedere.
The little house.

\conmee:
Aha.
And are you good boys at school?

\conmee:
O.~That is very good now.
And what is your name?

\Boy{1}:
Jack Sohan.

\conmee:
And your name?

\Boy{2}:
Ger.~Gallaher.

\conmee:
And the other little man?

\Boy{3}:
My name is Brunny Lynam.

\conmee:
O,~that is a very nice name to have.

\N1:
Father Conmee gave a letter from his breast to Master Brunny Lynam
and pointed to the red pillarbox at the corner of Fitzgibbon street.

\conmee:
But mind you don't post yourself into the box, little man.

\N1:
The boys sixeyed Father Conmee and laughed:

\Boys:
O, sir.
\gab{50}

\conmee:
Well, let me see if you can post a letter.

\N{1a}:
Master Brunny Lynam ran across the road
and put Father Conmee's letter to father provincial
into the mouth of the bright red letterbox.

\N1:
Father Conmee smiled and nodded and smiled |
and walked along Mountjoy square east.

\begin{mdframed}
    \N{20}:
    Mr Denis J Maginni, |
    professor of dancing \&c, |
    in silk hat,
    slate frockcoat with silk facings,
    white kerchief tie,
    tight lavender trousers,
    canary gloves
    and pointed patent boots, |
    walking with grave deportment |
    most respectfully took the curbstone |
    as he passed lady Maxwell
    at the corner of Dignam's court.
    %\gab{60}
\end{mdframed}

\conmeeint:
Is that not Mrs M'Guinness?

\N1:
Mrs M'Guinness,
stately, silverhaired,
bowed to Father Conmee from the farther footpath
along which she sailed.
And Father Conmee smiled and saluted.

\conmee:
How do you do?

\conmeeint:
A fine carriage she has.
Like Mary, queen of Scots, something.
And to think that she is a pawnbroker!
Well, now!
Such a \ldots
what should I say? \ldots
such a queenly mien.

\N1:
Father Conmee walked down Great Charles street
and glanced at the shutup free church on his left.
The reverend T.R.~Greene~B.A. will (D.V.) speak.

\conmeeint:
The incumbent they call him.
\gab{70}
He feels it incumbent on him to say a few words.
But one should be charitable.
Invincible ignorance.
They act according to their lights.

\N1:
Father Conmee turned the corner
and walked along the North Circular road.

\conmeeint:
It is a wonder that there was not a tramline
in such an important thoroughfare.
Surely, there ought to be.

\N{1a}:
A band of satchelled schoolboys crossed from Richmond street.
All raised untidy caps.
Father Conmee greeted them more than once benignly.

\conmeeint:
Christian brother boys.

\N1:
Father Conmee smelt incense on his right hand as he walked.

\conmeeint:
Saint Joseph's church, Portland row.
\gab{80}
For aged and virtuous females.

\N1:
Father Conmee raised his hat to the Blessed Sacrament.

\conmeeint:
Virtuous: but occasionally they are also badtempered.

\N1:
Near Aldborough house |
Father Conmee thought of that spendthrift nobleman.

\conmeeint:
And now it is an office or something.

\N1:
Father Conmee began to walk along the North Strand road |
and was saluted by Mr William Gallagher
who stood in the doorway of his shop.
Father Conmee saluted Mr William Gallagher |
and perceived the odours
that came from baconflitches and ample cools of butter.
He passed Grogan's the Tobacconist |
against which newsboards leaned |
and told of a dreadful catastrophe
\gab{90}
in New York.

\conmeeint:
In America those things are continually happening.
Unfortunate people to die like that, unprepared.
Still, an act of perfect contrition.

\N{1a}:
Father Conmee went by Daniel Bergin's publichouse
against the window of which |
two unlabouring men lounged.
They saluted him and were saluted.

\N1:
Father Conmee passed H.J.~O'Neill's funeral establishment |
where Corny Kelleher totted figures in the daybook |
while he chewed a blade of hay.
A constable on his beat saluted Father Conmee
and Father Conmee saluted the constable.
In Youkstetter's, the porkbutcher's,
Father Conmee observed pig's puddings,
\gab{100}
white and black and red,
lie neatly curled in tubes.

\N{1a}:
Moored under the trees of Charleville Mall |
Father Conmee saw a turfbarge,
a towhorse with pendent head,
a bargeman with a hat of dirty straw seated amidships,
smoking and staring at a branch of poplar above him.
It was idyllic:
and Father Conmee reflected on the providence of the Creator
who had made turf to be in bogs |
whence men might dig it out and bring it to town and hamlet |
to make fires in the houses of poor people.

\N1:
On Newcomen bridge |
the very reverend John Conmee~S.J. |
of saint Francis Xavier's church, upper Gardiner street,
stepped on to an outward bound tram.

\N{19}:
Off an inward bound tram |
\gab{110}
stepped the reverend Nicholas Dudley~C.C.
of saint Agatha's church, north William street,
on to Newcomen bridge.

\N1:
At Newcomen bridge |
Father Conmee stepped into an outward bound tram |
for he disliked to traverse on foot the dingy way past Mud Island.

\N1:
Father Conmee sat in a corner of the tramcar, |
a blue ticket tucked with care in the eye of one plump kid glove, |
while four shillings, a sixpence and five pennies |
chuted from his other plump glovepalm into his purse.
Passing the ivy church |
he reflected that the ticket inspector usually made his visit |
when one had carelessly thrown away the ticket.
The solemnity of the occupants of the car
\gab{120}
seemed to Father Conmee excessive |
for a journey so short and cheap.
Father Conmee liked cheerful decorum.

\N{1a}:
It was a peaceful day.
The gentleman with the glasses opposite Father Conmee
had finished explaining and looked down.
His wife, Father Conmee supposed.
A tiny yawn opened the mouth of the wife of the gentleman with the glasses.
She raised her small gloved fist,
yawned ever so gently,
tiptapping her small gloved fist on her opening mouth |
and smiled tinily, sweetly.

\N{1a}:
Father Conmee perceived her perfume in the car.
He perceived also that the awkward man at the other side of her |
was sitting on the edge of the seat.%
\gab{130}

\N{1b}:
Father Conmee at the altarrails |
placed the host with difficulty |
in the mouth of the awkward old man who had the shaky head.

\N{1a}:
At Annesley bridge |
the tram halted and, when it was about to go, |
an old woman rose suddenly from her place to alight.
The conductor pulled the bellstrap to stay the car for her.
She passed out with her basket and a marketnet:
and Father Conmee saw the conductor help her and net and basket down:
and Father Conmee thought
that, as she had nearly passed the end of the penny fare,

\conmeeint:
she is one of those good souls who have always to be told twice
\emph{bless you, my child},
that they have been absolved,
\emph{pray for me}.
But they have so many worries in life,
\gab{140}
so many cares, poor creatures.

\N1:
From the hoardings |
Mr Eugene Stratton grimaced with thick niggerlips at Father Conmee.

\N{1a}:
Father Conmee thought of the souls of black and brown and yellow men |
and of his sermon on saint Peter Claver~S.J. and the African mission |

\conmeeint:
and of the propagation of the faith |
and of the millions of black and brown and yellow souls |
that have not received the baptism of water |
when their last hour comes like a thief in the night.
That book by the Belgian jesuit,
\french{Le Nombre des Elus},
seems a reasonable plea.
Those are millions of human souls created by God in His Own likeness |
to whom the faith has not (D.V.) been brought.
\gab{150}
But they are God's souls, created by God.
It seems a pity that they should all be lost, |
a waste, if one might say.

\N1:
At the Howth road stop |
Father Conmee alighted,
was saluted by the conductor |
and saluted in his turn.

\N1:
The Malahide road was quiet.
It pleased Father Conmee, road and name.

\conmeeint:
The joybells were ringing in gay Malahide.
Lord Talbot de Malahide,
immediate hereditary lord admiral of Malahide and the seas adjoining.
Then came the call to arms |
and she was maid, wife and widow in one day.
Those were old worldish days,
loyal times in joyous townlands,
old times in the barony.%
\gab{160}

\N1:
Father Conmee, walking, thought of his little book
\Title{Old Times in the Barony}
and of the book that might be written about Jesuit houses
and of Mary Rochfort,
daughter of lord Molesworth,
first countess of Belvedere.

\conmeeint:
A listless lady, no more young,
walked alone the shore of lough Ennel,
Mary, first countess of Belvedere,
listlessly walking in the evening,
not startled when an otter plunged.
Who could know the truth?
Not the jealous lord Belvedere |
and not her confessor |
if she had not committed adultery fully,
\latin{eiaculatio seminis inter vas naturale mulieris},
with her husband's brother?
She would half confess if she had not all sinned |
as women do.
\gab{170}
Only God knows |
and she |
and he, her husband's brother.

\N{1a}:
Father Conmee thought of that tyrannous incontinence,
needed however for man's race on earth,
and of the ways of God which were not our ways.

\N{1b}:
Don John Conmee walked and moved in times of yore.
He was humane and honoured there.
He bore in mind secrets confessed
and he smiled at smiling noble faces in a beeswaxed drawingroom,
ceiled with full fruit clusters.
And the hands of a bride and of a bridegroom,
noble to noble,
were impalmed by Don John Conmee.

\conmeeint:
It is a charming day.

\N{1a}:
The lychgate of a field showed Father Conmee |
\gab{180}
breadths of cabbages,
curtseying to him with ample underleaves.
The sky showed him a flock of small white clouds |
going slowly down the wind.

\conmeeint:
\french{Moutonner}, the French say.
A just and homely word.

\N{1b}:
Father Conmee, reading his office, |
watched a flock of muttoning clouds over Rathcoffey.
His thinsocked ankles were tickled by the stubble of Clongowes field.
He walked there, reading in the evening,
and heard the cries of the boys' lines at their play,
young cries in the quiet evening.
He was their rector: his reign was mild.

\N{1a}:
Father Conmee drew off his gloves
and took his rededged breviary out.
\gab{190}
An ivory bookmark told him the page.

\conmeeint:
Nones.
I should have read that before lunch.
But lady Maxwell came.

\N1:
Father Conmee read in secret \latin{Pater} and \latin{Ave}
and crossed his breast.

\conmeeint:
\latin{Deus in adiutorium.}

\N1:
He walked calmly and read mutely the nones,
walking and reading till
he came to
\latin{Res} in

\conmeeint:
\latin{beati immaculati:
principium verborum tuorum veritas:
in eternum omnia indicia iustitiæ tuæ.}

\N{1a}:
A flushed young man came from a gap of a hedge |
and after him came a young woman
with wild nodding daisies in her hand.
\gab{200}
The young man raised his cap abruptly:
the young woman abruptly bent
and with slow care |
detached from her light skirt a clinging twig.

\N{1a}:
Father Conmee blessed both gravely
and turned a thin page of his breviary.

\conmeeint:
\latin{Sin:
principes persecuti sunt me gratis:
et a verbis tuis formidavit cor meum.}


\section*{Rock II: Corny Kelleher}

\N2:
Corny Kelleher closed his long daybook
and glanced with his drooping eye
at a pine coffinlid sentried in a corner.
He pulled himself erect,
went to it |
and, spinning it on its axle, |
viewed its shape and brass furnishings.
\gab{210}
Chewing his blade of hay |
he laid the coffinlid by and came to the doorway.
There |
he tilted his hatbrim to give shade to his eyes
and leaned against the doorcase,
looking idly out.

\begin{mdframed}
    \N1:
    Father John Conmee stepped into the Dollymount tram on Newcomen bridge.
\end{mdframed}

\N2:
Corny Kelleher locked his largefooted boots and gazed,
his hat downtilted,
chewing his blade of hay.

\N2:
Constable 57C,
on his beat,
stood to pass the time of day.

\constable:
That's a fine day, Mr Kelleher.

\corny:
Ay, \sout{Corny Kelleher said}.

\constable:
It's very close,
\gab{220}
\sout{the constable said.}

\N2:
Corny Kelleher sped a silent jet of hayjuice arching from his mouth

\begin{mdframed}
    \N3:
    while a generous white arm from a window in Eccles street |
    flung forth a coin.
\end{mdframed}

\corny:
What's the best news? \sout{he asked.}

\constable:
I seen that particular party last evening,

\N2:
the constable said with bated breath.


\section*{Rock III: For England, Home and Beauty}

\N3:
A onelegged sailor crutched himself round MacConnell's corner,
skirting Rabaiotti's icecream car,
and jerked himself up Eccles street.
Towards Larry O'Rourke,
\gab{230}
in shirtsleeves in his doorway,
he growled unamiably:

\sailor:
For England \ldots

\N3:
He swung himself violently forward past Katey and Boody Dedalus,
halted and growled:

\sailor:
Home and beauty.

\begin{mdframed}
    \N8:
    J.J.~O'Molloy's white careworn face was told |
    that Mr Lambert was in the warehouse with a visitor.
\end{mdframed}

\N3:
A stout lady stopped,
took a copper coin from her purse |
and dropped it into the cap held out to her.
The sailor grumbled thanks,
glanced sourly at the unheeding windows,
\gab{240}
sank his head
and swung himself forward four strides.

\N3:
He halted and growled angrily:

\sailor:
For England \ldots

\N3:
Two barefoot urchins,
sucking long liquorice laces,
halted near him,
gaping at his stump with their yellowslobbered mouths.

\N3:
He swung himself forward in vigorous jerks,
halted,
lifted his head towards a window
and bayed deeply:

\sailor:
Home and beauty.

\N3:
The gay sweet chirping whistling within |
went on a bar or two, ceased.
The blind of the window was drawn aside.
\gab{250}
A card \Title{Unfurnished Apartments}
slipped from the sash and fell.
A plump bare generous arm shone,
was seen,
held forth from a white petticoatbodice and taut shiftstraps.
A woman's hand flung forth a coin
over the area railings.
It fell on the path.

\N3:
One of the urchins ran to it,
picked it up and dropped it into the minstrel's cap, |
saying:

\boy{1}:
There, sir.


\section*{Rock IV: Katey, Boody, and Maggy Dedalus}

\N4:
Katey and Boody Dedalus shoved in the door
of the closesteaming kitchen.

\boody:
Did you put in the books?
\gab{260}
Boody asked.

\N4:
Maggy at the range |
rammed down a greyish mass |
beneath bubbling suds |
twice with her potstick |
and wiped her brow.

\maggy:
They wouldn't give anything on them,
\sout{she said.}

\begin{mdframed}
    \N{1b}:
    Father Conmee walked through Clongowes fields,
    his thinsocked ankles tickled by stubble.
\end{mdframed}

\boody:
Where did you try?
\sout{Boody asked.}

\maggy:
M'Guinness's.

\N4:
Boody stamped her foot and threw her satchel on the table.

\boody:
Bad cess to her big face!
\sout{she cried.}

\N4:
Katey went to the range
\gab{270}
and peered with squinting eyes.

\katey:
What's in the pot?
\sout{she asked.}

\maggy:
Shirts,
\sout{Maggy said.}

\N4:
Boody cried angrily:

\boody:
Crickey, is there nothing for us to eat?

\N4:
Katey,
lifting the kettlelid in a pad of her stained skirt,
asked:

\katey:
And what's in this?

\N4:
A heavy fume gushed in answer.

\maggy:
Peasoup,
\sout{Maggy said.}

\katey:
Where did you get it?
\sout{Katey asked.}

\maggy:
Sister Mary Patrick,
\gab{280}
\sout{Maggy said.}

\N{11}:
The lacquey rang his bell.

\stage{Barang!}

\N4:
Boody sat down at the table and said hungrily:

\boody:
Give us it here.

\N4:
Maggy poured yellow thick soup from the kettle into a bowl.
Katey,
sitting opposite Boody,
said quietly, |
as her fingertip lifted to her mouth |
random crumbs:

\katey:
A good job we have that much.
Where's Dilly?

\maggy:
Gone to meet father,
\sout{Maggy said.}

\N4:
Boody,
breaking big chunks of bread into the yellow soup,
\gab{290}
added:

\boody:
Our father who art not in heaven.

\N4:
Maggy,
pouring yellow soup in Katey's bowl,
exclaimed:

\maggy:
Boody!
For shame!

\begin{mdframed}
    \N{20}:
    A skiff,
    a crumpled throwaway,
    Elijah is coming,
    rode lightly down the Liffey,
    under Loopline bridge,
    shooting the rapids where water chafed around the bridgepiers,
    sailing eastward past hulls and anchorchains,
    between the Customhouse old dock and George's quay.
\end{mdframed}


\section*{Rock V: Blazes Boylan}

\N5:
The blond girl in Thornton's |
bedded the wicker basket with rustling fibre.
\gab{300}
Blazes Boylan handed her the bottle swathed in pink tissue paper |
and a small jar.

\boylan:
Put these in first, will you?
\sout{he said.}

\blondgirl:
Yes, sir,

\N5:
the blond girl said.

\blondgirl:
And the fruit on top.

\boylan:
That'll do, game ball,
\sout{Blazes Boylan said.}

\N5:
She bestowed fat pears neatly,
head by tail,
and among them ripe shamefaced peaches.

\N5:
Blazes Boylan walked here and there in new tan shoes
about the fruitsmelling shop,
lifting fruits,
young juicy crinkled and plump red tomatoes,
sniffing smells.

\begin{mdframed}
    \N{20}:
    Haitch Ee Ell Wye Apostrophe Ess \stage{[H.E.L.Y.'S]}
    filed before him, tallwhitehatted,
    %\gab{310}
    past Tangier lane,
    plodding towards their goal.
\end{mdframed}

\N5:
He turned suddenly from a chip of strawberries,
drew a gold watch from his fob |
and held it at its chain's length.

\boylan:
Can you send them by tram? Now?

\begin{mdframed}
    \N{10}:
    A darkbacked figure under Merchants' arch
    scanned books on the hawker's cart.
\end{mdframed}

\blondgirl:
Certainly, sir.
Is it in the city?

\boylan:
O, yes,
\sout{Blazes Boylan said.}
Ten minutes.

\N5:
The blond girl handed him a docket and pencil.

\blondgirl:
Will you write the address,
\gab{320}
sir?

\N5:
Blazes Boylan at the counter wrote
and pushed the docket to her.

\boylan:
Send it at once, will you?
\sout{he said.}
It's for an invalid.

\blondgirl:
Yes, sir. I will, sir.

\N5:
Blazes Boylan rattled merry money in his trousers' pocket.

\boylan:
What's the damage?
\sout{he asked.}

\N5:
The blond girl's slim fingers reckoned the fruits.

\N5:
Blazes Boylan looked into the cut of her blouse.

\boylanint:
A young pullet.

\N5:
He took a red carnation from the tall stemglass.

\boylan:
This for me?
\sout{he asked gallantly.}

\N5:
The blond girl glanced sideways at him,
\gab{330}
got up regardless,
with his tie a bit crooked,
blushing.

\blondgirl:
Yes, sir,
\sout{she said.}

\N5:
Bending archly |
she reckoned again fat pears and blushing peaches.

\N5:
Blazes Boylan looked in her blouse with more favour,
the stalk of the red flower between his smiling teeth.

\boylan:
May I say a word to your telephone, missy?
\sout{he asked roguishly.}


\section*{Rock VI: Stephen Dedalus + Almidano Artifoni}


\artifoni:
\italian{Ma!}

\N6:
Almidano Artifoni said.

\N6:
He gazed over Stephen's shoulder at Goldsmith's knobby poll.

\N6:
Two carfuls of tourists passed slowly,
\gab{340}
their women sitting fore,
gripping the handrests.

\StephenInt:
Palefaces.

\N6:
Men's arms frankly round their stunted forms.
They looked from Trinity
to the blind columned porch of the bank of Ireland
where pigeons roocoocooed.

\artifoni:
\italian{Anch'io ho avuto di queste idee,}
\sout{Almidano Artifoni said,}
\italian{Quand' ero giovine come lei.
Eppoi mi sono convinto che il mondo e una bestia.
Peccato.
Perche la sua voce \ldots
sarebbe un cespite di rendita, via.
Invece, lei si sacrifica.}

\Stephen:
\italian{Sacrifizio incruento,}

\N6:
Stephen said smiling,
swaying his ashplant in slow swingswong from its midpoint,
lightly.

\artifoni:
\italian{Speriamo,}
\sout{the round mustachioed face said pleasantly.}
\gab{350}
\italian{Ma, dia retta a me.
Ci rifletta.}

\N6:
By the stern stone hand of Grattan,
bidding halt,
an Inchicore tram unloaded straggling Highland soldiers of a band.

\Stephen:
\italian{Ci riflettero,}

\N6:
Stephen said,
glancing down the solid trouserleg.

\artifoni:
\italian{Ma, sul serio, eh?}
\sout{Almidano Artifoni said.}

\N6:
His heavy hand took Stephen's firmly.

\StephenInt:
Human eyes.

\N6:
They gazed curiously an instant
and turned quickly towards a Dalkey tram.

\artifoni:
\italian{Eccolo,}

\sout{Almidano Artifoni said in friendly haste.}

\artifoni:
\italian{Venga a trovarmi e ci pensi.
Addio, caro.}

\Stephen:
\italian{Arrivederla, maestro,}

\N6:
Stephen said,
\gab{360}
raising his hat when his hand was freed.

\Stephen:
\italian{Grazie.}

\artifoni:
\italian{Di che?}
\sout{Almidano Artifoni said.}
\italian{Scusi, eh?
Tante belle cose!}

\N6:
Almidano Artifoni,
holding up a baton of rolled music as a signal,
trotted on stout trousers after the Dalkey tram.
In vain he trotted,
signalling in vain |
among the rout of barekneed gillies |
smuggling implements of music through Trinity gates.


\section*{Rock VII: Miss Dunne}


\N7:
Miss Dunne hid the Capel street library copy
of \Title{The Woman in White} far back in her drawer |
and rolled a sheet of gaudy notepaper into her typewriter.%
\gab{370}

\dunneint:
Too much mystery business in it.
Is he in love with that one, Marion?
Change it and get another by Mary Cecil Haye.
 
\begin{mdframed}
    \N9:
    The disk shot down the groove,
    wobbled a while,
    ceased and ogled them: six.
\end{mdframed}

\N7:
Miss Dunne clicked on the keyboard:

\missdunne:
16 June 1904.

\begin{mdframed}
    \N{20}:
    Five tallwhitehatted sandwichmen
    between Monypeny's corner and the slab where Wolfe Tone's statue was not,
    eeled themselves turning
    Haitch Ee Ell Wye Apostrophe Ess \stage{[H.E.L.Y.'S]}
    and plodded back as they had come.
\end{mdframed}

\N7:
Then she stared at the large poster of Marie Kendall,
\gab{380}
charming soubrette,
and,
listlessly lolling,
scribbled on the jotter|
sixteens and capital esses.

\dunneint:
Mustard hair and dauby cheeks.
She's not nicelooking, is she?
The way she's holding up her bit of a skirt.
Wonder will that fellow be at the band tonight.
If I could get that dressmaker
to make a concertina skirt like Susy Nagle's.
They kick out grand.
Shannon and all the boatclub swells never took his eyes off her.
Hope to goodness
he won't keep me here till seven.

\N7:
The telephone rang rudely by her ear.

\missdunne:
Hello.
Yes, sir.
No, sir.
Yes, sir.
I'll ring them up after five.
Only those two, sir, for Belfast and Liverpool.
\gab{390}
All right, sir.
Then I can go after six if you're not back.
A quarter after.
Yes, sir.
Twentyseven and six.
I'll tell him.
Yes: one, seven, six.

\N7:
She scribbled three figures on an envelope.

\missdunne:
Mr Boylan!
Hello!
That gentleman from \Title{Sport} was in looking for you.
Mr Lenehan, yes.
He said he'll be in the Ormond at four.
No, sir.
Yes, sir.
I'll ring them up after five.


\section*{Rock VIII: Lambert and O'Molloy}


\N8:
Two pink faces turned in the flare of the tiny torch.

\lambert:
Who's that?

\N8:
Ned Lambert asked.

\lambert:
Is that Crotty?

\jjom:
Ringabella and Crosshaven,
\gab{400}

\N8:
a voice replied groping for foothold.

\lambert:
Hello, Jack, is that yourself?

\N8:
Ned Lambert said,
raising in salute his pliant lath
among the flickering arches.

\lambert:
Come on.
Mind your steps there.

\N8:
The vesta in the clergyman's uplifted hand consumed itself
in a long soft
flame and was let fall.
At their feet its red speck died:
and mouldy air
closed round them.

\revlove:
How interesting!

\N8:
a refined accent said in the gloom.

\lambert:
Yes, sir,
\sout{Ned Lambert said heartily.}
We are standing in the historic council chamber of saint Mary's abbey
where silken Thomas proclaimed himself a rebel in 1534.
This is the most historic spot in all Dublin.
O'Madden Burke is going to write something about it
\gab{410}
one of these days.
The old bank of Ireland was over the way
till the time of the union
and the original jews' temple was here too
before they built their synagogue over in Adelaide road.
You were never here before, Jack,
were you?

\jjom:
No, Ned.

\revlove:
He rode down through Dame walk,
\sout{the refined accent said,}
if my memory serves me.
The mansion of the Kildares was in Thomas court.

\lambert:
That's right,
\sout{Ned Lambert said.}
That's quite right, sir.

\revlove:
If you will be so kind then,
\sout{the clergyman said,}
the next time to allow
me perhaps \ldots

\lambert:
Certainly,
\sout{Ned Lambert said.}
\gab{420}
Bring the camera whenever you like.
I'll get those bags cleared away from the windows.
You can take it from here or from here.

\N8:
In the still faint light |
he moved about,
tapping with his lath |
the piled seedbags and points of vantage on the floor.

\begin{mdframed}
    \N{16}:
    From a long face
    a beard and gaze hung on a chessboard.
\end{mdframed}

\revlove:
I'm deeply obliged, Mr Lambert,
\sout{the clergyman said.}
I won't trespass on your valuable time \ldots

\lambert:
You're welcome, sir,
\sout{Ned Lambert said.}
Drop in whenever you like.
Next week, say.
Can you see?

\revlove:
Yes, yes.
\gab{430}
Good afternoon, Mr Lambert.
Very pleased to have met you.

\lambert:
Pleasure is mine, sir,
\sout{Ned Lambert answered.}

\N8:
He followed his guest to the outlet |
and then whirled his lath away among the pillars.
With J.J.~O'Molloy |
he came forth slowly into Mary's abbey |
where draymen were loading floats
with sacks of carob and palmnut meal, |
O'Connor, Wexford.

\N8:
He stood to read the card in his hand.

\lambert:
The reverend Hugh~C. Love, Rathcoffey.
Present address:
Saint Michael's, Sallins.
Nice young chap he is.
He's writing a book about the Fitzgeralds he told me.
He's well up in history, faith.

\begin{mdframed}
    \N{1a}:
    The young woman with slow care
    %\gab{440}
    detached from her light skirt a clinging twig.
\end{mdframed}

\jjom:
I thought you were at a new gunpowder plot,
\sout{J.J.~O'Molloy said.}

\N8:
Ned Lambert cracked his fingers in the air.

\lambert:
God!
\sout{he cried.}
I forgot to tell him that one about the earl of Kildare |
after he set fire to Cashel cathedral.
You know that one?
\emph{I'm bloody sorry I did it,}
says he,
\emph{but I declare to God I thought the archbishop was inside.}
He mightn't like it, though.
What?
God, I'll tell him anyhow.
That was the great earl, the Fitzgerald Mor.
Hot members they were |
all of them, the Geraldines.

\N8:
The horses he passed
started nervously
\gab{450}
under their slack harness.
He slapped a piebald haunch quivering near him and cried:

\lambert:
Woa, sonny!

\N8:
He turned to J.J.~O'Molloy and asked:

\lambert:
Well, Jack.
What is it?
What's the trouble?
Wait awhile.
Hold hard.

\N8:
With gaping mouth and head far back
he stood still |
and, after an instant,
sneezed loudly.

\lambert:
Chow!
\sout{he said.}
Blast you!

\jjom:
The dust from those sacks,
\sout{J.J.~O'Molloy said politely.}

\lambert:
No,
\sout{Ned Lambert gasped,}
I caught a \ldots
cold night before \ldots
blast your soul \ldots
\gab{460}
night before last \ldots
and there was a hell of a lot of draught \ldots

\N8:
He held his handkerchief ready for the coming \ldots

\lambert:
I was \ldots
Glasnevin this morning \ldots
poor little \ldots
what do you call him \ldots
Chow! \ldots
Mother of Moses!


\section*{Rock IX: Rochford and M'Coy}


\N9:
Tom Rochford took the top disk from the pile
he clasped against his claret waistcoat.

\rochford:
See?
\sout{he said.}
Say it's turn six.
In here, see.
Turn Now On.

\N9:
He slid it into the left slot for them.
It shot down the groove,
wobbled a while,
ceased,
ogling them:
six.

\begin{mdframed}
    \N{19}:
    Lawyers of the past, haughty, pleading,
    %\gab{470}
    beheld pass from the consolidated taxing office |
    to Nisi Prius court |
    Richie Goulding |
    carrying the costbag of Goulding, Collis and Ward |
    and heard rustling |
    from the admiralty division of king's bench |
    to the court of appeal |
    an elderly female with false teeth |
    smiling incredulously |
    and a black silk skirt of great amplitude.
\end{mdframed}

\rochford:
See?
\sout{he said.}
See now the last one I put in is over here:
Turns Over.
The impact. Leverage, see?

\N9:
He showed them the rising column of disks on the right.

\nosey:
Smart idea,

\N9:
Nosey Flynn said, snuffling.
\gab{480}

\nosey:
So a fellow coming in late |
can see what turn is on |
and what turns are over.

\rochford:
See?
\sout{Tom Rochford said.}

\N9:
He slid in a disk for himself:
and watched it
shoot,
wobble,
ogle,
stop:
four.
Turn Now On.

\lenehan:
I'll see him now in the Ormond,

\N9:
Lenehan said,

\lenehan:
and sound him.
One good turn deserves another.

\rochford:
Do,
\sout{Tom Rochford said.}
Tell him I'm Boylan with impatience.

\mcoy:
Goodnight,

\N9:
M'Coy said abruptly.

\mcoy:
When you two begin

\N9:
Nosey Flynn stooped towards the lever,
snuffling at it.

\nosey:
But how does it work here, Tommy?
\sout{he asked.}

\lenehan:
Tooraloo,
\sout{Lenehan said.}
\gab{490}
See you later.

\N9:
He followed M'Coy out
across the tiny square of Crampton court.

\lenehan:
He's a hero,
\sout{he said simply.}

\mcoy:
I know,
\sout{M'Coy said.}
The drain, you mean.

\lenehan:
Drain?
\sout{Lenehan said.}
It was down a manhole.

\N9:
They passed Dan Lowry's musichall |
where Marie Kendall,
charming soubrette,
smiled on them from a poster |
a dauby smile.

\N9:
Going down the path of Sycamore street
beside the Empire musichall |
Lenehan showed M'Coy how the whole thing was.

\lenehan:
One of those manholes like a bloody gaspipe |
and there was the poor devil stuck down in it,
half choked with sewer gas.
\gab{500}
Down went Tom Rochford anyhow,
booky's vest and all,
with the rope round him.
And be damned |
but he got the rope round the poor devil |
and the two were hauled up.

\lenehan:
The act of a hero,
\sout{he said.}

\N9:
At the Dolphin |
they halted to allow the ambulance car
to gallop past them for Jervis street.

\lenehan:
This way,
\sout{he said,
walking to the right.}
I want to pop into Lynam's
to see Sceptre's starting price.
What's the time by your gold watch and chain?

\N9:
M'Coy peered into Marcus Tertius Moses' sombre office,
then at O'Neill's clock.

\mcoy:
After three,
\sout{he said.}
\gab{510}
Who's riding her?

\lenehan:
O. Madden,
\sout{Lenehan said.}
And a game filly she is.

\N9:
While he waited in Temple bar |
M'Coy dodged a banana peel
with gentle pushes of his toe
from the path
to the gutter.

\mcoy:
Fellow might damn
easy get a nasty fall there
coming along tight in the dark.

\N{19}:
The gates of the drive opened wide
to give egress to the viceregal cavalcade.

\lenehan:
Even money,
\sout{Lenehan said returning.}
I knocked against Bantam Lyons in there |
going to back a bloody horse |
someone gave him that hasn't an earthly.
Through here.

\N9:
They went up the steps and under Merchants' arch.
\gab{520}

\begin{mdframed}
    \N{10}:
    A darkbacked figure scanned books on the hawker's cart.
\end{mdframed}

\lenehan:
There he is,
\sout{Lenehan said.}

\mcoy:
Wonder what he's buying,
\sout{M'Coy said,
glancing behind.}

\lenehan:
\Title{Leopoldo or the Bloom is on the Rye,}
\sout{Lenehan said.}

\mcoy:
He's dead nuts on sales,
\sout{M'Coy said.}
I was with him one day
and he bought a book 
from an old one in Liffey street for two bob.
There were fine plates in it worth double the money,
the stars and the moon and comets with long tails.
Astronomy it was about.

\N9:
Lenehan laughed.

\lenehan:
I'll tell you a damn good one about comets' tails,
\sout{he said.}
\gab{530}
Come over in the sun.

\N9:
They crossed to the metal bridge
and went along Wellington quay by the riverwall.

\begin{mdframed}
    \N{18}:
    Master Patrick Aloysius Dignam came out of Mangan's,
    late Fehrenbach's,
    carrying a pound and a half of porksteaks.
\end{mdframed}

\lenehan:
There was a long spread out at Glencree reformatory,
\sout{Lenehan said eagerly.}
The annual dinner, you know.
Boiled shirt affair.
The lord mayor was there,
Val Dillon it was,
and sir Charles Cameron and Dan Dawson spoke |
and there was music.
Bartell d'Arcy sang and Benjamin Dollard \ldots

\mcoy:
I know,
\sout{M'Coy broke in.}
\gab{540}
My missus sang there once.

\lenehan:
Did she?
\sout{Lenehan said.}

\begin{mdframed}
    \N3:
    A card \Title{Unfurnished Apartments} reappeared
    on the windowsash of number 7 Eccles street.
\end{mdframed}

\N9:
He checked his tale a moment
but broke out in a wheezy laugh.

\lenehan:
But wait till I tell you,
\sout{he said.}
Delahunt of Camden street had the catering
and yours truly was chief bottlewasher.
Bloom and the wife were there.
Lashings of stuff we put up:
port wine and sherry and curacao |
to which we did ample justice.
Fast and furious it was.
After liquids came solids.
Cold joints galore and mince pies \ldots

\mcoy:
I know,
\sout{M'Coy said.}
\gab{550}
The year the missus was there \ldots

\N9:
Lenehan linked his arm warmly.

\lenehan:
But wait till I tell you,
\sout{he said.}
We had a midnight lunch too |
after all the jollification |
and when we sallied forth |
it was blue o'clock the morning after the night before.
Coming home |
it was a gorgeous winter's night on the Featherbed Mountain.
Bloom and Chris Callinan were on one side of the car |
and I was with the wife on the other.
We started singing glees and duets:
\Title{Lo, the Early Beam of Morning.}
She was well primed |
with a good load of Delahunt's port under her bellyband.
Every jolt the bloody car gave |
I had her bumping up against me.
Hell's delights!
She has a fine pair, God bless her.
\gab{560}
Like that.

\N9:
He held his caved hands a cubit from him,
frowning:

\lenehan:
I was tucking the rug under her
and settling her boa all the time.
Know what I mean?

\N9:
His hands moulded ample curves of air.
He shut his eyes tight in delight,
his body shrinking,
and blew a sweet chirp from his lips.

\lenehan:
The lad stood to attention anyhow,
\sout{he said with a sigh.}
She's a gamey mare and no mistake.
Bloom was pointing out all the stars and the comets in the heavens |
to Chris Callinan and the jarvey:
the great bear and Hercules and the dragon,
and the whole jingbang lot.
But, by God, I was lost,
so to speak,
in the milky way.
\gab{570}
He knows them all, faith.
At last she spotted a weeny weeshy one miles away.
\emph{And what star is that, Poldy?}
says she.
By God, she had Bloom cornered.
\emph{That one, is it?}
says Chris Callinan,
\emph{Sure that's only what you might call a pinprick.}
By God, he wasn't far wide of the mark.

\N9:
Lenehan stopped and leaned on the riverwall,
panting with soft laughter.

\lenehan:
I'm weak,
\sout{he gasped.}

\N9:
M'Coy's white face smiled about it at instants and grew grave.
Lenehan walked on again.
He lifted his yachtingcap and scratched his hindhead rapidly.
\gab{580}
He glanced sideways in the sunlight at M'Coy.

\lenehan:
He's a cultured allroundman,
Bloom is,
\sout{he said seriously.}
He's not one of your common or garden \ldots
you know \ldots
There's a touch of the artist about old Bloom.


\section*{Rock X: Mr Bloom}


\N{10}:
Mr Bloom turned over idly pages of
\Title{The Awful Disclosures of Maria Monk},
then of Aristotle's \Title{Masterpiece}.

\BloomInt:
Crooked botched print.
Plates:
infants cuddled in a ball in bloodred wombs |
like livers of slaughtered cows.
Lots of them like that at this moment all over the world.
All butting with their skulls to get out of it.
Child born every minute somewhere.
Mrs Purefoy.%
\gab{590}

\N{10}:
He laid both books aside and glanced at the third:
\Title{Tales of the Ghetto}
by Leopold von Sacher Masoch.

\Bloom:
That I had,

\N{10}:
he said,
pushing it by.

\N{10}:
The shopman let two volumes fall on the counter.

\shopman:
Them are two good ones,
\sout{he said.}

\N{10}:
Onions of his breath came across the counter
out of his ruined mouth.
He bent to make a bundle of the other books,
hugged them against his unbuttoned waistcoat
and bore them off behind the dingy curtain.

\begin{mdframed}
    \N{20}:
    On O'Connell bridge
    many persons observed the grave deportment
    and gay apparel of Mr Denis J Maginni,
    %\gab{600}
    professor of dancing \&c.
\end{mdframed}

\N{10}:
Mr Bloom, alone,
looked at the titles.
\Title{Fair Tyrants} by James Lovebirch.

\BloomInt:
Know the kind that is.
Had it?
Yes.

\N{10}:
He opened it.

\BloomInt:
Thought so.

\N{10}:
A woman's voice behind the dingy curtain.

\BloomInt:
Listen: the man.

\BloomInt:
No:
she wouldn't like that much.
Got her it once.

\N{10}:
He read the other title:
\Title{Sweets of Sin.}

\BloomInt:
More in her line.
Let us see.

\N{10}:
He read where his finger opened.

\BloomInt:
\emph{All the dollarbills her husband gave her |
were spent in the stores
on wondrous gowns and costliest frillies.
For him!
For Raoul!}

\BloomInt:
Yes.
This.
\gab{610}
Here.
Try.

\BloomInt:
\emph{Her mouth glued on his in a luscious voluptuous kiss |
while his hands felt for the opulent curves inside her déshabillé.}

\BloomInt:
Yes.
Take this.
The end.

\BloomInt:
\emph{You are late,
he spoke hoarsely,
eying her with a suspicious glare.
The beautiful woman threw off her sabletrimmed wrap,
displaying her queenly shoulders and heaving embonpoint.
An imperceptible smile played round her perfect lips
as she turned to him calmly.}

\N{10}:
Mr Bloom read again:

\BloomInt:
\emph{The beautiful woman.}

\BloomInt:
Warmth showered gently over him,
cowing his flesh.
Flesh yielded amply amid rumpled clothes:
\gab{620}
whites of eyes swooning up.
His nostrils arched themselves for prey.
Melting breast ointments

\BloomInt:
\emph{For him!
For Raoul!}

\BloomInt:
Armpits' oniony sweat.
Fishgluey slime

\BloomInt:
\emph{Her heaving embonpoint!}

\BloomInt:
Feel!
Press!
Crushed!
Sulphur dung of lions!

\BloomInt:
Young!
Young!

\begin{mdframed}
    \N{19}:
    An elderly female,
    no more young,
    left the building of the courts of chancery,
    king's bench,
    exchequer
    and common pleas,
    having heard in the lord chancellor's court
    the case in lunacy of Potterton,
    in the admiralty division |
    the summons,
    exparte motion,
    of the owners of the Lady Cairns
    versus the owners of the barque Mona,
    in the court of appeal |
    reservation of judgment
    %\gab{630}
    in the case of Harvey versus
    the Ocean Accident and Guarantee Corporation.
\end{mdframed}

\N{10}:
Phlegmy coughs shook the air of the bookshop,
bulging out the dingy curtains.
The shopman's uncombed grey head came out
and his unshaven reddened face,
coughing.
He raked his throat rudely,
puked phlegm on the floor.
He put his boot on what he had spat,
wiping his sole along it,
and bent,
showing a rawskinned crown,
scantily haired.

\N{10}:
Mr Bloom beheld it.

\N{10}:
Mastering his troubled breath,
he said:

\Bloom:
I'll take this one.

\N{10}:
The shopman lifted eyes
\gab{640}
bleared with old rheum.

\shopman:
\Title{Sweets of Sin},

\N{10}:
he said,
tapping on it.

\shopman:
That's a good one.


\section*{Rock XI: Dilly and Simon Dedalus}


\N{11}:
The lacquey by the door of Dillon's auctionrooms
shook his handbell
twice again |
and viewed himself in the chalked mirror of the cabinet.

\N{11}:
Dilly Dedalus,
loitering by the curbstone,
heard the beats of the bell,
the cries of the auctioneer within.

\auctioneer:
Four and nine.
Those lovely curtains.
Five shillings.
Cosy curtains.
Selling new at two guineas.
Any advance on five shillings?
Going for five shillings.

\N{11}:
The lacquey lifted his handbell and shook it:

\stage{Barang!}%
\gab{650}

\begin{mdframed}
    \N{19}:
    Bang of the lastlap bell |
    spurred the halfmile wheelmen to their sprint.
    J.A.~Jackson,
    W.E.~Wylie,
    A.~Munro
    and H.T.~Gahan,
    their stretched necks wagging,
    negotiated the curve by the College library.
\end{mdframed}

\N{11}:
Mr Dedalus,
tugging a long moustache,
came round from Williams's row.
He halted near his daughter.

\dilly:
It's time for you,
\sout{she said.}

\simon:
Stand up straight for the love of the lord Jesus,
\sout{Mr Dedalus said.}
Are you trying to imitate your uncle John,
the cornetplayer,
head upon shoulder?
Melancholy God!

\N{11}:
Dilly shrugged her shoulders.
\gab{660}
Mr Dedalus placed his hands on them
and held them back.

\simon:
Stand up straight, girl,
\sout{he said.}
You'll get curvature of the spine.
Do you know what you look like?

\N{11}:
He let his head sink suddenly down and forward,
hunching his shoulders and dropping his underjaw.

\dilly:
Give it up, father,
\sout{Dilly said.}
All the people are looking at you.

\N{11}:
Mr Dedalus drew himself upright
and tugged again at his moustache.

\dilly:
Did you get any money?
\sout{Dilly asked.}

\simon:
Where would I get money?
\sout{Mr Dedalus said.}
There is no-one in Dublin would lend me fourpence.%
\gab{670}

\dilly:
You got some,

\N{11}:
Dilly said,
looking in his eyes.

\simon:
How do you know that?

\N{11}:
Mr Dedalus asked,
his tongue in his cheek.

\begin{mdframed}
    \N{12}:
    Mr Kernan,
    pleased with the order he had booked,
    walked boldly along James's street.
\end{mdframed}

\dilly:
I know you did,
\sout{Dilly answered.}
Were you in the Scotch house now?

\simon:
I was not, then,
\sout{Mr Dedalus said, smiling.}
Was it the little nuns taught you to be so saucy?
Here.

\N{11}:
He handed her a shilling.

\simon:
See if you can do anything with that,
\sout{he said.}

\dilly:
I suppose you got five,
\sout{Dilly said.}
\gab{680}
Give me more than that.

\simon:
Wait awhile,
\sout{Mr Dedalus said threateningly.}
You're like the rest of them, are you?
An insolent pack of little bitches since your poor mother died.
But wait awhile.
You'll all get a short shrift and a long day from me.
Low blackguardism!
I'm going to get rid of you.
Wouldn't care if I was stretched out stiff.
He's dead.
The man upstairs is dead.

\N{11}:
He left her and walked on.
Dilly followed quickly and pulled his coat.

\simon:
Well, what is it?

\N{11}:
he said, stopping.

\N{11}:
The lacquey rang his bell behind their backs.

\stage{Barang!}

\simon:
Curse your bloody blatant soul,
\gab{690}

\N{11}:
Mr Dedalus cried,
turning on him.

\N{11}:
The lacquey,
aware of comment,
shook the lolling clapper of his bell |
but feebly:

\bell:
Bang!

\N{11}:
Mr Dedalus stared at him.

\simon:
Watch him,
\sout{he said.}
It's instructive.
I wonder will he allow us to talk.

\dilly:
You got more than that, father,
\sout{Dilly said.}

\simon:
I'm going to show you a little trick,
\sout{Mr Dedalus said.}
I'll leave you all where Jesus left the jews.
Look, there's all I have.
I got two shillings from Jack Power |
and I spent twopence for a shave for the funeral.

\N{11}:
He drew forth a handful of copper coins,
\gab{700}
nervously.

\dilly:
Can't you look for some money somewhere?
\sout{Dilly said.}

\N{11}:
Mr Dedalus thought and nodded.

\simon:
I will,
\sout{he said gravely.}
I looked all along the gutter in O'Connell street.
I'll try this one now.

\dilly:
You're very funny,
\sout{Dilly said, grinning.}

\simon:
Here,

\N{11}:
Mr Dedalus said,
handing her two pennies.

\simon:
Get a glass of milk for yourself
and a bun or a something.
I'll be home shortly.

\N{11}:
He put the other coins in his pocket and started to walk on.

\begin{mdframed}
    \N{19}:
    The viceregal cavalcade passed,
    greeted by obsequious policemen,
    out of Parkgate.%
    %\gab{710}
\end{mdframed}

\dilly:
I'm sure you have another shilling,
\sout{Dilly said.}

\N{11}:
The lacquey banged loudly.

\N{11}:
Mr Dedalus amid the din walked off,
murmuring to himself |
with a pursing mincing mouth gently:

\simon:
The little nuns!
Nice little things!
O, sure they wouldn't do anything!
O, sure they wouldn't really!
Is it little sister Monica!


\section*{Rock XII: Mr Kernan}


\N{12}:
From the sundial
towards James's gate |
walked Mr Kernan,
pleased with the order he had booked for Pulbrook Robertson,
boldly along James's street,
past Shackleton's offices.
\gab{720}

\kernanint:
Got round him all right.

\kernan:
How do you do, Mr Crimmins?

\crimmins:
First rate, sir.

\kernan:
I was afraid you might be up in your other establishment in Pimlico.
How are things going?

\crimmins:
Just keeping alive.

\kernan:
Lovely weather we're having.

\crimmins:
Yes, indeed.

\kernan:
Good for the country.
Those farmers are always grumbling.
I'll just take a thimbleful of your best gin, Mr Crimmins.

\crimmins:
A small gin, sir.
Yes, sir.

\kernan:
Terrible affair that General Slocum explosion.

\crimmins:
Terrible, terrible!

\kernan:
A thousand casualties.
And heartrending scenes.
Men trampling down women and children.
Most brutal thing.

\crimmins:
What do they say was the cause?

\kernan:
Spontaneous combustion.
Most scandalous revelation.
Not a single lifeboat would float and the firehose all burst.

\crimmins:
What I can't understand
\gab{730}
is how the inspectors ever allowed a boat like that \ldots

\kernan:
Now, you're talking straight, Mr Crimmins.
You know why?
Palm oil.

\crimmins:
Is that a fact?

\kernan:
Without a doubt.

\crimmins:
Well now, look at that.
And America they say is the land of the free.
I thought we were bad here.

\kernanint:
I smiled at him.
\emph{America,}
I said quietly,
just like that.
\emph{What is it?
The sweepings of every country including our own.
Isn't that true?}

\crimmins:
That's a fact.

\kernan:
Graft,
my dear sir.
Well, of course, where there's money going |
there's always someone to pick it up.

\kernanint:
Saw him looking at my frockcoat.
Dress does it.
Nothing like a dressy appearance.
Bowls them over.

\begin{mdframed}
    \hskip 0.1em

    \cowley:
    Hello, Simon,

    \N{14}:
    Father Cowley said.
    %\gab{740}

    \cowley:
    How are things?

    \simon:
    Hello, Bob, old man,

    \N{14}:
    Mr Dedalus answered,
    stopping.

\end{mdframed}

\N{12}:
Mr Kernan halted and preened himself |
before the sloping mirror of Peter Kennedy, hairdresser.

\kernanint:
Stylish coat, beyond a doubt.
Scott of Dawson street.
Well worth the half sovereign I gave Neary for it.
Never built under three guineas.
Fits me down to the ground.
Some Kildare street club toff had it probably.
John Mulligan, the manager of the Hibernian bank,
gave me a very sharp eye yesterday
on Carlisle bridge as if he remembered me.

\kernanint:
Aham!
Must dress the character for those fellows.
Knight of the road.
Gentleman.
And now, Mr Crimmins,
may we have the honour of your custom again, sir.
\gab{750}
The cup that cheers but not inebriates,
as the old saying has it.

\begin{mdframed}
    \N{20}:
    North wall and sir John Rogerson's quay,
    with hulls and anchorchains,
    sailing westward,
    sailed by a skiff,
    a crumpled throwaway,
    rocked on the ferrywash,
    Elijah is coming.
\end{mdframed}

\N{12}:
Mr Kernan glanced in farewell at his image.

\kernanint:
High colour, of course.
Grizzled moustache.
Returned Indian officer.
Bravely he bore his stumpy body forward on spatted feet,
squaring his shoulders.
Is that Ned Lambert's brother over the way, Sam?
What?
Yes.
He's as like it as damn it.
No.
The windscreen of that motorcar in the sun there.
Just a flash like that.
Damn like him.%
\gab{760}

\kernanint:
Aham!

\N{12}:
Hot spirit of juniper juice warmed his vitals and his breath.

\kernanint:
Good drop of gin, that was.

\N{12}:
His frocktails winked in bright sunshine to his fat strut.

\kernanint:
Down there Emmet was hanged, drawn and quartered.
Greasy black rope.
Dogs licking the blood off the street
when the lord lieutenant's wife drove by in her noddy.

\kernanint:
Bad times those were.
Well, well.
Over and done with.
Great topers too.
Fourbottle men.

\kernanint:
Let me see.
Is he buried in saint Michan's?
Or no, there was a midnight burial in Glasnevin.
\gab{770}
Corpse brought in through a secret door in the wall.
Dignam is there now.
Went out in a puff.
Well, well.
Better turn down here.
Make a detour.

\N{12}:
Mr Kernan turned and walked down the slope of Watling street
by the corner of Guinness's visitors' waitingroom.
Outside the Dublin Distillers Company's stores
an outside car without fare or jarvey stood,
the reins knotted to the wheel.

\kernanint:
Damn dangerous thing.
Some Tipperary bosthoon
endangering the lives of the citizens.
Runaway horse.

\begin{mdframed}
    \N{20}:
    Denis Breen with his tomes,
    weary of having waited an hour in John Henry Menton's office,
    led his wife over O'Connell bridge,
    bound for the office
    %\gab{780}
    of Messrs Collis and Ward.
\end{mdframed}

\N{12}:
Mr Kernan approached Island street.

\kernanint:
Times of the troubles.
Must ask Ned Lambert to lend me
those reminiscences of sir Jonah Barrington.
When you look back on it all now
in a kind of retrospective arrangement.
Gaming at Daly's.
No cardsharping then.
One of those fellows got his hand nailed to the table by a dagger.
Somewhere here lord Edward Fitzgerald escaped from major Sirr.
Stables behind Moira house.

\kernanint:
Damn good gin that was.

\kernanint:
Fine dashing young nobleman.
Good stock, of course.
That ruffian, that sham squire, with his violet gloves |
gave him away.
Course they were on the wrong side.
\gab{790}
They rose in dark and evil days.
Fine poem that is:
Ingram.
They were gentlemen.
Ben Dollard does sing that ballad touchingly.
Masterly rendition.

\kernanint:
\song{At the Siege of Ross did my father fall.}

\N{12}:
A cavalcade in easy trot along Pembroke quay |
passed,
outriders leaping, leaping in their, in their saddles.
Frockcoats.
Cream sunshades.

\N{12}:
Mr Kernan hurried forward,
blowing pursily.

\kernanint:
His Excellency!
Too bad!
Just missed that by a hair.
Damn it!
What a pity!


\section*{Rock XIII: Stephen Dedalus and Old Russell}


\N{13}:
Stephen Dedalus watched through the webbed window |
\gab{800}
the lapidary's fingers prove a timedulled chain.
Dust webbed the window and the showtrays.
Dust darkened the toiling fingers with their vulture nails.
Dust slept on dull coils of bronze and silver,
lozenges of cinnabar,
on rubies,
leprous and winedark stones.

\StephenInt:
Born all in the dark wormy earth,
cold specks of fire,
evil, lights shining in the darkness.
Where fallen archangels flung the stars of their brows.
Muddy swinesnouts,
hands,
root and root,
gripe and wrest them.

\StephenInt:
She dances in a foul gloom where gum bums with garlic.
A sailorman, rustbearded, sips from a beaker rum and eyes her.
A long and seafed silent rut.
\gab{810}
She dances, capers,
wagging her sowish haunches and her hips,
on her gross belly flapping a ruby egg.

\N{13}:
Old Russell with a smeared shammy rag |
burnished again his gem,
turned it
and held it at the point of his Moses' beard.
Grandfather ape gloating on a stolen hoard.

\StephenInt:
And you who wrest old images from the burial earth?
The brainsick words of sophists:
Antisthenes.
A lore of drugs.
Orient and immortal wheat
standing from everlasting to everlasting.

\N{13}:
Two old women
fresh from their whiff of the briny |
trudged through Irishtown along London bridge road,
one with a sanded tired umbrella,
one with a midwife's bag
\gab{820}
in which eleven cockles rolled.

\N{13}:
The whirr of flapping leathern bands
and hum of dynamos from the powerhouse
urged Stephen to be on.

\StephenInt:
Beingless beings.
Stop!
Throb always without you
and the throb always within.
Your heart you sing of.
I between them.
Where?
Between two roaring worlds where they swirl, I.
Shatter them, one and both.
But stun myself too in the blow.
Shatter me |
you who can.
Bawd and butcher were the words.
I say!
Not yet awhile.
A look around.

\russell:
Yes, quite true.
Very large and wonderful and keeps famous time.
You say right, sir.
A Monday morning, 'twas so, indeed.

\N{13}:
Stephen went down Bedford row,
\gab{830}
the handle of the ash
clacking against his shoulderblade.
In Clohissey's window
a faded 1860 print of Heenan boxing Sayers |
held his eye.
Staring backers with square hats
stood round the roped prizering.
The heavyweights in tight loincloths proposed gently |
each to other his bulbous fists.
And they are throbbing:
heroes' hearts.

\N{13}:
He turned and halted by the slanted bookcart.

\huckster:
Twopence each,

\N{13}:
the huckster said.

\huckster:
Four for sixpence.


\StephenInt:
Tattered pages.
\Title{The Irish Beekeeper}.
\Title{Life and Miracles of the Curé of Ars}.
\Title{Pocket Guide to Killarney}.


\StephenInt:
I might find here one of my pawned schoolprizes.
\gab{840}
\latin{Stephano Dedalo, alumno optimo, palmam ferenti}.

\begin{mdframed}
    \N1:
    Father Conmee, having read his little hours,
    walked through the hamlet of Donnycarney,
    murmuring vespers.
\end{mdframed}

\StephenInt:
Binding too good probably.
What is this?
Eighth and ninth book of Moses.
Secret of all secrets.
Seal of King David.
Thumbed pages: read and read.
Who has passed here before me?
How to soften chapped hands.
Recipe for white wine vinegar.
How to win a woman's love.
For me this.
Say the following talisman three times with hands folded:

\StephenInt:
\hebrew{Se el yilo nebrakada femininum!
Amor me solo!
Sanktus!
Amen.}


\StephenInt:
Who wrote this?
\gab{850}
Charms and invocations of the most blessed abbot Peter Salanka
to all true believers divulged.
As good as any other abbot's charms,
as mumbling Joachim's.
Down, baldynoddle, or we'll wool your wool.

\dilly:
What are you doing here, Stephen?

\N{13}:
Dilly's high shoulders and shabby dress.

\StephenInt:
Shut the book quick.
Don't let see.

\Stephen:
What are you doing?
\sout{Stephen said.}

\StephenInt:
A Stuart face of nonesuch Charles,
lank locks falling at its sides.
It glowed as she crouched |
feeding the fire with broken boots.
I told her of Paris.
\gab{860}
Late lieabed under a quilt of old overcoats,
fingering a pinchbeck bracelet,
Dan Kelly's token.
\hebrew{Nebrakada femininum}.

\Stephen:
What have you there?
\sout{Stephen asked.}

\dilly:
I bought it from the other cart for a penny,
\sout{Dilly said, laughing nervously.}
Is it any good?

\StephenInt:
My eyes they say she has.
Do others see me so?
Quick, far and daring.
Shadow of my mind.

\N{13}:
He took the coverless book from her hand.

\StephenInt:
Chardenal's French primer.

\Stephen:
What did you buy that for?
\sout{he asked.}
To learn French?

\N{13}:
She nodded,
\gab{870}
reddening and closing tight her lips.

\StephenInt:
Show no surprise.
Quite natural.

\Stephen:
Here,
\sout{Stephen said.}
It's all right.
Mind Maggy doesn't pawn it on you.
I suppose all my books are gone.

\dilly:
Some,
\sout{Dilly said.}
We had to.

\StephenInt:
She is drowning.
Agenbite.
Save her.
Agenbite.
All against us.
She will drown me with her,
eyes and hair.
Lank coils of seaweed hair around me,
my heart,
my soul.
Salt green death.

\StephenInt:
We.

\StephenInt:
Agenbite of inwit.
Inwit's agenbite.

\StephenInt:
Misery!
\gab{880}
Misery!


\section*{Rock XIV: Simon Dedalus, Fr Cowley, and Ben Dollard}


\cowley:
Hello, Simon,

\N{14}:
Father Cowley said.

\cowley:
How are things?

\simon:
Hello, Bob, old man,

\N{14}:
Mr Dedalus answered,
stopping.

\N{14}:
They clasped hands loudly outside Reddy and Daughter's.
Father Cowley brushed his moustache often downward
with a scooping hand.

\simon:
What's the best news?
\sout{Mr Dedalus said.}

\cowley:
Why then not much,
\sout{Father Cowley said.}
I'm barricaded up, Simon,
with two men prowling around the house
trying to effect an entrance.

\simon:
Jolly,
\sout{Mr Dedalus said.}
Who is it?

\cowley:
O,
\sout{Father Cowley said.}
\gab{890}
A certain gombeen man of our acquaintance.

\simon:
With a broken back, is it?
\sout{Mr Dedalus asked.}

\cowley:
The same, Simon,
\sout{Father Cowley answered.}
Reuben of that ilk.
I'm just waiting for Ben Dollard.
He's going to say a word to long John
to get him to take those two men off.
All I want is a little time.

\N{14}:
He looked with vague hope up and down the quay,
a big apple bulging in his neck.

\simon:
I know,
\sout{Mr Dedalus said, nodding.}
Poor old bockedy Ben!
He's always doing a good turn for someone.
Hold hard!

\N{14}:
He put on his glasses
and gazed towards the metal bridge an instant.

\simon:
There he is, by God,
\sout{he said,}
\gab{900}
arse and pockets.

\N{14}:
Ben Dollard's loose blue cutaway and square hat above large slops |
crossed the quay in full gait
from the metal bridge.
He came towards them at an amble,
scratching actively behind his coattails.

\N{14}:
As he came near
Mr Dedalus greeted:

\simon:
Hold that fellow with the bad trousers.

\dollard:
Hold him now,

\N{14}:
Ben Dollard said.

\N{14}:
Mr Dedalus eyed with cold wandering scorn
various points of Ben Dollard's figure.
Then,
turning to Father Cowley with a nod,
he muttered sneeringly:

\simon:
That's a pretty garment, isn't it,
\gab{910}
for a summer's day?

\dollard:
Why, God eternally curse your soul,
\sout{Ben Dollard growled furiously,}
I threw out more clothes in my time than you ever saw.

\N{14}:
He stood beside them beaming,
on them first and on his roomy clothes
from points of which Mr Dedalus flicked fluff,
saying:

\simon:
They were made for a man in his health, Ben, anyhow.

\dollard:
Bad luck to the jewman that made them,
\sout{Ben Dollard said.}
Thanks be to God he's not paid yet.

\cowley:
And how is that \italian{basso profondo}, Benjamin?
\sout{Father Cowley asked.}

\begin{mdframed}
    \N{17}:
    Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell, |
    murmuring, glassy\-eyed, |
    %\gab{920}
    strode past the Kildare street club.
\end{mdframed}

\N{14}:
Ben Dollard frowned |
and, making suddenly a chanter's mouth, |
gave forth a deep note.

\dollard:
Aw!
\sout{he said.}

\simon:
That's the style,

\N{14}:
Mr Dedalus said,
nodding to its drone.

\dollard:
What about that?
\sout{Ben Dollard said.}
Not too dusty?
What?

\N{14}:
He turned to both.

\cowley:
That'll do,

\N{14}:
Father Cowley said,
nodding also.

\begin{mdframed}
    \N8:
    The reverend Hugh~C. Love
    walked from the old chapterhouse of saint Mary's abbey
    past James and Charles Kennedy's, rectifiers,
    attended by Geraldines tall and personable,
    %\gab{930}
    towards the Tholsel beyond the ford of hurdles.
\end{mdframed}

\N{14}:
Ben Dollard with a heavy list towards the shopfronts
led them forward,
his joyful fingers in the air.

\dollard:
Come along with me to the subsheriff's office,
\sout{he said.}
I want to show you the new beauty Rock has for a bailiff.
He's a cross between Lobengula and Lynchehaun.
He's well worth seeing, mind you.
Come along.
I saw John Henry Menton casually in the Bodega just now
and it will cost me a fall |
if I don't \ldots
Wait awhile \ldots
We're on the right lay, Bob, believe you me.

\cowley:
For a few days tell him,
\sout{Father Cowley said anxiously.}

\N{14}:
Ben Dollard halted and stared,
\gab{940}
his loud orifice open,
a dangling button of his coat wagging brightbacked from its thread |
as he wiped away the heavy shraums that clogged his eyes
to hear aright.

\dollard:
What few days?
\sout{he boomed.}
Hasn't your landlord distrained for rent?

\cowley:
He has,
\sout{Father Cowley said.}

\dollard:
Then our friend's writ is not worth the paper it's printed on,
\sout{Ben Dollard said.}
The landlord has the prior claim.
I gave him all the particulars.
29 Windsor avenue.
Love is the name?

\cowley:
That's right,
\sout{Father Cowley said.}
The reverend Mr Love.
He's a minister in the country somewhere.
But are you sure of that?

\dollard:
You can tell Barabbas from me,
\sout{Ben Dollard said,}
\gab{950}
that he can put that writ where Jacko put the nuts.

\N{14}:
He led Father Cowley boldly forward,
linked to his bulk.

\simon:
Filberts I believe they were,

\N{14}:
Mr Dedalus said,
as he dropped his
glasses on his coatfront,
following them.


\section*{Rock XV: Martin Cunningham, Mr Power, \ldots}


\cunningham:
The youngster will be all right,

\N{15}:
Martin Cunningham said,
as they passed
out of the Castleyard gate.

\N{15}:
The policeman touched his forehead.

\cunningham:
God bless you,
\sout{Martin Cunningham said, cheerily.}

\N{15}:
He signed to the waiting jarvey
\gab{960}
who chucked at the reins
and set on towards Lord Edward street.

\begin{mdframed}
    \N{20}:
    Bronze by gold,
    Miss Kennedy's head by Miss Douce's head,
    appeared above
    the crossblind of the Ormond hotel.
\end{mdframed}

\cunningham:
Yes,
\sout{Martin Cunningham said,
fingering his beard.}
I wrote to Father Conmee and laid the whole case before him.

\power:
You could try our friend,

\N{15}:
Mr Power suggested backward.

\cunningham:
Boyd?
\sout{Martin Cunningham said shortly.}
Touch me not.

\N{15}:
John Wyse Nolan,
lagging behind, reading the list,
came after them
quickly down Cork hill.

\N{15}:
On the steps of the City hall |
\gab{970}
Councillor Nannetti, descending,
hailed Alderman Cowley and Councillor Abraham Lyon |
ascending.

\N{15}:
The castle car wheeled empty into upper Exchange street.

\johnwyse:
Look here, Martin,

\N{15}:
John Wyse Nolan said,
overtaking them at the \Title{Mail} office.

\johnwyse:
I see Bloom put his name down for five shillings.

\cunningham:
Quite right,
\sout{Martin Cunningham said,
taking the list.}
And put down the
five shillings too.

\power:
Without a second word either,
\sout{Mr Power said.}

\cunningham:
Strange but true,
\sout{Martin Cunningham added.}

\N{15}:
John Wyse Nolan opened wide eyes.

\johnwyse:
I'll say there is much kindness in the jew,
\gab{980}
\sout{he quoted, elegantly.}

\N{15}:
They went down Parliament street.

\power:
There's Jimmy Henry,
\sout{Mr Power said,}
just heading for Kavanagh's.

\cunningham:
Righto,
\sout{Martin Cunningham said.}
Here goes.

\begin{mdframed}
    \N5:
    Outside \french{La Maison Claire}
    Blazes Boylan waylaid Jack Mooney's brother-in-law,
    humpy, tight,
    making for the liberties.
\end{mdframed}

\N{15}:
John Wyse Nolan fell back with Mr Power,
while Martin Cunningham took the elbow
of a dapper little man in a shower of hail suit,
who walked uncertainly,
with hasty steps past Micky Anderson's watches.

\johnwyse:
The assistant town clerk's corns are giving him some trouble,
\sout{John Wyse Nolan told Mr Power.}%
\gab{990}

\N{15}:
They followed round the corner
towards James Kavanagh's winerooms.
The empty castle car fronted them at rest in Essex gate.
Martin Cunningham,
speaking always,
showed often the list at which Jimmy Henry did not glance.

\johnwyse:
And long John Fanning is here too,
\sout{John Wyse Nolan said,}
as large as life.

\N{15}:
The tall form of long John Fanning
filled the doorway where he stood.

\cunningham:
Good day, Mr Subsheriff,

\N{15}:
Martin Cunningham said,
as all halted and greeted.%
\gab{1000}

\N{15}:
Long John Fanning made no way for them.
He removed his large Henry Clay decisively
and his large fierce eyes scowled intelligently over all their faces.

\longjohn:
Are the conscript fathers pursuing their peaceful deliberations?

\N{15}:
he said with rich acrid utterance
to the assistant town clerk.

\jimmyhenry:
Hell open to christians they were having,

\N{15}:
Jimmy Henry said pettishly,

\jimmyhenry:
about their damned Irish language.
Where was the marshal,
he wanted to know,
to keep order in the council chamber.
And old Barlow the macebearer
laid up with asthma,
no mace on the table,
nothing in order,
no quorum even,
\gab{1010}
and Hutchinson, the lord mayor, in Llandudno
and little Lorcan Sherlock doing \latin{locum tenens} for him.
Damned Irish language,
language of our forefathers.

\N{15}:
Long John Fanning blew a plume of smoke from his lips.

\N{15}:
Martin Cunningham spoke by turns,
twirling the peak of his beard,
to the assistant town clerk and the subsheriff,
while John Wyse Nolan held his peace.

\longjohn:
What Dignam was that?
\sout{long John Fanning asked.}

\N{15}:
Jimmy Henry made a grimace and lifted his left foot.

\jimmyhenry:
O, my corns!
\sout{he said plaintively.}
Come upstairs for goodness' sake till I sit down somewhere.
\gab{1020}
Uff!
Ooo!
Mind!

\N{15}:
Testily he made room for himself
beside long John Fanning's flank
and passed in and up the stairs.

\cunningham:
Come on up,

\N{15}:
Martin Cunningham said to the subsheriff.

\cunningham:
I don't think you knew him
or perhaps you did, though.

\N{15}:
With John Wyse Nolan |
Mr Power followed them in.

\power:
Decent little soul he was,

\N{15}:
Mr Power said
to the stalwart back of long John Fanning
ascending towards long John Fanning in the mirror.

\cunningham:
Rather lowsized.
Dignam of Menton's office that was,
\sout{Martin Cunningham said.}

\N{15}:
Long John Fanning could not remember him.%
\gab{1030}

\N{15}:
Clatter of horsehoofs sounded from the air.

\cunningham:
What's that?
\sout{Martin Cunningham said.}

\N{15}:
All turned where they stood.
John Wyse Nolan came down again.
From the cool shadow of the doorway |
he saw the horses pass Parliament street,
harness and glossy pasterns in sunlight shimmering.
Gaily they went past |
before his cool unfriendly eyes,
not quickly.
In saddles of the leaders,
leaping leaders,
rode outriders.

\cunningham:
What was it?

\N{15}:
Martin Cunningham asked,
as they went on up the staircase.

\johnwyse:
The lord lieutenantgeneral and general governor of Ireland,
\gab{1040}

\N{15}:
John Wyse Nolan answered from the stairfoot.


\section*{Rock XVI: Buck Mulligan and Haines}


\N{16}:
As they trod across the thick carpet
Buck Mulligan whispered behind his Panama to Haines:

\mulligan:
Parnell's brother.
There in the corner.

\N{16}:
They chose a small table near the window,
opposite a longfaced man
whose beard and gaze hung intently down on a chessboard.

\haines:
Is that he?

\N{16}:
Haines asked,
twisting round in his seat.

\mulligan:
Yes,
\sout{Mulligan said.}
That's John Howard, his brother, our city marshal.

\N{16}:
John Howard Parnell translated a white bishop quietly
\gab{1050}
and his grey claw went up again to his forehead |
whereat it rested.
An instant after,
under its screen,
his eyes looked quickly, ghostbright, at his foe |
and fell once more upon a working corner.

\haines:
I'll take a \french{mélange},

\N{16}:
Haines said to the waitress.

\mulligan:
Two \french{mélanges},
\sout{Buck Mulligan said.}
And bring us some scones and butter
and some cakes as well.

\N{16}:
When she had gone
he said, laughing:

\mulligan:
We call it D.B.C. because they have damn bad cakes.
O, but you missed
Dedalus on \Title{Hamlet}.

\N{16}:
Haines opened his newbought book.%
\gab{1060}

\haines:
I'm sorry,
\sout{he said.}
Shakespeare is the happy huntingground of all minds
that have lost their balance.

\begin{mdframed}
    \N3:
    The onelegged sailor growled at the area of 14 Nelson street:

    \sailor:
    \Title{England Expects \ldots}
\end{mdframed}

\N{16}:
Buck Mulligan's primrose waistcoat shook gaily to his laughter.

\mulligan:
You should see him,
\sout{he said,}
when his body loses its balance.
Wandering Aengus I call him.

\haines:
I am sure he has an \french{idée fixe},

\N{16}:
Haines said,
pinching his chin thoughtfully with thumb and forefinger.

\haines:
Now I am speculating what it would be likely to be.
\gab{1070}
Such persons always have.

\N{16}:
Buck Mulligan bent across the table gravely.

\mulligan:
They drove his wits astray,
\sout{he said,}
by visions of hell.
He will never capture the Attic note.
The note of Swinburne,
of all poets,
the white death and the ruddy birth.
That is his tragedy.
He can never be a poet.
The joy of creation \ldots

\haines:
Eternal punishment,
\sout{Haines said,
nodding curtly.}
I see.
I tackled him
this morning on belief.
There was something on his mind, I saw.
It's rather interesting
because professor Pokorny of Vienna
makes an interesting point out of that.

\N{16}:
Buck Mulligan's watchful eyes saw the waitress come.
\gab{1080}
He helped her to unload her tray.

\haines:
He can find no trace of hell in ancient Irish myth,
\sout{Haines said,
amid the cheerful cups.}
The moral idea seems lacking,
the sense of destiny, of retribution.
Rather strange he should have just that fixed idea.
Does he write anything for your movement?

\N{16}:
He sank two lumps of sugar deftly longwise
through the whipped cream.
Buck Mulligan slit a steaming scone in two
and plastered butter
over its smoking pith.
He bit off a soft piece hungrily.

\mulligan:
Ten years,
\sout{he said,
chewing and laughing.}
He is going to write something in ten years.%
\gab{1090}

\haines:
Seems a long way off,
\sout{Haines said,
thoughtfully lifting his spoon.}
Still, I shouldn't wonder if he did after all.

\N{16}:
He tasted a spoonful from the creamy cone of his cup.

\haines:
This is real Irish cream I take it,
\sout{he said with forbearance.}
I don't want to be imposed on.

\begin{mdframed}
    \N{20}:
    Elijah,
    skiff,
    light crumpled throwaway,
    sailed eastward by flanks of
    ships and trawlers,
    amid an archipelago of corks,
    beyond new Wapping street
    past Benson's ferry,
    and by the threemasted schooner \Title{Rosevean} from
    Bridgwater with bricks.
\end{mdframed}


\section*{Rock XVII: Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell}

\gab{1100}

\N{17}:
Almidano Artifoni walked past Holles street,
past Sewell's yard.
Behind him
Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell,
with stick\-umbrella\-dust\-coat dangling,
shunned the lamp before Mr Law Smith's house
and, crossing, walked along Merrion square.
Distantly behind him
a blind stripling tapped his way
by the wall of College park.

\N{17}:
Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell
walked as far as
Mr Lewis Werner's cheerful windows,
then turned and strode back along Merrion square,
his stick\-umbrella\-dust\-coat dangling.

\N{17}:
At the corner of Wilde's house |
he halted,
frowned at Elijah's name announced on the Metropolitan hall,
\gab{1110}
frowned at the distant pleasance of duke's lawn.
His eyeglass flashed frowning in the sun.
With ratsteeth bared he muttered:

\cashel:
\latin{Coactus volui}.

\N{17}:
He strode on for Clare street,
grinding his fierce word.

\N{17}:
As he strode past Mr Bloom's dental windows |
the sway of his dustcoat brushed rudely from its angle |
a slender tapping cane and swept onwards,
having buffeted a thewless body.
The blind stripling turned his sickly face after the striding form.

\stripling:
God's curse on you,

\N{17}:
he said sourly,

\stripling:
whoever you are!
You're blinder nor I am,
\gab{1120}
you bitch's bastard!


\section*{Rock XVIII: Master Dignam}


\N{18}:
Opposite Ruggy O'Donohoe's |
Master Patrick Aloysius Dignam,
pawing the pound and a half of Mangan's,
late Fehrenbach's,
porksteaks |
he had been sent for,
went along warm Wicklow street dawdling.
It was too blooming dull sitting in the parlour |
with Mrs Stoer and Mrs Quigley and Mrs MacDowell
and the blind down |
and they all at their sniffles |
and sipping sups of the superior tawny sherry |
uncle Barney brought from Tunney's.
And they eating crumbs of the cottage fruitcake,
jawing the whole blooming time and sighing.

\N{18}:
After Wicklow lane |
\gab{1130}
the window of Madame Doyle,
courtdress milliner,
stopped him.
He stood looking in at the two puckers |
stripped to their pelts and putting up their props.
From the sidemirrors |
two mourning Masters Dignam gaped silently.
Myler Keogh,
Dublin's pet lamb,
will meet sergeantmajor Bennett,
the Portobello bruiser,
for a purse of fifty sovereigns.

\masterdignam:
Gob, that'd be a good pucking match to see.
Myler Keogh, that's the chap sparring out to him with the green sash.
Two bar entrance, soldiers half price.
I could easy do a bunk on ma.

\N{18}:
Master Dignam on his left |
turned as he turned.

\masterdignam:
That's me in mourning.
When is it?
May the twentysecond.
Sure, the blooming thing is all over.

\N{18}:
He turned to the right |
and on his right |
Master Dignam turned, |
\gab{1140}
his cap awry,
his collar sticking up.
Buttoning it down,
his chin lifted, |
he saw the image of Marie Kendall,
charming soubrette,
beside the two puckers.

\masterdignam:
One of them mots that do be in the packets of fags |
Stoer smokes |
that his old fellow |
welted hell out of him for |
one time he found out.

\N{18}:
Master Dignam got his collar down and dawdled on.

\masterdignam:
The best pucker
going for strength was Fitzsimons.
One puck in the wind from that fellow |
would knock you into the middle of next week, man.
But the best pucker for science |
was Jem Corbet
before Fitzsimons knocked the stuffings out of him,
dodging and all.

\N{18}:
In Grafton street |
\gab{1150}
Master Dignam saw\ldots

\masterdignam:
a red flower in a toff's mouth |
and a swell pair of kicks on him |
and he listening to what the drunk was telling him |
and grinning all the time.

\masterdignam:
No Sandymount tram.

\N{18}:
Master Dignam walked along Nassau street,
shifted the porksteaks to his other hand.
His collar sprang up again and he tugged it down.

\masterdignam:
The blooming stud was too small
for the buttonhole of the shirt,
blooming end to it.

\N{18}:
He met schoolboys with satchels.

\masterdignam:
I'm not going tomorrow either,
stay away till Monday.

\N{18}:
He met other schoolboys.

\masterdignam:
Do they notice I'm in mourning?
Uncle Barney said he'd get it into the paper tonight.
Then they'll all see it in the paper
\gab{1160}
and read my name printed and pa's name.

\N{18}:
His face got all grey
instead of being red like it was |
and there was a fly walking over it |
up to his eye.

\masterdignam:
The scrunch that was
when they were screwing the screws into the coffin:
and the bumps when they were bringing it downstairs.

\masterdignam:
Pa was inside it |
and ma crying in the parlour |
and uncle Barney telling the men
how to get it round the bend.
A big coffin it was,
and high and heavylooking.
How was that?
The last night pa was boosed |
he was standing on the landing there |
bawling out for his boots to go out to Tunney's |
for to boose more |
and he looked butty and short in his shirt.
Never see him again.
Death, that is.
Pa is dead.
My father is dead.
\gab{1170}
He told me to be a good son to ma.
I couldn't hear the other things he said |
but I saw his tongue and his teeth trying to say it better.
Poor pa.
That was Mr Dignam, my father.
I hope he's in purgatory now |
because he went to confession to Father Conroy on Saturday night.


\section*{Rock XIX: The Viceregal Procession}


\N{19}:
William Humble,
earl of Dudley,
and lady Dudley,
accompanied by lieutenantcolonel Heseltine,
drove out after luncheon |
from the viceregal lodge.
In the following carriage |
were the honourable Mrs Paget,
Miss de Courcy
and the honourable Gerald Ward A.D.C. |
in attendance.

\N{19}:
The cavalcade passed out by the lower gate of Phoenix park |
\gab{1180}
saluted by obsequious policemen |
and proceeded past Kingsbridge |
along the northern quays.
The viceroy was most cordially greeted |
on his way through the metropolis.
At Bloody bridge |
Mr Thomas Kernan beyond the river |
greeted him vainly from afar.
Between Queen's and Whitworth bridges |
lord Dudley's viceregal carriages |
passed and were unsaluted by |
Mr Dudley White, B.L., M.A.,
who stood on Arran quay outside Mrs M.E. White's,
the pawnbroker's,
at the corner of Arran street west |
stroking his nose with his forefinger,
undecided whether he should arrive at Phibsborough more quickly |
by a triple change of tram
or by hailing a car
or on foot |
through Smithfield, Constitution hill and Broadstone terminus.
\gab{1190}
In the porch of Four Courts |
Richie Goulding
with the costbag of Goulding, Collis and Ward |
saw him with surprise.
Past Richmond bridge |
at the doorstep
of the office of Reuben J Dodd, solicitor,
agent for the Patriotic Insurance Company,
an elderly female about to enter |
changed her plan |
and retracing her steps by King's windows |
smiled credulously
on the representative of His Majesty.
From its sluice in Wood quay wall |
under Tom Devan's office |
Poddle river hung out in fealty |
a tongue of liquid sewage.
Above the crossblind of the Ormond hotel,
gold by bronze,
Miss Kennedy's head by Miss Douce's head
watched and admired.
On Ormond quay
Mr Simon Dedalus,
steering his way from the greenhouse for the subsheriff's office,
\gab{1200}
stood still in midstreet
and brought his hat low.
His Excellency graciously returned Mr Dedalus' greeting.
From Cahill's corner
the reverend Hugh~C. Love, M.A.,
made obeisance
unperceived,
mindful of lords deputies |
whose hands benignant
had held of yore rich advowsons.
On Grattan bridge
Lenehan and M'Coy,
taking leave of each other,
watched the carriages go by.
Passing by Roger Greene's office
and Dollard's big red printinghouse
Gerty MacDowell,
carrying the Catesby's cork lino letters |
for her father who was laid up,
knew by the style |
it was the lord and lady lieutenant |
but she couldn't see
what Her Excellency had on |
because the tram 
and Spring's big yellow furniture van
had to stop in front of her |
\gab{1210}
on account of its being the lord lieutenant.
Beyond Lundy Foot's |
from the shaded door of Kavanagh's winerooms |
John Wyse Nolan smiled with unseen coldness |
towards the lord lieutenantgeneral and general governor of Ireland.
The Right Honourable
William Humble,
earl of Dudley,
G.C.V.O.,
passed Micky Anderson's
all times ticking watches |
and Henry and James's wax smartsuited freshcheeked models,
the gentleman Henry,
\french{dernier cri} James.
Over against Dame gate |
Tom Rochford and Nosey Flynn
watched the approach of the cavalcade.
Tom Rochford,
seeing the eyes of lady Dudley fixed on him,
took his thumbs
quickly out of the pockets of his claret waistcoat
and doffed his cap to her.
\gab{1220}
A charming \french{soubrette},
great Marie Kendall, with dauby cheeks and lifted skirt |
smiled daubily from her poster
upon William Humble, earl of Dudley,
and upon lieutenantcolonel H. G. Heseltine,
and also upon the honourable Gerald Ward A.D.C.
From the window of the D.B.C. |
Buck Mulligan gaily,
and Haines gravely,
gazed down on the viceregal equipage |
over the shoulders of eager guests,
whose mass of forms
darkened the chessboard |
whereon John Howard Parnell looked intently.
In Fownes's street |
Dilly Dedalus,
straining her sight upward |
from Chardenal's first French primer,
saw sunshades spanned
and wheelspokes spinning in the glare.
John Henry Menton,
filling the doorway of Commercial Buildings,
\gab{1230}
stared from winebig oyster eyes,
holding a fat gold hunter watch |
not looked at |
in his fat left hand |
not feeling it.
Where the foreleg of King Billy's horse pawed the air |
Mrs Breen plucked her hastening husband back |
from under the hoofs of the outriders.
She shouted in his ear the tidings.
Understanding, he shifted his tomes to his left breast
and saluted the second carriage.
The honourable Gerald Ward A.D.C.,
agreeably surprised,
made haste to reply.

\N{20}:
At Ponsonby's corner
a jaded white flagon Haitch \stage{[H.]} halted |
and four tallhatted white flagons halted behind him,
Ee Ell Wye Apostrophe Ess \stage{[E.L.Y.'S]} |
while outriders pranced past
and carriages.

\N{20}:
Opposite Pigott's music warerooms |
Mr Denis J Maginni,
professor of dancing \&c,
gaily apparelled,
gravely walked,
\gab{1240}
outpassed by a viceroy and unobserved.

\N5:
By the provost's wall |
came jauntily Blazes Boylan,
stepping in tan shoes
and socks with skyblue clocks |
to the refrain of
\Title{My Girl's a Yorkshire Girl}.

\N5:
Blazes Boylan presented to the leaders' skyblue frontlets and high action |
a skyblue tie,
a widebrimmed straw hat at a rakish angle |
and a suit of indigo serge.
His hands in his jacket pockets forgot to salute 
but he offered to the three ladies |
the bold admiration of his eyes
and the red flower between his lips.

\N{19}:
As they drove along Nassau street |
His Excellency
drew the attention of his bowing consort |
to the programme of music
which was being discoursed in College park.
Unseen brazen highland laddies
blared and drumthumped
after the \french{cortège}:%
\gab{1250}


\begin{verse}
    But though she's a factory lass \\
    And wears no fancy clothes. \\
    Baraabum. \\
    Yet I've a sort of a \\
    Yorkshire relish for \\
    My little Yorkshire rose. \\
    Baraabum.
\end{verse}

\N{19}:
Thither of the wall |
the quartermile flat handicappers,
M.C.~Green,
H.~Shrift,
T.M.~Patey,
C.~Scaife,
J.B.~Jeffs,
G.N.~Morphy,
F.~Stevenson,
C.~Adderly
and W.C.~Huggard,
started in pursuit.
\gab{1260}
Striding past Finn's hotel
Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell
stared through a fierce eyeglass
across the carriages |
at the head of Mr M.E.~Solomons
in the window of the Austro-Hungarian viceconsulate.
Deep in Leinster street
by Trinity's postern |
a loyal king's man, Hornblower,
touched his tallyho cap.
As the glossy horses pranced by Merrion square |
Master Patrick Aloysius Dignam, waiting,
saw salutes being given to the gent with the topper |
and raised also his new black cap
with fingers greased by porksteak paper.
His collar too sprang up.
The viceroy,
on his way to inaugurate the Mirus bazaar |
in aid of funds for Mercer's hospital,
\gab{1270}
drove with his following
towards Lower Mount street.
He passed a blind stripling opposite Broadbent's.
In Lower Mount street |
a pedestrian in a brown macintosh,
eating dry bread,
passed swiftly and unscathed across the viceroy's path.
At the Royal Canal bridge,
from his hoarding,
Mr Eugene Stratton,
his blub lips agrin,
bade all comers welcome to Pembroke township.
At Haddington road corner |
two sanded women halted themselves,
an umbrella and a bag in which eleven cockles rolled |
to view with wonder |
the lord mayor and lady mayoress without his golden chain.
On Northumberland and Lansdowne roads |
His Excellency acknowledged punctually
salutes from rare male walkers,
the salute of two small schoolboys
at the garden gate of the house |
said to have been admired by the late queen |
\gab{1280}
when visiting the Irish capital |
with her husband, the prince consort, in 1849 |
and the salute of Almidano Artifoni's sturdy trousers |
swallowed by a closing door.
