\Nq:
I was just passing the time of day with old Troy of the D. M. P. at the
corner of Arbour hill there and be damned but a bloody sweep came along
and he near drove his gear into my eye. I turned around to let him have
the weight of my tongue when who should I see dodging along Stony Batter
only Joe Hynes.

\N1:
Lo, Joe, \sout{says I}. How are you blowing? Did you see that bloody
chimneysweep near shove my eye out with his brush?

\joe:
Soot's luck, \sout{says Joe}.
Who's the old ballocks you were talking to?

\N1:
Old Troy, \sout{says I}, was in the force.
I'm on two minds not to give that
fellow in charge for obstructing the thoroughfare with his brooms and
ladders.

\joe:
What are you doing round those parts? \sout{says Joe.}

\N1:
Devil a much, \sout{says I}.
There's a bloody big foxy thief beyond by the
garrison church at the corner of Chicken lane--old Troy was just giving
me a wrinkle about him--lifted any God's quantity of tea and sugar to pay
three bob a week said he had a farm in the county Down off a
hop-of-my-thumb by the name of Moses Herzog over there near Heytesbury
street.

\joe:
Circumcised? \sout{says Joe.}

\N1:
Ay, \sout{says I}.
A bit off the top. An old plumber named Geraghty. I'm
hanging on to his taw now for the past fortnight and I can't get a penny
out of him.

\joe:
That the lay you're on now? \sout{says Joe.}

\N1:
Ay, \sout{says I}.
How are the mighty fallen! Collector of bad and doubtful
debts. But that's the most notorious bloody robber you'd meet in a day's
walk and the face on him all pockmarks would hold a shower of rain. \emph{Tell
him} says he, \emph{I dare him}, says he, \emph{and I doubledare him to send you round
here again or if he does} says he, \emph{I'll have him summonsed up before the
court, so I will, for trading without a licence.} And he after stuffing
himself till he's fit to burst. Jesus, I had to laugh at the little jewy
getting his shirt out. \emph{He drink me my teas. He eat me my sugars. Because
he no pay me my moneys?}

\N2:
For nonperishable goods bought of Moses Herzog, of 13 Saint
Kevin's parade in the city of Dublin, Wood quay ward, merchant,
hereinafter called the vendor, and sold and delivered to Michael E.
Geraghty, esquire, of 29 Arbour hill in the city of Dublin, Arran quay
ward, gentleman, hereinafter called the purchaser, videlicet, five pounds
avoirdupois of first choice tea at three shillings and no pence per pound
avoirdupois and three stone avoirdupois of sugar, crushed crystal, at
threepence per pound avoirdupois, the said purchaser debtor to the said
vendor of one pound five shillings and sixpence sterling for value
received which amount shall be paid by said purchaser to said vendor in
weekly instalments every seven calendar days of three shillings and no
pence sterling: and the said nonperishable goods shall not be pawned or
pledged or sold or otherwise alienated by the said purchaser but shall be
and remain and be held to be the sole and exclusive property of the said
vendor to be disposed of at his good will and pleasure until the said
amount shall have been duly paid by the said purchaser to the said vendor
in the manner herein set forth as this day hereby agreed between the said
vendor, his heirs, successors, trustees and assigns of the one part and
the said purchaser, his heirs, successors, trustees and assigns of the
other part.

\joe:
Are you a strict t.t.? \sout{says Joe.}

\N1:
Not taking anything between drinks, \sout{says I}.

\joe:
What about paying our respects to our friend? \sout{says Joe.}

\N1:
Who? \sout{says I.}
Sure, he's out in John of God's off his head, poor man.

\joe:
Drinking his own stuff? \sout{says Joe.}

\N1:
Ay, \sout{says I}. Whisky and water on the brain.

\joe:
Come around to Barney Kiernan's, \sout{says Joe}.
I want to see the citizen.

\N1:
Barney mavourneen's be it, \sout{says I}.
Anything strange or wonderful, Joe?

\joe:
Not a word, \sout{says Joe}.
I was up at that meeting in the City Arms.

\N1:
What was that, Joe? \sout{says I.}

\joe:
Cattle traders, \sout{says Joe,}
about the foot and mouth disease. I want to
give the citizen the hard word about it.

\Nq:
So we went around by the Linenhall barracks and the back of the
courthouse talking of one thing or another. Decent fellow Joe when he has
it but sure like that he never has it. Jesus, I couldn't get over that
bloody foxy Geraghty, the daylight robber. For trading without a licence,
\sout{says he}.

\N2:
In Inisfail the fair there lies a land, the land of holy Michan. There
rises a watchtower beheld of men afar. There sleep the mighty dead as in
life they slept, warriors and princes of high renown. A pleasant land it
is in sooth of murmuring waters, fishful streams where sport the gurnard,
the plaice, the roach, the halibut, the gibbed haddock, the grilse,
the dab, the brill, the flounder, the pollock, the mixed coarse fish
generally and other denizens of the aqueous kingdom too numerous to be
enumerated.

\N3:
In the mild breezes of the west and of the east the lofty
trees wave in different directions their firstclass foliage, the wafty
sycamore, the Lebanonian cedar, the exalted planetree, the eugenic
eucalyptus and other ornaments of the arboreal world with which that
region is thoroughly well supplied. Lovely maidens sit in close proximity
to the roots of the lovely trees singing the most lovely songs while they
play with all kinds of lovely objects as for example golden ingots,
silvery fishes, crans of herrings, drafts of eels, codlings, creels of
fingerlings, purple seagems and playful insects.  And heroes voyage from
afar to woo them, from Eblana to Slievemargy, the peerless princes of
unfettered Munster and of Connacht the just and of smooth sleek Leinster
and of Cruahan's land and of Armagh the splendid and of the noble district
of Boyle, princes, the sons of kings.

\N2:
And there rises a shining palace whose crystal glittering roof is seen by
mariners who traverse the extensive sea in barks built expressly for that
purpose, and thither come all herds and fatlings and firstfruits of that
land for O'Connell Fitzsimon takes toll of them, a chieftain descended
from chieftains.

\N3:
Thither the extremely large wains bring foison of the
fields, flaskets of cauliflowers, floats of spinach, pineapple chunks,
Rangoon beans, strikes of tomatoes, drums of figs, drills of Swedes,
spherical potatoes and tallies of iridescent kale, York and Savoy, and
trays of onions, pearls of the earth, and punnets of mushrooms and
custard marrows and fat vetches and bere and rape and red green yellow
brown russet sweet big bitter ripe pomellated apples and chips of
strawberries and sieves of gooseberries, pulpy and pelurious, and
strawberries fit for princes and raspberries from their canes.

\Nq:
\emph{I dare him}, says he, \emph{and I doubledare him}.
Come out here, Geraghty,
you notorious bloody hill and dale robber!

\N2:
And by that way wend the herds innumerable of bellwethers and
flushed ewes and shearling rams and lambs and stubble geese and medium
steers and roaring mares and polled calves and longwoods and storesheep
and Cuffe's prime springers and culls and sowpigs and baconhogs and the
various different varieties of highly distinguished swine and Angus
heifers and polly bullocks of immaculate pedigree together with prime
premiated milchcows and beeves\sout{:}

\N3:
And there is ever heard a trampling,
cackling, roaring, lowing, bleating, bellowing, rumbling, grunting,
champing, chewing, of sheep and pigs and heavyhooved kine from
pasturelands of Lusk and Rush and Carrickmines and from the streamy vales
of Thomond, from the M'Gillicuddy's reeks the inaccessible and lordly
Shannon the unfathomable, and from the gentle declivities of the place of
the race of Kiar, their udders distended with superabundance of milk and
butts of butter and rennets of cheese and farmer's firkins and targets of
lamb and crannocks of corn and oblong eggs in great hundreds, various in
size, the agate with this dun.

\Nq:
So we turned into Barney Kiernan's and there, sure enough, was the citizen
up in the corner having a great confab with himself and that bloody
mangy mongrel, Garryowen, and he waiting for what the sky would drop
in the way of drink.

\N1:
There he is, \sout{says I,}
in his gloryhole, with his cruiskeen lawn and his
load of papers, working for the cause.

\Nq:
The bloody mongrel let a grouse out of him would give you the creeps.  Be
a corporal work of mercy if someone would take the life of that
bloody dog. I'm told for a fact he ate a good part of the breeches off a
constabulary man in Santry that came round one time with a blue paper
about a licence.

\citizen:
Stand and deliver, \sout{says he}.

\joe:
That's all right, citizen, \sout{says Joe}. Friends here.

\citizen:
Pass, friends, \sout{says he}.

\Nq:
Then he rubs his hand in his eye and says he:

\citizen:
What's your opinion of the times?

\Nq:
Doing the rapparee and Rory of the hill. But, begob, Joe was equal to
the occasion.

\joe:
I think the markets are on a rise, \sout{says he,}
\sout{sliding his hand down his fork.}

\Nq:
So begob the citizen claps his paw on his knee and he says:

\citizen:
Foreign wars is the cause of it.

\Nq:
And says Joe, sticking his thumb in his pocket:

\joe:
It's the Russians wish to tyrannise.

\N1:
Arrah, give over your bloody codding, Joe, \sout{says I}.
I've a thirst on me I
wouldn't sell for half a crown.

\joe:
Give it a name, citizen, \sout{says Joe}.

\citizen:
Wine of the country, \sout{says he}.

\joe:
What's yours? \sout{says Joe.}

\N1:
Ditto MacAnaspey, \sout{says I}.

\joe:
Three pints, Terry, \sout{says Joe}.
And how's the old heart, citizen? \sout{says he.}

\citizen:
Never better, \irish{a chara}, \sout{says he}.
What Garry? Are we going to win? Eh?

\Nq:
And with that he took the bloody old towser by the scruff of the neck
and, by Jesus, he near throttled him.

\N2:
The figure seated on a large boulder at the foot of a round tower
was that of a broadshouldered deepchested stronglimbed frankeyed
redhaired freelyfreckled shaggybearded widemouthed largenosed
longheaded deepvoiced barekneed brawnyhanded hairylegged ruddyfaced
sinewyarmed hero. From shoulder to shoulder he measured several ells and
his rocklike mountainous knees were covered, as was likewise the rest of
his body wherever visible, with a strong growth of tawny prickly hair in
hue and toughness similar to the mountain gorse
(\latin{Ulex Europeus}).

\N3:
The widewinged nostrils, from which bristles of the same tawny hue projected,
were of such capaciousness that within their cavernous obscurity the
fieldlark might easily have lodged her nest. The eyes in which a tear and
a smile strove ever for the mastery were of the dimensions of a goodsized
cauliflower. A powerful current of warm breath issued at regular intervals
from the profound cavity of his mouth while in rhythmic resonance the
loud strong hale reverberations of his formidable heart thundered
rumblingly causing the ground, the summit of the lofty tower and the still
loftier walls of the cave to vibrate and tremble.

\N2:
He wore a long unsleeved garment of recently flayed oxhide reaching to the
knees in a loose kilt and this was bound about his middle by a girdle of
plaited straw and rushes. Beneath this he wore trews of deerskin, roughly
stitched with gut. His nether extremities were encased in high Balbriggan
buskins dyed in lichen purple, the feet being shod with brogues of salted
cowhide laced with the windpipe of the same beast. From his girdle hung a
row of seastones which jangled at every movement of his portentous frame
and on these were graven with rude yet striking art the tribal images of
many Irish heroes and heroines of antiquity,

\N2:
Cuchulin,

\N3:
Conn of hundred battles,

\N2:
Niall of nine hostages,

\N3:
Brian of Kincora,

\N2:
the ardri Malachi,

\N3:
Art MacMurragh,

\N2:
Shane O'Neill,

\N3:
Father John Murphy,

\N2:
Owen Roe,

\N3:
Patrick Sarsfield,

\N2:
Red Hugh O'Donnell,

\N3:
Red Jim MacDermott,

\N2:
Soggarth Eoghan O'Growney,

\N3:
Michael Dwyer,

\N2:
Francy Higgins,

\N3:
Henry Joy M'Cracken,

\N2:
Goliath,

\N3:
Horace Wheatley,

\N2:
Thomas Conneff,

\N3:
Peg Woffington,

\N2:
the Village Blacksmith,

\N3:
Captain Moonlight,

\N2:
Captain Boycott,

\N3:
Dante Alighieri,

\N2:
Christopher Columbus,

\N3:
S. Fursa,

\N2:
S. Brendan,

\N3:
Marshal MacMahon,

\N2:
Charlemagne,

\N3:
Theobald Wolfe Tone,

\N2:
the Mother of the Maccabees,

\N3:
the Last of the Mohicans,

\N2:
the Rose of Castile,

\N3:
the Man for Galway,

\N2:
The Man that Broke the Bank at Monte Carlo,

\N3:
The Man in the Gap,

\N2:
The Woman Who Didn't,

\N3:
Benjamin Franklin,

\N2:
Napoleon Bonaparte,

\N3:
John L. Sullivan,

\N2:
Cleopatra,

\N3:
Savourneen Deelish,

\N2:
Julius Caesar,

\N3:
Paracelsus,

\N2:
sir Thomas Lipton,

\N3:
William Tell,

\N2:
Michelangelo Hayes,

\N3:
Muhammad,

\N2:
the Bride of Lammermoor,

\N3:
Peter the Hermit,

\N2:
Peter the Packer,

\N3:
Dark Rosaleen,

\N2:
Patrick W. Shakespeare,

\N3:
Brian Confucius,

\N2:
Murtagh Gutenberg,

\N3:
Patricio Velasquez,

\N2:
Captain Nemo,

\N3:
Tristan and Isolde,

\N2:
the first Prince of Wales,

\N3:
Thomas Cook and Son,

\N2:
the Bold Soldier Boy,

\N3:
Arrah na Pogue,

\N2:
Dick Turpin,

\N3:
Ludwig Beethoven,

\N2:
the Colleen Bawn,

\N3:
Waddler Healy,

\N2:
Angus the Culdee,

\N3:
Dolly Mount,

\N2:
Sidney Parade,

\N3:
Ben Howth,

\N2:
Valentine Greatrakes,

\N3:
Adam and Eve,

\N2:
Arthur Wellesley,

\N3:
Boss Croker,

\N2:
Herodotus,

\N3:
Jack the Giantkiller,

\N2:
Gautama Buddha,

\N3:
Lady Godiva,

\N2:
The Lily of Killarney,

\N3:
Balor of the Evil Eye,

\N2:
the Queen of Sheba,

\N3:
Acky Nagle,

\N2:
Joe Nagle,

\N3:
Alessandro Volta,

\N2:
Jeremiah O'Donovan Rossa,

\N3:
Don Philip O'Sullivan Beare.

\N2:
A couched spear of acuminated granite rested by him while at his feet
reposed a savage animal of the canine tribe whose stertorous gasps
announced that he was sunk in uneasy slumber, a supposition confirmed by
hoarse growls and spasmodic movements which his master repressed from time
to time by tranquilising blows of a mighty cudgel rudely fashioned out of
paleolithic stone.

\Nq:
So anyhow Terry brought the three pints Joe was standing and begob
the sight nearly left my eyes when I saw him land out a quid O, as true as
I'm telling you. A goodlooking sovereign.

\joe:
And there's more where that came from, \sout{says he}.

\N1:
Were you robbing the poorbox, Joe? \sout{says I.}

\joe:
Sweat of my brow, \sout{says Joe}.
'Twas the prudent member gave me the wheeze.

\N1:
I saw him before I met you, \sout{says I},
sloping around by Pill lane and
Greek street with his cod's eye counting up all the guts of the fish.

\N2:
Who comes through Michan's land, bedight in sable armour? O'Bloom,
the son of Rory: it is he. Impervious to fear is Rory's son: he
of the prudent soul.

\citizen:
For the old woman of Prince's street,
\sout{says the citizen}, the subsidised
organ. The pledgebound party on the floor of the house. And look at this
blasted rag, \sout{says he}. Look at this, \sout{says he}.
\emph{The Irish Independent}, if you
please, founded by Parnell to be the workingman's friend. Listen to the
births and deaths in the \emph{Irish all for Ireland Independent}, and I'll thank
you and the marriages.

\Nq:
And he starts reading them out:

\citizen:
Gordon, Barnfield crescent, Exeter;
Redmayne of Iffley, Saint Anne's on
Sea: the wife of William T Redmayne of a son. How's that, eh? Wright and
Flint, Vincent and Gillett to Rotha Marion daughter of Rosa and the late
George Alfred Gillett, 179 Clapham road, Stockwell, Playwood and
Ridsdale at Saint Jude's, Kensington by the very reverend Dr Forrest, dean
of Worcester. Eh? Deaths. Bristow, at Whitehall lane, London: Carr, Stoke
Newington, of gastritis and heart disease: Cockburn, at the Moat house,
Chepstow ...

\joe:
I know that fellow, \sout{says Joe},
from bitter experience.

\citizen:
Cockburn. Dimsey, wife of David Dimsey,
late of the admiralty: Miller,
Tottenham, aged eightyfive: Welsh, June 12, at 35 Canning street,
Liverpool, Isabella Helen. How's that for a national press, eh, my brown
son! How's that for Martin Murphy, the Bantry jobber?

\joe:
Ah, well, \sout{says Joe},
[\stage{handing round the boose.}]
Thanks be to God they had
the start of us. Drink that, citizen.

\citizen:
I will, \sout{says he}, honourable person.

\N1:
Health, Joe, \sout{says I}. And all down the form.

\Nq:
Ah! Ow! Don't be talking! I was blue mouldy for the want of that
pint. Declare to God I could hear it hit the pit of my stomach with a
click.

\N3:
And lo, as they quaffed their cup of joy, a godlike messenger came
swiftly in, radiant as the eye of heaven, a comely youth and behind him
there passed an elder of noble gait and countenance, bearing the sacred
scrolls of law and with him his lady wife a dame of peerless lineage,
fairest of her race.

\Nq:
Little Alf Bergan popped in round the door and hid behind Barney's
snug, squeezed up with the laughing. And who was sitting up there in the
corner that I hadn't seen snoring drunk blind to the world only Bob Doran.
I didn't know what was up and Alf kept making signs out of the door. And
begob what was it only that bloody old pantaloon Denis Breen in his
bathslippers with two bloody big books tucked under his oxter and the wife
hotfoot after him, unfortunate wretched woman, trotting like a poodle. I
thought Alf would split.

\bergan:
Look at him, \sout{says he}. Breen.
He's traipsing all round Dublin with a
postcard someone sent him with U. p: up on it to take a li ...

\Nq:
And he doubled up.

\N1:
Take a what? \sout{says I.}

\bergan:
Libel action \sout{says he},
for ten thousand pounds.

\N1:
O hell! \sout{says I.}

\Nq:
The bloody mongrel began to growl that'd put the fear of God in you
seeing something was up but the citizen gave him a kick in the ribs.

\citizen:
\irish{Bi i dho husht} \sout{says he}.

\joe:
Who? \sout{says Joe.}

\bergan:
Breen \sout{says Alf}.
He was in John Henry Menton's and then he went round
to Collis and Ward's and then Tom Rochford met him and sent him round
to the subsheriff's for a lark. O God, I've a pain laughing. U. p: up. The
long fellow gave him an eye as good as a process and now the bloody old
lunatic is gone round to Green street to look for a G man.

\joe:
When is long John going to hang that fellow in Mountjoy?
\sout{says Joe.}

\doran:
Bergan

\Nq:
says Bob Doran, waking up.

\doran:
Is that Alf Bergan?

\bergan:
Yes \sout{says Alf}. Hanging?
Wait till I show you. Here, Terry, give us a
pony. That bloody old fool! Ten thousand pounds. You should have seen long
John's eye. U. p ...

\Nq:
And he started laughing.

\doran:
Who are you laughing at? \sout{says Bob Doran.}
Is that Bergan?

\bergan:
Hurry up, Terry boy \sout{says Alf}.

\N2:
Terence O'Ryan heard him and straightway brought him a crystal
cup full of the foamy ebon ale which the noble twin brothers Bungiveagh
and Bungardilaun brew ever in their divine alevats, cunning as the sons of
deathless Leda. For they garner the succulent berries of the hop and mass
and sift and bruise and brew them and they mix therewith sour juices and
bring the must to the sacred fire and cease not night or day from their
toil, those cunning brothers, lords of the vat.

\N3:
Then did you, chivalrous Terence, hand forth, as to the manner born,
that nectarous beverage and you offered the crystal cup to him that
thirsted, the soul of chivalry, in beauty akin to the immortals.

\N2:
But he, the young chief of the O'Bergan's, could ill brook to be outdone
in generous deeds but gave therefor with gracious gesture a testoon
of costliest bronze. Thereon embossed in excellent smithwork was seen the
image of a queen of regal port, scion of the house of Brunswick, Victoria
her name, Her Most Excellent Majesty, by grace of God of the United
Kingdom of Great Britain and Ireland and of the British dominions beyond
the sea, queen, defender of the faith, Empress of India, even she, who
bore rule, a victress over many peoples, the wellbeloved, for they knew
and loved her from the rising of the sun to the going down thereof, the
pale, the dark, the ruddy and the ethiop.

\citizen:
What's that bloody freemason doing,
\sout{says the citizen,}
prowling up and down outside?

\joe:
What's that? \sout{says Joe.}

\bergan:
Here you are \sout{says Alf}, chucking out the rhino.
Talking about hanging,
I'll show you something you never saw. Hangmen's letters. Look at here.

[\stage{Bergan takes a bundle of wisps of letters and envelopes out of his pocket.}]

\N1:
Are you codding? \sout{says I.}

\bergan:
Honest injun \sout{says Alf}. Read them.

\Nq:
So Joe took up the letters.

\doran:
Who are you laughing at? says Bob Doran.

\Nq:
So I saw there was going to be a bit of a dust Bob's a queer chap
when the porter's up in him so says I just to make talk:

\N1:
How's Willy Murray those times, Alf?

\bergan:
I don't know, \sout{says Alf}
I saw him just now in Capel street with Paddy
Dignam. Only I was running after that ...

\joe:
You what?
\sout{says Joe,} [\stage{throwing down the letters.}]
With who?

\bergan:
With Dignam \sout{says Alf}.

\joe:
Is it Paddy? \sout{says Joe.}

\bergan:
Yes \sout{says Alf}. Why?

\joe:
Don't you know he's dead? \sout{says Joe.}

\bergan:
Paddy Dignam dead! \sout{says Alf.}

\joe:
Ay \sout{says Joe}.

\bergan:
Sure I'm after seeing him not five minutes ago
\sout{says Alf}, as plain as a pikestaff.

\doran:
Who's dead? \sout{says Bob Doran.}

\joe:
You saw his ghost then \sout{says Joe},
God between us and harm.

\bergan:
What? \sout{says Alf.} Good Christ, only five ...
What? ... And Willy Murray
with him, the two of them there near whatdoyoucallhim's ... What?
Dignam dead?

\doran:
What about Dignam? \sout{says Bob Doran.}
Who's talking about... ?

\bergan:
Dead! \sout{says Alf.}
He's no more dead than you are.

\joe:
Maybe so \sout{says Joe}.
They took the liberty of burying him this morning
anyhow.

\bergan:
Paddy? \sout{says Alf.}

\joe:
Ay \sout{says Joe}.
He paid the debt of nature, God be merciful to him.

\bergan:
Good Christ! \sout{says Alf.}

\Nq:
Begob he was what you might call flabbergasted.

\N3:
In the darkness spirit hands were felt to flutter and when prayer by
tantras had been directed to the proper quarter a faint but increasing
luminosity of ruby light became gradually visible, the apparition of the
etheric double being particularly lifelike owing to the discharge of jivic
rays from the crown of the head and face. Communication was effected
through the pituitary body and also by means of the orangefiery and
scarlet rays emanating from the sacral region and solar plexus. Questioned
by his earthname as to his whereabouts in the heavenworld he stated that
he was now on the path of präläya or return but was still submitted to
trial at the hands of certain bloodthirsty entities on the lower astral
levels.

\N2:
In reply to a question as to his first sensations in the great
divide beyond he stated that previously he had seen as in a glass darkly
but that those who had passed over had summit possibilities of atmic
development opened up to them. Interrogated as to whether life there
resembled our experience in the flesh he stated that he had heard from
more favoured beings now in the spirit that their abodes were equipped
with every modern home comfort such as talafana, alavatar, hatakalda,
wataklasat and that the highest adepts were steeped in waves of volupcy
of the very purest nature.

\N3:
Having requested a quart of buttermilk this was
brought and evidently afforded relief. Asked if he had any message
for the living he exhorted all who were still at the wrong side of Maya
to acknowledge the true path for it was reported in devanic circles that
Mars and Jupiter were out for mischief on the eastern angle where the
ram has power.

\N2:
It was then queried whether there were any special
desires on the part of the defunct and the reply was: \emph{We greet you,
friends of earth, who are still in the body. Mind C. K. doesn't pile it
on.} It was ascertained that the reference was to Mr Cornelius Kelleher,
manager of Messrs H. J. O'Neill's popular funeral establishment, a
personal friend of the defunct, who had been responsible for the carrying
out of the interment arrangements. Before departing he requested that it
should be told to his dear son Patsy that the other boot which he had been
looking for was at present under the commode in the return room and that
the pair should be sent to Cullen's to be soled only as the heels were
still good. He stated that this had greatly perturbed his peace of mind in
the other region and earnestly requested that his desire should be made
known.

\N3:
Assurances were given that the matter would be attended to and it was
intimated that this had given satisfaction.

\N2:
He is gone from mortal haunts: O'Dignam, sun of our morning. Fleet
was his foot on the bracken: Patrick of the beamy brow. Wail, Banba, with
your wind: and wail, O ocean, with your whirlwind.

\citizen:
There he is again.
\sout{says the citizen, staring out.}

\N1:
Who? \sout{says I.}

\citizen:
Bloom \sout{says he}.
He's on point duty up and down there for the last ten
minutes.

\Nq:
And, begob, I saw his physog do a peep in
and then slidder off again.

\Nq:
Little Alf was knocked bawways. Faith, he was.

\bergan:
Good Christ! \sout{says he.}
I could have sworn it was him.

\Nq:
And says Bob Doran, with the hat on the back of his poll, lowest
blackguard in Dublin when he's under the influence:

\doran:
Who said Christ is good?

\bergan:
I beg your parsnips \sout{says Alf}.

\doran:
Is that a good Christ, \sout{says Bob Doran,}
to take away poor little Willy
Dignam?

\bergan:
Ah, well. \sout{says Alf,}
[\stage{trying to pass it off.}]
He's over all his troubles.

\Nq:
But Bob Doran shouts out of him.

\doran:
He's a bloody ruffian, I say,
to take away poor little Willy Dignam.

\Nq:
Terry came down and tipped him the wink to keep quiet, that they
didn't want that kind of talk in a respectable licensed premises. And Bob
Doran starts doing the weeps about Paddy Dignam, true as you're there.

\doran:
The finest man, \sout{says he,}
[\stage{snivelling}] the finest purest character.

\Nq:
The tear is bloody near your eye. Talking through his bloody hat.
Fitter for him go home to the little sleepwalking bitch he married,
Mooney, the bumbailiff's daughter, mother kept a kip in Hardwicke street,
that used to be stravaging about the landings Bantam Lyons told me that
was stopping there at two in the morning without a stitch on her, exposing
her person, open to all comers, fair field and no favour.

\doran:
The noblest, the truest \sout{says he}.
And he's gone, poor little Willy, poor
little Paddy Dignam.

\N2:
And mournful and with a heavy heart he bewept the extinction of that
beam of heaven.

\Nq:
Old Garryowen started growling again at Bloom that was skeezing
round the door.

\citizen:
Come in, come on, he won't eat you,
\sout{says the citizen}.

\Nq:
So Bloom slopes in with his cod's eye on the dog and he asks Terry
was Martin Cunningham there.

\joe:
O, Christ M'Keown. \sout{says Joe,}
[\stage{reading one of the letters.}]
Listen to this, will you?

\Nq:
And he starts reading out one.


\joe:
\emph{7 Hunter Street, Liverpool.\\
To the High Sheriff of Dublin, Dublin.}

\joe:
\emph{Honoured sir i beg to offer my services in the abovementioned painful
case i hanged Joe Gann in Bootle jail on the 12 of Febuary 1900 and i
hanged ...}

\N1:
Show us, Joe \sout{says I}.

\joe:
\emph{... private Arthur Chace for fowl murder of Jessie Tilsit in
Pentonville prison and i was assistant when ...}

\N1:
Jesus \sout{says I}.

\joe:
\emph{... Billington executed the awful murderer Toad Smith ...}

\Nq:
The citizen made a grab at the letter.

\joe:
Hold hard \sout{says Joe},
\emph{I have a special nack of putting the noose once in
he can't get out hoping to be favoured I remain, honoured sir, my terms is
five ginnees.}

\emph{H. Rumbold, \\
Master Barber.}

\citizen:
And a barbarous bloody barbarian he is too,
\sout{says the citizen}.

\joe:
And the dirty scrawl of the wretch \sout{says Joe}.
Here, \sout{says he,} take them
to hell out of my sight, Alf.

[\stage{Enter Bloom.}]

\joe:
Hello, Bloom \sout{says he}, what will you have?

\Nq:
So they started arguing about the point, Bloom saying he wouldn't
and he couldn't and excuse him no offence and all to that and then he said
well he'd just take a cigar. Gob, he's a prudent member and no mistake.

\joe:
Give us one of your prime stinkers,
Terry \sout{says Joe}.

\Nq:
And Alf was telling us there was one chap sent in a mourning card
with a black border round it.

\bergan:
They're all barbers \sout{says he},
from the black country that would hang
their own fathers for five quid down and travelling expenses.

\Nq:
And he was telling us there's two fellows waiting below to pull his
heels down when he gets the drop and choke him properly and then they
chop up the rope after and sell the bits for a few bob a skull.

\N3:
In the dark land they bide, the vengeful knights of the razor. Their
deadly coil they grasp: yea, and therein they lead to Erebus whatsoever
wight hath done a deed of blood for I will on nowise suffer it even so
saith the Lord.

\Nq:
So they started talking about capital punishment and of course Bloom
comes out with the why and the wherefore and all the codology of the
business and the old dog smelling him all the time I'm told those jewies
does have a sort of a queer odour coming off them for dogs about I don't
know what all deterrent effect and so forth and so on.

\bergan:
There's one thing it hasn't a deterrent effect
on \sout{says Alf}.

\joe:
What's that? \sout{says Joe.}

\bergan:
The poor bugger's tool that's being hanged
\sout{says Alf}.

\joe:
That so? \sout{says Joe.}

\bergan:
God's truth \sout{says Alf}.
I heard that from the head warder that was in
Kilmainham when they hanged Joe Brady, the invincible. He told me when
they cut him down after the drop it was standing up in their faces like a
poker.

\joe:
Ruling passion strong in death \sout{says Joe},
as someone said.

\Bloom:
That can be explained by science \sout{says Bloom}.
It's only a natural
phenomenon, don't you see, because on account of the ...

\Nq:
And then he starts with his jawbreakers about phenomenon and
science and this phenomenon and the other phenomenon.

\N2:
The distinguished scientist Herr Professor Luitpold Blumenduft
tendered medical evidence to the effect that the instantaneous fracture of
the cervical vertebrae and consequent scission of the spinal cord would,
according to the best approved tradition of medical science, be calculated
to inevitably produce in the human subject a violent ganglionic stimulus
of the nerve centres of the genital apparatus, thereby causing the elastic
pores of the \latin{corpora cavernosa} to rapidly dilate in such a way as to
instantaneously facilitate the flow of blood to that part of the human
anatomy known as the penis or male organ resulting in the phenomenon which
has been denominated by the faculty a morbid upwards and outwards
philoprogenitive erection \latin{in articulo mortis per diminutionem
capitis}.

\Nq:
So of course the citizen was only waiting for the wink of the word and
he starts gassing out of him about the invincibles and the old guard and
the men of sixtyseven and who fears to speak of ninetyeight and Joe with
him about all the fellows that were hanged, drawn and transported for the
cause by drumhead courtmartial and a new Ireland and new this, that and
the other. Talking about new Ireland he ought to go and get a new dog so
he ought. Mangy ravenous brute sniffing and sneezing all round the place
and scratching his scabs. And round he goes to Bob Doran that was
standing Alf a half one sucking up for what he could get. So of course Bob
Doran starts doing the bloody fool with him:

\doran:
Give us the paw! Give the paw, doggy! Good old doggy! Give the paw
here! Give us the paw!

\Nq:
Arrah, bloody end to the paw he'd paw and Alf trying to keep him
from tumbling off the bloody stool atop of the bloody old dog and he
talking all kinds of drivel about training by kindness and thoroughbred
dog and intelligent dog: give you the bloody pip. Then he starts scraping
a few bits of old biscuit out of the bottom of a Jacobs' tin he told Terry
to bring. Gob, he golloped it down like old boots and his tongue hanging
out of him a yard long for more. Near ate the tin and all, hungry bloody
mongrel.

\Nq:
And the citizen and Bloom having an argument about the point, the
brothers Sheares and Wolfe Tone beyond on Arbour Hill and Robert
Emmet and die for your country, the Tommy Moore touch about Sara
Curran and she's far from the land. And Bloom, of course, with his
knockmedown cigar putting on swank with his lardy face. Phenomenon!
The fat heap he married is a nice old phenomenon with a back on her like a
ballalley. Time they were stopping up in the City Arms pisser Burke told
me there was an old one there with a cracked loodheramaun of a nephew and
Bloom trying to get the soft side of her doing the mollycoddle playing
bezique to come in for a bit of the wampum in her will and not eating meat
of a Friday because the old one was always thumping her craw and taking
the lout out for a walk. And one time he led him the rounds of Dublin and,
by the holy farmer, he never cried crack till he brought him home as drunk
as a boiled owl and he said he did it to teach him the evils of alcohol
and by herrings, if the three women didn't near roast him, it's a queer
story, the old one, Bloom's wife and Mrs O'Dowd that kept the hotel.
Jesus, I had to laugh at pisser Burke taking them off chewing the fat.
And Bloom with his \emph{but don't you see}? and
\emph{but on the other hand}. And sure,
more be token, the lout I'm told was in Power's after, the blender's,
round in Cope street going home footless in a cab five times in the week
after drinking his way through all the samples in the bloody
establishment. Phenomenon!

\citizen:
The memory of the dead

\Nq:
says the citizen taking up his pintglass and
glaring at Bloom.

\joe:
Ay, ay \sout{says Joe}.

\Bloom:
You don't grasp my point \sout{says Bloom}.
What I mean is ...

\citizen:
\irish{Sinn fein!} \sout{says the citizen.}
\irish{Sinn fein amhain!} The friends we love are
by our side and the foes we hate before us.

\N3:
The last farewell was affecting in the extreme. From the belfries far
and near the funereal deathbell tolled unceasingly while all around the
gloomy precincts rolled the ominous warning of a hundred muffled drums
punctuated by the hollow booming of pieces of ordnance. The deafening
claps of thunder and the dazzling flashes of lightning which lit up the
ghastly scene testified that the artillery of heaven had lent its
supernatural pomp to the already gruesome spectacle. A torrential rain
poured down from the floodgates of the angry heavens upon the bared heads
of the assembled multitude which numbered at the lowest computation five
hundred thousand persons.

\N2:
A posse of Dublin Metropolitan police
superintended by the Chief Commissioner in person maintained order in
the vast throng for whom the York street brass and reed band whiled away
the intervening time by admirably rendering on their blackdraped
instruments the matchless melody endeared to us from the cradle by
Speranza's plaintive muse. Special quick excursion trains and upholstered
charabancs had been provided for the comfort of our country cousins of
whom there were large contingents.

\N3:
Considerable amusement was caused
by the favourite Dublin streetsingers L-n-h-n and M-ll-g-n who sang \Title{The
Night before Larry was Stretched} in their usual mirth-provoking fashion.
Our two inimitable drolls did a roaring trade with their broadsheets among
lovers of the comedy element and nobody who has a corner in his heart for
real Irish fun without vulgarity will grudge them their hardearned
pennies. The children of the Male and Female Foundling Hospital who
thronged the windows overlooking the scene were delighted with this
unexpected addition to the day's entertainment and a word of praise is due
to the Little Sisters of the Poor for their excellent idea of affording
the poor fatherless and motherless children a genuinely instructive
treat.

\N2:
The viceregal houseparty which included many wellknown ladies was
chaperoned by Their Excellencies to the most favourable positions on the
grandstand while the picturesque foreign delegation known as the Friends
of the Emerald Isle was accommodated on a tribune directly opposite.
The delegation, present in full force,
consisted of Commendatore Bacibaci Beninobenone (the semiparalysed
\french{doyen} of the party who had
to be assisted to his seat by the aid of a powerful steam crane),

\N3:
Monsieur Pierrepaul Petitepatant,

\N2:
the Grandjoker Vladinmire Pokethankertscheff,

\N3:
the Archjoker Leopold Rudolph von Schwanzenbad-Hodenthaler,

\N2:
Countess Marha Viraga Kisaszony Putrapesthi,

\N3:
Hiram Y. Bomboost,

\N2:
Count Athanatos Karamelopulos,

\N3:
Ali Baba Backsheesh Rahat Lokum Effendi, 

\N2:
Senor Hidalgo Caballero Don Pecadillo y Palabras y Paternoster de la Malora de la Malaria, 

\N3:
Hokopoko Harakiri, 

\N2:
Hi Hung Chang, 

\N3:
Olaf Kobberkeddelsen, 

\N2:
Mynheer Trik van Trumps, 

\N3:
Pan Poleaxe Paddyrisky, 

\N2:
Goosepond Prhklstr Kratchinabritchisitch, 

\N3:
Borus Hupinkoff, 

\N2:
Herr Hurhaus·direktor·president Hans Chuechli-Steuerli, 

\N3:
National·gymnasium·museum·sanatorium·and·suspensoriums·ordinary·privat·docent-
general·history·special·professor·doctor Kriegfried Ueberallgemein.

\N2:
All the delegates without exception expressed themselves
in the strongest possible heterogeneous terms concerning the nameless
barbarity which they had been called upon to witness. An animated
altercation (in which all took part) ensued among the F. O. T. E. I.
as to whether the eighth or the ninth of March was the correct
date of the birth of Ireland's patron saint.

\N3:
In the course of the argument cannonballs,

\N2:
scimitars,

\N3:
boomerangs, 

\N2:
blunderbusses, 

\N3:
stinkpots, 

\N2:
meatchoppers, 

\N3:
umbrellas, 

\N2:
catapults, 

\N3:
knuckledusters, 

\N2:
sandbags, 

\N3:
lumps of pig iron were resorted to and blows were freely exchanged.

\N2:
The baby policeman, Constable MacFadden, summoned by special courier from
Booterstown, quickly restored order and with lightning promptitude
proposed the seventeenth of the month as a solution equally honourable for
both contending parties. The readywitted ninefooter's suggestion at once
appealed to all and was unanimously accepted. Constable MacFadden was
heartily congratulated by all the F.O.T.E.I., several of whom were
bleeding profusely.

\N3:
Commendatore Beninobenone having been extricated
from underneath the presidential armchair, it was explained by his legal
adviser Avvocato Pagamimi that the various articles secreted in his
thirtytwo pockets had been abstracted by him during the affray from the
pockets of his junior colleagues in the hope of bringing them to their
senses. The objects (which included several hundred ladies' and
gentlemen's gold and silver watches) were promptly restored to their
rightful owners and general harmony reigned supreme.

\N2:
Quietly, unassumingly Rumbold stepped on to the scaffold in faultless
morning dress and wearing his favourite flower,
the \latin{Gladiolus Cruentus}.
He announced his presence by that gentle Rumboldian cough which so
many have tried (unsuccessfully) to imitate--short, painstaking yet withal
so characteristic of the man. The arrival of the worldrenowned headsman
was greeted by a roar of acclamation from the huge concourse, the
viceregal ladies waving their handkerchiefs in their excitement while the
even more excitable foreign delegates cheered vociferously in a medley of
cries, \emph{hoch, banzai, eljen, zivio, chinchin, polla kronia, hiphip, vive,
Allah,} amid which the ringing \emph{evviva} of the delegate of the land of song
(a high double F recalling those piercingly lovely notes with which the
eunuch Catalani beglamoured our greatgreatgrandmothers) was easily
distinguishable.

\N3:
It was exactly seventeen o'clock. The signal for prayer
was then promptly given by megaphone and in an instant all heads were
bared, the commendatore's patriarchal sombrero, which has been in the
possession of his family since the revolution of Rienzi, being removed by
his medical adviser in attendance, Dr Pippi. The learned prelate who
administered the last comforts of holy religion to the hero martyr when
about to pay the death penalty knelt in a most christian spirit in a pool
of rainwater, his cassock above his hoary head, and offered up to the
throne of grace fervent prayers of supplication.

\N2:
Hand by the block stood
the grim figure of the executioner, his visage being concealed in a
tengallon pot with two circular perforated apertures through which
his eyes glowered furiously. As he awaited the fatal signal he
tested the edge of his horrible weapon by honing it upon his
brawny forearm or decapitated in rapid succession a flock of
sheep which had been provided by the admirers of his fell but necessary
office. On a handsome mahogany table near him were neatly arranged the
quartering knife, the various finely tempered disembowelling appliances
(specially supplied by the worldfamous firm of cutlers, Messrs John Round
and Sons, Sheffield), a terra cotta saucepan for the reception of the
duodenum, colon, blind intestine and appendix etc when successfully
extracted and two commodious milkjugs destined to receive the most
precious blood of the most precious victim.

\N3:
The housesteward of the
amalgamated cats' and dogs' home was in attendance to convey these
vessels when replenished to that beneficent institution. Quite an
excellent repast consisting of rashers and eggs, fried steak and onions,
done to a nicety, delicious hot breakfast rolls and invigorating tea had
been considerately provided by the authorities for the consumption
of the central figure of the tragedy who was in capital spirits
when prepared for death and evinced the keenest interest in the
proceedings from beginning to end but he, with an abnegation rare
in these our times, rose nobly to the occasion and expressed the
dying wish (immediately acceded to) that the meal should be
divided in aliquot parts among the members of the sick and indigent
roomkeepers' association as a token of his regard and esteem.

\N2:
The \latin{nec} and \latin{non plus ultra}
of emotion were reached when the blushing bride elect burst
her way through the serried ranks of the bystanders and flung herself upon
the muscular bosom of him who was about to be launched into eternity for
her sake. The hero folded her willowy form in a loving embrace murmuring
fondly \emph{Sheila, my own}. Encouraged by this use of her christian name she
kissed passionately all the various suitable areas of his person which the
decencies of prison garb permitted her ardour to reach. She swore to him
as they mingled the salt streams of their tears that she would ever
cherish his memory, that she would never forget her hero boy who went to
his death with a song on his lips as if he were but going to a hurling
match in Clonturk park. She brought back to his recollection the happy
days of blissful childhood together on the banks of Anna Liffey when they
had indulged in the innocent pastimes of the young and, oblivious of the
dreadful present, they both laughed heartily, all the spectators,
including the venerable pastor, joining in the general merriment.

\N3:
That monster audience simply rocked with delight. But anon they were overcome
with grief and clasped their hands for the last time. A fresh torrent of
tears burst from their lachrymal ducts and the vast concourse of people,
touched to the inmost core, broke into heartrending sobs, not the least
affected being the aged prebendary himself. Big strong men, officers of
the peace and genial giants of the royal Irish constabulary,
were making frank use of their handkerchiefs and it is safe to say
that there was not a dry eye in that record assemblage.

\N2:
A most romantic incident occurred when a handsome young Oxford graduate,
noted for his chivalry towards the fair sex, stepped forward and,
presenting his visiting card, bankbook and genealogical tree,
solicited the hand of the hapless young lady, requesting her to
name the day, and was accepted on the spot. Every lady in the
audience was presented with a tasteful souvenir of the occasion
in the shape of a skull and crossbones brooch, a timely and generous
act which evoked a fresh outburst of emotion: and when the gallant
young Oxonian (the bearer, by the way, of one of the most timehonoured
names in Albion's history) placed on the finger of his blushing
\french{fiancée}
an expensive engagement ring with emeralds set in the form of a
fourleaved shamrock the excitement knew no bounds.

\N3:
Nay, even the stern
provostmarshal, lieutenantcolonel Tomkin-Maxwell ffrenchmullan Tomlinson,
who presided on the sad occasion, he who had blown a considerable number
of sepoys from the cannonmouth without flinching, could not now restrain
his natural emotion. With his mailed gauntlet he brushed away a furtive
tear and was overheard, by those privileged burghers who happened to be
in his immediate \french{entourage},
to murmur to himself in a faltering undertone:

\tomlinson:
God blimey if she aint a clinker,
that there bleeding tart. Blimey it
makes me kind of bleeding cry, straight, it does, when I sees her cause I
thinks of my old mashtub what's waiting for me down Limehouse way.

\Nq:
So then the citizen begins talking about the Irish language and the
corporation meeting and all to that and the shoneens that can't speak
their own language and Joe chipping in because he stuck someone for
a quid and Bloom putting in his old goo with his twopenny stump that
he cadged off of Joe and talking about the Gaelic league and the
antitreating league and drink, the curse of Ireland. Antitreating
is about the size of it. Gob, he'd let you pour all manner of drink
down his throat till the Lord would call him before you'd ever
see the froth of his pint. And one night I went in with a fellow
into one of their musical evenings, song and dance about she could
get up on a truss of hay she could my Maureen Lay and there was a fellow
with a Ballyhooly blue ribbon badge spiffing out of him in Irish and a lot
of colleen bawns going about with temperance beverages and selling medals
and oranges and lemonade and a few old dry buns, gob, flahoolagh
entertainment, don't be talking. Ireland sober is Ireland free. And then
an old fellow starts blowing into his bagpipes and all the gougers
shuffling their feet to the tune the old cow died of. And one or two sky
pilots having an eye around that there was no goings on with the females,
hitting below the belt.

\Nq:
So howandever, as I was saying, the old dog seeing the tin was empty
starts mousing around by Joe and me. I'd train him by kindness, so I
would, if he was my dog. Give him a rousing fine kick now and again where
it wouldn't blind him.

\citizen:
Afraid he'll bite you?
\sout{says the citizen,} [\stage{jeering.}]

\N1:
No \sout{says I}. But he might take my leg for a lamppost.

\Nq:
So he calls the old dog over.

\citizen:
What's on you, Garry? \sout{says he.}

\Nq:
Then he starts hauling and mauling and talking to him in Irish and
the old towser growling, letting on to answer, like a duet in the opera.
Such growling you never heard as they let off between them. Someone that
has nothing better to do ought to write a letter \latin{pro bono publico} to the
papers about the muzzling order for a dog the like of that. Growling and
grousing and his eye all bloodshot from the drouth is in it and the
hydrophobia dropping out of his jaws.

\N2:
All those who are interested in the spread of human culture among
the lower animals (and their name is legion) should make a point of not
missing the really marvellous exhibition of cynanthropy given by the
famous old Irish red setter wolfdog formerly known by the \french{sobriquet} of
Garryowen and recently rechristened by his large circle of friends and
acquaintances Owen Garry. The exhibition, which is the result of years of
training by kindness and a carefully thoughtout dietary system, comprises,
among other achievements, the recitation of verse.

\N3:
Our greatest living
phonetic expert (wild horses shall not drag it from us!) has left no stone
unturned in his efforts to delucidate and compare the verse recited and has
found it bears a \emph{striking} resemblance (the italics are ours) to the ranns
of ancient Celtic bards. We are not speaking so much of those delightful
lovesongs with which the writer who conceals his identity under the
graceful pseudonym of the Little Sweet Branch has familiarised the
bookloving world but rather (as a contributor D. O. C. points out in an
interesting communication published by an evening contemporary) of the
harsher and more personal note which is found in the satirical effusions
of the famous Raftery and of Donal MacConsidine to say nothing of a more
modern lyrist at present very much in the public eye.

\N2:
We subjoin a
specimen which has been rendered into English by an eminent scholar
whose name for the moment we are not at liberty to disclose though
we believe that our readers will find the topical allusion rather
more than an indication. The metrical system of the canine original,
which recalls the intricate alliterative and isosyllabic rules of
the Welsh englyn, is infinitely more complicated but we believe our
readers will agree that the spirit has been well caught. Perhaps
it should be added that the effect is greatly increased if Owen's
verse be spoken somewhat slowly and indistinctly in a tone suggestive
of suppressed rancour.

\garryowen:
    The curse of my curses\\
    Seven days every day\\
    And seven dry Thursdays\\
    On you, Barney Kiernan,\\
    Has no sup of water\\
    To cool my courage,\\
    And my guts red roaring\\
    after Lowry's lights.

\Nq:
So he told Terry to bring some water for the dog and, gob, you could
hear him lapping it up a mile off. And Joe asked him would he have
another.

\citizen:
I will \sout{says he}, \irish{a chara},
to show there's no ill feeling.

\Nq:
Gob, he's not as green as he's cabbagelooking. Arsing around from
one pub to another, leaving it to your own honour, with old Giltrap's dog
and getting fed up by the ratepayers and corporators. Entertainment for
man and beast. And says Joe:

\joe:
Could you make a hole in another pint?

\N1:
Could a swim duck? \sout{says I.}

\joe:
Same again, Terry \sout{says Joe}.
Are you sure you won't have anything in the
way of liquid refreshment? \sout{says he.}

\Bloom:
Thank you, no \sout{says Bloom}.
As a matter of fact I just wanted to meet
Martin Cunningham, don't you see, about this insurance of poor Dignam's.
Martin asked me to go to the house. You see, he, Dignam, I mean, didn't
serve any notice of the assignment on the company at the time and
nominally under the act the mortgagee can't recover on the policy.

\joe:
Holy Wars \sout{says Joe},
[\stage{laughing}] that's a good one if old Shylock is
landed. So the wife comes out top dog, what?

\Bloom:
Well, that's a point \sout{says Bloom},
for the wife's admirers.

\joe:
Whose admirers? \sout{says Joe.}

\Bloom:
The wife's advisers, I mean \sout{says Bloom}.

\Nq:
Then he starts all confused mucking it up about mortgagor under the act
like the lord chancellor giving it out on the bench and for the benefit of
the wife and that a trust is created but on the other hand that Dignam
owed Bridgeman the money and if now the wife or the widow contested the
mortgagee's right till he near had the head of me addled with his
mortgagor under the act. He was bloody safe he wasn't run in himself under
the act that time as a rogue and vagabond only he had a friend in court.
Selling bazaar tickets or what do you call it royal Hungarian privileged
lottery. True as you're there. O, commend me to an israelite! Royal and
privileged Hungarian robbery.

\Nq:
So Bob Doran comes lurching around asking Bloom to tell Mrs
Dignam he was sorry for her trouble and he was very sorry about the
funeral and to tell her that he said and everyone who knew him said that
there was never a truer, a finer than poor little Willy that's dead to tell
her. Choking with bloody foolery. And shaking Bloom's hand doing the
tragic to tell her that. Shake hands, brother. You're a rogue and I'm
another.

\doran:
Let me, \sout{said he,}
so far presume upon our acquaintance which, however
slight it may appear if judged by the standard of mere time, is founded,
as I hope and believe, on a sentiment of mutual esteem as to request of
you this favour. But, should I have overstepped the limits of reserve
let the sincerity of my feelings be the excuse for my boldness.

\Bloom:
No, \sout{rejoined the other,}
I appreciate to the full the motives which
actuate your conduct and I shall discharge the office you entrust
to me consoled by the reflection that, though the errand be one of
sorrow, this proof of your confidence sweetens in some measure the
bitterness of the cup.

\doran:
Then suffer me to take your hand, \sout{said he}.
The goodness of your heart, I
feel sure, will dictate to you better than my inadequate words the
expressions which are most suitable to convey an emotion whose
poignancy, were I to give vent to my feelings, would deprive me even of
speech.

\Nq:
And off with him and out trying to walk straight. Boosed at five
o'clock. Night he was near being lagged only Paddy Leonard knew the bobby,
14A. Blind to the world up in a shebeen in Bride street after closing
time, fornicating with two shawls and a bully on guard, drinking porter
out of teacups. And calling himself a Frenchy for the shawls, Joseph
Manuo, and talking against the Catholic religion, and he serving mass in
Adam and Eve's when he was young with his eyes shut, who wrote the new
testament, and the old testament, and hugging and smugging. And the two
shawls killed with the laughing, picking his pockets, the bloody
fool and he spilling the porter all over the bed and the two shawls
screeching laughing at one another. \emph{How is your testament? Have you
got an old testament?} Only Paddy was passing there, I tell you what.
Then see him of a Sunday with his little concubine of a wife, and
she wagging her tail up the aisle of the chapel with her patent boots
on her, no less, and her violets, nice as pie, doing the little lady.
Jack Mooney's sister. And the old prostitute of a mother
procuring rooms to street couples. Gob, Jack made him toe the line. Told
him if he didn't patch up the pot, Jesus, he'd kick the shite out of him.

\Nq:
So Terry brought the three pints.

\joe:
Here. \sout{says Joe,}
[\stage{doing the honours.}] Here, citizen.

\citizen:
\irish{Slán leat} \sout{says he}.

\N1:
Fortune, Joe \sout{says I}. Good health, citizen.

\Nq:
Gob, he had his mouth half way down the tumbler already. Want a
small fortune to keep him in drinks.

\joe:
Who is the long fellow running for the mayoralty,
Alf? \sout{says Joe.}

\bergan:
Friend of yours \sout{says Alf}.

\joe:
Nannan? \sout{says Joe.} The mimber?

\bergan:
I won't mention any names \sout{says Alf}.

\joe:
I thought so \sout{says Joe}.
I saw him up at that meeting now with William
Field, M. P., the cattle traders.

\citizen:
Hairy Iopas, \sout{says the citizen,}
that exploded volcano, the darling of all
countries and the idol of his own.

\Nq:
So Joe starts telling the citizen about the foot and mouth disease and
the cattle traders and taking action in the matter and the citizen sending
them all to the rightabout and Bloom coming out with his sheepdip for the
scab and a hoose drench for coughing calves and the guaranteed remedy
for timber tongue. Because he was up one time in a knacker's yard.
Walking about with his book and pencil here's my head and my heels are
coming till Joe Cuffe gave him the order of the boot for giving lip to a
grazier. Mister Knowall. Teach your grandmother how to milk ducks.
Pisser Burke was telling me in the hotel the wife used to be in rivers of
tears some times with Mrs O'Dowd crying her eyes out with her eight inches
of fat all over her. Couldn't loosen her farting strings but old cod's eye
was waltzing around her showing her how to do it. What's your programme
today? Ay. Humane methods. Because the poor animals suffer and experts
say and the best known remedy that doesn't cause pain to the animal and
on the sore spot administer gently. Gob, he'd have a soft hand under a
hen.

\N3:
Ga Ga Gara. Klook Klook Klook. Black Liz is our hen. She lays eggs
for us. When she lays her egg she is so glad. Gara. Klook Klook Klook.
Then comes good uncle Leo. He puts his hand under black Liz and takes
her fresh egg. Ga ga ga ga Gara. Klook Klook Klook.

\joe:
Anyhow \sout{says Joe},
Field and Nannetti are going over tonight to London
to ask about it on the floor of the house of commons.

\Bloom:
Are you sure \sout{says Bloom},
the councillor is going? I wanted to see him,
as it happens.

\joe:
Well, he's going off by the mailboat
\sout{says Joe}, tonight.

\Bloom:
That's too bad \sout{says Bloom}.
I wanted particularly. Perhaps only Mr Field
is going. I couldn't phone. No. You're sure?

\joe:
Nannan's going too \sout{says Joe}.
The league told him to ask a question
tomorrow about the commissioner of police forbidding Irish games in the
park. What do you think of that, citizen?
\irish{The Sluagh na h-Eireann.}

\N2:
Mr Cowe Conacre (Multifarnham. Nat.):

\conacre:
Arising out of the question of my
honourable friend, the member for Shillelagh, may I ask the right
honourable gentleman whether the government has issued orders that these
animals shall be slaughtered though no medical evidence is forthcoming as
to their pathological condition?

\N3:
Mr Allfours (Tamoshant. Con.):

\allfours:
Honourable members are already in
possession of the evidence produced before a committee of the whole house.
I feel I cannot usefully add anything to that. The answer to the
honourable member's question is in the affirmative.

\N2:
Mr Orelli O'Reilly (Montenotte. Nat.): 

\oreilly:
Have similar orders been issued for
the slaughter of human animals who dare to play Irish games in the
Phoenix park?

\N3:
Mr Allfours: 

\allfours:
The answer is in the negative.

\N2:
Mr Cowe Conacre: 

\conacre:
Has the right honourable gentleman's famous
Mitchelstown telegram inspired the policy of gentlemen on the Treasury
bench? (O! O!)

\N3:
Mr Allfours: 

\allfours:
I must have notice of that question.

\N2:
Mr Staylewit (Buncombe. Ind.): 

\staylewit:
Don't hesitate to shoot.

[\stage{(Ironical opposition cheers.)}]

\N2:
The speaker:

\Speaker:
Order! Order!

[\stage{(The house rises. Cheers.)}]

\joe:
There's the man \sout{says Joe},
that made the Gaelic sports revival. There he
is sitting there. The man that got away James Stephens. The champion of
all Ireland at putting the sixteen pound shot. What was your best throw,
citizen?

\citizen:
\irish{Na bacleis}, \sout{says the citizen,}
letting on to be modest. There was a time
I was as good as the next fellow anyhow.

\joe:
Put it there, citizen \sout{says Joe}.
You were and a bloody sight better.

\bergan:
Is that really a fact? \sout{says Alf.}

\Bloom:
Yes \sout{says Bloom}.
That's well known. Did you not know that?

\Nq:
So off they started about Irish sports and shoneen games the like of lawn
tennis and about hurley and putting the stone and racy of the soil and
building up a nation once again and all to that. And of course Bloom had
to have his say too about if a fellow had a rower's heart violent
exercise was bad. I declare to my antimacassar if you took up a
straw from the bloody floor and if you said to Bloom: \emph{Look at, Bloom.
Do you see that straw? That's a straw.} Declare to my aunt he'd talk
about it for an hour so he would and talk steady.

\N2:
A most interesting discussion took place in the ancient hall of
\irish{Brian O'Ciarnain's} in \irish{Sraid na Bretaine Bheag},
under the auspices of \irish{Sluagh na h-Eireann},
on the revival of ancient Gaelic sports and the importance of
physical culture, as understood in ancient Greece and ancient Rome and
ancient Ireland, for the development of the race. The venerable president
of the noble order was in the chair and the attendance was of large
dimensions. After an instructive discourse by the chairman, a magnificent
oration eloquently and forcibly expressed, a most interesting and
instructive discussion of the usual high standard of excellence
ensued as to the desirability of the revivability of the ancient
games and sports of our ancient Panceltic forefathers.

\N3:
The wellknown and highly respected worker in the cause of our old
tongue, Mr Joseph M'Carthy Hynes, made an eloquent appeal for
the resuscitation of the ancient Gaelic sports and pastimes,
practised morning and evening by Finn MacCool, as calculated to revive the
best traditions of manly strength and prowess handed down to us from
ancient ages.

\N2:
L. Bloom, who met with a mixed reception of applause and
hisses, having espoused the negative the vocalist chairman brought the
discussion to a close, in response to repeated requests and hearty
plaudits from all parts of a bumper house, by a remarkably noteworthy
rendering of the immortal Thomas Osborne Davis' evergreen verses (happily
too familiar to need recalling here) \emph{A Nation Once Again} in the execution
of which the veteran patriot champion may be said without fear of
contradiction to have fairly excelled himself.

\N3:
The Irish Caruso-Garibaldi
was in superlative form and his stentorian notes were heard to the
greatest advantage in the timehonoured anthem sung as only our citizen
can sing it. His superb highclass vocalism, which by its superquality
greatly enhanced his already international reputation, was vociferously
applauded by the large audience among which were to be noticed many
prominent members of the clergy as well as representatives of the press
and the bar and the other learned professions. The proceedings then
terminated.

\N2:
Amongst the clergy present were
the very rev. William Delany, S. J., L. L. D.;

\N3:
the rt rev. Gerald Molloy, D. D.;

\N2:
the rev. P. J. Kavanagh, C. S. Sp.;

\N3:
the rev. T. Waters, C. C.; 

\N2:
the rev. John M. Ivers, P. P.; 

\N3:
the rev. P. J. Cleary, O. S. F.; 

\N2:
the rev. L. J. Hickey, O. P.; 

\N3:
the very rev. Fr. Nicholas, O. S. F. C.; 

\N2:
the very rev. B. Gorman, O. D. C.; 

\N3:
the rev. T. Maher, S. J.; 

\N2:
the very rev. James Murphy, S. J.; 

\N3:
the rev. John Lavery, V. F.; 

\N2:
the very rev. William Doherty, D. D.; 

\N3:
the rev. Peter Fagan, O. M.; 

\N2:
the rev. T. Brangan, O. S. A.; 

\N3:
the rev. J. Flavin, C. C.; 

\N2:
the rev. M. A. Hackett, C. C.; 

\N3:
the rev. W. Hurley, C. C.; 

\N2:
the rt rev. Mgr M'Manus, V. G.; 

\N3:
the rev. B. R. Slattery, O. M. I.; 

\N2:
the very rev. M. D. Scally, P. P.; 

\N3:
the rev. F. T. Purcell, O. P.; 

\N2:
the very rev. Timothy canon Gorman, P. P.; 

\N3:
the rev. J. Flanagan, C. C. 

\N2:
The laity included P. Fay, T. Quirke, etc., etc.

\bergan:
Talking about violent exercise \sout{says Alf},
were you at that Keogh-Bennett match?

\joe:
No \sout{says Joe}.

\bergan:
I heard So and So made a cool hundred quid
over it \sout{says Alf}.

\joe:
Who? Blazes? \sout{says Joe.}

\Nq:
And says Bloom:

\Bloom:
What I meant about tennis, for example,
is the agility and training the eye.

\bergan:
Ay, Blazes \sout{says Alf}.
He let out that Myler was on the beer to run up
the odds and he swatting all the time.

\citizen:
We know him, \sout{says the citizen}.
The traitor's son. We know what put
English gold in his pocket.

\joe:
True for you \sout{says Joe}.

\Nq:
And Bloom cuts in again about lawn tennis and the circulation of the
blood, asking Alf:

\Bloom:
Now, don't you think, Bergan?

\bergan:
Myler dusted the floor with him \sout{says Alf}.
Heenan and Sayers was only a
bloody fool to it. Handed him the father and mother of a beating. See the
little kipper not up to his navel and the big fellow swiping. God, he gave
him one last puck in the wind, Queensberry rules and all, made him puke
what he never ate.

\N3:
It was a historic and a hefty battle when Myler and Percy were
scheduled to don the gloves for the purse of fifty sovereigns. Handicapped
as he was by lack of poundage, Dublin's pet lamb made up for it by
superlative skill in ringcraft. The final bout of fireworks was a
gruelling for both champions. The welterweight sergeantmajor had
tapped some lively claret in the previous mixup during which Keogh
had been receivergeneral of rights and lefts, the artilleryman
putting in some neat work on the pet's nose, and Myler came on
looking groggy.

\N2:
The soldier got to business, leading off with a
powerful left jab to which the Irish gladiator retaliated by shooting
out a stiff one flush to the point of Bennett's jaw. The redcoat
ducked but the Dubliner lifted him with a left hook, the body punch being
a fine one. The men came to handigrips. Myler quickly became busy and got
his man under, the bout ending with the bulkier man on the ropes, Myler
punishing him.

\N3:
The Englishman, whose right eye was nearly closed, took
his corner where he was liberally drenched with water and when the bell
went came on gamey and brimful of pluck, confident of knocking out the
fistic Eblanite in jigtime. It was a fight to a finish and the best man
for it. The two fought like tigers and excitement ran fever high.

\N2:
The referee twice cautioned Pucking Percy for holding but the pet was tricky
and his footwork a treat to watch. After a brisk exchange of courtesies
during which a smart upper cut of the military man brought blood freely
from his opponent's mouth the lamb suddenly waded in all over his man and
landed a terrific left to Battling Bennett's stomach, flooring him flat.
It was a knockout clean and clever. Amid tense expectation the Portobello
bruiser was being counted out when Bennett's second Ole Pfotts Wettstein
threw in the towel and the Santry boy was declared victor to the frenzied
cheers of the public who broke through the ringropes and fairly mobbed him
with delight.

\bergan:
He knows which side his bread is buttered \sout{says Alf}.
I hear he's running
a concert tour now up in the north.

\joe:
He is \sout{says Joe}. Isn't he?

\Bloom:
Who? \sout{says Bloom.}
Ah, yes. That's quite true. Yes, a kind of summer tour,
you see. Just a holiday.

\joe:
Mrs B. is the bright particular star,
isn't she? \sout{says Joe.}

\Bloom:
My wife? \sout{says Bloom.} She's singing, yes.
I think it will be a success too.
He's an excellent man to organise. Excellent.

\Nq:
Hoho begob says I to myself says I. That explains the milk in the cocoanut
and absence of hair on the animal's chest. Blazes doing the tootle on the
flute. Concert tour. Dirty Dan the dodger's son off Island bridge that
sold the same horses twice over to the government to fight the Boers. Old
Whatwhat. I called about the poor and water rate, Mr Boylan. You what?
The water rate, Mr Boylan. You whatwhat? That's the bucko that'll
organise her, take my tip. 'Twixt me and you Caddareesh.

\N3:
Pride of Calpe's rocky mount, the ravenhaired daughter of Tweedy.
There grew she to peerless beauty where loquat and almond scent the air.
The gardens of Alameda knew her step: the garths of olives knew and
bowed. The chaste spouse of Leopold is she: Marion of the bountiful
bosoms.

\N2:
And lo, there entered one of the clan of the O'Molloy's, a comely hero
of white face yet withal somewhat ruddy, his majesty's counsel learned in
the law, and with him the prince and heir of the noble line of Lambert.

\bergan:
Hello, Ned.

\lambert:
Hello, Alf.

\joe:
Hello, Jack.

\jjom:
Hello, Joe.

\citizen:
God save you, \sout{says the citizen}.

\jjom:
Save you kindly \sout{says J. J}.
What'll it be, Ned?

\lambert:
Half one \sout{says Ned}.

\Nq:
So J. J. ordered the drinks.

\joe:
Were you round at the court? \sout{says Joe.}

\jjom:
Yes \sout{says J. J}.
He'll square that, Ned, \sout{says he}.

\lambert:
Hope so \sout{says Ned}.

\Nq:
Now what were those two at? J. J. getting him off the grand jury list
and the other give him a leg over the stile. With his name in Stubbs's.
Playing cards, hobnobbing with flash toffs with a swank glass in their
eye, adrinking fizz and he half smothered in writs and garnishee orders.
Pawning his gold watch in Cummins of Francis street where no-one would
know him in the private office when I was there with Pisser releasing his
boots out of the pop. What's your name, sir? Dunne says he. Ay, and done
says I. Gob, he'll come home by weeping cross one of those days, I'm
thinking.

\bergan:
Did you see that bloody lunatic Breen round there?
\sout{says Alf.} U. p: up.

\jjom:
Yes \sout{says J. J}.
Looking for a private detective.

\lambert:
Ay \sout{says Ned}.
And he wanted right go wrong to address the court only
Corny Kelleher got round him telling him to get the handwriting examined
first.

\bergan:
Ten thousand pounds. \sout{says Alf},
[\stage{laughing}]
God, I'd give anything to hear
him before a judge and jury.

\joe:
Was it you did it, Alf? \sout{says Joe.}
The truth, the whole truth and nothing
but the truth, so help you Jimmy Johnson.

\bergan:
Me? \sout{says Alf.}
Don't cast your nasturtiums on my character.

\joe:
Whatever statement you make \sout{says Joe},
will be taken down in evidence
against you.

\jjom:
Of course an action would lie \sout{says J. J}.
It implies that he is not
\latin{compos mentis}. U. p: up.

\bergan:
\latin{Compos} your eye!
\sout{says Alf,} [\stage{laughing}]
Do you know that he's balmy?
Look at his head. Do you know that some mornings he has to get his hat on
with a shoehorn.

\jjom:
Yes \sout{says J. J.},
but the truth of a libel is no defence to an indictment
for publishing it in the eyes of the law.

\joe:
Ha ha, Alf \sout{says Joe}.

\Bloom:
Still \sout{says Bloom}, on account of the poor woman,
I mean his wife.

\citizen:
Pity about her, \sout{says the citizen}.
Or any other woman marries a half and half.

\Bloom:
How half and half? \sout{says Bloom.}
Do you mean he ...

\citizen:
Half and half I mean, \sout{says the citizen}.
A fellow that's neither fish nor flesh.

\joe:
Nor good red herring \sout{says Joe}.

\citizen:
That what's I mean, \sout{says the citizen}.
A pishogue, if you know what that is.

\Nq:
Begob I saw there was trouble coming. And Bloom explaining he meant on
account of it being cruel for the wife having to go round after the
old stuttering fool. Cruelty to animals so it is to let that bloody
povertystricken Breen out on grass with his beard out tripping him,
bringing down the rain. And she with her nose cockahoop after she married
him because a cousin of his old fellow's was pewopener to the pope.
Picture of him on the wall with his Smashall Sweeney's moustaches, the
signior Brini from Summerhill, the eyetallyano, papal Zouave to the Holy
Father, has left the quay and gone to Moss street. And who was he, tell
us? A nobody, two pair back and passages, at seven shillings a week, and
he covered with all kinds of breastplates bidding defiance to the world.

\jjom:
And moreover \sout{says J. J.},
a postcard is publication. It was held to be
sufficient evidence of malice in the testcase Sadgrove v. Hole. In my
opinion an action might lie.

\Nq:
Six and eightpence, please. Who wants your opinion? Let us drink
our pints in peace. Gob, we won't be let even do that much itself.

\lambert:
Well, good health, Jack \sout{says Ned}.

\jjom:
Good health, Ned \sout{says J. J}.

\joe:
-There he is again \sout{says Joe}.

\bergan:
Where? \sout{says Alf.}

\Nq:
And begob there he was passing the door with his books under his
oxter and the wife beside him and Corny Kelleher with his wall eye looking
in as they went past, talking to him like a father, trying to sell him a
secondhand coffin.

\joe:
How did that Canada swindle case go off? \sout{says Joe.}

\jjom:
Remanded \sout{says J. J}.

\Nq:
One of the bottlenosed fraternity it was went by the name of James
Wought alias Saphiro alias Spark and Spiro, put an ad in the papers saying
he'd give a passage to Canada for twenty bob. What? Do you see any green
in the white of my eye? Course it was a bloody barney. What? Swindled
them all, skivvies and badhachs from the county Meath, ay, and his own
kidney too. J. J. was telling us there was an ancient Hebrew Zaretsky or
something weeping in the witnessbox with his hat on him, swearing by the
holy Moses he was stuck for two quid.

\joe:
Who tried the case? \sout{says Joe.}

\lambert:
Recorder \sout{says Ned}.

\bergan:
Poor old sir Frederick \sout{says Alf},
you can cod him up to the two eyes.

\lambert:
Heart as big as a lion \sout{says Ned}.
Tell him a tale of woe about arrears
of rent and a sick wife and a squad of kids and, faith, he'll dissolve in
tears on the bench.

\bergan:
Ay \sout{says Alf}.
Reuben J was bloody lucky he didn't clap him in the dock
the other day for suing poor little Gumley that's minding stones, for the
corporation there near Butt bridge.

\Nq:
And he starts taking off the old recorder letting on to cry:

\bergan:
A most scandalous thing! This poor hardworking man! How many
children? Ten, did you say?

\bergan:
Yes, your worship. And my wife has the typhoid.

\bergan:
And the wife with typhoid fever! Scandalous! Leave the court
immediately, sir. No, sir, I'll make no order for payment. How dare you,
sir, come up before me and ask me to make an order! A poor hardworking
industrious man! I dismiss the case.

\N3:
And whereas on the sixteenth day of the month of the oxeyed goddess and in
the third week after the feastday of the Holy and Undivided Trinity,
the daughter of the skies, the virgin moon being then in her first
quarter, it came to pass that those learned judges repaired them to the
halls of law. There master Courtenay, sitting in his own chamber,
gave his rede and master Justice Andrews, sitting without a jury
in the probate court, weighed well and pondered the claim of the
first chargeant upon the property in the matter of the will
propounded and final testamentary disposition IN RE the real and
personal estate of the late lamented Jacob Halliday, vintner, deceased,
versus Livingstone, an infant, of unsound mind, and another.

\N2:
And to the
solemn court of Green street there came sir Frederick the Falconer. And he
sat him there about the hour of five o'clock to administer the law of the
brehons at the commission for all that and those parts to be holden in
and for the county of the city of Dublin. And there sat with him the high
sinhedrim of the twelve tribes of Iar, for every tribe one man,
of the tribe of Patrick

\N3:
and of the tribe of Hugh

\N2:
and of the tribe of Owen

\N3:
and of the tribe of Conn

\N2:
and of the tribe of Oscar

\N3:
and of the tribe of Fergus

\N2:
and of the tribe of Finn

\N3:
and of the tribe of Dermot

\N2:
and of the tribe of Cormac

\N3:
and of the tribe of Kevin

\N2:
and of the tribe of Caolte

\N3:
and of the tribe of Ossian,

\N2:
there being in all twelve good men and true.

\N3:
And he conjured them by Him who died on rood that
they should well and truly try and true deliverance make in the
issue joined between their sovereign lord the king and the prisoner at
the bar and true verdict give according to the evidence so help them God
and kiss the book. And they rose in their seats, those twelve of Iar, and
they swore by the name of Him Who is from everlasting that they would do
His rightwiseness.

\N2:
And straightway the minions of the law led forth from
their donjon keep one whom the sleuthhounds of justice had apprehended in
consequence of information received. And they shackled him hand and foot
and would take of him ne bail ne mainprise but preferred a charge against
him for he was a malefactor.

\citizen:
Those are nice things, \sout{says the citizen,}
coming over here to Ireland filling the country with bugs.

\Nq:
So Bloom lets on he heard nothing and he starts talking with Joe, telling
him he needn't trouble about that little matter till the first but if he
would just say a word to Mr Crawford. And so Joe swore high and holy by
this and by that he'd do the devil and all.

\Bloom:
Because, you see \sout{says Bloom},
for an advertisement you must have
repetition. That's the whole secret.

\joe:
Rely on me \sout{says Joe}.

\citizen:
Swindling the peasants, \sout{says the citizen,}
and the poor of Ireland. We
want no more strangers in our house.

\Bloom:
O, I'm sure that will be all right,
Hynes \sout{says Bloom}. It's just that
Keyes, you see.

\joe:
Consider that done \sout{says Joe}.

\Bloom:
Very kind of you \sout{says Bloom}.

\citizen:
The strangers, \sout{says the citizen}.
Our own fault. We let them come in. We
brought them in. The adulteress and her paramour brought the Saxon
robbers here.

\jjom:
Decree \latin{nisi} \sout{says J. J}.

\Nq:
And Bloom letting on to be awfully deeply interested in nothing, a
spider's web in the corner behind the barrel, and the citizen scowling
after him and the old dog at his feet looking up to know who to bite and
when.

\citizen:
A dishonoured wife, \sout{says the citizen,}
that's what's the cause of all our
misfortunes.

\bergan:
And here she is

\Nq:
Says Alf that was giggling over the \Title{Police Gazette}
with Terry on the counter.

\bergan:
In all her warpaint.

\N1:
Give us a squint at her \sout{says I}.

\Nq:
And what was it only one of the smutty yankee pictures Terry
borrows off of Corny Kelleher. Secrets for enlarging your private parts.
Misconduct of society belle. Norman W. Tupper, wealthy Chicago
contractor, finds pretty but faithless wife in lap of officer Taylor.
Belle in her bloomers misconducting herself, and her fancyman feeling for
her tickles and Norman W. Tupper bouncing in with his peashooter just in
time to be late after she doing the trick of the loop with officer Taylor.

\joe:
O jakers, Jenny \sout{says Joe},
how short your shirt is!

\N1:
There's hair, Joe \sout{says I}.
Get a queer old tailend of corned beef off of
that one, what?

\Nq:
So anyhow in came John Wyse Nolan and Lenehan with him with a
face on him as long as a late breakfast.

\citizen:
Well, \sout{says the citizen,}
what's the latest from the scene of action? What
did those tinkers in the city hall at their caucus meeting decide about
the Irish language?

\N3:
O'Nolan, clad in shining armour, low bending made obeisance to the
puissant and high and mighty chief of all Erin and did him to wit of that
which had befallen, how that the grave elders of the most obedient city,
second of the realm, had met them in the tholsel, and there, after due
prayers to the gods who dwell in ether supernal, had taken solemn counsel
whereby they might, if so be it might be, bring once more into honour
among mortal men the winged speech of the seadivided Gael.

\citizen:
It's on the march, \sout{says the citizen}.
To hell with the bloody brutal
Sassenachs and their \french{patois}.

\Nq:
So J. J. puts in a word, doing the toff about one story was good till
you heard another and blinking facts and the Nelson policy, putting your
blind eye to the telescope and drawing up a bill of attainder to impeach a
nation, and Bloom trying to back him up moderation and botheration and
their colonies and their civilisation.

\citizen:
Their syphilisation, you mean, \sout{says the citizen}.
To hell with them! The
curse of a goodfornothing God light sideways on the bloody thicklugged
sons of whores' gets! No music and no art and no literature worthy of the
name. Any civilisation they have they stole from us. Tonguetied sons of
bastards' ghosts.

\jjom:
The European family
\sout{says J. J.} ...

\citizen:
They're not European, \sout{says the citizen}.
I was in Europe with Kevin Egan
of Paris. You wouldn't see a trace of them or their language anywhere in
Europe except in a \french{cabinet d'aisance}.

\Nq:
And says John Wyse:

\johnwyse:
Full many a flower is born to blush unseen.

\Nq:
And says Lenehan that knows a bit of the lingo:

\lenehan:
\french{Conspuez les anglais! Perfide Albion!}

\N2:
He said and then lifted he in his rude great brawny strengthy hands
the medher of dark strong foamy ale and, uttering his tribal slogan \irish{Lamh
Dearg Abu}, he drank to the undoing of his foes, a race of mighty valorous
heroes, rulers of the waves, who sit on thrones of alabaster silent as the
deathless gods.

\N1:
What's up with you, \sout{says I to Lenehan}.
You look like a fellow that had
lost a bob and found a tanner.

\lenehan:
Gold cup \sout{says he}.

\terry:
Who won, Mr Lenehan? \sout{says Terry.}

\lenehan:
\emph{Throwaway} \sout{says he}, at twenty to one.
A rank outsider. And the rest
nowhere.

\terry:
And Bass's mare? \sout{says Terry.}

\lenehan:
Still running \sout{says he}.
We're all in a cart. Boylan plunged two quid on
my tip \emph{Sceptre} for himself and a lady friend.

\terry:
I had half a crown myself \sout{says Terry},
on \emph{Zinfandel} that Mr Flynn gave
me. Lord Howard de Walden's.

\lenehan:
Twenty to one \sout{says Lenehan}.
Such is life in an outhouse. \emph{Throwaway},
says he. Takes the biscuit, and talking about bunions. Frailty, thy name
is \emph{Sceptre}.

\Nq:
So he went over to the biscuit tin Bob Doran left to see if there was
anything he could lift on the nod, the old cur after him backing his luck
with his mangy snout up. Old Mother Hubbard went to the cupboard.

\lenehan:
Not there, my child \sout{says he}.

\joe:
Keep your pecker up \sout{says Joe}.
She'd have won the money only for the
other dog.

\Nq:
And J. J. and the citizen arguing about law and history with Bloom
sticking in an odd word.

\Bloom:
Some people \sout{says Bloom},
can see the mote in others' eyes but they can't
see the beam in their own.

\citizen:
\irish{Raimeis}, \sout{says the citizen}.
There's no-one as blind as the fellow that
won't see, if you know what that means. Where are our missing
twenty millions of Irish should be here today instead of four,
our lost tribes? And our potteries and textiles, the finest in
the whole world! And our wool that was sold in Rome in the time
of Juvenal and our flax and our damask from the looms of Antrim
and our Limerick lace, our tanneries and our white flint glass
down there by Ballybough and our Huguenot poplin that we have since
Jacquard de Lyon and our woven silk and our Foxford tweeds and ivory
raised point from the Carmelite convent in New Ross, nothing like it in
the whole wide world. Where are the Greek merchants that came through the
pillars of Hercules, the Gibraltar now grabbed by the foe of mankind, with
gold and Tyrian purple to sell in Wexford at the fair of Carmen? Read
Tacitus and Ptolemy, even Giraldus Cambrensis. Wine, peltries,
Connemara marble, silver from Tipperary, second to none, our farfamed
horses even today, the Irish hobbies, with king Philip of Spain offering
to pay customs duties for the right to fish in our waters. What do the
yellowjohns of Anglia owe us for our ruined trade and our ruined hearths?
And the beds of the Barrow and Shannon they won't deepen with millions
of acres of marsh and bog to make us all die of consumption?

\johnwyse:
As treeless as Portugal we'll be soon,
\sout{says John Wyse,} or Heligoland
with its one tree if something is not done to reafforest the land.
Larches, firs, all the trees of the conifer family are going fast. I was
reading a report of lord Castletown's ...

\citizen:
Save them, \sout{says the citizen,}
the giant ash of Galway and the chieftain
elm of Kildare with a fortyfoot bole and an acre of foliage. Save the
trees of Ireland for the future men of Ireland on the fair hills of
Eire, O.

\lenehan:
Europe has its eyes on you \sout{says Lenehan}.

\N3:
The fashionable international world attended
\french{en masse} this afternoon
at the wedding of the chevalier Jean Wyse de Neaulan, grand high chief
ranger of the Irish National Foresters,
with Miss Fir Conifer of Pine Valley.

\N2:
Lady Sylvester Elmshade,

\N3:
Mrs Barbara Lovebirch, 

\N2:
Mrs Poll Ash, 

\N3:
Mrs Holly Hazeleyes, 

\N2:
Miss Daphne Bays, 

\N3:
Miss Dorothy Canebrake, 

\N2:
Mrs Clyde Twelvetrees, 

\N3:
Mrs Rowan Greene, 

\N2:
Mrs Helen Vinegadding, 

\N3:
Miss Virginia Creeper, 

\N2:
Miss Gladys Beech, 

\N3:
Miss Olive Garth, 

\N2:
Miss Blanche Maple, 

\N3:
Mrs Maud Mahogany, 

\N2:
Miss Myra Myrtle, 

\N3:
Miss Priscilla Elderflower, 

\N2:
Miss Bee Honeysuckle, 

\N3:
Miss Grace Poplar, 

\N2:
Miss O Mimosa San, 

\N3:
Miss Rachel Cedarfrond, 

\N2:
the Misses Lilian and Viola Lilac, 

\N3:
Miss Timidity Aspenall, 

\N2:
Mrs Kitty Dewey-Mosse, 

\N3:
Miss May Hawthorne, 

\N2:
Mrs Gloriana Palme, 

\N3:
Mrs Liana Forrest, 

\N2:
Mrs Arabella Blackwood

\N3:
and Mrs Norma Holyoake of Oakholme Regis
graced the ceremony by their presence.

\N3:
The bride who was given away by her father, the M'Conifer of
the Glands, looked exquisitely charming in a creation carried out in green
mercerised silk, moulded on an underslip of gloaming grey, sashed with a
yoke of broad emerald and finished with a triple flounce of darkerhued
fringe, the scheme being relieved by bretelles and hip insertions of acorn bronze.

\N2:
The maids of honour, Miss Larch Conifer and Miss Spruce Conifer,
sisters of the bride, wore very becoming costumes in the same tone, a
dainty \french{motif} of plume rose
being worked into the pleats in a pinstripe and
repeated capriciously in the jadegreen toques in the form of heron
feathers of paletinted coral. Senhor Enrique Flor presided at the
organ with his wellknown ability and, in addition to the prescribed
numbers of the nuptial mass, played a new and striking arrangement
of \Title{Woodman, spare that tree} at the conclusion of the
service.

\N3:
On leaving the church of Saint Fiacre \latin{in horto} after the papal
blessing the happy pair were subjected to a playful crossfire
of hazelnuts, beechmast, bayleaves, catkins of willow, ivytod,
hollyberries, mistletoe sprigs and quicken shoots. Mr and Mrs Wyse
Conifer Neaulan will spend a quiet honeymoon in the Black Forest.

\citizen:
And our eyes are on Europe, \sout{says the citizen}.
We had our trade with
Spain and the French and with the Flemings before those mongrels were
pupped, Spanish ale in Galway, the winebark on the winedark waterway.

\joe:
And will again \sout{says Joe}.

\citizen:
And with the help of the holy mother of God
we will again. \sout{says the
citizen,} [\stage{clapping his thigh}]
Our harbours that are empty will be full
again, Queenstown, Kinsale, Galway, Blacksod Bay, Ventry in the kingdom of
Kerry, Killybegs, the third largest harbour in the wide world with a fleet
of masts of the Galway Lynches and the Cavan O'Reillys and the
O'Kennedys of Dublin when the earl of Desmond could make a treaty with
the emperor Charles the Fifth himself. And will again \sout{says he}, when the
first Irish battleship is seen breasting the waves with our own flag to
the fore, none of your Henry Tudor's harps, no, the oldest flag afloat,
the flag of the province of Desmond and Thomond, three crowns on a blue
field, the three sons of Milesius.

\Nq:
And he took the last swig out of the pint. Moya. All wind and piss like
a tanyard cat. Cows in Connacht have long horns. As much as his bloody
life is worth to go down and address his tall talk to the assembled
multitude in Shanagolden where he daren't show his nose with the Molly
Maguires looking for him to let daylight through him for grabbing the
holding of an evicted tenant.

\johnwyse:
Hear, hear to that, \sout{says John Wyse}.
What will you have?

\lenehan:
An imperial yeomanry \sout{says Lenehan},
to celebrate the occasion.

\johnwyse:
Half one, Terry, \sout{says John Wyse,}
and a hands up. Terry! Are you asleep?

\terry:
Yes, sir \sout{says Terry}.
Small whisky and bottle of Allsop. Right, sir.

\Nq:
Hanging over the bloody paper with Alf looking for spicy bits instead
of attending to the general public. Picture of a butting match, trying to
crack their bloody skulls, one chap going for the other with his head down
like a bull at a gate. And another one:
\Title{Black Beast Burned In Omaha, Ga.}
A lot of Deadwood Dicks in slouch hats and they firing at a Sambo strung
up in a tree with his tongue out and a bonfire under him. Gob, they ought
to drown him in the sea after and electrocute and crucify him to make sure
of their job.

\lambert:
But what about the fighting navy \sout{says Ned},
that keeps our foes at bay?

\citizen:
I'll tell you what about it, \sout{says the citizen}.
Hell upon earth it is.
Read the revelations that's going on in the papers about flogging on the
training ships at Portsmouth. A fellow writes that calls himself \emph{Disgusted
One}.

\Nq:
So he starts telling us about corporal punishment and about the crew
of tars and officers and rearadmirals drawn up in cocked hats and the
parson with his protestant bible to witness punishment and a young lad
brought out, howling for his ma, and they tie him down on the buttend of a
gun.

\citizen:
A rump and dozen, \sout{says the citizen,}
was what that old ruffian sir John
Beresford called it but the modern God's Englishman calls it caning on the
breech.

\Nq:
And says John Wyse:

\johnwyse:
'Tis a custom more honoured in the breach
than in the observance.

\Nq:
Then he was telling us the master at arms comes along with a long
cane and he draws out and he flogs the bloody backside off of the poor lad
till he yells meila murder.

\citizen:
That's your glorious British navy,
\sout{says the citizen,}
that bosses the earth.

\citizen:
The fellows that never will be slaves,
with the only hereditary chamber on
the face of God's earth and their land in the hands of a dozen gamehogs
and cottonball barons. That's the great empire they boast about of drudges
and whipped serfs.

\joe:
On which the sun never rises \sout{says Joe}.

\citizen:
And the tragedy of it is, \sout{says the citizen,}
they believe it. The
unfortunate yahoos believe it.

\N2:
They believe in rod, the scourger almighty, creator of hell upon earth,
and in Jacky Tar, the son of a gun, who was conceived of unholy boast,
born of the fighting navy, suffered under rump and dozen, was scarified,
flayed and curried, yelled like bloody hell, the third day he arose again
from the bed, steered into haven, sitteth on his beamend till further
orders whence he shall come to drudge for a living and be paid.

\Bloom:
But \sout{says Bloom},
isn't discipline the same everywhere. I mean wouldn't
it be the same here if you put force against force?

\Nq:
Didn't I tell you? As true as I'm drinking this porter if he was at his
last gasp he'd try to downface you that dying was living.

\citizen:
We'll put force against force, \sout{says the citizen}.
We have our greater
Ireland beyond the sea. They were driven out of house and home in the
black 47. Their mudcabins and their shielings by the roadside were laid
low by the batteringram and the \Title{Times} rubbed its hands and told the
whitelivered Saxons there would soon be as few Irish in Ireland as
redskins in America. Even the Grand Turk sent us his piastres. But the
Sassenach tried to starve the nation at home while the land was full of
crops that the British hyenas bought and sold in Rio de Janeiro. Ay, they
drove out the peasants in hordes. Twenty thousand of them died in the
coffinships. But those that came to the land of the free remember the
land of bondage. And they will come again and with a vengeance, no
cravens, the sons of Granuaile, the champions of Kathleen ni Houlihan.

\Bloom:
Perfectly true \sout{says Bloom}.
But my point was ...

\lambert:
We are a long time waiting for that day,
citizen \sout{says Ned}. Since the
poor old woman told us that the French were on the sea and landed at
Killala.

\johnwyse:
Ay, \sout{says John Wyse}.
We fought for the royal Stuarts that reneged us
against the Williamites and they betrayed us. Remember Limerick and the
broken treatystone. We gave our best blood to France and Spain, the wild
geese. Fontenoy, eh? And Sarsfield and O'Donnell, duke of Tetuan in
Spain, and Ulysses Browne of Camus that was fieldmarshal to Maria Teresa.
But what did we ever get for it?

\citizen:
The French! \sout{says the citizen.}
Set of dancing masters! Do you know what
it is? They were never worth a roasted fart to Ireland. Aren't they
trying to make an \french{entente cordiale} now at Tay Pay's dinnerparty with
perfidious Albion? Firebrands of Europe and they always were.

\lenehan:
\french{Conspuez les français}

\Nq:
says Lenehan, nobbling his beer.

\joe:
And as for the Prooshians and the Hanoverians
\sout{says Joe}, haven't we had
enough of those sausageeating bastards on the throne from George the
elector down to the German lad and the flatulent old bitch that's dead?

\Nq:
Jesus, I had to laugh at the way he came out with that about the old one
with the winkers on her, blind drunk in her royal palace every night of
God, old Vic, with her jorum of mountain dew and her coachman carting her
up body and bones to roll into bed and she pulling him by the whiskers
and singing him old bits of songs about \Title{Ehren on the Rhine}
and come where the boose is cheaper.

\jjom:
Well \sout{says J. J}.
We have Edward the peacemaker now.

\citizen:
Tell that to a fool, \sout{says the citizen}.
There's a bloody sight more pox
than pax about that boyo. Edward Guelph-Wettin!

\joe:
And what do you think \sout{says Joe}, of the holy boys, the priests and
bishops of Ireland doing up his room in Maynooth in His Satanic Majesty's
racing colours and sticking up pictures of all the horses his jockeys
rode. The earl of Dublin, no less.

\bergan:
They ought to have stuck up all the women
he rode himself, \sout{says little
Alf}.

\Nq:
And says J. J.:

\jjom:
Considerations of space influenced
their lordships' decision.

\joe:
Will you try another, citizen? \sout{says Joe.}

\citizen:
Yes, sir \sout{says he}. I will.

\joe:
You? \sout{says Joe.}

\N1:
Beholden to you, Joe \sout{says I}.
May your shadow never grow less.

\joe:
Repeat that dose \sout{says Joe}.

\Nq:
Bloom was talking and talking with John Wyse and he quite excited with
his dunducketymudcoloured mug on him and his old plumeyes rolling about.

\Bloom:
Persecution \sout{says he},
all the history of the world is full of it.
Perpetuating national hatred among nations.

\johnwyse:
But do you know what a nation means?
\sout{says John Wyse.}

\Bloom:
Yes \sout{says Bloom}.

\johnwyse:
What is it? \sout{says John Wyse.}

\Bloom:
A nation? \sout{says Bloom.}
A nation is the same people living in the same place.

\lambert:
By God, then, \sout{says Ned},
[\stage{laughing}] if that's so I'm a nation for I'm
living in the same place for the past five years.

\Nq:
So of course everyone had the laugh at Bloom and says he, trying to
muck out of it:

\Bloom:
Or also living in different places.

\joe:
That covers my case \sout{says Joe}.

\citizen:
What is your nation if I may ask?
\sout{says the citizen.}

\Bloom:
Ireland \sout{says Bloom}.
I was born here. Ireland.

\Nq:
The citizen said nothing only cleared the spit out of his gullet and,
gob, he spat a Red bank oyster out of him right in the corner.

\citizen:
After you with the push, Joe \sout{says he}

\Nq:
taking out his handkerchief to
swab himself dry.

\joe:
Here you are, citizen \sout{says Joe}.
Take that in your right hand and repeat
after me the following words.

\N3:
The muchtreasured and intricately embroidered ancient Irish
facecloth attributed to Solomon of Droma and Manus Tomaltach og
MacDonogh, authors of the Book of Ballymote, was then carefully
produced and called forth prolonged admiration. No need to dwell on the
legendary beauty of the cornerpieces, the acme of art, wherein one can
distinctly discern each of the four evangelists in turn presenting to each
of the four masters his evangelical symbol, a bogoak sceptre, a North
American puma (a far nobler king of beasts than the British article, be it
said in passing), a Kerry calf and a golden eagle from Carrantuohill. The
scenes depicted on the emunctory field, showing our ancient duns and raths
and cromlechs and grianauns and seats of learning and maledictive stones,
are as wonderfully beautiful and the pigments as delicate as when the
Sligo illuminators gave free rein to their artistic fantasy long long ago
in the time of the Barmecides.

\N2:
Glendalough,

\N3:
the lovely lakes of Killarney,

\N2:
the ruins of Clonmacnois,

\N3:
Cong Abbey,

\N2:
Glen Inagh and the Twelve Pins,

\N3:
Ireland's Eye,

\N2:
the Green Hills of Tallaght,

\N3:
Croagh Patrick,

\N2:
the brewery of Messrs Arthur Guinness, Son and Company (Limited),

\N3:
Lough Neagh's banks,

\N2:
the vale of Ovoca,

\N3:
Isolde's tower,

\N2:
the Mapas obelisk, 

\N3:
Sir Patrick Dun's hospital, 

\N2:
Cape Clear, 

\N3:
the glen of Aherlow, 

\N2:
Lynch's castle, 

\N3:
the Scotch house, 

\N2:
Rathdown Union Workhouse at Loughlinstown, 

\N3:
Tullamore jail, 

\N2:
Castleconnel rapids, 

\N3:
Kilballymacshonakill, 

\N2:
the cross at Monasterboice, 

\N3:
Jury's Hotel, 

\N2:
S. Patrick's Purgatory, 

\N3:
the Salmon Leap, 

\N2:
Maynooth college refectory, 

\N3:
Curley's hole, 

\N2:
the three birthplaces of the first duke of Wellington, 

\N3:
the rock of Cashel, 

\N2:
the bog of Allen, 

\N3:
the Henry Street Warehouse, 

\N2:
Fingal's Cave

\N3:
\sout{--}all these moving scenes are still there for us
today rendered more beautiful still by the waters of sorrow which have
passed over them and by the rich incrustations of time.

\N1:
Show us over the drink \sout{says I}.
Which is which?

\joe:
That's mine \sout{says Joe},
as the devil said to the dead policeman.

\Bloom:
And I belong to a race too \sout{says Bloom},
that is hated and persecuted.
Also now. This very moment. This very instant.

\Nq:
Gob, he near burnt his fingers with the butt of his old cigar.

\Bloom:
Robbed \sout{says he}. Plundered. Insulted.
Persecuted. Taking what belongs
to us by right. At this very moment \sout{says he},
[\stage{putting up his fist}]
sold by auction in Morocco like slaves or cattle.

\citizen:
Are you talking about the new Jerusalem?
\sout{says the citizen.}

\Bloom:
I'm talking about injustice \sout{says Bloom}.

\johnwyse:
Right, \sout{says John Wyse}.
Stand up to it then with force like men.

\Nq:
That's an almanac picture for you. Mark for a softnosed bullet. Old
lardyface standing up to the business end of a gun. Gob, he'd adorn a
sweepingbrush, so he would, if he only had a nurse's apron on him. And
then he collapses all of a sudden, twisting around all the opposite, as
limp as a wet rag.

\Bloom:
But it's no use \sout{says he}.
Force, hatred, history, all that. That's not
life for men and women, insult and hatred. And everybody knows that it's
the very opposite of that that is really life.

\bergan:
What? \sout{says Alf.}

\Bloom:
Love \sout{says Bloom}.
I mean the opposite of hatred. I must go now,
[\stage{says he to John Wyse.}]
Just round to the court a moment to see if Martin is there.
If he comes just say I'll be back in a second. Just a moment.

\Nq:
Who's hindering you? And off he pops like greased lightning.

[\stage{Exit Bloom.}]

\citizen:
A new apostle to the gentiles, \sout{says the citizen}.
Universal love.

\johnwyse:
Well \sout{says John Wyse}.
Isn't that what we're told. Love your neighbour.

\citizen:
That chap? \sout{says the citizen.}
Beggar my neighbour is his motto. Love,
moya! He's a nice pattern of a Romeo and Juliet.

\N3:
Love loves to love love.

\N2:
Nurse loves the new chemist. 

\N3:
Constable 14A loves Mary Kelly. 

\N2:
Gerty MacDowell loves the boy that has the bicycle. 

\N3:
M. B. loves a fair gentleman. 

\N2:
Li Chi Han lovey up kissy Cha Pu Chow. 

\N3:
Jumbo, the elephant, loves Alice, the elephant. 

\N2:
Old Mr Verschoyle with the ear trumpet loves
old Mrs Verschoyle with the turnedin eye. 

\N3:
The man in the brown macintosh loves a lady who is dead. 

\N2:
His Majesty the King loves Her Majesty the Queen. 

\N3:
Mrs Norman W. Tupper loves officer Taylor. 

\N2:
You love a certain person. 

\N3:
And this person loves that other person because everybody loves somebody
but God loves everybody.

\N1:
Well, Joe \sout{says I},
your very good health and song. More power, citizen.

\joe:
Hurrah, there \sout{says Joe}.

\citizen:
The blessing of God and Mary and Patrick on you,
\sout{says the citizen}.

\Nq:
And he ups with his pint to wet his whistle.

\citizen:
We know those canters \sout{says he},
preaching and picking your pocket.
What about sanctimonious Cromwell and his ironsides that put the women
and children of Drogheda to the sword with the bible text \emph{God is love}
pasted round the mouth of his cannon? The bible! Did you read that skit in
the \Title{United Irishman} today
about that Zulu chief that's visiting England?

\joe:
What's that? \sout{says Joe.}

\Nq:
So the citizen takes up one of his paraphernalia papers and he starts
reading out:

\citizen:
A delegation of the chief cotton
magnates of Manchester was presented
yesterday to His Majesty the Alaki of Abeakuta by Gold Stick in Waiting,
Lord Walkup of Walkup on Eggs, to tender to His Majesty the heartfelt
thanks of British traders for the facilities afforded them in his
dominions. The delegation partook of luncheon at the conclusion
of which the dusky potentate, in the course of a happy speech,
freely translated by the British chaplain, the reverend Ananias
Praisegod Barebones, tendered his best thanks to Massa Walkup and
emphasised the cordial relations existing between Abeakuta and the
British empire, stating that he treasured as one of his dearest
possessions an illuminated bible, the volume of the word of God
and the secret of England's greatness, graciously presented to him by
the white chief woman, the great squaw Victoria, with a personal
dedication from the august hand of the Royal Donor. The Alaki then drank a
lovingcup of firstshot usquebaugh to the toast \emph{Black and White} from the
skull of his immediate predecessor in the dynasty Kakachakachak,
surnamed Forty Warts, after which he visited the chief factory of
Cottonopolis and signed his mark in the visitors' book, subsequently
executing a charming old Abeakutic wardance, in the course of which he
swallowed several knives and forks, amid hilarious applause from the girl
hands.

\lambert:
Widow woman \sout{says Ned}.
I wouldn't doubt her. Wonder did he put that
bible to the same use as I would.

\lenehan:
Same only more so \sout{says Lenehan}.
And thereafter in that fruitful land
the broadleaved mango flourished exceedingly.

\johnwyse:
Is that by Griffith?
\sout{says John Wyse.}

\citizen:
No, \sout{says the citizen}.
It's not signed Shanganagh. It's only
initialled: P.

\joe:
And a very good initial too \sout{says Joe}.

\citizen:
That's how it's worked, \sout{says the citizen}.
Trade follows the flag.

\jjom:
Well \sout{says J. J.},
if they're any worse than those Belgians in the Congo
Free State they must be bad. Did you read that report by a man what's this
his name is?

\citizen:
Casement, \sout{says the citizen}.
He's an Irishman.

\jjom:
Yes, that's the man \sout{says J. J}.
Raping the women and girls and flogging
the natives on the belly to squeeze all the red rubber they can out of
them.

\lenehan:
I know where he's gone \sout{says Lenehan},
[\stage{cracking his fingers.}]

\N1:
Who? \sout{says I.}

\lenehan:
Bloom \sout{says he}.
The courthouse is a blind. He had a few bob on
\emph{Throwaway} and he's gone to gather in the shekels.

\citizen:
Is it that whiteeyed kaffir?

\Nq:
says the citizen, that never backed a horse
in anger in his life?

\lenehan:
That's where he's gone \sout{says Lenehan}.
I met Bantam Lyons going to back
that horse only I put him off it and he told me Bloom gave him the tip.
Bet you what you like he has a hundred shillings to five on. He's the only
man in Dublin has it. A dark horse.

\joe:
He's a bloody dark horse himself \sout{says Joe}.

\N1:
Mind, Joe \sout{says I}. Show us the entrance out.

\terry:
There you are \sout{says Terry}.

\Nq:
Goodbye Ireland I'm going to Gort. So I just went round the back of
the yard to pumpship and begob (hundred shillings to five) while I was
letting off my (\emph{Throwaway} twenty to) letting off my load gob says I to
myself I knew he was uneasy in his (two pints off of Joe and one in
Slattery's off) in his mind to get off the mark to (hundred shillings is
five quid) and when they were in the (dark horse) pisser Burke
was telling me card party and letting on the child was sick (gob, must
have done about a gallon) flabbyarse of a wife speaking down the tube
\emph{she's better} or \emph{she's} (ow!) all a plan so he could vamoose with the
pool if he won or (Jesus, full up I was) trading without a licence (ow!)
Ireland my nation says he (hoik! phthook!) never be up to those
bloody (there's the last of it) Jerusalem (ah!) cuckoos.

\Nq:
So anyhow when I got back they were at it dingdong, John Wyse
saying it was Bloom gave the ideas for Sinn Fein to Griffith to put in his
paper all kinds of jerrymandering, packed juries and swindling the taxes
off of the government and appointing consuls all over the world to walk
about selling Irish industries. Robbing Peter to pay Paul. Gob, that puts
the bloody kybosh on it if old sloppy eyes is mucking up the show. Give us
a bloody chance. God save Ireland from the likes of that bloody
mouseabout. Mr Bloom with his argol bargol. And his old fellow before him
perpetrating frauds, old Methusalem Bloom, the robbing bagman, that
poisoned himself with the prussic acid after he swamping the country with
his baubles and his penny diamonds. Loans by post on easy terms. Any
amount of money advanced on note of hand. Distance no object. No security.
Gob, he's like Lanty MacHale's goat that'd go a piece of the road with
every one.

\johnwyse:
Well, it's a fact, \sout{says John Wyse}.
And there's the man now that'll tell
you all about it, Martin Cunningham.

\Nq:
Sure enough the castle car drove up with Martin on it and Jack Power
with him and a fellow named Crofter or Crofton, pensioner out of the
collector general's, an orangeman Blackburn does have on the registration
and he drawing his pay or Crawford gallivanting around the country at the
king's expense.

\N2:
Our travellers reached the rustic hostelry and alighted from their
palfreys.

\cunningham:
Ho, varlet!

\N2:
cried he, who by his mien seemed the leader of the party.

\cunningham:
Saucy knave! To us!

\N2:
So saying he knocked loudly with his swordhilt upon the open lattice.

\N3:
Mine host came forth at the summons, girding him with his tabard.

\terry:
Give you good den, my masters,

\N3:
said he with an obsequious bow.

\cunningham:
Bestir thyself, sirrah!

\N2:
cried he who had knocked.

\cunningham:
Look to our steeds.
And for ourselves give us of your best for ifaith we need it.

\terry:
Lackaday, good masters, \sout{said the host,}
my poor house has but a bare
larder. I know not what to offer your lordships.

\power:
How now, fellow?

\N2:
cried the second of the party, a man of pleasant countenance.

\power:
So servest thou the king's messengers, master Taptun?

\N3:
An instantaneous change overspread the landlord's visage.

\terry:
Cry you mercy, gentlemen,

\N3:
he said humbly.

\terry:
An you be the king's messengers

\N3:
(God shield His Majesty!)

\terry:
you shall not want for aught. The king's friends

\N3:
(God bless His Majesty!)

\terry:
shall not go afasting in my house I warrant me.

\crofton:
Then about!

\N2:
cried the traveller who had not spoken,
a lusty trencherman by his aspect.

\crofton:
Hast aught to give us?

\N3:
Mine host bowed again as he made answer:

\terry:
What say you, good masters,
to a squab pigeon pasty, some collops of
venison, a saddle of veal, widgeon with crisp hog's bacon, a boar's head
with pistachios, a bason of jolly custard, a medlar tansy and a flagon of
old Rhenish?

\crofton:
Gadzooks! \sout{cried the last speaker.}
That likes me well. Pistachios!

\power:
Aha! \sout{cried he of the pleasant countenance.}
A poor house and a bare
larder, quotha! 'Tis a merry rogue.

\Nq:
So in comes Martin asking where was Bloom.

\lenehan:
Where is he? \sout{says Lenehan.}
Defrauding widows and orphans.

\johnwyse:
Isn't that a fact, \sout{says John Wyse,}
what I was telling the citizen about
Bloom and the Sinn Fein?

\cunningham:
That's so \sout{says Martin}.
Or so they allege.

\bergan:
Who made those allegations? \sout{says Alf.}

\joe:
I \sout{says Joe}. I'm the alligator.

\johnwyse:
And after all, \sout{says John Wyse,}
why can't a jew love his country like the
next fellow?

\jjom:
Why not? \sout{says J. J.},
when he's quite sure which country it is.

\lambert:
Is he a jew or a gentile or a holy Roman
or a swaddler or what the hell
is he? \sout{says Ned.} Or who is he? No offence, Crofton.

\jjom:
Who is Junius? \sout{says J. J.}

\crofton:
We don't want him,

\Nq:
says Crofter the Orangeman or presbyterian.

\cunningham:
He's a perverted jew \sout{says Martin},
from a place in Hungary and it was he
drew up all the plans according to the Hungarian system. We know that in
the castle.

\power:
Isn't he a cousin of Bloom the dentist?
\sout{says Jack Power.}

\cunningham:
Not at all \sout{says Martin}.
Only namesakes. His name was Virag, the
father's name that poisoned himself. He changed it by deedpoll, the father
did.

\citizen:
That's the new Messiah for Ireland!
\sout{says the citizen.}
Island of saints and sages!

\cunningham:
Well, they're still waiting for their redeemer
\sout{says Martin}. For that
matter so are we.

\jjom:
Yes \sout{says J. J.},
and every male that's born they think it may be their
Messiah. And every jew is in a tall state of excitement, I believe, till
he knows if he's a father or a mother.

\lenehan:
Expecting every moment will be his next
\sout{says Lenehan}.

\lambert:
O, by God \sout{says Ned},
you should have seen Bloom before that son of his
that died was born. I met him one day in the south city markets buying a
tin of Neave's food six weeks before the wife was delivered.

\jjom:
\french{En ventre sa mère} \sout{says J. J}.

\citizen:
Do you call that a man?
\sout{says the citizen.}

\joe:
I wonder did he ever put it out of sight
\sout{says Joe}.

\power:
Well, there were two children born anyhow,
\sout{says Jack Power}.

\citizen:
And who does he suspect?
\sout{says the citizen.}

\Nq:
Gob, there's many a true word spoken in jest. One of those mixed
middlings he is. Lying up in the hotel Pisser was telling me once a month
with headache like a totty with her courses. Do you know what I'm telling
you? It'd be an act of God to take a hold of a fellow the like of that and
throw him in the bloody sea. Justifiable homicide, so it would. Then
sloping off with his five quid without putting up a pint of stuff like a
man. Give us your blessing. Not as much as would blind your eye.

\cunningham:
Charity to the neighbour \sout{says Martin}.
But where is he? We can't wait.

\citizen:
A wolf in sheep's clothing,
\sout{says the citizen}. That's what he is. Virag
from Hungary! Ahasuerus I call him. Cursed by God.

\lambert:
Have you time for a brief libation,
Martin? \sout{says Ned.}

\cunningham:
Only one \sout{says Martin}.
We must be quick. J. J. and S.

\lambert:
You, Jack? Crofton? Three half ones, Terry.

\citizen:
Saint Patrick would want to land again
at Ballykinlar and convert us,
\sout{says the citizen,} after allowing things like that to contaminate our
shores.

\cunningham:
Well. \sout{says Martin},
[\stage{rapping for his glass}] God bless all here is my
prayer.

\citizen:
Amen, \sout{says the citizen}.

\joe:
And I'm sure He will \sout{says Joe}.

\N2:
And at the sound of the sacring bell,
headed by a crucifer with acolytes,

\N3:
thurifers,

\N2:
boatbearers,

\N3:
readers,

\N2:
ostiarii,

\N3:
deacons and subdeacons,

\N2:
the blessed company drew nigh of mitred abbots
and priors and guardians and monks and friars:

\N3:
the monks of Benedict of Spoleto,

\N2:
Carthusians and Camaldolesi,

\N3:
Cistercians and Olivetans,

\N2:
Oratorians and Vallombrosans,

\N3:
and the friars of Augustine,

\N2:
Brigittines,

\N3:
Premonstratensians,

\N2:
Servi,

\N3:
Trinitarians,

\N2:
and the children of Peter Nolasco:

\N3:
and therewith from Carmel mount the children of
Elijah prophet led by Albert bishop and by Teresa of Avila,
calced and other:

\N2:
and friars, brown and grey,

\N3:
sons of poor Francis,

\N2:
capuchins,

\N3:
cordeliers,

\N2:
minimes and observants and the daughters of Clara:

\N3:
and the sons of Dominic, the friars preachers,

\N2:
and the sons of Vincent:

\N3:
and the monks of S. Wolstan:

\N2:
and Ignatius his children:

\N3:
and the confraternity of the christian brothers
led by the reverend brother Edmund Ignatius Rice.

\N2:
And after came all saints and martyrs,

\N3:
virgins and confessors:

\N2:
S. Cyr

\N3:
and S. Isidore Arator

\N2:
and S. James the Less 

\N3:
and S. Phocas of Sinope 

\N2:
and S. Julian Hospitator 

\N3:
and S. Felix de Cantalice 

\N2:
and S. Simon Stylites 

\N3:
and S. Stephen Protomartyr 

\N2:
and S. John of God 

\N3:
and S. Ferreol 

\N2:
and S. Leugarde 

\N3:
and S. Theodotus 

\N2:
and S. Vulmar 

\N3:
and S. Richard 

\N2:
and S. Vincent de Paul 

\N3:
and S. Martin of Todi 

\N2:
and S. Martin of Tours 

\N3:
and S. Alfred 

\N2:
and S. Joseph 

\N3:
and S. Denis 

\N2:
and S. Cornelius 

\N3:
and S. Leopold 

\N2:
and S. Bernard 

\N3:
and S. Terence 

\N2:
and S. Edward 

\N3:
and S. Owen Caniculus 

\N2:
and S. Anonymous 

\N3:
and S. Eponymous 

\N2:
and S. Pseudonymous 

\N3:
and S. Homonymous 

\N2:
and S. Paronymous 

\N3:
and S. Synonymous 

\N2:
and S. Laurence O'Toole 

\N3:
and S. James of Dingle and Compostella 

\N2:
and S. Columcille 

\N3:
and S. Columba 

\N2:
and S. Celestine 

\N3:
and S. Colman 

\N2:
and S. Kevin 

\N3:
and S. Brendan 

\N2:
and S. Frigidian 

\N3:
and S. Senan 

\N2:
and S. Fachtna 

\N3:
and S. Columbanus 

\N2:
and S. Gall 

\N3:
and S. Fursey 

\N2:
and S. Fintan 

\N3:
and S. Fiacre 

\N2:
and S. John Nepomuc 

\N3:
and S. Thomas Aquinas 

\N2:
and S. Ives of Brittany 

\N3:
and S. Michan 

\N2:
and S. Herman-Joseph 

\N3:
and the three patrons of holy youth S. Aloysius Gonzaga 

\N2:
and S. Stanislaus Kostka 

\N3:
and S. John Berchmans 

\N2:
and the saints Gervasius, Servasius and Bonifacius 

\N3:
and S. Bride 

\N2:
and S. Kieran 

\N3:
and S. Canice of Kilkenny 

\N2:
and S. Jarlath of Tuam 

\N3:
and S. Finbarr 

\N2:
and S. Pappin of Ballymun 

\N3:
and Brother Aloysius Pacificus 

\N2:
and Brother Louis Bellicosus 

\N3:
and the saints Rose of Lima 

\N2:
and of Viterbo 

\N3:
and S. Martha of Bethany 

\N2:
and S. Mary of Egypt 

\N3:
and S. Lucy 

\N2:
and S. Brigid 

\N3:
and S. Attracta 

\N2:
and S. Dympna 

\N3:
and S. Ita 

\N2:
and S. Marion Calpensis 

\N3:
and the Blessed Sister Teresa of the Child Jesus 

\N2:
and S. Barbara 

\N3:
and S. Scholastica 

\N2:
and S. Ursula with eleven thousand virgins.

\N3:
And all came with nimbi and aureoles and gloriae,

\N2:
bearing palms and harps and swords and olive crowns,

\N3:
in robes whereon were woven the blessed symbols of their efficacies,

\N2:
inkhorns,

\N3:
arrows,

\N2:
loaves,

\N3:
cruses,

\N2:
fetters,

\N3:
axes,

\N2:
trees,

\N3:
bridges,

\N2:
babes in a bathtub,

\N3:
shells,

\N2:
wallets,

\N3:
shears,

\N2:
keys,

\N3:
dragons,

\N2:
lilies,

\N3:
buckshot,

\N2:
beards,

\N3:
hogs,

\N2:
lamps,

\N3:
bellows,

\N2:
beehives,

\N3:
soupladles,

\N2:
stars,

\N3:
snakes,

\N2:
anvils,

\N3:
boxes of vaseline,

\N2:
bells,

\N3:
crutches,

\N2:
forceps,

\N3:
stags' horns,

\N2:
watertight boots,

\N3:
hawks,

\N2:
millstones,

\N3:
eyes on a dish,

\N2:
wax candles,

\N3:
aspergills,

\N2:
unicorns.

\N3:
And as they wended their way by Nelson's Pillar,

\N2:
Henry street,

\N3:
Mary street,

\N2:
Capel street,

\N3:
Little Britain street
chanting the introit \latin{in epiphania domini} which beginneth
\latin{surge, illuminare} and thereafter most sweetly the gradual
\latin{omnes} which saith \french{de saba venient}
they did divers wonders such as casting out devils,

\N2:
raising the dead to life,

\N3:
multiplying fishes,

\N2:
healing the halt and the blind,

\N3:
discovering various articles which had been mislaid,

\N2:
interpreting and fulfilling the scriptures,

\N3:
blessing and prophesying.

\N2:
And last, beneath
a canopy of cloth of gold came the reverend Father O'Flynn attended by
Malachi and Patrick. And when the good fathers had reached the appointed place,
the house of Bernard Kiernan and Co, limited, 8, 9 and 10 little Britain street,

\N3:
wholesale grocers,

\N2:
wine and brandy shippers,

\N3:
licensed for the sale of beer, wine and spirits for consumption on the premises,

\N2:
the celebrant blessed the house

\N3:
and censed the mullioned windows 

\N2:
and the groynes 

\N3:
and the vaults 

\N2:
and the arrises 

\N3:
and the capitals 

\N2:
and the pediments 

\N3:
and the cornices 

\N2:
and the engrailed arches 

\N3:
and the spires 

\N2:
and the cupolas 

\N3:
and sprinkled the lintels thereof with blessed water

\N2:
and prayed that God might bless that house as he had
blessed the house of Abraham and Isaac and Jacob and
make the angels of His light to inhabit therein.
And entering he blessed the viands and the beverages and the company of
all the blessed answered his prayers.

\N3:
\latin{Adiutorium nostrum in nomine Domini.}

\N2:
\latin{Qui fecit coelum et terram.}

\N3:
\latin{Dominus vobiscum.}

\N2:
\latin{Et cum spiritu tuo.}

\N2:
And he laid his hands upon that he blessed and gave thanks and he
prayed and they all with him prayed:

\All:
\latin{Deus, cuius verbo sanctificantur omnia,
benedictionem tuam effunde super
creaturas istas: et praesta ut quisquis eis secundum legem et voluntatem
Tuam cum gratiarum actione usus fuerit per invocationem sanctissimi
nominis Tui corporis sanitatem et animae tutelam Te auctore percipiat per
Christum Dominum nostrum.}

\power:
And so say all of us \sout{says Jack}.

\crofton:
Thousand a year, Lambert,
\sout{says Crofton or Crawford}.

\lambert:
Right. \sout{says Ned, taking up his John Jameson.}
And butter for fish.

\Nq:
I was just looking around to see who the happy thought would strike
when be damned but in he comes again letting on to be in a hell of a
hurry.

[\stage{Enter Bloom.}]

\Bloom:
I was just round at the courthouse \sout{says he},
looking for you. I hope I'm
not ...

\cunningham:
No \sout{says Martin}, we're ready.

\Nq:
Courthouse my eye and your pockets hanging down with gold and silver.
Mean bloody scut. Stand us a drink itself. Devil a sweet fear! There's
a jew for you! All for number one. Cute as a shithouse rat. Hundred to
five.

\citizen:
Don't tell anyone, \sout{says the citizen}.

\Bloom:
Beg your pardon \sout{says he}.

\cunningham:
Come on boys

\Nq:
says Martin, seeing it was looking blue.

\cunningham:
Come along now.

\citizen:
Don't tell anyone,

\Nq:
says the citizen, letting a bawl out of him.

\citizen:
It's a secret.

\Nq:
And the bloody dog woke up and let a growl.

\cunningham:
Bye bye all \sout{says Martin}.

\Nq:
And he got them out as quick as he could, Jack Power and Crofton or
whatever you call him and him in the middle of them letting on to be all
at sea and up with them on the bloody jaunting car.

\cunningham:
Off with you,

\Nq:
says Martin to the jarvey.

\N3:
The milkwhite dolphin tossed his mane and, rising in the golden poop
the helmsman spread the bellying sail upon the wind and stood off forward
with all sail set, the spinnaker to larboard. A many comely nymphs drew
nigh to starboard and to larboard and, clinging to the sides of the noble
bark, they linked their shining forms as doth the cunning wheelwright when
he fashions about the heart of his wheel the equidistant rays whereof each
one is sister to another and he binds them all with an outer ring and
giveth speed to the feet of men whenas they ride to a hosting or contend
for the smile of ladies fair. Even so did they come and set them, those
willing nymphs, the undying sisters. And they laughed, sporting in a
circle of their foam: and the bark clave the waves.

\Nq:
But begob I was just lowering the heel of the pint when I saw the
citizen getting up to waddle to the door, puffing and blowing with the
dropsy, and he cursing the curse of Cromwell on him, bell, book and candle
in Irish, spitting and spatting out of him and Joe and little Alf round
him like a leprechaun trying to peacify him.

\citizen:
Let me alone \sout{says he}.

\Nq:
And begob he got as far as the door and they holding him and he
bawls out of him:

\citizen:
Three cheers for Israel!

\Nq:
Arrah, sit down on the parliamentary side of your arse for Christ'
sake and don't be making a public exhibition of yourself. Jesus, there's
always some bloody clown or other kicking up a bloody murder about
bloody nothing. Gob, it'd turn the porter sour in your guts, so it would.

\Nq:
And all the ragamuffins and sluts of the nation round the door and Martin
telling the jarvey to drive ahead and the citizen bawling and Alf and
Joe at him to whisht and he on his high horse about the jews and the
loafers calling for a speech and Jack Power trying to get him to sit down
on the car and hold his bloody jaw and a loafer with a patch over his eye
starts singing \Title{If the man in the moon was a jew, jew, jew} and a slut
shouts out of her:

\ragamuffin:
Eh, mister! Your fly is open, mister!

\Nq:
And says he:

\Bloom:
Mendelssohn was a jew and Karl Marx and Mercadante and Spinoza.
And the Saviour was a jew and his father was a jew. Your God.

\cunningham:
He had no father \sout{says Martin}.
That'll do now. Drive ahead.

\citizen:
Whose God? \sout{says the citizen.}

\Bloom:
Well, his uncle was a jew \sout{says he}.
Your God was a jew. Christ was a jew
like me.

\Nq:
Gob, the citizen made a plunge back into the shop.

\citizen:
By Jesus \sout{says he},
I'll brain that bloody jewman for using the holy
name. By Jesus, I'll crucify him so I will.
Give us that biscuitbox here.

\joe:
Stop! Stop! \sout{says Joe.}

\N2:
A large and appreciative gathering of friends and acquaintances from
the metropolis and greater Dublin assembled in their thousands to bid
farewell to Nagyasagos uram Lipoti Virag, late of Messrs Alexander
Thom's, printers to His Majesty, on the occasion of his departure for the
distant clime of Szazharminczbrojugulyas-Dugulas (Meadow of
Murmuring Waters). The ceremony which went off with great \french{éclat} was
characterised by the most affecting cordiality.

\N3:
An illuminated scroll of
ancient Irish vellum, the work of Irish artists, was presented to the
distinguished phenomenologist on behalf of a large section of the
community and was accompanied by the gift of a silver casket, tastefully
executed in the style of ancient Celtic ornament, a work which reflects
every credit on the makers, Messrs Jacob \irish{agus} Jacob. The departing guest
was the recipient of a hearty ovation, many of those who were present
being visibly moved when the select orchestra of Irish pipes struck up the
wellknown strains of \Title{Come Back to Erin},
followed immediately by \Title{Rakoczsy's March}.

\N2:
Tarbarrels and bonfires were lighted along the coastline
of the four seas on the summits of
the Hill of Howth,

\N3:
Three Rock Mountain,

\N2:
Sugarloaf,

\N3:
Bray Head,

\N2:
the mountains of Mourne,

\N3:
the Galtees,

\N2:
the Ox and Donegal and Sperrin peaks,

\N3:
the Nagles and the Bograghs,

\N2:
the Connemara hills,

\N3:
the reeks of M Gillicuddy,

\N2:
Slieve Aughty,

\N3:
Slieve Bernagh

\N2:
and Slieve Bloom.

\N3:
Amid cheers that rent the welkin, responded to by answering cheers from a big
muster of henchmen on the distant Cambrian and Caledonian hills, the
mastodontic pleasureship slowly moved away saluted by a final floral
tribute from the representatives of the fair sex who were present in large
numbers while, as it proceeded down the river, escorted by a flotilla of
barges, the flags of the Ballast office and Custom House were dipped in
salute as were also those of the electrical power station at the
Pigeonhouse and the Poolbeg Light. \hungarian{Visszontlatasra, kedves baraton!
Visszontlatasra!} Gone but not forgotten.

\Nq:
Gob, the devil wouldn't stop him till he got hold of the bloody tin
anyhow and out with him and little Alf hanging on to his elbow and he
shouting like a stuck pig, as good as any bloody play in the Queen's royal
theatre:

\citizen:
Where is he till I murder him?

\Nq:
And Ned and J. J. paralysed with the laughing.

\N1:
Bloody wars \sout{says I},
I'll be in for the last gospel.

\Nq:
But as luck would have it the jarvey got the nag's head round the
other way and off with him.

\joe:
Hold on, citizen \sout{says Joe}. Stop!

\Nq:
Begob he drew his hand and made a swipe and let fly. Mercy of God the sun
was in his eyes or he'd have left him for dead. Gob, he near sent it
into the county Longford. The bloody nag took fright and the old mongrel
after the car like bloody hell and all the populace shouting and laughing
and the old tinbox clattering along the street.

\N3:
The catastrophe was terrific and instantaneous in its effect. The
observatory of Dunsink registered in all eleven shocks, all of the fifth
grade of Mercalli's scale, and there is no record extant of a similar
seismic disturbance in our island since the earthquake of 1534, the
year of the rebellion of Silken Thomas. The epicentre appears to have
been that part of the metropolis which constitutes the Inn's Quay
ward and parish of Saint Michan covering a surface of fortyone acres,
two roods and one square pole or perch.

\N2:
All the lordly residences in
the vicinity of the palace of justice were demolished and that noble
edifice itself, in which at the time of the catastrophe important
legal debates were in progress, is literally a mass of ruins beneath
which it is to be feared all the occupants have been buried alive.
From the reports of eyewitnesses it transpires that the seismic waves
were accompanied by a violent atmospheric perturbation of cyclonic
character.

\N3:
An article of headgear since ascertained to belong to the much
respected clerk of the crown and peace Mr George Fottrell and a silk
umbrella with gold handle with the engraved initials, crest, coat of arms
and house number of the erudite and worshipful chairman of quarter
sessions sir Frederick Falkiner, recorder of Dublin, have been discovered
by search parties in remote parts of the island respectively, the former
on the third basaltic ridge of the giant's causeway, the latter embedded
to the extent of one foot three inches in the sandy beach of Holeopen
bay near the old head of Kinsale.

\N2:
Other eyewitnesses depose that they
observed an incandescent object of enormous proportions hurtling through
the atmosphere at a terrifying velocity in a trajectory directed
southwest by west. Messages of condolence and sympathy are being
hourly received from all parts of the different continents and the
sovereign pontiff has been graciously pleased to decree that a
special \latin{missa pro defunctis} shall be celebrated simultaneously by
the ordinaries of each and every cathedral church of all the episcopal
dioceses subject to the spiritual authority of the Holy See in suffrage of
the souls of those faithful departed who have been so unexpectedly called
away from our midst.

\N3:
The work of salvage, removal of \french{débris},
human remains etc has been entrusted to Messrs Michael Meade and Son, 159 Great
Brunswick street, and Messrs T. and C. Martin, 77, 78, 79 and 80 North
Wall, assisted by the men and officers of the Duke of Cornwall's light
infantry under the general supervision of H. R. H., rear admiral, the
right honourable sir Hercules Hannibal Habeas Corpus Anderson, K. G.,
K. P., K. T., P. C., K. C. B., M. P, J. P., M. B., D. S. O., S. O. D.,
M. F. H., M. R. I. A., B. L., Mus. Doc., P. L. G., F. T. C. D.,
F. R. U. I., F. R. C. P. I. and F. R. C. S. I.

\Nq:
You never saw the like of it in all your born puff. Gob, if he got that
lottery ticket on the side of his poll he'd remember the gold cup,
he would so, but begob the citizen would have been lagged for assault
and battery and Joe for aiding and abetting. The jarvey saved his life
by furious driving as sure as God made Moses. What? O, Jesus, he did.
And he let a volley of oaths after him.

\citizen:
Did I kill him \sout{says he}, or what?

\Nq:
And he shouting to the bloody dog:

\citizen:
After him, Garry! After him, boy!

\Nq:
And the last we saw was the bloody car rounding the corner and old
sheepsface on it gesticulating and the bloody mongrel after it with his
lugs back for all he was bloody well worth to tear him limb from limb.
Hundred to five! Jesus, he took the value of it out of him,
I promise you.

\N2:
When, lo, there came about them all a great brightness and they
beheld the chariot wherein He stood ascend to heaven. And they beheld
Him in the chariot, clothed upon in the glory of the brightness, having
raiment as of the sun, fair as the moon and terrible that for awe they
durst not look upon Him. And there came a voice out of heaven, calling:
\emph{Elijah! Elijah!} And He answered with a main cry:
\emph{Abba! Adonai!} And they
beheld Him even Him, ben Bloom Elijah, amid clouds of angels ascend
to the glory of the brightness at an angle of fortyfive degrees over
Donohoe's in Little Green street like a shot off a shovel.

