\N2:
He turned northeast and crossed the firmer sand
\gab{160}
towards the Pigeonhouse.

\patrice:
\french{Qui vous a mis dans cette fichue position?}

\Stephen:
\french{C'est le pigeon, Joseph.}

\StephenInt:
Patrice, home on furlough, lapped warm milk with me in the bar MacMahon.
Son of the wild goose
[\stage{cheer}],
Kevin Egan of Paris.
My father's a bird,
he lapped the sweet \french{lait chaud} with pink young tongue,
plump bunny's face.
Lap, \french{lapin}.
He hopes to win in the \french{gros lots}.
About the nature of women he read in Michelet.
But he must send me \french{La Vie de Jesus} by M. Leo Taxil.
Lent it to his friend.

\patrice:
\french{C'est tordant, vous savez.
Moi, je suis socialiste.
Je ne crois pas en l'existence de Dieu.
\gab{170}
Faut pas le dire a mon père.}

\Stephen:
\french{Il croit?}

\patrice:
\french{Mon pere, oui.}

\StephenInt:
\german{Schluss.}
He laps.

\StephenInt:
My Latin quarter hat.
God, we simply must dress the character.
I want puce gloves.
You were a student, weren't you?
Of what in the other devil's name?
Paysayenn.
P.C.N., you know:
\french{physiques, chimiques et naturelles.}
Aha.
Eating your groatsworth of \french{mou en civet},
fleshpots of Egypt, elbowed by belching cabmen.
Just say in the most natural tone:
when I was in Paris;
\french{Boul' Mich'}, I used to.
Yes, used to carry punched tickets to prove an alibi
if they arrested you for murder somewhere.
\gab{180}
Justice.
On the night of the seventeenth of February 1904
the prisoner was seen by two witnesses.
Other fellow did it:
other me.
Hat, tie, overcoat, nose.
\french{Lui, c'est moi.}
You seem to have enjoyed yourself.

\StephenInt:
Proudly walking.
Whom were you trying to walk like?
Forget:
a dispossessed.
With mother's money order, eight shillings,
the banging door of the post office slammed in your face by the usher.
Hunger toothache.
\french{Encore deux minutes.}
Look clock.
Must get.
\french{Fermé.}
Hired dog!
Shoot him to bloody bits with a bang shotgun,
bits man spattered walls all brass buttons.
Bits all khrrrrklak in place clack back.
Not hurt?
\gab{190}
O,~that's all right.
Shake hands.
See what I meant, see?
O,~that's all right.
Shake a shake.
O,~that's all only all right.

\StephenInt:
You were going to do wonders, what?
Missionary to Europe after fiery Columbanus.
Fiacre and Scotus on their creepystools in heaven
spilt from their pintpots, loudlatinlaughing:
\latin{Euge! Euge!}
Pretending to speak broken English
as you dragged your valise, porter threepence,
across the slimy pier at Newhaven.
\french{Comment?}
Rich booty you brought back;
\french{Le Tutu,}
five tattered numbers of \french{Pantalon Blanc et Culotte Rouge};
a blue French telegram, curiosity to show:

\simon:
Mother dying come home father.

\buck:
The aunt thinks you killed your mother.
\gab{200}
That's why she won't.

\N1:
\begin{verse}
    Then here's a health to Mulligan's aunt \\
    And I'll tell you the reason why. \\
    She always kept things decent in \\
    The Hannigan famileye.
\end{verse}

\N1:
His feet marched in sudden proud rhythm over the sand furrows,
along by the boulders of the south wall.
He stared at them proudly,
piled stone mammoth skulls.
Gold light on sea, on sand, on boulders.
The sun is there, the slender trees, the lemon houses.

\N2:
Paris rawly waking, crude sunlight on her lemon streets.
Moist pith of farls of bread,
the froggreen wormwood,
\gab{210}
her matin incense,
court the air.
Belluomo rises from the bed of his wife's lover's wife,
the kerchiefed housewife is astir,
a saucer of acetic acid in her hand.
In Rodot's
Yvonne and Madeleine newmake their tumbled beauties,
shattering with gold teeth \french{chaussons} of pastry,
their mouths yellowed with the \french{pus} of \french{flan breton}.
Faces of Paris men go by, their wellpleased pleasers,
curled \spanish{conquistadores}.

\StephenInt:
Noon slumbers.
Kevin Egan rolls gunpowder cigarettes
through fingers smeared with printer's ink,
sipping his green fairy
as Patrice his white.
About us gobblers fork spiced beans down their gullets.

\kevin:
\french{Un demi setier!}

\StephenInt:
A jet of coffee steam from the burnished caldron.
She serves me at his beck.

\kevin:
\french{Il est irlandais.}
\gab{220}

\waitress:
\french{Hollandais?}

\kevin:
\french{Non fromage.
Deux irlandais, nous,
Irlande, vous savez ah, oui!}
She thought you wanted a cheese \french{hollandais}.
Your postprandial, do you know that word?
Postprandial.
There was a fellow I knew once in Barcelona, queer fellow,
used to call it his postprandial.
Well:
\irish{slainte!}

\StephenInt:
Around the slabbed tables |
the tangle of wined breaths and grumbling gorges.
His breath hangs over our saucestained plates,
the green fairy's fang thrusting between his lips.
Of Ireland, the Dalcassians,
of hopes, conspiracies,
of Arthur Griffith now,
Æ,
pimander, good shepherd of men.
To yoke me as his yokefellow,
our crimes our common cause.

\kevin:
You're your father's son.
I know the voice.

\StephenInt:
His fustian shirt, sanguineflowered,
\gab{230}
trembles its Spanish tassels at his secrets.

\kevin:
M. Drumont, famous journalist, Drumont,
know what he called queen Victoria?
Old hag with the yellow teeth.
\french{Vieille ogresse} with the \french{dents jaunes}.
Maud Gonne, beautiful woman,
\french{la Patrie,}
M. Millevoye,
Félix Faure, know how he died?
Licentious men.
The froeken, \french{bonne à tout faire},
who rubs male nakedness in the bath at Upsala.
\french{Moi faire}, she said, \french{tous les messieurs}.
Not this \french{monsieur}, I said.
Most licentious custom.
Bath a most private thing.
I wouldn't let my brother, not even my own brother,
most lascivious thing.

\StephenInt:
Green eyes, I see you.
Fang, I feel.
Lascivious people.

\StephenInt:
The blue fuse burns deadly between hands and burns clear.
Loose tobaccoshreds catch fire:
\gab{240}
a flame and acrid smoke light our corner.
Raw facebones under his peep of day boy's hat.
How the head centre got away,
authentic version.
Got up as a young bride,
man, veil, orangeblossoms,
drove out the road to Malahide.
Did, faith.
Of lost leaders, the betrayed, wild escapes.
Disguises, clutched at, gone, not here.

\kevin:
Spurned lover.
I was a strapping young gossoon at that time, I tell you.
I'll show you my likeness one day.
I was, faith.

\StephenInt:
Lover,
for her love
he prowled with colonel Richard Burke,
tanist of his sept,
under the walls of Clerkenwell
and, crouching,
saw a flame of vengeance hurl them upward in the fog.
Shattered glass and toppling masonry.
In gay Paree he hides,
Egan of Paris,
unsought by any save by me.
\gab{250}
Making his day's stations,
the dingy printingcase,
his three taverns,
the Montmartre lair he sleeps short night in,
\french{rue de la Goutte-d'Or,}
damascened with flyblown faces of the gone.
Loveless, landless, wifeless.
She is quite nicey comfy without her outcast man,
madame in \french{rue Gît-le-Coeur},
canary and two buck lodgers.
Peachy cheeks, a zebra skirt, frisky as a young thing's.
Spurned and undespairing.

\kevin:
Tell Pat you saw me, won't you?
I wanted to get poor Pat a job one time.
\french{Mon fils}, soldier of France.
I taught him to sing
\Title{The Boys of Kilkenny are stout roaring blades}.
Know that old lay?
I taught Patrice that.
Old Kilkenny:
saint Canice, Strongbow's castle on the Nore.
Goes like this.
\song{O, O.}
\gab{260}
He takes me, Napper Tandy, by the hand.

\kevin:
\begin{verse}
    O, O the boys of Kilkenny ...
\end{verse}

\StephenInt:
Weak wasting hand on mine.
They have forgotten Kevin Egan,
not he them.
Remembering thee, O Sion.

\N1:
He had come nearer the edge of the sea and wet sand slapped his boots.
The new air greeted him, harping in wild nerves,
wind of wild air
of seeds of brightness.

\StephenInt:
Here,
I am not walking out to the Kish lightship,
am I?

\N1:
He stood suddenly,
his feet beginning to sink slowly in the quaking soil.

\StephenInt:
Turn back.

\N1:
Turning, he scanned the shore south,
\gab{270}
his feet sinking again slowly in new sockets.

\StephenInt:
The cold domed room of the tower waits.
Through the barbacans |
the shafts of light are moving ever,
slowly ever as my feet are sinking,
creeping duskward over the dial floor.
Blue dusk, nightfall, deep blue night.
In the darkness of the dome
they wait, their pushedback chairs,
my obelisk valise,
around a board of abandoned platters.
Who to clear it?
He has the key.
I will not sleep there when this night comes.
A shut door of a silent tower,
entombing their--blind bodies,
the panthersahib and his pointer.
Call:
no answer.

\N1:
He lifted his feet up from the suck
and turned back by the mole of boulders.

\StephenInt:
Take all, keep all.
My soul walks with me, form of forms.
So in the moon's midwatches
\gab{280}
I pace the path above the rocks,
in sable silvered, hearing Elsinore's tempting flood.

\StephenInt:
The flood is following me.
I can watch it flow past from here.
Get back then by the Poolbeg road to the strand there.

\N1:
He climbed over the sedge and eely oarweeds
and sat on a stool of rock, resting his ashplant in a grike.

\N1:
A bloated carcass of a dog lay lolled on bladderwrack.
Before him the gunwale of a boat, sunk in sand.

\StephenInt:
\french{Un coche ensablé} Louis Veuillot called Gautier's prose.
These heavy sands are language |
tide and wind have silted here.
And these, the stoneheaps of dead builders, a warren of weasel rats.
Hide gold there.
Try it.
You have some.
Sands and stones.
\gab{290}
Heavy of the past.
Sir Lout's toys.
Mind you don't get one bang on the ear.
I'm the bloody well gigant
rolls all them bloody well boulders,
bones for my steppingstones.
Feefawfum.
I zmellz de bloodz odz an Iridzman.
