%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\input{defs}
% \input{largeprint} \renewcommand{\gab}[1]{\marginpar[\tiny\raggedleft #1]{\tiny #1}}

\title{\Huge Proteus}
\author{George~V. Reilly\\
\\
{\small for the}\\
Wild Geese Players of Seattle\\
{\emph{www.WildGeeseSeattle.org}}\\
\\
{\small Chapter~3 of \emph{Ulysses} by James Joyce}\\
{\small Adapted from the 1922~edition at Project Gutenberg}
\\
\\
{\small First Draft}}
\date{March \nth{27}, 2016}
\raggedbottom

\begin{document}

% Page i, Title Page
\maketitle
\thispagestyle{empty}
\pagebreak

% Page ii, Cast
\begin{tabular}{lp{10cm}}
    \multicolumn{2}{c}{\Large \textsc{The Cast}} \\
\\
    \multicolumn{2}{c}{\large \textit{TBD}} \\
Stephen Dedalus \\
N1 \\
N2 \\
Simon Dedalus \\
Richie Goulding \\
Walter Goulding \\
Patrice Egan \\
Kevin Egan \\
Mrs Egan \\
\end{tabular}

\thispagestyle{empty}
\newpage


% Page 1, first page of script proper
\setcounter{page}{1}


\StephenInt:
Ineluctable modality of the visible:
at least that if no more, thought through my eyes.
Signatures of all things I am here to read,
seaspawn and seawrack,
the nearing tide,
that rusty boot.
Snotgreen, bluesilver, rust:
coloured signs.
Limits of the diaphane.

\N1:
But he adds:

\StephenInt:
in bodies.
Then he was aware of them bodies before of them coloured.
How?
By knocking his sconce against them, sure.
Go easy.
Bald he was and a millionaire,
\italian{maestro di color che sanno}.
Limit of the diaphane in.
Why in?
Diaphane, adiaphane.
If you can put your five fingers through it it is a gate,
if not a door.
Shut your eyes and see.

\N1:
Stephen closed his eyes to hear his boots
\gab{10}
crush crackling wrack and shells.

\StephenInt:
You are walking through it howsomever.
I am, a stride at a time.
A very short space of time through very short times of space.
Five, six: the \german{Nacheinander}.
Exactly: and that is the ineluctable modality of the audible.
Open your eyes.
No.
Jesus!
If I fell over a cliff that beetles o'er his base,
fell through the \german{Nebeneinander} ineluctably!
I am getting on nicely in the dark.
My ash sword hangs at my side.
Tap with it: they do.
My two feet in his boots are at the ends of his legs, \german{nebeneinander}.
Sounds solid:
made by the mallet of \spanish{los demiurgos}.
Am I walking into eternity along Sandymount strand?
Crush, crack, crick, crick.
Wild sea money.
Dominie Deasy kens
\gab{20}
them a'.

\N2:
\begin{verse}
    Won't you come to Sandymount, \\
    Madeline the mare?
\end{verse}

\StephenInt:
Rhythm begins, you see.
I hear.
Acatalectic tetrameter of iambs marching.
No, agallop:
\emph{Deline the mare}.

\StephenInt:
Open your eyes now.
I will.
One moment.
Has all vanished since?
If I open and am for ever in the black adiaphane.
\italian{Basta!}
I will see if I can see.

\StephenInt:
See now.
There all the time without you:
and ever shall be,
world without end.

\N2:
They came down the steps from Leahy's terrace prudently,
\german{Frauenzimmer}:
and down the shelving shore flabbily,
\gab{30}
their splayed feet sinking in the silted sand.

\StephenInt:
Like me, like Algy,
coming down to our mighty mother.

\N2:
Number one swung lourdily her midwife's bag,
the other's gamp poked in the beach.

\N1:
From the liberties, out for the day.
Mrs Florence MacCabe,
relict of the late Patk MacCabe, deeply lamented,
of Bride Street.

\StephenInt:
One of her sisterhood lugged me squealing into life.
Creation from nothing.
What has she in the bag?
A misbirth with a trailing navelcord,
hushed in ruddy wool.
The cords of all link back,
strandentwining cable of all flesh.
That is why mystic monks.
Will you be as gods?
Gaze in your
\greek{Omphalos}.
Hello!
Kinch here.
Put me on to Edenville.
Aleph, alpha:
nought, nought, one.
\gab{40}

\StephenInt:
Spouse and helpmate of Adam Kadmon:
Heva, naked Eve.
She had no navel.
Gaze.
Belly without blemish, bulging big,
a buckler of taut vellum,
no, whiteheaped corn,
orient and immortal,
standing from everlasting to everlasting.
Womb of sin.

\StephenInt:
Wombed in sin darkness I was too, made not begotten.
By them, the man with my voice and my eyes
and a ghostwoman with ashes on her breath.
They clasped and sundered, did the coupler's will.
From before the ages
He willed me
and now may not will me away or ever.
\latin{A lex eterna} stays about Him.
Is that then the divine substance wherein Father and Son are consubstantial?
\gab{50}
Where is poor dear Arius to try conclusions?
Warring his life long upon the contransmagnificandjewbangtantiality.
Illstarred heresiarch'
In a Greek watercloset he breathed his last:
euthanasia.
With beaded mitre and with crozier,
stalled upon his throne,
widower of a widowed see,
with upstiffed omophorion,
with clotted hinderparts.

\N1:
Airs romped round him, nipping and eager airs.

\StephenInt:
They are coming,
waves.
The whitemaned seahorses, champing,
brightwindbridled,
the steeds of Mananaan.

\StephenInt:
I mustn't forget his letter for the press.
And after?
The Ship, half twelve.
By the way
go easy with that money like a good young imbecile.

\StephenInt:
Yes, I must.
\gab{60}

\N1:
His pace slackened.

\StephenInt:
Here.
Am I going to aunt Sara's or not?
My consubstantial father's voice.

\simon:
Did you see anything of your artist brother Stephen lately?
No?
Sure he's not down in Strasburg terrace with his aunt Sally?
Couldn't he fly a bit higher than that, eh?
And and and and tell us, Stephen,
how is uncle Si?
O, weeping God, the things I married into!
De boys up in de hayloft.
The drunken little costdrawer
and his brother, the cornet player.
Highly respectable gondoliers!
And skeweyed Walter sirring his father, no less!
Sir.
Yes, sir.
No, sir.
Jesus wept:
and no wonder, by Christ!

\StephenInt:
I pull the wheezy bell of their shuttered cottage:
\gab{70}
and wait.
They take me for a dun,
peer out from a coign of vantage.

\walter:
It's Stephen, sir.

\richie:
Let him in.
Let Stephen in.

\StephenInt:
A bolt drawn back and Walter welcomes me.

\walter:
We thought you were someone else.

\N2:
In his broad bed
nuncle Richie, pillowed and blanketed,
extends over the hillock of his knees a sturdy forearm.

\StephenInt:
Cleanchested.
He has washed the upper moiety.

\richie:
Morrow, nephew.

\N2:
He lays aside the lapboard whereon he drafts his bills of costs
\gab{80}
for the eyes of master Goff and master Shapland Tandy,
filing consents and common searches and a writ of \latin{duces tecum}.
A bogoak frame over his bald head:
Wilde's \latin{requiescat}.
The drone of his misleading whistle brings Walter back.

\walter:
Yes, sir?

\richie:
Malt for Richie and Stephen, tell mother.
Where is she?

\walter:
Bathing Crissie, sir.

\StephenInt:
Papa's little bedpal.
Lump of love.

\Stephen:
No, uncle Richie ...

\richie:
Call me Richie.
\gab{90}
Damn your lithia water.
It lowers.
Whusky!

\Stephen:
Uncle Richie, really ...

\richie:
Sit down or by the law Harry
I'll knock you down.

\N1:
Walter squints vainly for a chair.

\walter:
He has nothing to sit down on, sir.

\richie:
He has nowhere to put it, you mug.
Bring in our chippendale chair.
Would you like a bite of something?
None of your damned lawdeedaw airs here.
The rich of a rasher fried with a herring?
Sure?
So much the better.
We have nothing in the house but backache pills.

\richie:
\italian{All'erta!}

\N1:
He drones bars of Ferrando's \italian{aria di sortita}.
\gab{100}

\richie:
The grandest number, Stephen,
in the whole opera.
Listen.

\N1:
His tuneful whistle sounds again, finely shaded,
with rushes of the air,
his fists bigdrumming on his padded knees.

\StephenInt:
This wind is sweeter.

\StephenInt:
Houses of decay, mine, his and all.
You told the Clongowes gentry
you had an uncle a judge
and an uncle a general in the army.
Come out of them, Stephen.
Beauty is not there.
Nor in the stagnant bay of Marsh's library
where you read the fading prophecies of Joachim Abbas.
For whom?
The hundredheaded rabble of the cathedral close.
A hater of his kind ran from them to the wood of madness,
\gab{110}
his mane foaming in the moon, his eyeballs stars.
Houyhnhnm, horsenostrilled.
The oval equine faces,
Temple, Buck Mulligan, Foxy Campbell, Lanternjaws.
Abbas father,--
furious dean, what offence laid fire to their brains?
Paff!
\latin{Descende, calve, ut ne amplius decalveris.}
A garland of grey hair on his comminated head
see him me clambering down to the footpace (\latin{descende!}),
clutching a monstrance, basiliskeyed.
Get down, baldpoll!
A choir gives back menace and echo,
assisting about the altar's horns,
the snorted Latin of jackpriests
moving burly in their albs,
tonsured and oiled and gelded,
fat with the fat of kidneys of wheat.

\StephenInt:
And at the same instant
\gab{120}
perhaps a priest round the corner is elevating it.
Dringdring!
And two streets off
another locking it into a pyx.
Dringadring!
And in a ladychapel
another taking housel all to his own cheek.
Dringdring!
Down, up, forward, back.
Dan Occam thought of that,
invincible doctor.
A misty English morning
the imp hypostasis tickled his brain.
Bringing his host down and kneeling
he heard twine with his second bell
the first bell in the transept
(he is lifting his)
and, rising, heard
(now I am lifting)
their two bells
(he is kneeling)
twang in diphthong.

\StephenInt:
Cousin Stephen, you will never be a saint.
Isle of saints.
You were awfully holy, weren't you?
You prayed to the Blessed Virgin
that you might not have a red nose.
\gab{130}
You prayed to the devil in Serpentine avenue
that the fubsy widow in front might lift her clothes
still more from the wet street.
\italian{O si, certo!}
Sell your soul for that, do, dyed rags pinned round a squaw.
More tell me, more still!!
On the top of the Howth tram alone crying to the rain:
Naked women!
\emph{Naked women!}
What about that, eh?

\StephenInt:
What about what?
What else were they invented for?

\StephenInt:
Reading two pages apiece of seven books every night, eh?
I was young.
You bowed to yourself in the mirror,
stepping forward to applause earnestly, striking face.
Hurray for the Goddamned idiot!
Hray!
No-one saw:
tell no-one.
Books you were going to write with letters for titles.
Have you read his \Title{F}?
O yes, but I prefer \Title{Q}.
\gab{140}
Yes, but \Title{W} is wonderful.
O yes, \Title{W}.
Remember your epiphanies written on green oval leaves,
deeply deep,
copies to be sent if you died to all the great libraries of the world,
including Alexandria?
Someone was to read them there after a few thousand years,
a mahamanvantara.
Pico della Mirandola like.
Ay, very like a whale.
When one reads these strange pages of one long gone
one feels that one is at one with one who once ...

\N2:
The grainy sand had gone from under his feet.
His boots trod again a damp crackling mast,
razorshells, squeaking pebbles,
that on the unnumbered pebbles beats,
wood sieved by the shipworm,
lost Armada.
Unwholesome sandflats waited to suck his treading soles,
\gab{150}
breathing upward sewage breath,
a pocket of seaweed smouldered in seafire under a midden of man's ashes.
He coasted them, walking warily.
A porterbottle stood up,
stogged to its waist, in the cakey sand dough.
A sentinel:
isle of dreadful thirst.
Broken hoops on the shore;
at the land a maze of dark cunning nets;
farther away chalkscrawled backdoors
and on the higher beach
a dryingline with two crucified shirts.
Ringsend:
wigwams of brown steersmen and master mariners.
Human shells.

\N2:
He halted.

\StephenInt:
I have passed the way to aunt Sara's.
Am I not going there?
Seems not.
No-one about.

\N2:
He turned northeast and crossed the firmer sand
\gab{160}
towards the Pigeonhouse.

\patrice:
\french{Qui vous a mis dans cette fichue position?}

\Stephen:
\french{C'est le pigeon, Joseph.}

\StephenInt:
Patrice, home on furlough, lapped warm milk with me in the bar MacMahon.
Son of the wild goose, Kevin Egan of Paris.
My father's a bird,
he lapped the sweet \french{lait chaud} with pink young tongue,
plump bunny's face.
Lap, \french{lapin}.
He hopes to win in the \french{gros lots}.
About the nature of women he read in Michelet.
But he must send me \french{La Vie de Jesus} by M. Leo Taxil.
Lent it to his friend.

\patrice:
\french{C'est tordant, vous savez.
Moi, je suis socialiste.
Je ne crois pas en l'existence de Dieu.
\gab{170}
Faut pas le dire a mon père.}

\Stephen:
\french{Il croit?}

\patrice:
\french{Mon pere, oui.}

\StephenInt:
\german{Schluss.}
He laps.

\StephenInt:
My Latin quarter hat.
God, we simply must dress the character.
I want puce gloves.
You were a student, weren't you?
Of what in the other devil's name?
Paysayenn.
P.C.N., you know:
\french{Physiques, Chimiques et Naturelles.}
Aha.
Eating your groatsworth of \french{mou en civet},
fleshpots of Egypt, elbowed by belching cabmen.
Just say in the most natural tone:
when I was in Paris;
\french{Boul' Mich'}, I used to.
Yes, used to carry punched tickets to prove an alibi
if they arrested you for murder somewhere.
\gab{180}
Justice.
On the night of the seventeenth of February 1904
the prisoner was seen by two witnesses.
Other fellow did it:
other me.
Hat, tie, overcoat, nose.
\french{Lui, c'est moi.}
You seem to have enjoyed yourself.

\StephenInt:
Proudly walking.
Whom were you trying to walk like?
Forget:
a dispossessed.
With mother's money order, eight shillings,
the banging door of the post office slammed in your face by the usher.
Hunger toothache.
\french{Encore deux minutes.}
Look clock.
Must get.
\french{Ferme.}
Hired dog!
Shoot him to bloody bits with a bang shotgun,
bits man spattered walls all brass buttons.
Bits all khrrrrklak in place clack back.
Not hurt?
\gab{190}
O, that's all right.
Shake hands.
See what I meant, see?
O, that's all right.
Shake a shake.
O, that's all only all right.

\StephenInt:
You were going to do wonders, what?
Missionary to Europe after fiery Columbanus.
Fiacre and Scotus on their creepystools in heaven
spilt from their pintpots, loudlatinlaughing:
\latin{Euge! Euge!}
Pretending to speak broken English
as you dragged your valise, porter threepence,
across the slimy pier at Newhaven.
\french{Comment?}
Rich booty you brought back;
\french{Le Tutu,}
five tattered numbers of \french{Pantalon Blanc et Culotte Rouge};
a blue French telegram, curiosity to show:

\simon:
Mother dying come home father.

\buck:
The aunt thinks you killed your mother.
\gab{200}
That's why she won't.

\N1:
\begin{verse}
    Then here's a health to Mulligan's aunt \\
    And I'll tell you the reason why. \\
    She always kept things decent in \\
    The Hannigan famileye.
\end{verse}

\N1:
His feet marched in sudden proud rhythm over the sand furrows,
along by the boulders of the south wall.
He stared at them proudly,
piled stone mammoth skulls.
Gold light on sea, on sand, on boulders.
The sun is there, the slender trees, the lemon houses.

\N2:
Paris rawly waking, crude sunlight on her lemon streets.
Moist pith of farls of bread,
the froggreen wormwood,
\gab{210}
her matin incense,
court the air.
Belluomo rises from the bed of his wife's lover's wife,
the kerchiefed housewife is astir,
a saucer of acetic acid in her hand.
In Rodot's
Yvonne and Madeleine newmake their tumbled beauties,
shattering with gold teeth \french{chaussons} of pastry,
their mouths yellowed with the \french{pus} of \french{flan breton}.
Faces of Paris men go by, their wellpleased pleasers,
curled conquistadores.

\StephenInt:
Noon slumbers.
Kevin Egan rolls gunpowder cigarettes
through fingers smeared with printer's ink,
sipping his green fairy
as Patrice his white.
About us gobblers fork spiced beans down their gullets.

\kevin:
\french{Un demi setier!}

\StephenInt:
A jet of coffee steam from the burnished caldron.
She serves me at his beck.

\kevin:
\french{Il est irlandais.
\gab{220}
Hollandais?
Non fromage.
Deux irlandais, nous,
Irlande, vous savez ah, oui!}
She thought you wanted a cheese \french{hollandais}.
Your postprandial, do you know that word?
Postprandial.
There was a fellow I knew once in Barcelona, queer fellow,
used to call it his postprandial.
Well:
\irish{Slainte!}

\StephenInt:
Around the slabbed tables
the tangle of wined breaths and grumbling gorges.
His breath hangs over our saucestained plates,
the green fairy's fang thrusting between his lips.
Of Ireland, the Dalcassians,
of hopes, conspiracies,
of Arthur Griffith now,
A~E,
pimander, good shepherd of men.
To yoke me as his yokefellow,
our crimes our common cause.

\kevin:
You're your father's son.
I know the voice.

\StephenInt:
His fustian shirt, sanguineflowered,
\gab{230}
trembles its Spanish tassels at his secrets.

\kevin:
M. Drumont, famous journalist, Drumont,
know what he called queen Victoria?
Old hag with the yellow teeth.
\french{Vieille ogresse} with the \french{dents jaunes}.
Maud Gonne, beautiful woman,
\french{la patrie,}
M. Millevoye,
Felix Faure, know how he died?
Licentious men.
The froeken, \french{bonne a tout faire},
who rubs male nakedness in the bath at Upsala.
\french{Moi faire}, she said, \french{tous les messieurs}.
Not this \french{monsieur}, I said.
Most licentious custom.
Bath a most private thing.
I wouldn't let my brother, not even my own brother,
most lascivious thing.
Green eyes, I see you.
Fang, I feel.
Lascivious people.

\StephenInt:
The blue fuse burns deadly between hands and burns clear.
Loose tobaccoshreds catch fire:
\gab{240}
a flame and acrid smoke light our corner.
Raw facebones under his peep of day boy's hat.
How the head centre got away,
authentic version.
Got up as a young bride,
man, veil, orangeblossoms,
drove out the road to Malahide.
Did, faith.
Of lost leaders, the betrayed, wild escapes.
Disguises, clutched at, gone, not here.

\kevin:
Spurned lover.
I was a strapping young gossoon at that time, I tell you.
I'll show you my likeness one day.
I was, faith.

\StephenInt:
Lover,
for her love
he prowled with colonel Richard Burke,
tanist of his sept,
under the walls of Clerkenwell
and, crouching,
saw a flame of vengeance hurl them upward in the fog.
Shattered glass and toppling masonry.
In gay Paree he hides,
Egan of Paris,
unsought by any save by me.
\gab{250}
Making his day's stations,
the dingy printingcase,
his three taverns,
the Montmartre lair he sleeps short night in,
rue de la Goutte-d'Or,
damascened with flyblown faces of the gone.
Loveless, landless, wifeless.
She is quite nicey comfy without her outcast man,
madame in rue Git-le-Coeur,
canary and two buck lodgers.
Peachy cheeks, a zebra skirt, frisky as a young thing's.
Spurned and undespairing.

\mrsegan:
Tell Pat you saw me, won't you?
I wanted to get poor Pat a job one time.
\french{Mon fils}, soldier of France.
I taught him to sing
\Title{The Boys of Kilkenny are Stout Roaring Blades}.
Know that old lay?
I taught Patrice that.
Old Kilkenny:
saint Canice, Strongbow's castle on the Nore.
Goes like this.
\song{O, O.}
\gab{260}
He takes me, Napper Tandy, by the hand.

\mrsegan:
\begin{verse}
    O, O the boys of Kilkenny ...
\end{verse}

\StephenInt:
Weak wasting hand on mine.
They have forgotten Kevin Egan,
not he them.
Remembering thee, O Sion.

\N1:
He had come nearer the edge of the sea and wet sand slapped his boots.
The new air greeted him, harping in wild nerves,
wind of wild air
of seeds of brightness.

\StephenInt:
Here,
I am not walking out to the Kish lightship,
am I?

\N1:
He stood suddenly,
his feet beginning to sink slowly in the quaking soil.
Turn back.

\N1:
Turning, he scanned the shore south,
\gab{270}
his feet sinking again slowly in new sockets.

\StephenInt:
The cold domed room of the tower waits.
Through the barbacans
the shafts of light are moving ever,
slowly ever as my feet are sinking,
creeping duskward over the dial floor.
Blue dusk, nightfall, deep blue night.
In the darkness of the dome
they wait, their pushedback chairs,
my obelisk valise,
around a board of abandoned platters.
Who to clear it?
He has the key.
I will not sleep there when this night comes.
A shut door of a silent tower,
entombing their--blind bodies,
the panthersahib and his pointer.
Call:
no answer.

\N1:
He lifted his feet up from the suck
and turned back by the mole of boulders.

\StephenInt:
Take all, keep all.
My soul walks with me, form of forms.
So in the moon's midwatches
\gab{280}
I pace the path above the rocks,
in sable silvered, hearing Elsinore's tempting flood.

\StephenInt:
The flood is following me.
I can watch it flow past from here.
Get back then by the Poolbeg road to the strand there.

\N1:
He climbed over the sedge and eely oarweeds
and sat on a stool of rock, resting his ashplant in a grike.

\N1:
A bloated carcass of a dog lay lolled on bladderwrack.
Before him the gunwale of a boat, sunk in sand.

\StephenInt:
\french{Un coche ensable} Louis Veuillot called Gautier's prose.
These heavy sands are language tide and wind have silted here.
And these, the stoneheaps of dead builders, a warren of weasel rats.
Hide gold there.
Try it.
You have some.
Sands and stones.
\gab{290}
Heavy of the past.
Sir Lout's toys.
Mind you don't get one bang on the ear.
I'm the bloody well gigant
rolls all them bloody well boulders,
bones for my steppingstones.
Feefawfum.
I zmellz de bloodz odz an Iridzman.

\N1:
A point, live dog,
grew into sight running across the sweep of sand.

\StephenInt:
Lord, is he going to attack me?
Respect his liberty.
You will not be master of others or their slave.
I have my stick.
Sit tight.

\N2:
From farther away,
walking shoreward across from the crested tide,
figures, two.

\StephenInt:
The two maries.
They have tucked it safe mong the bulrushes.
Peekaboo.
I see you.
No, the dog.
He is running back to them.
Who?

\StephenInt:
Galleys of the Lochlanns ran here to beach,
\gab{300}
in quest of prey,
their bloodbeaked prows riding low on a molten pewter surf.
Dane vikings,
torcs of tomahawks aglitter on their breasts
when Malachi wore the collar of gold.
A school of turlehide whales stranded in hot noon, spouting,
hobbling in the shallows.
Then from the starving cagework city
a horde of jerkined dwarfs,
my people, with flayers' knives,
running, scaling, hacking in green blubbery whalemeat.
Famine, plague and slaughters.
Their blood is in me, their lusts my waves.
I moved among them on the frozen Liffey,
that I, a changeling, among the spluttering resin fires.
I spoke to no-one:
none to me.

\N1:
The dog's bark ran towards him, stopped, ran back.
\gab{310}

\StephenInt:
Dog of my enemy.
I just simply stood pale, silent, bayed about.
\latin{Terribilia meditans.}
A primrose doublet, fortune's knave, smiled on my fear.
For that are you pining,
the bark of their applause?
Pretenders:
live their lives.
The Bruce's brother,
Thomas Fitzgerald, silken knight,
Perkin Warbeck, York's false scion,
in breeches of silk of whiterose ivory,
wonder of a day,
and Lambert Simnel, with a tail of nans and sutlers, a scullion crowned.
All kings' sons.
Paradise of pretenders then and now.
He saved men from drowning
and you shake at a cur's yelping.
But the courtiers who mocked Guido in \italian{Or san Michele}
were in their own house.
House of ...
We don't want any of your medieval abstrusiosities.
\gab{320}
Would you do what he did?
A boat would be near, a lifebuoy.
\german{Naturlich}, put there for you.
Would you or would you not?
The man that was drowned nine days ago off Maiden's rock.
They are waiting for him now.
The truth, spit it out.
I would want to.
I would try.
I am not a strong swimmer.
Water cold soft.
When I put my face into it in the basin at Clongowes.
Can't see!
Who's behind me?
Out quickly, quickly!
Do you see the tide flowing quickly in on all sides,
sheeting the lows of sand quickly, shellcocoacoloured?
If I had land under my feet.
I want his life still to be his, mine to be mine.
A drowning man.
His human eyes scream to me out of horror of his death.
I ...
With him together down ...
I could not save her.
\gab{330}
Waters:
bitter death:
lost.

\StephenInt:
A woman and a man.
I see her skirties.
Pinned up, I bet.

\N1:
Their dog ambled about a bank of dwindling sand,
trotting, sniffing on all sides.
Looking for something lost in a past life.
Suddenly he made off like a bounding hare,
ears flung back, chasing the shadow of a lowskimming gull.
The man's shrieked whistle struck his limp ears.
He turned, bounded back, came nearer, trotted on twinkling shanks.
On a field tenney
a buck, trippant, proper, unattired.
At the lacefringe of the tide
he halted with stiff forehoofs, seawardpointed ears.
His snout lifted
barked at the wavenoise, herds of seamorse.
They serpented towards his feet,
curling, unfurling many crests,
\gab{340}
every ninth,
breaking, plashing,
from far,
from farther out,
waves and waves.

\StephenInt:
Cocklepickers.

\N1:
They waded a little way in the water and, stooping,
soused their bags and, lifting them again, waded out.
The dog yelped
running to them,
reared up and pawed them,
dropping on all fours,
again reared up at them with mute bearish fawning.
Unheeded
he kept by them
as they came towards the drier sand,
a rag of wolf's tongue redpanting from his jaws.
His speckled body ambled ahead of them
and then loped off at a calf's gallop.
The carcass lay on his path.
He stopped, sniffed, stalked round it,
brother, nosing closer, went round it,
sniffling rapidly like a dog
all over the dead dog's bedraggled fell.

\StephenInt:
Dogskull, dogsniff,
\gab{350}
eyes on the ground,
moves to one great goal.
Ah, poor dogsbody!
Here lies poor dogsbody's body.

\man:
Tatters!
Out of that, you mongrel!

\N1:
The cry brought him skulking back to his master
and a blunt bootless kick sent him unscathed
across a spit of sand, crouched in flight.
He slunk back in a curve.

\StephenInt:
Doesn't see me.

\N1:
Along by the edge of the mole
he lolloped, dawdled, smelt a rock.
and from under a cocked hindleg
pissed against it.
He trotted forward and, lifting again his hindleg,
pissed quick short at an unsmelt rock.
The simple pleasures of the poor.
His hindpaws then scattered the sand:
then his forepaws dabbled and delved.
\gab{360}
Something he buried there, his grandmother.
He rooted in the sand,
dabbling, delving
and stopped to listen to the air,
scraped up the sand
again with a fury of his claws,
soon ceasing, a pard, a panther,
got in spousebreach,
vulturing the dead.

\StephenInt:
After he woke me last night same dream or was it?
Wait.
Open hallway.
Street of harlots.
Remember.
Haroun al Raschid.
I am almosting it.
That man led me, spoke.
I was not afraid.
The melon he had he held against my face.
Smiled:
creamfruit smell.
That was the rule, said.
In.
Come.
Red carpet spread.
You will see who.

\N2:
Shouldering their bags they trudged, the red Egyptians.
\gab{370}
His blued feet out of turnedup trousers
slapped the clammy sand,
a dull brick muffler strangling his unshaven neck.
With woman steps she followed:
the ruffian and his strolling mort.
Spoils slung at her back.
Loose sand and shellgrit crusted her bare feet.
About her windraw face hair trailed.
Behind her lord, his helpmate, bing awast to Romeville.
When night hides her body's flaws
calling under her brown shawl from an archway where dogs have mired.
Her fancyman is treating two Royal Dublins in O'Loughlin's of Blackpitts.
Buss her, wap in rogues' rum lingo,
for, O, my dimber wapping dell!
A shefiend's whiteness under her rancid rags.
Fumbally's lane that night:
the tanyard smells.
\gab{380}

\N2:
\begin{verse}
    White thy fambles, red thy gan \\
    And thy quarrons dainty is. \\
    Couch a hogshead with me then. \\
    In the darkmans clip and kiss.
\end{verse}

\StephenInt:
Morose delectation Aquinas tunbelly calls this, \latin{frate porcospino}.
Unfallen Adam rode and not rutted.
Call away let him:
\emph{Thy quarrons dainty is}.
Language no whit worse than his.
Monkwords, marybeads jabber on their girdles:
roguewords, tough nuggets patter in their pockets.

\StephenInt:
Passing now.

\StephenInt:
A side eye at my Hamlet hat.
\gab{390}
If I were suddenly naked here as I sit?
I am not.
Across the sands of all the world,
followed by the sun's flaming sword,
to the west, trekking to evening lands.
She trudges, schlepps,
trains, drags, trascines her load.
A tide westering, moondrawn, in her wake.
Tides, myriadislanded, within her,
blood not mine,
\greek{oinopa ponton},
a winedark sea.
Behold the handmaid of the moon.
In sleep the wet sign calls her hour, bids her rise.
Bridebed, childbed,
bed of death,
ghostcandled.
\latin{Omnis caro ad te veniet.}
He comes, pale vampire, through storm
his eyes, his bat sails bloodying the sea,
mouth to her mouth's kiss.

\StephenInt:
Here.
Put a pin in that chap, will you?
My tablets.
Mouth to her kiss.

\StephenInt:
No.
Must be two of em.
\gab{400}
Glue em well.
Mouth to her mouth's kiss.

\N2:
His lips lipped and mouthed fleshless lips of air:
mouth to her moomb.

\StephenInt:
Oomb, allwombing tomb.

\N2:
His mouth moulded issuing breath, unspeeched:

\StephenInt:
ooeeehah:
roar of cataractic planets, globed, blazing,
roaring wayawayawayawayaway.
Paper.
The banknotes, blast them.
Old Deasy's letter.
Here.
Thanking you for the hospitality tear the blank end off.

\N1:
Turning his back to the sun
he bent over far to a table of rock and scribbled words.

\StephenInt:
That's twice I forgot to take slips from the library counter.

\N1:
His shadow lay over the rocks as he bent, ending.

\StephenInt:
Why not endless till the farthest star?
Darkly they are there behind this light,
darkness shining in the brightness,
delta of Cassiopeia, worlds.
\gab{410}
Me sits there with his augur's rod of ash,
in borrowed sandals,
by day beside a livid sea,
unbeheld, in violet night
walking beneath a reign of uncouth stars.
I throw this ended shadow from me, manshape ineluctable, call it back.
Endless, would it be mine, form of my form?
Who watches me here?
Who ever anywhere will read these written words?
Signs on a white field.
Somewhere to someone in your flutiest voice.
The good bishop of Cloyne took the veil of the temple
out of his shovel hat:
veil of space with coloured emblems
hatched on its field.
Hold hard.
Coloured on a flat:
yes, that's right.
Flat I see, then think distance,
near, far, flat I see, east, back.
Ah, see now!
Falls back suddenly, frozen in stereoscope.
\gab{420}
Click does the trick.
You find my words dark.
Darkness is in our souls do you not think?
Flutier.
Our souls, shamewounded by our sins, cling to us yet more,
a woman to her lover clinging, the more the more.

\StephenInt:
She trusts me, her hand gentle, the longlashed eyes.
Now where the blue hell am I bringing her beyond the veil?
Into the ineluctable modality
of the ineluctable visuality.
She, she, she.
What she?
The virgin at Hodges Figgis' window on Monday looking in
for one of the alphabet books you were going to write.
Keen glance you gave her.
Wrist through the braided jesse of her sunshade.
She lives in Leeson park with a grief and kickshaws, a lady of letters.
\gab{430}
Talk that to someone else,
Stevie:
a pickmeup.
Bet she wears those curse of God stays
suspenders and yellow stockings, darned with lumpy wool.
Talk about apple dumplings,
\italian{Piuttosto.}
Where are your wits?

\StephenInt:
Touch me.
Soft eyes.
Soft soft soft hand.
I am lonely here.
O, touch me soon, now.
What is that word known to all men?
I am quiet here alone.
Sad too.
Touch, touch me.

\N1:
He lay back at full stretch over the sharp rocks,
cramming the scribbled note and pencil into a pock his hat.
His hat down on his eyes.

\StephenInt:
That is Kevin Egan's movement I made,
nodding for his nap, sabbath sleep.

\kevin:
\latin{Et vidit Deus.
Et erant valde bona.}
\gab{440}
\french{Alo!
Bonjour.}

\StephenInt:
Welcome as the flowers in May.
Under its leaf he watched through peacocktwittering lashes
the southing sun.
I am caught in this burning scene.
Pan's hour, the faunal noon.
Among gumheavy serpentplants, milkoozing fruits,
where on the tawny waters leaves lie wide.
Pain is far.

\StephenInt:
\emph{And no more turn aside and brood.}

\N1:
His gaze brooded on his broadtoed boots, a buck's castoffs, \german{Nebeneinander}.
He counted the creases of rucked leather wherein another's foot had nested warm.

\StephenInt:
The foot that beat the ground in tripudium,
foot I dislove.
But you were delighted when Esther Osvalt's shoe went on you:
girl I knew in Paris.
\french{Tiens, quel petit pied!}
\gab{450}
Staunch friend, a brother soul:
Wilde's love that dare not speak its name.
His arm:
Cranly's arm.
He now will leave me.
And the blame?
As I am.
As I am.
All or not at all.

\N2:
In long lassoes from the Cock lake
the water flowed full,
covering greengoldenly lagoons of sand,
rising, flowing.

\StephenInt:
My ashplant will float away.
I shall wait.
No, they will pass on, passing,
chafing against the low rocks, swirling, passing.
Better get this job over quick.
Listen:
a fourworded wavespeech:
seesoo, hrss, rsseeiss, ooos.
Vehement breath of waters
amid seasnakes, rearing horses, rocks.
In cups of rocks it slops:
flop, slop, slap:
bounded in barrels.
And, spent, its speech ceases.
It flows purling,
widely flowing,
floating foampool,
\gab{460}
flower unfurling.

\N2:
Under the upswelling tide
he saw the writhing weeds lift languidly
and sway reluctant arms,
hising up their petticoats,
in whispering water swaying and upturning coy silver fronds.
Day by day:
night by night:
lifted, flooded and let fall.
Lord, they are weary;
and, whispered to, they sigh.
Saint Ambrose heard it,
sigh of leaves and waves,
waiting, awaiting the fullness of their times,
\latin{diebus ac noctibus iniurias patiens ingemiscit}.
To no end gathered;
vainly then released, forthflowing,
wending back:
loom of the moon.
Weary too in sight of lovers,
lascivious men,
a naked woman shining in her courts,
she draws a toil of waters.

\StephenInt:
Five fathoms out there.
Full fathom five thy father lies.
\gab{470}
At one, he said.
Found drowned.
High water at Dublin bar.
Driving before it a loose drift of rubble,
fanshoals of fishes, silly shells.
A corpse rising saltwhite from the undertow,
bobbing a pace a pace a porpoise landward.
There he is.
Hook it quick.
Pull.
Sunk though he be beneath the watery floor.
We have him.
Easy now.

\StephenInt:
Bag of corpsegas sopping in foul brine.
A quiver of minnows,
fat of a spongy titbit,
flash through the slits of his buttoned trouserfly.
God becomes man
becomes fish
becomes barnacle goose
becomes featherbed mountain.
Dead breaths I living breathe,
tread dead dust,
devour a urinous offal from all dead.
\gab{480}
Hauled stark over the gunwale
he breathes upward the stench of his green grave,
his leprous nosehole snoring to the sun.

\StephenInt:
A seachange this, brown eyes saltblue.
Seadeath, mildest of all deaths known to man.
Old Father Ocean.
\french{Prix de Paris}:
beware of imitations.
Just you give it a fair trial.
We enjoyed ourselves immensely.

\StephenInt:
Come.
I thirst.
Clouding over.
No black clouds anywhere, are there?
Thunderstorm.
Allbright he falls, proud lightning of the intellect,
\latin{Lucifer, dico, qui nescit occasum}.
No.
My cockle hat and staff and hismy sandal shoon.
Where?
To evening lands.
Evening will find itself.

\N2:
He took the hilt of his ashplant,
lunging with it softly, dallying still.

\StephenInt:
Yes, evening will find itself in me,
without me.
\gab{490}
All days make their end.
By the way next when is it Tuesday
will be the longest day.
Of all the glad new year,
mother, the rum tum tiddledy tum.
Lawn Tennyson, gentleman poet.
\italian{Gia}.
For the old hag with the yellow teeth.
And Monsieur Drumont,
gentleman journalist.
\italian{Gia}.
My teeth are very bad.
Why, I wonder.
Feel.
That one is going too.
Shells.
Ought I go to a dentist, I wonder, with that money?
That one.
This.
Toothless Kinch, the superman.
Why is that, I wonder,
or does it mean something perhaps?

\StephenInt:
My handkerchief.
He threw it.
I remember.
Did I not take it up?

\N2:
His hand groped vainly in his pockets.

\StephenInt:
No, I didn't.
Better buy one.

\N1:
He laid the dry snot picked from his nostril
\gab{500}
on a ledge of rock,
carefully.

\StephenInt:
For the rest let look who will.

\StephenInt:
Behind.
Perhaps there is someone.

\N1:
He turned his face over a shoulder, rere regardant.
Moving through the air high spars of a threemaster,
her sails brailed up on the crosstrees,
homing, upstream, silently moving,
a silent ship.

\end{document}
