\N1:
A point, live dog,
grew into sight running across the sweep of sand.

\StephenInt:
Lord, is he going to attack me?
Respect his liberty.
You will not be master of others or their slave.
I have my stick.
Sit tight.

\N2:
From farther away,
walking shoreward across from the crested tide,
figures, two.

\StephenInt:
The two maries.
They have tucked it safe mong the bulrushes.
Peekaboo.
I see you.
No, the dog.
He is running back to them.
Who?

\StephenInt:
Galleys of the Lochlanns ran here to beach,
\gab{300}
in quest of prey,
their bloodbeaked prows riding low on a molten pewter surf.
Dane vikings,
torcs of tomahawks aglitter on their breasts
when Malachi wore the collar of gold.
A school of turlehide whales stranded in hot noon, spouting,
hobbling in the shallows.
Then from the starving cagework city
a horde of jerkined dwarfs,
my people, with flayers' knives,
running, scaling, hacking in green blubbery whalemeat.
Famine, plague and slaughters.
Their blood is in me, their lusts my waves.
I moved among them on the frozen Liffey,
that I, a changeling, among the spluttering resin fires.
I spoke to no-one:
none to me.

\N1:
The dog's bark ran towards him, stopped, ran back.
\gab{310}

\StephenInt:
Dog of my enemy.
I just simply stood pale, silent, bayed about.
\latin{Terribilia meditans.}
A primrose doublet, fortune's knave, smiled on my fear.
For that are you pining,
the bark of their applause?
Pretenders:
live their lives.
The Bruce's brother,
Thomas Fitzgerald, silken knight,
Perkin Warbeck, York's false scion,
in breeches of silk of whiterose ivory,
wonder of a day,
and Lambert Simnel, with a tail of nans and sutlers, a scullion crowned.
All kings' sons.
Paradise of pretenders then and now.
He saved men from drowning
and you shake at a cur's yelping.
But the courtiers who mocked Guido in \italian{Or san Michele}
were in their own house.
House of ...
We don't want any of your medieval abstrusiosities.
\gab{320}
Would you do what he did?
A boat would be near, a lifebuoy.
\german{Naturlich}, put there for you.
Would you or would you not?
The man that was drowned nine days ago off Maiden's rock.
They are waiting for him now.
The truth, spit it out.
I would want to.
I would try.
I am not a strong swimmer.
Water cold soft.
When I put my face into it in the basin at Clongowes.
Can't see!
Who's behind me?
Out quickly, quickly!
Do you see the tide flowing quickly in on all sides,
sheeting the lows of sand quickly, shellcocoacoloured?
If I had land under my feet.
I want his life still to be his, mine to be mine.
A drowning man.
His human eyes scream to me out of horror of his death.
I ...
With him together down ...
I could not save her.
\gab{330}
Waters:
bitter death:
lost.

\StephenInt:
A woman and a man.
I see her skirties.
Pinned up, I bet.

\N1:
Their dog ambled about a bank of dwindling sand,
trotting, sniffing on all sides.
Looking for something lost in a past life.
Suddenly he made off like a bounding hare,
ears flung back, chasing the shadow of a lowskimming gull.
The man's shrieked whistle struck his limp ears.
He turned, bounded back, came nearer, trotted on twinkling shanks.
On a field tenney
a buck, trippant, proper, unattired.
At the lacefringe of the tide
he halted with stiff forehoofs, seawardpointed ears.
His snout lifted
barked at the wavenoise, herds of seamorse.
They serpented towards his feet,
curling, unfurling many crests,
\gab{340}
every ninth,
breaking, plashing,
from far,
from farther out,
waves and waves.

\StephenInt:
Cocklepickers.

\N1:
They waded a little way in the water and, stooping,
soused their bags and, lifting them again, waded out.
The dog yelped
running to them,
reared up and pawed them,
dropping on all fours,
again reared up at them with mute bearish fawning.
Unheeded
he kept by them
as they came towards the drier sand,
a rag of wolf's tongue redpanting from his jaws.
His speckled body ambled ahead of them
and then loped off at a calf's gallop.
The carcass lay on his path.
He stopped, sniffed, stalked round it,
brother, nosing closer, went round it,
sniffling rapidly like a dog
all over the dead dog's bedraggled fell.

\StephenInt:
Dogskull, dogsniff,
\gab{350}
eyes on the ground,
moves to one great goal.
Ah, poor dogsbody!
Here lies poor dogsbody's body.

\man:
Tatters!
Out of that, you mongrel!

\N1:
The cry brought him skulking back to his master
and a blunt bootless kick sent him unscathed
across a spit of sand, crouched in flight.
He slunk back in a curve.

\StephenInt:
Doesn't see me.

\N1:
Along by the edge of the mole
he lolloped, dawdled, smelt a rock.
and from under a cocked hindleg
pissed against it.
He trotted forward and, lifting again his hindleg,
pissed quick short at an unsmelt rock.
The simple pleasures of the poor.
His hindpaws then scattered the sand:
then his forepaws dabbled and delved.
\gab{360}
Something he buried there, his grandmother.
He rooted in the sand,
dabbling, delving
and stopped to listen to the air,
scraped up the sand
again with a fury of his claws,
soon ceasing, a pard, a panther,
got in spousebreach,
vulturing the dead.

\StephenInt:
After he woke me last night same dream or was it?
Wait.
Open hallway.
Street of harlots.
Remember.
Haroun al Raschid.
I am almosting it.
That man led me, spoke.
I was not afraid.
The melon he had he held against my face.
Smiled:
creamfruit smell.
That was the rule, said.
In.
Come.
Red carpet spread.
You will see who.

\N2:
Shouldering their bags they trudged, the red Egyptians.
\gab{370}
His blued feet out of turnedup trousers
slapped the clammy sand,
a dull brick muffler strangling his unshaven neck.
With woman steps she followed:
the ruffian and his strolling mort.
Spoils slung at her back.
Loose sand and shellgrit crusted her bare feet.
About her windraw face hair trailed.
Behind her lord, his helpmate, bing awast to Romeville.
When night hides her body's flaws
calling under her brown shawl from an archway where dogs have mired.
Her fancyman is treating two Royal Dublins in O'Loughlin's of Blackpitts.
Buss her, wap in rogues' rum lingo,
for, O, my dimber wapping dell!
A shefiend's whiteness under her rancid rags.
Fumbally's lane that night:
the tanyard smells.
\gab{380}

\N2:
\begin{verse}
    White thy fambles, red thy gan \\
    And thy quarrons dainty is. \\
    Couch a hogshead with me then. \\
    In the darkmans clip and kiss.
\end{verse}

\StephenInt:
Morose delectation Aquinas tunbelly calls this, \latin{frate porcospino}.
Unfallen Adam rode and not rutted.
Call away let him:
\emph{Thy quarrons dainty is}.
Language no whit worse than his.
Monkwords, marybeads jabber on their girdles:
roguewords, tough nuggets patter in their pockets.

\StephenInt:
Passing now.

\StephenInt:
A side eye at my Hamlet hat.
\gab{390}
If I were suddenly naked here as I sit?
I am not.
Across the sands of all the world,
followed by the sun's flaming sword,
to the west, trekking to evening lands.
She trudges, schlepps,
trains, drags, trascines her load.
A tide westering, moondrawn, in her wake.
Tides, myriadislanded, within her,
blood not mine,
\greek{oinopa ponton},
a winedark sea.
Behold the handmaid of the moon.
In sleep the wet sign calls her hour, bids her rise.
Bridebed, childbed,
bed of death,
ghostcandled.
\latin{Omnis caro ad te veniet.}
He comes, pale vampire, through storm
his eyes, his bat sails bloodying the sea,
mouth to her mouth's kiss.

\StephenInt:
Here.
Put a pin in that chap, will you?
My tablets.
Mouth to her kiss.

\StephenInt:
No.
Must be two of em.
\gab{400}
Glue em well.
Mouth to her mouth's kiss.

\N2:
His lips lipped and mouthed fleshless lips of air:
mouth to her moomb.

\StephenInt:
Oomb, allwombing tomb.

\N2:
His mouth moulded issuing breath, unspeeched:

\StephenInt:
ooeeehah:
roar of cataractic planets, globed, blazing,
roaring wayawayawayawayaway.
Paper.
The banknotes, blast them.
Old Deasy's letter.
Here.
Thanking you for the hospitality tear the blank end off.

\N1:
Turning his back to the sun
he bent over far to a table of rock and scribbled words.

\StephenInt:
That's twice I forgot to take slips from the library counter.

\N1:
His shadow lay over the rocks as he bent, ending.

\StephenInt:
Why not endless till the farthest star?
Darkly they are there behind this light,
darkness shining in the brightness,
delta of Cassiopeia, worlds.
\gab{410}
Me sits there with his augur's rod of ash,
in borrowed sandals,
by day beside a livid sea,
unbeheld, in violet night
walking beneath a reign of uncouth stars.
I throw this ended shadow from me, manshape ineluctable, call it back.
Endless, would it be mine, form of my form?
Who watches me here?
Who ever anywhere will read these written words?
Signs on a white field.
Somewhere to someone in your flutiest voice.
The good bishop of Cloyne took the veil of the temple
out of his shovel hat:
veil of space with coloured emblems
hatched on its field.
Hold hard.
Coloured on a flat:
yes, that's right.
Flat I see, then think distance,
near, far, flat I see, east, back.
Ah, see now!
Falls back suddenly, frozen in stereoscope.
\gab{420}
Click does the trick.
You find my words dark.
Darkness is in our souls do you not think?
Flutier.
Our souls, shamewounded by our sins, cling to us yet more,
a woman to her lover clinging, the more the more.

\StephenInt:
She trusts me, her hand gentle, the longlashed eyes.
Now where the blue hell am I bringing her beyond the veil?
Into the ineluctable modality
of the ineluctable visuality.
She, she, she.
What she?
The virgin at Hodges Figgis' window on Monday looking in
for one of the alphabet books you were going to write.
Keen glance you gave her.
Wrist through the braided jesse of her sunshade.
She lives in Leeson park with a grief and kickshaws, a lady of letters.
\gab{430}
Talk that to someone else,
Stevie:
a pickmeup.
Bet she wears those curse of God stays
suspenders and yellow stockings, darned with lumpy wool.
Talk about apple dumplings,
\italian{Piuttosto.}
Where are your wits?

\StephenInt:
Touch me.
Soft eyes.
Soft soft soft hand.
I am lonely here.
O, touch me soon, now.
What is that word known to all men?
I am quiet here alone.
Sad too.
Touch, touch me.

\N1:
He lay back at full stretch over the sharp rocks,
cramming the scribbled note and pencil into a pock his hat.
His hat down on his eyes.

\StephenInt:
That is Kevin Egan's movement I made,
nodding for his nap, sabbath sleep.

\kevin:
\latin{Et vidit Deus.
Et erant valde bona.}
\gab{440}
\french{Alo!
Bonjour.}

\StephenInt:
Welcome as the flowers in May.
Under its leaf he watched through peacocktwittering lashes
the southing sun.
I am caught in this burning scene.
Pan's hour, the faunal noon.
Among gumheavy serpentplants, milkoozing fruits,
where on the tawny waters leaves lie wide.
Pain is far.

\StephenInt:
\emph{And no more turn aside and brood.}

\N1:
His gaze brooded on his broadtoed boots, a buck's castoffs, \german{Nebeneinander}.
He counted the creases of rucked leather wherein another's foot had nested warm.

\StephenInt:
The foot that beat the ground in tripudium,
foot I dislove.
But you were delighted when Esther Osvalt's shoe went on you:
girl I knew in Paris.
\french{Tiens, quel petit pied!}
\gab{450}
Staunch friend, a brother soul:
Wilde's love that dare not speak its name.
His arm:
Cranly's arm.
He now will leave me.
And the blame?
As I am.
As I am.
All or not at all.

\N2:
In long lassoes from the Cock lake
the water flowed full,
covering greengoldenly lagoons of sand,
rising, flowing.

\StephenInt:
My ashplant will float away.
I shall wait.
No, they will pass on, passing,
chafing against the low rocks, swirling, passing.
Better get this job over quick.
Listen:
a fourworded wavespeech:
seesoo, hrss, rsseeiss, ooos.
Vehement breath of waters
amid seasnakes, rearing horses, rocks.
In cups of rocks it slops:
flop, slop, slap:
bounded in barrels.
And, spent, its speech ceases.
It flows purling,
widely flowing,
floating foampool,
\gab{460}
flower unfurling.

\N2:
Under the upswelling tide
he saw the writhing weeds lift languidly
and sway reluctant arms,
hising up their petticoats,
in whispering water swaying and upturning coy silver fronds.
Day by day:
night by night:
lifted, flooded and let fall.
Lord, they are weary;
and, whispered to, they sigh.
Saint Ambrose heard it,
sigh of leaves and waves,
waiting, awaiting the fullness of their times,
\latin{diebus ac noctibus iniurias patiens ingemiscit}.
To no end gathered;
vainly then released, forthflowing,
wending back:
loom of the moon.
Weary too in sight of lovers,
lascivious men,
a naked woman shining in her courts,
she draws a toil of waters.

\StephenInt:
Five fathoms out there.
Full fathom five thy father lies.
\gab{470}
At one, he said.
Found drowned.
High water at Dublin bar.
Driving before it a loose drift of rubble,
fanshoals of fishes, silly shells.
A corpse rising saltwhite from the undertow,
bobbing a pace a pace a porpoise landward.
There he is.
Hook it quick.
Pull.
Sunk though he be beneath the watery floor.
We have him.
Easy now.

\StephenInt:
Bag of corpsegas sopping in foul brine.
A quiver of minnows,
fat of a spongy titbit,
flash through the slits of his buttoned trouserfly.
God becomes man
becomes fish
becomes barnacle goose
becomes featherbed mountain.
Dead breaths I living breathe,
tread dead dust,
devour a urinous offal from all dead.
\gab{480}
Hauled stark over the gunwale
he breathes upward the stench of his green grave,
his leprous nosehole snoring to the sun.

\StephenInt:
A seachange this, brown eyes saltblue.
Seadeath, mildest of all deaths known to man.
Old Father Ocean.
\french{Prix de Paris}:
beware of imitations.
Just you give it a fair trial.
We enjoyed ourselves immensely.

\StephenInt:
Come.
I thirst.
Clouding over.
No black clouds anywhere, are there?
Thunderstorm.
Allbright he falls, proud lightning of the intellect,
\latin{Lucifer, dico, qui nescit occasum}.
No.
My cockle hat and staff and hismy sandal shoon.
Where?
To evening lands.
Evening will find itself.

\N2:
He took the hilt of his ashplant,
lunging with it softly, dallying still.

\StephenInt:
Yes, evening will find itself in me,
without me.
\gab{490}
All days make their end.
By the way next when is it Tuesday
will be the longest day.
Of all the glad new year,
mother, the rum tum tiddledy tum.
Lawn Tennyson, gentleman poet.
\italian{Gia}.
For the old hag with the yellow teeth.
And Monsieur Drumont,
gentleman journalist.
\italian{Gia}.
My teeth are very bad.
Why, I wonder.
Feel.
That one is going too.
Shells.
Ought I go to a dentist, I wonder, with that money?
That one.
This.
Toothless Kinch, the superman.
Why is that, I wonder,
or does it mean something perhaps?

\StephenInt:
My handkerchief.
He threw it.
I remember.
Did I not take it up?

\N2:
His hand groped vainly in his pockets.

\StephenInt:
No, I didn't.
Better buy one.

\N1:
He laid the dry snot picked from his nostril
\gab{500}
on a ledge of rock,
carefully.

\StephenInt:
For the rest let look who will.

\StephenInt:
Behind.
Perhaps there is someone.

\N1:
He turned his face over a shoulder, rere regardant.
Moving through the air high spars of a threemaster,
her sails brailed up on the crosstrees,
homing, upstream, silently moving,
a silent ship.
