%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\input{defs}

\title{Circe\\
\small first draft}
\author{George V. Reilly\\
for the Wild Geese Players of Seattle\\
{\emph{www.WildGeeseSeattle.org}}}
\date{February 28th, 2009}
\raggedbottom

\begin{document}

\maketitle

\pagebreak

\commentary{
At the end of \Title{The Oxen of the Sun}, after Burke's had closed,
Stephen Dedalus and Vincent Lynch set off for Nighttown
in search of a brothel.
Stephen is ratarsed with drink and absinthe.
Leopold Bloom, taking a paternal interest in Stephen,
follows them at a distance.

An enormously long chapter full of hallucinations and nightmares,
where Bloom confronts many of his unspoken desires
and emerges strengthened.

In Homer's \Title{Odyssey},
Circe is a sorceress who drugs Odysseus' men and turns them into swine.
Odysseus rescues his men from Circe and becomes her lover.

Bloom's innermost fantasies and wishes.
Nightmare.
Stephen and Bloom sink to the bottom.

Burgess and Kenner both argue that the hallucinations
are coming from without,
that Bloom himself does not experience them.
Burgess: ``They are hallucinations from the author's brain,
not his hero's, but there is nothing vague or shadowy about them.''
}

\stage{The Mabbot street entrance of Nighttown,
before which stretches an uncobbled tramsiding set with skeleton tracks,
red and green will-o'-the-wisps and danger signals.
Rows of grimy houses with gaping doors.
Rare lamps with faint rainbow fins.
Round rabaiotti's halted ice gondola
stunted men and women squabble.
They grab wafers between which are wedged
lumps of coral and copper snow.
Sucking, they scatter slowly.
Children.
The swancomb of the gondola, highreared, forges on through the murk,
white and blue under a lighthouse.
Whistles call and answer.}

\Calls:
\gab{10}
Wait, my love, and I'll be with you.

\Answers:
Round behind the stable.

\stage{(A deafmute idiot with goggle eyes, his shapeless mouth dribbling,
jerks past, shaken in Saint Vitus' dance.
A chain of children's hands imprisons him.)}

\Children:
Kithogue! Salute!

\Idiot:
\stage{(lifts a palsied left arm and gurgles)} Grhahute!%
\gab{20}

\Children:
Where's the great light?

\Idiot:
\stage{(gobbling)} Ghaghahest.

\stage{(They release him.
He jerks on.
A pigmy woman swings on a rope slung between two railings, counting.
A form sprawled against a dustbin and muffled by its arm and hat snores,
groans, grinding growling teeth, and snores again.
On a step a gnome totting among a rubbishtip
crouches to shoulder a sack of rags and bones.
A crone standing by with a smoky oillamp
\gab{30}
rams her last bottle in the maw of his sack.
He heaves his booty, tugs askew his peaked cap and hobbles off mutely.
The crone makes back for her lair, swaying her lamp.
A bandy child, asquat on the doorstep with a paper shuttlecock,
crawls sidling after her in spurts, clutches her skirt, scrambles up.
A drunken navvy grips with both hands the railings of an area, lurching heavily.
At a corner two night watch in shouldercapes,
their hands upon their staffholsters, loom tall.
A plate crashes:
a woman screams: a child wails.
Oaths of a man roar, mutter, cease.
Figures wander, lurk, peer from warrens.
In a room lit by a candle stuck in a bottleneck
\gab{40}
a slut combs out the tatts from the hair of a scrofulous child.
Cissy Caffrey's voice, still young, sings shrill from a lane.)}

\Cissy:
\begin{verse}
    I gave it to Molly\\
    Because she was jolly,\\
    The leg of the duck,\\
    The leg of the duck.
\end{verse}

\stage{(Private Carr and Private Compton, swaggersticks tight in their oxters,
as they march unsteadily rightaboutface and
burst together from their mouths a volleyed fart.
Laughter of men from the lane.
\gab{50}
A hoarse virago retorts.)}

\Virago:
Signs on you, hairy arse.
More power the Cavan girl.

\Cissy:
More luck to me.
Cavan, Cootehill and Belturbet.
\stage{(she sings)}

\begin{verse}
    I gave it to Nelly\\
    To stick in her belly,\\
    The leg of the duck,\\
    The leg of the duck.
\end{verse}


\stage{(Private Carr and Private Compton turn and counterretort,
\gab{60}
their tunics bloodbright in a lampglow,
black sockets of caps on their blond cropped polls.
Stephen Dedalus and Lynch pass through the crowd close to the redcoats.)}

\Compton:
\stage{(jerks his finger)} Way for the parson.

\Carr:
\stage{(turns and calls)} What ho, parson!

\Cissy:
\stage{(her voice soaring higher)}

\begin{verse}
\gab{70}%
    She has it, she got it,\\
    Wherever she put it,\\
    The leg of the duck.
\end{verse}


\stage{(Stephen, flourishing the ashplant in his left hand,
chants with joy the introit for paschal time.
Lynch, his jockeycap low on his brow, attends him,
a sneer of discontent wrinkling his face.)}

\Stephen:
\latin{Vidi aquam egredientem de templo a latere dextro.
Alleluia.}

\stage{(the famished snaggletusks of an elderly bawd protrude from a doorway.)}

\Bawd:
\gab{80}
\stage{(her voice whispering huskily)} Sst! Come here till I tell you.
Maidenhead inside. Sst!

\Stephen:
\stage{(altius aliquantulum)}
\latin{Et omnes ad quos pervenit aqua ista.}

\Bawd:
\stage{(spits in their trail her jet of venom)}
Trinity medicals.
Fallopian tube.
All prick and no pence.

\stage{(Edy Boardman, sniffling, crouched with Bertha Supple,
draws her shawl across her nostrils.)}

\Edy:
\gab{90}
\stage{(bickering)}
And says the one:
I seen you up Faithful place with your squarepusher,
the greaser off the railway, in his cometobed hat.
Did you, says I.
That's not for you to say, says I.
You never seen
me in the mantrap with a married highlander, says I.
The likes of her!
Stag that one is!
Stubborn as a mule!
And her walking with two fellows the one time,
Kilbride, the enginedriver, and lancecorporal Oliphant.

\Stephen:
\stage{(triumphaliter)}
\latin{Salvi facti sunt.}

\stage{(He flourishes his ashplant, shivering the lamp image,
shattering light over the world.
\gab{100}
A liver and white spaniel on the prowl slinks after him, growling.
Lynch scares it with a kick.)}

\Lynch:
So that?

\Stephen:
\stage{(looks behind)}
So that gesture, not music not odour, would be a universal language,
the gift of tongues rendering visible not the lay sense
but the first entelechy, the structural rhythm.

\Lynch:
Pornosophical philotheology.
Metaphysics in Mecklenburgh street!

\Stephen:
\gab{110}
We have shrewridden Shakespeare and henpecked Socrates.
Even the allwisest Stagyrite was bitted, bridled and mounted by a light of love.

\Lynch:
Ba!

\Stephen:
Anyway, who wants two gestures to illustrate a loaf and a jug?
This movement illustrates the loaf and jug of bread or wine in Omar.
Hold my stick.

\Lynch:
Damn your yellow stick.
\gab{120}
Where are we going?

\Stephen:
Lecherous lynx, to \french{la belle dame sans merci}, Georgina Johnson,
\latin{ad deam qui laetificat iuventutem meam}.

\stage{(Stephen thrusts the ashplant on him and slowly holds out his hands,
his head going back till both hands are a span from his breast, down turned,
in planes intersecting, the fingers about to part, the left being higher.)}

\Lynch:
Which is the jug of bread? It skills not.
That or the customhouse.
Illustrate thou.
\gab{130}
Here take your crutch and walk.

\stage{(They pass.
Tommy Caffrey scrambles to a gaslamp and, clasping, climbs in spasms.
From the top spur he slides down.
Jacky Caffrey clasps to climb.
The navvy lurches against the lamp.
The twins scuttle off in the dark.
The navvy, swaying, presses a forefinger against a wing of his nose and
ejects from the farther nostril a long liquid jet of snot.
Shouldering the lamp he staggers away through the crowd with his flaring cresset.}

\stage{Snakes of river fog creep slowly.
From drains, clefts, cesspools, middens
arise on all sides stagnant fumes.
A glow leaps in the south
beyond the seaward reaches of the river.
\gab{140}
The navvy, staggering forward, cleaves the crowd
and lurches towards the tramsiding.}

\stage{On the farther side under the railway bridge Bloom appears,
flushed, panting, cramming bread and chocolate into a sidepocket.
From Gillen's hairdresser's window a composite portrait
shows him gallant Nelson's image.
A concave mirror at the side presents to him lovelorn longlost lugubru Booloohoom.
Grave Gladstone sees him level, Bloom for Bloom.
He passes, struck by the stare of truculent Wellington,
but in the convex mirror grin unstruck
the bonham eyes and fatchuck cheekchops of jollypoldy the rixdix doldy.}

\stage{At Antonio Rabaiotti's door Bloom halts,
\gab{150}
sweated under the bright arclamp.
He disappears.
In a moment he reappears and hurries on.)}

\Bloom:
Fish and taters. N.~g. Ah!

\stage{(He disappears into Olhausen's, the porkbutcher's,
under the downcoming rollshutter.
A few moments later he emerges from under the shutter,
puffing Poldy, blowing Bloohoom.
In each hand he holds a parcel,
one containing a lukewarm pig's crubeen,
the other a cold sheep's trotter, sprinkled with wholepepper.
He gasps, standing upright.
\gab{160}
Then bending to one side he presses a parcel against his ribs and groans.)}

\Bloom:
Stitch in my side.
Why did I run?

\stage{(He takes breath with care and goes forward slowly towards the lampset siding.
The glow leaps again.)}

\Bloom:
What is that? A flasher? Searchlight.

\stage{(He stands at Cormack's corner, watching)}

\Bloom:
\latin{Aurora borealis} or a steel foundry?
Ah, the brigade, of course.
\gab{170}
South side anyhow.
Big blaze. Might be his house.
Beggar's bush.
We're safe.
\stage{(He hums cheerfully)}
London's burning, London's burning! On fire, on fire!
\stage{(He catches sight of the navvy lurching through the crowd
at the farther side of Talbot street)}
I'll miss him.
Run. Quick. Better cross here.

\stage{(He darts to cross the road.
Urchins shout.)}

\Urchins:
Mind out, mister!
\stage{(Two cyclists, with lighted paper lanterns aswing,
swim by him, grazing him, their bells rattling.)}

\Bells:
\gab{180}
Haltyaltyaltyall.

\Bloom:
\stage{(Halts erect, stung by a spasm)} Ow!

\stage{(He looks round, darts forward suddenly.
Through rising fog a dragon sandstrewer, travelling at caution,
slews heavily down upon him, its huge red headlight winking,
its trolley hissing on the wire.
The motorman bangs his footgong.)}

\Gong:
Bang Bang Bla Bak Blud Bugg Bloo.

\stage{(The brake cracks violently.
\gab{190}
Bloom, raising a policeman's whitegloved hand,
blunders stifflegged out of the track.
The motorman, thrown forward, pugnosed, on the guidewheel,
yells as he slides past over chains and keys.)}

\Motorman:
Hey, shitbreeches, are you doing the hat trick?

\Bloom:
\stage{(Bloom trickleaps to the curbstone and halts again.
He brushes a mudflake from his cheek with a parcelled hand.)}
No thoroughfare.
Close shave that but cured the stitch.
Must take up Sandow's exercises again.
\gab{200}
On the hands down.
Insure against street accident too.
The Providential.
\stage{(he feels his trouser pocket)}
Poor mamma's panacea.
Heel easily catch in track or bootlace in a cog.
Day the wheel of the black Maria peeled off my shoe at Leonard's corner.
Third time is the charm.
Shoe trick.
Insolent driver.
I ought to report him.
Tension makes them nervous.
Might be the fellow balked me this morning with that horsey woman.
Same style of beauty.
Quick of him all the same.
The stiff walk.
True word spoken in jest.
That awful cramp in Lad lane.
Something poisonous I ate.
Emblem of luck.
Why? Probably lost cattle.
Mark of the beast.
\stage{(he closes his eyes an instant)}
Bit light in the head.
Monthly or effect of the other.
\gab{210}
Brainfogfag.
That tired feeling.
Too much for me now.
Ow!

\stage{(A sinister figure leans on plaited legs against O'Beirne's wall,
a visage unknown, injected with dark mercury.
From under a wideleaved sombrero the figure regards him with evil eye.)}

\Bloom:
\spanish{Buenas noches, senorita Blanca.
Que calle es esta?}

\Figure:
\stage{(impassive, raises a signal arm)}
Password. \irish{Sraid Mabbot.}

\Bloom:
Haha.
\gab{220}
\french{Merci.}
Esperanto.
\irish{Slan leath.}
\stage{(he mutters)}
Gaelic league spy, sent by that fireeater.

\stage{(He steps forward.
A sackshouldered ragman bars his path.
He steps left, ragsackman left.)}

\Bloom:
I beg.
\stage{(He swerves, sidles, stepaside, slips past and on.)}

\Bloom:
\gab{230}
Keep to the right, right, right.
If there is a signpost planted by the Touring Club at Stepaside
who procured that public boon?
I who lost my way and contributed to the columns of the \Title{Irish Cyclist}
the letter headed \Title{In darkest Stepaside}.
Keep, keep, keep to the right.
Rags and bones at midnight.
A fence more likely.
First place murderer makes for.
Wash off his sins of the world.

\stage{(Jacky Caffrey, hunted by Tommy Caffrey, runs full tilt against Bloom.)}

\Bloom:
O
\gab{240}

\stage{(Shocked, on weak hams, he halts.
Tommy and Jacky vanish there, there.
Bloom pats with parcelled hands watchfob, pocketbookpocket, pursepoke,
sweets of sin, potatosoap.)}

\Bloom:
Beware of pickpockets.
Old thieves' dodge.
Collide. Then snatch your purse.

\stage{(The retriever approaches sniffing, nose to the ground.
A sprawled form sneezes.
A stooped bearded figure appears garbed in the long caftan of an elder in Zion
and a smokingcap with magenta tassels.
Horned spectacles hang down at the wings of the nose.
\gab{250}
Yellow poison streaks are on the drawn face.)}

\Rudolph:
Second halfcrown waste money today.
I told you not go with drunken goy ever.
So you catch no money.

\Bloom:
\stage{(hides the crubeen and trotter behind his back and,
crestfallen, feels warm and cold feetmeat)}
\german{Ja, ich weiss, papachi.}

\Rudolph:
What you making down this place? Have you no soul?
\stage{(with feeble vulture talons he feels the silent face of Bloom)}
\gab{260}
Are you not my son Leopold, the grandson of Leopold?
Are you not my dear son Leopold who left the house of his father
and left the god of his fathers Abraham and Jacob?

\Bloom:
\stage{(with precaution)}
I suppose so, father.
Mosenthal. All that's left of him.

\Rudolph:
\stage{(severely)}
One night they bring you home drunk as dog after spend your good money.
What you call them running chaps?

\Bloom:
\stage{(In youth's smart blue Oxford suit with white vestslips,
narrowshouldered, in brown Alpine hat,
\gab{270}
wearing gent's sterling silver Waterbury keyless watch
and double curb Albert with seal attached,
one side of him coated with stiffening mud)}
Harriers, father.
Only that once.

\Rudolph:
Once! Mud head to foot.
Cut your hand open. Lockjaw.
They make you kaputt, Leopoldleben.
You watch them chaps.

\Bloom:
\stage{(weakly)}
they challenged me to a sprint.
It was muddy. I slipped.

\Rudolph:
\stage{(with contempt)}
\yiddish{Goim nachez!}
Nice spectacles for your poor mother!

\Bloom:
\gab{280}
Mamma!

\Ellen:
\stage{(in pantomime dame's stringed mobcap,
widow Twankey's crinoline and bustle,
blouse with muttonleg sleeves buttoned behind,
grey mittens and cameo brooch,
her plaited hair in a crispine net,
appears over the staircase banisters,
a slanted candlestick in her hand,
and cries out in shrill alarm)}
O blessed Redeemer, what have they done to him!
My smelling salts!
\stage{(She hauls up a reef of skirt and ransacks the
pouch of her striped blay petticoat.
A phial, an Agnus Dei,
a shrivelled potato and a celluloid doll fall out.)}
Sacred Heart of Mary, where were you at all at all?
\gab{290}

\stage{(Bloom, mumbling, his eyes downcast,
begins to bestow his parcels in his filled pockets but desists, muttering.)}

\Voice:
\stage{(sharply)} Poldy!

\Bloom:
Who? \stage{(he ducks and wards off a blow clumsily)}
At your service.

\stage{(He looks up.
Beside her mirage of datepalms
a handsome woman in turkish costume stands before him.
Opulent curves fill out her scarlet trousers and jacket, slashed with gold.
A wide yellow cummerbund girdles her.
\gab{300}
A white yashmak, violet in the night, covers her face,
leaving free only her large dark eyes and raven hair.)}

\Bloom:
Molly!

\Marion:
Welly?
Mrs~Marion from this out, my dear man, when you speak to me.
\stage{(satirically)} Has poor little hubby cold feet waiting so long?

\Bloom:
\stage{(shifts from foot to foot)}
No, no. Not the least little bit.

\stage{
\gab{310}
(He breathes in deep agitation, swallowing gulps of air, questions,
hopes, crubeens for her supper, things to tell her, excuse, desire, spellbound.
A coin gleams on her forehead.
On her feet are jewelled toerings.
Her ankles are linked by a slender fetterchain.
Beside her a camel, hooded with a turreting turban, waits.
A silk ladder of innumerable rungs climbs to his bobbing howdah.
He ambles near with disgruntled hindquarters.
Fiercely she slaps his haunch, her goldcurb wristbangles angriling,
scolding him in Moorish.)}

\Marion:
Nebrakada! Femininum!

\stage{
\gab{320}
(The camel, lifting a foreleg, plucks from a tree a large mango fruit,
offers it to his mistress, blinking, in his cloven hoof,
then droops his head and, grunting, with uplifted neck, fumbles to kneel.
Bloom stoops his back for leapfrog.)}

\Bloom:
I can give you ...
I mean as your business menagerer ...
Mrs~Marion ...
if you  ...

\Marion:
So you notice some change?
\stage{(her hands passing slowly over her trinketed stomacher,
a slow friendly mockery in her eyes)}
O Poldy, Poldy, you are a poor old stick in the mud!
\gab{330}
Go and see life.
See the wide world.

\Bloom:
I was just going back for that lotion whitewax, orangeflower water.
Shop closes early on Thursday.
But the first thing in the morning.
\stage{(he pats divers pockets)}
This moving kidney. Ah!

\stage{(He points to the south, then to the east.
A cake of new clean lemon soap arises, diffusing light and perfume.)}

\Soap:
\begin{verse}
    We're a capital couple are Bloom and I.\\
    He brightens the earth. I polish the sky.
\end{verse}

\stage{
\gab{340}
(The freckled face of Sweny, the druggist, appears in the disc of the soapsun.)}

\Sweny:
Three and a penny, please.

\Bloom:
Yes. For my wife.
Mrs~Marion. Special recipe.

\Marion:
\stage{(softly)} Poldy!

\Bloom:
Yes, ma'am?

\Marion:
\gab{350}
\italian{Ti trema un poco il cuore?}

\stage{(In disdain she saunters away, plump as a pampered pouter pigeon,
humming the duet from \italian{Don Giovanni}.)}

\Bloom:
Are you sure about that \italian{Voglio}?
I mean the pronunciati ...

\stage{(He follows, followed by the sniffing terrier.
The elderly bawd seizes his sleeve,
the bristles of her chinmole glittering.)}

\Bawd:
Ten shillings a maidenhead.
Fresh thing was never touched. Fifteen.
There's no-one in it
\gab{360}
only her old father that's dead drunk.

\stage{(She points.
In the gap of her dark den
furtive, rainbedraggled, Bridie Kelly stands.)}

\Bridie:
Hatch street.
Any good in your mind?

\stage{(With a squeak she flaps her bat shawl and runs.
A burly rough pursues with booted strides.
He stumbles on the steps, recovers, plunges into gloom.
Weak squeaks of laughter are heard, weaker.)}

\Bawd:
\stage{(her wolfeyes shining)}
He's getting his pleasure.
You won't get a virgin in the flash houses.
Ten shillings.
\gab{370}
Don't be all night before the polis in plain clothes sees us.
Sixtyseven is a bitch.

\stage{(Leering, Gerty MacDowell limps forward.
She draws from behind, ogling,
and shows coyly her bloodied clout.)}

\Gerty:
With all my worldly goods I thee and thou.
\stage{(she murmurs)} You did that.
I hate you.

\Bloom:
I? When?
You're dreaming.
I never saw you.

\Bawd:
Leave the gentleman alone, you cheat.
\gab{380}
Writing the gentleman false letters.
Streetwalking and soliciting.
Better for your mother
take the strap to you at the bedpost, hussy like you.

\Gerty:
\stage{(to Bloom)}
When you saw all the secrets of my bottom drawer.
\stage{(she paws his sleeve, slobbering)}
Dirty married man!
I love you for doing that to me.

\stage{(She glides away crookedly.
Mrs~Breen in man's frieze overcoat with loose bellows pockets,
stands in the causeway, her roguish eyes wideopen,
smiling in all her herbivorous buckteeth.)}

\MrsBreen:
Mr~...
\gab{390}

\Bloom:
\stage{(coughs gravely)}
Madam, when we last had this pleasure by letter dated the sixteenth instant ...

\MrsBreen:
Mr~Bloom! You down here in the haunts of sin!
I caught you nicely! Scamp!

\Bloom:
\stage{(hurriedly)}
Not so loud my name.
Whatever do you think of me?
Don't give me away.
Walls have ears.
How do you do? It's ages since I.
You're looking splendid.
\gab{400}
Absolutely it.
Seasonable weather we are having this time of year.
Black refracts heat.
Short cut home here.
Interesting quarter.
Rescue of fallen women.
Magdalen asylum.
I am the secretary ...

\MrsBreen:
\stage{(holds up a finger)}
Now, don't tell a big fib!
I know somebody won't like that.
O just wait till I see Molly!
\stage{(slily)}
Account for yourself this very sminute or woe betide you!

\Bloom:
\stage{(looks behind)}
She often said she'd like to visit.
Slumming.
The exotic, you see.
Negro servants in livery too if she had money.
Othello black brute.
Eugene Stratton.
\gab{410}
Even the bones and cornerman at the Livermore christies.
Bohee brothers.
Sweep for that matter.

\stage{(Tom and Sam Bohee, coloured coons in white duck suits, scarlet socks,
upstarched sambo chokers and large scarlet asters in their buttonholes, leap out.
Each has his banjo slung.
Their paler smaller negroid hands jingle the twingtwang wires.
Flashing white kaffir eyes and tusks
they rattle through a breakdown in clumsy clogs,
twinging, singing, back to back, toe heel, heel toe,
with smackfatclacking nigger lips.)}

\Bohees:
\begin{verse}
\gab{420}%
    There's someone in the house with Dina\\
    There's someone in the house, I know,\\
    There's someone in the house with Dina\\
    Playing on the old banjo.
\end{verse}

\stage{(They whisk black masks from raw babby faces:
then, chuckling, chortling, trumming, twanging,
they diddle diddle cakewalk dance away.)}

\Bloom:
\stage{(with a sour tenderish smile)}
A little frivol, shall we, if you are so inclined?
Would you like me perhaps to embrace you
just for a fraction of a second?

\MrsBreen:
\gab{430}
\stage{(screams gaily)}
O, you ruck! You ought to see yourself!

\Bloom:
For old sake' sake.
I only meant a square party, a mixed marriage
mingling of our different little conjugials.
You know I had a soft corner for you.
\stage{(gloomily)}
'Twas I sent you that valentine of the dear gazelle.

\MrsBreen:
Glory Alice, you do look a holy show! Killing simply.
\stage{(she puts out her hand inquisitively)}
What are you hiding behind your back?
Tell us, there's a dear.

\Bloom:
\gab{440}
\stage{(seizes her wrist with his free hand)}
Josie Powell that was, prettiest deb in Dublin.
How time flies by! Do you remember, harking back
in a retrospective arrangement, Old Christmas night,
Georgina Simpson's housewarming while they were playing the Irving Bishop game,
finding the pin blindfold and thoughtreading?
Subject, what is in this snuffbox?

\MrsBreen:
You were the lion of the night
with your seriocomic recitation and you looked the part.
You were always a favourite with the ladies.

\Bloom:
\stage{(squire of dames, in dinner jacket with wateredsilk facings,
\gab{450}
blue masonic badge in his buttonhole, black bow and mother-of-pearl studs,
a prismatic champagne glass tilted in his hand)}
Ladies and gentlemen, I give you Ireland, home and beauty.

\MrsBreen:
The dear dead days beyond recall.
Love's old sweet song.

\Bloom:
\stage{(meaningfully dropping his voice)}
I confess I'm teapot with curiosity to find out
whether some person's something is a little teapot at present.

\MrsBreen:
\stage{(gushingly)}
Tremendously teapot!
\gab{460}
London's teapot and I'm simply teapot all over me!
\stage{(she rubs sides with him)}
After the parlour mystery games and the crackers from the tree
we sat on the staircase ottoman.
Under the mistletoe.
Two is company.

\Bloom:
\stage{(wearing a purple napoleon hat with an amber halfmoon,
his fingers and thumb passing slowly down to her soft moist meaty palm
which she surrenders gently)}
The witching hour of night.
I took the splinter out of this hand, carefully, slowly.
\stage{(tenderly, as he slips on her finger a ruby ring)}
\italian{Là ci darem la mano.}

\MrsBreen:
\gab{470}
\stage{(in a onepiece evening frock executed in moonlight blue,
a tinsel sylph's diadem on her brow with her dancecard
fallen beside her moonblue satin slipper,
curves her palm softly, breathing quickly)}
\italian{Voglio e non.}
You're hot! You're scalding!
The left hand nearest the heart.

\Bloom:
When you made your present choice
they said it was beauty and the beast.
I can never forgive you for that.
\stage{(his clenched fist at his brow)}
Think what it means.
All you meant to me then.
\stage{(hoarsely)}
Woman, it's breaking me!

\stage{(Denis Breen, whitetallhatted, with Wisdom Hely's sandwich-boards,
\gab{480}
shuffles past them in carpet slippers, his dull beard thrust out,
muttering to right and left.
Little Alf Bergan, cloaked in the pall of the ace of spades,
dogs him to left and right, doubled in laughter.)}

\AlfBergan:
\stage{(points jeering at the sandwichboards)}
U. p:
Up.

\MrsBreen:
\stage{(to Bloom)}
High jinks below stairs.
\stage{(she gives him the glad eye)}
Why didn't you kiss the spot to make it well?
You wanted to.

\Bloom:
\stage{(shocked)}
\gab{490}
Molly's best friend! Could you?

\MrsBreen:
\stage{(her pulpy tongue between her lips, offers a pigeon kiss)}
Hnhn.
The answer is a lemon.
Have you a little present for me there?

\Bloom:
\stage{(offhandedly)}
Kosher.
A snack for supper.
The home without potted meat is incomplete.
I was at \Title{Leah}.
Mrs~Bandmann Palmer.
Trenchant exponent of Shakespeare.
Unfortunately threw away the programme.
Rattling good place round there for pigs' feet.
Feel.

\stage{(Richie Goulding, three ladies' hats pinned on his head,
appears weighted to one side by the black legal bag of
\gab{500}
Collis and Ward on which a skull and crossbones are painted in white limewash.
He opens it and shows it full of polonies,
kippered herrings, findon haddies and tightpacked pills.)}

\Richie:
Best value in Dub.

\stage{(Bald Pat, bothered beetle, stands on the curbstone,
folding his napkin, waiting to wait.)}

\Pat:
\stage{(advances with a tilted dish of spillspilling gravy)}
Steak and kidney.
Bottle of lager.
Hee hee hee.
\gab{510}
Wait till I wait.

\Richie:
Goodgod.
Inev erate inall ...

\stage{(With hanging head he marches doggedly forward.
The navvy, lurching by, gores him with his flaming pronghorn.)}

\Richie:
\stage{(with a cry of pain, his hand to his back)}
Ah! Bright's! Lights!

\Bloom:
\stage{(points to the navvy)}
A spy.
Don't attract attention.
I hate stupid crowds.
I am not on pleasure bent.
I am in a grave predicament.

\MrsBreen:
\gab{520}
Humbugging and deluthering as per usual with your cock and bull story.

\Bloom:
I want to tell you a little secret about how I came to be here.
But you must never tell.
Not even Molly.
I have a most particular reason.

\MrsBreen:
\stage{(all agog)}
O, not for worlds.

\Bloom:
Let's walk on.
Shall us?

\MrsBreen:
Let's.
\gab{530}

\stage{(The bawd makes an unheeded sign.
Bloom walks on with Mrs~Breen.
The terrier follows, whining piteously, wagging his tail.)}

\Bawd:
Jewman's melt!

\Bloom:
\stage{(in an oatmeal sporting suit, a sprig of woodbine in the lapel,
tony buff shirt, shepherd's plaid Saint Andrew's cross scarftie,
white spats, fawn dustcoat on his arm, tawny red brogues,
fieldglasses in bandolier and a grey billycock hat)}
Do you remember a long long time,
years and years ago, just after Milly,
Marionette we called her,
\gab{540}
was weaned when we all went together to Fairyhouse races, was it?

\MrsBreen:
\stage{(in smart saxe tailormade, white velours hat and spider veil)}
Leopardstown.

\Bloom:
I mean, Leopardstown.
And Molly won seven shillings on a three year old named Nevertell
and coming home along by Foxrock
in that old fiveseater shanderadan of a waggonette
you were in your heyday then and
you had on that new hat of white velours with a surround of molefur that Mrs~Hayes
\gab{550}
advised you to buy because it was marked down to nineteen and eleven,
a bit of wire and an old rag of velveteen,
and I'll lay you what you like she did it on purpose ...

\MrsBreen:
She did, of course, the cat!
Don't tell me! Nice adviser!

\Bloom:
Because it didn't suit you one quarter as well as
the other ducky little tammy toque with the bird of paradise wing in it
that I admired on you and you honestly looked just too fetching in it
though it was a pity to kill it, you cruel naughty creature,
little mite of a thing with a heart the size of a
fullstop.%
\gab{560}

\MrsBreen:
\stage{(squeezes his arm, simpers)}
Naughty cruel I was!

\Bloom:
\stage{(low, secretly, ever more rapidly)}
And Molly was eating a sandwich of spiced beef
out of Mrs~Joe Gallaher's lunch basket.
Frankly, though she had her advisers or admirers,
I never cared much for her style.
She was ...

\MrsBreen:
Too ...

\Bloom:
Yes.
\gab{570}
And Molly was laughing because Rogers and Maggot O'Reilly
were mimicking a cock as we passed a farmhouse and
Marcus Tertius Moses, the tea merchant,
drove past us in a gig with his daughter,
Dancer Moses was her name,
and the poodle in her lap bridled up
and you asked me if I ever heard or read or knew or came across ...

\MrsBreen:
\stage{(eagerly)}
Yes, yes, yes, yes, yes, yes, yes.

\stage{(She fades from his side.
Followed by the whining dog he walks on towards hellsgates.
In an archway a standing woman, bent forward, her feet apart, pisses cowily.
Outside a shuttered pub a bunch of loiterers
\gab{580}
listen to a tale which their brokensnouted gaffer rasps out with raucous humour.
An armless pair of them flop wrestling, growling, in maimed sodden playfight.)}

\Gaffer:
\stage{(crouches, his voice twisted in his snout)}
And when Cairns came down from the scaffolding in Beaver street
what was he after doing it
into only into the bucket of porter that was there
waiting on the shavings for Derwan's plasterers.

\Loiterers:
\stage{(guffaw with cleft palates)}
O jays!

\stage{(Their paintspeckled hats wag.
\gab{590}
Spattered with size and lime of their lodges
they frisk limblessly about him.)}

\Bloom:
Coincidence too.
They think it funny.
Anything but that.
Broad daylight.
Trying to walk.
Lucky no woman.

\Loiterers:
Jays, that's a good one.
Glauber salts.
O jays, into the men's porter.

\stage{(Bloom passes.
Cheap whores, singly, coupled, shawled, dishevelled,
call from lanes, doors, corners.)}

\Whores:
\begin{verse}
\gab{600}%
    Are you going far, queer fellow?\\
    How's your middle leg?\\
    Got a match on you?\\
    Eh, come here till I stiffen it for you.
\end{verse}

\stage{(He plodges through their sump towards the lighted street beyond.
From a bulge of window curtains a gramophone rears a battered brazen trunk.
In the shadow a shebeenkeeper haggles with the navvy and the two redcoats.)}

\Navvy:
\stage{(belching)}
Where's the bloody house?

\Shebeenkeeper:
\gab{610}
Purdon street.
Shilling a bottle of stout.
Respectable woman.

\Navvy:
\stage{(gripping the two redcoats, staggers forward with them)}
Come on, you British army!

\Carr:
\stage{(behind his back)}
He aint half balmy.

\Compton:
\stage{(laughs)}
What ho!

\Carr:
\stage{(to the Navvy)}
\gab{620}
Portobello barracks canteen.
You ask for Carr. Just Carr.

\Navvy:
\stage{(shouts)}
    We are the boys. Of Wexford.

\Compton:
Say! What price the sergeantmajor?

\Carr:
Bennett? He's my pal.
I love old Bennett.

\Navvy:
\stage{(shouts)}
\begin{verse}
\gab{630}%
    The galling chain.\\
    And free our native land.
\end{verse}

\stage{(He staggers forward, dragging them with him.
Bloom stops, at fault.
The dog approaches, his tongue outlolling, panting)}

\Bloom:
Wildgoose chase this.
Disorderly houses.
Lord knows where they are gone.
Drunks cover distance double quick.
Nice mixup.
Scene at Westland row.
Then jump in first class with third ticket.
Then too far.
Train with engine behind.
Might have taken me to Malahide or a siding for the night or collision.
Second drink does it.
Once is a dose.
What am I following him for?
Still, he's the best of that lot.
\gab{640}
If I hadn't heard about Mrs~Beaufoy Purefoy
I wouldn't have gone and wouldn't have met.
Kismet.
He'll lose that cash.
Relieving office here.
Good biz for cheapjacks, organs.
What do ye lack? Soon got, soon gone.
Might have lost my life too
with that mangongwheeltracktrolleyglarejuggernaut
only for presence of mind.
Can't always save you, though.
If I had passed Truelock's window that day
two minutes later would have been shot.
Absence of body.
Still if bullet only went through my coat
get damages for shock, five hundred pounds.
What was he?
Kildare street club toff.
God help his gamekeeper.

\stage{(He gazes ahead, reading on the wall
a scrawled chalk legend \Title{Wet Dream} and a phallic design.)}
\gab{650}%
Odd! Molly drawing on the frosted carriagepane at Kingstown.
What's that like?
\stage{(Gaudy dollwomen loll in the lighted doorways,
in window embrasures, smoking birdseye cigarettes.
The odour of the sicksweet weed floats towards him
in slow round ovalling wreaths.)}

\Wreaths:
Sweet are the sweets.
Sweets of sin.

\Bloom:
My spine's a bit limp.
Go or turn?
And this food? Eat it and get all pigsticky.
Absurd I am.
Waste of money.
One and eightpence too much.
\stage{(The retriever drives a cold snivelling muzzle against his hand, wagging his tail.)}
Strange how they take to me.
Even that brute today.
\gab{660}
Better speak to him first.
Like women they like \french{rencontres}.
Stinks like a polecat.
\french{Chacun son goût.}
He might be mad.
Dogdays.
Uncertain in his movements.
Good fellow! Fido!
Good fellow! Garryowen!
\stage{(The wolfdog sprawls on his back, wriggling obscenely with begging paws,
his long black tongue lolling out.)}
Influence of his surroundings.
Give and have done with it.
Provided nobody.
\stage{(Calling encouraging words he shambles back
with a furtive poacher's tread,
dogged by the setter into a dark stalestunk corner.
He unrolls one parcel and goes to dump the crubeen softly
but holds back and feels the trotter.)}
Sizeable for threepence.
But then I have it in my left hand.
Calls for more effort.
\gab{670}
Why? Smaller from want of use.
O, let it slide.
Two and six.

\stage{(With regret he lets the unrolled crubeen and trotter slide.
The mastiff mauls the bundle clumsily and gluts himself
with growling greed, crunching the bones.
Two raincaped watch approach, silent, vigilant.
They murmur together.)}

\Watch:
Bloom.
Of Bloom.
For Bloom.
Bloom.

\stage{(Each lays hand on Bloom's shoulder.)}

\FirstWatch:
Caught in the act.
\gab{680}
Commit no nuisance.

\Bloom:
\stage{(stammers)}
I am doing good to others.

\stage{(A covey of gulls, storm petrels, rises hungrily from Liffey slime
with Banbury cakes in their beaks.)}

\Gulls:
Kaw kave kankury kake.

\Bloom:
The friend of man.
Trained by kindness.

\stage{(He points.
Bob Doran, toppling from a high barstool, sways over the munching spaniel.)%
\gab{690}
}

\Doran:
Towser.
Give us the paw.
Give the paw.

\stage{(The bulldog growls, his scruff standing, a gobbet of pig's knuckle
between his molars through which rabid scumspittle dribbles.
Bob Doran fills silently into an area.)}

\SecondWatch:
Prevention of cruelty to animals.

\Bloom:
\stage{(enthusiastically)}
A noble work!
I scolded that tramdriver on Harold's cross bridge
for illusing the poor horse with his harness scab.
\gab{700}
Bad French I got for my pains.
Of course it was frosty and the last tram.
All tales of circus life are highly demoralising.

\stage{(Signor Maffei, passionpale,
in liontamer's costume with diamond studs in his shirtfront,
steps forward, holding a circus paperhoop,
a curling carriagewhip and a revolver
with which he covers the gorging boarhound.)}

\Maffei:
\stage{(with a sinister smile)}
Ladies and gentlemen, my educated greyhound.
It was I broke in the bucking broncho Ajax
with my patent spiked saddle for carnivores.
Lash under the belly with a knotted thong.
\gab{710}
Block tackle and a strangling pulley will bring your lion to heel,
no matter how fractious, even \latin{leo ferox} there, the Libyan maneater.
A redhot crowbar and some liniment rubbing on the burning part
produced Fritz of Amsterdam, the thinking hyena.
\stage{(he glares)}
I possess the Indian sign.
The glint of my eye does it with these breastsparklers.
\stage{(with a bewitching smile)}
I now introduce Mademoiselle Ruby, the pride of the ring.

\FirstWatch:
Come.
Name and address.

\Bloom:
I have forgotten for the moment.
Ah, yes!
\gab{720}
\stage{(he takes off his high grade hat, saluting)}
Dr Bloom, Leopold, dental surgeon.
You have heard of von Blum Pasha.
Umpteen millions.
\german{Donnerwetter!}
Owns half Austria. Egypt. Cousin.

\FirstWatch:
Proof.

\stage{(A card falls from inside the leather headband of Bloom's hat.)}

\Bloom:
\stage{(in red fez, cadi's dress coat with broad green sash,
wearing a false badge of the Legion of Honour,
picks up the card hastily and offers it)}
Allow me.
My club is the Junior Army and Navy.
\gab{730}
Solicitors:
Messrs John Henry Menton, 27 Bachelor's Walk.

\FirstWatch:
\stage{(reads)}
Henry Flower.
No fixed abode.
Unlawfully watching and besetting.

\SecondWatch:
An alibi.
You are cautioned.

\Bloom:
\stage{(produces from his heartpocket a crumpled yellow flower)}
This is the flower in question.
It was given me by a man I don't know his name.
\stage{(plausibly)}
\gab{740}
You know that old joke, rose of Castile.
Bloom.
The change of name.
Virag.
\stage{(he murmurs privately and confidentially)}
We are engaged you see, sergeant.
Lady in the case.
Love entanglement.
\stage{(he shoulders the second watch gently)}
Dash it all.
It's a way we gallants have in the navy.
Uniform that does it.
\stage{(he turns gravely to the first watch)}
Still, of course, you do get your Waterloo sometimes.
Drop in some evening and have a glass of old Burgundy.
\stage{(to the second watch gaily)}
I'll introduce you, inspector.
She's game.
Do it in the shake of a lamb's tail.

\stage{(A dark mercurialised face appears, leading a veiled figure.)}

\DarkMercury:
The Castle is looking for him.
\gab{750}
He was drummed out of the army.

\Martha:
\stage{(Thickveiled, a crimson halter round her neck,
a copy of the \Title{Irish Times} in her hand, in tone of reproach, pointing)}
Henry! Leopold!
Lionel, thou lost one!
Clear my name.

\FirstWatch:
\stage{(sternly)} Come to the station.

\Bloom:
\stage{(scared, hats himself, steps back,
then, plucking at his heart and lifting his right forearm on the square,
he gives the sign and dueguard of fellowcraft)}
No, no, worshipful master, light of love.
\gab{760}
Mistaken identity.
The Lyons mail.
Lesurques and Dubosc.
You remember the Childs fratricide case.
We medical men.
By striking him dead with a hatchet.
I am wrongfully accused.
Better one guilty escape than ninetynine wrongfully condemned.

\Martha:
\stage{(sobbing behind her veil)}
Breach of promise.
My real name is Peggy Griffin.
He wrote to me that he was miserable.
I'll tell my brother, the Bective rugger fullback, on you, heartless flirt.

\Bloom:
\stage{(behind his hand)}
She's drunk.
The woman is inebriated.
\stage{(he murmurs vaguely the pass of Ephraim)}
\gab{770}
Shitbroleeth.

\SecondWatch:
\stage{(tears in his eyes, to Bloom)}
You ought to be thoroughly well ashamed of yourself.

\Bloom:
Gentlemen of the jury, let me explain.
A pure mare's nest.
I am a man misunderstood.
I am being made a scapegoat of.
I am a respectable married man, without a stain on my character.
I live in Eccles street.
My wife, I am the daughter of a most distinguished commander,
a gallant upstanding gentleman,
what do you call him, Majorgeneral Brian Tweedy,
one of Britain's fighting men who helped to win our battles.
\gab{780}
Got his majority for the heroic defence of Rorke's Drift.

\FirstWatch:
Regiment.

\Bloom:
\stage{(turns to the gallery)}
The royal Dublins, boys, the salt of the earth, known the world over.
I think I see some old comrades in arms up there among you.
The R.D.F., with our own Metropolitan police,
guardians of our homes, the pluckiest lads and the finest body of men,
as physique, in the service of our sovereign.

\Voice:
\gab{790}
Turncoat! Up the Boers! Who booed Joe Chamberlain?

\Bloom:
\stage{(his hand on the shoulder of the first watch)}
My old dad too was a J.P.
I'm as staunch a Britisher as you are, sir.
I fought with the colours for king and country
in the absentminded war under general Gough in the park
and was disabled at Spion Kop and Bloemfontein,
was mentioned in dispatches.
I did all a white man could.
\stage{(with quiet feeling)}
Jim Bludso.
Hold her nozzle again the bank.

\FirstWatch:
Profession or trade.
\gab{800}

\Bloom:
Well, I follow a literary occupation, author-journalist.
In fact we are just bringing out a collection of prize stories
of which I am the inventor, something that is an entirely new departure.
I am connected with the British and Irish press.
If you ring up ...

\stage{(Myles Crawford strides out jerkily, a quill between his teeth.
His scarlet beak blazes within the aureole of his straw hat.
He dangles a hank of Spanish onions in one hand and holds with the other hand
a telephone receiver nozzle to his ear.)}

\Myles:
\gab{810}
\stage{(his cock's wattles wagging)}
Hello, seventyseven eightfour. Hello.
\Title{Freeman's Urinal} and \Title{Weekly Arsewipe} here.
Paralyse Europe.
You which? Bluebags?
Who writes? Is it Bloom?

\stage{(Mr~Philip Beaufoy, palefaced, stands in the witnessbox,
in accurate morning dress, outbreast pocket with peak of handkerchief showing,
creased lavender trousers and patent boots.
He carries a large portfolio labelled \Title{Matcham's Masterstrokes}.)}

\Beaufoy:
\stage{(drawls})
No, you aren't.
Not by a long shot if I know it.
I don't see it that's all.
\gab{820}
No born gentleman, no-one with the most rudimentary promptings of a gentleman
would stoop to such particularly loathsome conduct.
One of those, my lord.
A plagiarist.
A soapy sneak masquerading as a litterateur.
It's perfectly obvious that with the most inherent baseness
he has cribbed some of my bestselling copy, really gorgeous stuff,
a perfect gem, the love passages in which are beneath suspicion.
The Beaufoy books of love and great possessions,
with which your lordship is doubtless familiar,
are a household word throughout the kingdom.

\Bloom:
\stage{(murmurs with hangdog meekness glum)}
That bit about the laughing witch
\gab{830}
hand in hand I take exception to, if I may ...

\Beaufoy:
\stage{(his lip upcurled, smiles superciliously on the court)}
You funny ass, you!
You're too beastly awfully weird for words!
I don't think you need over excessively disincommodate yourself in that regard.
My literary agent Mr~J.B. Pinker is in attendance.
I presume, my lord, we shall receive the usual witnesses' fees, shan't we?
We are considerably out of pocket over this bally pressman johnny,
this jackdaw of Rheims, who has not even been to a university.

\Bloom:
\stage{(indistinctly)}
\gab{840}
University of life. Bad art.

\Beaufoy:
\stage{(shouts)}
It's a damnably foul lie, showing the moral rottenness of the man!
\stage{(he extends his portfolio)}
We have here damning evidence, the \latin{corpus delicti}, my lord,
a specimen of my maturer work disfigured by the hallmark of the beast.

\GalleryVoice:
\begin{verse}
    Moses, Moses, king of the jews,\\
    Wiped his arse in the \Title{Daily News}.
\end{verse}

\Bloom:
\stage{(bravely)} Overdrawn.
\gab{850}

\Beaufoy:
You low cad!
You ought to be ducked in the horsepond, you rotter!
\stage{(to the court)}
Why, look at the man's private life!
Leading a quadruple existence!
Street angel and house devil.
Not fit to be mentioned in mixed society!
The archconspirator of the age!

\Bloom:
\stage{(to the court)}
And he, a bachelor, how ...

\FirstWatch:
The King versus Bloom.
Call the woman Driscoll.

\Crier:
\gab{860}
Mary Driscoll, scullerymaid!

\stage{(Mary Driscoll, a slipshod servant girl, approaches.
She has a bucket on the crook of her arm and a scouringbrush in her hand.)}

\SecondWatch:
Another!
Are you of the unfortunate class?

\Driscoll:
\stage{(indignantly)}
I'm not a bad one.
I bear a respectable character and was four months in my last place.
I was in a situation, six pounds a year and my chances with Fridays out
and I had to leave owing to his carryings on.

\FirstWatch:
\gab{870}
What do you tax him with?

\Driscoll:
He made a certain suggestion
but I thought more of myself as poor as I am.

\Bloom:
\stage{(in housejacket of ripplecloth, flannel trousers,
heelless slippers, unshaven, his hair rumpled:
softly)}
I treated you white.
I gave you mementos, smart emerald garters far above your station.
Incautiously I took your part when you were accused of pilfering.
There's a medium in all things.
Play cricket.

\Driscoll:
\stage{(excitedly)}
\gab{880}
As God is looking down on me this night if ever I laid a hand to them oysters!

\FirstWatch:
The offence complained of?
Did something happen?

\Driscoll:
He surprised me in the rere of the premises, Your honour,
when the missus was out shopping one morning with a request for a safety pin.
He held me and I was discoloured in four places as a result.
And he interfered twict with my clothing.

\Bloom:
She counterassaulted.
\gab{890}

\Driscoll:
\stage{(scornfully)}
I had more respect for the scouringbrush, so I had.
I remonstrated with him, Your lord, and he remarked:
keep it quiet.

\stage{(General laughter.)}

\Fottrell:
\stage{(clerk of the crown and peace, resonantly)}
Order in court!
The accused will now make a bogus statement.

\stage{(Bloom, pleading not guilty and holding a fullblown waterlily,
begins a long unintelligible speech.
They would hear what counsel had to say
in his stirring address to the grand jury.
\gab{900}
He was down and out but, though branded as a black sheep,
if he might say so, he meant to reform,
to retrieve the memory of the past in a purely sisterly way
and return to nature as a purely domestic animal.
A sevenmonths' child, he had been carefully brought up
and nurtured by an aged bedridden parent.
There might have been lapses of an erring father
but he wanted to turn over a new leaf
and now, when at long last in sight of the whipping post,
to lead a homely life in the evening of his days,
permeated by the affectionate surroundings of the heaving bosom of the family.
An acclimatised Britisher, he had seen that summer eve
\gab{910}
from the footplate of an engine cab of the Loop line railway company
while the rain refrained from falling glimpses, as it were,
through the windows of loveful households in Dublin city and urban district
of scenes truly rural of happiness of the better land
with Dockrell's wallpaper at one and ninepence a dozen,
innocent Britishborn bairns lisping prayers to the Sacred Infant,
youthful scholars grappling with their pensums
or model young ladies playing on the pianoforte
or anon all with fervour reciting the family rosary
round the crackling Yulelog
while in the boreens and green lanes
the colleens with their swains strolled
what times the strains of the organtoned melodeon
\gab{920}
Britannia metalbound with four acting stops and twelvefold bellows,
a sacrifice, greatest bargain ever...)}

\stage{(Renewed laughter.
He mumbles incoherently.
Reporters complain that they cannot hear.)}

\LongShortHand:
\stage{(without looking up from their notebooks)}
Loosen his boots.

\MacHugh:
\stage{(from the presstable, coughs and calls)}
Cough it up, man.
Get it out in bits.

\stage{(The crossexamination proceeds re Bloom and the bucket.
A large bucket.
Bloom himself.
Bowel trouble.
\gab{930}
In Beaver street.
Gripe, yes.
Quite bad.
A plasterer's bucket.
By walking stifflegged.
Suffered untold misery.
Deadly agony.
About noon.
Love or burgundy.
Yes, some spinach.
Crucial moment.
He did not look in the bucket nobody.
Rather a mess.
Not completely.
A \Title{Titbits} back number.)}

\stage{(Uproar and catcalls.
Bloom in a torn frockcoat stained with whitewash,
dinged silk hat sideways on his head,
a strip of stickingplaster across his nose, talks inaudibly.)}

\JJOM:
\stage{(in barrister's grey wig and stuffgown,
speaking with a voice of pained protest)}
\gab{940}
This is no place for indecent levity at the
expense of an erring mortal disguised in liquor.
We are not in a beargarden nor at an Oxford rag
nor is this a travesty of justice.
My client is an infant, a poor foreign immigrant
who started scratch as a stowaway and is now trying to turn an honest penny.
The trumped up misdemeanour was due to a momentary aberration of heredity,
brought on by hallucination, such familiarities as the alleged guilty occurrence
being quite permitted in my client's native place, the land of the Pharaoh.
\latin{Prima facie}, I put it to you that there was no attempt at carnally knowing.
Intimacy did not occur and the offence complained of by Driscoll,
that her virtue was solicited, was not repeated.
I would deal in especial with atavism.
\gab{950}
There have been cases of shipwreck and somnambulism in my client's family.
If the accused could speak he could a tale unfold—one
of the strangest that have ever been narrated between the covers of a book.
He himself, my lord, is a physical wreck from cobbler's weak chest.
His submission is that he is of Mongolian extraction
and irresponsible for his actions.
Not all there, in fact.

\Bloom:
\stage{(Barefoot, pigeonbreasted, in lascar's vest and trousers,
apologetic toes turned in, opens his tiny mole's eyes and looks about him dazedly,
passing a slow hand across his forehead.
Then he hitches his belt sailor fashion
and with a shrug of oriental obeisance salutes the court,
\gab{960}
pointing one thumb heavenward.)}
Him makee velly muchee fine night.
\stage{(he begins to lilt simply)}
\begin{verse}
    Li li poo lil chile\\
    Blingee pigfoot evly night\\
    Payee two shilly ...
\end{verse}

\stage{(he is howled down.)}

\JJOM:
\stage{(hotly to the populace)} This is a lonehand fight.
By Hades, I will not have any client of mine gagged and badgered
in this fashion by a pack of curs and laughing hyenas.
The Mosaic code has superseded the law of the jungle.
\gab{970}
I say it and I say it emphatically,
without wishing for one moment to defeat the ends of justice,
accused was not accessory before the act and prosecutrix has not been tampered with.
The young person was treated by defendant as if she were his very own daughter.
\stage{(Bloom takes J.J.~O'Molloy's hand and raises it to his lips.)}
I shall call rebutting evidence to prove up to the hilt
that the hidden hand is again at its old game.
When in doubt persecute Bloom.
My client, an innately bashful man, would be the last man in the world
to do anything ungentlemanly which injured modesty could object to
or cast a stone at a girl who took the wrong turning when some dastard,
responsible for her condition, had worked his own sweet will on her.
\gab{980}
He wants to go straight.
I regard him as the whitest man I know.
He is down on his luck at present owing to the mortgaging of his extensive property
at Agendath Netaim in faraway Asia Minor, slides of which will now be shown.
\stage{(to Bloom)}
I suggest that you will do the handsome thing.

\Bloom:
A penny in the pound.

\stage{(The image of the lake of Kinnereth with blurred cattle cropping in silver haze
is projected on the wall.
Moses Dlugacz, ferreteyed albino, in blue dungarees, stands up in the gallery,
holding in each hand an orange citron and a pork kidney.)}

\Dlugacz:
\gab{990}
\stage{(hoarsely)}
Bleibtreustrasse, Berlin, W.13.

\stage{(J.J.~O'Molloy steps on to a low plinth
and holds the lapel of his coat with solemnity.
His face lengthens, grows pale and bearded, with sunken eyes,
the blotches of phthisis and hectic cheekbones of John~F. Taylor.
He applies his handkerchief to his mouth
and scrutinises the galloping tide of rosepink blood.)}

\JJOM:
\stage{(almost voicelessly)}
Excuse me.
I am suffering from a severe chill, have recently come from a sickbed.
A few wellchosen words.
\stage{(he assumes the avine head, foxy moustache
\gab{1000}
and proboscidal eloquence of Seymour Bushe.)}
When the angel's book comes to be opened
if aught that the pensive bosom has inaugurated of
soultransfigured and of soultransfiguring deserves to live
I say accord the prisoner at the bar the sacred benefit of the doubt.

\stage{(A paper with something written on it is handed into court.)}

\Bloom:
\stage{(in court dress)}
Can give best references.
Messrs Callan, Coleman.
Mr~Wisdom Hely~J.P.
My old chief Joe Cuffe.
Mr~V.B.~Dillon, ex lord mayor of Dublin.
I have moved in the charmed circle of the highest ...
Queens of Dublin society.
\stage{(carelessly)}
\gab{1010}
I was just chatting this afternoon
at the viceregal lodge to my old pals, sir Robert and lady Ball,
astronomer royal at the levee.
Sir Bob, I said ...

\Yelverton:
\stage{(in lowcorsaged opal balldress and elbowlength ivory gloves,
wearing a sabletrimmed brickquilted dolman, a comb of brilliants
and panache of osprey in her hair)}
Arrest him, constable.
He wrote me an anonymous letter in prentice backhand
when my husband was in the North Riding of Tipperary on the Munster circuit,
signed James Lovebirch.
He said that he had seen from the gods my peerless globes as I sat in a box
of the \Title{Theatre Royal} at a command performance of \french{La Cigale}.
\gab{1020}
I deeply inflamed him, he said.
He made improper overtures to me to misconduct
myself at half past four p.m.
on the following Thursday, Dunsink time.
He offered to send me through the post a work of fiction by Monsieur Paul de
Kock, entitled \Title{The Girl with the Three Pairs of Stays}.

\Bellingham:
\stage{(in cap and seal coney mantle, wrapped up to the nose,
steps out of her brougham and scans through tortoiseshell quizzing-glasses
which she takes from inside her huge opossum muff)}
Also to me.
Yes, I believe it is the same objectionable person.
Because he closed my carriage door outside sir Thornley Stoker's
\gab{1030}
one sleety day during the cold snap of February ninetythree
when even the grid of the wastepipe
and the ballstop in my bath cistern were frozen.
Subsequently he enclosed a bloom of edelweiss culled on the heights,
as he said, in my honour.
I had it examined by a botanical expert
and elicited the information that it was
a blossom of the homegrown potato plant
purloined from a forcingcase of the model farm.

\Yelverton:
Shame on him!

\stage{(A crowd of sluts and ragamuffins surges forward)}

\Ragamuffins:
\stage{(screaming)}
\gab{1040}
Stop thief!
Hurrah there, Bluebeard!
Three cheers for Ikey Mo!

\SecondWatch:
\stage{(produces handcuffs)}
Here are the darbies.

\Bellingham:
He addressed me in several handwritings with fulsome compliments
as a Venus in furs and alleged profound pity for my frostbound coachman Palmer
while in the same breath he expressed himself
as envious of his earflaps and fleecy sheepskins
and of his fortunate proximity to my person,
when standing behind my chair wearing my livery
and the armorial bearings of the Bellingham escutcheon garnished sable,
\gab{1050}
a buck's head couped or.
He lauded almost extravagantly my nether extremities,
my swelling calves in silk hose drawn up to the limit,
and eulogised glowingly my other hidden treasures in priceless lace
which, he said, he could conjure up.
He urged me (stating that he felt it his mission in life to urge me)
to defile the marriage bed, to commit adultery at the earliest possible opportunity.

\Talboys:
\stage{(in amazon costume, hard hat,
jackboots cockspurred, vermilion waistcoat,
fawn musketeer gauntlets with braided drums, long train held up
and hunting crop with which she strikes her welt constantly)}
Also me.
\gab{1060}
Because he saw me on the polo ground of the Phoenix park
at the match All Ireland versus the Rest of Ireland.
My eyes, I know, shone divinely as I watched Captain Slogger Dennehy
of the Inniskillings win the final chukkar on his darling cob \Title{Centaur}.
This plebeian Don Juan observed me from behind a hackney car
and sent me in double envelopes an obscene photograph,
such as are sold after dark on Paris boulevards, insulting to any lady.
I have it still.
It represents a partially nude senorita, frail and lovely
(his wife, as he solemnly assured me, taken by him from nature),
practising illicit intercourse with a muscular torero, evidently a blackguard.
He urged me to do likewise, to misbehave, to sin with officers of the garrison.
\gab{1070}
He implored me to soil his letter in an unspeakable manner,
to chastise him as he richly deserves, to bestride and ride him,
to give him a most vicious horsewhipping.

\Bellingham:
Me too.

\Yelverton:
Me too.

\stage{(Several highly respectable Dublin ladies hold up improper letters
received from Bloom.)}

\Talboys:
\gab{1080}
\stage{(stamps her jingling spurs in a sudden paroxysm of fury)}
I will, by the God above me.
I'll scourge the pigeonlivered cur as long as I can stand over him.
I'll flay him alive.

\Bloom:
\stage{(his eyes closing, quails expectantly)}
Here?
\stage{(he squirms)}
Again!
\stage{(he pants cringing)}
I love the danger.

\Talboys:
Very much so!
I'll make it hot for you.
I'll make you dance Jack Latten for that.

\Bellingham:
\gab{1090}
Tan his breech well, the upstart!
Write the stars and stripes on it!

\Yelverton:
Disgraceful!
There's no excuse for him!
A married man!

\Bloom:
All these people.
I meant only the spanking idea.
A warm tingling glow without effusion.
Refined birching to stimulate the circulation.

\Talboys:
\stage{(laughs derisively)}
o, did you, my fine fellow?
Well, by the living God, you'll get the surprise of your life now,
believe me, the most unmerciful hiding a man ever bargained for.
\gab{1100}
You have lashed the dormant tigress in my nature into fury.

\Bellingham:
\stage{(shakes her muff and quizzing-glasses vindictively)}
make him smart, Hanna dear.
Give him ginger.
Thrash the mongrel within an inch of his life.
The cat-o'-nine-tails.
Geld him.
Vivisect him.

\Bloom:
\stage{(shuddering, shrinking, joins his hands: with hangdog mien)}
O cold! O shivery!
It was your ambrosial beauty.
Forget, forgive. Kismet.
Let me off this once.
\stage{(he offers the other cheek)}

\Yelverton:
\gab{1110}
\stage{(severely)}
Don't do so on any account, Mrs~Talboys!
He should be soundly trounced!

\Talboys:
\stage{(unbuttoning her gauntlet violently)}
i'll do no such thing.
Pigdog and always was ever since he was pupped!
To dare address me!
I'll flog him black and blue in the public streets.
I'll dig my spurs in him up to the rowel.
He is a wellknown cuckold.
\stage{(she swishes her huntingcrop savagely in the air)}
Take down his trousers without loss of time.
Come here, sir! Quick! Ready?

\Bloom:
\gab{1120}
\stage{(trembling, beginning to obey)}
The weather has been so warm.

\stage{(Davy Stephens, ringletted, passes with a bevy of barefoot newsboys.)}

\DavyStephens:
\Title{Messenger of the Sacred Heart}
and \Title{Evening Telegraph} with Saint Patrick's Day supplement.
Containing the new addresses of all the cuckolds in Dublin.

\stage{(The very reverend Canon O'Hanlon in cloth of gold cope
elevates and exposes a marble timepiece.
Before him Father Conroy and the Reverend John Hughes S.J. bend low.)%
\gab{1130}
}

\Timepiece:
\stage{(unportalling)}
\begin{verse}
    Cuckoo.\\
    Cuckoo.\\
    Cuckoo.
\end{verse}


\stage{(The brass quoits of a bed are heard to jingle.)}

\Quoits:
Jigjag. Jigajiga. Jigjag.

\stage{(A panel of fog rolls back rapidly, revealing rapidly in the jurybox
the faces of Martin Cunningham, foreman, silkhatted,
\gab{1140}
Jack Power, Simon Dedalus, Tom Kernan, Ned Lambert,
John Henry Menton, Myles Crawford,
Lenehan, Paddy Leonard, Nosey Flynn, M'Coy
and the featureless face of a Nameless One.)}

\NamelessOne:
Bareback riding.
Weight for age.
Gob, he organised her.

\Jurors:
\stage{(all their heads turned to his voice)}
Really?

\NamelessOne:
\stage{(snarls)}
Arse over tip.
Hundred shillings to five.

\Jurors:
\gab{1150}
\stage{(all their heads lowered in assent)}
Most of us thought as much.

\FirstWatch:
He is a marked man.
Another girl's plait cut.
Wanted:
Jack the Ripper.
A thousand pounds reward.

\SecondWatch:
\stage{(awed, whispers)}
And in black.
A mormon.
Anarchist.

\Crier:
\stage{(loudly)}
Whereas Leopold Bloom of no fixed abode is a wellknown dynamitard,
forger, bigamist, bawd and cuckold
and a public nuisance to the citizens of Dublin
\gab{1160}
and whereas at this commission of assizes the most honourable ...

\stage{(His Honour, sir Frederick Falkiner, recorder of Dublin,
in judicial garb of grey stone rises from the bench, stonebearded.
He bears in his arms an umbrella sceptre.
From his forehead arise starkly the Mosaic ramshorns.)}

\Recorder:
I will put an end to this white slave traffic
and rid Dublin of this odious pest.
Scandalous!
\stage{(he dons the black cap)}
Let him be taken, Mr~Subsheriff, from the dock where he now stands
and detained in custody in Mountjoy prison during His Majesty's pleasure
\gab{1170}
and there be hanged by the neck until he is dead
and therein fail not at your peril or may the Lord have mercy on your soul.
Remove him.
\stage{(a black skullcap descends upon his head.)}

\stage{(the subsheriff Long John Fanning appears, smoking a pungent Henry Clay.)}

\LongJohn:
\stage{(scowls and calls with rich rolling utterance)}
Who'll hang Judas Iscariot?

\stage{(H. Rumbold, master barber, in a bloodcoloured jerkin and tanner's apron,
a rope coiled over his shoulder, mounts the block.
A life preserver and a nailstudded bludgeon are stuck in his belt.
He rubs grimly his grappling hands,
\gab{1180}
knobbed with knuckledusters.)}

\Rumbold:
\stage{(to the recorder with sinister familiarity)}
Hanging Harry, your Majesty, the Mersey terror.
Five guineas a jugular.
Neck or nothing.

\stage{(The bells of George's church toll slowly, loud dark iron.)}

\Bells:
Heigho! Heigho!

\Bloom:
\stage{(desperately)}
Wait. Stop.
Gulls.
Good heart.
I saw.
Innocence.
Girl in the monkeyhouse.
Zoo.
Lewd chimpanzee.
\stage{(breathlessly)}
Pelvic basin.
Her artless blush unmanned me.
\stage{(overcome with emotion)}
\gab{1190}
I left the precincts.
\stage{(he turns to a figure in the crowd, appealing)}
Hynes, may I speak to you?
You know me.
That three shillings you can keep.
If you want a little more ...

\Hynes:
\stage{(coldly)}
You are a perfect stranger.

\SecondWatch:
\stage{(points to the corner)}
The bomb is here.

\FirstWatch:
Infernal machine with a time fuse.

\Bloom:
\gab{1200}
No, no.
Pig's feet.
I was at a funeral.

\FirstWatch:
\stage{(draws his truncheon)} Liar!

\stage{(The beagle lifts his snout, showing the grey scorbutic face of Paddy Dignam.
He has gnawed all.
He exhales a putrid carcasefed breath.
He grows to human size and shape.
His dachshund coat becomes a brown mortuary habit.
His green eye flashes bloodshot.
Half of one ear, all the nose and both thumbs are ghouleaten.)}

\Dignam:
\stage{(in a hollow voice)}
It is true.
It was my funeral.
\gab{1210}
Doctor Finucane pronounced life extinct
when I succumbed to the disease from natural causes.

\stage{(He lifts his mutilated ashen face moonwards and bays lugubriously.)}

\Bloom:
\stage{(in triumph)}
You hear?

\Dignam:
Bloom, I am Paddy Dignam's spirit.
List, list, O list!

\Bloom:
The voice is the voice of Esau.
\gab{1220}

\SecondWatch:
\stage{(blesses himself)}
How is that possible?

\FirstWatch:
It is not in the penny catechism.

\Dignam:
By metempsychosis. Spooks.

\Voice:
O rocks.

\Dignam:
\stage{(earnestly)}
\gab{1230}
Once I was in the employ of Mr~J.H.~Menton,
solicitor, commissioner for oaths and affidavits, of 27~Bachelor's Walk.
Now I am defunct, the wall of the heart hypertrophied.
Hard lines.
The poor wife was awfully cut up.
How is she bearing it?
Keep her off that bottle of sherry.
\stage{(he looks round him)}
A lamp.
I must satisfy an animal need.
That buttermilk didn't agree with me.

\stage{(The portly figure of John O'Connell, caretaker, stands forth,
holding a bunch of keys tied with crape.
Beside him stands Father Coffey, chaplain,
toadbellied, wrynecked, in a surplice and bandanna nightcap,
holding sleepily a staff of twisted poppies.)}

\FrCoffey:
\gab{1240}
\stage{(yawns, then chants with a hoarse croak)}
Namine.
Jacobs.
Vobiscuits.
Amen.

\OConnell:
\stage{(foghorns stormily through his megaphone)}
Dignam, Patrick~T, deceased.

\Dignam:
\stage{(with pricked up ears, winces)}
Overtones.
\stage{(he wriggles forward and places an ear to the ground)}
My master's voice!

\OConnell:
Burial docket letter number U.P. eightyfive thousand.
Field seventeen.
House of Keys.
\gab{1250}
Plot, one hundred and one.

\stage{(Paddy Dignam listens with visible effort, thinking,
his tail stiffpointed, his ears cocked.)}

\Dignam:
Pray for the repose of his soul.

\stage{(He worms down through a coalhole,
his brown habit trailing its tether over rattling pebbles.
After him toddles an obese grandfather rat
on fungus turtle paws under a grey carapace.
Dignam's voice, muffled, is heard baying under ground:
\emph{Dignam's dead and gone below.}
Tom Rochford, robinredbreasted, in cap and breeches,
\gab{1260}%
jumps from his twocolumned machine.)}

\Rochford:
\stage{(a hand to his breastbone, bows)}
Reuben J.
A florin I find him.
\stage{(he fixes the manhole with a resolute stare)}
My turn now on.
Follow me up to Carlow.

\stage{(He executes a daredevil salmon leap in the air and is engulfed in the coalhole.
Two discs on the columns wobble, eyes of nought.
All recedes.
Bloom plodges forward again through the sump.
Kisses chirp amid the rifts of fog a piano sounds.
He stands before a lighted house, listening.
The kisses, winging from their bowers fly about him,
\gab{1270}
twittering, warbling, cooing.)}

\Kisses:
\stage{(warbling)} Leo!
\stage{(twittering)} Icky licky micky sticky for Leo!
\stage{(cooing)} Coo coocoo!
Yummyyum, Womwom!
\stage{(warbling)} Big comebig!
Pirouette!
Leopopold!
\stage{(twittering)} Leeolee!
\stage{(warbling)} O Leo!

\stage{(They rustle, flutter upon his garments, alight, bright giddy flecks, silvery sequins.)}

\Bloom:
A man's touch.
Sad music.
Church music.
Perhaps here.

\stage{(Zoe Higgins, a young whore in a sapphire slip,
closed with three bronze buckles, a slim black velvet fillet round her throat,
\gab{1280}
nods, trips down the steps and accosts him.)}

\Zoe:
Are you looking for someone?
He's inside with his friend.

\Bloom:
Is this Mrs~Mack's?

\Zoe:
No, eightyone.
Mrs~Cohen's.
You might go farther and fare worse.
Mother Slipperslapper.
\stage{(familiarly)}
She's on the job herself tonight with the vet
her tipster that gives her all the winners and pays for her son in Oxford.
Working overtime but her luck's turned today.
\gab{1290}
\stage{(suspiciously)}
You're not his father, are you?

\Bloom:
Not I!

\Zoe:
You both in black.
Has little mousey any tickles tonight?

\stage{(His skin, alert, feels her fingertips approach.
A hand glides over his left thigh.)}

\Zoe:
How's the nuts?

\Bloom:
\gab{1300}
Off side.
Curiously they are on the right.
Heavier, I suppose.
One in a million my tailor, Mesias, says.

\Zoe:
\stage{(in sudden alarm)}
You've a hard chancre.

\Bloom:
Not likely.

\Zoe:
I feel it.

\stage{(Her hand slides into his left trouser pocket
and brings out a hard black shrivelled potato.
\gab{1310}
She regards it and Bloom with dumb moist lips.)}

\Bloom:
A talisman.
Heirloom.

\Zoe:
For Zoe?
For keeps?
For being so nice, eh?

\stage{(She puts the potato greedily into a pocket then links his arm,
cuddling him with supple warmth.
He smiles uneasily.
Slowly, note by note, oriental music is played.
He gazes in the tawny crystal of her eyes, ringed with kohol.
His smile softens.)}

\Zoe:
\gab{1320}
You'll know me the next time.

\Bloom:
\stage{(forlornly)}
I never loved a dear gazelle but it was sure to ...

\stage{(Gazelles are leaping, feeding on the mountains.
Near are lakes.
Round their shores file shadows black of cedargroves.
Aroma rises, a strong hairgrowth of resin.
It burns, the orient, a sky of sapphire, cleft by the bronze flight of eagles.
Under it lies the womancity nude, white, still, cool, in luxury.
A fountain murmurs among damask roses.
Mammoth roses murmur of scarlet winegrapes.
A wine of shame, lust, blood exudes, strangely murmuring.)%
\gab{1330}
}

\Zoe:
\stage{(Murmuring singsong with the music, her odalisk lips lusciously
smeared with salve of swinefat and rosewater)}
\hebrew{Schorach ani wenowach, benoith Hierushaloim.}

\Bloom:
\stage{(fascinated)}
I thought you were of good stock by your accent.

\Zoe:
And you know what thought did?

\stage{(She bites his ear gently with little goldstopped teeth,
sending on him a cloying breath of stale garlic.
\gab{1340}
The roses draw apart, disclose a sepulchre of the gold of kings
and their mouldering bones.)}

\Bloom:
\stage{(draws back, mechanically caressing her right bub with a flat awkward hand)}
Are you a Dublin girl?

\Zoe:
\stage{(catches a stray hair deftly and twists it to her coil)}
No bloody fear.
I'm English.
Have you a swaggerroot?

\Bloom:
\stage{(as before)}
Rarely smoke, dear.
Cigar now and then.
Childish device.
\stage{(lewdly)}
\gab{1350}
The mouth can be better engaged than with a cylinder of rank weed.

\Zoe:
Go on.
Make a stump speech out of it.

\Bloom:
\stage{(in workman's corduroy overalls, black gansy with red floating tie and apache cap)}
Mankind is incorrigible.
Sir Walter Ralegh brought from the new world that potato and that weed,
the one a killer of pestilence by absorption,
the other a poisoner of the ear, eye, heart, memory, will understanding, all.
That is to say he brought the poison a hundred years
before another person whose name I forget brought the food.
\gab{1360}
Suicide. Lies. All our habits.
Why, look at our public life!

\stage{(Midnight chimes from distant steeples.)}

\Chimes:
Turn again, Leopold!
Lord mayor of Dublin!

\Bloom:
\stage{(in alderman's gown and chain)}
Electors of Arran Quay, Inns Quay, Rotunda, Mountjoy and North Dock,
better run a tramline, I say, from the cattlemarket to the river.
That's the music of the future.
That's my programme.
\latin{Cui bono?}
But our bucaneering Vanderdeckens in their phantom ship of finance ...%
\gab{1370}

\Elector:
Three times three for our future chief magistrate!

\stage{(The Aurora Borealis of the torchlight procession leaps.)}

\Torchbearers:
Hooray!

\stage{(Several wellknown burgesses, city magnates and freemen of the city
shake hands with Bloom and congratulate him.
Timothy Harrington, late thrice Lord Mayor of Dublin,
imposing in mayoral scarlet, gold chain and white silk tie,
confers with councillor lorcan sherlock, \latin{locum tenens}.
\gab{1380}
They nod vigorously in agreement.)}

\Harrington:
\stage{(in scarlet robe with mace, gold mayoral chain and large white silk scarf)}
That alderman sir Leo Bloom's speech be printed at the expense of the ratepayers.
That the house in which he was born be ornamented with a commemorative tablet
and that the thoroughfare hitherto known as Cow Parlour off Cork street
be henceforth designated Boulevard Bloom.

\Lorcan:
Carried unanimously.

\Bloom:
\stage{(impassionedly)}
\gab{1390}
These flying Dutchmen or lying Dutchmen as they recline
in their upholstered poop, casting dice, what reck they?
Machines is their cry, their chimera, their panacea.
Laboursaving apparatuses, supplanters, bugbears,
manufactured monsters for mutual murder,
hideous hobgoblins produced by a horde of capitalistic lusts upon our prostituted labour.
The poor man starves while they are grassing their royal mountain stags
or shooting peasants and phartridges in their purblind pomp of pelf and power.
But their reign is rover for rever and ever and ev ...

\stage{(Prolonged applause.
Venetian masts, maypoles and festal arches spring up.
A streamer bearing the legends \irish{Cead Mile Failte} and
\hebrew{Mah Ttob Melek Israel} spans the street.
\gab{1400}
All the windows are thronged with sightseers, chiefly ladies.
Along the route the regiments of the Royal Dublin Fusiliers,
the King's Own Scottish Borderers, the Cameron Highlanders
and the Welsh Fusiliers standing to attention, keep back the crowd.
Boys from High school are perched on the lampposts, telegraph poles,
windowsills, cornices, gutters, chimneypots, railings, rainspouts,
whistling and cheering.
The pillar of the cloud appears.
A fife and drum band is heard in the distance playing the Kol Nidre.
The beaters approach with imperial eagles hoisted,
trailing banners and waving oriental palms.
\gab{1410}
The chryselephantine papal standard rises high,
surrounded by pennons of the civic flag.
The van of the procession appears headed by John Howard Parnell, city marshal,
in a chessboard tabard, the Athlone poursuivant and Ulster King of Arms.
They are followed by the Right Honourable Joseph Hutchinson,
lord mayor of Dublin, his lordship the lord mayor of Cork,
their worships the mayors of Limerick, Galway, Sligo and Waterford,
twentyeight Irish representative peers, sirdars, grandees and maharajahs
bearing the cloth of estate, the Dublin Metropolitan Fire Brigade,
the chapter of the saints of finance in their plutocratic order of precedence,
the bishop of Down and Connor,
\gab{1420}
His Eminence Michael cardinal Logue, archbishop of Armagh, primate of all Ireland,
His Grace, the most reverend Dr William Alexander,
archbishop of Armagh, primate of all Ireland,
the chief rabbi, the presbyterian moderator,
the heads of the baptist, anabaptist, methodist and Moravian chapels
and the honorary secretary of the society of friends.
After them march the guilds and trades and trainbands with flying colours:
coopers, bird fanciers, millwrights, newspaper canvassers, law scriveners,
masseurs, vintners, trussmakers, chimneysweeps, lard refiners,
tabinet and poplin weavers, farriers, Italian warehousemen,
\gab{1430}
church decorators, bootjack manufacturers, undertakers,
silk mercers, lapidaries, salesmasters, corkcutters,
assessors of fire losses, dyers and cleaners, export bottlers,
fellmongers, ticketwriters, heraldic seal engravers,
horse repository hands, bullion brokers, cricket and archery outfitters,
riddlemakers, egg and potato factors, hosiers and glovers, plumbing contractors.
After them march gentlemen of the bedchamber,
Black Rod, Deputy Garter, Gold Stick, the master of horse,
the lord great chamberlain, the earl marshal,
the high constable carrying the sword of state,
saint Stephen's iron crown, the chalice and bible.
\gab{1440}
Four buglers on foot blow a sennet.
Beefeaters reply, winding clarions of welcome.
Under an arch of triumph Bloom appears, bareheaded,
in a crimson velvet mantle trimmed with ermine,
bearing Saint Edward's staff the orb and sceptre with the dove, the curtana.
He is seated on a milkwhite horse with long flowing crimson tail,
richly caparisoned, with golden headstall.
Wild excitement.
The ladies from their balconies throw down rosepetals.
The air is perfumed with essences.
The men cheer.
Bloom's boys run amid the bystanders with branches of hawthorn and wrenbushes.)}

\BloomsBoys:
\gab{1450}
\begin{verse}
    The wren, the wren,\\
    The king of all birds,\\
    Saint Stephen's his day\\
    Was caught in the furze.
\end{verse}


\Blacksmith \stage{(murmurs)}
For the honour of God!
And is that Bloom?
He scarcely looks thirtyone.

\Flagger:
That's the famous Bloom now, the world's greatest reformer.
Hats off!

\stage{(All uncover their heads.
\gab{1460}
Women whisper eagerly.)}

\Millionairess:
\stage{(richly)}
Isn't he simply wonderful?

\Noblewoman:
\stage{(nobly)}
All that man has seen!

\Feminist:
\stage{(masculinely)}
And done!

\Bellhanger:
A classic face!
He has the forehead of a thinker.

\stage{(Bloom's weather.
A sunburst appears in the northwest.)}

\DownConnor:
\gab{1470}
I here present your undoubted emperor-president and king-chairman,
the most serene and potent and very puissant ruler of this realm.
God save Leopold the First!

\All:
God save Leopold the First!

\Bloom:
\stage{(in dalmatic and purple mantle,
to the bishop of Down and Connor, with dignity)}
Thanks, somewhat eminent sir.

\WillArmagh:
\stage{(in purple stock and shovel hat)}
\gab{1480}
Will you to your power cause law and mercy to be executed
in all your judgments in Ireland and territories thereunto belonging?

\Bloom:
\stage{(placing his right hand on his testicles, swears)}
So may the Creator deal with me.
All this I promise to do.

\MikeArmagh:
\stage{(pours a cruse of hairoil over Bloom's head)}
\latin{Gaud\-ium magnum annuntio vobis.
Habemus carneficem.}
Leopold, Patrick, Andrew, David, George, be thou anointed!

\stage{
\gab{1490}
(Bloom assumes a mantle of cloth of gold and puts on a ruby ring.
He ascends and stands on the stone of destiny.
The representative peers put on at the same time their twentyeight crowns.
Joybells ring in Christ church, Saint Patrick's, George's and gay Malahide.
Mirus bazaar fireworks go up from all sides with symbolical phallopyrotechnic designs.
The peers do homage, one by one, approaching and genuflecting.)}

\Peers:
I do become your liege man of life and limb to earthly worship.

\stage{
(Bloom holds up his right hand on which sparkles the Koh-i-Noor diamond.
\gab{1500}
His palfrey neighs.
Immediate silence.
Wireless intercontinental and interplanetary transmitters are set for reception of message.)}

\Bloom:
My subjects!
We hereby nominate our faithful charger Copula Felix hereditary Grand Vizier
and announce that we have this day repudiated our former spouse
and have bestowed our royal hand upon the princess Selene,
the splendour of night.

\stage{(The former morganatic spouse of Bloom is hastily removed in the black maria.
The princess Selene, in moonblue robes, a silver crescent on her head,
\gab{1510}
descends from a Sedan chair, borne by two giants.
An outburst of cheering.)}

\Parnell:
\stage{(raises the royal standard)}
Illustrious Bloom!
Successor to my famous brother!

\Bloom:
\stage{(embraces John Howard Parnell)}
We thank you from our heart, John,
for this right royal welcome to green Erin,
the promised land of our common ancestors.

\stage{(The freedom of the city is presented to him embodied in a charter.
The keys of Dublin, crossed on a crimson cushion, are given to him.%
\gab{1520}
He shows all that he is wearing green socks.)}

\Kernan:
You deserve it, your honour.

\Bloom:
On this day twenty years ago we overcame the hereditary enemy at Ladysmith.
Our howitzers and camel swivel guns played on his lines with telling effect.
Half a league onward!
They charge!
All is lost now!
Do we yield?
No!
We drive them headlong!
Lo! We charge!
Deploying to the left our light horse swept across the heights of Plevna
and, uttering their warcry \latin{Bonafide Sabaoth},
\gab{1530}
sabred the Saracen gunners to a man.

\Typesetters:
Hear! Hear!

\JohnWyse:
There's the man that got away James Stephens.

\Bluecoat:
Bravo!

\OldResident:
You're a credit to your country, sir, that's what you are.

\AppleWoman:
He's a man like Ireland wants.
\gab{1540}

\Bloom:
My beloved subjects, a new era is about to dawn.
I, Bloom, tell you verily it is even now at hand.
Yea, on the word of a Bloom, ye shall ere long
enter into the golden city which is to be, the new Bloomusalem
in the Nova Hibernia of the future.

\stage{(Thirtytwo workmen, wearing rosettes, from all the counties of Ireland,
under the guidance of Derwan the builder, construct the new Bloomusalem.
It is a colossal edifice with crystal roof,
built in the shape of a huge pork kidney, containing forty thousand rooms.
In the course of its extension
\gab{1550}
several buildings and monuments are demolished.
Government offices are temporarily transferred to railway sheds.
Numerous houses are razed to the ground.
The inhabitants are lodged in barrels and boxes,
All marked in red with the letters:
L.B.
Several paupers fill from a ladder.
A part of the walls of Dublin, crowded with loyal sightseers, collapses.)}

\Sightseers:
\stage{(dying)}
\latin{Morituri te salutant.}
\stage{(they die)}

\stage{(A man in a brown macintosh springs up through a trapdoor.
He points an elongated finger at Bloom.)}

\Macintosh:
\gab{1560}
Don't you believe a word he says.
That man is Leopold M'Intosh, the notorious fireraiser.
His real name is Higgins.

\Bloom:
Shoot him!
Dog of a christian!
So much for M'Intosh!

\stage{(A cannonshot.
The man in the macintosh disappears.
Bloom with his sceptre strikes down poppies.
The instantaneous deaths of many powerful enemies,
graziers, members of parliament, members of standing committees, are reported.
Bloom's bodyguard distribute maundy money, commemoration medals,
loaves and fishes, temperance badges, expensive Henry Clay cigars,
\gab{1570}
free cowbones for soup, rubber preservatives in sealed envelopes
tied with gold thread, butter scotch, pineapple rock,
\french{billets doux} in the form of cocked hats,
readymade suits, porringers of toad in the hole,
bottles of Jeyes' Fluid, purchase stamps, 40 days' indulgences,
spurious coins, dairyfed pork sausages, theatre passes,
season tickets available for all tramlines,
coupons of the royal and privileged Hungarian lottery, penny dinner counters,
cheap reprints of the World's Twelve Worst Books:
Froggy and Fritz (politic),
Care of the Baby (infantilic),
50 Meals for 7/6 (culinic),
Was Jesus a Sun Myth? (historic),
\gab{1580}
Expel That Pain (medic),
Infant's Compendium of the Universe (cosmic),
Let's All Chortle (hilaric),
Canvasser's Vade Mecum (journalic),
Loveletters of Mother Assistant (erotic),
Who's Who in Space (astric),
Songs that Reached Our Heart (melodic),
Pennywise's Way to Wealth (parsimonic).
A general rush and scramble.
Women press forward to touch the hem of Bloom's robe.
The lady Gwendolen Dubedat bursts through the throng, leaps on his horse
and kisses him on both cheeks amid great acclamation.
A magnesium flashlight photograph is taken.
Babes and sucklings are held up.)}

\Women:
\gab{1590}
Little father!
Little father!

\Sucklings:
\begin{verse}
    Clap clap hands till Poldy comes home,\\
    Cakes in his pocket for Leo alone.
\end{verse}

\stage{(Bloom, bending down, pokes Baby Boardman gently in the stomach.)}

\BabyBoardman:
\stage{(hiccups, curdled milk flowing from his mouth)}
Hajajaja.

\Bloom:
\stage{(shaking hands with a blind stripling)}
\gab{1600}
My more than Brother!
\stage{(placing his arms round the shoulders of an old couple)}
Dear old friends!
\stage{(he plays pussy fourcorners with ragged boys and girls)}
Peep! Bopeep!
\stage{(he wheels twins in a perambulator)}
Ticktacktwo wouldyousetashoe?
\stage{(he performs juggler's tricks, draws red, orange, yellow,
green, blue, indigo and violet silk handkerchiefs from his mouth)}
Roygbiv.
32 feet per second.
\stage{(he consoles a widow)}
Absence makes the heart grow younger.
\stage{(he dances the highland fling with grotesque antics)}
Leg it, ye devils!
\stage{(he kisses the bedsores of a palsied veteran)}
Honourable wounds!
\stage{(he trips up a fit policeman)}
U.P:
up. U.P: up.
\stage{(he whispers in the ear of a blushing waitress and laughs kindly)}
Ah, naughty, naughty!
\gab{1610}
\stage{(he eats a raw turnip offered him by Maurice Butterly, farmer)}
Fine! Splendid!
\stage{(he refuses to accept three shillings offered him by Joseph Hynes, journalist)}
My dear fellow, not at all!
\stage{(he gives his coat to a beggar)}
Please accept.
\stage{(he takes part in a stomach race with elderly male and female cripples)}
Come on, boys!
Wriggle it, girls!

\Citizen:
\stage{(choked with emotion, brushes aside a tear in his emerald muffler)}
May the good God bless him!

\stage{(The rams' horns sound for silence.
The standard of Zion is hoisted.)%
\gab{1620}
}

\Bloom:
\stage{(uncloaks impressively, revealing obesity, unrolls a paper and reads solemnly)}
\hebrew{Aleph Beth Ghimel Daleth Hagadah Tephilim Kosher Yom Kippur
Hanukah Roschaschana Beni Brith Bar Mitzvah Mazzoth Askenazim
Meshuggah Talith.}

\stage{(An official translation is read by Jimmy Henry, assistant town clerk.)}

\JimmyHenry:
The Court of Conscience is now open.
His Most Catholic Majesty will now administer open air justice.
\gab{1630}
Free medical and legal advice, solution of doubles and other problems.
All cordially invited.
Given at this our loyal city of Dublin in the year~I of the Paradisiacal Era.

\PaddyLeonard:
What am I to do about my rates and taxes?

\Bloom:
Pay them, my friend.

\PaddyLeonard:
Thank you.

\NoseyFlynn:
Can I raise a mortgage on my fire insurance?%
\gab{1640}

\Bloom:
\stage{(obdurately)}
Sirs, take notice that by the law of torts you are bound over
in your own recognisances for six months in the sum of five pounds.

\JJOM:
A Daniel did I say?
Nay! A Peter O'Brien!

\NoseyFlynn:
Where do I draw the five pounds?

\PisserBurke:
For bladder trouble?

\Bloom:
\gab{1650}
\begin{verse}
\latin{Acid. nit. hydrochlor. dil.}, 20 minims\\
\latin{Tinct. nux vom.}, 5 minims\\
\latin{Extr. taraxel. iiq.}, 30 minims.\\
\latin{Aq. dis. ter in die.}
\end{verse}

\ChrisCallinan:
What is the parallax of the subsolar ecliptic of Aldebaran?

\Bloom:
Pleased to hear from you, Chris. K.~II.

\JoeHynes:
Why aren't you in uniform?%
\gab{1660}

\Bloom:
When my progenitor of sainted memory
wore the uniform of the Austrian despot in a dank prison
where was yours?

\BenDollard:
Pansies?

\Bloom:
Embellish (beautify) suburban gardens.

\BenDollard:
When twins arrive?

\Bloom:
\gab{1670}
Father (pater, dad) starts thinking.

\LarryORourke:
An eightday licence for my new premises.
You remember me, sir Leo, when you were in number seven.
I'm sending around a dozen of stout for the missus.

\Bloom:
\stage{(coldly)}
You have the advantage of me.
Lady Bloom accepts no presents.

\Crofton:
This is indeed a festivity.

\Bloom:
\gab{1680}
\stage{(solemnly)}
You call it a festivity.
I call it a sacrament.

\AlexanderKeyes:
When will we have our own house of keys?

\Bloom:
I stand for the reform of municipal morals and the plain ten commandments.
New worlds for old.
Union of all, jew, moslem and gentile.
Three acres and a cow for all children of nature.
Saloon motor hearses.
Compulsory manual labour for all.
All parks open to the public day and night.
Electric dishscrubbers.
Tuberculosis, lunacy, war and mendicancy must now cease.
\gab{1690}
General amnesty, weekly carnival with masked licence,
bonuses for all, esperanto the universal language with universal brotherhood.
No more patriotism of barspongers and dropsical impostors.
Free money, free rent, free love and a free lay church in a free lay state.

\OMaddenBurke:
Free fox in a free henroost.

\DavyByrne:
\stage{(yawning)} Iiiiiiiiiaaaaaaach!

\Bloom:
Mixed races and mixed marriage.

\Lenehan:
\gab{1700}
What about mixed bathing?

\stage{(Bloom explains to those near him his schemes for social regeneration.
All agree with him.
The keeper of the Kildare street museum appears,
dragging a lorry on which are the shaking statues of several naked goddesses,
Venus Callipyge, Venus Pandemos, Venus Metempsychosis,
and plaster figures, also naked, representing the new nine muses,
Commerce, Operatic Music, Amor, Publicity, Manufacture, Liberty of Speech,
Plural Voting, Gastronomy, Private Hygiene, Seaside Concert Entertainments,
Painless Obstetrics
\gab{1710}
and Astronomy for the People.)}

\FatherFarley:
He is an episcopalian, an agnostic, an anythingarian
seeking to overthrow our holy faith.

\MrsRiordan:
\stage{(tears up her will)}
I'm disappointed in you! You bad man!

\MotherGrogan:
\stage{(removes her boot to throw it at Bloom)}
You beast!
You abominable person!

\NoseyFlynn:
Give us a tune, Bloom.
\gab{1720}
One of the old sweet songs.

\Bloom:
\stage{(with rollicking humour)}
\begin{verse}
    I vowed that I never would leave her,\\
    She turned out a cruel deceiver.\\
    With my tooraloom tooraloom tooraloom tooraloom.
\end{verse}

\HoppyHolohan:
Good old Bloom!
There's nobody like him after all.

\PaddyLeonard:
Stage Irishman!

\Bloom:
\gab{1730}
What railway opera is like a tramline in Gibraltar?
The Rows of Casteele.
\stage{(laughter.)}

\Lenehan:
Plagiarist!
Down with Bloom!

\VeiledSibyl:
\stage{(enthusiastically)}
I'm a Bloomite and I glory in it.
I believe in him in spite of all.
I'd give my life for him, the funniest man on earth.

\Bloom:
\stage{(winks at the bystanders)}
I bet she's a bonny lassie.

\TheodorePurefoy:
\gab{1740}
\stage{(in fishingcap and oilskin jacket)}
He employs a mechanical device to frustrate the sacred ends of nature.

\VeiledSibyl:
\stage{(stabs herself)}
My hero god!
\stage{(she dies)}

\stage{(Many most attractive and enthusiastic women also commit suicide
by stabbing, drowning, drinking prussic acid, aconite, arsenic,
opening their veins, refusing food, casting themselves under steamrollers,
from the top of Nelson's Pillar, into the great vat of Guinness's brewery,
asphyxiating themselves by placing their heads in gasovens,
hanging themselves in stylish garters,
\gab{1750}
leaping from windows of different storeys.)}

\Dowie:
\stage{(violently)}
Fellowchristians and antiBloomites,
the man called Bloom is from the roots of hell,
a disgrace to christian men.
A fiendish libertine from his earliest years
this stinking goat of Mendes
gave precocious signs of infantile debauchery,
recalling the cities of the plain, with a dissolute granddam.
This vile hypocrite, bronzed with infamy,
is the white bull mentioned in the Apocalypse.
A worshipper of the Scarlet Woman, intrigue is the very breath of his nostrils.
The stake faggots and the caldron of boiling oil are for him.
\gab{1760}
Caliban!

\Mob:
Lynch him!
Roast him!
He's as bad as Parnell was.
Mr~Fox!

\stage{(Mother Grogan throws her boot at Bloom.
Several shopkeepers from upper and lower Dorset street
throw objects of little or no commercial value,
hambones, condensed milk tins, unsaleable cabbage, stale bread,
sheep's tails, odd pieces of fat.)}

\Bloom:
\stage{(excitedly)}
This is midsummer madness, some ghastly joke again.
By heaven, I am guiltless as the unsunned snow!
It was my brother Henry.
He is my double.
\gab{1770}
He lives in number 2 Dolphin's Barn.
Slander, the viper, has wrongfully accused me.
Fellowcountrymen, \irish{sgeul i mbarr bata coisde gan capall}.
I call on my old friend, Dr Malachi Mulligan, sex specialist,
to give medical testimony on my behalf.

\DrMulligan:
\stage{(in motor jerkin, green motorgoggles on his brow)}
Dr Bloom is bisexually abnormal.
He has recently escaped from Dr Eustace's
private asylum for demented gentlemen.
Born out of bedlock hereditary epilepsy is present,
the consequence of unbridled lust.
Traces of elephantiasis have been discovered among his ascendants.
There are marked symptoms of chronic exhibitionism.
\gab{1780}
Ambidexterity is also latent.
He is prematurely bald from selfabuse,
perversely idealistic in consequence,
a reformed rake, and has metal teeth.
In consequence of a family complex
he has temporarily lost his memory
and I believe him to be more sinned against than sinning.
I have made a pervaginal examination
and, after application of the acid test to
5427 anal, axillary, pectoral and pubic hairs,
I declare him to be \latin{virgo intacta}.

\stage{(Bloom holds his high grade hat over his genital organs.)}

\DrMadden:
Hypsospadia is also marked.
In the interest of coming generations
I suggest that the parts affected should be preserved
\gab{1790}
in spirits of wine in the national teratological museum.

\DrCrotthers:
I have examined the patient's urine.
It is albuminoid.
Salivation is insufficient, the patellar reflex intermittent.

\DrPunchCostello:
The \latin{fetor judaicus} is most perceptible.

\DrDixon:
\stage{(reads a bill of health)}
Professor Bloom is a finished example of the new womanly man.
His moral nature is simple and lovable.
Many have found him a dear man, a dear person.
\gab{1800}
He is a rather quaint fellow on the whole,
coy though not feebleminded in the medical sense.
He has written a really beautiful letter, a poem in itself,
to the court missionary of the Reformed Priests' Protection Society
which clears up everything.
He is practically a total abstainer
and I can affirm that he sleeps on a straw litter
and eats the most Spartan food, cold dried grocer's peas.
He wears a hairshirt of pure Irish manufacture winter and summer
and scourges himself every Saturday.
He was, I understand, at one time
a firstclass misdemeanant in Glencree reformatory.
Another report states that he was a very posthumous child.
I appeal for clemency in the name of
the most sacred word our vocal organs have ever been called upon to speak.
\gab{1810}
He is about to have a baby.

\stage{(General commotion and compassion.
Women faint.
A wealthy American makes a street collection for Bloom.
Gold and silver coins, blank cheques,
banknotes, jewels, treasury bonds, maturing bills of exchange, I.O.U's,
wedding rings, watchchains, lockets,
necklaces and bracelets are rapidly collected.)}

\Bloom:
O, I so want to be a mother.

\MrsThornton:
\stage{(in nursetender's gown)}
Embrace me tight, dear.
You'll be soon over it.
Tight, dear.%
\gab{1820}

\stage{(Bloom embraces her tightly and bears eight male yellow and white children.
They appear on a redcarpeted staircase adorned with expensive plants.
All the octuplets are handsome, with valuable metallic faces,
wellmade, respectably dressed and wellconducted,
speaking five modern languages fluently
and interested in various arts and sciences.
Each has his name printed in legible letters on his shirtfront:
Nasodoro, Goldfinger, Chrysostomos, Maindoree,
Silversmile, Silberselber, Vifargent, Panargyros.
They are immediately appointed to positions of
high public trust in several different countries
\gab{1830}
as managing directors of banks, traffic managers of railways,
chairmen of limited liability companies, vicechairmen of hotel syndicates.)}

\Voice:
Bloom, are you the Messiah ben Joseph or ben David?

\Bloom:
\stage{(darkly)}
You have said it.

\BrotherBuzz:
Then perform a miracle like Father Charles.

\BantamLyons:
Prophesy who will win the Saint Leger.%
\gab{1840}

\stage{(Bloom walks on a net, covers his left eye with his left ear,
passes through several walls, climbs Nelson's Pillar,
hangs from the top ledge by his eyelids,
eats twelve dozen oysters (shells included),
heals several sufferers from king's evil,
contracts his face so as to resemble many historical personages,
Lord Beaconsfield, Lord Byron, Wat Tyler,
Moses of Egypt, Moses Maimonides, Moses Mendelssohn, Henry Irving,
Rip van Winkle, Kossuth, Jean Jacques Rousseau, Baron Leopold Rothschild,
Robinson Crusoe, Sherlock Holmes, Pasteur,
turns each foot simultaneously in different directions,
\gab{1850}
bids the tide turn back,
eclipses the sun by extending his little finger.)}

\Brini:
\stage{(in papal zouave's uniform, steel cuirasses as breastplate,
armplates, thighplates, legplates, large profane moustaches and brown paper mitre)}
\latin{Leopoldi autem generatio.}
Moses begat Noah
and Noah begat Eunuch
and Eunuch begat O'Halloran
and O'Halloran begat Guggenheim
and Guggenheim begat Agendath
and Agendath begat Netaim
and Netaim begat Le Hirsch
and Le Hirsch begat Jesurum
and Jesurum begat MacKay
and MacKay begat Ostrolopsky
and Ostrolopsky begat Smerdoz
and Smerdoz begat Weiss
\gab{1860}
and Weiss begat Schwarz
and Schwarz begat Adrianopoli
and Adrianopoli begat Aranjuez
and Aranjuez begat Lewy Lawson
and Lewy Lawson begat Ichabudonosor
and Ichabudonosor begat O'Donnell Magnus
and O'Donnell Magnus begat Christbaum
and Christbaum begat ben Maimun
and ben Maimun begat Dusty Rhodes
and Dusty Rhodes begat Benamor
and Benamor begat Jones-Smith
and Jones-Smith begat Savorgnanovich
and Savorgnanovich begat Jasperstone
and Jasperstone begat Vingtetunieme
and Vingtetunieme begat Szombathely
and Szombathely begat Virag
and Virag begat Bloom
\latin{et vocabitur nomen eius Emmanuel}.

\Deadhand:
\gab{1870}
\stage{(writes on the wall)}
Bloom is a cod.

\Crab:
\stage{(in bushranger's kit)}
What did you do in the cattlecreep behind Kilbarrack?

\FemaleInfant:
\stage{(shakes a rattle)}
And under Ballybough bridge?

\Hollybush:
And in the devil's glen?

\Bloom:
\stage{(blushes furiously all over from frons to nates,
\gab{1880}
three tears filling from his left eye)}
Spare my past.

\EvictedTenants:
\stage{(in bodycoats, kneebreeches, with donnybrook fair shillelaghs)}
Sjambok him!

\stage{(Bloom with asses' ears seats himself in the pillory with crossed arms,
his feet protruding.
He whistles \italian{Don Giovanni, a cenar teco}.
Artane orphans, joining hands, caper round him.
Girls of the Prison Gate Mission, joining hands,
caper round in the opposite direction.)}

\ArtaneOrphans:
\begin{verse}
\gab{1890}%
    You hig, you hog, you dirty dog!\\
    You think the ladies love you!
\end{verse}

\PrisonGateGirls:
\begin{verse}
    If you see Kay\\
    Tell him he may\\
    See you in tea\\
    Tell him from me.
\end{verse}

\Hornblower:
\stage{(in ephod and huntingcap, announces)}
And he shall carry the sins of the people to Azazel,
the spirit which is in the wilderness,
and to Lilith, the nighthag.
And they shall stone him and defile him, yea,
\gab{1900}
all from Agendath Netaim and from Mizraim, the land of Ham.

\stage{(All the people cast soft pantomime stones at Bloom.
Many bonafide travellers and ownerless dogs come near him and defile him.
Mastiansky and Citron approach in gaberdines, wearing long earlocks.
They wag their beards at Bloom.)}

\MastianskyCitron:
Belial!
Laemlein of Istria, the false Messiah!
Abulafia!
Recant!

\stage{(George R Mesias, Bloom's tailor, appears,
a tailor's goose under his arm, presenting a bill)}

\Mesias:
\gab{1910}
To alteration one pair trousers eleven shillings.

\Bloom:
\stage{(rubs his hands cheerfully)}
Just like old times.
Poor Bloom!

\stage{(Reuben J Dodd, blackbearded Iscariot, bad shepherd,
bearing on his shoulders the drowned corpse of his son,
approaches the pillory.)%
\gab{1920}
}

\ReubenJ:
\stage{(whispers hoarsely)}
The squeak is out.
A split is gone for the flatties.
Nip the first rattler.

\FireBrigade:
Pflaap!

\BrotherBuzz:
\stage{(Invests Bloom in a yellow habit
with embroidery of painted flames and high pointed hat.
He places a bag of gunpowder round his neck
and hands him over to the civil power, saying)}
Forgive him his trespasses.

\stage{(Lieutenant Myers of the Dublin Fire Brigade
\gab{1930}
by general request sets fire to Bloom.
Lamentations.)}

\Citizen:
Thank heaven!

\Bloom:
\stage{(in a seamless garment marked I.H.S.
stands upright amid phoenix flames)}
Weep not for me, O daughters of Erin.

\stage{(He exhibits to Dublin reporters traces of burning.)}

\stage{(The daughters of Erin, in black garments,
with large prayerbooks and long lighted candles in their hands,
kneel down and pray.)}

\DaughtersErin:
\gab{1940}
\begin{verse}
    Kidney of Bloom, pray for us\\
    Flower of the Bath, pray for us\\
    Mentor of Menton, pray for us\\
    Canvasser for the Freeman, pray for us\\
    Charitable Mason, pray for us\\
    Wandering Soap, pray for us\\
    Sweets of Sin, pray for us\\
    Music without Words, pray for us\\
    Reprover of the Citizen, pray for us\\
\gab{1950}%
    Friend of all Frillies, pray for us\\
    Midwife Most Merciful, pray for us\\
    Potato Preservative against Plague and Pestilence, pray for us.
\end{verse}

\stage{(A choir of six hundred voices, conducted by Vincent O'Brien,
sings the chorus from Handel's Messiah
\Title{Alleluia for the Lord God Omnipotent Reigneth,}
accompanied on the organ by Joseph Glynn.
Bloom becomes mute, shrunken, carbonised.)}

\Zoe:
Talk away till you're black in the face.

\Bloom:
\stage{(in caubeen with clay pipe stuck in the band, dusty brogues,
\gab{1960}
an emigrant's red handkerchief bundle in his hand,
leading a black bogoak pig by a sugaun, with a smile in his eye)}
Let me be going now, woman of the house,
for by all the goats in Connemara
I'm after having the father and mother of a bating.
\stage{(with a tear in his eye)}
All insanity.
Patriotism, sorrow for the dead, music, future of the race.
To be or not to be.
Life's dream is o'er.
End it peacefully.
They can live on.
\stage{(he gazes far away mournfully)}
I am ruined.
A few pastilles of aconite.
The blinds drawn.
A letter.
Then lie back to rest.
\stage{(he breathes softly)}
No more.
I have lived.
Fare. Farewell.

\Zoe:
\stage{(stiffly, her finger in her neckfillet)}
Honest?
Till the next time.
\gab{1970}
\stage{(she sneers)}
Suppose you got up the wrong side of the bed or came too quick with your best girl.
O, I can read your thoughts!

\Bloom:
\stage{(bitterly)}
Man and woman, love, what is it?
A cork and bottle.
I'm sick of it.
Let everything rip.

\Zoe:
\stage{(in sudden sulks)}
I hate a rotter that's insincere.
Give a bleeding whore a chance.

\Bloom:
\stage{(repentantly)}
\gab{1980}
I am very disagreeable.
You are a necessary evil.
Where are you from?
London?

\Zoe:
\stage{(glibly)}
Hog's Norton where the pigs plays the organs.
I'm Yorkshire born.
\stage{(she holds his hand which is feeling for her nipple)}
I say, Tommy Tittlemouse.
Stop that and begin worse.
Have you cash for a short time?
Ten shillings?

\Bloom:
\stage{(smiles, nods slowly)}
More, houri, more.

\Zoe:
And more's mother?
\gab{1990}
\stage{(she pats him offhandedly with velvet paws)}
Are you coming into the musicroom to see our new pianola?
Come and I'll peel off.

\Bloom:
\stage{(feeling his occiput dubiously with the unparalleled embarrassment
of a harassed pedlar gauging the symmetry of her peeled pears)}
Somebody would be dreadfully jealous if she knew.
The greeneyed monster.
\stage{(earnestly)}
You know how difficult it is.
I needn't tell you.

\Zoe:
\stage{(flattered)}
What the eye can't see the heart can't grieve for.
\stage{(she pats him)}
Come.

\Bloom:
\gab{2000}
Laughing witch!
The hand that rocks the cradle.

\Zoe:
Babby!

\Bloom:
\stage{(in babylinen and pelisse,
bigheaded, with a caul of dark hair,
fixes big eyes on her fluid slip
and counts its bronze buckles with a chubby finger,
his moist tongue lolling and lisping)}
One two tlee:
tlee tlwo tlone.

\Buckles:
Love me.
Love me not.
Love me.

\Zoe:
\gab{2010}
Silent means consent.
\stage{(with little parted talons she captures his hand,
her forefinger giving to his palm the passtouch of secret monitor,
luring him to doom.)}
Hot hands cold gizzard.

\stage{(He hesitates amid scents, music, temptations.
She leads him towards the steps,
drawing him by the odour of her armpits,
the vice of her painted eyes,
the rustle of her slip in whose sinuous folds
lurks the lion reek of all the male brutes that have possessed her.)}

\MaleBrutes:
\stage{(exhaling sulphur of rut and dung
and ramping in their loosebox, faintly roaring,
their drugged heads swaying to and fro)}
\gab{2020}
Good!

\stage{(Zoe and Bloom reach the doorway where two sister whores are seated.
They examine him curiously from under their pencilled brows and smile to his hasty bow.
He trips awkwardly.)}

\Zoe:
\stage{(her lucky hand instantly saving him)}
Hoopsa!
Don't fall upstairs.

\Bloom:
The just man falls seven times.
\stage{(he stands aside at the threshold)}
After you is good manners.

\Zoe:
Ladies first, gentlemen after.
\gab{2020}

\stage{(She crosses the threshold.
He hesitates.
She turns and, holding out her hands, draws him over.
He hops.
On the antlered rack of the hall hang a man 's hat and waterproof.
Bloom uncovers himself but, seeing them,
frowns, then smiles, preoccupied.
A door on the return landing is flung open.
A man in purple shirt and grey trousers, brownsocked,
passes with an ape's gait, his bald head and goatee beard upheld,
hugging a full waterjugjar, his twotailed black braces dangling at heels.
Averting his face quickly Bloom bends to examine on the halltable
the spaniel eyes of a running fox:
then, his lifted head sniffing, follows Zoe into the musicroom.
\gab{2040}
A shade of mauve tissuepaper dims the light of the chandelier.
Round and round a moth flies, colliding, escaping.
The floor is covered with an oilcloth mosaic of jade and azure and cinnabar rhomboids.
Footmarks are stamped over it in all senses, heel to heel, heel to hollow,
toe to toe, feet locked, a morris of shuffling feet without body phantoms,
all in a scrimmage higgledypiggledy.
The walls are tapestried with a paper of yewfronds and clear glades.
In the grate is spread a screen of peacock feathers.
Lynch squats crosslegged on the hearthrug of matted hair, his cap back to the front.
With a wand he beats time slowly.
\gab{2050}
Kitty Ricketts, a bony pallid whore in navy costume,
doeskin gloves rolled back from a coral wristlet, a chain purse in her hand,
sits perched on the edge of the table swinging her leg
and glancing at herself in the gilt mirror over the mantelpiece.
A tag of her corsetlace hangs slightly below her jacket.
Lynch indicates mockingly the couple at the piano.)}

\Kitty:
\stage{(coughs behind her hand)}
She's a bit imbecillic.
\stage{(she signs with a waggling forefinger)}
Blemblem.
\stage{(Lynch lifts up her skirt and white petticoat with his wand
she settles them down quickly.)}
Respect yourself.
\stage{(she hiccups, then bends quickly her sailor hat
\gab{2060}
under which her hair glows, red with henna)}
O, excuse!

\Zoe:
More limelight, Charley.
\stage{(she goes to the chandelier and turns the gas full cock)}

\Kitty:
\stage{(peers at the gasjet)}
What ails it tonight?

\Lynch:
\stage{(deeply)}
Enter a ghost and hobgoblins.

\Zoe:
Clap on the back for Zoe.
\gab{2070}

\stage{(The wand in Lynch's hand flashes:
a brass poker.
Stephen stands at the pianola on which sprawl his hat and ashplant.
With two fingers he repeats once more the series of empty fifths.
Florry Talbot, a blond feeble goosefat whore
in a tatterdemalion gown of mildewed strawberry,
lolls spreadeagle in the sofacorner,
her limp forearm pendent over the bolster, listening.
A heavy stye droops over her sleepy eyelid.)}

\Kitty:
\stage{(hiccups again with a kick of her horsed foot)}
O, excuse!

\Zoe:
\gab{2080}
\stage{(promptly)}
Your boy's thinking of you.
Tie a knot on your shift.

\stage{(Kitty Ricketts bends her head.
Her boa uncoils, slides, glides over her shoulder, back, arm, chair to the ground.
Lynch lifts the curled caterpillar on his wand.
She snakes her neck, nestling.
Stephen glances behind at the squatted figure with its cap back to the front.)}

\Stephen:
As a matter of fact it is of no importance
whether Benedetto Marcello found it or made it.
The rite is the poet's rest.
It may be an old hymn to Demeter
or also illustrate \latin{Coela enarrant gloriam Domini}.
It is susceptible of nodes or modes as far apart as hyperphrygian and mixolydian
\gab{2090}
and of texts so divergent as priests haihooping round
David's that is Circe's or what am I saying Ceres' altar
and David's tip from the stable to his chief bassoonist
about the alrightness of his almightiness.
\french{Mais nom de nom}, that is another pair of trousers.
\stage{Jetez la gourme.}
\stage{faut que jeunesse se passe.}
\stage{(he stops, points at Lynch's cap, smiles, laughs)}
Which side is your knowledge bump?

\Cap:
\stage{(with saturnine spleen)}
Bah!
It is because it is.
Woman's reason.
Jewgreek is greekjew.
Extremes meet.
Death is the highest form of life.
Bah!

\Stephen:
You remember fairly accurately all my errors, boasts, mistakes.
\gab{2100}
How long shall I continue to close my eyes to disloyalty?
Whetstone!

\Cap:
Bah!

\Stephen:
Here's another for you.
\stage{(he frowns)}
The reason is because the fundamental and the dominant
are separated by the greatest possible interval which ...

\Cap:
Which?
Finish.
You can't.

\Stephen:
\gab{2110}
\stage{(with an effort)}
Interval which.
Is the greatest possible ellipse.
Consistent with.
The ultimate return.
The octave.
Which.

\Cap:
Which?

\stage{(Outside the gramophone begins to blare \Title{The Holy City}.)}

\Stephen:
\stage{(abruptly)}
What went forth to the ends of the world to traverse not itself,
God, the sun, Shakespeare, a commercial traveller,
having itself traversed in reality itself becomes that self.
Wait a moment.
Wait a second.
Damn that fellow's noise in the street.
\gab{2120}
Self which it itself was ineluctably preconditioned to become.
\stage{Ecco!}

\Lynch:
\stage{(with a mocking whinny of laughter grins at Bloom and Zoe Higgins)}
What a learned speech, eh?

\Zoe:
\stage{(briskly)}
God help your head, he knows more than you have forgotten.

\stage{(With obese stupidity Florry Talbot regards Stephen.)}

\Florry:
They say the last day is coming this summer.

\Kitty:
\gab{2130}
No!

\Zoe:
\stage{(explodes in laughter)}
Great unjust God!

\Florry:
\stage{(offended)}
Well, it was in the papers about Antichrist.
O, my foot's tickling.

\stage{(Ragged barefoot newsboys, jogging a wagtail kite, patter past, yelling.)}

\Newsboys:
Stop press edition.
\gab{2140}
Result of the rockinghorse races.
Sea serpent in the royal canal.
Safe arrival of Antichrist.

\stage{(Stephen turns and sees Bloom.)}

\Stephen:
A time, times and half a time.

\stage{(Reuben J Antichrist, wandering jew,
a clutching hand open on his spine, stumps forward.
Across his loins is slung a pilgrim's wallet from which
protrude promissory notes and dishonoured bills.
Aloft over his shoulder he bears a long boatpole
from the hook of which the sodden huddled mass of his only son,
saved from Liffey waters, hangs from the slack of its breeches.
\gab{2150}
A hobgoblin in the image of Punch Costello, hipshot, crookbacked,
hydrocephalic, prognathic with receding forehead and Ally Sloper nose,
tumbles in somersaults through the gathering darkness.)}

\All:
What?

\Hobgoblin:
\stage{(his jaws chattering, capers to and fro, goggling his eyes,
squeaking, kangaroohopping with outstretched clutching arms,
then all at once thrusts his lipless face through the fork of his thighs)}
\french{Il vient!
C'est moi!
L'homme qui rit!
L'homme primigène!}
\gab{2160}
\stage{(he whirls round and round with dervish howls)}
\french{Sieurs et dames, faites vos jeux!}
\stage{(He crouches juggling.
Tiny roulette planets fly from his hands.)}
\french{Les jeux sont faits!}
\stage{(the planets rush together, uttering crepitant cracks) }
\french{Rien va plus!}
\stage{(The planets, buoyant balloons, sail swollen up and away.
He springs off into vacuum.)}

\Florry:
\stage{(sinking into torpor, crossing herself secretly)}
The end of the world!

\stage{(A female tepid effluvium leaks out from her.
Nebulous obscurity occupies space.
Through the drifting fog without the gramophone
blares over coughs and feetshuffling.)}

\Gramophone:
\gab{2170}
\begin{verse}
    Jerusalem!\\
    Open your gates and sing\\
    Hosanna ...
\end{verse}

\stage{(A rocket rushes up the sky and bursts.
A white star fills from it,
proclaiming the consummation of all things and second coming of Elijah.
Along an infinite invisible tightrope taut from zenith to nadir
the End of the World, a twoheaded octopus in gillie's kilts,
busby and tartan filibegs, whirls through the murk,
head over heels, in the form of the Three Legs of Man.)}

\EndOfWorld:
\gab{2180}
\stage{(with a Scotch accent)}
Wha'll dance the keel row, the keel row, the keel row?

\stage{(Over the possing drift and choking breathcoughs,
Elijah's voice, harsh as a corncrake's, jars on high.
Perspiring in a loose lawn surplice with funnel sleeves he is seen,
vergerfaced, above a rostrum about which the banner of old glory is draped.
He thumps the parapet.)}

\Elijah:
No yapping, if you please, in this booth.
Jake Crane, Creole Sue, Dove Campbell, Abe Kirschner,
do your coughing with your mouths shut.
\gab{2190}
Say, I am operating all this trunk line.
Boys, do it now.
God's time is 12.25.
Tell mother you'll be there.
Rush your order and you play a slick ace.
Join on right here.
Book through to eternity junction, the nonstop run.
Just one word more.
Are you a god or a doggone clod?
If the second advent came to Coney Island are we ready?
Florry Christ, Stephen Christ, Zoe Christ, Bloom Christ, Kitty Christ, Lynch Christ,
it's up to you to sense that cosmic force.
Have we cold feet about the cosmos?
No. Be on the side of the angels.
Be a prism.
You have that something within, the higher self.
You can rub shoulders with a Jesus, a Gautama, an Ingersoll.
Are you all in this vibration?
\gab{2200}
I say you are.
You once nobble that, congregation,
and a buck joyride to heaven becomes a back number.
You got me?
It's a lifebrightener, sure.
The hottest stuff ever was.
It's the whole pie with jam in.
It's just the cutest snappiest line out.
It is immense, supersumptuous.
It restores.
It vibrates.
I know and I am some vibrator.
Joking apart and, getting down to bedrock, A.J.~Christ Dowie
and the harmonial philosophy, have you got that?
O.K.
Seventyseven west sixtyninth street.
Got me?
That's it.
You call me up by sunphone any old time.
Bumboosers, save your stamps.
\stage{(he shouts)}
Now then our glory song.
All join heartily in the singing.
Encore!
\stage{(he sings)}
Jeru ...

\Gramophone:
\gab{2210}
\stage{(drowning his voice)}
Whorusalaminyourhighhohhhh ...
\stage{(the disc rasps gratingly against the needle)}

\ThreeWhores:
\stage{(covering their ears, squawk)}
Ahhkkk!

\Elijah:
\stage{(in rolledup shirtsleeves, black in the face,
shouts at the top of his voice, his arms uplifted)}
Big Brother up there, Mr~President,
you hear what I done just been saying to you.
Certainly, I sort of believe strong in you, Mr~President.
I certainly am thinking now
Miss Higgins and Miss Ricketts got religion way inside them.
\gab{2220}
Certainly seems to me I don't never see no wusser scared female
than the way you been, Miss Florry, just now as I done seed you.
Mr~President, you come long and help me save our sisters dear.
\stage{(he winks at his audience)}
Our Mr~President, he twig the whole lot and he aint saying nothing.

\KittyKate:
I forgot myself.
In a weak moment I erred and did what I did on Constitution hill.
I was confirmed by the bishop and enrolled in the brown scapular.
My mother's sister married a Montmorency.
It was a working plumber was my ruination when I was pure.

\ZoeFanny:
\gab{2230}
I let him larrup it into me for the fun of it.

\FlorryTeresa:
It was in consequence of a portwine beverage on top of Hennessy's three star.
I was guilty with Whelan when he slipped into the bed.

\Stephen:
In the beginning was the word, in the end the world without end.
Blessed be the eight beatitudes.

\stage{(The beatitudes, Dixon, Madden, Crotthers, Costello, Lenehan, Bannon,
Mulligan and Lynch in white surgical students' gowns, four abreast,
\gab{2240}
goosestepping, tramp fist past in noisy marching)}

\Beatitudes:
\stage{(incoherently)}
Beer beef battledog buybull businum barnum buggerum bishop.

\Lyster:
\stage{(in quakergrey kneebreeches and broadbrimmed hat, says discreetly)}
He is our friend.
I need not mention names.
Seek thou the light.

\stage{(He corantos by.
Best enters in hairdresser's attire, shinily laundered, his locks in curlpapers.
He leads John Eglinton who wears a mandarin's kimono of Nankeen yellow,
lizardlettered, and a high pagoda hat.)%
\gab{2250}
}

\Best:
\stage{(smiling, lifts the hat and displays a shaven poll
from the crown of which bristles a pigtail toupee tied with an orange topknot)}
I was just beautifying him, don't you know.
A thing of beauty, don't you know,
Yeats says, or I mean, Keats says.

\JohnEglinton:
\stage{(produces a greencapped dark lantern and flashes it
towards a corner:
with carping accent)}
Esthetics and cosmetics are for the boudoir.
I am out for truth.
Plain truth for a plain man.
Tanderagee wants the facts and means to get them.%
\gab{2260}

\stage{(In the cone of the searchlight behind the coalscuttle, ollave, holyeyed,
the bearded figure of Mananaun Maclir broods, chin on knees.
He rises slowly.
A cold seawind blows from his druid mouth.
About his head writhe eels and elvers.
He is encrusted with weeds and shells.
His right hand holds a bicycle pump.
His left hand grasps a huge crayfish by its two talons.)}

\MananaunMaclir:
\stage{(with a voice of waves)}
Aum! Hek! Wal! Ak! Lub! Mor! Ma!
White yoghin of the gods.
Occult pimander of Hermes Trismegistos.
\stage{(with a voice of whistling seawind)}
\gab{2270}
Punarjanam patsypunjaub!
I won't have my leg pulled.
It has been said by one:
beware the left, the cult of Shakti.
\stage{(with a cry of stormbirds)}
Shakti Shiva, darkhidden Father!
\stage{(He smites with his bicycle pump the crayfish in his left hand.
On its cooperative dial glow the twelve signs of the zodiac.
He wails with the vehemence of the ocean.)}
Aum! Baum! Pyjaum!
I am the light of the homestead!
I am the dreamery creamery butter.

\stage{(A skeleton judashand strangles the light.
The green light wanes to mauve.
The gasjet wails whistling.)}

\Gasjet:
Pooah! Pfuiiiiiii!
\gab{2280}

\stage{(Zoe runs to the chandelier and, crooking her leg, adjusts the mantle.)}

\Zoe:
Who has a fag as I'm here?

\Lynch:
\stage{(tossing a cigarette on to the table)}
Here.

\Zoe:
\stage{(her head perched aside in mock pride)}
Is that the way to hand the \french{pot} to a lady?
\stage{(She stretches up to light the cigarette over the flame,
twirling it slowly, showing the brown tufts of her armpits.
\gab{2290}
Lynch with his poker lifts boldly a side of her slip.
Bare from her garters up her flesh appears under the sapphire a nixie's green.
She puffs calmly at her cigarette.)}
Can you see the beautyspot of my behind?

\Lynch:
I'm not looking

\Zoe:
\stage{(makes sheep's eyes)}
No? You wouldn't do a less thing.
Would you suck a lemon?

\stage{(Squinting in mock shame she glances with sidelong meaning at Bloom,
\gab{2300}
then twists round towards him, pulling her slip free of the poker.
Blue fluid again flows over her flesh.
Bloom stands, smiling desirously, twirling his thumbs.
Kitty Ricketts licks her middle finger with her spittle and,
gazing in the mirror, smooths both eyebrows.
Lipoti Virag, basilicogrammate,
chutes rapidly down through the chimneyflue and struts
two steps to the left on gawky pink stilts.
He is sausaged into several overcoats and wears a brown macintosh
under which he holds a roll of parchment.
In his left eye flashes the monocle of
Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell.
On his head is perched an Egyptian pshent.
Two quills project over his ears.)%
\gab{2310}
}

\Virag:
\stage{(heels together, bows)}
My name is Virag Lipoti, of Szombathely.
\stage{(he coughs thoughtfully, drily)}
Promiscuous nakedness is much in evidence hereabouts, eh?
Inadvertently her backview revealed the fact that she is not wearing
those rather intimate garments of which you are a particular devotee.
The injection mark on the thigh I hope you perceived?
Good.

\Bloom:
Granpapachi. But ...

\Virag:
Number two on the other hand, she of the cherry rouge and coiffeuse white,%
\gab{2320}
whose hair owes not a little to our tribal elixir of gopherwood,
is in walking costume and tightly staysed by her sit, I should opine.
Backbone in front, so to say.
Correct me but I always understood that the act so performed by skittish humans
with glimpses of lingerie appealed to you in virtue of its exhibitionististicicity.
In a word. Hippogriff.
Am I right?

\Bloom:
She is rather lean.

\Virag:
\stage{(not unpleasantly)}
Absolutely!
Well observed and those pannier pockets of the skirt
\gab{2330}
and slightly pegtop effect are devised to suggest bunchiness of hip.
A new purchase at some monster sale for which a gull has been mulcted.
Meretricious finery to deceive the eye.
Observe the attention to details of dustspecks.
Never put on you tomorrow what you can wear today.
Parallax!
\stage{(with a nervous twitch of his head)}
Did you hear my brain go snap?
Pollysyllabax!

\Bloom:
\stage{(an elbow resting in a hand, a forefinger against his cheek)}
She seems sad.

\Virag:
\stage{(cynically, his weasel teeth bared yellow, draws down his left eye
with a finger and barks hoarsely)}
\gab{2340}
Hoax! Beware of the flapper and bogus mournful.
Lily of the alley.
All possess bachelor's button discovered by Rualdus Columbus.
Tumble her. Columble her. Chameleon.
\stage{(more genially)}
Well then, permit me to draw your attention to item number three.
There is plenty of her visible to the naked eye.
Observe the mass of oxygenated vegetable matter on her skull.
What ho, she bumps!
The ugly duckling of the party, longcasted and deep in keel.

\Bloom:
\stage{(regretfully)}
When you come out without your gun.

\Virag:
We can do you all brands, mild, medium and strong.
\gab{2350}
Pay your money, take your choice.
How happy could you be with either ...

\Bloom:
With ...?

\Virag:
\stage{(his tongue upcurling)}
Lyum! Look.
Her beam is broad.
She is coated with quite a considerable layer of fat.
Obviously mammal in weight of bosom
you remark that she has in front well to the fore
two protuberances of very respectable dimensions,
inclined to fall in the noonday soupplate,
while on her rere lower down are two additional protuberances,
suggestive of potent rectum and tumescent for palpation,
\gab{2360}
which leave nothing to be desired save compactness.
Such fleshy parts are the product of careful nurture.
When coopfattened their livers reach an elephantine size.
Pellets of new bread with fennygreek and gumbenjamin
swamped down by potions of green tea endow them during
their brief existence with natural pincushions of quite colossal blubber.
That suits your book, eh?
Fleshhotpots of Egypt to hanker after.
Wallow in it.
Lycopodium.
\stage{(his throat twitches)}
Slapbang!
There he goes again.

\Bloom:
The stye I dislike.

\Virag:
\gab{2370}
\stage{(arches his eyebrows)}
Contact with a goldring, they say.
\latin{argumentum ad feminam},
as we said in old Rome and ancient Greece
in the consulship of Diplodocus and Ichthyosauros.
For the rest Eve's sovereign remedy.
Not for sale.
Hire only.
Huguenot.
\stage{(he twitches)}
It is a funny sound.
\stage{(he coughs encouragingly)}
But possibly it is only a wart.
I presume you shall have remembered what I will have taught you on that head?
Wheatenmeal with honey and nutmeg.

\Bloom:
\stage{(reflecting)}
Wheatenmeal with lycopodium and syllabax.
This searching ordeal.
It has been an unusually fatiguing day, a chapter of accidents.
\gab{2380}
Wait.
I mean, wartsblood spreads warts, you said ...

\Virag:
\stage{(severely, his nose hardhumped, his side eye winking)}
Stop twirling your thumbs and have a good old thunk.
See, you have forgotten.
Exercise your mnemotechnic.
\italian{La causa è santa.}
Tara. Tara.
\stage{(aside)}
He will surely remember.

\Bloom:
Rosemary also did I understand you to say or willpower over parasitic tissues.
Then nay no I have an inkling.
The touch of a deadhand cures.
Mnemo?%
\gab{2390}

\Virag:
\stage{(excitedly)}
I say so. I say so.
E'en so.
Technic.
\stage{(he taps his parchmentroll energetically)}
This book tells you how to act with all descriptive particulars.
Consult index for agitated fear of aconite,
melancholy of muriatic, priapic pulsatilla.
Virag is going to talk about amputation.
Our old friend caustic.
They must be starved.
Snip off with horsehair under the denned neck.
But, to change the venue to the Bulgar and the Basque,
have you made up your mind whether you like or dislike women in male habiliments?
\stage{(with a dry snigger)}
You intended to devote an entire year to the study
of the religious problem and the summer months of 1886
\gab{2400}
to square the circle and win that million.
Pomegranate!
From the sublime to the ridiculous is but a step.
Pyjamas, let us say?
Or stockingette gussetted knickers, closed?
Or, put we the case, those complicated combinations, camiknickers?
\stage{(he crows derisively)}
Keekeereekee!

\stage{(Bloom surveys uncertainly the three whores
then gazes at the veiled mauve light, hearing the everflying moth.)}

\Bloom:
I wanted then to have now concluded.
Nightdress was never.
Hence this.
But tomorrow is a new day will be.
Past was is today.
What now is will then morrow
\gab{2410}
as now was be past yester.

\Virag:
\stage{(prompts in a pig's whisper)}
Insects of the day spend their brief existence in reiterated coition,
lured by the smell of the inferiorly pulchritudinous fumale
possessing extendified pudendal nerve in dorsal region.
Pretty Poll!
\stage{(his yellow parrotbeak gabbles nasally)}
They had a proverb in the Carpathians in or about the year
five thousand five hundred and fifty of our era.
One tablespoonful of honey will attract friend Bruin
more than half a dozen barrels of first choice malt vinegar.
Bear's buzz bothers bees.
But of this apart.
At another time we may resume.
We were very pleased, we others.
\gab{2420}
\stage{(he coughs and, bending his brow,
rubs his nose thoughtfully with a scooping hand)}
You shall find that these night insects follow the light.
An illusion for remember their complex unadjustable eye.
For all these knotty points see the seventeenth book
of my Fundamentals of Sexology or the Love Passion
which Doctor L.B. says is the book sensation of the year.
Some, to example, there are again whose movements are automatic.
Perceive.
That is his appropriate sun.
Nightbird nightsun nighttown.
Chase me, Charley!
\stage{(he blows into Bloom's ear)}
Buzz!

\Bloom:
Bee or bluebottle too other day butting shadow on wall dazed self
then me wandered dazed down shirt good job I ...%
\gab{2430}

\Virag:
\stage{(his face impassive, laughs in a rich feminine key)}
Splendid!
Spanish fly in his fly or mustard plaster on his dibble.
\stage{(he gobbles gluttonously with turkey wattles)}
Bubbly jock! Bubbly jock!
Where are we?
Open Sesame!
Cometh forth!
\stage{(he unrolls his parchment rapidly and reads,
his glowworm's nose running backwards over the letters which he claws)}
Stay, good friend.
I bring thee thy answer.
Redbank oysters will shortly
be upon us.
I'm the best o'cook.
Those succulent bivalves may help us and
the truffles of Perigord, tubers dislodged through mister omnivorous porker,
were unsurpassed in cases of nervous debility or viragitis.
\gab{2440}
Though they stink yet they sting.
\stage{(he wags his head with cackling raillery)}
Jocular.
With my eyeglass in my ocular.
\stage{(he sneezes)}
Amen!

\Bloom:
\stage{(absently)}
Ocularly woman's bivalve case is worse.
Always open sesame.
The cloven sex.
Why they fear vermin, creeping things.
Yet Eve and the serpent contradicts.
Not a historical fact.
Obvious analogy to my idea.
Serpents too are gluttons for woman's milk.
Wind their way through miles of omnivorous forest to sucksucculent her breast dry.
Like those bubblyjocular Roman matrons one reads of in Elephantuliasis.

\Virag:
\gab{2450}
\stage{(his mouth projected in hard wrinkles, eyes stonily forlornly closed,
psalms in outlandish monotone)}
That the cows with their those distended udders
that they have been the the known ...

\Bloom:
I am going to scream.
I beg your pardon.
Ah? So.
\stage{(he repeats)}
Spontaneously to seek out the saurian's lair in order to
entrust their teats to his avid suction.
Ant milks aphis.
\stage{(profoundly)}
Instinct rules the world.
In life.
In death.

\Virag:
\stage{(head askew, arches his back and hunched wingshoulders,
\gab{2460}
peers at the moth out of blear bulged eyes, points a horning claw and cries)}
Who's moth moth?
Who's dear Gerald?
Dear Ger, that you?
O dear, he is Gerald.
O, I much fear he shall be most badly burned.
Will some pleashe pershon not now impediment
so catastrophics mit agitation of firstclass tablenumpkin?
\stage{(he mews)}
Puss puss puss puss!
\stage{(he sighs, draws back and stares sideways down with dropping underjaw)}
Well, well.
He doth rest anon.
\stage{(he snaps his jaws suddenly on the air)}

\Moth:
\begin{verse}
    I'm a tiny tiny thing\\
\gab{2470}%
    Ever flying in the spring\\
    Round and round a ringaring.\\
    Long ago I was a king\\
    Now I do this kind of thing\\
    On the wing, on the wing!\\
    Bing!
\end{verse}
\stage{(he rushes against the mauve shade, flapping noisily)}
Pretty pretty pretty pretty pretty pretty petticoats.

\stage{(From left upper entrance with two gliding steps
Henry Flower comes forward to left front centre.
He wears a dark mantle and drooping plumed sombrero.
\gab{2480}
He carries a silverstringed inlaid dulcimer and a longstemmed
bamboo Jacob's pipe, its clay bowl fashioned as a female head.
He wears dark velvet hose and silverbuckled pumps.
He has the romantic Saviour's face with flowing locks, thin beard and moustache.
His spindlelegs and sparrow feet are those of the tenor Mario, prince of Candia.
He settles down his goffered ruffs
and moistens his lips with a passage of his amorous tongue.)}

\Henry:
\stage{(in a low dulcet voice, touching the strings of his guitar)}
There is a flower that bloometh.%
\gab{2490}

\stage{(Virag truculent, his jowl set, stares at the lamp.
Grave Bloom regards Zoe's neck.
Henry gallant turns with pendant dewlap to the piano.)}

\Stephen:
\stage{(to himself)}
Play with your eyes shut.
Imitate pa.
Filling my belly with husks of swine.
Too much of this.
I will arise and go to my.
Expect this is the.
Steve, thou art in a parlous way.
Must visit old Deasy or telegraph.
Our interview of this morning has left on me a deep impression.
Though our ages.
Will write fully tomorrow.
I'm partially drunk, by the way.
\stage{(he touches the keys again)}
\gab{2500}
Minor chord comes now.
Yes. Not much however.

\stage{(Almidano Artifoni holds out a batonroll of music with vigorous moustachework.)}

\Artifoni:
\italian{Ci rifletta.
Lei rovina tutto.}

\Florry:
Sing us something.
Love's old sweet song.

\Stephen:
No voice.
I am a most finished artist.
Lynch, did I show you the letter about the lute?

\Florry:
\gab{2510}
\stage{(smirking)}
The bird that can sing and won't sing.

\stage{(The Siamese twins, Philip Drunk and Philip Sober,
two Oxford dons with lawnmowers, appear in the window embrasure.
Both are masked with Matthew Arnold's face.)}

\PhilipSober:
Take a fool's advice.
All is not well.
Work it out with the buttend of a pencil, like a good young idiot.
Three pounds twelve you got, two notes,
one sovereign, two crowns, if youth but knew.
Mooney's en ville, Mooney's sur mer, the Moira, Larchet's,
Holles street hospital, Burke's.
Eh? I am watching you.%
\gab{2520}

\PhilipDrunk:
\stage{(impatiently)}
Ah, bosh, man.
Go to hell!
I paid my way.
If I could only find out about octaves.
Reduplication of personality.
Who was it told me his name?
\stage{(his lawnmower begins to purr)}
Aha, yes.
\greek{Zoe mou sas agapo.}
Have a notion I was here before.
When was it not Atkinson his card I have somewhere.
Mac Somebody.
Unmack I have it.
He told me about, hold on, Swinburne, was it, no?

\Florry:
And the song?

\Stephen:
\gab{2530}
Spirit is willing but the flesh is weak.

\Florry:
Are you out of Maynooth?
You're like someone I knew once.

\Stephen:
Out of it now.
\stage{(to himself)}
Clever.

\PhilipDrunk and \PhilipSober:
\stage{(their lawnmowers purring with a rigadoon of grasshalms)}
Clever ever.
Out of it out of it.
By the bye have you the book, the thing, the ashplant?
Yes, there it, yes.
Cleverever outofitnow.
Keep in condition.
Do like us.

\Zoe:
\gab{2540}
There was a priest down here two nights ago
to do his bit of business with his coat buttoned up.
You needn't try to hide, I says to him.
I know you've a Roman collar.

\Virag:
Perfectly logical from his standpoint.
Fall of man.
\stage{(harshly, his pupils waxing)}
To hell with the pope!
Nothing new under the sun.
I am the Virag who disclosed the Sex Secrets of Monks and Maidens.
Why I left the Church of Rome.
Read the Priest, the Woman and the Confessional.
Penrose.
Flipperty Jippert.
\stage{(he wriggles)}
Woman, undoing with sweet pudor her belt of rushrope,
offers her allmoist yoni to man's lingam.
\gab{2550}
Short time after man presents woman with pieces of jungle meat.
Woman shows joy and covers herself with featherskins.
Man loves her yoni fiercely with big lingam, the stiff one.
\stage{(he cries)}
\latin{Coactus volui.}
Then giddy woman will run about.
Strong man grapses woman's wrist.
Woman squeals, bites, spucks.
Man, now fierce angry, strikes woman's fat yadgana.
\stage{(he chases his tail)}
Piffpaff!
Popo!
\stage{(he stops, sneezes)}
Pchp!
\stage{(he worries his butt)}
Prrrrrht!

\Lynch:
I hope you gave the good father a penance.
Nine glorias for shooting a bishop.

\Zoe:
\gab{2560}
\stage{(spouts walrus smoke through her nostrils)}
He couldn't get a connection.
Only, you know, sensation.
A dry rush.

\Bloom:
Poor man!

\Zoe:
\stage{(lightly)}
Only for what happened him.

\Bloom:
How?

\Virag:
\stage{(A diabolic rictus of black luminosity contracting his visage,
\gab{2570}
cranes his scraggy neck forward.
He lifts a mooncalf nozzle and howls.)}
\yiddish{Verfluchte Goim!}
He had a father, forty fathers.
He never existed.
Pig God!
He had two left feet.
He was Judas Iacchia, a Libyan eunuch, the pope's bastard.
\stage{(he leans out on tortured forepaws, elbows bent rigid,
his eye agonising in his flat skullneck and yelps over the mute world)}
A son of a whore.
Apocalypse.

\Kitty:
And Mary Shortall that was in the lock with the pox
she got from Jimmy Pidgeon in the blue caps
had a child off him that couldn't swallow
and was smothered with the convulsions in the mattress
\gab{2580}
and we all subscribed for the funeral.

\PhilipDrunk:
\stage{(gravely)}
\french{Qui vous a mis dans cette fichue position, Philippe?}

\PhilipSober:
\stage{(gaily)}
\french{C'était le sacré pigeon, Philippe.}

\stage{(Kitty unpins her hat and sets it down calmly, patting her henna hair.
And a prettier, a daintier head of winsome curls was never seen on a whore's shoulders.
Lynch puts on her hat.
She whips it off.)}

\Lynch:
\stage{(laughs)}
\gab{2590}
And to such delights has Metchnikoff inoculated anthropoid apes.

\Florry:
\stage{(nods)}
Locomotor ataxy.

\Zoe:
\stage{(gaily)}
O, my dictionary.

\Lynch:
Three wise virgins.

\Virag:
\stage{(agueshaken, profuse yellow spawn foaming over his bony epileptic lips)}
She sold lovephiltres, whitewax, orangeflower.
Panther, the Roman centurion, polluted her with his genitories.
\gab{2600}
\stage{(he sticks out a flickering phosphorescent scorpion tongue, his hand on his fork)}
Messiah!
He burst her tympanum.
\stage{(with gibbering baboon's cries he jerks his hips in the cynical spasm)}
Hik! Hek! Hak! Hok! Huk! Kok! Kuk!

\stage{(Ben Jumbo Dollard, rubicund, musclebound, hairynostrilled, hugebearded,
cabbageeared, shaggychested, shockmaned, fat-papped, stands forth,
his loins and genitals tightened into a pair of black bathing bagslops.)}

\BenDollard:
\stage{(nakkering castanet bones in his huge padded paws,
yodels jovially in base barreltone)}
\gab{2610}
When love absorbs my ardent soul.

\stage{(The virgins Nurse Callan and Nurse Quigley burst through the ringkeepers
and the ropes and mob him with open arms.)}

\Virgins:
\stage{(gushingly)}
Big Ben!
Ben my Chree!

\Voice:
Hold that fellow with the bad breeches.

\BenDollard:
\stage{(smites his thigh in abundant laughter)}
Hold him now.

\Henry:
\stage{(caressing on his breast a severed female head, murmurs)}
\gab{2620}
Thine heart, mine love.
\stage{(he plucks his lutestrings)}
When first I saw ...

\Virag:
\stage{(sloughing his skins, his multitudinous plumage moulting)}
Rats!
\stage{(he yawns, showing a coalblack throat,
and closes his jaws by an upward push of his parchmentroll)}
After having said which I took my departure.
Farewell.
Fare thee well.
\yiddish{Dreck!}

\stage{(Henry Flower combs his moustache and beard rapidly with a pocketcomb
and gives a cow's lick to his hair.
Steered by his rapier, he glides to the door, his wild harp slung behind him.
Virag reaches the door in two ungainly stilthops,
\gab{2630}
his tail cocked, and deftly claps sideways on the wall
a pusyellow flybill, butting it with his head.)}

\Flybill:
K.~II.
Post No Bills.
Strictly confidential.
Dr Hy Franks.

\Henry:
All is lost now.

\stage{(Virag unscrews his head in a trice and holds it under his arm.)}

\ViragsHead:
Quack!

\stage{(Exeunt severally.)}

\Stephen:
\gab{2640}
\stage{(over his shoulder to Zoe)}
You would have preferred the fighting parson who founded the protestant error.
But beware Antisthenes, the dog sage, and the last end of Arius Heresiarchus.
The agony in the closet.

\Lynch:
All one and the same God to her.

\Stephen:
\stage{(devoutly)}
And sovereign Lord of all things.

\Florry:
\stage{(to Stephen)}
I'm sure you're a spoiled priest.
Or a monk.

\Lynch:
\gab{2650}
He is.
A cardinal's son.

\Stephen:
Cardinal sin.
Monks of the screw.

\stage{(His Eminence Simon Stephen cardinal Dedalus, primate of all Ireland,
appears in the doorway, dressed in red soutane, sandals and socks.
Seven dwarf simian acolytes, also in red, cardinal sins,
uphold his train, peeping under it.
He wears a battered silk hat sideways on his head.
His thumbs are stuck in his armpits and his palms outspread.
Round his neck hangs a rosary of corks ending on his breast in a corkscrew cross.
\gab{2660}
Releasing his thumbs, he invokes grace from on high with large wave gestures
and proclaims with bloated pomp:)}

\Cardinal:
\begin{verse}
    Conservio lies captured\\
    He lies in the lowest dungeon\\
    With manacles and chains around his limbs\\
    Weighing upwards of three tons.
\end{verse}

\stage{(He looks at all for a moment, his right eye closed tight, his left cheek puffed out.
Then, unable to repress his merriment, he rocks to and fro,
\gab{2670}
arms akimbo, and sings with broad rollicking humour:)}

\begin{verse}
    O, the poor little fellow\\
    Hihihihihis legs they were yellow\\
    He was plump, fat and heavy and brisk as a snake\\
    But some bloody savage\\
    To graize his white cabbage\\
    He murdered Nell Flaherty's duckloving drake.
\end{verse}

\stage{(A multitude of midges swarms white over his robe.
He scratches himself with crossed arms at his ribs, grimacing, and exclaims:)}

I'm suffering the agony of the damned.
By the hoky fiddle, thanks be to Jesus those funny little chaps are not unanimous.
\gab{2680}
If they were they'd walk me off the face of the bloody globe.

\stage{(His head aslant he blesses curtly with fore and middle fingers,
imparts the Easter kiss and doubleshuffles off comically,
swaying his hat from side to side,
shrinking quickly to the size of his trainbearers.
The dwarf acolytes, giggling, peeping, nudging,
ogling, easterkissing, zigzag behind him.
His voice is heard mellow from afar, merciful male, melodious:)}

\begin{verse}
    Shall carry my heart to thee,\\
    Shall carry my heart to thee,\\
\gab{2690}%
    And the breath of the balmy night\\
    Shall carry my heart to thee!
\end{verse}

\stage{(The trick doorhandle turns.)}

\Doorhandle:
Theeee!

\Zoe:
The devil is in that door.

\stage{(A male form passes down the creaking staircase
and is heard taking the waterproof and hat from the rack.
Bloom starts forward involuntarily and,
half closing the door as he passes,
takes the chocolate from his pocket
\gab{2700}%
and offers it nervously to Zoe.)}

\Zoe:
\stage{(sniffs his hair briskly)}
Hmmm!
Thank your mother for the rabbits.
I'm very fond of what I like.

\Bloom:
\stage{(hearing a male voice in talk with the whores on the doorstep, pricks his ears)}
If it were he?
After?
Or because not?
Or the double event?

\Zoe:
\stage{(tears open the silverfoil)}
Fingers was made before forks.
\stage{(she breaks off and nibbles a piece gives a piece to Kitty Ricketts
and then turns kittenishly to Lynch)}
\gab{2710}
No objection to French lozenges?
\stage{(He nods.
She taunts him.)}
Have it now or wait till you get it?
\stage{(He opens his mouth, his head cocked.
She whirls the prize in left circle.
His head follows.
She whirls it back in right circle.
He eyes her.)}
Catch!

\stage{(She tosses a piece.
With an adroit snap he catches it and bites it through with a crack.)}

\Kitty:
\stage{(chewing)}
The engineer I was with at the bazaar does have lovely ones.
Full of the best liqueurs.
And the viceroy was there with his lady.
The gas we had on the Toft's hobbyhorses.
I'm giddy still.

\Bloom:
\gab{2720}
\stage{(In Svengali's fur overcoat, with folded arms and Napoleonic forelock,
frowns in ventriloquial exorcism with piercing eagle glance towards the door.
Then rigid with left foot advanced he makes a swift pass
with impelling fingers and gives the sign of past master,
drawing his right arm downwards from his left shoulder.)}
Go, go, go, I conjure you, whoever you are!

\stage{(A male cough and tread are heard passing through the mist outside.
Bloom's features relax.
He places a hand in his waistcoat, posing calmly.
Zoe offers him chocolate.)}

\Bloom:
\gab{2730}
\stage{(solemnly)}
Thanks.

\Zoe:
Do as you're bid.
Here!

\stage{(a firm heelclacking tread is heard on the stairs.)}

\Bloom:
\stage{(takes the chocolate)}
Aphrodisiac?
Tansy and pennyroyal.
But I bought it.
Vanilla calms or?
Mnemo.
Confused light confuses memory.
Red influences lupus.
Colours affect women's characters, any they have.
This black makes me sad.
Eat and be merry for tomorrow.
\stage{(he eats)}
Influence taste too, mauve.
But it is so long since I.
Seems new.
Aphro.
That priest.
\gab{2740}
Must come.
Better late than never.
Try truffles at Andrews.

\stage{(The door opens.
Bella Cohen, a massive whoremistress, enters.
She is dressed in a threequarter ivory gown,
fringed round the hem with tasselled selvedge,
and cools herself flirting a black horn fan like Minnie Hauck in \title{Carmen}.
On her left hand are wedding and keeper rings.
Her eyes are deeply carboned.
She has a sprouting moustache.
Her olive face is heavy, slightly sweated and fullnosed with orangetainted nostrils.
She has large pendant beryl eardrops.)}

\Bella:
My word!
\gab{2750}
I'm all of a mucksweat.

\stage{(She glances round her at the couples.
Then her eyes rest on Bloom with hard insistence.
Her large fan winnows wind towards her heated faceneck and embonpoint.
Her falcon eyes glitter.)}

\Fan:
\stage{(flirting quickly, then slowly)}
Married, I see.

\Bloom:
Yes.
Partly, I have mislaid ...

\Fan:
\stage{(half opening, then closing)}
And the missus is master.
Petticoat government.%
\gab{2760}

\Bloom:
\stage{(looks down with a sheepish grin)}
That is so.

\Fan:
\stage{(folding together, rests against her left eardrop)}
Have you forgotten me?

\Bloom:
Nes. Yo.

\Fan:
\stage{(folded akimbo against her waist)}
Is me her was you dreamed before?
Was then she him you us since knew?
Am all them and the same now we?

\stage{(Bella approaches,
\gab{2770}
gently tapping with the fan.)}

\Bloom:
\stage{(wincing)}
Powerful being.
In my eyes read that slumber which women love.

\Fan:
\stage{(tapping)}
We have met.
You are mine.
It is fate.

\Bloom:
\stage{(cowed)}
Exuberant female.
Enormously I desiderate your domination.
I am exhausted, abandoned, no more young.
I stand, so to speak, with an unposted letter
bearing the extra regulation fee
before the too late box of the general postoffice of human life.
\gab{2780}
The door and window open at a right angle
cause a draught of thirtytwo feet per second
according to the law of falling bodies.
I have felt this instant a twinge of sciatica in my left glutear muscle.
It runs in our family.
Poor dear papa, a widower, was a regular barometer from it.
He believed in animal heat.
A skin of tabby lined his winter waistcoat.
Near the end, remembering king David and the Sunamite,
he shared his bed with Athos, faithful after death.
A dog's spittle as you probably ...
\stage{(he winces)}
Ah!

\Richie:
\stage{(bagweighted, passes the door)}
Mocking is catch.
Best value in Dub.
Fit for a prince's.
\gab{2790}
Liver and kidney.

\Fan:
\stage{(tapping)}
All things end.
Be mine.
Now.

\Bloom:
\stage{(undecided)}
All now?
I should not have parted with my talisman.
Rain, exposure at dewfall on the searocks, a peccadillo at my time of life.
Every phenomenon has a natural cause.

\Fan:
\stage{(points downwards slowly)}
You may.

\Bloom:
\stage{(looks downwards and perceives her unfastened bootlace)}
\gab{2800}
We are observed.

\Fan:
\stage{(points downwards quickly)}
You must.

\Bloom:
\stage{(with desire, with reluctance)}
I can make a true black knot.
Learned when I served my time and worked the mail order line for Kellett's.
Experienced hand.
Every knot says a lot.
Let me.
In courtesy.
I knelt once before today.
Ah!

\stage{(Bella raises her gown slightly and, steadying her pose,
lifts to the edge of a chair a plump buskined hoof
\gab{2810}
and a full pastern, silksocked.
Bloom, stifflegged, aging, bends over her hoof
and with gentle fingers draws out and in her laces.)}

\Bloom:
\stage{(murmurs lovingly)}
To be a shoefitter in Manfield's was my love's young dream,
the darling joys of sweet buttonhooking,
to lace up crisscrossed to kneelength the dressy kid footwear satinlined,
so incredibly impossibly small, of Clyde Road ladies.
Even their wax model Raymonde I visited daily
to admire her cobweb hose and stick of rhubarb toe, as worn in Paris.

\Hoof:
Smell my hot goathide.
\gab{2820}
Feel my royal weight.

\Bloom:
\stage{(crosslacing)}
Too tight?

\Hoof:
If you bungle, Handy Andy, I'll kick your football for you.

\Bloom:
Not to lace the wrong eyelet as I did the night of the bazaar dance.
Bad luck.
Hook in wrong tache of her ... person you mentioned.
That night she met ... Now!

\stage{(He knots the lace.
Bella places her foot on the floor.
Bloom raises his head.
\gab{2830}
Her heavy face, her eyes strike him in midbrow.
His eyes grow dull, darker and pouched, his nose thickens.)}

\Bloom:
\stage{(mumbles)}
Awaiting your further orders we remain, gentlemen, ...

\Bello:
\stage{(with a hard basilisk stare, in a baritone voice)}
Hound of dishonour!

\Bloom:
\stage{(infatuated)}
Empress!

\Bello:
\stage{(his heavy cheekchops sagging)}
Adorer of the adulterous rump!

\Bloom:
\gab{2840}
\stage{(plaintively)}
Hugeness!

\Bello:
Dungdevourer!

\Bloom:
\stage{(with sinews semiflexed)}
Magmagnificence!

\Bello:
Down!
\stage{(he taps her on the shoulder with his fan)}
Incline feet forward!
Slide left foot one pace back!
You will fall.
You are falling.
On the hands down!

\Bloom:
\gab{2850}
\stage{(her eyes upturned in the sign of admiration, closing, yaps)}
Truffles!

\stage{(With a piercing epileptic cry she sinks on all fours,
grunting, snuffling, rooting at his feet:
then lies, shamming dead, with eyes shut tight, trembling eyelids,
bowed upon the ground in the attitude of most excellent master.)}

\Bello:
\stage{(with bobbed hair, purple gills, fit moustache rings round his shaven mouth,
in mountaineer's puttees, green silverbuttoned coat,
sport skirt and alpine hat with moorcock's feather,
his hands stuck deep in his breeches pockets,
places his heel on her neck and grinds it in)}
\gab{2860}
Footstool!
Feel my entire weight.
Bow, bondslave, before the throne of your despot's glorious heels
so glistening in their proud erectness.

\Bloom:
\stage{(enthralled, bleats)}
I promise never to disobey.

\Bello:
\stage{(laughs loudly)}
Holy smoke!
You little know what's in store for you.
I'm the Tartar to settle your little lot and break you in!
I'll bet Kentucky cocktails all round I shame it out of you, old son.
Cheek me, I dare you.
If you do tremble in anticipation of heel discipline
to be inflicted in gym costume.%
\gab{2870}

\stage{(Bloom creeps under the sofa and peers out through the fringe.)}

\Zoe:
\stage{(widening her slip to screen her)}
She's not here.

\Bloom:
\stage{(closing her eyes)}
She's not here.

\Florry:
\stage{(hiding her with her gown)}
She didn't mean it, Mr~Bello.
She'll be good, sir.

\Kitty:
Don't be too hard on her, Mr~Bello.
\gab{2880}
Sure you won't, ma'amsir.

\Bello:
\stage{(coaxingly)}
Come, ducky dear, I want a word with you, darling,
just to administer correction.
Just a little heart to heart talk, sweety.
\stage{(Bloom puts out her timid head)}
There's a good girly now.
\stage{(Bello grabs her hair violently and drags her forward)}
I only want to correct you for your own good on a soft safe spot.
How's that tender behind?
O, ever so gently, pet.
Begin to get ready.

\Bloom:
\stage{(fainting)}
Don't tear my ...

\Bello:
\gab{2890}
\stage{(savagely)}
The nosering, the pliers, the bastinado, the hanging hook,
the knout I'll make you kiss while the flutes play like the Nubian slave of old.
You're in for it this time!
I'll make you remember me for the balance of your natural life.
\stage{(his forehead veins swollen, his face congested)}
I shall sit on your ottoman saddleback every morning after
my thumping good breakfast of Matterson's fat hamrashers
and a bottle of Guinness's porter.
\stage{(he belches)}
And suck my thumping good Stock Exchange cigar
while I read the \Title{Licensed Victualler's Gazette.}
Very possibly I shall have you slaughtered and skewered in my stables
and enjoy a slice of you with crisp crackling from the baking tin
\gab{2900}
basted and baked like sucking pig with rice and lemon or currant sauce.
It will hurt you.
\stage{(He twists her arm.
Bloom squeals, turning turtle.)}

\Bloom:
Don't be cruel, nurse!
Don't!

\Bello:
\stage{(twisting)}
Another!

\Bloom:
\stage{(screams)}
O, it's hell itself!
Every nerve in my body aches like mad!

\Bello:
\stage{(shouts)}
Good, by the rumping jumping general!
\gab{2910}
That's the best bit of news I heard these six weeks.
Here, don't keep me waiting, damn you!
\stage{(he slaps her face)}

\Bloom:
\stage{(whimpers)}
You're after hitting me.
I'll tell ...

\Bello:
Hold him down, girls, till I squat on him.

\Zoe:
Yes.
Walk on him!
I will.

\Florry:
I will.
\gab{2920}
Don't be greedy.

\Kitty:
No, me.
Lend him to me.

\stage{(The brothel cook, Mrs~Keogh, wrinkled, greybearded, in a greasy bib,
men's grey and green socks and brogues, floursmeared,
a rollingpin stuck with raw pastry in her bare red arm and hand,
appears at the door.)}

\MrsKeogh:
\stage{(ferociously)}
Can I help?
\stage{(They hold and pinion Bloom.)}

\Bello:
\gab{2930}
\stage{(squats with a grunt on Bloom's upturned face,
puffing cigarsmoke, nursing a fat leg)}
I see Keating Clay is elected vicechairman of the Richmond asylum
and by the by Guinness's preference shares are at sixteen three quarters.
Curse me for a fool that didn't buy that lot Craig and Gardner told me about.
Just my infernal luck, curse it.
And that Goddamned outsider \Title{Throwaway} at twenty to one.
\stage{(he quenches his cigar angrily on Bloom's ear)}
Where's that Goddamned cursed ashtray?

\Bloom:
\stage{(goaded, buttocksmothered)}
O! O!
Monsters!
Cruel one!

\Bello:
\gab{2940}
Ask for that every ten minutes. Beg.
Pray for it as you never prayed before.
\stage{(he thrusts out a figged fist and foul cigar)}
Here, kiss that.
Both. Kiss.
\stage{(he throws a leg astride and,
pressing with horseman's knees, calls in a hard voice)}
Gee up!
A cockhorse to Banbury cross.
I'll ride him for the Eclipse stakes.
\stage{(he bends sideways and squeezes his mount's testicles roughly, shouting)}
Ho! Off we pop!
I'll nurse you in proper fashion.
\stage{(he horserides cockhorse, leaping in the, in the saddle)}
The lady goes a pace a pace and the coachman goes a trot a trot
and the gentleman goes a gallop a gallop a gallop a gallop.

\Florry:
\gab{2950}
\stage{(pulls at Bello)}
Let me on him now.
You had enough.
I asked before you.

\Zoe:
\stage{(pulling at Florry)}
Me. Me.
Are you not finished with him yet, suckeress?

\Bloom:
\stage{(stifling)}
Can't.

\Bello:
Well, I'm not.
Wait.
\stage{(he holds in his breath)}
Curse it.
Here.
This bung's about burst.
\stage{(he uncorks himself behind:
then, contorting his features, farts loudly)}
Take that!
\stage{(he recorks himself)}
Yes, by Jingo, sixteen three quarters.%
\gab{2960}

\Bloom:
\stage{(a sweat breaking out over him)}
Not man.
\stage{(he sniffs)}
Woman.

\Bello:
\stage{(stands up)}
No more blow hot and cold.
What you longed for has come to pass.
Henceforth you are unmanned and mine in earnest, a thing under the yoke.
Now for your punishment frock.
You will shed your male garments, you understand, Ruby Cohen?
and don the shot silk luxuriously rustling over head and shoulders.
And quickly too!

\Bloom:
\stage{(shrinks)}
\gab{2970}
Silk, mistress said!
O crinkly!
Scrapy!
Must I tiptouch it with my nails?

\Bello:
\stage{(points to his whores)}
As they are now so will you be, wigged,
singed, perfumesprayed, ricepowdered, with smoothshaven armpits.
Tape measurements will be taken next your skin.
You will be laced with cruel force into vicelike corsets
of soft dove coutille with whalebone busk to the diamondtrimmed pelvis,
the absolute outside edge, while your figure,
plumper than when at large, will be restrained in nettight frocks,
pretty two ounce petticoats and fringes and things stamped,
of course, with my houseflag,
\gab{2980}
creations of lovely lingerie for Alice and nice scent for Alice.
Alice will feel the pullpull.
Martha and Mary will be a little chilly at first
in such delicate thighcasing but the frilly flimsiness of lace
round your bare knees will remind you ...

\Bloom:
\stage{(charming soubrette with dauby cheeks, mustard hair
and large male hands and nose, leering mouth)}
I tried her things on only twice, a small prank, in Holles street.
When we were hard up I washed them to save the laundry bill.
My own shirts I turned.
It was the purest thrift.

\Bello:
\stage{(jeers)}
Little jobs that make mother pleased, eh?
\gab{2990}
And showed off coquettishly in your domino at the mirror
behind closedrawn blinds your unskirted thighs and hegoat's udders
in various poses of surrender, eh?
Ho! Ho!
I have to laugh!
That secondhand black operatop shift and short trunkleg naughties
all split up the stitches at her last rape
that Mrs~Miriam Dandrade sold you from the Shelbourne hotel, eh?

\Bloom:
Miriam.
Black.
Demimondaine.

\Bello:
\stage{(guffaws)}
Christ Almighty it's too tickling, this!
You were a nicelooking Miriam when you clipped off your backgate hairs
\gab{3000}
and lay swooning in the thing across the bed as Mrs~Dandrade
about to be violated by lieutenant Smythe-Smythe, Mr~Philip Augustus Blockwell~M.P.,
signor Laci Daremo, the robust tenor, blueeyed Bert, the liftboy,
Henri Fleury of Gordon Bennett fame, Sheridan, the quadroon Croesus,
the varsity wetbob eight from old Trinity, Ponto, her splendid Newfoundland
and Bobs, dowager duchess of Manorhamilton.
\stage{(he guffaws again)}
Christ, wouldn't it make a Siamese cat laugh?

\Bloom:
\stage{(her hands and features working)}
It was Gerald converted me to be a true corsetlover
\gab{3010}
when I was female impersonator in the High School play \Title{Vice Versa}.
It was dear Gerald.
He got that kink, fascinated by sister's stays.
Now dearest Gerald uses pinky greasepaint and gilds his eyelids.
Cult of the beautiful.

\Bello:
\stage{(with wicked glee)}
Beautiful!
Give us a breather!
When you took your seat with womanish care,
lifting your billowy flounces, on the smoothworn throne.

\Bloom:
Science.
To compare the various joys we each enjoy.
\stage{(earnestly)}
And really it's better the position ...
\gab{3020}
because often I used to wet ...

\Bello:
\stage{(sternly)}
No insubordination!
The sawdust is there in the corner for you.
I gave you strict instructions, didn't I?
Do it standing, sir!
I'll teach you to behave like a jinkleman!
If I catch a trace on your swaddles.
Aha!
By the ass of the Dorans you'll find I'm a martinet.
The sins of your past are rising against you.
Many. Hundreds.

\SinsPast:
\stage{(in a medley of voices)}
He went through a form of clandestine marriage
with at least one woman in the shadow of the Black church.
Unspeakable messages he telephoned mentally to Miss Dunn
\gab{3030}
at an address in D'Olier street
while he presented himself indecently to the instrument in the callbox.
By word and deed he frankly encouraged a nocturnal strumpet
to deposit fecal and other matter
in an unsanitary outhouse attached to empty premises.
In five public conveniences he wrote pencilled messages
offering his nuptial partner to all strongmembered males.
And by the offensively smelling vitriol works
did he not pass night after night
by loving courting couples
to see if and what and how much he could see?
Did he not lie in bed, the gross boar,
gloating over a nauseous fragment of wellused toilet paper
presented to him by a nasty harlot,
stimulated by gingerbread and a postal order?%
\gab{3040}

\Bello:
\stage{(whistles loudly)}
Say!
What was the most revolting piece of obscenity in all your career of crime?
Go the whole hog.
Puke it out!
Be candid for once.

\stage{(Mute inhuman faces throng forward, leering, vanishing, gibbering, Booloohoom,
Poldy Kock, Bootlaces a penny, Cassidy's hag, blind stripling,
Larry rhinoceros, the girl, the woman, the whore, the other, the ...)}

\Bloom:
Don't ask me!
Our mutual faith.
Pleasants street.
I only thought the half of the ...
I swear on my sacred oath ...%
\gab{3050}

\Bello:
\stage{(peremptorily)}
Answer.
Repugnant wretch!
I insist on knowing.
Tell me something to amuse me, smut or a bloody good ghoststory
or a line of poetry, quick, quick, quick!
Where? How?
What time?
With how many?
I give you just three seconds.
One! Two! Thr ...

\Bloom:
\stage{(docile, gurgles)}
I rererepugnosed in rerererepugnant

\Bello:
\stage{(imperiously)}
O, get out, you skunk!
Hold your tongue!
Speak when you're spoken to.%
\gab{3060}

\Bloom:
\stage{(bows)}
Master!
Mistress!
Mantamer!

\stage{(He lifts his arms.
His bangle bracelets fill.)}

\Bello:
\stage{(satirically)}
By day you will souse and bat our smelling underclothes
also when we ladies are unwell, and swab out our latrines
with dress pinned up and a dishclout tied to your tail.
Won't that be nice?
\stage{(he places a ruby ring on her finger)}
And there now!
With this ring I thee own.
Say, \emph{thank you, mistress}.

\Bloom:
\gab{3070}
Thank you, mistress.

\Bello:
You will make the beds, get my tub ready,
empty the pisspots in the different rooms,
including old Mrs~Keogh's the cook's, a sandy one.
Ay, and rinse the seven of them well, mind, or lap it up like champagne.
Drink me piping hot.
Hop!
You will dance attendance or I'll lecture you on your misdeeds,
Miss Ruby, and spank your bare bot right well, miss,
with the hairbrush.
You'll be taught the error of your ways.
At night your wellcreamed braceletted hands will wear fortythreebutton gloves
newpowdered with talc and having delicately scented fingertips.
\gab{3080}
For such favours knights of old laid down their lives.
\stage{(he chuckles)}
My boys will be no end charmed to see you so ladylike,
the colonel, above all,
when they come here the night before the wedding
to fondle my new attraction in gilded heels.
First I'll have a go at you myself.
A man I know on the turf named Charles Alberta Marsh
(I was in bed with him just now
and another gentleman out of the Hanaper and Petty Bag office)
is on the lookout for a maid of all work at a short knock.
Swell the bust.
Smile.
Droop shoulders.
What offers?
\stage{(he points)}
For that lot.
Trained by owner to fetch and carry, basket in mouth.
\stage{(he bares his arm and plunges it elbowdeep in Bloom's vulva)}
There's fine depth for you!
\gab{3090}
What, boys?
That give you a hardon?
\stage{(he shoves his arm in a bidder's face)}
Here wet the deck and wipe it round!

\Bidder:
A florin.

\stage{(Dillon's lacquey rings his handbell.)}

\Lacquey:
Barang!

\Voice:
One and eightpence too much.

\Marsh:
Must be virgin.
\gab{3100}
Good breath.
Clean.

\Bello:
\stage{(gives a rap with his gavel)}
Two bar.
Rockbottom figure and cheap at the price.
Fourteen hands high.
Touch and examine his points.
Handle him.
This downy skin, these soft muscles, this tender flesh.
If I had only my gold piercer here!
And quite easy to milk.
Three newlaid gallons a day.
A pure stockgetter, due to lay within the hour.
His sire's milk record was a thousand gallons of whole milk in forty weeks.
Whoa my jewel!
Beg up!
Whoa!
\stage{(he brands his initial C on Bloom's croup)}
So!
Warranted Cohen!
What advance on two bob, gentlemen?

\DarkvisagedMan:
\gab{3110}
\stage{(in disguised accent)}
Hoondert punt sterlink.

\Voices:
\stage{(subdued)}
For the Caliph.
Haroun Al Raschid.

\Bello:
\stage{(gaily)}
Right.
Let them all come.
The scanty, daringly short skirt,
riding up at the knee to show a peep of white pantalette,
is a potent weapon and transparent stockings, emeraldgartered,
with the long straight seam trailing up beyond the knee,
appeal to the better instincts of the \french{blasé} man about town.
Learn the smooth mincing walk on four inch Louis Quinze heels,
the Grecian bend with provoking croup,
\gab{3120}
the thighs fluescent, knees modestly kissing.
Bring all your powers of fascination to bear on them.
Pander to their Gomorrahan vices.

\Bloom:
\stage{(bends his blushing face into his armpit
and simpers with forefinger in mouth)}
O, I know what you're hinting at now!

\Bello:
What else are you good for, an impotent thing like you?
\stage{(he stoops and, peering, pokes with his fan rudely
under the fat suet folds of Bloom's haunches)}
Up! Up!
Manx cat!
What have we here?
Where's your curly teapot gone to or who docked it on you, cockyolly?
Sing, birdy, sing.
\gab{3130}
It's as limp as a boy of six's doing his pooly behind a cart.
Buy a bucket or sell your pump.
\stage{(loudly)}
Can you do a man's job?

\Bloom:
Eccles street ...

\Bello:
\stage{(sarcastically)}
I wouldn't hurt your feelings for the world
but there's a man of brawn in possession there.
The tables are turned, my gay young fellow!
He is something like a fullgrown outdoor man.
Well for you, you muff, if you had that weapon
with knobs and lumps and warts all over it.
He shot his bolt, I can tell you!
\gab{3140}
Foot to foot, knee to knee, belly to belly, bubs to breast!
He's no eunuch.
A shock of red hair he has sticking out of him behind like a furzebush!
Wait for nine months, my lad!
Holy ginger, it's kicking and coughing up and down in her guts already!
That makes you wild, don't it?
Touches the spot?
\stage{(he spits in contempt)}
Spittoon!

\Bloom:
I was indecently treated, I ...
Inform the police.
Hundred pounds.
Unmentionable.
I ...

\Bello:
Would if you could, lame duck.
A downpour we want not your drizzle.

\Bloom:
\gab{3150}
To drive me mad!
Moll!
I forgot!
Forgive!
Moll ...
We ...
Still ...

\Bello:
\stage{(ruthlessly)}
No, Leopold Bloom, all is changed by woman's will
since you slept horizontal in Sleepy Hollow your night of twenty years.
Return and see.

\stage{(Old Sleepy Hollow calls over the wold.)}

\SleepyHollow:
Rip van Wink!
Rip van Winkle!

\Bloom:
\stage{(in tattered mocassins with a rusty fowlingpiece,
\gab{3160}
tiptoeing, fingertipping,
his haggard bony bearded face peering through the diamond panes, cries out)}
I see her!
It's she!
The first night at Mat Dillon's!
But that dress, the green!
And her hair is dyed gold and he ...

\Bello:
\stage{(laughs mockingly)}
That's your daughter, you owl, with a Mullingar student.

\stage{(Milly Bloom, fairhaired, greenvested, slimsandalled,
her blue scarf in the seawind simply swirling,
breaks from the arms of her lover and calls,
her young eyes wonderwide.)}

\Milly:
\gab{3170}
My!
It's Papli!
But, O Papli, how old you've grown!

\Bello:
Changed, eh?
Our whatnot, our writingtable where we never wrote,
aunt Hegarty's armchair, our classic reprints of old masters.
A man and his menfriends are living there in clover.
The \Title{Cuckoos' Rest}!
Why not?
How many women had you, eh, following them up dark streets, flatfoot,
exciting them by your smothered grunts, what, you male prostitute?
Blameless dames with parcels of groceries.
Turn about.
Sauce for the goose, my gander~O.

\Bloom:
\gab{3180}
They ... I~...

\Bello:
\stage{(cuttingly)}
Their heelmarks will stamp the Brusselette carpet you bought at Wren's auction.
In their horseplay with Moll the romp to find the buck flea in her breeches
they will deface the little statue
you carried home in the rain for art for art' sake.
They will violate the secrets of your bottom drawer.
Pages will be torn from your handbook of astronomy to make them pipespills.
And they will spit in your ten shilling brass fender from Hampton Leedom's.

\Bloom:
\gab{3190}
Ten and six.
The act of low scoundrels.
Let me go.
I will return.
I will prove ...

\Voice:
Swear!

\stage{(Bloom clenches his fists and crawls forward,
a bowieknife between his teeth.)}

\Bello:
As a paying guest or a kept man?
Too late.
You have made your secondbest bed and others must lie in it.
Your epitaph is written.
You are down and out and don't you forget it, old bean.%
\gab{3200}

\Bloom:
Justice!
All Ireland versus one!
Has nobody ...?
\stage{(he bites his thumb)}

\Bello:
Die and be damned to you if you have any sense of decency or grace about you.
I can give you a rare old wine that'll send you skipping to hell and back.
Sign a will and leave us any coin you have!
If you have none see you damn well get it, steal it, rob it!
We'll bury you in our shrubbery jakes where you'll be dead and dirty with old Cuck Cohen,
my stepnephew I married, the bloody old gouty procurator and sodomite
with a crick in his neck, and my other ten or eleven husbands,
\gab{3210}
whatever the buggers' names were, suffocated in the one cesspool.
\stage{(he explodes in a loud phlegmy laugh)}
We'll manure you, Mr~Flower!
\stage{(he pipes scoffingly)}
Byby, Poldy!
Byby, Papli!

\Bloom:
\stage{(clasps his head)}
My willpower!
Memory!
I have sinned!
I have suff ...
\stage{(he weeps tearlessly)}

\Bello:
\stage{(sneers)}
Crybabby!
Crocodile tears!

\stage{(Bloom, broken, closely veiled for the sacrifice, sobs, his face to the earth.
\gab{3220}
The passing bell is heard.
Darkshawled figures of the circumcised,
in sackcloth and ashes, stand by the wailing wall.
M. Shulomowitz, Joseph Goldwater, Moses Herzog, Harris Rosenberg,
M. Moisel, J. Citron, Minnie Watchman,
P. Mastiansky, the reverend Leopold Abramovitz, chazen.
With swaying arms they wail in pneuma over the recreant Bloom.)}

\Circumcised:
\stage{(in dark guttural chant as they cast dead sea fruit upon him, no flowers)}
\hebrew{Shema Israel Adonai Elohenu Adonai Echad.}

\Voices:
\stage{(sighing)}
\gab{3230}
So he's gone.
Ah yes.
Yes, indeed.
Bloom?
Never heard of him.
No?
Queer kind of chap.
There's the widow.
That so?
Ah, yes.

\stage{(From the suttee pyre the flame of gum camphire ascends.
The pall of incense smoke screens and disperses.
Out of her oakframe a nymph with hair unbound,
lightly clad in teabrown artcolours, descends from her grotto
and passing under interlacing yews stands over Bloom.)}

\Yews:
\stage{(their leaves whispering)}
Sister.
Our sister.
Ssh!

\Nymph:
\stage{(softly)}
\gab{3240}
Mortal!
\stage{(kindly)}
Nay, dost not weepest!

\Bloom:
\stage{(crawls jellily forward under the boughs,
streaked by sunlight, with dignity)}
This position.
I felt it was expected of me.
Force of habit.

\Nymph:
Mortal!
You found me in evil company, highkickers,
coster picnicmakers, pugilists, popular generals,
immoral panto boys in fleshtights and the nifty shimmy dancers,
La Aurora and Karini, musical act, the hit of the century.
I was hidden in cheap pink paper that smelt of rock oil.
I was surrounded by the stale smut of clubmen,
stories to disturb callow youth, ads for transparencies,
truedup dice and bustpads, proprietary articles
\gab{3250}
and why wear a truss with testimonial from ruptured gentleman.
Useful hints to the married.

\Bloom:
\stage{(lifts a turtle head towards her lap)}
We have met before.
On another star.

\Nymph:
\stage{(sadly)}
Rubber goods.
Neverrip brand as supplied to the aristocracy.
Corsets for men.
I cure fits or money refunded.
Unsolicited testimonials for Professor Waldmann's wonderful chest exuber.
My bust developed four inches in three weeks, reports Mrs~Gus Rublin with photo.

\Bloom:
\gab{3260}
You mean \Title{Photo Bits}?

\Nymph:
I do.
You bore me away, framed me in oak and tinsel,
set me above your marriage couch.
Unseen, one summer eve, you kissed me in four places.
And with loving pencil you shaded my eyes, my bosom and my shame.

\Bloom:
\stage{(humbly kisses her long hair)}
Your classic curves, beautiful immortal,
I was glad to look on you, to praise you, a thing of beauty,
almost to pray.

\Nymph:
During dark nights I heard your praise.%
\gab{3270}

\Bloom:
\stage{(quickly)}
Yes, yes.
You mean that I ...
Sleep reveals the worst side of everyone, children perhaps excepted.
I know I fell out of bed or rather was pushed.
Steel wine is said to cure snoring.
For the rest there is that English invention,
pamphlet of which I received some days ago, incorrectly addressed.
It claims to afford a noiseless, inoffensive vent.
\stage{(he sighs)}
'Twas ever thus.
Frailty, thy name is marriage.

\Nymph:
\stage{(her fingers in her ears)}
And words.
They are not in my dictionary.

\Bloom:
\gab{3280}
You understood them?

\Yews:
Ssh!

\Nymph:
\stage{(covers her face with her hands)}
What have I not seen in that chamber?
What must my eyes look down on?

\Bloom:
\stage{(apologetically)}
I know.
Soiled personal linen, wrong side up with care.
The quoits are loose.
From Gibraltar by long sea long ago.

\Nymph:
\gab{3290}
\stage{(bends her head)}
Worse, worse!

\Bloom:
\stage{(reflects precautiously)}
That antiquated commode.
It wasn't her weight.
She scaled just eleven stone nine.
She put on nine pounds after weaning.
It was a crack and want of glue.
Eh?
And that absurd orangekeyed utensil which has only one handle.

\stage{(The sound of a waterfall is heard in bright cascade.)}

\Waterfall:
\begin{verse}
    Poulaphouca Poulaphouca\\
\gab{3300}%
    Poulaphouca Poulaphouca.
\end{verse}

\Yews:
\stage{(mingling their boughs)}
Listen.
Whisper.
She is right, our sister.
We grew by Poulaphouca waterfall.
We gave shade on languorous summer days.

\JohnWyse:
\stage{(in the background, in Irish National Forester's uniform,
doffs his plumed hat)}
Prosper!
Give shade on languorous days, trees of Ireland!

\Yews:
\stage{(murmuring)}
Who came to Poulaphouca with the High School excursion?
Who left his nutquesting classmates to seek our shade?

\Bloom:
\gab{3310}
\stage{(scared)}
High School of Poula?
Mnemo?
Not in full possession of faculties.
Concussion.
Run over by tram.

\Echo:
Sham!

\Bloom:
\stage{(pigeonbreasted, bottleshouldered, padded,
in nondescript juvenile grey and black striped suit,
too small for him, white tennis shoes,
bordered stockings with turnover tops and a red schoolcap with badge)}
I was in my teens, a growing boy.
A little then sufficed, a jolting car,
the mingling odours of the ladies' cloakroom and lavatory,
\gab{3320}
the throng penned tight on the old Royal stairs
(for they love crushes, instinct of the herd,
and the dark sexsmelling theatre unbridles vice),
even a pricelist of their hosiery.
And then the heat.
There were sunspots that summer.
End of school.
And tipsycake.
Halcyon days.

\stage{(Halcyon Days, High School boys in blue and white football jerseys and shorts,
Master Donald Turnbull, Master Abraham Chatterton, Master Owen Goldberg,
Master Jack Meredith, Master Percy Apjohn,
stand in a clearing of the trees and shout to Master Leopold Bloom.)}

\HalcyonDays:
\gab{3330}
Mackerel!
Live us again.
Hurray!
\stage{(they cheer)}

\Bloom:
\stage{(hobbledehoy, warmgloved, mammamufflered,
starred with spent snowballs, struggles to rise)}
Again!
I feel sixteen!
What a lark!
Let's ring all the bells in Montague street.
\stage{(he cheers feebly)}
Hurray for the High School!

\Echo:
Fool!

\Yews:
\stage{(rustling)}
She is right, our sister.
Whisper.
\gab{3340}
\stage{(Whispered kisses are heard in all the wood.
Faces of hamadryads peep out from the boles
and among the leaves and break, blossoming into bloom.)}
Who profaned our silent shade?

\Nymph:
\stage{(coyly, through parting fingers)}
There?
In the open air?

\Yews:
\stage{(sweeping downward)}
Sister, yes.
And on our virgin sward.

\Waterfall:
\begin{verse}
    Poulaphouca Poulaphouca\\
    Phoucaphouca Phoucaphouca.
\end{verse}

\Nymph:
\gab{3350}
\stage{(with wide fingers)}
O, infamy!

\Bloom:
I was precocious.
Youth.
The fauna.
I sacrificed to the god of the forest.
The flowers that bloom in the spring.
It was pairing time.
Capillary attraction is a natural phenomenon.
Lotty Clarke, flaxenhaired,
I saw at her night toilette through illclosed curtains
with poor papa's operaglasses:
The wanton ate grass wildly.
She rolled downhill at Rialto bridge
to tempt me with her flow of animal spirits.
She climbed their crooked tree and I ...
A saint couldn't resist it.
The demon possessed me.
Besides, who saw?

\stage{(Staggering Bob, a whitepolled calf,
\gab{3360}
thrusts a ruminating head with humid nostrils through the foliage.)}

\StaggeringBob:
\stage{(large teardrops rolling from his prominent eyes, snivels)}
Me. Me see.

\Bloom:
Simply satisfying a need I ...
\stage{(with pathos)}
No girl would when I went girling.
Too ugly.
They wouldn't play ...

\stage{(High on Ben Howth through rhododendrons a nannygoat passes,
plumpuddered, buttytailed, dropping currants.)}

\Nannygoat:
\stage{(bleats)}
\gab{3370}
Megeggaggegg!
Nannannanny!

\Bloom:
\stage{(hatless, flushed, covered with burrs of thistledown and gorsespine)}
Regularly engaged.
Circumstances alter cases.
\stage{(he gazes intently downwards on the water)}
Thirtytwo head over heels per second.
Press nightmare.
Giddy Elijah.
Fall from cliff.
Sad end of government printer's clerk.

\stage{(Through silversilent summer air the dummy of Bloom, rolled in a mummy,
rolls roteatingly from the Lion's Head cliff into the purple waiting waters.)}

\Dummymummy:
\gab{3380}
Bbbbblllllblblblblobschbg!

\stage{(Far out in the bay between Bailey and Kish Lights
the \title{Erin's King} sails,
sending a broadening plume of coalsmoke from her funnel towards the land.)}

\CouncillorNannetii:
\stage{(alone on deck, in dark alpaca, yellowkitefaced,
his hand in his waistcoat opening, declaims)}
When my country takes her place among the nations of the earth,
then, and not till then, let my epitaph be written.
I have ...

\Bloom:
Done.
Prff!
\gab{3390}

\Nymph:
\stage{(loftily)}
We immortals, as you saw today,
have not such a place and no hair there either.
We are stonecold and pure.
We eat electric light.
\stage{(she arches her body in lascivious crispation,
placing her forefinger in her mouth)}
Spoke to me.
Heard from behind.
How then could you ...?

\Bloom:
\stage{(pawing the heather abjectly)}
O, I have been a perfect pig.
Enemas too I have administered.
One third of a pint of quassia to which add a tablespoonful of rocksalt.
Up the fundament.
With Hamilton Long's syringe, the ladies' friend.%
\gab{3400}

\Nymph:
In my presence.
The powderpuff.
\stage{(she blushes and makes a knee)}
And the rest!

\Bloom:
\stage{(dejected)}
Yes.
\latin{Peccavi!}
I have paid homage on that living altar where the back changes name.
\stage{(with sudden fervour)}
For why should the dainty scented jewelled hand, the hand that rules ...?

\stage{(Figures wind serpenting in slow woodland pattern
around the treestems, cooeeing)}

\VoiceKitty:
\gab{3410}
\stage{(in the thicket)}
Show us one of them cushions.

\VoiceFlorry:
Here.

\stage{(A grouse wings clumsily through the underwood.)}

\VoiceLynch:
\stage{(in the thicket)}
Whew!
Piping hot!

\VoiceZoe:
\stage{(from the thicket)}
Came from a hot place.

\VoiceVirag:
\gab{3420}%
\stage{(a birdchief, bluestreaked and feathered in war panoply with his assegai,
striding through a crackling canebrake over beechmast and acorns)}
Hot! Hot!
Ware Sitting Bull!

\Bloom:
It overpowers me.
The warm impress of her warm form.
Even to sit where a woman has sat, especially with divaricated thighs,
as though to grant the last favours,
most especially with previously well uplifted white sateen coatpans.
So womanly, full.
It fills me full.

\Waterfall:
\begin{verse}
    Phillaphulla Poulaphouca\\
\gab{3430}%
    Poulaphouca Poulaphouca.
\end{verse}

\Yews:
Ssh!
Sister, speak!

\Nymph:
\stage{(eyeless, in nun's white habit, coif and hugewinged wimple,
softly, with remote eyes)}
Tranquilla convent.
Sister Agatha.
Mount Carmel.
The apparitions of Knock and Lourdes.
No more desire.
\stage{(she reclines her head, sighing)}
Only the ethereal.
Where dreamy creamy gull waves o'er the waters dull.

\stage{(Bloom half rises.
His back trouserbutton snaps.)}

\Button:
\gab{3440}
Bip!

\stage{(Two sluts of the Coombe dance rainily by, shawled, yelling flatly.)}

\Sluts:
\begin{verse}
    O, Leopold lost the pin of his drawers\\
    He didn't know what to do,\\
    To keep it up,\\
    To keep it up.
\end{verse}


\Bloom:
\stage{(coldly)}
You have broken the spell.
The last straw.
If there were only ethereal where would you all be, postulants and novices?
\gab{3450}
Shy but willing like an ass pissing.

\Yews:
\stage{(their silverfoil of leaves precipitating,
their skinny arms aging and swaying)}
Deciduously!

\Nymph:
\stage{(her features hardening, gropes in the folds of her habit)}
Sacrilege!
To attempt my virtue!
\stage{(a large moist stain appears on her robe)}
Sully my innocence!
You are not fit to touch the garment of a pure woman.
\stage{(she clutches again in her robe)}
Wait.
Satan, you'll sing no more lovesongs.
Amen. Amen. Amen. Amen.
\gab{3460}
\stage{(she draws a poniard and,
clad in the sheathmail of an elected knight of nine,
strikes at his loins)}
Nekum!

\Bloom:
\stage{(starts up, seizes her hand)}
Hoy!
Nebrakada!
Cat o' nine lives!
Fair play, madam.
No pruningknife.
The fox and the grapes, is it?
What do you lack with your barbed wire?
Crucifix not thick enough?
\stage{(he clutches her veil)}
A holy abbot you want or Brophy, the lame gardener,
or the spoutless statue of the watercarrier,
or good mother Alphonsus, eh Reynard?

\Nymph:
\stage{(with a cry flees from him unveiled, her plaster cast cracking,
a cloud of stench escaping from the cracks)}
\gab{3470}
Poli ...!

\Bloom:
\stage{(calls after her)}
As if you didn't get it on the double yourselves.
No jerks and multiple mucosities all over you.
I tried it.
Your strength our weakness.
What's our studfee?
What will you pay on the nail?
You fee mendancers on the Riviera, I read.
\stage{(the fleeing nymph raises a keen)}
Eh?
I have sixteen years of black slave labour behind me.
And would a jury give me five shillings alimony tomorrow, eh?
Fool someone else, not me.
\stage{(he sniffs)}
Rut. Onions. Stale. Sulphur. Grease.

\stage{(The figure of Bella Cohen stands before him.)}

\Bella:
\gab{3480}
You'll know me the next time.

\Bloom:
\stage{(composed, regards her)}
\french{Passée}.
Mutton dressed as lamb.
Long in the tooth and superfluous hair.
A raw onion the last thing at night would benefit your complexion.
And take some double chin drill.
Your eyes are as vapid as the glasseyes of your stuffed fox.
They have the dimensions of your other features, that's all.
I'm not a triple screw propeller.

\Bella:
\stage{(contemptuously)}
You're not game, in fact.
\stage{(her sowcunt barks)}
Fbhracht!%
\gab{3490}

\Bloom:
\stage{(contemptuously)}
Clean your nailless middle finger first,
your bully's cold spunk is dripping from your cockscomb.
Take a handful of hay and wipe yourself.

\Bella:
I know you, canvasser!
Dead cod!

\Bloom:
I saw him, kipkeeper!
Pox and gleet vendor!

\Bella:
\stage{(turns to the piano)}
\gab{3500}
Which of you was playing the dead march from \Title{Saul}?

\Zoe:
Me.
Mind your cornflowers.
\stage{(she darts to the piano and bangs chords on it with crossed arms)}
The cat's ramble through the slag.
\stage{(she glances back)}
Eh?
Who's making love to my sweeties?
\stage{(she darts back to the table)}
What's yours is mine and what's mine is my own.

\stage{(Kitty, disconcerted, coats her teeth with the silver paper.
Bloom approaches Zoe.)}

\Bloom:
\stage{(gently)}
Give me back that potato, will you?

\Zoe:
\gab{3510}
Forfeits, a fine thing and a superfine thing.

\Bloom:
\stage{(with feeling)}
It is nothing, but still, a relic of poor mamma.

\Zoe:
\begin{verse}
    Give a thing and take it back\\
    God'll ask you where is that\\
    You'll say you don't know\\
    God'll send you down below.
\end{verse}

\Bloom:
There is a memory attached to it.
\gab{3520}
I should like to have it.

\Stephen:
To have or not to have that is the question.

\Zoe:
Here.
\stage{(she hauls up a reef of her slip, revealing her bare thigh,
and unrolls the potato from the top of her stocking)}
Those that hides knows where to find.

\Bella:
\stage{(frowns)}
Here.
This isn't a musical peepshow.
And don't you smash that piano.
Who's paying here?

\stage{(She goes to the pianola.
\gab{3530}
Stephen fumbles in his pocket and,
taking out a banknote by its corner, hands it to her.)}

\Stephen:
\stage{(with exaggerated politeness)}
This silken purse I made out of the sow's ear of the public.
Madam, excuse me.
If you allow me.
\stage{(he indicates vaguely Lynch and Bloom)}
We are all in the same sweepstake, Kinch and Lynch.
\french{Dans ce bordel où tenons nostre état.}

\Lynch:
\stage{(calls from the hearth)}
Dedalus!
Give her your blessing for me.

\Stephen:
\stage{(hands Bella a coin)}
\gab{3540}
Gold.
She has it.

\Bella:
\stage{(looks at the money, then at Stephen, then at Zoe, Florry and Kitty)}
Do you want three girls?
It's ten shillings here.

\Stephen:
\stage{(delightedly)}
A hundred thousand apologies.
\stage{(he fumbles again and takes out and hands her two crowns)}
Permit, \french{brevi manu}, my sight is somewhat troubled.

\stage{(Bella goes to the table to count the money
while Stephen talks to himself in monosyllables.
Zoe bends over the table.
Kitty leans over Zoe's neck.
\gab{3550}
Lynch gets up, rights his cap and, clasping Kitty's waist,
adds his head to the group.)}

\Florry:
\stage{(strives heavily to rise)}
Ow!
My foot's asleep.
\stage{(She limps over to the table.
Bloom approaches.)}

\BZKLaB:
\stage{(chattering and squabbling)}
The gentleman ...
ten shillings ...
paying for the three ...
allow me a moment ...
this gentleman pays separate ...
who's touching it?
... ow! ...
mind who you're pinching ...
are you staying the night or a short time?...
who did?...
you're a liar, excuse me ...
the gentleman paid down like a gentleman ...
\gab{3560}
drink ...
it's long after eleven.

\Stephen:
\stage{(at the pianola, making a gesture of abhorrence)}
No bottles!
What, eleven?
A riddle!

\Zoe:
\stage{(lifting up her pettigown and folding a half sovereign
into the top of her stocking)}
Hard earned on the flat of my back.

\Lynch:
\stage{(lifting Kitty from the table)}
Come!

\Kitty:
Wait.
\gab{3570}
\stage{(she clutches the two crowns)}

\Florry:
And me?

\Lynch:
Hoopla!
\stage{(he lifts her, carries her and bumps her down on the sofa.)}

\Stephen:
\begin{verse}
    The fox crew, the cocks flew,\\
    The bells in heaven\\
    Were striking eleven.\\
\gab{3580}%
    'Tis time for her poor soul\\
    To get out of heaven.
\end{verse}

\Bloom:
\stage{(quietly lays a half sovereign on the table between Bella and Florry)}
So. Allow me.
\stage{(he takes up the poundnote)}
Three times ten.
We're square.

\Bella:
\stage{(admiringly)}
You're such a slyboots, old cocky.
I could kiss you.

\Zoe:
\stage{(points)}
Him?
Deep as a drawwell.
\stage{(Lynch bends Kitty back over the sofa and kisses her.
Bloom goes with the poundnote to Stephen.)%
\gab{3590}
}

\Bloom:
This is yours.

\Stephen:
How is that?
The distrait or absentminded beggar.
\stage{(He fumbles again in his pocket and draws out a handful of coins.
An object falls.)}
That fell.

\Bloom:
\stage{(stooping, picks up and hands a box of matches)}
This.

\Stephen:
Lucifer.
Thanks.

\Bloom:
\gab{3600}
\stage{(quietly)}
You had better hand over that cash to me to take care of.
Why pay more?

\Stephen:
\stage{(hands him all his coins)}
Be just before you are generous.

\Bloom:
I will but is it wise?
\stage{(he counts)}
One, seven, eleven, and five.
Six. Eleven.
I don't answer for what you may have lost.

\Stephen:
Why striking eleven?
Proparoxyton.
Moment before the next Lessing says.
Thirsty fox.
\gab{3610}
\stage{(he laughs loudly)}
Burying his grandmother.
Probably he killed her.

\Bloom:
That is one pound six and eleven.
One pound seven, say.

\Stephen:
Doesn't matter a rambling damn.

\Bloom:
No, but ...

\Stephen:
\stage{(comes to the table)}
Cigarette, please.
\stage{(Lynch tosses a cigarette from the sofa to the table)}
\gab{3620}
And so Georgina Johnson is dead and married.
\stage{(A cigarette appears on the table.
Stephen looks at it)}
Wonder.
Parlour magic.
Married.
Hm.
\stage{(he strikes a match and proceeds
to light the cigarette with enigmatic melancholy)}

\Lynch:
\stage{(watching him)}
You would have a better chance of lighting it
if you held the match nearer.

\Stephen:
\stage{(brings the match near his eye)}
Lynx eye.
Must get glasses.
Broke them yesterday.
Sixteen years ago.
Distance.
The eye sees all flat.
\stage{(He draws the match away.
It goes out.)}
\gab{3630}
Brain thinks.
Near: far.
Ineluctable modality of the visible.
\stage{(he frowns mysteriously)}
Hm.
Sphinx.
The beast that has twobacks at midnight.
Married.

\Zoe:
It was a commercial traveller married her and took her away with him.

\Florry:
\stage{(nods)}
Mr~Lambe from London.

\Stephen:
Lamb of London, who takest away the sins of our world.

\Lynch:
\stage{(embracing Kitty on the sofa, chants deeply)}
\gab{3640}
\latin{Dona nobis pacem.}

\stage{(The cigarette slips from Stephen's fingers.
Bloom picks it up and throws it in the grate.)}

\Bloom:
Don't smoke.
You ought to eat.
Cursed dog I met.
\stage{(to Zoe)}
You have nothing?

\Zoe:
Is he hungry?

\Stephen:
\stage{(extends his hand to her smiling
and chants to the air of the bloodoath in the \title{Dusk of the Gods})%
\gab{3650}
}

\begin{verse}
\german{%
    Hangende Hunger,\\
    Fragende Frau,\\
    Macht uns alle kaputt.}
\end{verse}


\Zoe:
\stage{(tragically)}
Hamlet, I am thy father's gimlet!
\stage{(she takes his hand)}
Blue eyes beauty I'll read your hand.
\stage{(she points to his forehead)}
No wit, no wrinkles.
\stage{(she counts)}
Two, three, Mars, that's courage.
\stage{(Stephen shakes his head)}
No kid.

\Lynch:
Sheet lightning courage.
The youth who could not shiver and shake.
\gab{3660}
\stage{(to Zoe)}
Who taught you palmistry?

\Zoe:
\stage{(turns)}
Ask my ballocks that I haven't got.
\stage{(to Stephen)}
I see it in your face.
The eye, like that.
\stage{(she frowns with lowered head)}

\Lynch:
\stage{(laughing, slaps Kitty behind twice)}
Like that.
Pandybat.

\stage{(Twice loudly a pandybat cracks, the coffin of the pianola flies open,
the bald little round Jack-in-the-box head of Father Dolan springs up.)}

\FatherDolan:
\gab{3670}
Any boy want flogging?
Broke his glasses?
Lazy idle little schemer.
See it in your eye.

\stage{(mild, benign, rectorial, reproving, the head of Don John Conmee
rises from the pianola coffin.)}

\DonJohnConmee:
Now, Father Dolan!
Now.
I'm sure that Stephen is a very good little boy!

\Zoe:
\stage{(examining Stephen's palm)}
Woman's hand.

\Stephen:
\stage{(murmurs)}
\gab{3680}
Continue.
Lie.
Hold me. Caress.
I never could read His handwriting except His criminal thumbprint on the haddock.

\Zoe:
What day were you born?

\Stephen:
Thursday.
Today.

\Zoe:
Thursday's child has far to go.
\stage{(she traces lines on his hand)}
Line of fate.
Influential friends.

\Florry:
\stage{(pointing)}
\gab{3690}
Imagination.

\Zoe:
Mount of the moon.
You'll meet with a ...
\stage{(she peers at his hands abruptly)}
I won't tell you what's not good for you.
Or do you want to know?

\Bloom:
\stage{(detaches her fingers and offers his palm)}
More harm than good.
Here.
Read mine.

\Bella:
Show.
\stage{(she turns up Bloom's hand)}
I thought so.
Knobby knuckles for the women.

\Zoe:
\gab{3700}
\stage{(peering at Bloom's palm)}
Gridiron.
Travels beyond the sea and marry money.

\Bloom:
Wrong.

\Zoe:
\stage{(quickly)}
O, I see.
Short little finger.
Henpecked husband.
That wrong?

\stage{(Black Liz, a huge rooster hatching in a chalked circle,
rises, stretches her wings and clucks.)}

\BlackLiz:
Gara.
Klook.
Klook.
Klook.
\gab{3710}
\stage{(she sidles from her newlaid egg and waddles off)}

\Bloom:
\stage{(points to his hand)}
That weal there is an accident.
Fell and cut it twentytwo years ago.
I was sixteen.

\Zoe:
I see, says the blind man.
Tell us news.

\Stephen:
See?
Moves to one great goal.
I am twentytwo.
Sixteen years ago he was twentytwo too.
Sixteen years ago I twentytwo tumbled.
Twentytwo years ago he sixteen fell off his hobbyhorse.
\gab{3720}
\stage{(he winces)}
Hurt my hand somewhere.
Must see a dentist.
Money?

\stage{(Zoe whispers to Florry.
They giggle.
Bloom releases his hand and writes idly
on the table in backhand, pencilling slow curves.)}

\Florry:
What?

\stage{(A hackneycar, number Three Hundred and Twentyfour,
with a gallantbuttocked mare,
driven by James Barton, Harmony avenue, Donnybrook, trots past.
Blazes Boylan and Lenehan sprawl swaying on the sideseats.
The Ormond boots crouches behind on the axle.
\gab{3730}
Sadly over the crossblind Lydia Douce and Mina Kennedy gaze.)}

\Boots:
\stage{(jogging, mocks them with thumb and wriggling wormfingers)}
Haw haw have you the horn?

\stage{(Bronze by gold they whisper.)}

\Zoe:
\stage{(to Florry)}
Whisper.

\stage{(They whisper again)}

\stage{(Over the well of the car Blazes Boylan leans,
his boater straw set sideways, a red flower in his mouth.
Lenehan in yachtsman's cap and white shoes
\gab{3740}
officiously detaches a long hair from Blazes Boylan's coat shoulder.)}

\Lenehan:
Ho!
What do I here behold?
Were you brushing the cobwebs off a few quims?

\Boylan:
\stage{(seated, smiles)}
Plucking a turkey.

\Lenehan:
A good night's work.

\Boylan:
\stage{(holding up four thick bluntungulated fingers, winks)}
\gab{3750}
Blazes Kate!
Up to sample or your money back.
\stage{(he holds out a forefinger)}
Smell that.

\Lenehan:
\stage{(smells gleefully)}
Ah!
Lobster and mayonnaise.
Ah!

\ZoeAndFlorry:
\stage{(laugh together)}
Ha ha ha ha.

\Boylan:
\stage{(jumps surely from the car and calls loudly for all to hear)}
Hello, Bloom!
Mrs Bloom dressed yet?

\Bloom:
\stage{(in flunkey's prune plush coat and kneebreeches,
\gab{3760}
buff stockings and powdered wig)}
I'm afraid not, sir.
The last articles ...

\Boylan:
\stage{(tosses him sixpence)}
Here, to buy yourself a gin and splash.
\stage{(he hangs his hat smartly on a peg of Bloom's antlered head)}
Show me in.
I have a little private business with your wife, you understand?

\Bloom:
Thank you, sir.
Yes, sir.
Madam Tweedy is in her bath, sir.

\Marion:
He ought to feel himself highly honoured.
\stage{(she plops splashing out of the water)}
Raoul darling, come and dry me.
I'm in my pelt.
\gab{3770}
Only my new hat and a carriage sponge.

\Boylan:
\stage{(a merry twinkle in his eye)}
Topping!

\Bella:
What?
What is it?

\stage{(Zoe whispers to her.)}

\Marion:
Let him look, the pishogue!
Pimp!
And scourge himself!
I'll write to a powerful prostitute or Bartholomona, the bearded woman,
to raise weals out on him an inch thick
\gab{3780}
and make him bring me back a signed and stamped receipt.

\Boylan:
\stage{(clasps himself)}
Here, I can't hold this little lot much longer.
\stage{(he strides off on stiff cavalry legs)}

\Bella:
\stage{(laughing)}
Ho ho ho ho.

\Boylan:
\stage{(to Bloom, over his shoulder)}
You can apply your eye to the keyhole
and play with yourself while I just go through her a few times.

\Bloom:
\gab{3790}
Thank you, sir.
I will, sir.
May I bring two men chums to witness the deed and take a snapshot?
\stage{(he holds out an ointment jar)}
Vaseline, sir?
Orangeflower ...?
Lukewarm water ...?

\Kitty:
\stage{(from the sofa)}
Tell us, Florry.
Tell us.
What.

\stage{(Florry whispers to her.
Whispering lovewords murmur, liplapping loudly, poppysmic plopslop.)}

\MinaKennedy:
\stage{(her eyes upturned)}
O, it must be like the scent of geraniums and lovely peaches!
\gab{3800}
O, he simply idolises every bit of her!
Stuck together!
Covered with kisses!

\LydiaDouce:
\stage{(her mouth opening)}
Yumyum.
O, he's carrying her round the room doing it!
Ride a cockhorse.
You could hear them in Paris and New York.
Like mouthfuls of strawberries and cream.

\Kitty:
\stage{(laughing)}
Hee hee hee.

\BoylansVoice:
\stage{(sweetly, hoarsely, in the pit of his stomach)}
Ah!
Godblazegrukbrukarchkhrasht!%
\gab{3810}

\MarionsVoice:
\stage{(hoarsely, sweetly, rising to her throat)}
O!
Weeshwashtkissinapooisthnapoohuck?

\Bloom:
\stage{(his eyes wildly dilated, clasps himself)}
Show! Hide! Show!
Plough her!
More!
Shoot!

\BellaZoeFlorryKitty:
Ho ho!
Ha ha!
Hee hee!

\Lynch:
\stage{(points)}
The mirror up to nature.
\stage{(he laughs)}
\gab{3820}
Hu hu hu hu hu!

\stage{(Stephen and Bloom gaze in the mirror.
The face of William Shakespeare,
beardless, appears there, rigid in facial paralysis,
crowned by the reflection of the reindeer antlered hatrack in the hall.)}

\Shakespeare:
\stage{(in dignified ventriloquy)}
'Tis the loud laugh bespeaks the vacant mind.
\stage{(to Bloom)}
Thou thoughtest as how thou wastest invisible.
Gaze.
\stage{(he crows with a black capon's laugh)}
Iagogo!
How my Oldfellow chokit his Thursdaymornun.
Iagogogo!

\Bloom:
\gab{3830}
\stage{(smiles yellowly at the three whores)}
When will I hear the joke?

\Zoe:
Before you're twice married and once a widower.

\Bloom:
Lapses are condoned.
Even the great Napoleon when measurements
were taken next the skin after his death ...

\stage{(Mrs~Dignam, widow woman, her snubnose and cheeks flushed with deathtalk,
tears and Tunney's tawny sherry, hurries by in her weeds,
her bonnet awry, rouging and powdering her cheeks, lips and nose,
a pen chivvying her brood of cygnets.
\gab{3840}
Beneath her skirt appear her late husband's
everyday trousers and turnedup boots, large eights.
She holds a Scottish Widows' insurance policy and a large marquee umbrella
under which her brood run with her,
Patsy hopping on one shod foot,
his collar loose, a hank of porksteaks dangling,
Freddy whimpering,
Susy with a crying cod's mouth,
Alice struggling with the baby.
She cuffs them on, her streamers flaunting aloft.)}

\Freddy:
Ah, ma, you're dragging me along!

\Susy:
\gab{3850}
Mamma, the beeftea is fizzing over!

\Shakespeare:
\stage{(with paralytic rage)}
Weda seca whokilla farst.

\stage{(The face of Martin Cunningham, bearded,
refeatures Shakespeare's beardless face.
The marquee umbrella sways drunkenly, the children run aside.
Under the umbrella appears Mrs~Cunningham in merry widow hat and kimono gown.
She glides sidling and bowing, twirling japanesily.)}

\MrsCunningham:
\stage{(sings)}
\gab{3860}
\begin{verse}
    And they call me the jewel of Asia!
\end{verse}

\MartinCunningham:
\stage{(gazes on her, impassive)}
Immense!
Most bloody awful demirep!

\Stephen:
\latin{Et exaltabuntur cornua iusti.}
Queens lay with prize bulls.
Remember Pasiphae for whose lust my grandoldgrossfather made the first confessionbox.
Forget not Madam Grissel Steevens nor the suine scions of the house of Lambert.
And Noah was drunk with wine.
And his ark was open.

\Bella:
\gab{3870}
None of that here.
Come to the wrong shop.

\Lynch:
Let him alone.
He's back from Paris.

\Zoe:
\stage{(runs to Stephen and links him)}
O go on!
Give us some parleyvoo.

\stage{(Stephen claps hat on head and leaps over to the fireplace
where he stands with shrugged shoulders, finny hands outspread,
a painted smile on his face.)}

\Lynch:
\stage{(pommelling on the sofa)}
\gab{3880}
Rmm Rmm Rmm Rrrrrrmmmm.

\Stephen:
\stage{(gabbles with marionette jerks)}
Thousand places of entertainment to expense your evenings
with lovely ladies saling gloves and other things
perhaps hers heart beerchops perfect fashionable house
very eccentric where lots cocottes beautiful dressed much about princesses
like are dancing cancan and walking there parisian clowneries
extra foolish for bachelors foreigns the same if talking a poor english
how much smart they are on things love and sensations voluptuous.
Misters very selects for is pleasure must to visit heaven and hell show
with mortuary candles and they tears silver which occur every night.
\gab{3890}
Perfectly shocking terrific of religion's things mockery seen in universal world.
All chic womans which arrive full of modesty then disrobe and squeal loud
to see vampire man debauch nun very fresh young with \french{dessous troublants}.
\stage{(he clacks his tongue loudly)}
\french{Ho, là là!
Ce pif qu'il a!}

\Lynch:
\french{Vive le vampire!}

\Whores:
Bravo!
Parleyvoo!

\Stephen:
\stage{(grimacing with head back, laughs loudly, clapping himself)}
\gab{3900}
Great success of laughing.
Angels much prostitutes like and holy apostles big damn ruffians.
\french{Demimondaines} nicely handsome sparkling of diamonds very amiable costumed.
Or do you are fond better what belongs they
moderns pleasure turpitude of old mans?
\stage{(he points about him with grotesque gestures which Lynch and the whores reply to)}
Caoutchouc statue woman reversible or lifesize tompeeptom of virgins nudities
very lesbic the kiss five ten times.
Enter, gentleman, to see in mirror every positions trapezes
all that machine there besides also if desire act
awfully bestial butcher's boy pollutes in warm veal liver
or omlet on the belly \french{pièce de Shakespeare}.

\Bella:
\gab{3910}
\stage{(clapping her belly sinks back on the sofa, with a shout of laughter)}
An omelette on the ...
Ho! ho! ho! ho! ...
omelette on the ...

\Stephen:
\stage{(mincingly)}
I love you, sir darling.
Speak you englishman tongue for \french{double entente cordiale}.
O yes, \french{mon loup}.
How much cost?
Waterloo.
Watercloset.
\stage{(he ceases suddenly and holds up a forefinger)}

\Bella:
\stage{(laughing)}
Omelette ...

\Whores:
\stage{(laughing)}
\gab{3920}
Encore! Encore!

\Stephen:
Mark me.
I dreamt of a watermelon.

\Zoe:
Go abroad and love a foreign lady.

\Lynch:
Across the world for a wife.

\Florry:
Dreams goes by contraries.

\Stephen:
\stage{(extends his arms)}
\gab{3930}
It was here.
Street of harlots.
In Serpentine avenue Beelzebub showed me her, a fubsy widow.
Where's the red carpet spread?

\Bloom:
\stage{(approaching Stephen)}
Look ...

\Stephen:
No, I flew.
My foes beneath me.
And ever shall be.
World without end.
\stage{(he cries)}
\latin{Pater!}
Free!

\Bloom:
I say, look ...

\Stephen:
Break my spirit, will he?
\gab{3940}
\french{O merde alors!}
\stage{(he cries, his vulture talons sharpened)}
\spanish{Holà!}
Hillyho!

\stage{(Simon Dedalus' voice hilloes in answer, somewhat sleepy but ready.)}

\Simon:
That's all right.
\stage{(he swoops uncertainly through the air,
wheeling, uttering cries of heartening, on strong ponderous buzzard wings)}
Ho, boy!
Are you going to win?
Hoop!
Pschatt!
Stable with those halfcastes.
Wouldn't let them within the bawl of an ass.
Head up!
Keep our flag flying!
An eagle gules volant in a field argent displayed.
Ulster king at arms!
Haihoop!
\stage{(he makes the beagle's call, giving tongue)}
\gab{3950}
Bulbul!
Burblblburblbl!
Hai, boy!

\stage{(The fronds and spaces of the wallpaper file rapidly across country.
A stout fox, drawn from covert, brush pointed, having buried his grandmother,
runs swift for the open, brighteyed, seeking badger earth, under the leaves.
The pack of staghounds follows, nose to the ground,
sniffing their quarry, beaglebaying, burblbrbling to be blooded.
Ward Union huntsmen and huntswomen live with them, hot for a kill.
From Six Mile Point, Flathouse, Nine Mile Stone follow the footpeople with knotty sticks,
hayforks, salmongaffs, lassos, flockmasters with stockwhips,
bearbaiters with tomtoms, toreadors with bullswords, greynegroes waving torches.
\gab{3960}
The crowd bawls of dicers, crown and anchor players, thimbleriggers, broadsmen.
Crows and touts, hoarse bookies in high wizard hats clamour deafeningly.)}

\Crowd:
Card of the races. Racing card!\\
Ten to one the field!\\
Tommy on the clay here! Tommy on the clay!\\
Ten to one bar one! Ten to one bar one!\\
Try your luck on Spinning Jenny!\\
Ten to one bar one!%
\gab{3970}\\
Sell the monkey, boys! Sell the monkey!\\
I'll give ten to one!\\
Ten to one bar one!

\stage{(A dark horse, riderless, bolts like a phantom past the winningpost,
his mane moonfoaming, his eyeballs stars.
The field follows, a bunch of bucking mounts.
Skeleton horses, Sceptre, Maximum the Second, Zinfandel,
the duke of Westminster's Shotover, Repulse,
the duke of Beaufort's Ceylon, prix de Paris.
Dwarfs ride them, rustyarmoured, leaping, leaping in their, in their saddles.
Last in a drizzle of rain on a brokenwinded isabelle nag,
\gab{3980}
Cock of the North, the favourite, honey cap, green jacket,
orange sleeves, Garrett Deasy up, gripping the reins, a hockeystick at the ready.
His nag on spavined whitegaitered feet jogs along the rocky road.)}

\OrangeLodges:
\stage{(jeering)}
Get down and push, mister.
Last lap!
You'll be home the night!

\GarrettDeasy:
\stage{(bolt upright, his nailscraped face plastered with postagestamps,
brandishes his hockeystick, his blue eyes flashing
in the prism of the chandelier as his mount lopes by at schooling gallop)}
\latin{Per vias rectas!}

\stage{(A yoke of buckets leopards all over him and his rearing nag
\gab{3990}
a torrent of mutton broth with dancing coins
of carrots, barley, onions, turnips, potatoes.)}

\GreenLodges:
Soft day, sir John!
Soft day, your honour!

\stage{(Private Carr, Private Compton and Cissy Caffrey pass beneath the windows,
singing in discord.)}

\Stephen:
Hark!
Our friend noise in the street.

\Zoe:
\stage{(holds up her hand)}
\gab{4000}
Stop!

\CarrComptonCissy:
\begin{verse}
    Yet I've a sort a\\
    Yorkshire relish for ...
\end{verse}

\commentary{The Dance of Death}

\Zoe:
That's me.
\stage{(she claps her hands)}
Dance!
Dance!
\stage{(she runs to the pianola)}
Who has twopence?

\Bloom:
Who'll ...?

\Lynch:
\stage{(handing her coins)}
\gab{4010}
Here.

\Stephen:
\stage{(cracking his fingers impatiently)}
Quick!
Quick!
Where's my augur's rod?
\stage{(he runs to the piano and takes his ashplant,
beating his foot in tripudium)}

\Zoe:
\stage{(turns the drumhandle)}
There.

\stage{(She drops two pennies in the slot.
Gold, pink and violet lights start forth.
The drum turns purring in low hesitation waltz.
Professor Goodwin, in a bowknotted periwig, in court dress,
wearing a stained Inverness cape, bent in two from incredible age,
totters across the room, his hands fluttering.
\gab{4020}
He sits tinily on the pianostool and lifts
and beats handless sticks of arms on the keyboard,
nodding with damsel's grace, his bowknot bobbing)}

\Zoe:
\stage{(twirls round herself, heeltapping)}
Dance.
Anybody here for there?
Who'll dance?
Clear the table.

\stage{(The pianola with changing lights plays in waltz time
the prelude of \title{My Girl's a Yorkshire Girl}.
Stephen throws his ashplant on the table and seizes Zoe round the waist.
Florry and Bella push the table towards the fireplace.
Stephen, arming Zoe with exaggerated grace, begins to waltz her round the room.
Bloom stands aside.
\gab{4030}
Her sleeve filling from gracing arms reveals a white fleshflower of vaccination.
Between the curtains Professor Maginni inserts a leg
on the toepoint of which spins a silk hat.
With a deft kick he sends it spinning to his crown and jauntyhatted skates in.
He wears a slate frockcoat with claret silk lapels,
a gorget of cream tulle, a green lowcut waistcoat, stock collar with white kerchief, 
tight lavender trousers, patent pumps and canary gloves.
In his buttonhole is an immense dahlia.
He twirls in reversed directions a clouded cane, then wedges it tight in his oxter.
He places a hand lightly on his breastbone,
\gab{4040}
bows, and fondles his flower and buttons.)}

\Maginni:
The poetry of motion, art of calisthenics.
No connection with Madam Legget Byrne's or Levenston's.
Fancy dress balls arranged.
Deportment.
The Katty Lanner step.
So. Watch me!
My terpsichorean abilities.
\stage{(he minuets forward three paces on tripping bee's feet)}
\french{Tout le monde en avant!
Révérence!
Tout le monde en place!}

\stage{(The prelude ceases.
Professor Goodwin, beating vague arms shrivels, sinks,
his live cape filling about the stool.
The air in firmer waltz time sounds.
Stephen and Zoe circle freely.
The lights change,
\gab{4050}
glow, fide gold rosy violet.)}

\Pianola:
\begin{verse}
    Two young fellows were talking about their girls, girls, girls,\\
    Sweethearts they'd left behind ...
\end{verse}

\stage{(From a corner the morning hours run out, goldhaired, slimsandalled,
in girlish blue, waspwaisted, with innocent hands.
Nimbly they dance, twirling their skipping ropes.
The hours of noon follow in amber gold.
Laughing, linked, high haircombs flashing,
they catch the sun in mocking mirrors, lifting their arms.)}

\Maginni:
\stage{(clipclaps glovesilent hands)}
\gab{4060}
\french{Carré!
Avant deux!}
Breathe evenly!
\french{Balancé!}

\stage{(The morning and noon hours waltz in their places, turning,
advancing to each other, shaping their curves, bowing visavis.
Cavaliers behind them arch and suspend their arms,
with hands descending to, touching, rising from their shoulders.)}

\Hours:
You may touch my.

\Cavaliers:
May I touch your?

\Hours:
\gab{4070}
O, but lightly!

\Cavaliers:
O, so lightly!

\Pianola:
\begin{verse}
    My little shy little lass has a waist.
\end{verse}


\stage{(Zoe and Stephen turn boldly with looser swing.
The twilight hours advance from long landshadows, dispersed, lagging, languideyed,
their cheeks delicate with cipria and false faint bloom.
They are in grey gauze with dark bat sleeves that flutter in the land breeze.)}

\Maginni:
\french{Avant huit!
\gab{4080}
Traversé!
Salut!
Cours de mains!
Croisé!}

\stage{(The night hours, one by one, steal to the last place.
Morning, noon and twilight hours retreat before them.
They are masked, with daggered hair and bracelets of dull bells.
Weary they curchycurchy under veils.)}

\Bracelets:
Heigho!
Heigho!

\Zoe:
\stage{(twirling, her hand to her brow)}
O!

\Maginni:
\french{Les tiroirs!
\gab{4090}
Chaîne de dames!
La corbeille!
Dos à dos!}

\stage{(Arabesquing wearily they weave a pattern on the floor, weaving,
unweaving, curtseying, twirling, simply swirling.)}

\Zoe:
I'm giddy!

\stage{(She frees herself, droops on a chair.
Stephen seizes Florry and turns with her.)}

\Maginni:
\french{Boulangère!
Les ronds!
Les ponts!
Chevaux de bois!
Escargots!}

\stage{(Twining, receding, with interchanging hands the night hours
link each each with arching arms in a mosaic of movements.
\gab{4100}
Stephen and Florry turn cumbrously.)}

\Maginni:
\french{Dansez avec vos dames!
Changez de dames!
Donnez le petit bouquet à votre dame!
Remerciez!}

\Pianola:
\begin{verse}
    Best, best of all,
    Baraabum!
\end{verse}

\Kitty:
\stage{(jumps up)}
O, they played that on the hobbyhorses at the \Title{Mirus} bazaar!

\stage{(She runs to Stephen.
\gab{4110}
He leaves Florry brusquely and seizes Kitty.
A screaming bittern's harsh high whistle shrieks.
Groangrousegurgling Toft's cumbersome whirligig turns slowly
the room right roundabout the room.)}

\Pianola:
\begin{verse}
    My girl's a Yorkshire girl.
\end{verse}

\Zoe:
Yorkshire through and through.
Come on all!

\stage{(she seizes Florry and waltzes her.)}

\Stephen:
\french{Pas seul!}
\gab{4120}

\stage{(He wheels Kitty into Lynch's arms,
snatches up his ashplant from the table and takes the floor.
All wheel whirl waltz twirl
Bloombella Kittylynch Florryzoe jujuby women.
Stephen with hat ashplant frogsplits in middle highkicks
with skykicking mouth shut hand clasp part under thigh.
With clang tinkle boomhammer tallyho Hornblower blue green yellow
flashes Toft's cumbersome turns with hobbyhorse riders from gilded snakes dangled,
bowels fandango leaping spurn soil foot and fall again.)}

\Pianola:
\begin{verse}
\gab{4130}%
    Though she's a factory lass\\
    And wears no fancy clothes.
\end{verse}


\stage{(Closeclutched swift swifter with glareblareflare scudding
they scootlootshoot lumbering by.
Baraabum!)}

\Tutti:
Encore!
Bis!
Bravo!
Encore!

\Simon:
Think of your mother's people!

\Stephen:
Dance of death.

\stage{(Bang fresh barang bang of lacquey's bell,
\gab{4140}
horse, nag, steer, piglings, Conmee on Christass,
lame crutch and leg sailor in cockboat armfolded
ropepulling hitching stamp hornpipe through and through.
Baraabum!
On nags hogs bellhorses Gadarene swine Corny in coffin steel shark stone
onehandled Nelson two trickies Frauenzimmer plumstained from pram falling bawling.
Gum he's a champion.
Fuseblue peer from barrel rev. evensong Love on hackney jaunt
Blazes blind coddoubled bicyclers
Dilly with snowcake no fancy clothes.
Then in last switchback lumbering up and down bump mashtub
sort of viceroy and reine relish for tublumber bumpshire rose.
\gab{4150}
Baraabum!)}

\stage{(The couples fall aside.
Stephen whirls giddily.
Room whirls back.
Eyes closed he totters.
Red rails fly spacewards.
Stars all around suns turn roundabout.
Bright midges dance on walls.
He stops dead.)}

\Stephen:
Ho!

\stage{(Stephen's mother, emaciated, rises stark through the floor,
in leper grey with a wreath of faded orangeblossoms and a torn bridal veil,
her face worn and noseless, green with gravemould.
Her hair is scant and lank.
\gab{4160}
She fixes her bluecircled hollow eyesockets on Stephen
and opens her toothless mouth uttering a silent word.
A choir of virgins and confessors sing voicelessly.)}

\Choir:
\begin{verse}
\latin{%
    Liliata rutilantium te confessorum ...\\
    Iubilantium te virginum ...}
\end{verse}

\stage{(From the top of a tower Buck Mulligan,
in particoloured jester's dress of puce and yellow
and clown's cap with curling bell, stands gaping at her,
a smoking buttered split scone in his hand.)}

\BuckMulligan:
She's beastly dead.
The pity of it!
\gab{4170}
Mulligan meets the afflicted mother.
\stage{(he upturns his eyes)}
Mercurial Malachi!

\Mother:
\stage{(with the subtle smile of death's madness)}
I was once the beautiful May Goulding.
I am dead.

\Stephen:
\stage{(horrorstruck)}
Lemur, who are you?
No. What bogeyman's trick is this?

\BuckMulligan:
\stage{(shakes his curling capbell)}
The mockery of it!
Kinch dogsbody killed her bitchbody.
She kicked the bucket.
\stage{(tears of molten butter fall from his eyes on to the scone)}
Our great sweet mother!
\gab{4180}
\greek{Epi oinopa ponton.}

\Mother:
\stage{(comes nearer, breathing upon him softly her breath of wetted ashes)}
All must go through it, Stephen.
More women than men in the world.
You too.
Time will come.

\Stephen:
\stage{(choking with fright, remorse and horror)}
They say I killed you, mother.
He offended your memory.
Cancer did it, not I.
Destiny.

\Mother:
\stage{(a green rill of bile trickling from a side of her mouth)}
You sang that song to me.
\gab{4190}
\Title{Love's bitter mystery.}

\Stephen:
\stage{(eagerly)}
Tell me the word, mother, if you know now.
The word known to all men.

\Mother:
Who saved you the night you jumped into the train at Dalkey with Paddy Lee?
Who had pity for you when you were sad among the strangers?
Prayer is allpowerful.
Prayer for the suffering souls in the Ursuline manual and forty days' indulgence.
Repent, Stephen.

\Stephen:
The ghoul!
\gab{4200}
Hyena!

\Mother:
I pray for you in my other world.
Get Dilly to make you that boiled rice every night after your brainwork.
Years and years I loved you, O, my son, my firstborn, when you lay in my womb.

\Zoe:
\stage{(fanning herself with the grate fan)}
I'm melting!

\Florry:
\stage{(points to Stephen)}
Look!
He's white.

\Bloom:
\stage{(goes to the window to open it more)}
\gab{4210}
Giddy.

\Mother:
\stage{(with smouldering eyes)}
Repent!
O, the fire of hell!

\Stephen:
\stage{(panting)}
His noncorrosive sublimate!
The corpsechewer!
Raw head and bloody bones.

\Mother:
\stage{(her face drawing near and nearer, sending out an ashen breath)}
Beware!
\stage{(she raises her blackened withered right arm slowly
towards Stephen's breast with outstretched finger)}
Beware God's hand!

\stage{(A green crab with malignant red eyes sticks deep
\gab{4220}
its grinning claws in Stephen's heart.)}

\Stephen:
\stage{(strangled with rage, his features drawn grey and old)}
Shite!

\Bloom:
\stage{(at the window)}
What?

\Stephen:
\french{Ah non, par exemple!}
The intellectual imagination!
With me all or not at all.
\latin{Non serviam!}

\Florry:
Give him some cold water.
Wait.
\gab{4230}
\stage{(she rushes out)}

\Mother:
\stage{(wrings her hands slowly, moaning desperately)}
O Sacred Heart of Jesus, have mercy on him!
Save him from hell, O Divine Sacred Heart!

\Stephen:
No! No! No!
Break my spirit, all of you, if you can!
I'll bring you all to heel!

\Mother:
\stage{(in the agony of her deathrattle)}
Have mercy on Stephen, Lord, for my sake!
Inexpressible was my anguish when expiring with love, grief and agony
\gab{4240}
on Mount Calvary.

\Stephen:
\german{Nothung!}

\stage{(He lifts his ashplant high with both hands and smashes the chandelier.
Time's livid final flame leaps and, in the following darkness,
ruin of all space, shattered glass and toppling masonry.)}

\Gasjet:
Pwfungg!

\Bloom:
Stop!

\Lynch:
\gab{4250}
\stage{(rushes forward and seizes Stephen's hand)}
Here!
Hold on!
Don't run amok!

\Bella:
Police!

\stage{(Stephen, abandoning his ashplant, his head and arms thrown back stark,
beats the ground and flies from the room, past the whores at the door.)}

\Bella:
\stage{(screams)}
After him!

\stage{(The two whores rush to the halldoor.
\gab{4260}
Lynch and Kitty and Zoe stampede from the room.
They talk excitedly.
Bloom follows, returns.)}

\Whores:
\stage{(jammed in the doorway, pointing)}
Down there.

\Zoe:
\stage{(pointing)}
There.
There's something up.

\Bella:
Who pays for the lamp?
\stage{(she seizes Bloom's coattail)}
Here, you were with him.
The lamp's broken.

\Bloom:
\gab{4270}
\stage{(rushes to the hall, rushes back)}
What lamp, woman?

\AWhore:
He tore his coat.

\Bella:
\stage{(her eyes hard with anger and cupidity, points)}
Who's to pay for that?
Ten shillings.
You're a witness.

\Bloom:
\stage{(snatches up Stephen's ashplant)}
Me? Ten shillings?
Haven't you lifted enough off him?
Didn't he ...?

\Bella:
\gab{4280}
\stage{(loudly)}
Here, none of your tall talk.
This isn't a brothel.
A tenshilling house.

\Bloom:
\stage{(His head under the lamp, pulls the chain.
Puling, the gasjet lights up a crushed mauve purple shade.
He raises the ashplant.)}
Only the chimney's broken.
Here is all he ...

\Bella:
\stage{(shrinks back and screams)}
Jesus!
Don't!

\Bloom:
\stage{(warding off a blow)}
To show you how he hit the paper.
\gab{4290}
There's not sixpenceworth of damage done.
Ten shillings!

\Florry:
\stage{(with a glass of water, enters)}
Where is he?

\Bella:
Do you want me to call the police?

\Bloom:
O, I know.
Bulldog on the premises.
But he's a Trinity student.
Patrons of your establishment.
Gentlemen that pay the rent.
\stage{(he makes a masonic sign)}
Know what I mean?
Nephew of the vice-chancellor.
You don't want a scandal.%
\gab{4300}

\Bella:
\stage{(angrily)}
Trinity.
Coming down here ragging after the boatraces and paying nothing.
Are you my commander here or?
Where is he?
I'll charge him!
Disgrace him, I will!
\stage{(she shouts)}
Zoe! Zoe!

\Bloom:
\stage{(urgently)}
And if it were your own son in Oxford?
\stage{(warningly)}
I know.

\Bella:
\stage{(almost speechless)}
Who are.
Incog!

\Zoe:
\stage{(in the doorway)}
\gab{4310}
There's a row on.

\Bloom:
What? Where?
\stage{(he throws a shilling on the table and starts)}
That's for the chimney.
Where?
I need mountain air.

\stage{(He hurries out through the hall.
The whores point.
Florry follows, spilling water from her tilted tumbler.
On the doorstep all the whores clustered talk volubly,
pointing to the right where the fog has cleared off.
From the left arrives a jingling hackney car.
It slows to in front of the house.
Bloom at the halldoor perceives Corny Kelleher
who is about to dismount from the car with two silent lechers.
\gab{4320}
He averts his face.
Bella from within the hall urges on her whores.
They blow ickylickysticky yumyum kisses.
Corny Kelleher replies with a ghastly lewd smile.
The silent lechers turn to pay the jarvey.
Zoe and Kitty still point right.
Bloom, parting them swiftly, draws his caliph's hood and poncho
and hurries down the steps with sideways face.
Incog Haroun al Raschid he flits behind the silent lechers
and hastens on by the railings with fleet step of a pard
strewing the drag behind him, torn envelopes drenched in aniseed.
The ashplant marks his stride.
A pack of bloodhounds, led by Hornblower of Trinity
brandishing a dogwhip in tallyho cap and an old pair of grey trousers,
\gab{4330}
follow from far, picking up the scent, nearer,
baying, panting, at fault, breaking away, throwing their tongues,
biting his heels, leaping at his tail.
He walks, runs, zigzags, gallops, lugs laid back.
He is pelted with gravel, cabbagestumps, biscuitboxes, eggs,
potatoes, dead codfish, woman's slipperslappers.
After him freshfound the hue and cry
zigzag gallops in hot pursuit of follow my leader:
65~C, 66~C, Night Watch, John Henry Menton, Wisdom Hely,
V.B.~Dillon, Councillor Nannetti, Alexander Keyes,
Larry O'Rourke, Joe Cuffe, Mrs~O'Dowd, Pisser Burke,
the Nameless One, Mrs~Riordan, the Citizen, Garryowen,
Whodoyoucallhim, Strangeface,
\gab{4340}
Fellowthatsolike, Sawhimbefore, Chapwithawen,
Chris Callinan, sir Charles Cameron, Benjamin Dollard,
Lenehan, Bartell d'Arcy, Joe Hynes, Red Murray, editor Brayden,
T.M.~Healy, Mr~Justice Fitzgibbon, John Howard Parnell,
the Reverend Tinned Salmon, Professor Joly, Mrs~Breen, Denis Breen,
Theodore Purefoy, Mina Purefoy, the Westland Row postmistress,
C.P.~M'Coy, friend of Lyons, Hoppy Holohan,
maninthestreet, othermaninthestreet, Footballboots,
pugnosed driver, rich protestant lady,
Davy Byrne, Mrs~Ellen M'Guinness,
Mrs~Joe Gallaher, George Lidwell, Jimmy Henry on corns,
superintendent Laracy, Father Cowley,
\gab{4350}
Crofton out of the Collector-general's,
Dan Dawson, dental surgeon Bloom with tweezers, Mrs~Bob Doran, Mrs~Kennefick,
Mrs~Wyse Nolan, John Wyse Nolan,
HandsomemarriedwomanrubbedagainstwidebehindinClonskeatram,
the bookseller of \Title{sweets of sin}, Miss Dubedatandshedidbedad,
Mesdames Gerald and Stanislaus Moran of Roebuck,
the managing clerk of Drimmie's, Wetherup, colonel Hayes,
Mastiansky, Citron, Penrose, Aaron Figatner, Moses Herzog,
Michael~E Geraghty, Inspector Troy, Mrs~Galbraith,
the constable off Eccles street corner, old doctor Brady with stethoscope,
the mystery man on the beach, a retriever,
\gab{4360}
Mrs~Miriam Dandrade and all her lovers.)}

\HueCry:
\stage{(helterskelterpelterwelter)}
He's Bloom!
Stop Bloom!
Stopabloom!
Stopperrobber!
Hi! Hi!
Stophim on the corner!

\stage{(At the corner of Beaver street beneath the scaffolding
Bloom panting stops on the fringe of the noisy quarrelling knot,
a lot not knowing a jot
what hi! hi!
row and wrangle round the whowhat brawlaltogether.)}

\Stephen:
\stage{(with elaborate gestures, breathing deeply and slowly)}
\gab{4370}
You are my guests.
Uninvited.
By virtue of the fifth of George and seventh of Edward.
History to blame.
Fabled by mothers of memory.

\Carr:
\stage{(to Cissy Caffrey)}
Was he insulting you?

\Stephen:
Addressed her in vocative feminine.
Probably neuter.
Ungenitive.

\Voices:
No, he didn't.
I seen him.
The girl there.
He was in Mrs~Cohen's.
What's up?
Soldier and civilian.

\Cissy:
\gab{4380}
I was in company with the soldiers and they left me
to do, you know, and the young man run up behind me.
But I'm faithful to the man that's treating me
though I'm only a shilling whore.

\Voices:
Shes faithfultheman.

\Stephen:
\stage{(catches sight of Lynch's and Kitty's heads)}
Hail, Sisyphus.
\stage{(he points to himself and the others)}
Poetic.
Uropoetic.

\Cissy:
Yes, to go with him.
\gab{4390}
And me with a soldier friend.

\Compton:
He doesn't half want a thick ear, the blighter.
Biff him one, Harry.

\Carr:
\stage{(to Cissy)}
Was he insulting you while me and him was having a piss?

\LordTennyson:
\stage{(gentleman poet in Union Jack blazer and cricket flannels,
bareheaded, flowingbearded)}
Theirs not to reason why.

\Compton:
Biff him, Harry.

\Stephen:
\gab{4400}
\stage{(to Private Compton)}
I don't know your name but you are quite right.
Doctor Swift says one man in armour will beat ten men in their shirts.
Shirt is synechdoche.
Part for the whole.

\Cissy:
\stage{(to the crowd)}
No, I was with the privates.

\Stephen:
\stage{(amiably)}
Why not?
The bold soldier boy.
In my opinion every lady for example \ldots.

\Carr:
\stage{(his cap awry, advances to Stephen)}
\gab{4410}
Say, how would it be, governor, if I was to bash in your jaw?

\Stephen:
\stage{(looks up to the sky)}
How?
Very unpleasant.
Noble art of selfpretence.
Personally, I detest action.
\stage{(he waves his hand)}
Hand hurts me slightly.
\french{Enfin ce sont vos oignons.}
\stage{(to Cissy Caffrey)}
Some trouble is on here.
What is it precisely?

\DollyGray:
\stage{(from her balcony waves her handkerchief,
giving the sign of the heroine of Jericho)}
Rahab.
Cook's son, goodbye.
Safe home to Dolly.
Dream of the girl you left behind
\gab{4420}
and she will dream of you.

\stage{(The soldiers turn their swimming eyes.)}

\Bloom:
\stage{(elbowing through the crowd, plucks Stephen's sleeve vigorously)}
Come now, professor, that carman is waiting.

\Stephen:
\stage{(turns)}
Eh?
\stage{(he disengages himself)}
Why should I not speak to him or to any human being
who walks upright upon this oblate orange?
\stage{(he points his finger)}
I'm not afraid of what I can talk to if I see his eye.
Retaining the perpendicular.
\stage{(he staggers a pace back)}

\Bloom:
\gab{4430}
\stage{(propping him)}
Retain your own.

\Stephen:
\stage{(laughs emptily)}
My centre of gravity is displaced.
I have forgotten the trick.
Let us sit down somewhere and discuss.
Struggle for life is the law of existence but but human philirenists,
notably the tsar and the king of England, have invented arbitration.
\stage{(he taps his brow)}
But in here it is I must kill the priest and the king.

\BiddyClap:
Did you hear what the professor said?
He's a professor out of the college.

\CuntyKate:
\gab{4440}
I did.
I heard that.

\BiddyClap:
He expresses himself with such marked refinement of phraseology.

\CuntyKate:
Indeed, yes.
And at the same time with such apposite trenchancy.

\Carr:
\stage{(pulls himself free and comes forward)}
What's that you're saying about my king?

\stage{(Edward the Seventh appears in an archway.
He wars a white jersey on which an image of the Sacred Heart
\gab{4450}
is stitched with the insignia of Garter and Thistle,
Golden Fleece, Elephant of Denmark, Skinner's and Probyn's horse,
Lincoln's Inn bencher
and ancient and honourable artillery company of Massachusetts.
He sucks a red jujube.
He is robed as a grand elect perfect and sublime mason
with trowel and apron, marked \title{made in Germany}.
In his left hand he holds a plasterer's bucket on which
is printed \french{Défense d'uriner}.
A roar of welcome greets him.)}

\EdwardSeventh:
\stage{(slowly, solemnly but indistinctly)}
Peace, perfect peace.
For identification, bucket in my hand.
\gab{4460}
Cheerio, boys.
\stage{(he turns to his subjects)}
We have come here to witness a clean straight fight
and we heartily wish both men the best of good luck.
Mahak makar a bak.
\stage{(he shakes hands with Private Carr, Private Compton,
Stephen, Bloom and Lynch)}

\stage{(General applause.
Edward the Seventh lifts his bucket graciously in acknowledgment.)}

\Carr:
\stage{(to Stephen)}
Say it again.

\Stephen:
\stage{(nervous, friendly, pulls himself up)}
I understand your point of view though I have no king myself for the moment.
\gab{4470}
This is the age of patent medicines.
A discussion is difficult down here.
But this is the point.
You die for your country.
Suppose.
\stage{(he places his arm on Private Carr's sleeve)}
Not that I wish it for you.
But I say: Let my country die for me.
Up to the  present it has done so.
I didn't want it to die.
Damn death.
Long live life!

\EdwardSeventh:
\stage{(levitates over heaps of slain,
in the garb and with the halo of Joking Jesus,
a white jujube in his phosphorescent face)}
\begin{verse}
    My methods are new and are causing surprise.\\
    To make the blind see I throw dust in their eyes.
\end{verse}

\Stephen:
\gab{4480}
Kings and unicorns!
\stage{(he fills back a pace)}
Come somewhere and we'll ...
What was that girl saying? ...

\Compton:
Eh, Harry, give him a kick in the knackers.
Stick one into Jerry.

\Bloom:
\stage{(to the privates, softly)}
He doesn't know what he's saying.
Taken a little more than is good for him.
Absinthe.
Greeneyed monster.
I know him.
He's a gentleman, a poet.
It's all right.

\Stephen:
\stage{(nods, smiling and laughing)}
\gab{4490}
Gentleman, patriot, scholar and judge of impostors.

\Carr:
I don't give a bugger who he is.

\Compton:
We don't give a bugger who he is.

\Stephen:
I seem to annoy them.
Green rag to a bull.

\stage{(Kevin Egan of Paris in black Spanish tasselled shirt
and peep-o'-day boy's hat signs to Stephen.)}

\KevinEgan:
\gab{4500}
H'lo!
\french{Bonjour!}
The \french{vieille ogresse} with the \french{dents jaunes}.

\stage{(Patrice Egan peeps from behind, his rabbitface nibbling a quince leaf.)}

\Patrice:
\french{Socialiste!}

\DonEmile:
\stage{(in medieval hauberk, two wild geese volant on his helm,
with noble indignation points a mailed hand against the privates)}
Werf those eykes to footboden, big grand
porcos of johnyellows todos covered of gravy!

\Bloom:
\gab{4510}
\stage{(to Stephen)}
Come home.
You'll get into trouble.

\Stephen:
\stage{(swaying)}
I don't avoid it.
He provokes my intelligence.

\BiddyClap:
One immediately observes that he is of patrician lineage.

\Virago:
Green above the red, says he.
Wolfe Tone.

\Bawd:
The red's as good as the green.
And better.
Up the soldiers!
Up King Edward!%
\gab{4520}

\ARough:
\stage{(laughs)}
Ay! Hands up to De Wet.

\Citizen:
\stage{(with a huge emerald muffler and shillelagh, calls)}
\begin{verse}
    May the God above\\
    Send down a dove\\
    With teeth as sharp as razors\\
    To slit the throats\\
    Of the English dogs\\
\gab{4530}%
    That hanged our Irish leaders.
\end{verse}


\CroppyBoy:
\stage{(the ropenoose round his neck, gripes in his issuing bowels with both hands)}
\begin{verse}
    I bear no hate to a living thing,\\
    But I love my country beyond the king.
\end{verse}


\RumboldDemonBarber:
\stage{(accompanied by two blackmasked assistants,
advances with gladstone bag which he opens)}
Ladies and gents, cleaver purchased by Mrs~Pearcy to slay Mogg.
Knife with which Voisin dismembered the wife of a compatriot
and hid remains in a sheet in the cellar,
\gab{4540}
the unfortunate female's throat being cut from ear to ear.
Phial containing arsenic retrieved from body of Miss Barron
which sent Seddon to the gallows.

\stage{(He jerks the rope.
The assistants leap at the victim's legs and drag him downward,
grunting the croppy boy's tongue protrudes violently.)}

\CroppyBoy:
\begin{verse}
    Horhot ho hray hor hother's hest.
\end{verse}


\stage{(He gives up the ghost.
A violent erection of the hanged sends gouts of sperm spouting
through his deathclothes on to the cobblestones.
Mrs~Bellingham, Mrs~Yelverton Barry
\gab{4550}
and the Honourable Mrs~Mervyn Talboys
rush forward with their handkerchiefs to sop it up.)}

\Rumbold:
I'm near it myself.
\stage{(he undoes the noose)}
Rope which hanged the awful rebel.
Ten shillings a time.
As applied to Her Royal Highness.
\stage{(he plunges his head into the gaping belly of the hanged
and draws out his head again clotted with coiled and smoking entrails)}
My painful duty has now been done.
God save the king!

\EdwardSeventh:
\stage{(dances slowly, solemnly, rattling his bucket, and
\gab{4560}
sings with soft contentment)}
\begin{verse}
    On coronation day, on coronation day,\\
    O, won't we have a merry time,\\
    Drinking whisky, beer and wine!
\end{verse}

\Carr:
Here. What are you saying about my king?

\Stephen:
\stage{(throws up his hands)}
O, this is too monotonous!
Nothing.
He wants my money and my life, though want must be his master,
for some brutish empire of his.
Money I haven't.
\stage{(he searches his pockets vaguely)}
\gab{4570}
Gave it to someone.

\Carr:
Who wants your bleeding money?

\Stephen:
\stage{(tries to move off)}
Will someone tell me where I am least likely to meet these necessary evils?
\french{Ça se voit aussi à Paris.}
Not that I ...
But, by Saint Patrick ...!

\stage{(The women's heads coalesce.
Old Gummy Granny in sugarloaf hat appears seated on a toadstool,
the deathflower of the potato blight on her breast.)%
\gab{4580}
}

\Stephen:
Aha!
I know you, gammer!
Hamlet, revenge!
The old sow that eats her farrow!

\GummyGranny:
\stage{(rocking to and fro)}
Ireland's sweetheart, the king of Spain's daughter, alanna.
Strangers in my house, bad manners to them!
\stage{(she keens with banshee woe)}
Ochone!
Ochone!
Silk of the kine!
\stage{(she wails)}
You met with poor old Ireland and how does she stand?

\Stephen:
How do I stand you?
\gab{4590}
The hat trick!
Where's the third person of the Blessed Trinity?
Soggarth Aroon?
The reverend Carrion Crow.

\Cissy:
\stage{(shrill)}
Stop them from fighting!

\ARough:
Our men retreated.

\Carr:
\stage{(tugging at his belt)}
I'll wring the neck of any fucker says a word against my fucking king.

\Bloom:
\stage{(terrified)}
\gab{4600}
He said nothing.
Not a word.
A pure misunderstanding.

\Compton:
Go it, Harry.
Do him one in the eye.
He's a proBoer.

\Stephen:
Did I?
When?

\Bloom:
\stage{(to the redcoats)}
We fought for you in South Africa, Irish missile troops.
Isn't that history?
Royal Dublin Fusiliers.
Honoured by our monarch.

\Navvy:
\stage{(staggering past)}
O, yes!
O God, yes!
O, make the kwawr a krowawr!
O! Bo!%
\gab{4610}

\stage{(Casqued halberdiers in armour thrust forward a pentice of gutted spearpoints.
Major Tweedy, moustached like Turko the terrible,
in bearskin cap with hackleplume and accoutrements,
with epaulettes, gilt chevrons and sabretaches,
his breast bright with medals, toes the line.
He gives the pilgrim warrior's sign of the knights templars.)}

\MajorTweedy:
\stage{(growls gruffly)}
Rorke's Drift!
Up, guards, and at them!
Mahar shalal hashbaz.

\Citizen:
\gab{4620}
\irish{Erin go bragh!}

\stage{(Major Tweedy and the Citizen exhibit to each other medals,
decorations, trophies of war, wounds.
Both salute with fierce hostility.)}

\Carr:
I'll do him in.

\Compton:
\stage{(waves the crowd back)}
Fair play, here.
Make a bleeding butcher's shop of the bugger.

\stage{(Massed bands blare \title{Garryowen} and \title{God Save the King}.)%
\gab{4630}
}

\Cissy:
They're going to fight.
For me!

\CuntyKate:
The brave and the fair.

\BiddyClap:
Methinks yon sable knight will joust it with the best.

\CuntyKate:
\stage{(blushing deeply)}
Nay, madam.
The gules doublet and merry saint George for me!

\Stephen:
\gab{4640}
\begin{verse}
    The harlot's cry from street to street\\
    Shall weave Old Ireland's windingsheet.
\end{verse}

\Carr:
\stage{(loosening his belt, shouts)}
I'll wring the neck of any fucking bastard
says a word against my bleeding fucking king.

\Bloom:
\stage{(shakes Cissy Caffrey's shoulders)}
Speak, you!
Are you struck dumb?
You are the link between nations and generations.
Speak, woman, sacred lifegiver!

\Cissy:
\gab{4650}
\stage{(alarmed, seizes Private Carr's sleeve)}
Amn't I with you?
Amn't I your girl?
Cissy's your girl.
\stage{(she cries)}
Police!

\Stephen:
\stage{(ecstatically, to Cissy Caffrey)}
\begin{verse}
    White thy fambles, red thy gan\\
    And thy quarrons dainty is.
\end{verse}

\Voices:
Police!

\DistantVoices:
Dublin's burning!
Dublin's burning!
\gab{4660}
On fire, on fire!

\stage{(Brimstone fires spring up.
Dense clouds roll past.
Heavy Gatling guns boom.
Pandemonium.
Troops deploy.
Gallop of hoofs.
Artillery.
Hoarse commands.
Bells clang.
Backers shout.
Drunkards bawl.
Whores screech.
Foghorns hoot.
Cries of valour.
Shrieks of dying.
Pikes clash on cuirasses.
Thieves rob the slain.
Birds of prey, winging from the sea,
rising from marshlands, swooping from eyries, hover screaming,
gannets, cormorants, vultures, goshawks, climbing woodcocks, peregrines,
merlins, blackgrouse, sea eagles, gulls, albatrosses, barnacle geese.
The midnight sun is darkened.
The earth trembles.
\gab{4670}
The dead of Dublin from Prospect and Mount Jerome
in white sheepskin overcoats and black goatfell cloaks
arise and appear to many.
A chasm opens with a noiseless yawn.
Tom Rochford, winner, in athlete's singlet and breeches,
arrives at the head of the National Hurdle Handicap and leaps into the void.
He is followed by a race of runners and leapers.
In wild attitudes they spring from the brink.
Their bodies plunge.
Factory lasses with fancy clothes toss redhot Yorkshire baraabombs.
Society ladies lift their skirts above their heads to protect themselves.
Laughing witches in red cutty sarks ride through the air on broomsticks.
Quakerlyster plasters blisters.
\gab{4680}
It rains dragons' teeth.
Armed heroes spring up from furrows.
They exchange in amity the pass of knights of the red cross
and fight duels with cavalry sabres:
Wolfe Tone against Henry Grattan,
Smith O'Brien against Daniel O'Connell,
Michael Davitt against Isaac Butt,
Justin M'Carthy against Parnell,
Arthur Griffith against John Redmond,
John O'Leary against Lear O'Johnny,
Lord Edward Fitzgerald against Lord Gerald Fitzedward,
the O'Donoghue of the Glens against the Glens of the O'Donoghue.
On an eminence, the centre of the earth, rises the feldaltar of Saint Barbara.
Black candles rise from its gospel and epistle horns.
\gab{4690}
From the high barbacans of the tower
two shafts of light fall on the smokepalled altarstone.
On the altarstone Mrs~Mina Purefoy, goddess of unreason, lies,
naked, fettered, a chalice resting on her swollen belly.
Father Malachi O'Flynn in a lace petticoat and reversed chasuble,
his two left feet back to the front, celebrates camp mass.
The reverend Mr~Hugh C~Haines Love~M.A.
in a plain cassock and mortarboard,
his head and collar back to the front,
holds over the celebrant's head an open umbrella.)}

\FrMalachiOFlynn:
\latin{Introibo ad altare diaboli.}

\HainesLove:
\gab{4700}
To the devil which hath made glad my young days.

\FrMalachiOFlynn:
\stage{(takes from the chalice and elevates a blooddripping host)}
\latin{Corpus meum.}

\HainesLove:
\stage{(raises high behind the celebrant's petticoat,
revealing his grey bare hairy buttocks between which a carrot is stuck)}
My body.

\VoiceAllDamned:
Htengier Tnetopinmo Dog Drol eht rof,
Aiulella!

\stage{(From on high the voice of Adonai calls.)}

\Adonai:
\gab{4710}
Dooooooooooog!

\VoiceAllBlessed:
Alleluia, for the Lord God Omnipotent reigneth!

\stage{(From on high the voice of Adonai calls.)}

\Adonai:
Goooooooooood!

\stage{(In strident discord peasants and townsmen of Orange and Green factions
sing \title{Kick the Pope} and \title{Daily, daily sing to Mary}.)}

\Carr:
\stage{(with ferocious articulation)}
\gab{4720}
I'll do him in, so help me fucking Christ!
I'll wring the bastard fucker's bleeding blasted fucking windpipe!

\stage{(The retriever, nosing on the fringe of the crowd, barks noisily.)}

\Bloom:
\stage{(runs to Lynch)}
Can't you get him away?

\Lynch:
He likes dialectic, the universal language.
Kitty!
\stage{(to Bloom)}
Get him away, you.
He won't listen to me.

\stage{(He drags Kitty away.)}

\Stephen:
\stage{(points)}
\gab{4730}
\latin{Exit Judas.
Et laqueo se suspendit.}

\Bloom:
\stage{(runs to Stephen)}
Come along with me now before worse happens.
Here's your stick.

\Stephen:
Stick, no.
Reason.
This feast of pure reason.

\GummyGranny:
\stage{(thrusts a dagger towards Stephen's hand)}
Remove him, acushla.
At 8.35~a.m. you will be in heaven and Ireland will be free.
\stage{(she prays)}
O good God, take him!

\Cissy:
\gab{4740}
\stage{(pulling Private Carr)}
Come on, you're boosed.
He insulted me but I forgive him.
\stage{(shouting in his ear)}
I forgive him for insulting me.

\Bloom:
\stage{(over Stephen's shoulder)}
Yes, go.
You see he's incapable.

\Carr:
\stage{(breaks loose)}
I'll insult him.

\stage{(He rushes towards Stephen, fist outstretched, and strikes him in the face.
Stephen totters, collapses, falls, stunned.
He lies prone, his face to the sky, his hat rolling to the wall.
Bloom follows and picks it up.)%
\gab{4750}
}

\MajorTweedy:
\stage{(loudly)}
Carbine in bucket!
Cease fire!
Salute!

\Retriever:
\stage{(barking furiously)}
Ute ute ute ute ute ute ute ute.

\Crowd:
Let him up!
Don't strike him when he's down!
Air!
Who?
The soldier hit him.
He's a professor.
Is he hurted?
Don't manhandle him!
He's fainted!

\Hag:
What call had the redcoat to strike the gentleman and he under the influence.
\gab{4760}
Let them go and fight the Boers!

\Bawd:
Listen to who's talking!
Hasn't the soldier a right to go with his girl?
He gave him the coward's blow.

\stage{(They grab at each other's hair, claw at each other and spit)}

\Retriever:
\stage{(barking)}
Wow wow wow.

\Bloom:
\stage{(shoves them back, loudly)}
Get back, stand back!

\Compton:
\stage{(tugging his comrade)}
\gab{4770}
Here.
Bugger off, Harry.
Here's the cops!

\stage{(Two raincaped watch, tall, stand in the group.)}

\FirstWatch:
What's wrong here?

\Compton:
We were with this lady.
And he insulted us.
And assaulted my chum.
\stage{(the retriever barks)}
Who owns the bleeding tyke?

\Cissy:
\stage{(with expectation)}
Is he bleeding!

\Man:
\stage{(rising from his knees)}
\gab{4780}
No.
Gone off.
He'll come to all right.

\Bloom:
\stage{(glances sharply at the man)}
Leave him to me.
I can easily ...

\SecondWatch:
Who are you?
Do you know him?

\Carr:
\stage{(lurches towards the watch)}
He insulted my lady friend.

\Bloom:
\stage{(angrily)}
You hit him without provocation.
I'm a witness.
Constable, take his regimental number.

\SecondWatch:
\gab{4790}
I don't want your instructions in the discharge of my duty.

\Compton:
\stage{(pulling his comrade)}
Here, bugger off Harry.
Or Bennett'll shove you in the lockup.

\Carr:
\stage{(staggering as he is pulled away)}
God fuck old Bennett.
He's a white\-arsed bugger.
I don't give a shit for him.

\FirstWatch:
\stage{(takes out his notebook)}
What's his name?

\Bloom:
\gab{4800}
\stage{(peering over the crowd)}
I just see a car there.
If you give me a hand a second, sergeant ...

\FirstWatch:
Name and address.

\stage{(Corny Kelleker, weepers round his hat, a death wreath in his hand,
appears among the bystanders.)}

\Bloom:
\stage{(quickly)}
O, the very man!
\stage{(he whispers)}
Simon Dedalus' son.
A bit sprung.
Get those policemen to move those loafers back.

\SecondWatch:
\gab{4810}
Night, Mr~Kelleher.

\Corny:
\stage{(to the watch, with drawling eye)}
That's all right.
I know him.
Won a bit on the races.
Gold cup.
\Title{Throwaway}.
\stage{(he laughs)}
Twenty to one.
Do you follow me?

\FirstWatch:
\stage{(turns to the crowd)}
Here, what are you all gaping at?
Move on out of that.

\stage{(The crowd disperses slowly, muttering, down the lane.)}

\Corny:
\gab{4820}
Leave it to me, sergeant.
That'll be all right.
\stage{(he laughs, shaking his head)}
We were often as bad ourselves, ay or worse.
What? Eh, what?

\FirstWatch:
\stage{(laughs)}
I suppose so.

\Corny:
\stage{(nudges the second watch)}
Come and wipe your name off the slate.
\stage{(he lilts, wagging his head)}
With my tooraloom tooraloom tooraloom tooraloom.
What, eh, do you follow me?

\SecondWatch:
\stage{(genially)}
\gab{4830}
Ah, sure we were too.

\Corny:
\stage{(winking)}
Boys will be boys.
I've a car round there.

\SecondWatch:
All right, Mr~Kelleher.
Good night.

\Corny:
I'll see to that.

\Bloom:
\stage{(shakes hands with both of the watch in turn)}
Thank you very much, gentlemen.
Thank you.
\stage{(he mumbles confidentially)}
We don't want any scandal, you understand.
\gab{4840}
Father is a wellknown highly respected citizen.
Just a little wild oats, you understand.

\FirstWatch:
O. I understand, sir.

\SecondWatch:
That's all right, sir.

\FirstWatch:
It was only in case of corporal injuries
I'd have to report it at the station.

\Bloom:
\stage{(nods rapidly)}
Naturally.
Quite right.
Only your bounden duty.

\SecondWatch:
\gab{4850}
It's our duty.

\Corny:
Good night, men.

\Watch:
\stage{(saluting together)}
Night, gentlemen.

\stage{(They move off with slow heavy tread)}

\Bloom:
\stage{(blows)}
Providential you came on the scene.
You have a car? ...

\Corny:
\stage{(laughs, pointing his thumb over his right shoulder
\gab{4860}
to the car brought up against the scaffolding)}
Two commercials that were standing fizz in Jammet's.
Like princes, faith.
One of them lost two quid on the race.
Drowning his grief.
And were on for a go with the jolly girls.
So I landed them up on Behan's car and down to nighttown.

\Bloom:
I was just going home by Gardiner street when I happened to ...

\Corny:
\stage{(laughs)}
Sure they wanted me to join in with the mots.
No, by God, says I.
Not for old stagers like myself and yourself.
\stage{(he laughs again and leers with lacklustre eye)}
\gab{4870}
Thanks be to God we have it in the house, what, eh, do you follow me?
Hah, hah, hah!

\Bloom:
\stage{(tries to laugh)}
He, he, he!
Yes.
Matter of fact I was just visiting an old friend of mine there,
Virag, you don't know him (poor fellow, he's laid up for the past week)
and we had a liquor together and I was just making my way home ...

\stage{(The horse neighs.)}

\Horse:
Hohohohohohoh!
Hohohohome!

\Corny:
\gab{4880}
Sure it was Behan our jarvey there that told me
after we left the two commercials in Mrs~Cohen's
and I told him to pull up and got off to see.
\stage{(he laughs)}
Sober hearsedrivers a speciality.
Will I give him a lift home?
Where does he hang out?
Somewhere in Cabra, what?

\Bloom:
No, in Sandycove, I believe, from what he let drop.

\stage{(Stephen, prone, breathes to the stars.
Corny Kelleher, asquint, drawls at the horse.
Bloom, in gloom, looms down.)}

\Corny:
\stage{(scratches his nape)}
\gab{4890}
Sandycove!
\stage{(he bends down and calls to Stephen)}
Eh!
\stage{(he calls again)}
Eh!
He's covered with shavings anyhow.
Take care they didn't lift anything off him.

\Bloom:
No, no, no.
I have his money and his hat here and stick.

\Corny:
Ah, well, he'll get over it.
No bones broken.
Well, I'll shove along.
\stage{(he laughs)}
I've a rendezvous in the morning.
Burying the dead.
Safe home!

\Horse:
\stage{(neighs)}
Hohohohohome.

\Bloom:
\gab{4900}
Good night.
I'll just wait and take him along in a few ...

\stage{(Corny Kelleher returns to the outside car and mounts it.
The horse harness jingles.)}

\Corny:
\stage{(from the car, standing)}
Night.

\Bloom:
Night.

\stage{(The jarvey chucks the reins and raises his whip encouragingly.
The car and horse back slowly, awkwardly, and turn.
Corny Kelleher on the sideseat sways his head to and fro
\gab{4910}
in sign of mirth at Bloom's plight.
The jarvey joins in the mute pantomimic merriment nodding from the farther seat.
Bloom shakes his head in mute mirthful reply.
With thumb and palm Corny Kelleher reassures
that the two bobbies will allow the sleep to continue
for what else is to be done.
With a slow nod Bloom conveys his gratitude
as that is exactly what Stephen needs.
The car jingles tooraloom round the corner of the tooraloom lane.
Corny Kelleher again reassuralooms with his hand.
Bloom with his hand assuralooms Corny Kelleher that he is reassuraloomtay.
The tinkling hoofs and jingling harness grow fainter
with their tooralooloo looloo lay.
\gab{4920}
Bloom, holding in his hand Stephen's hat,
festooned with shavings, and ashplant, stands irresolute.
Then he bends to him and shakes him by the shoulder.)}

\Bloom:
Eh! Ho!
\stage{(There is no answer. He bends again)}
Mr~Dedalus!
\stage{(there is no answer)}
The name if you call.
Somnambulist.
\stage{(he bends again and hesitating,
brings his mouth near the face of the prostrate form)}
Stephen!
\stage{(There is no answer. He calls again.)}
Stephen!

\Stephen:
\stage{(frowns)}
\gab{4930}
Who?
Black panther. Vampire.
\stage{(he sighs and stretches himself, then murmurs thickly with prolonged vowels)}
\begin{verse}
    Who ... drive ... Fergus now\\
    And pierce ... wood's woven shade ...?
\end{verse}

\stage{(He turns on his left side, sighing, doubling himself together.)}

\Bloom:
Poetry.
Well educated.
Pity.
\stage{(he bends again and undoes the buttons of Stephen's waistcoat)}
To breathe.
\stage{(he brushes the woodshavings from Stephen's clothes with light hand and fingers)}
One pound seven.
Not hurt anyhow.
\stage{(he listens)}
What?

\Stephen:
\gab{4940}
\stage{(murmurs)}
\begin{verse}
    ... shadows ... the woods\\
    ... white breast... dim sea.
\end{verse}

\stage{(He stretches out his arms, sighs again and curls his body.
Bloom, holding the hat and ashplant, stands erect.
A dog barks in the distance.
Bloom tightens and loosens his grip on the ashplant.
He looks down on Stephen's face and form.)}

\Bloom:
\stage{(communes with the night)}
Face reminds me of his poor mother.
In the shady wood.
The deep white breast.
Ferguson, I think I caught.
\gab{4950}
A girl.
Some girl.
Best thing could happen him.
\stage{(he murmurs)}
... swear that I will always hail, ever conceal, never reveal,
any part or parts, art or arts ...
\stage{(he murmurs)}
... in the rough sands of the sea ...
a cabletow's length from the shore ...
where the tide ebbs ...
and flows ...

\stage{(Silent, thoughtful, alert he stands on guard,
his fingers at his lips in the attitude of secret master.
Against the dark wall a figure appears slowly,
a fairy boy of eleven, a changeling, kidnapped,
dressed in an Eton suit with glass shoes and a little bronze helmet,
holding a book in his hand.
He reads from right to left inaudibly, smiling, kissing the page.)%
\gab{4960}
}

\Bloom:
\stage{(wonderstruck, calls inaudibly)}
Rudy!

\Rudy:
\stage{(Gazes, unseeing, into Bloom's eyes and goes on reading, kissing, smiling.
He has a delicate mauve face.
On his suit he has diamond and ruby buttons.
In his free left hand he holds a slim ivory cane with a violet bowknot.
A white lambkin peeps out of his waistcoat pocket.)}

\end{document}
