%--------------------------------------------------
\commentary{1. \textsc{Circe} --- Cardinal Fantasy}{2:41:06}
%--------------------------------------------------

The value of \verb|\parindent| is \texttt{\the\parindent}. Here are the other values:

\begin{itemize}
\item The value of \verb|\baselineskip| is \texttt{\the\baselineskip}
\item The value of \verb|\parskip| is \texttt{\the\parskip}
\end{itemize}

\N{1a}:
\gab{2640}
Stephen Dedalus |
over his shoulder |
to Zoe Higgins,
a young whore.

\Stephen:
You would have preferred the fighting parson |
who founded the protestant error.
But beware Antisthenes,
the dog sage,
and the last end of Arius Heresiarchus.
The agony in the closet.

\N{1b}:
Vincent Lynch,
a medical student.

\Lynch:
All one and the same God to her.

\Stephen:
\stage{(devoutly)}
And sovereign Lord of all things.

\N{1a}:
Florry Talbot,
a blond feeble goose·fat whore.

\Florry:
\stage{(to Stephen)}
I'm sure you're a spoiled priest.
Or a monk.

\Lynch:
\gab{2650}
He is.
A cardinal's son.

\Stephen:
Cardinal sin.
Monks of the screw.

\N{1b}:
His Eminence |
Simon Stephen |
cardinal Dedalus,
primate of all Ireland,
appears in the doorway,
dressed in red soutane,
sandals and socks.
Seven dwarf simian acolytes,
also in red,
cardinal sins,
uphold his train,
peeping under it.
He wears a battered silk hat sideways on his head.
His thumbs are stuck in his armpits |
and his palms outspread.
Round his neck |
hangs a rosary of corks |
ending on his breast |
in a corkscrew cross.
\gab{2660}%
Releasing his thumbs,
he invokes grace from on high |
with large wave gestures |
and proclaims with bloated pomp.

\Cardinal:
\begin{verse}
    Conservio lies captured\\
    He lies in the lowest dungeon\\
    With manacles and chains around his limbs\\
    Weighing upwards of three tons.
\end{verse}

\N{1b}:
He looks at all for a moment,
his right eye closed tight,
his left cheek puffed out.
Then,
unable to repress his merriment,
he rocks to and fro,
\gab{2670}%
arms akimbo,
and sings with broad rollicking humour.

\Cardinal:
\begin{verse}
    O, the poor little fellow\\
    Hi·hi·hi·hi·his legs they were yellow\\
    He was plump, fat and heavy and brisk as a snake\\
    But some bloody savage\\
    To graize his white cabbage\\
    He murdered Nell Flaherty's duck·loving drake.
\end{verse}

\N{1b}:
A multitude of midges swarms white over his robe.
He scratches himself with crossed arms at his ribs,
grimacing,
and exclaims.

\Cardinal:
I'm suffering the agony of the damned.
By the hoky fiddle,
thanks be to Jesus |
those funny little chaps are not unanimous.
\gab{2680}%
If they were,
they'd walk me off the face of the bloody globe.

\N{1b}:
His head aslant,
he blesses curtly with fore and middle fingers,
imparts the Easter kiss |
and double·shuffles off comically,
swaying his hat from side to side,
shrinking quickly to the size of his train·bearers.
The dwarf acolytes,
giggling,
peeping,
nudging,
ogling,
easter·kissing,
zig·zag behind him.
His voice is heard mellow from afar,
merciful male,
melodious.

\Cardinal:
\begin{verse}
    Shall carry my heart to thee,\\
    Shall carry my heart to thee,\\
\gab{2690}%
    And the breath of the balmy night\\
    Shall carry my heart to thee!
\end{verse}

\N{1a}:
The trick door·handle turns.

\Doorhandle:
Theeee!

\Zoe:
The devil is in that door.

\N{1a}:
A male form passes down the creaking staircase |
and is heard |
taking the waterproof and hat from the rack.
Leopold Bloom starts forward involuntarily |
and,
half closing the door as he passes,
takes the chocolate from his pocket |
\gab{2700}%
and offers it nervously to Zoe.

\N{1b}:
Zoe sniffs his hair briskly.

\Zoe:
Hmmm!
Thank your mother for the rabbits.
I'm very fond of what I like.

\N{1a}:
Bloom,
hearing a male voice |
in talk with the whores on the door·step,
pricks his ears.

\Bloom:
If it were he?
After?
Or because not?
Or the double event?

\N{1b}:
Zoe tears open the silver·foil.

\Zoe:
Fingers was made before forks.

\N{1b}:
She breaks off and nibbles a piece,
gives a piece to Kitty Ricketts |
and then turns kittenishly to Lynch.
\gab{2710}

\Zoe:
No objection to French lozenges?

\N{1b}:
Lynch nods.
Zoe taunts him.

\Zoe:
Have it now or wait till you get it?

\N{1b}:
Lynch opens his mouth,
his head cocked.
Zoe whirls the prize in left circle.
His head follows.
She whirls it back in right circle.
He eyes her.

\Zoe:
Catch!

\N{1b}:
Zoe tosses a piece.
With an adroit snap,
Lynch catches it |
and bites it through |
with a crack.

\Kitty:
\stage{(chewing)}
The engineer I was with at the bazaar |
does have lovely ones.
Full of the best liqueurs.
And the viceroy was there |
with his lady.
The gas we had |
on the Toft's hobby·horses.
I'm giddy still.

\gab{2720}
\N{1a}:
Bloom in Svengali's fur overcoat,
with folded arms and Napoleonic forelock,
frowns in ventriloquial exorcism |
with piercing eagle glance |
towards the door.
Then,
rigid with left foot advanced,
he makes a swift pass with impelling fingers |
and gives the sign of past master,
drawing his right arm downwards from his left shoulder.

\Bloom:
Go,
go,
go,
I conjure you,
whoever you are!

\N{1a}:
A male cough and tread are heard |
passing through the mist outside.
Bloom's features relax.
He places a hand in his waistcoat,
posing calmly.
Zoe offers him chocolate.

\Bloom:
\gab{2730}
\stage{(solemnly)}
Thanks.

\Zoe:
Do as you're bid.
Here!

\N{1b}:
A firm heel·clacking tread is heard on the stairs.

\N{1a}:
Bloom takes the chocolate.

\Bloom:
Aphrodisiac?
Tansy and penny·royal.
But I bought it.
Vanilla calms or?
Mnemo.
Confused light confuses memory.
Red influences lupus.
Colours affect women's characters,
any they have.
This black makes me sad.
Eat and be merry for tomorrow.
\stage{(he eats)}
Influence taste too,
mauve.
But it is so long since I\@.
Seems new.
Aphro.
That priest.
\gab{2740}%
Must come.
Better late than never.
Try truffles at Andrews.

\N{1b}:
The door opens.
Bella Cohen,
a massive whore·mistress,
enters.
She is dressed in a three·quarter ivory gown,
fringed round the hem with tasselled selvedge,
and cools herself |
flirting a black horn fan |
like Minnie Hauck in \Title{Carmen}.
On her left hand |
are wedding and keeper rings.
Her eyes are deeply carboned.
She has a sprouting moustache.
Her olive face is heavy,
slightly sweated and full·nosed |
with orange·tainted nostrils.
She has large pendant beryl eardrops.

\Bella:
My word!
\gab{2750}%
I'm all of a muck·sweat.

\N{1b}:
She glances round her at the couples.
Then her eyes rest on Bloom |
with hard insistence.

\N{1a}:
Her large fan winnows wind |
towards her heated face·neck and embonpoint.

\N{1b}:
Her falcon eyes glitter.

\N{1a}:
The fan,
flirting quickly,
then slowly.

\Fan:
Married,
I see.

\Bloom:
Yes.
Partly,
I have mislaid…

\N{1a}:
The fan,
half opening,
then closing.

\Fan:
And the missus is master.
Petticoat government.%
\gab{2760}

\N{1a}:
Bloom looks down with a sheepish grin.

\Bloom:
That is so.

\N{1a}:
The fan,
folding together,
rests against her left eardrop.

\Fan:
Have you forgotten me?

\Bloom:
Nes.
Yo.

\N{1a}:
The fan,
folded akimbo against her waist.

\Fan:
Is me her |
was you dreamed before?
Was then she him |
you us |
since knew?
Am all them |
and the same now |
we?

\N{1b}:
Bella approaches,
\gab{2770}%
gently tapping with the fan.

\N{1a}:
Bloom winces.

\Bloom:
Powerful being.
In my eyes |
read that slumber which women love.

\N{1a}:
The fan taps.

\Fan:
We have met.
You are mine.
It is fate.

\N{1a}:
Bloom is cowed.

\Bloom:
Exuberant female.
Enormously I desiderate your domination.
I am exhausted,
abandoned,
no more young.
I stand,
so to speak,
with an unposted letter |
bearing the extra regulation fee |
before the too late box |
of the general post·office of human life.
\gab{2780}%
The door and window open at a right angle |
cause a draught of thirty·two feet per second |
according to the law of falling bodies.
I have felt this instant |
a twinge of sciatica |
in my left glutear muscle.
It runs in our family.
Poor dear papa,
a widower,
was a regular barometer from it.
He believed in animal heat.
A skin of tabby |
lined his winter waistcoat.
Near the end,
remembering king David and the Sunamite,
he shared his bed with Athos,
faithful after death.
A dog's spittle as you probably…
\stage{(he winces)}
Ah!

\N{1b}:
Richie Goulding,
bag·weighted,
passes the door.

\Richie:
Mocking is catch.
Best value in Dub.
Fit for a prince's.
\gab{2790}%
Liver and kidney.

\N{1a}:
The fan taps.

\Fan:
All things end.
Be mine.
Now.

\N{1a}:
Bloom is undecided.

\Bloom:
All now?
I should not have parted with my talisman.
Rain,
exposure at dewfall on the sea·rocks,
a peccadillo at my time of life.
Every phenomenon has a natural cause.

\N{1a}:
The fan points downwards slowly.

\Fan:
You may.

\N{1a}:
Bloom looks downwards |
and perceives her unfastened boot·lace.
\gab{2800}

\Bloom:
We are observed.

\N{1a}:
The fan points downwards quickly.

\Fan:
You must.

\N{1a}:
Bloom,
with desire,
with reluctance.

\Bloom:
I can make a true black knot.
Learned when I served my time |
and worked the mail order line for Kellett's.
Experienced hand.
Every knot says a lot.
Let me.
In courtesy.
I knelt once before today.
Ah!

\N{1b}:
Bella raises her gown slightly |
and,
steadying her pose,
lifts to the edge of a chair |
a plump buskined hoof |
\gab{2810}%
and a full pastern,
silk·socked.

\N{1a}:
Bloom,
stiff·legged,
aging,
bends over her hoof |
and with gentle fingers |
draws out and in her laces.
He murmurs lovingly.

\Bloom:
To be a shoe·fitter in Manfield's |
was my love's young dream,
the darling joys of sweet button·hooking,
to lace up criss·crossed to knee·length |
the dressy kid footwear satin·lined,
so incredibly impossibly small,
of Clyde Road ladies.
Even their wax model Raymonde,
I visited daily |
to admire her cobweb hose |
and stick of rhubarb toe,
as worn in Paris.

\Hoof[1b]:
Smell my hot goat·hide.
\gab{2820}%
Feel my royal weight.

\N{1a}:
Bloom,
cross·lacing.

\Bloom:
Too tight?

\Hoof:
If you bungle,
Handy Andy,
I'll kick your football for you.

\Bloom:
Not to lace the wrong eyelet |
as I did the night of the bazaar dance.
Bad luck.
Hook in wrong tache of her…
person you mentioned.
That night she met…
Now!

\N{1a}:
He knots the lace.

%--------------------------------------------------
\commentary{2. \textsc{Unmanned} --- Bello Fantasy}{2:52:03}
%--------------------------------------------------

\N{2b}:
Bella places \emph{her} foot on the floor.

\N{2a}:
Bloom raises \emph{his} head.
\gab{2830}

\N{2b}:
\emph{Her} heavy face,
\emph{her} eyes strike \emph{him} in midbrow.

\N{2a}:
\emph{His} eyes grow dull,
darker and pouched,
\emph{his} nose thickens.

\Bloom:
\stage{(mumbles)}
Awaiting your further orders |
we remain,
gentlemen,
…

\N{2b}:
\emph{Bello},
with a hard basilisk stare,
in a baritone voice.

\Bello:
Hound of dishonour!

\N{2a}:
Bloom,
infatuated.

\Bloom:
Empress!

\N{2b}:
\emph{Bello},
\emph{his} heavy cheek·chops sagging.

\Bello:
Adorer of the adulterous rump!

\Bloom:
\gab{2840}
\stage{(plaintively)}
Hugeness!

\Bello:
Dung·devourer!

\N{2a}:
Bloom,
with sinews semi·flexed.

\Bloom:
Mag·magnificence!

\Bello:
Down!

\N{2b}:
Bello taps Bloom |
on \emph{her} shoulder |
with \emph{his} fan.

\Bello:
Incline feet forward!
Slide left foot |
one pace back!
You will fall.
You are falling.
On the hands |
down!

\N{2a}:
Bloom,
her eyes upturned |
in the sign of admiration,
closing,
yaps.

\Bloom:
Truffles!

\N{2a}:
With a piercing epileptic cry,
she sinks on all fours,
grunting,
snuffling,
rooting at his feet:
then lies,
shamming dead,
with eyes shut tight,
trembling eyelids,
bowed upon the ground |
in the attitude of most excellent master.

\N{2b}:
Bello,
with bobbed hair,
purple gills,
fit moustache rings round his shaven mouth,
in mountaineer's puttees,
green silver·buttoned coat,
sport skirt |
and alpine hat with moorcock's feather,
his hands stuck deep in his breeches pockets,
places his heel on her neck |
and grinds it in.
\gab{2860}

\Bello:
Foot·stool!
Feel my entire weight.
Bow,
bond·slave,
before the throne of your despot's glorious heels |
so glistening in their proud erectness.

\N{2a}:
Bloom,
enthralled,
bleats.

\Bloom:
I promise never to disobey.

\N{2b}:
Bello laughs loudly.

\Bello:
\stage{(laughs)}
Holy smoke!
You little know what's in store for you.
I'm the Tartar |
to settle your little lot  |
and break you in!
I'll bet Kentucky cocktails all round,
I shame it out of you,
old son.
Cheek me,
I dare you.
If you do,
tremble in anticipation of heel discipline |
to be inflicted in gym costume.%
\gab{2870}

\N{2a}:
Bloom creeps under the sofa |
and peers out through the fringe.

\N{2a}:
Zoe,
widening her slip to screen Bloom.

\Zoe:
She's not here.

\N{2a}:
Bloom,
closing her eyes.

\Bloom:
She's not here.

\N{2a}:
Florry,
hiding her with her gown.

\Florry:
She didn't mean it,
Mr~Bello.
She'll be good,
sir.

\Kitty:
Don't be too hard on her,
Mr~Bello.
\gab{2880}%
Sure you won't,
ma'am·sir.

\Bello:
\stage{(coaxingly)}
Come,
ducky dear,
I want a word with you,
darling,
just to administer correction.
Just a little heart to heart talk,
sweety.

\N{2a}:
Bloom puts out her timid head.

\Bello:
There's a good girly now.

\N{2b}:
Bello grabs her hair violently and drags her forward.

\Bello:
I only want to correct you for your own good |
on a soft safe spot.
How's that tender behind?
O,
ever so gently,
pet.
Begin to get ready.

\N{2a}:
Bloom,
fainting.

\Bloom:
Don't tear my…

\Bello:
\gab{2890}
\stage{(savagely)}
The nose·ring,
the pliers,
the bastinado,
the hanging hook,
the knout |
I'll make you kiss |
while the flutes play |
like the Nubian slave of old.
You're in for it this time!
I'll make you remember me |
for the balance of your natural life.

\N{2b}:
His forehead veins swollen,
his face congested.

\Bello:
I shall sit on your ottoman saddle·back every morning |
after my thumping good breakfast |
of Matterson's fat ham·rashers |
and a bottle of Guinness's porter.

\N{2b}:
Bello belches.

\Bello:
\stage{(belching)}
And suck my thumping good Stock Exchange cigar |
while I read the \Title{Licensed Victualler's Gazette}.
Very possibly |
I shall have you slaughtered and skewered |
in my stables |
and enjoy a slice of you |
with crisp crackling from the baking tin |
\gab{2900}%
basted and baked like sucking pig |
with rice and lemon |
or currant sauce.
It will hurt you.

\N{2b}:
He twists her arm.

\N{2a}:
Bloom squeals,
turning turtle.

\Bloom:
\stage{(squealing)}
Don't be cruel,
nurse!
Don't!

\N{2b}:
Bello,
twisting.

\Bello:
Another!

\N{2a}:
Bloom screams.

\Bloom:
\stage{(screaming)}
O,
it's hell itself!
Every nerve in my body aches like mad!

\Bello:
\stage{(shouts)}
Good,
by the rumping jumping general!
\gab{2910}%
That's the best bit of news I heard |
these six weeks.
Here,
don't keep me waiting,
damn you!

\N{2b}:
Bello slaps her face.

\N{2a}:
Bloom whimpers.

\Bloom:
You're after hitting me.
I'll tell…

\Bello:
Hold him down,
girls,
till I squat on him.

\Zoe:
Yes.
Walk on him!
I will.

\Florry:
I will.
\gab{2920}%
Don't be greedy.

\Kitty:
No,
me.
Lend him to me.

\N{2b}:
The brothel cook,
Mrs~Keogh,
wrinkled,
grey·bearded,
in a greasy bib,
men's grey and green socks |
and brogues,
flour·smeared,
a rolling·pin stuck with raw pastry |
in her bare red arm and hand,
appears at the door.

\MrsKeogh:
\stage{(ferociously)}
Can I help?

\N{2b}:
They hold and pinion Bloom.

\N{2b}:
Bello,
squats with a grunt on Bloom's upturned face,
\gab{2930}%
puffing cigar·smoke,
nursing a fat leg.

\Bello:
I see Keating Clay is elected vice·chairman of the Richmond asylum |
and by~the~by |
Guinness's preference shares are at sixteen three·quarters.
Curse me for a fool |
that didn't buy that lot |
Craig and Gardner told me about.
Just my infernal luck,
curse it.
And that Goddamned outsider \Title{Throwaway} |
at twenty to one.

\N{2b}:
He quenches his cigar angrily on Bloom's ear.

\Bello:
Where's that Goddamned cursed ashtray?

\N{2a}:
Bloom,
goaded,
buttock·smothered.

\Bloom:
O!
O!
Monsters!
Cruel one!

\Bello:
\gab{2940}%
Ask for that every ten minutes.
Beg.
Pray for it |
as you never prayed before.

\N{2b}:
He thrusts out a figged fist and foul cigar.

\Bello:
Here,
kiss that.
Both.
Kiss.

\N{2b}:
He throws a leg astride |
and,
pressing with horseman's knees,
calls in a hard voice.

\Bello:
Gee up!
A cock·horse to Banbury cross.
I'll ride him for the Eclipse stakes.

\N{2b}:
He bends sideways |
and squeezes his mount's testicles roughly,
shouting.

\Bello:
Ho!
Off we pop!
I'll nurse you in proper fashion.

\N{2b}:
He horse·rides cock·horse,
leaping in the,
in the saddle.

\Bello:
The lady goes a~pace a~pace |
and the coachman goes a~trot a~trot |
and the gentleman goes a~gallop a~gallop a~gallop a~gallop.

\N{2a}:
Florry pulls at Bello.%
\gab{2950}

\Florry:
Let me on him now.
You had enough.
I asked before you.

\N{2a}:
Zoe,
pulling at Florry.

\Zoe:
Me.
Me.
Are you not finished with him yet,
suckeress?

\N{2a}:
Bloom,
stifling.

\Bloom:
Can't.

\Bello:
Well,
I'm not.
Wait.

\N{2b}:
He holds in his breath.

\Bello:
Curse it.
Here.
This bung's about burst.

\N{2b}:
Bello uncorks himself behind:
then,
contorting his features,
farts loudly.

\Bello:
\stage{Pffffft!}
Take that!

\N{2b}:
He recorks himself.

\Bello:
Yes,
by Jingo,
sixteen three·quarters.%
\gab{2960}

\N{2a}:
Bloom,
a sweat breaking out over him.

\Bloom:
Not man.

\N{2a}:
He sniffs.

\Bloom:
Woman.

\N{2b}:
Bello stands up.

\Bello:
No more blow hot and cold.
What you longed for |
has come to pass.
Henceforth |
you are unmanned |
and mine in earnest,
a thing under the yoke.
Now for your punishment frock.
You will shed your male garments,
you understand,
Ruby Cohen?
and don the shot silk |
luxuriously rustling over head and shoulders.
And quickly too!

\N{2a}:
Bloom shrinks.
\gab{2970}

\Bloom:
Silk,
mistress said!
O~crinkly!
Scrapey!
Must I tip·touch it with my nails?

\N{2b}:
Bello points to his whores.

\Bello:
As they are now |
so will you be,
wigged,
singed,
perfume·sprayed,
rice·powdered,
with smooth·shaven armpits.
Tape measurements will be taken next your skin.
You will be laced with cruel force |
into vice·like corsets of soft dove coutille |
with whale·bone busk |
to the diamond·trimmed pelvis,
the absolute outside edge,
while your figure,
plumper than when at large,
will be restrained in net·tight frocks,
pretty two·ounce petticoats and fringes |
and things stamped,
of course,
with my house flag,
\gab{2980}%
creations of lovely lingerie for Alice |
and nice scent for Alice.
Alice will feel the pull·pull.
Martha and Mary will be a little chilly at first |
in such delicate thigh·casing |
but the frilly flimsiness of lace round your bare knees |
will remind you…

\N{2a}:
Bloom,
charming soubrette with dauby cheeks,
mustard hair |
and large male hands and nose,
leering mouth.

\Bloom:
I tried her things on |
only twice,
a small prank,
in Holles street.
When we were hard up,
I washed them to save the laundry bill.
My own shirts |
I turned.
It was the purest thrift.

\N{2b}:
Bello jeers.

\Bello:
Little jobs that make mother pleased,
eh?
\gab{2990}%
And showed off coquettishly |
in your domino at the mirror |
behind close·drawn blinds |
your unskirted thighs and he·goat's udders |
in various poses of surrender,
eh?
Ho!
Ho!
I have to laugh!
That second·hand black opera·top shift |
and short trunk·leg naughties |
all split up the stitches |
at her last rape |
that Mrs~Miriam Dandrade sold you |
from the Shelbourne hotel,
eh?

\Bloom:
Miriam.
Black.
Demi·mondaine.

\N{2b}:
Bello guffaws.

\Bello:
\stage{(guffaws)}
Christ Almighty,
it's too tickling,
this!
You were a nice·looking Miriam |
when you clipped off your back·gate hairs |
\gab{3000}%
and lay swooning in the thing across the bed |
as Mrs~Dandrade |
about to be violated by lieutenant Smythe-Smythe;
Mr~Philip Augustus Blockwell~M.P.;
signor Laci Daremo,
the robust tenor;
blue·eyed Bert,
the lift·boy;
Henri Fleury of Gordon Bennett fame;
Sheridan,
the quadroon Croesus;
the varsity wet·bob eight from old Trinity;
Ponto,
her splendid Newfoundland;
and Bobs,
dowager duchess of Manorhamilton.

\N{2b}:
Bello guffaws again.

\Bello:
\stage{(guffaws)}
Christ,
wouldn't it make a Siamese cat laugh?

\N{2a}:
Bloom,
her hands and features working.

\Bloom:
It was Gerald |
converted me to be a true corset lover |
\gab{3010}%
when I was female impersonator |
in the High School play \Title{Vice Versa}.
It was dear Gerald.
He got that kink,
fascinated by sister's stays.
Now dearest Gerald uses pinky grease·paint |
and gilds his eyelids.
Cult of the beautiful.

\N{2b}:
Bello with wicked glee.

\Bello:
Beautiful!
Give us a breather!
When you took your seat with womanish care,
lifting your billowy flounces,
on the smooth·worn throne.

\Bloom:
Science.
To compare the various joys we each enjoy.
\stage{(earnestly)}
And really |
it's better the position…
\gab{3020}%
because often I used to wet…

\N{2b}:
Bello,
sternly.

\Bello:
No insubordination!
The sawdust is there in the corner for you.
I gave you strict instructions,
didn't I?
Do it standing,
sir!
I'll teach you to behave like a jinkleman!
If I catch a trace on your swaddles.
Aha!
By the ass of the Dorans,
you'll find I'm a martinet.
The sins of your past are rising against you.
Many.
Hundreds.


\N{2a}:
The Sins of the Past,
in a medley of voices.

\SinsPast:\\
-- He went through a form of clandestine marriage |
with at least one woman |
in the shadow of the Black church.\\
-- Unspeakable messages |
he telephoned mentally to Miss Dunn |
\gab{3030}%
at an address in D'Olier street |
while he presented himself indecently |
to the instrument in the callbox.\\
-- By word and deed,
he frankly encouraged a nocturnal strumpet |
to deposit fecal and other matter |
in an unsanitary outhouse attached to empty premises.\\
-- In five public conveniences,
he wrote pencilled messages |
offering his nuptial partner |
to all strong·membered males.\\
-- And by the offensively smelling vitriol works,
did he not pass night after night |
by loving courting couples |
to see if and what and how much |
he could see?\\
-- Did he not lie in bed,
the gross boar,
gloating over a nauseous fragment of well·used toilet paper |
presented to him by a nasty harlot,
stimulated by ginger·bread and a postal order?%
\gab{3040}

\N{2b}:
Bello whistles loudly.

\Bello:
Say!
What was the most revolting piece of obscenity |
in all your career of crime?
Go the whole hog.
Puke it out!
Be candid for once.

\N{2b}:
Mute inhuman faces throng forward,
leering,
vanishing,
gibbering,
Boo·loo·hoom,
Poldy Kock,
Bootlaces a penny,
Cassidy's hag,
blind stripling,
Larry rhinoceros,
the girl,
the woman,
the whore,
the other,
the…

\Bloom:
Don't ask me!
Our mutual faith.
Pleasants street.
I only thought the half of the…
I swear on my sacred oath…%
\gab{3050}

\N{2b}:
Bello,
peremptorily.

\Bello:
Answer.
Repugnant wretch!
I insist on knowing.
Tell me something to amuse me,
smut |
or a bloody good ghost·story |
or a line of poetry,
quick,
quick,
quick!
Where?
How?
What time?
With how many?
I give you just three seconds.
One!
Two!
Thr---

\N{2a}:
Bloom,
docile,
gurgles.

\Bloom:
I re·re·re·pugnosed in re·re·re·repugnant

\N{2b}:
Bello,
imperiously.

\Bello:
O,
get out,
you skunk!
Hold your tongue!
Speak when you're spoken to.%
\gab{3060}

\N{2a}:
Bloom bows.

\Bloom:
Master!
Mistress!
Man tamer!

\N{2a}:
He lifts his arms.
His bangle bracelets fall.

\N{2b}:
Bello,
satirically.

\Bello:
By day |
you will souse and bat our smelling underclothes |
also when we ladies are unwell,
and swab out our latrines |
with dress pinned up |
and a dish·clout tied to your tail.
Won't that be nice?

\N{2b}:
Bello places a ruby ring on her finger.

\Bello:
And there now!
With this ring |
I thee own.
Say,
\emph{thank you,
mistress}.

\Bloom:
\gab{3070}
Thank you,
mistress.

\Bello:
You will make the beds,
get my tub ready,
empty the piss·pots in the different rooms,
including old Mrs~Keogh's the cook's,
a sandy one.
Ay,
and rinse the seven of them well,
mind,
or lap it up like champagne.
Drink me |
piping hot.
Hop!
You will dance attendance |
or I'll lecture you on your misdeeds,
Miss Ruby,
and spank your bare bot right well,
miss,
with the hair·brush.
You'll be taught the error of your ways.
At night |
your well·creamed braceletted hands |
will wear forty·three·button gloves |
new·powdered with talc |
and having delicately scented finger·tips.
\gab{3080}%
For such favours,
knights of old laid down their lives.
\stage{(he chuckles)}
My boys will be no end charmed |
to see you so ladylike,
the colonel,
above all,
when they come here |
the night before the wedding |
to fondle my new attraction in gilded heels.
First,
I'll have a go at you myself.
A man I know on the turf |
named Charles Alberta Marsh
(I was in bed with him just now |
and another gentleman |
out of the Hanaper and Petty Bag office)
is on the lookout |
for a maid of all work |
at a short knock.
Swell the bust.
Smile.
Droop shoulders.
What offers?

\N{2b}:
Bello points.

\Bello:
For that lot.
Trained by owner to fetch and carry,
basket in mouth.

\N{2b}:
Bello bares his arm |
and plunges it elbow·deep in Bloom's vulva.

\Bello:
There's fine depth for you!
\gab{3090}
What,
boys?
That give you a hard·on?

\N{2b}:
Bello shoves his arm in a bidder's face.

\Bello:
Here,
wet the deck and wipe it round!

\Bidder:
A florin.

\N{2a}:
Dillon's lacquey rings his handbell.

\Lacquey:
Barang!

\Voice:
One and eight·pence too much.

\Marsh[2b]:
Must be virgin.
\gab{3100}
Good breath.
Clean.

\N{2b}:
Bello gives a rap with his gavel.

\Bello:
Two bar.
Rock·bottom figure and cheap at the price.
Fourteen hands high.
Touch and examine his points.
Handle him.
This downy skin,
these soft muscles,
this tender flesh.
If I had only my gold piercer here!
And quite easy to milk.
Three new·laid gallons a day.
A pure stock·getter,
due to lay within the hour.
His sire's milk record |
was a thousand gallons of whole milk in forty weeks.
Whoa my jewel!
Beg up!
Whoa!

\N{2b}:
He brands his initial~C |
on Bloom's croup.

\Bello:
So!
Warranted Cohen!
What advance on two bob,
gentlemen?

\N{2b}:
A Dark·visaged Man,
in disguised accent.
\gab{3110}

\DarkvisagedMan:
Hoondert punt sterlink.

\Voices:
\stage{(subdued)}
For the Caliph.
Haroun Al Raschid.

\Bello:
\stage{(gaily)}
Right.
Let them all come.
The scanty,
daringly short skirt,
riding up at the knee |
to show a peep of white pantalette,
is a potent weapon |
and transparent stockings,
emerald·gartered,
with the long straight seam trailing up beyond the knee,
appeal to the better instincts |
of the \french{blasé} man about town.
Learn the smooth mincing walk |
on four·inch Louis Quinze heels,
the Grecian bend with provoking croup,
\gab{3120}%
the thighs fluescent,
knees modestly kissing.
Bring all your powers of fascination |
to bear on them.
Pander to their Gomorrahan vices.

\N{2a}:
Bloom bends his blushing face into his armpit |
and simpers with forefinger in mouth.

\Bloom:
O,
I know what you're hinting at now!

\Bello:
What else are you good for,
an impotent thing like you?

\N{2b}:
He stoops and,
peering,
pokes with his fan rudely |
under the fat suet folds of Bloom's haunches.

\Bello:
Up!
Up!
Manx cat!
What have we here?
Where's your curly teapot gone to |
or who docked it on you,
cocky·olly?
Sing,
birdy,
sing.
\gab{3130}%
It's as limp |
as a boy of six's |
doing his pooly behind a cart.
Buy a bucket or sell your pump.
\stage{(loudly)}
Can you do a man's job?

\Bloom:
Eccles street…

\Bello:
\stage{(sarcastically)}
I wouldn't hurt your feelings for the world |
but there's a man of brawn in possession there.
The tables are turned,
my gay young fellow!
He is something like a full·grown outdoor man.
Well for you,
you muff,
if you had that weapon |
with knobs and lumps and warts all over it.
He shot his bolt,
I can tell you!
\gab{3140}%
Foot to foot,
knee to knee,
belly to belly,
bubs to breast!
He's no eunuch.
A shock of red hair he has |
sticking out of him behind like a furze·bush!
Wait for nine months,
my lad!
Holy ginger,
it's kicking and coughing up and down |
in her guts already!
That makes you wild,
don't it?
Touches the spot?

\N{2b}:
He spits in contempt.

\Bello:
Spittoon!

\Bloom:
I was indecently treated,
I…
Inform the police.
Hundred pounds.
Unmentionable.
I…

\Bello:
Would if you could,
lame duck.
A downpour we want,
not your drizzle.

\Bloom:
\gab{3150}
To drive me mad!
Moll!
I forgot!
Forgive!
Moll…
We…
Still…

\N{2b}:
Bello,
ruthlessly.

\Bello:
No,
Leopold Bloom,
all is changed by woman's will |
since you slept horizontal in Sleepy Hollow |
your night of twenty years.
Return and see.

\N{2b}:
Old Sleepy Hollow calls over the wold.

\SleepyHollow:
Rip van Wink!
Rip van Winkle!

\N{2a}:
Bloom,
in tattered mocassins |
with a rusty fowling·piece,
\gab{3160}%
tiptoeing,
finger·tipping,
his haggard bony bearded face peering through the diamond panes,
cries out.

\Bloom:
I see her!
It's she!
The first night at Mat Dillon's!
But that dress,
the green!
And her hair is dyed gold and he…

\N{2b}:
Bello laughs mockingly.

\Bello:
That's your daughter,
you owl,
with a Mullingar student.

\N{2a}:
Milly Bloom,
fair·haired,
green·vested,
slim·sandalled,
her blue scarf in the sea·wind simply swirling,
breaks from the arms of her lover |
and calls,
her young eyes wonder·wide.

\Milly:
\gab{3170}
My!
It's Papli!
But,
O~Papli,
how old you've grown!

\Bello:
Changed,
eh?
Our what·not,
our writing table where we never wrote,
aunt Hegarty's armchair,
our classic reprints of old masters.
A man and his men·friends are living there in clover.
The \Title{Cuckoos' Rest}!
Why not?
How many women had you,
eh,
following them up dark streets,
flat·foot,
exciting them by your smothered grunts,
what,
you male prostitute?
Blameless dames with parcels of groceries.
Turn about.
Sauce for the goose,
my gander~O\@.

\Bloom:
\gab{3180}
They… I~…

\Bello:
\stage{(cuttingly)}
Their heel·marks will stamp the Brusselette carpet |
you bought at Wren's auction.
In their horse·play with Moll |
the romp to find the buck flea in her breeches,
they will deface the little statue |
you carried home in the rain |
for art for art' sake.
They will violate the secrets of your bottom drawer.
Pages will be torn from your handbook of astronomy |
to make them pipe·spills.
And they will spit |
in your ten shilling brass fender from Hampton Leedom's.

\Bloom:
\gab{3190}
Ten and six.
The act of low scoundrels.
Let me go.
I will return.
I will prove…

\Voice:
Swear!

\N{2a}:
Bloom clenches his fists and crawls forward,
a bowie·knife between his teeth.

\Bello:
As a paying guest or a kept man?
Too late.
You have made your second·best bed |
and others must lie in it.
Your epitaph is written.
You are down and out |
and don't you forget it,
old bean.%
\gab{3200}

\Bloom:
Justice!
All Ireland versus one!
Has nobody…?

\N{2a}:
He bites his thumb.

\Bello:
Die and be damned to you |
if you have any sense of decency or grace about you.
I can give you a rare old wine |
that'll send you skipping to hell and back.
Sign a will |
and leave us any coin you have!
If you have none,
see you damn well get it,
steal it,
rob it!
We'll bury you in our shrubbery jakes |
where you'll be dead and dirty  |
with old Cuck Cohen,
my step·nephew I married,
the bloody old gouty procurator and sodomite |
with a crick in his neck,
and my other ten or eleven husbands,
\gab{3210}%
whatever the buggers' names were,
suffocated in the one cesspool.

\N{2b}:
He explodes in a loud phlegmy laugh.

\Bello:
\stage{(laughs)}
We'll manure you,
Mr~Flower!

\N{2b}:
He pipes scoffingly.

\Bello:
Byby,
Poldy!
Byby,
Papli!

\N{2a}:
Bloom,
clasps his head.

\Bloom:
My willpower!
Memory!
I have sinned!
I have suff---

\N{2a}:
He weeps tearlessly.

\Bello:
\stage{(sneers)}
Crybabby!
Crocodile tears!

%--------------------------------------------------
\commentary{3. \textsc{Nymph} --- Nymph Fantasy}{3:16:39}
%--------------------------------------------------

\N{3a}:
Bloom,
broken,
closely veiled for the sacrifice,
sobs,
his face to the earth.
\gab{3220}%
The passing bell is heard.
Dark·shawled figures of the circumcised,
in sackcloth and ashes,
stand by the wailing wall.
M.~Shulomowitz,
Joseph Goldwater,
Moses Herzog,
Harris Rosenberg,
M.~Moisel,
J.~Citron,
Minnie Watchman,
P.~Mastiansky,
the reverend Leopold Abramovitz,
chazen.
With swaying arms,
they wail in pneuma over the recreant Bloom.

\N{3b}:
The Circumcised,
in dark guttural chant |
as they cast dead sea fruit upon him,
no flowers.

\Circumcised:
\hebrew{Shema Israel Adonai Elohenu Adonai Echad.}

\Voices:
\gab{3230}
\stage{(sighing)}\\*
-- So he's gone.\\
-- Ah yes.\\
-- Yes, indeed.\\
-- Bloom?\\
-- Never heard of him.\\
-- No?\\
-- Queer kind of chap.\\
-- There's the widow.\\
-- That so?\\
-- Ah, yes.

\N{3b}:
From the suttee pyre,
the flame of gum camphire ascends.
The pall of incense smoke screens and disperses.
Out of her oak·frame,
a nymph with hair unbound,
lightly clad in tea·brown art colours,
descends from her grotto |
and passing under interlacing yews,
stands over Bloom.

\N{3b}:
The Yews,
their leaves whispering.

\Yews:
Sister.
Our sister.
Ssh!

\Nymph[3b]:
\stage{(softly)}
\gab{3240}
Mortal!
\stage{(kindly)}
Nay,
dost not weepest!

\N{3a}:
Bloom crawls jellily forward under the boughs,
streaked by sunlight,
with dignity.

\Bloom:
This position.
I felt it was expected of me.
Force of habit.

\Nymph:
Mortal!
You found me in evil company,
high·kickers,
coster picnic·makers,
pugilists,
popular generals,
immoral panto boys in flesh·tights |
and the nifty shimmy dancers,
La Aurora and Karini,
musical act,
the hit of the century.
I was hidden in cheap pink paper that smelt of rock oil.
I was surrounded by the stale smut of club·men,
stories to disturb callow youth,
ads for transparencies,
trued·up dice and bust·pads,
proprietary articles |
\gab{3250}%
and why wear a truss |
with testimonial from ruptured gentleman.
Useful hints to the married.

\N{3a}:
Bloom lifts a turtle head towards her lap.

\Bloom:
We have met before.
On another star.

\Nymph:
\stage{(sadly)}
Rubber goods.
Never·rip brand as supplied to the aristocracy.
Corsets for men.
I cure fits or money refunded.
Unsolicited testimonials for Professor Waldmann's wonderful chest exuber.
My bust developed four inches in three weeks,
reports Mrs~Gus Rublin with photo.

\Bloom:
\gab{3260}
You mean \Title{Photo Bits}?

\Nymph:
I do.
You bore me away,
framed me in oak and tinsel,
set me above your marriage couch.
Unseen,
one summer eve,
you kissed me in four places.
And with loving pencil,
you shaded my eyes,
my bosom and my shame.

\N{3a}:
Bloom humbly kisses her long hair.

\Bloom:
Your classic curves,
beautiful immortal,
I was glad to look on you,
to praise you,
a thing of beauty,
almost to pray.

\Nymph:
During dark nights,
I heard your praise.%
\gab{3270}

\Bloom:
\stage{(quickly)}
Yes,
yes.
You mean that I…
Sleep reveals the worst side of everyone,
children perhaps excepted.
I know I fell out of bed |
or rather was pushed.
Steel wine is said to cure snoring.
For the rest,
there is that English invention,
pamphlet of which |
I received some days ago,
incorrectly addressed.
It claims to afford a noiseless,
inoffensive vent.
\stage{(he sighs)}
'Twas ever thus.
Frailty,
thy name is marriage.

\N{3b}:
The Nymph,
her fingers in her ears.

\Nymph:
And words.
They are not in my dictionary.

\Bloom:
\gab{3280}
You understood them?

\Yews[3b]:
Ssh!

\N{3b}:
The Nymph covers her face with her hands.

\Nymph:
What have I not seen in that chamber?
What must my eyes look down on?

\Bloom:
\stage{(apologetically)}
I know.
Soiled personal linen,
wrong side up |
with care.
The quoits are loose.
From Gibraltar,
by long sea,
long ago.

\N{3b}:
The Nymph bends her head.
\gab{3290}

\Nymph:
Worse,
worse!

\N{3a}:
Bloom reflects precautiously.

\Bloom:
That antiquated commode.
It wasn't her weight.
She scaled just eleven stone nine.
She put on nine pounds after weaning.
It was a crack and want of glue.
Eh?
And that absurd orange·keyed utensil which has only one handle.

\N{3b}:
The sound of a waterfall |
is heard in bright cascade.

\Waterfall:
\begin{verse}
    Poulaphouca Poulaphouca\\
\gab{3300}%
    Poulaphouca Poulaphouca.
\end{verse}

\N{3b}:
The Yews,
mingling their boughs.

\Yews:\\*
-- Listen.\\
-- Whisper.\\
-- She is right, our sister.\\
-- We grew by Poulaphouca waterfall.\\
-- We gave shade on languorous summer days.

\N{3a}:
John Wyse Nolan in the background,
in Irish National Forester's uniform,
doffs his plumed hat.

\JohnWyse:
Prosper!
Give shade on languorous days,
trees of Ireland!

\Yews:
\stage{(murmuring)}\\*
-- Who came to Poulaphouca with the High School excursion?\\
-- Who left his nut·questing classmates to seek our shade?

\Bloom:
\gab{3310}
\stage{(scared)}
High School of Poula?
Mnemo?
Not in full possession of faculties.
Concussion.
Run over by tram.

\Echo:
Sham!

\N{3a}:
Bloom pigeon·breasted,
bottle·shouldered,
padded,
in nondescript juvenile grey and black striped suit,
too small for him,
white tennis shoes,
bordered stockings with turnover tops |
and a red school·cap with badge.

\Bloom:
I was in my teens,
a growing boy.
A little |
then sufficed,
a jolting car,
the mingling odours of the ladies' cloakroom and lavatory,
\gab{3320}%
the throng penned tight |
on the old Royal stairs
(for they love crushes,
instinct of the herd,
and the dark sex·smelling theatre unbridles vice),
even a price·list of their hosiery.
And then the heat.
There were sunspots that summer.
End of school.
And tipsy·cake.
Halcyon days.

\N{3b}:
Halcyon Days,
High School boys in blue and white football jerseys and shorts,
Master Donald Turnbull,
Master Abraham Chatterton,
Master Owen Goldberg,
Master Jack Meredith,
Master Percy Apjohn,
stand in a clearing of the trees |
and shout to Master Leopold Bloom.

\HalcyonDays:
\gab{3330}\\*
-- Mackerel!\\
-- Live us again.\\
-- Hurray!\\
\stage{(they cheer)}

\N{3a}:
Bloom hobble·de·hoy,
warm·gloved,
mamma·mufflered,
starred with spent snowballs,
struggles to rise.

\Bloom:
Again!
I feel sixteen!
What a lark!
Let's ring all the bells in Montague street.
\stage{(he cheers feebly)}
Hurray for the High School!

\Echo:
Fool!

\Yews:
\stage{(rustling)}\\*
-- She is right, our sister.\\
-- Whisper.%
\gab{3340}

\N{3b}:
Whispered kisses are heard in all the wood.
Faces of hamadryads peep out from the boles |
and among the leaves |
and break,
blossoming into bloom.

\Yews:
Who profaned our silent shade?

\N{3b}:
The Nymph coyly,
through parting fingers.

\Nymph:
There?
In the open air?

\N{3b}:
The Yews,
sweeping downward.

\Yews:\\*
-- Sister, yes.\\
-- And on our virgin sward.

\Waterfall:
\begin{verse}
    Poulaphouca Poulaphouca\\
    Phoucaphouca Phoucaphouca.
\end{verse}

\N{3b}:
The Nymph,
with wide fingers.
\gab{3350}

\Nymph:
O,
infamy!

\Bloom:
I was precocious.
Youth.
The fauna.
I sacrificed to the god of the forest.
The flowers that bloom in the spring.
It was pairing time.
Capillary attraction is a natural phenomenon.
Lotty Clarke,
flaxen·haired,
I saw at her night toilette through ill·closed curtains |
with poor papa's opera glasses:
The wanton ate grass wildly.
She rolled downhill at Rialto bridge |
to tempt me with her flow of animal spirits.
She climbed their crooked tree and I…
A saint couldn't resist it.
The demon possessed me.
Besides,
who saw?

\N{3b}:
Staggering Bob,
a white·polled calf,
\gab{3360}%
thrusts a ruminating head with humid nostrils |
through the foliage.
Staggering Bob,
large teardrops rolling from his prominent eyes,
snivels.

\StaggeringBob:
Me.
Me see.

\Bloom:
Simply satisfying a need,
I…
\stage{(with pathos)}
No girl would |
when I went girling.
Too ugly.
They wouldn't play…

\N{3b}:
High on Ben Howth,
through rhododendrons |
a nanny·goat passes,
plump·uddered,
butty·tailed,
dropping currants.

\Nannygoat:
\stage{(bleats)}
\gab{3370}
Megeggaggegg!
Nannannanny!

\N{3a}:
Bloom hatless,
flushed,
covered with burrs of thistle·down and gorse·spine.

\Bloom:
Regularly engaged.
Circumstances alter cases.

\N{3a}:
Bloom gazes intently downwards on the water.

\Bloom:
Thirty·two head over heels per second.
Press nightmare.
Giddy Elijah.
Fall from cliff.
Sad end of government printer's clerk.

\N{3a}:
Through silver·silent summer air |
the dummy of Bloom,
rolled in a mummy,
rolls roteatingly from the Lion's Head cliff |
into the purple waiting waters.

\Dummymummy[3b]:
\gab{3380}
Bbbbblllllblblblblobschbg!

\N{3b}:
Far out in the bay |
between Bailey and Kish Lights,
the \Title{Erin's King} sails,
sending a broadening plume of coal·smoke |
from her funnel towards the land.

\N{3b}:
Councillor Nannetti,
alone on deck,
in dark alpaca,
yellow·kite·faced,
his hand in his waistcoat opening,
declaims.

\CouncillorNannetti:
When my country takes her place among the nations of the earth,
then,
and not till then,
let my epitaph be written.
I have…

\Bloom:
Done.
Prff!
\stage{(farts)}%
\gab{3390}

\Nymph:
\stage{(loftily)}
We immortals,
as you saw today,
have not such a place |
and no hair there either.
We are stone·cold and pure.
We eat electric light.

\N{3b}:
The Nymph arches her body in lascivious crispation,
placing her forefinger in her mouth.

\Nymph:
Spoke to me.
Heard from behind.
How then could you…?

\N{3a}:
Bloom |
pawing the heather abjectly.

\Bloom:
O,
I have been a perfect pig.
Enemas too |
I have administered.
One third of a pint of quassia |
to which add a tablespoonful of rocksalt.
Up the fundament.
With Hamilton Long's syringe,
the ladies' friend.%
\gab{3400}

\Nymph:
In my presence.
The powder·puff.

\N{3b}:
The Nymph blushes and makes a knee.

\Nymph:
And the rest!

\N{3a}:
Bloom,
dejected.

\Bloom:
Yes.
\latin{Peccavi!}
I have paid homage on that living altar |
where the back changes name.

\N{3a}:
Bloom,
with sudden fervour.

\Bloom:
For why should the dainty scented jewelled hand,
the hand that rules…?

\N{3b}:
Figures wind serpenting in slow woodland pattern |
around the tree·stems,
coo·ee·ing.

\N{3b}:
The Voice of Kitty in the thicket.
\gab{3410}

\Kitty:
Show us one of them cushions.

\N{3a}:
The Voice of Florry.

\Florry:
Here.

\N{3b}:
A grouse wings clumsily through the under·wood.

\N{3b}:
The Voice of Lynch in the thicket.

\Lynch:
Whew!
Piping hot!

\N{3b}:
The Voice of Zoe from the thicket.

\Zoe:
Came from a hot place.

\N{3b}:
The Voice of Virag,
\gab{3420}%
a bird chief,
blue·streaked and feathered in war panoply |
with his assegai,
striding through a crackling cane brake |
over beech·mast and acorns.

\Virag:
Hot!
Hot!
Ware Sitting Bull!

\Bloom:
It overpowers me.
The warm impress of her warm form.
Even to sit where a woman has sat,
especially with divaricated thighs,
as though to grant the last favours,
most especially with previously well uplifted white sateen coat·pans.
So womanly,
full.
It fills me full.

\Waterfall[3b]:
\begin{verse}
    Phillaphulla Poulaphouca\\
\gab{3430}%
    Poulaphouca Poulaphouca.
\end{verse}

\Yews[3b]:\\*
-- Ssh!\\
-- Sister, speak!

\N{3b}:
The Nymph,
eyeless,
in nun's white habit,
coif and huge·winged wimple,
softly,
with remote eyes.

\Nymph:
Tranquilla convent.
Sister Agatha.
Mount Carmel.
The apparitions of Knock and Lourdes.
No more desire.

\N{3b}:
The Nymph reclines her head,
sighing.

\Nymph:
Only the ethereal.
Where dreamy creamy gull |
waves o'er the waters dull.

\N{3a}:
Bloom half rises.
His back trouser button snaps.

\Button:
\gab{3440}
Bip!

\N{3b}:
Two sluts of the Coombe dance rainily by,
shawled,
yelling flatly.

\Sluts:
\begin{verse}
    O, Leopold lost the pin of his drawers\\
    He didn't know what to do,\\
    To keep it up,\\
    To keep it up.
\end{verse}

\Bloom:
\stage{(coldly)}
You have broken the spell.
The last straw.
If there were only ethereal,
where would you all be,
postulants and novices?
\gab{3450}%
Shy but willing |
like an ass pissing.

\N{3b}:
The Yews,
their silver·foil of leaves precipitating,
their skinny arms aging and swaying.

\Yews:
Deciduously!

\N{3b}:
The Nymph,
her features hardening,
gropes in the folds of her habit.

\Nymph:
Sacrilege!
To attempt my virtue!

\N{3b}:
A large moist stain appears on her robe.

\Nymph:
Sully my innocence!
You are not fit to touch the garment of a pure woman.

\N{3b}:
The Nymph clutches again in her robe.

\Nymph:
Wait.
Satan,
you'll sing no more lovesongs.
Amen.
Amen.
Amen.
Amen.
\gab{3460}

\N{3b}:
She draws a poniard |
and,
clad in the sheath·mail of an elected knight of nine,
strikes at his loins.

\Nymph:
Nekum!

\N{3a}:
Bloom starts up,
seizes her hand.

\Bloom:
Hoy!
Nebrakada!
Cat o' nine lives!
Fair play,
madam.
No pruning knife.
The fox and the grapes,
is it?
What do you lack with your barbed wire?
Crucifix not thick enough?

\N{3a}:
He clutches her veil.

\Bloom:
A holy abbot you want |
or Brophy,
the lame gardener,
or the spoutless statue of the water·carrier,
or good mother Alphonsus,
eh Reynard?

\N{3b}:
The Nymph with a cry |
flees from him unveiled,
her plaster cast cracking,
a cloud of stench escaping from the cracks.
\gab{3470}

\Nymph:
Poli---!

\N{3a}:
Bloom calls after her.

\Bloom:
As if you didn't get it on the double yourselves.
No jerks and multiple mucosities all over you.
I tried it.
Your strength,
our weakness.
What's our stud·fee?
What will you pay on the nail?
You fee men·dancers on the Riviera,
I read.

\N{3b}:
The fleeing nymph raises a keen.

\Bloom:
Eh?
I have sixteen years of black slave labour behind me.
And would a jury give me five shillings alimony tomorrow,
eh?
Fool someone else,
not me.

\N{3a}:
Bloom sniffs.

\Bloom:
Rut.
Onions.
Stale.
Sulphur.
Grease.

%--------------------------------------------------
\commentary{4. \textsc{Luck and Money}}{3:32:50}
%--------------------------------------------------

\N{4b}:
The figure of Bella Cohen stands before him.

\Bella:
\gab{3480}
You'll know me the next time.

\N{4a}:
Bloom composed,
regards her.

\Bloom:
\french{Passée}.
Mutton dressed as lamb.
Long in the tooth and superfluous hair.
A raw onion |
the last thing at night |
would benefit your complexion.
And take some double chin drill.
Your eyes are as vapid |
as the glass·eyes of your stuffed fox.
They have the dimensions of your other features,
that's all.
I'm not a triple screw propeller.

\N{4b}:
Bella,
contemptuously.

\Bella:
You're not game,
in fact.

\N{4b}:
Her sow·cunt barks.

\Bella:
Fbhracht!%
\gab{3490}

\N{4a}:
Bloom,
contemptuously.

\Bloom:
Clean your nail·less middle finger first,
your bully's cold spunk is dripping from your cocks·comb.
Take a handful of hay and wipe yourself.

\Bella:
I know you,
canvasser!
Dead cod!

\Bloom:
I saw him,
kip·keeper!
Pox and gleet vendor!

\N{4b}:
Bella turns to the piano.
\gab{3500}

\Bella:
Which of you was playing the dead march from \Title{Saul}?

\Zoe[4a]:
Me.
Mind your cornflowers.

\N{4a}:
Zoe darts to the piano and bangs chords on it with crossed arms.

\Zoe:
The cat's ramble through the slag.

\N{4a}:
She glances back.

\Zoe:
Eh?
Who's making love to my sweeties?

\N{4a}:
She darts back to the table.

\Zoe:
What's yours is mine |
and what's mine is my own.

\N{4b}:
Kitty,
disconcerted,
coats her teeth with the silver paper.

\N{4a}:
Bloom approaches Zoe.

\Bloom:
\stage{(gently)}
Give me back that potato,
will you?

\Zoe:
\gab{3510}
Forfeits,
a fine thing and a super·fine thing.

\Bloom:
\stage{(with feeling)}
It is nothing,
but still,
a relic of poor mamma.

\Zoe:
\begin{verse}
    Give a thing and take it back\\
    God'll ask you where is that\\
    You'll say you don't know\\
    God'll send you down below.
\end{verse}

\Bloom:
There is a memory attached to it.
\gab{3520}%
I should like to have it.

\Stephen[4a]:
To have or not to have |
that is the question.

\Zoe:
Here.

\N{4a}:
Zoe hauls up a reef of her slip,
revealing her bare thigh,
and unrolls the potato from the top of her stocking.

\Zoe:
Those that hides,
knows where to find.

\Bella:
\stage{(frowns)}
Here.
This isn't a musical peep·show.
And don't you smash that piano.
Who's paying here?

\N{4b}:
She goes to the pianola.
\gab{3530}

\N{4a}:
Stephen fumbles in his pocket |
and,
taking out a bank·note by its corner,
hands it to her,
saying with exaggerated politeness.

\Stephen:
This silken purse |
I made out of the sow's ear of the public.
Madam,
excuse me.
If you allow me.

\N{4a}:
Stephen indicates vaguely |
Lynch and Bloom.

\Stephen:
We are all in the same sweepstake,
Kinch and Lynch.
\french{Dans ce bordel où tenons nostre état.}

\N{4b}:
Lynch calls from the hearth.

\Lynch:
Dedalus!
Give her your blessing for me.

\N{4a}:
Stephen hands Bella a coin.
\gab{3540}

\Stephen:
Gold.
She has it.

\N{4b}:
Bella looks at the money,
then at Stephen,
then at Zoe,
Florry and Kitty.

\Bella:
Do you want three girls?
It's ten shillings here.

\Stephen:
\stage{(delightedly)}
A hundred thousand apologies.

\N{4a}:
Stephen fumbles again |
and takes out |
and hands her two crowns.

\Stephen:
Permit,
\french{brevi manu},
my sight is somewhat troubled.

\N{4a}:
Bella goes to the table to count the money |
while Stephen talks to himself in monosyllables.
Zoe bends over the table.
Kitty leans over Zoe's neck.
\gab{3550}%
Lynch gets up,
rights his cap |
and,
clasping Kitty's waist,
adds his head to the group.

\N{4b}:
Florry strives heavily to rise.

\Florry:
Ow!
My foot's asleep.

\N{4b}:
Florry limps over to the table.

\N{4a}:
Bloom approaches.

\N{4b}:
Bella,
Zoe,
Kitty,
Lynch,
Bloom are chattering and squabbling.

\Zoe:
The gentleman…

\Bella:
ten shillings…

\Lynch:
paying for the three…

\Kitty:
who's touching it?

\Bloom:
allow me a moment…

\Zoe:
this gentleman pays separate…

\Kitty:
… ow!…
mind who you're pinching…

\Bella:
are you staying the night or a short time?…

\Lynch:
who did?…

\Kitty:
you're a liar,
excuse me…

\Zoe:
the gentleman paid down like a gentleman…
\gab{3560}

\Lynch:
drink…

\Bella:
it's long after eleven.

\N{4a}:
Stephen at the pianola,
making a gesture of abhorrence.

\Stephen:
No bottles!
What,
eleven?
A riddle!

\N{4b}:
Zoe,
lifting up her pettigown |
and folding a half sovereign into the top of her stocking.

\Zoe:
Hard earned |
on the flat of my back.

\N{4b}:
Lynch,
lifting Kitty from the table.

\Lynch:
Come!

\Kitty:
Wait.
\gab{3570}

\N{4a}:
Kitty clutches the two crowns.

\Florry:
And me?

\Lynch:
Hoopla!

\N{4b}:
Lynch lifts Florry,
carries her and bumps her down on the sofa.

\Stephen:
\begin{verse}
    The fox crew, the cocks flew,\\
    The bells in heaven\\
    Were striking eleven.\\
\gab{3580}%
    'Tis time for her poor soul\\
    To get out of heaven.
\end{verse}

\N{4a}:
Bloom quietly lays a half sovereign |
on the table between Bella and Florry.

\Bloom:
So.
Allow me.

\N{4a}:
Bloom takes up the pound note.

\Bloom:
Three times ten.
We're square.

\Bella:
\stage{(admiringly)}
You're such a sly·boots,
old cocky.
I could kiss you.

\Zoe:
\stage{(points)}
Him?
Deep as a draw well.

\N{4b}:
Lynch bends Kitty back over the sofa |
and kisses her.

\N{4a}:
Bloom goes with the pound note to Stephen.
\gab{3590}

\Bloom:
This is yours.

\Stephen:
How is that?
The distrait or absent·minded beggar.

\N{4b}:
Stephen fumbles again in his pocket |
and draws out a handful of coins.
An object falls.

\Stephen:
That fell.

\N{4a}:
Bloom,
stooping,
picks up and hands a box of matches.

\Bloom:
This.

\Stephen:
Lucifer.
Thanks.

\Bloom:
\gab{3600}
\stage{(quietly)}
You had better hand over that cash to me |
to take care of.
Why pay more?

\N{4b}:
Stephen hands him all his coins.

\Stephen:
Be just |
before you are generous.

\Bloom:
I will |
but is it wise?
\stage{(he counts)}
One,
seven,
eleven,
and five.
Six.
Eleven.
I don't answer for what you may have lost.

\Stephen:
Why striking eleven?
Pro·pa·rox·y·ton.
Moment before the next |
Lessing says.
Thirsty fox.
\gab{3610}
\stage{(he laughs loudly)}
Burying his grandmother.
Probably he killed her.

\Bloom:
That is one pound |
six and eleven.
One pound seven,
say.

\Stephen:
Doesn't matter a rambling damn.

\Bloom:
No,
but…

\N{4b}:
Stephen comes to the table.

\Stephen:
Cigarette,
please.

\N{4b}:
Lynch tosses a cigarette |
from the sofa to the table.
\gab{3620}

\Stephen:
And so Georgina Johnson is dead and married.

\N{4b}:
A cigarette appears on the table.
Stephen looks at it.

\Stephen:
Wonder.
Parlour magic.
Married.
Hm.

\N{4b}:
Stephen strikes a match |
and proceeds to light the cigarette |
with enigmatic melancholy.

\N{4b}:
Lynch,
watching him.

\Lynch:
You would have a better chance of lighting it,
if you held the match nearer.

\N{4b}:
Stephen brings the match near his eye.

\Stephen:
Lynx eye.
Must get glasses.
Broke them yesterday.
Sixteen years ago.
Distance.
The eye sees all flat.

\N{4b}:
Stephen draws the match away.
It goes out.
\gab{3630}

\Stephen:
Brain thinks.
Near:
far.
Ineluctable modality of the visible.

\N{4b}:
Stephen frowns mysteriously.

\Stephen:
Hm.
Sphinx.
The beast that has two backs at midnight.
Married.

\Zoe:
It was a commercial traveller married her |
and took her away with him.

\Florry:
\stage{(nods)}
Mr~Lambe from London.

\Stephen:
Lamb of London,
who takest away the sins of our world.

\N{4b}:
Lynch,
embracing Kitty on the sofa,
chants deeply.
\gab{3640}

\Lynch:
\latin{Dona nobis pacem.}

\N{4a}:
The cigarette slips from Stephen's fingers.
Bloom picks it up and throws it in the grate.

\Bloom:
Don't smoke.
You ought to eat.
Cursed dog I met.

\N4a:
Bloom, to Zoe.

\Bloom:
You have nothing?

\Zoe:
Is he hungry?

\N{4b}:
Stephen extends his hand to her |
smiling |
and chants to the air of the blood·oath |
in the \Title{Dusk of the Gods}.
\gab{3650}

\Stephen:
\begin{verse}
\german{%
    Hangende Hunger,\\
    Fragende Frau,\\
    Macht uns alle kaputt.}
\end{verse}

\Zoe:
\stage{(tragically)}
Hamlet,
I am thy father's gimlet!

\N{4a}:
She takes his hand.

\Zoe:
Blue eyes beauty,
I'll read your hand.

\N{4a}:
Zoe points to his forehead.

\Zoe:
No wit,
no wrinkles.

\N{4a}:
She counts.

\Zoe:
Two,
three,
Mars,
that's courage.

\N{4a}:
Stephen shakes his head.

\Zoe:
No kid.

\Lynch:
Sheet lightning courage.
The youth who could not shiver and shake.
\gab{3660}

\N4b:
Lynch, to Zoe.

\Lynch:
Who taught you palmistry?

\N{4a}:
Zoe turns.

\Zoe:
Ask my ballocks |
that I haven't got.

\N{4a}:
Zoe,
to Stephen.

\Zoe:
I see it in your face.
The eye,
like that.

\N{4a}:
Zoe frowns with lowered head.

\N{4b}:
Lynch,
laughing,
slaps Kitty behind twice.
\stage{(slap,
slap)}

\Lynch:
Like that.
Pandy·bat.

\N{4b}:
Twice loudly a pandy·bat cracks,
the coffin of the pianola flies open,
the bald little round Jack-in-the-box head of Father Dolan |
springs up.

\FatherDolan:
\gab{3670}
Any boy want flogging?
Broke his glasses?
Lazy idle little schemer.
See it in your eye.

\N{4a}:
Mild,
benign,
rectorial,
reproving,
the head of Don John Conmee |
rises from the pianola coffin.

\DonJohnConmee:
Now,
Father Dolan!
Now.
I'm sure that Stephen is a very good little boy!

\N{4a}:
Zoe,
examining Stephen's palm.

\Zoe:
Woman's hand.

\Stephen:
\stage{(murmurs)}
\gab{3680}
Continue.
Lie.
Hold me.
Caress.
I never could read His handwriting |
except His criminal thumbprint on the haddock.

\Zoe:
What day were you born?

\Stephen:
Thursday.
Today.

\Zoe:
Thursday's child has far to go.

\N{4a}:
Zoe traces lines on his hand.

\Zoe:
Line of fate.
Influential friends.

\Florry:
\stage{(pointing)}
\gab{3690}
Imagination.

\Zoe:
Mount of the moon.
You'll meet with a…

\N{4a}:
She peers at his hands abruptly.

\Zoe:
I won't tell you |
what's not good for you.
Or do you want to know?

\N{4a}:
Bloom detaches her fingers and offers his palm.

\Bloom:
More harm than good.
Here.
Read mine.

\Bella:
Show.

\N{4b}:
Bella turns up Bloom's hand.

\Bella:
I thought so.
Knobby knuckles for the women.

\N{4a}:
Zoe,
peering at Bloom's palm.
\gab{3700}

\Zoe:
Gridiron.
Travels beyond the sea and marry money.

\Bloom:
Wrong.

\Zoe:
\stage{(quickly)}
O,
I see.
Short little finger.
Henpecked husband.
That wrong?

\N{4b}:
Black Liz,
a huge rooster hatching in a chalked circle,
rises,
stretches her wings and clucks.

\BlackLiz:
Gara.
Klook.
Klook.
Klook.
\gab{3710}

\N{4b}:
She sidles from her newlaid egg and waddles off.

\N{4a}:
Bloom points to his hand.

\Bloom:
That weal there is an accident.
Fell and cut it |
twenty·two years ago.
I was sixteen.

\Zoe:
I see,
says the blind man.
Tell us news.

\Stephen:
See?
Moves to one great goal.
I am twenty·two.
Sixteen years ago |
he was twenty·two too.
Sixteen years ago |
I twenty·two tumbled.
Twenty·two years ago |
he sixteen fell off his hobby·horse.
\gab{3720}
\stage{(he winces)}
Hurt my hand somewhere.
Must see a dentist.
Money?

\N{4a}:
Zoe whispers to Florry.
They giggle.
Bloom releases his hand |
and writes idly on the table in backhand,
pencilling slow curves.

\Florry:
What?

%--------------------------------------------------
\commentary{5. \textsc{Voyeur} --- Cuckold Fantasy, Shakespeare}{3:43:43}
%--------------------------------------------------

\N{5b}:
A hackney car,
number Three Hundred and Twenty·four,
with a gallant·butt\-ocked mare,
driven by James Barton,
Harmony Avenue,
Donnybrook,
trots past.
Blazes Boylan and Lenehan sprawl |
swaying on the side·seats.
The Ormond boots crouches behind on the axle.
\gab{3730}%
Sadly over the cross·blind,
Lydia Douce and Mina Kennedy gaze.
The Boots,
jogging,
mocks them with thumb and wriggling worm·fingers.

\Boots:
Haw,
haw,
have you the horn?

\N{5b}:
Bronze by gold,
they whisper.

\N{5a}:
Zoe to Florry.

\Zoe:
Whisper.

\N{5a}:
They whisper again.

\N{5b}:
Over the well of the car,
Blazes Boylan leans,
his boater straw set sideways,
a red flower in his mouth.
Lenehan,
in yachts·man's cap and white shoes,
\gab{3740}%
officiously detaches a long hair |
from Blazes Boylan's coat shoulder.

\Lenehan:
Ho!
What do I here behold?
Were you brushing the cobwebs off a few quims?

\Boylan:
\stage{(seated, smiles)}
Plucking a turkey.

\Lenehan:
A good night's work.

\N{5b}:
Boylan,
holding up four thick blunt·ungulated fingers,
winks.
\gab{3750}

\Boylan:
Blazes Kate!
Up to sample or your money back.

\N{5b}:
Boylan holds out a forefinger.

\Boylan:
Smell that.

\N{5b}:
Lenehan smells gleefully.

\Lenehan:
Ah!
Lobster and mayonnaise.
Ah!

\ZoeAndFlorry:
\stage{(laugh together)}
Ha ha ha ha.

\N{5b}:
Boylan jumps surely from the car |
and calls loudly for all to hear.

\Boylan:
Hello,
Bloom!
Mrs Bloom dressed yet?

\N{5a}:
Bloom,
in flunkey's prune plush coat and knee·breeches,
\gab{3760}%
buff stockings and powdered wig.

\Bloom:
I'm afraid not,
sir.
The last articles…

\N{5b}:
Boylan tosses him six·pence.

\Boylan:
Here,
to buy yourself a gin and splash.

\N{5b}:
Boylan hangs his hat smartly on a peg of Bloom's antlered head.

\Boylan:
Show me in.
I have a little private business with your wife,
you understand?

\Bloom:
Thank you,
sir.
Yes,
sir.
Madam Tweedy is in her bath,
sir.

\Marion[5a]:
He ought to feel himself highly honoured.

\N{5a}:
Marion plops splashing out of the water.

\Marion:
Raoul darling,
come and dry me.
I'm in my pelt.
\gab{3770}%
Only my new hat and a carriage sponge.

\N{5b}:
Boylan,
a merry twinkle in his eye.

\Boylan:
Topping!

\Bella:
What?
What is it?

\N{5a}:
Zoe whispers to her.

\Marion:
Let him look,
the pishogue!
Pimp!
And scourge himself!
I'll write to a powerful prostitute |
or Bartholomona,
the bearded woman,
to raise weals out on him |
an inch thick |
\gab{3780}%
and make him bring me back |
a signed and stamped receipt.

\N{5b}:
Boylan clasps himself.

\Boylan:
Here,
I can't hold this little lot much longer.

\N{5b}:
Boylan strides off on stiff cavalry legs.

\N{5b}:
Bella laughs.

\Bella:
Ho ho ho ho.

\N{5b}:
Boylan to Bloom,
over his shoulder.

\Boylan:
You can apply your eye to the keyhole |
and play with yourself |
while I just go through her a few times.

\Bloom:
\gab{3790}
Thank you,
sir.
I will,
sir.
May I bring two men chums |
to witness the deed and take a snapshot?

\N{5a}:
Bloom holds out an ointment jar.

\Bloom:
Vaseline,
sir?
Orange·flower…?
Lukewarm water…?

\N{5b}:
Kitty,
from the sofa.

\Kitty:
Tell us,
Florry.
Tell us.
What.

\N{5a}:
Florry whispers to her.
Whispering love·words murmur,
lip·lapping loudly,
pop·pys·mic plop·slop.

\N{5b}:
Mina Kennedy,
her eyes upturned.

\MinaKennedy:
O,
it must be like the scent of geraniums and lovely peaches!
\gab{3800}%
O,
he simply idolises every bit of her!
Stuck together!
Covered with kisses!

\N{5a}:
Lydia Douce,
her mouth opening.

\LydiaDouce:
Yum·yum.
O,
he's carrying her round the room doing it!
Ride a cock·horse.
You could hear them in Paris and New York.
Like mouthfuls of strawberries and cream.

\Kitty:
\stage{(laughing)}
Hee hee hee.

\BoylansVoice:
\stage{(sweetly,
hoarsely,
in the pit of his stomach)}
Ah!
God·blaze·gruk·bruk·arch·khrasht!%
\gab{3810}

\MarionsVoice:
\stage{(hoarsely,
sweetly,
rising to her throat)}
O!
Weesh·washt·kiss·in·a·poo·isth·na·poo·huck?

\N{5a}:
Bloom,
his eyes wildly dilated,
clasps himself.

\Bloom:
Show!
Hide!
Show!
Plough her!
More!
Shoot!

\BellaZoeFlorryKitty:
Ho ho!
Ha ha!
Hee hee!

\Lynch:
\stage{(points)}
The mirror up to nature.
\stage{(he laughs)}
\gab{3820}%
Hu hu hu hu hu!

\N{5a}:
Stephen and Bloom gaze in the mirror.
The face of William Shakespeare,
beardless,
appears there,
rigid in facial paralysis,
crowned by the reflection of the reindeer antlered hatrack in the hall.

\N{5b}:
Shakespeare,
in dignified ventriloquy.

\Shakespeare:
'Tis the loud laugh |
bespeaks the vacant mind.

\N{5b}:
To Bloom.

\Shakespeare:
Thou thoughtest as how thou wastest invisible.
Gaze.

\N{5b}:
He crows with a black capon's laugh.

\Shakespeare:
Iago·go!
How my Oldfellow chokit his Thursday mornun.
Iago·go·go!

\N{5b}:
Bloom smiles yellowly at the three whores.
\gab{3830}

\Bloom:
When will I hear the joke?

\Zoe:
Before you're twice married and once a widower.

\Bloom:
Lapses are condoned.
Even the great Napoleon |
when measurements were taken next the skin |
after his death…

\N{5b}:
Mrs~Dignam,
widow woman,
her snub·nose and cheeks flushed with death·talk,
tears and Tunney's tawny sherry,
hurries by in her weeds,
her bonnet awry,
rouging and powdering her cheeks,
lips and nose,
a pen chivvying her brood of cygnets.
\gab{3840}%
Beneath her skirt |
appear her late husband's everyday trousers and turned·up boots,
large eights.
She holds a Scottish Widows' insurance policy |
and a large marquee umbrella |
under which |
her brood run with her,
Patsy hopping on one shod foot,
his collar loose,
a hank of pork·steaks dangling,
Freddy whimpering,
Susy with a crying cod's mouth,
Alice struggling with the baby.
She cuffs them on,
her streamers flaunting aloft.

\Freddy[5a]:
Ah,
ma,
you're dragging me along!

\Susy[5b]:
\gab{3850}
Mamma,
the beef·tea is fizzing over!

\N{5b}:
Shakespeare,
with paralytic rage.

\Shakespeare:
Weda seca who·killa farst.

\N{5b}:
The face of Martin Cunningham,
bearded,
refeatures Shakespeare's beardless face.
The marquee umbrella sways drunkenly,
the children run aside.
Under the umbrella |
appears Mrs~Cunningham |
in merry widow hat and kimono gown.
She glides sidling and bowing,
twirling japanesily.

\MrsCunningham:
\stage{(sings)}
\gab{3860}
\begin{verse}
    And they call me the jewel of Asia!
\end{verse}

\N{5b}:
Martin Cunningham gazes on her,
impassive.

\MartinCunningham:
Immense!
Most bloody awful demi·rep!

%--------------------------------------------------
\commentary{6. \textsc{Dance of Death}}{3:50:26}
%--------------------------------------------------

\Stephen:
\latin{Et exalt·a·bun·tur cornua iusti.}
Queens lay with prize bulls.
Remember Pasiphae |
for whose lust |
my grand·old·gross·father |
made the first confession box.
Forget not Madam Grissel Steevens |
nor the suine scions of the house of Lambert.
And Noah was drunk with wine.
And his ark was open.

\Bella:
\gab{3870}
None of that here.
Come to the wrong shop.

\Lynch:
Let him alone.
He's back from Paris.

\N{6a}:
Zoe runs to Stephen and links him.

\Zoe:
O~go on!
Give us some parley·voo.

\N{6a}:
Stephen claps hat on head |
and leaps over to the fireplace |
where he stands with shrugged shoulders,
finny hands outspread,
a painted smile on his face.

\N{6b}:
Lynch,
pommelling on the sofa.
\gab{3880}

\Lynch:
Rmm Rmm Rmm Rrrrrrmmmm.

\N{6a}:
Stephen gabbles with marionette jerks.

\Stephen:
Thousand places of entertainment to expense your evenings |
with lovely ladies saling gloves and other things |
perhaps hers heart beer·chops perfect fashionable house |
very eccentric |
where lots cocottes |
beautiful dressed |
much about princesses like |
are dancing cancan |
and walking there parisian clowneries |
extra foolish for bachelors foreigns |
the same if talking a poor english |
how much smart they are |
on things love |
and sensations voluptuous.
Misters very selects for is pleasure |
must to visit heaven and hell |
show with mortuary candles |
and they tears silver |
which occur every night.
\gab{3890}%
Perfectly shocking |
terrific of religion's things |
mockery seen in universal world.
All chic womans |
which arrive full of modesty |
then disrobe and squeal loud |
to see vampire man debauch nun |
very fresh young |
with \french{dessous troublants}.

\N{6a}:
Stephen clacks his tongue loudly.

\Stephen:
\french{Ho,
là là!
Ce pif qu'il a!}

\Lynch:
\french{Vive le vampire!}

\Whores:
Bravo!
Parley·voo!

\N{6a}:
Stephen,
grimacing with head back,
laughs loudly,
clapping himself.
\gab{3900}

\Stephen:
Great success of laughing.
Angels much prostitutes like |
and holy apostles |
big damn ruffians.
\french{Demi·mondaines} nicely handsome |
sparkling of diamonds |
very amiable costumed.
Or do you |
are fond better |
what belongs they |
moderns pleasure |
turpitude of old mans?

\N{6a}:
Stephen points about him |
with grotesque gestures |
which Lynch and the whores reply to.

\Stephen:
Caou·tchouc statue woman |
reversible or lifesize tompe·ep·tom of virgins nudities |
very lesbic |
the kiss |
five ten times.
Enter,
gentleman,
to see in mirror |
every positions trapezes |
all that machine there besides |
also if desire |
act awfully bestial |
butcher's boy pollutes |
in warm veal |
liver or omlet on the belly |
\french{pièce de Shakespeare}.

\N{6b}:
Bella,
clapping her belly |
sinks back on the sofa,
with a shout of laughter.
\gab{3910}

\Bella:
An omelette on the…
Ho!
ho!
ho!
ho!…
omelette on the…

\Stephen:
\stage{(mincingly)}
I love you,
sir darling.
Speak you englishman tongue for \french{double entente cordiale}.
O~yes,
\french{mon loup}.
How much cost?
Water·loo.
Water·closet.

\N{6a}:
Stephen ceases suddenly and holds up a forefinger.

\Bella:
\stage{(laughing)}
Omelette…

\Whores:
\stage{(laughing)}
\gab{3920}
Encore!
Encore!

\Stephen:
Mark me.
I dreamt of a water·melon.

\Zoe:
Go abroad and love a foreign lady.

\Lynch:
Across the world for a wife.

\Florry:
Dreams goes by contraries.

\N{6a}:
Stephen extends his arms.
\gab{3930}

\Stephen:
It was here.
Street of harlots.
In Serpentine avenue,
Beelzebub showed me her,
a fubsy widow.
Where's the red carpet spread?

\N{6a}:
Bloom,
approaching Stephen.

\Bloom:
Look…

\Stephen:
No,
I flew.
My foes beneath me.
And ever shall be.
World without end.
\stage{(he cries)}
\latin{Pater!}
Free!

\Bloom:
I say,
look…

\Stephen:
Break my spirit,
will he?
\gab{3940}%
\french{O~merde alors!}

\N{6a}:
Stephen cries,
his vulture talons sharpened.

\Stephen:
Break my spirit,
will he?
\spanish{Holà!}
Hilly·ho!

\N{6b}:
Simon Dedalus' voice hilloes in answer,
somewhat sleepy but ready.

\Simon:
That's all right.

\N{6b}:
Simon swoops uncertainly through the air,
wheeling,
uttering cries of heartening,
on strong ponderous buzzard wings.

\Simon:
Ho,
boy!
Are you going to win?
Hoop!
Pschatt!
Stable with those half·castes.
Wouldn't let them |
within the bawl of an ass.
Head up!
Keep our flag flying!
An eagle gules volant |
in a field argent displayed.
Ulster king at arms!
Haihoop!

\N{6b}:
Simon makes the beagle's call,
giving tongue.
\gab{3950}

\Simon:
Bulbul!
Burblblburblbl!
Hai,
boy!

\N{6a}:
The fronds and spaces of the wallpaper |
file rapidly across country.
A stout fox,
drawn from covert,
brush pointed,
having buried his grandmother,
runs swift for the open,
bright·eyed,
seeking badger earth,
under the leaves.
The pack of stag·hounds follows,
nose to the ground,
sniffing their quarry,
beagle·baying,
burblbrbling to be blooded.
Ward Union huntsmen and huntswomen live with them,
hot for a kill.
From Six Mile Point,
Flathouse,
Nine Mile Stone,
follow the foot·people with knotty sticks,
hay·forks,
salmon·gaffs,
lassos,
flock·masters with stock·whips,
bear·baiters with tomtoms,
toreadors with bull·swords,
grey·negroes waving torches.
\gab{3960}

\N{6b}:
The crowd bawls of dicers,
crown and anchor players,
thimble·riggers,
broads·men.
Crows and touts,
hoarse bookies in high wizard hats |
clamour deafeningly.

\Crowd:\\*
-- Card of the races. Racing card!\\
-- Ten to one the field!\\
-- Tommy on the clay here! Tommy on the clay!\\
-- Ten to one bar one! Ten to one bar one!\\
-- Try your luck on Spinning Jenny!\\
-- Ten to one bar one!%
\gab{3970}\\
-- Sell the monkey, boys! Sell the monkey!\\
-- I'll give ten to one!\\
-- Ten to one bar one!

\N{6a}:
A dark horse,
riderless,
bolts like a phantom past the winning post,
his mane moon·foaming,
his eyeballs stars.
The field follows,
a bunch of bucking mounts.
Skeleton horses,
Sceptre,
Maximum the Second,
Zinfandel,
the duke of Westminster's Shot·over,
Repulse,
the duke of Beaufort's Ceylon,
prix de Paris.
Dwarfs ride them,
rusty·armoured,
leaping,
leaping in their,
in their saddles.
Last in a drizzle of rain |
on a broken·winded isabelle nag,
\gab{3980}%
Cock of the North,
the favourite,
honey cap,
green jacket,
orange sleeves,
Garrett Deasy up,
gripping the reins,
a hockey stick at the ready.
His nag on spavined white·gaitered feet |
jogs along the rocky road.

\OrangeLodges[6b]:
\stage{(jeering)}
Get down and push,
mister.
Last lap!
You'll be home the night!

\N{6a}:
Garrett Deasy,
bolt upright,
his nail·scraped face plastered with postage·stamps,
brandishes his hockey·stick,
his blue eyes flashing in the prism of the chandelier |
as his mount lopes by at schooling gallop.

\GarrettDeasy:
\latin{Per vias rectas!}

\N{6a}:
A yoke of buckets |
leopards all over him and his rearing nag |
\gab{3990}%
a torrent of mutton broth |
with dancing coins |
of carrots,
barley,
onions,
turnips,
potatoes.

\GreenLodges[6b]:
Soft day,
sir John!
Soft day,
your honour!

\N{6a}:
Private Carr,
Private Compton |
and Cissy Caffrey |
pass beneath the windows,
singing in discord.

\Stephen:
Hark!
Our friend noise in the street.

\N{6a}:
Zoe holds up her hand.
\gab{4000}

\Zoe:
Stop!

\CarrComptonCissy[6b]:
\begin{verse}
    Yet I've a sort a\\
    Yorkshire relish for…
\end{verse}

\Zoe:
That's me.
\stage{(she claps her hands)}
Dance!
Dance!

\N{6a}:
Zoe runs to the pianola.

\Zoe:
Who has two·pence?

\Bloom:
Who'll…?

\N{6b}:
Lynch,
handing her coins.
\gab{4010}

\Lynch:
Here.

\N{6a}:
Stephen,
cracking his fingers impatiently.

\Stephen:
Quick!
Quick!
Where's my augur's rod?

\N{6a}:
Stephen runs to the piano and takes his ash·plant,
beating his foot in tri·pu·di·um.

\N{6a}:
Zoe turns the drum·handle.

\Zoe:
There.

\N{6a}:
She drops two pennies in the slot.
Gold,
pink and violet lights start forth.
The drum turns purring |
in low hesitation waltz.

\N{6b}:
Professor Goodwin,
in a bow·knotted periwig,
in court dress,
wearing a stained Inverness cape,
bent in two from incredible age,
totters across the room,
his hands fluttering.
\gab{4020}%
He sits tinily on the piano·stool |
and lifts |
and beats handless sticks of arms |
on the keyboard,
nodding with damsel's grace,
his bow·knot bobbing.

\N{6a}:
Zoe twirls round herself,
heel·tapping.

\Zoe:
Dance.
Anybody here for there?
Who'll dance?
Clear the table.

\N{6a}:
The pianola with changing lights |
plays in waltz time |
the prelude of \Title{My Girl's a Yorkshire Girl}.
Stephen throws his ash·plant on the table |
and seizes Zoe round the waist.
Florry and Bella push the table towards the fireplace.
Stephen,
arming Zoe with exaggerated grace,
begins to waltz her round the room.
Bloom stands aside.
\gab{4030}%
Her sleeve falling from gracing arms |
reveals a white flesh·flower of vaccination.

\N{6b}:
Between the curtains,
Professor Maginni inserts a leg |
on the toe·point of which |
spins a silk hat.
With a deft kick,
he sends it spinning to his crown |
and jaunty·hatted |
skates in.
He wears a slate frock·coat with claret silk lapels,
a gorget of cream tulle,
a green low·cut waistcoat,
stock collar with white kerchief,
tight lavender trousers,
patent pumps |
and canary gloves.
In his button·hole |
is an immense dahlia.
He twirls in reversed directions |
a clouded cane,
then wedges it tight in his oxter.
He places a hand lightly on his breast·bone,
\gab{4040}%
bows,
and fondles his flower and buttons.

\Maginni[6b]:
The poetry of motion,
art of calisthenics.
No connection with Madam Legget Byrne's or Levenston's.
Fancy dress balls arranged.
Deportment.
The Katty Lanner step.
So.
Watch me!
My terpsichorean abilities.

\N{6b}:
Professor Maginni minuets forward three paces |
on tripping bee's feet.

\Maginni:
\french{Tout le monde en avant!
Révérence!
Tout le monde en place!}

\N{6b}:
The prelude ceases.
Professor Goodwin,
beating vague arms |
shrivels,
sinks,
his live cape filling about the stool.
The air in firmer waltz time |
sounds.

\N{6a}:
Stephen and Zoe circle freely.
The lights change,
\gab{4050}%
glow,
fade |
gold rosy violet.

\Pianola[6b]:
\begin{verse}
    Two young fellows were talking about their girls, girls, girls,\\
    Sweethearts they'd left behind…
\end{verse}

\N{6a}:
From a corner,
the morning hours run out,
gold·haired,
slim·sandalled,
in girlish blue,
wasp·waisted,
with innocent hands.
Nimbly they dance,
twirling their skipping ropes.
The hours of noon follow in amber gold.
Laughing,
linked,
high hair·combs flashing,
they catch the sun in mocking mirrors,
lifting their arms.

\N{6b}:
Maginni clip·claps glove·silent hands.
\gab{4060}

\Maginni:
\french{Carré!
Avant deux!}
Breathe evenly!
\french{Balancé!}

\N{6a}:
The morning and noon hours waltz in their places,
turning,
advancing to each other,
shaping their curves,
bowing vis·a·vis.

\N{6b}:
Cavaliers behind them |
arch and suspend their arms,
with hands descending to,
touching,
rising from their shoulders.

\Hours:
You may touch my.

\Cavaliers:
May I touch your?

\Hours:
\gab{4070}
O,
but lightly!

\Cavaliers:
O,
so lightly!

\Pianola:
\begin{verse}
    My little shy little lass has a waist.
\end{verse}

\N{6a}:
Zoe and Stephen turn boldly |
with looser swing.
The twilight hours advance from long land·shadows,
dispersed,
lagging,
languid·eyed,
their cheeks delicate with cipria and false faint bloom.
They are in grey gauze |
with dark bat sleeves |
that flutter in the land breeze.

\Maginni:
\french{Avant huit!
\gab{4080}
Traversé!
Salut!
Cours de mains!
Croisé!}

\N{6a}:
The night hours,
one by one,
steal to the last place.
Morning,
noon and twilight hours retreat before them.
They are masked,
with daggered hair and bracelets of dull bells.
Weary |
they curchy·curchy under veils.

\Bracelets:
Heigho!
Heigho!

\N{6a}:
Zoe,
twirling,
her hand to her brow.

\Zoe:
O!

\Maginni:
\french{Les tiroirs!
\gab{4090}%
Chaîne de dames!
La corbeille!
Dos à dos!}

\N{6a}:
Arabesquing wearily |
they weave a pattern on the floor,
weaving,
unweaving,
curtseying,
twirling,
simply swirling.

\Zoe:
I'm giddy!

\N{6a}:
Zoe frees herself,
droops on a chair.
Stephen seizes Florry and turns with her.

\Maginni:
\french{Boulangère!
Les ronds!
Les ponts!
Chevaux de bois!
Escargots!}

\N{6a}:
Twining,
receding,
with interchanging hands,
the night hours link each |
each with arching arms |
in a mosaic of movements.
\gab{4100}%
Stephen and Florry turn cumbrously.

\Maginni:
\french{Dansez avec vos dames!
Changez de dames!
Donnez le petit bouquet à votre dame!
Remerciez!}

\Pianola[6b]:
\begin{verse}
    Best, best of all,
    Baraabum!
\end{verse}

\N{6a}:
Kitty jumps up.

\Kitty:
O,
they played that on the hobby·horses at the \Title{Mirus} bazaar!

\N{6a}:
Kitty runs to Stephen.
\gab{4110}%
He leaves Florry brusquely and seizes Kitty.
A screaming bittern's harsh high whistle |
shrieks.
Groan·grouse·gurgling |
Toft's cumbersome whirligig turns slowly the room |
right roundabout the room.

\Pianola:
\begin{verse}
    My girl's a Yorkshire girl.
\end{verse}

\Zoe:
Yorkshire through and through.
Come on all!

\N{6a}:
Zoe seizes Florry and waltzes her.

\Stephen:
\french{Pas seul!}
\gab{4120}

\N{6a}:
Stephen wheels Kitty into Lynch's arms,
snatches up his ash·plant from the table |
and takes the floor.

\N{6b}:
All wheel |
whirl |
waltz |
twirl |
Bloom·bella |
Kitty·lynch |
Florry·zoe |
jujuby women.

\N{6a}:
Stephen with hat ash·plant |
frog splits in middle high kicks |
with sky·kicking mouth |
shut hand clasp part under thigh.
With clang tinkle boom·hammer tally·ho |
Hornblower blue green yellow |
flashes Toft's cumbersome turns |
with hobby·horse riders from gilded snakes dangled,
bowels fandango leaping |
spurn soil foot and fall again.

\Pianola:
\begin{verse}
\gab{4130}%
    Though she's a factory lass\\
    And wears no fancy clothes.
\end{verse}

\N{6a}:
Close·clutched swift swifter with glare·blare·flare scudding |
they scoot·loot·shoot lumbering by.
Baraabum!

\N{6b}:
Tutti:

\All:
Encore!
Bis!
Bravo!
Encore!

\Simon[6b]:
Think of your mother's people!

\Stephen:
Dance of death.

\N{6a}:
Bang fresh barang bang of lacquey's bell,
\gab{4140}%
horse,
nag,
steer,
piglings,
Conmee on Christ ass,
lame crutch and leg sailor in cock·boat |
arm·folded rope·pulling hitching stamp |
hornpipe through and through.
Baraabum!

\N{6b}:
On nags hogs bell·horses |
Gadarene swine |
Corny in coffin steel |
shark stone |
one·handled Nelson |
two trickies |
Frauen·zimmer |
plum·stained from pram |
falling bawling.
Gum he's a champion.
Fuse·blue peer from barrel |
Rev.~Evensong Love on hackney jaunt |
Blazes blind |
cod·doubled bicyclers |
Dilly with snowcake |
no fancy clothes.
Then in last switchback |
lumbering up and down |
bump mashtub |
sort of viceroy |
and reine |
relish for tub·lumber |
bump·shire rose.
\gab{4150}%
Baraabum!

\N{6a}:
The couples fall aside.
Stephen whirls giddily.
Room whirls back.
Eyes closed |
he totters.
Red rails fly space·wards.
Stars all around,
suns turn roundabout.
Bright midges dance on walls.
He stops dead.

\Stephen:
Ho!

\N{6b}:
Stephen's mother,
emaciated,
rises stark through the floor,
in leper grey |
with a wreath of faded orange·blossoms |
and a torn bridal veil,
her face worn and nose·less,
green with grave·mould.
Her hair is scant and lank.
\gab{4160}%
She fixes her blue·circled hollow eye·sockets on Stephen |
and opens her toothless mouth,
uttering a silent word.
A choir of virgins and confessors sing voicelessly.

\Choir:
\begin{verse}
\latin{%
    Liliata rutilantium te confessorum…\\
    Iubilantium te virginum…}
\end{verse}

\N{6a}:
From the top of a tower,
Buck Mulligan,
in particoloured jester's dress of puce and yellow |
and clown's cap with curling bell,
stands gaping at her,
a smoking buttered split scone in his hand.

\BuckMulligan:
She's beastly dead.
The pity of it!
\gab{4170}%
Mulligan meets the afflicted mother.

\N{6a}:
Mulligan upturns his eyes.

\BuckMulligan:
Mercurial Malachi!

\N{6b}:
Stephen's mother with the subtle smile of death's madness.

\Mother:
I was once the beautiful May Goulding.
I am dead.

\Stephen:
\stage{(horrorstruck)}
Lemur,
who are you?
No.
What bogeyman's trick is this?

\N{6a}:
Buck Mulligan shakes his curling capbell.

\BuckMulligan:
The mockery of it!
Kinch dogs·body killed her bitch·body.
She kicked the bucket.

\N{6a}:
Tears of molten butter |
fall from his eyes |
on to the scone.

\BuckMulligan:
Our great sweet mother!
\gab{4180}%
\greek{Epi oinopa ponton.}

\N{6b}:
Stephen's mother comes nearer,
breathing upon him softly |
her breath of wetted ashes.

\Mother:
All must go through it,
Stephen.
More women than men in the world.
You too.
Time will come.

\N{6a}:
Stephen choking with fright,
remorse and horror.

\Stephen:
They say I killed you,
mother.
He offended your memory.
Cancer did it,
not I\@.
Destiny.

\N{6b}:
Stephen's mother,
a green rill of bile |
trickling from a side of her mouth.

\Mother:
You sang that song to me.
\gab{4190}%
\Title{Love's bitter mystery.}

\Stephen:
\stage{(eagerly)}
Tell me the word,
mother,
if you know now.
The word known to all men.

\Mother:
Who saved you |
the night you jumped into the train at Dalkey |
with Paddy Lee?
Who had pity for you |
when you were sad among the strangers?
Prayer is all·powerful.
Prayer for the suffering souls in the Ursuline manual |
and forty days' indulgence.
Repent,
Stephen.

\Stephen:
The ghoul!
\gab{4200}%
Hyena!

\Mother:
I pray for you in my other world.
Get Dilly to make you that boiled rice |
every night after your brainwork.
Years and years |
I loved you,
O,
my son,
my first·born,
when you lay in my womb.

\N{6a}:
Zoe,
fanning herself with the grate fan.

\Zoe:
I'm melting!

\N{6b}:
Florry points to Stephen.

\Florry:
Look!
He's white.

\N{6a}:
Bloom goes to the window |
to open it more.
\gab{4210}

\Bloom:
Giddy.

\N{6b}:
Stephen's mother,
with smouldering eyes.

\Mother:
Repent!
O,
the fire of hell!

\Stephen:
\stage{(panting)}
His noncorrosive sublimate!
The corpse chewer!
Raw head and bloody bones.

\N{6b}:
Stephen's mother,
her face drawing near and nearer,
sending out an ashen breath.

\Mother:
Beware!

\N{6b}:
Stephen's mother raises her blackened withered right arm |
slowly towards Stephen's breast |
with outstretched finger.

\Mother:
Beware God's hand!

\N{6b}:
A green crab with malignant red eyes |
\gab{4220}%
sticks deep its grinning claws |
in Stephen's heart.

\N{6a}:
Stephen,
strangled with rage,
his features drawn grey and old.

\Stephen:
Shite!

\N{6a}:
Bloom,
at the window.

\Bloom:
What?

\Stephen:
\french{Ah non,
par exemple!}
The intellectual imagination!
With me |
all or not at all.
\latin{Non serviam!}

\Florry:
Give him some cold water.
Wait.
\gab{4230}

\N{6b}:
Florry rushes out.

\N{6b}:
Stephen's mother wrings her hands slowly,
moaning desperately.

\Mother:
O~Sacred Heart of Jesus,
have mercy on him!
Save him from hell,
O~Divine Sacred Heart!

\Stephen:
No!
No!
No!
Break my spirit,
all of you,
if you can!
I'll bring you all to heel!

\N{6b}:
Stephen's mother,
in the agony of her death·rattle.

\Mother:
Have mercy on Stephen,
Lord,
for my sake!
Inexpressible was my anguish |
when expiring with love,
grief and agony
\gab{4240}%
on Mount Calvary.

\Stephen:
\german{Nothung!}

\N{6a}:
Stephen lifts his ash·plant high with both hands |
and smashes the chandelier.
Time's livid final flame leaps |
and,
in the following darkness,
ruin of all space,
shattered glass and toppling masonry.

\Gasjet[6b]:
Pwfungg!

\Bloom:
Stop!

\N{6b}:
Lynch rushes forward and seizes Stephen's hand.
\gab{4250}

\Lynch:
Here!
Hold on!
Don't run amok!

\Bella:
Police!

\N{6b}:
Stephen,
abandoning his ash·plant,
his head and arms thrown back stark,
beats the ground |
and flies from the room,
past the whores at the door.

\Bella:
\stage{(screams)}
After him!

\N{6b}:
The two whores rush to the hall·door.
\gab{4260}%
Lynch and Kitty and Zoe stampede from the room.
They talk excitedly.

\N{6a}:
Bloom follows,
returns.

\N{6b}:
The whores,
jammed in the door·way,
pointing.

\Whores:
Down there.

\N{6a}:
Zoe,
pointing.

\Zoe:
There.
There's something up.

\Bella:
Who pays for the lamp?

\N{6b}:
Bella seizes Bloom's coat·tail.

\Bella:
Here,
you were with him.
The lamp's broken.

\N{6a}:
Bloom rushes to the hall,
rushes back.
\gab{4270}

\Bloom:
What lamp,
woman?

\Kitty:
He tore his coat.

\N{6b}:
Bella,
her eyes hard with anger and cupidity,
points.

\Bella:
Who's to pay for that?
Ten shillings.
You're a witness.

\N{6a}:
Bloom snatches up Stephen's ash·plant.

\Bloom:
Me?
Ten shillings?
Haven't you lifted enough off him?
Didn't he…?

\Bella:
\gab{4280}
\stage{(loudly)}
Here,
none of your tall talk.
This isn't a brothel.
A ten·shilling house.

\N{6a}:
Bloom,
his head under the lamp,
pulls the chain.
Puling,
the gasjet lights up a crushed mauve purple shade.
He raises the ash·plant.

\Bloom:
Only the chimney's broken.
Here is all he…

\N{6b}:
Bella shrinks back and screams.

\Bella:
Jesus!
Don't!

\N{6a}:
Bloom,
warding off a blow.

\Bloom:
To show you how he hit the paper.
\gab{4290}%
There's not six·pence·worth of damage done.
Ten shillings!

\N{6b}:
Florry,
with a glass of water,
enters.

\Florry:
Where is he?

\Bella:
Do you want me to call the police?

\Bloom:
O,
I know.
Bulldog on the premises.
But he's a Trinity student.
Patrons of your establishment.
Gentlemen that pay the rent.

\N{6a}:
Bloom makes a masonic sign.

\Bloom:
Know what I mean?
Nephew of the vice-chancellor.
You don't want a scandal.%
\gab{4300}

\Bella:
\stage{(angrily)}
Trinity.
Coming down here ragging |
after the boat·races |
and paying nothing.
Are you my commander here |
or?
Where is he?
I'll charge him!
Disgrace him,
I will!
\stage{(she shouts)}
Zoe!
Zoe!

\Bloom:
\stage{(urgently)}
And if it were your own son in Oxford?
\stage{(warningly)}
I know.

\Bella:
\stage{(almost speechless)}
Who are.
Incog!

\N{6a}:
Zoe,
in the doorway.
\gab{4310}

\Zoe:
There's a row on.

\Bloom:
What?
Where?

\N{6a}:
Bloom throws a shilling on the table |
and starts.

\Bloom:
That's for the chimney.
Where?
I need mountain air.

\N{6a}:
Bloom hurries out through the hall.
The whores point.
Florry follows,
spilling water from her tilted tumbler.
On the door·step,
all the whores clustered |
talk volubly,
pointing to the right |
where the fog has cleared off.
From the left,
arrives a jingling hackney car.
It slows to |
in front of the house.
Bloom at the hall·door |
perceives Corny Kelleher |
who is about to dismount from the car |
with two silent lechers.
\gab{4320}%
He averts his face.

\N{6b}:
Bella from within the hall |
urges on her whores.
They blow icky·licky·sticky yum·yum kisses.
Corny Kelleher replies with a ghastly lewd smile.
The silent lechers turn to pay the jarvey.
Zoe and Kitty still point right.

\N{6a}:
Bloom,
parting them swiftly,
draws his caliph's hood and poncho |
and hurries down the steps |
with sideways face.
Incog Haroun al Raschid |
he flits behind the silent lechers |
and hastens on by the railings |
with fleet step of a pard |
strewing the drag behind him,
torn envelopes drenched in aniseed.
The ash·plant marks his stride.

\N{6b}:
A pack of blood·hounds,
led by Hornblower of Trinity |
brandishing a dog·whip |
in tally·ho cap |
and an old pair of grey trousers,
\gab{4330}%
follow from far,
picking up the scent,
nearer,
baying,
panting,
at fault,
breaking away,
throwing their tongues,
biting his heels,
leaping at his tail.

\N{6a}:
Bloom walks,
runs,
zig·zags,
gallops,
lugs laid back.
He is pelted with gravel,
cabbage stumps,
biscuit boxes,
eggs,
potatoes,
dead codfish,
woman's slipper slappers.

\N{6b}:
After him |
fresh found |
the hue and cry |
zig·zag gallops in hot pursuit |
of follow my leader:
65~C,
66~C,
Night Watch,
John Henry Menton,
Wisdom Hely,
V.B.~Dillon,
Councillor Nannetti,
Alexander Keyes,
Larry O'Rourke,
Joe Cuffe,
Mrs~O'Dowd,
Pisser Burke,
the Nameless One,
Mrs~Riordan,
the Citizen,
Garryowen,
Who·do·you·call·him,
Strange·face,
---
\gab{4340}

\N{6a}:
Fellow·that·so·like,
Saw·him·before,
Chap·with·a·wen,
Chris Callinan,
sir Charles Cameron,
Benjamin Dollard,
Lenehan,
Bartell d'Arcy,
Joe Hynes,
Red Murray,
editor Brayden,
T.M.~Healy,
Mr~Justice Fitzgibbon,
John Howard Parnell,
the Reverend Tinned Salmon,
Professor Joly,
Mrs~Breen,
Denis Breen,
Theodore Purefoy,
Mina Purefoy,
the Westland Row post·mistress,
---

\N{6b}:
C.P.~M'Coy,
friend of Lyons,
Hoppy Holohan,
man·in·the·street,
other·man·in·the·street,
Football boots,
pug·nosed driver,
rich protestant lady,
Davy Byrne,
Mrs~Ellen M'Guinness,
Mrs~Joe Gallaher,
George Lidwell,
Jimmy Henry on corns,
superintendent Laracy,
Father Cowley,
\gab{4350}%
Crofton out of the Collector-general's,
Dan Dawson,
dental surgeon Bloom with tweezers,
Mrs~Bob Doran,
Mrs~Kennefick,
Mrs~Wyse Nolan,
John Wyse Nolan,
---

\N{6a}:
Handsome·married·woman·rubbed·against·wide·behind·in·Clonskea·tram,
the book·seller of \Title{Sweets of Sin},
Miss Dubedat·and·she·did·bedad,
Mesdames Gerald and Stanislaus Moran of Roebuck,
the managing clerk of Drimmie's,
Wetherup,
colonel Hayes,
---

\N{6b}:
Mastiansky,
Citron,
Penrose,
Aaron Figatner,
Moses Herzog,
Michael~E Geraghty,
Inspector Troy,
Mrs~Galbraith,
the constable off Eccles street corner,
old doctor Brady with stethoscope,
the mystery man on the beach,
a retriever,
---
\gab{4360}

\N{6a}:
Mrs~Miriam Dandrade and all her lovers.

\N{6b}:
The Hue and Cry,
helter·skelter·pelter·welter.

\HueCry:
He's Bloom!
Stop Bloom!
Stop·a·bloom!
Stopper·robber!
Hi!
Hi!
Stop·him on the corner!

%--------------------------------------------------
\commentary{7. \textsc{Defiance}}{4:18:06}
%--------------------------------------------------

\N{7a}:
At the corner of Beaver street,
beneath the scaffolding,
Bloom panting |
stops on the fringe of the noisy quarrelling knot,
a lot |
not knowing a jot |
what hi!
hi!
row and wrangle |
round the who·what brawl·altogether.

\N{7a}:
Stephen,
with elaborate gestures,
breathing deeply and slowly.
\gab{4370}

\Stephen:
You are my guests.
Uninvited.
By virtue of the fifth of George |
and seventh of Edward.
History to blame.
Fabled by mothers of memory.

\N{7b}:
Private Carr,
to Cissy Caffrey.

\Carr:
Was he insulting you?

\Stephen:
Addressed her in vocative feminine.
Probably neuter.
Ungenitive.

\Voices:\\*
-- No, he didn't.\\
-- I seen him.\\
-- The girl there.\\
-- He was in Mrs~Cohen's.\\
-- What's up?\\
-- Soldier and civilian.

\Cissy[7b]:
\gab{4380}%
I was in company with the soldiers |
and they left me |
to do,
you know,
and the young man run up behind me.
But I'm faithful to the man that's treating me,
though I'm only a shilling whore.

\Voices:
Shes·faithful·the·man.

\N{7a}:
Stephen catches sight of Lynch's and Kitty's heads.

\Stephen:
Hail,
Sisyphus.

\N{7a}:
Stephen points to himself and the others.

\Stephen:
Poetic.
Uropoetic.

\Cissy:
Yes,
to go with him.
\gab{4390}%
And me with a soldier friend.

\Compton[7b]:
He doesn't half want a thick ear,
the blighter.
Biff him one,
Harry.

\N{7b}:
Private Carr,
to Cissy.

\Carr:
Was he insulting you |
while me and him was having a piss?

\N{7a}:
Lord Tennyson,
gentleman poet in Union Jack blazer and cricket flannels,
bare·headed,
flowing·bearded.

\LordTennyson:
Theirs not to reason why.

\Compton:
Biff him,
Harry.

\N{7a}:
Stephen,
to Private Compton.
\gab{4400}

\Stephen:
I don't know your name |
but you are quite right.
Doctor Swift says one man in armour |
will beat ten men in their shirts.
Shirt is synechdoche.
Part for the whole.

\N{7b}:
Cissy Caffrey,
to the crowd.

\Cissy:
No,
I was with the privates.

\Stephen:
\stage{(amiably)}
Why not?
The bold soldier boy.
In my opinion,
every lady for example….

\N{7b}:
Private Carr,
his cap awry,
advances to Stephen.
\gab{4410}

\Carr:
Say,
how would it be,
governor,
if I was to bash in your jaw?

\N{7a}:
Stephen looks up to the sky.

\Stephen:
How?
Very unpleasant.
Noble art of self·pretence.
Personally,
I detest action.

\N{7a}:
Stephen waves his hand.

\Stephen:
Hand hurts me slightly.
\french{Enfin ce sont vos oignons.}

\N{7a}:
To Cissy Caffrey.

\Stephen:
Some trouble is on here.
What is it precisely?

\N{7b}:
Dolly Gray from her balcony |
waves her handkerchief,
giving the sign of the heroine of Jericho.

\DollyGray:
Rahab.
Cook's son,
goodbye.
Safe home to Dolly.
Dream of the girl you left behind |
\gab{4420}%
and she will dream of you.

\N{7b}:
The soldiers turn their swimming eyes.

\N{7a}:
Bloom,
elbowing through the crowd,
plucks Stephen's sleeve vigorously.

\Bloom:
Come now,
professor,
that car·man is waiting.

\N{7a}:
Stephen turns.

\Stephen:
Eh?

\N{7a}:
Stephen disengages himself.

\Stephen:
Why should I not speak to him |
or to any human being |
who walks upright upon this oblate orange?

\N{7a}:
Stephen points his finger.

\Stephen:
I'm not afraid of what I can talk to |
if I see his eye.
Retaining the perpendicular.

\N{7a}:
Stephen staggers a pace back.

\N{7a}:
Bloom,
propping him.
\gab{4430}

\Bloom:
Retain your own.

\N{7a}:
Stephen laughs emptily.

\Stephen:
My centre of gravity is displaced.
I have forgotten the trick.
Let us sit down somewhere and discuss.
Struggle for life |
is the law of existence |
but |
but human philirenists,
notably the tsar and the king of England,
have invented arbitration.

\N{7a}:
Stephen taps his brow.

\Stephen:
But in here |
it is I |
must kill the priest and the king.

\BiddyClap[7a]:
Did you hear what the professor said?
He's a professor out of the college.

\CuntyKate[7a]:
\gab{4440}
I did.
I heard that.

\BiddyClap[7a]:
He expresses himself with such marked refinement of phraseology.

\CuntyKate[7a]:
Indeed,
yes.
And at the same time with such apposite trenchancy.

\N{7b}:
Private Carr pulls himself free and comes forward.

\Carr:
What's that you're saying about my king?

\N{7b}:
Edward the Seventh appears in an arch·way.
He wears a white jersey |
on which |
an image
of the Sacred Heart |
\gab{4450}%
is stitched |
with the insignia of Garter and Thistle,
Golden Fleece,
Elephant of Denmark,
Skinner's and Probyn's horse,
Lincoln's Inn bencher,
and ancient and honourable artillery company of Massachusetts.
He sucks a red jujube.
He is robed as a grand elect perfect and sublime mason |
with trowel and apron,
marked \Title{made in Germany}.
In his left hand,
he holds a plasterer's bucket |
on which is printed \french{Défense d'uriner}.
A roar of welcome greets him.

\N{7a}:
Edward the Seventh,
slowly,
solemnly |
but indistinctly.

\EdwardSeventh:
Peace,
perfect peace.
For identification,
bucket in my hand.
\gab{4460}%
Cheerio,
boys.

\N{7a}:
Edward the Seventh turns to his subjects.

\EdwardSeventh:
We have come here to witness a clean straight fight |
and we heartily wish both men |
the best of good luck.
Mahak makar a bak.

\N{7a}:
He shakes hands with Private Carr,
Private Compton,
Stephen,
Bloom
and Lynch.
\stage{General applause.}
Edward the Seventh lifts his bucket graciously |
in acknowledgment.

\N{7b}:
Private Carr to Stephen.

\Carr:
Say it again.

\N{7a}:
Stephen,
nervous,
friendly,
pulls himself up.

\Stephen:
I understand your point of view |
though I have no king myself |
for the moment.
\gab{4470}%
This is the age of patent medicines.
A discussion is difficult down here.
But this is the point.
You die for your country.
Suppose.

\N{7a}:
Stephen places his arm on Private Carr's sleeve.

\Stephen:
Not that I wish it for you.
But I say:
Let my country die for me.
Up to the  present |
it has done so.
I didn't want it to die.
Damn death.
Long live life!

\N{7a}:
Edward the Seventh levitates over heaps of slain,
in the garb |
and with the halo |
of Joking Jesus,
a white jujube in his phosphorescent face.

\EdwardSeventh:
\begin{verse}
    My methods are new and are causing surprise.\\
    To make the blind see |
        I throw dust in their eyes.
\end{verse}

\Stephen:
\gab{4480}
Kings and unicorns!

\N{7a}:
Stephen falls back a pace.

\Stephen:
Come somewhere and we'll… 
What was that girl saying?…

\Compton[7b]:
Eh,
Harry,
give him a kick in the knackers.
Stick one into Jerry.

\N{7a}:
Bloom,
to the privates,
softly.

\Bloom:
He doesn't know what he's saying.
Taken a little more than is good for him.
Absinthe.
Green·eyed monster.
I know him.
He's a gentleman,
a poet.
It's all right.

\N{7a}:
Stephen nods,
smiling and laughing.
\gab{4490}

\Stephen:
Gentleman,
patriot,
scholar and judge of impostors.

\Carr:
I don't give a bugger who he is.

\Compton:
We don't give a bugger who he is.

\Stephen:
I seem to annoy them.
Green rag to a bull.

\N{7b}:
Kevin Egan of Paris |
in black Spanish tasselled shirt |
and peep-o'-day boy's hat |
signs to Stephen.

\KevinEgan:
\gab{4500}
H'lo!
\french{Bonjour!}
The \french{vieille ogresse} with the \french{dents jaunes}.

\N{7b}:
Patrice Egan peeps from behind,
his rabbit·face nibbling a quince leaf.

\Patrice:
\french{Socialiste!}

\N{7a}:
Don Emile Patrizio Franz Rupert Pope Hennessy,
in medieval hauberk,
two \emph{wild geese} volant on his helm,
with noble indignation |
points a mailed hand against the privates.

\DonEmile:
Werf those eykes to foot·boden,
big grand porcos of john·yellows |
todos covered of gravy!

\Bloom:
\gab{4510}
\stage{(to Stephen)}
Come home.
You'll get into trouble.

\N{7a}:
Stephen,
swaying.

\Stephen:
I don't avoid it.
He provokes my intelligence.

\BiddyClap[7b]:
One immediately observes that he is of patrician lineage.

\Virago[7b]:
Green above the red,
says he.
Wolfe Tone.

\Bawd[7b]:
The red's as good as the green.
And better.
Up the soldiers!
Up King Edward!%
\gab{4520}

\ARough[7b]:
\stage{(laughs)}
Ay!
Hands up to De Wet.

\N{7b}:
The Citizen,
with a huge emerald muffler and shillelagh,
calls.

\Citizen:
\begin{verse}
    May the God above\\
    Send down a dove\\
    With teeth as sharp as razors\\
    To slit the throats\\
    Of the English dogs\\
\gab{4530}%
    That hanged our Irish leaders.
\end{verse}

\N{7a}:
The Croppy Boy,
the rope·noose round his neck,
gripes in his issuing bowels |
with both hands.

\CroppyBoy:
\begin{verse}
    I bear no hate to a living thing,\\
    But I love my country beyond the king.
\end{verse}

\N{7b}:
Rumbold,
Demon Barber,
accompanied by two black·masked assistants,
advances with gladstone bag |
which he opens.

\RumboldDemonBarber:
Ladies and gents,
cleaver purchased by Mrs~Pearcy to slay Mogg.
Knife with which |
Voisin dismembered the wife of a compatriot |
and hid remains in a sheet in the cellar,
\gab{4540}%
the unfortunate female's throat |
being cut from ear to ear.
Phial containing arsenic |
retrieved from body of Miss Barron |
which sent Seddon to the gallows.

\N{7b}:
Rumbold jerks the rope.
The assistants leap at the victim's legs |
and drag him downward,
grunting.
The Croppy Boy's tongue protrudes violently.

\CroppyBoy:
\begin{verse}
    Hor·hot ho hray hor hother's hest.
\end{verse}

\N{7b}:
The Croppy Boy gives up the ghost.
A violent erection of the hanged |
sends gouts of sperm spouting through his death·clothes |
on to the cobble·stones.
Mrs~Bellingham,
Mrs~Yelverton Barry,
\gab{4550}%
and the Honourable Mrs~Mervyn Talboys |
rush forward with their handkerchiefs |
to sop it up.

\Rumbold:
I'm near it myself.

\N{7b}:
Rumbold undoes the noose.

\Rumbold:
Rope which hanged the awful rebel.
Ten shillings a time.
As applied to Her Royal Highness.

\N{7b}:
Rumbold plunges his head |
into the gaping belly of the hanged |
and draws out his head again |
clotted with coiled and smoking entrails.

\Rumbold:
My painful duty has now been done.
God save the king!

\N{7a}:
Edward the Seventh dances slowly,
solemnly,
rattling his bucket,
\gab{4560}%
and sings with soft contentment.

\EdwardSeventh:
\begin{verse}
    On coronation day, on coronation day,\\
    O, won't we have a merry time,\\
    Drinking whisky, beer and wine!
\end{verse}

\Carr[7b]:
Here.
What are you saying about my king?

\N{7a}:
Stephen throws up his hands.

\Stephen:
O,
this is too monotonous!
Nothing.
He wants my money and my life,
though want must be his master,
for some brutish empire of his.
Money I haven't.

\N{7a}:
Stephen searches his pockets vaguely.
\gab{4570}

\Stephen:
Gave it to someone.

\Carr:
Who wants your bleeding money?

\N{7a}:
Stephen tries to move off.

\Stephen:
Will someone tell me |
where I am least likely |
to meet these necessary evils?
\french{Ça se voit aussi à Paris.}
Not that I…
But,
by Saint Patrick…!

\N{7b}:
The women's heads coalesce.
Old Gummy Granny in sugar·loaf hat appears |
seated on a toadstool,
the death flower of the potato blight on her breast.%
\gab{4580}

\Stephen:
Aha!
I know you,
gammer!
Hamlet,
revenge!
The old sow that eats her farrow!

\N{7b}:
Old Gummy Granny,
rocking to and fro.

\GummyGranny:
Ireland's sweetheart,
the king of Spain's daughter,
alanna.
Strangers in my house,
bad manners to them!

\N{7b}:
She keens with banshee woe.

\GummyGranny:
Ochone!
Ochone!
Silk of the kine!

\N{7b}:
She wails.

\GummyGranny:
You met with poor old Ireland |
and how does she stand?

\Stephen:
How do I stand you?
\gab{4590}%
The hat trick!
Where's the third person of the Blessed Trinity?
Soggarth Aroon?
The reverend Carrion Crow.

\Cissy[7b]:
\stage{(shrill)}
Stop them from fighting!

\ARough[7a]:
Our men retreated.

\N{7b}:
Private Carr,
tugging at his belt.

\Carr:
I'll wring the neck of any fucker |
says a word against my fucking king.

\N{7a}:
Bloom,
terrified.
\gab{4600}

\Bloom:
He said nothing.
Not a word.
A pure misunderstanding.

\Compton[7b]:
Go it,
Harry.
Do him one in the eye.
He's a proBoer.

\Stephen:
Did I?
When?

\N{7a}:
Bloom,
to the redcoats.

\Bloom:
We fought for you in South Africa,
Irish missile troops.
Isn't that history?
Royal Dublin Fusiliers.
Honoured by our monarch.

\N{7b}:
The Navvy,
staggering past.

\Navvy:
O,
yes!
O~God,
yes!
O,
make the kwawr a krowawr!
O!
Bo!%
\gab{4610}

\N{7b}:
Casqued halberdiers in armour |
thrust forward a pentice of gutted spearpoints.
Major Tweedy,
moustached like Turko the terrible,
in bearskin cap |
with hackle·plume and accoutrements,
with epaulettes,
gilt chevrons |
and sabretaches,
his breast bright with medals,
toes the line.
He gives the pilgrim warrior's |
sign of the knights templars.

\N{7b}:
Major Tweedy growls gruffly.

\MajorTweedy:
Rorke's Drift!
Up,
guards,
and at them!
Mahar shalal hashbaz.

\Citizen[7b]:
\gab{4620}
\irish{Erin go bragh!}

\N{7a}:
Major Tweedy and the Citizen |
exhibit to each other |
medals,
decorations,
trophies of war,
wounds.
Both salute with fierce hostility.

\Carr[7b]:
I'll do him in.

\N{7b}:
Private Compton waves the crowd back.

\Compton:
Fair play,
here.
Make a bleeding butcher's shop |
of the bugger.

\N{7a}:
Massed bands blare \Title{Garryowen} and \Title{God Save the King}.%
\gab{4630}

\Cissy[7b]:
They're going to fight.
For me!

\CuntyKate[7a]:
The brave and the fair.

\BiddyClap[7a]:
Methinks yon sable knight |
will joust it with the best.

\N{7a}:
Cunty Kate,
blushing deeply.

\CuntyKate:
Nay,
madam.
The gules doublet and merry saint George |
for me!

\Stephen:
\gab{4640}
\begin{verse}
    The harlot's cry from street to street\\
    Shall weave Old Ireland's winding sheet.
\end{verse}

\N{7b}:
Private Carr,
loosening his belt,
shouts.

\Carr:
I'll wring the neck of any fucking bastard |
says a word against my bleeding fucking king.

\N{7a}:
Bloom shakes Cissy Caffrey's shoulders.

\Bloom:
Speak,
you!
Are you struck dumb?
You are the link between nations and generations.
Speak,
woman,
sacred lifegiver!

\N{7b}:
Cissy,
alarmed,
seizes Private Carr's sleeve.
\gab{4650}

\Cissy:
Amn't I with you?
Amn't I your girl?
Cissy's your girl.
\stage{(she cries)}
Police!

\N{7a}:
Stephen,
ecstatically,
to Cissy Caffrey.

\Stephen:
\begin{verse}
    White thy fambles,
        red thy gan\\
    And thy quarrons |
        dainty is.
\end{verse}

\Voices:
Police!

\DistantVoices:
Dublin's burning!
Dublin's burning!
\gab{4660}%
On fire,
on fire!

\N{7b}:
Brimstone fires spring up.
Dense clouds roll past.
Heavy Gatling guns boom.
Pandemonium.
Troops deploy.
Gallop of hoofs.
Artillery.
Hoarse commands.
Bells clang.
Backers shout.
Drunkards bawl.
Whores screech.
Foghorns hoot.
Cries of valour.
Shrieks of dying.
Pikes clash on cuirasses.
Thieves rob the slain.

\N{7a}:
Birds of prey,
winging from the sea,
rising from marshlands,
swooping from eyries,
hover screaming,
gannets,
cormorants,
vultures,
goshawks,
climbing woodcocks,
peregrines,
merlins,
black·grouse,
sea eagles,
gulls,
albatrosses,
barnacle geese.

\N{7b}:
The midnight sun is darkened.
The earth trembles.
\gab{4670}%
The dead of Dublin |
from Prospect and Mount Jerome |
in white sheepskin overcoats and black goat·fell cloaks |
arise |
and appear to many.
A chasm opens |
with a noiseless yawn.

\N{7a}:
Tom Rochford,
winner,
in athlete's singlet and breeches,
arrives at the head of the National Hurdle Handicap |
and leaps into the void.
He is followed by a race of runners and leapers.
In wild attitudes |
they spring from the brink.
Their bodies plunge.

\N{7b}:
Factory lasses with fancy clothes |
toss red·hot Yorkshire baraa·bombs.
Society ladies lift their skirts above their heads |
to protect themselves.
Laughing witches in red cutty sarks |
ride through the air on broomsticks.
Quaker·lyster plasters blisters.
\gab{4680}%
It rains dragons' teeth.

\N{7a}:
Armed heroes spring up from furrows.
They exchange in amity |
the pass of knights of the red cross |
and fight duels with cavalry sabres:
Wolfe Tone against Henry Grattan,

\N{7b}:
Smith O'Brien against Daniel O'Connell,

\N{7a}:
Michael Davitt against Isaac Butt,

\N{7b}:
Justin M'Carthy against Parnell,

\N{7a}:
Arthur Griffith against John Redmond,

\N{7b}:
John O'Leary against Lear O'Johnny,

\N{7a}:
Lord Edward FitzGerald against Lord Gerald FitzEdward,

\N{7b}:
the O'Donoghue of the Glens against the Glens of the O'Donoghue.

\N{7a}:
On an eminence,
the centre of the earth,
rises the feld·altar of Saint Barbara.
Black candles rise |
from its gospel and epistle horns.
\gab{4690}%
From the high barbacans of the tower,
two shafts of light fall on the smoke·palled altar·stone.
On the altar·stone |
Mrs~Mina Purefoy,
goddess of unreason,
lies,
naked,
fettered,
a chalice resting on her swollen belly.

\N{7b}:
Father Malachi O'Flynn |
in a lace petticoat and reversed chasuble,
his two left feet back to the front,
celebrates camp mass.
The reverend Mr~Hugh C~Haines Love~\textsc{m.a.}\ 
in a plain cassock and mortarboard,
his head and collar back to the front,
holds over the celebrant's head |
an open umbrella.

\FrMalachiOFlynn[7a]:
\latin{Introibo ad altare diaboli.}

\HainesLove[7b]:
\gab{4700}
To the devil |
which hath made glad my young days.

\N{7a}:
Father Malachi O'Flynn takes from the chalice |
and elevates a blood·dripping host.

\FrMalachiOFlynn:
\latin{Corpus meum.}

\N{7b}:
The Reverend Mr Haines Love |
raises high behind the celebrant's petticoat,
revealing his grey bare hairy buttocks |
between which |
a carrot is stuck.

\HainesLove:
My body.

\VoiceAllDamned[7b]:
Htengier Tnetopinmo Dog Drol eht rof,
Aiulella!

\N{7b}:
From on high |
the voice of Adonai calls.

\Adonai:
\gab{4710}
Dooooooooooog!

\VoiceAllBlessed[7a]:
Alleluia,
for the Lord God Omnipotent reigneth!

\N{7a}:
From on high |
the voice of Adonai calls.

\Adonai:
Goooooooooood!

\N{7b}:
In strident discord |
peasants and townsmen of Orange and Green factions |
sing \Title{Kick the Pope}
and \Title{Daily, daily sing to Mary}.

\N{7b}:
Private Carr,
with ferocious articulation.
\gab{4720}

\Carr:
I'll do him in,
so help me fucking Christ!
I'll wring the bastard fucker's |
bleeding blasted fucking windpipe!

\N{7b}:
The retriever,
nosing on the fringe of the crowd,
barks noisily.

\N{7a}:
Bloom runs to Lynch.

\Bloom:
Can't you get him away?

\Lynch:
He likes dialectic,
the universal language.
Kitty!

\N{7b}:
To Bloom.

\Lynch:
Get him away,
you.
He won't listen to me.

\N{7b}:
Lynch drags Kitty away.

\N{7a}:
Stephen points.
\gab{4730}

\Stephen:
Exit Judas.
\latin{Et laqueo se suspendit.}

\N{7a}:
Bloom runs to Stephen.

\Bloom:
Come along with me now |
before worse happens.
Here's your stick.

\Stephen:
Stick,
no.
Reason.
This feast of pure reason.

\N{7b}:
Old Gummy Granny |
thrusts a dagger towards Stephen's hand.

\GummyGranny:
Remove him,
acushla.
At 8.35~\textsc{a.m.},
you will be in heaven |
and Ireland will be free.

\N{7b}:
She prays.

\GummyGranny:
O~good God,
take him!

\N{7b}:
Cissy Caffrey,
pulling Private Carr.
\gab{4740}

\Cissy:
Come on,
you're boosed.
He insulted me |
but I forgive him.

\N{7b}:
Shouting in his ear.

\Cissy:
I forgive him for insulting me.

\N{7a}:
Bloom over Stephen's shoulder.

\Bloom:
Yes,
go.
You see he's incapable.

\N{7b}:
Private Carr breaks loose.

\Carr:
I'll insult him.

\N{7b}:
Private Carr rushes towards Stephen,
fist outstretched,
and strikes him in the face!

\N{7a}:
Stephen totters,
collapses,
falls,
stunned.
He lies prone,
his face to the sky,
his hat rolling to the wall.
Bloom follows and picks it up.%
\gab{4750}

\MajorTweedy[7a]:
\stage{(loudly)}
Carbine in bucket!
Cease fire!
Salute!

\Retriever:
\stage{(barking furiously)}
Ute ute ute ute ute ute ute ute.

\Crowd[7b]:\\*
-- Let him up!\\
-- Don't strike him when he's down!\\
-- Air!\\
-- Who?\\
-- The soldier hit him.\\
-- He's a professor.\\
-- Is he hurted?\\
-- Don't manhandle him!\\
-- He's fainted!

\Hag[7a]:
What call had the redcoat to strike the gentleman |
and he under the influence.
\gab{4760}%
Let them go and fight the Boers!

\Bawd[7b]:
Listen to who's talking!
Hasn't the soldier a right |
to go with his girl?
He gave him the coward's blow.

\N{7b}:
They grab at each other's hair,
claw at each other |
and spit.

\Retriever:
\stage{(barking)}
Wow wow wow.

\N{7a}:
Bloom shoves them back,
loudly.

\Bloom:
Get back,
stand back!

\N{7b}:
Private Compton,
tugging his comrade.
\gab{4770}

\Compton:
Here.
Bugger off,
Harry.
Here's the cops!

%--------------------------------------------------
\commentary{8. \textsc{Rescue}}{4:39:13}
%--------------------------------------------------

\N{8a}:
Two rain·caped watch,
tall,
stand in the group.

\FirstWatch[8a]:
What's wrong here?

\Compton:
We were with this lady.
And he insulted us.
And assaulted my chum.
\stage{(the retriever barks)}
Who owns the bleeding tyke?

\Cissy:
\stage{(with expectation)}
Is he bleeding!

\N{8b}:
A Man,
rising from his knees.
\gab{4780}

\Man1:
No.
Gone off.
He'll come to all right.

\N{8a}:
Bloom glances sharply at the man.

\Bloom:
Leave him to me.
I can easily…

\SecondWatch[8b]:
Who are you?
Do you know him?

\N{8b}:
Private Carr lurches towards the Watch.

\Carr:
He insulted my lady friend.

\Bloom:
\stage{(angrily)}
You hit him without provocation.
I'm a witness.
Constable,
take his regimental number.

\SecondWatch[8b]:
\gab{4790}
I don't want your instructions in the discharge of my duty.

\N{8b}:
Private Compton,
pulling his comrade.

\Compton:
Here,
bugger off,
Harry.
Or Bennett'll shove you in the lock·up.

\N{8b}:
Private Carr,
staggering as he is pulled away.

\Carr:
God fuck old Bennett.
He's a white·arsed bugger.
I don't give a shit for him.

\N{8a}:
The First Watch takes out his notebook.

\FirstWatch:
What's his name?

\N{8a}:
Bloom,
peering over the crowd.
\gab{4800}

\Bloom:
I just see a car there.
If you give me a hand a second,
sergeant…

\FirstWatch:
Name and address.

\N{8b}:
Corny Kelleher,
weepers round his hat,
a death wreath in his hand,
appears among the bystanders.

\Bloom:
\stage{(quickly)}
O,
the very man!
\stage{(he whispers)}
Simon Dedalus' son.
A bit sprung.
Get those policemen to move those loafers back.

\SecondWatch[8b]:
\gab{4810}
Night,
Mr~Kelleher.

\N{8b}:
Corny Kelleher,
to the watch,
with drawling eye.

\Corny:
That's all right.
I know him.
Won a bit on the races.
Gold cup.
\Title{Throwaway}.
\stage{(he laughs)}
Twenty to one.
Do you follow me?

\N{8a}:
The First Watch turns to the crowd.

\FirstWatch:
Here,
what are you all gaping at?
Move on out of that.

\N{8b}:
The crowd disperses slowly,
muttering,
down the lane.

\Corny:
\gab{4820}
Leave it to me,
sergeant.
That'll be all right.
\stage{(he laughs,
shaking his head)}
We were often as bad ourselves,
ay or worse.
What?
Eh,
what?

\FirstWatch:
\stage{(laughs)}
I suppose so.

\N{8b}:
Corny Kelleher nudges the second watch.

\Corny:
Come and wipe your name off the slate.

\N{8b}:
He lilts,
wagging his head.

\Corny:
With my tooraloom tooraloom tooraloom tooraloom.
What,
eh,
do you follow me?

\SecondWatch:
\stage{(genially)}
\gab{4830}
Ah,
sure |
we were too.

\Corny:
\stage{(winking)}
Boys will be boys.
I've a car round there.

\SecondWatch:
All right,
Mr~Kelleher.
Good night.

\Corny:
I'll see to that.

\N{8a}:
Bloom shakes hands with both of the watch |
in turn.

\Bloom:
Thank you very much,
gentlemen.
Thank you.

\N{8a}:
Bloom mumbles confidentially.

\Bloom:
We don't want any scandal,
you understand.
\gab{4840}%
Father is a well·known highly respected citizen.
Just a little wild oats,
you understand.

\FirstWatch:
O\@.
I understand,
sir.

\SecondWatch:
That's all right,
sir.

\FirstWatch:
It was only in case of corporal injuries,
I'd have to report it |
at the station.

\N{8a}:
Bloom nods rapidly.

\Bloom:
Naturally.
Quite right.
Only your bounden duty.

\SecondWatch:
\gab{4850}
It's our duty.

\Corny:
Good night,
men.

\N{8a}:
The Watch,
saluting together.

\Watch:
Night,
gentlemen.

\N{8a}:
They move off |
with slow heavy tread.

\Bloom:
\stage{(blows)}
Providential |
you came on the scene.
You have a car?…

\N{8b}:
Corny Kelleher laughs,
pointing his thumb over his right shoulder |
\gab{4860}%
to the car |
brought up against the scaffolding.

\Corny:
Two commercials that were standing fizz in Jammet's.
Like princes,
faith.
One of them lost two quid on the race.
Drowning his grief.
And were on for a go |
with the jolly girls.
So I landed them up on Behan's car |
and down to night·town.

\Bloom:
I was just going home by Gardiner street |
when I happened to…

\Corny:
\stage{(laughs)}
Sure |
they wanted me to join in with the mots.
No,
by God,
says I\@.
Not for old stagers |
like myself and yourself.

\N{8b}:
He laughs again |
and leers with lacklustre eye.

\Corny:
\gab{4870}
Thanks be to God |
we have it in the house,
what,
eh,
do you follow me?
Hah,
hah,
hah!

\Bloom:
\stage{(tries to laugh)}
He,
he,
he!
Yes.
Matter of fact |
I was just visiting an old friend of mine there,
Virag,
you don't know him
(poor fellow,
he's laid up for the past week)
and we had a liquor together |
and I was just making my way home…

\N{8b}:
The horse neighs.

\Horse:
Ho·ho·ho·ho·ho·hoh!
Ho·ho·ho·home!

\Corny:
\gab{4880}
Sure |
it was Behan our jarvey there |
that told me |
after we left the two commercials in Mrs~Cohen's |
and I told him to pull up |
and got off to see.
\stage{(he laughs)}
Sober hearse drivers a speciality.
Will I give him a lift home?
Where does he hang out?
Somewhere in Cabra,
what?

\Bloom:
No,
in Sandycove,
I believe,
from what he let drop.

\N{8a}:
Stephen,
prone,
breathes to the stars.

\N{8b}:
Corny Kelleher,
asquint,
drawls at the horse.

\N{8a}:
Bloom,
in gloom,
looms down.

\N{8b}:
Corny Kelleher scratches his nape.
\gab{4890}

\Corny:
Sandycove!

\N{8b}:
He bends down and calls to Stephen.

\Corny:
Eh!

\N{8b}:
Corny Kelleher calls again.

\Corny:
Eh!
He's covered with shavings anyhow.
Take care |
they didn't lift anything off him.

\Bloom:
No,
no,
no.
I have his money and his hat here |
and stick.

\Corny:
Ah,
well,
he'll get over it.
No bones broken.
Well,
I'll shove along.
\stage{(he laughs)}
I've a rendezvous in the morning.
Burying the dead.
Safe home!

\Horse:
\stage{(neighs)}
Ho·ho·ho·ho·home.

\Bloom:
\gab{4900}
Good night.
I'll just wait |
and take him along in a few…

\N{8b}:
Corny Kelleher returns to the outside car |
and mounts it.
The horse harness jingles.

\N{8b}:
Corny Kelleher,
from the car,
standing.

\Corny:
Night.

\Bloom:
Night.

\N{8b}:
The jarvey chucks the reins |
and raises his whip encouragingly.
The car and horse back slowly,
awkwardly,
and turn.
Corny Kelleher on the side·seat |
sways his head to and fro |
\gab{4910}%
in sign of mirth |
at Bloom's plight.
The jarvey joins in the mute pantomimic merriment |
nodding from the farther seat.

\N{8a}:
Bloom shakes his head |
in mute mirthful reply.

\N{8b}:
With thumb and palm,
Corny Kelleher reassures |
that the two bobbies will allow the sleep to continue,
for |
what else is to be done.

\N{8a}:
With a slow nod,
Bloom conveys his gratitude,
as that is exactly what Stephen needs.

\N{8b}:
The car jingles tooraloom |
round the corner of the tooraloom lane.
Corny Kelleher again reassuralooms |
with his hand.

\N{8a}:
Bloom with his hand |
assuralooms Corny Kelleher |
that he is reassuraloomtay.

\N{8b}:
The tinkling hoofs and jingling harness grow fainter |
with their tooralooloo looloo lay.
\gab{4920}

\N{8a}:
Bloom,
holding in his hand Stephen's hat,
festooned with shavings,
and ash·plant,
stands irresolute.
Then he bends to him and shakes him by the shoulder.

\Bloom:
Eh!
Ho!

\N{8a}:
There is no answer.
He bends again.

\Bloom:
Mr~Dedalus!

\N{8a}:
There is no answer.

\Bloom:
The name if you call.
Somnambulist.

\N{8a}:
Bloom bends again |
and hesitating,
brings his mouth |
near the face of the prostrate form.

\Bloom:
Stephen!

\N{8a}:
There is no answer.
He calls again.

\Bloom:
Stephen!

\N{8a}:
Stephen frowns.
\gab{4930}

\Stephen:
Who?
Black panther.
Vampire.

\N{8a}:
Stephen sighs and stretches himself,
then murmurs thickly |
with prolonged vowels.

\Stephen:
\begin{verse}
    Who… drive… Fergus now\\
    And pierce… wood's woven shade…?
\end{verse}

\N{8a}:
Stephen turns on his left side,
sighing,
doubling himself together.

\Bloom:
Poetry.
Well educated.
Pity.

\N{8a}:
Bloom bends again |
and undoes the buttons of Stephen's waistcoat.

\Bloom:
To breathe.

\N{8a}:
Bloom brushes the wood·shavings from Stephen's clothes |
with light hand and fingers.

\Bloom:
One pound seven.
Not hurt anyhow.

\N{8a}:
Bloom listens.

\Bloom:
What?

\Stephen:
\gab{4940}
\stage{(murmurs)}
\begin{verse}
   … shadows… the woods\\
   … white breast… dim sea.
\end{verse}

\N{8a}:
Stephen stretches out his arms,
sighs again and curls his body.
Bloom,
holding the hat and ash·plant,
stands erect.
A dog barks in the distance.
Bloom tightens and loosens his grip on the ash·plant.
He looks down on Stephen's face and form.

\N{8a}:
Bloom communes with the night.

\Bloom:
Face reminds me of his poor mother.
In the shady wood.
The deep white breast.
Ferguson,
I think I caught.
\gab{4950}%
A girl.
Some girl.
Best thing could happen him.
\stage{(he murmurs)}
… swear that I will always hail,
ever conceal,
never reveal,
any part or parts,
art or arts…
\stage{(he murmurs)}
… in the rough sands of the sea…
a cable·tow's length from the shore…
where the tide ebbs…
and flows…

\N{8a}:
Silent,
thoughtful,
alert,
he stands on guard,
his fingers at his lips |
in the attitude of secret master.

\N{8b}:
Against the dark wall,
a figure appears slowly,
a fairy boy of eleven,
a changeling,
kidnapped,
dressed in an Eton suit |
with glass shoes |
and a little bronze helmet,
holding a book in his hand.
He reads from right to left |
inaudibly,
smiling,
kissing the page.%
\gab{4960}

\N{8a}:
Bloom,
wonder·struck,
calls inaudibly.

\Bloom:
Rudy!

\N{8b}:
Rudy gazes,
unseeing,
into Bloom's eyes |
and goes on reading,
kissing,
smiling.
He has a delicate mauve face.
On his suit |
he has diamond and ruby buttons.
In his free left hand,
he holds a slim ivory cane |
with a violet bow·knot.
A white lambkin peeps out of his waistcoat pocket.

% \rte{4:48:51}
