%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\input{defs}
% \input{largeprint} \renewcommand{\gab}[1]{\marginpar{\tiny #1}}

\title{Circe\\
Part~I\\
\small second draft}
\author{George V. Reilly\\
for the Wild Geese Players of Seattle\\
{\emph{www.WildGeeseSeattle.org}}}
\date{April 6th, 2009}
\raggedbottom

\begin{document}

\maketitle

\pagebreak

\commentary{
In Homer's \Title{Odyssey},
Circe is a sorceress who drugs Odysseus' men and turns them into swine.
Odysseus rescues his men from Circe and becomes her lover.

At the end of \Title{The Oxen of the Sun}, after Burke's had closed,
Stephen Dedalus and Vincent Lynch set off for Nighttown
in search of a brothel.
Stephen is ratarsed with drink and absinthe.
Leopold Bloom, taking a paternal interest in Stephen,
follows them at a distance.

An enormously long chapter full of hallucinations and nightmares,
where Stephen and Bloom sink to the bottom,
and Bloom confronts many of his unspoken desires and emerges strengthened.

Burgess and Kenner both argue that the hallucinations
are coming from without,
that Bloom himself does not experience them.
Burgess: ``They are hallucinations from the author's brain,
not his hero's, but there is nothing vague or shadowy about them.''
}

\N1:
The Mabbot street entrance of Nighttown,
before which stretches an uncobbled tram siding set with skeleton tracks,
red and green will-o'-the-wisps and danger signals.
Rows of grimy houses with gaping doors.
Rare lamps with faint rainbow fins.
Round Rabaiotti's halted ice gondola,
stunted men and women squabble.
They grab wafers, between which are wedged
lumps of coral and copper snow.
Sucking, they scatter slowly.
Children.
The swancomb of the gondola, high reared, forges on through the murk,
white and blue under a lighthouse.
Whistles call and answer.

\Calls:
\gab{10}
Wait, my love, and I'll be with you.

\Answers:
Round behind the stable.

\N1: A deafmute idiot with goggle eyes, his shapeless mouth dribbling,
jerks past, shaken in Saint Vitus' dance.

\N2:
A chain of children's hands imprisons him.

\Children:
Kithogue! Salute!

\N1:
The Idiot lifts a palsied left arm and gurgles.

\Idiot:
Grhahute!%
\gab{20}

\Children:
Where's the great light?

\Idiot:
\stage{(gobbling)} Ghaghahest.

\N2:
They release him.
He jerks on.
A pigmy woman swings on a rope slung between two railings, counting.
A form sprawled against a dustbin and muffled by its arm and hat snores,
groans, grinding growling teeth, and snores again.
On a step, a gnome, totting among a rubbishtip,
crouches to shoulder a sack of rags and bones.
A crone, standing by with a smoky oillamp,
\gab{30}
rams her last bottle in the maw of his sack.
He heaves his booty, tugs askew his peaked cap, and hobbles off mutely.
The crone makes back for her lair, swaying her lamp.
A bandy child, asquat on the doorstep with a paper shuttlecock,
crawls sidling after her in spurts, clutches her skirt, scrambles up.

\N1:
A drunken navvy grips with both hands the railings of an area, lurching heavily.
At a corner, two night watch in shouldercapes,
their hands upon their staffholsters, loom tall.
A plate crashes: a woman screams: a child wails.
Oaths of a man roar, mutter, cease.
Figures wander, lurk, peer from warrens.
In a room lit by a candle stuck in a bottleneck,
\gab{40}
a slut combs out the tatts from the hair of a scrofulous child.

\N2:
Cissy Caffrey's voice, still young, sings shrill from a lane.

\Cissy:
\begin{verse}
    I gave it to Molly\\
    Because she was jolly,\\
    The leg of the duck,\\
    The leg of the duck.
\end{verse}

\N2:
Private Carr and Private Compton, swaggersticks tight in their oxters,
as they march unsteadily, rightaboutface, and
burst together from their mouths a volleyed fart.
Laughter of men from the lane.
\gab{50}

\N1:
A hoarse virago retorts.

\Virago:
Signs on you, hairy arse.
More power the Cavan girl.

\Cissy:
More luck to me.
Cavan, Cootehill, and Belturbet.
\stage{(she sings)}

\begin{verse}
    I gave it to Nelly\\
    To stick in her belly,\\
    The leg of the duck,\\
    The leg of the duck.
\end{verse}

\N2:
Private Carr and Private Compton turn and counterretort,
\gab{60}
their tunics bloodbright in a lampglow,
black sockets of caps on their blond cropped polls.

\N1:
Stephen Dedalus and Lynch pass through the crowd close to the redcoats.

\N2:
Private Compton jerks his finger.

\Compton:
Way for the parson.

\N2:
Private Carr turns and calls.

\Carr:
What ho, parson!

\Cissy:
\stage{(her voice soaring higher)}

\begin{verse}
\gab{70}%
    She has it, she got it,\\
    Wherever she put it,\\
    The leg of the duck.
\end{verse}

\N1:
Lynch, his jockeycap low on his brow, attends Stephen,
a sneer of discontent wrinkling his face.

\N2:
Stephen, flourishing the ashplant in his left hand,
chants with joy the introit for paschal time.

\Stephen:
\latin{Vidi aquam egredientem de templo a latere dextro.
Alleluia.}

\N1:
The famished snaggletusks of an elderly bawd protrude from a doorway.

\Bawd:
\gab{80}
\stage{(her voice whispering huskily)}
Sst! Come here till I tell you.
Maidenhead inside. Sst!

\Stephen:
% \stage{(altius aliquantulum)}
\latin{Et omnes ad quos pervenit aqua ista.}

\N1:
The bawd spits in their trail her jet of venom.

\Bawd:
Trinity medicals.
Fallopian tube.
All prick and no pence.

\N2:
Edy Boardman, sniffling, crouched with Bertha Supple,
draws her shawl across her nostrils.

\Edy:
\gab{90}
\stage{(bickering)}
And says the one:
I seen you up Faithful place with your squarepusher,
the greaser off the railway, in his come-to-bed hat.
Did you, says I.
That's not for you to say, says I.
You never seen
me in the mantrap with a married highlander, says I.
The likes of her!
Stag that one is!
Stubborn as a mule!
And her walking with two fellows the one time,
Kilbride, the engine driver, and lance-corporal Oliphant.

\Stephen:
% \stage{(triumphaliter)}
\latin{Salvi facti sunt.}

\N1:
He flourishes his ashplant, shivering the lamp image,
shattering light over the world.
\gab{100}
A liver and white spaniel on the prowl slinks after him, growling.

\N2:
Lynch scares it with a kick.

\Lynch:
So that?

\Stephen:
% \stage{(looks behind)}
So that gesture, not music not odour, would be a universal language,
the gift of tongues rendering visible not the lay sense
but the first entelechy, the structural rhythm.

\Lynch:
Pornosophical philotheology.
Metaphysics in Mecklenburgh street!

\Stephen:
\gab{110}
We have shrewridden Shakespeare and henpecked Socrates.
Even the allwisest Stagyrite was bitted, bridled and mounted by a light of love.

\Lynch:
Ba!

\Stephen:
Anyway, who wants two gestures to illustrate a loaf and a jug?
This movement illustrates the loaf and jug of bread or wine in Omar.
Hold my stick.

\Lynch:
Damn your yellow stick.
\gab{120}
Where are we going?

\Stephen:
Lecherous lynx, to \french{la belle dame sans merci}, Georgina Johnson,
\latin{ad deam qui laetificat iuventutem meam}.

\N1:
Stephen thrusts the ashplant on him and slowly holds out his hands,
his head going back till both hands are a span from his breast, down turned,
in planes intersecting, the fingers about to part, the left being higher.

\Lynch:
Which is the jug of bread? It skills not.
That or the customhouse.
Illustrate thou.
\gab{130}
Here take your crutch and walk.

\N2:
They pass.
Tommy Caffrey scrambles to a gaslamp and, clasping, climbs in spasms.
From the top spur he slides down.
Jacky Caffrey clasps to climb.
The navvy lurches against the lamp.
The twins scuttle off in the dark.
The navvy, swaying, presses a forefinger against a wing of his nose and
ejects from the farther nostril a long liquid jet of snot.
Shouldering the lamp, he staggers away through the crowd with his flaring cresset.

\N1:
Snakes of river fog creep slowly.
From drains, clefts, cesspools, middens
arise on all sides stagnant fumes.
A glow leaps in the south
beyond the seaward reaches of the river.
\gab{140}
The navvy, staggering forward, cleaves the crowd
and lurches towards the tram siding.

\N2:
On the farther side under the railway bridge Bloom appears,
flushed, panting, cramming bread and chocolate into a side pocket.
From Gillen's hairdresser's window, a composite portrait
shows him gallant Nelson's image.
A concave mirror at the side presents to him lovelorn longlost lugubru Booloohoom.
Grave Gladstone sees him level, Bloom for Bloom.
He passes, struck by the stare of truculent Wellington,
but in the convex mirror grin, unstruck,
the bonham eyes and fatchuck cheekchops of jollypoldy, the rixdix doldy.

\N2:
At Antonio Rabaiotti's door, Bloom halts,
\gab{150}
sweated under the bright arclamp.
He disappears.
In a moment, he reappears and hurries on.

\Bloom:
Fish and taters. N.~g. Ah!

\N1:
He disappears into Olhausen's, the porkbutcher's,
under the downcoming rollshutter.
A few moments later, he emerges from under the shutter,
puffing Poldy, blowing Bloohoom.
In each hand, he holds a parcel,
one containing a lukewarm pig's crubeen,
the other, a cold sheep's trotter, sprinkled with wholepepper.
He gasps, standing upright.
\gab{160}
Then bending to one side, he presses a parcel against his ribs and groans.

\Bloom:
Stitch in my side.
Why did I run?

\N1:
He takes breath with care and goes forward slowly towards the lampset siding.
The glow leaps again.

\Bloom:
What is that? A flasher? Searchlight.

\N2:
He stands at Cormack's corner, watching.

\Bloom:
\latin{Aurora borealis} or a steel foundry?
Ah, the brigade, of course.
\gab{170}
South side anyhow.
Big blaze. Might be his house.
Beggar's bush.
We're safe.
\stage{(He hums cheerfully)}
London's burning, London's burning! On fire, on fire!

\N2:
He catches sight of the navvy lurching through the crowd
at the farther side of Talbot street.

\Bloom:
I'll miss him.
Run. Quick. Better cross here.

\N1:
He darts to cross the road.
Urchins shout.

\Urchins:
Mind out, mister!

\N1:
Two cyclists, with lighted paper lanterns aswing,
swim by him, grazing him, their bells rattling.

\Bells:
\gab{180}
Halty-alty-alty-all.

\N2:
Bloom halts erect, stung by a spasm.

\Bloom:
Ow!

\N1:
He looks round, darts forward suddenly.
Through rising fog a dragon sandstrewer, travelling at caution,
slews heavily down upon him, its huge red headlight winking,
its trolley hissing on the wire.
The motorman bangs his footgong.

\Gong:
Bang Bang Bla Bak Blud Bugg Bloo.

\N2:
The brake cracks violently.
\gab{190}
Bloom, raising a policeman's whitegloved hand,
blunders stifflegged out of the track.

\N1:
The motorman, thrown forward, pugnosed, on the guidewheel,
yells as he slides past over chains and keys.

\Motorman:
Hey, shitbreeches, are you doing the hat trick?

\N2:
Bloom trickleaps to the curbstone and halts again.
He brushes a mudflake from his cheek with a parcelled hand.

\Bloom:
No thoroughfare.
Close shave that but cured the stitch.
Must take up Sandow's exercises again.
\gab{200}
On the hands down.
Insure against street accident too.
The Providential.

\N1:
Bloom feels his trouser pocket.

\Bloom:
Poor mamma's panacea.
Heel easily catch in track or bootlace in a cog.
Day the wheel of the black Maria peeled off my shoe at Leonard's corner.
Third time is the charm.
Shoe trick.
Insolent driver.
I ought to report him.
Tension makes them nervous.
Might be the fellow balked me this morning with that horsey woman.
Same style of beauty.
Quick of him all the same.
The stiff walk.
True word spoken in jest.
That awful cramp in Lad lane.
Something poisonous I ate.
Emblem of luck.
Why? Probably lost cattle.
Mark of the beast.

\N1:
Bloom closes his eyes an instant.

\Bloom:
Bit light in the head.
Monthly or effect of the other.
\gab{210}
Brain fog fag.
That tired feeling.
Too much for me now.
Ow!

\N2:
A sinister figure leans on plaited legs against O'Beirne's wall,
a visage unknown, injected with dark mercury.
From under a wideleaved sombrero, the figure regards him with evil eye.

\Bloom:
\spanish{Buenas noches, senorita Blanca.
Que calle es esta?}

\N1:
The figure, impassive, raises a signal arm.

\Figure:
Password. \irish{Sráid Mabbot.}

\Bloom:
Haha.
\gab{220}
\french{Merci.}
Esperanto.
\irish{Slán leath.}
\stage{(he mutters)}
Gaelic league spy, sent by that fire eater.

\N2:
He steps forward.
A sackshouldered ragman bars his path.
He steps left, ragsackman left.

\Bloom:
I beg.

\N2:
He swerves, sidles, stepaside, slips past and on.

\Bloom:
\gab{230}
Keep to the right, right, right.
If there is a signpost planted by the Touring Club at Stepaside
who procured that public boon?
I who lost my way and contributed to the columns of the \Title{Irish Cyclist}
the letter headed \Title{In darkest Stepaside}.
Keep, keep, keep to the right.
Rags and bones at midnight.
A fence more likely.
First place murderer makes for.
Wash off his sins of the world.

\N1:
Jacky Caffrey, hunted by Tommy Caffrey, runs full tilt against Bloom.

\Bloom:
O.
\gab{240}

\N1:
Shocked, on weak hams, he halts.
Tommy and Jacky vanish there, there.
Bloom pats with parcelled hands watch fob, pocket book pocket, purse poke,
\Title{Sweets of Sin}, potato soap.

\Bloom:
Beware of pickpockets.
Old thieves' dodge.
Collide.
Then snatch your purse.

\N2:
The retriever approaches sniffing, nose to the ground.
A sprawled form sneezes.

\N1:
A stooped bearded figure appears, garbed in the long caftan of an elder in Zion
and a smokingcap with magenta tassels.
Horned spectacles hang down at the wings of the nose.
\gab{250}
Yellow poison streaks are on the drawn face of Rudolph Bloom.

\Rudolph:
Second half-crown waste money today.
I told you not go with drunken goy ever.
So you catch no money.

\N2:
Bloom hides the crubeen and trotter behind his back and,
crestfallen, feels warm and cold feetmeat.

\Bloom:
\german{Ja, ich weiss, Papachi.}

\Rudolph:
What you making down this place?
Have you no soul?

\N1:
With feeble vulture talons, he feels the silent face of Bloom.
\gab{260}

\Rudolph:
Are you not my son Leopold, the grandson of Leopold?
Are you not my dear son Leopold, who left the house of his father
and left the god of his fathers, Abraham and Jacob?

\Bloom:
\stage{(with precaution)}
I suppose so, father.
Mosenthal.
All that's left of him.

\Rudolph:
\stage{(severely)}
One night they bring you home drunk as dog after spend your good money.
What you call them running chaps?

\N2:
Bloom, in youth's smart blue Oxford suit with white vestslips,
narrowshouldered, in brown Alpine hat,
\gab{270}
wearing gent's sterling silver Waterbury keyless watch
and double curb Albert with seal attached,
one side of him coated with stiffening mud.

\Bloom:
Harriers, father.
Only that once.

\Rudolph:
Once! Mud head to foot.
Cut your hand open. Lockjaw.
They make you kaputt, Leopoldleben.
You watch them chaps.

\Bloom:
\stage{(weakly)}
They challenged me to a sprint.
It was muddy.
I slipped.

\Rudolph:
\stage{(with contempt)}
\yiddish{Goim nachez!}
Nice spectacles for your poor mother!

\Bloom:
\gab{280}
Mamma!

\N1:
Ellen Bloom, in pantomime dame's stringed mobcap,
widow Twankey's crinoline and bustle,
blouse with muttonleg sleeves buttoned behind,
grey mittens and cameo brooch,
her plaited hair in a crispine net,
appears over the staircase banisters,
a slanted candlestick in her hand,
and cries out in shrill alarm.

\Ellen:
O blessed Redeemer, what have they done to him!
My smelling salts!

\N2:
She hauls up a reef of skirt and ransacks the
pouch of her striped blay petticoat.
A phial, an Agnus Dei,
a shrivelled potato, and a celluloid doll fall out.

\Ellen:
Sacred Heart of Mary, where were you, at all, at all?
\gab{290}

\N1:
Bloom, mumbling, his eyes downcast,
begins to bestow his parcels in his filled pockets, but desists, muttering.

\Marion:
\stage{(sharply)} Poldy!

\Bloom:
Who?

\N1:
He ducks and wards off a blow clumsily.

\Bloom:
At your service.

\N2:
He looks up.
Beside her mirage of datepalms,
a handsome woman in turkish costume stands before him.
Opulent curves fill out her scarlet trousers and jacket, slashed with gold.
A wide yellow cummerbund girdles her.
\gab{300}
A white yashmak, violet in the night, covers her face,
leaving free only her large dark eyes and raven hair.

\Bloom:
Molly!

\Marion[1]:
Welly?
Mrs~Marion from this out, my dear man, when you speak to me.
\stage{(satirically)}
Has poor little hubby cold feet waiting so long?

\N2:
Bloom shifts from foot to foot.

\Bloom:
No, no. Not the least little bit.

\N1:
\gab{310}
He breathes in deep agitation, swallowing gulps of air, questions, hopes,
cru\-beens for her supper, things to tell her, excuse, desire, spellbound.
A coin gleams on her forehead.
On her feet are jewelled toerings.
Her ankles are linked by a slender fetter chain.
Beside her, a camel, hooded with a turreting turban, waits.
A silk ladder of innumerable rungs climbs to his bobbing howdah.
He ambles near with disgruntled hindquarters.
Fiercely, she slaps his haunch, her goldcurb wrist bangles angriling,
scolding him in Moorish.

\Marion:
Nebrakada! Femininum!

\N2:
\gab{320}
The camel, lifting a foreleg, plucks from a tree a large mango fruit,
offers it to his mistress, blinking, in his cloven hoof,
then droops his head and, grunting, with uplifted neck, fumbles to kneel.
Bloom stoops his back for leapfrog.

\Bloom:
I can give you \ldots
I mean as your business menagerer \ldots
Mrs~Marion \ldots
if you \ldots

\Marion:
So you notice some change?

\N1:
Her hands pass slowly over her trinketed stomacher,
a slow friendly mockery in her eyes.

\Marion:
O Poldy, Poldy, you are a poor old stick in the mud!
\gab{330}
Go and see life.
See the wide world.

\Bloom:
I was just going back for that lotion whitewax, orangeflower water.
Shop closes early on Thursday.
But the first thing in the morning.

\N1:
He pats diverse pockets.

\Bloom:
This moving kidney. Ah!

\N2:
He points to the south, then to the east.

\N1:
A cake of new clean lemon soap arises, diffusing light and perfume.

\Soap:
\begin{verse}
    We're a capital couple, are Bloom and I.\\
    He brightens the earth. I polish the sky.
\end{verse}

\N1:
The freckled face of Sweny, the druggist,
\gab{340}
appears in the disc of the soapsun.

\Sweny:
Three and a penny, please.

\Bloom:
Yes. For my wife.
Mrs~Marion. Special recipe.

\Marion:
\stage{(softly)} Poldy!

\Bloom:
Yes, ma'am?

\Marion:
\gab{350}
\italian{Ti trema un poco il cuore?}

\N2:
In disdain she saunters away, plump as a pampered pouter pigeon,
humming the duet from \italian{Don Giovanni}.

\Bloom:
Are you sure about that \italian{Voglio}?
I mean the pronunciash---

\N2:
He follows, followed by the sniffing terrier.

\N1:
The elderly bawd seizes his sleeve,
the bristles of her chinmole glittering.

\Bawd:
Ten shillings a maidenhead.
Fresh thing was never touched. Fifteen.
There's no-one in it,
\gab{360}
only her old father that's dead drunk.

\N1:
She points.
In the gap of her dark den,
furtive, rainbedraggled, Bridie Kelly stands.

\Bridie:
Hatch street.
Any good in your mind?

\N2:
With a squeak, she flaps her bat shawl and runs.
A burly rough pursues with booted strides.
He stumbles on the steps, recovers, plunges into gloom.
Weak squeaks of laughter are heard, weaker.

\N1:
The Bawd, her wolfeyes shining.

\Bawd:
He's getting his pleasure.
You won't get a virgin in the flash houses.
Ten shillings.
\gab{370}
Don't be all night before the polis in plain clothes sees us.
Sixtyseven is a bitch.

\N2:
Leering, Gerty MacDowell limps forward.
She draws from behind, ogling,
and shows coyly her bloodied clout.

\Gerty:
With all my worldly goods I thee and thou.
\stage{(she murmurs)} You did that.
I hate you.

\Bloom:
I? When?
You're dreaming.
I never saw you.

\Bawd:
Leave the gentleman alone, you cheat.
\gab{380}
Writing the gentleman false letters.
Streetwalking and soliciting.
Better for your mother
take the strap to you at the bedpost, hussy like you.

\Gerty:
% \stage{(to Bloom)}
When you saw all the secrets of my bottom drawer.

\N2:
She paws his sleeve, slobbering.

\Gerty:
Dirty married man!
I love you for doing that to me.

\N2:
She glides away crookedly.

\N1:
Mrs~Breen in man's frieze overcoat with loose bellows pockets,
stands in the causeway, her roguish eyes wideopen,
smiling in all her herbivorous buckteeth.

\MrsBreen:
Mr~\ldots
\gab{390}

\Bloom:
\stage{(coughs gravely)}
Madam, when we last had this pleasure by letter dated the sixteenth instant \ldots

\MrsBreen:
Mr~Bloom! You down here in the haunts of sin!
I caught you nicely! Scamp!

\Bloom:
\stage{(hurriedly)}
Not so loud my name.
Whatever do you think of me?
Don't give me away.
Walls have ears.
How do you do? It's ages since I.
You're looking splendid.
\gab{400}
Absolutely it.
Seasonable weather we are having this time of year.
Black refracts heat.
Short cut home here.
Interesting quarter.
Rescue of fallen women.
Magdalen asylum.
I am the secretary \ldots

\MrsBreen:
\stage{(holds up a finger)}
Now, don't tell a big fib!
I know somebody won't like that.
O just wait till I see Molly!
\stage{(slily)}
Account for yourself this very sminute or woe betide you!

\Bloom:
\stage{(looks behind)}
\gab{408}
She often said she'd like to visit.
Slumming.
The exotic, you see.
% Negro servants in livery too if she had money.
% Othello black brute.
% Eugene Stratton.
% \gab{410}
% Even the bones and cornerman at the Livermore christies.
% Bohee brothers.
% Sweep for that matter.
% 
% \stage{(Tom and Sam Bohee, coloured coons in white duck suits, scarlet socks,
% upstarched sambo chokers and large scarlet asters in their buttonholes, leap out.
% Each has his banjo slung.
% Their paler smaller negroid hands jingle the twingtwang wires.
% Flashing white kaffir eyes and tusks
% they rattle through a breakdown in clumsy clogs,
% twinging, singing, back to back, toe heel, heel toe,
% with smackfatclacking nigger lips.)}
% 
% \Bohees:
% \begin{verse}
% \gab{420}%
%     There's someone in the house with Dina\\
%     There's someone in the house, I know,\\
%     There's someone in the house with Dina\\
%     Playing on the old banjo.
% \end{verse}
% 
% \stage{(They whisk black masks from raw babby faces:
% then, chuckling, chortling, trumming, twanging,
% they diddle diddle cakewalk dance away.)}
%
% \Bloom:
\stage{(with a sour tenderish smile)}
\gab{428}
A little frivol, shall we, if you are so inclined?
Would you like me perhaps to embrace you
just for a fraction of a second?

\MrsBreen:
% \gab{430}
\stage{(screams gaily)}
O, you ruck! You ought to see yourself!

\Bloom:
For old sake' sake.
I only meant a square party, a mixed marriage
mingling of our different little conjugials.
You know I had a soft corner for you.
\stage{(gloomily)}
'Twas I sent you that valentine of the dear gazelle.

\MrsBreen:
Glory Alice, you do look a holy show!
Killing simply.
\stage{(she puts out her hand inquisitively)}
What are you hiding behind your back?
Tell us, there's a dear.

\Bloom:
\gab{440}
\stage{(seizes her wrist with his free hand)}
Josie Powell that was, prettiest deb in Dublin.
How time flies by!
Do you remember, harking back
in a retrospective arrangement, Old Christmas night,
Georgina Simpson's housewarming while they were playing the Irving Bishop game,
finding the pin blindfold and thought reading?
Subject, what is in this snuffbox?

\MrsBreen:
You were the lion of the night
with your seriocomic recitation and you looked the part.
You were always a favourite with the ladies.

\N2:
Bloom, squire of dames, in dinner jacket with wateredsilk facings,
\gab{450}
blue masonic badge in his buttonhole, black bow and mother-of-pearl studs,
a prismatic champagne glass tilted in his hand.

\Bloom:
Ladies and gentlemen, I give you Ireland, home and beauty.

\MrsBreen:
The dear dead days beyond recall.
Love's old sweet song.

\Bloom:
\stage{(meaningfully dropping his voice)}
I confess I'm teapot with curiosity to find out
whether some person's something is a little teapot at present.

\MrsBreen:
\stage{(gushingly)}
Tremendously teapot!
\gab{460}
London's teapot and I'm simply teapot all over me!
\stage{(she rubs sides with him)}
After the parlour mystery games and the crackers from the tree
we sat on the staircase ottoman.
Under the mistletoe.
Two is company.

\N2:
Bloom, wearing a purple napoleon hat with an amber halfmoon,
his fingers and thumb passing slowly down to her soft moist meaty palm,
which she surrenders gently.

\Bloom:
The witching hour of night.
I took the splinter out of this hand, carefully, slowly.
\stage{(tenderly, as he slips on her finger a ruby ring)}
\italian{Là ci darem la mano.}

\N1:
Mrs Breen, in a onepiece evening frock executed in moonlight blue,
\gab{470}
a tinsel sylph's diadem on her brow,
with her dancecard fallen beside her moonblue satin slipper,
curves her palm softly, breathing quickly.

\MrsBreen:
\italian{Voglio e non.}
You're hot! You're scalding!
The left hand nearest the heart.

\Bloom:
When you made your present choice
they said it was beauty and the beast.
I can never forgive you for that.
\stage{(his clenched fist at his brow)}
Think what it means.
All you meant to me then.
\stage{(hoarsely)}
Woman, it's breaking me!

\N1:
Denis Breen, white-tall-hatted, with Wisdom Hely's sandwich-boards,
\gab{480}
shuffles past them in carpet slippers, his dull beard thrust out,
muttering to right and left.

\N2:
Little Alf Bergan, cloaked in the pall of the ace of spades,
dogs him to left and right, doubled in laughter.
He points jeering at the sandwich boards.

\AlfBergan:
U. p: Up.

\MrsBreen:
\stage{(to Bloom)}
High jinks below stairs.

\N1:
She gives him the glad eye.

\MrsBreen:
Why didn't you kiss the spot to make it well?
You wanted to.

\Bloom:
\stage{(shocked)}
\gab{490}
Molly's best friend! Could you?

\N1:
Mrs Breen, her pulpy tongue between her lips, offers a pigeon kiss.

\MrsBreen:
Hnhn.
The answer is a lemon.
Have you a little present for me there?

\Bloom:
\stage{(offhandedly)}
Kosher.
A snack for supper.
The home without potted meat is incomplete.
I was at \Title{Leah}.
Mrs~Bandmann Palmer.
Trenchant exponent of Shakespeare.
Unfortunately threw away the programme.
Rattling good place round there for pigs' feet.
Feel.

\N2:
Richie Goulding, three ladies' hats pinned on his head,
appears weighted to one side by the black legal bag of
\gab{500}
Collis and Ward on which a skull and crossbones are painted in white limewash.
He opens it and shows it full of polonies,
kippered herrings, findon haddies and tightpacked pills.

\Richie:
Best value in Dub.

\N1:
Bald Pat, bothered beetle, stands on the curbstone,
folding his napkin, waiting to wait.
He advances with a tilted dish of spillspilling gravy.

\Pat:
Steak and kidney.
Bottle of lager.
Hee hee hee.
\gab{510}
Wait till I wait.

\Richie[2]:
Goodgod.
Inev erate inall \ldots

\N2:
With hanging head he marches doggedly forward.

\N1:
The navvy, lurching by, gores him with his flaming pronghorn.

\N2:
Richie Goulding, with a cry of pain, his hand to his back.

\Richie:
Ah! Bright's! Lights!

\N1:
Bloom points to the navvy.

\Bloom:
A spy.
Don't attract attention.
I hate stupid crowds.
I am not on pleasure bent.
I am in a grave predicament.

\MrsBreen:
\gab{520}
Humbugging and deluthering as per usual with your cock and bull story.

\Bloom:
I want to tell you a little secret about how I came to be here.
But you must never tell.
Not even Molly.
I have a most particular reason.

\MrsBreen:
\stage{(all agog)}
O, not for worlds.

\Bloom:
Let's walk on.
Shall us?

\MrsBreen:
Let's.
\gab{530}

\N1:
The bawd makes an unheeded sign.

\N2:
Bloom walks on with Mrs~Breen.
The terrier follows, whining piteously, wagging his tail.

\Bawd[1]:
Jewman's melt!

\N2:
Bloom, in an oatmeal sporting suit, a sprig of woodbine in the lapel,
tony buff shirt, shepherd's plaid Saint Andrew's cross scarftie,
white spats, fawn dustcoat on his arm, tawny red brogues,
fieldglasses in bandolier and a grey billycock hat.

\Bloom:
Do you remember a long long time,
years and years ago, just after Milly,
Marionette we called her,
\gab{540}
was weaned when we all went together to Fairyhouse races, was it?

\N1:
Mrs Breen, in smart saxe tailormade, white velours hat and spider veil.

\MrsBreen:
Leopardstown.

\Bloom:
I mean, Leopardstown.
And Molly won seven shillings on a three year old named Nevertell
and coming home along by Foxrock
in that old fiveseater shanderadan of a waggonette
you were in your heyday then and
you had on that new hat of white velours with a surround of molefur that Mrs~Hayes
\gab{550}
advised you to buy because it was marked down to nineteen and eleven,
a bit of wire and an old rag of velveteen,
and I'll lay you what you like she did it on purpose \ldots

\MrsBreen:
She did, of course, the cat!
Don't tell me! Nice adviser!

\Bloom:
Because it didn't suit you one quarter as well as
the other ducky little tammy toque with the bird of paradise wing in it
that I admired on you and you honestly looked just too fetching in it
though it was a pity to kill it, you cruel naughty creature,
little mite of a thing with a heart the size of a
fullstop.%
\gab{560}

\MrsBreen:
\stage{(squeezes his arm, simpers)}
Naughty cruel I was!

\Bloom:
\stage{(low, secretly, ever more rapidly)}
And Molly was eating a sandwich of spiced beef
out of Mrs~Joe Gallaher's lunch basket.
Frankly, though she had her advisers or admirers,
I never cared much for her style.
She was \ldots

\MrsBreen:
Too \ldots

\Bloom:
Yes.
\gab{570}
And Molly was laughing because Rogers and Maggot O'Reilly
were mimicking a cock as we passed a farmhouse and
Marcus Tertius Moses, the tea merchant,
drove past us in a gig with his daughter,
Dancer Moses was her name,
and the poodle in her lap bridled up
and you asked me if I ever heard or read or knew or came across \ldots

\MrsBreen:
\stage{(eagerly)}
Yes, yes, yes, yes, yes, yes, yes.

\N1:
She fades from his side.

\N2:
Followed by the whining dog, he walks on towards hells gates.
In an archway, a standing woman, bent forward, her feet apart, pisses cowily.
Outside a shuttered pub a bunch of loiterers
\gab{580}
listen to a tale which their broken-snouted gaffer rasps out with raucous humour.
An armless pair of them flop wrestling, growling, in maimed sodden playfight.

\N1:
The gaffer crouches, his voice twisted in his snout.

\Gaffer:
And when Cairns came down from the scaffolding in Beaver street,
what was he after doing it into,
only into the bucket of porter that was there,
waiting on the shavings for Derwan's plasterers.

\N2:
The loiterers guffaw with cleft palates.

\Loiterers:
O jays!

\N2:
Their paintspeckled hats wag.
\gab{590}
Spattered with size and lime of their lodges
they frisk limblessly about him.

\Bloom:
Coincidence too.
They think it funny.
Anything but that.
Broad daylight.
Trying to walk.
Lucky no woman.

\Loiterers:
Jays, that's a good one.
Glauber salts.
O jays, into the men's porter.

\N2:
Bloom passes.

\N1:
Cheap whores, singly, coupled, shawled, dishevelled,
call from lanes, doors, corners.

\Whores:
\begin{verse}
\gab{600}%
    Are you going far, queer fellow?\\
    How's your middle leg?\\
    Got a match on you?\\
    Eh, come here till I stiffen it for you.
\end{verse}

\N1:
He plodges through their sump towards the lighted street beyond.
From a bulge of window curtains, a gramophone rears a battered brazen trunk.
In the shadow, a shebeen keeper haggles with the navvy and the two redcoats.

\N2:
The navvy asks.

\Navvy:
\stage{(belching)}
Where's the bloody house?

\ShebeenKeeper[1]:
\gab{610}
Purdon street.
Shilling a bottle of stout.
Respectable woman.

\N2:
The navvy, gripping the two redcoats, staggers forward with them.

\Navvy:
Come on, you British army!

\N1:
Private Carr, behind his back.

\Carr:
He aint half balmy.

\N2:
Private Compton laughs.

\Compton:
\stage{(laughs)}
What ho!

\Carr:
\stage{(to the Navvy)}
\gab{620}
Portobello barracks canteen.
You ask for Carr. Just Carr.

\Navvy:
\stage{(shouts)}
    We are the boys. Of Wexford.

\Compton:
Say! What price the sergeantmajor?

\Carr:
Bennett? He's my pal.
I love old Bennett.

\Navvy:
\stage{(shouts)}
\begin{verse}
\gab{630}%
    The galling chain.\\
    And free our native land.
\end{verse}

\N2:
The navvy staggers forward, dragging them with him.

\N1:
Bloom stops, at fault.
The dog approaches, his tongue outlolling, panting.

\Bloom:
Wildgoose chase this.
Disorderly houses.
Lord knows where they are gone.
Drunks cover distance double quick.
Nice mixup.
Scene at Westland row.
Then jump in first class with third ticket.
Then too far.
Train with engine behind.
Might have taken me to Malahide or a siding for the night or collision.
Second drink does it.
Once is a dose.
What am I following him for?
Still, he's the best of that lot.
\gab{640}
If I hadn't heard about Mrs~Beaufoy Purefoy,
I wouldn't have gone and wouldn't have met.
Kismet.
He'll lose that cash.
Relieving office here.
Good biz for cheapjacks, organs.
What do ye lack? Soon got, soon gone.
Might have lost my life too
with that man-gong-wheel-track-trolley-glare-juggernaut
only for presence of mind.
Can't always save you, though.
If I had passed Truelock's window that day,
two minutes later would have been shot.
Absence of body.
Still, if bullet only went through my coat
get damages for shock, five hundred pounds.
What was he?
Kildare street club toff.
God help his gamekeeper.

\N1:
He gazes ahead, reading on the wall
a scrawled chalk legend \Title{Wet Dream} and a phallic design.

\Bloom:
\gab{650}%
Odd! Molly drawing on the frosted carriagepane at Kingstown.
What's that like?

\N2:
Gaudy dollwomen loll in the lighted doorways,
in window embrasures, smoking birdseye cigarettes.
The odour of the sicksweet weed floats towards him
in slow round ovalling wreaths.

\Wreaths:
Sweet are the sweets.
Sweets of sin.

\Bloom:
My spine's a bit limp.
Go or turn?
And this food? Eat it and get all pigsticky.
Absurd I am.
Waste of money.
One and eightpence too much.

\N1:
The retriever drives a cold snivelling muzzle against his hand, wagging his tail.

\Bloom:
Strange how they take to me.
Even that brute today.
\gab{660}
Better speak to him first.
Like women, they like \french{rencontres}.
Stinks like a polecat.
\french{Chacun son goût.}
He might be mad.
Dogdays.
Uncertain in his movements.
Good fellow! Fido!
Good fellow! Garryowen!

\N1:
The wolfdog sprawls on his back, wriggling obscenely with begging paws,
his long black tongue lolling out.

\Bloom:
Influence of his surroundings.
Give and have done with it.
Provided nobody.

\N1:
Calling encouraging words, he shambles back with a furtive poacher's tread,
dogged by the setter, into a dark stalestunk corner.
He unrolls one parcel and goes to dump the crubeen softly
but holds back and feels the trotter.

\Bloom:
Sizeable for threepence.
But then I have it in my left hand.
Calls for more effort.
\gab{670}
Why? Smaller from want of use.
O, let it slide.
Two and six.

\N1:
With regret, he lets the unrolled crubeen and trotter slide.
The mastiff mauls the bundle clumsily and gluts himself
with growling greed, crunching the bones.

\N2:
Two raincaped watch approach, silent, vigilant.
They murmur together.

\Watch:
Bloom.
Of Bloom.
For Bloom.
Bloom.

\N2:
Each lays hand on Bloom's shoulder.

\FirstWatch[1]:
Caught in the act.
\gab{680}
Commit no nuisance.

\Bloom:
\stage{(stammers)}
I am doing good to others.

\N1:
A covey of gulls, storm petrels, rises hungrily from Liffey slime
with Banbury cakes in their beaks.

\Gulls:
Kaw kave kankury kake.

\Bloom:
The friend of man.
Trained by kindness.

\N2:
He points.

\N1:
Bob Doran, toppling from a high barstool, sways over the munching spaniel.%
\gab{690}

\Doran:
Towser.
Give us the paw.
Give the paw.

\N1:
The bulldog growls, his scruff standing, a gobbet of pig's knuckle
between his molars through which rabid scumspittle dribbles.
Bob Doran falls silently into an area.

\SecondWatch[2]:
Prevention of cruelty to animals.

\Bloom:
\stage{(enthusiastically)}
A noble work!
I scolded that tramdriver on Harold's cross bridge
for ill-using the poor horse with his harness scab.
\gab{700}
Bad French I got for my pains.
Of course it was frosty and the last tram.
All tales of circus life are highly demoralising.

\N1:
Signor Maffei, passionpale,
in liontamer's costume with diamond studs in his shirtfront,
steps forward, holding a circus paperhoop,
a curling carriagewhip, and a revolver,
with which he covers the gorging boarhound.

\Maffei:
\stage{(with a sinister smile)}
Ladies and gentlemen, my educated greyhound.
It was I broke in the bucking broncho Ajax
with my patent spiked saddle for carnivores.
Lash under the belly with a knotted thong.
\gab{710}
Block tackle and a strangling pulley will bring your lion to heel,
no matter how fractious, even \latin{leo ferox} there, the Libyan maneater.
A redhot crowbar and some liniment rubbing on the burning part
produced Fritz of Amsterdam, the thinking hyena.
\stage{(he glares)}
I possess the Indian sign.
The glint of my eye does it with these breastsparklers.
\stage{(with a bewitching smile)}
I now introduce Mademoiselle Ruby, the pride of the ring.

\FirstWatch[1]:
Come.
Name and address.

\Bloom:
I have forgotten for the moment.
Ah, yes!
\gab{720}

\N2:
He takes off his high grade hat, saluting.

\Bloom:
Dr Bloom, Leopold, dental surgeon.
You have heard of von Blum Pasha.
Umpteen millions.
\german{Donnerwetter!}
Owns half Austria. Egypt. Cousin.

\FirstWatch:
Proof.

\N2:
A card falls from inside the leather headband of Bloom's hat.

\N1:
Bloom, in red fez, cadi's dress coat with broad green sash,
wearing a false badge of the Legion of Honour,
picks up the card hastily and offers it.

\Bloom:
Allow me.
My club is the Junior Army and Navy.
\gab{730}
Solicitors:
Messrs John Henry Menton, 27 Bachelor's Walk.

\FirstWatch:
\stage{(reads)}
Henry Flower.
No fixed abode.
Unlawfully watching and besetting.

\SecondWatch[2]:
An alibi.
You are cautioned.

\N1:
Bloom produces from his heartpocket a crumpled yellow flower.

\Bloom:
This is the flower in question.
It was given me by a man I don't know his name.
\stage{(plausibly)}
\gab{740}
You know that old joke, rose of Castile.
Bloom.
The change of name.
Virag.
\stage{(he murmurs privately and confidentially)}
We are engaged you see, sergeant.
Lady in the case.
Love entanglement.
\stage{(he shoulders the second watch gently)}
Dash it all.
It's a way we gallants have in the navy.
Uniform that does it.
\stage{(he turns gravely to the first watch)}
Still, of course, you do get your Waterloo sometimes.
Drop in some evening and have a glass of old Burgundy.
\stage{(to the second watch gaily)}
I'll introduce you, inspector.
She's game.
Do it in the shake of a lamb's tail.

\N2:
A dark mercurialised face appears, leading a veiled figure.

\DarkMercury:
The Castle is looking for him.
\gab{750}
He was drummed out of the army.

\N2:
Martha, thickveiled, a crimson halter round her neck,
a copy of the \Title{Irish Times} in her hand, in tone of reproach, pointing.

\Martha:
Henry! Leopold!
Lionel, thou lost one!
Clear my name.

\FirstWatch[1]:
\stage{(sternly)} Come to the station.

\N2:
Bloom, scared, hats himself, steps back,
then, plucking at his heart and lifting his right forearm on the square,
he gives the sign and dueguard of fellowcraft.

\Bloom:
No, no, worshipful master, light of love.
\gab{760}
Mistaken identity.
The Lyons mail.
Lesurques and Dubosc.
You remember the Childs fratricide case.
We medical men.
By striking him dead with a hatchet.
I am wrongfully accused.
Better one guilty escape than ninetynine wrongfully condemned.

\Martha:
\stage{(sobbing behind her veil)}
Breach of promise.
My real name is Peggy Griffin.
He wrote to me that he was miserable.
I'll tell my brother, the Bective rugger fullback, on you, heartless flirt.

\Bloom:
\stage{(behind his hand)}
She's drunk.
The woman is inebriated.

\N1:
He murmurs vaguely the pass of Ephraim.
\gab{770}

\Bloom:
Shitbroleeth.

\N2:
The Second Watch, tears in his eyes, says to Bloom.

\SecondWatch:
You ought to be thoroughly well ashamed of yourself.

\Bloom:
Gentlemen of the jury, let me explain.
A pure mare's nest.
I am a man misunderstood.
I am being made a scapegoat of.
I am a respectable married man, without a stain on my character.
I live in Eccles street.
My wife, I am the daughter of a most distinguished commander,
a gallant upstanding gentleman,
what do you call him, Majorgeneral Brian Tweedy,
one of Britain's fighting men who helped to win our battles.
\gab{780}
Got his majority for the heroic defence of Rorke's Drift.

\FirstWatch:
Regiment.

\N1:
Bloom turns to the gallery.

\Bloom:
The royal Dublins, boys, the salt of the earth, known the world over.
I think I see some old comrades in arms up there among you.
The R.D.F., with our own Metropolitan police,
guardians of our homes, the pluckiest lads and the finest body of men,
as physique, in the service of our sovereign.

\Voice:
\gab{790}
Turncoat! Up the Boers! Who booed Joe Chamberlain?

\Bloom:
\stage{(his hand on the shoulder of the first watch)}
My old dad too was a J.P.
I'm as staunch a Britisher as you are, sir.
I fought with the colours for king and country
in the absentminded war under general Gough in the park
and was disabled at Spion Kop and Bloemfontein,
was mentioned in dispatches.
I did all a white man could.
\stage{(with quiet feeling)}
Jim Bludso.
Hold her nozzle again the bank.

\FirstWatch:
Profession or trade.
\gab{800}

\Bloom:
Well, I follow a literary occupation, author-journalist.
In fact we are just bringing out a collection of prize stories
of which I am the inventor, something that is an entirely new departure.
I am connected with the British and Irish press.
If you ring up \ldots

\N2:
Myles Crawford strides out jerkily, a quill between his teeth.
His scarlet beak blazes within the aureole of his straw hat.
He dangles a hank of Spanish onions in one hand and holds with the other hand
a telephone receiver nozzle to his ear.

\Myles:
\gab{810}
% \stage{(his cock's wattles wagging)}
Hello, seventyseven eightfour. Hello.
\Title{Freeman's Urinal} and \Title{Weekly Arsewipe} here.
Paralyse Europe.
You which? Bluebags?
Who writes? Is it Bloom?

\N1:
Mr~Philip Beaufoy, palefaced, stands in the witnessbox,
in accurate morning dress, outbreast pocket with peak of handkerchief showing,
creased lavender trousers and patent boots.
He carries a large portfolio labelled \Title{Matcham's Masterstrokes}.

\Beaufoy:
\stage{(drawls})
No, you aren't.
Not by a long shot if I know it.
I don't see it that's all.
\gab{820}
No born gentleman, no-one with the most rudimentary promptings of a gentleman
would stoop to such particularly loathsome conduct.
One of those, my lord.
A plagiarist.
A soapy sneak masquerading as a litterateur.
It's perfectly obvious that with the most inherent baseness
he has cribbed some of my bestselling copy, really gorgeous stuff,
a perfect gem, the love passages in which are beneath suspicion.
The Beaufoy books of love and great possessions,
with which your lordship is doubtless familiar,
are a household word throughout the kingdom.

\N2:
Bloom murmurs with hangdog meekness glum.

\Bloom:
That bit about the laughing witch
\gab{830}
hand in hand I take exception to, if I may \ldots

\N1:
Beaufoy, his lip upcurled, smiles superciliously on the court.

\Beaufoy:
You funny ass, you!
You're too beastly awfully weird for words!
I don't think you need over excessively disincommodate yourself in that regard.
My literary agent Mr~J.B. Pinker is in attendance.
I presume, my lord, we shall receive the usual witnesses' fees, shan't we?
We are considerably out of pocket over this bally pressman johnny,
this jackdaw of Rheims, who has not even been to a university.

\Bloom:
\stage{(indistinctly)}
\gab{840}
University of life. Bad art.

\Beaufoy:
\stage{(shouts)}
It's a damnably foul lie, showing the moral rottenness of the man!
% \stage{(he extends his portfolio)}
We have here damning evidence, the \latin{corpus delicti}, my lord,
a specimen of my maturer work disfigured by the hallmark of the beast.

\GalleryVoice:
\begin{verse}
    Moses, Moses, king of the jews,\\
    Wiped his arse in the \Title{Daily News}.
\end{verse}

\Bloom:
\stage{(bravely)} Overdrawn.
\gab{850}

\Beaufoy:
You low cad!
You ought to be ducked in the horsepond, you rotter!
\stage{(to the court)}
Why, look at the man's private life!
Leading a quadruple existence!
Street angel and house devil.
Not fit to be mentioned in mixed society!
The archconspirator of the age!

\Bloom:
\stage{(to the court)}
And he, a bachelor, how \ldots

\FirstWatch:
The King versus Bloom.
Call the woman Driscoll.

\Crier:
\gab{860}
Mary Driscoll, scullerymaid!

\N1:
Mary Driscoll, a slipshod servant girl, approaches.
She has a bucket on the crook of her arm and a scouringbrush in her hand.

\SecondWatch:
Another!
Are you of the unfortunate class?

\Driscoll[1]:
\stage{(indignantly)}
I'm not a bad one.
I bear a respectable character and was four months in my last place.
I was in a situation, six pounds a year and my chances with Fridays out
and I had to leave owing to his carryings on.

\FirstWatch:
\gab{870}
What do you tax him with?

\Driscoll:
He made a certain suggestion
but I thought more of myself as poor as I am.

\N2:
Bloom, in housejacket of ripplecloth, flannel trousers,
heelless slippers, unshaven, his hair rumpled.

\Bloom:
\stage{(softly)}
I treated you white.
I gave you mementos, smart emerald garters far above your station.
Incautiously I took your part when you were accused of pilfering.
There's a medium in all things.
Play cricket.

\Driscoll:
\stage{(excitedly)}
\gab{880}
As God is looking down on me this night if ever I laid a hand to them oysters!

\FirstWatch:
The offence complained of?
Did something happen?

\Driscoll:
He surprised me in the rere of the premises, Your honour,
when the missus was out shopping one morning with a request for a safety pin.
He held me and I was discoloured in four places as a result.
And he interfered twict with my clothing.

\Bloom:
She counterassaulted.
\gab{890}

\Driscoll:
\stage{(scornfully)}
I had more respect for the scouringbrush, so I had.
I remonstrated with him, Your lord, and he remarked:
keep it quiet.

\stage{(General laughter.)}

\N1:
George Fottrell, clerk of the crown and peace.

\Fottrell:
\stage{(resonantly)}
Order in court!
The accused will now make a bogus statement.

\N2:
Bloom, pleading not guilty and holding a fullblown waterlily,
begins a long unintelligible speech.
They would hear what counsel had to say
in his stirring address to the grand jury.
\gab{900}
He was down and out but, though branded as a black sheep,
if he might say so, he meant to reform,
to retrieve the memory of the past in a purely sisterly way
and return to nature as a purely domestic animal.
A seven-months' child, he had been carefully brought up
and nurtured by an aged bedridden parent.
There might have been lapses of an erring father
but he wanted to turn over a new leaf
and now, when at long last in sight of the whipping post,
to lead a homely life in the evening of his days,
permeated by the affectionate surroundings of the heaving bosom of the family.
An acclimatised Britisher, he had seen that summer eve
\gab{910}
from the footplate of an engine cab of the Loop line railway company
while the rain refrained from falling glimpses, as it were,
through the windows of loveful households in Dublin city and urban district
of scenes truly rural of happiness of the better land
with Dockrell's wallpaper at one and ninepence a dozen,
innocent British-born bairns lisping prayers to the Sacred Infant,
youthful scholars grappling with their pensums
or model young ladies playing on the pianoforte
or anon all with fervour reciting the family rosary
round the crackling Yulelog
while in the boreens and green lanes
the colleens with their swains strolled
what times the strains of the organ-toned melodeon
\gab{920}
Britannia metal-bound with four acting stops and twelvefold bellows,
a sacrifice, greatest bargain ever\ldots

\N1:
Renewed laughter.
He mumbles incoherently.
Reporters complain that they cannot hear.

\LongShortHand:
% \stage{(without looking up from their notebooks)}
Loosen his boots.

\MacHugh:
% \stage{(from the presstable, coughs and calls)}
Cough it up, man.
Get it out in bits.

\N1:
The crossexamination proceeds re Bloom and the bucket.
A large bucket.
Bloom himself.
Bowel trouble.
\gab{930}
In Beaver street.
Gripe, yes.
Quite bad.
A plasterer's bucket.
By walking stifflegged.
Suffered untold misery.
Deadly agony.
About noon.
Love or burgundy.
Yes, some spinach.
Crucial moment.
He did not look in the bucket nobody.
Rather a mess.
Not completely.
A \Title{Titbits} back number.

\N2:
Uproar and catcalls.
Bloom in a torn frockcoat stained with whitewash,
dinged silk hat sideways on his head,
a strip of stickingplaster across his nose, talks inaudibly.

\N1:
J.J.~O'Molloy, in barrister's grey wig and stuffgown,
speaks with a voice of pained protest.

\JJOM:
\gab{940}
This is no place for indecent levity at the
expense of an erring mortal disguised in liquor.
We are not in a beargarden nor at an Oxford rag
nor is this a travesty of justice.
My client is an infant, a poor foreign immigrant
who started scratch as a stowaway and is now trying to turn an honest penny.
The trumped up misdemeanour was due to a momentary aberration of heredity,
brought on by hallucination, such familiarities as the alleged guilty occurrence
being quite permitted in my client's native place, the land of the Pharaoh.
\latin{Prima facie}, I put it to you that there was no attempt at carnally knowing.
\gab{948}
Intimacy did not occur and the offence complained of by Driscoll,
that her virtue was solicited, was not repeated.
%I would deal in especial with atavism.
%\gab{950}
% There have been cases of shipwreck and somnambulism in my client's family.
% If the accused could speak he could a tale unfold—one
% of the strangest that have ever been narrated between the covers of a book.
% He himself, my lord, is a physical wreck from cobbler's weak chest.
% His submission is that he is of Mongolian extraction
% and irresponsible for his actions.
% Not all there, in fact.
% 
% \Bloom:
% \stage{(Barefoot, pigeonbreasted, in lascar's vest and trousers,
% apologetic toes turned in, opens his tiny mole's eyes and looks about him dazedly,
% passing a slow hand across his forehead.
% Then he hitches his belt sailor fashion
% and with a shrug of oriental obeisance salutes the court,
% \gab{960}
% pointing one thumb heavenward.)}
% Him makee velly muchee fine night.
% \stage{(he begins to lilt simply)}
% \begin{verse}
%     Li li poo lil chile\\
%     Blingee pigfoot evly night\\
%     Payee two shilly \ldots
% \end{verse}
% 
% \stage{(he is howled down.)}
% 
% \JJOM:
% \stage{(hotly to the populace)}
% This is a lonehand fight.
% By Hades, I will not have any client of mine gagged and badgered
% in this fashion by a pack of curs and laughing hyenas.
% The Mosaic code has superseded the law of the jungle.
\gab{970}
I say it and I say it emphatically,
without wishing for one moment to defeat the ends of justice,
accused was not accessory before the act and prosecutrix has not been tampered with.
The young person was treated by defendant as if she were his very own daughter.
\stage{(Bloom takes J.J.~O'Molloy's hand and raises it to his lips.)}
I shall call rebutting evidence to prove up to the hilt
that the hidden hand is again at its old game.
When in doubt persecute Bloom.
My client, an innately bashful man, would be the last man in the world
to do anything ungentlemanly which injured modesty could object to
or cast a stone at a girl who took the wrong turning when some dastard,
responsible for her condition, had worked his own sweet will on her.
\gab{980}
He wants to go straight.
I regard him as the whitest man I know.
He is down on his luck at present owing to the mortgaging of his extensive property
at Agendath Netaim in faraway Asia Minor, slides of which will now be shown.
\stage{(to Bloom)}
I suggest that you will do the handsome thing.

\Bloom:
A penny in the pound.

\N1:
The image of the lake of Kinnereth with blurred cattle cropping in silver haze
is projected on the wall.

\N2:
Moses Dlugacz, ferreteyed albino, in blue dungarees, stands up in the gallery,
holding in each hand an orange citron and a pork kidney.

\Dlugacz:
\gab{990}
\stage{(hoarsely)}
Bleibtreustrasse, Berlin, W.13.

\N1:
J.J.~O'Molloy steps on to a low plinth
and holds the lapel of his coat with solemnity.
His face lengthens, grows pale and bearded, with sunken eyes,
the blotches of phthisis and hectic cheekbones of John~F. Taylor.
He applies his handkerchief to his mouth
and scrutinises the galloping tide of rosepink blood.

\JJOM:
\stage{(almost voicelessly)}
Excuse me.
I am suffering from a severe chill, have recently come from a sickbed.
A few wellchosen words.

\N1:
He assumes the avine head, foxy moustache
\gab{1000}
and proboscidal eloquence of Seymour Bushe.

\JJOM:
When the angel's book comes to be opened
if aught that the pensive bosom has inaugurated of
soultransfigured and of soultransfiguring deserves to live
I say accord the prisoner at the bar the sacred benefit of the doubt.

\N2:
A paper with something written on it is handed into court.

\Bloom:
% \stage{(in court dress)}
Can give best references.
Messrs Callan, Coleman.
Mr~Wisdom Hely~J.P.
My old chief Joe Cuffe.
Mr~V.B.~Dillon, ex lord mayor of Dublin.
I have moved in the charmed circle of the highest \ldots
Queens of Dublin society.
\stage{(carelessly)}
\gab{1010}
I was just chatting this afternoon
at the viceregal lodge to my old pals, sir Robert and lady Ball,
astronomer royal at the levee.
Sir Bob, I said \ldots

\N1:
Mrs Yelverton Barry, in lowcorsaged opal balldress and elbowlength ivory gloves,
wearing a sabletrimmed brickquilted dolman, a comb of brilliants
and panache of osprey in her hair.

\Yelverton:
Arrest him, constable.
He wrote me an anonymous letter in prentice backhand
when my husband was in the North Riding of Tipperary on the Munster circuit,
signed James Lovebirch.
He said that he had seen from the gods my peerless globes as I sat in a box
of the \Title{Theatre Royal} at a command performance of \french{La Cigale}.
\gab{1020}
I deeply inflamed him, he said.
He made improper overtures to me to misconduct
myself at half past four p.m.
on the following Thursday, Dunsink time.
He offered to send me through the post a work of fiction by Monsieur Paul de
Kock, entitled \Title{The Girl with the Three Pairs of Stays}.

\N2:
Mrs Bellingham, in cap and seal coney mantle, wrapped up to the nose,
steps out of her brougham and scans through tortoiseshell quizzing-glasses
which she takes from inside her huge opossum muff.

\Bellingham:
Also to me.
Yes, I believe it is the same objectionable person.
Because he closed my carriage door outside sir Thornley Stoker's
\gab{1030}
one sleety day during the cold snap of February ninetythree
when even the grid of the wastepipe
and the ballstop in my bath cistern were frozen.
Subsequently he enclosed a bloom of edelweiss culled on the heights,
as he said, in my honour.
I had it examined by a botanical expert
and elicited the information that it was
a blossom of the homegrown potato plant
purloined from a forcingcase of the model farm.

\Yelverton:
Shame on him!

\N1:
A crowd of sluts and ragamuffins surges forward.

\Ragamuffins:
\stage{(screaming)}
\gab{1040}
Stop thief!
Hurrah there, Bluebeard!
Three cheers for Ikey Mo!

\N2:
The Second Watch produces handcuffs.

\SecondWatch:
Here are the darbies.

\Bellingham:
He addressed me in several handwritings with fulsome compliments
as a Venus in furs and alleged profound pity for my frostbound coachman Palmer
while in the same breath he expressed himself
as envious of his earflaps and fleecy sheepskins
and of his fortunate proximity to my person,
when standing behind my chair wearing my livery
and the armorial bearings of the Bellingham escutcheon garnished sable,
\gab{1050}
a buck's head couped or.
He lauded almost extravagantly my nether extremities,
my swelling calves in silk hose drawn up to the limit,
and eulogised glowingly my other hidden treasures in priceless lace
which, he said, he could conjure up.
He urged me (stating that he felt it his mission in life to urge me)
to defile the marriage bed, to commit adultery at the earliest possible opportunity.

\N1:
The Honourable Mrs Mervyn Talboys, in amazon costume, hard hat,
jackboots cockspurred, vermilion waistcoat,
fawn musketeer gauntlets with braided drums, long train held up
and hunting crop with which she strikes her welt constantly.

\Talboys:
Also me.
\gab{1060}
Because he saw me on the polo ground of the Phoenix park
at the match All Ireland versus the Rest of Ireland.
My eyes, I know, shone divinely as I watched Captain Slogger Dennehy
of the Inniskillings win the final chukkar on his darling cob \Title{Centaur}.
This plebeian Don Juan observed me from behind a hackney car
and sent me in double envelopes an obscene photograph,
such as are sold after dark on Paris boulevards, insulting to any lady.
I have it still.
It represents a partially nude senorita, frail and lovely
(his wife, as he solemnly assured me, taken by him from nature),
practising illicit intercourse with a muscular torero, evidently a blackguard.
He urged me to do likewise, to misbehave, to sin with officers of the garrison.
\gab{1070}
He implored me to soil his letter in an unspeakable manner,
to chastise him as he richly deserves, to bestride and ride him,
to give him a most vicious horsewhipping.

\Bellingham:
Me too.

\Yelverton:
Me too.

\N2:
Several highly respectable Dublin ladies hold up improper letters
received from Bloom.

\N1:
The Honourable Mrs Mervyn Talboys stamps her jingling spurs
in a sudden paroxysm of fury.

\Talboys:
\gab{1080}
I will, by the God above me.
I'll scourge the pigeonlivered cur as long as I can stand over him.
I'll flay him alive.

\N2:
Bloom, his eyes closing, quails expectantly.

\Bloom:
Here?

\N2:
He squirms.

\Bloom:
Again!

\N2:
He pants cringing.

\Bloom:
I love the danger.

\Talboys:
Very much so!
I'll make it hot for you.
I'll make you dance Jack Latten for that.

\Bellingham:
\gab{1090}
Tan his breech well, the upstart!
Write the stars and stripes on it!

\Yelverton:
Disgraceful!
There's no excuse for him!
A married man!

\Bloom:
All these people.
I meant only the spanking idea.
A warm tingling glow without effusion.
Refined birching to stimulate the circulation.

\Talboys:
\stage{(laughs derisively)}
O, did you, my fine fellow?
Well, by the living God, you'll get the surprise of your life now,
believe me, the most unmerciful hiding a man ever bargained for.
\gab{1100}
You have lashed the dormant tigress in my nature into fury.

\N1:
Mrs Bellingham shakes her muff and quizzing-glasses vindictively.

\Bellingham:
Make him smart, Hanna dear.
Give him ginger.
Thrash the mongrel within an inch of his life.
The cat-o'-nine-tails.
Geld him.
Vivisect him.

\N2:
Bloom, shuddering, shrinking, joins his hands: with hangdog mien.

\Bloom:
O cold! O shivery!
It was your ambrosial beauty.
Forget, forgive. Kismet.
Let me off this once.

\N2:
He offers the other cheek.

\Yelverton:
\gab{1110}
\stage{(severely)}
Don't do so on any account, Mrs~Talboys!
He should be soundly trounced!

\N1:
The Honourable Mrs Mervyn Talboys unbuttons her gauntlet violently.

\Talboys:
I'll do no such thing.
Pigdog and always was ever since he was pupped!
To dare address me!
I'll flog him black and blue in the public streets.
I'll dig my spurs in him up to the rowel.
He is a wellknown cuckold.
\stage{(she swishes her huntingcrop savagely in the air)}
Take down his trousers without loss of time.
Come here, sir! Quick! Ready?

\N2:
Bloom trembles, beginning to obey.

\Bloom:
\gab{1120}
The weather has been so warm.

\N1:
Davy Stephens, ringletted, passes with a bevy of barefoot newsboys.

\DavyStephens:
\Title{Messenger of the Sacred Heart}
and \Title{Evening Telegraph} with Saint Patrick's Day supplement.
Containing the new addresses of all the cuckolds in Dublin.

\N2:
The very reverend Canon O'Hanlon in cloth of gold cope
elevates and exposes a marble timepiece.
Before him Father Conroy and the Reverend John Hughes S.J. bend low.%
\gab{1130}

\Timepiece:
\stage{(unportalling)}
\begin{verse}
    Cuckoo.\\
    Cuckoo.\\
    Cuckoo.
\end{verse}

\N1:
The brass quoits of a bed are heard to jingle.

\Quoits:
Jigjag. Jigajiga. Jigjag.

\N2:
A panel of fog rolls back rapidly, revealing rapidly in the jurybox
the faces of Martin Cunningham, foreman, silk-hatted;
\gab{1140}
Jack Power; Simon Dedalus; Tom Kernan; Ned Lambert;
John Henry Menton; Myles Crawford;
Lenehan; Paddy Leonard; Nosey Flynn; M'Coy;
and the featureless face of a Nameless One.

\NamelessOne:
Bareback riding.
Weight for age.
Gob, he organised her.

\N2:
The jurors turn their heads to his voice.

\Jurors:
% \stage{(all their heads turned to his voice)}
Really?

\NamelessOne:
\stage{(snarls)}
Arse over tip.
Hundred shillings to five.

\Jurors:
\gab{1150}
% \stage{(all their heads lowered in assent)}
Most of us thought as much.

\FirstWatch[1]:
He is a marked man.
Another girl's plait cut.
Wanted:
Jack the Ripper.
A thousand pounds reward.

\N2:
The Second Watch, awed, whispers.

\SecondWatch:
And in black.
A Mormon.
Anarchist.

\Crier:
\stage{(loudly)}
Whereas Leopold Bloom of no fixed abode is a well-known dynamitard,
forger, bigamist, bawd, and cuckold,
and a public nuisance to the citizens of Dublin,
\gab{1160}
and whereas at this commission of assizes, the most honourable \ldots

\N1:
His Honour, sir Frederick Falkiner, recorder of Dublin,
in judicial garb of grey stone rises from the bench, stonebearded.
He bears in his arms an umbrella sceptre.
From his forehead arise starkly the Mosaic ramshorns.

\Recorder:
I will put an end to this white slave traffic
and rid Dublin of this odious pest.
Scandalous!

\N1:
He dons the black cap.

\Recorder:
Let him be taken, Mr~Subsheriff, from the dock where he now stands
and detained in custody in Mountjoy prison during His Majesty's pleasure
\gab{1170}
and there be hanged by the neck until he is dead
and therein fail not at your peril or may the Lord have mercy on your soul.
Remove him.

\N1:
A black skullcap descends upon his head.

\N2:
The subsheriff Long John Fanning appears, smoking a pungent Henry Clay.
He scowls and calls with rich rolling utterance.

\LongJohn:
Who'll hang Judas Iscariot?

\N1:
H. Rumbold, master barber, in a bloodcoloured jerkin and tanner's apron,
a rope coiled over his shoulder, mounts the block.
A life preserver and a nailstudded bludgeon are stuck in his belt.
He rubs grimly his grappling hands,
\gab{1180}
knobbed with knuckledusters.
He addresses the recorder with sinister familiarity.

\Rumbold:
Hanging Harry, your Majesty, the Mersey terror.
Five guineas a jugular.
Neck or nothing.

\N2:
The bells of George's church toll slowly, loud dark iron.

\Bells:
Heigho! Heigho!

\Bloom:
\stage{(desperately)}
Wait. Stop.
Gulls.
Good heart.
I saw.
Innocence.
Girl in the monkeyhouse.
Zoo.
Lewd chimpanzee.
\stage{(breathlessly)}
Pelvic basin.
Her artless blush unmanned me.
\stage{(overcome with emotion)}
\gab{1190}
I left the precincts.
\stage{(he turns to a figure in the crowd, appealing)}
Hynes, may I speak to you?
You know me.
That three shillings you can keep.
If you want a little more \ldots

\Hynes[1]:
\stage{(coldly)}
You are a perfect stranger.

\N2:
The Second Watch points to the corner.

\SecondWatch:
The bomb is here.

\FirstWatch:
Infernal machine with a time fuse.

\Bloom:
\gab{1200}
No, no.
Pig's feet.
I was at a funeral.

\N1:
The First Watch draws his truncheon.

\FirstWatch:
Liar!

\N2:
The beagle lifts his snout, showing the grey scorbutic face of Paddy Dignam.
He has gnawed all.
He exhales a putrid carcase-fed breath.
He grows to human size and shape.
His dachshund coat becomes a brown mortuary habit.
His green eye flashes bloodshot.
Half of one ear, all the nose and both thumbs are ghoul-eaten.

\N1:
Paddy Dignam says in a hollow voice.

\Dignam:
It is true.
It was my funeral.
\gab{1210}
Doctor Finucane pronounced life extinct
when I succumbed to the disease from natural causes.

\N2:
He lifts his mutilated ashen face moonwards and bays lugubriously.

\Bloom:
\stage{(in triumph)}
You hear?

\Dignam:
Bloom, I am Paddy Dignam's spirit.
List, list, O list!

\Bloom:
The voice is the voice of Esau.
\gab{1220}

\N2:
The Second Watch blesses himself.

\SecondWatch:
How is that possible?

\FirstWatch:
It is not in the penny catechism.

\Dignam:
By metempsychosis. Spooks.

\Voice:
O rocks.

\Dignam:
\stage{(earnestly)}
\gab{1230}
Once I was in the employ of Mr~J.H.~Menton,
solicitor, commissioner for oaths and affidavits, of 27~Bachelor's Walk.
Now I am defunct, the wall of the heart hypertrophied.
Hard lines.
The poor wife was awfully cut up.
How is she bearing it?
Keep her off that bottle of sherry.
\stage{(he looks round him)}
A lamp.
I must satisfy an animal need.
That buttermilk didn't agree with me.

\N1:
The portly figure of John O'Connell, caretaker, stands forth,
holding a bunch of keys tied with crape.

\N2:
Beside him stands Father Coffey, chaplain,
toadbellied, wrynecked, in a surplice and bandanna nightcap,
holding sleepily a staff of twisted poppies.

\FrCoffey:
\gab{1240}
\stage{(yawns, then chants with a hoarse croak)}
Namine.
Jacobs.
Vobiscuits.
Amen.

\N1:
John O'Connell foghorns stormily through his megaphone.

\OConnell:
Dignam, Patrick~T, deceased.

\N2:
Paddy Dignam with pricked up ears, winces.

\Dignam:
Overtones.

\N2:
He wriggles forward and places an ear to the ground.

\Dignam:
My master's voice!

\OConnell:
Burial docket letter number U.P. eightyfive thousand.
Field seventeen.
House of Keys.
\gab{1250}
Plot, one hundred and one.

\N2:
Paddy Dignam listens with visible effort, thinking,
his tail stiffpointed, his ears cocked.

\Dignam:
Pray for the repose of his soul.

\N2:
He worms down through a coalhole,
his brown habit trailing its tether over rattling pebbles.
After him toddles an obese grandfather rat
on fungus turtle paws under a grey carapace.
Dignam's voice, muffled, is heard baying under ground:
\emph{Dignam's dead and gone below.}

\N1:
Tom Rochford, robin-red-breasted, in cap and breeches,
\gab{1260}%
jumps from his two-columned machine.
With a hand to his breastbone, he bows.

\Rochford:
Reuben J.
A florin I find him.

\N1:
He fixes the manhole with a resolute stare.

\Rochford:
My turn now on.
Follow me up to Carlow.

\N1:
He executes a daredevil salmon leap in the air and is engulfed in the coalhole.
Two discs on the columns wobble, eyes of nought.
All recedes.

\N2:
Bloom plodges forward again through the sump.
Kisses chirp amid the rifts of fog.
A piano sounds.
He stands before a lighted house, listening.

\N1:
The kisses, winging from their bowers fly about him,
\gab{1270}
twittering, warbling, cooing.

\Kisses:
\stage{(warbling)} Leo!
\stage{(twittering)} Icky licky micky sticky for Leo!
\stage{(cooing)} Coo coocoo!
Yummyyum, Womwom!
\stage{(warbling)} Big comebig!
Pirouette!
Leopopold!
\stage{(twittering)} Leeolee!
\stage{(warbling)} O Leo!

\N1:
They rustle, flutter upon his garments,
alight, bright giddy flecks, silvery sequins.

\Bloom:
A man's touch.
Sad music.
Church music.
Perhaps here.

\N2:
Zoe Higgins, a young whore in a sapphire slip,
closed with three bronze buckles, a slim black velvet fillet round her throat,
\gab{1280}
nods, trips down the steps and accosts him.

\Zoe:
Are you looking for someone?
He's inside with his friend.

\Bloom:
Is this Mrs~Mack's?

\Zoe:
No, eightyone.
Mrs~Cohen's.
You might go farther and fare worse.
Mother Slipperslapper.
\stage{(familiarly)}
She's on the job herself tonight with the vet
her tipster that gives her all the winners and pays for her son in Oxford.
Working overtime but her luck's turned today.
\gab{1290}
\stage{(suspiciously)}
You're not his father, are you?

\Bloom:
Not I!

\Zoe:
You both in black.
Has little mousey any tickles tonight?

\N1:
His skin, alert, feels her fingertips approach.
A hand glides over his left thigh.

\Zoe:
How's the nuts?

\Bloom:
\gab{1300}
Off side.
Curiously they are on the right.
Heavier, I suppose.
One in a million my tailor, Mesias, says.

\Zoe:
\stage{(in sudden alarm)}
You've a hard chancre.

\Bloom:
Not likely.

\Zoe:
I feel it.

\N2:
Her hand slides into his left trouser pocket
and brings out a hard black shrivelled potato.
\gab{1310}
She regards it and Bloom with dumb moist lips.

\Bloom:
A talisman.
Heirloom.

\Zoe:
For Zoe?
For keeps?
For being so nice, eh?

\N2:
She puts the potato greedily into a pocket then links his arm,
cuddling him with supple warmth.
He smiles uneasily.
Slowly, note by note, oriental music is played.
He gazes in the tawny crystal of her eyes, ringed with kohol.
His smile softens.

\Zoe:
\gab{1320}
You'll know me the next time.

\Bloom:
\stage{(forlornly)}
I never loved a dear gazelle but it was sure to \ldots

\N1:
Gazelles are leaping, feeding on the mountains.
Near are lakes.
Round their shores file shadows black of cedargroves.
Aroma rises, a strong hairgrowth of resin.
It burns, the orient, a sky of sapphire, cleft by the bronze flight of eagles.
Under it lies the womancity nude, white, still, cool, in luxury.
A fountain murmurs among damask roses.
Mammoth roses murmur of scarlet winegrapes.
A wine of shame, lust, blood exudes, strangely murmuring.%
\gab{1330}

\N2:
Murmuring singsong with the music, her odalisk lips lusciously
smeared with salve of swinefat and rosewater.

\Zoe:
\hebrew{Schorach ani wenowach, benoith Hierushaloim.}

\Bloom:
\stage{(fascinated)}
I thought you were of good stock by your accent.

\Zoe:
And you know what thought did?

\N2:
She bites his ear gently with little goldstopped teeth,
sending on him a cloying breath of stale garlic.
\gab{1340}
The roses draw apart, disclose a sepulchre of the gold of kings
and their mouldering bones.

\N1:
Bloom draws back, mechanically caressing her right bub with a flat awkward hand.

\Bloom:
Are you a Dublin girl?

\N2:
Zoe catches a stray hair deftly and twists it to her coil.

\Zoe:
No bloody fear.
I'm English.
Have you a swaggerroot?

\Bloom:
% \stage{(as before)}
Rarely smoke, dear.
Cigar now and then.
Childish device.
\stage{(lewdly)}
\gab{1350}
The mouth can be better engaged than with a cylinder of rank weed.

\Zoe:
Go on.
Make a stump speech out of it.

\stage{Zoe freezes.}

\par

\N1:
Bloom, in workman's corduroy overalls,
black gansy, with red floating tie and apache cap.

\Bloom:
Mankind is incorrigible.
Sir Walter Ralegh brought from the new world that potato and that weed,
the one a killer of pestilence by absorption,
the other a poisoner of the ear, eye, heart, memory, will understanding, all.
That is to say he brought the poison a hundred years
before another person whose name I forget brought the food.
\gab{1360}
Suicide. Lies. All our habits.
Why, look at our public life!

\N2:
Midnight chimes from distant steeples.

\Chimes:
Turn again, Leopold!
Lord mayor of Dublin!

\N1:
Bloom, in alderman's gown and chain.

\Bloom:
Electors of Arran Quay, Inns Quay, Rotunda, Mountjoy and North Dock,
better run a tramline, I say, from the cattlemarket to the river.
That's the music of the future.
That's my programme.
\latin{Cui bono?}
But our bucaneering Vanderdeckens in their phantom ship of finance \ldots%
\gab{1370}

\Elector:
Three times three for our future chief magistrate!

\N2:
The Aurora Borealis of the torchlight procession leaps.

\Torchbearers:
Hooray!

\N2:
Several wellknown burgesses, city magnates and freemen of the city
shake hands with Bloom and congratulate him.
Timothy Harrington, late thrice Lord Mayor of Dublin,
imposing in mayoral scarlet, gold chain and white silk tie,
confers with councillor Lorcan Sherlock, \latin{locum tenens}.
\gab{1380}
They nod vigorously in agreement.

\N1:
Late Lord Mayor Harrington, in scarlet robe with mace, 
gold mayoral chain and large white silk scarf.

\Harrington:
That alderman sir Leo Bloom's speech be printed at the expense of the ratepayers.
That the house in which he was born be ornamented with a commemorative tablet
and that the thoroughfare hitherto known as Cow Parlour off Cork street
be henceforth designated Boulevard Bloom.

\Lorcan[2]:
Carried unanimously.

\Bloom:
\stage{(impassionedly)}
\gab{1390}
These flying Dutchmen or lying Dutchmen as they recline
in their upholstered poop, casting dice, what reck they?
Machines is their cry, their chimera, their panacea.
Laboursaving apparatuses, supplanters, bugbears,
manufactured monsters for mutual murder,
hideous hobgoblins produced by
a horde of capitalistic lusts upon our prostituted labour.
The poor man starves while they are grassing their royal mountain stags
or shooting peasants and phartridges in their purblind pomp of pelf and power.
But their reign is rover for rever and ever and ev \ldots

\N1:
Prolonged applause.
Venetian masts, maypoles and festal arches spring up.
A streamer bearing the legends \irish{Céad Mile Fáilte} and
\hebrew{Mah Ttob Melek Israel} spans the street.
\gab{1400}
All the windows are thronged with sightseers, chiefly ladies.
Along the route the regiments of the Royal Dublin Fusiliers,
the King's Own Scottish Borderers, the Cameron Highlanders
and the Welsh Fusiliers standing to attention, keep back the crowd.
Boys from High school are perched on the lampposts, telegraph poles,
windowsills, cornices, gutters, chimneypots, railings, rainspouts,
whistling and cheering.

\N2:
The pillar of the cloud appears.
A fife and drum band is heard in the distance playing the Kol Nidre.
The beaters approach with imperial eagles hoisted,
trailing banners and waving oriental palms.
\gab{1410}
The chryselephantine papal standard rises high,
surrounded by pennons of the civic flag.

\N1:
The van of the procession appears headed by John Howard Parnell, city marshal,
in a chessboard tabard, the Athlone poursuivant and Ulster King of Arms.
They are followed by the Right Honourable Joseph Hutchinson,
lord mayor of Dublin, his lordship the lord mayor of Cork,
their worships the mayors of Limerick, Galway, Sligo and Waterford,
twentyeight Irish representative peers, sirdars, grandees and maharajahs
bearing the cloth of estate, the Dublin Metropolitan Fire Brigade,

\N2:
the chapter of the saints of finance in their plutocratic order of precedence,
the bishop of Down and Connor,
\gab{1420}

\N1:
His Eminence Michael cardinal Logue,
archbishop of Armagh, primate of all Ireland,

\N2:
His Grace, the most reverend Dr William Alexander,
archbishop of Armagh, primate of all Ireland,

\N1:
the Chief Rabbi, the Presbyterian moderator,
the heads of the Baptist, Anabaptist, Methodist and Moravian chapels
and the honorary secretary of the Society of Friends.

\N2:
After them, march the guilds and trades and trainbands with flying colours:
coopers, bird fanciers, millwrights, newspaper canvassers, law scriveners,

\N1:
masseurs, vintners, trussmakers, chimneysweeps, lard refiners,
tabinet and poplin weavers, farriers, Italian warehousemen,
\gab{1430}

\N2:
church decorators, bootjack manufacturers, undertakers,
silk mercers, lapidaries, salesmasters, corkcutters,
assessors of fire losses, dyers and cleaners, export bottlers,

\N1:
fellmongers, ticketwriters, heraldic seal engravers,
horse repository hands, bullion brokers, cricket and archery outfitters,
riddlemakers, egg and potato factors, hosiers and glovers, plumbing contractors.

\N2:
After them march gentlemen of the bedchamber,
Black Rod, Deputy Garter, Gold Stick, the master of horse,
the lord great chamberlain, the earl marshal,
the high constable carrying the sword of state,
saint Stephen's iron crown, the chalice and bible.
\gab{1440}
Four buglers on foot blow a sennet.
Beefeaters reply, winding clarions of welcome.

\N1:
Under an arch of triumph Bloom appears, bareheaded,
in a crimson velvet mantle trimmed with ermine,
bearing Saint Edward's staff the orb and sceptre with the dove, the curtana.
He is seated on a milkwhite horse with long flowing crimson tail,
richly caparisoned, with golden headstall.
Wild excitement.

\N2:
The ladies from their balconies throw down rosepetals.
The air is perfumed with essences.
The men cheer.

\N1:
Bloom's boys run amid the bystanders with branches of hawthorn and wrenbushes.

\BloomsBoys:
\gab{1450}
\begin{verse}
    The wren, the wren,\\
    The king of all birds,\\
    Saint Stephen's his day\\
    Was caught in the furze.
\end{verse}

\N2:
A blacksmith murmurs.

\Blacksmith:
For the honour of God!
And is that Bloom?
He scarcely looks thirtyone.

\Flagger[1]:
That's the famous Bloom now, the world's greatest reformer.
Hats off!

\N2:
All uncover their heads.
\gab{1460}
Women whisper eagerly.

\N1:
A millionairess, richly.

\Millionairess:
Isn't he simply wonderful?

\N1:
A noblewoman, nobly.

\Noblewoman:
All that man has seen!

\N1:
A feminist, masculinely.

\Feminist:
And done!

\Bellhanger[2]:
A classic face!
He has the forehead of a thinker.

\N1:
Bloom's weather.
A sunburst appears in the northwest.

\DownConnor[2]:
\gab{1470}
I here present your undoubted emperor-president and king-chairman,
the most serene and potent and very puissant ruler of this realm.
God save Leopold the First!

\All:
God save Leopold the First!

\N1:
Bloom, in dalmatic and purple mantle,
to the bishop of Down and Connor, with dignity.

\Bloom:
Thanks, somewhat eminent sir.

\N2:
William, Archbishop of Armagh, in purple stock and shovel hat.

\WillArmagh:
\gab{1480}
Will you to your power cause law and mercy to be executed
in all your judgments in Ireland and territories thereunto belonging?

\N1:
Bloom, placing his right hand on his testicles, swears.

\Bloom:
So may the Creator deal with me.
All this I promise to do.

\N2:
Michael, Archbishop of Armagh, pours a cruse of hairoil over Bloom's head.

\MikeArmagh:
\latin{Gaud\-ium magnum annuntio vobis.
Habemus carneficem.}
Leopold, Patrick, Andrew, David, George, be thou anointed!

\N1:
\gab{1490}
Bloom assumes a mantle of cloth of gold and puts on a ruby ring.
He ascends and stands on the stone of destiny.
The representative peers put on at the same time their twentyeight crowns.
Joybells ring in Christ church, Saint Patrick's, George's and gay Malahide.
Mirus bazaar fireworks go up from all sides with symbolical phallopyrotechnic designs.

\N2:
The peers do homage, one by one, approaching and genuflecting.

\Peers:
I do become your liege man of life and limb to earthly worship.

\N1:
Bloom holds up his right hand on which sparkles the Koh-i-Noor diamond.
\gab{1500}
His palfrey neighs.
Immediate silence.
Wireless intercontinental and interplanetary transmitters are set for reception of message.

\Bloom:
My subjects!
We hereby nominate our faithful charger Copula Felix hereditary Grand Vizier
and announce that we have this day repudiated our former spouse
and have bestowed our royal hand upon the princess Selene,
the splendour of night.

\N1:
The former morganatic spouse of Bloom is hastily removed in the black maria.
The princess Selene, in moonblue robes, a silver crescent on her head,
\gab{1510}
descends from a Sedan chair, borne by two giants.
An outburst of cheering.

\N2:
John Howard Parnell raises the royal standard.

\Parnell:
Illustrious Bloom!
Successor to my famous brother!

\N1:
Bloom embraces John Howard Parnell.

\Bloom:
We thank you from our heart, John,
for this right royal welcome to green Erin,
the promised land of our common ancestors.

\N1:
The freedom of the city is presented to him embodied in a charter.
The keys of Dublin, crossed on a crimson cushion, are given to him.%
\gab{1520}
He shows all that he is wearing green socks.

\Kernan[2]:
You deserve it, your honour.

\Bloom:
On this day twenty years ago we overcame the hereditary enemy at Ladysmith.
Our howitzers and camel swivel guns played on his lines with telling effect.
Half a league onward!
They charge!
All is lost now!
Do we yield?
No!
We drive them headlong!
Lo! We charge!
Deploying to the left our light horse swept across the heights of Plevna
and, uttering their warcry \latin{Bonafide Sabaoth},
\gab{1530}
sabred the Saracen gunners to a man.

\Typesetters[1]:
Hear! Hear!

\JohnWyse[2]:
There's the man that got away James Stephens.

\Bluecoat[1]:
Bravo!

\OldResident[2]:
You're a credit to your country, sir, that's what you are.

\AppleWoman[1]:
He's a man like Ireland wants.
\gab{1540}

\Bloom:
My beloved subjects, a new era is about to dawn.
I, Bloom, tell you verily it is even now at hand.
Yea, on the word of a Bloom, ye shall ere long
enter into the golden city which is to be, the new Bloomusalem
in the Nova Hibernia of the future.

\N2:
Thirtytwo workmen, wearing rosettes, from all the counties of Ireland,
under the guidance of Derwan the builder, construct the new Bloomusalem.
It is a colossal edifice with crystal roof,
built in the shape of a huge pork kidney, containing forty thousand rooms.
In the course of its extension
\gab{1550}
several buildings and monuments are demolished.
Government offices are temporarily transferred to railway sheds.
Numerous houses are razed to the ground.
The inhabitants are lodged in barrels and boxes,
all marked in red with the letters: L.B.
Several paupers fall from a ladder.

\N1:
A part of the walls of Dublin, crowded with loyal sightseers, collapses.

\Sightseers:
\stage{(dying)}
\latin{Morituri te salutant.}

% \stage{(they die)}

\N1:
The sightseers die.

\N2:
A man in a brown macintosh springs up through a trapdoor.
He points an elongated finger at Bloom.

\Macintosh:
\gab{1560}
Don't you believe a word he says.
That man is Leopold M'Intosh, the notorious fireraiser.
His real name is Higgins.

\Bloom:
Shoot him!
Dog of a christian!
So much for M'Intosh!

\N1:
A cannonshot.
The man in the macintosh disappears.
Bloom with his sceptre strikes down poppies.
The instantaneous deaths of many powerful enemies,
graziers, members of parliament, members of standing committees, are reported.

\N2:
Bloom's bodyguard distribute maundy money, commemoration medals,
loaves and fishes, temperance badges, expensive Henry Clay cigars,
\gab{1570}
free cowbones for soup, rubber preservatives in sealed envelopes
tied with gold thread, butter scotch, pineapple rock,
\french{billets doux} in the form of cocked hats,
readymade suits, porringers of toad in the hole,

\N1:
bottles of Jeyes' Fluid, purchase stamps, 40 days' indulgences,
spurious coins, dairyfed pork sausages, theatre passes,
season tickets available for all tramlines,
coupons of the royal and privileged Hungarian lottery, penny dinner counters,

\N2:
cheap reprints of the World's Twelve Worst Books:

\N1:
Froggy and Fritz (politic),

\N2:
Care of the Baby (infantilic),

\N1:
50 Meals for 7/6 (culinic),

\N2:
Was Jesus a Sun Myth? (historic),
\gab{1580}

\N1:
Expel That Pain (medic),

\N2:
Infant's Compendium of the Universe (cosmic),

\N1:
Let's All Chortle (hilaric),

\N2:
Canvasser's Vade Mecum (journalic),

\N1:
Loveletters of Mother Assistant (erotic),

\N2:
Who's Who in Space (astric),

\N1:
Songs that Reached Our Heart (melodic),

\N2:
Pennywise's Way to Wealth (parsimonic).

\N1:
A general rush and scramble.
Women press forward to touch the hem of Bloom's robe.
The lady Gwendolen Dubedat bursts through the throng, leaps on his horse
and kisses him on both cheeks amid great acclamation.
A magnesium flashlight photograph is taken.
Babes and sucklings are held up.

\Women[2]:
\gab{1590}
Little father!
Little father!

\Sucklings[1]:
\begin{verse}
    Clap clap hands till Poldy comes home,\\
    Cakes in his pocket for Leo alone.
\end{verse}

\N2:
Bloom, bending down, pokes Baby Boardman gently in the stomach.
Baby Boardman hiccups, curdled milk flowing from his mouth.

\BabyBoardman:
Hajajaja.

\N1:
Bloom, shaking hands with a blind stripling:

\Bloom:
\gab{1600}
My more than Brother!

\N1:
Placing his arms round the shoulders of an old couple:

\Bloom:
Dear old friends!

\N1:
He plays pussy fourcorners with ragged boys and girls:

\Bloom:
Peep! Bopeep!

\N1:
He wheels twins in a perambulator:

\Bloom:
Ticktacktwo wouldyousetashoe?

\N1:
He performs juggler's tricks, draws red, orange, yellow,
green, blue, indigo and violet silk handkerchiefs from his mouth:

\Bloom:
Roygbiv.
32 feet per second.

\N1:
He consoles a widow:

\Bloom:
Absence makes the heart grow younger.

\N1:
He dances the highland fling with grotesque antics:

\Bloom:
Leg it, ye devils!

\N1:
He kisses the bedsores of a palsied veteran:

\Bloom:
Honourable wounds!

\N1:
He trips up a fit policeman:

\Bloom:
U.P: up.
U.P: up.

\N1:
He whispers in the ear of a blushing waitress and laughs kindly:

\Bloom:
Ah, naughty, naughty!
\gab{1610}

\N1:
He eats a raw turnip offered him by Maurice Butterly, farmer:

\Bloom:
Fine! Splendid!

\N1:
He refuses to accept three shillings offered him by Joseph Hynes, journalist:

\Bloom:
My dear fellow, not at all!

\N1:
He gives his coat to a beggar:

\Bloom:
Please accept.

\N1:
He takes part in a stomach race with elderly male and female cripples:

\Bloom:
Come on, boys!
Wriggle it, girls!

\N2:
The Citizen, choked with emotion, brushes aside a tear in his emerald muffler.

\Citizen:
May the good God bless him!

\N2:
The rams' horns sound for silence.
The standard of Zion is hoisted.%
\gab{1620}

\N1:
Bloom uncloaks impressively, revealing obesity, unrolls a paper and reads solemnly.

\Bloom:
\hebrew{Aleph, Beth, Ghimel, Daleth, Hagadah, Tephilim, Kosher, Yom Kippur,
Hanukah, Roschaschana, Beni Brith, Bar Mitzvah, Mazzoth, Askenazim,
Meshuggah, Talith.}

\N2:
An official translation is read by Jimmy Henry, assistant town clerk.

\JimmyHenry:
The Court of Conscience is now open.
His Most Catholic Majesty will now administer open air justice.
\gab{1630}
Free medical and legal advice, solution of doubles and other problems.
All cordially invited.
Given at this our loyal city of Dublin in the year~I of the Paradisiacal Era.

\PaddyLeonard[1]:
What am I to do about my rates and taxes?

\Bloom:
Pay them, my friend.

\PaddyLeonard[1]:
Thank you.

\NoseyFlynn[2]:
Can I raise a mortgage on my fire insurance?%
\gab{1640}

\Bloom:
\stage{(obdurately)}
Sirs, take notice that by the law of torts you are bound over
in your own recognisances for six months in the sum of five pounds.

\JJOM[1]:
A Daniel did I say?
Nay! A Peter O'Brien!

\NoseyFlynn[2]:
Where do I draw the five pounds?

% \PisserBurke:
% For bladder trouble?
% 
% \Bloom:
% \gab{1650}
% \begin{verse}
% \latin{Acid. nit. hydrochlor. dil.}, 20 minims\\
% \latin{Tinct. nux vom.}, 5 minims\\
% \latin{Extr. taraxel. iiq.}, 30 minims.\\
% \latin{Aq. dis. ter in die.}
% \end{verse}

\ChrisCallinan[1]:
What is the parallax of the subsolar ecliptic of Aldebaran?

\Bloom:
Pleased to hear from you, Chris. K.~II.

\JoeHynes[2]:
Why aren't you in uniform?%
\gab{1660}

\Bloom:
When my progenitor of sainted memory
wore the uniform of the Austrian despot in a dank prison,
where was yours?

\BenDollard[1]:
Pansies?

\Bloom:
Embellish (beautify) suburban gardens.

\BenDollard[1]:
When twins arrive?

\Bloom:
\gab{1670}
Father (pater, dad) starts thinking.

\LarryORourke[2]:
An eightday licence for my new premises.
You remember me, sir Leo, when you were in number seven.
I'm sending around a dozen of stout for the missus.

\Bloom:
\stage{(coldly)}
You have the advantage of me.
Lady Bloom accepts no presents.

\Crofton[1]:
This is indeed a festivity.

\Bloom:
\gab{1680}
\stage{(solemnly)}
You call it a festivity.
I call it a sacrament.

\AlexanderKeyes[2]:
When will we have our own house of keys?

\Bloom:
I stand for the reform of municipal morals and the plain Ten Commandments.
New worlds for old.
Union of all, Jew, Moslem and Gentile.
Three acres and a cow for all children of nature.
Saloon motor hearses.
Compulsory manual labour for all.
All parks open to the public day and night.
Electric dishscrubbers.
Tuberculosis, lunacy, war and mendicancy must now cease.
\gab{1690}
General amnesty, weekly carnival with masked licence,
bonuses for all, esperanto the universal language with universal brotherhood.
No more patriotism of barspongers and dropsical impostors.
Free money, free rent, free love and a free lay church in a free lay state.

\OMaddenBurke[1]:
Free fox in a free henroost.

\DavyByrne[2]:
\stage{(yawning)} Iiiiiiiiiaaaaaaach!

\Bloom:
Mixed races and mixed marriage.

\Lenehan[1]:
\gab{1700}
What about mixed bathing?

\N2:
Bloom explains to those near him his schemes for social regeneration.
All agree with him.
The keeper of the Kildare street museum appears,
dragging a lorry on which are the shaking statues of several naked goddesses,
Venus Callipyge, Venus Pandemos, Venus Metempsychosis,
and plaster figures, also naked, representing the new nine muses,
Commerce, Operatic Music, Amor, Publicity, Manufacture, Liberty of Speech,
Plural Voting, Gastronomy, Private Hygiene, Seaside Concert Entertainments,
Painless Obstetrics
\gab{1710}
and Astronomy for the People.

\FatherFarley[1]:
He is an episcopalian, an agnostic, an anythingarian
seeking to overthrow our holy faith.

\N2:
Mrs Riordan tears up her will.

\MrsRiordan:
I'm disappointed in you! You bad man!

\N1:
Mother Grogan removes her boot to throw it at Bloom.

\MotherGrogan:
You beast!
You abominable person!

\NoseyFlynn[2]:
Give us a tune, Bloom.
\gab{1720}
One of the old sweet songs.

\Bloom:
\stage{(with rollicking humour)}
\begin{verse}
    I vowed that I never would leave her,\\
    She turned out a cruel deceiver.\\
    With my tooraloom tooraloom tooraloom tooraloom.
\end{verse}

\HoppyHolohan[1]:
Good old Bloom!
There's nobody like him after all.

\PaddyLeonard[2]:
Stage Irishman!

\Bloom:
\gab{1730}
What railway opera is like a tramline in Gibraltar?
The Rows of Casteele.

\stage{(Laughter.)}

\Lenehan[1]:
Plagiarist!
Down with Bloom!

\VeiledSibyl[2]:
\stage{(enthusiastically)}
I'm a Bloomite and I glory in it.
I believe in him in spite of all.
I'd give my life for him, the funniest man on earth.

\N1:
Bloom winks at the bystanders.

\Bloom:
I bet she's a bonny lassie.

\N2:
Theodore Purefoy, in fishingcap and oilskin jacket.

\TheodorePurefoy:
\gab{1740}
He employs a mechanical device to frustrate the sacred ends of nature.

\N1:
The Veiled Sibyl stabs herself.

\VeiledSibyl:
My hero god!

\N1:
She dies.

\N2:
Many most attractive and enthusiastic women also commit suicide
by stabbing, drowning, drinking prussic acid, aconite, arsenic,
opening their veins, refusing food, casting themselves under steamrollers,
from the top of Nelson's Pillar, into the great vat of Guinness's brewery,
asphyxiating themselves by placing their heads in gasovens,
hanging themselves in stylish garters,
\gab{1750}
leaping from windows of different storeys.

\Dowie[1]:
\stage{(violently)}
Fellowchristians and antiBloomites,
the man called Bloom is from the roots of hell,
a disgrace to christian men.
A fiendish libertine from his earliest years
this stinking goat of Mendes
gave precocious signs of infantile debauchery,
recalling the cities of the plain, with a dissolute granddam.
This vile hypocrite, bronzed with infamy,
is the white bull mentioned in the Apocalypse.
A worshipper of the Scarlet Woman,
intrigue is the very breath of his nostrils.
The stake faggots and the caldron of boiling oil are for him.
\gab{1760}
Caliban!

\Mob[2]:
Lynch him!
Roast him!
He's as bad as Parnell was.
Mr~Fox!

\N1:
Mother Grogan throws her boot at Bloom.
Several shopkeepers from upper and lower Dorset street
throw objects of little or no commercial value,
hambones, condensed milk tins, unsaleable cabbage, stale bread,
sheep's tails, odd pieces of fat.

\Bloom:
\stage{(excitedly)}
This is midsummer madness, some ghastly joke again.
By heaven, I am guiltless as the unsunned snow!
It was my brother Henry.
He is my double.
\gab{1770}
He lives in number~2, Dolphin's Barn.
Slander, the viper, has wrongfully accused me.
Fellowcountrymen, \irish{sgeul i mbarr bata coisde gan capall}.
I call on my old friend, Dr Malachi Mulligan, sex specialist,
to give medical testimony on my behalf.

\N2:
Dr Mulligan in motor jerkin, green motorgoggles on his brow.

\DrMulligan:
Dr Bloom is bisexually abnormal.
He has recently escaped from Dr Eustace's
private asylum for demented gentlemen.
Born out of bedlock, hereditary epilepsy is present,
the consequence of unbridled lust.
Traces of elephantiasis have been discovered among his ascendants.
There are marked symptoms of chronic exhibitionism.
\gab{1780}
Ambidexterity is also latent.
He is prematurely bald from selfabuse,
perversely idealistic in consequence,
a reformed rake, and has metal teeth.
In consequence of a family complex,
he has temporarily lost his memory,
and I believe him to be more sinned against than sinning.
I have made a pervaginal examination
and, after application of the acid test to
5427 anal, axillary, pectoral and pubic hairs,
I declare him to be \latin{virgo intacta}.

\N1:
Bloom holds his high grade hat over his genital organs.

\DrMadden[2]:
Hypsospadia is also marked.
In the interest of coming generations,
I suggest that the parts affected should be preserved
\gab{1790}
in spirits of wine in the national teratological museum.

\DrCrotthers[2]:
I have examined the patient's urine.
It is albuminoid.
Salivation is insufficient, the patellar reflex, intermittent.

\DrPunchCostello[2]:
The \latin{fetor judaicus} is most perceptible.

\N2:
Dr Dixon reads a bill of health.

\DrDixon:
Professor Bloom is a finished example of the new womanly man.
His moral nature is simple and lovable.
Many have found him a dear man, a dear person.
\gab{1800}
He is a rather quaint fellow on the whole,
coy though not feebleminded in the medical sense.
He has written a really beautiful letter, a poem in itself,
to the court missionary of the Reformed Priests' Protection Society
which clears up everything.
He is practically a total abstainer
and I can affirm that he sleeps on a straw litter
and eats the most Spartan food, cold dried grocer's peas.
He wears a hairshirt of pure Irish manufacture, winter and summer,
and scourges himself every Saturday.
He was, I understand, at one time
a firstclass misdemeanant in Glencree reformatory.
Another report states that he was a very posthumous child.
I appeal for clemency in the name of
the most sacred word our vocal organs have ever been called upon to speak.
\gab{1810}
He is about to have a baby.

\N1:
General commotion and compassion.
Women faint.
A wealthy American makes a street collection for Bloom.
Gold and silver coins, blank cheques,
banknotes, jewels, treasury bonds, maturing bills of exchange, I.O.U's,
wedding rings, watchchains, lockets,
necklaces and bracelets are rapidly collected.

\Bloom:
O, I so want to be a mother.

\N2:
Mrs Thornton, in nursetender's gown.

\MrsThornton:
Embrace me tight, dear.
You'll be soon over it.
Tight, dear.%
\gab{1820}

\N1:
Bloom embraces her tightly and bears eight male yellow and white children.
They appear on a redcarpeted staircase adorned with expensive plants.
All the octuplets are handsome, with valuable metallic faces,
wellmade, respectably dressed and wellconducted,
speaking five modern languages fluently
and interested in various arts and sciences.
Each has his name printed in legible letters on his shirtfront:
Nasodoro, Goldfinger, Chrysostomos, Maindoree,
Silversmile, Silberselber, Vifargent, Panargyros.
They are immediately appointed to positions of
high public trust in several different countries
\gab{1830}
as managing directors of banks, traffic managers of railways,
chairmen of limited liability companies, vicechairmen of hotel syndicates.

\Voice:
Bloom, are you the Messiah ben Joseph or ben David?

\Bloom:
\stage{(darkly)}
You have said it.

\BrotherBuzz[2]:
Then perform a miracle like Father Charles.

\BantamLyons[2]:
Prophesy who will win the Saint Leger.%
\gab{1840}

\N1:
Bloom walks on a net, covers his left eye with his left ear,
passes through several walls, climbs Nelson's Pillar,
hangs from the top ledge by his eyelids,
eats twelve dozen oysters (shells included),
heals several sufferers from king's evil,
contracts his face so as to resemble many historical personages,
Lord Beaconsfield, Lord Byron, Wat Tyler,
Moses of Egypt, Moses Maimonides, Moses Mendelssohn, Henry Irving,
Rip van Winkle, Kossuth, Jean Jacques Rousseau, Baron Leopold Rothschild,
Robinson Crusoe, Sherlock Holmes, Pasteur,
turns each foot simultaneously in different directions,
\gab{1850}
bids the tide turn back,
eclipses the sun by extending his little finger.

% \Brini:
% \stage{(in papal zouave's uniform, steel cuirasses as breastplate,
% armplates, thighplates, legplates, large profane moustaches and brown paper mitre)}
% \latin{Leopoldi autem generatio.}
% Moses begat Noah
% and Noah begat Eunuch
% and Eunuch begat O'Halloran
% and O'Halloran begat Guggenheim
% and Guggenheim begat Agendath
% and Agendath begat Netaim
% and Netaim begat Le Hirsch
% and Le Hirsch begat Jesurum
% and Jesurum begat MacKay
% and MacKay begat Ostrolopsky
% and Ostrolopsky begat Smerdoz
% and Smerdoz begat Weiss
% \gab{1860}
% and Weiss begat Schwarz
% and Schwarz begat Adrianopoli
% and Adrianopoli begat Aranjuez
% and Aranjuez begat Lewy Lawson
% and Lewy Lawson begat Ichabudonosor
% and Ichabudonosor begat O'Donnell Magnus
% and O'Donnell Magnus begat Christbaum
% and Christbaum begat ben Maimun
% and ben Maimun begat Dusty Rhodes
% and Dusty Rhodes begat Benamor
% and Benamor begat Jones-Smith
% and Jones-Smith begat Savorgnanovich
% and Savorgnanovich begat Jasperstone
% and Jasperstone begat Vingtetunieme
% and Vingtetunieme begat Szombathely
% and Szombathely begat Virag
% and Virag begat Bloom
% \latin{et vocabitur nomen eius Emmanuel}.

% \Deadhand:
% \gab{1870}
% \stage{(writes on the wall)}
% Bloom is a cod.

\Crab:
\gab{1872}
% \stage{(in bushranger's kit)}
What did you do in the cattlecreep behind Kilbarrack?

\FemaleInfant:
% \stage{(shakes a rattle)}
And under Ballybough bridge?

\Hollybush:
And in the devil's glen?

\N1:
Bloom blushes furiously all over from frons to nates,
\gab{1880}
three tears filling from his left eye.

\Bloom:
Spare my past.

\N2:
The Irish Evicted Tenants in bodycoats, kneebreeches,
with Donnybrook fair shillelaghs:

\EvictedTenants:
Sjambok him!

\N1:
Bloom with asses' ears seats himself in the pillory with crossed arms,
his feet protruding.
He whistles \italian{Don Giovanni, a cenar teco}.

\N2:
The Artane Orphans, joining hands, caper round him.
Girls of the Prison Gate Mission, joining hands,
caper round in the opposite direction.

\ArtaneOrphans:
\begin{verse}
\gab{1890}%
    You hig, you hog, you dirty dog!\\
    You think the ladies love you!
\end{verse}

\PrisonGateGirls[1]:
\begin{verse}
    If you see Kay\\
    Tell him he may\\
    See you in tea\\
    Tell him from me.
\end{verse}

\N2:
Hornblower, in ephod and huntingcap, announces.

\Hornblower:
And he shall carry the sins of the people to Azazel,
the spirit which is in the wilderness,
and to Lilith, the nighthag.
And they shall stone him and defile him, yea,
\gab{1900}
all from Agendath Netaim and from Mizraim, the land of Ham.

\N2:
All the people cast soft pantomime stones at Bloom.
Many bonafide travellers and ownerless dogs come near him and defile him.

\N1:
Mastiansky and Citron approach in gaberdines, wearing long earlocks.
They wag their beards at Bloom.

\MastianskyCitron:
Belial!
Laemlein of Istria, the false Messiah!
Abulafia!
Recant!

\N2:
George R Mesias, Bloom's tailor, appears,
a tailor's goose under his arm, presenting a bill.

\Mesias:
\gab{1910}
To alteration, one pair trousers, eleven shillings.

\Bloom:
\stage{(rubs his hands cheerfully)}
Just like old times.
Poor Bloom!

\N1:
Reuben J Dodd, blackbearded Iscariot, bad shepherd,
bearing on his shoulders the drowned corpse of his son,
approaches the pillory.%
\gab{1920}

\ReubenJ:
\stage{(whispers hoarsely)}
The squeak is out.
A split is gone for the flatties.
Nip the first rattler.

\FireBrigade:
Pflaap!

\N1:
Brother Buzz invests Bloom in a yellow habit
with embroidery of painted flames and high pointed hat.
He places a bag of gunpowder round his neck
and hands him over to the civil power, saying.

\BrotherBuzz:
Forgive him his trespasses.

\N2:
Lieutenant Myers of the Dublin Fire Brigade
\gab{1930}
by general request sets fire to Bloom.
Lamentations.

\Citizen[1]:
Thank heaven!

\N1:
Bloom in a seamless garment marked I.H.S.
stands upright amid phoenix flames.

\Bloom:
Weep not for me, O daughters of Erin.

\N1:
He exhibits to Dublin reporters traces of burning.

\N2:
The daughters of Erin, in black garments,
with large prayerbooks and long lighted candles in their hands,
kneel down and pray.

\DaughtersErin:
\gab{1940}
\begin{verse}
    Kidney of Bloom, \emph{pray for us}\\
    Flower of the Bath, \emph{pray for us}\\
    Mentor of Menton, \emph{pray for us}\\
    Canvasser for the Freeman, \emph{pray for us}\\
    Charitable Mason, \emph{pray for us}\\
    Wandering Soap, \emph{pray for us}\\
    Sweets of Sin, \emph{pray for us}\\
    Music without Words, \emph{pray for us}\\
    Reprover of the Citizen, \emph{pray for us}\\
\gab{1950}%
    Friend of all Frillies, \emph{pray for us}\\
    Midwife Most Merciful, \emph{pray for us}\\
    Potato Preservative against Plague and Pestilence, \emph{pray for us}.
\end{verse}

\N2:
A choir of six hundred voices, conducted by Vincent O'Brien,
sings the chorus from Handel's Messiah
\Title{Alleluia for the Lord God Omnipotent Reigneth,}
accompanied on the organ by Joseph Glynn.

\N1:
Bloom becomes mute, shrunken, carbonised.

\par

\Zoe[2]:
Talk away till you're black in the face.

\N1:
Bloom, in caubeen with clay pipe stuck in the band, dusty brogues,
\gab{1960}
an emigrant's red handkerchief bundle in his hand,
leading a black bogoak pig by a sugaun, with a smile in his eye.

\Bloom:
Let me be going now, woman of the house,
for by all the goats in Connemara,
I'm after having the father and mother of a bating.

\N1:
With a tear in his eye.

\Bloom:
All insanity.
Patriotism, sorrow for the dead, music, future of the race.
To be or not to be.
Life's dream is o'er.
End it peacefully.
They can live on.

\N1:
He gazes far away mournfully.

\Bloom:
I am ruined.
A few pastilles of aconite.
The blinds drawn.
A letter.
Then lie back to rest.

\N1:
He breathes softly.

\Bloom:
No more.
I have lived.
Fare. Farewell.

\N2:
Zoe, stiffly, her finger in her neckfillet.

\Zoe:
Honest?
Till the next time.
\gab{1970}
\stage{(she sneers)}
Suppose you got up the wrong side of the bed or came too quick with your best girl.
O, I can read your thoughts!

\Bloom:
\stage{(bitterly)}
Man and woman, love, what is it?
A cork and bottle.
I'm sick of it.
Let everything rip.

\Zoe:
\stage{(in sudden sulks)}
I hate a rotter that's insincere.
Give a bleeding whore a chance.

\Bloom:
\stage{(repentantly)}
\gab{1980}
I am very disagreeable.
You are a necessary evil.
Where are you from?
London?

\Zoe:
\stage{(glibly)}
Hog's Norton where the pigs plays the organs.
I'm Yorkshire born.

\N2:
She holds his hand which is feeling for her nipple.

\Zoe:
I say, Tommy Tittlemouse.
Stop that and begin worse.
Have you cash for a short time?
Ten shillings?

\N1:
Bloom smiles, nods slowly.

\Bloom:
More, houri, more.

\Zoe:
And more's mother?
\gab{1990}

\N2:
She pats him offhandedly with velvet paws.

\Zoe:
Are you coming into the musicroom to see our new pianola?
Come and I'll peel off.

\N1:
Bloom, feeling his occiput dubiously with the unparalleled embarrassment
of a harassed pedlar gauging the symmetry of her peeled pears.

\Bloom:
Somebody would be dreadfully jealous if she knew.
The greeneyed monster.
\stage{(earnestly)}
You know how difficult it is.
I needn't tell you.

\Zoe:
\stage{(flattered)}
What the eye can't see, the heart can't grieve for.
% \stage{(she pats him)}
Come.

\Bloom:
\gab{2000}
Laughing witch!
The hand that rocks the cradle.

\Zoe:
Babby!

\N1:
Bloom, in babylinen and pelisse,
bigheaded, with a caul of dark hair,
fixes big eyes on her fluid slip
and counts its bronze buckles with a chubby finger,
his moist tongue lolling and lisping.

\Bloom:
One two tlee:
tlee tlwo tlone.

\Buckles[2]:
Love me.
Love me not.
Love me.

\Zoe:
\gab{2010}
Silent means consent.

\N2:
With little parted talons she captures his hand,
her forefinger giving to his palm the passtouch of secret monitor,
luring him to doom.

\Zoe:
Hot hands, cold gizzard.

\N1:
He hesitates amid scents, music, temptations.
She leads him towards the steps,
drawing him by the odour of her armpits,
the vice of her painted eyes,
the rustle of her slip in whose sinuous folds
lurks the lion reek of all the male brutes that have possessed her.

\N2:
The Male Brutes, exhaling sulphur of rut and dung
and ramping in their loosebox, faintly roaring,
their drugged heads swaying to and fro.

\MaleBrutes:
\gab{2020}
Good!

\N1:
Zoe and Bloom reach the doorway where two sister whores are seated.
They examine him curiously from under their pencilled brows and smile to his hasty bow.
He trips awkwardly.

\N2:
Zoe, her lucky hand instantly saving him.

\Zoe:
Hoopsa!
Don't fall upstairs.

\Bloom:
The just man falls seven times.

\N1:
He stands aside at the threshold.

\Bloom:
After you is good manners.

\Zoe:
Ladies first, gentlemen after.
\gab{2020}

\N2:
She crosses the threshold.
He hesitates.
She turns and, holding out her hands, draws him over.
He hops.
On the antlered rack of the hall hang a man's hat and waterproof.

\N1:
Bloom uncovers himself but, seeing them,
frowns, then smiles, preoccupied.
A door on the return landing is flung open.
A man in purple shirt and grey trousers, brownsocked,
passes with an ape's gait, his bald head and goatee beard upheld,
hugging a full waterjugjar, his twotailed black braces dangling at heels.
Averting his face quickly Bloom bends to examine on the halltable
the spaniel eyes of a running fox:
then, his lifted head sniffing, follows Zoe into the musicroom.
\gab{2040}

\N2:
A shade of mauve tissuepaper dims the light of the chandelier.
Round and round a moth flies, colliding, escaping.
The floor is covered with an oilcloth mosaic of jade and azure and cinnabar rhomboids.
Footmarks are stamped over it in all senses, heel to heel, heel to hollow,
toe to toe, feet locked, a morris of shuffling feet without body phantoms,
all in a scrimmage higgledypiggledy.
The walls are tapestried with a paper of yewfronds and clear glades.
In the grate is spread a screen of peacock feathers.

\N1:
Lynch squats crosslegged on the hearthrug of matted hair, his cap back to the front.
With a wand, he beats time slowly.
\gab{2050}
Kitty Ricketts, a bony pallid whore in navy costume,
doeskin gloves rolled back from a coral wristlet, a chain purse in her hand,
sits perched on the edge of the table swinging her leg
and glancing at herself in the gilt mirror over the mantelpiece.
A tag of her corsetlace hangs slightly below her jacket.
Lynch indicates mockingly the couple at the piano.

\N2:
Kitty coughs behind her hand.

\Kitty:
She's a bit imbecillic.

\N2:
She signs with a waggling forefinger.

\Kitty:
Blemblem.

\N1:
Lynch lifts up her skirt and white petticoat with his wand.

\N2:
Kitty settles them down quickly.

\Kitty:
Respect yourself.

\N2:
She hiccups, then bends quickly her sailor hat,
\gab{2060}
under which her hair glows, red with henna.

\Kitty:
O, excuse!

\Zoe:
More limelight, Charley.

\N1:
Zoe goes to the chandelier and turns the gas full cock.

\N2:
Kitty peers at the gasjet.

\Kitty:
What ails it tonight?

\N1:
Lynch intones.

\Lynch:
\stage{(deeply)}
Enter a ghost and hobgoblins.

\Zoe:
Clap on the back for Zoe.
\gab{2070}

\N1:
The wand in Lynch's hand flashes: a brass poker.
Stephen stands at the pianola on which sprawl his hat and ashplant.
With two fingers he repeats once more the series of empty fifths.

\N2:
Florry Talbot, a blond feeble goosefat whore
in a tatterdemalion gown of mildewed strawberry,
lolls spreadeagle in the sofacorner,
her limp forearm pendent over the bolster, listening.
A heavy stye droops over her sleepy eyelid.

\N1:
Kitty hiccups again with a kick of her horsed foot.

\Kitty:
O, excuse!

\Zoe:
\gab{2080}
\stage{(promptly)}
Your boy's thinking of you.
Tie a knot on your shift.

\N1:
Kitty Ricketts bends her head.
Her boa uncoils, slides, glides over her shoulder, back, arm, chair to the ground.
Lynch lifts the curled caterpillar on his wand.
She snakes her neck, nestling.

\N2:
Stephen glances behind at the squatted figure with its cap back to the front.

\Stephen:
As a matter of fact it is of no importance
whether Benedetto Marcello found it or made it.
The rite is the poet's rest.
It may be an old hymn to Demeter
or also illustrate \latin{Coela enarrant gloriam Domini}.
It is susceptible of nodes or modes as far apart as hyperphrygian and mixolydian
\gab{2090}
and of texts so divergent as priests haihooping round David's,
that is Circe's, or what am I saying, Ceres' altar,
and David's tip from the stable to his chief bassoonist
about the alrightness of his almightiness.
\french{Mais nom de nom}, that is another pair of trousers.
\french{Jetez la gourme.}
\french{Faut que jeunesse se passe.}

\N2:
He stops, points at Lynch's cap, smiles, laughs.

\Stephen:
Which side is your knowledge bump?

\N1:
Lynch's cap, with saturnine spleen.

\Cap:
Bah!
It is because it is.
Woman's reason.
Jewgreek is greekjew.
Extremes meet.
Death is the highest form of life.
Bah!

\Stephen:
You remember fairly accurately all my errors, boasts, mistakes.
\gab{2100}
How long shall I continue to close my eyes to disloyalty?
Whetstone!

\Cap[1]:
Bah!

\Stephen:
Here's another for you.
\stage{(he frowns)}
The reason is because the fundamental and the dominant
are separated by the greatest possible interval which \ldots

\Cap:
Which?
Finish.
You can't.

\Stephen:
\gab{2110}
\stage{(with an effort)}
Interval which.
Is the greatest possible ellipse.
Consistent with.
The ultimate return.
The octave.
Which.

\Cap:
Which?

\N2:
Outside the gramophone begins to blare \Title{The Holy City}.

\Stephen:
\stage{(abruptly)}
What went forth to the ends of the world to traverse not itself,
God, the sun, Shakespeare, a commercial traveller,
having itself traversed in reality itself becomes that self.
Wait a moment.
Wait a second.
Damn that fellow's noise in the street.
\gab{2120}
Self which it itself was ineluctably preconditioned to become.
\latin{Ecco!}

\N1:
Lynch, with a mocking whinny of laughter, grins at Bloom and Zoe Higgins.

\Lynch:
What a learned speech, eh?

\Zoe:
\stage{(briskly)}
God help your head, he knows more than you have forgotten.

\N2:
With obese stupidity, Florry Talbot regards Stephen.

\Florry:
They say the last day is coming this summer.

\Kitty:
\gab{2130}
No!

\Zoe:
\stage{(explodes in laughter)}
Great unjust God!

\Florry:
\stage{(offended)}
Well, it was in the papers about Antichrist.
O, my foot's tickling.

\N1:
Ragged barefoot newsboys, jogging a wagtail kite, patter past, yelling.

\Newsboys:
Stop press edition.
\gab{2140}
Result of the rockinghorse races.
Sea serpent in the royal canal.
Safe arrival of Antichrist.

\N2:
Stephen turns and sees Bloom.

\Stephen:
A time, times, and half a time.

\N1:
Reuben J Antichrist, wandering Jew,
a clutching hand open on his spine, stumps forward.
Across his loins is slung a pilgrim's wallet from which
protrude promissory notes and dishonoured bills.
Aloft over his shoulder, he bears a long boatpole
from the hook of which the sodden huddled mass of his only son,
saved from Liffey waters, hangs from the slack of its breeches.
\gab{2150}

\N2:
A hobgoblin in the image of Punch Costello, hipshot, crookbacked,
hydrocephalic, prognathic with receding forehead and Ally Sloper nose,
tumbles in somersaults through the gathering darkness.

\All:
What?

\N2:
The Hobgoblin, his jaws chattering, capers to and fro, goggling his eyes,
squeaking, kangaroohopping with outstretched clutching arms,
then all at once thrusts his lipless face through the fork of his thighs.

\Hobgoblin:
\french{Il vient!
C'est moi!
L'homme qui rit!
L'homme primigène!}
\gab{2160}

\N2:
He whirls round and round with dervish howls.

\Hobgoblin:
\french{Sieurs et dames, faites vos jeux!}

\N2:
He crouches juggling.
Tiny roulette planets fly from his hands.

\Hobgoblin:
\french{Les jeux sont faits!}

\N2:
The planets rush together, uttering crepitant cracks.

\Hobgoblin:
\french{Rien va plus!}

\N2:
The planets, buoyant balloons, sail swollen up and away.
He springs off into vacuum.

\N1:
Florry, sinking into torpor, crossing herself secretly.

\Florry:
The end of the world!

\N1:
A female tepid effluvium leaks out from her.
Nebulous obscurity occupies space.

\N2:
Through the drifting fog without, the gramophone blares over coughs and feetshuffling.

\Gramophone:
\gab{2170}
\begin{verse}
    Jerusalem!\\
    Open your gates and sing\\
    Hosanna \ldots
\end{verse}

\N1:
A rocket rushes up the sky and bursts.
A white star fills from it,
proclaiming the consummation of all things and second coming of Elijah.

\N2:
Along an infinite invisible tightrope taut from zenith to nadir,
the End of the World, a twoheaded octopus in gillie's kilts,
busby and tartan filibegs, whirls through the murk,
head over heels, in the form of the Three Legs of Man.

\EndOfWorld:
\gab{2180}
\stage{(with a Scotch accent)}
Wha'll dance the keel row, the keel row, the keel row?

\N1:
Over the possing drift and choking breathcoughs,
Elijah's voice, harsh as a corncrake's, jars on high.
Perspiring in a loose lawn surplice with funnel sleeves, he is seen,
vergerfaced, above a rostrum about which the banner of old glory is draped.
He thumps the parapet.

\Elijah:
No yapping, if you please, in this booth.
Jake Crane, Creole Sue, Dove Campbell, Abe Kirschner,
do your coughing with your mouths shut.
\gab{2190}
Say, I am operating all this trunk line.
Boys, do it now.
God's time is 12.25.
Tell mother you'll be there.
Rush your order and you play a slick ace.
Join on right here.
Book through to eternity junction, the nonstop run.
Just one word more.
Are you a god or a doggone clod?
If the second advent came to Coney Island, are we ready?
Florry Christ, Stephen Christ, Zoe Christ, Bloom Christ, Kitty Christ, Lynch Christ,
it's up to you to sense that cosmic force.
Have we cold feet about the cosmos?
No. Be on the side of the angels.
Be a prism.
You have that something within, the higher self.
You can rub shoulders with a Jesus, a Gautama, an Ingersoll.
Are you all in this vibration?
\gab{2200}
I say you are.
You once nobble that, congregation,
and a buck joyride to heaven becomes a back number.
You got me?
It's a lifebrightener, sure.
The hottest stuff ever was.
It's the whole pie with jam in.
It's just the cutest snappiest line out.
It is immense, supersumptuous.
It restores.
It vibrates.
I know and I am some vibrator.
Joking apart and, getting down to bedrock, A.J.~Christ Dowie
and the harmonial philosophy, have you got that?
O.K.
Seventyseven west sixtyninth street.
Got me?
That's it.
You call me up by sunphone any old time.
Bumboosers, save your stamps.
\stage{(he shouts)}
Now then our glory song.
All join heartily in the singing.
Encore!
\stage{(he sings)}
Jeru \ldots

\N2:
The Gramophone, drowning his voice.
\gab{2210}

\Gramophone:
Whorusalaminyourhighhohhhh \ldots

\N2:
The disc rasps gratingly against the needle.

\N1:
The Three Whores, covering their ears, squawk.

\Zoe, \Kitty, and \Florry:
Ahhkkk!

\N1:
Elijah, in rolledup shirtsleeves, black in the face,
shouts at the top of his voice, his arms uplifted.

\Elijah:
Big Brother up there, Mr~President,
you hear what I done just been saying to you.
Certainly, I sort of believe strong in you, Mr~President.
I certainly am thinking now
Miss Higgins and Miss Ricketts got religion way inside them.
\gab{2220}
Certainly seems to me I don't never see no wusser scared female
than the way you been, Miss Florry, just now as I done seed you.
Mr~President, you come long and help me save our sisters dear.

\N1:
He winks at his audience.

\Elijah:
Our Mr~President, he twig the whole lot and he aint saying nothing.

\KittyKate[2]:
I forgot myself.
In a weak moment I erred and did what I did on Constitution hill.
I was confirmed by the bishop and enrolled in the brown scapular.
My mother's sister married a Montmorency.
It was a working plumber was my ruination when I was pure.

\ZoeFanny[2]:
\gab{2230}
I let him larrup it into me for the fun of it.

\FlorryTeresa[2]:
It was in consequence of a portwine beverage on top of Hennessy's three star.
I was guilty with Whelan, when he slipped into the bed.

\Stephen[1]:
In the beginning was the word, in the end the world without end.
Blessed be the eight beatitudes.

\N2:
The beatitudes, Dixon, Madden, Crotthers, Costello, Lenehan, Bannon,
Mulligan, and Lynch, in white surgical students' gowns, four abreast,
\gab{2240}
goosestepping, tramp fast past in noisy marching.

\Beatitudes:
\stage{(incoherently)}
Beer beef battledog buybull businum barnum buggerum bishop.

\N1:
Lyster, in quakergrey kneebreeches and broadbrimmed hat, says discreetly.

\Lyster:
He is our friend.
I need not mention names.
Seek thou the light.

\N2:
He corantos by.

\N1:
Best enters in hairdresser's attire, shinily laundered, his locks in curlpapers.
He leads John Eglinton who wears a mandarin's kimono of Nankeen yellow,
lizardlettered, and a high pagoda hat.
\gab{2250}
Best, smiling, lifts the hat and displays a shaven poll,
from the crown of which bristles a pigtail toupee tied with an orange topknot.

\Best:
I was just beautifying him, don't you know.
A thing of beauty, don't you know,
Yeats says, or I mean, Keats says.

\N2:
John Eglinton produces a greencapped dark lantern
and flashes it towards a corner, saying with carping accent.

\JohnEglinton:
Esthetics and cosmetics are for the boudoir.
I am out for truth.
Plain truth for a plain man.
Tanderagee wants the facts and means to get them.%
\gab{2260}

\N1:
In the cone of the searchlight, behind the coalscuttle,
ollave, holyeyed, the bearded figure of Mananaun Maclir broods, chin on knees.
He rises slowly.
A cold seawind blows from his druid mouth.
About his head writhe eels and elvers.
He is encrusted with weeds and shells.
His right hand holds a bicycle pump.
His left hand grasps a huge crayfish by its two talons.
Mananaun Maclir cries with a voice of waves.

\MananaunMaclir:
Aum! Hek! Wal! Ak! Lub! Mor! Ma!
White yoghin of the gods.
Occult pimander of Hermes Trismegistos.

\N1:
With a voice of whistling seawind.
\gab{2270}

\MananaunMaclir:
Punarjanam patsypunjaub!
I won't have my leg pulled.
It has been said by one:
beware the left, the cult of Shakti.

\N1:
With a cry of stormbirds.

\MananaunMaclir:
Shakti Shiva, darkhidden Father!

\N1:
He smites with his bicycle pump the crayfish in his left hand.
On its cooperative dial glow the twelve signs of the zodiac.
He wails with the vehemence of the ocean.

\MananaunMaclir:
Aum! Baum! Pyjaum!
I am the light of the homestead!
I am the dreamery creamery butter.

\N2:
A skeleton judashand strangles the light.
The green light wanes to mauve.

\N1:
The gasjet wails whistling.

\Gasjet:
Pooah! Pfuiiiiiii!
\gab{2280}

\N2:
Zoe runs to the chandelier and, crooking her leg, adjusts the mantle.

\Zoe:
Who has a fag as I'm here?

\N1:
Lynch tosses a cigarette on to the table.

\Lynch:
Here.

\N2:
Zoe, her head perched aside in mock pride.

\Zoe:
Is that the way to hand the \french{pot} to a lady?

\N2:
She stretches up to light the cigarette over the flame,
twirling it slowly, showing the brown tufts of her armpits.
\gab{2290}
Lynch with his poker lifts boldly a side of her slip.
Bare from her garters up her flesh appears under the sapphire a nixie's green.
She puffs calmly at her cigarette.

\Zoe:
Can you see the beautyspot of my behind?

\Lynch:
I'm not looking.

\N2:
Zoe makes sheep's eyes.

\Zoe:
No? You wouldn't do a less thing.
Would you suck a lemon?

\N2:
Squinting in mock shame, she glances with sidelong meaning at Bloom,
\gab{2300}
then twists round towards him, pulling her slip free of the poker.
Blue fluid again flows over her flesh.
Bloom stands, smiling desirously, twirling his thumbs.
Kitty Ricketts licks her middle finger with her spittle and,
gazing in the mirror, smooths both eyebrows.

\N1:
Lipoti Virag, basilicogrammate,
chutes rapidly down through the chimneyflue and struts
two steps to the left on gawky pink stilts.
He is sausaged into several overcoats and wears a brown macintosh
under which he holds a roll of parchment.
In his left eye, flashes the monocle of
Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell.
On his head is perched an Egyptian pshent.
Two quills project over his ears.%
\gab{2310}
Virag, heels together, bows.

\Virag:
My name is Virag Lipoti, of Szombathely.

\N1:
He coughs thoughtfully, drily.

\Virag:
Promiscuous nakedness is much in evidence hereabouts, eh?
Inadvertently her backview revealed the fact that she is not wearing
those rather intimate garments of which you are a particular devotee.
The injection mark on the thigh I hope you perceived?
Good.

\Bloom:
Granpapachi. But \ldots

\Virag:
Number two on the other hand, she of the cherry rouge and coiffeuse white,%
\gab{2320}
whose hair owes not a little to our tribal elixir of gopherwood,
is in walking costume and tightly staysed by her sit, I should opine.
Backbone in front, so to say.
Correct me but I always understood that the act so performed by skittish humans
with glimpses of lingerie, appealed to you in virtue of its exhibitionististicicity.
In a word. Hippogriff.
Am I right?

\Bloom:
She is rather lean.

\N1:
Virag, not unpleasantly.

\Virag:
Absolutely!
Well observed and those pannier pockets of the skirt
\gab{2330}
and slightly pegtop effect are devised to suggest bunchiness of hip.
A new purchase at some monster sale for which a gull has been mulcted.
Meretricious finery to deceive the eye.
Observe the attention to details of dustspecks.
Never put on you tomorrow what you can wear today.
Parallax!

\N1:
With a nervous twitch of his head.

\Virag:
Did you hear my brain go snap?
Pollysyllabax!

\N2:
Bloom, an elbow resting in a hand, a forefinger against his cheek.

\Bloom:
She seems sad.

\N1:
Virag, cynically, his weasel teeth bared yellow, draws down his left eye
with a finger and barks hoarsely.

\Virag:
\gab{2340}
Hoax! Beware of the flapper and bogus mournful.
Lily of the alley.
All possess bachelor's button discovered by Rualdus Columbus.
Tumble her. Columble her. Chameleon.
\stage{(more genially)}
Well then, permit me to draw your attention to item number three.
There is plenty of her visible to the naked eye.
Observe the mass of oxygenated vegetable matter on her skull.
What ho, she bumps!
The ugly duckling of the party, longcasted and deep in keel.

\Bloom:
\stage{(regretfully)}
When you come out without your gun.

\Virag:
We can do you all brands, mild, medium, and strong.
\gab{2350}
Pay your money, take your choice.
How happy could you be with either \ldots

\Bloom:
With \ldots?

\N1:
Virag, his tongue upcurling.

\Virag:
Lyum! Look.
Her beam is broad.
She is coated with quite a considerable layer of fat.
Obviously mammal in weight of bosom,
you remark that she has in front, well to the fore,
two protuberances of very respectable dimensions,
inclined to fall in the noonday soupplate,
while on her rere, lower down, are two additional protuberances,
suggestive of potent rectum and tumescent for palpation,
\gab{2360}
which leave nothing to be desired, save compactness.
Such fleshy parts are the product of careful nurture.
When coop-fattened, their livers reach an elephantine size.
Pellets of new bread with fennygreek and gumbenjamin
swamped down by potions of green tea,
endow them during their brief existence
with natural pincushions of quite colossal blubber.
That suits your book, eh?
Fleshhotpots of Egypt to hanker after.
Wallow in it.
Lycopodium.

\N1:
His throat twitches.

\Virag:
Slapbang!
There he goes again.

\Bloom:
The stye I dislike.

\N1:
Virag arches his eyebrows.
\gab{2370}

\Virag:
Contact with a goldring, they say.
\latin{Argumentum ad feminam},
as we said in old Rome and ancient Greece
in the consulship of Diplodocus and Ichthyosauros.
For the rest, Eve's sovereign remedy.
Not for sale.
Hire only.
Huguenot.

\N1:
He twitches.

\Virag:
It is a funny sound.

\N1:
He coughs encouragingly.

\Virag:
But possibly it is only a wart.
I presume you shall have remembered what I will have taught you on that head?
Wheatenmeal with honey and nutmeg.

\N2:
Bloom, reflecting.

\Bloom:
Wheatenmeal with lycopodium and syllabax.
This searching ordeal.
It has been an unusually fatiguing day, a chapter of accidents.
\gab{2380}
\stage{to Virag}
Wait.
I mean, wartsblood spreads warts, you said \ldots

\N1:
Virag, severely, his nose hardhumped, his side eye winking.

\Virag:
Stop twirling your thumbs and have a good old thunk.
See, you have forgotten.
Exercise your mnemotechnic.
\italian{La causa è santa.}
Tara. Tara.
\stage{(aside)}
He will surely remember.

\Bloom:
Rosemary also did I understand you to say or willpower over parasitic tissues.
Then nay, no, I have an inkling.
The touch of a deadhand cures.
Mnemo?%
\gab{2390}

\Virag:
\stage{(excitedly)}
I say so. I say so.
E'en so.
Technic.

\N1:
He taps his parchment roll energetically.

\Virag:
This book tells you how to act with all descriptive particulars.
Consult index for agitated fear of aconite,
melancholy of muriatic, priapic pulsatilla.
Virag is going to talk about amputation.
Our old friend caustic.
They must be starved.
Snip off with horsehair under the denned neck.
But, to change the venue to the Bulgar and the Basque,
have you made up your mind
whether you like or dislike women in male habiliments?

\N1:
With a dry snigger.

\Virag:
You intended to devote an entire year to the study
of the religious problem and the summer months of 1886
\gab{2400}
to square the circle and win that million.
Pomegranate!
From the sublime to the ridiculous is but a step.
Pyjamas, let us say?
Or stockingette gussetted knickers, closed?
Or, put we the case, those complicated combinations, camiknickers?

\N1:
He crows derisively.

\Virag:
Keekeereekee!

\N2:
Bloom surveys uncertainly the three whores,
then gazes at the veiled mauve light, hearing the everflying moth.

\Bloom:
I wanted then to have now concluded.
Nightdress was never.
Hence this.
But tomorrow is a new day will be.
Past was is today.
What now is will then morrow
\gab{2410}
as now was be past yester.

\N1:
Virag prompts in a pig's whisper.

\Virag:
Insects of the day spend their brief existence in reiterated coition,
lured by the smell of the inferiorly pulchritudinous fumale
possessing extendified pudendal nerve in dorsal region.
Pretty Poll!

\N1:
His yellow parrotbeak gabbles nasally.

\Virag:
They had a proverb in the Carpathians, in or about the year
five thousand five hundred and fifty of our era.
One tablespoonful of honey will attract friend Bruin
more than half a dozen barrels of first choice malt vinegar.
Bear's buzz bothers bees.
But of this apart.
At another time we may resume.
We were very pleased, we others.
\gab{2420}

\N1:
He coughs and, bending his brow,
rubs his nose thoughtfully with a scooping hand.

\Virag:
You shall find that these night insects follow the light.
An illusion for remember their complex unadjustable eye.
For all these knotty points see the seventeenth book
of my Fundamentals of Sexology or the Love Passion
which Doctor L.B. says is the book sensation of the year.
Some, to example, there are again whose movements are automatic.
Perceive.
That is his appropriate sun.
Nightbird nightsun nighttown.
Chase me, Charley!

\N1:
He blows into Bloom's ear.

\Virag:
Buzz!

\Bloom:
Bee or bluebottle too other day butting shadow on wall dazed self
then me wandered dazed down shirt good job I \ldots%
\gab{2430}

\N1:
Virag, his face impassive, laughs in a rich feminine key.

\Virag:
Splendid!
Spanish fly in his fly or mustard plaster on his dibble.

\N1:
He gobbles gluttonously with turkey wattles.

\Virag:
Bubbly jock! Bubbly jock!
Where are we?
Open Sesame!
Cometh forth!

\N1:
He unrolls his parchment rapidly and reads,
his glowworm's nose running backwards over the letters which he claws.

\Virag:
Stay, good friend.
I bring thee thy answer.
Redbank oysters will shortly
be upon us.
I'm the best o'cook.
Those succulent bivalves may help us and
the truffles of Perigord, tubers dislodged through mister omnivorous porker,
were unsurpassed in cases of nervous debility or viragitis.
\gab{2440}
Though they stink, yet they sting.

\N1:
He wags his head with cackling raillery.

\Virag:
Jocular.
With my eyeglass in my ocular.

\N1:
He sneezes.

\Virag:
Amen!

\Bloom:
\stage{(absently)}
Ocularly woman's bivalve case is worse.
Always open sesame.
The cloven sex.
Why they fear vermin, creeping things.
Yet Eve and the serpent contradicts.
Not a historical fact.
Obvious analogy to my idea.
Serpents too are gluttons for woman's milk.
Wind their way through miles of omnivorous forest to sucksucculent her breast dry.
Like those bubblyjocular Roman matrons one reads of in Elephantuliasis.

\N1:
Virag, his mouth projected in hard wrinkles, eyes stonily forlornly closed,
\gab{2450}
psalms in outlandish monotone.

\Virag:
That the cows with their those distended udders
that they have been the the known \ldots

\Bloom:
I am going to scream.
I beg your pardon.
Ah? So.
% \stage{(he repeats)}
Spontaneously to seek out the saurian's lair in order to
entrust their teats to his avid suction.
Ant milks aphis.
\stage{(profoundly)}
Instinct rules the world.
In life.
In death.

\N1:
Virag, head askew, arches his back and hunched wingshoulders,
\gab{2460}
peers at the moth out of blear bulged eyes, points a horning claw and cries.

\Virag:
Who's moth moth?
Who's dear Gerald?
Dear Ger, that you?
O dear, he is Gerald.
O, I much fear he shall be most badly burned.
Will some pleashe pershon not now impediment
so catastrophics mit agitation of firstclass tablenumpkin?

\N1:
He mews.

\Virag:
Puss puss puss puss!

\N1:
He sighs, draws back, and stares sideways down with dropping underjaw.

\Virag:
Well, well.
He doth rest anon.

\N1:
He snaps his jaws suddenly on the air.

\Moth[2]:
\begin{verse}
    I'm a tiny tiny thing\\
\gab{2470}%
    Ever flying in the spring\\
    Round and round a ringaring.\\
    Long ago I was a king\\
    Now I do this kind of thing\\
    On the wing, on the wing!\\
    Bing!
\end{verse}

\N2:
The moth rushes against the mauve shade, flapping noisily.

\Moth:
Pretty pretty pretty pretty pretty pretty petticoats.

\N2:
From left upper entrance with two gliding steps,
Henry Flower comes forward to left front centre.
He wears a dark mantle and drooping plumed sombrero.
\gab{2480}
He carries a silverstringed inlaid dulcimer and
a longstemmed bamboo Jacob's pipe, its clay bowl fashioned as a female head.
He wears dark velvet hose and silverbuckled pumps.
He has the romantic Saviour's face with flowing locks, thin beard and moustache.
His spindlelegs and sparrow feet are those of the tenor Mario, prince of Candia.
He settles down his goffered ruffs
and moistens his lips with a passage of his amorous tongue.
He says in a low dulcet voice, touching the strings of his guitar.

\Henry:
There is a flower that bloometh.%
\gab{2490}

\N2:
Virag truculent, his jowl set, stares at the lamp.
Grave Bloom regards Zoe's neck.
Henry gallant turns with pendant dewlap to the piano.

\N1:
Stephen, to himself.

\Stephen:
Play with your eyes shut.
Imitate pa.
Filling my belly with husks of swine.
Too much of this.
I will arise and go to my.
Expect this is the.
Steve, thou art in a parlous way.
Must visit old Deasy or telegraph.
Our interview of this morning has left on me a deep impression.
Though our ages.
Will write fully tomorrow.
I'm partially drunk, by the way.

\N1:
He touches the keys again.
\gab{2500}

\Stephen:
Minor chord comes now.
Yes. Not much however.

\N2:
Almidano Artifoni holds out a batonroll of music with vigorous moustachework.

\Artifoni:
\italian{Ci rifletta.
Lei rovina tutto.}

\Florry[1]:
Sing us something.
Love's old sweet song.

\Stephen:
No voice.
I am a most finished artist.
Lynch, did I show you the letter about the lute?

\Florry:
\gab{2510}
\stage{(smirking)}
The bird that can sing and won't sing.

\N2:
The Siamese twins, Philip Drunk and Philip Sober,
two Oxford dons with lawnmowers, appear in the window embrasure.
Both are masked with Matthew Arnold's face.

\PhilipSober[2]:
Take a fool's advice.
All is not well.
Work it out with the buttend of a pencil, like a good young idiot.
Three pounds twelve you got, two notes,
one sovereign, two crowns, if youth but knew.
Mooney's en ville, Mooney's sur mer, the Moira, Larchet's,
Holles street hospital, Burke's.
Eh? I am watching you.%
\gab{2520}

\PhilipDrunk[2]:
\stage{(impatiently)}
Ah, bosh, man.
Go to hell!
I paid my way.
If I could only find out about octaves.
Reduplication of personality.
Who was it told me his name?

\N2:
His lawnmower begins to purr.

\PhilipDrunk:
Aha, yes.
\greek{Zoe mou sas agapo.}
Have a notion I was here before.
When was it not Atkinson his card I have somewhere.
Mac Somebody.
Unmack I have it.
He told me about, hold on, Swinburne, was it, no?

\Florry:
And the song?

\Stephen:
\gab{2530}
Spirit is willing but the flesh is weak.

\Florry:
Are you out of Maynooth?
You're like someone I knew once.

\Stephen:
Out of it now.
\stage{(to himself)}
Clever.

\stage{(Philip Drunk and Philip Sober,
their lawnmowers purring with a rigadoon of grasshalms.)}

\PhilipDrunk\ and \PhilipSober:
Clever ever.
Out of it out of it.
By the bye have you the book, the thing, the ashplant?
Yes, there it, yes.
Cleverever outofitnow.
Keep in condition.
Do like us.

\Zoe:
\gab{2540}
There was a priest down here two nights ago
to do his bit of business with his coat buttoned up.
You needn't try to hide, I says to him.
I know you've a Roman collar.

\Virag:
Perfectly logical from his standpoint.
Fall of man.

\N1:
Harshly, his pupils waxing.

\Virag:
To hell with the pope!
Nothing new under the sun.
I am the Virag who disclosed the Sex Secrets of Monks and Maidens.
Why I left the Church of Rome.
Read the Priest, the Woman and the Confessional.
Penrose.
Flipperty Jippert.

\N1:
He wriggles.

\Virag:
Woman, undoing with sweet pudor her belt of rushrope,
offers her allmoist yoni to man's lingam.
\gab{2550}
Short time after man presents woman with pieces of jungle meat.
Woman shows joy and covers herself with featherskins.
Man loves her yoni fiercely with big lingam, the stiff one.

\N1:
He cries.

\Virag:
\latin{Coactus volui.}
Then giddy woman will run about.
Strong man grapses woman's wrist.
Woman squeals, bites, spucks.
Man, now fierce angry, strikes woman's fat yadgana.

\N1:
He chases his tail.

\Virag:
Piffpaff!
Popo!

\N1:
He stops, sneezes.

\Virag:
Pchp!

\N1:
He worries his butt.

\Virag:
Prrrrrht!

\Lynch[2]:
I hope you gave the good father a penance.
Nine glorias for shooting a bishop.

\N1:
Zoe spouts walrus smoke through her nostrils.
\gab{2560}

\Zoe:
He couldn't get a connection.
Only, you know, sensation.
A dry rush.

\Bloom:
Poor man!

\Zoe:
\stage{(lightly)}
Only for what happened him.

\Bloom:
How?

\N1:
A diabolic rictus of black luminosity contracting his visage,
\gab{2570}
cranes his scraggy neck forward.
He lifts a mooncalf nozzle and howls.

\Virag:
\yiddish{Verfluchte Goim!}
He had a father, forty fathers.
He never existed.
Pig God!
He had two left feet.
He was Judas Iacchia, a Libyan eunuch, the pope's bastard.

\N1:
He leans out on tortured forepaws, elbows bent rigid,
his eye agonising in his flat skullneck and yelps over the mute world.

\Virag:
A son of a whore.
Apocalypse.

\Kitty[2]:
And Mary Shortall that was in the lock with the pox
she got from Jimmy Pidgeon in the blue caps
had a child off him that couldn't swallow
and was smothered with the convulsions in the mattress
\gab{2580}
and we all subscribed for the funeral.

\N2:
Philip Drunk, gravely.

\PhilipDrunk:
\french{Qui vous a mis dans cette fichue position, Philippe?}

\N2:
Philip Sober, gaily.

\PhilipSober:
\french{C'était le sacré pigeon, Philippe.}

\N2:
Kitty unpins her hat and sets it down calmly, patting her henna hair.
And a prettier, a daintier head of winsome curls was never seen on a whore's shoulders.
Lynch puts on her hat.
She whips it off.

\N1:
Lynch laughs.
\gab{2590}

\Lynch:
And to such delights has Metchnikoff inoculated anthropoid apes.

\N2:
Florry nods.

\Florry:
Locomotor ataxy.

\Zoe:
\stage{(gaily)}
O, my dictionary.

\Lynch:
Three wise virgins.

\N1:
Virag, ague-shaken, profuse yellow spawn foaming over his bony epileptic lips.

\Virag:
She sold lovephiltres, whitewax, orangeflower.
Panther, the Roman centurion, polluted her with his genitories.
\gab{2600}

\N1:
He sticks out a flickering phosphorescent scorpion tongue, his hand on his fork.

\Virag:
Messiah!
He burst her tympanum.

\N1:
With gibbering baboon's cries, he jerks his hips in the cynical spasm.

\Virag:
Hik! Hek! Hak! Hok! Huk! Kok! Kuk!

\N2:
Ben Jumbo Dollard, rubicund, musclebound, hairynostrilled, hugebearded,
cabbageeared, shaggychested, shockmaned, fat-papped, stands forth,
his loins and genitals tightened into a pair of black bathing bagslops.
Dollard, nakkering castanet bones in his huge padded paws,
yodels jovially in base barreltone.

\BenDollard:
\gab{2610}
When love absorbs my ardent soul.

\N1:
The virgins Nurse Callan and Nurse Quigley burst through the ringkeepers
and the ropes and mob him with open arms.

\Virgins:
\stage{(gushingly)}
Big Ben!
Ben my Chree!

\Voice:
Hold that fellow with the bad breeches.

\N2:
Dollard, smites his thigh in abundant laughter.

\BenDollard:
Hold him now.

\N1:
Henry Flower, caressing on his breast a severed female head, murmurs.
\gab{2620}

\Henry:
Thine heart, mine love.

\N1:
He plucks his lutestrings.

\Henry:
When first I saw \ldots

\N1:
Virag, sloughing his skins, his multitudinous plumage moulting.

\Virag:
Rats!

\N1:
He yawns, showing a coalblack throat,
and closes his jaws by an upward push of his parchment roll.

\Virag:
After having said which I took my departure.
Farewell.
Fare thee well.
\yiddish{Dreck!}

\N2:
Henry Flower combs his moustache and beard rapidly with a pocketcomb
and gives a cow's lick to his hair.
Steered by his rapier, he glides to the door, his wild harp slung behind him.

\N1:
Virag reaches the door in two ungainly stilthops,
\gab{2630}
his tail cocked, and deftly claps sideways on the wall
a pus-yellow flybill, butting it with his head.

\Flybill[2]:
K.~II.
Post No Bills.
Strictly confidential.
Dr Hy Franks.

\Henry:
All is lost now.

\N1:
Virag unscrews his head in a trice and holds it under his arm.

\ViragsHead:
Quack!

\All:
Exeunt severally.

\end{document}
