\N2:
He had reached the open backdoor of All Hallows.
Stepping into the porch
he doffed his hat,
took the card from his pocket
and tucked it again behind the leather headband.

\BloomInt:
Damn it.
\gab{320}
I might have tried to work M'Coy for a pass to Mullingar.

\N2:
Same notice on the door.
Sermon by the very reverend John Conmee SJ
on saint Peter Claver SJ and the African Mission.

\BloomInt:
Prayers for the conversion of Gladstone they had too
when he was almost unconscious.
The protestants are the same.
Convert Dr~William~J. Walsh~DD to the true religion.
Save China's millions.
Wonder how they explain it to the heathen Chinee.
Prefer an ounce of opium.
Celestials.
Rank heresy for them.
Buddha their god lying on his side in the museum.
Taking it easy
with hand under his cheek.
Josssticks burning.
Not like Ecce Homo.
Crown of thorns and cross.
Clever idea
Saint Patrick the shamrock.
\gab{330}
Chopsticks?
Conmee:
Martin Cunningham knows him:
distinguishedlooking.
Sorry I didn't work him about getting Molly into the choir
instead of that Father Farley who looked a fool but wasn't.
They're taught that.
He's not going out in bluey specs
with the sweat rolling off him to baptise blacks,
is he?
The glasses would take their fancy, flashing.
Like to see them sitting round in a ring with blub lips,
entranced, listening.
Still life.
Lap it up like milk, I suppose.

\N2:
The cold smell of sacred stone called him.
He trod the worn steps,
pushed the swingdoor
and entered softly by the rere.

\BloomInt:
Something going on: some sodality.
\gab{340}
Pity so empty.
Nice discreet place to be next some girl.
Who is my neighbour?
Jammed by the hour to slow music.
That woman at midnight mass.
Seventh heaven.

\N2:
Women knelt in the benches with crimson halters round their necks,
heads bowed.
A batch knelt at the altarrails.
The priest went along by them, murmuring,
holding the thing in his hands.
He stopped at each, took out a communion,
shook a drop or two

\BloomInt:
(are they in water?)

\N2:
off it and put it neatly into her mouth.
Her hat and head sank.
Then the next one.
Her hat sank at once.
Then the next one:
a small old woman.
The priest bent down to put it into her mouth,
murmuring all the time.

\BloomInt:
Latin.

\N2:
The next one.

\BloomInt:
Shut your eyes and open your mouth.
\gab{350}
What?
\latin{Corpus}:
body.
Corpse.
Good idea the Latin.
Stupefies them first.
Hospice for the dying.
They don't seem to chew it:
only swallow it down.
Rum idea:
eating bits of a corpse.
Why the cannibals cotton to it.

\N2:
He stood aside
watching their blind masks pass down the aisle,
one by one,
and seek their places.
He approached a bench and seated himself in its corner,
nursing his hat and newspaper.

\BloomInt:
These pots we have to wear.
We ought to have hats modelled on our heads.

\N2:
They were about him here and there,
with heads still bowed in their crimson halters,
waiting for it to melt in their stomachs.

\BloomInt:
Something like those mazzoth:
it's that sort of bread:
unleavened shewbread.
Look at them.
Now I bet it makes them feel happy.
Lollipop.
\gab{360}
It does.
Yes, bread of angels it's called.
There's a big idea behind it,
kind of kingdom of God is within you feel.
First communicants.
Hokypoky penny a lump.
Then feel all like one family party,
same in the theatre, all in the same swim.
They do.
I'm sure of that.
Not so lonely.
In our confraternity.
Then come out a bit spreeish.
Let off steam.
Thing is if you really believe in it.
Lourdes cure,
waters of oblivion,
and the Knock apparition,
statues bleeding.
Old fellow asleep near that confessionbox.
Hence those snores.
Blind faith.
Safe in the arms of kingdom come.
Lulls all pain.
Wake this time next year.

\N2:
He saw the priest stow the communion cup away,
well in,
and kneel an instant before it,
\gab{370}
showing a large grey bootsole from under the lace affair he had on.

\BloomInt:
Suppose he lost the pin of his.
He wouldn't know what to do to.
Bald spot behind.
Letters on his back:
I.N.R.I?
No:
I.H.S.
Molly told me one time I asked her.
I have sinned:
or no:
I have suffered, it is.
And the other one?
Iron nails ran in.

\BloomInt:
Meet one Sunday after the rosary.
Do not deny my request.
Turn up with a veil and black bag.
Dusk and the light behind her.
She might be here with a ribbon round her neck
and do the other thing all the same on the sly.
Their character.
That fellow that turned queen's evidence on the invincibles
he used to receive the,
Carey was his name,
the communion every morning.
This very church.
Peter Carey, yes.
\gab{380}
No, Peter Claver I am
thinking of.
Denis Carey.
And just imagine that.
Wife and six children at home.
And plotting that murder all the time.
Those crawthumpers,
now that's a good name for them,
there's always something shiftylooking about them.
They're not straight men of business either.
O, no, she's not here:
the flower:
no, no.
By the way, did I tear up that envelope?
Yes: under the bridge.

\N2:
The priest was rinsing out the chalice:
then he tossed off the dregs smartly.

\BloomInt:
Wine.
Makes it more aristocratic than
for example if he drank
what they are used to |
Guinness's porter or some temperance beverage |
Wheatley's Dublin hop bitters |
or Cantrell and Cochrane's ginger ale (aromatic).
Doesn't give them any of it:
\gab{390}
shew wine:
only the other.
Cold comfort.
Pious fraud but quite right:
otherwise they'd have one old booser worse than another coming along,
cadging for a drink.
Queer the whole atmosphere of the.
Quite right.
Perfectly right that is.

\N2:
Mr Bloom looked back towards the choir.

\BloomInt:
Not going to be any music.
Pity.
Who has the organ here I wonder?
Old Glynn he knew how to make that instrument talk,
the \italian{vibrato}:
fifty pounds a year they say he had in Gardiner street.
Molly was in fine voice that day,
the \latin{Stabat Mater} of Rossini.
Father Bernard Vaughan's sermon first.
Christ or Pilate?
Christ,
but don't keep us all night over it.
Music they wanted.
Footdrill stopped.
Could hear a pin drop.
\gab{400}
I told her to pitch her voice against that corner.
I could feel the thrill in the air,
the full,
the people looking up:

\priest:
\latin{Quis est homo.}

\BloomInt:
Some of that old sacred music splendid.
Mercadante: seven last words.
Mozart's twelfth mass:
\latin{Gloria} in that.
Those old popes keen on music,
on art and statues and pictures of all kinds.
Palestrina for example too.
They had a gay old time while it lasted.
Healthy too,
chanting,
regular hours,
then brew liqueurs.
Benedictine.
Green Chartreuse.
Still, having eunuchs in their choir
that was coming it a bit thick.
What kind of voice is it?
Must be curious to hear after their own strong basses.
Connoisseurs.
Suppose they wouldn't feel anything after.
\gab{410}
Kind of a placid.
No worry.
Fall into flesh, don't they?
Gluttons, tall, long legs.
Who knows?
Eunuch.
One way out of it.

\N2:
He saw the priest bend down and kiss the altar
and then face about and bless all the people.
All crossed themselves and stood up.
Mr Bloom glanced about him
and then stood up,
looking over the risen hats.

\BloomInt:
Stand up at the gospel of course.

\N2:
Then all settled down on their knees again
and he sat back quietly in his bench.
The priest came down from the altar,
holding the thing out from him,
and he and the massboy answered each other in Latin.
Then the priest knelt down and began to read off a card:

\priest:
O God, our refuge and our strength ...
\gab{420}

\N2:
Mr Bloom put his face forward to catch the words.

\BloomInt:
English.
Throw them the bone.
I remember slightly.
How long since your last mass?
Glorious and immaculate virgin.
Joseph, her spouse.
Peter and Paul.
More interesting if you understood what it was all about.
Wonderful organisation certainly,
goes like clockwork.
Confession.
Everyone wants to.
Then I will tell you all.
Penance.
Punish me, please.
Great weapon in their hands.
More than doctor or solicitor.
Woman dying to.
And I schschschschschsch.
And did you chachachachacha?
And why did you?
Look down at her ring to find an excuse.
Whispering gallery walls have ears.
Husband learn to his surprise.
God's little joke.
\gab{430}
Then out she comes.
Repentance skindeep.
Lovely shame.
Pray at an altar.
Hail Mary and Holy Mary.
Flowers, incense, candles melting.
Hide her blushes.
Salvation army blatant imitation.
Reformed prostitute will address the meeting.
How I found the Lord.
Squareheaded chaps those must be in Rome:
they work the whole show.
And don't they rake in the money too?
Bequests also:
to the P.P. for the time being in his absolute discretion.
Masses for the repose of my soul to be said publicly with open doors.
Monasteries and convents.
The priest in that Fermanagh will case in the witnessbox.
No browbeating him.
He had his answer pat for everything.
Liberty and exaltation of our holy mother the church.
\gab{440}
The doctors of the church:
they mapped out the whole theology of it.

\N2:
The priest prayed:

\priest:
Blessed Michael,
archangel,
defend us in the hour of conflict.
Be our safeguard against the wickedness and snares of the devil
(may God restrain him, we humbly pray!):
and do thou,
O prince of the heavenly host,
by the power of God
thrust Satan down to hell and with him
those other wicked spirits who wander through the world
for the ruin of souls.

\N2:
The priest and the massboy stood up and walked off.

\BloomInt:
All over.

\N2:
The women remained behind:

\BloomInt:
thanksgiving.

\BloomInt:
Better be shoving along.
\gab{450}
Brother Buzz.
Come around with the plate perhaps.
Pay your Easter duty.

\N1:
He stood up.

\BloomInt:
Hello.
Were those two buttons of my waistcoat open all the time?
Women enjoy it.
Never tell you.
But we.
Excuse, miss, there's a (whh!)
just a (whh!) fluff.
Or their skirt behind, placket unhooked.
Glimpses of the moon.
Annoyed if you don't.
Why didn't you tell me before.
Still like you better untidy.
Good job it wasn't farther south.

\N1:
He passed, discreetly buttoning,
down the aisle and out through the main door into the light.
He stood a moment unseeing
by the cold black marble bowl
while before him and behind |
two worshippers dipped furtive hands in the low tide of holy water.
\gab{460}
Trams:
a car of Prescott's dyeworks:
a widow in her weeds.

\BloomInt:
Notice because I'm in mourning myself.

\N1:
He covered himself.
