\BloomInt:
M'Coy.
Get rid of him quickly.
Take me out of my way.
Hate company when you.

\mccoy:
Hello, Bloom.
Where are you off to?

\Bloom:
Hello, M'Coy.
Nowhere in particular.

\mccoy:
How's the body?

\Bloom:
Fine.
How are you?

\mccoy:
Just keeping alive,
\sout{M'Coy said.}

\N1:
His eyes on the black tie and clothes
he asked with low respect:

\mccoy:
Is there any ... no trouble I hope?
I see you're ...

\Bloom:
O, no, \sout{Mr Bloom said.}
Poor Dignam, you know.
The funeral is today.

\mccoy:
To be sure, poor fellow.
So it is.
What time?

\BloomInt:
A photo it isn't.
A badge maybe.

\Bloom:
E ... eleven,
\sout{Mr Bloom answered.}

\mccoy:
I must try to get out there,
\sout{M'Coy said.}
Eleven, is it?
I only heard it last night.
Who was telling me?
Holohan.
You know Hoppy?

\Bloom:
I know.

\N1:
Mr Bloom gazed across the road at the outsider
drawn up before the door of the Grosvenor.
The porter hoisted the valise up on the well.

\N2:
She stood still, waiting,
while the man, husband, brother,
like her,
searched his pockets for change.

\BloomInt:
Stylish kind of coat with that roll collar,
warm for a day like this,
looks like blanketcloth.
Careless stand of her
with her hands in those patch pockets.
Like that haughty creature at the polo match.
Women all for caste till you touch the spot.
Handsome is and handsome does.
Reserved about to yield.
The honourable Mrs
and Brutus is an honourable man.
Possess her once
take the starch out of her.

\mccoy:
I was with Bob Doran,
he's on one of his periodical bends,
and what do you call him Bantam Lyons.
Just down there in Conway's
we were.

\mccoy:
Doran Lyons in Conway's.

\N2:
She raised a gloved hand to her hair.

\mccoy:
In came Hoppy.
Having a wet.

\N2:
Drawing back his head
and gazing far from beneath
his vailed eyelids
he saw the bright fawn skin shine in the glare,
the braided drums.

\BloomInt:
Clearly I can see today.
Moisture about gives long sight perhaps.
Talking of one thing or another.
Lady's hand.
Which side will she get up?

\mccoy:
And he said: \Quotation{Sad thing about our poor friend Paddy!}
\Quotation{What Paddy?} I said.
\Quotation{Poor little Paddy Dignam}, he said.

\BloomInt:
Off to the country:
Broadstone probably.
High brown boots with laces dangling.
Wellturned foot.
What is he foostering over that change for?
Sees me looking.
Eye out for other fellow always.
Good fallback.
Two strings to her bow.

\mccoy:
\Quotation{Why}? I said.
\Quotation{What's wrong with him?} I said.

\BloomInt:
Proud:
rich:
silk stockings.

\Bloom:
Yes,
\sout{Mr Bloom said.}

\N1:
He moved a little to the side of M'Coy's talking head.

\BloomInt:
Getting up in a minute.

\mccoy:
\Quotation{What's wrong with him?} He said.
\Quotation{He's dead}, he said.
And, faith, he filled up.
\Quotation{Is it Paddy Dignam?} I said.
I couldn't believe it when I heard it.
I was with him no later than Friday last
or Thursday was it
in the Arch.
\Quotation{Yes}, he said.
\Quotation{He's gone.
He died on Monday, poor fellow.}

\BloomInt:
Watch!
Watch!
Silk flash rich stockings white.
Watch!

\N1:
A heavy tramcar honking its gong
slewed between.

\BloomInt:
Lost it.
Curse your noisy pugnose.
Feels locked out of it.
Paradise and the peri.
Always happening like that.
The very moment.
Girl in Eustace street hallway
Monday was it
settling her garter.
Her friend covering the display of.
\french{Ésprit de corps}.
Well, what are you gaping at?

\Bloom:
Yes, yes,

\N1:
Mr Bloom said after a dull sigh.

\Bloom:
Another gone.

\mccoy:
One of the best,
\sout{M'Coy said.}

\N1:
The tram passed.

\N2:
They drove off towards the Loop Line bridge,
her rich gloved hand on the steel grip.
Flicker, flicker:
the laceflare of her hat in the sun:
flicker, flick.

\mccoy:
Wife well, I suppose?

\N1:
M'Coy's changed voice said.

\Bloom:
O, yes,
\sout{Mr Bloom said.}
Tiptop, thanks.

\N2:
He unrolled the newspaper baton idly and read idly:

\BloomInt:
\begin{verse}
    What is home without \\
    Plumtree's potted meat? \\
    Incomplete \\
    With it an abode of bliss.
\end{verse}

\mccoy:
My missus has just got an engagement.
At least it's not settled yet.

\BloomInt:
Valise tack again.
By the way no harm.
I'm off that, thanks.

\N1:
Mr Bloom turned his largelidded eyes
with unhasty friendliness.

\Bloom:
My wife too,
\sout{he said.}
She's going to sing at a swagger affair in the Ulster Hall, Belfast,
on the twenty-fifth.

\mccoy:
That so?
\sout{M'Coy said.}
Glad to hear that, old man.
Who's getting it up?

\BloomInt:
Mrs Marion Bloom.
Not up yet.
Queen was in her bedroom eating bread and.
No book.
Blackened court cards laid along her thigh by sevens.
Dark lady and fair man.
Letter.
Cat furry black ball.
Torn strip of envelope.

\BloomInt:
\begin{verse}
    Love's \\
    Old \\
    Sweet \\
    Song \\
    Comes lo-ove's old ...
\end{verse}

\Bloom:
It's a kind of a tour,
don't you see,
\sout{Mr Bloom said thoughtfully.}

\BloomInt:
\Quotation{sweeeet song}.
There's a committee formed.

\Bloom:
Part shares and part profits.

\N2:
M'Coy nodded,
picking at his moustache stubble.

\mccoy:
O, well,
\sout{he said.}
That's good news.

\N2:
He moved to go.

\mccoy:
Well, glad to see you looking fit,
\sout{he said.}
Meet you knocking around.

\Bloom:
Yes,
\sout{Mr Bloom said.}

\mccoy:
Tell you what,
\sout{M'Coy said.}
You might put down my name at the funeral,
will you?
I'd like to go but I mightn't be able, you see.
There's a drowning case at Sandycove may turn up
and then the coroner and myself would have to go down
if the body is found.
You just shove in my name if I'm not there,
will you?

\Bloom:
I'll do that,

\N1:
Mr Bloom said,
moving to get off.

\Bloom:
That'll be all right.

\mccoy:
Right,
\sout{M'Coy said brightly.}
Thanks, old man.
I'd go if I possibly could.
Well, tolloll.
Just C.P. M'Coy will do.

\Bloom:
That will be done,

\N1:
Mr Bloom answered firmly.

\BloomInt:
Didn't catch me napping that wheeze.
The quick touch.
Soft mark.
I'd like my job.
Valise I have a particular fancy for.
Leather.
Capped corners,
rivetted edges,
double action lever lock.
Bob Cowley lent him his for the Wicklow regatta concert last year
and never heard tidings of it from that good day to this.
