\N1:
By lorries along sir John Rogerson's quay
Mr Bloom walked soberly,
past Windmill lane,
Leask's the linseed crusher,
the postal telegraph office.

\BloomInt:
Could have given that address too.

\N1:
And past the sailors' home.
He turned from the morning noises of the quayside
and walked through Lime street.
By Brady's cottages
a boy for the skins lolled,
his bucket of offal linked,
smoking a chewed fagbutt.
A smaller girl with scars of eczema on her forehead
eyed him,
listlessly holding her battered caskhoop.

\BloomInt:
Tell him if he smokes he won't grow.
O let him!
His life isn't such a bed of roses.
Waiting outside pubs to bring da home.
Come home to ma, da.
Slack hour: won't be many there.

\N1:
He crossed Townsend street,
passed the frowning face of Bethel.

\BloomInt:
El, yes:
house of: Aleph, Beth.

\N1:
And past Nichols' the undertaker.

\BloomInt:
At eleven it is.
Time enough.
Daresay Corny Kelleher bagged the job for O'Neill's.
Singing with his eyes shut.
Corny.
Met her once in the park.
In the dark.
What a lark.
Police tout.
Her name and address she then told
with my tooraloom tooraloom tay.
O, surely he bagged it.
Bury him cheap in a whatyoumaycall.
With my tooraloom, tooraloom, tooraloom, tooraloom.

\N2:
In Westland row
he halted before the window of the Belfast and Oriental Tea Company
and read the legends of leadpapered packets:

\BloomInt:
choice blend, finest quality, family tea.
Rather warm.
Tea.
Must get some from Tom Kernan.
Couldn't ask him at a funeral, though.

\N2:
While his eyes still read blandly
he took off his hat
quietly inhaling his hairoil
and sent his right hand with slow grace over his brow and hair.

\BloomInt:
Very warm morning.

\N2:
Under their dropped lids
his eyes found the tiny bow of the leather headband inside his high grade ha--.

\BloomInt:
Just there.

\N2:
His right hand came down into the bowl of his hat.
His fingers found quickly a card behind the headband
and transferred it to his waistcoat pocket.

\BloomInt:
So warm.

\N1:
His right hand once more more slowly went over his brow and hair. 
hen he put on his hat again, relieved:
and read again:

\BloomInt:
choice blend, made of the finest Ceylon brands.
The far east.
Lovely spot it must be:
the garden of the world,
big lazy leaves to float about on,
cactuses, flowery meads,
snaky lianas they call them.
Wonder is it like that.
Those Cinghalese lobbing about in the sun in \italian{Dolce Far Niente},
not doing a hand's turn all day.
Sleep six months out of twelve.
Too hot to quarrel.
Influence of the climate.
Lethargy.
Flowers of idleness.
The air feeds most.
Azotes.
Hothouse in Botanic gardens.
Sensitive plants.
Waterlilies.
Petals too tired to.
Sleeping sickness in the air.
Walk on roseleaves.
Imagine trying to eat tripe and cowheel.
Where was the chap I saw in that picture somewhere?
Ah yes, in the dead sea floating on his back,
reading a book with a parasol open.
Couldn't sink if you tried:
so thick with salt.
Because the weight of the water,
no, the weight of the
body in the water
is equal to the weight of the what?
Or is it the volume
is equal to the weight?
It's a law something like that.
Vance in High school cracking his fingerjoints, teaching.
The college curriculum.
Cracking curriculum.
What is weight really when you say the weight?
Thirtytwo feet per second per second.
Law of falling bodies:
per second per second.
They all fall to the ground.
The earth.
It's the force of gravity of the earth
is the weight.

\N1:
He turned away and sauntered across the road.

\BloomInt:
How did she walk with her sausages?
Like that something.

\N1:
As he walked he took the folded \Title{Freeman} from his sidepocket,
unfolded it, rolled it lengthwise in a baton
and tapped it at each sauntering step against his trouserleg.

\BloomInt:
Careless air: just drop in to see.
Per second per second.
Per second for every second
it means.

\N1:
From the curbstone
he darted a keen glance through the door of the postoffice.

\BloomInt:
Too late box.
Post here.
No-one.
In.

\N1:
He handed the card through the brass grill.

\Bloom:
Are there any letters for me?
\sout{he asked}.

\N1:
While the postmistress searched a pigeonhole
he gazed at the recruiting poster with soldiers of all arms on parade:
and held the tip of his baton against his nostrils,
smelling freshprinted rag paper.

\BloomInt:
No answer probably.
Went too far last time.

\N1:
The postmistress handed him back through the grill
his card with a letter.
He thanked her and glanced rapidly at the typed envelope.

\N2:
Henry Flower Esq,
c/o P. O. Westland Row,
City.

\BloomInt:
Answered anyhow.

\N1:
He slipped card and letter into his sidepocket,
reviewing again the soldiers on parade.

\BloomInt:
Where's old Tweedy's regiment?
Castoff soldier.
There: bearskin cap and hackle plume.
No, he's a grenadier.
Pointed cuffs.
There he is: royal Dublin fusiliers.
Redcoats.
Too showy.
That must be why the women go after them.
Uniform.
Easier to enlist and drill.
Maud Gonne's letter about taking them off O'Connell street at night:
disgrace to our Irish capital.
Griffith's paper is on the same tack now:
an army rotten with venereal disease:
overseas or halfseasover empire.
Half baked they look:
hypnotised like.
Eyes front.
Mark time.
Table: able.
Bed: ed.
The King's own.
Never see him dressed up as a fireman or a bobby.
A mason, yes.

\N2:
He strolled out of the postoffice and turned to the right.

\BloomInt:
Talk:
as if that would mend matters.

\N2:
His hand went into his pocket
and a forefinger felt its way under the flap of the envelope,
ripping it open in jerks.

\BloomInt:
Women will pay a lot of heed, I don't think.

\N2:
His fingers drew forth the letter
the letter and crumpled the envelope in his pocket.

\BloomInt:
Something pinned on:
photo perhaps.
Hair?
No.
