\BloomInt:
How goes the time?
Quarter past.
Time enough yet.
Better get that lotion made up.
Where is this?
Ah yes, the last time.
Sweny's in Lincoln place.
Chemists rarely move.
Their green and gold beaconjars too heavy to stir.
Hamilton Long's, founded in the year of the flood.
Huguenot churchyard near there.
Visit some day.

\N1:
He walked southward along Westland row.

\BloomInt:
But the recipe is in the other trousers.
O, and I forgot that latchkey too.
Bore
this funeral affair.
O well, poor fellow,
it's not his fault.
When was it I got it made up last?
Wait.
I changed a sovereign I remember.
First of the month it must have been
or the second.
\gab{470}
O, he can look it up in the prescriptions book.

\N1:
The chemist turned back page after page.

\BloomInt:
Sandy shrivelled smell he seems to have.
Shrunken skull.
And old.
Quest for the philosopher's stone.
The alchemists.
Drugs age you after mental excitement.
Lethargy then.
Why?
Reaction.
A lifetime in a night.
Gradually changes your character.
Living all the day among herbs, ointments, disinfectants.
All his alabaster lilypots.
Mortar and pestle.
Aq.\
Dist.\
Fol.\
Laur.\
Te Virid.
Smell almost cure you like the dentist's doorbell.
Doctor Whack.
He ought to physic himself a bit.
Electuary or emulsion.
The first fellow that picked an herb to cure himself had a bit of pluck.
Simples.
Want to be careful.
\gab{480}
Enough stuff here to chloroform you.
Test: turns blue litmus paper red.
Chloroform.
Overdose of laudanum.
Sleeping draughts.
Lovephiltres.
Paragoric poppysyrup bad for cough.
Clogs the pores or the phlegm.
Poisons the only cures.
Remedy where you least expect it.
Clever of nature.

\sweny:
About a fortnight ago, sir?

\Bloom:
Yes,
\sout{Mr Bloom said.}

\N1:
He waited by the counter,
inhaling slowly the keen reek of drugs,
the dusty dry smell of sponges and loofahs.

\BloomInt:
Lot of time taken up telling your
aches and pains.

\Bloom:
Sweet almond oil and tincture of benzoin,
\sout{Mr Bloom said,}
\gab{490}
and then orangeflower water ...

\BloomInt:
It certainly did make her skin so delicate white like wax.

\Bloom:
And white wax also,
\sout{he said.}

\BloomInt:
Brings out the darkness of her eyes.
Looking at me,
the sheet up to her eyes,
Spanish,
smelling herself,
when I was fixing the links in my cuffs.
Those homely recipes are often the best:
strawberries for the teeth:
nettles and rainwater:
oatmeal they say steeped in buttermilk.
Skinfood.
One of the old queen's sons,
duke of Albany was it?
had only one skin.
Leopold, yes.
Three we have.
Warts, bunions and pimples to make it worse.
But you want a perfume too.

\martha:
What perfume does your?
\gab{500}

\BloomInt:
\french{Peau d'Espagne}.

\Molly:
That orangeflower water is so fresh.

\BloomInt:
Nice smell these soaps have.
Pure curd soap.
Time to get a bath round the corner.
Hammam.
Turkish.
Massage.
Dirt gets rolled up in your navel.
Nicer if a nice girl did it.
Also I think I.
Yes I.
Do it in the bath.
Curious longing I.
Water to water.
Combine business with pleasure.
Pity no time for massage.
Feel fresh then all the day.
Funeral be rather glum.

\sweny:
Yes, sir,
\sout{the chemist said.}
That was two and nine.
Have you brought a bottle?

\Bloom:
No,
\sout{Mr Bloom said.}
Make it up, please.
I'll call later in the day
and I'll take one of these soaps.
\gab{510}
How much are they?

\sweny:
Fourpence, sir.

\N1:
Mr Bloom raised a cake to his nostrils.

\BloomInt:
Sweet lemony wax.

\Bloom:
I'll take this one,
\sout{he said.}
That makes three and a penny.

\sweny:
Yes, sir,
\sout{the chemist said.}
You can pay all together, sir, when you come back.

\Bloom:
Good,
\sout{Mr Bloom said.}

\N1:
He strolled out of the shop,
the newspaper baton under his armpit,
the coolwrappered soap in his left hand.

\N1:
At his armpit |
Bantam Lyons' voice and hand said:

\lyons:
Hello, Bloom.
What's the best news?
\gab{520}
Is that today's?
Show us a minute.

\BloomInt:
Shaved off his moustache again, by Jove!
Long cold upper lip.
To look younger.
He does look balmy.
Younger than I am.

\N1:
Bantam Lyons's yellow blacknailed fingers unrolled the baton.

\BloomInt:
Wants a wash too.
Take off the rough dirt.
Good morning, have you used Pears' soap?
Dandruff on his shoulders.
Scalp wants oiling.

\lyons:
I want to see about that French horse that's running today,
\sout{Bantam Lyons said.}
Where the bugger is it?

\N1:
He rustled the pleated pages,
jerking his chin on his high collar.

\BloomInt:
Barber's itch.
Tight collar | he'll lose his hair.
Better leave him the paper and get shut of him.
\gab{530}

\Bloom:
You can keep it,
\sout{Mr Bloom said.}

\lyons:
Ascot.
Gold cup.
Wait,
\sout{Bantam Lyons muttered.}
Half a mo.
Maximum the second.

\Bloom:
I was just going to throw it away,
\sout{Mr Bloom said.}

\N1:
Bantam Lyons raised his eyes suddenly and leered weakly.

\lyons:
What's that?

\N1:
his sharp voice said.

\Bloom:
I say you can keep it,
\sout{Mr Bloom answered.}
I was going to throw it away that moment.

\N1:
Bantam Lyons doubted an instant, leering:
then thrust the outspread sheets back on Mr Bloom's arms.
\gab{540}

\lyons:
I'll risk it,
\sout{he said.}
Here, thanks.

\N1:
He sped off towards Conway's corner.

\BloomInt:
God speed | scut.

\N1:
Mr Bloom folded the sheets again to a neat square
and lodged the soap in it,
smiling.

\BloomInt:
Silly lips of that chap.
Betting.
Regular hotbed of it lately.
Messenger boys stealing to put on sixpence.
Raffle for large tender turkey.
Your Christmas dinner for threepence.
Jack Fleming embezzling to gamble
then smuggled off to America.
Keeps a hotel now.
They never come back.
Fleshpots of Egypt.

\N2:
He walked cheerfully towards the mosque of the baths.

\BloomInt:
Remind you of a mosque,
\gab{550}
redbaked bricks, the minarets.
College sports today I see.

\N2:
He eyed the horseshoe poster over the gate of college park:
cyclist doubled up like a cod in a pot.

\BloomInt:
Damn bad ad.
Now if they had made it round like a wheel.
Then the spokes:
sports, sports, sports: and the hub big:
college.
Something to catch the eye.

\BloomInt:
There's Hornblower standing at the porter's lodge.
Keep him on hands:
might take a turn in there on the nod.
How do you do, Mr Hornblower?
How do you do, sir?

\BloomInt:
Heavenly weather really.
If life was always like that.
Cricket weather.
Sit around under sunshades.
Over after over.
Out.
They can't play it here.
Duck for six wickets.
\gab{560}
Still Captain Culler broke a window in the Kildare street club
with a slog to square leg.
Donnybrook fair more in their line.
And the skulls we were acracking when M'Carthy took the floor.
Heatwave.
Won't last.
Always passing,
the stream of life,
which in the stream of life we trace
is dearer than them all.

\BloomInt:
Enjoy a bath now:
clean trough of water,
cool enamel,
the gentle tepid stream.
This is my body.

\N1:
He foresaw his pale body reclined in it at full,
naked,
in a womb of warmth,
oiled by scented melting soap,
softly laved.
He saw his trunk and limbs riprippled over
and sustained,
buoyed lightly upward,
lemonyellow:
his navel, bud of flesh:
\gab{570}
and saw the dark tangled curls of his bush
floating,
floating hair of the stream
around the limp father of thousands,
a languid floating flower.
