\N1:
Bronze by gold |
heard the hoofirons,
steelyringing

\boots:
Imper·thn·thn thn·thn·thn.

\simon:
Chips,
picking chips off rocky thumbnail,
chips.

\MissK:
Horrid!

\N2:
And gold flushed more.

\N3:
A husky fife·note blew.

\Bloom:
Blew.
Blue bloom is on the

\N2:
Gold pinnacled hair.

\Bloom:
A jumping rose on satiny breast of satin,
rose of Castile.

\MissD:
Trilling, trilling:
Idolores.

\lenehan:
Peep!
Who's in the ...
\gab{10}
peep·of·gold?

\N1:
Tink cried to bronze in pity.

\N2:
And a call,
pure,
long and throbbing.

\Bloom:
Long·in·dying call.

\Bloom:
Decoy.
Soft word.
But look:
the bright stars fade.
Notes chirruping answer.
O rose!
Castile.
The morn is breaking.

\boylan:
Jingle jingle jaunted jingling.

\Bloom:
Coin rang.
Clock clacked.

\lenehan:
Avowal.
\french{Sonnez}.
I could.
Rebound of garter.
Not leave thee.
Smack.
\french{La cloche!}
Thigh smack.
Avowal.
Warm.
Sweetheart, goodbye!

\boylan:
Jingle.

\Bloom:
Bloo.

\N1:
Boomed crashing chords.
\gab{20}

\Bloom:
When love absorbs.
War!
War!
The tympanum.

\Bloom:
A sail!
A veil awave upon the waves.

\Bloom:
Lost.
Throstle fluted.
All is lost now.

\boylan:
Horn.
Hawhorn.

\N4:
When first he saw.

\Bloom:
Alas!

\Bloom:
Full tup.
Full throb.

\N4:
Warbling.

\Bloom:
Ah, lure!
Alluring.

\Bloom:
Martha!
Come!

\N1:
Clap·clop.
Clip·clap.
Clappy·clap.

\goulding:
Good·god he·never·heard in·all.

\N4:
Deaf bald Pat brought pad knife
\gab{30}
took up.

\Bloom:
A moonlit nightcall:
far:
far.

\Bloom:
I feel so sad.
P.S. So lonely blooming.

\N3:
Listen!

\Bloom:
The spiked and winding cold seahorn.
Have you the?
Each,
and for other,
plash and silent roar.

\Bloom:
Pearls:
when she.
Liszt's rhapsodies.

\N1:
Hissss.

\Bloom:
You don't?

\Bloom:
Did not:
no, no:
believe:
Lid·lyd.

\N1:
With a cock |
with a carra.

\Bloom:
Black.
Deep·sounding.

\simon:
Do, Ben, do.

\Bloom:
Wait while you wait.
Hee hee.
\gab{40}
Wait while you hee.

\Bloom:
But wait!

\Bloom:
Low in dark middle earth.
Embedded ore.

\dollard:
\latin{Namine·damine.}

\Bloom:
All gone.
All fallen.

\simon:
Tiny,
her tremulous fernfoils of maidenhair.

\N2:
Amen!
He gnashed in fury.

\N1:
Fro.
To, fro.
A baton cool protruding.

\Bloom:
Bronze·lydia by Mina·gold.

\N1:
By bronze,
by gold,
in oceangreen of shadow.
Bloom.
Old Bloom.

\N3:
One rapped,
one tapped,
\gab{50}
with a carra,
with a cock.

\Bloom:
Pray for him!
Pray, good people!

\dollard:
His gouty fingers nakkering.

\simon:
Big Ben·a·ben.
Big Ben·ben.

\Bloom:
Last rose Castile of summer |
left bloom |
I feel so sad |
alone.

\Bloom:
Pwee!
Little wind piped wee.

\Bloom:
True men.
Lid |
Ker |
Cow |
De |
and Doll.
Ay, ay.
Like you men.
Will lift your tschink with tschunk.

\N3:
Fff!
Oo!

\Bloom:
Where bronze from anear?
Where gold from afar?
Where hoofs?

\N3:
Rrrpr.
Kraa.
\gab{60}
Kraandl.

\emmet:
\emph{Then not till then.}
My epp·ri·pff·taph.
Be pfr·writt.

\Bloom:
Done.

\Bloom:
Begin!

\begin{center}
  $\clubsuit$~~~$\clubsuit$~~~$\clubsuit$
\end{center}

\pagebreak[2]

% -------------------------------------------

\N1:
Bronze by gold,
miss Douce's head by miss Kennedy's head,
over the crossblind of the Ormond bar |
heard the viceregal hoofs go by,
ringing steel.

\MissK:
Is that her?

\N2:
asked miss Kennedy.

\N1:
Miss Douce said

\MissD:
yes,
sitting with his ex,
pearl grey and \french{eau de Nil}

\MissK:
Exquisite contrast,
\sout{miss Kennedy said.}

\N1:
When all agog |
miss Douce said eagerly:

\MissD:
Look at the fellow
\gab{70}
in the tall silk.

\MissK:
Who?
Where?

\N2:
gold asked more eagerly.

\MissD:
In the second carriage.

\N1:
miss Douce's wet lips said,
laughing in the sun.

\MissD:
He's looking.
Mind till I see.

\N1:
She darted,
bronze,
to the backmost corner,
flattening her face against the pane |
in a halo of hurried breath.

\N1:
Her wet lips tittered:

\MissD:
He's killed looking back.
\stage{[laughs]}
\sout{She laughed:}

\MissD:
O wept!
Aren't men frightful idiots?

\N2:
With sadness.
\gab{80}

\N2:
Miss Kennedy sauntered sadly from bright light,
twining a loose hair behind an ear.
Sauntering sadly,
gold no more,
she twisted twined a hair.
Sadly |
she twined in sauntering |
gold hair behind a curving ear.

\MissK:
It's them has the fine times,

\N2:
sadly then she said.

\N3:
A man.

\N3:
Bloo·who went by by Moulang's pipes |
bearing in his breast the sweets of sin,
by Wine's antiques,
in memory bearing sweet sinful words,
by Carroll's dusky battered plate,
for Raoul.

\N4:
The boots to them,
them in the bar,
them barmaids came.
For them unheeding him |
\gab{90}
he banged on the counter |
his tray of chattering china.
And

\boots:
There's your teas,
\sout{he said.}

\N2:
Miss Kennedy with manners |
transposed the teatray |
down to an upturned lithia crate,
safe from eyes,
low.

\boots:
What is it?

\N4:
loud boots unmannerly asked.

\MissD:
Find out,

\N1:
miss Douce retorted,
leaving her spyingpoint.

\boots:
Your \french{beau},
is it?

\N1:
A haughty bronze replied:

\MissD:
I'll complain to Mrs de Massey on you |
if I hear any more of your impertinent insolence.

\boots:
Imper·thn·thn thn·thn·thn,

\N4:
boots·snout sniffed rudely,
\gab{100}
as he retreated |
as she threatened |
as he had come.

\N3:
Bloom.

\N1:
On her flower frowning |
miss Douce said:

\MissD:
Most aggravating |
that young brat is.
If he doesn't conduct himself |
I'll wring his ear for him |
a yard long.

\N2:
Ladylike in exquisite contrast.

\MissK:
Take no notice,

\N2:
miss Kennedy rejoined.

\N2:
She poured in a teacup |
tea,
then back in the teapot |
tea.
They cowered under their reef of counter,
waiting on footstools,
crates upturned,
waiting for their teas to draw.
\gab{110}
They pawed their blouses,
both of black satin,
two and nine a yard,
waiting for their teas to draw,
and two and seven.

\N1:
Yes,
bronze from anear,
by gold from afar,
heard steel from anear,
hoofs ring from afar,
and heard steelhoofs |
ringhoof |
ringsteel.

\MissD:
Am I awfully sunburnt?

\N1:
Miss bronze unbloused her neck.

\MissK:
No,
\sout{said Miss Kennedy.}
It gets brown after.
Did you try the borax with the cherry laurel water?

\N1:
Miss Douce halfstood |
to see her skin askance in the barmirror |
gilded·lettered where hock and claret glasses shimmered |
and in their midst |
a shell.
\gab{120}

\MissD:
And leave it to my hands,
\sout{she said}

\MissK:
Try it with the glycerine,
\sout{miss Kennedy advised.}

\N1:
Bidding her neck and hands adieu |
miss Douce

\MissD:
Those things only bring out a rash,

\N1:
replied,
reseated.

\MissD:
I asked that old fogey in Boyd's |
for something for my skin.

\N2:
Miss Kennedy,
pouring now fulldrawn tea,
grimaced and prayed:

\MissK:
O, don't remind me of him |
for mercy' sake!

\MissD:
But wait till I tell you.

\N1:
miss Douce entreated.

\N2:
Sweet tea |
miss Kennedy having poured with milk |
plugged both two ears
\gab{130}
with little fingers.

\MissK:
No,
don't,
\sout{she cried.}

\MissK:
I won't listen,
\sout{she cried.}

\N3:
But Bloom?

\N1:
Miss Douce grunted in snuffy fogey's tone:

\MissD:
For your what?
says he.

\N2:
Miss Kennedy unplugged her ears |
to hear,
to speak:
but said,
but prayed again:

\MissK:
Don't let me think of him or I'll expire.
The hideous old wretch!
That night in the Antient Concert Rooms.

\N2:
She sipped distastefully her brew,
hot tea,
\gab{140}
a sip,
sipped sweet tea.

\MissD:
Here he was.

\N1:
miss Douce said,
cocking her bronze head three quarters,
ruffling her nosewings.

\MissD:
Hufa!
Hufa!

\N2:
Shrill shriek of laughter |
sprang from miss Kennedy's throat.

\N1:
Miss Douce huffed and snorted down her nostrils |
that quivered imper·thn·thn |
like a snout in quest.

\MissK:
O!
\stage{[shrieking]}
\sout{miss Kennedy cried.}
Will you ever forget his goggle eye?

\N1:
Miss Douce chimed in in deep bronze laughter,
shouting:

\MissD:
And your other eye!

\N3:
Bloo·whose dark eye |
read Aaron Figatner's name.

\BloomInt:
Why do I always think Figather?
Gathering figs,
I think.
\gab{150}
And Prosper Loré's huguenot name.

\N3:
By Bassi's blessed virgins |
Bloom's dark eyes went by.

\BloomInt:
Bluerobed,
white under,
come to me.
God they believe she is:
or goddess.
Those today.
I could not see.
That fellow spoke.
A student.
After with Dedalus' son.
He might be Mulligan.
All comely virgins.
That brings those rakes of fellows in:
her white.

\N3:
By went his eyes.

\BloomInt:
The sweets of sin.
Sweet are the sweets.

\N3:
Of sin.

\N2:
In a giggling peal |
young gold·bronze voices blended,
Douce with Kennedy

\MissesDK:
your other eye.

\N1:
They threw young heads back,
bronze giggle·gold,
to let free·fly their laughter,
screaming,

\MissesDK:
your other,
\gab{160}

\N2:
signals to each other,
high piercing notes.

\MissesDK:
Ah,

\N2:
panting,
sighing.

\N1:
Sighing,

\MissesDK:
ah,

\N1:
fordone |
their mirth died down.

\N2:
Miss Kennedy lipped her cup again,
raised,
drank a sip and giggle·giggled.

\N1:
Miss Douce,
bending over the teatray,
ruffled again her nose
and rolled droll fattened eyes.

\N2:
Again Kenny·giggles,
stooping,
her fair pinnacles of hair,
stooping,
her tortoise napecomb showed,
spluttered out of her mouth her tea,
choking in tea and laughter,
coughing with choking,
crying:

\MissK:
O greasy eyes!
Imagine being married to a man like that,
\sout{she cried.}
With his bit
\gab{170}
of beard!

\N1:
Douce gave full vent to a splendid yell,
a full yell of full woman,
delight,
joy,
indignation.

\MissD:
Married to the greasy nose!
\sout{she yelled.}

\N1:
Shrill,
with deep laughter,
after,
gold after bronze,
they urged each |
each to peal after peal,
ringing in changes,
bronze·gold,
gold·bronze,
shrill·deep,
to laughter after laughter.
And then laughed more.

\MissK:
Greasy I knows.

\N2:
Exhausted,
breathless,
their shaken heads they laid,
braided and pinnacled by glossy·combed,
against the counter·ledge.
All flushed

\MissesDK:
(O!),

\N2:
panting,
sweating

\MissesDK:
(O!),

\N2:
all breathless.

\N3:
Married to Bloom,
\gab{180}
to grease·a·bloom.

\MissD:
O saints above!

\N1:
miss Douce said,
sighed above her jumping rose.

\MissD:
I wished I hadn't laughed so much.
I feel all wet.

\MissK:
O, miss Douce!
\sout{miss Kennedy protested.}
You horrid thing!

\N2:
And flushed yet more

\MissK:
(you horrid!)

\N2:
more goldenly.

\N3:
By Cantwell's offices |
roved Grease·a·bloom,
by Ceppi's virgins,
bright of their oils.

\BloomInt:
Nannetti's father hawked those things about,
wheedling at doors as~I\@.
Religion pays.
Must see him for that par.
Eat first.
I want.
Not yet.
At four,
she said.
Time ever passing.
Clockhands turning.
On.
Where eat?
The Clarence,
Dolphin.
On.
For Raoul.
Eat.
If I net five guineas with those ads.
\gab{190}
The violet silk petticoats.
Not yet.
The sweets of sin.

\N2:
Flushed less,
still less,
goldenly paled.

\N3:
Into their bar |
strolled Mr Dedalus.

\N3:
Chips,
picking chips off one of his rocky thumbnails.
Chips.

\N3:
He strolled.

\simon:
O,
welcome back,
miss Douce.

\N3:
He held her hand.
Enjoyed her holidays?

\MissD:
Tiptop.

\N3:
He hoped she had nice weather in Rostrevor.

\MissD:
Gorgeous.
\sout{she said.}
Look at the holy show I am.
Lying out on the strand
all day.

\N1:
Bronze whiteness.
\gab{200}

\simon:
That was exceedingly naughty of you.

\N3:
Mr Dedalus told her and pressed her hand indulgently.

\simon:
Tempting poor simple males.

\N1:
Miss Douce of satin douced her arm away.

\MissD:
O go away!
\sout{she said.}
You're very simple,
I don't think.

\N3:
He was.

\simon:
Well now,
I am,
\sout{he mused.}
I looked so simple in the cradle |
they christened me simple Simon.

\MissD:
You must have been a doaty.
\sout{miss Douce made answer.}
And what did the doctor order today?

\simon:
Well now,
\sout{he mused,}
whatever you say yourself.
\gab{210}
I think I'll trouble you |
for some fresh water
and a half glass of whisky.

\N4:
Jingle.

\MissD:
With the greatest alacrity,
\sout{miss Douce agreed.}

\N1:
With grace of alacrity |
towards the mirror |
gilt Cantrell and Cochrane's |
she turned herself.
With grace |
she tapped a measure of gold whisky
from her crystal keg.

\N3:
Forth from the skirt of his coat |
Mr Dedalus brought pouch and pipe.

\N1:
Alacrity |
she served.

\N3:
He blew through the flue |
two husky fife·notes.

\simon:
By Jove,
\sout{he mused.}
I often wanted to see the Mourne mountains.
Must be a great tonic in the air down there.
\gab{220}
But a long threatening comes at last,
they say.
Yes.
Yes.

\N3:
Yes.
He fingered shreds of hair,
her maidenhair,
her mermaid's,
into the bowl.
Chips.
Shreds.
Musing.
Mute.

\N3:
None nought said nothing.
Yes.

\N1:
Gaily miss Douce polished a tumbler,
trilling:

\MissD:
\song{O, Idolores,
queen of the eastern seas!}

\lenehan:
Was Mr Lidwell in today?

\N4:
In came Lenehan.
Round him peered Lenehan.

\N3:
Mr Bloom reached Essex bridge.
Yes,
Mr Bloom crossed bridge of Yessex.

\BloomInt:
To Martha I must write.
Buy paper.
Daly's.
Girl there civil.
\gab{230}

\N3:
Bloom.
Old Bloom.
Blue bloom is on the rye.

\MissD:
He was in at lunchtime,
\sout{miss Douce said.}

\N4:
Lenehan came forward.

\lenehan:
Was Mr Boylan looking for me?

\sout{He asked.
She answered:}

\MissD:
Miss Kennedy,
was Mr Boylan in |
while I was upstairs?

\sout{She asked.}

\N2:
Miss voice of Kennedy answered,
a second teacup poised,
her gaze upon a page:

\MissK:
No.
He was not.

\N2:
Miss gaze of Kennedy,
heard,
not seen,
read on.
\gab{240}

\N4:
Lenehan |
round the sandwich·bell |
wound his round body |
round.

\lenehan:
Peep!
Who's in the corner?

\N4:
No glance of Kennedy rewarding him |
he yet made overtures.

\N2:
To mind her stops.
To read only the black ones:
round o and crooked ess.

\N4:
Jingle jaunty jingle.

\N2:
Girl·gold she read and did not glance.

\MissKInt:
Take no notice.

\N2:
She took no notice |
while he read by rote |
a solfa fable for her,
plappering flatly:

\lenehan:
Ah fox met ah stork.
Said thee fox too thee stork:
Will you put your bill down inn my troath
and pull upp ah bone?

\N4:
He droned in vain.
\gab{250}

\N1:
Miss Douce turned to her tea aside.

\N4:
He sighed aside:

\lenehan:
Ah me!
O my!

\N4:
He greeted Mr Dedalus and got a nod.

\lenehan:
Greetings from the famous son of a famous father.

\simon:
Who may he be?
\sout{Mr Dedalus asked.}

\N4:
Lenehan opened most genial arms.

\lenehan:
Who?

\lenehan:
Who may he be?
\sout{he asked.}
Can you ask?
Stephen,
the youthful bard.

\N3:
Dry.

\N3:
Mr Dedalus,
famous father,
laid by his dry filled pipe.

\simon:
I see,
\gab{260}
\sout{he said.}
I didn't recognise him for the moment.
I hear he is keeping very select company.
Have you seen him lately?

\N4:
He had.

\lenehan:
I quaffed the nectarbowl with him this very day,
\sout{said Lenehan.}
In Mooney's \french{en ville}
and in Mooney's \french{sur mer}.
He had received the rhino for the labour of his muse.

\N4:
He smiled at bronze's teabathed lips,
at listening lips and eyes:

\lenehan:
The \french{élite} of Erin hung upon his lips.
The ponderous pundit,
Hugh MacHugh,
Dublin's most brilliant scribe and editor |
and that minstrel boy of the wild wet west |
who is known by
the euphonious appellation |
of the O'Madden
\gab{270}
Burke.

\N3:
After an interval |
Mr Dedalus raised his grog and

\simon:
That must have been highly diverting,
\sout{said he.}
I see.

\N3:
He see.
He drank.
With faraway mourning mountain eye.
Set down his glass.

\N3:
He looked towards the saloon door.

\simon:
I see you have moved the piano.

\MissD:
The tuner was in today.
\sout{miss Douce replied,}
tuning it for the smoking concert |
and I never heard such an exquisite player.

\simon:
Is that a fact?

\MissD:
Didn't he,
miss Kennedy?
\gab{280}
The real classical,
you know.
And blind too,
poor fellow.
Not twenty |
I'm sure he was.

\simon:
Is that a fact?
\sout{Mr Dedalus said.}

\N3:
He drank and strayed away.

\MissD:
So sad to look at his face.

\N1:
miss Douce condoled.

\stripling:
God's curse on bitch's bastard.

\N1:
Tink to her pity |
cried a diner's bell.

\N4:
To the door of the bar and diningroom |
came bald Pat,
came bothered Pat,
came Pat,
waiter of Ormond.
Lager for diner.

\N1:
Lager without alacrity |
she served.

\N4:
With patience  |
Lenehan waited for Boylan |
with impatience,
for jinglejaunty
\gab{290}
blazes boy.

\N3:
Upholding the lid |
he
[\stage{Simon}]

\N1:
(who?)

\N3:
gazed in the coffin

\N1:
(coffin?)

\N3:
at the oblique triple

\N1:
(piano!)

\N3:
wires.
He pressed

\N1:
(the same who pressed indulgently her hand),

\N3:
soft pedalling,
a triple of keys |
to see the thicknesses of felt advancing,
to hear the muffled hammerfall in action.

\BloomInt:
Two sheets cream vellum paper |
one reserve |
two envelopes |
when I was
in Wisdom Hely's

\N3:
wise Bloom |
in Daly's Henry Flower bought.

\BloomInt:
Are you not happy in your home?
Flower to console me and a pin cuts lo.
Means something,
language of flow.
Was it a daisy?
Innocence that is.
Respectable girl meet after mass.
Thanks awfully muchly.

\N3:
Wise Bloom eyed on the door |
a poster,
\gab{300}
a swaying mermaid smoking |
mid nice waves.

\BloomInt:
Smoke mermaids,
coolest whiff of all.
Hair streaming:
lovelorn.
For some man.
For Raoul.

\N3:
He eyed |
and saw afar on Essex bridge |
a gay hat riding on a jaunting car.

\BloomInt:
It is.
Again.
Third time.
Coincidence.

\N4:
Jingling on supple rubbers |
it jaunted from the bridge to Ormond quay.

\BloomInt:
Follow.
Risk it.
Go quick.
At four.
Near now.
Out.

\shopgirl:
Twopence,
sir.

\N2:
the shopgirl dared to say.

\Bloom:
Aha ...
I was forgetting ...
Excuse ...

\shopgirl:
And four.

\BloomInt:
At four |
she.

\N2:
Winsomely |
she on Bloo·him·whom smiled.

\N3:
Bloo smi qui go.
Ternoon.

\BloomInt:
Think you're the only pebble on the beach?
\gab{310}
Does that to all.
For men.

\N2:
In drowsy silence |
gold bent on her page.

\N3:
From the saloon |
a call came,
long in dying.
That was a tuningfork the tuner had |
that he forgot |
that he now struck.
A call again.
That he now poised |
that it now throbbed.

\simon:
You hear?

\N3:
It throbbed,
pure,
purer,
softly and softlier,
its buzzing prongs.
Longer in dying call.

\N4:
Pat paid for diner's popcorked bottle:
and over tumbler, tray and popcorked bottle |
ere he went |
he whispered,
bald and bothered,
with Miss Douce.

\simon:
\song{The bright stars fade ...}
\gab{320}

\N3:
A voiceless song sang from within,
singing:

\simon:
\song{...
the morn is breaking.}

\N3:
A duodene of birdnotes |
chirruped bright treble answer |
under sensitive hands.
Brightly the keys,
all twinkling,
linked,
all harpsichording,
called to a voice |
to sing the strain of dewy morn,
of youth,
of love's leavetaking,
life's,
love's morn.

\simon:
\song{the dewdrops pearl ...}

\N4:
Lenehan's lips over the counter |
lisped a low whistle of decoy.

\lenehan:
But look this way,
\sout{he said,}
rose of Castile.

\N4:
Jingle jaunted by the curb |
\gab{330}
and stopped.

\N2:
She rose and closed her reading,
rose of Castile:
fretted,
forlorn,
dreamily rose.

\lenehan:
Did she fall or was she pushed?
\sout{he asked her.}

\N2:
She answered,
slighting:

\MissK:
Ask no questions and you'll hear no lies.

\N2:
Like lady,
ladylike.

\N4:
Blazes Boylan's smart tan shoes |
creaked on the barfloor |
where he strode.

\N2:
Yes,
gold from anear |
by bronze from afar.

\N4:
Lenehan heard |
and knew |
and hailed him:

\lenehan:
See the conquering hero comes.
\gab{340}

\N3:
Between the car and window,
warily walking,
went Bloom,
unconquered hero.

\BloomInt:
See me |
he might.
The seat he sat on:
warm.

\N4:
Black wary he·cat |
walked towards Richie Goulding's legal bag,
lifted aloft,
saluting.

\simon:
\song{And I from thee ...}

\boylan:
I heard you were round,

\N4:
said Blazes Boylan.

\N4:
He touched to fair miss Kennedy |
a rim of his slanted straw.

\N2:
She smiled on him.

\N1:
But sister bronze outsmiled her,
preening for him |
her richer hair,
a bosom
and a rose.

\N4:
Smart Boylan bespoke potions.

\boylan:
What's your cry?
\gab{350}
Glass of bitter?
Glass of bitter, please,
and a sloe·gin for me.
Wire in yet?

\BloomInt:
Not yet.
At four |
she.
Who said four?

\N2:
Cowley's red lugs and bulging apple |
in the door of the sheriff's office.

\BloomInt:
Avoid.
Goulding a chance.
What is he doing in the Ormond?
Car waiting.
Wait.

\Bloom:
Hello.

\goulding:
Where off to?

\Bloom:
Something to eat?

\goulding:
I too was just.

\Bloom:
In here.

\goulding:
What,
Ormond?

\Bloom:
Best value in Dublin.

\goulding:
Is that so?

\Bloom:
Dining·room.

\Bloom:
Sit tight there.

\BloomInt:
See,
not be seen.

\goulding:
I think I'll join you.
Come on.

\N1:
Richie led on.

\N3:
Bloom followed bag.

\BloomInt:
Dinner fit for a prince.

\N1:
Miss Douce reached high to take a flagon,
\gab{360}
stretching her satin arm,
her bust,
that all but burst,
so high.

\lenehan:
O!
O!

\N4:
jerked Lenehan,
gasping at each stretch.

\lenehan:
O!

\N1:
But easily she seized her prey |
and led it low in triumph.

\boylan:
Why don't you grow?

\N4:
asked Blazes Boylan.

\N1:
She·bronze,
dealing from her oblique jar |
thick syrupy liquor for his lips,
looked as it flowed

\BloomInt:
(flower in his coat:
who gave him?),

\N1:
and syrupped with her voice:

\MissD:
Fine goods in small parcels.

\N1:
That is to say |
she.
Neatly |
she poured slow·syrupy sloe.

\boylan:
Here's fortune,
\gab{370}
\sout{Blazes said.}

\N4:
He pitched a broad coin down.
Coin rang.

\lenehan:
Hold on,
\sout{said Lenehan,}
till I ...

\lenehan:
Fortune.

\N4:
he wished,
lifting his bubbled ale.

\lenehan:
Sceptre will win in a canter.

\boylan:
I plunged a bit,

\N4:
said Boylan |
winking and drinking.

\boylan:
Not on my own,
you know.
Fancy of a friend of mine.

\N4:
Lenehan still drank |
and grinned at his tilted ale |
and

\N1:
at miss Douce's lips |
that all but hummed,
not shut,
the oceansong |
her lips had trilled.

\MissD:
\song{Idolores.
The eastern seas.}

\N4:
Clock whirred.
\gab{380}

\N2:
Miss Kennedy passed their way

\BloomInt:
(flower,
wonder who gave),

\N2:
bearing away teatray.

\N4:
Clock clacked.

\N1:
Miss Douce took Boylan's coin,
struck boldly the cashregister.

\N3:
It clanged.

\N4:
Clock clacked.

\N1:
Fair one of Egypt |
teased and sorted in the till |
and hummed and handed coins in change.

\BloomInt:
Look to the west.

\N4:
A clack.

\BloomInt:
For me.

\boylan:
What time is that?
Four?

\N4:
O'clock.

\N4:
Lenehan,
small eyes a·hunger on her humming,
bust a·humming,
tugged Blazes Boylan's elbow·sleeve.

\lenehan:
Let's hear the time,
\sout{he said.}

\N1:
The bag of Goulding, Collis, Ward |
\gab{390}
led Bloom |
by rye·bloom flowered tables.

\N3:
Aimless |
he chose with agitated aim,
bald Pat attending,
a table near the door.

\BloomInt:
Be near.
At four.
Has he forgotten?
Perhaps a trick.
Not come:
whet appetite.
I couldn't do.
Wait,
wait.

\N4:
Pat,
waiter,
waited.

\N1:
Sparkling bronze azure |
eyed Blazure's skyblue bow and eyes.

\lenehan:
Go on,

\N4:
pressed Lenehan.

\lenehan:
There's no-one.
He never heard.

\simon:
\song{...
to Flora's lips did hie.}

\N3:
High,
a high note pealed in the treble clear.

\N1:
Bronze·douce |
communing with her rose that sank and rose |
sought Blazes Boylan's flower and eyes.

\lenehan:
Please,
please.
\gab{400}

\N4:
He pleaded over returning phrases of avowal.

\simon:
\song{I could not leave thee ...}

\MissD:
Afterwits,

\N1:
Miss Douce promised coyly.

\lenehan:
No,
now,
\sout{urged Lenehan.}
\french{Sonnez la cloche!}
O do!
There's no-one.

\N1:
She looked.
Quick.
Miss Kenn out of earshot.
Sudden bent.

\N4:
Two kindling faces watched her bend.

\N3:
Quavering |
the chords strayed from the air,
found it again,
lost chord,
and lost and found it,
faltering.

\lenehan:
Go on!
Do!
\french{Sonnez!}

\N1:
Bending,
she nipped a peak of skirt above her knee.
\gab{410}
Delayed.
Taunted them still,
bending,
suspending,
with wilful eyes.

\lenehan:
\french{Sonnez!}

\N1:
Smack.
She set free |
sudden in rebound |
her nipped elastic garter |
smack·warm against her smackable |
a woman's warm·hosed thigh.

\lenehan:
\french{La cloche!}
\sout{cried gleeful Lenehan.}
Trained by owner.
No sawdust there.

\N1:
She smile·smirked supercilious

\MissDInt:
(wept!
aren't men?),

\N1:
but,
lightward gliding,
mild |
she smiled on Boylan.

\MissD:
You're the essence of vulgarity,

\N1:
she in gliding said.

\N4:
Boylan,
eyed,
eyed.
Tossed to fat lips |
his chalice,
drank off his chalice tiny,
sucking the last fat violet syrupy drops.
\gab{420}
His spellbound eyes went after,
after her gliding head |
as it went down the bar by mirrors,
gilded arch for ginger ale,
hock and claret glasses shimmering,
a spiky shell,
where it concerted,
mirrored,
bronze with sunnier bronze.

\N3:
Yes,
bronze from anearby.

\simon:
\song{...
sweetheart,
goodbye!}

\boylan:
I'm off,

\N4:
said Boylan with impatience.

\N4:
He slid his chalice brisk away,
grasped his change.

\lenehan:
Wait a shake,

\N4:
begged Lenehan,
drinking quickly.

\lenehan:
I wanted to tell you.
Tom Rochford ...

\boylan:
Come on to blazes,
\gab{430}

\N4:
said Blazes Boylan,
going.

\N4:
Lenehan gulped to go.

\lenehan:
Got the horn or what?
\sout{he said.}
Wait.
I'm coming.

\N4:
He followed the hasty creaking shoes |
but stood by nimbly by the threshold,
saluting forms,
a bulky with a slender.

\lenehan:
How do you do,
Mr Dollard?

\dollard:
Eh?
How do?
How do?

\N4:
Ben Dollard's vague bass answered,
turning an instant from Father Cowley's woe.

\dollard:
He won't give you any trouble,
Bob.
Alf Bergan will speak to the long fellow.
We'll put a barley·straw in that Judas Iscariot's ear
this time.

\N3:
Sighing |
Mr Dedalus came through the saloon,
\gab{440}
a finger soothing an eyelid.

\dollard:
Hoho,
we will,

\N4:
Ben Dollard yodled jollily.

\dollard:
Come on,
Simon.
Give us a ditty.
We heard the piano.

\N4:
Bald Pat,
bothered waiter,
waited for drink orders.

\N1:
Power for Richie.

\N3:
And Bloom?

\BloomInt:
Let me see.
Not make him walk twice.
His corns.
Four now.
How warm this black is.
Course nerves a bit.
Refracts (is it?) heat.

\Bloom:
Let me see.
Cider.
Yes,
bottle of cider.

\simon:
What's that?
\sout{Mr Dedalus said.}
I was only vamping,
man.

\dollard:
Come on,
come on,
\sout{Ben Dollard called.}
Begone dull care.
Come,
Bob.

\N4:
He ambled Dollard,
bulky slops,
before them
\gab{450}

\simon:
(hold that fellow with the:
hold him now)

\N4:
into the saloon.
He plumped him |
Dollard |
on the stool.
His gouty paws plumped chords.
Plumped,
stopped abrupt.

\N4:
Bald Pat in the doorway |
met tealess gold returning.
Bothered,
he wanted Power and cider.

\N1:
Bronze by the window,
watched,
bronze from afar.

\N4:
Jingle a tinkle jaunted.

\N3:
Bloom heard a jing,
a little sound.

\BloomInt:
He's off.

\N3:
Light sob of breath |
Bloom sighed on the silent blue·hued flowers.

\N4:
Jingling.

\BloomInt:
He's gone.

\N4:
Jingle.

\BloomInt:
Hear.

\simon:
\Title{Love and War}, Ben,
\sout{Mr Dedalus said.}
God be with old times.

\N1:
Miss Douce's brave eyes,
unregarded,
\gab{460}
turned from the crossblind,
smitten by sunlight.

\MissDInt:
Gone.

\N1:
Pensive

\MissDInt:
(who knows?),

\N1:
smitten

\MissDInt:
(the smiting light),

\N1:
she lowered the dropblind with a sliding cord.
She drew down pensive

\MissDInt:
(why did he go so quick |
when I?)

\N1:
about her bronze,
over the bar |

\N2:
where bald stood by sister gold,
inexquisite contrast,
contrast inexquisite nonexquisite,
slow cool dim seagreen |
sliding depth of shadow,
\french{eau de Nil}.

\cowley:
Poor old Goodwin was the pianist that night,

\N2:
Father Cowley reminded them.

\cowley:
There was a slight difference of opinion |
between himself and the Collard grand.

\N3:
There was.

\simon:
A symposium all his own,
\gab{470}

\N3:
Mr Dedalus said.

\simon:
The devil wouldn't stop him,
He was a crotchety old fellow |
in the primary stage of drink.

\dollard:
God,
do you remember?

\N4:
Ben bulky Dollard said,
turning from the punished keyboard.

\dollard:
And by Japers |
I had no wedding garment.

\N1:
They laughed |
all three.

\simon:
He had no wed.

\N1:
All trio laughed.

\dollard:
No wedding garment.

\simon:
Our friend Bloom turned in handy that night,
\sout{Mr Dedalus said.}
Where's my pipe,
by the way?

\N3:
He wandered back to the bar |
to the lost chord pipe.

\N4:
Bald Pat carried
two diners' drinks,
Richie and Poldy.

\N2:
And Father Cowley laughed again.

\cowley:
I saved the situation,
\gab{480}
Ben,
I think.

\dollard:
You did,
\sout{averred Ben Dollard.}
I remember those tight trousers too.
That was a brilliant idea,
Bob.

\N2:
Father Cowley blushed to his brilliant purply lobes.

\N2:
He saved the situa.
Tight trou.
Brilliant ide.

\cowley:
I knew he was on the rocks,
\sout{he said.}
The wife was playing the piano |
in the coffee palace on Saturdays |
for a very trifling consideration |
and who was it gave me the wheeze |
she was doing the other business?
Do you remember?
We had to search all Holles street to find them |
till the chap in Keogh's gave us the number.
Remember?

\N4:
Ben remembered,
\gab{490}
his broad visage wondering.

\dollard:
By God,
she had some luxurious opera·cloaks
and things there.

\N3:
Mr Dedalus wandered back,
pipe in hand.

\dollard:
Merrion square style.
Ball·dresses,
by God,
and court dresses.
He wouldn't take any money either.
What?
Any God's quantity of cocked hats
and boleros
and trunk·hose.
What?

\simon:
Ay, ay,
\sout{Mr Dedalus nodded.}
Mrs Marion Bloom has left off clothes
of all descriptions.

\N4:
Jingle jaunted down the quays.
Blazes sprawled on bounding tyres.

\N2:
Liver and bacon.
Steak and kidney pie.
Right, sir.
Right, Pat.

\BloomInt:
Mrs Marion.
Met him pike hoses.
\gab{500}
Smell of burn.
Of Paul de Kock.
Nice name he.

\dollard:
What's this her name was?
A buxom lassy.
Marion ...

\simon:
Tweedy.

\dollard:
Yes.
Is she alive?

\simon:
And kicking.

\dollard:
She was a daughter of ...

\simon:
Daughter of the regiment.

\dollard:
Yes,
begad.
I remember the old drum·major.

\N3:
Mr Dedalus struck,
whizzed,
lit,
puffed savoury puff after

\dollard:
Irish?
I don't know,
faith.
\gab{510}
Is she,
Simon?

\N3:
Puff after stiff,
a puff,
strong,
savoury,
crackling.

\simon:
Buccinator muscle is ...
What?
...
Bit rusty ...
O,
she is ...
My Irish Molly,
O.

\N3:
He puffed a pungent plumy blast.

\simon:
From the rock of Gibraltar...
all the way.

\N2:
They pined in depth of ocean shadow,
gold by the beer·pull,
bronze by ma\-ra\-schi\-no,
thoughtful all two.
Mina Kennedy,
4 Lismore terrace,
Drumcondra |
with Idolores,
a queen,
Dolores,
silent.

\N4:
Pat served,
uncovered dishes.

\N3:
Leopold cut liver·slices.
As said before |
he ate with relish the inner organs,
\gab{520}
nutty gizzards,
fried cods' roes |
while

\N1:
Richie Goulding, |
Collis, Ward |
ate steak and kidney,
steak |
then kidney,
bite by bite of pie he ate |
Bloom ate |
they ate.

\N3:
Bloom with Goulding,
married in silence,
ate.
Dinners fit for princes.

\N4:
By Bachelor's walk |
jogjaunty |
jingled |
Blazes Boylan,
bachelor,
in sun
in heat,
mare's glossy rump atrot,
with flick of whip,
on bounding tyres:
sprawled,
warm·seated,
Boylan impatience,
ardent·bold.
Horn.
Have you the?
Horn.
Have you the?
Haw haw horn.

\N4:
Over their voices |
Dollard bassooned attack,
booming over bombarding chords:

\dollard:
\song{When love absorbs my ardent soul ...}
\gab{530}

\N4:
Roll of Ben·soul·benjamin |
rolled to the quivery love·shivery roof·panes.

\cowley:
War!
War!
\sout{cried Father Cowley.}
You're the warrior.

\dollard:
So I am,

\N4:
Ben Warrior laughed.

\dollard:
I was thinking of your landlord.
Love or money.

\N4:
He stopped.
He wagged huge beard,
huge face |
over his blunder huge.

\simon:
Sure,
you'd burst the tympanum of her ear,
man,

\N3:
Mr Dedalus said through smoke aroma,

\simon:
with an organ like yours.

\N4:
In bearded abundant laughter |
Dollard shook upon the keyboard.
He would.

\cowley:
Not to mention another membrane,
\gab{540}
\sout{Father Cowley added.}
Half time, Ben.
\italian{Amoroso ma non troppo}.
Let me there.

\N2:
Miss Kennedy served two gentlemen with tankards of cool stout.
She passed a remark.

\N1:
It was indeed,
first gentleman said,
beautiful weather.
They drank cool stout.
Did she know where the lord lieutenant was going?

\N4:
And heard steelhoofs ringhoof ring.

\N2:
No,
she couldn't say.
But it would be in the paper.

\N1:
O, she need not trouble.

\N2:
No trouble.
She waved about her outspread \Title{Independent},
searching,

\MissKInt:
the lord lieutenant,

\N2:
her pinnacles of hair slowmoving,

\MissKInt:
lord lieuten.

\N1:
Too much trouble,
first gentleman said.

\N2:
O, not in the least.

\MissKInt:
Way he looked that.
Lord lieutenant.

\N4:
Gold by bronze
heard iron steel.%
\gab{550}

\dollard:
\song{............
my ardent soul \\
I care not foror the morrow.}

\N3:
In liver gravy |
Bloom mashed |
mashed potatoes.

\BloomInt:
\Title{Love and War} someone is.
Ben Dollard's famous.
Night he ran round to us |
to borrow a dress suit for that concert.
Trousers tight as a drum on him.
Musical porkers.
Molly did laugh when he went out.
Threw herself back across the bed,
screaming,
kicking.
With all his belongings on show.
O saints above,
I'm drenched!
O, the women in the front row!
O, I never laughed so many!
Well,
of course |
that's what gives him the base barrel·tone.
For instance |
eunuchs.
\gab{560}
Wonder who's playing.
Nice touch.
Must be Cowley.
Musical.
Knows whatever note you play.
Bad breath he has,
poor chap.
Stopped.

\N1:
Miss Douce,
engaging,
Lydia Douce,
bowed to suave solicitor,
George Lidwell,
gentleman,
entering.

\lidwell:
Good afternoon.

\N1:
She gave her moist (a lady's) hand to his firm clasp.

\MissD:
Afternoon.

\N1:
Yes,
she was back.
To the old dingdong again.

\MissD:
Your friends are inside,
Mr Lidwell.

\N4:
George Lidwell,
suave,
solicited,
held a lydia·hand.

\N3:
Bloom ate liv |
as said before.

\BloomInt:
Clean here at least.
That chap in the Burton,
gummy with gristle.
\gab{570}
No-one here:
Goulding and I\@.
Clean tables,
flowers,
mitres of napkins.
Pat to and fro.
Bald Pat.
Nothing to do.
Best value in Dub.

\BloomInt:
Piano again.
Cowley it is.
Way he sits in to it,
like one together,
mutual understanding.
Tiresome shapers scraping fiddles,
eye on the bow·end,
sawing the cello,
remind you of toothache.
Her high long snore.
Night we were in the box.
Trombone under |
blowing like a grampus,
between the acts,
other brass chap unscrewing,
emptying spittle.
Conductor's legs too,
bags·trousers,
jiggedy jiggedy.
Do right to hide them.

\N4:
Jiggedy jingle jaunty jaunty.

\BloomInt:
Only the harp.
Lovely.
Gold glowering light.
Girl touched it.
\gab{580}
Poop of a lovely.
Gravy's rather good fit for a.
Golden ship.
Erin.
The harp that
once or twice.
Cool hands.
Ben Howth,
the rhododendrons.
We are their harps.
I\@.
He.
Old.
Young.

\simon:
Ah,
I couldn't,
man,

\N3:
Mr Dedalus said,
shy,
listless.

\N4:
Strongly.

\dollard:
Go on,
blast you,
\sout{Ben Dollard growled.}
Get it out in bits.

\cowley:
\italian{M'appari},
Simon,
\sout{Father Cowley said.}

\N2:
Down stage |
he [\stage{Cowley}] strode some paces,
grave,
tall in affliction,
his long arms out·held.
Hoarsely |
the apple of his throat hoarsed softly.
Softly |
he sang to a dusty seascape there:
\gab{590}
\song{A Last Farewell}.
A headland,
a ship,
a sail upon the billows.
Farewell.
A lovely girl,
her veil awave upon the wind |
upon the headland,
wind around her.

\N2:
Cowley sang:

\cowley:
\song{M'appari tutt'amor: \\
Il mio sguardo l'incontr ...}

\N2:
She waved,
unhearing Cowley,
her veil,
to one departing,
dear one,
to wind,
love,
speeding sail,
return.

\dollard:
Go on,
Simon.

\simon:
Ah,
sure,
my dancing days are done,
Ben ...
Well ...

\N3:
Mr Dedalus laid his pipe to rest beside the tuningfork |
\gab{600}
and,
sitting,
touched the obedient keys.

\cowley:
No,
Simon,

\N2:
Father Cowley turned.

\cowley:
Play it in the original.
One flat.

\N3:
The keys,
obedient,
rose higher,
told,
faltered,
confessed,
confused.

\N2:
Up stage |
strode Father Cowley.

\cowley:
Here,
Simon,
I'll accompany you,
\sout{he said}.
Get up.

\N4:
By Graham Lemon's pineapple rock,
by Elvery's elephant
jingly jogged.

\N3:
Steak,
kidney,
liver,
mashed,
at meat fit for princes |
sat princes Bloom and Goulding.
Princes at meat |
they raised and drank,
Power and cider.

\goulding:
Most beautiful tenor air ever written,
\gab{610}

\N1:
Richie said:

\goulding:
\italian{Sonnambula}

\N1:
He heard Joe Maas sing that one night.

\goulding:
Ah,
what M'Guckin!
Yes.
In his way.
Choirboy style.
Maas was the boy.

\BloomInt:
Massboy.

\goulding:
A lyrical tenor if you like.
Never forget it.
Never.

\N3:
Tenderly |
Bloom over liverless bacon |
saw the tightened features strain.

\BloomInt:
Backache he.
Bright's bright eye.
Next item on the programme.
Paying the piper.
Pills,
pounded bread,
worth a guinea a box.
Stave it off awhile.
Sings too:
\song{Down among the dead men.}
Appropriate.
Kidney pie.
Sweets to the.
Not making much hand of it.
Best value in.
Characteristic of him.
Power.
Particular about his drink.
Flaw in the glass,
fresh Vartry water.
Fecking matches from counters to save.
\gab{620}
Then squander a sovereign
in dribs and drabs.
And when he's wanted |
not a farthing.
Screwed |
refusing to pay his fare.
Curious types.

\N1:
Never would Richie forget that night.
As long as he lived:
never.
In the gods of the old Royal |
with little Peake.
And when the first note.

\N1:
Speech paused on Richie's lips.

\BloomInt:
Coming out with a whopper now.
Rhapsodies about damn all.
Believes his own lies.
Does really.
Wonderful liar.
But want a good memory.

\Bloom:
Which air is that?
\sout{asked Leopold Bloom.}

\goulding:
\song{All is lost now.}

\N1:
Richie cocked his lips a·pout.
\gab{630}
A low incipient note |
sweet banshee murmured:
all.

\BloomInt:
A thrush.
A throstle.
His breath,
bird·sweet,
good teeth he's proud of,
fluted with plaintive woe.
Is lost.
Rich sound.
Two notes in one there.
Blackbird I heard |
in the hawthorn valley.
Taking my motives |
he twined and turned them.
All most too new call is lost in all.
Echo.
How sweet the answer.
How is that done?
All lost now.
Mournful he whistled.
Fall,
surrender,
lost.

\N3:
Bloom bent leopold ear,
turning a fringe of doyley down under the vase.

\BloomInt:
Order.
Yes,
I remember.
Lovely air.
In sleep |
she went to him.
Innocence in the moon.
Brave.
Don't know their danger.
Still |
hold her back.
Call name.
\gab{640}
Touch water.
Jingle jaunty.
Too late.
She longed to go.
That's why.
Woman.
As easy stop the sea.
Yes:
all is lost.

\Bloom:
A beautiful air,

\N3:
said Bloom lost Leopold.

\Bloom:
I know it well.

\N1:
Never in all his life had Richie Goulding.

\BloomInt:
He knows it well too.
Or he feels.
Still harping on his daughter.
Wise child that knows her father,
Dedalus said.
Me?

\N3:
Bloom askance over liverless saw.

\BloomInt:
Face of the all is lost.
Rollicking
Richie once.
Jokes old stale now.
Wagging his ear.
Napkin·ring in his eye.
Now begging letters he sends his son with.
Cross·eyed Walter |
sir |
I did sir.
Wouldn't trouble |
only I was expecting some money.
Apologise.

\BloomInt:
Piano again.
\gab{650}
Sounds better than last time I heard.
Tuned probably.
Stopped again.

\N4:
Dollard and Cowley still urged the lingering singer |
out with it.

\dollard:
With it,
Simon.

\cowley:
It,
Simon.

\simon:
Ladies and gentlemen,
I am most deeply obliged by your kind
solicitations.

\cowley:
It,
Simon.

\simon:
I have no money |
but if you will lend me your attention |
I shall endeavour |
to sing to you |
of a heart bowed down.

\N1:
By the sandwich·bell |
\gab{660}
in screening shadow |
Lydia,
her bronze and rose,
a lady's grace,
gave and withheld:

\N2:
as in cool glaucous \french{eau de Nil} |
Mina |
to tankards two |
her pinnacles of gold.

\N2:
The harping chords of prelude closed.
A chord,
longdrawn,
expectant,
drew a voice away.

\simon:
\song{When first I saw that form endearing ...}

\N1:
Richie turned.

\goulding:
Si Dedalus' voice,
\sout{he said.}

\N1:
Brain·tipped,
cheek touched with flame,
they listened |
feeling that flow endearing |
flow over skin |
limbs |
human heart |
soul |
spine.

\N3:
Bloom signed to Pat,

\BloomInt:
bald Pat is a waiter |
hard of hearing,
\gab{670}

\N3:
to set ajar the door of the bar.
The door of the bar.

\BloomInt:
So.
That will do.

\N4:
Pat,
waiter,
waited,
waiting
to hear,
for he was hard of hear |
by the door.

\simon:
\song{... Sorrow from me seemed to depart.}

\N2:
Through the hush of air |
a voice sang to them,
low,
not rain,
not leaves
in murmur,
like no voice of strings |
or reeds |
or what·do·you·call·them dulcimers |
touching their still ears with words,
still hearts of their |
each his remembered lives.

\N2:
Good,
good to hear:
sorrow from them each |
seemed to from both depart |
when first they heard.
When first they saw,
lost Richie |
Poldy, |
mercy of beauty,
heard from a person |
wouldn't expect it in the least,
her first merciful
\gab{680}
love·soft oft·loved word.

\BloomInt:
Love that is singing:
love's old sweet song.

\N3:
Bloom unwound slowly |
the elastic band of his packet.

\N4:
Love's old sweet
\french{sonnez la}
gold.

\N3:
Bloom wound a skein round four fork·fingers,
stretched it,
relaxed,
and wound it round his troubled double,
fourfold,
in octave,
gyved them fast.

\simon:
\song{Full of hope and all delighted ...}

\BloomInt:
Tenors get women by the score.
Increase their flow.
Throw flower at his feet.
When will we meet?
My head |
it simply.
Jingle |
all delighted.
He can't sing for tall hats.
Your head |
it simply swurls.
Perfumed for him.
What perfume does your wife?
I want to know.
Jing.
Stop.
Knock.
Last look at mirror always |
before she answers the door.
The hall.
\gab{690}
There?
How do you?
I do well.
There?
What?
Or?
Phial of cachous,
kissing comfits,
in her satchel.
Yes?
Hands felt for the opulent.

\N2:
Alas! |
The voice rose,
sighing,
changed:
loud,
full,
shining,
proud.

\simon:
\song{But alas, 'twas idle dreaming ...}

\BloomInt:
Glorious tone he has still.
Cork air softer |
also their brogue.
Silly man!
Could have made oceans of money.
Singing wrong words.
Wore out his wife:
now sings.
But hard to tell.
Only the two themselves.
If he doesn't break down.
Keep a trot for the avenue.
His hands and feet sing too.
Drink.
Nerves overstrung.
Must be abstemious to sing.
Jenny Lind soup:
stock,
sage,
raw eggs,
half pint of cream.
\gab{700}
For creamy dreamy.

\N2:
Tenderness |
it welled:
slow,
swelling,
full |
it throbbed.

\BloomInt:
That's the chat.
Ha,
give!
Take!

\N2:
Throb,
a throb,
a pulsing proud erect.

\BloomInt:
Words?
Music?
No:
it's what's behind.

\N3:
Bloom looped,
unlooped,
noded,
disnoded.

\BloomInt:
Bloom.
Flood of warm jamjam |
lick·it·up |
secretness |
flowed to |
flow in |
music out,
in desire,
dark to lick |
flow invading.
Tipping her |
tepping her |
tapping her |
topping her.
Tup.
Pores to dilate dilating.
Tup.
The joy |
the feel |
the warm |
the.
Tup.
To pour o'er sluices |
pouring gushes.
Flood,
gush,
flow,
joy·gush,
tup·throb.
Now!
Language of love.

\simon:
\song{... ray of hope is ...}
\gab{710}

\N1:
Beaming.
Lydia for Lidwell |
squeak scarcely hear |
so ladylike |
the muse unsqueaked |
a ray of hope.

\BloomInt:
\Title{Martha} it is.
Coincidence.
Just going to write.
Lionel's song.
Lovely name you have.
Can't write.
Accept my little pres.
Play on her heartstrings |
pursestrings too.
She's a.
I called you naughty boy.
Still the name:
Martha.
How strange!
Today.

\N2:
The voice of Lionel returned,
weaker but unwearied.
It sang again to Richie |
Poldy |
Lydia |
Lidwell |
also sang to Pat |
open mouth |
ear |
waiting to wait.
How first he saw that form endearing,
how sorrow seemed to part,
how look,
form,
word |
charmed him |
\gab{720}
Gould |
Lidwell,
won Pat Bloom's heart.

\BloomInt:
Wish I could see his face,
though.
Explain better.
Why the barber in Drago's |
always looked my face |
when I spoke his face in the glass.
Still hear it better here |
than in the bar |
though farther.

\simon:
\song{Each graceful look ...}

\BloomInt:
First night when first I saw her |
at Mat Dillon's in Terenure.
Yellow,
black lace she wore.
Musical chairs.
We two the last.
Fate.
After her.
Fate.

\BloomInt:
Round and round slow.
Quick round.
We two.
All looked.
Halt.
Down she
sat.
All ousted looked.
Lips laughing.
Yellow knees.

\simon:
\song{Charmed my eye ...}

\BloomInt:
Singing.
\gab{730}
\Title{Waiting} |
she sang.
I turned her music.
Full voice of perfume |
of what perfume |
does your lilac·trees.
Bosom I saw,
both full,
throat warbling.
First I saw.
She thanked me.
Why did she |
me?
Fate.
Spanishy eyes.
Under a pear·tree |
alone patio |
this hour in old Madrid |
one side in shadow |
Dolores |
she·dolores.
At me.
Luring.
Ah, |
alluring.

\simon:
\song{Martha!
Ah, Martha!}

\N2:
Quitting all languor |
Lionel cried in grief,
in cry of passion dominant |
to love |
to return |
with deepening |
yet with rising |
chords of harmony.
In cry of lionel |
loneliness that she should know,
must Martha feel.
For only her |
he waited.
Where?
Here |
there |
try there |
here |
all try |
where.
Somewhere.

\simon:
\song{Co-ome, thou lost one!%
\gab{740}\\
Co-ome, thou dear one!}

\N2:
Alone.
One love.
One hope.
One comfort me.
Martha,
chest·note,
return!

\simon:
\song{Come!}

\N2:
It soared,
a bird,
it held its flight,
a swift pure cry,
soar silver orb |
it leaped serene,
speeding,
sustained,
to come,
don't spin it out too long |
long breath |
he breath |
long life,
soaring high,
high resplendent,
aflame,
crowned,
high in the effulgence symbolistic,
high, of the etherial bosom,
high, of the high vast irradiation |
everywhere |
all soaring |
all around about |
the all,
the endless·ness·ness·ness%
\gab{750}\\
...

\simon:
\song{To me!}

\N2:
Sio·pold!

\BloomInt:
Consumed.

\BloomInt:
Come.
Well sung.
All clapped.
She ought to.
Come.
To me,
to him,
to her,
you too,
me,
us.

\lenehan:
Bravo!

\N1:
Clap·clap.

\dollard:
Good man,
Simon.

\N3:
Clappy·clap·clap.

\cowley:
Encore!

\N1:
Clap·clip·clap clap.

\lenehan:
Sound as a bell.
Bravo,
Simon!

\N3:
Clap·clop·clap.

\N1:
Encore,
enclap,

\N3:
said,
cried,
clapped all,
Ben Dollard,
Lydia Douce,
George
Lidwell,
Pat,
Mina Kennedy,
two gentlemen with two tankards,
Cowley,
first gent with tank and bronze miss Douce
\gab{760}
and gold Miss Mina.

\N4:
Blazes Boylan's smart tan shoes creaked on the barfloor,
said before.
Jingle by monuments |
of sir John Gray,
Horatio one·handled Nelson,
reverend father Theobald Mathew,
jaunted,
as said before |
just now.
Atrot,
in heat,
heat·seated.
\french{Cloche.
Sonnez la.
Cloche.
Sonnez la.}
Slower the mare went |
up the hill by the Rotunda,
Rutland square.
Too slow for Boylan,
blazes Boylan,
impatience Boylan,
joggled the mare.

\N2:
An after·clang of Cowley's chords closed,
died on the air |
made richer.

\N1:
And Richie Goulding drank his Power |

\N3:
and Leopold Bloom his cider drank,

\N4:
Lidwell his Guinness,

\N1:
second gentleman said they would partake of two more tankards |
if she did not mind.
\gab{770}

\N2:
Miss Kennedy smirked,
disserving,
coral lips,
at first,
at second.
She did not mind.

\dollard:
Seven days in jail,
\sout{Ben Dollard said,}
on bread and water.
Then you'd sing,
Simon,
like a garden thrush.

\N3:
Lionel Simon,
singer,
laughed.

\N2:
Father Bob Cowley played.

\N2:
Mina Kennedy served.

\N1:
Second gentleman paid.

\N4:
Tom Kernan strutted in.

\N1:
Lydia,
admired,
admired.

\N3:
But Bloom sang dumb.

\N1:
Admiring.

\N1:
Richie,
admiring,
descanted on that man's glorious voice.
He remembered one night long ago.

\goulding:
Never forget that night.
Si sang

\simon:
\song{'Twas rank and fame}

\goulding:
in Ned Lambert's 'twas.
\gab{780}

\N1:
Good God |
he never heard |
in all his life |
a note like that |
he never did

\simon:
\song{then false one we had better part}

\N1:
so clear so |
God he never heard

\simon:
\song{since love lives not}

\N1:
a clinking voice |
lives not |

\goulding:
ask Lambert |
he can tell you too.

\N1:
Goulding,
a flush struggling in his pale,

\N2:
told Mr Bloom,

\N1:
face of the night,

\goulding:
Si in Ned Lambert's,

\N1:
Dedalus house,

\goulding:
sang

\simon:
\song{'Twas rank and fame}

\N1:
He,
Mr Bloom,
listened |
while he,
Richie Goulding, |
told him,
Mr Bloom, |
of the night |
he,
Richie, |
heard him,
Si Dedalus, |
sing

\simon:
\song{'Twas rank and fame}

\N1:
in his,
Ned Lambert's,
house.

\BloomInt:
Brothers-in-law:
relations.
We never speak as we pass by.
Rift in the lute |
I think.
Treats him with scorn.
\gab{790}
See.
He admires him all the more.
The night Si sang.
The human voice,
two tiny silky chords,
wonderful,
more than all others.

\BloomInt:
That voice was a lamentation.
Calmer now.
It's in the silence after
you feel you hear.
Vibrations.
Now silent air.

\N3:
Bloom ungyved his crisscrossed hands |
and |
with slack fingers |
plucked the slender catgut thong.
He drew and plucked.
It buzz,
it twanged.

\N1:
While Goulding talked of Barraclough's voice production,

\N4:
while Tom Kernan,
harking back in a retrospective sort of arrangement |
talked to 

\N2:
listening Father Cowley,
who played a voluntary,
who nodded as he played.

\N4:
While big Ben Dollard talked with Simon Dedalus,
lighting,
\gab{800}
who nodded as he smoked,
who smoked.

\BloomInt:
Thou lost one.
All songs on that theme.

\N3:
Yet more |
Bloom stretched his string.

\BloomInt:
Cruel it seems.
Let people get fond of each other:
lure them on.
Then tear asunder.
Death.
Explos.
Knock on the head.
Out·to·hell·out·of·that.
Human life.
Dignam.
Ugh,
that rat's tail wriggling!
Five bob I gave.
\latin{Corpus paradisum}.
Corncrake croaker:
belly like a poisoned pup.
Gone.
They sing.
Forgotten.
I too.
And one day she with.
Leave her:
get tired.
Suffer then.
Snivel.
Big spanishy eyes goggling at nothing.
Her wavy·avy·eavy·heavy·eavy·evy·evy hair un comb:'d.

\BloomInt:
Yet too much happy |
bores.
\gab{810}

\N3:
He stretched more,
more.

\BloomInt:
Are you not happy in your?

\N3:
Twang.
It snapped.

\N4:
Jingle into Dorset street.

\N1:
Miss Douce withdrew her satiny arm,
reproachful,
pleased.

\MissD:
Don't make half so free,
\sout{said she,}
till we are better acquainted.

\N4:
George Lidwell told her really and truly:

\N1:
but she did not believe.

\N4:
First gentleman told Mina that was so.

\N2:
She asked him was that so.

\N4:
And second tankard told her so.

\N3:
That that was so.

\N1:
Miss Douce,
miss Lydia,
did not believe:

\N2:
miss Kennedy,
Mina,
did not believe:

\N4:
George Lidwell,
no:

\N1:
miss Dou did not:
the first,
the first:

\N4:
gent with the tank:
\gab{820}
believe,
no, no:
did not,

\N2:
miss Kenn:

\N1:
Lid·lydia·well:

\N4:
the tank.

\BloomInt:
Better write it here.
Quills in the post·office chewed and twisted.

\N4:
Bald Pat |
at a sign |
drew nigh.
A pen and ink.
He went.
A pad.
He went.
A pad to blot.
He heard,
deaf Pat.

\Bloom:
Yes.

\N3:
Mr Bloom said,
teasing the curling catgut line.

\Bloom:
It certainly is.

\BloomInt:
Few lines will do.
My present.

\Bloom:
All that Italian florid music is.
Who is this wrote?
Know the name |
you know better.

\BloomInt:
Take out
sheet notepaper,
envelope:
unconcerned.

\Bloom:
It's so characteristic.

\goulding:
Grandest number in the whole opera.

\Bloom:
It is.
\sout{Bloom said.}

\BloomInt:
Numbers it is.
\gab{830}
All music |
when you come to think.
Two multiplied by two |
divided by half |
is twice one.
Vibrations:
chords those are.
One plus two plus six |
is seven.
Do anything you like |
with figures juggling.
Always find out |
this equal to that.
Symmetry under a cemetery wall.
He doesn't see my mourning.
Callous:
all for his own gut.
Muse·mathematics.
And you think you're listening to the etherial.
But suppose you said it like:
Martha,
seven times nine |
minus x |
is thirty·five thousand.
Fall quite flat.
It's on account of the sounds |
it is.

\BloomInt:
Instance he's playing now.
Improvising.
Might be what you like,
till you hear the words.
Want to listen sharp.
Hard.
Begin all right:
then hear chords a bit off:
feel lost a bit.
\gab{840}
In and out of sacks,
over barrels,
through wirefences,
obstacle race.
Time makes the tune.
Question of mood you're in.
Still |
always nice to hear.
Except scales up and down,
girls learning.
Two together |
nextdoor neighbours.
Ought to invent dummy pianos for that.
\german{Blumenlied} |
I bought for her.
The name.
Playing it slow,
a girl,
night I came home,
the girl.
Door of the stables near Cecilia street.
Milly |
no taste.
Queer |
because we both,
I mean.

\N4:
Bald deaf Pat |
brought quite flat pad |
ink.
Pat set with ink pen |
quite flat pad.
Pat took plate dish knife fork.
Pat went.

\simon:
It was the only language

\N3:
Mr Dedalus said to Ben.

\simon:
[I] heard them as a boy in Ringabella,
Crosshaven,
Ringabella,
\gab{850}
singing their barcaroles.
Queenstown harbour full of Italian ships.
Walking,
you know,
Ben,
in the moonlight with those earthquake hats.
Blending their voices.
God,
such music,
Ben.
Heard as a boy.
Cross Ringabella haven moon·carole.

\N3:
Sour pipe removed |
he held a shield of hand |
beside his lips |
that cooed a moonlight nightcall,
clear from anear,
a call from afar,
replying.

\N3:
Down the edge of his \Title{Freeman} baton |
ranged Bloom's,

\N1:
your other eye,

\N3:
scanning for

\BloomInt:
where did I see that.
Callan,
Coleman,
Dignam Patrick.
Heigho!
Heigho!
Fawcett.
Aha!
Just I was looking ...

\BloomInt:
Hope he's not looking,
cute as a rat.

\N3:
He held unfurled his \Title{Freeman}

\BloomInt:
Can't see now.
Remember write Greek ees.
\gab{860}

\N3:
Bloom dipped,

\BloomInt:
Bloo mur:
dear sir.

\N3:
Dear Henry wrote:

\BloomInt:
dear Mady.
Got your lett and flow.
Hell did I put?
Some pock or oth.
It is utterl imposs.
Underline \emph{imposs}.
To write today.

\BloomInt:
Bore this.

\N3:
Bored Bloom tambourined gently with

\BloomInt:
I am just reflecting

\N3:
fingers on flat pad |
Pat brought.

\BloomInt:
On.
Know what I mean.
No,
change that ee.
Accep my poor litt pres |
enclos.
Ask her |
no answ.
Hold on.
Five |
Dig.
Two |
about here.
Penny |
the gulls.
Elijah is com.
Seven |
Davy Byrne's.
Is eight about.
Say half a crown.
My poor little pres:
p.o.\
two and six.
Write me a long.
Do you despise?
Jingle,
have you the?
So excited.
Why do you call me naught?
You naughty too?
\gab{870}
O,
Mairy lost the string of her.
Bye for today.
Yes,
yes,
will tell you.
Want to.
To keep it up.
Call me that other.
Other world she wrote.
My patience are exhaust.
To keep it up.
You must believe.
Believe.
The tank.
It.
Is.
True.

\BloomInt:
Folly am I writing?
Husbands don't.
That's marriage does,
their wives.
Because I'm away from.
Suppose.
But how?
She must.
Keep young.
If she found out.
Card in my high grade ha.
No,
not tell all.
Useless pain.
If they don't see.
Woman.
Sauce for the gander.

\N4:
A hackney car,
number three hundred and twenty·four,
driver Barton James |
of number one Harmony avenue,
Donnybrook,
on which sat a fare,
a young gentleman,
\gab{880}
stylishly dressed in an indigo·blue serge suit |
made by George Robert Mesias,
tailor and cutter,
of number five Eden quay,
and wearing a straw hat |
very dressy,
bought of John Plasto |
of number one Great Brunswick street,
hatter.

\BloomInt:
Eh?
This is the jingle that joggled and jingled.

\N4:
By Dlugacz' pork·shop |
bright tubes of Agendath |
trotted a gallant·buttocked mare.

\goulding:
Answering an ad?

\N1:
keen Richie's eyes asked Bloom.

\Bloom:
Yes.
\sout{Mr Bloom said.}
Town traveller.
Nothing doing,
I expect.

\BloomInt:
Bloom mur:
best references.
But Henry wrote:
it will excite me.
You know how.
In haste.
Henry.
Greek ee.
Better add postscript.
What is he playing now?
Improvising.
\gab{890}
Intermezzo.
P.S\@.
The rum tum tum.
How will you pun?
You punish me?
Crooked skirt swinging,
whack by.
Tell me |
I want to.
Know.
O\@.
Course if I didn't |
I wouldn't ask.
La la la ree.
Trails off there |
sad in minor.
Why minor sad?
Sign H\@.
They like sad tail at end.
P.P.S\@.
La la la ree.
I feel so sad today.
La ree.
So lonely.
Dee.

\N3:
He blotted quick |
on pad of Pat.

\BloomInt:
Envel.
Address.
Just copy out of paper.

\N3:
Murmured:

\BloomInt:
Messrs Callan,
Coleman and Co,
limited.

\N3:
Henry wrote:

\N1:
    Miss Martha Clifford\\
    \-\hspace{3cm}    c/o P.O.\\
    \-\hspace{2cm} Dolphin's Barn Lane\\
    \-\hspace{3cm}    Dublin
\gab{900}

\BloomInt:
Blot over the other |
so he can't read.
There.
Right.
Idea prize |
titbit.
Something detective read off blotting·pad.
Payment at the rate of |
guinea per col.
Matcham often thinks the laughing witch.
Poor Mrs Purefoy.
U.P: up.

\BloomInt:
Too poetical that about the sad.
Music did that.
Music hath charms.
Shakespeare said.
Quotations every day in the year.
To be or not to be.
Wisdom while you wait.

\BloomInt:
In Gerard's rosery of Fetter lane |
he walks,
greyed·auburn.
One life is all.
One body.
Do.
But do.

\BloomInt:
Done anyhow.
Postal order,
stamp.
Post·office lower down.
Walk now.
Enough.
\gab{910}
Barney Kiernan's |
I promised to meet them.
Dislike that job.

\BloomInt:
House of mourning.
Walk.

\Bloom:
Pat!

\BloomInt:
Doesn't hear.
Deaf beetle he is.

\BloomInt:
Car near there now.
Talk.
Talk.

\Bloom:
Pat!

\BloomInt:
Doesn't.
Settling those napkins.
Lot of ground |
he must cover in the day.
Paint face behind on him |
then he'd be two.
Wish they'd sing more.
Keep my mind off.

\N4:
Bald Pat who is bothered |
mitred the napkins.

\BloomInt:
Pat is a waiter |
hard of his hearing.
Pat is a waiter |
who waits while you wait.
Hee hee hee hee.
He waits while you wait.
Hee hee.
A waiter is he.
Hee hee hee hee.
He waits while you wait.
While you wait |
if you wait |
he will wait |
while you wait.
Hee hee hee hee.
Hoh.
Wait while you wait.

\N1:
Douce now.
\gab{920}
Douce Lydia.
Bronze and rose.

\N1:
She had a gorgeous,
simply gorgeous,
time.
And look at the lovely shell she brought.

\N1:
To the end of the bar |
to him she bore lightly |
the spiked and winding seahorn |
that he,
George Lidwell,
solicitor,
might hear.

\MissD:
Listen!

\N1:
she bade him.

\N4:
Under Tom Kernan's gin·hot words |
the accompanist wove music slow.

\kernan:
Authentic fact.
How Walter Bapty lost his voice.
Well, sir,
the husband took him by the throat.
\emph{Scoundrel},
said he,
\emph{you'll sing no more lovesongs}.
He did,
faith,
sir Tom.

\N2:
Bob Cowley wove.

\BloomInt:
Tenors get wom.

\N2:
Cowley lay back.

\N1:
Ah,
now he heard,
she holding it to his ear.
\gab{930}

\MissD:
Hear!

\N1:
He heard.

\lidwell:
Wonderful.

\N1:
She held it to her own.

\N2:
And through the sifted light |
pale gold in contrast glided.
To hear.

\stripling:
Tap

\N3:
Bloom |
through the bardoor |
saw a shell held at their ears.
He heard more faintly that |
that they heard,
each for herself alone,
then each for other,
hearing the plash of waves,
loudly,
a silent roar.

\N2:
Bronze by a weary gold,
anear,
afar,
they listened.

\BloomInt:
Her ear too |
is a shell,
the peeping lobe there.
Been to the seaside.
Lovely seaside girls.
Skin tanned raw.
Should have put on cold·cream first |
make it brown.
\gab{940}
Buttered toast.
O and that lotion |
mustn't forget.
Fever near her mouth.
Your head |
it simply.
Hair braided over:
shell with seaweed.
Why do they hide their ears with seaweed hair?
And Turks |
the mouth,
why?
Her eyes over the sheet.
Yashmak.
Find the way in.
A cave.
No admittance except on business.

\BloomInt:
The sea |
they think they hear.
Singing.
A roar.
The blood it is.
Souse
in the ear sometimes.
Well,
it's a sea.
Corpuscle islands.

\lidwell:
Wonderful really.
So distinct.
Again.

\N4:
George Lidwell held its murmur,
hearing:
then laid it by,
gently.

\lidwell:
What are the wild waves saying?

\N1:
he asked her,
smiled.

\N1:
Charming,
sea·smiling and unanswering |
\gab{950}
Lydia on Lidwell smiled.

\stripling:
Tap

\N4:
By Larry O'Rourke's,
by Larry,
bold Larry O',
Boylan swayed
and Boylan turned.

\N2:
From the forsaken shell |
miss Mina glided to her tankards waiting.

\N1:
No,
she was not so lonely |
archly miss Douce's head |
let Mr Lidwell know.
Walks in the moonlight by the sea.
No,
not alone.
With whom?
She nobly answered:
with a gentleman friend.

\N2:
Bob Cowley's twinkling fingers in the treble |
played again.
The landlord has the prior.
A little time.
Long John.
Big Ben.
Lightly |
he played a light bright tinkling measure |
for tripping ladies,
\gab{960}
arch and smiling,
and for their gallants,
gentlemen friends.
One:
one,
one,
one,
one:
two,
one,
three,
four.

\BloomInt:
Sea,
wind,
leaves,
thunder,
waters,
cows lowing,
the cattle·market,
cocks,
hens don't crow,
snakes hissss.
There's music everywhere.
Ruttledge's door:
ee creaking.
No,
that's noise.
Minuet of \Title{Don Giovanni} |
he's playing now.
Court dresses of all descriptions in castle chambers |
dancing.
Misery.
Peasants outside.
Green starving faces eating dock·leaves.
Nice that is.
Look:
look,
look,
look,
look,
look:
you look at us.

\BloomInt:
That's joyful I can feel.
Never have written it.
Why?
My joy is other joy.
But both are joys.
\gab{970}
Yes,
joy it must be.
Mere fact of music |
shows you are.
Often thought she was in the dumps |
till she began to lilt.
Then know.

\BloomInt:
M'Coy valise.
My wife and your wife.
Squealing cat.
Like tearing silk.
Tongue when she talks |
like the clapper of a bellows.
They can't manage men's intervals.
Gap in their voices too.
Fill me.
I'm warm,
dark,
open.
Molly in \latin{quis est homo}:
Mercadante.
My ear against the wall to hear.
Want a woman who can deliver the goods.

\N4:
Jog jig jogged |
stopped.
Dandy tan shoe of dandy Boylan socks |
sky·blue clocks |
came light to earth.

\BloomInt:
O,
look we are so!
Chamber music.
Could make a kind of pun on that.
\gab{980}
It is a kind of music |
I often thought |
when she.
Acoustics |
that is.
Tinkling.
Empty vessels make most noise.
Because the acoustics,
the resonance changes |
according as the weight of the water |
is equal to the law of falling water.
Like those rhapsodies of Liszt's,
Hungarian,
gipsy·eyed.
Pearls.
Drops.
Rain.
Diddle·iddle addle·addle ooddle·ooddle.
Hissss.
Now.
Maybe now.
Before.

\BloomInt:
One rapped on a door,
one tapped with a knock,
did he knock Paul de Kock |
with a loud proud knocker |
with a cock |
carra·carra·carra cock.
Cock·cock.

\stripling:
Tap

\cowley:
\italian{Qui sdegno},
Ben
\gab{990}

\kernan:
No, Ben,

\N4:
Tom Kernan interfered.

\kernan:
\Title{The Croppy Boy}.
Our native Doric.

\simon:
Ay do,
Ben,
\sout{Mr Dedalus said.}
Good men and true.

\N2:
Do,
do,
they begged in one.

\BloomInt:
I'll go.
Here,
Pat,
return.
Come.
He came,
he came,
he did not stay.
To me.
How much?

\cowley:
What key?
Six sharps?

\dollard:
F sharp major,
\sout{Ben Dollard said}

\N2:
Bob Cowley's outstretched talons |
gripped the black deep·sounding chords.

\Bloom:
Must go
\gab{1000}

\N3:
prince Bloom told Richie prince.

\goulding:
No,
\sout{Richie said.}

\Bloom:
Yes,
must.

\BloomInt:
Got money somewhere.
He's on for a razzle backache spree.
Much?
He see·hears lip·speech.

\Bloom:
One and nine.
Penny for yourself.

\BloomInt:
Here.
Give him twopence tip.
Deaf,
bothered.
But perhaps he has wife and family waiting,
waiting Patty come home.
Hee hee hee hee.
Deaf wait |
while they wait.

\BloomInt:
But wait.
But hear.
Chords dark.
Lugu·gu·gubrious.
Low.
In a cave of the dark middle earth.
Embedded ore.
Lump·music.

\N2:
The voice of dark age,
of unlove,
earth's fatigue made grave approach,
and painful,
come from afar,
from hoary mountains,
called on good men and true.
The priest he sought.
With him |
would he speak a word.

\stripling:
Tap
\gab{1010}

\BloomInt:
Ben Dollard's voice.
Base barrel·tone.
Doing his level best to say it.
Croak of vast man·less moonless wo·moon·less marsh.
Other comedown.
Big ships' chandler's business |
he did once.
Remember:
rosiny ropes,
ships' lanterns.
Failed to the tune of ten thousand pounds.
Now in the Iveagh home.
Cubicle number so and so.
Number one Bass |
did that for him.

\N2:
The priest's at home.
A false priest's servant |
bade him welcome.
Step in.
The holy father.
With bows |
a traitor servant.
Curlycues of chords.

\BloomInt:
Ruin them.
Wreck their lives.
Then build them cubicles |
to end their days in.
Hushaby.
Lullaby.
Die,
dog.
Little dog,
die.

\N2:
The voice of warning,
\gab{1020}
solemn warning,
told them |
the youth had entered a lonely hall,
told them |
how solemn fell his footsteps there,
told them |
the gloomy chamber,
the vested priest sitting to shrive.

\BloomInt:
Decent soul.
Bit addled now.
Thinks he'll win |
in \Title{Answers} poets' picture puzzle.
We hand you crisp five pound note.
Bird sitting hatching in a nest.
Lay of the last minstrel |
he thought it was.
See blank tee |
what domestic animal?
Tee dash ar |
most courageous mariner.
Good voice he has still.
No eunuch yet |
with all his belongings.

\BloomInt:
Listen.

\N3:
Bloom listened.

\N1:
Richie Goulding listened.

\N4:
And by the door
deaf Pat,
bald Pat,
tipped Pat,
listened.

\N2:
The chords harped slower.
\gab{1030}

\N2:
The voice of penance and of grief |
came slow,
embellished,
tremulous.
Ben's contrite beard confessed.

\dollard:
\latin{In nomine Domini},

\N2:
in God's name.
He knelt.
He beat his hand upon his breast,
confessing:

\dollard:
\latin{mea culpa}.

\BloomInt:
Latin again.
That holds them like birdlime.
Priest with the communion corpus for those women.
Chap in the mortuary,
coffin or coffey,
\latin{corpus·nomine}.
Wonder where that rat is by now.
Scrape.

\stripling:
Tap

\N4:
They listened.
Tankards and miss Kennedy.
George Lidwell,
eyelid well expressive,
full·busted satin.
Kernan.
Si.

\N2:
The sighing voice of sorrow sang.
\gab{1040}
His sins.

\N2:
Since Easter |
he had cursed three times.

\N1:
You bitch's bast.

\N2:
And once at mass·time |
he had gone to play.
Once by the church·yard |
he had passed |
and for his mother's rest |
he had not prayed.
A boy.
A croppy boy.

\N1:
Bronze,
listening,
by the beer·pull gazed far away.
Soulfully.

\BloomInt:
Doesn't half know I'm.
Molly |
great dab at seeing anyone |
looking.

\N3:
Bronze gazed far sideways.

\BloomInt:
Mirror there.
Is that best side of her face?
They always know.
Knock at the door.
Last tip to titivate.

\BloomInt:
Cock·carra·carra.

\BloomInt:
What do they think |
when they hear music?
Way to catch rattle\-snakes.
Night |
Michael Gunn gave us the box.
\gab{1050}
Tuning up.
Shah of Persia liked that best.
Remind him of home sweet home.
Wiped his nose in curtain too.
Custom his country|
perhaps.
That's music too.
Not as bad as it sounds.
Tootling.
Brasses |
braying asses through uptrunks.
Double·basses |
helpless,
gashes in their sides.
Woodwinds |
mooing cows.
Semi·grand |
open crocodile |
music hath jaws.
Woodwind like Goodwin's name.

\BloomInt:
She looked fine.
Her crocus dress she wore lowcut,
belongings on show.
Clove her breath was always |
in theatre |
when she bent to ask a question.
Told her what Spinoza says |
in that book of poor papa's.
Hypnotised,
listening.
Eyes like that.
She bent.
Chap in dress·circle |
staring down into her with his opera·glass |
for all he was worth.
\gab{1060}
Beauty of music |
you must hear twice.
Nature woman |
half a look.
God made the country |
man the tune.
Met him pike hoses.
Philosophy.
O rocks!

\N2:
All gone.
All fallen.
At the siege of Ross |
his father,
at Gorey |
all his brothers fell.
To Wexford,
we are the boys of Wexford,
he would.
Last of his name and race.

\BloomInt:
I |
too.
Last of my race.
Milly |
young student.
Well,
my fault perhaps.
No son.
Rudy.
Too late now.
Or if not?
If not?
If still?

\N2:
He bore no hate.

\BloomInt:
Hate.
Love.
Those are names.
Rudy.
Soon I am old.

\N4:
Big Ben |
his voice unfolded.
\gab{1070}

\N1:
Great voice |
Richie Goulding said,
a flush struggling in his pale,
to Bloom |
soon old.
But when |
was young?

\BloomInt:
Ireland comes now.
My country above the king.
She listens.
Who fears to speak of nineteen four?
Time to be shoving.
Looked enough.

\dollard:
\song{Bless me, father,}

\N4:
Dollard the croppy cried.

\dollard:
\song{Bless me and let me go.}

\stripling:
Tap

\N3:
Bloom looked,
unblessed |
to go.

\BloomInt:
Got up to kill:
on eighteen bob a week.
Fellows shell out the dibs.
Want to keep your weather·eye open.
Those girls,
those lovely.
By the sad sea waves.
Chorus·girl's romance.
Letters read out for breach of promise.
From Chick·a·biddy's owny Mumpsy·pum.
Laughter in court.
\gab{1080}
Henry.
I never signed it.
The lovely name you.

\N2:
Low sank the music,
air and words.
Then hastened.
The false priest |
rustling soldier from his cassock.
A yeoman captain.

\BloomInt:
They know it all by heart.
The thrill they itch for.
Yeoman cap.

\stripling:
Tap.
Tap.

\N1:
Thrilled |
she listened,
bending in sympathy to hear.

\BloomInt:
Blank face.
Virgin should say:
or fingered only.
Write something on it:
page.
If not |
what becomes of them?
Decline,
despair.
Keeps them young.
Even admire themselves.
See.
Play on her.
Lip blow.
Body of white woman,
a flute alive.
Blow gentle.
Loud.
Three holes,
all women.
Goddess |
I didn't see.
They want it.
\gab{1090}
Not too much polite.
That's why he gets them.
Gold in your pocket,
brass in your face.
Say something.
Make her hear.
With look to look.
Songs without words.
Molly,
that hurdy·gurdy boy.
She knew he meant the monkey was sick.
Or because so like the Spanish.
Understand animals too that way.
Solomon did.
Gift of nature.

\BloomInt:
Ventriloquise.
My lips closed.
Think in my stom.
What?

\BloomInt:
Will?
You?
I\@.
Want.
You.
To.

\N2:
With hoarse rude fury |
the yeoman cursed,
swelling in apoplectic bitch's bastard.
A good thought,
boy,
to come.
One hour's your time to live,
your last.

\stripling:
Tap.
Tap.
\gab{1100}

\BloomInt:
Thrill now.
Pity they feel.
To wipe away a tear |
for martyrs that want to,
dying to,
die.
For all things dying,
for all things born.
Poor Mrs Purefoy.
Hope she's over.
Because their wombs.

\N1:
A liquid of womb |
of woman eyeball |
gazed under a fence of lashes,
calmly,
hearing.

\BloomInt:
See real beauty of the eye |
when she not speaks.
On yonder river.
At each slow satiny heaving bosom's wave |
(her heaving embon) |
red rose |
rose slowly |
sank red rose.
Heartbeats:
her breath:
breath that is life.
And all the tiny tiny fernfoils |
trembled of maidenhair.

\BloomInt:
But look.
The bright stars fade.
O rose!
Castile.
The morn.
Ha.
\gab{1110}
Lidwell.
For him then not for.
Infatuated.
I like that?
See her from here though.
Popped corks,
splashes of beerfroth,
stacks of empties.

\N1:
On the smooth jutting beer·pull |
laid Lydia hand,
lightly,
plumply,

\BloomInt:
leave it to my hands.
All lost in pity for croppy.

\N1:
Fro,
to:
to,
fro:
over the polished knob

\BloomInt:
(she knows his eyes,
my eyes,
her eyes)

\N1:
her thumb and finger passed in pity:
passed,
reposed |
and, gently touching,
then slid so smoothly,
slowly down,
a cool firm white enamel baton |
protruding through their sliding ring.

\BloomInt:
With a cock |
with a carra.

\stripling:
Tap.
Tap.
Tap.

\N2:
I hold this house.
Amen.
\gab{1120}
He gnashed in fury.
Traitors swing.

\N2:
The chords consented.
Very sad thing.
But had to be.

\BloomInt:
Get out before the end.
Thanks,
that was heavenly.
Where's my hat.
Pass by her.
Can leave that \Title{Freeman}.
Letter I have.
Suppose she were the?
No.
Walk,
walk,
walk.
Like Cashel Boylo |
Connoro Coylo |
Tisdall Maurice |
Tisntdall Farrell.
Waaaaaaalk.

\Bloom:
Well,
I must be.

\goulding:
Are you off?

\Bloom:
Yrf·mst·byes.
Blm·stup.

\goulding:
O'er rye·high blue.

\BloomInt:
Ow.

\N3:
Bloom stood up.

\BloomInt:
Soap feeling rather sticky behind.
Must have sweated:
music.
That lotion,
remember.

\Bloom:
Well,
so long.

\BloomInt:
High grade.
Card inside.
Yes.

\N3:
By deaf Pat in the doorway |
\gab{1130}
straining ear |
Bloom passed.

\N2:
At Geneva barrack |
that young man died.
At Passage |
was his body laid.
Dolor!
O,
he dolores!
The voice of the mournful chanter |
called to dolorous prayer.

\N3:
By rose,
by satiny bosom,
by the fondling hand,
by slops,
by empties,
by popped corks,
greeting in going,
past eyes and maidenhair,
bronze and faint gold in deep·sea·shadow,
went Bloom,
soft Bloom,
I feel so lonely Bloom.

\stripling:
[\stage{Louder}]
Tap.
Tap.
Tap

\N2:
Pray for him,
prayed the bass of Dollard.
You who hear in peace.
Breathe a prayer,
drop a tear,
good men,
good people.
\gab{1140}
He was the croppy boy.

\N3:
Scaring eavesdropping boots |
croppy boots·boy |
Bloom in the Ormond hallway |
heard the growls and roars of bravo,
fat backslapping,
their boots all treading,
boots not the boots the boy.

\BloomInt:
General chorus off for a swill to wash it down.
Glad I avoided.

\simon:
Come on,
Ben,
\sout{Simon Dedalus cried.}
By God,
you're as good as ever you were.

\kernan:
Better,

\N4:
said Tom·gin Kernan.

\kernan:
Most trenchant rendition of that ballad,
upon my soul and honour It is.

\cowley:
Lablache,
\gab{1150}
\sout{said Father Cowley.}

\N4:
Ben Dollard bulkily cachuchad towards the bar,
mightily praisefed |
and all big roseate,
on heavyfooted feet,
his gouty fingers |
nakkering castagnettes in the air.

\N2:
Big Ben·a·ben Dollard.
Big Ben·ben.
Big Ben·ben.

\N2:
Rrr.

\N3:
And deep·moved all,
Simon trumping compassion from foghorn nose,
all laughing |
they brought him forth,
Ben Dollard,
in right good cheer.

\lidwell:
You're looking rubicund,
\sout{George Lidwell said.}

\N1:
Miss Douce composed her rose to wait.

\simon:
Ben machree,
\gab{1160}

\N3:
said Mr Dedalus,
clapping Ben's fat back shoulderblade.

\simon:
Fit as a fiddle |
only he has a lot of adipose tissue |
concealed about his person.

\N2:
Rrrrrrrsss.

\dollard:
Fat of death,
Simon,

\N4:
Ben Dollard growled.

\N1:
Richie rift in the lute alone sat:
Goulding,
Collis,
Ward.
Uncertainly he waited.

\N4:
Unpaid Pat too.

\stripling:
[\stage{Louder}]
Tap.
Tap.
Tap.
Tap

\N2:
Miss Mina Kennedy brought near her lips |
to ear of tankard one.

\N2:
Mr Dollard,
they murmured low.

\N1:
Dollard,
murmured tankard.

\N1:
Tank one believed:
\gab{1170}
miss Kenn when she:
that doll he was:
she doll:
the tank.

\N1:
He murmured that he knew the name.
The name was familiar to him,
that is to say.
That was to say he had heard the name of.
Dollard,
was it?

\N2:
Dollard,
yes.

\MissK:
Yes,

\N2:
her lips said more loudly,

\MissK:
Mr Dollard.
He sang that song lovely,
\sout{murmured Mina.}
Mr Dollard.
And \Title{The Last Rose of Summer} was a lovely song.

\N2:
Mina loved that song.

\N1:
Tankard loved the song that Mina.

\N3:
'Tis the last rose of summer |
dollard left |
bloom felt |
wind wound |
round inside.

\BloomInt:
Gassy thing that cider:
binding too.
\gab{1180}
Wait.
Post·office near Reuben J's |
one and eightpence too.
Get shut of it.
Dodge round by Greek street.
Wish I hadn't promised to meet.
Freer in air.
Music.
Gets on your nerves.
Beer·pull.
Her hand that rocks the cradle rules the.
Ben Howth.
That rules
the world.

\N3:
Far.
Far.
Far.
Far.

\stripling:
[\stage{Louder}]
Tap.
Tap.
Tap.
Tap.

\N3:
Up the quay |
went Lionel·leopold,
naughty Henry with letter for Mady,
with sweets of sin |
with frillies for Raoul |
with met him pike hoses |
went Poldy on.

\N1:
Tap |
blind walked
\gab{1190}
tapping |
by the tap |
the curbstone tapping,
tap by tap.

\BloomInt:
Cowley,
he stuns himself with it:
kind of drunkenness.
Better give way |
only half way |
the way of a man with a maid.
Instance enthusiasts.
All ears.
Not lose a demisemiquaver.
Eyes shut.
Head nodding in time.
Dotty.
You daren't budge.
Thinking strictly prohibited.
Always talking shop.
Fiddle·faddle about notes.

\BloomInt:
All a kind of attempt to talk.
Unpleasant when it stops |
because you never know exac.
Organ in Gardiner street.
Old Glynn |
fifty quid a year.
Queer up there in the cock·loft,
alone,
with stops and locks and keys.
Seated all day at the organ.
Maunder on for hours,
talking to himself |
or the other fellow blowing the bellows.
\gab{1200}
Growl angry,
then shriek cursing
(want to have wadding or something in his |
no don't |
she cried),
then all of a soft |
sudden wee |
little wee |
little pipy wind.

\N3:
Pwee!
A wee little wind |
piped eeee.
In Bloom's little wee.

\simon:
Was he?

\N3:
Mr Dedalus said,
returning with fetched pipe.

\simon:
I was with him this morning |
at poor little Paddy Dignam's ...

\lidwell:
Ay,
the Lord have mercy on him.

\simon:
By the bye |
there's a tuningfork in there |
on the ...

\stripling:
Tap.
Tap.
Tap.
Tap

\lidwell:
The wife has a fine voice.
Or had.
What?
\sout{Lidwell asked.}

\MissD:
O,
that must be the tuner,
\gab{1210}

\N1:
Lydia said to Simon·lionel

\MissD:
first I saw,
forgot it when he was here.

\N1:
Blind he was |
she told George Lidwell |
second I saw.
And played so exquisitely,
treat to hear.
Exquisite contrast:
bronze·lid,
mina·gold.

\dollard:
Shout!

\N4:
Ben Dollard shouted,
pouring.

\dollard:
Sing out!

\cowley:
'lldo!
\sout{cried Father Cowley.}

\N2:
Rrrrrr.

\BloomInt:
I feel I want ...

\stripling:
[\stage{Louder}]
Tap.
Tap.
Tap.
Tap.
Tap

\simon:
Very,

\N3:
Mr Dedalus said,
staring hard at a headless sardine.

\N1:
Under the sandwich·bell |
\gab{1220}
lay on a bier of bread |
one last,
one lonely,
last sardine of summer.
Bloom alone.

\simon:
Very,
\sout{he stared.}
The lower register,
for choice.

\stripling:
[\stage{Louder}]
Tap.
Tap.
Tap.
Tap.
Tap.
Tap.
[\stage{Softer}]
Tap.
Tap.

\N3:
Bloom went by Barry's.

\BloomInt:
Wish I could.
Wait.
That wonder·worker if I had.
Twenty·four solicitors in that one house.
Counted them.
Litigation.
Love one another.
Piles of parchment.
Messrs Pick and Pocket have power of attorney.
Goulding,
Collis,
Ward.

\BloomInt:
But for example |
the chap that wallops the big drum.
His vocation:
Mickey Rooney's band.
Wonder how it first struck him.
Sitting at home |
after pig's cheek and cabbage |
nursing it in the armchair.
\gab{1230}
Rehearsing his band part.
Pom.
Pompedy.
Jolly for the wife.
Asses' skins.
Welt them through life,
then wallop after death.
Pom.
Wallop.
Seems to be what you call yashmak |
or I mean kismet.
Fate.

\stripling:
Tap.
Tap.

\N1:
A stripling,
blind,
with a tapping cane |
came tap·tap·tapping
by Daly's window |
where a mermaid hair |
all streaming
(but he couldn't see)
blew whiffs of a mermaid
(blind couldn't),
mermaid,
coolest whiff of all.

\BloomInt:
Instruments.
A blade of grass,
shell of her hands,
then blow.
Even comb and tissue·paper |
you can knock a tune out of.
Molly in her shift |
in Lombard street west,
hair down.
I suppose each kind of trade |
made its own,
don't you see?
\gab{1240}
Hunter with a horn.
Haw.
Have you the?
\french{Cloche.
Sonnez la.}
Shepherd his pipe.
Pwee little wee.
Policeman a whistle.
Locks and keys!
Sweep!
Four o'clock's |
all's well!
Sleep!
All is lost now.
Drum?
Pompedy.
Wait.
I know.
Towncrier,
bum·bailiff.
Long John.
Waken the dead.
Pom.
Dignam.
Poor little \latin{nomine·domine}.
Pom.
It is music.
I mean of course |
it's all pom pom pom |
very much what they call
\italian{da capo}.
Still you can hear.
As we march,
we march along,
march along.
Pom.

\BloomInt:
I must really.
Fff.
Now |
if I did that at a banquet.
Just a question of custom |
shah of Persia.
Breathe a prayer,
drop a tear.
All the same |
he must have been a bit of a natural |
not to see it was a yeoman cap.
Muffled up.
Wonder who was that chap at the grave |
in the brown macin.
\gab{1250}
O,
the whore of the lane!

\N4:
A frowsy whore |
with black straw sailor hat askew |
came glazily in the day |
along the quay |
towards Mr Bloom.

\N3:
When first he saw |
that form endearing?

\BloomInt:
Yes,
it is.
I feel so lonely.
Wet night in the lane.
Horn.
Who had the?
Heehaw shesaw.
Off her beat here.
What is she?
Hope she.
Psst!
Any chance of your wash.
Knew Molly.
Had me decked.
Stout lady does be with you |
in the brown costume.
Put you off your stroke,
that.
Appointment we made |
knowing we'd never,
well hardly ever.
Too dear |
too near to home sweet home.
Sees me,
does she?
Looks a fright in the day.
Face like dip.
Damn her.
\gab{1260}
O,
well,
she has to live like the rest.
Look in here.

\N3:
In Lionel Marks's antique saleshop window |
haughty Henry |
Lionel Leopold |
dear Henry Flower |
earnestly Mr Leopold Bloom envisaged |
battered candlesticks |
melodeon |
oozing maggoty blow·bags.

\BloomInt:
Bargain:
six bob.
Might learn to play.
Cheap.
Let her pass.
Course everything is dear |
if you don't want it.
That's what good salesman is.
Make you buy |
what he wants to sell.
Chap sold me the Swedish razor |
he shaved me with.
Wanted to charge me |
for the edge he gave it.
She's passing now.
Six bob.

\BloomInt:
Must be the cider or perhaps the burgund.

\N1:
Near bronze from anear |
near gold from afar |
they chinked their clinking glasses all,
brighteyed and gallant,
\gab{1270}
before bronze Lydia's tempting
last rose of summer,
rose of Castile.
First Lid,
De,
Cow,
Ker,
Doll,
a fifth:
Lidwell,
Si Dedalus,
Bob Cowley,
Kernan |
and big Ben Dollard.

\stripling:
Tap.

\N1:
A youth entered a lonely Ormond hall.

\N3:
Bloom viewed a gallant pictured hero |
in Lionel Marks's window.

\BloomInt:
Robert Emmet's last words.
Seven last words.
Of Meyerbeer that is.

\dollard:
True men like you men.

\simon:
Ay,
ay,
Ben.

\dollard:
Will lift your glass with us.

\N4:
They lifted.

\stage{Glasses clinking.}

\N1:
Tschink.
\gab{1280}

\N4:
Tschunk.

\stripling:
Tip.

\N1:
An unseeing stripling stood in the door.
He saw not bronze.
He saw not gold.
Nor Ben |
nor Bob |
nor Tom |
nor Si |
nor George |
nor tanks |
nor Richie |
nor Pat.
Hee hee hee hee.
He did not see.

\N3:
Sea·bloom,
grease·a·bloom |
viewed last words.

\BloomInt:
Softly.

\emmet:
\emph{When my country takes her place among.}

\N3:
Prrprr.

\BloomInt:
Must be the bur.

\N3:
Fff!
Oo.
Rrpr.

\emmet:
\emph{Nations of the earth.}

\BloomInt:
No-one behind.
She's passed.

\emmet:
\emph{Then and not till then.}

\BloomInt:
Tram.

\N3:
Kran kran kran.

\BloomInt:
Good oppor.
\gab{1290}
Coming.

\N3:
Krandl·kran·kran.

\BloomInt:
I'm sure it's the burgund.
Yes.
One,
two.

\emmet:
\emph{Let my epitaph be.}

\N3:
Kraaaaaa.

\emmet:
\emph{Written.
I have.}

\N3:
Pp·rr·pff·rr·ppffff.

\BloomInt:
Done.
