\deasy:
Now then,

\N1:
Mr Deasy said, rising.

\N1:
He came to the table, pinning together his sheets.
\gab{320}

\N2:
Stephen stood up.

\deasy:
I have put the matter into a nutshell, \sout{Mr Deasy said.}
It's about the foot and mouth disease.
Just look through it.
There can be no two opinions on the matter.

\StephenInt: \stage{[Paraphrasing]}
May I trespass on your valuable space.
That doctrine of \french{laissez faire} which so often in our history.
Our cattle trade.
The way of all our old industries.
Liverpool ring which jockeyed the Galway harbour scheme.
European conflagration.
Grain supplies through the narrow waters of the channel.
The pluterperfect imperturbability of the department of agriculture.
Pardoned a classical allusion.
Cassandra.
By a woman who was no better than she should be.
\gab{330}
To come to the point at issue.

\deasy:
I don't mince words, do I?

\N1:
Mr Deasy asked as Stephen read on.

\StephenInt: \stage{paraphrasing}
Foot and mouth disease.
Known as Koch's preparation.
Serum and virus.
Percentage of salted horses.
Rinderpest.
Emperor's horses at Murzsteg, lower Austria.
Veterinary surgeons.
Mr Henry Blackwood Price.
Courteous offer a fair trial.
Dictates of common sense.
Allimportant question.
In every sense of the word take the bull by the horns.
Thanking you for the hospitality of your columns.

\deasy:
I want that to be printed and read, \sout{Mr Deasy said}.
You will see at the next outbreak they will put an embargo on Irish cattle.
And it can be cured.
It is cured.
My cousin, Blackwood Price, writes to me
\gab{340}
it is regularly treated and cured in Austria by cattledoctors there.
They offer to come over here.
I am trying to work up influence with the department.
Now I'm going to try publicity.
I am surrounded by difficulties,
by ... intrigues by ... backstairs influence by ...

\N1:
He raised his forefinger and beat the air oldly before his voice spoke.

\deasy:
Mark my words, Mr Dedalus, \sout{he said.}
England is in the hands of the jews.
In all the highest places: her finance, her press.
And they are the signs of a nation's decay.
Wherever they gather they eat up the nation's vital strength.
I have seen it coming these years.
As sure as we are standing here
the jew merchants are already at their work of destruction.
\gab{350}
Old England is dying.

\N1:
He stepped swiftly off, his eyes coming to blue life
as they passed a broad sunbeam.
He faced about and back again.

\deasy:
Dying, \sout{he said again},
if not dead by now.

\StephenInt:
\begin{verse}
    The harlot's cry from street to street \\
    Shall weave old England's windingsheet.
\end{verse}

\N1:
His eyes open wide in vision stared sternly across the sunbeam
in which he halted.

\Stephen:
A merchant, \sout{Stephen said,}
is one who buys cheap and sells dear,
jew or gentile,
\gab{360}
is he not?

\deasy:
They sinned against the light,
\sout{Mr Deasy said gravely.}
And you can see the darkness in their eyes.
And that is why they are wanderers on the earth to this day.

\StephenInt:
On the steps of the Paris stock exchange
the goldskinned men quoting prices on their gemmed fingers.
Gabble of geese.
They swarmed loud, uncouth about the temple,
their heads thickplotting under maladroit silk hats.
Not theirs: these clothes, this speech, these gestures.
Their full slow eyes belied the words,
the gestures eager and unoffending,
but knew the rancours massed about them
and knew their zeal was vain.
Vain patience to heap and hoard.
\gab{370}
Time surely would scatter all.
A hoard heaped by the roadside:
plundered and passing on.
Their eyes knew their years of wandering
and, patient, knew the dishonours of their flesh.

\Stephen:
Who has not?
\sout{Stephen said.}

\deasy:
What do you mean?
\sout{Mr Deasy asked.}

\N2:
He came forward a pace and stood by the table.
His underjaw fell sideways open uncertainly.

\StephenInt:
Is this old wisdom?
He waits to hear from me.

\Stephen:
History, \sout{Stephen said,}
is a nightmare from which I am trying to awake.

\N1:
From the playfield the boys raised a shout.
A whirring whistle: goal.

\StephenInt:
What if that nightmare gave you a back kick?

\deasy:
The ways of the Creator are not our ways,
\sout{Mr Deasy said.}
\gab{380}
All human history moves towards one great goal,
the manifestation of God.

\N2:
Stephen jerked his thumb towards the window, saying:

\Stephen:
That is God.

\All:
Hooray! Ay! Whrrwhee!

\deasy:
What?
\sout{Mr Deasy asked.}

\Stephen:
A shout in the street,

\N2:
Stephen answered, shrugging his shoulders.

\N1:
Mr Deasy looked down
and held for awhile the wings of his nose
tweaked between his fingers.
Looking up again he set them free.

\deasy:
I am happier than you are, \sout{he said.}
We have committed many errors and many sins.
A woman brought sin into the world.
\gab{390}
For a woman who was no better than she should be,
Helen, the runaway wife of Menelaus,
ten years the Greeks made war on Troy.
A faithless wife first brought the strangers to our shore here,
MacMurrough's wife and her leman, O'Rourke, prince of Breffni.
A woman too brought Parnell low.
Many errors, many failures but not the one sin.
I am a struggler now at the end of my days.
But I will fight for the right till the end.

\begin{verse}
    For Ulster will fight \\
    And Ulster will be right.
\end{verse}

\N2:
Stephen raised the sheets in his hand.

\Stephen:
Well, sir, he began ...
\gab{400}

\deasy:
I foresee, \sout{Mr Deasy said,}
that you will not remain here very long at this work.
You were not born to be a teacher, I think.
Perhaps I am wrong.

\Stephen:
A learner rather, \sout{Stephen said.}

\StephenInt:
And here what will you learn more?

\N1:
Mr Deasy shook his head.

\deasy:
Who knows? \sout{he said.}
To learn one must be humble.
But life is the great teacher.

\N2:
Stephen rustled the sheets again.

\Stephen:
As regards these, \sout{he began.}

\deasy:
Yes, \sout{Mr Deasy said.}
You have two copies there.
\gab{410}
If you can have them published at once.

\StephenInt:
\Title{Telegraph}.
\Title{Irish Homestead}.

\Stephen:
I will try, \sout{Stephen said,}
and let you know tomorrow.
I know two editors slightly.

\deasy:
That will do, \sout{Mr Deasy said briskly.}
I wrote last night to Mr Field, M.P.
There is a meeting of the cattletraders' association today
at the City Arms hotel.
I asked him to lay my letter before the meeting.
You see if you can get it into your two papers.
What are they?

\Stephen:
\Title{The Evening Telegraph} ...

\deasy:
That will do, \sout{Mr Deasy said.}
\gab{420}
There is no time to lose.
Now I have to answer that letter from my cousin.

\Stephen:
Good morning, sir,

\N2:
Stephen said, putting the sheets in his pocket.

\Stephen:
Thank you.

\deasy:
Not at all,

\N1:
Mr Deasy said as he searched the papers on his desk.

\deasy:
I like to break a lance with you, old as I am.

\Stephen:
Good morning, sir,

\N2:
Stephen said again, bowing to his bent back.

\N2:
He went out by the open porch and down the gravel path under the trees,
hearing the cries of voices and crack of sticks from the playfield.
The lions couchant on the pillars as he passed out through the gate:

\StephenInt:
Toothless terrors.
Still I will help him in his fight.
\gab{430}
Mulligan will dub me a new name: the bullockbefriending bard.

\deasy:
Mr Dedalus!

\StephenInt:
Running after me.
No more letters, I hope.

\deasy:
Just one moment.

\Stephen:
Yes, sir,

\N2:
Stephen said, turning back at the gate.

\N1:
Mr Deasy halted, breathing hard and swallowing his breath.

\deasy:
I just wanted to say, \sout{he said.}
Ireland, they say, has the honour of being the only country
which never persecuted the jews.
Do you know that?
No.
And do you know why?

\N1:
He frowned sternly on the bright air.
\gab{440}

\Stephen:
Why, sir?

\N2:
Stephen asked, beginning to smile.

\deasy:
Because she never let them in,
\sout{Mr Deasy said solemnly.}

\N1:
A coughball of laughter leaped from his throat
dragging after it a rattling chain of phlegm.
He turned back quickly, coughing, laughing,
his lifted arms waving to the air.

\deasy:
She never let them in,

\N1:
he cried again through his laughter
as he stamped on gaitered feet over the gravel of the path.

\deasy:
That's why.

\N1:
On his wise shoulders through the checkerwork of leaves
the sun flung spangles, dancing coins.

