\N1:
Sargent who alone had lingered
came forward slowly, showing an open copybook.
His thick hair and scraggy neck gave witness of unreadiness
and through his misty glasses weak eyes looked up pleading.
On his cheek, dull and bloodless,
a soft stain of ink lay, dateshaped,
recent and damp as a snail's bed.

\N1:
He held out his copybook.
The word \emph{sums} was written on the headline.
Beneath were sloping figures
and at the foot a crooked signature with blind loops and a blot.
\gab{130}
Cyril Sargent: his name and seal.

\sargent:
Mr Deasy told me to write them out all again, \sout{he said},
and show them to you, sir.

\N2:
Stephen touched the edges of the book.

\StephenInt:
Futility.

\Stephen:
Do you understand how to do them now?
\sout{he asked}.

\sargent:
Numbers eleven to fifteen,
\sout{Sargent answered}.
Mr Deasy said I was to copy them off the board, sir.

\Stephen:
Can you do them yourself?
\sout{Stephen asked}.

\sargent:
No, sir.

\StephenInt:
Ugly and futile:
lean neck and thick hair
and a stain of ink, a snail's bed.
Yet someone had loved him,
\gab{140}
borne him in her arms and in her heart.
But for her the race of the world would have trampled him underfoot,
a squashed boneless snail.
She had loved his weak watery blood drained from her own.
Was that then real?
The only true thing in life?
His mother's prostrate body
the fiery Columbanus in holy zeal bestrode.
She was no more:
the trembling skeleton of a twig burnt in the fire,
an odour of rosewood and wetted ashes.
She had saved him from being trampled underfoot
and had gone, scarcely having been.
A poor soul gone to heaven:
and on a heath beneath winking stars
a fox, red reek of rapine in his fur,
with merciless bright eyes scraped in the earth,
listened, scraped up the earth,
listened, scraped and scraped.
\gab{150}

\N2:
Sitting at his side Stephen solved out the problem.

\mulligan:
He proves by algebra that Shakespeare's ghost is Hamlet's grandfather.

\N2:
Sargent peered askance through his slanted glasses.
Hockeysticks rattled in the lumberroom:
the hollow knock of a ball and calls from the field.

\StephenInt:
Across the page the symbols moved in grave morrice,
in the mummery of their letters,
wearing quaint caps of squares and cubes.
Give hands, traverse, bow to partner:
so:
imps of fancy of the Moors.
Gone too from the world,
Averroes and Moses Maimonides,
dark men in mien and movement,
flashing in their mocking mirrors the obscure soul of the world,
a darkness shining in brightness
\gab{160}
which brightness could not comprehend.

\Stephen:
Do you understand now?
Can you work the second for yourself?

\sargent:
Yes, sir.

\N1:
In long shaky strokes Sargent copied the data.
Waiting always for a word of help
his hand moved faithfully the unsteady symbols,
a faint hue of shame flickering behind his dull skin.

\StephenInt:
\latin{Amor matris}: subjective and objective genitive.
With her weak blood and wheysour milk
she had fed him and hid from sight of others his swaddling bands.

\StephenInt:
Like him was I, these sloping shoulders, this gracelessness.
My childhood bends beside me.
Too far for me to lay a hand there once or lightly.
Mine is far and his secret as our eyes.
\gab{170}
Secrets, silent, stony sit in the dark palaces of both our hearts:
secrets weary of their tyranny:
tyrants, willing to be dethroned.

\N1:
The sum was done.

\Stephen:
It is very simple,

\N2:
Stephen said as he stood up.

\sargent:
Yes, sir.
Thanks, \sout{Sargent answered}.

\N1:
He dried the page with a sheet of thin blottingpaper
and carried his copybook back to his bench.

\Stephen:
You had better get your stick and go out to the others,

\N2:
Stephen said as he followed towards the door the boy's graceless form.

\sargent:
Yes, sir.
\gab{180}

\N1:
In the corridor his name was heard, called from the playfield.

\deasy:
Sargent!

\Stephen:
Run on, \sout{Stephen said}.
Mr Deasy is calling you.
