\N2:
He stood in the porch and watched the laggard hurry towards the scrappy field
where sharp voices were in strife.
They were sorted in teams
and Mr Deasy came away stepping over wisps of grass with gaitered feet.
When he had reached the schoolhouse
voices again contending called to him.
He turned his angry white moustache.

\deasy:
What is it now?

\N1:
he cried continually without listening.

\Stephen:
Cochrane and Halliday are on the same side, sir,
\gab{190}
\sout{Stephen said}.

\deasy:
Will you wait in my study for a moment, \sout{Mr Deasy said},
till I restore order here.


\N2:
And as he stepped fussily back across the field
his old man's voice cried sternly:

\deasy:
What is the matter?
What is it now?

\N2:
Their sharp voices cried about him on all sides:
their many forms closed round him,
the garish sunshine bleaching the honey of his illdyed head.

\N1:
Stale smoky air hung in the study
with the smell of drab abraded leather of its chairs.
\gab{200}

\StephenInt:
As on the first day he bargained with me here.
As it was in the beginning, is now.
On the sideboard the tray of Stuart coins, base treasure of a bog:
and ever shall be.
And snug in their spooncase of purple plush, faded,
the twelve apostles having preached to all the gentiles:
world without end.

\N1:
A hasty step over the stone porch and in the corridor.
Blowing out his rare moustache
Mr Deasy halted at the table.

\deasy:
First, our little financial settlement, \sout{he said}.

\N1:
He brought out of his coat a pocketbook bound by a leather thong.
It slapped open and he took from it two notes, one of joined halves,
and laid them carefully on the table.
\gab{210}

\deasy:
Two,

\N1:
he said, strapping and stowing his pocketbook away.

\N1:
And now his strongroom for the gold.
Stephen's embarrassed hand moved over the shells heaped in the cold stone mortar:

\StephenInt:
whelks and money cowries and leopard shells:
and this, whorled as an emir's turban,
and this, the scallop of saint James.
An old pilgrim's hoard, dead treasure, hollow shells.

\N1:
A sovereign fell, bright and new, on the soft pile of the tablecloth.

\deasy:
Three,

\N1:
Mr Deasy said, turning his little savingsbox about in his hand.

\deasy:
These are handy things to have.
See.
This is for sovereigns.
This is for shillings.
Sixpences, halfcrowns.
And here crowns.
\gab{220}
See.

\N1:
He shot from it two crowns and two shillings.

\deasy:
Three twelve, \sout{he said}. I think you'll find that's right.

\Stephen:
Thank you, sir,

\N2:
Stephen said, gathering the money together with shy haste
and putting it all in a pocket of his trousers.

\deasy:
No thanks at all, \sout{Mr Deasy said}. You have earned it.

\N2:
Stephen's hand, free again, went back to the hollow shells.

\StephenInt:
Symbols too of beauty and of power.
A lump in my pocket: symbols soiled by greed and misery.

\deasy:
Don't carry it like that,
\sout{Mr Deasy said}.
You'll pull it out somewhere and lose it.
You just buy one of these machines.
\gab{230}
You'll find them very handy.

\StephenInt:
Answer something.

\Stephen:
Mine would be often empty,
\sout{Stephen said}.

\StephenInt:
The same room and hour, the same wisdom:
and I the same.
Three times now.
Three nooses round me here.
Well?
I can break them in this instant if I will.

\deasy:
Because you don't save,
\sout{Mr Deasy said, pointing his finger.}
You don't know yet what money is.
Money is power.
When you have lived as long as I have.
I know, I know.
If youth but knew.
But what does Shakespeare say?
\Quotation{Put but money in thy purse}.

\Stephen:
Iago,
\gab{240}
\sout{Stephen murmured}.

\N2:
He lifted his gaze from the idle shells to the old man's stare.

\deasy:
He knew what money was, \sout{Mr Deasy said}.
He made money.
A poet, yes, but an Englishman too.
Do you know what is the pride of the English?
Do you know what is the proudest word you will ever hear from an Englishman's mouth?

\StephenInt:
The seas' ruler.
His seacold eyes looked on the empty bay:
it seems history is to blame:
on me and on my words, unhating.

\Stephen:
That on his empire, \sout{Stephen said,} the sun never sets.

\deasy:
Ba! \sout{Mr Deasy cried}.
That's not English.
A French Celt said that.

\N1:
He tapped his savingsbox against his thumbnail.
\gab{250}

\deasy:
I will tell you, \sout{he said solemnly,} what is his proudest boast.
\emph{I paid my way.}

\StephenInt:
Good man, good man.

\deasy:
\emph{I paid my way.
I never borrowed a shilling in my life.}
Can you feel that?
\emph{I owe nothing.}
Can you?

\StephenInt:
Mulligan, nine pounds, three pairs of socks, one pair brogues, ties.
Curran, ten guineas.
McCann, one guinea.
Fred Ryan, two shillings.
Temple, two lunches.
Russell, one guinea,
Cousins, ten shillings,
Bob Reynolds, half a guinea,
Koehler, three guineas,
Mrs MacKernan, five weeks' board.
The lump I have is useless.

\Stephen:
For the moment, no,
\gab{260}
\sout{Stephen answered}.

\N1:
Mr Deasy laughed with rich delight, putting back his savingsbox.

\deasy:
I knew you couldn't, \sout{he said joyously.}
But one day you must feel it.
We are a generous people but we must also be just.

\Stephen:
I fear those big words, \sout{Stephen said,}
which make us so unhappy.

\N1:
Mr Deasy stared sternly for some moments over the mantelpiece
at the shapely bulk of a man in tartan filibegs:
Albert Edward, prince of Wales.

\deasy:
You think me an old fogey and an old tory, \sout{his thoughtful voice said.}
I saw three generations since O'Connell's time.
I remember the famine in~'46.
Do you know that the Orange lodges agitated for repeal of the union
\gab{270}
twenty years before O'Connell did
or before the prelates of your communion denounced him as a demagogue?
You Fenians forget some things.

\StephenInt:
Glorious, pious and immortal memory.
The lodge of Diamond in Armagh
the splendid behung with corpses of papishes.
Hoarse, masked and armed, the planters' covenant.
The black north and true blue bible.
Croppies lie down.

\N2:
Stephen sketched a brief gesture.

\deasy:
I have rebel blood in me too, \sout{Mr Deasy said}.
On the spindle side.
But I am descended from sir John Blackwood who voted for the union.
We are all Irish,
\gab{280}
all kings' sons.

\Stephen:
Alas, \sout{Stephen said}.

\deasy:
\latin{Per vias rectas}, \sout{Mr Deasy said firmly}, was his motto.
He voted for it and put on his topboots to ride to Dublin from the Ards of Down to do so.

\StephenInt:
\begin{verse}
    Lal the ral the ra \\
    The rocky road to Dublin.
\end{verse}

\StephenInt:
A gruff squire on horseback with shiny topboots.
Soft day, sir John!
Soft day, your honour!
... Day!
... Day!
... Two topboots jog dangling on to Dublin.
Lal the ral the ra.
Lal the ral the raddy.

\deasy:
That reminds me, \sout{Mr Deasy said}.
You can do me a favour, Mr Dedalus,
with some of your literary friends.
\gab{290}
I have a letter here for the press.
Sit down a moment.
I have just to copy the end.

\N1:
He went to the desk near the window,
pulled in his chair twice
and read off some words from the sheet on the drum of his typewriter.

\deasy:
Sit down.
Excuse me,

\N1:
he said over his shoulder,

\deasy:
\emph{The Dictates of Common Sense}.
Just a moment.

\N1:
He peered from under his shaggy brows at the manuscript by his elbow
and, muttering, began to prod the stiff buttons of the keyboard slowly,
sometimes blowing as he screwed up the drum to erase an error.

\N2:
Stephen seated himself noiselessly before the princely presence.
Framed around the walls
images of vanished horses stood in homage,
\gab{300}
their meek heads poised in air:
lord Hastings' Repulse,
the duke of Westminster's Shotover,
the duke of Beaufort's Ceylon,
\Title{Prix de Paris}, 1866.
Elfin riders sat them, watchful of a sign.
He saw their speeds, backing king's colours,
and shouted with the shouts of vanished crowds.

\deasy:
Full stop,

\N1:
Mr Deasy bade his keys.

\deasy:
But prompt ventilation of this allimportant question ...

\StephenInt:
Where Cranly led me to get rich quick,
hunting his winners among the mudsplashed brakes,
amid the bawls of bookies on their pitches and reek of the canteen,
over the motley slush.
Fair Rebel!
Fair Rebel!
Even money the favourite:
ten to one the field.
\gab{310}
Dicers and thimbleriggers we hurried by after the hoofs,
the vying caps and jackets
and past the meatfaced woman,
a butcher's dame,
nuzzling thirstily her clove of orange.

\N1:
Shouts rang shrill from the boys' playfield and a whirring whistle.

\StephenInt:
Again: a goal.
I am among them, among their battling bodies in a medley,
the joust of life.
You mean that knockkneed mother's darling who seems to be slightly crawsick?
Jousts.
Time shocked rebounds, shock by shock.
Jousts, slush and uproar of battles, the frozen deathspew of the slain,
a shout of spearspikes baited with men's bloodied guts.

