\Stephen:
You, Cochrane, what city sent for him?

\cochrane:
Tarentum, sir.

\Stephen:
Very good. Well?

\cochrane:
There was a battle, sir.

\Stephen:
Very good.
Where?

\N1:
The boy's blank face asked the blank window.

\StephenInt:
Fabled by the daughters of memory.
And yet it was in some way if not as memory fabled it.
A phrase, then, of impatience, thud of Blake's wings of excess.
I hear the ruin of all space, shattered glass and toppling masonry,
and time one livid final flame.
\gab{10}
What's left us then?

\cochrane:
I forget the place, sir.
279 B.C.

\Stephen:
Asculum,

\N2:
Stephen said, glancing at the name and date in the gorescarred book.

\cochrane:
Yes, sir.
And he said: \latin{Another victory like that and we are done for.}

\StephenInt:
That phrase the world had remembered.
A dull ease of the mind.
From a hill above a corpsestrewn plain
a general speaking to his officers, leaned upon his spear.
Any general to any officers.
They lend ear.

\Stephen:
You, Armstrong, \sout{Stephen said}.
What was the end of Pyrrhus?

\armstrong:
End of Pyrrhus, sir?

\comyn:
I know, sir.
\gab{20}
Ask me, sir,

\N1:
Comyn said.

\Stephen:
Wait.
You, Armstrong.
Do you know anything about Pyrrhus?

\N1:
A bag of figrolls lay snugly in Armstrong's satchel.
He curled them between his palms at whiles and swallowed them softly.
Crumbs adhered to the tissue of his lips.

\StephenInt:
A sweetened boy's breath.
Welloff people, proud that their eldest son was in the navy.
Vico road, Dalkey.

\armstrong:
Pyrrhus, sir?
Pyrrhus, a pier.

\N2:
All laughed.
Mirthless high malicious laughter.
Armstrong looked round at his classmates, silly glee in profile.

\StephenInt:
In a moment they will laugh more loudly,
aware of my lack of rule and of the fees their papas pay.

\Stephen:
Tell me now,

\N2:
Stephen said, poking the boy's shoulder with the book,
\gab{30}

\Stephen:
what is a pier.

\armstrong:
A pier, sir, \sout{Armstrong said}.
A thing out in the water.
A kind of a bridge.
Kingstown pier, sir.

\N2:
Some laughed again: mirthless but with meaning.
Two in the back bench whispered.

\StephenInt:
Yes.
They knew: had never learned nor ever been innocent.
All.

\N2:
With envy he watched their faces: Edith, Ethel, Gerty, Lily.
Their likes: their breaths, too,
sweetened with tea and jam,
their bracelets tittering in the struggle.

\Stephen:
Kingstown pier, \sout{Stephen said}.
Yes, a disappointed bridge.

\N1:
The words troubled their gaze.
\gab{40}

\comyn:
How, sir?
\sout{Comyn asked}.
A bridge is across a river.

\StephenInt:
For Haines's chapbook.
No-one here to hear.
Tonight deftly amid wild drink and talk,
to pierce the polished mail of his mind.
What then?
A jester at the court of his master, 
indulged and disesteemed,
winning a clement master's praise.
Why had they chosen all that part?
Not wholly for the smooth caress.
For them too history was a tale like any other too often heard,
their land a pawnshop.

Had Pyrrhus not fallen by a beldam's hand in Argos
or Julius Caesar not been knifed to death.
They are not to be thought away.
Time has branded them
\gab{50}
and fettered they are lodged in the room of the infinite possibilities
they have ousted.
But can those have been possible
seeing that they never were?
Or was that only possible which came to pass?
Weave, weaver of the wind.

\All:
Tell us a story, sir.

\All:
O, do, sir. A ghoststory.

\Stephen:
Where do you begin in this?

\N2:
Stephen asked, opening another book.

\comyn:
\Quotation{Weep no more}, \sout{Comyn said}.

\Stephen:
Go on then, Talbot.

\talbot:
And the story, sir?

\Stephen:
After, \sout{Stephen said}.
\gab{60}
Go on, Talbot.

\N2:
A swarthy boy opened a book
and propped it nimbly under the breastwork of his satchel.
He recited jerks of verse with odd glances at the text:

\talbot:
\begin{verse}
    Weep no more, woful shepherds, weep no more \\
    For Lycidas, your sorrow, is not dead, \\
    Sunk though he be beneath the watery floor ...
\end{verse}

\StephenInt:
It must be a movement then, an actuality of the possible as possible.

\N1:
Aristotle's phrase formed itself within the gabbled verses
and floated out into the studious silence of the library of Saint Genevieve
where he had read,
sheltered from the sin of Paris,
night by night.
\gab{70}
By his elbow a delicate Siamese conned a handbook of strategy.

\StephenInt:
Fed and feeding brains about me:
under glowlamps, impaled, with faintly beating feelers:
and in my mind's darkness a sloth of the underworld,
reluctant, shy of brightness, shifting her dragon scaly folds.
Thought is the thought of thought.
Tranquil brightness.
The soul is in a manner all that is:
the soul is the form of forms.
Tranquility sudden, vast, candescent: form of forms.

\N2:
Talbot repeated:

\begin{samepage}
\talbot:
\begin{verse}
    Through the dear might of him that walked the waves, \\
    Through the dear might ...
\end{verse}
\end{samepage}

\Stephen:
Turn over, \sout{Stephen said quietly}.
\gab{80}
I don't see anything.

\talbot:
What, sir?

\N1:
Talbot asked simply, bending forward.

\N2:
His hand turned the page over.
He leaned back and went on again, having just remembered.

\StephenInt:
Of him that walked the waves.
Here also over these craven hearts his shadow lies
and on the scoffer's heart and lips and on mine.
It lies upon their eager faces who offered him a coin of the tribute.
To Caesar what is Caesar's, to God what is God's.
A long look from dark eyes,
a riddling sentence to be woven and woven on the church's looms.
Ay.

\StephenInt:
\begin{verse}
    Riddle me, riddle me, randy ro. \\
    My father gave me seeds to sow.
\end{verse}

\N2:
Talbot slid his closed book into his satchel.
\gab{90}

\Stephen:
Have I heard all?
\sout{Stephen asked}.

\talbot:
Yes, sir. Hockey at ten, sir.

\armstrong:
Half day, sir.
Thursday.

\Stephen:
Who can answer a riddle?
\sout{Stephen asked}.

\N2:
They bundled their books away, pencils clacking, pages rustling.
Crowding together they strapped and buckled their satchels, all gabbling gaily:

\comyn:
A riddle, sir?
Ask me, sir.

\armstrong:
O, ask me, sir.

\cochrane:
A hard one, sir.
\gab{100}

\begin{samepage}
\Stephen:
This is the riddle,
\sout{Stephen said}:

\begin{verse}
    The cock crew, \\
    The sky was blue: \\
    The bells in heaven \\
    Were striking eleven. \\
    'Tis time for this poor soul \\
    To go to heaven.
\end{verse}
\end{samepage}

\Stephen:
What is that?

\comyn:
What, sir?

\armstrong:
Again, sir.
We didn't hear.
\gab{110}

\N2:
Their eyes grew bigger as the lines were repeated.
After a silence Cochrane said:

\cochrane:
What is it, sir?
We give it up.

\N1:
Stephen, his throat itching, answered:

\Stephen:
The fox burying his grandmother under a hollybush.

\N1:
He stood up and gave a shout of nervous laughter
to which their cries echoed dismay.

\N2:
A stick struck the door and a voice in the corridor called:

\deasy:
Hockey!

\N2:
They broke asunder,
\gab{120}
sidling out of their benches, leaping them.
Quickly they were gone
and from the lumberroom came the rattle of sticks
and clamour of their boots and tongues.
