\N1:
Martin Cunningham, first,
poked his silkhatted head into the creaking carriage
and, entering deftly, seated himself.
Mr Power stepped in after him,
curving his height with care.

\power:
Come on, Simon.

\Bloom:
After you,

\N1:
Mr Bloom said.

\N2:
Mr Dedalus covered himself quickly and got in, saying:

\simon:
Yes, yes.

\cunningham:
Are we all here now?

\N2:
Martin Cunningham asked.

\cunningham:
Come along, Bloom.

\N1:
Mr Bloom entered and sat in the vacant place.
He pulled the door to after him
and slammed it twice till it shut tight.
\gab{10}
He passed an arm through the armstrap
and looked seriously from the open carriagewindow
at the lowered blinds of the avenue.
One dragged aside:
an old woman peeping.

\BloomInt:
Nose whiteflattened against the pane.
Thanking her stars she was passed over.
Extraordinary the interest they take in a corpse.
Glad to see us go |
we give them such trouble coming.
Job seems to suit them.
Huggermugger in corners.
Slop about in slipperslappers for fear he'd wake.
Then getting it ready.
Laying it out.
Molly and Mrs Fleming making the bed.
Pull it more to your side.
Our windingsheet.
Never know who will touch you dead.
Wash and shampoo.
I believe they clip the nails and the hair.
Keep a bit in an envelope.
\gab{20}
Grows all the same after.
Unclean job.

\N1:
All waited.
Nothing was said.

\BloomInt:
Stowing in the wreaths probably.
I am sitting on something hard.
Ah, that soap:
in my hip pocket.
Better shift it out of that.
Wait for an opportunity.

\N1:
All waited.
Then wheels were heard from in front, turning:
then nearer:
then horses' hoofs.
A jolt.
Their carriage began to move, creaking and swaying.
Other hoofs and creaking wheels started behind.
The blinds of the avenue passed
and number nine with its craped knocker, door ajar.
At walking pace.

\N1:
They waited still, their knees jogging, till they had turned
and were passing along the tramtracks.
\gab{30}

\BloomInt:
Tritonville road.
Quicker.

\N1:
The wheels
rattled rolling over the cobbled causeway
and the crazy glasses shook rattling in the doorframes.

\power:
What way is he taking us?

\N2:
Mr Power asked through both windows.

\cunningham:
Irishtown,
\sout{Martin Cunningham said.}
Ringsend.
Brunswick street.

\N2:
Mr Dedalus nodded, looking out.

\Bloom:
That's a fine old custom,
\sout{he said.}
I am glad to see it has not died out.

\N1:
All watched awhile through their windows |
caps and hats lifted by passers.

\BloomInt:
Respect.

\N1:
The carriage swerved from the tramtrack to the smoother road past Watery lane.
Mr Bloom at gaze saw a lithe young man,
clad in mourning, a wide hat.

\Bloom:
There's a friend of yours gone by, Dedalus,
\gab{40}
\sout{he said.}

\simon:
Who is that?

\Bloom:
Your son and heir.

\simon:
Where is he?

\N2:
Mr Dedalus said,
stretching over across.

\N1:
The carriage, passing the open drains and mounds of rippedup roadway
before the tenement houses,
lurched round the corner
and, swerving back to the tramtrack, rolled on noisily with chattering wheels.
Mr Dedalus fell back, saying:

\simon:
Was that Mulligan cad with him?
His \latin{fidus achates!}

\Bloom:
No,
\sout{Mr Bloom said.}
\gab{50}
He was alone.

\simon:
Down with his aunt Sally, I suppose,
\sout{Mr Dedalus said,}
the Goulding faction,
the drunken little costdrawer
and Crissie, papa's little lump of dung,
the wise child that knows her own father.

\N1:
Mr Bloom smiled joylessly on Ringsend road.

\BloomInt:
Wallace Bros:
the bottleworks:
Dodder bridge.

\BloomInt:
Richie Goulding and the legal bag.
Goulding, Collis and Ward he calls the firm.
His jokes are getting a bit damp.
Great card he was.
Waltzing in Stamer street with Ignatius Gallaher on a Sunday morning,
the landlady's two hats pinned on his head.
Out on the rampage all night.
Beginning to tell on him now:
that backache of his, I fear.
\gab{60}
Wife ironing his back.
Thinks he'll cure it with pills.
All breadcrumbs they are.
About six hundred per cent profit.

\simon:
He's in with a lowdown crowd,

\N2:
Mr Dedalus snarled.

\simon:
That Mulligan is a contaminated bloody double-dyed ruffian |
by all accounts.
His name stinks all over Dublin.
But with the help of God and His blessed mother |
I'll make it my business to write a letter
one of those days |
to his mother or his aunt or whatever she is |
that will open her eye as wide as a gate.
I'll tickle his catastrophe, believe you me.

\N2:
He cried above the clatter of the wheels:

\simon:
I won't have her bastard of a nephew ruin my son.
\gab{70}
A counter\-jumper's son.
Selling tapes in my cousin, Peter Paul M'Swiney's.
Not likely.

\N2:
He ceased.

\N1:
Mr Bloom glanced from his angry moustache
to Mr Power's mild face
and Martin Cunningham's eyes and beard,
gravely shaking.

\BloomInt:
Noisy selfwilled man.
Full of his son.
He is right.
Something to hand on.
If little Rudy had lived.
See him grow up.
Hear his voice in the house.
Walking beside Molly in an Eton suit.
My son.
Me in his eyes.
Strange feeling it would be.
From me.
Just a chance.
Must have been that morning in Raymond terrace |
she was at the window watching the two dogs at it |
by the wall of the cease to do evil.
And the sergeant grinning up.
She had that cream gown on |
with the rip she never stitched.
\gab{80}
Give us a touch, Poldy.
God, I'm dying for it.
How life begins.

\BloomInt:
Got big then.
Had to refuse the Greystones concert.
My son inside her.
I could have helped him on in life.
I could.
Make him independent.
Learn German too.

\power:
Are we late?
\sout{Mr Power asked.}

\cunningham:
Ten minutes,

\N2:
Martin Cunningham said, looking at his watch.

\BloomInt:
Molly.
Milly.
Same thing watered down.
Her tomboy oaths.
O~jumping Jupiter!
Ye gods and little fishes!
Still, she's a dear girl.
Soon be a woman.
Mullingar.
Dearest Papli.
Young student.
Yes, yes:
a woman too.
Life,
\gab{90}
life.

\N1:
The carriage heeled over and back,
their four trunks swaying.

\power:
Corny might have given us a more commodious yoke,
\sout{Mr Power said.}

\simon:
He might,
\sout{Mr Dedalus said,}
if he hadn't that squint troubling him.
Do you follow me?

\N2:
He closed his left eye.
Martin Cunningham began to brush away crustcrumbs from under his thighs.

\cunningham:
What is this,
\sout{he said,}
in the name of God?
Crumbs?

\power:
Someone seems to have been making a picnic party here lately,
\sout{Mr Power said.}

\N1:
All raised their thighs and eyed with disfavour
\gab{100}
the mildewed buttonless leather of the seats.
Mr Dedalus, twisting his nose, frowned downward and said:

\simon:
Unless I'm greatly mistaken.
What do you think, Martin?

\cunningham:
It struck me too,
\sout{Martin Cunningham said.}

\N1:
Mr Bloom set his thigh down.

\BloomInt:
Glad I took that bath.
Feel my feet quite clean.
But I wish Mrs Fleming had darned these socks better.

\N2:
Mr Dedalus sighed resignedly.

\simon:
After all,
\sout{he said,}
it's the most natural thing in the world.

\cunningham:
Did Tom Kernan turn up?

\N1:
Martin Cunningham asked,
twirling the peak of his beard
\gab{110}
gently.

\Bloom:
Yes,
\sout{Mr Bloom answered.}
He's behind with Ned Lambert and Hynes.

\power:
And Corny Kelleher himself?
\sout{Mr Power asked.}

\cunningham:
At the cemetery,
\sout{Martin Cunningham said.}

\Bloom:
I met M'Coy this morning,
\sout{Mr Bloom said.}
He said he'd try to come.

\N1:
The carriage halted short.

\simon:
What's wrong?

\power:
We're stopped.

\cunningham:
Where are we?

\N1:
Mr Bloom put his head out of the window.

\Bloom:
The grand canal,
\gab{120}
\sout{he said.}

\BloomInt:
Gasworks.
Whooping cough they say it cures.
Good job Milly never got it.
Poor children!
Doubles them up black and blue in convulsions.
Shame really.
Got off lightly with illnesses compared.
Only measles.
Flaxseed tea.
Scarlatina, influenza epidemics.
Canvassing for death.
Don't miss this chance.
Dogs' home over there.
Poor old Athos!
[\stage{Virag}]
\emph{Be good to Athos, Leopold, is my last wish.}
Thy will be done.
We obey them in the grave.
A dying scrawl.
He took it to heart, pined away.
Quiet brute.
Old men's dogs usually are.

\N1:
A raindrop spat on his hat.
He drew back and saw an instant of shower |
spray dots over the grey flags.

\BloomInt:
Apart.
Curious.
\gab{130}
Like through a colander.
I thought it would.
My boots were creaking |
I remember now.

\Bloom:
The weather is changing,
\sout{he said quietly.}

\cunningham:
A pity it did not keep up fine,
\sout{Martin Cunningham said.}

\power:
Wanted for the country,
\sout{Mr Power said.}
There's the sun again coming out.

\N2:
Mr Dedalus, peering through his glasses towards the veiled sun,
hurled a mute curse at the sky.

\simon:
It's as uncertain as a child's bottom,
\sout{he said.}

\cunningham:
We're off again.

\N1:
The carriage turned again its stiff wheels
\gab{140}
and their trunks swayed gently.
Martin Cunningham twirled more quickly the peak of his beard.

\cunningham:
Tom Kernan was immense last night,
\sout{he said.}
And Paddy Leonard taking him off to his face.

\power:
O, draw him out, Martin,
\sout{Mr Power said eagerly.}
Wait till you hear him, Simon,
on Ben Dollard's singing of \Title{The Croppy Boy}.

\cunningham:
Immense,
\sout{Martin Cunningham said pompously.}
\emph{His singing of that simple ballad, Martin,
is the most trenchant rendering
I ever heard in the whole course of my experience.}

\power:
Trenchant,
\sout{Mr Power said laughing.}
He's dead nuts on that.
And the retrospective arrangement.%
\gab{150}

\cunningham:
Did you read Dan Dawson's speech?
\sout{Martin Cunningham asked.}

\simon:
I did not then,
\sout{Mr Dedalus said.}
Where is it?

\cunningham:
In the paper this morning.

\N1:
Mr Bloom took the paper from his inside pocket.

\BloomInt:
That book I must change for her.

\simon:
No, no,
\sout{Mr Dedalus said quickly.}
Later on please.

\N1:
Mr Bloom's glance travelled down the edge of the paper,
scanning the deaths:

\BloomInt:
Callan, Coleman, Dignam, Fawcett, Lowry, Naumann,
Peake, what Peake is that?
is it the chap was in Crosbie and Alleyne's?
no, Sexton, Urbright.%
\gab{160}

\N1:
Inked characters fast fading on the frayed breaking paper.

\BloomInt:
Thanks to the Little Flower.
Sadly missed.
To the inexpressible grief of his.
Aged 88 after a long and tedious illness.
Month's mind:
Quinlan.
On whose soul Sweet Jesus have mercy.

\N2:
\begin{verse}
    It is now a month since dear Henry fled \\
    To his home up above in the sky \\
    While his family weeps and mourns his loss \\
    Hoping some day to meet him on high.
\end{verse}

\BloomInt:
I tore up the envelope?
Yes.
Where did I put her letter after I read it in the bath?

\N1:
He patted his waistcoat-pocket.

\BloomInt:
There all right.
Dear Henry fled.
Before my patience
\gab{170}
are exhausted.

\BloomInt:
National school.
Meade's yard.
The hazard.
Only two there now.
Nodding.
Full as a tick.
Too much bone in their skulls.
The other trotting round with a fare.
An hour ago I was passing there.
The jarvies raised their hats.

\N1:
A pointsman's back straightened itself upright suddenly |
against a tramway standard |
by Mr Bloom's window.

\BloomInt:
Couldn't they invent something automatic
so that the wheel itself much handier?
Well but that fellow would lose his job then?
Well but then another fellow would get a job making the new invention?

\BloomInt:
Antient concert rooms.
\gab{180}
Nothing on there.
A man in a buff suit with a crape armlet.
Not much grief there.
Quarter mourning.
People in law perhaps.

\N2:
They went past the bleak pulpit of saint Mark's,
under the railway bridge,
past the Queen's theatre:
in silence.

\BloomInt:
Hoardings:
Eugene Stratton,
Mrs Bandmann Palmer.
Could I go to see \Title{Leah} tonight, I wonder.
I said I.
Or the \Title{Lily of Killarney}?
Elster Grimes Opera Company.
Big powerful change.
Wet bright bills for next week.
\Title{Fun on the Bristol}.
Martin Cunningham could work a pass for the Gaiety.
Have to stand a drink or two.
As broad as it's long.

\BloomInt:
He's coming in the afternoon.
\gab{190}
Her songs.

\BloomInt:
Plasto's.
Sir Philip Crampton's memorial fountain bust.
Who was he?

\cunningham:
How do you do?

\N1:
Martin Cunningham said,
raising his palm to his brow in salute.

\power:
He doesn't see us,
\sout{Mr Power said.}
Yes, he does.
How do you do?

\simon:
Who?
\sout{Mr Dedalus asked.}

\power:
Blazes Boylan,
\sout{Mr Power said.}
There he is airing his quiff.

\BloomInt:
Just that moment I was thinking.

\N2:
Mr Dedalus bent across to salute.
From the door of the Red Bank |
the white disc of a straw hat flashed reply:

\BloomInt:
spruce figure:
passed.

\N1:
Mr Bloom reviewed the nails of his left hand,
\gab{200}
then those of his right hand.

\BloomInt:
The nails, yes.
Is there anything more in him that \sout{they} she sees?
Fascination.
Worst man in Dublin.
That keeps him alive.
They sometimes feel what a person is.
Instinct.
But a type like that.
My nails.
I am just looking at them:
well pared.
And after:
thinking alone.
Body getting a bit softy.
I would notice that:
from remembering.
What causes that?
I suppose the skin can't contract quickly enough when the flesh falls off.
But the shape is there.
The shape is there still.
Shoulders.
Hips.
Plump.
Night of the dance dressing.
Shift stuck between the cheeks behind.

\N1:
He clasped his hands between his knees
and, satisfied,
sent his vacant glance
\gab{210}
over their faces.

\N2:
Mr Power asked:

\power:
How is the concert tour getting on, Bloom?

\Bloom:
O, very well,
\sout{Mr Bloom said.}
I hear great accounts of it.
It's a good idea, you see ...

\power:
Are you going yourself?

\Bloom:
Well no,
\sout{Mr Bloom said.}
In point of fact |
I have to go down to the county Clare on some private business.
You see |
the idea is to tour the chief towns.
What you lose on one |
you can make up on the other.

\cunningham:
Quite so,
\sout{Martin Cunningham said.}
Mary Anderson is up there now.
Have you
\gab{220}
good artists?

\Bloom:
Louis Werner is touring her,
\sout{Mr Bloom said.}
O~yes, we'll have all topnobbers.
J.C.~Doyle and John MacCormack I~hope and.
The best, in fact.

\power:
And \french{Madame},
\sout{Mr Power said smiling.}
Last but not least.

\N1:
Mr Bloom unclasped his hands in a gesture of soft politeness and clasped them.

\BloomInt:
Smith O'Brien.
Someone has laid a bunch of flowers there.
Woman.
Must be his deathday.
For many happy returns.

\N1:
The carriage wheeling by Farrell's statue
united noiselessly their unresisting knees.

\man:
Oot:

\N2:
a dullgarbed old man from the curbstone tendered his wares,
his mouth opening:
\gab{230}

\man:
oot.

\man:
Four bootlaces for a penny.

\BloomInt:
Wonder why he was struck off the rolls.
Had his office in Hume street.
Same house as Molly's namesake,
Tweedy, crown solicitor for Waterford.
Has that silk hat ever since.
Relics of old decency.
Mourning too.
Terrible comedown, poor wretch!
Kicked about like snuff at a wake.
O'Callaghan on his last legs.

\BloomInt:
And \french{Madame}.
Twenty past eleven.
Up.
Mrs Fleming is in to clean.
Doing her hair, humming.
\italian{Voglio e non vorrei.}
No.
\italian{Vorrei e non.}
Looking at the tips of her hairs to see if they are split.
\italian{Mi trema un poco il.}
Beautiful on that \italian{tre} her voice is:
weeping tone.
\gab{240}
A thrush.
A throstle.
There is a word throstle that expresses that.

\N2:
His eyes passed lightly over Mr Power's goodlooking face.

\BloomInt:
Greyish over the ears.
\french{Madame}:
smiling.
I smiled back.
A smile goes a long way.
Only politeness perhaps.
Nice fellow.
Who knows is that true about the woman he keeps?
Not pleasant for the wife.
Yet they say,
who was it told me,
there is no carnal.
You would imagine that would get played out pretty quick.
Yes, it was Crofton met him one evening |
bringing her a pound of rumpsteak.
What is this she was?
Barmaid in Jury's.
Or the Moira, was it?
