\BloomInt:
Paltry funeral:
coach and three carriages.
It's all the same.
Pallbearers, gold reins, requiem mass, firing a volley.
Pomp of death.

\N1:
Beyond the hind carriage
\gab{500}
a hawker stood by his barrow of cakes and fruit.

\BloomInt:
Simnel cakes those are, stuck together:
cakes for the dead.
Dogbiscuits.
Who ate them?
Mourners coming out.

\N1:
He followed his companions.

\N2:
Mr Kernan and Ned Lambert followed,
Hynes walking after them.
Corny Kelleher stood by the opened hearse and took out the two wreaths.
He handed one to the boy.

\BloomInt:
Where is that child's funeral disappeared to?

\N1:
A team of horses passed from Finglas with toiling plodding tread,
dragging through the funereal silence a creaking waggon |
on which lay a granite block.
The waggoner marching at their head saluted.

\BloomInt:
Coffin now.
Got here before us, dead as he is.
\gab{510}
Horse looking round at it with his plume skeowways.
Dull eye:
collar tight on his neck,
pressing on a bloodvessel or something.
Do they know what they cart out here every day?
Must be twenty or thirty funerals every day.
Then Mount Jerome for the protestants.
Funerals all over the world
everywhere every minute.
Shovelling them under by the cartload doublequick.
Thousands every hour.
Too many in the world.

\N1:
Mourners came out through the gates:

\BloomInt:
woman and a girl.

\BloomInt:
Leanjawed harpy, hard woman at a bargain, her bonnet awry.

\BloomInt:
Girl's face stained with dirt and tears,
holding the woman's arm, looking up at her for a sign to cry.

\BloomInt:
Fish's face,
\gab{520}
bloodless and livid.

\N2:
The mutes shouldered the coffin and bore it in through the gates.

\BloomInt:
So much dead weight.
Felt heavier myself stepping out of that bath.
First the stiff:
then the friends of the stiff.

\N2:
Corny Kelleher and the boy followed with their wreaths.

\BloomInt:
Who is that beside them?
Ah, the brother-in-law.

\N2:
All walked after.

\N1:
Martin Cunningham whispered:

\cunningham:
I was in mortal agony with you talking of suicide before Bloom.

\power:
What?

\N2:
Mr Power whispered.

\power:
What?
How so?

\cunningham:
His father poisoned himself,
\sout{Martin Cunningham whispered.}
Had the Queen's hotel in Ennis.
\gab{530}
You heard him say he was going to Clare.
Anniversary.

\power:
O God!
\sout{Mr Power whispered.}
First I heard of it.
Poisoned himself?

\N2:
He glanced behind him to where
a face with dark thinking eyes followed towards the cardinal's mausoleum.
Speaking.

\Bloom:
Was he insured?
\sout{Mr Bloom asked.}

\kernan:
I believe so,

\N1:
Mr Kernan answered.

\kernan:
But the policy was heavily mortgaged.
Martin is trying to get the youngster into Artane.

\Bloom:
How many children did he leave?

\kernan:
Five.
Ned Lambert says he'll try to get one of the girls into Todd's.

\Bloom:
A sad case,
\sout{Mr Bloom said gently.}
\gab{540}
Five young children.

\kernan:
A great blow to the poor wife,
\sout{Mr Kernan added.}

\Bloom:
Indeed yes,
\sout{Mr Bloom agreed.}

\BloomInt:
Has the laugh at him now.

\N1:
He looked down at the boots |
he had blacked and polished.

\BloomInt:
She had outlived him.
Lost her husband.
More dead for her than for me.
One must outlive the other.
Wise men say.
There are more women than men in the world.
Condole with her.
Your terrible loss.
I hope you'll soon follow him.
For Hindu widows only.
She would marry another.
Him?
No.
Yet who knows after.
Widowhood not the thing since the old queen died.
Drawn on a guncarriage.
\gab{550}
Victoria and Albert.
Frogmore memorial mourning.
But in the end she put a few violets in her bonnet.
Vain in her heart of hearts.
All for a shadow.
Consort not even a king.
Her son was the substance.
Something new to hope for |
not like the past she wanted back, waiting.
It never comes.
One must go first:
alone, under the ground:
and lie no more in her warm bed.

\lambert:
How are you, Simon?

\N1:
Ned Lambert said softly,
clasping hands.

\lambert:
Haven't seen you for a month of Sundays.

\simon:
Never better.
How are all in Cork's own town?

\lambert:
I was down there for the Cork park races on Easter Monday,
\sout{Ned Lambert said.}
Same old six and eightpence.
\gab{560}
Stopped with Dick Tivy.

\simon:
And how is Dick, the solid man?

\lambert:
Nothing between himself and heaven,
\sout{Ned Lambert answered.}

\simon:
By the holy Paul!
\sout{Mr Dedalus said in subdued wonder.}
Dick Tivy bald?

\lambert:
Martin is going to get up a whip for the youngsters,
\sout{Ned Lambert said, pointing ahead.}
A few bob a skull.
Just to keep them going till the insurance is cleared up.

\simon:
Yes, yes,
\sout{Mr Dedalus said dubiously.}
Is that the eldest boy in front?

\lambert:
Yes,
\sout{Ned Lambert said,}
with the wife's brother.
John Henry Menton is behind.
He put down his name for a quid.

\simon:
I'll engage he did,
\sout{Mr Dedalus said.}
\gab{570}
I often told poor Paddy he ought to mind that job.
John Henry is not the worst in the world.

\lambert:
How did he lose it?
\sout{Ned Lambert asked.}
Liquor, what?

\simon:
Many a good man's fault,
\sout{Mr Dedalus said with a sigh.}

\N1:
They halted about the door of the mortuary chapel.
Mr Bloom stood behind the boy with the wreath
looking down at his sleekcombed hair and
at the slender furrowed neck inside his brandnew collar.

\BloomInt:
Poor boy!
Was he there when the father?
Both unconscious.
Lighten up at the last moment and recognise for the last time.
All he might have done.
I owe three shillings to O'Grady.
Would he understand?

\N1:
The mutes bore the coffin into the chapel.
\gab{580}

\BloomInt:
Which end is his head?

\N1:
After a moment he followed the others in,
blinking in the screened light.

\N2:
The coffin lay on its bier before the chancel,
four tall yellow candles at its corners.

\BloomInt:
Always in front of us.

\N2:
Corny Kelleher, laying a wreath at each fore corner,
beckoned to the boy to kneel.
The mourners knelt here and there in prayingdesks.

\N1:
Mr Bloom stood behind near the font
and, when all had knelt,
dropped carefully his unfolded newspaper from his pocket
and knelt his right knee upon it.
He fitted his black hat gently on his left knee
and, holding its brim, bent over piously.

\N2:
A server bearing a brass bucket with something in it |
came out through a door.
The whitesmocked priest came after him,
\gab{590}
tidying his stole with one hand,
balancing with the other a little book against his toad's belly.

\BloomInt:
Who'll read the book?
I, said the rook.

\N2:
They halted by the bier
and the priest began to read out of his book
with a fluent croak.

\BloomInt:
Father Coffey.
I knew his name was like a coffin.
\latin{Domine-Namine.}
Bully about the muzzle he looks.
Bosses the show.
Muscular christian.
Woe betide anyone that looks crooked at him:
priest.
Thou art Peter.
Burst sideways like a sheep in clover |
Dedalus says he will.
With a belly on him like a poisoned pup.
Most amusing expressions that man finds.
Hhhn:
burst sideways.%
\gab{600}

\coffey:
\latin{Non intres in judicium cum servo tuo, Domine.}

\BloomInt:
Makes them feel more important to be prayed over in Latin.
Requiem mass.
Crape weepers.
Blackedged notepaper.
Your name on the altarlist.
Chilly place this.
Want to feed well, sitting in there all the morning
in the gloom kicking his heels
waiting for the next please.
Eyes of a toad too.
What swells him up that way?
Molly gets swelled after cabbage.
Air of the place maybe.
Looks full up of bad gas.
Must be an infernal lot of bad gas round the place.
Butchers, for instance:
they get like raw beefsteaks.
Who was telling me?
Mervyn Browne.
Down in the vaults of saint Werburgh's |
lovely old organ hundred and fifty |
\gab{610}
they have to bore a hole in the coffins sometimes |
to let out the bad gas and burn it.
Out it rushes:
blue.
One whiff of that and you're a doner.

\BloomInt:
My kneecap is hurting me.
Ow.
That's better.

\N2:
The priest took a stick with a knob at the end of it |
out of the boy's bucket and shook it over the coffin.
Then he walked to the other end and shook it again.
Then he came back and put it back in the bucket.

\BloomInt:
As you were before you rested.
It's all written down:
he has to do it.

\coffey:
\latin{Et ne nos inducas in tentationem.}

\N2:
The server piped the answers in the treble.

\BloomInt:
I often thought it would be better to have boy servants.
\gab{620}
Up to fifteen or so.
After that, of course ...

\BloomInt:
Holy water that was, I expect.
Shaking sleep out of it.
He must be fed up with that job,
shaking that thing over all the corpses they trot up.
What harm if he could see what he was shaking it over.
Every mortal day a fresh batch:
middleaged men, old women, children,
women dead in childbirth,
men with beards, baldheaded businessmen,
consumptive girls with little sparrows' breasts.
All the year round |
he prayed the same thing over them all and shook water on top of them:
sleep.
On Dignam now.

\coffey:
\latin{In paradisum.}

\BloomInt:
Said he was going to paradise or is in paradise.
Says that over everybody.
\gab{630}
Tiresome kind of a job.
But he has to say something.

\N2:
The priest closed his book and went off, followed by the server.

\N1:
Corny Kelleher opened the sidedoors and the gravediggers came in,
hoisted the coffin again, carried it out and shoved it on their cart.
Corny Kelleher gave one wreath to the boy and one to the brother-in-law.
All followed them out of the sidedoors into the mild grey air.
Mr Bloom came last |
folding his paper again into his pocket.
He gazed gravely at the ground |
till the coffincart wheeled off to the left.
The metal wheels ground the gravel
with a sharp grating cry |
and the pack of blunt boots followed the trundled barrow
along a lane of sepulchres.

\BloomInt:
The ree |
the ra |
the ree |
the ra |
the roo.
\gab{640}
Lord, I mustn't lilt here.

\simon:
The O'Connell circle,
\sout{Mr Dedalus said about him.}

\N1:
Mr Power's soft eyes went up to the apex of the lofty cone.

\power:
He's at rest,
\sout{he said,}
in the middle of his people, old Dan O'.
But his heart is buried in Rome.
How many broken hearts are buried here, Simon!

\simon:
Her grave is over there, Jack,
\sout{Mr Dedalus said.}
I'll soon be stretched beside her.
Let Him take me whenever He likes.

\N2:
Breaking down, he began to weep to himself quietly,
stumbling a little in his walk.
Mr Power took his arm.

\power:
She's better where she is,
\sout{he said kindly.}

\simon:
I suppose so,
\sout{Mr Dedalus said with a weak gasp.}
\gab{650}
I suppose she is in heaven |
if there is a heaven.

\N1:
Corny Kelleher stepped aside from his rank and allowed the mourners to plod by.

\kernan:
Sad occasions,

\N2:
Mr Kernan began politely.

\N1:
Mr Bloom closed his eyes and sadly twice bowed his head.

\kernan:
The others are putting on their hats,
\sout{Mr Kernan said.}
I suppose we can do so too.
We are the last.
This cemetery is a treacherous place.

\N1:
They covered their heads.

\kernan:
The reverend gentleman read the service too quickly, don't you think?

\N2:
Mr Kernan said with reproof.
\gab{660}

\N1:
Mr Bloom nodded gravely |
looking in the quick bloodshot eyes.

\BloomInt:
Secret eyes, secretsearching.
Mason, I think:
not sure.
Beside him again.
We are the last.
In the same boat.
Hope he'll say something else.

\N2:
Mr Kernan added:

\kernan:
The service of the Irish church used in Mount Jerome
is simpler, more impressive |
I must say.

\N1:
Mr Bloom gave prudent assent.

\BloomInt:
The language of course was another thing.

\N2:
Mr Kernan said with solemnity:

\kernan:
\emph{I am the resurrection and the life}.
\gab{670}
That touches a man's inmost heart.

\Bloom:
It does,
\sout{Mr Bloom said.}

\BloomInt:
Your heart perhaps |
but what price the fellow
in the six feet by two |
with his toes to the daisies?
No touching that.
Seat of the affections.
Broken heart.
A pump after all, pumping thousands of gallons of blood every day.
One fine day it gets bunged up:
and there you are.
Lots of them lying around here:
lungs, hearts, livers.
Old rusty pumps:
damn the thing else.
The resurrection and the life.
Once you are dead |
you are dead.
That last day idea.
Knocking them all up out of their graves.
Come forth, Lazarus!
And he came fifth and lost the job.
Get up!
Last day!
Then every fellow mousing around for his liver and his lights
and the rest of his traps.
\gab{680}
Find damn all of himself that morning.
Pennyweight of powder in a skull.
Twelve grammes |
one pennyweight.
Troy measure.

\N1:
Corny Kelleher fell into step at their side.

\corny:
Everything went off A1,
\sout{he said.}
What?

\N1:
He looked on them from his drawling eye.

\BloomInt:
Policeman's shoulders.
With your tooraloom tooraloom.

\kernan:
As it should be,
\sout{Mr Kernan said.}

\corny:
What?
Eh?
\sout{Corny Kelleher said.}

\N2:
Mr Kernan assured him.
