\N2:
They passed under the hugecloaked Liberator's form.

\N1:
Martin Cunningham nudged Mr Power.
\gab{250}

\cunningham:
Of the tribe of Reuben,

\N1:
he said.

\N2:
A tall blackbearded figure, bent on a stick,
stumping round the corner of Elvery's Elephant house,
showed them a curved hand open on his spine.

\power:
In all his pristine beauty,

\N1:
Mr Power said.

\N2:
Mr Dedalus looked after the stumping figure and said mildly:

\simon:
The devil break the hasp of your back!

\N1:
Mr Power, collapsing in laughter,
shaded his face from the window |
as the carriage passed Gray's statue.

\cunningham:
We have all been there,
\sout{Martin Cunningham said broadly.}

\N1:
His eyes met Mr Bloom's eyes.
\gab{260}
He caressed his beard, adding:

\cunningham:
Well, nearly all of us.

\N2:
Mr Bloom began to speak with sudden eagerness to his companions' faces.

\Bloom:
That's an awfully good one that's going the rounds
about Reuben J and the son.

\power:
About the boatman?
\sout{Mr Power asked.}

\Bloom:
Yes.
Isn't it awfully good?

\simon:
What is that?
\sout{Mr Dedalus asked.}
I didn't hear it.

\Bloom:
There was a girl in the case,
\sout{Mr Bloom began,}
and he determined to send him
to the Isle of Man out of harm's way
\gab{270}
but when they were both ...

\simon:
What?
\sout{Mr Dedalus asked.}
That confirmed bloody hobble\-dehoy is it?

\Bloom:
Yes,
\sout{Mr Bloom said.}
They were both on the way to the boat and he tried to drown ...

\simon:
Drown Barabbas!
\sout{Mr Dedalus cried.}
I wish to Christ he did!

\N2:
Mr Power sent a long laugh down his shaded nostrils.

\Bloom:
No,
\sout{Mr Bloom said,}
the son himself ...

\N1:
Martin Cunningham thwarted his speech rudely:

\cunningham:
Reuben and the son were piking it down the quay next the river
on their way to the Isle of Man boat
and the young chiseller suddenly got loose
and over the wall with him
\gab{280}
into the Liffey.

\simon:
For God's sake!
\sout{Mr Dedalus exclaimed in fright.}
Is he dead?

\cunningham:
Dead!
\sout{Martin Cunningham cried.}
Not he!
A boatman got a pole and fished him out by the slack of the breeches
and he was landed up to the father on the quay
more dead than alive.
Half the town was there.

\Bloom:
Yes,
\sout{Mr Bloom said.}
But the funny part is ...

\cunningham:
And Reuben J,
\sout{Martin Cunningham said,}
gave the boatman a florin for saving his son's life.

\N2:
A stifled sigh came from under Mr Power's hand.

\cunningham:
O, he did,
\sout{Martin Cunningham affirmed.}
Like a hero.
A silver florin.

\Bloom:
Isn't it awfully good?
\gab{290}
\sout{Mr Bloom said eagerly.}

\simon:
One and eightpence too much,
\sout{Mr Dedalus said drily.}

\N2:
Mr Power's choked laugh
[\stage{laughter}]
burst quietly in the carriage.

\BloomInt:
Nelson's pillar.

\woman:
Eight plums a penny!
Eight for a penny!

\cunningham:
We had better look a little serious,
\sout{Martin Cunningham said.}

\N2:
Mr Dedalus sighed.

\simon:
Ah then indeed,
\sout{he said,}
poor little Paddy wouldn't grudge us a laugh.
Many a good one he told himself.

\power:
The Lord forgive me!

\N2:
Mr Power said,
wiping his wet eyes with his fingers.

\power:
Poor Paddy!
\gab{300}
I little thought a week ago |
when I saw him last |
and he was in his usual health |
that I'd be driving after him like this.
He's gone from us.

\simon:
As decent a little man as ever wore a hat,
\sout{Mr Dedalus said.}
He went very suddenly.

\cunningham:
Breakdown,
\sout{Martin Cunningham said.}
Heart.

\N1:
He tapped his chest sadly.

\BloomInt:
Blazing face:
redhot.
Too much John Barleycorn.
Cure for a red nose.
Drink like the devil till it turns adelite.
A lot of money he spent colouring it.

\N1:
Mr Power gazed at the passing houses
\gab{310}
with rueful apprehension.

\power:
He had a sudden death, poor fellow,
\sout{he said.}

\Bloom:
The best death,
\sout{Mr Bloom said.}

\N1:
Their wide open eyes looked at him.

\Bloom:
No suffering,
\sout{he said.}
A moment and all is over.
Like dying in sleep.

\N1:
No-one spoke.

\BloomInt:
Dead side of the street this.
Dull business by day,
land agents, temperance hotel,
Falconer's railway guide, civil service college,
Gill's, catholic club, the industrious blind.
Why?
Some reason.
Sun or wind.
At night too.
Chummies and slaveys.
Under the patronage of the late Father Mathew.
\gab{320}
Foundation stone for Parnell.
Breakdown.
Heart.

\N2:
White horses with white frontlet plumes came round the Rotunda corner,
galloping.
A tiny coffin flashed by.

\BloomInt:
In a hurry to bury.
A mourning coach.
Unmarried.
Black for the married.
Piebald for bachelors.
Dun for a nun.

\cunningham:
Sad,
\sout{Martin Cunningham said.}
A child.

\BloomInt:
A dwarf's face, mauve and wrinkled like little Rudy's was.
Dwarf's body, weak as putty, in a whitelined deal box.
Burial friendly society pays.
Penny a week for a sod of turf.
Our.
Little.
Beggar.
Baby.
Meant nothing.
Mistake of nature.
If it's healthy |
it's from the mother.
If not |
from the man.
Better luck
\gab{330}
next time.

\simon:
Poor little thing,
\sout{Mr Dedalus said.}
It's well out of it.

\N2:
The carriage climbed more slowly the hill of Rutland square.

\BloomInt:
Rattle his bones.
Over the stones.
Only a pauper.
Nobody owns.

\cunningham:
In the midst of life,
\sout{Martin Cunningham said.}

\power:
But the worst of all,
\sout{Mr Power said,}
is the man who takes his own life.

\N1:
Martin Cunningham drew out his watch briskly,
coughed and put it back.

\power:
The greatest disgrace to have in the family,
\sout{Mr Power added.}

\cunningham:
Temporary insanity, of course,
\sout{Martin Cunningham said decisively.}
We must take
\gab{340}
a charitable view of it.

\simon:
They say a man who does it is a coward,
\sout{Mr Dedalus said.}

\cunningham:
It is not for us to judge,
\sout{Martin Cunningham said.}

\N1:
Mr Bloom, about to speak, closed his lips again.

\BloomInt:
Martin Cunningham's large eyes.

\BloomInt:
Looking away now.
Sympathetic human man he is.
Intelligent.
Like Shakespeare's face.
Always a good word to say.
They have no mercy on that here or infanticide.
Refuse christian burial.
They used to drive a stake of wood through his heart in the grave.
As if it wasn't broken already.
Yet sometimes they repent too late.
Found in the riverbed clutching rushes.
He looked at me.
And that awful drunkard of a wife of his.
\gab{350}
Setting up house for her time after time
and then pawning the furniture on him every Saturday almost.
Leading him the life of the damned.
Wear the heart out of a stone, that.
Monday morning.
Start afresh.
Shoulder to the wheel.
Lord, she must have looked a sight that night |
Dedalus told me he was in there.
Drunk about the place and capering with Martin's umbrella.

\BloomInt:
\begin{verse}
    And they call me the jewel of Asia, \\
    Of Asia, \\
    The geisha.
\end{verse}

\BloomInt:
He looked away from me.
He knows.
Rattle his bones.

\BloomInt:
That afternoon of the inquest.
The redlabelled bottle on the table.
The room in the hotel with hunting pictures.
\gab{360}
Stuffy it was.
Sunlight through the slats of the Venetian blind.
The coroner's sunlit ears, big and hairy.
Boots giving evidence.
Thought he was asleep first.
Then saw like yellow streaks on his face.
Had slipped down to the foot of the bed.
Verdict:
overdose.
Death by misadventure.
The letter.
For my son Leopold.

\BloomInt:
No more pain.
Wake no more.
Nobody owns.

\N2:
The carriage rattled swiftly along Blessington street.
Over the stones.

\cunningham:
We are going the pace, I think,
\sout{Martin Cunningham said.}

\power:
God grant he doesn't upset us on the road,
\sout{Mr Power said.}

\cunningham:
I hope not,
\sout{Martin Cunningham said.}
That will be a great race tomorrow in Germany.
\gab{370}
The Gordon Bennett.

\simon:
Yes, by Jove,
\sout{Mr Dedalus said.}
That will be worth seeing, faith.

\N2:
As they turned into Berkeley street |
a streetorgan near the Basin sent over and after them |
a rollicking rattling song of the halls.

\man:
Has anybody here seen Kelly?
Kay ee double ell wy.

\BloomInt:
Dead March from \Title{Saul}.

\man:
He's as bad
as old Antonio.
He left me on my ownio.
Pirouette!

\BloomInt:
The \latin{Mater Misericordiae}.
Eccles street.
My house down there.
Big place.
Ward for incurables there.
Very encouraging.
Our Lady's Hospice for the dying.
Deadhouse handy underneath.
Where old Mrs Riordan died.
They look terrible the women.
Her feeding cup and rubbing her mouth with the spoon.
\gab{380}
Then the screen round her bed for her to die.
Nice young student that was dressed that bite the bee gave me.
He's gone over to the lying-in hospital they told me.
From one extreme to the other.

\N2:
The carriage galloped round a corner:
stopped.

\cunningham:
What's wrong now?

\N1:
A divided drove of branded cattle passed the windows,
lowing,
slouching by on padded hoofs,
whisking their tails slowly on their clotted bony croups.
Outside them and through them
ran raddled sheep
bleating their fear.

\power:
Emigrants,
\sout{Mr Power said.}

\drover:
Huuuh!

\N1:
the drover's voice cried,
\gab{390}
his switch sounding on their flanks.

\drover:
Huuuh!
out of that!

\BloomInt:
Thursday, of course.
Tomorrow is killing day.
Springers.
Cuffe sold them |
about twentyseven quid each.
For Liverpool probably.
Roastbeef for old England.
They buy up all the juicy ones.
And then the fifth quarter lost:
all that raw stuff, hide, hair, horns.
Comes to a big thing in a year.
Dead meat trade.
Byproducts of the slaughterhouses for tanneries, soap, margarine.
Wonder if that dodge works now |
getting dicky meat off the train at Clonsilla.

\N1:
The carriage moved on through the drove.

\Bloom:
I can't make out why the corporation doesn't run a tramline
\gab{400}
from the parkgate to the quays,
\sout{Mr Bloom said.}
All those animals could be taken in trucks down to the boats.

\cunningham:
Instead of blocking up the thoroughfare,
\sout{Martin Cunningham said.}
Quite right.
They ought to.

\Bloom:
Yes,
\sout{Mr Bloom said,}
and another thing I often thought,
is to have municipal funeral trams like they have in Milan, you know.
Run the line out to the cemetery gates and have special trams,
hearse and carriage and all.
Don't you see what I mean?

\simon:
O, that be damned for a story,
\sout{Mr Dedalus said.}
Pullman car and saloon diningroom.%
\gab{410}

\power:
A poor lookout for Corny,
\sout{Mr Power added.}

\Bloom:
Why?
\sout{Mr Bloom asked, turning to Mr Dedalus.}
Wouldn't it be more decent than galloping two abreast?

\simon:
Well, there's something in that,
\sout{Mr Dedalus granted.}

\cunningham:
And,
\sout{Martin Cunningham said,}
we wouldn't have scenes like that
when the hearse capsized round Dunphy's
and upset the coffin on to the road.

\power:
That was terrible,
\sout{Mr Power's shocked face said,}
and the corpse fell about the road.
Terrible!

\simon:
First round Dunphy's,
\sout{Mr Dedalus said, nodding.}
Gordon Bennett cup.

\cunningham:
Praises be to God!
\gab{420}
\sout{Martin Cunningham said piously.}

\BloomInt:
Bom!
Upset.
A coffin bumped out on to the road.
Burst open.
Paddy Dignam shot out and rolling over stiff in the dust
in a brown habit too large for him.
Red face:
grey now.
Mouth fallen open.
Asking what's up now.
Quite right to close it.
Looks horrid open.
Then the insides decompose quickly.
Much better to close up all the orifices.
Yes, also.
With wax.
The sphincter loose.
Seal up all.

\power:
Dunphy's,

\N2:
Mr Power announced
as the carriage turned right.

\BloomInt:
Dunphy's corner.
Mourning coaches drawn up, drowning their grief.
A pause by the wayside.
Tiptop position for a pub.
Expect we'll pull up here on the way back to drink his health.
\gab{430}
Pass round the consolation.
Elixir of life.

\BloomInt:
But suppose now it did happen.
Would he bleed if a nail say cut him in the knocking about?
He would and he wouldn't, I suppose.
Depends on where.
The circulation stops.
Still |
some might ooze out of an artery.
It would be better to bury them in red:
a dark red.

\N2:
In silence they drove along Phibsborough road.
An empty hearse trotted by, coming from the cemetery:

\BloomInt:
looks relieved.

\BloomInt:
Crossguns bridge:
the royal canal.

\N1:
Water rushed roaring through the sluices.
A man stood on his
dropping barge, between clamps of turf.
\gab{440}
On the towpath by the lock |
a slacktethered horse.
Aboard of the \Title{Bugabu}.

\N1:
Their eyes watched him.
On the slow weedy waterway
he had floated on his raft coastward over Ireland |
drawn by a haulage rope |
past beds of reeds,
over slime, mudchoked bottles, carrion dogs.

\BloomInt:
Athlone, Mullingar, Moyvalley,
I could make a walking tour to see Milly by the canal.
Or cycle down.
Hire some old crock, safety.
Wren had one the other day at the auction |
but a lady's.
Developing waterways.
James M'Cann's hobby to row me o'er the ferry.
Cheaper transit.
By easy stages.
Houseboats.
Camping out.
Also hearses.
To heaven by water.
Perhaps I will without writing.
Come as a surprise, Leixlip, Clonsilla.
\gab{450}
Dropping down lock by lock to Dublin.
With turf from the midland bogs.
Salute.

\N1:
He lifted his brown straw hat,
saluting Paddy Dignam.

\N2:
They drove on past Brian Boroimhe house.

\BloomInt:
Near it now.

\power:
I wonder how is our friend Fogarty getting on,
\sout{Mr Power said.}

\simon:
Better ask Tom Kernan,
\sout{Mr Dedalus said.}

\cunningham:
How is that?
\sout{Martin Cunningham said.}
Left him weeping, I suppose?

\simon:
Though lost to sight,
\sout{Mr Dedalus said,}
to memory dear.

\N2:
The carriage steered left for Finglas road.

\BloomInt:
The stonecutter's yard on the right.
Last lap.

\N1:
Crowded on the spit of land
silent shapes appeared,
\gab{460}
white, sorrowful, holding out calm hands,
knelt in grief, pointing.
Fragments of shapes, hewn.
In white silence:
appealing.

\BloomInt:
The best obtainable.
Thos.~H.~Dennany, monumental builder and sculptor.

\BloomInt:
Passed.

\N2:
On the curbstone before Jimmy Geary, the sexton's,
an old tramp sat, grumbling,
emptying the dirt and stones out of his huge dustbrown yawning boot.

\BloomInt:
After life's journey.

\N1:
Gloomy gardens then went by:
one by one:
gloomy houses.

\N2:
Mr Power pointed.

\power:
That is where Childs was murdered,
\sout{he said.}
The last house.

\simon:
So it is,
\sout{Mr Dedalus said.}
A gruesome case.
\gab{470}
Seymour Bushe got him off.
Murdered his brother.
Or so they said.

\power:
The crown had no evidence,
\sout{Mr Power said.}

\cunningham:
Only circumstantial,
\sout{Martin Cunningham added.}
That's the maxim of the law.
Better for ninetynine guilty to escape
than for one innocent person to be wrongfully condemned.

\N2:
They looked.
Murderer's ground.
It passed darkly.
Shuttered, tenantless, unweeded garden.

\BloomInt:
Whole place gone to hell.
Wrongfully condemned.
Murder.
The murderer's image in the eye of the murdered.
They love reading about it.
Man's head found in a garden.
Her clothing consisted of.
How she met her death.
Recent outrage.
The weapon used.
Murderer is still at large.
Clues.
\gab{480}
A shoelace.
The body to be exhumed.
Murder will out.

\BloomInt:
Cramped in this carriage.
She mightn't like me to come that way without letting her know.
Must be careful about women.
Catch them once with their pants down.
Never forgive you after.
Fifteen.

\N1:
The high railings of Prospect rippled past their gaze.
Dark poplars, rare white forms.
Forms more frequent,
white shapes thronged amid the trees,
white forms and fragments streaming by mutely,
sustaining vain gestures on the air.

\N1:
The felly harshed against the curbstone:
stopped.
\gab{490}
Martin Cunningham put out his arm
and, wrenching back the handle,
shoved the door open with his knee.
He stepped out.
Mr Power and Mr Dedalus followed.

\BloomInt:
Change that soap now.

\N1:
Mr Bloom's hand unbuttoned his hip pocket swiftly
and transferred the paperstuck soap to his inner handkerchief pocket.
He stepped out of the carriage,
replacing the newspaper |
his other hand still held.
