\menton:
Who is that chap behind with Tom Kernan?

\N1:
John Henry Menton asked.
\gab{690}

\menton:
I know his face.

\N2:
Ned Lambert glanced back.

\lambert:
Bloom,
\sout{he said,}
Madame Marion Tweedy that was,
is, I mean,
the soprano.
She's his wife.

\menton:
O, to be sure,
\sout{John Henry Menton said.}
I haven't seen her for some time.
She was a finelooking woman.
I danced with her, wait, fifteen seventeen golden years ago,
at Mat Dillon's in Roundtown.
And a good armful she was.

\N1:
He looked behind through the others.

\menton:
What is he?
\sout{he asked.}
\gab{700}
What does he do?
Wasn't he in the stationery line?
I fell foul of him one evening, I remember, at bowls.

\N2:
Ned Lambert smiled.

\lambert:
Yes, he was,
\sout{he said,}
in Wisdom Hely's.
A traveller for blottingpaper.

\menton:
In God's name,
\sout{John Henry Menton said,}
what did she marry a coon like that for?
She had plenty of game in her then.

\lambert:
Has still,
\sout{Ned Lambert said.}
He does some canvassing for ads.

\N1:
John Henry Menton's large eyes stared ahead.

\N2:
The barrow turned into a side lane.
A portly man, ambushed among the grasses, raised his hat in homage.
The gravediggers touched their caps.

\power:
John O'Connell,
\sout{Mr Power said pleased.}
\gab{710}
He never forgets a friend.

\N2:
Mr O'Connell shook all their hands in silence.
Mr Dedalus said:

\simon:
I am come to pay you another visit.

\oconnell:
My dear Simon,

\N2:
the caretaker answered in a low voice.

\oconnell:
I don't want your custom at all.

\N2:
Saluting Ned Lambert and John Henry Menton |
he walked on at Martin Cunningham's side |
puzzling two long keys at his back.

\oconnell:
Did you hear that one,
\sout{he asked them,}
about Mulcahy from the Coombe?

\cunningham:
I did not,
\sout{Martin Cunningham said.}

\N2:
They bent their silk hats in concert and Hynes inclined his ear.
The caretaker hung his thumbs in the loops of his gold watchchain |
\gab{720}
and spoke in a discreet tone to their vacant smiles.

\oconnell:
They tell the story,
\sout{he said,}
that two drunks came out here one foggy evening
to look for the grave of a friend of theirs.
They asked for Mulcahy from the Coombe and were told where he was buried.
After traipsing about in the fog they found the grave sure enough.
One of the drunks spelt out the name:
Terence Mulcahy.
The other drunk was blinking up at a statue of Our Saviour
the widow had got put up.

\N2:
The caretaker blinked up at one of the sepulchres they passed.
He resumed:

\oconnell:
And, after blinking up at the sacred figure,
\gab{730}
\emph{Not a bloody bit like the man,} says he.
\emph{That's not Mulcahy,} says he,
\emph{whoever done it.}

\N2:
Rewarded by smiles |
he fell back and spoke with Corny Kelleher,
accepting the dockets given him,
turning them over and scanning them |
as he walked.

\cunningham:
That's all done with a purpose,

\N1:
Martin Cunningham explained to Hynes.

\hynes:
I know,

\N1:
Hynes said.

\hynes:
I know that.

\cunningham:
To cheer a fellow up,
\sout{Martin Cunningham said.}
It's pure goodheartedness:
damn the thing else.

\N1:
Mr Bloom admired the caretaker's prosperous bulk.

\BloomInt:
All want to be on good terms with him.
\gab{740}
Decent fellow, John O'Connell, real good sort.
Keys:
like Keyes's ad:
no fear of anyone getting out.
No passout checks.
\latin{Habeas corpus.}
I must see about that ad after the funeral.
Did I write Ballsbridge on the envelope I took to cover |
when she disturbed me writing to Martha?
Hope it's not chucked in the dead letter office.
Be the better of a shave.
Grey sprouting beard.
That's the first sign |
when the hairs come out grey.
And temper getting cross.
Silver threads among the grey.
Fancy being his wife.
Wonder he had the gumption to propose to any girl.
Come out and live in the graveyard.
Dangle that before her.
It might thrill her first.
Courting death ...
Shades of night hovering here with all the dead stretched about.
The shadows of the tombs when churchyards yawn |
\gab{750}
and Daniel O'Connell |
must be a descendant I suppose |
who is this used to say |
he was a queer breedy man |
great catholic all the same |
like a big giant in the dark.
Will o' the wisp.
Gas of graves.
Want to keep her mind off it |
to conceive at all.
Women especially are so touchy.
Tell her a ghost story in bed to make her sleep.
Have you ever seen a ghost?
Well, I have.
It was a pitchdark night.
The clock was on the stroke of twelve.
Still they'd kiss all right if properly keyed up.
Whores in Turkish graveyards.
Learn anything if taken young.
You might pick up a young widow here.
Men like that.
Love among the tombstones.
Romeo.
Spice of pleasure.
In the midst of death |
we are in life.
Both ends meet.
Tantalising for the poor dead.
\gab{760}
Smell of grilled beefsteaks to the starving.
Gnawing their vitals.
Desire to grig people.
Molly wanting to do it at the window.
Eight children he has anyway.

\BloomInt:
He has seen a fair share go under in his time,
lying around him field after field.
Holy fields.
More room if they buried them standing.
Sitting or kneeling |
you couldn't.
Standing?
His head might come up some day above ground in a landslip with his hand pointing.
All honeycombed the ground must be:
oblong cells.
And very neat he keeps it too:
trim grass and edgings.
His garden |
Major Gamble calls Mount Jerome.
Well, so it is.
Ought to be flowers of sleep.
Chinese cemeteries with giant poppies growing produce the best opium |
Mastiansky told me.
\gab{770}
The Botanic Gardens are just over there.
It's the blood sinking in the earth gives new life.
Same idea those jews |
they said killed the christian boy.
Every man his price.
Well preserved fat corpse, gentleman, epicure,
invaluable for fruit garden.
A bargain.
By carcass of William Wilkinson,
auditor and accountant, lately deceased,
three pounds thirteen and six.
With thanks.

\BloomInt:
I daresay the soil would be quite fat with corpsemanure, bones, flesh, nails.
Charnelhouses.
Dreadful.
Turning green and pink decomposing.
Rot quick in damp earth.
The lean old ones tougher.
Then a kind of a tallowy kind of a cheesy.
Then begin to get black, black treacle oozing out of them.
Then dried up.
Deathmoths.
\gab{780}
Of course the cells or whatever they are go on living.
Changing about.
Live for ever practically.
Nothing to feed on |
feed on themselves.

\BloomInt:
But they must breed a devil of a lot of maggots.
Soil must be simply swirling with them.
Your head |
it simply swurls.
Those pretty little seaside gurls.
He looks cheerful enough over it.
Gives him a sense of power |
seeing all the others go under first.
Wonder how he looks at life.
Cracking his jokes too:
warms the cockles of his heart.
The one about the bulletin.
Spurgeon went to heaven 4 a.m. this morning.
11 p.m. (closing time).
Not arrived yet.
Peter.
The dead themselves
the men anyhow would like to hear an odd joke |
\gab{790}
or the women to know what's in fashion.
A juicy pear or ladies' punch, hot, strong and sweet.
Keep out the damp.
You must laugh sometimes so better do it that way.
Gravediggers in \Title{Hamlet}.
Shows the profound knowledge of the human heart.
Daren't joke about the dead for two years at least.
\latin{De mortuis nil nisi prius.}
Go out of mourning first.
Hard to imagine his funeral.
Seems a sort of a joke.
Read your own obituary notice they say you live longer.
Gives you second wind.
New lease of life.

\oconnell:
How many have-you for tomorrow?

\N1:
the caretaker asked.

\corny:
Two,

\N2:
Corny Kelleher said.

\corny:
Half ten and eleven.

\N1:
The caretaker put the papers in his pocket.
The barrow had ceased to trundle.
\gab{800}
The mourners split and moved to each side of the hole,
stepping with care round the graves.
The gravediggers bore the coffin
and set its nose on the brink,
looping the bands round it.

\BloomInt:
Burying him.
We come to bury Caesar.
His ides of March or June.
He doesn't know who is here nor care.
Now who is that lankylooking galoot over there in the macintosh?
Now who is he |
I'd like to know?
Now I'd give a trifle to know who he is.
Always someone turns up you never dreamt of.
A fellow could live on his lonesome all his life.
Yes, he could.
Still he'd have to get someone to sod him after he died |
though he could dig his own grave.
We all do.
Only man buries.
No, ants too.
\gab{810}
First thing strikes anybody.
Bury the dead.
Say Robinson Crusoe was true to life.
Well then Friday buried him.
Every Friday buries a Thursday if you come to look at it.

\BloomInt:
\begin{verse}
    O, poor Robinson Crusoe! \\
    How could you possibly do so?
\end{verse}

\BloomInt:
Poor Dignam!
His last lie on the earth in his box.
When you think of them all
it does seem a waste of wood.
All gnawed through.
They could invent a handsome bier with a kind of panel sliding,
let it down that way.
Ay but they might object to be buried out of another fellow's.
They're so particular.
Lay me in my native earth.
Bit of clay from the holy land.
Only a mother and deadborn child ever buried in the one coffin.
\gab{820}
I see what it means.
I see.
To protect him as long as possible |
even in the earth.
The Irishman's house is his coffin.
Embalming in catacombs, mummies the same idea.

\N1:
Mr Bloom stood far back, his hat in his hand,
counting the bared heads.

\BloomInt:
Twelve.
I'm thirteen.
No.
The chap in the macintosh is thirteen.
Death's number.
Where the deuce did he pop out of?
He wasn't in the chapel, that I'll swear.
Silly superstition that about thirteen.

\BloomInt:
Nice soft tweed Ned Lambert has in that suit.
Tinge of purple.
I had one like that when we lived in Lombard street west.
Dressy fellow he was once.
Used to change three suits in the day.
\gab{830}
Must get that grey suit of mine turned by Mesias.
Hello.
It's dyed.
His wife |
I forgot he's not married |
or his landlady |
ought to have picked out those threads for him.

\N1:
The coffin dived out of sight,
eased down by the men straddled on the gravetrestles.
They struggled up and out:
and all uncovered.

\BloomInt:
Twenty.

\N1:
Pause.

\BloomInt:
If we were all suddenly somebody else.

\N1:
Far away a donkey brayed.

\BloomInt:
Rain.
No such ass.
Never see a dead one, they say.
Shame of death.
They hide.
Also poor papa went away.

\N1:
Gentle sweet air blew round the bared heads in a whisper.
Whisper.
The boy by the gravehead held his wreath with both hands
\gab{840}
staring quietly in the black open space.
Mr Bloom moved behind the portly kindly caretaker.

\BloomInt:
Wellcut frockcoat.
Weighing them up |
perhaps to see which will go next.
Well, it is a long rest.
Feel no more.
It's the moment you feel.
Must be damned unpleasant.
Can't believe it at first.
Mistake must be:
someone else.
Try the house opposite.
Wait, I wanted to.
I haven't yet.
Then darkened deathchamber.
Light they want.
Whispering around you.
Would you like to see a priest?
Then rambling and wandering.
Delirium all you hid all your life.
The death struggle.
His sleep is not natural.
Press his lower eyelid.
Watching |
is his nose pointed |
is his jaw sinking |
are the soles of his feet yellow.
Pull the pillow away and finish it off on the floor
\gab{850}
since he's doomed.
Devil in that picture of sinner's death |
showing him a woman.
Dying to embrace her in his shirt.
Last act of \italian{Lucia}.
\emph{Shall I nevermore behold thee?}
Bam!
He expires.
Gone at last.
People talk about you a bit:
forget you.
Don't forget to pray for him.
Remember him in your prayers.
Even Parnell.
Ivy day dying out.
Then they follow:
dropping into a hole, one after the other.

\BloomInt:
We are praying now for the repose of his soul.
Hoping you're well and not in hell.
Nice change of air.
Out of the fryingpan of life into the fire of purgatory.

\BloomInt:
Does he ever think of the hole waiting for himself?
\gab{860}
They say you do when you shiver in the sun.
Someone walking over it.
Callboy's warning.
Near you.
Mine over there towards Finglas, the plot I bought.
Mamma, poor mamma, and little Rudy.

\N2:
The gravediggers took up their spades
and flung heavy clods of clay in on the coffin.
Mr Bloom turned away his face.

\BloomInt:
And if he was alive all the time?
Whew!
By jingo, that would be awful!
No, no:
he is dead, of course.
Of course he is dead.
Monday he died.
They ought to have some law to pierce the heart and make sure |
or an electric clock or a telephone in the coffin |
and some kind of a canvas airhole.
Flag of distress.
Three days.
Rather long to keep them in summer.
\gab{870}
Just as well to get shut of them |
as soon as you are sure there's no.

\N2:
The clay fell softer.

\BloomInt:
Begin to be forgotten.
Out of sight, out of mind.

\N2:
The caretaker moved away a few paces and put on his hat.

\BloomInt:
Had enough of it.

\N2:
The mourners took heart of grace,
one by one, covering themselves without show.
Mr Bloom put on his hat
and saw the portly figure make its way deftly through the maze of graves.
Quietly, sure of his ground,
he traversed the dismal fields.

\BloomInt:
Hynes jotting down something in his notebook.

\BloomInt:
Ah, the names.
But he knows them all.
No:
coming to me.

\hynes:
I am just taking the names,

\N1:
Hynes said below his breath.
\gab{880}

\hynes:
What is your christian name?
I'm not sure.

\Bloom:
L,
\sout{Mr Bloom said.}
Leopold.
And you might put down M'Coy's name too.
He asked me to.

\hynes:
Charley,

\N1:
Hynes said writing.

\hynes:
Charley,
I know.
He was on the \Title{Freeman} once.

\BloomInt:
So he was |
before he got the job in the morgue under Louis Byrne.
Good idea a postmortem for doctors.
Find out what they imagine they know.
He died of a Tuesday.
Got the run.
Levanted with the cash of a few ads.
Charley, you're my darling.
That was why he asked me to.
O well, does no harm.
I saw to that, M'Coy.
Thanks, old chap:
much obliged.
Leave him under an obligation:
\gab{890}
costs nothing.

\hynes:
And tell us,
\sout{Hynes said,}
do you know that fellow in the,
fellow was over there in the ...

\N1:
He looked around.

\Bloom:
Macintosh.
Yes, I saw him,
\sout{Mr Bloom said.}
Where is he now?

\hynes:
M'Intosh,

\N1:
Hynes said scribbling.

\hynes:
M'Intosh,
I don't know who he is.
Is that his name?

\N1:
He moved away, looking about him.

\Bloom:
No,

\N1:
Mr Bloom began, turning and stopping.

\Bloom:
I say, Hynes!

\BloomInt:
Didn't hear.
What?
Where has he disappeared to?
Not a sign.
Well of all the.
Has anybody here seen?
\gab{900}
Kay ee double ell.
Become invisible.
Good Lord, what became of him?

\N2:
A seventh gravedigger came beside Mr Bloom |
to take up an idle spade.

\Bloom:
O, excuse me!

\N1:
He stepped aside nimbly.

\N2:
Clay, brown, damp, began to be seen in the hole.
It rose.

\BloomInt:
Nearly over.

\N2:
A mound of damp clods rose more, rose,
and the gravediggers rested their spades.
All uncovered again for a few instants.
The boy propped his wreath against a corner:
the brother-in-law |
his on a lump.
The gravediggers put on their caps
and carried their earthy spades towards the barrow.
\gab{910}
Then knocked the blades lightly on the turf:
clean.
One bent to pluck from the haft a long tuft of grass.
One, leaving his mates, walked slowly on with shouldered weapon,
its blade blueglancing.
Silently at the gravehead |
another coiled the coffinband.

\BloomInt:
His navelcord.

\N1:
The brother-in-law, turning away,
placed something in his free hand.

\BloomInt:
Thanks in silence.

\gravedigger:
Sorry, sir:
trouble.

\BloomInt:
Headshake.

\brotherInLaw:
I know that.
For yourselves just.

\N2:
The mourners moved away slowly without aim, by devious paths,
staying at whiles to read a name on a tomb.

\hynes:
Let us go round by the chief's grave,
\sout{Hynes said.}
We have time.

\power:
Let us,
\gab{920}
\sout{Mr Power said.}

\N1:
They turned to the right, following their slow thoughts.
With awe |
Mr Power's blank voice spoke:

\power:
Some say he is not in that grave at all.
That the coffin was filled with stones.
That one day he will come again.

\N2:
Hynes shook his head.

\hynes:
Parnell will never come again,
\sout{he said.}
He's there, all that was mortal of him.
Peace to his ashes.

\N1:
Mr Bloom walked unheeded along his grove
by saddened angels, crosses, broken pillars,
family vaults,
stone hopes praying with upcast eyes,
old Ireland's hearts and hands.

\BloomInt:
More sensible
\gab{930}
to spend the money on some charity for the living.
Pray for the repose of the soul of.
Does anybody really?
Plant him and have done with him.
Like down a coalshoot.
Then lump them together to save time.
All souls' day.
Twentyseventh I'll be at his grave.
Ten shillings for the gardener.
He keeps it free of weeds.
Old man himself.
Bent down double with his shears clipping.
Near death's door.
Who passed away.
Who departed this life.
As if they did it of their own accord.
Got the shove, all of them.
Who kicked the bucket.
More interesting if they told you what they were.
So and So, wheelwright.
I travelled for cork lino.
I paid five shillings in the pound.
Or a woman's with her saucepan.
I cooked good Irish stew.
\gab{940}
Eulogy in a country churchyard it ought to be |
that poem of |
whose is it |
Wordsworth or Thomas Campbell.
Entered into rest |
the protestants put it.
Old Dr Murren's.
The great physician called him home.
Well it's God's acre for them.
Nice country residence.
Newly plastered and painted.
Ideal spot to have a quiet smoke and read the \Title{Church Times}.
Marriage ads |
they never try to beautify.
Rusty wreaths hung on knobs, garlands of bronzefoil.
Better value that |
for the money.
Still, the flowers are more poetical.
The other gets rather tiresome, never withering.
Expresses nothing.
Immortelles.

\N1:
A bird sat tamely perched on a poplar branch.

\BloomInt:
Like stuffed.
Like the wedding present |
alderman Hooper gave us.
\gab{950}
Hoo!
Not a budge out of him.
Knows there are no catapults to let fly at him.
Dead animal even sadder.
Silly-Milly burying the little dead bird in the kitchen matchbox,
a daisychain and bits of broken chainies on the grave.

\BloomInt:
The Sacred Heart that is:
showing it.
Heart on his sleeve.
Ought to be sideways |
and red it should be painted |
like a real heart.
Ireland was dedicated to it or whatever that.
Seems anything but pleased.
Why this infliction?
Would birds come then and peck |
like the boy with the basket of fruit |
but he said no |
because they ought to have been afraid of the boy.
Apollo that was.

\BloomInt:
How many!
All these here |
once walked round Dublin.
\gab{960}
Faithful departed.
As you are now |
so once were we.

\BloomInt:
Besides how could you remember everybody?
Eyes, walk, voice.
Well, the voice, yes:
gramophone.
Have a gramophone in every grave or keep it in the house.
After dinner on a Sunday.
Put on poor old greatgrandfather.
Kraahraark!
Hellohellohello amawfullyglad kraark awfullygladaseeagain hellohello amawf krpthsth.
Remind you of the voice
like the photograph reminds you of the face.
Otherwise you couldn't remember the face after fifteen years, say.
For instance who?
For instance some fellow that died
when I was in Wisdom Hely's.

\N1:
Rtststr!
A rattle of pebbles.
\gab{970}

\BloomInt:
Wait.
Stop!

\N1:
He looked down intently into a stone crypt.

\BloomInt:
Some animal.
Wait.
There he goes.

\N1:
An obese grey rat toddled along the side of the crypt, moving the pebbles.

\BloomInt:
An old stager:
greatgrandfather:
he knows the ropes.

\N1:
The grey alive crushed itself in under the plinth,
wriggled itself in under it.

\BloomInt:
Good hidingplace for treasure.

\BloomInt:
Who lives there?
Are laid the remains of Robert Emery.
Robert Emmet was buried here by torchlight, wasn't he?
Making his rounds.

\BloomInt:
Tail gone now.

\BloomInt:
One of those chaps would make short work of a fellow.
\gab{980}
Pick the bones clean |
no matter who it was.
Ordinary meat for them.
A corpse is meat gone bad.
Well and what's cheese?
Corpse of milk.
I read in that \Title{Voyages in China} |
that the Chinese say a white man smells like a corpse.
Cremation better.
Priests dead against it.
Devilling for the other firm.
Wholesale burners and Dutch oven dealers.
Time of the plague.
Quicklime feverpits to eat them.
Lethal chamber.
Ashes to ashes.
Or bury at sea.
Where is that Parsee tower of silence?
Eaten by birds.
Earth, fire, water.
Drowning they say |
is the pleasantest.
See your whole life in a flash.
But being brought back to life |
no.
Can't bury in the air however.
Out of a flying machine.
\gab{990}
Wonder does the news go about whenever a fresh one is let down.
Underground communication.
We learned that from them.
Wouldn't be surprised.
Regular square feed for them.
Flies come before he's well dead.
Got wind of Dignam.
They wouldn't care about the smell of it.
Saltwhite crumbling mush of corpse:
smell, taste like raw white turnips.

\N1:
The gates glimmered in front:
still open.

\BloomInt:
Back to the world again.
Enough of this place.
Brings you a bit nearer every time.
Last time I was here |
was Mrs Sinico's funeral.
Poor papa too.
The love that kills.
And even scraping up the earth at night with a lantern |
like that case I read of |
to get at fresh buried females |
or even putrefied with running gravesores.
Give you the creeps after a bit.
\gab{1000}
I will appear to you after death.
You will see my ghost after death.
My ghost will haunt you after death.
There is another world after death |
named hell.
I do not like that other world |
she wrote.
No more do I.
Plenty to see and hear and feel yet.
Feel live warm beings near you.
Let them sleep in their maggoty beds.
They are not going to get me this innings.
Warm beds:
warm fullblooded life.

\N2:
Martin Cunningham emerged from a sidepath, talking gravely.

\BloomInt:
Solicitor, I think.
I know his face.
Menton, John Henry, solicitor,
commissioner for oaths and affidavits.
Dignam used to be in his office.
Mat Dillon's long ago.
Jolly Mat.
Convivial evenings.
Cold fowl, cigars, the Tantalus glasses.
Heart of gold really.
\gab{1010}
Yes, Menton.
Got his rag out that evening on the bowling-green
because I sailed inside him.
Pure fluke of mine:
the bias.
Why he took such a rooted dislike to me.
Hate at first sight.
Molly and Floey Dillon linked under the lilactree, laughing.
Fellow always like that, mortified if women are by.

\BloomInt:
Got a dinge in the side of his hat.
Carriage probably.

\Bloom:
Excuse me, sir,

\N1:
Mr Bloom said beside them.

\N2:
They stopped.

\Bloom:
Your hat is a little crushed,

\N1:
Mr Bloom said pointing.

\N2:
John Henry Menton stared at him for an instant without moving.

\cunningham:
There,

\N2:
Martin Cunningham helped, pointing also.
\gab{1020}
John Henry Menton took off his hat,
bulged out the dinge
and smoothed the nap with care on his coatsleeve.
He clapped the hat on his head again.

\cunningham:
It's all right now,
\sout{Martin Cunningham said.}

\N2:
John Henry Menton jerked his head down in acknowledgment.

\menton:
Thank you,
\sout{he said shortly.}

\N1:
They walked on towards the gates.
Mr Bloom, chapfallen, drew behind a few paces so as not to overhear.

\BloomInt:
Martin laying down the law.
Martin could wind a sappyhead like that round his little finger,
without his seeing it.%
\gab{1030}

\BloomInt:
Oyster eyes.
Never mind.
Be sorry after perhaps |
when it dawns on him.
Get the pull over him that way.

\BloomInt:
Thank you.
How grand we are this morning!

