\N1:
\gab{243}
Two letters and a card lay on the hallfloor.
He stooped and gathered them.

\BloomInt:
Mrs Marion Bloom.

\N1:
His quickened heart slowed at once.

\BloomInt:
Bold hand.
Mrs Marion.

\Molly:
Poldy!

\N1:
Entering the bedroom
he halfclosed his eyes
and walked through warm yellow twilight
towards her tousled head.

\Molly:
Who are the letters for?

\N1:
He looked at them.
\gab{250}

\BloomInt:
Mullingar.
Milly.

\Bloom:
A letter for me from Milly,
\sout{he said carefully,}
and a card to you.
And a letter for you.

\N1:
He laid her card and letter
on the twill bedspread |
near the curve of her knees.

\Bloom:
Do you want the blind up?

\N1:
Letting the blind up by gentle tugs halfway |
his backward eye saw her glance at the letter
and tuck it under her pillow.

\Bloom:
That do?

\N1:
he asked,
turning.

\N2:
She was reading the card,
propped on her elbow.

\Molly:
She got the things,
\gab{260}
\sout{she said}.

\N2:
He waited till she had laid the card aside
and curled herself back slowly with a snug sigh.

\Molly:
Hurry up with that tea,
\sout{she said}.
I'm parched.

\Bloom:
The kettle is boiling,
\sout{he said}.

\N2:
But he delayed to clear the chair:
her striped petticoat, tossed soiled linen:
and lifted all in an armful on to the foot of the bed.

\N2:
As he went down the kitchen stairs |
she called:

\Molly:
Poldy!

\Bloom:
What?

\Molly:
Scald the teapot.
\gab{270}

\N2:
On the boil sure enough:
a plume of steam from the spout.
He scalded and rinsed out the teapot and put in four full spoons of tea,
tilting the kettle then to let the water flow in.
Having set it to draw |
he took off the kettle,
crushed the pan flat on the live coals
and watched the lump of butter slide and melt.
While he unwrapped the kidney
the cat mewed hungrily against him.

\BloomInt:
Give her too much meat |
she won't mouse.
Say they won't eat pork.
Kosher.
Here.

\N2:
He let the bloodsmeared paper fall to her
and dropped the kidney amid the sizzling butter sauce.

\BloomInt:
Pepper.

\N2:
He sprinkled it through his fingers ringwise from the chipped eggcup.

\N2:
Then he slit open his letter,
\gab{280}
glancing down the page and over.

\Milly:
Thanks:
new tam:
Mr Coghlan:
lough Owel picnic:
young student:
Blazes Boylan's seaside girls.

\N2:
The tea was drawn.
He filled his own moustachecup, sham crown

\BloomInt:
Derby, smiling.
Silly Milly's birthday gift.
Only five she was then.
No, wait: four.
I gave her the amberoid necklace she broke.
Putting pieces of folded brown paper in the letterbox for her.

\N2:
He smiled, pouring.

\BloomInt:
\begin{quote}
    O, Milly Bloom, you are my darling. \\
    You are my lookingglass from night to morning. \\
    I'd rather have you without a farthing \\
    Than Katey Keogh with her ass and garden.
\gab{290}
\end{quote}

\BloomInt:
Poor old professor Goodwin.
Dreadful old case.
Still he was a courteous old chap.
Oldfashioned way he used to bow Molly off the platform.
And the little mirror in his silk hat.
The night Milly brought it into the parlour.

\Milly:
O, look what I found in professor Goodwin's hat!

\BloomInt:
All we laughed.
Sex breaking out even then.
Pert little piece she was.

\N2:
He prodded a fork into the kidney and slapped it over:
then fitted the teapot on the tray.
Its hump bumped as he took it up.

\BloomInt:
Everything on it?
Bread and butter,
four, sugar, spoon, her cream.
Yes.

\N2:
He carried it upstairs,
his thumb hooked in the teapot handle.

\N1:
Nudging the door open with his knee
\gab{300}
he carried the tray in
and set it on the chair by the bedhead.

\Molly:
What a time you were!
\sout{she said}.

\N1:
She set the brasses jingling as she raised herself briskly,
an elbow on the pillow.
He looked calmly down on her bulk and between her large soft bubs,
sloping within her nightdress like a shegoat's udder.
The warmth of her couched body rose on the air,
mingling with the fragrance of the tea she poured.

\N1:
A strip of torn envelope peeped from under the dimpled pillow.
In the act of going |
he stayed to straighten the bedspread.

\Bloom:
Who was the letter from?
\gab{310}
\sout{he asked}.

\BloomInt:
Bold hand. Marion.

\Molly:
O, Boylan,
\sout{she said}.
He's bringing the programme.

\Bloom:
What are you singing?

\Molly:
\Title{La ci darem} with J.~C.~Doyle,
\sout{she said},
and \Title{Love's Old Sweet Song}.

\N1:
Her full lips, drinking, smiled.

\BloomInt:
Rather stale smell that incense leaves next day.
Like foul flowerwater.

\Bloom:
Would you like the window open a little?

\N1:
She doubled a slice of bread into her mouth, asking:

\Molly:
What time is the funeral?

\Bloom:
Eleven, I think,
\gab{320}
\sout{he answered}.
I didn't see the paper.

\N1:
Following the pointing of her finger |
he took up a leg of her soiled drawers from the bed.

\BloomInt:
No?

\N1:
Then, a twisted grey garter looped round a stocking:
rumpled, shiny sole.

\Molly:
No: that book.

\BloomInt:
Other stocking.
Her petticoat.

\Molly:
It must have fell down,
\sout{she said}.

\N1:
He felt here and there.

\BloomInt:
\italian{Voglio e non vorrei}.
Wonder if she pronounces that right:
\italian{Voglio}.
Not in the bed.
Must have slid down.

\N1:
He stooped and lifted the valance.
The book,
fallen,
sprawled against the bulge of the
orangekeyed chamberpot.
\gab{330}

\Molly:
Show here,
\sout{she said}.
I put a mark in it.
There's a word I wanted to ask you.

\N2:
She swallowed a draught of tea from her cup held by nothandle
and, having wiped her fingertips smartly on the blanket,
began to search the text with the hairpin
till she reached the word.

\Bloom:
Met him what?
\sout{he asked}.

\Molly:
Here,
\sout{she said}.
What does that mean?

\N2:
He leaned downward and read |
near her polished thumbnail.

\Bloom:
Metempsychosis?

\Molly:
Yes.
\gab{340}
Who's he when he's at home?

\Bloom:
Metempsychosis,
\sout{he said, frowning}.
It's Greek: from the Greek.
That means the transmigration of souls.

\Molly:
O, rocks!
\sout{she said}.
Tell us in plain words.

\N2:
He smiled, glancing askance at her mocking eyes.

\BloomInt:
The same young eyes.
The first night after the charades.
Dolphin's Barn.

\N2:
He turned over the smudged pages.

\Voice:
\Title{Ruby: The Pride of the Ring.}

\BloomInt:
Hello.
Illustration.
Fierce Italian with carriagewhip.
Must be Ruby pride of the |
on the floor naked.
Sheet kindly lent.

\Voice:
\italian{The monster Maffei desisted and flung his victim from him with an oath.}

\BloomInt:
Cruelty behind it all.
Doped animals.
Trapeze at Hengler's.
Had to look the other way.
Mob gaping.
\gab{350}
Break your neck and we'll break our sides.
Families of them.
Bone them young so they metamspychosis.
That we live after death.
Our souls.
That a man's soul after he dies.
Dignam's soul …

\Bloom:
Did you finish it?
\sout{he asked}.

\Molly:
Yes,
\sout{she said}.
There's nothing smutty in it.
Is she in love with the first fellow all the time?

\Bloom:
Never read it.
Do you want another?

\Molly:
Yes.
Get another of Paul de Kock's.
Nice name he has.

\N2:
She poured more tea into her cup,
watching it flow sideways.

\BloomInt:
Must get that Capel street library book renewed
\gab{360}
or they'll write to Kearney, my guarantor.
Reincarnation:
that's the word.

\Bloom:
Some people believe,
\sout{he said}
that we go on living in another body after death,
that we lived before.
They call it reincarnation.
That we all lived before on the earth thousands of years ago
or some other planet.
They say we have forgotten it.
Some say they remember their past lives.

\N2:
The sluggish cream wound curdling spirals through her tea.

\BloomInt:
Better remind her of the word:
metempsychosis.
An example would be better.
An example?

\BloomInt:
The \Title{Bath of the Nymph} over the bed.
Given away with the Easter number of \Title{Photo Bits}:
splendid masterpiece in art colours.
\gab{370}
Tea before you put milk in.
Not unlike her with her hair down:
slimmer.
Three and six I gave for the frame.
She said it would look nice over the bed.
Naked nymphs:
Greece:
and for instance all the people that lived then.

\N2:
He turned the pages back.

\Bloom:
Metempsychosis
\sout{he said}
is what the ancient Greeks called it.
They used to believe
you could be changed into an animal or a tree,
for instance.
What they called nymphs, for example.

\N2:
Her spoon ceased to stir up the sugar.
She gazed straight before her,
inhaling through her arched nostrils.

\Molly:
There's a smell of burn,
\gab{380}
\sout{she said}.
Did you leave anything on the fire?

\Bloom:
The kidney!
\sout{he cried suddenly}.

\N1:
He fitted the book roughly into his inner pocket |
and, stubbing his toes against the broken commode,
hurried out towards the smell,
stepping hastily down the stairs with a flurried stork's legs.
Pungent smoke shot up in an angry jet from a side of the pan.
By prodding a prong of the fork under the kidney |
he detached it and turned it turtle on its back.

\BloomInt:
Only a little burnt.

\N1:
He tossed it off the pan on to a plate
and let the scanty brown gravy trickle over it.

\BloomInt:
Cup of tea now.

\N1:
He sat down, cut and buttered a slice of the loaf.
He shore away the burnt flesh and flung it to the cat.
\gab{390}
Then he put a forkful into his mouth,
chewing with discernment |
the toothsome pliant meat.

\BloomInt:
Done to a turn.

\N1:
A mouthful of tea.
Then he cut away dies of bread,
sopped one in the gravy and put it in his mouth.

\BloomInt:
What was that about some young student and a picnic?

\N1:
He creased out the letter at his side,
reading it slowly as he chewed,
sopping another die of bread in the gravy and
raising it to his mouth.

\Milly:
Dearest Papli.
Thanks ever so much for the lovely birthday present.
It suits me splendid.
Everyone says I am quite the belle in my new tam.
I got mummy's Iovely box of creams and am writing.
\gab{400}
They are lovely.
I am getting on swimming in the photo business now.
Mr Coghlan took one of me and Mrs.
Will send when developed.
We did great biz yesterday.
Fair day and all the beef to the heels were in.
We are going to lough Owel on Monday with a few friends to make a scrap picnic.
Give my love to mummy and to yourself a big kiss and thanks.
I hear them at the piano downstairs.
There is to be a concert in the Greville Arms on Saturday.
There is a young student comes here some evenings named Bannon |
his cousins or something are big swells
and he sings Boylan's
(I was on the pop of writing Blazes Boylan's)
song about those seaside girls.
Tell him silly Milly sends my best respects.
I must now close with fondest love.
\gab{410}
Your fond daughter,
\emph{Milly}.

\Milly:
P. S.
Excuse bad writing |
am in hurry.
Byby.
M.

\BloomInt:
Fifteen yesterday.
Curious, fifteenth of the month too.
Her first birthday away from home.
Separation.
Remember the summer morning she was born,
running to knock up Mrs Thornton in Denzille street.
Jolly old woman.
Lot of babies she must have helped into the world.
She knew from the first |
poor little Rudy wouldn't live.
Well, God is good, sir.
She knew at once.
\gab{420}
He would be eleven now if he had lived.

\N1:
His vacant face stared pityingly at the postscript.

\BloomInt:
Excuse bad writing.
Hurry.
Piano downstairs.
Coming out of her shell.
Row with her in the XL
Cafe about the bracelet.
Wouldn't eat her cakes or speak or look.
Saucebox.

\N1:
He sopped other dies of bread in the gravy
and ate piece after piece of kidney.

\BloomInt:
Twelve and six a week.
Not much.
Still, she might do worse.
Music hall stage.
Young student.

\N1:
He drank a draught of cooler tea to wash down his meal.
Then he read the letter again:
twice.

\BloomInt:
O, well: she knows how to mind herself.
But if not?
No, nothing has happened.
Of course it might.
Wait in any case till it does.
A wild piece of goods.
\gab{430}
Her slim legs running up the staircase.
Destiny.
Ripening now.

\BloomInt:
Vain: very.

\N1:
He smiled with troubled affection at the kitchen window.

\BloomInt:
Day I caught her in the street |
pinching her cheeks to make them red.
Anemic a little.
Was given milk too long.
On the \Title{Erin's King} that day round the Kish.
Damned old tub pitching about.
Not a bit funky.
Her pale blue scarf loose in the wind with her hair.

\BloomInt:
\begin{verse}
    All dimpled cheeks and curls, \\
    Your head it simply swirls.
\end{verse}

\BloomInt:
Seaside girls.
Torn envelope.
Hands stuck in his trousers' pockets,
jarvey off for the day, singing.
Friend of the family.
\gab{440}
\emph{Swurls}, he says.
Pier with lamps, summer evening, band,

\BloomInt:
\begin{verse}
    Those girls, those girls, \\
    Those lovely seaside girls.
\end{verse}

\BloomInt:
Milly too.
Young kisses:
the first.
Far away now past.
Mrs Marion.
Reading,
lying back now, counting the strands of her hair,
smiling, braiding.

\N1:
A soft qualm, regret,
flowed down his backbone,
increasing.

\BloomInt:
Will happen, yes.
Prevent.
Useless: can't move.
Girl's sweet light lips.
Will happen too.

\N1:
He felt the flowing qualm spread over him.

\BloomInt:
Useless to move now.
Lips kissed, kissing, kissed.
\gab{450}
Full gluey woman's lips.

\BloomInt:
Better where she is down there: away.
Occupy her.
Wanted a dog to pass the time.
Might take a trip down there.
August bank holiday, only two and six return.
Six weeks off, however.
Might work a press pass.
Or through M'Coy.

\N2:
The cat,
having cleaned all her fur,
returned to the meatstained paper,
nosed at it
and stalked to the door.
She looked back at him, mewing.

\BloomInt:
Wants to go out.
Wait before a door |
sometime it will open.
Let her wait.
Has the fidgets.
Electric.
Thunder in the air.
Was washing at her ear
with her back to the fire too.

\N2:
He felt heavy, full:
then a gentle loosening of his bowels.
\gab{460}
He stood up,
undoing the waistband of his trousers.
The cat mewed to him.

\Bloom:
Miaow!
\sout{he said in answer}.
Wait till I'm ready.

\BloomInt:
Heaviness: hot day coming.
Too much trouble to fag up the stairs to the landing.

\N2:
A paper.
He liked to read at stool.

\BloomInt:
Hope no ape comes knocking just as I'm.

\N2:
In the tabledrawer he found an old number of \Title{Titbits}.
He folded it under his armpit,
went to the door and opened it.
The cat went up in soft bounds.

\BloomInt:
Ah, wanted to go upstairs,
curl up in a ball on the bed.

\N2:
Listening, he heard her voice:
\gab{470}

\Molly:
Come, come, pussy.
Come.

\N2:
He went out through the backdoor into the garden:
stood to listen towards the next garden.
No sound.

\BloomInt:
Perhaps hanging clothes out to dry.

\N2:
The maid was in the garden.

\Bloom:
Fine morning.

\N2:
He bent down to regard a lean file of spearmint growing by the wall.

\BloomInt:
Make a summerhouse here.
Scarlet runners.
Virginia creepers.
Want to manure the whole place over,
scabby soil.
A coat of liver of sulphur.
All soil like that without dung.
Household slops.
Loam, what is this that is?
The hens in the next garden:
their droppings are very good top dressing.
Best of all though are the cattle,
\gab{480}
especially when they are fed on those oilcakes.
Mulch of dung.
Best thing to clean ladies' kid gloves.
Dirty cleans.
Ashes too.
Reclaim the whole place.
Grow peas in that corner there.
Lettuce.
Always have fresh greens then.
Still gardens have their drawbacks.
That bee or bluebottle here Whitmonday.

\N2:
He walked on.

\BloomInt:
Where is my hat, by the way?
Must have put it back on the peg.
Or hanging up on the floor.
Funny I don't remember that.
Hallstand too full.
Four umbrellas, her raincloak.
Picking up the letters.
Drago's shopbell ringing.
Queer I was just thinking that moment.
Brown brillantined hair over his collar.
Just had a wash and brushup.
Wonder have I time for a bath this morning.
\gab{490}
Tara street.
Chap in the paybox there |
got away James Stephens,
they say.
O'Brien.

\BloomInt:
Deep voice that fellow Dlugacz has.
Agendath |
what is it?
\emph{Now, my miss}.
Enthusiast.

\N2:
He kicked open the crazy door of the jakes.

\BloomInt:
Better be careful not to get these trousers dirty for the funeral.

\N2:
He went in, bowing his head under the low lintel.
Leaving the door ajar,
amid the stench of mouldy limewash and stale cobwebs
he undid his braces.
Before sitting down
he peered through a chink up at the nextdoor windows.
The king was in his countinghouse.
Nobody.

\N2:
Asquat on the cuckstool |
\gab{500}
he folded out his paper,
turning its pages over on his bared knees.

\BloomInt:
Something new and easy.
No great hurry.
Keep it a bit.

\Voice:
Our prize titbit:
\Title{Matcham's Masterstroke}.
Written by Mr~Philip Beaufoy,
Playgoers' Club, London.
Payment at the rate of one guinea a column has been made to the writer.

\BloomInt:
Three and a half.
Three pounds three.
Three pounds, thirteen and six.

\N2:
Quietly he read, restraining himself, the first column
and, yielding but resisting, began the second.
Midway, his last resistance yielding,
he allowed his bowels to ease themselves quietly as he read,
reading still patiently

\BloomInt:
that slight constipation of yesterday quite gone.
Hope it's not too big |
bring on piles again.
\gab{510}
No, just right.
So.
Ah!
Costive.
One tabloid of cascara sagrada.

\N2:
Life might be so.
It did not move or touch him
but it was something quick and neat.

\BloomInt:
Print anything now.
Silly season.

\N2:
He read on, seated calm above his own rising smell.

\BloomInt:
Neat certainly.

\Voice:
\emph{Matcham often thinks of the masterstroke
by which he won the laughing witch who now.}

\BloomInt:
Begins and ends morally.

\Voice:
\emph{Hand in hand.}

\BloomInt:
Smart.

\N2:
He glanced back through what he had read
and, while feeling his water flow quietly,
he envied kindly Mr Beaufoy who had written it
and received payment of three pounds, thirteen and six.

\BloomInt:
Might manage a sketch.
By Mr~and~Mrs L.~M.~Bloom.
Invent a story for some proverb.
Which?
Time I used to try jotting down on my cuff what she said dressing.
\gab{520}
Dislike dressing together.
Nicked myself shaving.
Biting her nether lip,
hooking the placket of her skirt.

\BloomInt:
Timing her.
9.15.

\Molly:
Did Roberts pay you yet?

\BloomInt:
9.20.

\Molly:
What had Gretta Conroy on?

\BloomInt:
9.23.

\Molly:
What possessed me to buy this comb?

\BloomInt:
9.24.

\Molly:
I'm swelled after that cabbage.

\BloomInt:
A speck of dust on the patent leather of her boot.

\N2:
Rubbing smartly in turn each welt against her stockinged calf.

\BloomInt:
Morning after the bazaar dance
when May's band played Ponchielli's dance of the hours.
Explain that:
morning hours, noon,
then evening coming on,
then night hours.
Washing her teeth.
That was the first night.
Her head dancing.
Her fansticks clicking.

\Molly:
Is that Boylan well off?

\Bloom:
He has money.
Why?

\Molly:
I noticed he had a good rich smell off his breath dancing.

\BloomInt:
No use humming then.
\gab{530}
Allude to it.
Strange kind of music that |
last night.
The mirror was in shadow.
She rubbed her handglass briskly on her woollen vest
against her full wagging bub.
Peering into it.
Lines in her eyes.
It wouldn't pan out somehow.

\BloomInt:
Evening hours,
girls in grey gauze.
Night hours then:
black with daggers and eyemasks.
Poetical idea:
pink, then golden, then grey, then black.
Still, true to life also.
Day: then the night.

\N2:
He tore away half the prize story sharply and wiped himself with it.
Then he girded up his trousers, braced and buttoned himself.
He pulled back the jerky shaky door of the jakes
and came forth from the gloom into the air.
\gab{540}

\N2:
In the bright light,
lightened and cooled in limb,
he eyed carefully his black trousers:
the ends, the knees, the houghs of the knees.

\BloomInt:
What time is the funeral?
Better find out in the paper.

\N2:
A creak and a dark whirr in the air high up.
The bells of George's church.
They tolled the hour:
loud dark iron.

\N1:
\begin{verse}
    Heigho! Heigho! \\
    Heigho! Heigho! \\
    Heigho! Heigho!
\end{verse}

\BloomInt:
Quarter to.

\N2:
There again:
the overtone following through the air,
a third.
\gab{550}

\BloomInt:
Poor Dignam!
