\N1:
\gab{72}
On the doorstep he felt in his hip pocket for the latchkey.

\BloomInt:
Not there.
In the trousers I left off.
Must get it.
Potato I have.
Creaky wardrobe.
No use disturbing her.
She turned over sleepily that time.

\N1:
He pulled the halldoor to |
after him |
very quietly, more,
till the footleaf dropped gently over the threshold,
a limp lid.
Looked shut.

\BloomInt:
All right till I come back anyhow.

\N1:
He crossed to the bright side,
avoiding the loose cellarflap of number seventyfive.
The sun was nearing the steeple of George's church.

\BloomInt:
Be a warm day I fancy.
Specially in these black clothes |
feel it more.
Black conducts, reflects, (refracts is it?), the heat.
\gab{80}
But I couldn't go in that light suit.
Make a picnic of it.

\N1:
His eyelids sank quietly often as he walked in happy warmth.
Boland's breadvan delivering with trays

\BloomInt:
our daily but she prefers yesterday's loaves |
turnovers crisp crowns hot.
Makes you feel young.
Somewhere in the east:
early morning:
set off at dawn.
Travel round in front of the sun,
steal a day's march on him.
Keep it up for ever never grow a day older technically.
Walk along a strand,
strange land,
come to a city gate,
sentry there,
old ranker too, old
Tweedy's big moustaches,
leaning on a long kind of a spear.
Wander through awned streets.
Turbaned faces going by.
Dark caves of carpet shops,
big man, Turko the terrible,
seated crosslegged, smoking a coiled pipe.
\gab{90}
Cries of sellers in the streets.
Drink water scented with fennel, sherbet.
Dander along all day.
Might meet a robber or two.
Well, meet him.
Getting on to sundown.
The shadows of the mosques among the pillars:
priest with a scroll rolled up.
A shiver of the trees, signal, the evening wind.
I pass on.
Fading gold sky.
A mother watches me from her doorway.
She calls her children home in their dark language.
High wall:
beyond strings twanged.
Night sky, moon, violet,
colour of Molly's new garters.
Strings.
Listen.
A girl playing one of those instruments |
what do you call them:
dulcimers.
I pass.

\BloomInt:
Probably not a bit like it really.
Kind of stuff you read:
in the track of the sun.
Sunburst on the titlepage.
\gab{100}

\N1:
He smiled, pleasing himself.

\BloomInt:
What Arthur Griffith said about the headpiece over the \Title{Freeman} leader:
a homerule sun rising up in the northwest
from the laneway behind the Bank of Ireland.

\N1:
He prolonged his pleased smile.

\BloomInt:
Ikey touch that:
homerule sun
rising up in the north-west.

\N2:
He approached Larry O'Rourke's.
From the cellar grating
floated up the flabby gush of porter.
Through the open doorway
the bar squirted out whiffs of ginger, teadust, biscuitmush.

\BloomInt:
Good house, however:
just the end of the city traffic.
For instance M'Auley's down there:
n. g. as position.
Of course
if they ran a tramline along the North Circular from the cattlemarket to the quays |
\gab{110}
value would go up like a shot.

\BloomInt:
Baldhead over the blind.
Cute old codger.
No use canvassing him for an ad.
Still he knows his own business best.
There he is, sure enough, my bold Larry,
leaning against the sugarbin in his shirtsleeves
watching the aproned curate swab up with mop and bucket.
Simon Dedalus takes him off to a tee |
with his eyes screwed up.

\ORourke:
Do you know what I'm going to tell you?

\Simon:
What's that, Mr O'Rourke?

\ORourke:
Do you know what?
The Russians, they'd only be an eight o'clock breakfast for the Japanese.

\BloomInt:
Stop and say a word:
about the funeral perhaps.
Sad thing about poor Dignam, Mr O'Rourke.

\N2:
Turning into Dorset street
\gab{120}
he said freshly in greeting through the doorway:

\Bloom:
Good day, Mr O'Rourke.

\ORourke:
Good day to you.

\Bloom:
Lovely weather, sir.

\ORourke:
'Tis all that.

\BloomInt:
Where do they get the money?
Coming up redheaded curates from the county Leitrim,
rinsing empties and old man in the cellar.
Then, lo and behold,
they blossom out as Adam Findlaters or Dan Tallons.
Then thin of the competition.
General thirst.
\emph{Good puzzle would be cross Dublin without passing a pub.}
\gab{130}
Save it they can't.
Off the drunks perhaps.
Put down three and carry five.
What is that, a bob here and there, dribs and drabs.
On the wholesale orders perhaps.
Doing a double shuffle with the town travellers.
Square it you with the boss and we'll split the job, see?

\BloomInt:
How much would that tot to off the porter in the month?
Say ten barrels of stuff.
Say he got ten per cent off.
O more.
Fifteen.

\N2:
He passed Saint Joseph's National school.
Brats' clamour.

\BloomInt:
Windows open.
Fresh air helps memory.

\N2:
Or a lilt.
Ahbeesee defeegee kelomen opeecue rustyouvee doubleyou.

\BloomInt:
Boys are they?
Yes.
Inishturk.
Inishark.
Inishboffin.
At their joggerfry.
Mine.
Slieve Bloom.

\N1:
He halted before Dlugacz's window,
\gab{140}
staring at the hanks of sausages,
polonies, black and white.

\BloomInt:
Fifteen multiplied by.

\N1:
The figures whitened in his mind, unsolved:
displeased, he let them fade.
The shiny links,
packed with forcemeat,
fed his gaze
and he breathed in tranquilly
the lukewarm breath of cooked spicy pigs' blood.

\N1:
A kidney oozed bloodgouts on the willowpatterned dish:
the last.
He stood by the nextdoor girl at the counter.
Would she buy it too,
calling the items from a slip in her hand?

\BloomInt:
Chapped: washingsoda.

\NextDoorGirl:
And a pound and a half of Denny's sausages.

\N1:
His eyes rested on her vigorous hips.

\BloomInt:
Woods his name is.
Wonder what he does.
Wife is oldish.
New blood.
No followers allowed.
\gab{150}
Strong pair of arms.
Whacking a carpet on the clothesline.
She does whack it, by George.
The way her crooked skirt swings at each whack.

\N1:
The ferreteyed porkbutcher folded the sausages
he had snipped off
with blotchy fingers, sausagepink.

\BloomInt:
Sound meat there:
like a stallfed heifer.

\N1:
He took a page up from the pile of cut sheets:

\BloomInt:
the model farm at Kinnereth on the lakeshore of Tiberias.
Can become ideal winter sanatorium.
Moses Montefiore.
I thought he was.
Farmhouse, wall round it,
blurred cattle cropping.

\N1:
He held the page from him:

\BloomInt:
interesting:
read it
nearer,

\N1:
the title,
the blurred cropping cattle,
the page rustling.

\BloomInt:
A young white heifer.
Those mornings in the cattlemarket,
the beasts lowing in their pens,
branded sheep,
\gab{160}
flop and fall of dung,
the breeders in
hobnailed boots trudging through the litter,
slapping a palm on a ripemeated hindquarter,
there's a prime one,
unpeeled switches in their hands.

\N1:
He held the page aslant patiently,
bending his senses and his will,
his soft subject gaze at rest.

\BloomInt:
The crooked skirt swinging,
whack by whack by whack.

\N2:
The porkbutcher snapped two sheets from the pile,
wrapped up her prime sausages
and made a red grimace.

\Dlugacz:
Now, my miss,
\sout{he said}.

\N2:
She tendered a coin, smiling boldly, holding her thick wrist out.

\Dlugacz:
Thank you, my miss.
And one shilling threepence change.
For you, please?
\gab{170}

\N2:
Mr Bloom pointed quickly.

\BloomInt:
To catch up and walk behind her if she went slowly,
behind her moving hams.
Pleasant to see first thing in the morning.
Hurry up, damn it.
Make hay while the sun shines.

\N2:
She stood outside the shop in sunlight and sauntered lazily to the right.
He sighed down his nose:

\BloomInt:
they never understand.
Sodachapped hands.
Crusted toenails too.
Brown scapulars in tatters, defending her both ways.

\N2:
The sting of disregard glowed to weak pleasure within his breast.

\BloomInt:
For another:
a constable off duty cuddling her in Eccles lane.
They like them sizeable.
Prime sausage.
O please, Mr Policeman, I'm lost in the wood.

\Dlugacz:
Threepence, please.
\gab{180}

\N2:
His hand accepted the moist tender gland and slid it into a sidepocket.
Then it fetched up three coins from his trousers' pocket
and laid them on the rubber prickles.
They lay, were read quickly and quickly slid,
disc by disc, into the till.

\Dlugacz:
Thank you, sir.
Another time.

\N1:
A speck of eager fire from foxeyes thanked him.
He withdrew his gaze after an instant.

\BloomInt:
No: better not: another time.

\Bloom:
Good morning,

\N1:
he said, moving away.

\Dlugacz:
Good morning, sir.

\BloomInt:
No sign.
Gone.
What matter?
\gab{190}

\N1:
He walked back along Dorset street, reading gravely.

\BloomInt:
Agendath Netaim: planters' company.
To purchase waste sandy tracts from Turkish government
and plant with eucalyptus trees.
Excellent for shade, fuel and construction.
Orangegroves and immense melonfields north of Jaffa.
You pay eighty marks and they plant a dunam of land for you
with olives, oranges, almonds or citrons.
Olives cheaper: oranges need artificial irrigation.
Every year you get a sending of the crop.
Your name entered for life as owner in the book of the union.
Can pay ten down and the balance in yearly instalments.
Bleibtreustrasse~34, Berlin, W.~15.

\BloomInt:
Nothing doing.
Still an idea behind it.
\gab{200}

\N1:
He looked at the cattle, blurred in silver heat.

\BloomInt:
Silverpowdered olivetrees.
Quiet long days:
pruning, ripening.
Olives are packed in jars, eh?
I have a few left from Andrews.
Molly spitting them out.
Knows the taste of them now.
Oranges in tissue paper packed in crates.
Citrons too.
Wonder is poor Citron still in Saint Kevin's parade.
And Mastiansky with the old cither.
Pleasant evenings we had then.
Molly in Citron's basketchair.
Nice to hold, cool waxen fruit, hold in the hand,
lift it to the nostrils and smell the perfume.
Like that, |
heavy, sweet, wild perfume.
Always the same, year after year.
They fetched high prices too, Moisel told me.
Arbutus place: Pleasants street: pleasant old times.
\gab{210}
Must be without a flaw, he said.
Coming all that way:
Spain, Gibraltar, Mediterranean, the Levant.
Crates lined up on the quayside at Jaffa,
chap ticking them off in a book,
navvies handling them barefoot in soiled dungarees.
There's whatdoyoucallhim out of.

\Bloom:
How do you?

\BloomInt:
Doesn't see.
Chap you know just to salute |
bit of a bore.
His back is like that Norwegian captain's.
Wonder if I'll meet him today.
Watering cart.
To provoke the rain.
On earth as it is in heaven.

\N1:
A cloud began to cover the sun slowly, wholly.
Grey.
Far.

\BloomInt:
No, not like that.
A barren land, bare waste.
Vulcanic lake, the dead sea:
no fish, weedless, sunk deep in the earth.
\gab{220}
No wind could lift those waves,
grey metal, poisonous foggy waters.
Brimstone they called it |
raining down:
the cities of the plain:
Sodom, Gomorrah, Edom.
All dead names.

\N2:
A dead sea in a dead land, grey and old.
Old now.
It bore the oldest, the first race.

\N1:
A bent hag crossed from Cassidy's,
clutching a naggin bottle by the neck.

\BloomInt:
The oldest people.
Wandered far away over all the earth,
captivity to captivity,
multiplying, dying,
being born everywhere.

\N2:
It lay there now.
Now it could bear no more.
Dead:
an old woman's:
the grey sunken cunt of the world.

\N1:
Desolation.

\N2:
Grey horror seared his flesh.
\gab{230}
Folding the page into his pocket
he turned into Eccles street,
hurrying homeward.
Cold oils slid along his veins,
chilling his blood:
age crusting him with a salt cloak.

\BloomInt:
Well, I am here now.
Yes, I am here now.
Morning mouth |
bad images.
Got up wrong side of the bed.
Must begin again those Sandow's exercises.
On the hands down.
Blotchy brown brick houses.
Number eighty still unlet.
Why is that?
Valuation is only twenty-eight.
Towers, Battersby, North, MacArthur:
parlour windows plastered with bills.
Plasters on a sore eye.
To smell the gentle smoke of tea,
fume of the pan, sizzling butter.
Be near her ample bedwarmed flesh.
Yes, yes.

\N1:
Quick warm sunlight came running from Berkeley road,
\gab{240}
swiftly,
in slim sandals,
along the brightening footpath.

\BloomInt:
Runs, she runs to meet me,
a girl with gold hair on the wind.
