% Abbreviated reading for Folio Seattle

...

\Nnovel:
The three girl friends were seated on the rocks,
enjoying the evening scene |
and the air |
which was fresh but not too chilly.%
\gab{10}
Many a time and oft |
were they wont to come there |
to that favourite nook |
to have a cosy chat
beside the sparkling waves |
and discuss matters feminine,
Cissy Caffrey
and Edy Boardman with the baby in the push·car |
and Tommy and Jacky Caffrey,
two little curly·headed boys,
dressed in sailor suits |
with caps to match |
and the name H.M.S. Belleisle printed on both.
For Tommy and Jacky Caffrey were twins,
scarce four years old |
and very noisy and spoiled twins sometimes |
but for all that |
darling little fellows |
with bright merry faces and endearing ways about them.

...

\gertyNovel:
Gerty MacDowell |
who was seated near her companions,
lost in thought,
gazing far away into the distance |
was,%
\gab{80}
in very truth,
as fair a specimen |
of winsome Irish girlhood |
as one could wish to see.
She was pronounced beautiful |
by all who knew her

...

\gertyNovel:
The paly light of evening |
falls upon a face infinitely sad and wistful.
Gerty MacDowell yearns in vain.

\gertyRomantic:
Yes,
she had known from the very first |
that her daydream of a marriage

...

\gertyRomantic:
was not to be.
He was too young to understand.
\gab{200}
He would not believe in love,
a woman's birthright.

...

\gertyRomantic:
Strength of character |
had never been Reggy Wylie's strong point |
and he who would woo and win Gerty MacDowell |
must be a man among men.

\gertyNovel:
But waiting,
always waiting to be asked |
and it was leap year too |
and would soon be over.
No prince charming is her beau ideal |
to lay a rare and wondrous love at her feet |
\gab{210}

\gertySex:
but rather a manly man |
with a strong quiet face |
who had not found his ideal,
perhaps his hair slightly flecked with grey,
and who would understand,
take her in his sheltering arms,
strain her to him |
in all the strength of his deep passionate nature |
and comfort her |
with a long long kiss.
It would be like heaven.

\gertyRomantic:
For such a one |
she yearns this balmy summer eve.
With all the heart of her |
she longs to be his only,
his affianced bride |
for riches
for poor,
in sickness
in health,
till death us two part,
from this |
to this day forward.

...

\gertyNovel:
He would be tall with broad shoulders
(she had always admired tall men for a husband)
with glistening white teeth
under his carefully trimmed sweeping moustache |
and they would go on the continent
for their honeymoon
(three wonderful weeks!)
and then,
when they settled down in a nice snug and cosy little homely house,
every morning |
they would both have brekky,%
\gab{240}
simple but perfectly served,
for their own two selves |
and before he went out to business |
he would give his dear little wifey |
a good hearty hug |
and gaze for a moment |
deep down into her eyes.

...

\gertyReal:
\gab{360}
A delicate pink
crept into her pretty cheek |
but she was determined to let them see |
so she just lifted her skirt |
a little but just enough |
and took good aim |
and gave the ball a jolly good kick |
and it went ever so far |
and the two twins after it |
down towards the shingle.

\gertyJudgy:
Pure jealousy |
of course |
it was nothing else |
to draw attention |
on account of the gentleman opposite looking.

\gertySex:
She felt the warm flush,
a danger signal always with Gerty MacDowell,
surging and flaming into her cheeks.

\gertyRomantic:
Till then |
they had only exchanged glances |
of the most casual |
but now |
under the brim of her new hat |
she ventured a look at him |
and the face that met her gaze there in the twilight,
wan and strangely drawn,
seemed to her |%
\gab{370}
the saddest |
she had ever seen.

\Nrelig:
Through the open window of the church |
the fragrant incense was wafted |
and with it |
the fragrant names of her |
who was conceived without stain of original sin,
spiritual vessel,
pray for us,
honourable vessel,
pray for us,
vessel of singular devotion,
pray for us,
mystical rose.
And careworn hearts were there |
and toilers for their daily bread |
and many who had erred and wandered,
their eyes wet with contrition |
but for all that |
bright with hope |
for the reverend father Father Hughes
had told them |
what the great saint Bernard said in his famous prayer |
of Mary,
the most pious Virgin's intercessory power |
that it was not recorded in any age |
that those who implored her powerful protection |%
\gab{380}
were ever abandoned by her.

...

\gertyNovel:
He was in deep mourning,
she could see that,
and the story of a haunting sorrow |
was written on his face.
She would have given worlds |
to know what it was.

\gertySex:
He was looking up so intently,
so still,
and he saw her kick the ball |
and perhaps he could see |
the bright steel buckles of her shoes |
if she swung them like that |
thoughtfully with the toes down.
She was glad that something told her |
to put on the transparent stockings |
thinking Reggy Wylie might be out |
but that was far away.

\gertyRomantic:
Here |
was that of which she had so often dreamed.
It was he who mattered |
and there was joy on her face |
because she wanted him |
because she felt instinctively |
that he was like no-one else.%
\gab{430}
The very heart of the girl·woman |
went out to him,
her dream·husband,
because she knew on the instant |
it was him.

\gertyNovel:
If he had suffered,
more sinned against than sinning,
or even,
even,
if he had been himself a sinner,
a wicked man,
she cared not.
Even if he was a protestant or methodist |
she could convert him easily |
if he truly loved her.
There were wounds |
that wanted healing with heart·balm.

\gertySex:
She was a womanly woman

\gertyJudgy:
not like other flighty girls unfeminine |
he had known,
those cyclists showing off what they hadn't got

\gertySex:
and she just yearned to know all,
to forgive all |
if she could make him |
fall in love with her,
make him forget |
the memory of the past.
Then mayhap |
he would embrace her gently,
like a real man,
crushing her soft body to him,%
\gab{440}
and love her,
his ownest girlie,
for herself alone.

\Nrelig:
Refuge of sinners.
Comfortress of the afflicted.
\latin{Ora pro nobis}.
Well has it been said |
that whosoever |
prays to her with faith and constancy |
can never be lost or cast away:
and fitly |
is she too |
a haven of refuge |
for the afflicted |
because of the seven dolours |
which transpierced her own heart.

...

\N1:
Cissy and Edy shouted after the twins to come back |
because they were afraid |
the tide might come in on them |
and be drowned.

...

\N1:
Cissy Caffrey caught the two twins |
and she was itching |
to give them a ringing good clip on the ear |
but she didn't |
because she thought |
he might be watching

\gertyJudgy:
but she never made a bigger mistake in all her life

\gertySex:
because Gerty could see without looking |
that he never took his eyes off of her

\Nrelig:
and then Canon O'Hanlon
handed the thurible
back to Father Conroy |
and knelt down |
looking up at the Blessed Sacrament |
and the choir
began to sing the \latin{tantum ergo}

\gertySex:
and she just swung her foot |
in and out in time

\Nrelig:
as the music rose and fell |
to the \latin{tantumer gosa cramen tum}.

...

\gertySex:
and that was what he was looking at,
transparent,

...

\N1:
Cissy came up along the strand |
with the two twins
and their ball

\gertyJudgy:
with her hat anyhow on her |
to one side after her run |
and she did look a streel |
tugging the two kids |
along with the flimsy blouse |
she bought only a fortnight before |
like a rag on her back |
and a bit of her petticoat
hanging like a caricature.

\gertyNovel:
Gerty just took off her hat for a moment |
to settle her hair |

...

\gertySex:
She could almost see |
the swift answering flash of admiration in his eyes |
that set her tingling
in every nerve.
She put on her hat |
so that she could see from underneath
the brim |
and swung her buckled shoe faster |
for her breath caught |
as she caught the expression
in his eyes.
He was eying her |
as a snake eyes its prey.
Her woman's instinct told her |
that she had raised the devil in him |
and at the thought |
a burning scarlet
swept from throat to brow |
till the lovely colour of her face
became a glorious rose.%
\gab{520}

...

\gertySex:
she could see the gentleman
winding his watch |
and listening to the works |
and she swung her leg more |
in and out |
in time.
It was getting darker |
but he could see |
and he was looking all the time |
that he was winding the watch |
or whatever he was doing to it |
and then he put it back |
and put his hands back into his pockets.%
\gab{560}
She felt a kind of a sensation |
rushing all over her

\gertyReal:
and she knew by the feel of her scalp |
and that irritation against her stays |
that that thing must be coming on |
because the last time too |
was when she clipped her hair |
on account of the moon.

\gertySex:
His dark eyes fixed themselves on her again |
drinking in her every contour, |
literally worshipping at her shrine.
If ever there was undisguised admiration |
in a man's passionate gaze |
it was there |
plain to be seen |
on that man's face.

\gertyNovel:
It is for you,
Gertrude MacDowell,
and you know it.

...

\gertyRomantic:
and oft·times |
the beauty of poetry,
so sad in its transient loveliness,
had misted her eyes |
with silent tears |
for she felt
that the years were slipping by for her,
one by one,

\gertyNovel:
and |
but for that one shortcoming |%
\gab{650}
she knew she need fear no competition

\gertyReal:
and that was an accident coming down Dalkey hill |
and she always tried to conceal it.

\gertyNovel:
But it must end,
she felt.
If she saw that magic lure in his eyes |
there would be no holding back for her.
Love laughs at locksmiths.
She would make the great sacrifice.
Her every effort |
would be to share his thoughts.

...

\gertyNovel:
Perhaps it was an old flame |
he was in mourning for |
from the days beyond recall.
She thought she understood.
She would try to understand him |
because men were so different.
The old love was waiting,
waiting with little white hands stretched out,
with blue appealing eyes.%
\gab{670}
Heart of mine!

\gertyRomantic:
She would follow,
her dream of love,
the dictates of her heart |
that told her |
he was her all in all,
the only man in all the world |
for her |
for love was the master guide.

\gertySex:
Nothing else mattered.
Come what might |
she would be wild,
untrammelled,
free.

\Nrelig:
Canon O'Hanlon
put the Blessed Sacrament
back into the tabernacle |
and genuflected |
and the choir sang \latin{laudate dominum omnes gentes} |
and then he locked the tabernacle door |
because the benediction was over |
and Father Conroy handed him his hat |
to put on

\N1:
and cross·cat Edy asked |
wasn't she coming |
but Jacky Caffrey called out:

\jacky:
O,
look,
Cissy!

\N1:
And they all looked |
was it sheet lightning |%
\gab{680}
but Tommy saw it too |
over the trees beside the church,
blue |
and then green
and purple.

\cissy:
It's fireworks.
\sout{Cissy Caffrey said.}

\N1:
And they all ran down the strand |
to see over the houses
and the church,
helter·skelter,
Edy with the push·car |
with baby Boardman in it |
and Cissy holding Tommy and Jacky by the hand |
so they wouldn't fall running.

\cissy:
Come on,
Gerty.
\sout{Cissy called.}
It's the bazaar fireworks.

\gertyReal:
But Gerty was adamant.
She had no intention
of being at their beck and call.
If they could run like rossies |
she could sit |
so she said she could see |
from where she was.

\gertySex:
The eyes that were fastened upon her |
set her pulses tingling.%
\gab{690}
She looked at him a moment,
meeting his glance,
and a light broke in upon her.
White·hot passion was in that face,
passion silent as the grave,
and it had made her |
his.

\gertyNovel:
At last |
they were left alone |
without the others to pry and pass remarks |
and she knew |
he could be trusted to the death,
steadfast,
a sterling man,
a man of inflexible honour |
to his fingertips.

\gertySex:
His hands and face were working |
and a tremour went over her.
She leaned back far |
to look up
where the fireworks were |
and she caught her knee in her hands |
so as not to fall back looking up |
and there was no-one to see |
only him |
and her |
when she revealed |
all her graceful beautifully shaped legs like that,
supply soft and delicately rounded,
and she seemed to hear |
the panting of his heart,
his hoarse breathing,%
\gab{700}
because she knew too |
about the passion of men like that,
hot·blooded,

...

\N1:
And Jacky Caffrey shouted |
to look,
there was another

\gertySex:
and she leaned back |
and the garters were blue to match |
on account of the transparent

\N1:
and they all saw it |
and they all shouted to look,
look,
there it was

\gertySex:
and she leaned back ever so far |
to see the fireworks |
and something queer was flying through the air,
a soft thing,
to and fro,
dark.
And she saw a long Roman candle |
going up over the trees,%
\gab{720}
up,
up,

\N1:
and,
in the tense hush,
they were all breathless with excitement |
as it went higher and higher

\gertySex:
and she had to lean back more and more |
to look up after it,
high,
high,
almost out of sight,
and her face was suffused |
with a divine,
an entrancing blush from straining back |
and he could see her other things too,

...

\gertySex:
and she let him |
and she saw that he saw |
and then it went so high |
it went out of sight a moment |
and she was trembling in every limb |
from being bent so far back |
that he had a full view |
high up above her knee |
where no-one ever |
not even on the swing |
or wading |
and she wasn't ashamed |
and he wasn't either |%
\gab{730}
to look in that immodest way like that |
because he couldn't resist |
the sight of the wondrous revealment |
half offered like those skirt·dancers |
behaving so immodest before gentlemen looking |
and he kept on looking,
looking.

...

\gertySex:
a little strangled cry,
wrung from her,
that cry
that has rung through the ages.

\N1:
And then a rocket
sprang and bang |
shot blind blank |
and

\gertySex:
O!

\N1:
then the Roman candle burst |
and it was like a sigh of

\gertySex:
O!

\N1:
and everyone cried

\All:
O!
O!

\N1:
in raptures |
and it gushed out of it |
a stream of rain gold hair threads |
and they shed |
and

\All:
ah!

\N1:
they were
all greeny dewy stars falling with golden,%
\gab{740}

\gertySex:
O so lovely,
O,
soft,
sweet,
soft!

\Nnovel:
Then all melted away |
dewily in the grey air:
all was silent.

\gertySex:
Ah!

\Nnovel:
She glanced at him |
as she bent forward quickly,
a pathetic little glance of piteous protest,
of shy reproach |
under which he coloured like a girl |
He was leaning back against the rock behind.

\N2:
Leopold Bloom
(for it is he)
stands silent,
with bowed head |
before those young guileless eyes.

\Nnovel:
What a brute he had been!
At it again?
A fair unsullied soul had called to him |
and,
wretch that he was,
how had he answered?
An utter cad
he had been!
He of all men!

\N2:
But there was an infinite store of mercy |
in those eyes,
for him too |
a word of pardon |
even though he had erred |
and sinned |
and wandered.

\gertyNovel:
Should a girl tell?%
\gab{750}
No,
a thousand times no.
That was their secret,
only theirs,
alone in the hiding twilight

\N2:
and there was none to know or tell |
save the little bat |
that flew so softly through the evening |
to and fro |
and little bats don't tell.

\N1:
Cissy Caffrey whistled,
imitating the boys in the football field |
to show what a great person she was:
and then she cried:

\cissy:
Gerty!
Gerty!
We're going.
Come on.
We can see from farther up.

\gertyNovel:
Gerty had an idea,
one of love's little ruses.
She slipped a hand |
into her kerchief pocket |
and took out the wadding |
and waved in reply |
of course 
without letting him |
and then slipped it back.

...

\gertyNovel:
She half smiled at him wanly,
a sweet forgiving smile,
a smile that verged on tears,
and then they parted.

\N1:
Slowly,
without looking back |
she went down the uneven strand to Cissy,
to Edy |
to Jacky and Tommy Caffrey,
to little baby Boardman.
It was darker now |
and there were stones and bits of wood on the strand |
and slippy seaweed.
She walked with a certain quiet dignity |
characteristic of her |
but with care |%
\gab{770}
and very slowly because—%
because Gerty MacDowell was ...

\BloomInt:
Tight boots?
No.
She's lame!
O!

\N2:
Mr Bloom watched her |
as she limped away.

\BloomCurrent:
Poor girl!
That's why
she's left on the shelf |
and the others did a sprint.
Thought something was wrong |
by the cut of her jib.
Jilted beauty.

\BloomAbstract:
A defect is ten times worse in a woman.
But makes them polite.

\BloomCurrent:
Glad I didn't know it |
when she was on show.
Hot little devil |
all the same.
I wouldn't mind.
Curiosity |
like a nun |
or a negress |
or a girl with glasses.
That squinty one is delicate.
Near her monthlies,
I expect,
makes them feel ticklish.

\BloomOther:
I have such a bad headache today.

\BloomToday:
Where did I put the letter?
Yes,
all right.

...

\BloomToday:
Damned glad |
I didn't do it in the bath this morning |
over her silly |
I will punish you letter.
Made up for that tram·driver this morning.
That gouger M'Coy stopping me |
to say nothing.
And his wife |
engagement in the country |
valise,
voice like a pickaxe.

...

\BloomCurrent:
Daresay |
she felt I.
When you feel like that |
you often meet what you feel.
Liked me or what?

\BloomAbstract:
Dress |
they look at.
Always know a fellow courting:
collars and cuffs.%
\gab{830}
Well |
cocks and lions do the same |
and stags.
Same time |
might prefer a tie undone |
or something.
Trousers?

\BloomCurrent:
Suppose I |
when I was?
No.
Gently does it.
Dislike rough and tumble.
Kiss in the dark
and never tell.
Saw something in me.
Wonder what.
Sooner have me
as I am |
than some poet chap |
with bears·grease plastery hair,
love·lock over his dexter optic.

...

\BloomCurrent:
Ought to attend to my appearance |
my age.
Didn't let her see me in profile.

\BloomAbstract:
Still,
you never know.
Pretty girls and ugly men marrying.
Beauty and the beast.

\BloomHist:
Besides I can't be so |
if Molly.

\BloomCurrent:
Took off her hat to show her hair.
Wide brim.
Bought to hide her face,
meeting someone might know her,
bend down |
or carry a bunch of flowers to smell.
Hair |
strong in rut.
\gab{840}

\BloomHist:
Ten bob |
I got for Molly's combings |
when we were on the rocks
in Holles street.
Why not?

\BloomToday:
Suppose he gave her money.

\BloomHist:
Why not?
All a prejudice.
She's worth ten,
fifteen,
more,
a pound.
What?
I think so.
All that for nothing.
Bold hand:
Mrs Marion.
Did I forget
to write address on that letter |
like the postcard I sent to Flynn?
And the day I went to Drimmie's |
without a necktie.
Wrangle with Molly
it was put me off.
No,
I remember.
Richie Goulding:
he's another.
Weighs on his mind.

\BloomToday:
Funny |
my watch stopped at half past four.

\BloomAbstract:
Dust.
Shark liver oil
they use to clean.
Could do it myself.
Save.

\BloomToday:
Was that just when he,
she?

\BloomToday:
O,
he did.
Into her.
She did.
Done.

\BloomInt:
Ah!%
\gab{850}

\N2:
Mr Bloom with careful hand |
recomposed his wet shirt.

\BloomCurrent:
O Lord,
that little limping devil.
Begins to feel cold and clammy.
After·effect not pleasant.

\BloomAbstract:
Still |
you have to get rid of it someway.
They don't care.
Complimented |
perhaps.
Go home to nicey bread and milky |
and say night prayers with the kiddies.
Well,
aren't they?
See her as she is |
spoil all.
Must have the stage setting,
the rouge,
costume,
position,
music.
The name too.
\french{Amours} of actresses.
Nell Gwynn,
Mrs Bracegirdle,
Maud Branscombe.
Curtain up.
Moonlight silver effulgence.
Maiden discovered with pensive bosom.
Little sweetheart |
come and kiss me.
Still,
I feel.
The strength it gives a man.
That's the secret of it.%
\gab{860}

...

\BloomCurrent:
Didn't look back |
when she was going down the strand.
Wouldn't give that satisfaction.

\BloomHist:
Those girls,
those girls,
those lovely seaside girls.

\BloomCurrent:
Fine eyes she had,
clear.

\BloomAbstract:
It's the white of the eye brings that out |
not so much the pupil.

\BloomCurrent:
Did she know |
what I?
Course.
Like a cat |
sitting beyond a dog's jump.

\BloomHist:
Women never meet one |
like that Wilkins in the high school |
drawing a picture of Venus |
with all his belongings on show.%
\gab{910}
Call that innocence?
Poor idiot!
His wife has her work cut out for her.

\BloomAbstract:
Never see them sit |
on a bench marked \Title{Wet Paint}.
Eyes all over them.
Look under the bed |
for what's not there.
Longing to get the fright of their lives.
Sharp as needles |
they are.

...

\BloomCurrent:
Still she was game.
Lord,
I am wet.
Devil you are.
Swell of her calf.
Transparent stockings,
stretched to breaking point.%
\gab{930}

...

\N2:
A monkey puzzle rocket burst,
spluttering in darting crackles.
Zrads and zrads,
zrads,
zrads.
And Cissy and Tommy and Jacky ran out to see |
and Edy after with the push·car |
and then Gerty |
beyond the curve of the rocks.

\BloomInt:
Will she?
Watch!
Watch!
See!
Looked round.
She smelt an onion.

\BloomAbstract:
Darling,
I saw,
your.
I saw all.

\BloomInt:
Lord!

\BloomCurrent:
Did me good |
all the same.

...

\BloomToday:
Mrs Beaufoy,
Purefoy.
Must call to the hospital.
Wonder |
is nurse Callan there still.%
\gab{960}

\BloomHist:
She used to look over some nights |
when Molly was in the Coffee Palace.
That young doctor O'Hare |
I noticed her brushing his coat.
And Mrs Breen |
and Mrs Dignam |
once like that too,
marriageable.
Worst of all at night |
Mrs Duggan told me in the City Arms.
Husband rolling in drunk,
stink of pub off him |
like a polecat.
Have that in your nose in the dark,
whiff of stale boose.
Then ask in the morning:

\BloomOther:
was I drunk last night?

\BloomAbstract:
Bad policy however |
to fault the husband.
Chickens come home to roost.
They stick by one another like glue.
Maybe the women's fault also.

\BloomHist:
That's where Molly can knock spots off them.
It's the blood of the south.
Moorish.
Also the form,
the figure.
Hands felt for the opulent.
Just compare |
for instance |
those others.%
\gab{970}
Wife locked up at home,
skeleton in the cupboard.

\BloomOther:
Allow me to introduce my.

\BloomHist:
Then they trot you out |
some kind of a nondescript,
wouldn't know what to call her.
Always see a fellow's weak point |
in his wife.

\BloomAbstract:
Still there's destiny in it,
falling in love.
Have their own secrets between them.
Chaps that would go to the dogs |
if some woman didn't take them in hand.
Then little chits of girls,
height of a shilling in coppers,
with little hubbies.
As God made them |
he matched them.
Sometimes children turn out well enough.
Twice nought | makes one.
Or old rich chap of seventy |
and blushing bride.
Marry in May |
and repent in December.

\BloomCurrent:
This wet is very unpleasant.
Stuck.
Well |
the foreskin is not back.
Better detach.%
\gab{980}

\BloomInt:
Ow!

\BloomAbstract:
Other hand |
a six·footer with a wifey up to his watch·pocket.
Long and the short of it.
Big he |
and little she.

...

\BloomCurrent:
Wait.
Hm.
Hm.
Yes.
That's her perfume.
Why she waved her hand.
I leave you this |
to think of me |
when I'm far away on the pillow.
What is it?
Heliotrope?
No.
Hyacinth?
Hm.
Roses,
I think.
She'd like scent of that kind.
Sweet and cheap:
soon sour.%
\gab{1010}

\BloomHist:
Why Molly likes opoponax.
Suits her,
with a little jessamine mixed.
Her high notes and her low notes.
At the dance |
night she met him,
dance of the hours.
Heat brought it out.
She was wearing her black |
and it had the perfume of the time before.
Good conductor,
is it?
Or bad?
Light too.
Suppose there's some connection.
For instance |
if you go into a cellar |
where it's dark.
Mysterious thing too.

\BloomCurrent:
Why did I smell it only now?
Took its time in coming |
like herself,
slow but sure.
Suppose it's ever so many millions of tiny grains blown across.

\BloomAbstract:
Yes,
it is.
Because those spice islands,
Cinghalese this morning,
smell them leagues off.
Tell you what it is.
It's like a fine fine veil
or web they have |
all over the skin,
fine |
like |
what do you call it |
gossamer,%
\gab{1020}
and they're always spinning it out of them,
fine as anything,
like rainbow colours |
without knowing it.

\BloomHist:
Clings to everything she takes off.
Vamp of her stockings.
Warm shoe.
Stays.
Drawers:
little kick,
taking them off.
By·by till next time.
Also |
the cat likes to sniff in her shift
on the bed.
Know her smell in a thousand.
Bathwater too.
Reminds me of strawberries and cream.
Wonder where it is really.
There |
or the armpits |
or under the neck.
Because you get it |
out of all holes and corners.

\BloomAbstract:
Hyacinth perfume |
made of oil of ether or something.
Muskrat.
Bag under their tails.
One grain |
pour off odour for years.
Dogs at each other behind.

\BloomOther:
Good evening.

---Evening.

---How do you sniff?

---Hm.

---Hm.

---Very well,
thank you.

\BloomAbstract:
Animals go by that.%
\gab{1030}
Yes now,
look at it that way.
We're the same.
Some women,
instance,
warn you off |
when they have their period.
Come near.
Then get a hogo |
you could hang your hat on.
Like what?
Potted herrings gone stale or.
Boof!
Please keep off the grass.

\BloomAbstract:
Perhaps they get a man smell off us.
What though?
Cigary gloves |
long John had on his desk |
the other day.
Breath?
What you eat and drink |
gives that.
No.
Man·smell,
I mean.
Must be connected with that |
because priests |
that are supposed to be |
are different.
Women buzz round it |
like flies round treacle.
Railed off the altar |
get on to it |
at any cost.
The tree of forbidden priest.

\BloomOther:
O,
father,
will you?

---Let me be the first to.

\BloomAbstract:
That diffuses itself all through the body,
permeates.%
\gab{1040}
Source of life.
And it's extremely curious |
the smell.
Celery sauce.

...

\Nnovel:
A last lonely candle |
wandered up the sky |
from Mirus bazaar |
in search of funds for Mercer's hospital |
and broke,
drooping,
and shed a cluster of violet |
but one white |
stars.
They floated,
fell:
they faded.
The shepherd's hour:
the hour of folding:
hour of tryst.
From house to house,
giving his ever·welcome double knock,%
\gab{1170}
went the nine o'clock postman,
the glow·worm's lamp at his belt |
gleaming here and there |
through the laurel hedges.
And among the five young trees |
a hoisted lint·stock |
lit the lamp at Leahy's terrace.
By screens of lighted windows,
by equal gardens |
a shrill voice went crying,
wailing:
\Title{Evening Telegraph,
stop press edition!
Result of the Gold Cup Race!}
and from the door of Dignam's house |
a boy ran out and called.

\N2:
Twittering |
the bat flew here,
flew there.
Far out over the sands |
the coming surf crept,
grey.
Howth settled for slumber,
tired of long days,
of yum·yum rhododendrons
(he was old)
and felt gladly |
the night breeze lift,
ruffle his fell of ferns.
He lay |
but opened a red eye
unsleeping,
deep and slowly breathing,
slumberous but awake.%
\gab{1180}
And far on Kish bank |
the anchored light·ship twinkled,
winked at Mr Bloom.

...
