%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\input{defs}
% \input{largeprint} \renewcommand{\gab}[1]{\marginpar[\tiny\raggedleft #1]{\tiny #1}}


\title{\Huge Ithaca}
\author{George~V. Reilly\\
{\small for the}\\
Wild Geese Players of Seattle\\
{\emph{www.WildGeeseSeattle.org}}\\
\\
{\small Chapter~17 of \emph{Ulysses} by James Joyce}\\
{\small Adapted from the 1922~edition at Project Gutenberg}
\\
{\small Final Draft}}
\date{May \nth{21}, 2012}
\raggedbottom

\begin{document}

% Page i, Title Page
\maketitle
\thispagestyle{empty}
\pagebreak

% Page ii, Cast
\begin{tabular}{ll}
\multicolumn{1}{c}{\large Cast} \\
\\
Bloom \\
Stephen \\
Q1---Catechist \#1\\
Q2---Catechist \#2 \\
Adverts \\
Factual \\
Household \\
Memories \\
Philosophy \\
Places \\
Poetry \\
Religious \\
Science \\
Theater \\
Molly
\end{tabular}

\newpage


% Page 1, first page of script proper
\setcounter{page}{1}

\acting{Bloom and Stephen walk/talk}

\Q1
What parallel courses |
did Bloom and Stephen follow |
returning?

\Places
Starting united |
both at normal walking pace from Beresford place |
they followed in the order named |
Lower and Middle Gardiner streets |
and Mountjoy square, west:
then, at reduced pace, each bearing left, |
Gardiner's place by an inadvertence |
as far as the farther corner of Temple street: |
then, at reduced pace with interruptions of halt, |
bearing right, Temple street, |
north, as far as Hardwicke place.
Approaching, disparate, at relaxed walking pace |
they crossed both the circus before George's church diametrically, |
the chord in any circle being less than the arc which it subtends.%
\gab{10}


\Q2
Of what |
did the duumvirate deliberate |
during their itinerary?

\Factual
Music, literature,
Ireland, Dublin, Paris, |
friendship,
woman, prostitution, |
diet, |
the influence of gaslight |
or the light of arc and glowlamps on the growth of adjoining paraheliotropic trees, |
exposed corporation emergency dustbuckets, |
the Roman catholic church, ecclesiastical celibacy, |
the Irish nation, jesuit education, |
careers, the study of medicine, |
the past day, |
the maleficent influence of the presabbath, |
Stephen's collapse.

\Q1
Did Bloom discover common factors of similarity |
between their respective like and unlike reactions to experience?

\Philosophy
Both were sensitive to artistic impressions,
\gab{20}
musical in preference to plastic or pictorial.
Both preferred a continental to an insular manner of life,
a cisatlantic to a transatlantic place of residence.
Both indurated by early domestic training
and an inherited tenacity of heterodox resistance |
professed their disbelief
in many orthodox religious, national, social and ethical doctrines.
Both admitted the alternately stimulating and obtunding influence
of heterosexual magnetism.


\Q2
Were their views on some points |
divergent?

\Factual
Stephen dissented openly from Bloom's views |
on the importance of dietary and civic selfhelp |
while Bloom dissented tacitly from Stephen's views |
on the eternal affirmation of the spirit of man in literature.
\gab{30}
Bloom assented covertly to Stephen's
rectification of the anachronism |
involved in assigning the date of the conversion of the Irish nation
to christianity from druidism |
by Patrick son of Calpornus, son of Potitus, son of Odyssus, |
sent by pope Celestine~I in the year~432 in the reign of Leary |
to the year~260 or thereabouts in the reign of Cormac MacArt
\paren{(died 266~A.D.),} |
suffocated by imperfect deglutition of aliment |
at Sletty and interred at Rossnaree.
The collapse |
which Bloom ascribed to gastric inanition and
certain chemical compounds of varying degrees of adulteration and alcoholic strength, |
accelerated by mental exertion
and the velocity of rapid circular motion in a relaxing atmosphere, |
\gab{40}
Stephen attributed to the reapparition of a matutinal cloud |
\paren{(perceived by both from two different points of observation Sandycove and Dublin)} |
at first no bigger than a woman's hand.


\Q1
Was there one point |
on which their views were equal and negative?

\Science
The influence of gaslight or electric light
on the growth of adjoining paraheliotropic trees.


\Q2
Had Bloom discussed similar subjects |
during nocturnal perambulations in the past?

\Places
In~1884 |
with Owen Goldberg and Cecil Turnbull |
at night on public thoroughfares
between Longwood avenue and Leonard's corner |
and Leonard's corner and Synge street |
\gab{50}
and Synge street and Bloomfield avenue.
In~1885 |
with Percy Apjohn in the evenings, |
reclined against the wall
between Gibraltar villa and Bloomfield house in Crumlin,
barony of Uppercross.
In~1886 |
occasionally |
with casual acquaintances and prospective purchasers on doorsteps,
in front parlours, in third class railway carriages of suburban lines.
In~1888 |
frequently |
with major Brian Tweedy and his daughter Miss Marion Tweedy,
together and separately on the lounge in Matthew Dillon's house in Roundtown.
Once in~1892 and once in~1893 |
with Julius
\paren{(Juda)}
Mastiansky, |
on both occasions in the parlour of his
\paren{(Bloom's)}
house in Lombard street, west.


\Q1
What reflection |
concerning the irregular sequence of dates |
1884,~1885, 1886, 1888, 1892, 1893,~1904 |
\gab{60}
did Bloom make before their arrival at their destination?

\Philosophy
He reflected that the progressive extension
of the field of individual development and experience |
was regressively accompanied by a restriction of
the converse domain of interindividual relations.


\Q2
As in what ways?

\Religious
From inexistence to existence |
he came to many and was as one received: |
existence with existence |
he was with any as any with any: |
from existence to nonexistence gone |
he would be by all as none perceived.

\acting{arrive at house, Bloom get key out of pocket}

\Q1
What act |
did Bloom make on their arrival at their destination?
\gab{70}

\Factual
At the housesteps of the \nth{4}~of the equidifferent uneven numbers,
number~7 Eccles street,
he inserted his hand mechanically
into the back pocket of his trousers |
to obtain his latchkey.


\Q2
Was it there?

\Factual
It was in the corresponding pocket of the trousers |
which he had worn on the day but one preceding.


\Q1
Why was he doubly irritated?

\Bloom
Because he had forgotten |
and because he remembered that
he had reminded himself twice |
not to forget.


\Q2
What were then |
\gab{80}
the alternatives before the,
premeditatedly
\paren{(respectively)}
and inadvertently,
keyless couple?

\Philosophy
To enter or not to enter.
To knock or not to knock.


\acting{Bloom enact as best he can, climb, jump}

\Q1
Bloom's decision?

\Science
A~stratagem.
Resting his feet on the dwarf wall,
he climbed over the area railings,
compressed his hat on his head,
grasped two points at the lower union of rails and stiles,
lowered his body gradually |
by its length of five feet nine inches and a half |
to within two feet ten inches of the area pavement |
and allowed his body to move freely in space |
by separating himself from the railings |
and crouching in preparation for the impact of the fall.


\acting{Bloom enact as best he can, fall}

\Q1
Did he fall?
\gab{90}

\Science
By his body's known weight |
of eleven stone and four pounds in avoirdupois measure, |
as certified by the graduated machine for periodical selfweighing |
in the premises of Francis Froedman, |
pharmaceutical chemist of 19~Frederick street, north, |
on the last feast of the Ascension, |
to wit, the twelfth day of May of the bissextile year |
one thousand nine hundred and four of the christian era |
\paren{(jewish era | five thousand six hundred and sixtyfour, |
mohammadan era | one thousand three hundred and twentytwo),} |
golden number~5,
epact~13,
solar cycle~9,
dominical letters~C~B,
Roman indiction~2,
Julian period~6617,
MCMIV.


\acting{Bloom enact as best he can}

\Q1
Did he rise |
uninjured by concussion?
\gab{100}

\Science
Regaining new stable equilibrium |
he rose uninjured though concussed by the impact, |
raised the latch of the area door
by the exertion of force at its freely moving flange |
and by leverage of the first kind applied at its fulcrum, |
gained retarded access to the kitchen through the subadjacent scullery, |
ignited a lucifer match by friction, |
set free inflammable coal gas by turning on the ventcock, |
lit a high flame which, |
by regulating, |
he reduced to quiescent candescence |
and lit finally a portable candle.


\acting{both enact as best they can}

\Q2
What discrete succession of images |
did Stephen meanwhile perceive?

\Science
Reclined against the area railings |
he perceived through the transparent kitchen panes |
a man regulating a gasflame of 14~CP,
\gab{110}
a man lighting a candle of 1~CP,
a man removing in turn each of his two boots,
a man leaving the kitchen holding a candle.


\Q1
Did the man reappear elsewhere?

\Factual
After a lapse of four minutes |
the glimmer of his candle was discernible
through the semitransparent semicircular glass fanlight over the halldoor.
The halldoor turned gradually on its hinges.
In the open space of the doorway |
the man reappeared without his hat, with his candle.


\Q2
Did Stephen obey his sign?

\Factual
Yes,
entering softly,
he helped to close and chain the door |
and followed softly along the hallway |
\gab{120}
the man's back and listed feet and lighted candle |
past a lighted crevice of doorway on the left
and carefully down a turning staircase of more than five steps |
into the kitchen of Bloom's house.


\Q1
What did Bloom do?

\Science
He extinguished the candle |
by a sharp expiration of breath upon its flame, |
drew two spoonseat deal chairs to the hearthstone, |
one for Stephen with its back to the area window, |
the other for himself when necessary, |
knelt on one knee, |
composed in the grate a pyre of crosslaid resintipped sticks
and various coloured papers |
and irregular polygons of best Abram coal |
at twentyone shillings a ton
from the yard of Messrs Flower and M'Donald of 14~D'Olier street, |
kindled it at three projecting points of paper
\gab{130}
with one ignited lucifer match, |
thereby releasing the potential energy
contained in the fuel |
by allowing its carbon and hydrogen elements
to enter into free union with the oxygen of the air.


\acting{speaking of his memories of hearths, litany}

\Q2
Of what similar apparitions |
did Stephen think?

\Stephen
Of others elsewhere in other times |
who, kneeling on one knee or on two, |
had kindled fires for him, |
of Brother Michael |
in the infirmary of the college of the Society of Jesus
at Clongowes Wood, Sallins, in the county of Kildare:
of his father, Simon Dedalus, |
in an unfurnished room of his first residence in Dublin,
number thirteen Fitzgibbon street:
of his godmother Miss Kate Morkan |
in the house of her dying sister Miss Julia Morkan
\gab{140}
at 15~Usher's Island:
of his aunt Sara, |
wife of Richie
\paren{(Richard)}
Goulding,
in the kitchen of their lodgings at 62~Clanbrassil street:
of his mother Mary, |
wife of Simon Dedalus,
in the kitchen of number twelve North Richmond street |
on the morning of the feast of Saint Francis Xavier~1898:
of the dean of studies, Father Butt, |
in the physics' theatre of university College,
16~Stephen's Green, north:
of his sister Dilly
\paren{(Delia)} |
in his father's house in Cabra.


\acting{Stephen ``raise gaze''}

\Q2
What did Stephen see |
on raising his gaze to the height of a yard from the fire |
towards the opposite wall?

\Factual
Under a row of five coiled spring housebells |
\gab{150}
a curvilinear rope, |
stretched between two holdfasts athwart
across the recess beside the chimney pier, |
from which hung four smallsized square handkerchiefs
folded unattached consecutively in adjacent rectangles |
and one pair of ladies' grey hose
with Lisle suspender tops and feet in their habitual position |
clamped by three erect wooden pegs |
two at their outer extremities |
and the third at their point of junction.


\acting{Bloom look at range}

\Q1
What did Bloom see on the range?
\gab{160}

\Household
On the right
\paren{(smaller)}
hob |
a blue enamelled saucepan:
on the left
\paren{(larger)}
hob |
a black iron kettle.


\acting{Bloom enact as best he can}

\Q1
What did Bloom do at the range?

\Household
He removed the saucepan to the left hob, |
rose and carried the iron kettle to the sink |
in order to tap the current |
by turning the faucet to let it flow.


\Q2
Did it flow?

\Science
Yes.
From Roundwood reservoir in county Wicklow |
of a cubic capacity of 2400~million gallons, |
percolating through a subterranean aqueduct
of filter mains of single and double pipeage |
constructed at an initial plant cost
of 5~pounds per linear yard |
by way of the Dargle, Rathdown, Glen of the Downs and Callowhill
to the 26~acre reservoir at Stillorgan, |
a distance of 22~statute miles, |
and thence,
through a system of relieving tanks, |
by a gradient of 250~feet |
to the city boundary at Eustace bridge, upper Leeson street, |
\gab{170}
though from prolonged summer drouth
and daily supply of 12½~million gallons |
the water had fallen below the sill of the overflow weir |
for which reason |
the borough surveyor and waterworks engineer,
Mr~Spencer Harty,~C.E., |
on the instructions of the waterworks committee |
had prohibited the use of municipal water
for purposes other than those of consumption |
\paren{(envisaging the possibility of recourse being had
to the impotable water of the Grand and Royal canals
as in~1893)} |
particularly as the South Dublin Guardians, |
notwithstanding their ration of 15~gallons per day per pauper |
supplied through a 6~inch meter, |
had been convicted of a wastage of 20,000~gallons per night |
by a reading of their meter |
\gab{180}
on the affirmation of the law agent of the corporation,
Mr~Ignatius Rice, solicitor, |
thereby acting to the detriment of another section of the public, |
selfsupporting taxpayers, solvent, sound.

\Q1
What |
in water |
did Bloom, |
waterlover, drawer of water, watercarrier, |
returning to the range, |
admire?

\Household
Its universality:

\Philosophy
its democratic equality and constancy to its nature |
in seeking its own level:

\Places
its vastness |
in the ocean of Mercator's projection:

\Religious
its unplumbed profundity |
in the Sundam trench of the Pacific exceeding 8000~fathoms:

\Factual
the restlessness of its waves and surface particles |
visiting in turn all points of its seaboard:

\Memories
the independence of its units:
the variability of states of sea:

\Poetry
its hydrostatic quiescence |
in calm:
\gab{190}

\Science
its hydrokinetic turgidity |
in neap and spring tides:

\Household
its subsidence |
after devastation:

\Memories
its sterility |
in the circumpolar icecaps, arctic and antarctic:

\Factual
its climatic and commercial significance:

\Places
its preponderance |
of 3~to~1 over the dry land of the globe:

\Religious
its indisputable hegemony |
extending in square leagues
over all the region below the subequatorial tropic of Capricorn:

\Factual
the multisecular stability of its primeval basin:

\Science
its luteofulvous bed:

\Memories
its capacity |
to dissolve and hold in solution
all soluble substances |
including millions of tons of the most precious metals:

\Household
its slow erosions |
of peninsulas and islands, |

\Places
its persistent formation of homothetic islands,
peninsulas and downwardtending promontories:

\Poetry
its alluvial deposits:

\Philosophy
its weight and volume and density:
\gab{200}

\Factual
its imperturbability |
in lagoons and highland tarns:

\Religious
its gradation of colours |
in the torrid and temperate and frigid zones:

\Memories
its vehicular ramifications |
in continental lakecontained streams |
and confluent oceanflowing rivers with their tributaries |
and transoceanic currents, gulfstream, north and south equatorial courses:

\Science
its violence |
in seaquakes, waterspouts, Artesian wells, |
eruptions, torrents, eddies, freshets, |
spates, groundswells, watersheds, waterpartings, |
geysers, cataracts, whirlpools, maelstroms, |
inundations, deluges, cloudbursts:

\Places
its vast circumterrestrial ahorizontal curve:

\Household
its secrecy |
in springs and latent humidity, |
revealed by rhabdomantic or hygrometric instruments |
\gab{210}
and exemplified by the well
by the hole in the wall at Ashtown gate, |
saturation of air,
distillation of dew:

\Philosophy
the simplicity of its composition, |
two constituent parts of hydrogen |
with one constituent part of oxygen:

\Religious
its healing virtues:

\Memories
its buoyancy |
in the waters of the Dead Sea:

\Factual
its persevering penetrativeness |
in runnels, gullies, inadequate dams, leaks on shipboard:

\Poetry
its properties for cleansing, quenching thirst and fire, nourishing vegetation:

\Science
its infallibility |
as paradigm and paragon:

\Places
its metamorphoses |
as vapour, mist, cloud, rain, sleet, snow, hail:

\Household
its strength |
in rigid hydrants:

\Religious
its variety of forms |
in loughs and bays |
and gulfs and bights and guts and lagoons |
and atolls and archipelagos |
and sounds and fjords and minches |
and tidal estuaries and arms of sea:

\Science
its solidity |
in glaciers, icebergs, icefloes:
\gab{220}

\Memories
its docility |
in working hydraulic millwheels, |
turbines, dynamos, electric power stations, |
bleachworks, tanneries, scutchmills:

\Poetry
its utility |
in canals, |
rivers, if navigable, |
floating and graving docks:

\Places
its potentiality |
derivable from harnessed tides or watercourses |
falling from level to level:

\Philosophy
its submarine fauna and flora
\paren{(anacoustic, photophobe),} |
numerically, if not literally, |
the inhabitants of the globe:

\Household
its ubiquity |
as constituting 90~percent of the human body:

\Factual
the noxiousness of its effluvia |
in lacustrine marshes, pestilential fens,
faded flowerwater,
stagnant pools in the waning moon.


\acting{Bloom wash hands}

\Q1
Having set the half\/filled kettle on the now burning coals, |
why did he return to the stillflowing tap?%
\gab{230}

\Household
To wash his soiled hands |
with a partially consumed tablet of Barrington's lemonflavoured soap, |
to which paper still adhered, |
\paren{(bought thirteen hours previously for fourpence and still unpaid for),} |
in fresh cold neverchanging everchanging water |
and dry them,
face and hands, |
in a long redbordered holland cloth |
passed over a wooden revolving roller.


\acting{Stephen declines to wash hands}

\Q2
What reason did Stephen give |
for declining Bloom's offer?

\Philosophy
That he was hydrophobe, |
hating partial contact by immersion |
or total by submersion in cold water, |
\paren{(his last bath having taken place in the month of October of the preceding year),} |
disliking the aqueous substances of glass and crystal, |
\gab{240}
distrusting aquacities of thought and language.


\acting{Bloom inner thoughts}

\Q1
What impeded Bloom |
from giving Stephen counsels of hygiene and prophylactic |
to which should be added suggestions |
concerning a preliminary wetting of the head and contraction of the muscles |
with rapid splashing of the face and neck and thoracic and epigastric region |
in case of sea or river bathing, |
the parts of the human anatomy most sensitive to cold |
being the nape, stomach and thenar or sole of foot?

\Bloom
The incompatibility of aquacity |
with the erratic originality of genius.


\Q2
What additional didactic counsels |
did he similarly repress?

\Bloom
Dietary:
concerning the respective percentage of protein and caloric energy |
in bacon, salt ling and butter, |
\gab{250}
the absence of the former in the lastnamed |
and the abundance of the latter in the firstnamed.


\Q1
Which seemed to the host |
to be the predominant qualities of his guest?

\Bloom
Confidence in himself, |
an equal and opposite power of abandonment and recuperation.


\Q2
What concomitant phenomenon |
took place in the vessel of liquid |
by the agency of fire?

\Science
The phenomenon of ebullition.
Fanned by a constant updraught of ventilation |
between the kitchen and the chimneyflue, |
ignition was communicated
from the faggots of precombustible fuel |
to polyhedral masses of bituminous coal, |
\gab{260}
containing in compressed mineral form |
the foliated fossilised decidua of primeval forests |
which had in turn derived their vegetative existence from the sun, |
primal source of heat
\paren{(radiant),} |
transmitted through omnipresent luminiferous diathermanous ether.
Heat
\paren{(convected),} |
a mode of motion developed by such combustion, |
was constantly and increasingly conveyed |
from the source of calorification |
to the liquid contained in the vessel, |
being radiated through the uneven unpolished dark surface of the metal iron, |
in part reflected,
in part absorbed,
in part transmitted, |
gradually raising the temperature of the water |
from normal to boiling point,
a rise in temperature expressible |
as the result of an expenditure of 72~thermal units needed |
\gab{270}
to raise 1~pound of water |
from 50~degrees to 212~degrees Fahrenheit.


\Q1
What announced the accomplishment |
of this rise in temperature?

\Science
A~double falciform ejection of water vapour |
from under the kettlelid |
at both sides simultaneously.


\Q2
For what personal purpose |
could Bloom have applied the water |
so boiled?

\Factual
To shave himself.


\Q1
What advantages |
attended shaving by night?

\Memories
A~softer beard:
a softer brush |
if intentionally allowed
to remain from shave to shave in its agglutinated lather:
a softer skin |
if unexpectedly encountering female acquaintances
\gab{280}
in remote places at incustomary hours:
quiet reflections upon the course of the day:
a cleaner sensation |
when awaking after a fresher sleep |
since matutinal noises, premonitions and perturbations, |
a clattered milkcan, a postman's double knock,
a paper read, reread while lathering, relathering the same spot, |
a shock, a shoot,
with thought of aught he sought though fraught with nought |
might cause a faster rate of shaving and a nick |
on which incision |
plaster with precision |
cut and humected and applied |
adhered:
which was to be done.


\Q2
Why did absence of light |
disturb him less than presence of noise?

\Memories
Because of the surety of the sense of touch |
in his firm full masculine feminine passive active hand.%
\gab{290}


\Q1
What quality |
did it
\paren{(his hand)}
possess |
but with what counteracting influence?

\Memories
The operative surgical quality |
but that he was reluctant to shed human blood |
even when the end justified the means, |
preferring, in their natural order, |
heliotherapy, psychophysicotherapeutics, osteopathic surgery.


\acting{Bloom open dresser drawer}

\Q2
What lay under exposure |
on the lower, middle and upper shelves of the kitchen dresser, |
opened by Bloom?

\Household
On the lower shelf |
five vertical breakfast plates, six horizontal breakfast saucers |
on which rested inverted breakfast cups, |
a moustachecup, uninverted, |
and saucer of Crown Derby, |
\gab{300}
four white goldrimmed eggcups, |
an open shammy purse displaying coins, mostly copper, |
and a phial of aromatic
\paren{(violet)}
comfits.
On the middle shelf |
a chipped eggcup containing pepper, a drum of table salt, |
four conglomerated black olives in oleaginous paper, |
an empty pot of Plumtree's potted meat, |
an oval wicker basket bedded with fibre and containing one Jersey pear, |
a halfempty bottle of William Gilbey and Co's white invalid port, |
half disrobed of its swathe of coralpink tissue paper, |
a packet of Epps's soluble cocoa, |
five ounces of Anne Lynch's choice tea at 2/-~per lb in a crinkled leadpaper bag, |
a cylindrical canister containing the best crystallised lump sugar, |
two onions, |
one, the larger, Spanish, entire, |
\gab{310}
the other, smaller, Irish, bisected with augmented surface and more redolent, |
a jar of Irish Model Dairy's cream, |
a jug of brown crockery containing a naggin and a quarter
of soured adulterated milk, |
converted by heat into water, acidulous serum and semisolidified curds, |
which added to the quantity subtracted for Mr~Bloom's and Mrs~Fleming's breakfasts, |
made one imperial pint, |
the total quantity originally delivered, |
two cloves,
a halfpenny
and a small dish containing a slice of fresh ribsteak.
On the upper shelf |
a battery of jamjars
\paren{(empty)}
of various sizes and proveniences.


\acting{Bloom notice betting tickets}

\Q1
What attracted his attention |
lying on the apron of the dresser?

\Science
Four polygonal fragments of two lacerated scarlet betting tickets, |
\gab{320}
numbered 8~87, 88~6.


\Q2
What reminiscences |
temporarily corrugated his brow?

\Memories
Reminiscences of coincidences, |
truth stranger than fiction, |
preindicative of the result of the Gold Cup flat handicap, |
the official and definitive result of which he had read |
in the \Title{Evening Telegraph},
late pink edition, |
in the cabman's shelter, at Butt bridge.


\Q1
Where had previous intimations of the result, |
effected or projected, |
been received by him?

\Memories
In Bernard Kiernan's licensed premises |
8,~9 and~10~little Britain street: |
in David Byrne's licensed premises, 14~Duke street: |
\gab{330}
in O'Connell street lower, outside Graham Lemon's |
when a dark man had placed in his hand a throwaway |
\paren{(subsequently thrown away),} |
advertising Elijah, restorer of the church in Zion: |

\Places
in Lincoln place
outside the premises of F.W.~Sweny and~Co (Limited), dispensing chemists, |
when, |
when Frederick~M. (Bantam) Lyons
had rapidly and successively requested, perused and restituted
the copy of the current issue of the \Title{Freeman's Journal} and \Title{National Press}
which he had been about to throw away |
\paren{(subsequently thrown away),} |
he had proceeded towards the oriental edifice of the Turkish and Warm Baths,
11~Leinster street, |
with the light of inspiration shining in his countenance |
and bearing in his arms the secret of the race, |
\gab{340}
graven in the language of prediction.

\Q2
What qualifying considerations |
allayed his perturbations?

\Science
The difficulties of interpretation |
since the significance of any event
followed its occurrence |
as variably as the acoustic report followed the electrical discharge |
and of counterestimating against an actual loss |
by failure to interpret the total sum of possible losses |
proceeding originally from a successful interpretation.


\Q1
His mood?

\Bloom
He had not risked,
he did not expect,
he had not been disappointed,
he was satisfied.%
\gab{350}


\Q2
What satisfied him?

\Bloom
To have sustained no positive loss.
To have brought a positive gain to others.
Light to the gentiles.


\acting{Bloom enact as best he can}

\Q1
How did Bloom prepare a collation for a gentile?

\Science
He poured into two teacups |
two level spoonfuls, four in all, of Epps's soluble cocoa |
and proceeded |
according to the directions for use printed on the label, |
to each adding |
after sufficient time for infusion |
the prescribed ingredients for diffusion |
in the manner and in the quantity prescribed.


\Q2
What supererogatory marks of special hospitality |
did the host show his guest?%
\gab{360}

\Factual
Relinquishing his symposiarchal right
to the moustache cup of imitation Crown Derby |
presented to him by his only daughter,
Millicent
\paren{(Milly),} |
he substituted a cup identical with that of his guest |
and served extraordinarily to his guest |
and, in reduced measure, to himself |
the viscous cream ordinarily reserved
for the breakfast of his wife Marion
\paren{(Molly).}


\acting{both enact as best they can}

\Q1
Was the guest conscious of |
and did he acknowledge |
these marks of hospitality?

\Factual
His attention was directed to them by his host jocosely, |
and he accepted them seriously |
as they drank in jocoserious silence |
Epps's massproduct, |
the creature cocoa.%
\gab{370}


\Q2
Were there marks of hospitality |
which he contemplated but suppressed, |
reserving them for another |
and for himself on future occasions |
to complete the act begun?

\Bloom
The reparation of a fissure of the length of 1½ inches |
in the right side of his guest's jacket.
A~gift to his guest of one of the four lady's handkerchiefs, |
if and when ascertained to be in a presentable condition.


\acting{both drink cocoa}

\Q1
Who drank more quickly?

\Science \stage{racing}
Bloom,
having the advantage of ten seconds at the initiation |
and taking, |
from the concave surface of a spoon |
along the handle of which a steady flow of heat was conducted, |
\gab{380}
three sips to his opponent's one, |
six to two, |
nine to three.


\Q2
What cerebration accompanied his frequentative act?

\Philosophy
Concluding by inspection |
but erroneously |
that his silent companion was engaged in mental composition |
he reflected on the pleasures derived
from literature of instruction rather than of amusement |
as he himself had applied to the works of William Shakespeare more than once |
for the solution of difficult problems in imaginary or real life.


\Q1
Had he found their solution?

\Philosophy
In spite of careful and repeated reading of certain classical passages, |
aided by a glossary, |
\gab{390}
he had derived imperfect conviction from the text, |
the answers not bearing in all points.


\Q2
What lines concluded his first piece of original verse |
written by him, potential poet, |
at the age of~11 in~1877 |
on the occasion of the offering of three prizes |
of~10/-, 5/- and~2/6 respectively |
for competition |
by the \Title{Shamrock}, |
a weekly newspaper?

\Poetry

\begin{verse}
    An ambition to squint\\
    At my verses in print\\
    Makes me hope that for these you'll find room?.\\
    If you so condescend\\
    Then please place at the end%
\gab{400}\\
    The name of yours truly, L.~Bloom.
\end{verse}

\Q1
Did he find four separating forces |
between his temporary guest and him?

\Philosophy
Name, age, race, creed.


\Q2
What anagrams had he made on his name |
in youth?

\Poetry

\begin{verse}
    Leopold Bloom\\
    Ellpodbomool\\
    Molldopeloob\\
    Bollopedoom\\
    Old Ollebo, M.P.
\end{verse}


\Q1
What acrostic |
upon the abbreviation of his first name |
\gab{410}
had he
\paren{(kinetic poet)} |
sent to Miss Marion (Molly) Tweedy |
on the 14 February 1888?

\Poetry

\begin{verse}
    {\textbf{P}}oets oft have sung in rhyme\\
    {\textbf{O}}f music sweet their praise divine.\\
    {\textbf{L}}et them hymn it nine times nine.\\
    {\textbf{D}}earer far than song or wine.\\
    {\textbf{Y}}ou are mine. The world is mine.
\end{verse}


\Q2
What had prevented him from completing a topical song |
\paren{(music by R.G.~Johnston)} |
on the events of the past, |
or fixtures for the actual, |
years, |
entitled \Title{If Brian Boru could but come back and see old Dublin now}, |
commissioned by Michael Gunn,
\gab{420}
lessee of the Gaiety Theatre, 46,~47, 48,~49 South King street, |
and to be introduced into the sixth scene,
the valley of diamonds, |
of the second edition
\paren{(30~January 1893)}
of the grand annual Christmas pantomime \Title{Sinbad the Sailor} |
\paren{(produced by R.~Shelton 26~December~1892,
written by Greenleaf Whittier,
scenery by George~A. Jackson and Cecil Hicks,
costumes by Mrs~and~Miss Whelan
under the personal supervision of Mrs~Michael Gunn,
ballets by Jessie Noir,
harlequinade by Thomas Otto)} |
and sung by Nelly Bouverist, principal girl?

\Memories
Firstly, |
oscillation between events of imperial and of local interest, |
the anticipated diamond jubilee of Queen Victoria
\paren{(born 1820, acceded 1837)} |
and the posticipated opening of the new municipal fish market:%
\gab{430}

\Factual
Secondly, |
apprehension of opposition from extreme circles |
on the questions of the respective visits
of Their Royal Highnesses the duke and duchess of York
\paren{(real)} |
and of His Majesty King Brian Boru
\paren{(imaginary)}:

\Science
Thirdly, |
a conflict between professional etiquette
and professional emulation |
concerning the recent erections of the Grand Lyric Hall on Burgh Quay
and the Theatre Royal in Hawkins street:

\Philosophy
Fourthly, |
distraction resultant from compassion |
for Nelly Bouverist's non-intellectual, non-political, non-topical
expression of countenance and concupiscence |
caused by Nelly Bouverist's revelations of white articles |
of non-intellectual, non-political, non-topical underclothing |
\gab{440}
while she
\paren{(Nelly Bouverist)}
was in the articles:

\Religious
Fifthly, |
the difficulties of the selection of appropriate music and humorous allusions |
from \Title{Everybody's Book of Jokes} |
\paren{(1000 pages and a laugh in every one)}:

\Places
Sixthly, |
the rhymes, homophonous and cacophonous, |
associated with the names of the new lord mayor, Daniel Tallon, |
the new high sheriff, Thomas Pile |
and the new solicitorgeneral, Dunbar Plunket Barton.


\Q1
What relation |
existed between their ages?

\Science
16~years before in~1888 |
when Bloom was of Stephen's present age |
Stephen was~6.
16~years after in~1920 |
when Stephen would be of Bloom's present age |
Bloom would be~54.
In~1936 |
when Bloom would be~70 and Stephen~54 |
their ages |
initially in the ratio of 16~to~0 |
would be as 17½~to~13½, |
\gab{450}
the proportion increasing and the disparity diminishing |
according as arbitrary future years were added, |
for |
if the proportion existing in~1883 |
had continued immutable, |
conceiving that to be possible, |
till then~1904 |
when Stephen was~22 |
Bloom would be~374 |
and in~1920 |
when Stephen would be~38, |
as Bloom then was, |
Bloom would be~646 |
while in~1952 |
when Stephen would have attained the maximum postdiluvian age of~70 |
Bloom, being 1190~years alive |
having been born in the year~714, |
would have surpassed by 221~years the maximum antediluvian age, |
that of Methusalah, 969~years, |
while, if Stephen would continue to live |
until he would attain that age in the year 3072~A.D., |
Bloom would have been obliged to have been alive 83,300~years, |
\gab{460}
having been obliged to have been born in the year 81,396~B.C.


\Q2
What events |
might nullify these calculations?

\Philosophy
The cessation of existence of both or either, |
the inauguration of a new era or calendar, |
the annihilation of the world and consequent extermination of the human species, |
inevitable but impredictable.


\Q1
How many previous encounters |
proved their preexisting acquaintance?

\Memories
Two.
The first |
in the lilacgarden of Matthew Dillon's house, |
Medina Villa, Kimmage road, Roundtown, |
in~1887, |
in the company of Stephen's mother, |
Stephen being then of the age of~5 |
and reluctant to give his hand in salutation.
\gab{470}

\Places
The second |
in the coffeeroom of Breslin's hotel |
on a rainy Sunday in the January of~1892, |
in the company of Stephen's father and Stephen's granduncle, |
Stephen being then 5~years older.


\Q2
Did Bloom accept the invitation to dinner |
given then by the son and afterwards seconded by the father?

\Bloom
Very gratefully, |
with grateful appreciation, |
with sincere appreciative gratitude, |
in appreciatively grateful sincerity of regret, |
he declined.


\Q1
Did their conversation on the subject of these reminiscences |
reveal a third connecting link between them?

\Places
Mrs Riordan
\paren{(Dante),} |
a widow of independent means, |
had resided in the house of Stephen's parents |
\gab{480}
from 1~September 1888 to 29~December 1891 |
and had also resided during the years~1892, 1893 and~1894 |
in the City Arms Hotel |
owned by Elizabeth O'Dowd of 54~Prussia street |
where, during parts of the years~1893 and~1894, |
she had been a constant informant of Bloom |
who resided also in the same hotel, |
being at that time a clerk
in the employment of Joseph Cuffe of 5~Smithfield |
for the superintendence of sales |
in the adjacent Dublin Cattle market on the North Circular road.


\Q2
Had he performed |
any special corporal work of mercy for her?

\Memories
He had sometimes propelled her on warm summer evenings, |
an infirm widow of independent, if limited, means, |
in her convalescent bathchair |
with slow revolutions of its wheels |
\gab{490}
as far as the corner of the North Circular road
opposite Mr Gavin Low's place of business |
where she had remained for a certain time |
scanning through his onelensed binocular fieldglasses |
unrecognisable citizens on tramcars, |
roadster bicycles equipped with inflated pneumatic tyres, |
hackney carriages, tandems, private and hired landaus, dogcarts, ponytraps and brakes |
passing from the city to the Phoenix Park |
and vice versa.


\Q1
Why could he then support |
that his vigil |
with the greater equanimity?

\Memories
Because in middle youth |
he had often sat observing |
through a rondel of bossed glass of a multicoloured pane |
the spectacle offered with continual changes of the thoroughfare without, |
\gab{500}
pedestrians, quadrupeds, velocipedes, vehicles, |
passing slowly, quickly, evenly, round and round and round |
the rim of a round and round precipitous globe.


\Q2
What distinct different memories had each of her |
now eight years deceased?

\Bloom
The older, |
her bezique cards and counters, |
her Skye terrier, |
her suppositious wealth, |
her lapses of responsiveness and incipient catarrhal deafness:

\Stephen
the younger, |
her lamp of colza oil before the statue of the Immaculate Conception, |
her green and maroon brushes for Charles Stewart Parnell and for Michael Davitt, |
her tissue papers.


\Q1
Were there no means still remaining to him |
to achieve the rejuvenation which these reminiscences |
divulged to a younger companion |
\gab{510}
rendered the more desirable?

\Science
The indoor exercises, |
formerly intermittently practised, subsequently abandoned, |
prescribed in Eugen Sandow's
\Title{Physical Strength and How to Obtain It} |
which, designed particularly for commercial men engaged in sedentary occupations, |
were to be made with mental concentration
in front of a mirror |
so as to bring into play the various families of muscles |
and produce successively |
a pleasant rigidity, |
a more pleasant relaxation |
and the most pleasant repristination of juvenile agility.


\Q2
Had any special agility been his |
in earlier youth?

\Memories
Though ringweight lifting had been beyond his strength |
\gab{520}
and the full circle gyration beyond his courage |
yet as a High school scholar |
he had excelled in his stable and protracted execution
of the half lever movement on the parallel bars |
in consequence of his abnormally developed abdominal muscles.


\Q1
Did either openly allude |
to their racial difference?

\Factual
Neither.


\acting{maybe have everyone look at who ``he'' is}

\Q2
What, |
reduced to their simplest reciprocal form, |
were Bloom's thoughts about Stephen's thoughts about Bloom |
and about Stephen's thoughts about Bloom's thoughts about Stephen?

\Philosophy
He thought that he thought that he was a jew |
\gab{530}
whereas he knew that he knew |
that he knew that he was not.


\Q1
What, |
the enclosures of reticence removed, |
were their respective parentages?

\Factual[~(legal)]
Bloom, |
only born male transubstantial heir of Rudolf Virag
\paren{(subsequently Rudolph Bloom)} |
of Szombathely, Vienna, Budapest, Milan, London and Dublin |
and of Ellen Higgins, |
second daughter of Julius Higgins
\paren{(born Karoly)}
and Fanny Higgins
\paren{(born Hegarty).}
Stephen, eldest surviving male consubstantial heir |
of Simon Dedalus of Cork and Dublin |
and of Mary, daughter of Richard and Christina Goulding
\paren{(born Grier).}


\Q2
Had Bloom and Stephen been baptised, |
and where and by whom,
\gab{540}
cleric or layman?

\Religious
Bloom
\paren{(three times),} |
by the reverend Mr Gilmer Johnston~M.A., alone, |
in the protestant church of Saint Nicholas Without, Coombe, |
by James O'Connor, Philip Gilligan and James Fitzpatrick, together, |
under a pump in the village of Swords, |
and by the reverend Charles Malone~C.C., |
in the church of the Three Patrons, Rathgar.
Stephen
\paren{(once)} |
by the reverend Charles Malone~C.C., alone, |
in the church of the Three Patrons, Rathgar.


\Q1
Did they find their educational careers similar?

\Factual
Substituting Stephen for Bloom |
Stoom would have passed successively through a dame's school and the high school.
\gab{550}
Substituting Bloom for Stephen |
Blephen would have passed successively
through the preparatory, junior, middle and senior grades of the intermediate |
and through the matriculation, first arts, second arts and arts degree courses
of the royal university.


\Q2
Why did Bloom refrain from stating |
that he had frequented the university of life?

\Memories
Because of his fluctuating incertitude |
as to whether this observation |
had or had not been already made by him to Stephen |
or by Stephen to him.


\Q1
What two temperaments |
did they individually represent?

\Bloom
The scientific.
\gab{560}

\Stephen
The artistic.


\Q2
What proofs did Bloom adduce |
to prove that his tendency was towards applied, |
rather than towards pure, |
science?

\Science
Certain possible inventions |
of which he had cogitated when reclining
in a state of supine repletion to aid digestion, |
stimulated by his appreciation of the importance
of inventions now common but once revolutionary, |
for example, |
the aeronautic parachute,
the reflecting telescope,
the spiral corkscrew,  |
the safety pin,
the mineral water siphon,
the canal lock with winch and sluice, |
the suction pump.


\Q1
Were these inventions principally intended |
for an improved scheme of kindergarten?%
\gab{570}

\Science
Yes, |
rendering obsolete |
popguns, elastic airbladders, games of hazard, catapults.
They comprised astronomical kaleidoscopes |
exhibiting the twelve constellations of the zodiac from Aries to Pisces, |
miniature mechanical orreries, |
arithmetical gelatine lozenges, |
geometrical to correspond with zoological biscuits, |
globemap playing balls, |
historically costumed dolls.


\Q2
What also stimulated him |
in his cogitations?

\Adverts
The financial success achieved by Ephraim Marks and Charles~A. James, |
the former by his 1d~bazaar at 42~George's street, south, |
the latter at his 6½d shop and world's fancy fair and waxwork exhibition |
at 30~Henry street, admission~2d, children~1d:
\gab{580}
and the infinite possibilities |
hitherto unexploited |
of the modern art of advertisement |
if condensed in triliteral monoideal symbols, |
vertically of maximum visibility
\paren{(divined),} |
horizontally of maximum legibility
\paren{(deciphered)} |
and of magnetising efficacy to arrest involuntary attention, |
to interest,
to convince,
to decide.


\Q1
Such as?

\Adverts
K.~11. Kino's 11/-~Trousers.\\
House of Keys. Alexander~J. Keyes.


\Q1
Such as not?

\Adverts
Look at this long candle.
Calculate when it burns out |
and you receive gratis 1~pair of our special non-compo boots, |
\gab{590}
guaranteed 1~candle power.
Address:
Barclay and Cook, 18~Talbot street.\\
Bacilikil
\paren{(Insect Powder).}\\
Veribest
\paren{(Boot Blacking).}\\
Uwantit
\paren{(Combined pocket twoblade penknife |
with corkscrew, nailfile and pipecleaner).}


\Q1
Such as never?

\Adverts
What is home |
without Plumtree's Potted Meat?\\
Incomplete.\\
With it |
an abode of bliss.\\
Manufactured by George Plumtree, 23~Merchants' quay, Dublin, |
\gab{600}
put up in 4~oz pots, |
and inserted by
Councillor Joseph~P. Nannetti,~M.P., Rotunda Ward, 19~Hardwicke street, |
under the obituary notices and anniversaries of deceases.
The name on the label is Plumtree.
A~plumtree in a meatpot,
registered trade mark.
Beware of imitations.
Peatmot.
Trumplee.
Moutpat.
Plamtroo.


\Q2
Which example |
did he adduce |
to induce Stephen |
to deduce that originality, |
though producing its own reward, |
does not invariably conduce to success?

\Memories
His own ideated and rejected project of an illuminated showcart, |
drawn by a beast of burden, |
in which two smartly dressed girls were to be seated |
engaged in writing.%
\gab{610}


\acting{slowly like screenplay}

\Q1
What suggested scene |
was then constructed by Stephen?

\Stephen
Solitary hotel in mountain pass.
Autumn.
Twilight.
Fire lit.
In dark corner young man seated.
Young woman enters.
Restless.
Solitary.
She sits.
She goes to window.
She stands.
She sits.
Twilight.
She thinks.
On solitary hotel paper she writes.
She thinks.
She writes.
She sighs.
Wheels and hoofs.
She hurries out.
He comes from his dark corner.
He seizes solitary paper.
He holds it towards fire.
Twilight.
He reads.
Solitary.


\acting{slowly}

\Q2
What?

\Stephen
In sloping, upright and backhands:
Queen's Hotel, Queen's Hotel, Queen's Hotel.
\gab{620}
Queen's Ho…


\acting{sadly}

\Q1
What suggested scene |
was then reconstructed by Bloom?

\Memories
The Queen's Hotel, Ennis, county Clare, |
where Rudolph Bloom
\paren{(Rudolf Virag)}
died |
on the evening of the 27~June 1886,
at some hour unstated, |
in consequence of an overdose of monkshood
\paren{(aconite)}
selfadministered |
in the form of a neuralgic liniment |
composed of 2~parts of aconite liniment
to 1~of chloroform liniment |
\paren{(purchased by him at 10.20~a.m.\ on the morning of 27~June 1886 |
at the medical hall of Francis Dennehy, 17~Church street, Ennis)} |
after having, though not in consequence of having, |
purchased at 3.15~p.m.\ on the afternoon of 27~June 1886 |
a new boater straw hat, extra smart |
\paren{(after having, though not in consequence of having, |
\gab{630}
purchased at the hour and in the place aforesaid, |
the toxin aforesaid),} |
at the general drapery store of
James Cullen, 4~Main street, Ennis.


\Q2
Did he attribute this homonymity |
to information |
or coincidence |
or intuition?

\Factual
Coincidence.


\Q1
Did he depict the scene verbally |
for his guest to see?

\Philosophy
He preferred himself to see another's face |
and listen to another's words |
by which potential narration was realised |
and kinetic temperament relieved.


\acting{but talking of future}

\Q2
Did he see only a second coincidence |
in the second scene narrated to him, |
described by the narrator as \Title{A Pisgah Sight of Palestine} |
\gab{640}
or \Title{The Parable of the Plums}?

\Memories
It, with the preceding scene |
and with others unnarrated but existent by implication, |
to which add essays on various subjects or moral apothegms |
\paren{(e.g.\ \Title{My Favourite Hero} or \Title{Procrastination is the Thief of Time})} |
composed during schoolyears, |
seemed to him to contain in itself |
and in conjunction with the personal equation |
certain possibilities of financial, social, personal and sexual success, |
whether specially collected and selected as model pedagogic themes |
\paren{(of cent per cent merit)} |
for the use of preparatory and junior grade students |
or contributed in printed form, |
following the precedent of Philip Beaufoy |
or Doctor Dick |
or Heblon's \Title{Studies in Blue}, |
\gab{650}
to a publication of certified circulation and solvency |
or employed verbally as intellectual stimulation for sympathetic auditors, |
tacitly appreciative of successful narrative |
and confidently augurative of successful achievement, |
during the increasingly longer nights |
gradually following the summer solstice on the day but three following, |
\latin{videlicet}, Tuesday, 21~June
\paren{(S.~Aloysius Gonzaga),} |
sunrise 3.33~a.m., sunset 8.29~p.m.


\Q1
Which domestic problem |
as much as, if not more than, any other |
frequently engaged his mind?

\Philosophy
What to do with our wives.


\Q2
What had been |
his hypothetical singular solutions?%
\gab{660}

\Factual
Parlour games |
\paren{(dominos, halma, tiddledywinks, spilikins, cup and ball, |
nap, spoil five, bezique, twentyfive, beggar my neighbour, |
draughts, chess or backgammon)}:

\Science
embroidery, darning or knitting for the policeaided clothing society:

\Religious
musical duets, mandoline and guitar, piano and flute, guitar and piano:

\Philosophy
legal scrivenery or envelope addressing:

\Poetry
biweekly visits to variety entertainments:

\Places
commercial activity as pleasantly commanding and pleasingly obeyed mistress |
proprietress in a cool dairy shop or warm cigar divan:

\Household
the clandestine satisfaction of erotic irritation in masculine brothels, |
state inspected and medically controlled:

\Memories
social visits,
at regular infrequent prevented intervals |
\gab{670}
and with regular frequent preventive superintendence, |
to and from female acquaintances of recognised respectability in the vicinity:

\Poetry
courses of evening instruction |
specially designed to render liberal instruction agreeable.


\Q1
What instances of deficient mental development in his wife |
inclined him in favour of the lastmentioned
\paren{(ninth)}
solution?

\Household
In disoccupied moments |
she had more than once |
covered a sheet of paper with signs and hieroglyphics |
which she stated were Greek and Irish and Hebrew characters.

\Places
She had interrogated constantly at varying intervals |
as to the correct method of writing the capital initial
of the name of a city in Canada, Quebec.
\gab{680}

\Philosophy
She understood little of political complications, internal, |
or balance of power, external.

\Memories
In calculating the addenda of bills |
she frequently had recourse to digital aid.

\Religious
After completion of laconic epistolary compositions |
she abandoned the implement of calligraphy in the encaustic pigment, |
exposed to the corrosive action of copperas, green vitriol and nutgall.

\Factual
Unusual polysyllables of foreign origin |
she interpreted phonetically or by false analogy or by both:

\Science
\latin{metempsychosis}
\paren{(met him pike hoses),} |

\Poetry
\latin{alias} \stage{a-lias}
\paren{(a mendacious person mentioned in sacred scripture).}


\Q2
What compensated |
in the false balance of her intelligence |
for these and such deficiencies of judgment |
regarding persons, places and things?

\Philosophy
The false apparent parallelism
\gab{690}
of all perpendicular arms of all balances, |
proved true by construction. |
The counterbalance of her proficiency of judgment regarding one person, |
proved true by experiment.


\Q1
How had he attempted to remedy |
this state of comparative ignorance?

\Factual
Variously.
By leaving in a conspicuous place |
a certain book open at a certain page:
by assuming in her,
when alluding explanatorily,
latent knowledge:
by open ridicule in her presence |
of some absent other's ignorant lapse.


\Q2
With what success |
had he attempted direct instruction?

\Memories
She followed not all, |
a part of the whole, |
gave attention with interest, |
comprehended with surprise, |
\gab{700}
with care repeated, |
with greater difficulty remembered, |
forgot with ease, |
with misgiving reremembered, |
rerepeated with error.


\Q1
What system had proved more effective?

\Philosophy
Indirect suggestion |
implicating self\/interest.


\Q2
Example?

\Philosophy
She disliked umbrella with rain, |
he liked woman with umbrella, |
she disliked new hat with rain, |
he liked woman with new hat, |
he bought new hat with rain, |
she carried umbrella with new hat.


\Q1
Accepting the analogy implied in his guest's parable |
which examples of postexilic eminence |
\gab{710}
did he adduce?

\Philosophy
Three seekers of the pure truth, |
Moses of Egypt, |
Moses Maimonides, author of \Title{More Nebukim}
\paren{(Guide of the Perplexed)} |
and Moses Mendelssohn of such eminence |
that from Moses
\paren{(of Egypt)}
to Moses
\paren{(Mendelssohn)} |
there arose none like Moses
\paren{(Maimonides).}


\Q2
What statement was made, |
under correction, |
by Bloom |
concerning a fourth seeker of pure truth, |
by name Aristotle, |
mentioned, |
with permission, |
by Stephen?

\Philosophy
That the seeker mentioned |
had been a pupil of a rabbinical philosopher, |
name uncertain.


\Q1
Were other anapocryphal illustrious sons of the law
\gab{720}
and children of a selected or rejected race |
mentioned?

\Philosophy
Felix Bartholdy Mendelssohn
\paren{(composer),} |
Baruch Spinoza
\paren{(philosopher),} |
Mendoza
\paren{(pugilist),} |
Ferdinand Lassalle
\paren{(reformer, duellist).}


\Q2
What fragments of verse |
from the ancient Hebrew and ancient Irish languages |
were cited with modulations of voice and translation of texts |
by guest to host |
and by host to guest?

\Stephen
By Stephen:
\irish{suil, suil, suil arun, suil go siocair agus suil go cuin} |
\paren{(walk, walk, walk your way, walk in safety, walk with care).}

\Bloom
By Bloom:
\hebrew{kifeloch, harimon rakatejch m'baad l'zamatejch} |
\paren{(thy temple amid thy hair
\gab{730}
is as a slice of pomegranate).}


\Q1
How was a glyphic comparison of the phonic symbols of both languages |
made in substantiation of the oral comparison?

\Philosophy
By juxtaposition.
On the penultimate blank page of a book of inferior literary style, |
entituled \Title{Sweets of Sin} |
\paren{(produced by Bloom and so manipulated
that its front cover came in contact with the surface of the table)} |
with a pencil
\paren{(supplied by Stephen)}
Stephen wrote the Irish characters
for \irish{gee}, \irish{eh}, \irish{dee}, \irish{em}, simple and modified, |
and Bloom in turn
wrote the Hebrew characters \hebrew{ghimel}, \hebrew{aleph}, \hebrew{daleth}
and
\paren{(in the absence of \hebrew{mem})} |
a substituted \hebrew{qoph}, |
explaining their arithmetical values as ordinal and cardinal numbers, |
\gab{740}
\latin{videlicet} 3,~1, 4, and~100.


\Q2
Was the knowledge possessed by both |
of each of these languages, |
the extinct and the revived, |
theoretical or practical?

\Philosophy
Theoretical, |
being confined to certain grammatical rules of accidence and syntax |
and practically |
excluding vocabulary.


\Q1
What points of contact |
existed between these languages |
and between the peoples who spoke them?

\Religious
The presence of guttural sounds, diacritic aspirations,
epenthetic and servile letters in both languages:
their antiquity, |
both having been taught on the plain of Shinar |
242~years after the deluge |
in the seminary instituted by Fenius Farsaigh, |
descendant of Noah, progenitor of Israel, |
\gab{750}
and ascendant of Heber and Heremon, progenitors of Ireland:
their archaeological, genealogical, hagiographical, |
exegetical, homiletic, toponomastic, historical and religious literatures |
comprising the works of rabbis and culdees, |
Torah, Talmud
\paren{(Mischna and Ghemara),}
Massor, Pentateuch, |
Book of the Dun Cow, Book of Ballymote, Garland of Howth, Book of Kells:
their dispersal, persecution, survival and revival:
the isolation of their synagogical and ecclesiastical rites |
in ghetto
\paren{(S.~Mary's Abbey)} |
and masshouse
\paren{(Adam and Eve's tavern)}:
the proscription of their national costumes |
in penal laws and jewish dress acts:
the restoration in Chanah David of Zion |
and the possibility of
\gab{760}
Irish political autonomy or devolution.


\Q2
What anthem |
did Bloom chant partially |
in anticipation of that multiple, ethnically irreducible consummation?

\Bloom
     \hebrew{Kolod balejwaw pnimah\\
     Nefesch, jehudi, homijah.}


\Q1
Why was the chant arrested |
at the conclusion of this first distich?

\Factual
In consequence of defective mnemotechnic.


\Q2
How did the chanter compensate for this deficiency?

\Factual
By a periphrastic version of the general text.


\Q1
In what common study |
did their mutual reflections merge?

\Science
The increasing simplification
\gab{770}
traceable from the Egyptian epigraphic hieroglyphs |
to the Greek and Roman alphabets |
and the anticipation of modern stenography and telegraphic code |
in the cuneiform inscriptions
\paren{(Semitic)} |
and the virgular quinquecostate ogham writing
\paren{(Celtic).}


\acting{stephen signs name}

\Q2
Did the guest comply with his host's request?

\Factual
Doubly, by appending his signature in Irish and Roman characters.


\Q1
What was Stephen's auditive sensation?

\Memories
He heard in a profound ancient male unfamiliar melody |
the accumulation of the past.


\Q2
What was Bloom's visual sensation?

\Factual
He saw in a quick young male familiar form |
\gab{780}
the predestination of a future.


\Q1
What were Stephen's and Bloom's quasisimultaneous volitional quasisensations
of concealed identities?

\Factual
Visually, Stephen's:
The traditional figure of hypostasis, |
depicted by Johannes Damascenus, Lentulus Romanus and Epiphanius Monachus |
as leucodermic, sesquipedalian with winedark hair.
Auditively, Bloom's:
The traditional accent of the ecstasy of catastrophe.


\Q2
What future careers had been possible for Bloom |
in the past and with what exemplars?

\Religious
In the church, Roman, Anglican or Nonconformist:
exemplars, |
the very reverend John Conmee~S.J., |
\gab{790}
the reverend T.~Salmon,~D.D., provost of Trinity college, |
Dr Alexander~J. Dowie.

\Factual
At the bar, English or Irish:
exemplars, |
Seymour Bushe,~K.C., |
Rufus Isaacs,~K.C. |

\Poetry
On the stage |
modern or Shakespearean:
exemplars, |
Charles Wyndham, high comedian |
Osmond Tearle
\paren{(died~1901),}
exponent of Shakespeare.


\Q1
Did the host encourage his guest to chant in a modulated voice |
a strange legend on an allied theme?

\Science
Reassuringly, their place, |
where none could hear them talk, being secluded, |
reassured, the decocted beverages, |
allowing for subsolid residual sediment of a mechanical mixture, |
water plus sugar plus cream plus cocoa, |
having been consumed.%
\gab{800}


\Q2
Recite the first (major) part |
of this chanted legend.

\Poetry

\begin{verse}
    Little Harry Hughes and his schoolfellows all\\
    Went out for to play ball.\\
    And the very first ball little Harry Hughes played\\
    He drove it o'er the jew's garden wall.\\
    And the very second ball little Harry Hughes played\\
    He broke the jew's windows all.
\end{verse}


\Q1
How did the son of Rudolph receive this first part?

\Factual
With unmixed feeling.
\gab{810}
Smiling, a jew |
he heard with pleasure and saw the
unbroken kitchen window.


\Q2
Recite the second part (minor) |
of the legend.

\Poetry

\begin{verse}
    Then out there came the jew's daughter\\
    And she all dressed in green.\\
    ``Come back, come back, you pretty little boy,\\
    And play your ball again.''

    ``I can't come back and I won't come back\\
    Without my schoolfellows all.\\
    For if my master he did hear\\
    He'd make it a sorry ball.''%
\gab{820}

    She took him by the lilywhite hand\\
    And led him along the hall\\
    Until she led him to a room\\
    Where none could hear him call.

    She took a penknife out of her pocket\\
    And cut off his little head.\\
    And now he'll play his ball no more\\
    For he lies among the dead.
\end{verse}


\Q1
How did the father of Millicent receive this second part?

\Factual
With mixed feelings.
\gab{830}
Unsmiling, he heard and saw with wonder |
a jew's daughter, all dressed in green.


\Q2
Condense Stephen's commentary.

\Stephen
One of all, the least of all, |
is the victim predestined.
Once by inadvertence |
twice by design |
he challenges his destiny.
It comes when he is abandoned |
and challenges him reluctant |
and, as an apparition of hope and youth, |
holds him unresisting.
It leads him to a strange habitation, |
to a secret infidel apartment, |
and there, implacable, |
immolates him, consenting.


\Q1
Why was the host
\paren{(victim predestined)} |
sad?

\Bloom
He wished that a tale of a deed should be told of a deed not by him |
should by him not be told.%
\gab{840}


\Q2
Why was the host
\paren{(reluctant, unresisting)} |
still?

\Science
In accordance with the law of the conservation of energy.


\Q1
Why was the host
\paren{(secret infidel)} |
silent?

\Factual
He weighed the possible evidences for and against ritual murder:

\Poetry
the incitations of the hierarchy, |
the superstition of the populace, |

\Memories
the propagation of rumour in continued fraction of veridicity, |
the envy of opulence, |
the influence of retaliation, |

\Household
the sporadic reappearance of atavistic delinquency, |
the mitigating circumstances of fanaticism, |
hypnotic suggestion |
and somnambulism.


\Q2
From which
\paren{(if any)}
of these mental or physical disorders |
\gab{850}
was he not totally immune?

\Science
From hypnotic suggestion:
once, waking, |
he had not recognised his sleeping apartment:
more than once, waking, |
he had been for an indefinite time incapable of moving or uttering sounds.
From somnambulism:
once, sleeping, |
his body had risen, crouched and crawled in the direction of a heatless fire |
and, having attained its destination, |
there, curled, unheated, in night attire had lain, |
sleeping.


\Q1
Had this latter or any cognate phenomenon |
declared itself in any member of his family?

\Memories
Twice, in Holles street and in Ontario terrace, |
\gab{860}
his daughter Millicent (Milly) |
at the ages of~6 and 8~years |
had uttered in sleep an exclamation of terror |
and had replied to the interrogations |
of two figures in night attire |
with a vacant mute expression.


\Q2
What other infantile memories |
had he of her?

\Memories
15~June 1889.
A~querulous newborn female infant |
crying to cause and lessen congestion.
A~child renamed Padney Socks |
she shook with shocks her moneybox:
counted his three free moneypenny buttons, |
one, tloo, tlee:
a doll, a boy, a sailor she cast away:
blond, born of two dark, she had blond ancestry, |
remote, a violation, Herr Hauptmann Hainau, Austrian army, |
proximate, a hallucination,
\gab{870}
lieutenant Mulvey, British navy.


\Q1
What endemic characteristics were present?

\Science
Conversely |
the nasal and frontal formation was derived |
in a direct line of lineage |
which, though interrupted, |
would continue at distant intervals |
to more distant intervals |
to its most distant intervals.


\Q2
What memories had he |
of her adolescence?

\Factual[~(legal)]
She relegated her hoop and skippingrope to a recess.
On the duke's lawn, |
entreated by an English visitor, |
she declined to permit him to make and take away her photographic image |
\paren{(objection not stated).}
On the South Circular road |
in the company of Elsa Potter, |
followed by an individual of sinister aspect, |
\gab{880}
she went half way down Stamer street and turned abruptly back |
\paren{(reason of change not stated).}
On the vigil of the \nth{15}~anniversary of her birth |
she wrote a letter from Mullingar, county Westmeath, |
making a brief allusion to a local student |
\paren{(faculty and year not stated).}


\Q1
Did that first division,
portending a second division, |
afflict him?

\Philosophy
Less than he had imagined,
more than he had hoped.


\Q2
What second departure |
was contemporaneously perceived by him |
similarly, if differently?

\Factual
A~temporary departure of his cat.


\Q1
Why similarly,
why differently?

\Philosophy
Similarly, because actuated by a secret purpose |
\gab{890}
the quest of a new male
\paren{(Mullingar student)} |
or of a healing herb
\paren{(valerian).}
Differently, |
because of different possible returns to the inhabitants or to the habitation.


\Q2
In other respects |
were their differences similar?

\Philosophy
In passivity, in economy, in the instinct of tradition, in unexpectedness.


\Q1
As?

\Philosophy
Inasmuch as leaning |
she sustained her blond hair for him |
to ribbon it for her |
\paren{(cf.\ \stage{compare} neckarching cat).}
Moreover, on the free surface of the lake in Stephen's green |
amid inverted reflections of trees |
her uncommented spit,
describing concentric circles of waterrings, |
indicated by the constancy of its permanence |
the locus of a somnolent prostrate fish |
\gab{900}
\paren{(cf.\ mousewatching cat).}
Again, |
in order to remember the date, combatants,
issue and consequences of a famous military engagement |
she pulled a plait of her hair |
\paren{(cf.\ earwashing cat).}
Furthermore, silly Milly, |
she dreamed of having had an unspoken unremembered conversation |
with a horse whose name had been Joseph |
to whom
\paren{(which)}
she had offered a tumblerful of lemonade |
which it
\paren{(he)}
had appeared to have accepted |
\paren{(cf.\ hearthdreaming cat).}
Hence, in passivity, in economy, in the instinct of tradition, in unexpectedness, |
their differences were similar.


\Q2
In what way |
had he utilised gifts |
(1)~an owl, |
(2)~a clock, |
given as matrimonial auguries, |
\gab{910}
to interest and to instruct her?

\Science
As object lessons to explain:
1)~the nature and habits of oviparous animals, |
the possibility of aerial flight, |
certain abnormalities of vision, |
the secular process of imbalsamation:
2)~the principle of the pendulum, |
exemplified in bob, wheelgear and regulator, |
the translation in terms of human or social regulation |
of the various positions of clockwise moveable indicators on an unmoving dial, |
the exactitude of the recurrence per hour |
of an instant in each hour when the longer and the
shorter indicator were at the same angle of inclination, |
\latin{videlicet},
$5\frac{5}{11}$~minutes past each hour per hour |
in arithmetical progression.


\Q1
In what manners |
\gab{920}
did she reciprocate?

\Memories
She remembered:
on the \nth{27}~anniversary of his birth |
she presented to him |
a breakfast moustachecup of imitation Crown Derby porcelain ware.

\Household
She provided:
at quarter day or thereabouts |
if or when purchases had been made by him not for her |
she showed herself attentive to his necessities, anticipating his desires.

\Poetry
She admired:
a natural phenomenon having been explained by him to her |
she expressed the immediate desire to possess |
without gradual acquisition |
a fraction of his science, |
the moiety, the quarter, a thousandth part.


\acting{Bloom invite, pointing upstairs}

\Q2
What proposal |
did Bloom, diambulist, |
father of Milly, somnambulist, |
make to Stephen,
\gab{930}
noctambulist?

\Factual[~(legal)]
To pass in repose |
the hours intervening
between Thursday
\paren{(proper)}
and Friday
\paren{(normal)} |
on an extemporised cubicle
in the apartment immediately above the kitchen |
and immediately adjacent
to the sleeping apartment of his host and hostess.


\acting{both look upstairs}

\Q1
What various advantages |
would or might have resulted from |
a prolongation of such an extemporisation?

\Stephen
For the guest:
security of domicile and seclusion of study.

\Bloom
For the host:
rejuvenation of intelligence,
vicarious satisfaction.

\Molly
For the hostess:
disintegration of obsession,
acquisition of correct Italian pronunciation.


\Q2
Why might these several provisional contingencies |
\gab{940}
between a guest and a hostess |
not necessarily preclude or be precluded by |
a permanent eventuality of reconciliatory union
between a schoolfellow and a jew's daughter?

\Religious
Because the way to daughter led through mother,
the way to mother through daughter.


\acting{Bloom question Stephen, who shakes head}

\Q1
To what inconsequent polysyllabic question of his host |
did the guest return a monosyllabic negative answer?

\Places
If he had known the late Mrs Emily Sinico, |
accidentally killed at Sydney Parade railway station, 14 October 1903.


\Q2
What inchoate corollary statement |
\gab{950}
was consequently suppressed by the host?

\Factual[~(legal)]
A~statement explanatory of his absence
on the occasion of the interment
of Mrs Mary Dedalus
\paren{(born Goulding),}
26 June 1903, |
vigil of the anniversary of the decease of Rudolph Bloom
\paren{(born Virag).}


\acting{Stephen shake head \emph{no} again}

\Q1
Was the proposal of asylum |
accepted?

\Factual
Promptly, inexplicably, with amicability, gratefully |
it was declined.


\Q2
What exchange of money |
took place between host and guest?

\Factual
The former returned to the latter, without interest, |
a sum of money,
%(£1-7-0),
one pound seven shillings sterling, |
advanced by the latter to the former.


\acting{Look as if proposing and agreeing all this}

\Q1
What counterproposals |
\gab{960}
were alternately advanced, accepted, modified, declined, |
restated in other terms, reaccepted, ratified, reconfirmed?

\Places
To inaugurate a prearranged course of Italian instruction, |
place the residence of the instructed.
To inaugurate a course of vocal instruction, |
place the residence of the instructress.
To inaugurate a series of static semistatic and peripatetic intellectual dialogues, |
places the residence of both speakers |
\paren{(if both speakers were resident in the same place),}

\Poetry
the Ship hotel and tavern, 6~Lower Abbey street
\paren{(W.~and~E. Connery, proprietors),} |
the National Library of Ireland, 10~Kildare street, |
the National Maternity Hospital, 29,~30 and 31~Holles street, |
a public garden, |
the vicinity of a place of worship, |
a conjunction of two or more public thoroughfares, |
\gab{970}
the point of bisection of a right line drawn between their residences |
\paren{(if both speakers were resident in different places).}


\Q2
What rendered problematic for Bloom |
the realisation of these mutually selfexcluding propositions?

\Memories
The irreparability of the past:
once |
at a performance of Albert Hengler's circus in the Rotunda, Rutland square, Dublin, |
an intuitive particoloured clown in quest of paternity |
had penetrated from the ring to a place in the auditorium |
where Bloom, solitary, was seated |
and had publicly declared to an exhilarated audience |
that he
\paren{(Bloom)}
was his
\paren{(the clown's)}
papa.
The imprevidibility of the future:
once in the summer of 1898 |
\gab{980}
he
\paren{(Bloom)}
had marked a florin~\paren{(2/-)}
with three notches on the milled edge |
and tendered it in payment of an account |
due to and received by |
J.~and~T. Davy, family grocers, 1~Charlemont Mall, Grand Canal, |
for circulation on the waters of civic finance, |
for possible, circuitous or direct, return.


\Q1
Was the clown |
Bloom's son?

\Factual
No.


\Q2
Had Bloom's coin returned?

\Factual
Never.


\Q1
Why would a recurrent frustration |
the more depress him?

\Philosophy
Because at the critical turningpoint of human existence |
\gab{990}
he desired to amend many social conditions, |
the product of inequality and avarice and international animosity.


\Q2
He believed then |
that human life was infinitely perfectible, |
eliminating these conditions?

\Philosophy
There remained the generic conditions imposed by natural, |
as distinct from human law, |
as integral parts of the human whole:

\Memories
the necessity of destruction to procure alimentary sustenance:

\Places
the painful character of the ultimate functions of separate existence,
the agonies of birth and death:

\Household
the monotonous menstruation of simian and
\paren{(particularly)}
human females |
extending from the age of puberty to the menopause:
\gab{1000}

\Factual
inevitable accidents at sea, in mines and factories:

\Science
certain very painful maladies and their resultant surgical operations, |
innate lunacy and congenital criminality, |
decimating epidemics:

\Poetry
catastrophic cataclysms which make terror the basis of human mentality:

\Religious
seismic upheavals |
the epicentres of which are located in densely populated regions:

\Science
the fact of vital growth,
through convulsions of metamorphosis,
from infancy through maturity to decay.


\Q1
Why did he desist from speculation?

\Philosophy
Because it was a task for a superior intelligence |
to substitute other more acceptable phenomena |
in the place of the less acceptable phenomena to be removed.%
\gab{1010}


\Q2
Did Stephen participate in his dejection?

\Philosophy
He affirmed his significance as a conscious rational animal |
proceeding syllogistically from the known to the unknown |
and a conscious rational reagent
between a micro and a macrocosm |
ineluctably constructed upon the incertitude of the void.


\Q1
Was this affirmation apprehended by Bloom?

\Factual
Not verbally.
Substantially.


\Q2
What comforted his misapprehension?

\Bloom
That as a competent keyless citizen |
he had proceeded energetically from the unknown to the known |
\gab{1020}
through the incertitude of the void.


\acting{enact}

\Q1
In what order of precedence,
with what attendant ceremony |
was the exodus from the house of bondage
to the wilderness of inhabitation |
effected?

\Theater
\begin{center}
Lighted Candle in Stick\\
borne by\\
\textsc{Bloom}\\
Diaconal Hat on Ashplant\\
borne by\\
\textsc{Stephen}
\end{center}


\Q2
With what intonation \latin{secreto} |
of what commemorative psalm?

\Religious
The \nth{113},
\latin{modus peregrinus}\/
\gab{1030}
\latin{In exitu Israel de Egypto}\/
\latin{domus Jacob de populo barbaro}.


\acting{enact}

\Q1
What did each do |
at the door of egress?

\Theater
Bloom set the candlestick on the floor.
Stephen put the hat on his head.


\Q2
For what creature |
was the door of egress |
a door of ingress?

\Factual
For a cat.


\acting{both gaze up at stars}

\Q1
What spectacle confronted them |
when they, first the host, then the guest, |
emerged silently, doubly dark, from obscurity |
by a passage from the rere of the house |
into the penumbra of the garden?

\Poetry
The heaventree of stars hung with humid nightblue fruit.


\Q2
With what meditations |
\gab{1040}
did Bloom accompany his demonstration to his companion |
of various constellations?

\Science
Meditations of evolution increasingly vaster:
of the moon invisible in incipient lunation, approaching perigee:
of the infinite lattiginous scintillating uncondensed milky way, |
discernible by daylight by an observer |
placed at the lower end of a cylindrical vertical shaft 5000~ft deep |
sunk from the surface towards the centre of the earth:
of Sirius
\paren{(alpha in Canis Maior)} |
10~lightyears
%(57,000,000,000,000 miles)
\paren{(57~trillion miles)}
distant |
and in volume 900~times the dimension of our planet:
of Arcturus:
of the precession of equinoxes:
of Orion with belt and sextuple sun theta and nebula |
in which 100~of our solar systems could be contained:
\gab{1050}

\Poetry
of moribund and of nascent new stars such as Nova in~1901:
of our system plunging towards the constellation of Hercules:
of the parallax or parallactic drift of so-called fixed stars,
in reality evermoving wanderers |
from immeasurably remote eons to infinitely remote futures |
in comparison with which |
the years, threescore and ten, of allotted human life |
formed a parenthesis of infinitesimal brevity.


\Q1
Were there obverse meditations of involution |
increasingly less vast?

\Science
Of the eons of geological periods recorded in the stratifications of the earth:
of the myriad minute entomological organic existences concealed |
in cavities of the earth, beneath removable stones, in hives and mounds, |
\gab{1060}
of microbes, germs, bacteria, bacilli, spermatozoa:
of the incalculable trillions of billions of millions of imperceptible molecules |
contained by cohesion of molecular affinity in a single pinhead:

\Poetry
of the universe of human serum constellated with red and white bodies, |
themselves universes of void space constellated with other bodies, |
each, in continuity, its universe of divisible component bodies |
of which each was again divisible in divisions of redivisible component bodies, |
dividends and divisors ever diminishing without actual division |
till, if the progress were carried far enough, |
nought nowhere was never reached.


\Q2
Why did he not elaborate these calculations |
\gab{1070}
to a more precise result?

\Science
Because some years previously in~1886 |
when occupied with the problem of the quadrature of the circle |
he had learned of the existence of a number
computed to a relative degree of accuracy |
to be of such magnitude and of so many places,
e.g., the \nth{9}~power of the \nth{9}~power of~9, |
that, the result having been obtained, |
33~closely printed volumes of 1000~pages |
each of innumerable quires and reams of India paper |
would have to be requisitioned |
in order to contain the complete tale of its printed integers |
of units, tens, hundreds, thousands, tens of thousands,
hundreds of thousands, millions, tens of millions, hundreds of millions, billions, |
the nucleus of the nebula of every digit of every series |
\gab{1080}
containing succinctly the potentiality of being raised |
to the utmost kinetic elaboration
of any power of any of its powers.


\Q1
Did he find the problems |
of the inhabitability of the planets and their satellites |
by a race, given in species, |
and of the possible social and moral redemption of said race |
by a redeemer, |
easier of solution?

\Science
Of a different order of difficulty.
Conscious that the human organism, |
normally capable of sustaining an atmospheric pressure of 19~tons, |
when elevated to a considerable altitude in the terrestrial atmosphere |
suffered with arithmetical progression of intensity, |
according as the line of demarcation between troposphere and stratosphere |
\gab{1090}
was approximated from nasal hemorrhage, impeded respiration and vertigo, |
when proposing this problem for solution,
he had conjectured as a working hypothesis |
which could not be proved impossible |
that a more adaptable and differently anatomically constructed race of beings |
might subsist otherwise |
under Martian, Mercurial, Veneral, Jovian, Saturnian, Neptunian
or Uranian sufficient and equivalent conditions, |
though an apogean humanity of beings |
created in varying forms with finite differences |
resulting similar to the whole and to one another |
would probably |
there as here |
remain inalterably and inalienably attached to vanities, |
to vanities of vanities |
and to all that is vanity.%
\gab{1100}


\Q2
And the problem of possible redemption?

\Philosophy
The minor was proved by the major.


\Q1
Which various features of the constellations |
were in turn considered?

\Poetry
The various colours significant of various degrees of vitality
\paren{(white, yellow, crimson, vermilion, cinnabar)}:

\Places
their degrees of brilliancy:

\Memories
their magnitudes revealed up to and including the~\nth{7}:

\Religious
their positions:

\Factual
the waggoner's star:

\Household
Walsingham way:

\Science
the chariot of David:

\Philosophy
the annular cinctures of Saturn:

\Poetry
the condensation of spiral nebulae into suns:

\Memories
the interdependent gyrations of double suns:

\Places
the independent synchronous discoveries of Galileo, Simon Marius,
Piazzi, Le Verrier, Herschel, Galle:
\gab{1110}

\Religious
the systematisations attempted by Bode and Kepler |
of cubes of distances and squares of times of revolution:

\Household
the almost infinite compressibility of hirsute comets |
and their vast elliptical egressive and reentrant orbits |
from perihelion to aphelion:

\Factual
the sidereal origin of meteoric stones:

\Philosophy
the Libyan floods on Mars
about the period of the birth of the younger astroscopist:

\Memories
the annual recurrence of meteoric showers |
about the period of the feast of S.~Lawrence
\paren{(martyr, 10~August)}:

\Household
the monthly recurrence known as the new moon with the old moon in her arms:

\Religious
the posited influence of celestial on human bodies:

\Science
the appearance of a star
\paren{(\nth{1}~magnitude)}
of exceeding brilliancy dominating by night and day |
\gab{1120}
\paren{(a new luminous sun |
generated by the collision and amalgamation in incandescence of two nonluminous exsuns)}
about the period of the birth of William Shakespeare |
over delta in the recumbent neversetting constellation of Cassiopeia |
and of a star
\paren{(\nth{2}~magnitude)}
of similar origin but of lesser brilliancy |
which had appeared in and disappeared from |
the constellation of the Corona Septentrionalis |
about the period of the birth of Leopold Bloom |
and of other stars of
\paren{(presumably)}
similar origin |
which had
\paren{(effectively or presumably)}
appeared in and disappeared from |
the constellation of Andromeda about the period |
of the birth of Stephen Dedalus, |
and in and from the constellation of Auriga some years after |
the birth and death of Rudolph Bloom, junior, |
\gab{1130}
and in and from other constellations |
some years before or after the birth or death of other persons:

\Poetry
the attendant phenomena of eclipses, solar and lunar, |
from immersion to emersion,
abatement of wind, transit of shadow,
taciturnity of winged creatures,
emergence of nocturnal or crepuscular animals,
persistence of infernal light,
obscurity of terrestrial waters,
pallor of human beings.


\Q2
His
\paren{(Bloom's)}
logical conclusion, |
having weighed the matter |
and allowing for possible error?

\Poetry
That it was not a heaventree, not a heavengrot, not a heavenbeast, not a heavenman.
\gab{1140}
That it was a Utopia, |
there being no known method from the known to the unknown:
an infinity renderable equally finite |
by the suppositious apposition of one or more bodies |
equally of the same and of different magnitudes:
a mobility of illusory forms immobilised in space,
remobilised in air:
a past which possibly had ceased to exist as a present |
before its probable spectators had entered actual present existence.


\Q1
Was he more convinced |
of the esthetic value of the spectacle?

\Poetry
Indubitably |
in consequence of the reiterated examples of poets |
in the delirium of the frenzy of attachment |
or in the abasement of rejection |
invoking ardent sympathetic constellations |
or the frigidity of the satellite of their planet.%
\gab{1150}


\Q2
Did he then accept |
as an article of belief |
the theory of astrological influences upon sublunary disasters?

\Science
It seemed to him |
as possible of proof |
as of confutation |
and the nomenclature employed in its selenographical charts |
as attributable to verifiable intuition |
as to fallacious analogy:

\Poetry
the lake of dreams, the sea of rains, the gulf of dews, the ocean of fecundity.


\Q1
What special affinities |
appeared to him to exist between the moon and woman?

\Philosophy
Her antiquity in preceding and surviving successive tellurian generations:
her nocturnal predominance:
\gab{1160}
her satellitic dependence:
her luminary reflection:
her constancy under all her phases,
rising and setting by her appointed times,
waxing and waning:
the forced invariability of her aspect:
her indeterminate response to inaffirmative interrogation:
her potency over effluent and refluent waters:

\Poetry
her power to enamour, to mortify, to invest with beauty, |
to render insane, to incite to and aid delinquency:
the tranquil inscrutability of her visage:
the terribility of her isolated dominant implacable resplendent propinquity:
her omens of tempest and of calm:
the stimulation of her light, her motion and her presence:
the admonition of her craters, her arid seas, her silence:
her splendour, when visible:
her attraction,
\gab{1170}
when invisible.


\acting{both look up at window}

\Q2
What visible luminous sign |
attracted Bloom's, who attracted Stephen's, |
gaze?

\Household
In the second storey
\paren{(rere)}
of his
\paren{(Bloom's)}
house |
the light of a paraffin oil lamp with oblique shade |
projected on a screen of roller blind |
supplied by Frank O'Hara, |
window blind, curtain pole and revolving shutter manufacturer, |
16 Aungier street.


\acting{Bloom explain it's Molly}

\Q1
How did he elucidate the mystery |
of an invisible attractive person, |
his wife Marion (Molly) Bloom, |
denoted by a visible splendid sign, |
a lamp?

\Poetry
With indirect and direct verbal allusions or affirmations:
with subdued affection and admiration:
\gab{1180}
with description:
with impediment:
with suggestion.


\acting{contemplate each other}

\Q2
Both then were silent?

\Poetry
Silent, each contemplating the other in both mirrors |
of the reciprocal flesh of their-his-not-his fellowfaces.


\acting{both urinate and look up at window}

\Q1
Were they indefinitely inactive?

\Science
At Stephen's suggestion,
at Bloom's instigation |
both, |
first Stephen, then Bloom, |
in penumbra urinated, |
their sides contiguous, |
their organs of micturition reciprocally rendered invisible by manual circumposition, |
their gazes, |
first Bloom's, then Stephen's, |
elevated to the projected luminous and semiluminous shadow.%
\gab{1190}


\Q2
Similarly?

\Science
The trajectories of their, |
first sequent, then simultaneous, |
urinations were dissimilar:

\Memories
Bloom's longer, less irruent, |
in the incomplete form of the bifurcated penultimate alphabetical letter, |
who in his ultimate year at High School
\paren{(1880)} | |
had been capable of attaining the point of greatest altitude |
against the whole concurrent strength of the institution,
210~scholars:
Stephen's higher, more sibilant, |
who in the ultimate hours of the previous day |
had augmented by diuretic consumption |
an insistent vesical pressure.


\Q1
What different problems presented themselves to each |
concerning the invisible audible collateral organ of the other?%
\gab{1200}

\Science
To Bloom:
the problems of irritability, tumescence, rigidity,
reactivity, dimension, sanitariness, pilosity.

\Religious
To Stephen:
the problem of the sacerdotal integrity of Jesus circumcised |
\paren{(1~January, holiday of obligation to hear mass
and abstain from unnecessary servile work)}
and the problem as to whether the divine prepuce, |
the carnal bridal ring of the holy Roman catholic apostolic church, |
conserved in Calcata, |
were deserving of simple hyperduly |
or of the fourth degree of latria |
accorded to the abscission of such divine excrescences as hair and toenails.


\Q2
What celestial sign |
\gab{1210}
was by both simultaneously observed?

\Science
A~star precipitated with great apparent velocity |
across the firmament from Vega in the Lyre above the zenith |
beyond the stargroup of the Tress of Berenice |
towards the zodiacal sign of Leo.


\acting{enact insert key, pull bolt, open door}

\Q1
How did the centripetal remainer |
afford egress to the centrifugal departer?

\Science + \Factual
By inserting the barrel of an arruginated male key
in the hole of an unstable female lock,
obtaining a purchase on the bow of the key |
and turning its wards from right to left,
withdrawing a bolt from its staple,
pulling inward spasmodically an obsolescent unhinged door |
and revealing an aperture for free egress and free ingress.


\acting{enact shaking hands with straight arms}

\Q2
How did they take leave, one of the other, in separation?%
\gab{1220}

\Science + \Factual
Standing perpendicular at the same door |
and on different sides of its base,
the lines of their valedictory arms,
meeting at any point and forming any angle less than the sum of two right angles.


\acting{enact hearing bells}

\Q1
What sound accompanied the union of their tangent, |
the disunion of their
\paren{(respectively)}
centrifugal and centripetal hands?

\Places + \Poetry
The sound of the peal of the hour of the night |
by the chime of the bells in the church of Saint George.


\Q2
What echoes of that sound |
were by both and each heard?

\Religious
By Stephen:

\latin{Liliata rutilantium.
\gab{1230}
Turma circumdet.
Iubilantium te virginum.
Chorus excipiat.}

\Factual
By Bloom:

\begin{verse}
    Heigho, heigho,\\
    Heigho, heigho.
\end{verse}


\Q1
Where were the several members of the company |
which with Bloom that day at the bidding of that peal |
had travelled from Sandymount in the south |
to Glasnevin in the north?

\Turns[~(men)]
Martin Cunningham
\paren{(in bed),}
Jack Power
\paren{(in bed),}
Simon Dedalus
\paren{(in bed),}
Ned Lambert
\paren{(in bed),}
Tom Kernan
\paren{(in bed),}
Joe Hynes
\paren{(in bed),}
John Henry Menton
\paren{(in bed),}
\gab{1240}
Bernard Corrigan
\paren{(in bed),}
Patsy Dignam
\paren{(in bed),}
Paddy Dignam
\paren{(in the grave).}


\acting{Stephen leave (become reader)}

\Q2
Alone, |
what did Bloom hear?

\Poetry
The double reverberation of retreating feet on the heavenborn earth,
the double vibration of a jew's harp in the resonant lane.


\Q1
Alone, |
what did Bloom feel?

\Science
The cold of interstellar space,
thousands of degrees below freezing point |
or the absolute zero of Fahrenheit, Centigrade or Reaumur:
the incipient intimations of proximate dawn.


\Q2
Of what |
did bellchime and handtouch and footstep and lonechill |
\gab{1250}
remind him?

\Religious
Of companions now in various manners in different places |
defunct:
Percy Apjohn
\paren{(killed in action, Modder River),} |
Philip Gilligan
\paren{(phthisis, Jervis Street hospital),} |
Matthew~F. Kane
\paren{(accidental drowning, Dublin Bay),} |
Philip Moisel
\paren{(pyemia, Heytesbury street),} |
Michael Hart
\paren{(phthisis, Mater Misericordiae hospital),} |
Patrick Dignam
\paren{(apoplexy, Sandymount).}


\Q1
What prospect of what phenomena |
inclined him to remain?

\Science
The disparition of three final stars,
the diffusion of daybreak,
the apparition of a new solar disk.


\Q2
Had he ever been a spectator of those phenomena?

\Memories
Once, in 1887, |
after a protracted performance of charades |
\gab{1260}
in the house of Luke Doyle, Kimmage, |
he had awaited with patience the apparition of the diurnal phenomenon, |
seated on a wall,
his gaze turned in the direction of Mizrach, the east.


\Q1
He remembered the initial paraphenomena?

\Poetry
More active air,
a matutinal distant cock,
ecclesiastical clocks at various points,
avine music,
the isolated tread of an early wayfarer,
the visible diffusion of the light of an invisible luminous body,
the first golden limb of the resurgent sun perceptible low on the horizon.


\acting{Bloom breathe in, walk, reenter}

\Q2
Did he remain?

\Factual[~(legal)]
With deep inspiration |
he returned,
\gab{1270}
retraversing the garden, reentering the passage, reclosing the door.
With brief suspiration |
he reassumed the candle,
reascended the stairs,
reapproached the door of the front room, hallfloor,
and reentered.


\acting{Bloom hit head hard on angle of sideboard}

\Q1
What suddenly arrested his ingress?

\Science
The right temporal lobe of the hollow sphere of his cranium |
came into contact with a solid timber angle |
where, an infinitesimal but sensible fraction of a second later, |
a painful sensation was located |
in consequence of antecedent sensations |
transmitted and registered.


\Q2
Describe the alterations |
effected in the disposition of the articles of
furniture.%
\gab{1280}

\Household
A~sofa upholstered in prune plush had been translocated |
from opposite the door to the ingleside near the compactly furled Union Jack |
\paren{(an alteration which he had frequently intended to execute)}:
the blue and white checker inlaid majolica-topped table |
had been placed opposite the door in the place vacated by the prune plush sofa:
the walnut sideboard |
\paren{(a projecting angle of which had momentarily arrested his ingress)}
had been moved from its position beside the door |
to a more advantageous but more perilous position in front of the door:
two chairs had been moved from right and left of the ingleside |
to the position originally occupied
by the blue and white checker inlaid majolica-topped table.%
\gab{1290}


\Q1
Describe them.

\Philosophy
One:
a squat stuffed easychair,
with stout arms extended and back slanted to the rere,
which, repelled in recoil,
had then upturned an irregular fringe of a rectangular rug |
and now displayed on its amply upholstered seat |
a centralised diffusing and diminishing discolouration.

\Poetry
The other:
a slender splayfoot chair of glossy cane curves,
placed directly opposite the former,
its frame from top to seat and from seat to base |
being varnished dark brown,
its seat being a bright circle of white plaited rush.


\Q2
What significances |
attached to these two chairs?

\Household
Significances of similitude, of posture, of symbolism,
\gab{1300}
of circumstantial evidence, of testimonial supermanence.


\Q1
What occupied the position |
originally occupied by the sideboard?

\Household
A~vertical piano
\paren{(Cadby)}
with exposed keyboard,
its closed coffin supporting a pair of long yellow ladies' gloves |
and an emerald ashtray containing four consumed matches,
a partly consumed cigarette and two discoloured ends of cigarettes,
its musicrest supporting the music |
in the key of G~natural for voice and piano |
of \Title{Love's Old Sweet Song} |
\paren{(words by G.~Clifton Bingham,
composed by J.L.~Molloy,
sung by Madam Antoinette Sterling)}
open at the last page |
with the final indications
\latin{ad libitum}, \italian{forte}, pedal,
\italian{animato}, sustained pedal, \italian{ritirando},
\gab{1310}
close.


\acting{Bloom lift candle, feel temple, bend, etc}

\Q2
With what sensations |
did Bloom contemplate in rotation these objects?

\Turns
With strain, |
elevating a candlestick:
with pain, |
feeling on his right temple a contused tumescence:
with attention, |
focussing his gaze
on a large dull passive and a slender bright active:
with solicitation, |
bending and downturning the upturned rugfringe:
with amusement, |
remembering Dr Malachi Mulligan's scheme of colour |
containing the gradation of green:
with pleasure, |
repeating the words and antecedent act |
and perceiving through various channels of internal sensibility |
the consequent and concomitant tepid pleasant diffusion of gradual discolouration.


\acting{Bloom actually enact with incense cone}

\Q1
His next proceeding?
\gab{1320}

\Science
From an open box on the majolica-topped table |
he extracted a black diminutive cone, one inch in height,
placed it on its circular base on a small tin plate,
placed his candlestick on the right corner of the mantelpiece,
produced from his waistcoat a folded page of prospectus
\paren{(illustrated)}
entitled Agendath Netaim,
unfolded the same,
examined it superficially,
rolled it into a thin cylinder,
ignited it in the candleflame,
applied it when ignited to the apex of the cone |
till the latter reached the stage of rutilance,
placed the cylinder in the basin of the candlestick |
disposing its unconsumed part
in such a manner as to facilitate total combustion.


\acting{Bloom actually enact with incense cone}

\Q2
What followed this operation?
\gab{1330}

\Science
The truncated conical crater summit of the diminutive volcano |
emitted a vertical and serpentine fume |
redolent of aromatic oriental incense.


\Q1
What homothetic objects,
other than the candlestick,
stood on the mantelpiece?

\Household
A~timepiece of striated Connemara marble,
stopped at the hour of 4.46~a.m.\ on the 21~March 1896, |
matrimonial gift of Matthew Dillon:
a dwarf tree of glacial arborescence under a transparent bellshade, |
matrimonial gift of Luke and Caroline Doyle:
an embalmed owl, |
matrimonial gift of Alderman John Hooper.


\acting{Bloom look at tree}

\Q2
What interchanges of looks |
took place between these three objects and
Bloom?%
\gab{1340}

\Household
In the mirror of the giltbordered pierglass |
the undecorated back of the dwarf tree
regarded the upright back of the embalmed owl.
Before the mirror |
the matrimonial gift of Alderman John Hooper |
with a clear melancholy wise bright
motionless compassionate gaze |
regarded Bloom |
while Bloom with obscure tranquil profound motionless compassionated gaze |
regarded the matrimonial gift of Luke and Caroline Doyle.


\acting{Bloom look in mirror}

\Q1
What composite asymmetrical image in the mirror |
then attracted his attention?

\Philosophy
The image of a solitary
\paren{(ipsorelative)}
\gab{1350}
mutable
\paren{(aliorelative)}
man.


\Q2
Why solitary
\paren{(ipsorelative)?}

\Poetry

\begin{verse}
    Brothers and sisters had he none.\\
    Yet that man's father was his grandfather's son.
\end{verse}


\Q1
Why mutable
\paren{(aliorelative)?}

\Philosophy
From infancy to maturity |
he had resembled his maternal procreatrix.
From maturity to senility |
he would increasingly resemble his paternal procreator.


\acting{Bloom notice upside down books}

\Q2
What final visual impression |
was communicated to him by the mirror?

\Science
The optical reflection of several inverted volumes improperly arranged |
and not in the order of their common letters |
with scintillating titles on the two bookshelves opposite.%
\gab{1360}


\acting{last long turn read by ``memories''}

\Q1
Catalogue these books.

\Turns
\Title{Thom's Dublin Post Office Directory, 1886}.\\
Denis Florence M'Carthy's \Title{Poetical Works}
\paren{(copper beechleaf bookmark at p.~5).}\\
Shakespeare's \Title{Works}
\paren{(dark crimson morocco, goldtooled).}\\
\Title{The Useful Ready Reckoner}
\paren{(brown cloth).}\\
\Title{The Secret History of the Court of Charles~II}
\paren{(red cloth, tooled binding).}\\
\Title{The Child's Guide}
\paren{(blue cloth).}\\
\Title{The Beauties of Killarney}
\paren{(wrappers).}\\
\Title{When We Were Boys} by William O'Brien~M.P.
\gab{1370}
\paren{(green cloth, slightly faded, envelope bookmark at p.~217).}\\
\Title{Thoughts from Spinoza}
\paren{(maroon leather).}\\
\Title{The Story of the Heavens} by Sir Robert Ball
\paren{(blue cloth).}\\
Ellis's \Title{Three Trips to Madagascar}
\paren{(brown cloth, title obliterated).}\\
\Title{The Stark-Munro Letters} by A.~Conan Doyle,
property of the City of Dublin Public Library, 106~Capel street, |
lent 21~May
\paren{(Whitsun Eve)}
1904, |
due 4~June 1904, 13~days overdue
\paren{(black cloth binding, bearing white letternumber ticket).}\\
\Title{Voyages in China} by ``Viator''
\paren{(recovered with brown paper, red ink title).}\\
\Title{Philosophy of the Talmud}
\gab{1380}
\paren{(sewn pamphlet).}\\
Lockhart's \Title{Life of Napoleon}
\paren{(cover wanting, marginal annotations, minimising victories,
aggrandising defeats of the protagonist).}\\
\Title{Soll und Haben} by Gustav Freytag
\paren{(black boards, Gothic characters, cigarette coupon bookmark at p.~24).}\\
Hozier's \Title{History of the Russo-Turkish War}
\paren{(brown cloth, 2~volumes, with gummed label,
Garrison Library, Governor's Parade, Gibraltar, on verso of cover).}\\
\Title{Laurence Bloomfield in Ireland} by William Allingham
\paren{(second edition, green cloth, gilt trefoil design, |
previous owner's name on recto of flyleaf erased).}%
\gab{1390}\\
\Title{A Handbook of Astronomy}
\paren{(cover, brown leather, detached, 5~plates,
antique letterpress long primer, author's footnotes nonpareil,
marginal clues brevier, captions small pica).}\\
\Title{The Hidden Life of Christ}
\paren{(black boards).}\\
\Title{In the Track of the Sun}
\paren{(yellow cloth, titlepage missing, recurrent title intestation).}\\
\Title{Physical Strength and How to Obtain It} by Eugen Sandow
\paren{(red cloth).}\\
\Title{Short but yet Plain Elements of Geometry}
written in French by F.~Ignat.\ Pardies
and rendered into Engliſh by John Harris~D.D. London, |
printed for R.~Knaplock at the Biſhop's Head, MDCCXI \stage{1711}, |
\gab{1400}
with dedicatory epiſtle to his worthy friend Charles Cox, eſquire, |
Member of Parliament for the burgh of Southwark |
and having ink calligraphed statement on the flyleaf |
certifying that the book was the property of Michael Gallagher, |
dated this \nth{10}~day of May~1822 |
and requeſting the perſon who should find it, |
if the book should be loſt or go aſtray, |
to reſtore it to Michael Gallagher, carpenter, |
Dufery Gate, Enniſcorthy, county Wicklow, |
the fineſt place in the world.


\acting{Bloom put books right way up}

\Q2
What reflections occupied his mind |
during the process of reversion of the inverted volumes?

\Philosophy
The necessity of order,
\gab{1410}
a place for everything and everything in its place:
the deficient appreciation of literature possessed by females:
the incongruity of an apple incuneated in a tumbler |
and of an umbrella inclined in a closestool:
the insecurity of hiding any secret document |
behind, beneath or between the pages of a book.


\Q1
Which volume was the largest in bulk?

\Factual
Hozier's \Title{History of the Russo-Turkish war.}


\Q2
What |
among other data |
did the second volume of the work in question contain?

\Memories
The name of a decisive battle
\paren{(forgotten),}
frequently remembered by a decisive officer,
\gab{1420}
major Brian Cooper Tweedy
\paren{(remembered).}


\Q1
Why,
firstly and secondly,
did he not consult the work in question?

\Bloom
Firstly, in order to exercise mnemotechnic:
secondly, because after an interval of amnesia,
when, seated at the central table,
about to consult the work in question,
he remembered by mnemotechnic |
the name of the military engagement, |
Plevna.


\Q2
What caused him consolation |
in his sitting posture?

\Household
The candour, nudity, pose, tranquility, youth, grace, sex,
counsel of a statue erect in the centre of the table, |
an image of Narcissus purchased by auction
from P.A.~Wren, 9~Bachelor's Walk.


\acting{Bloom look irritated by collar}

\Q1
What caused him irritation |
in his sitting posture?
\gab{1430}

\Science
Inhibitory pressure of collar
\paren{(size~17)}
and waistcoat
\paren{(5~buttons),} |
two articles of clothing superfluous in the costume of mature males |
and inelastic to alterations of mass by expansion.


\acting{Bloom take off collar, waistcoat and unbutton}

\Q2
How was the irritation allayed?

\Factual
He removed his collar,
with contained black necktie and collapsible stud,
from his neck to a position on the left of the table.

\Science
He unbuttoned successively |
in reversed direction |
waistcoat, trousers, shirt and vest |
along the medial line of irregular incrispated black hairs |
extending in triangular convergence |
from the pelvic basin over the circumference of the abdomen
and umbilicular fossicle along the medial line of nodes |
\gab{1440}
to the intersection of the sixth pectoral vertebrae,
thence produced both ways at right angles and terminating in circles |
described about two equidistant points,
right and left, |
on the summits of the mammary prominences.
He unbraced successively |
each of six minus one braced trouser buttons,
arranged in pairs,
of which |
one incomplete.


\acting{parentheses read by ``memories''}

\Q1
What involuntary actions followed?

\Science
He compressed between 2~fingers |
the flesh circumjacent to a cicatrice |
in the left infracostal region below the diaphragm |
resulting from a sting |
inflicted 2~weeks and 3~days previously
\paren{(23~May 1904)}
by a bee.
He scratched imprecisely with his right hand,
\gab{1450}
though insensible of prurition, |
various points and surfaces of his partly exposed, wholly abluted skin.

\Factual
He inserted his left hand into the left lower pocket of his waistcoat |
and extracted and replaced a silver coin
\paren{(1~shilling),}
placed there
\paren{(presumably)}
on the occasion
\paren{(17 October 1903)}
of the interment of Mrs Emily Sinico, Sydney Parade.


\Q2
Compile the budget for 16~June 1904.

\newcommand{\fjnl}{\Title{Freeman's J$^{\textit{nl}}$}}

\Roger

\begin{tabular}{l r l r}
    \multicolumn{2}{c}{\emph{Debit}}        & \multicolumn{2}{c}{\emph{Credit}}     \\
                                & £--s--d   &                           & £--s--d   \\
1 Pork Kidney                   & 0--0--3   & Cash in hand              & 0--4--9   \\
1 Copy \fjnl                    & 0--0--1   & Commission recd.\ \fjnl   & 1--7--6   \\
1 Bath And Gratification        & 0--1--6   & Loan (Stephen Dedalus)    & 1--7--0   \\
Tramfare                        & 0--0--1   \\
1 In Memoriam Patrick Dignam    & 0--5--0   \\
2 Banbury cakes                 & 0--0--1   \\
1 Lunch                         & 0--0--7   \\
1 Renewal fee for book          & 0--1--0   \\
1 Packet Notepaper \& Envelopes & 0--0--2   \\
1 Dinner and Gratification      & 0--2--0   \\
1 Postal Order and Stamp        & 0--2--8   \\
Tramfare                        & 0--0--1   \\
1 Pig's Foot                    & 0--0--4   \\
1 Sheep's Trotter               & 0--0--3   \\
1 Cake Fry's Plain Chocolate    & 0--0--1   \\
1 Square Soda Bread             & 0--0--4   \\
1 Coffee and Bun                & 0--0--4   \\
Loan (Stephen Dedalus) refunded & 1--7--0   \\
Balance                         & 0-16--6   \\

                                \cline{2-2}                            \cline{4-4}\\
Total                           &£ 2-19--3  & Total                    &£ 2-19--3\\
\end{tabular}


\acting{Bloom do the toenail smell thing}

\Q1
Did the process of divestiture |
continue?

\Factual
Sensible of a benignant persistent ache in his footsoles |
\gab{1480}
he extended his foot to one side |
and observed the creases, protuberances and salient points |
caused by foot pressure in the course of walking repeatedly
in several different directions, |
then,
inclined, |
he disnoded the laceknots, unhooked and loosened the laces, |

\Science
took off each of his two boots for the second time, |
detached the partially moistened right sock |
through the fore part of which the nail of his great toe had again effracted, |
raised his right foot |
and,
having unhooked a purple elastic sock suspender, |
took off his right sock, |
placed his unclothed right foot on the margin of the seat of his chair, |
picked at and gently lacerated the protruding part of the great toenail, |
raised the part lacerated to his nostrils |
and inhaled the odour of the quick, |
\gab{1490}
then,
with satisfaction, |
threw away the lacerated ungual fragment.


\Q2
Why |
with satisfaction?

\Memories
Because the odour inhaled corresponded to other odours |
inhaled of other ungual fragments, |
picked and lacerated by Master Bloom, |
pupil of Mrs Ellis's juvenile school, |
patiently each night |
in the act of brief genuflection and nocturnal prayer and ambitious meditation.


\Q1
In what ultimate ambition |
had all concurrent and consecutive ambitions |
now coalesced?

\Factual
Not to inherit by right of primogeniture, gavelkind or borough English, |
or possess in perpetuity |
\gab{1500}
an extensive demesne of a sufficient number of acres, roods and perches, |
statute land measure
\paren{(valuation 42~pounds),} |
of grazing turbary |
surrounding a baronial hall |
with gatelodge and carriage drive |
nor, |
on the other hand, |
a terracehouse or semidetached villa, |
described as \latin{Rus in Urbe} or \latin{Qui si sana}, |
but |
to purchase by private treaty in fee simple |
a thatched bungalowshaped 2~storey dwellinghouse of southerly aspect, |
surmounted by vane and lightning conductor, |
connected with the earth, |
with porch covered by parasitic plants
\paren{(ivy or Virginia creeper),} |
halldoor, olive green,
with smart carriage finish and neat doorbrasses, |
stucco front with gilt tracery at eaves and gable, |
rising, if possible, upon a gentle eminence |
with agreeable prospect from balcony with stone pillar parapet |
\gab{1510}
over unoccupied and unoccupyable interjacent pastures |
and standing in 5~or~6 acres of its own ground, |
at such a distance from the nearest public thoroughfare |
as to render its houselights visible at night |
above and through a quickset hornbeam hedge of topiary cutting, |
situate at a given point not less than 1~statute mile
from the periphery of the metropolis, |
within a time limit of not more than 15~minutes from tram or train line |
\paren{(e.g., Dundrum, south, or Sutton, north, |
both localities equally reported by trial
to resemble the terrestrial poles |
in being favourable climates for phthisical subjects),} |

\Household
the premises to be held under feefarm grant, lease 999~years, |
the messuage to consist of 1~drawingroom with baywindow
\paren{(2~lancets),}
\gab{1520}
thermometer affixed, |
1~sittingroom, 4~bedrooms, 2~servants' rooms, |
tiled kitchen with close range and scullery, |
lounge hall fitted with linen wallpresses, |
fumed oak sectional bookcase
containing the Encyclopaedia Britannica and New Century Dictionary, |
transverse obsolete medieval and oriental weapons, |
dinner gong, alabaster lamp, bowl pendant, |
vulcanite automatic telephone receiver with adjacent directory, |
handtufted Axminster carpet with cream ground and trellis border, |
loo table with pillar and claw legs, |
hearth with massive firebrasses
and ormolu mantel chronometer clock, guaranteed timekeeper with cathedral chime, |
barometer with hygrographic chart, |
comfortable lounge settees
\gab{1530}
and corner fitments,
upholstered in ruby plush with good springing and sunk centre, |
three banner Japanese screen and cuspidors |
\paren{(club style, rich winecoloured leather, |
gloss renewable with a minimum of labour by use of linseed oil and vinegar)} |
and pyramidically prismatic central chandelier lustre, |
bentwood perch with fingertame parrot
\paren{(expurgated language),} |
embossed mural paper at 10/-~per dozen |
with transverse swags of carmine floral design and top crown frieze, |
staircase, three continuous flights at successive right angles,
of varnished cleargrained oak, treads and risers, newel, balusters and handrail, |
with steppedup panel dado,
dressed with camphorated wax:

\Science
bathroom, hot and cold supply, reclining and shower:
\gab{1540}
water closet on mezzanine provided with opaque singlepane oblong window, |
tipup seat, bracket lamp, brass tierod and brace,
armrests, footstool and artistic oleograph on inner face of door:
ditto, plain:
servants' apartments with separate sanitary and hygienic necessaries |
for cook, general and betweenmaid |
\paren{(salary, rising by biennial unearned increments of 2~pounds, |
with comprehensive fidelity insurance, |
annual bonus
    \paren{(1~pound)} |
and retiring allowance
    \paren{(based on the 65~system)}
after 30~years' service),} |
pantry, buttery, larder, refrigerator, |
outoffices, coal and wood cellarage with winebin
\paren{(still and sparkling vintages)} |
for distinguished guests, if entertained to dinner
\paren{(evening dress),} |
carbon monoxide gas supply throughout.%
\gab{1550}


\Q2
What additional attractions |
might the grounds contain?

\Household
As addenda, |
a tennis and fives court, |
a shrubbery, |
a glass summerhouse with tropical palms, equipped in the best botanical manner, |
a rockery with waterspray, |
a beehive arranged on humane principles, |
oval flowerbeds in rectangular grassplots
set with eccentric ellipses of scarlet and chrome tulips, |
blue scillas, crocuses, polyanthus,
sweet William, sweet pea, |
lily of the valley |
\paren{(bulbs obtainable from sir James~W. Mackey (Limited) |
wholesale and retail seed and bulb merchants and nurserymen, |
agents for chemical manures, 23~Sackville street, upper),} |
an orchard, kitchen garden and vinery |
protected against illegal trespassers by glasstopped mural enclosures, |
\gab{1560}
a lumbershed with padlock for various inventoried implements.


\Q1
As?

\Household
Eeltraps, lobsterpots, fishingrods, hatchet, steelyard, grindstone, |
clodcrusher, swatheturner, carriagesack, telescope ladder, 10~tooth rake, |
washing clogs, hay\-tedder, tumbling rake, billhook, paintpot, |
brush, hoe and so on.

\Q2
What improvements |
might be subsequently introduced?

\Household
A~rabbitry and fowlrun, a dovecote, a botanical conservatory, |
2~hammocks
\paren{(lady's and gentleman's),} |
a sundial shaded and sheltered by laburnum or lilac trees, |
\gab{1570}
an exotically harmonically accorded Japanese tinkle gatebell |
affixed to left lateral gatepost, |
a capacious waterbutt, |
a lawnmower with side delivery and grassbox, |
a lawnsprinkler with hydraulic hose.


\Q1
What facilities of transit |
were desirable?

\Places
When citybound |
frequent connection by train or tram |
from their respective intermediate station or terminal.
When countrybound |
velocipedes, |
a chainless freewheel roadster cycle with side basketcar attached, |
or draught conveyance, |
a donkey with wicker trap |
or smart phaeton with good working solidungular cob
\paren{(roan gelding, 14~h).}


\Q2
What might be the name |
of this erigible or erected residence?

\Bloom
Bloom Cottage.
Saint Leopold's.
\gab{1580}
Flowerville.


\acting{Bloom imagine}

\Q1
Could Bloom of 7~Eccles street |
foresee Bloom of Flowerville?

\Philosophy
In loose allwool garments with Harris tweed cap, price~8/6, |
and useful garden boots with elastic gussets and wateringcan, |
planting aligned young firtrees, syringing, pruning, staking, sowing hayseed, |
trundling a weedladen wheelbarrow without excessive fatigue at sunset |
amid the scent of newmown hay, |
ameliorating the soil, multiplying wisdom, achieving longevity.


\acting{Bloom imagine}

\Q2
What syllabus of intellectual pursuits |
was simultaneously possible?

\Philosophy
Snapshot photography,
comparative study of religions,
folklore relative to various amatory and superstitious practices,
\gab{1590}
contemplation of the celestial constellations.


\Q1
What lighter recreations?

\Poetry
Outdoor:
garden and fieldwork, |
cycling on level macadamised causeways |
ascents of moderately high hills, |
natation in secluded fresh water |
and unmolested river boating in secure wherry
or light curricle with kedge anchor |
on reaches free from weirs and rapids |
\paren{(period of estivation),} |
vespertinal perambulation
or equestrian circumprocession with inspection of sterile landscape |
and contrastingly agreeable cottagers' fires of smoking peat turves |
\paren{(period of hibernation).}

\Household
Indoor:
discussion in tepid security |
of unsolved historical and criminal problems:
\gab{1600}
lecture of unexpurgated exotic erotic masterpieces:
house carpentry with toolbox
containing hammer, awl nails, screws, tintacks,
gimlet, tweezers, bullnose plane and turnscrew.


\Q2
Might he become a gentleman farmer |
of field produce and live stock?

\Household
Not impossibly, |
with 1~or~2 stripper cows, |
1~pike of upland hay and requisite farming implements, |
e.g., an end-to-end churn,
a turnip pulper etc.


\Q1
What would be |
his civic functions and social status |
among the county families and landed gentry?

\Poetry[~(emotion)]
Arranged successively in ascending powers of hierarchical order, |
that of gardener, groundsman, cultivator, breeder, |
and at the zenith of his career, |
resident magistrate or justice of the peace |
\gab{1610}
with a family crest and coat of arms
and appropriate classical motto
\paren{(\latin{Semper paratus}),} |
duly recorded in the court directory |
\paren{(Bloom, Leopold~P., M.P., P.C., K.P., L.L.D.\
    \paren{(\latin{honoris causa}),} |
Bloomville, Dundrum)} |
and mentioned in court and fashionable intelligence |
\paren{(Mr~and Mrs Leopold Bloom have left Kingstown for England).}


\Q2
What course of action |
did he outline for himself |
in such capacity?

\Factual[~(legal)]
A~course that lay between undue clemency and excessive rigour:
the dispensation in a heterogeneous society of arbitrary classes, |
incessantly rearranged in terms of greater and lesser social inequality, |
of unbiassed homogeneous indisputable justice, |
\gab{1620}
tempered with mitigants of the widest possible latitude |
but exactable to the uttermost farthing |
with confiscation of estate, real and personal, to the crown.
Loyal to the highest constituted power in the land, |
actuated by an innate love of rectitude |
his aims would be the strict maintenance of public order, |
the repression of many abuses though not of all simultaneously |
\paren{(every measure of reform or retrenchment |
being a preliminary solution to be contained by fluxion in the final solution),} |
the upholding of the letter of the law
\paren{(common, statute and law merchant)} |
against all traversers in covin |
and trespassers acting in contravention of bylaws and regulations, |
all resuscitators
\paren{(by trespass and petty larceny of kindlings)} |
\gab{1630}
of venville rights, obsolete by desuetude, |
all orotund instigators of international persecution, |
all perpetuators of international animosities, |
all menial molestors of domestic conviviality, |
all recalcitrant violators of domestic connubiality.


\Q1
Prove that he had loved rectitude |
from his earliest youth.

\Religious
To Master Percy Apjohn at High School in~1880 |
he had divulged his disbelief in the tenets of the Irish
\paren{(protestant)}
church |
\paren{(to which his father Rudolf Virag
    \paren{(later Rudolph Bloom)}
had been converted from the Israelitic faith and communion in~1865 |
by the Society for promoting Christianity among the jews)}
subsequently abjured by him in favour of Roman catholicism |
at the epoch of and with a view to his matrimony in~1888.
\gab{1640}

\Science
To Daniel Magrane and Francis Wade in~1882 during a juvenile friendship |
\paren{(terminated by the premature emigration of the former)} |
he had advocated during nocturnal perambulations |
the political theory of colonial
\paren{(e.g.\ Canadian)}
expansion |
and the evolutionary theories of Charles Darwin, |
expounded in \Title{The Descent of Man} and \Title{The Origin of Species}.

\Factual
In~1885 he had publicly expressed his adherence |
to the collective and national economic programme |
advocated by James Fintan Lalor, John Fisher Murray, John Mitchel,
J.F.X.~O'Brien and others, |
the agrarian policy of Michael Davitt, |
the constitutional agitation of Charles Stewart Parnell
\paren{(M.P.~for Cork City),} |
\gab{1650}
the programme of peace, retrenchment and reform |
of William Ewart Gladstone
\paren{(M.P.~for Midlothian,~N.B.)} |

\Memories
and,
in support of his political convictions, |
had climbed up into a secure position |
amid the ramifications of a tree on Northumberland road |
to see the entrance
\paren{(2~February 1888)}
into the capital |
of a demonstrative torchlight procession of 20,000~torchbearers, |
divided into 120~trade corporations, |
bearing 2000~torches in escort of the marquess of Ripon and
\paren{(honest)}
John Morley.


\Q2
How much and how |
did he propose to pay for this country residence?

\Factual[~(legal)]
As per prospectus of
the Industrious Foreign Acclimatised Nationalised
Friendly State-aided Building Society
\paren{(incorporated~1874),} |
a maximum of 60~pounds per annum, |
being $\frac{1}{6}$~of an assured income,
\gab{1660}
derived from giltedged securities, |
representing at 5\%~simple interest on capital of 1200~pounds |
\paren{(estimate of price at 20~years' purchase),} |
of which to be paid on acquisition and the balance in the form of annual rent, |
viz.\ \stage{namely} 800~pounds plus 2½\%~interest on the same, |
repayable quarterly in equal annual instalments |
until extinction by amortisation of loan advanced |
for purchase within a period of 20~years, |
amounting to an annual rental of 64~pounds, headrent included, |
the titledeeds to remain in possession of the lender or lenders |
with a saving clause envisaging forced sale,
foreclosure and mutual compensation |
in the event of protracted failure to pay the terms assigned, |
otherwise the mess\-uage to become the absolute property of the tenant occupier |
\gab{1670}
upon expiry of the period of years stipulated.


\Q1
What rapid but insecure means to opulence |
might facilitate immediate purchase?

\Philosophy
A~private wireless telegraph
which would transmit by dot and dash system |
the result of a national equine handicap
\paren{(flat or steeplechase)} |
of 1~or more miles and furlongs |
won by an outsider at odds of 50~to~1 |
at 3~hr~8~m~p.m.\ at Ascot
\paren{(Greenwich time),} |
the message being received and available for betting purposes |
in Dublin at~2.59~p.m.\
\paren{(Dunsink time).}

\Poetry
The unexpected discovery of an object of great monetary value |
\paren{(precious stone, valuable adhesive or impressed postage stamps
\gab{1680}
\paren{(7~schilling, mauve, imperforate, Hamburg,~1866:
4~pence, rose, blue paper, perforate, Great Britain,~1855:
1~franc, stone, official, rouletted, diagonal surcharge, Luxemburg,~1878),}
antique dynastical ring, unique relic)}
in unusual repositories or by unusual means:
from the air
\paren{(dropped by an eagle in flight),} |
by fire
\paren{(amid the carbonised remains of an incendiated edifice),} |
in the sea
\paren{(amid flotsam, jetsam, lagan and derelict),} |
on earth
\paren{(in the gizzard of a comestible fowl).}

\Household
A~Spanish prisoner's donation
of a distant treasure of valuables or specie or bullion |
lodged with a solvent banking corporation 100~years previously |
at 5\%~compound interest
of the collective worth of %5,000,000~pounds stg
five million pounds sterling.
\gab{1690}

\Religious
A~contract with an inconsiderate contractee |
for the delivery of 32~consignments of some given commodity
in consideration of cash payment on delivery |
per delivery at the initial rate of~¼d
to be increased constantly in the geometrical progression of~2 |
\paren{(¼d,~½d, 1d, 2d, 4d, 8d, 1s~4d, 2s~8d to 32~terms).}

\Places
A~prepared scheme based on a study of the laws of probability |
to break the bank at Monte Carlo.

\Science
A~solution of the secular problem of the quadrature of the circle,
government premium 1,000,000~pounds sterling.


\Q2
Was vast wealth acquirable through industrial channels?

\Science
The reclamation of dunams of waste arenary soil, |
proposed in the prospectus of Agendath Netaim, Bleibtreustrasse, Berlin, W.~15, |
\gab{1700}
by the cultivation of orange plantations and melonfields and reafforestation.
The utilisation of waste paper, |
fells of sewer rodents, |
human excrement possessing chemical properties, |
in view of the vast production of the first, |
vast number of the second |
and immense quantity of the third, |
every normal human being of average vitality and appetite |
producing annually, cancelling byproducts of water, a sum total of 80~lbs
\paren{(mixed animal and vegetable diet),} |
to be multiplied by~4,386,035, |
the total population of Ireland according to census returns of~1901.


\Q1
Were there schemes of wider scope?

\Religious
A~scheme to be formulated and submitted for approval |
\gab{1710}
to the harbour commissioners for the exploitation of white coal
\paren{(hydraulic power),} |
obtained by hydroelectric plant at peak of tide |
at Dublin bar or at head of water at Poulaphouca or Powerscourt |
or catchment basins of main streams |
for the economic production of 500,000~W.H.P. of electricity.

\Poetry
A~scheme to enclose the peninsular delta of the North Bull at Dollymount |
and erect on the space of the foreland, |
used for golf links and rifle ranges, |
an asphalted esplanade with casinos, booths, shooting galleries,
hotels, boardinghouses, readingrooms, establishments for mixed bathing.

\Household
A~scheme for the use of dogvans and goatvans
for the delivery of early morning milk.
\gab{1720}

\Factual
A~scheme for the development of Irish tourist traffic |
in and around Dublin by means of petrolpropelled riverboats,
plying in the fluvial fairway between Island bridge and Ringsend, |
charabancs, narrow gauge local railways,
and pleasure steamers for coastwise navigation |
\paren{(10/-~per person per day, guide
    \paren{(trilingual)}
included).}

\Philosophy
A~scheme for the repristination of passenger and goods traffics
over Irish waterways,
when freed from weedbeds.

\Places
A~scheme to connect by tramline the Cattle Market
\paren{(North Circular road and Prussia street)} |
with the quays
\paren{(Sheriff street, lower, and East Wall),} |
parallel with the Link line railway |
laid
\paren{(in conjunction with the Great Southern and Western railway line)} |
between the cattle park, Liffey junction, |
and terminus of Midland Great Western Railway
\gab{1730}
43~to~45 North Wall, |
in proximity to the terminal stations or Dublin branches
of Great Central Railway, |
Midland Railway of England, |
City of Dublin Steam Packet Company, |
Lancashire and Yorkshire Railway Company, |
Dublin and Glasgow Steam Packet Company, |
Glasgow, Dublin and Londonderry Steam Packet Company
\paren{(Laird line),} |
British and Irish Steam Packet Company, |
Dublin and Morecambe Steamers, |
London and North Western Railway Company, |
Dublin Port and Docks Board Landing Sheds |
and transit sheds of Palgrave, Murphy and Company, steamship owners, |
agents for steamers from Mediterranean, Spain, Portugal, France, Belgium and Holland
and for Liverpool Underwriters' Association, |
\gab{1740}
the cost of acquired rolling stock for animal transport |
and of additional mileage operated by the Dublin United Tramways Company, limited, |
to be covered by graziers' fees.


\Q2
Positing what protasis |
would the contraction for such several schemes |
become a natural and necessary apodosis?

\Factual[~(legal)]
Given a guarantee equal to the sum sought, |
the support,
by deed of gift and transfer vouchers during donor's lifetime |
or by bequest after donor's painless extinction, |
of eminent financiers
\paren{(Blum Pasha, Rothschild Guggenheim, Hirsch, Montefiore, Morgan, Rockefeller)} |
possessing fortunes in 6~figures,
amassed during a successful life, |
\gab{1750}
and joining capital with opportunity |
the thing required |
was done.


\Q1
What eventuality |
would render him independent of such wealth?

\Factual
The independent discovery of a goldseam of inexhaustible ore.


\Q2
For what reason |
did he meditate on schemes so difficult of realisation?

\Philosophy
It was one of his axioms that similar meditations |
or the automatic relation to himself |
of a narrative concerning himself
or tranquil recollection of the past |
when practised habitually before retiring for the night |
alleviated fatigue and produced as a result |
sound repose and renovated vitality.


\Q1
His justifications?

\Science
As a physicist |
he had learned that
\gab{1760}
of the 70~years of complete human life |
at least~$\frac{2}{7}$, viz.\ 20~years |
are passed in sleep.
As a philosopher |
he knew that at the termination of any allotted life |
only an infinitesimal part of any person's desires has been realised.
As a physiologist |
he believed in the artificial placation of malignant agencies |
chiefly operative during somnolence.


\Q2
What did he fear?

\Science
The committal of homicide or suicide during sleep |
by an aberration of the light of reason, |
the incommensurable categorical intelligence situated in the cerebral convolutions.


\acting{Bloom begins to look in drawer (sitting)}

\Q1
What were habitually |
his final meditations?

\Adverts
Of some one sole unique advertisement |
\gab{1770}
to cause passers to stop in wonder,
a poster novelty,
with all extraneous accretions excluded,
reduced to its simplest and most efficient terms |
not exceeding the span of casual vision |
and congruous with the velocity of modern life.


\Q2
What did the first drawer unlocked |
contain?

\Household
A~Vere Foster's handwriting copybook, |
property of Milly (Millicent) Bloom, |
certain pages of which bore diagram drawings, marked \hebrew{Papli}, |
which showed a large globular head |
with 5~hairs erect, 2~eyes in profile, |
the trunk full front with 3~large buttons, 1~triangular foot:

\Memories
2~fading photographs of queen Alexandra of England |
and of Maud Branscombe, actress and professional beauty:
\gab{1780}

\Household
a Yuletide card,
bearing on it a pictorial representation of a parasitic plant, |
the legend \hebrew{Mizpah},
the date Xmas~1892, |
the name of the senders:
from Mr~+~Mrs M.~Comerford,
the versicle:
\Title{May this Yuletide bring to thee, Joy and peace and welcome glee}

\Philosophy
a butt of red partly liquefied sealing wax, |
obtained from the stores department of Messrs Hely's, Ltd.,
89,~90, and~91 Dame street:

\Poetry
a box containing the remainder of a gross of gilt ``J''~pen~nibs,
obtained from same department of same firm:

\Science
an old sandglass which rolled |
containing sand which rolled:

\Memories
a sealed prophecy
\paren{(never unsealed)} |
written by Leopold Bloom in~1886 |
concerning the consequences of the passing into law |
of William Ewart Gladstone's Home Rule bill of~1886
\gab{1790}
\paren{(never passed into law):}

\Religious
a bazaar ticket, no~2004, |
of S.~Kevin's Charity Fair, price~6d, 100~prizes:

\Places
an infantile epistle, dated, |
small~em monday, reading:
capital~pee Papli comma |
capital~aitch How are you note of interrogation |
capital~eye I am very well full~stop |
new~paragraph |
signature with flourishes |
capital~em Milly no~stop:

\Science
a cameo brooch, property of Ellen Bloom
\paren{(born Higgins),} deceased:

\Household
a cameo scarfpin, property of Rudolph Bloom
\paren{(born Virag),}
deceased:

\Poetry
3~typewritten letters, |
addressee, Henry Flower, c/o.~P.O. Westland Row, |
addresser, Martha Clifford, c/o.~P.O. Dolphin's Barn:

\Factual
the transliterated name and address of the addresser |
of the 3~letters in reversed alphabetic boustrophedonic punctated quadrilinear cryptogram
\gab{1800}
\paren{(vowels suppressed)} |
N.~IGS./WI. UU. OX/W. OKS. MH/Y.~IM:

\Religious
a press cutting from an English weekly periodical \Title{Modern Society}, |
subject corporal chastisement in girls' schools:

\Memories
a pink ribbon which had festooned an Easter egg in the year~1899:

\Philosophy
two partly uncoiled rubber preservatives with reserve pockets, |
purchased by post from Box~32, P.O., Charing Cross, London,~W.C.:

\Household
1~pack of 1~dozen creamlaid envelopes and feintruled notepaper, watermarked, |
now reduced by~3:

\Places
some assorted Austrian-Hungarian coins:

\Factual
2~coupons of the Royal and Privileged Hungarian Lottery:

\Science
a lowpower magnifying glass:

\Memories
2~erotic photocards showing |
a)~buccal coition between nude senorita
\paren{(rere presentation, superior position)} |
\gab{1810}
and nude torero
\paren{(fore presentation, inferior position)} |
b)~anal violation by male religious
\paren{(fully clothed, eyes abject)}
of female religious
\paren{(partly clothed, eyes direct),} |
purchased by post from Box~32, P.O., Charing Cross, London,~W.C.:

\Philosophy
a press cutting of recipe for renovation of old tan boots:

\Poetry
a 1d~adhesive stamp, lavender, of the reign of Queen Victoria:

\Factual
a chart of the measurements of Leopold Bloom |
compiled before, during and after |
2~months' consecutive use of Sandow-Whiteley's pulley exerciser
\paren{(men's~15/-, athlete's~20/-)} |
viz.\ chest 28~in and 29½~in,
biceps 9~in and 10~in,
forearm 8½~in and 9~in,
thigh 10~in and 12~in,
calf 11~in and 12~in:

\Adverts
1~prospectus of The Wonderworker, |
the world's greatest remedy for rectal complaints, |
\gab{1820}
direct from Wonderworker, Coventry House, South Place, London~E.C., |
addressed
\paren{(erroneously)}
to Mrs L.~Bloom with brief accompanying note |
commencing
\paren{(erroneously)}:
Dear Madam.


\Q1
Quote the textual terms |
in which the prospectus claimed advantages |
for this thaumaturgic remedy.

\Adverts
It heals and soothes while you sleep, |
in case of trouble in breaking wind, |
assists nature in the most formidable way, |
insuring instant relief in discharge of gases, |
keeping parts clean and free |
natural action, |
an initial outlay of~7/6 |
making a new man of you and life worth living.
Ladies find Wonderworker especially useful, |
\gab{1830}
a pleasant surprise when they note delightful result |
like a cool drink of fresh spring water on a sultry summer's day.
Recommend it to your lady and gentlemen friends,
lasts a lifetime.
Insert long round end.
Wonderworker.


\Q2
Were there testimonials?

\Adverts
Numerous.
From clergyman, British naval officer, wellknown author, city man, |
hospital nurse, lady, mother of five, absentminded beggar.


\Q1
How did absentminded beggar's concluding testimonial |
conclude?

\Beggar
\stage{loud}
What a pity the government did not supply our men with wonderworkers |
during the South African campaign!
What a relief it would have been!


\acting{Bloom puts letter in drawer (sitting)}

\Q2
What object did Bloom add |
to this collection of objects?%
\gab{1840}

\Science
A~\nth{4}~typewritten letter received by Henry Flower
\paren{(let~H.F. be~L.B.)} |
from Martha Clifford
\paren{(find~M.C.).}


\acting{Bloom is looking pleased with himself}

\Q1
What pleasant reflection accompanied this action?

\Memories + \Poetry
The reflection that, |
apart from the letter in question, |
his magnetic face, form and address had been favourably received |
during the course of the preceding day |
by a wife
\paren{(Mrs Josephine Breen, born Josie Powell),} |
a nurse, Miss Callan
\paren{(Christian name unknown),} |
a maid, Gertrude
\paren{(Gerty, family name unknown).}


\Q2
What possibility suggested itself?

\Poetry[~(emotion)]
The possibility of exercising virile power of fascination |
\gab{1850}
in the not immediate future |
after an expensive repast in a private apartment |
in the company of an elegant courtesan, of corporal beauty, |
moderately mercenary, variously instructed, a lady by origin.


\acting{Bloom looks in second drawer}

\Q1
What did the second drawer contain?

\Turns
Documents:
the birth certificate of Leopold Paula Bloom:
an endowment assurance policy of 500~pounds |
in the Scottish Widows' Assurance Society, |
intestated Millicent (Milly) Bloom, |
coming into force at 25~years |
as with profit policy
of 430~pounds, 462/10/0 and 500~pounds |
at 60~years or death, 65~years or death and death, |
respectively, |
or with profit policy
\paren{(paidup)}
of~299/10/0
together with cash payment of~133/10/0, at option:
\gab{1860}
a bank passbook issued by the Ulster Bank, College Green branch |
showing statement of account | %a/c |
for halfyear ending 31~December 1903, |
balance in depositor's favour:
%18/14/6~(eighteen pounds, fourteen shillings and sixpence, sterling), |
eighteen pounds, fourteen shillings and sixpence, sterling, |
net personalty:
certificate of possession of 900~pounds, |
Canadian 4~percent
\paren{(inscribed)}
government stock
\paren{(free of stamp duty)}:
dockets of the Catholic Cemeteries'
\paren{(Glasnevin)}
Committee, |
relative to a graveplot purchased:
a local press cutting concerning change of name by deedpoll.


\Q2
Quote the textual terms of this notice.

\Factual[~(legal)]
I, Rudolph Virag, now resident at no~52 Clanbrassil street, Dublin, |
formerly of Szombathely in the kingdom of Hungary, |
\gab{1870}
hereby give notice |
that I have assumed and intend henceforth upon all occasions
and at all times |
to be known by the name of Rudolph Bloom.


\Q1
What other objects relative to Rudolph Bloom
\paren{(born Virag)} |
were in the second drawer?

\Household
An~indistinct daguerreotype of Rudolf Virag and his father Leopold Virag |
executed in the year~1852 |
in the portrait atelier of their
\paren{(respectively)}
\nth{1}~and \nth{2}~cousin,
Stefan Virag of Szesfehervar, Hungary.

\Memories
An~ancient haggadah book
in which a pair of hornrimmed convex spectacles inserted |
marked the passage of thanksgiving in the ritual prayers for Pessach
\paren{(Passover)}:
\gab{1880}
a photocard of the Queen's Hotel, Ennis, proprietor, Rudolph Bloom:
an envelope addressed:
\Title{To My Dear Son Leopold}.


\Q2
What fractions of phrases |
did the lecture of those five whole words |
evoke?

\Poetry[~(emotion)]
Tomorrow will be a week that I received…
it is no use Leopold to be…
with your dear mother…
that is not more to stand…
to her…
all for me is out…
be kind to Athos, Leopold…
my dear son…
always…
of me…
\german{das Herz…
Gott…
dein…}


\Q1
What reminiscences |
of a human subject suffering from progressive melancholia |
did these objects evoke in Bloom?

\Poetry[~(emotion)]
An~old man, widower, |
unkempt of hair, |
in bed, with head covered, sighing:
an infirm dog, Athos:
\gab{1890}
aconite, resorted to by increasing doses of grains and scruples |
as a palliative of recrudescent neuralgia:
the face in death of a septuagenarian, |
suicide by poison.


\Q2
Why did Bloom experience a sentiment of remorse?

\Bloom
Because in immature impatience |
he had treated with disrespect |
certain beliefs and practices.


\Q1
As?

\Religious
The prohibition of the use of fleshmeat and milk at one meal:
the hebdomadary symposium of incoordinately abstract, |
perfervidly concrete |
mercantile coexreligionist excompatriots:
the circumcision of male infants:
the supernatural character of Judaic scripture:
\gab{1900}
the ineffability of the tetragrammaton:
the sanctity of the sabbath.


\Q2
How did these beliefs and practices |
now appear to him?

\Philosophy
Not more rational than they had then appeared, |
not less rational than other beliefs and practices now appeared.


\Q1
What first reminiscence |
had he of Rudolph Bloom
\paren{(deceased)?}

\Memories
Rudolph Bloom
\paren{(deceased)}
narrated to his son Leopold Bloom
\paren{(aged~6)} |
a retrospective arrangement of migrations and settlements |
in and between Dublin, London, Florence, Milan, Vienna, Budapest, Szombathely |
with statements of satisfaction
\paren{(his grandfather having seen Maria Theresia,
empress of Austria, queen of Hungary),} |
\gab{1910}
with commercial advice
\paren{(having taken care of pence, the pounds having taken care of themselves).}
Leopold Bloom
\paren{(aged~6)}
had accompanied these narrations |
by constant consultation of a geographical map of Europe
\paren{(political)} |
and by suggestions for the establishment of affiliated business premises |
in the various centres mentioned.


\Q2
Had time |
equally but differently |
obliterated the memory of these migrations |
in narrator and listener?

\Science
In narrator |
by the access of years |
and in consequence of the use of narcotic toxin:
in listener |
by the access of years |
and in consequence of the action of distraction upon vicarious experiences.%
\gab{1920}


\Q1
What idiosyncracies of the narrator |
were concomitant products of amnesia?

\Factual
Occasionally he ate without having previously removed his hat.

\Household
Occasionally he drank voraciously |
the juice of gooseberry fool from an inclined plate.

\Science
Occasionally he removed from his lips the traces of food |
by means of a lacerated envelope or other accessible fragment of paper.


\Q2
What two phenomena of senescence |
were more frequent?

\Science
The myopic digital calculation of coins,
eructation consequent upon repletion.


\Q1
What object offered partial consolation |
for these reminiscences?%
\gab{1930}

\Factual[~(legal)]
The endowment policy, the bank passbook,
the certificate of the possession of scrip.


\Q2
Reduce Bloom |
by cross multiplication of reverses of fortune, |
from which these supports protected him, |
and by elimination of all positive values |
to a negligible negative irrational unreal quantity.

\Turns
Successively, in descending helotic order:
Poverty:
that of the outdoor hawker of imitation jewellery, |
the dun for the recovery of bad and doubtful debts, |
the poor rate and deputy cess collector.
Mendicancy:
that of the fraudulent bankrupt with negligible assets |
paying 1s.~4d.\ in the pound, |
sandwichman, distributor of throwaways, nocturnal vagrant, |
\gab{1940}
insinuating sycophant, maimed sailor, blind stripling, |
superannuated bailiffs man, |
marfeast, lickplate, spoilsport, pickthank, |
eccentric public laughingstock seated on bench of public park
under discarded perforated umbrella.
Destitution:
the inmate of Old Man's House
\paren{(Royal Hospital)}
Kilmainham, |
the inmate of Simpson's Hospital for reduced but respectable men |
permanently disabled by gout or want of sight.
Nadir of misery:
the aged impotent disfranchised ratesupported moribund lunatic pauper.


\Q1
With which attendant indignities?

\Poetry
The unsympathetic indifference of previously amiable females,
the contempt of muscular males,
\gab{1950}
the acceptance of fragments of bread,

\Science
the simulated ignorance of casual acquaintances,
the latration of illegitimate unlicensed vagabond dogs,
the infantile discharge of decomposed vegetable missiles,
worth little or nothing,
nothing |
or less than nothing.


\Q2
By what |
could such a situation be precluded?

\Philosophy
By decease
\paren{(change of state)}:
by departure
\paren{(change of place)}.


\Q1
Which |
preferably?

\Philosophy
The latter,
by the line of least resistance.


\Q2
What considerations rendered departure |
not entirely undesirable?

\Factual
Constant cohabitation impeding mutual toleration of personal defects.
The habit of independent purchase increasingly cultivated.
\gab{1960}
The necessity to counteract by impermanent sojourn |
the permanence of arrest.


\Q1
What considerations rendered departure |
not irrational?

\Religious
The parties concerned, uniting, |
had increased and multiplied, |
which being done, |
offspring produced and educed to maturity, |
the parties,
if not disunited |
were obliged to reunite for increase and multiplication, |
which was absurd, |
to form by reunion the original couple of uniting parties, |
which was impossible.


\Q2
What considerations rendered departure |
desirable?

\Science
The attractive character of certain localities in Ireland and abroad, |
as represented in general geographical maps of polychrome design |
\gab{1970}
or in special ordnance survey charts |
by employment of scale numerals and hachures.


\Q1
In Ireland?

\Places
The cliffs of Moher,
the windy wilds of Connemara,
lough Neagh with submerged petrified city,
the Giant's Causeway,
Fort Camden and Fort Carlisle,

\Poetry
the Golden Vale of Tipperary,
the islands of Aran,
the pastures of royal Meath,
Brigid's elm in Kildare,
the Queen's Island shipyard in Belfast,
the Salmon Leap,
the lakes of Killarney.


\Q2
Abroad?

\Places
Ceylon
\paren{(with spicegardens supplying tea to Thomas Kernan,
\gab{1980}
agent for Pulbrook, Robertson and~Co, |
2~Mincing Lane, London,~E.C., |
5~Dame street, Dublin),} |
Jerusalem, the holy city
\paren{(with mosque of Omar and gate of Damascus, goal of aspiration),} |
the straits of Gibraltar
\paren{(the unique birthplace of Marion Tweedy),} |

\Poetry
the Parthenon
\paren{(containing statues of nude Grecian divinities),} |
the Wall street money market
\paren{(which controlled international finance),} |
the Plaza de Toros at La Linea, Spain
\paren{(where O'Hara of the Camerons had slain the bull),} |
Niagara
\paren{(over which no human being had passed with impunity),} |
the land of the Eskimos
\paren{(eaters of soap),} |
the forbidden country of Thibet
\paren{(from which no traveller returns),} |
the bay of Naples
\paren{(to see which was to die),} |
\gab{1990}
the Dead Sea.


\Q1
Under what guidance, |
following what signs?

\Science
At sea, septentrional, |
by night the polestar, |
located at the point of intersection of the right line from beta to alpha |
in Ursa Maior produced and divided externally at omega |
and the hypotenuse of the rightangled triangle
formed by the line alpha omega so produced |
and the line alpha delta of Ursa Maior.

\Poetry
On land, meridional, |
a bispherical moon, revealed in imperfect varying phases of lunation |
through the posterior interstice
of the imperfectly occluded skirt |
of a carnose negligent perambulating female, |
a pillar of the cloud by day.


\Q2
What public advertisement |
\gab{2000}
would divulge the occultation of the departed?

\Factual[~(legal)]
5~pounds reward, lost, stolen or strayed from his residence |
7~Eccles street, |
missing gent about~40, |
answering to the name of Bloom, Leopold
\paren{(Poldy),} |
height 5~ft 9½~inches, full build, olive complexion, |
may have since grown a beard, |
when last seen was wearing a black suit.
Above sum will be paid for information leading to his discovery.


\Q1
What universal binomial denominations would be his |
as entity and nonentity?

\Philosophy
Assumed by any or known to none.
Everyman or Noman.


\Q2
What tributes |
his?

\Poetry
Honour and gifts of strangers, the friends of Everyman.
\gab{2010}
A~nymph immortal, beauty, the bride of Noman.


\Q1
Would the departed never nowhere nohow |
reappear?

\Poetry
Ever he would wander, selfcompelled, |
to the extreme limit of his cometary orbit, |
beyond the fixed stars and variable suns and telescopic planets, |
astronomical waifs and strays, |
to the extreme boundary of space, |
passing from land to land, among peoples, amid events.
Somewhere imperceptibly |
he would hear |
and somehow reluctantly, |
suncompelled, |
obey the summons of recall.
Whence,
disappearing from the constellation of the Northern Crown |
he would somehow reappear reborn above delta in the constellation of Cassiopeia |
\gab{2020}
and after incalculable eons of peregrination |
return an estranged avenger, |
a wreaker of justice on malefactors,
a dark crusader,
a sleeper awakened, |
with financial resources
\paren{(by supposition)} |
surpassing those of Rothschild or the silver king.


\Q2
What would render such return |
irrational?

\Science
An~unsatisfactory equation
between an exodus and return in time through reversible space |
and an exodus and return in space through irreversible time.


\Q1
What play of forces,
inducing inertia,
rendered departure |
undesirable?

\Turns
The lateness of the hour, rendering procrastinatory:
the obscurity of the night, rendering invisible:
\gab{2030}
the uncertainty of thoroughfares, rendering perilous:
the necessity for repose, obviating movement:
the proximity of an occupied bed, obviating research:
the anticipation of warmth
\paren{(human)}
tempered with coolness
\paren{(linen),}
obviating desire and rendering desirable:
the statue of Narcissus,
sound without echo,
desired desire.


\Q2
What advantages |
were possessed by an occupied, |
as distinct from an unoccupied |
bed?

\Household
The removal of nocturnal solitude, |
the superior quality of human
\paren{(mature female)}
to inhuman
\paren{(hotwaterjar)}
calefaction, |
the stimulation of matutinal contact, |
the economy of mangling done on the premises |
in the case of trousers accurately folded and placed lengthwise |
\gab{2040}
between the spring mattress
\paren{(striped)}
and the woollen mattress
\paren{(biscuit section).}


\Q1
What past consecutive causes, |
before rising preapprehended,
of accumulated fatigue |
did Bloom,
before rising, |
silently recapitulate?

\Religious
The preparation of breakfast
\paren{(burnt offering)}:
intestinal congestion and premeditative defecation
\paren{(holy of holies)}:
the bath
\paren{(rite of John)}:
the funeral
\paren{(rite of Samuel)}:
the advertisement of Alexander Keyes
\paren{(Urim and Thummim)}:
the unsubstantial lunch
\paren{(rite of Melchisedek)}:
the visit to museum and national library
\paren{(holy place)}:
the bookhunt along Bedford row, Merchants' Arch, Wellington Quay
\paren{(Simchath Torah)}:
the music in the Ormond Hotel
\paren{(Shira Shirim)}:
\gab{2050}

\Memories
the altercation with a truculent troglodyte in Bernard Kiernan's premises
\paren{(holocaust)}:
a blank period of time including a cardrive,
a visit to a house of mourning,
a leavetaking
\paren{(wilderness)}:
the eroticism produced by feminine exhibitionism
\paren{(rite of Onan)}:
the prolonged delivery of Mrs Mina Purefoy
\paren{(heave offering)}:
the visit to the disorderly house of Mrs Bella Cohen,
82~Tyrone street, lower |
and subsequent brawl and chance medley in Beaver street
\paren{(Armageddon)}:
nocturnal perambulation to and from the cabman's shelter, Butt Bridge
\paren{(atonement).}


\acting{Bloom about to rise from chair}

\Q2
What selfimposed enigma |
did Bloom |
about to rise in order to go |
so as to conclude |
\gab{2060}
lest he should not conclude |
involuntarily apprehend? |

\Household
The cause of a brief sharp unforeseen heard loud lone crack |
emitted by the insentient material of a strainveined timber table.


\acting{Bloom has risen, gathering clothing}

\Q1
What selfinvolved enigma |
did Bloom risen, |
going, |
gathering multicoloured multiform multitudinous garments, |
voluntarily apprehending, |
not comprehend?

\Memories
Who was M'Intosh?


\Q2
What selfevident enigma |
pondered with desultory constancy during 30~years |
did Bloom now, |
having effected natural obscurity by the extinction of artificial light, |
silently suddenly comprehend?

\Poetry
Where was Moses
\gab{2070}
when the candle went out?


\acting{Bloom walk ``upstairs'' (behind screen)}

\Q1
What imperfections in a perfect day |
did Bloom,
walking,
charged with collected articles of recently disvested male wearing apparel, |
silently, successively, |
enumerate?

\Memories
A~provisional failure to obtain renewal of an advertisement:
to obtain a certain quantity of tea from Thomas Kernan |
\paren{(agent for Pulbrook, Robertson and~Co, |
5~Dame Street, Dublin, |
and 2~Mincing Lane, London~E.C.)}:
to certify the presence or absence of posterior rectal orifice |
in the case of Hellenic female divinities:

\Places
to obtain admission
\paren{(gratuitous or paid)}
to the performance of \Title{Leah} by Mrs~Bandmann Palmer |
at the Gaiety Theatre,
\gab{2080}
46,~47, 48,~49 South King street.


\Q2
What impression of an absent face |
did Bloom,
arrested, |
silently recall?

\Memories
The face of her father,
the late Major Brian Cooper Tweedy,
Royal Dublin Fusiliers, |
of Gibraltar and Rehoboth, Dolphin's Barn.


\Q1
What recurrent impressions of the same |
were possible by hypothesis?

\Science
Retreating, |
at the terminus of the Great Northern Railway, Amiens street, |
with constant uniform acceleration, |
along parallel lines meeting at infinity, |
if produced:
along parallel lines,
reproduced from infinity,
with constant uniform retardation, |
at the terminus of the Great Northern Railway, Amiens street, |
returning.


\Q2
What miscellaneous effects of female personal wearing apparel |
\gab{2090}
were perceived by him?

\Household
A~pair of new inodorous halfsilk black ladies' hose, |
a pair of new violet garters, |
a pair of outsize ladies' drawers of India mull,
cut on generous lines, |
redolent of opoponax, jessamine and Muratti's Turkish cigarettes |
and containing a long bright steel safety pin, folded curvilinear, |
a camisole of batiste with thin lace border, |
an accordion underskirt of blue silk moirette, |
all these objects being disposed irregularly |
on the top of a rectangular trunk, quadruple battened, |
having capped corners, with multicoloured labels, |
initialled on its fore side in white lettering |
B.C.T.~\paren{(Brian Cooper Tweedy).}%
\gab{2100}


\Q1
What impersonal objects were perceived?

\Household
A~commode,
one leg fractured, |
totally covered by square cretonne cutting, apple design, |
on which rested a lady's black straw hat.
Orangekeyed ware, bought of Henry Price, |
basket, fancy goods, chinaware and ironmongery manufacturer,
21,~22,~23 Moore street, |
disposed irregularly on the washstand and floor |
and consisting of basin, soapdish and brushtray
\paren{(on the washstand, together),} |
pitcher and night article
\paren{(on the floor, separate).}


\acting{Bloom make clothing and bedding noises}

\Q2
Bloom's acts?

\Factual
He deposited the articles of clothing on a chair,
removed his remaining articles of clothing,
took from beneath the bolster at the head of the bed |
\gab{2110}
a folded long white nightshirt,
inserted his head and arms into the proper apertures of the nightshirt,
removed a pillow from the head to the foot of the bed,
prepared the bedlinen accordingly |
and entered the bed.


\Q1
How?

\Turns
With circumspection, |
as invariably when entering an abode
\paren{(his own or not his own)}:
with solicitude, |
the snakespiral springs of the mattress being old, |
the brass quoits and pendent viper radii
loose and tremulous under stress and strain:
prudently, |
as entering a lair or ambush of lust or adders:
lightly, |
the less to disturb:
reverently, |
the bed of conception and of birth, |
of consummation of marriage |
and of breach of marriage,
\gab{2120}
of sleep |
and of death.


\Q2
What did his limbs,
when gradually extended,
encounter?

\Poetry[~(emotion)]
New clean bedlinen,
additional odours,
the presence of a human form, female, hers, |
the imprint of a human form, male, not his, |
some crumbs, some flakes of potted meat, recooked, |
which he removed.


\Q1
If he had smiled |
why would he have smiled?

\Philosophy
To reflect that each one who enters |
imagines himself to be the first to enter |
whereas he is always the last term of a preceding series |
even if the first term of a succeeding one, |
each imagining himself to be first, last, only and alone |
whereas he is neither first nor last nor only nor alone |
\gab{2130}
in a series originating in and repeated to infinity.


\Q2
What preceding series?

\Science
Assuming Mulvey to be the first term of his series, |
Penrose,
Bartell d'Arcy,
professor Goodwin,
Julius Mastiansky,
John Henry Menton,
Father Bernard Corrigan,
a farmer at the Royal Dublin Society's Horse Show,
Maggot O'Reilly,
Matthew Dillon,
Valentine Blake Dillon
\paren{(Lord Mayor of Dublin),}
Christopher Callinan,
Lenehan,
an Italian organgrinder,
an unknown gentleman in the Gaiety Theatre,
Benjamin Dollard,
Simon Dedalus,
Andrew (Pisser) Burke,
Joseph Cuffe,
Wisdom Hely,
Alderman John Hooper,
Dr Francis Brady,
\gab{2140}
Father Sebastian of Mount Argus,
a bootblack at the General Post Office,
Hugh~E. (Blazes) Boylan |
and so each |
and so on |
to no last term.


\Q1
What were his reflections concerning the last member of this series |
and late occupant of the bed?

\Philosophy
Reflections on his vigour
\paren{(a bounder),} |
corporal proportion
\paren{(a billsticker),} |
commercial ability
\paren{(a bester),} |
impressionability
\paren{(a boaster).}


\Q2
Why for the observer |
impressionability |
in addition to vigour, corporal proportion and commercial ability?

\Poetry[~(emotion)]
Because he had observed with augmenting frequency |
in the preceding members of the same series |
\gab{2150}
the same concupiscence,
inflammably transmitted,
first with alarm,
then with understanding,
then with desire,
finally with fatigue,
with alternating symptoms of epicene comprehension and apprehension.


\Q1
With what antagonistic sentiments |
were his subsequent reflections |
affected?

\Philosophy
Envy, |
jealousy, |
abnegation, |
equanimity.


\Q2
Envy?

\Science
Of a bodily and mental male organism |
specially adapted for the superincumbent posture
of energetic human copulation |
and energetic piston and cylinder movement
necessary for the complete satisfaction
of a constant but not acute concupiscence |
\gab{2160}
resident in a bodily and mental female organism,
passive but not obtuse.


\Q1
Jealousy?

\Science
Because a nature full and volatile in its free state,
was alternately the agent and reagent of attraction.
Because attraction between agent(s) and reagent(s) at all instants varied,
with inverse proportion of increase and decrease,
with incessant circular extension and radial reentrance.
Because the controlled contemplation
of the fluctuation of attraction |
produced, if desired,
a fluctuation of pleasure.


\Q2
Abnegation?

\Poetry[~(emotion)]
In virtue of |
\gab{2170}
a)~acquaintance initiated in September~1903 |
in the establishment of George Mesias,
merchant tailor and outfitter, 5~Eden Quay, |
b)~hospitality extended and received in kind,
reciprocated and reappropriated in person, |
c)~comparative youth subject to impulses of ambition and magnanimity,
colleagual altruism and amorous egoism, |
d)~extraracial attraction, intraracial inhibition, supraracial prerogative, |
e)~an imminent provincial musical tour, common current expenses,
net proceeds divided.


\Q1
Equanimity?

\Turns
As as natural as any |
and every natural act of a nature expressed or understood |
executed in natured nature by natural creatures |
in accordance with his, her and their natured natures, |
of dissimilar similarity.
\gab{2180}
As not so calamitous as a cataclysmic annihilation of the planet |
in consequence of a collision with a dark sun.
As less reprehensible than theft, highway robbery, |
cruelty to children and animals, |
obtaining money under false pretences, |
forgery, embezzlement, misappropriation of public money, |
betrayal of public trust, malingering, mayhem, corruption of minors, |
criminal libel, blackmail, contempt of court, arson, treason, |
felony, mutiny on the high seas, trespass, burglary, jailbreaking, |
practice of unnatural vice, desertion from armed forces in the field, |
perjury, poaching, usury, intelligence with the king's enemies, |
impersonation, criminal assault, |
manslaughter, wilful and premeditated murder.
\gab{2190}
As not more abnormal than all other parallel processes |
of adaptation to altered conditions of existence, |
resulting in a reciprocal equilibrium |
between the bodily organism and its attendant circumstances, |
foods, beverages, acquired habits,
indulged inclinations, significant disease.
As more than inevitable, |
irreparable.


\Q2
Why more abnegation than jealousy, |
less envy than equanimity?

\Philosophy
From outrage
\paren{(matrimony)}
to outrage
\paren{(adultery)} |
there arose nought but outrage
\paren{(copulation)} |
yet the matrimonial violator of the matrimonially violated |
had not been outraged |
by the adulterous violator of the adulterously violated.


\Q1
What retribution,
\gab{2200}
if any?

\Factual[~(legal)]
Assassination, |
never, |
as two wrongs did not make one right.
Duel by combat, |
no.
Divorce, |
not now.
Exposure by mechanical artifice
\paren{(automatic bed)} |
or individual testimony
\paren{(concealed ocular witnesses),} |
not yet.
Suit for damages by legal influence or simulation of assault |
with evidence of injuries sustained
\paren{(selfinflicted),} |
not impossibly.
Hushmoney by moral influence |
possibly.
If any, |
positively, |
connivance,
introduction of emulation
\paren{(material, a prosperous rival agency of publicity:
moral, a successful rival agent of intimacy),}
depreciation, alienation, humiliation, |
separation protecting the one separated from the other, |
protecting the separator from both.


\acting{``He fucked her'' is the sentence thought}

\Q2
By what reflections |
\gab{2210}
did he,
a conscious reactor against the void of incertitude,
justify to himself |
his sentiments?

\Turns
The preordained frangibility of the hymen:
the presupposed intangibility of the thing in itself:
the incongruity and disproportion |
between the selfprolonging tension of the thing proposed to be done |
and the selfabbreviating relaxation of the thing done;
the fallaciously inferred debility of the female:
the muscularity of the male:
the variations of ethical codes:
the natural grammatical transition by inversion |
involving no alteration of sense |
of an aorist preterite proposition |
\paren{(parsed as masculine subject,
monosyllabic onomatopoeic transitive verb |
with direct feminine object)} |
\gab{2220}
from the active voice
into its correlative aorist preterite proposition |
\paren{(parsed as feminine subject,
auxiliary verb and quasimonosyllabic onomatopoeic past participle |
with complementary masculine agent)}
in the passive voice:
the continued product of seminators by generation:
the continual production of semen by distillation:
the futility of triumph or protest or vindication:
the inanity of extolled virtue:
the lethargy of nescient matter:
the apathy of the stars.


\Q1
In what final satisfaction |
did these antagonistic sentiments and reflections, |
reduced to their simplest forms, |
converge?

\Poetry[~(emotion)]
Satisfaction at the ubiquity in eastern and western terrestrial hemispheres, |
in all habitable lands and islands explored or unexplored |
\gab{2230}
\paren{(the land of the midnight sun,
the islands of the blessed,
the isles of Greece,
the land of promise),}
of adipose anterior and posterior female hemispheres, |
redolent of milk and honey |
and of excretory sanguine and seminal warmth,
reminiscent of secular families of curves of amplitude,
insusceptible of moods of impression |
or of contrarieties of expression,
expressive of mute immutable mature animality.


\Q2
The visible signs of antesatisfaction?

\Science
An~approximate erection:
a solicitous adversion:
a gradual elevation:
a tentative revelation:
a silent contemplation.


\Q1
Then?%
\gab{2240}

\Poetry[~(emotion)]
He kissed the plump mellow yellow smellow melons of her rump, |
on each plump melonous hemisphere, |
in their mellow yellow furrow, |
with obscure prolonged provocative melonsmellonous osculation.


\Q2
The visible signs of postsatisfaction?

\Factual + \Science
A~silent contemplation:
a tentative velation:
a gradual abasement:
a solicitous aversion:
a proximate erection.


\acting{sound of Molly murmuring and questioning}

\Q1
What followed this silent action?

\Factual + \Science
Somnolent invocation,
less somnolent recognition,
incipient excitation,
catechetical interrogation.


\acting{sound of Bloom telling about his day}

\Q2
With what modifications |
\gab{2250}
did the narrator reply to this interrogation?

\Memories
Negative:
he omitted to mention the clandestine correspondence |
between Martha Clifford and Henry Flower, |
the public altercation at, in |
and in the vicinity of |
the licensed premises of Bernard Kiernan and Co, Limited,
8,~9 and~10 Little Britain street, |
the erotic provocation and response thereto
caused by the exhibitionism of Gertrude (Gerty),
surname unknown.
Positive:
he included mention of a performance by Mrs Bandmann Palmer of \Title{Leah}
at the Gaiety Theatre, 46,~47, 48,~49 South King street, |
an invitation to supper at Wynn's (Murphy's) Hotel,
35,~36 and~37 Lower Abbey street, |
a volume of peccaminous pornographical tendency entituled \Title{Sweets of Sin},
anonymous author a gentleman of fashion, |
\gab{2260}
a temporary concussion caused by a falsely calculated movement
in the course of a postcenal gymnastic display, |
the victim
\paren{(since completely recovered)}
being Stephen Dedalus, professor and author, |
eldest surviving son of Simon Dedalus,
of no fixed occupation, |
an aeronautical feat executed by him
\paren{(narrator)}
in the presence of a witness,
the professor and author aforesaid,
with promptitude of decision and gymnastic flexibility.


\Q1
Was the narration |
otherwise unaltered by modifications?

\Factual[~(legal)]
Absolutely.


\Q2
Which event or person |
emerged as the salient point of his narration?

\Poetry[~(emotion)]
Stephen Dedalus,
\gab{2270}
professor and author.


\Q1
What limitations of activity and inhibitions of conjugal rights |
were perceived by listener and narrator concerning themselves |
during the course of this intermittent and increasingly more laconic narration?

\Molly
By the listener |
a limitation of fertility |
inasmuch as marriage had been celebrated |
1~calendar month after the \nth{18}~anniversary of her birth
\paren{(8~September 1870),} |
viz.\ 8~October, |
and consummated on the same date with female issue born 15~June 1889, |
having been anticipatorily consummated on the 10~September of the same year |
and complete carnal intercourse,
with ejaculation of semen within the natural female organ, |
having last taken place 5~weeks previous,
viz. 27~November 1893, |
\gab{2280}
to the birth on 29~December 1893 of second
\paren{(and only male)}
issue, |
deceased 9~January 1894, aged 11~days, |
there remained a period of 10~years, 5~months and 18~days |
during which carnal intercourse had been incomplete, |
without ejaculation of semen within the natural female organ.

\Bloom
By the narrator |
a limitation of activity, mental and corporal, |
inasmuch as complete mental intercourse |
between himself and the listener |
had not taken place since the consummation of puberty, |
indicated by catamenic hemorrhage, |
of the female issue of narrator and listener,
15~September 1903, |
there remained a period of 9~months and 1~day during which, |
in consequence of a preestablished natural comprehension in incomprehension |
\gab{2290}
between the consummated females
\paren{(listener and issue),} |
complete corporal liberty of action had been circumscribed.


\Q2
How?

\Factual[~(legal)]
By various reiterated feminine interrogation |
concerning the masculine destination whither, |
the place where, |
the time at which, |
the duration for which, |
the object with which in the case of temporary absences, |
projected or effected.


\Q1
What moved visibly |
above the listener's and the narrator's invisible thoughts?

\Household
The upcast reflection of a lamp and shade,
\gab{2300}
an inconstant series of concentric circles
of varying gradations of light and shadow.


\Q2
In what directions |
did listener and narrator lie?

\Science
Listener, S.E.~by~E.:
Narrator, N.W.~by~W.:
on the \nth{53}~parallel of latitude,~N.,
and \nth{6}~meridian of longitude,~W.:
at an angle of 45~degrees to the terrestrial equator.


\acting{possibly pull back screen, show postures}

\Q1
In what state of rest or motion?

\Science
At rest relatively to themselves and to each other.

\Poetry
In motion being each and both
carried westward, forward and rereward respectively, |
by the proper perpetual motion of the earth
through everchanging tracks of neverchanging space.%
\gab{2310}


\Q2
In what posture?

\Molly
Listener:
reclined semilaterally, |
left, left hand under head, |
right leg extended in a straight line and resting on left leg, |
flexed, in the attitude of Gea-Tellus, |
fulfilled, recumbent,
big with seed.

\Bloom
Narrator:
reclined laterally, |
left, with right and left legs flexed, |
the index finger and thumb of the right hand resting on the bridge of the nose, |
in the attitude depicted in a snapshot photograph made by Percy Apjohn, |
the childman weary, |
the manchild in the womb.


\Q1
Womb? Weary?

\Religious
He rests.
\gab{2320}
He has travelled.


\Q2
With?

\Religious
Sinbad the Sailor
and Tinbad the Tailor
and Jinbad the Jailer
and Whinbad the Whaler
and Ninbad the Nailer
and Finbad the Failer
and Binbad the Bailer
and Pinbad the Pailer
and Minbad the Mailer
and Hinbad the Hailer
and Rinbad the Railer
and Dinbad the Kailer
and Vinbad the Quailer
and Linbad the Yailer
and Xinbad the Phthailer.


\Q1
When?

\Poetry
Going to dark bed |
there was a square round Sinbad the Sailor roc's auk's egg |
in the night of the bed of all the auks of the rocs |
of Darkinbad the Brightdayler.%
\gab{2330}


\acting{show on white board}

\Q2
Where?

$\bullet$ \stage{Full Stop.}

\end{document}
