\subsection{[Roger as Bloom]}

\BloomInt:
Pineapple rock, lemon platt, butter scotch.
A sugarsticky girl |
shovelling scoopfuls of creams for a christian brother.
Some school treat.
Bad for their tummies.
Lozenge and comfit manufacturer to His Majesty the King.
God.
Save.
Our.
Sitting on his throne |
sucking red jujubes |
white.

\N1:
A sombre Y.M.C.A. young man,
watchful among the warm sweet fumes of Graham Lemon's,
placed a throwaway in a hand of Mr Bloom.

\N2:
Heart to heart talks.

\BloomInt:
Bloo ...
Me?
No.

\BloomInt:
Blood of the Lamb.


\N1:
His slow feet walked him riverward, reading.
\gab{10}

\N2:
Are you saved?
All are washed in the blood of the lamb.

\BloomInt:
God wants blood victim.
Birth, hymen, martyr, war,
foundation of a building,
sacrifice,
kidney burntoffering,
druids' altars.

\N2:
Elijah is coming.
Dr John Alexander Dowie |
restorer of the church in Zion |
is coming.

\N2:
\Quote{%
    Is coming! Is coming!! Is coming!!!\\
    All heartily welcome.}

\BloomInt:
Paying game.
Torry and Alexander last year.
Polygamy.
His wife will put the stopper on that.
Where was that ad |
some Birmingham firm |
the luminous crucifix.
Our Saviour.
Wake up in the dead of night |
and see him on the wall, hanging.
Pepper's ghost idea.
\gab{20}
Iron nails ran in.

\BloomInt:
Phosphorus |
it must be done with.
If you leave a bit of codfish for instance.
I could see the bluey silver over it.
Night I went down to the pantry in the kitchen.
Don't like all the smells in it |
waiting to rush out.
What was it she wanted?
The Malaga raisins.
Thinking of Spain.
Before Rudy was born.
The phosphorescence, that bluey greeny.
Very good for the brain.

\N1:
From Butler's monument house corner |
he glanced along Bachelor's walk.

\BloomInt:
Dedalus' daughter there |
still outside Dillon's auctionrooms.
Must be selling off some old furniture.
Knew her eyes at once |
from the father.
Lobbing about |
waiting for him.
\gab{30}
Home always breaks up when the mother goes.
Fifteen children he had.
Birth every year almost.
That's in their theology |
or the priest won't give the poor woman
the confession,
the absolution.
Increase and multiply.
Did you ever hear such an idea?
Eat you out of house and home.
No families themselves to feed.
Living on the fat of the land.
Their butteries and larders.
I'd like to see them do the black fast |
Yom Kippur.
Crossbuns.
One meal and a collation |
for fear he'd collapse on the altar.
A housekeeper of one of those fellows |
if you could pick it out of her.
Never pick it out of her.
Like getting l.s.d.\@ \stage{[pounds, shilling, and pence]} out of him.
Does himself well.
No guests.
All for number one.
Watching his water.
Bring your own bread and butter.
\gab{40}
His reverence:
mum's the word.

\BloomInt:
Good Lord, that poor child's dress is in flitters.
Underfed she looks too.
Potatoes and marge,
marge and potatoes.
It's after |
they feel it.
Proof of the pudding.
Undermines the constitution.

\N1:
As he set foot on O'Connell bridge |
a puffball of smoke plumed up from the parapet.

\BloomInt:
Brewery barge with export stout.
England.
Sea air sours it, I heard.
Be interesting some day |
get a pass through Hancock |
to see the brewery.
Regular world in itself.
Vats of porter |
wonderful.
Rats get in too.
Drink themselves |
bloated as big as a collie |
floating.
Dead drunk on the porter.
Drink till they puke again like christians.
Imagine drinking that!
Rats: vats.
\gab{50}
Well, of course,
if we knew all the things.

\N1:
Looking down |
he saw flapping strongly,
wheeling between the gaunt quaywalls,
gulls.

\BloomInt:
Rough weather outside.
If I threw myself down?
Reuben~J's son |
must have swallowed a good bellyful of that sewage.
One and eightpence too much.
Hhhhm.
It's the droll way |
he comes out with the things.
Knows how to tell a story too.

\subsection{[George as Bloom]}

\N1:
They wheeled lower.

\BloomInt:
Looking for grub.
Wait.

\N1:
He threw down among them |
a crumpled paper ball.

\BloomInt:
Elijah |
thirtytwo feet per sec |
is com.
Not a bit.

\N1:
The ball bobbed unheeded on the wake of swells,
floated under by the bridgepiers.

\BloomInt:
Not such damn fools.
Also the day I threw that stale cake out of the Erin's King |
\gab{60}
picked it up in the wake fifty yards astern.
Live by their wits.

\N1:
They wheeled,
flapping.

\BloomInt:
\Quote{%
    The hungry famished gull \\
    flaps o'er the waters dull.}

\BloomInt:
That is how poets write, the similar sounds.
But then Shakespeare has no rhymes:
blank verse.
The flow of the language it is.
The thoughts.
Solemn.

\BloomInt:
\Quote{%
    Hamlet, I am thy father's spirit\\
    Doomed for a certain time to walk the earth.}

\Woman1:
Two apples a penny!
Two for a penny!

\N1:
His gaze passed over the glazed apples |
\gab{70}
serried on her stand.

\BloomInt:
Australians they must be |
this time of year.
Shiny peels:
polishes them up
with a rag or a handkerchief.

\BloomInt:
Wait.
Those poor birds.

\N1:
He halted again and bought from the old applewoman |
two Banbury cakes for a penny |
and broke the brittle paste
and threw its fragments down into the Liffey.

\BloomInt:
See that?

\N1:
The gulls swooped silently,
two,
then all from their heights,
pouncing on prey.

\BloomInt:
Gone.
Every morsel.

\N1:
Aware of their greed and cunning |
he shook the powdery crumb from his hands.

\BloomInt:
They never expected that.
Manna.
Live on fish,
fishy flesh they have,
all seabirds,
gulls,
seagoose.
Swans from Anna Liffey swim down here sometimes
\gab{80}
to preen themselves.
No accounting for tastes.
Wonder what kind is swanmeat.
Robinson Crusoe had to live on them.

\N1:
They wheeled |
flapping weakly.

\BloomInt:
I'm not going to throw any more.
Penny quite enough.
Lot of thanks I get.
Not even a caw.
They spread foot and mouth disease too.
If you cram a turkey |
say on chestnutmeal |
it tastes like that.
Eat pig like pig.
But then why is it that saltwater fish are not salty?
How is that?

\N1:
His eyes sought answer from the river |
and saw a rowboat rock at anchor |
on the treacly swells |
lazily its plastered board.

\N2:
\Quote{%
    Kino's%
    \gab{90}\\
    11/- \stage{[eleven shillings]}\\
    Trousers}


\BloomInt:
Good idea that.
Wonder if he pays rent to the corporation.
How can you own water really?
It's always flowing in a stream,
never the same,
which in the stream of life we trace.
Because life is a stream.
All kinds of places are good for ads.
That quack doctor for the clap used to be stuck up |
in all the greenhouses.
Never see it now.
Strictly confidential.
Dr Hy Franks.
Didn't cost him a red |
like Maginni the dancing master |
self advertisement.
Got fellows to stick them up
or stick them up himself |
for that matter |
on the q.t. |
running in to loosen a button.
\gab{100}
Flybynight.
Just the place too.
\emph{Post no bills.
Post no pills.}
Some chap with a dose burning him.

\BloomInt:
If he ...?

\BloomInt:
O!

\BloomInt:
Eh?

\BloomInt:
No ...
No.

\BloomInt:
No, no.
I don't believe it.
He wouldn't surely?

\BloomInt:
No, no.

\N1:
Mr Bloom moved forward, raising his troubled eyes.

\BloomInt:
Think no more about that.
After one.
Timeball on the ballast\-office is down.
Dunsink time.
Fascinating little book that is of sir Robert Ball's.
\gab{110}
Parallax.
I never exactly understood.
There's a priest.
Could ask him.
Par | it's Greek:
parallel, parallax.
Met him pike hoses |
she called it |
till I told her about the transmigration.
O rocks!

\N1:
Mr Bloom smiled

\BloomInt:
O rocks

\N1:
at two windows of the ballast\-office.

\BloomInt:
She's right |
after all.
Only big words for ordinary things |
on account of the sound.
She's not exactly witty.
Can be rude too.
Blurt out what I was thinking.
Still, I don't know.
She used to say Ben Dollard had a base barreltone voice.
He has legs like barrels |
and you'd think he was singing into a barrel.
Now, isn't that wit.
They used to call him big Ben.
Not half as witty as calling him base barreltone.
\gab{120}
Appetite like an albatross.
Get outside of a baron of beef.
Powerful man he was |
at stowing away number one Bass.
Barrel of Bass.
See?
It all works out.

\subsection{[Iain as Bloom]}

\N2:
A procession of whitesmocked sandwichmen |
marched slowly towards him |
along the gutter,
scarlet sashes across their boards.

\BloomInt:
Bargains.
Like that priest they are |
this morning:
we have sinned:
we have suffered.

\N2:
He read the scarlet letters on their five tall white hats:

\BloomInt:
H.E.L.Y.S\@.
\stage{[H—E—L—Y—apostrophe~S]}
Wisdom Hely's.

\N2:
Y |
lagging behind |
drew a chunk of bread from under his foreboard,
crammed it into his mouth |
and munched as he walked.

\BloomInt:
Our staple food.
Three bob a day,
walking along the gutters,
street after street.
Just keep skin and bone together,
bread and skilly.
\gab{130}
They are not Boyl:
no, M'Glade's men.
Doesn't bring in any business either.
I suggested to him about a transparent showcart
with two smart girls sitting inside |
writing letters, copybooks, envelopes, blottingpaper.
I bet that would have caught on.
Smart girls writing something |
catch the eye at once.
Everyone dying to know what she's writing.
Get twenty of them round you |
if you stare at nothing.
Have a finger in the pie.
Women too.
Curiosity.
Pillar of salt.
Wouldn't have it of course |
because he didn't think of it himself first.
Or the inkbottle I suggested |
with a false stain of black celluloid.
His ideas for ads |
like Plumtree's potted under the obituaries,
cold meat department.
You can't lick 'em.
What?
Our envelopes.
\gab{140}
Hello, Jones, where are you going?
Can't stop, Robinson,
I am hastening to purchase the only reliable ink\-eraser \emph{Kansell},
sold by Hely's Ltd,
85~Dame street.
Well out of that ruck |
I~am.
Devil of a job it was |
collecting accounts of those convents.
Tranquilla convent.
That was a nice nun there,
really sweet face.
Wimple suited her small head.
Sister?
Sister?
I am sure she was crossed in love |
by her eyes.
Very hard to bargain with that sort of a woman.
I disturbed her at her devotions that morning.
But glad to communicate with the outside world.
Our great day,
she said.
Feast of Our Lady of Mount Carmel.
Sweet name too:
caramel.
She knew I,
I think she knew |
by the way she.
If she had married |
she would have changed.
\gab{150}
I suppose they really were short of money.
Fried everything in the best butter |
all the same.
No lard for them.
My heart's broke |
eating dripping.
They like buttering themselves |
in and out.
Molly tasting it,
her veil up.
Sister?
Pat Claffey,
the pawnbroker's daughter.
It was a nun they say |
invented barbed wire.

\N2:
He crossed Westmoreland street |
when apostrophe~S had plodded by.

\BloomInt:
Rover cycleshop.
Those races are on today.
How long ago is that?
Year Phil Gilligan died.
We were in Lombard street west.
Wait:
was in Thom's.
Got the job in Wisdom Hely's |
year we married.
Six years.
Ten years ago:
ninetyfour he died |
yes that's right |
the big fire at Arnott's.
Val Dillon was lord mayor.
The Glencree dinner.
\gab{160}
Alderman Robert O'Reilly emptying the port into his soup |
before the flag fell.
Bobbob lapping it for the inner alderman.
Couldn't hear what the band played.
For what we have already received |
may the Lord make us.
Milly was a kiddy then.
Molly had that elephantgrey dress with the braided frogs.
Mantailored with selfcovered buttons.
She didn't like it |
because I sprained my ankle |
first day she wore |
choir picnic at the Sugarloaf.
As if that.
Old Goodwin's tall hat |
done up with some sticky stuff.
Flies' picnic too.
Never put a dress on her back like it.
Fitted her like a glove,
shoulders and hips.
Just beginning to plump it out well.
Rabbitpie we had that day.
People looking after her.


\BloomInt:
Happy.
\gab{170}
Happier then.
Snug little room that was |
with the red wallpaper.
Dockrell's, one and ninepence a dozen.
Milly's tubbing night.
American soap I bought:
elderflower.
Cosy smell of her bathwater.
Funny she looked |
soaped all over.
Shapely too.
Now photography.
Poor papa's daguerreotype atelier |
he told me of.
Hereditary taste.

\N2:
He walked along the curbstone.

\BloomInt:
Stream of life.
What was the name of that priestylooking chap |
was always squinting in |
when he passed?
Weak eyes,
woman.
Stopped in Citron's |
saint Kevin's parade.
Pen something.
Pendennis?
My memory is getting.
Pen ...?
Of course it's years ago.
Noise of the trams probably.
Well, if he couldn't remember the dayfather's name |
\gab{180}
that he sees every day.

\BloomInt:
Bartell d'Arcy was the tenor,
just coming out then.
Seeing her home after practice.
Conceited fellow with his waxedup moustache.
Gave her that song
\Title{Winds that blow from the south}.

\BloomInt:
Windy night that was |
I went to fetch her there |
was that lodge meeting on |
about those lottery tickets |
after Goodwin's concert |
in the supperroom or oakroom of the Mansion house.
He and I behind.
Sheet of her music blew out of my hand
against the High school railings.
Lucky it didn't.
Thing like that |
spoils the effect of a night for her.
Professor Goodwin linking her in front.
Shaky on his pins,
poor old sot.
His farewell concerts.
Positively last appearance on any stage.
\gab{190}
May be for months and may be for never.
Remember her laughing at the wind,
her blizzard collar up.
Corner of Harcourt road |
remember that gust.
Brrfoo!
Blew up all her skirts |
and her boa nearly smothered old Goodwin.
She did get flushed in the wind.
Remember when we got home |
raking up the fire |
and frying up those pieces of lap of mutton for her supper |
with the Chutney sauce she liked.
And the mulled rum.
Could see her in the bedroom from the hearth |
unclamping the busk of her stays:
white.

\BloomInt:
Swish and soft flop |
her stays made on the bed.
Always warm from her.
Always liked to let her self out.
Sitting there after till near two |
taking out her hairpins.
\gab{200}
Milly tucked up in beddyhouse.
Happy.
Happy.
That was the night ...

\subsection{[Mark as Bloom]}

\josie:
O, Mr Bloom, how do you do?

\Bloom:
O, how do you do, Mrs Breen?

\josie:
No use complaining.
How is Molly those times?
Haven't seen her for ages.

\Bloom:
In the pink,
\sout{Mr Bloom said gaily.}
Milly has a position down in Mullingar, you know.

\josie:
Go away!
Isn't that grand for her?

\Bloom:
Yes.
In a photographer's there.
Getting on like a house on fire.
How are all your charges?%
\gab{210}

\josie:
All on the baker's list,
\sout{Mrs Breen said.}

\BloomInt:
How many has she?
No other in sight.

\josie:
You're in black, I see.
You have no ...

\Bloom:
No,
\sout{Mr Bloom said.}
I have just come from a funeral.

\BloomInt:
Going to crop up all day, I foresee.
Who's dead,
when and what did he die of?
Turn up like a bad penny.

\josie:
O, dear me,
\sout{Mrs Breen said.}
I hope it wasn't any near relation.

\BloomInt:
May as well get her sympathy.

\Bloom:
Dignam,
\sout{Mr Bloom said.}
An old friend of mine.
He died quite suddenly,
poor fellow.
Heart trouble, I believe.
\gab{220}
Funeral was this morning.

\BloomInt:
\song{%
    Your funeral's tomorrow\\
    While you're coming through the rye.\\
    Diddlediddle dumdum\\
    Diddlediddle ...}

\josie:
Sad to lose the old friends,

\N1:
Mrs Breen's womaneyes said melancholily.

\BloomInt:
Now that's quite enough about that.
Just:
quietly:
husband.

\Bloom:
And your lord and master?

\N2:
Mrs Breen turned up her two large eyes.

\BloomInt:
Hasn't lost them anyhow.

\josie:
O, don't be talking!
\sout{she said.}
He's a caution to rattlesnakes.
He's in there now with his lawbooks |
finding out the law of libel.
\gab{230}
He has me heartscalded.
Wait till I show you.

\N2:
Hot mockturtle vapour and steam of newbaked jampuffs rolypoly |
poured out from Harrison's.
The heavy noonreek tickled the top of Mr Bloom's gullet.

\BloomInt:
Want to make good pastry, |
butter, best flour, Demerara sugar, |
or they'd taste it with the hot tea.
Or is it from her?

\N1:
A barefoot arab stood over the grating,
breathing in the fumes.

\BloomInt:
Deaden the gnaw of hunger that way.
Pleasure or pain is it?
Penny dinner.
Knife and fork
chained to the table.

\BloomInt:
Opening her handbag, chipped leather.
Hatpin:
ought to have a guard on those things.
Stick it in a chap's eye in the tram.
\gab{240}
Rummaging.
Open.
Money.
Please take one.
Devils if they lose sixpence.
Raise Cain.
Husband barging.
Where's the ten shillings I gave you on Monday?
Are you feeding your little brother's family?
Soiled handkerchief:
medicinebottle.
Pastille that was fell.
What is she?
...

\josie:
There must be a new moon out,
\sout{she said.}
He's always bad then.
Do you know what he did last night?

\N1:
Her hand ceased to rummage.
Her eyes fixed themselves on him,
wide in alarm,
yet smiling.

\Bloom:
What?
\sout{Mr Bloom asked.}

\BloomInt:
Let her speak.
Look straight in her eyes.
\gab{250}
I believe you.
Trust me.

\josie:
Woke me up in the night,
\sout{she said.}
Dream he had,
a nightmare.

\BloomInt:
Indiges.

\josie:
Said the ace of spades was walking up the stairs.

\Bloom:
The ace of spades!
\sout{Mr Bloom said.}

\N1:
She took a folded postcard from her handbag.

\josie:
Read that,
\sout{she said.}
He got it this morning.

\Bloom:
What is it?

\N2:
Mr Bloom asked,
taking the card.

\Bloom:
U.P.?

\josie:
U.P.: up,
\sout{she said.}
Someone taking a rise out of him.
It's a great shame for them |
whoever he is.

\Bloom:
Indeed it is,
\gab{260}
\sout{Mr Bloom said.}

\N1:
She took back the card,
sighing.

\josie:
And now he's going round to Mr Menton's office.
He's going to take an action for ten thousand pounds,
he says.

\N1:
She folded the card into her untidy bag and snapped the catch.

\BloomInt:
Same blue serge dress she had two years ago,
the nap bleaching.
Seen its best days.
Wispish hair over her ears.
And that dowdy toque:
three old grapes to take the harm out of it.
Shabby genteel.
She used to be a tasty dresser.
Lines round her mouth.
Only a year or so older than Molly.

\BloomInt:
See the eye that woman gave her,
passing.
Cruel.
The unfair sex.

\N2:
He looked still at her,
\gab{270}
holding back behind his look |
his discontent.

\BloomInt:
Pungent mockturtle oxtail mulligatawny.
I'm hungry too.
Flakes of pastry on the gusset of her dress:
daub of sugary flour stuck to her cheek.
Rhubarb tart with liberal fillings,
rich fruit interior.
Josie Powell that was.
In Luke Doyle's long ago.
Dolphin's Barn,
the charades.
U.P.:
up.

\BloomInt:
Change the subject.

\Bloom:
Do you ever see anything of Mrs Beaufoy?
\sout{Mr Bloom asked.}

\josie:
Mina Purefoy?
\sout{she said.}

\BloomInt:
Philip Beaufoy I was thinking.
Playgoers' Club.
Matcham often thinks of the masterstroke.
Did I pull the chain?
Yes.
The last act.

\Bloom:
Yes.
\gab{280}

\josie:
I just called to ask on the way in |
is she over it.
She's in the lying-in hospital in Holles street.
Dr Horne got her in.
She's three days bad now.

\Bloom:
O,
\sout{Mr Bloom said.}
I'm sorry to hear that.

\josie:
Yes,
\sout{Mrs Breen said.}
And a houseful of kids at home.
It's a very stiff birth,
the nurse told me.

\Bloom:
O,
\sout{Mr Bloom said.}

\N2:
His heavy pitying gaze absorbed her news.
His tongue clacked in compassion.

\Bloom:
Dth!
Dth!

\Bloom:
I'm sorry to hear that,
\sout{he said.}
Poor thing!
Three days!
That's terrible for her.%
\gab{290}

\N1:
Mrs Breen nodded.

\josie:
She was taken bad on the Tuesday ...

\N2:
Mr Bloom touched her funnybone gently,
warning her:

\Bloom:
Mind!
Let this man pass.

\N1:
A bony form strode along the curbstone |
from the river |
staring with a rapt gaze into the sunlight |
through a heavystringed glass.
Tight as a skullpiece |
a tiny hat gripped his head.
From his arm |
a folded dustcoat, a stick and an umbrella |
dangled to his stride.

\Bloom:
Watch him,
\sout{Mr Bloom said.}
He always walks outside the lampposts.
Watch!%
\gab{300}

\josie:
Who is he |
if it's a fair question?
\sout{Mrs Breen asked.}
Is he dotty?

\Bloom:
His name is Cashel Boyle O'Connor Fitzmaurice Tisdall Farrell,
\sout{Mr Bloom said smiling.}
Watch!

\josie:
He has enough of them,
\sout{she said.}
Denis will be like that |
one of these days.

\N1:
She broke off suddenly.

\josie:
There he is,
\sout{she said.}
I must go after him.
Goodbye.
Remember me to Molly,
won't you?

\Bloom:
I will,
\sout{Mr Bloom said.}

\subsection{[Irene as Bloom]}

\N2:
He watched her dodge through passers towards the shopfronts.

\N1:
Denis Breen in skimpy frockcoat and blue canvas shoes |
\gab{310}
shuffled out of Harrison's |
hugging two heavy tomes to his ribs.

\BloomInt:
Blown in from the bay.
Like old times.

\N1:
He suffered her to overtake him without surprise |
and thrust his dull grey beard towards her,
his loose jaw wagging as he spoke earnestly.

\BloomInt:
Meshuggah.
Off his chump.

\N2:
Mr Bloom walked on again easily,
seeing ahead of him in sunlight |
the tight skullpiece,
the dangling stickumbrelladustcoat.

\BloomInt:
Going the two days.
Watch him!
Out he goes again.
One way of getting on in the world.
And that other old mosey lunatic in those duds.
Hard time |
she must have with him.

\BloomInt:
U.P.:
up.
\gab{320}
I'll take my oath |
that's Alf Bergan or Richie Goulding.
Wrote it for a lark in the Scotch house |
I bet anything.
Round to Menton's office.
His oyster eyes staring at the postcard.
Be a feast for the gods.

\N1:
He passed the \Title{Irish Times}.

\BloomInt:
There might be other answers lying there.
Like to answer them all.
Good system for criminals.
Code.
At their lunch now.
Clerk with the glasses there |
doesn't know me.
O, leave them there to simmer.
Enough bother |
wading through fortyfour of them.
Wanted,
smart lady typist |
to aid gentleman in literary work.
I called you naughty darling |
because I do not like that other world.
Please tell me what is the meaning.
Please tell me what perfume does your wife.
Tell me who made the world.
The way they spring those questions on you.
\gab{330}
And the other one |
Lizzie Twigg.
My literary efforts have had the good fortune |
to meet with the approval of the eminent poet~A.E.\@
(Mr Geo.\@ Russell).
No time to do her hair |
drinking sloppy tea with a book of poetry.

\BloomInt:
Best paper |
by long chalks |
for a small ad.
Got the provinces now.
Cook and general,
exc.\@ \stage{[excellent]} cuisine,
housemaid kept.
Wanted live man for spirit counter.
Resp.\@ \stage{[respectable]} girl (R.C.\@ \stage{[Roman Catholic]})
wishes to hear of post in fruit or pork shop.
James Carlisle made that.
Six and a half per cent dividend.
Made a big deal on Coates's shares.
Ca'canny.
Cunning old Scotch hunks.
All the toady news.
Our gracious and popular vicereine.
Bought the \Title{Irish Field} now.
Lady Mountcashel has quite recovered after her confinement |
\gab{340}
and rode out with the Ward Union staghounds |
at the enlargement yesterday |
at Rathoath.
Uneatable fox.
Pothunters too.
Fear injects juices |
make it tender enough for them.
Riding astride.
Sit her horse like a man.
Weight\-carrying huntress.
No sidesaddle or pillion for her,
not for Joe.
First to the meet |
and in at the death.
Strong as a brood mare |
some of those horsey women.
Swagger around livery stables.
Toss off a glass of brandy neat |
while you'd say knife.
That one at the Grosvenor this morning.
Up with her on the car:
wish\-swish.
Stonewall or fivebarred gate |
put her mount to it.
Think that pugnosed driver did it out of spite.
Who is this she was like?
O yes!
Mrs Miriam Dandrade |
that sold me
\gab{350}
her old wraps and black underclothes |
in the Shelbourne hotel.
Divorced Spanish American.
Didn't take a feather out of her |
my handling them.
As if I was her clotheshorse.
Saw her in the viceregal party |
when Stubbs the park ranger got me in |
with Whelan of the \Title{Express}.
Scavenging what the quality left.
High tea.
Mayonnaise |
I poured on the plums |
thinking it was custard.
Her ears ought to have tingled for a few weeks after.
Want to be a bull for her.
Born courtesan.
No nursery work for her,
thanks.

\BloomInt:
Poor Mrs Purefoy!
Methodist husband.
Method in his madness.
Saffron bun and milk and soda lunch |
in the educational dairy.
Y.M.C.A\@.
Eating with a stopwatch,
thirtytwo chews to the minute.
\gab{360}
And still his muttonchop whiskers grew.
Supposed to be well connected.
Theodore's cousin in Dublin Castle.
One tony relative in every family.
Hardy annuals |
he presents her with.
Saw him out at the Three Jolly Topers |
marching along bareheaded |
and his eldest boy |
carrying one in a marketnet.
The squallers.
Poor thing!
Then having to give the breast year after year |
all hours of the night.
Selfish those t.t's are.
Dog in the manger.
Only one lump of sugar in my tea,
if you please.

\subsection{[Wendy as Bloom]}

\N1:
He stood at Fleet street crossing.

\BloomInt:
Luncheon interval.
A sixpenny at Rowe's?
Must look up that ad in the national library.
An eightpenny in the Burton.
\gab{370}
Better.
On my way.

\N1:
He walked on past Bolton's Westmoreland house.

\BloomInt:
Tea.
Tea.
Tea.
I forgot to tap Tom Kernan.

\Bloom:
Sss.
Dth, dth, dth!

\BloomInt:
Three days imagine |
groaning on a bed with a vinegared handkerchief round her forehead,
her belly swollen out.
Phew!
Dreadful simply!
Child's head too big:
forceps.
Doubled up inside her |
trying to butt its way out blindly,
groping for the way out.
Kill me that would.
Lucky |
Molly got over hers lightly.
They ought to invent something to stop that.
Life with hard labour.
Twilight sleep idea:
queen Victoria was given that.
Nine |
she had.
A good layer.
Old woman that lived in a shoe |
she had so many children.
\gab{380}
Suppose he was consumptive.
Time someone thought about it |
instead of gassing about the |
what was it |
the pensive bosom of the silver effulgence.
Flapdoodle to feed fools on.
They could easily have big establishments |
whole thing quite painless |
out of all the taxes |
give every child born |
five quid at compound interest |
up to twentyone |
five per cent is a hundred shillings |
and five |
tiresome pounds |
multiply by twenty |
decimal system |
encourage people to put by money |
save hundred and ten and a bit |
twentyone years |
want to work it out on paper |
come to a tidy sum |
more than you think.

\BloomInt:
Not stillborn of course.
They are not even registered.
Trouble for nothing.%
\gab{390}

\BloomInt:
Funny sight |
two of them together,
their bellies out.
Molly and Mrs Moisel.
Mothers' meeting.
Phthisis retires for the time being,
then returns.
How flat they look all of a sudden |
after.
Peaceful eyes.
Weight off their mind.
Old Mrs Thornton was a jolly old soul.
All my babies,
she said.
The spoon of pap in her mouth |
before she fed them.
O, that's nyumnyum.
Got her hand crushed by old Tom Wall's son.
His first bow to the public.
Head like a prize pumpkin.
Snuffy Dr Murren.
People knocking them up at all hours.
For God' sake, doctor.
Wife in her throes.
Then keep them waiting months for their fee.
To attendance on your wife.
No gratitude in people.
Humane doctors,
\gab{400}
most of them.

\N2:
Before the huge high door of the Irish house of parliament |
a flock of pigeons flew.

\BloomInt:
Their little frolic after meals.
Who will we do it on?
I pick the fellow in black.
Here goes.
Here's good luck.
Must be thrilling from the air.
Apjohn, myself and Owen Goldberg |
up in the trees near Goose green |
playing the monkeys.
Mackerel |
they called me.

\N2:
A squad of constables debouched from College street,
marching in Indian file.

\BloomInt:
Goosestep.
Foodheated faces,
sweating helmets,
patting their truncheons.
After their feed |
with a good load of fat soup under their belts.
Policeman's lot is oft a happy one.

\N2:
They split up in groups and scattered,
saluting,
towards their beats.

\BloomInt:
Let out to graze.
\gab{410}
Best moment to attack one in pudding time.
A punch in his dinner.

\N2:
A squad of others,
marching irregularly,
rounded Trinity railings |
making for the station.

\BloomInt:
Bound for their troughs.
Prepare to receive cavalry.
Prepare to receive soup.

\N1:
He crossed under Tommy Moore's roguish finger.

\BloomInt:
They did right to put him up over a urinal:
meeting of the waters.
Ought to be places for women.
Running into cakeshops.
Settle my hat straight.
\Title{There is not in this wide world a vallee}.
Great song of Julia Morkan's.
Kept her voice |
up to the very last.
Pupil of Michael Balfe's,
wasn't she?

\subsection{[Joseph as Bloom]}

\N1:
He gazed after the last broad tunic.

\BloomInt:
Nasty customers to tackle.
Jack Power could a tale unfold:
father a~G~man.
\gab{420}
If a fellow gave them trouble being lagged |
they let him have it |
hot and heavy in the bridewell.
Can't blame them after all |
with the job they have |
especially the young hornies.
That horsepoliceman |
the day Joe Chamberlain was given his degree in Trinity |
he got a run for his money.
My word he did!
His horse's hoofs |
clattering after us down Abbey street.
Lucky |
I had the presence of mind to dive into Manning's |
or I was souped.
He did come a wallop,
by George.
Must have cracked his skull on the cobblestones.
I oughtn't to have got myself swept along with those medicals.
And the Trinity jibs in their mortarboards.
Looking for trouble.
Still |
I got to know that young Dixon |
who dressed that sting for me in the Mater |
\gab{430}
and now he's in Holles street |
where Mrs Purefoy.
Wheels within wheels.
Police whistle in my ears still.
All skedaddled.
Why he fixed on me.
Give me in charge.
Right here it began.

\Man1:
Up the Boers!

\Man2:
Three cheers for De Wet!

\Man3:
We'll hang Joe Chamberlain on a sourapple tree.

\BloomInt:
Silly billies:
mob of young cubs yelling their guts out.
Vinegar hill.
The Butter exchange band.
Few years' time |
half of them magistrates and civil servants.
War comes on:
into the army helterskelter:
same fellows used to.
\gab{440}
Whether on the scaffold high.

\BloomInt:
Never know who you're talking to.
Corny Kelleher |
he has Harvey Duff in his eye.
Like that Peter or Denis or James Carey |
that blew the gaff on the invincibles.
Member of the corporation too.
Egging raw youths on |
to get in the know |
all the time drawing secret service pay from the castle.
Drop him like a hot potato.
Why those plainclothes men are always courting slaveys.
Easily twig a man used to uniform.
Squarepushing up
against a backdoor.
Maul her a bit.
Then the next thing on the menu.
And who is the gentleman does be visiting there?
Was the young master saying anything?
Peeping Tom through the keyhole.
Decoy duck.
Hotblooded young student fooling round |
\gab{450}
her fat arms ironing.

\policeman:
Are those yours,
Mary?

\mary:
I don't wear such things ...
Stop or I'll tell the missus on you.
Out half the night.

\policeman:
There are great times coming,
Mary.
Wait till you see.

\mary:
Ah,
gelong with your great times coming.

\BloomInt:
Barmaids too.
Tobaccoshopgirls.

\BloomInt:
James Stephens' idea was the best.
He knew them.
Circles of ten |
so that a fellow couldn't round on more than his own ring.
Sinn Fein.
Back out |
you get the knife.
Hidden hand.
Stay in.
The firing squad.
Turnkey's daughter got him out of Richmond,
off from Lusk.
\gab{460}
Putting up in the Buckingham Palace hotel |
under their very noses.
Garibaldi.

\BloomInt:
You must have a certain fascination:
Parnell.
Arthur Griffith is a squareheaded fellow |
but he has no go in him for the mob.
Or gas about our lovely land.
Gammon and spinach.
Dublin Bakery Company's tearoom.
Debating societies.
That republicanism |
is the best form of government.
That the language question |
should take precedence of the economic question.
Have your daughters inveigling them to your house.
Stuff them up with meat and drink.
Michaelmas goose.
Here's a good lump of thyme seasoning under the apron for you.
Have another quart of goosegrease before it gets too cold.
\gab{470}
Half\/fed enthusiasts.
Penny roll and a walk with the band.
No grace for the carver.
The thought that the other chap pays |
best sauce in the world.
Make themselves thoroughly at home.
Show us over those apricots,
meaning peaches.
The not far distant day.
Homerule sun rising up in the northwest.

\subsection{[Bill as Bloom]}

\N1:
His smile faded as he walked,
a heavy cloud hiding the sun slowly,
shadowing Trinity's surly front.
Trams passed one another, |
ingoing,
outgoing,
clanging.

\BloomInt:
Useless words.
Things go on same,
day after day:
squads of police marching out,
back:
trams in,
out.
Those two loonies mooching about.
Dignam carted off.
Mina Purefoy |
swollen belly on a bed |
groaning to have a child tugged out of her.
\gab{480}
One born every second |
somewhere.
Other dying every second.
Since I fed the birds |
five minutes.
Three hundred kicked the bucket.
Other three hundred born,
washing the blood off,
all are washed in the blood of the lamb,
bawling maaaaaa.

\BloomInt:
Cityful passing away,
other cityful coming,
passing away too:
other coming on,
passing on.
Houses, lines of houses,
streets,
miles of pavements,
piledup bricks,
stones.
Changing hands.
This owner,
that.
Landlord never dies |
they say.
Other steps into his shoes |
when he gets his notice to quit.
They buy the place up with gold |
and still they have all the gold.
Swindle in it |
somewhere.
Piled up in cities,
worn away age after age.
Pyramids in sand.
Built on bread and onions.
Slaves |
Chinese wall.
\gab{490}
Babylon.
Big stones left.
Round towers.
Rest rubble,
sprawling suburbs,
jerrybuilt.
Kerwan's mushroom houses |
built of breeze.
Shelter,
for the night.

\BloomInt:
No-one is anything.

\BloomInt:
This is the very worst hour of the day.
Vitality.
Dull, gloomy:
hate this hour.
Feel as if I had been eaten and spewed.

\BloomInt:
Provost's house.
The reverend Dr Salmon:
tinned salmon.
Well tinned in there.
Like a mortuary chapel.
Wouldn't live in it |
if they paid me.
Hope they have liver and bacon today.
Nature abhors a vacuum.

\N2:
The sun freed itself slowly |
and lit glints of light among the silverware |
opposite in Walter Sexton's window |
\gab{500}
by which John Howard Parnell passed,
unseeing.

\BloomInt:
There he is:
the brother.
Image of him.
Haunting face.
Now that's a coincidence.
Course |
hundreds of times you think of a person and don't meet him.
Like a man walking in his sleep.
No-one knows him.
Must be a corporation meeting today.
They say he never put on the city marshal's uniform |
since he got the job.
Charley Kavanagh used to come out on his high horse, |
cocked hat, |
puffed, powdered and shaved.
Look at the woebegone walk of him.
Eaten a bad egg.
Poached eyes on ghost.
I have a pain.
Great man's brother:
his brother's brother.
He'd look nice on the city charger.
Drop into the D.B.C.\@ probably |
for his coffee,
\gab{510}
play chess there.
His brother used men as pawns.
Let them all go to pot.
Afraid to pass a remark on him.
Freeze them up with that eye of his.
That's the fascination:
the name.
All a bit touched.
Mad Fanny and his other sister Mrs Dickinson |
driving about with scarlet harness.
Bolt upright |
like surgeon M'Ardle.
Still |
David Sheehy beat him for south Meath.
Apply for the Chiltern Hundreds and retire into public life.
The patriot's banquet.
Eating orangepeels in the park.
Simon Dedalus said |
when they put him in parliament |
that Parnell would come back from the grave |
and lead him out of the house of commons by the arm.

\subsection{[Claudia as Bloom]}

\AErussell:
Of the twoheaded octopus,
\gab{520}
one of whose heads is the head |
upon which the ends of the world have forgotten to come |
while the other speaks with a Scotch accent.
The tentacles ...

\N2:
They passed from behind Mr Bloom along the curbstone.

\BloomInt:
Beard and bicycle.
Young woman.

\BloomInt:
And there he is too.
Now that's really a coincidence:
second time.
Coming events cast their shadows before.
With the approval of the eminent poet,
Mr Geo.\@ Russell.
That might be Lizzie Twigg with him.
A.E.:
what does that mean?
Initials perhaps.
Albert Edward,
Arthur Edmund,
Alphonsus Eb~Ed~El Esquire.
What was he saying?
The ends of the world with a Scotch accent.
\gab{530}
Tentacles:
octopus.
Something occult:
symbolism.
Holding forth.
She's taking it all in.
Not saying a word.
To aid gentleman in literary work.

\N1:
His eyes followed the high figure in homespun,
beard and bicycle, |
a listening woman at his side.

\BloomInt:
Coming from the vegetarian.
Only weggebobbles and fruit.
Don't eat a beefsteak.
If you do |
the eyes of that cow will pursue you through all eternity.
They say it's healthier.
Windandwatery though.
Tried it.
Keep you on the run all day.
Bad as a bloater.
Dreams all night.
Why do they call that thing they gave me |
nutsteak?
Nutarians.
Fruitarians.
To give you the idea |
you are eating rumpsteak.
\gab{540}
Absurd.
Salty too.
They cook in soda.
Keep you sitting by the tap all night.

\BloomInt:
Her stockings are loose over her ankles.
I detest that:
so tasteless.
Those literary etherial people they are all.
Dreamy,
cloudy,
symbolistic.
Esthetes they are.
I wouldn't be surprised |
if it was that kind of food you see |
produces the |
like waves of the brain |
the poetical.
For example |
one of those policemen sweating Irish stew into their shirts |
you couldn't squeeze a line of poetry out of him.
Don't know what poetry is even.
Must be in a certain mood.

\BloomInt:
\song{%
    The dreamy cloudy gull\\
    Waves o'er the waters dull.}%
\gab{550}

\N1:
He crossed at Nassau street corner |
and stood before the window of Yeates and Son,
pricing the fieldglasses.

\BloomInt:
Or will I drop into old Harris's
and have a chat with young Sinclair?
Wellmannered fellow.
Probably at his lunch.
Must get those old glasses of mine set right.
Goerz lenses |
six guineas.
Germans making their way everywhere.
Sell on easy terms to capture trade.
Undercutting.
Might chance on a pair |
in the railway lost property office.
Astonishing the things people leave behind them |
in trains and cloakrooms.
What do they be thinking about?
Women too.
Incredible.
Last year |
travelling to Ennis |
had to pick up that farmer's daughter's bag |
and hand it to her at Limerick junction.
\gab{560}
Unclaimed money too.
There's a little watch up there |
on the roof of the bank |
to test those glasses by.

\N1:
His lids came down on the lower rims of his irides.

\BloomInt:
Can't see it.
If you imagine it's there |
you can almost see it.
Can't see it.

\N1:
He faced about |
and, standing between the awnings, |
held out his right hand at arm's length |
towards the sun.

\BloomInt:
Wanted to try that often.
Yes:
completely.

\N1:
The tip of his little finger blotted out the sun's disk.

\BloomInt:
Must be the focus where the rays cross.
If I had black glasses.
Interesting.
There was a lot of talk about those sunspots |
when we were in Lombard street west.
Looking up from the back garden.
Terrific explosions they are.
There will be a total eclipse this year:
\gab{570}
autumn some time.

\BloomInt:
Now that I come to think of it |
that ball falls at Greenwich time.
It's the clock |
is worked by an electric wire from Dunsink.
Must go out there |
some first Saturday of the month.
If I could get an introduction to professor Joly |
or learn up something about his family.
That would do to:
man always feels complimented.
Flattery where least expected.
Nobleman proud to be descended from some king's mistress.
His foremother.
Lay it on with a trowel.
Cap in hand goes through the land.
Not go in and blurt out |
what you know you're not to:
what's parallax?
Show this gentleman the door.

\BloomInt:
Ah.

\N1:
His hand fell to his side again.
\gab{580}

\BloomInt:
Never know anything about it.
Waste of time.
Gasballs spinning about,
crossing each other,
passing.
Same old dingdong always.
Gas:
then solid:
then world:
then cold:
then dead shell drifting around,
frozen rock,
like that pineapple rock.
The moon.
Must be a new moon out,
she said.
I believe there is.

\subsection{[Lynne as Bloom]}

\N1:
He went on by la Maison Claire.

\BloomInt:
Wait.
The full moon was the night we were |
Sunday fortnight exactly there is a new moon.
Walking down by the Tolka.
Not bad for a Fairview moon.
She was humming.
The young May moon |
she's beaming,
love.
He other side of her.
Elbow, arm.
He.
\gab{590}
Glowworm's la-amp is gleaming, love.
Touch.
Fingers.
Asking.
Answer.
Yes.

\BloomInt:
Stop.
Stop.
If it was |
it was.
Must.

\N1:
Mr Bloom,
quickbreathing,
slowlier walking |
passed Adam court.

\N1:
With a keep quiet relief |
his eyes took note |

\BloomInt:
this is the street here |
middle of the day |
of Bob Doran's bottle shoulders.
On his annual bend,
M'Coy said.
They drink in order to say or do something
or \french{cherchez la femme}.
Up in the Coombe with chummies and streetwalkers |
and then the rest of the year |
sober as a judge.

\BloomInt:
Yes.
Thought so.
Sloping into the Empire.
Gone.
Plain soda would do him good.
Where Pat Kinsella had his Harp theatre |
\gab{600}
before Whitbred ran the Queen's.
Broth of a boy.
Dion Boucicault business |
with his harvestmoon face in a poky bonnet.
Three Purty Maids from School.
How time flies, eh?
Showing long red pantaloons under his skirts.
Drinkers, drinking, laughed spluttering,
their drink against their breath.
More power, Pat.
Coarse red:
fun for drunkards:
guffaw and smoke.
Take off that white hat.
His parboiled eyes.
Where is he now?
Beggar somewhere.
The harp that once did starve us all.

\BloomInt:
I was happier then.
Or was that I?
Or am I now I?
Twentyeight I was.
She twentythree.
When we left Lombard street west |
something changed.
Could never like it again |
after Rudy.
Can't bring back time.
\gab{610}
Like holding water in your hand.
Would you go back to then?
Just beginning then.
Would you?
Are you not happy in your home |
you poor little naughty boy?
Wants to sew on buttons for me.
I must answer.
Write it in the library.

\N1:
Grafton street |
gay with housed awnings |
lured his senses.

\N2:
Muslin prints, silkdames and dowagers,
jingle of harnesses,
hoofthuds lowringing in the baking causeway.

\BloomInt:
Thick feet that woman has |
in the white stockings.
Hope the rain mucks them up on her.
Countrybred chawbacon.
All the beef to the heels were in.
Always gives a woman clumsy feet.
Molly looks out of plumb.

\N1:
He passed, dallying,
\gab{620}
the windows of Brown Thomas, silk mercers.

\N2:
Cascades of ribbons.
Flimsy China silks.
A tilted urn poured from its mouth |
a flood of bloodhued poplin:

\BloomInt:
lustrous blood.
The huguenots brought that here.
\italian{La causa è santa}!
Tara Tara.
Great chorus that.
Taree tara.
Must be washed in rainwater.
Meyerbeer.
Tara:
bom bom bom.

\BloomInt:
Pincushions.
I'm a long time threatening to buy one.
Sticking them all over the place.
Needles in window curtains.

\N1:
He bared slightly his left forearm.

\BloomInt:
Scrape: nearly gone.
Not today anyhow.
Must go back for that lotion.
For her birthday perhaps.
June\-july\-aug\-september eighth.
Nearly three months off.
Then she mightn't like it.
Women won't pick up pins.
\gab{630}
Say it cuts lo.

\N2:
Gleaming silks,
petticoats on slim brass rails,
rays of flat silk stockings.

\BloomInt:
Useless to go back.
Had to be.
Tell me all.

\BloomInt:
High voices.

\N2:
Sunwarm silk.
Jingling harnesses.

\BloomInt:
All for a woman,
home and houses,
silkwebs, silver,
rich fruits spicy from Jaffa.
Agendath Netaim.
Wealth of the world.

\N1:
A warm human plumpness settled down on his brain.
His brain yielded.

\N2:
Perfume of embraces all him assailed.
With hungered flesh obscurely,
he mutely craved to adore.

\BloomInt:
Duke street.
\gab{640}
Here we are.
Must eat.
The Burton.
Feel better then.

\N1:
He turned Combridge's corner, still pursued.

\N2:
Jingling, hoofthuds.
Perfumed bodies, warm, full.
All kissed, yielded:
in deep summer fields,
tangled pressed grass,
in trickling hallways of tenements,
along sofas, creaking beds.

\Woman1:
Jack,
love!

\Woman1:
Darling!

\Woman3:
Kiss me,
Reggy!

\Woman1:
My boy!

\Woman2:
Love!

\subsection{[Leon as Bloom]}

\N1:
His heart astir |
\gab{650}
he pushed in the door of the Burton restaurant.

\N2:
Stink gripped his trembling breath:
pungent meatjuice, slush of greens.

\BloomInt:
See the animals feed.

\BloomInt:
Men, men, men.

\N2:
Perched on high stools by the bar,
hats shoved back,
at the tables |
calling for more bread no charge,
swilling,
wolfing gobfuls of sloppy food,
their eyes bulging,
wiping wetted moustaches.
A pallid suetfaced young man |
polished his tumbler knife fork and spoon with his napkin.

\BloomInt:
New set of microbes.

\N2:
A man with an infant's saucestained napkin tucked round him |
shovelled gurgling soup down his gullet.
A man spitting back on his plate:

\BloomInt:
halfmasticated gristle:
gums:
no teeth to chewchewchew it.
\gab{660}
Chump chop from the grill.
Bolting to get it over.
Sad booser's eyes.
Bitten off more than he can chew.
Am I like that?
See ourselves as others see us.
Hungry man is an angry man.
Working tooth and jaw.
Don't!
O! A bone!
That last pagan king of Ireland Cormac |
in the schoolpoem |
choked himself at Sletty southward of the Boyne.
Wonder what he was eating.
Something galoptious.
Saint Patrick converted him to Christianity.
Couldn't swallow it all however.

\Man1:
Roast beef and cabbage.

\Man2:
One stew.

\BloomInt:
Smells of men.
\gab{670}
Spaton sawdust,
sweetish warmish cigarette smoke,
reek of plug,
spilt beer,
men's beery piss,
the stale of ferment.

\N1:
His gorge rose.

\BloomInt:
Couldn't eat a morsel here.
Fellow sharpening knife and fork |
to eat all before him,
old chap picking his tootles.
Slight spasm, full, chewing the cud.
Before and after.
Grace after meals.
Look on this picture |
then on that.
Scoffing up stewgravy with sopping sippets of bread.
Lick it off the plate, man!
Get out of this.

\N1:
He gazed round the stooled and tabled eaters,
tightening the wings of his nose.

\Man1:
Two stouts here.
\gab{680}

\Man2:
One corned and cabbage.

\BloomInt:
That fellow ramming a knifeful of cabbage down |
as if his life depended on it.
Good stroke.
Give me the fidgets to look.
Safer to eat from his three hands.
Tear it limb from limb.
Second nature to him.
Born with a silver knife in his mouth.
That's witty, I think.
Or no.
Silver means born rich.
Born with a knife.
But then the allusion is lost.

\N1:
An illgirt server gathered sticky clattering plates.
Rock, the head bailiff,
standing at the bar |
blew the foamy crown from his tankard.
Well up:
it splashed yellow near his boot.

\N2:
A diner, knife and fork upright,
elbows on table,
ready for a second helping |
stared towards the foodlift |
\gab{690}
across his stained square of newspaper.

\BloomInt:
Other chap telling him something |
with his mouth full.
Sympathetic listener.
Table talk.
I munched hum un thu Unchster Bunk un Munchday.
Ha?
Did you, faith?

\N1:
Mr Bloom raised two fingers doubtfully to his lips.
His eyes said:

\Bloom:
Not here.
Don't see him.

\BloomInt:
Out.
I hate dirty eaters.

\N1:
He backed towards the door.

\BloomInt:
Get a light snack in Davy Byrne's.
Stopgap.
Keep me going.
Had a good breakfast.

\Man1:
Roast and mashed here.

\Man2:
Pint of stout.
\gab{700}

\BloomInt:
Every fellow for his own, tooth and nail.
Gulp.
Grub.
Gulp.
Gobstuff.

\N1:
He came out into clearer air and turned back towards Grafton street.

\BloomInt:
Eat or be eaten.
Kill!
Kill!

\BloomInt:
Suppose that communal kitchen |
years to come perhaps.
All trotting down with porringers and tommycans to be filled.
Devour contents in the street.
John Howard Parnell |
example the provost of Trinity |
every mother's son |
don't talk of your provosts and provost of Trinity |
women and children |
cabmen |
priests parsons |
fieldmarshals archbishops.
From Ailesbury road,
Clyde road,
artisans' dwellings,
north Dublin union,
lord mayor in his gingerbread coach,
old queen in a bathchair.
\gab{710}
My plate's empty.
After you |
with our incorporated drinkingcup.
Like sir Philip Crampton's fountain.
Rub off the microbes with your handkerchief.
Next chap rubs on a new batch with his.
Father O'Flynn would make hares of them all.
Have rows |
all the same.
All for number one.
Children fighting for the scrapings of the pot.
Want a souppot as big as the Phoenix park.
Harpooning flitches and hindquarters out of it.
Hate people all round you.
City Arms hotel |
\french{table d'hôte} she called it.
Soup, joint and sweet.
Never know whose thoughts you're chewing.
Then who'd wash up all the plates and forks?
Might be all feeding on tabloids that time.
Teeth getting worse and worse.

\BloomInt:
After all |
there's a lot in that vegetarian fine flavour of things |
\gab{720}
from the earth |
garlic of course |
it stinks |
after Italian organgrinders |
crisp of onions |
mushrooms truffles.
Pain to the animal too.
Pluck and draw fowl.
Wretched brutes there at the cattlemarket |
waiting for the poleaxe to split their skulls open.
Moo.
Poor trembling calves.
Meh.
Staggering bob.
Bubble and squeak.
Butchers' buckets wobbly lights.
Give us that brisket off the hook.
Plup.
Rawhead and bloody bones.
Flayed glasseyed sheep hung from their haunches,
sheepsnouts bloodypapered snivelling nosejam on sawdust.
Top and lashers going out.
Don't maul them pieces, young one.

\BloomInt:
Hot fresh blood they prescribe for decline.
Blood always needed.
Insidious.
\gab{730}
Lick it up smokinghot, thick sugary.
Famished ghosts.

\BloomInt:
Ah, I'm hungry.

\N1:
He entered Davy Byrne's.
