\BloomInt:
Ah, I'm hungry.

\N1:
He entered Davy Byrne's.

\BloomInt:
Moral pub.
He doesn't chat.
Stands a drink now and then.
But in leapyear once in four.
Cashed a cheque for me once.

\BloomInt:
What will I take now?

\N1:
He drew his watch.

\BloomInt:
Let me see now.
Shandygaff?

\nosey:
Hello, Bloom!

\N2:
Nosey Flynn said from his nook.

\Bloom:
Hello,
Flynn.

\nosey:
How's things?

\Bloom:
Tiptop ...
\gab{740}
Let me see.
I'll take a glass of burgundy and ...
let me see.

\BloomInt:
Sardines on the shelves.
Almost taste them by looking.
Sandwich?
Ham and his descendants |
musterred and bred there.
Potted meats.
What is home |
without Plumtree's potted meat?
Incomplete.
What a stupid ad!
Under the obituary notices they stuck it.
All up a plumtree.
Dignam's potted meat.
Cannibals would with lemon and rice.
White missionary |
too salty.
Like pickled pork.
Expect the chief consumes the parts of honour.
Ought to be tough from exercise.
His wives in a row to watch the effect.
\emph{There was a right royal old nigger.
Who ate or something |
the somethings of the reverend Mr MacTrigger.}
With it an abode of bliss.
Lord knows what concoction.
\gab{750}
Cauls |
mouldy tripes |
windpipes |
faked and minced up.
Puzzle |
find the meat.
Kosher.
No meat and milk together.
Hygiene |
that was what they call now.
Yom Kippur fast |
spring cleaning of inside.
Peace and war depend on some fellow's digestion.
Religions.
Christmas turkeys and geese.
Slaughter of innocents.
Eat drink and be merry.
Then casual wards full after.
Heads bandaged.
Cheese digests all but itself.
Mighty cheese.

\Bloom:
Have you a cheese sandwich?

\curate:
Yes, sir.

\BloomInt:
Like a few olives too if they had them.
Italian I prefer.
Good glass of burgundy;
take away that.
Lubricate.
A nice salad,
cool as a cucumber,
Tom Kernan can dress.
Puts gusto into it.
Pure olive oil.
\gab{760}
Milly served me that cutlet with a sprig of parsley.
Take one Spanish onion.
God made food, |
the devil the cooks.
Devilled crab.

\nosey:
Wife well?

\Bloom:
Quite well,
thanks ...
A cheese sandwich, then.
Gorgonzola,
have you?

\curate:
Yes, sir.

\N2:
Nosey Flynn sipped his grog.

\nosey:
Doing any singing those times?

\BloomInt:
Look at his mouth.
Could whistle in his own ear.
Flap ears to match.
Music.
Knows as much about it as my coachman.
Still |
better tell him.
Does no harm.
\gab{770}
Free ad.

\Bloom:
She's engaged for a big tour |
end of this month.
You may have heard perhaps.

\nosey:
No.
O, that's the style.
Who's getting it up?

\N2:
The curate served.

\Bloom:
How much is that?

\curate:
Seven d., sir ...
Thank you, sir.

\N1:
Mr Bloom cut his sandwich into slender strips.

\BloomInt:
\emph{Mr MacTrigger.}
Easier than the dreamy creamy stuff.
\emph{His five hundred wives.
Had the time of their lives.}

\curate:
Mustard, sir?
\gab{780}

\Bloom:
Thank you.

\N1:
He studded under each lifted strip |
yellow blobs.

\BloomInt:
\emph{Their lives.}
I have it.
\emph{It grew bigger and bigger and bigger.}

\Bloom:
Getting it up?
\sout{he said.}
Well, it's like a company idea, you see.
Part shares and part profits.

\nosey:
Ay, now I remember,

\N2:
Nosey Flynn said,
putting his hand in his pocket to scratch his groin.

\nosey:
Who is this was telling me?
Isn't Blazes Boylan mixed up in it?

\N1:
A warm shock of air heat of mustard |
haunched on Mr Bloom's heart.
He raised his eyes and met the stare of a bilious clock.

\BloomInt:
\gab{790}
Two.
Pub clock |
five minutes fast.
Time going on.
Hands moving.
Two.
Not yet.

\N1:
His midriff yearned then upward,
sank within him,
yearned more longly,
longingly.

\BloomInt:
Wine.

\N1:
He smellsipped the cordial juice |
and, bidding his throat strongly to speed it, |
set his wineglass delicately down.

\Bloom:
Yes,
\sout{he said.}
He's the organiser |
in point of fact.

\BloomInt:
No fear:
no brains.

\N2:
Nosey Flynn snuffled and scratched.

\BloomInt:
Flea having a good square meal.

\nosey:
He had a good slice of luck,
\gab{800}
Jack Mooney was telling me,
over that boxingmatch |
Myler Keogh won again that soldier in the Portobello barracks.
By God,
he had the little kipper down in the county Carlow |
he was telling me ...

\BloomInt:
Hope that dewdrop doesn't come down into his glass.
No,
snuffled it up.

\nosey:
For near a month, man,
before it came off.
Sucking duck eggs |
by God |
till further orders.
Keep him off the boose, see?
O, by God,
Blazes is a hairy chap.

\N2:
Davy Byrne came forward from the hindbar |
in tuckstitched shirtsleeves,
cleaning his lips with two wipes of his napkin.
\gab{810}

\BloomInt:
Herring's blush.
Whose smile upon each feature |
plays with such and such replete.
Too much fat on the parsnips.

\nosey:
And here's himself and pepper on him,
\sout{Nosey Flynn said.}
Can you give us a good one for the Gold cup?

\davybyrne:
I'm off that, Mr Flynn,

\N2:
Davy Byrne answered.

\davybyrne:
I never put anything on a horse.

\nosey:
You're right there,
\sout{Nosey Flynn said.}

\N1:
Mr Bloom ate his strips of sandwich,
fresh clean bread,
with relish of disgust |
pungent mustard,
the feety savour of green cheese.
Sips of his wine soothed his palate.
\gab{820}

\BloomInt:
Not logwood that.
Tastes fuller this weather |
with the chill off.

\BloomInt:
Nice quiet bar.
Nice piece of wood in that counter.
Nicely planed.
Like the way it curves there.

\davybyrne:
I wouldn't do anything at all in that line,
\sout{Davy Byrne said.}
It ruined many a man,
the same horses.

\BloomInt:
Vintners' sweepstake.
Licensed for the sale of beer, wine and spirits |
for consumption on the premises.
Heads I win |
tails you lose.

\nosey:
True for you,
\sout{Nosey Flynn said.}
Unless you're in the know.
There's no straight sport going now.
Lenehan gets some good ones.
He's giving Sceptre today.
\gab{830}
Zinfandel's the favourite,
lord Howard de Walden's,
won at Epsom.
Morny Cannon is riding him.
I could have got seven to one against Saint Amant |
a fortnight before.

\davybyrne:
That so?
\sout{Davy Byrne said ...}

\N2:
He went towards the window |
and, taking up the pettycash book,
scanned its pages.

\nosey:
I could, faith,
\sout{Nosey Flynn said, snuffling.}
That was a rare bit of horseflesh.
Saint Frusquin was her sire.
She won in a thunderstorm,
Rothschild's filly,
with wadding in her ears.
Blue jacket and yellow cap.
Bad luck to big Ben Dollard and his John O'Gaunt.
He put me off it.
Ay.

\N2:
He drank resignedly from his tumbler,
\gab{840}
running his fingers down the flutes.

\nosey:
Ay,
\sout{he said, sighing.}

\N1:
Mr Bloom,
champing, standing,
looked upon his sigh.

\BloomInt:
Nosey numbskull.
Will I tell him that horse |
Lenehan?
He knows already.
Better let him forget.
Go and lose more.
Fool and his money.
Dewdrop coming down again.
Cold nose he'd have |
kissing a woman.
Still |
they might like.
Prickly beards they like.
Dogs' cold noses.
Old Mrs Riordan with the rumbling stomach's |
Skye terrier in the City Arms hotel.
Molly fondling him in her lap.
O, the big doggybowwowsywowsy!

\BloomInt:
Wine soaked and softened |
rolled pith of bread
\gab{850}
mustard a moment |
mawkish cheese.
Nice wine it is.
Taste it better because I'm not thirsty.
Bath of course does that.
Just a bite or two.
Then about six o'clock |
I can.
Six.
Six.
Time will be gone then.
She ...

\N1:
Mild fire of wine |
kindled his veins.

\BloomInt:
I wanted that badly.
Felt so off colour.

\N1:
His eyes unhungrily saw shelves of tins:
sardines, gaudy lobsters' claws.

\BloomInt:
All the odd things people pick up for food.
Out of shells,
periwinkles with a pin,
off trees,
snails out of the ground |
the French eat,
out of the sea with bait on a hook.
Silly fish learn nothing in a thousand years.
If you didn't know |
risky putting anything into your mouth.
Poisonous berries.
Johnny Magories.
Roundness you think good.
\gab{860}
Gaudy colour warns you off.
One fellow told another and so on.
Try it on the dog first.
Led on by the smell or the look.
Tempting fruit.
Ice cones.
Cream.
Instinct.
Orangegroves for instance.
Need artificial irrigation.
\german{Bleibtreustrasse.}
Yes but what about oysters.
Unsightly |
like a clot of phlegm.
Filthy shells.
Devil to open them too.
Who found them out?
Garbage, sewage they feed on.
Fizz and Red bank oysters.
Effect on the sexual.
Aphrodis.
He was in the Red Bank this morning.
Was he |
oysters |
old fish at table |
perhaps he |
young flesh in bed |
no |
June has no ar |
no oysters.
But there are people like things high.
Tainted game.
Jugged hare.
First catch your hare.
Chinese eating eggs fifty years old,
blue and green again.
Dinner of thirty courses.
\gab{870}
Each dish harmless |
might mix inside.
Idea for a poison mystery.
That archduke Leopold was it |
no yes |
or was it Otto |
one of those Habsburgs?
Or who was it |
used to eat the scruff off his own head?
Cheapest lunch in town.
Of course |
aristocrats,
then the others copy to be in the fashion.
Milly too |
rock oil and flour.
Raw pastry |
I like myself.
Half the catch of oysters |
they throw back in the sea to keep up the price.
Cheap |
no-one would buy.
Caviare.
Do the grand.
Hock in green glasses.
Swell blowout.
Lady this.
Powdered bosom pearls.
The \french{élite.
Crème de la crème.}
They want special dishes to pretend they're.
Hermit with a platter of pulse |
keep down the stings of the flesh.
Know me |
come eat with me.
\gab{880}
Royal sturgeon |
high sheriff,
Coffey, the butcher,
right to venisons of the forest from his ex.
Send him back the half of a cow.
Spread I saw down in the Master of the Rolls' kitchen area.
Whitehatted \french{chef} like a rabbi.
Combustible duck.
Curly cabbage \french{à la duchesse de Parme}.
Just as well |
to write it on the bill of fare |
so you can know what you've eaten.
Too many drugs spoil the broth.
I know it myself.
Dosing it with Edwards' desiccated soup.
Geese stuffed silly for them.
Lobsters boiled alive.
Do ptake some ptarmigan.
Wouldn't mind being a waiter in a swell hotel.
Tips,
evening dress,
halfnaked ladies.
May I tempt you to a little more filleted lemon sole,
miss Dubedat?
Yes, do bedad.
And she did bedad.
Huguenot name I expect |
that.
\gab{890}
A miss Dubedat lived in Killiney,
I remember.
\french{Du, de la}, French.
Still |
it's the same fish,
perhaps old Micky Hanlon of Moore street ripped the guts out of |
making money hand over fist |
finger in fishes' gills |
can't write his name on a cheque |
think he was painting the landscape |
with his mouth twisted.
Moooikill |
A Aitcha Ha |
ignorant as a kish of brogues,
worth fifty thousand pounds.

\N2:
Stuck on the pane |
two flies buzzed,
stuck.

\N1:
Glowing wine on his palate |
lingered |
swallowed.

\BloomInt:
Crushing in the winepress grapes of Burgundy.
Sun's heat it is.
Seems to a secret touch telling me memory.

\N1:
Touched his sense |
moistened |
remembered.

\BloomInt:
Hidden under wild ferns on Howth |
below us bay sleeping:
sky.
\gab{900}
No sound.
The sky.
The bay purple by the Lion's head.
Green by Drumleck.
Yellowgreen towards Sutton.
Fields of undersea,
the lines faint brown in grass,
buried cities.
Pillowed on my coat |
she had her hair,
earwigs in the heather |
scrub my hand under her nape,
you'll toss me all.
O wonder!
Coolsoft with ointments |
her hand touched me,
caressed:
her eyes upon me did not turn away.
Ravished over her I lay,
full lips full open,
kissed her mouth.
Yum.
Softly she gave me in my mouth |
the seedcake warm and chewed.
Mawkish pulp her mouth had mumbled |
sweetsour of her spittle.
Joy:
I ate it:
joy.
Young life,
her lips that gave me pouting.
Soft warm sticky gumjelly lips.
Flowers her eyes were,
take me,
willing eyes.
\gab{910}
Pebbles fell.
She lay still.
A goat.
No-one.
High on Ben Howth rhododendrons |
a nannygoat walking surefooted, dropping currants.
Screened under ferns she laughed |
warmfolded.
Wildly I lay on her,
kissed her:
eyes, her lips,
her stretched neck beating,
woman's breasts full in her blouse of nun's veiling,
fat nipples upright.
Hot I tongued her.
She kissed me.
I was kissed.
All yielding |
she tossed
my hair.
Kissed, she kissed me.

\BloomInt:
Me.
And me now.

\N2:
Stuck,
the flies buzzed.

\N1:
His downcast eyes |
followed the silent veining of the oaken slab.

\BloomInt:
Beauty:
\gab{920}
it curves:
curves are beauty.
Shapely goddesses,
Venus, Juno:
curves the world admires.
Can see them library museum |
standing in the round hall,
naked goddesses.
Aids to digestion.
They don't care what man looks.
All to see.
Never speaking.
I mean to say to fellows like Flynn.
Suppose she did Pygmalion and Galatea |
what would she say first?
Mortal!
Put you in your proper place.
Quaffing nectar at mess with gods,
golden dishes,
all ambrosial.
Not like a tanner lunch we have,
boiled mutton,
carrots and turnips,
bottle of Allsop.
Nectar, imagine it |
drinking electricity:
gods' food.
Lovely forms of women |
sculped Junonian.
Immortal lovely.
And we stuffing food in one hole and out behind:
food, chyle, blood, dung, earth, food:
have to feed it |
like stoking an engine.
\gab{930}
They have no.
Never looked.
I'll look today.
Keeper won't see.
Bend down |
let something drop |
see if she.

\N1:
Dribbling |
a quiet message from his bladder came |
to go |
to do |
not to do |
there to do.
A man and ready |
he drained his glass to the lees |
and walked,

\N1:
to men too they gave themselves,
manly conscious,
lay with men lovers,
a youth enjoyed her,
to the yard.

\N2:
When the sound of his boots had ceased |
Davy Byrne said from his book:

\davybyrne:
What is this he is?
Isn't he in the insurance line?

\nosey:
He's out of that long ago,
\sout{Nosey Flynn said.}
\gab{940}
He does canvassing for the \Title{Freeman}.

\davybyrne:
I know him well to see,
\sout{Davy Byrne said.}
Is he in trouble?

\nosey:
Trouble?
\sout{Nosey Flynn said.}
Not that I heard of.
Why?

\davybyrne:
I noticed he was in mourning.

\nosey:
Was he?
\sout{Nosey Flynn said.}
So he was, faith.
I asked him how was all at home.
You're right, by God.
So he was.

\davybyrne:
I never broach the subject,
\sout{Davy Byrne said humanely,}
if I see a gentleman is in trouble that way.
It only brings it up fresh in their minds.

\nosey:
It's not the wife anyhow,
\sout{Nosey Flynn said.}
I met him the day before yesterday
and he coming out of that Irish farm dairy |
\gab{950}
John Wyse Nolan's wife has in Henry street |
with a jar of cream in his hand |
taking it home to his better half.
She's well nourished, I tell you.
Plovers on toast.

\davybyrne:
And is he doing for the \Title{Freeman}?
\sout{Davy Byrne said.}

\N2:
Nosey Flynn pursed his lips.

\nosey:
He doesn't buy cream on the ads |
he picks up.
You can make bacon of that.

\davybyrne:
How so?

\N2:
Davy Byrne asked,
coming from his book.

\N2:
Nosey Flynn made swift passes in the air with juggling fingers.
He winked.

\nosey:
He's in the craft,
\gab{960}
\sout{he said.}

\davybyrne:
-Do you tell me so?
\sout{Davy Byrne said.}

\nosey:
Very much so,
\sout{Nosey Flynn said.}
Ancient free and accepted order.
He's an excellent brother.
Light, life and love, by God.
They give him a leg up.
I was told that by a,
well,
I won't say who.

\davybyrne:
Is that a fact?

\nosey:
O, it's a fine order,
\sout{Nosey Flynn said.}
They stick to you when you're down.
I know a fellow |
was trying to get into it.
But they're as close as damn it.
By God |
they did right to keep the women out of it.

\N2:
Davy Byrne smiled\-yawned\-nodded all in one:

\davybyrne:
Iiiiiichaaaaaaach!
\gab{970}

\nosey:
There was one woman,
\sout{Nosey Flynn said,}
hid herself in a clock |
to find out what they do be doing.
But be damned |
but they smelt her out |
and swore her in on the spot |
a master mason.
That was one of the saint Legers of Doneraile.

\N2:
Davy Byrne,
sated after his yawn,
said with tearwashed eyes:

\davybyrne:
And is that a fact?
Decent quiet man he is.
I often saw him in here |
and I never once saw him,
you know,
over the line.

\nosey:
God Almighty couldn't make him drunk,
\sout{Nosey Flynn said firmly.}
Slips off when the fun gets too hot.
Didn't you see him look at his watch?
Ah, you weren't there.
\gab{980}
If you ask him to have a drink |
first thing he does |
he outs with the watch |
to see what he ought to imbibe.
Declare to God |
he does.

\davybyrne:
There are some like that,
\sout{Davy Byrne said.}
He's a safe man,
I'd say.

\nosey:
He's not too bad,
\sout{Nosey Flynn said, snuffling it up.}
He's been known to put his hand down too |
to help a fellow.
Give the devil his due.
O, Bloom has his good points.
But there's one thing he'll never do.

\N2:
His hand scrawled a dry pen signature beside his grog.

\davybyrne:
I know,
\sout{Davy Byrne said.}

\nosey:
Nothing in black and white,
\sout{Nosey Flynn said.}

\N1:
Paddy Leonard and Bantam Lyons came in.
Tom Rochford followed frowning,
\gab{990}
a plaining hand on his claret waistcoat.

\bantam:
Day, Mr Byrne.

\davybyrne:
Day, gentlemen.

\N2:
They paused at the counter.

\leonard:
Who's standing?

\N1:
Paddy Leonard asked.

\nosey:
I'm sitting anyhow,
\sout{Nosey Flynn answered.}

\leonard:
Well, what'll it be?
\sout{Paddy Leonard asked.}

\bantam:
I'll take a stone ginger,

\N1:
Bantam Lyons said.

\leonard:
How much?
\sout{Paddy Leonard cried.}
Since when, for God' sake?
What's yours, Tom?

\nosey:
How is the main drainage?
\gab{1000}
\sout{Nosey Flynn asked, sipping.}

\N1:
For answer |
Tom Rochford pressed his hand to his breastbone and hiccupped.

\rochford:
Would I trouble you for a glass of fresh water, Mr Byrne?
\sout{he said.}

\davybyrne:
Certainly, sir.

\N1:
Paddy Leonard eyed his alemates.

\leonard:
Lord love a duck,
\sout{he said.}
Look at what I'm standing drinks to!
Cold water and gingerpop!
Two fellows that would suck whisky off a sore leg.
He has some bloody horse up his sleeve for the Gold cup.
A dead snip.

\nosey:
Zinfandel is it?
\sout{Nosey Flynn asked.}

\N1:
Tom Rochford spilt powder from a twisted paper |
\gab{1010}
into the water set before him.

\rochford:
That cursed dyspepsia,

\N1:
he said before drinking.

\davybyrne:
Breadsoda is very good,
\sout{Davy Byrne said.}

\N1:
Tom Rochford nodded and drank.

\nosey:
Is it Zinfandel?

\bantam:
Say nothing!

\N1:
Bantam Lyons winked.

\bantam:
I'm going to plunge five bob on my own.

\leonard:
Tell us |
if you're worth your salt |
and be damned to you,
\sout{Paddy Leonard said.}
Who gave it to you?

\N1:
Mr Bloom on his way out |
raised three fingers in greeting.
\gab{1020}

\nosey:
So long!
\sout{Nosey Flynn said.}

\N1:
The others turned.

\bantam:
That's the man now |
that gave it to me,
\sout{Bantam Lyons whispered.}

\leonard:
Prrwht!
\sout{Paddy Leonard said with scorn.}
Mr Byrne, sir,
we'll take two of your small Jamesons after that and a ...

\davybyrne:
Stone ginger,
\sout{Davy Byrne added civilly.}

\leonard:
Ay,
\sout{Paddy Leonard said.}
A suckingbottle for the baby.

\N1:
Mr Bloom walked towards Dawson street,
his tongue brushing his teeth smooth.

\BloomInt:
Something green it would have to be:
spinach, say.
Then with those Rontgen rays |
\gab{1030}
searchlight you could.

\N1:
At Duke lane |
a ravenous terrier choked up a sick knuckly cud on the cobblestones |
and lapped it with new zest.

\BloomInt:
Surfeit.
Returned with thanks |
having fully digested the contents.
First sweet |
then savoury.

\N1:
Mr Bloom coasted warily.

\BloomInt:
Ruminants.
His second course.
Their upper jaw they move.
Wonder if Tom Rochford will do anything with that invention of his?
Wasting time explaining it to Flynn's mouth.
Lean people long mouths.
Ought to be a hall or a place where inventors could go in and invent free.
Course then you'd have all the cranks pestering.

\N1:
He hummed,
prolonging in solemn echo |
the closes of the bars:

\Bloom:
\song{%
    Don Giovanni,
\gab{1040}
        a cenar teco\\
    M'invitasti.}


\BloomInt:
Feel better.
Burgundy.
Good pick me up.
Who distilled first?
Some chap in the blues.
Dutch courage.
That \Title{Kilkenny People} in the national library now |
I must.

\N1:
Bare clean closestools |
waiting in the window of William Miller, plumber,
turned back his thoughts.

\BloomInt:
They could:
and watch it all the way down,
swallow a pin |
sometimes come out of the ribs years after,
tour round the body |
changing biliary duct |
spleen |
squirting liver |
gastric juice |
coils of intestines like pipes.
But the poor buffer
would have to stand all the time |
with his insides entrails on show.
\gab{1050}
Science.

\Bloom:
\italian{A cenar teco}.

\BloomInt:
What does that \italian{teco} mean?
Tonight perhaps.

\BloomInt:
\song{%
    Don Giovanni,\\
    Thou hast me invited\\
    To come to supper tonight,\\
    The rum the rumdum.}

\BloomInt:
Doesn't go properly.

\BloomInt:
Keyes: two months if I get Nannetti to.
That'll be two pounds ten |
about two pounds eight.
Three |
Hynes owes me.
Two eleven.
Prescott's dyeworks van over there.
If I get Billy Prescott's ad:
two fifteen.
Five guineas about.
\gab{1060}
On the pig's back.

\BloomInt:
Could buy one of those silk petticoats for Molly,
colour of her new garters.

\BloomInt:
Today.
Today.
Not think.

\BloomInt:
Tour the south then.
What about English wateringplaces?
Brighton, Margate.
Piers by moonlight.
Her voice floating out.
Those lovely seaside girls.

\N2:
Against John Long's |
a drowsing loafer lounged in heavy thought,
gnawing a crusted knuckle.

\BloomInt:
Handy man wants job.
Small wages.
Will eat anything.

\N1:
Mr Bloom turned at Gray's confectioner's window of unbought tarts |
and passed the reverend Thomas Connellan's bookstore.
\gab{1070}

\BloomInt:
\Title{Why I left the Church of Rome?}
\Title{Birds' Nest.}
Women run him.
They say they used to give pauper children soup |
to change to protestants in the time of the potato blight.
Society over the way |
papa went to |
for the conversion of poor jews.
Same bait.
Why we left the church of Rome?

\N2:
A blind stripling stood |
tapping the curbstone with his slender cane.

\BloomInt:
No tram in sight.
Wants to cross.

\Bloom:
Do you want to cross?
\sout{Mr Bloom asked.}

\N2:
The blind stripling did not answer.
His wallface frowned weakly.
He moved his head uncertainly.

\Bloom:
You're in Dawson street,
\sout{Mr Bloom said.}
\gab{1080}
Molesworth street is opposite.
Do you want to cross?
There's nothing in the way.

\N2:
The cane moved out |
trembling |
to the left.

\N1:
Mr Bloom's eye followed its line |
and saw again |
the dyeworks' van drawn up before Drago's.

\BloomInt:
Where I saw his brillantined hair |
just when I was.
Horse drooping.
Driver in John Long's.
Slaking his drouth.

\Bloom:
There's a van there,
\sout{Mr Bloom said,}
but it's not moving.
I'll see you across.
Do you want to go to Molesworth street?

\stripling:
Yes,

\N2:
the stripling answered.

\stripling:
South Frederick street.

\Bloom:
Come,
\sout{Mr Bloom said.}

\N1:
He touched the thin elbow gently:
\gab{1090}
then took the limp seeing hand |
to guide it forward.

\BloomInt:
Say something to him.
Better not do the condescending.
They mistrust what you tell them.
Pass a common remark.

\Bloom:
The rain kept off.

\BloomInt:
No answer.

\BloomInt:
Stains on his coat.
Slobbers his food, I suppose.
Tastes all different for him.
Have to be spoonfed first.
Like a child's hand, his hand.
Like Milly's was.
Sensitive.
Sizing me up |
I daresay |
from my hand.
Wonder if he has a name.
Van.
Keep his cane clear of the horse's legs:
tired drudge get his doze.
\gab{1100}
That's right.
Clear.
Behind a bull:
in front of a horse.

\stripling:
Thanks, sir.

\BloomInt:
Knows I'm a man.
Voice.

\Bloom:
Right now?
First turn to the left.

\N2:
The blind stripling tapped the curbstone and went on his way,
drawing his cane back,
feeling again.

\N1:
Mr Bloom walked behind the eyeless feet,
a flatcut suit of herringbone tweed.

\BloomInt:
Poor young fellow!
How on earth did he know that van was there?
Must have felt it.
See things in their forehead perhaps:
kind of sense of volume.
Weight or size of it,
something blacker than the dark.
Wonder would he feel it
if something was removed.
Feel a gap.
\gab{1110}
Queer idea of Dublin he must have,
tapping his way round by the stones.
Could he walk in a beeline |
if he hadn't that cane?
Bloodless pious face |
like a fellow going in to be a priest.

\BloomInt:
Penrose!
That was that chap's name.

\BloomInt:
Look at all the things they can learn to do.
Read with their fingers.
Tune pianos.
Or we are surprised they have any brains.
Why we think a deformed person or a hunchback clever |
if he says something we might say.
Of course the other senses are more.
Embroider.
Plait baskets.
People ought to help.
Workbasket I could buy for Molly's birthday.
Hates sewing.
Might take an objection.
\gab{1120}
Dark men they call them.

\BloomInt:
Sense of smell must be stronger too.
Smells on all sides, bunched together.
Each street different smell.
Each person too.
Then the spring, the summer:
smells.
Tastes?
They say you can't taste wines |
with your eyes shut or a cold in the head.
Also smoke in the dark |
they say get no pleasure.

\BloomInt:
And with a woman, for instance.
More shameless not seeing.
That girl passing the Stewart institution,
head in the air.
Look at me.
I have them all on.
Must be strange not to see her.
Kind of a form in his mind's eye.
The voice, temperatures:
when he touches her with his fingers |
must almost see the lines, the curves.
His hands on her hair, for instance.
Say it was black, for instance.
Good.
We call it black.
\gab{1130}
Then passing over her white skin.
Different feel perhaps.
Feeling of white.

\BloomInt:
Postoffice.
Must answer.
Fag today.
Send her a postal order |
two shillings, half a crown.
Accept my little present.
Stationer's just here too.
Wait.
Think over it.

\N1:
With a gentle finger |
he felt ever so slowly |
the hair combed back above his ears.

\BloomInt:
Again.
Fibres of fine fine straw.

\N1:
Then gently |
his finger felt the skin of his right cheek.

\BloomInt:
Downy hair there too.
Not smooth enough.
The belly is the smoothest.
No-one about.
There he goes into Frederick street.
Perhaps to Levenston's dancing academy piano.
Might be settling my braces.

\N1:
Walking by Doran's publichouse |
\gab{1140}
he slid his hand between his waistcoat and trousers |
and, pulling aside his shirt gently,
felt a slack fold of his belly.

\BloomInt:
But I know it's whitey yellow.
Want to try in the dark to see.

\N1:
He withdrew his hand and pulled his dress to.

\BloomInt:
Poor fellow!
Quite a boy.
Terrible.
Really terrible.
What dreams would he have,
not seeing?
Life a dream for him.
Where is the justice being born that way?
All those women and children |
excursion beanfeast |
burned and drowned in New York.
Holocaust.
Karma they call that |
transmigration for sins you did in a past life |
the reincarnation |
met him pike hoses.
Dear, dear, dear.
Pity, of course:
but somehow you can't cotton on to them someway.%
\gab{1150}

\BloomInt:
Sir Frederick Falkiner going into the freemasons' hall.
Solemn as Troy.
After his good lunch in Earlsfort terrace.
Old legal cronies cracking a magnum.
Tales of the bench and assizes and annals of the bluecoat school.
I sentenced him to ten years.
I suppose he'd turn up his nose at that stuff I drank.
Vintage wine for them,
the year marked on a dusty bottle.
Has his own ideas of justice in the recorder's court.
Wellmeaning old man.
Police chargesheets crammed with cases |
get their percentage manufacturing crime.
Sends them to the rightabout.
The devil on moneylenders.
Gave Reuben~J. a great strawcalling.
Now he's really what they call a dirty jew.
Power those judges have.
\gab{1160}
Crusty old topers in wigs.
Bear with a sore paw.
And may the Lord have mercy on your soul.

\BloomInt:
Hello, placard.
Mirus bazaar.
His Excellency the lord lieutenant.
Sixteenth.
Today it is.
In aid of funds for Mercer's hospital.
\Title{The Messiah} was first given for that.
Yes.
Handel.
What about going out there:
Ballsbridge.
Drop in on Keyes.
No use sticking to him like a leech.
Wear out my welcome.
Sure to know someone on the gate.

\N1:
Mr Bloom came to Kildare street.

\BloomInt:
First I must.
Library.

\BloomInt:
Straw hat in sunlight.
Tan shoes.
Turnedup trousers.
It is.
It is.

\N1:
His heart quopped softly.

\BloomInt:
To the right.
Museum.
Goddesses.

\N1:
He swerved to the right.%
\gab{1170}

\BloomInt:
Is it?
Almost certain.
Won't look.
Wine in my face.
Why did I?
Too heady.
Yes,
it is.
The walk.
Not see.
Get on.

\N1:
Making for the museum gate |
with long windy steps |
he lifted his eyes.

\BloomInt:
Handsome building.
Sir Thomas Deane designed.
Not following me?

\BloomInt:
Didn't see me perhaps.
Light in his eyes.

\N1:
The flutter of his breath |
came forth in short sighs.

\BloomInt:
Quick.
Cold statues:
quiet there.
Safe in a minute.

\BloomInt:
No.
Didn't see me.
After two.
Just at the gate.

\BloomInt:
My heart!

\N1:
His eyes beating |
looked steadfastly at cream curves of stone.
\gab{1180}

\BloomInt:
Sir Thomas Deane was the Greek architecture.


\BloomInt:
Look for something |
I.

\N1:
His hasty hand went quick into a pocket,
took out,
read |
unfolded |
Agendath Netaim.

\BloomInt:
Where did I?

\BloomInt:
Busy looking.

\N1:
He thrust back quick |
Agendath.

\BloomInt:
Afternoon |
she said.

\BloomInt:
I am looking for that.
Yes, that.
Try all pockets.
Handker.
\Title{Freeman.}
Where did I?
Ah, yes.
Trousers.
Potato.
Purse.
Where did I?

\BloomInt:
Hurry.
\gab{1190}
Walk quietly.
Moment more.
My heart.

\N1:
His hand looking for the

\BloomInt:
where did I put

\N1:
found in his hip pocket

\BloomInt:
soap |
lotion |
have to call |
tepid |
paper stuck.
Ah |
soap there |
I |
yes.
Gate.

\BloomInt:
Safe!
