\histstyle{Arval}

\N1:
Deshil Holles Eamus.

\N{1,2}:
Deshil Holles Eamus.

\N{1,2,3}:
Deshil Holles Eamus.

\N1:
Send us bright one,
light one,
Horhorn,
quickening and wombfruit.

\N{1,2}:
Send us bright one,
light one,
Horhorn,
quickening and wombfruit.

\N{1,2,3}:
Send us bright one,
light one,
Horhorn,
quickening and wombfruit.

\Nurses:
Hoopsa boyaboy hoopsa!

\Nurses:
Hoopsa boyaboy hoopsa!

\Nurses:
Hoopsa boyaboy hoopsa!

\histstyle{Sallust (omitted in 2008)}

Universally |
that person's acumen is esteemed very little perceptive |
concerning whatsoever matters are being held |
as most profitably |
by mortals with sapience endowed to be studied |
who is ignorant of |
that which the most in doctrine erudite |
\gab{10}
and certainly by reason of that |
in them high mind's ornament |
deserving of veneration |
constantly maintain when |
by general consent |
they affirm |
that other circumstances being equal |
by no exterior splendour |
is the prosperity of a nation |
more efficaciously asserted |
than by the measure |
of how far forward |
may have progressed the tribute of its solicitude |
for that proliferent continuance |
which of evils |
the original |
if it be absent when fortunately present |
constitutes the certain sign |
of omnipotent nature's incorrupted benefaction.

For who is there |
who anything of some significance |
has apprehended |
but is conscious |
that that exterior splendour |
may be the surface of a downwardtending lutulent reality |
or on the contrary |
\gab{20}
anyone so is there unilluminated |
as not to perceive |
that as no nature's boon |
can contend against the bounty of increase |
so it behoves every most just citizen |
to become the exhortator and admonisher |
of his semblables |
and to tremble |
lest what had in the past |
been by the nation excellently commenced |
might be in the future |
not with similar excellence accomplished |
if an inverecund habit |
shall have gradually traduced the honourable |
by ancestors transmitted customs |
to that thither of profundity |
that that one was audacious |
excessively who would have the hardihood to rise |
affirming that no more odious offence |
can for anyone be |
than to oblivious neglect |
to consign that evangel simultaneously command and promise |
which on all mortals |
\gab{30}
with prophecy of abundance or with diminution's menace |
that exalted of reiteratedly procreating function |
ever irrevocably enjoined?

\histstyle{Malmesbury}

\N1:
It is not why |
therefore |
we shall wonder |
if,
as the best historians relate,
among the Celts,
who nothing |
that was not in its nature admirable |
admired,
the art of medicine |
shall have been highly honoured.
Not to speak |
of hostels,
leperyards,
sweating chambers,
plaguegraves,
their greatest doctors,
the O'Shiels,
the O'Hickeys,
the O'Lees,
have sedulously set down |
the divers methods |
by which the sick and the relapsed |
found again health |
whether the malady had been |
the trembling withering or loose boyconnell flux.
\gab{40}

[omitted in 2008]

Certainly in every public work |
which in it |
anything of gravity contains preparation |
should be with importance commensurate |
and therefore a plan was by them adopted
(whether by having preconsidered |
or as the maturation of experience |
it is difficult in being said |
which the discrepant opinions of subsequent inquirers |
are not up to the present |
congrued to render manifest)
whereby maternity was |
so far from all accident possibility removed |
that whatever care the patient |
in that all hardest of woman hour |
chiefly required |
and not solely for the copiously opulent |
but also for her |
who not being sufficiently moneyed |
scarcely and often not even scarcely |
could subsist valiantly |
and for an inconsiderable emolument |
was provided.
To her |
\gab{50}
nothing already then and thence·forward |
was anyway able to be molestful |
for this chiefly felt |
all citizens except with proliferent mothers |
prosperity at all not to can be |
and as they had received eternity |
gods mortals generation to befit them |
her beholding,
when the case was so hoving itself,
parturient |
in vehicle thereward carrying desire immense |
among all one another |
was impelling on of her |
to be received into that domicile.
O thing of prudent nation |
not merely in being seen |
but also even in being related worthy of being praised |
that they her |
by anticipation |
went seeing mother,
that she by them |
suddenly to be about to be cherished |
had been begun she felt!

\week{1}

\histstyle{Anglo-Saxon}

[omitted in 2008]

Before born |
bliss babe had.
\gab{60}
Within womb |
won he worship.
Whatever in that one case done |
commodiously done was.
A couch by midwives |
attended with wholesome food reposeful,
cleanest swaddles as though forthbringing |
were now done and by wise foresight set:
but to this |
no less of what drugs there is need |
and surgical implements which are pertaining to her case |
not omitting aspect |
of all very distracting spectacles in various latitudes |
by our terrestrial orb |
offered together with images,
divine and human,
the cogitation of which by sejunct females |
is to tumescence conducive |
or eases issue in the high sun·bright well·built fair home of mothers |
when,
ostensibly far gone and reproductitive,
it is come |
by her thereto to lie in,
\gab{70}
her term up.

\week{2}

\stage{Enter Bloom}

\N2:
Some man that wayfaring was |
stood by housedoor at night's oncoming.
Of Israel's folk was that man |
that on earth wandering far |
had fared.
Stark ruth of man |
his errand that him lone led |
till that house.

\week{3}

\N3:
Of that house |
A.~Horne is lord.
Seventy beds keeps he there |
teeming mothers are wont |
that they lie for to thole |
and bring forth bairns hale |
so God's angel to Mary quoth.
Watchers tway there walk,
white sisters in ward sleepless.
Smarts they still,
sickness soothing:
in twelve moons thrice |
an hundred.
Truest bedthanes they twain are,
for Horne holding wariest ward.

\week{4}

\N1:
In ward wary |
\gab{80}
the watcher hearing come |
that man mildhearted |
eft rising with swire |
ywimpled to him |
her gate wide undid.
Lo,
levin leaping |
lightens in eyeblink |
Ireland's westward welkin.
Full she drad |
that God the Wreaker |
all mankind would fordo with water |
for his evil sins.
Christ's rood made she on breastbone |
and him drew |
that he would rathe infare |
under her thatch.
That man |
her will wotting worthful |
went in Horne's house.

\week{5}

\stage{Enter Nurse Callan}

\N2:
Loth to irk |
in Horne's hall |
hat holding |
the seeker stood.
On her stow |
he ere was living |
with dear wife and lovesome daughter |
that then |
over land and seafloor |
nine years had long outwandered.

[omitted in 2008]

Once her |
in townhithe meeting |
he to her |
bow had not doffed.
Her to forgive now |
he craved with good ground |
of her allowed that |
that of him swift·seen face,
\gab{90}
hers,
so young then had looked.
Light swift |
her eyes kindled,
bloom of blushes |
his word winning.

\week{6}

As her eyes then |
ongot his weeds swart |
therefor sorrow she feared.
Glad after she was |
that ere adread was.
Her he asked |
if O'Hare Doctor tidings sent |
from far coast |
and she with grameful sigh |
him answered |
that O'Hare Doctor |
in heaven was.
Sad was the man |
that word to hear |
that him so heavied |
in bowels ruthful.
All she there told him,
ruing death |
for friend so young,
algate sore unwilling |
God's rightwiseness to withsay.
She said |
that he had a fair sweet death |
through God His goodness |
with masspriest to be shriven,
\gab{100}
holy housel |
and sick men's oil to his limbs.
The man then right earnest asked the nun |
of which death the dead man was died |
and the nun answered him and said |
that he was died in Mona Island |
through bellycrab |
three year agone come Childermas |
and she prayed to God the Allruthful |
to have his dear soul in his undeathliness.
He heard her sad words,
in held hat |
sad staring.
So stood they there both |
awhile in wanhope |
sorrowing one with other.

\week{7}

\histstyle{Everyman}

Therefore,
everyman,
look to that last end |
that is thy death |
and the dust that gripeth |
on every man that is born of woman |
for |
as he came naked forth from his mother's womb |
so naked shall he wend him |
at the last |
for to go as he came.%
\gab{110}

\week{8}

\N3:
The man that was come in to the house |
then spoke to the nursingwoman |
and he asked her |
how it fared |
with the woman that lay there in childbed.
The nursingwoman answered him |
and said that that woman was in throes |
now full three days |
and that it would be a hard birth |
unneth to bear |
but that now in a little |
it would be.
She said thereto |
that she had seen many births of women |
but never was none so hard |
as was that woman's birth.
Then she set it all forth to him |
for because she knew the man |
that time was |
had lived nigh that house.
The man hearkened to her words |
for he felt with wonder |
women's woe in the travail that they have of motherhood |
\gab{120}
and he wondered to look on her face |
that was a fair face for any man to see |
but yet was she left after long years |
a handmaid.
Nine twelve bloodflows |
chiding her childless.

\week{9}

\histstyle{Mandeville}

\N1:
And whiles they spake |
the door of the castle was opened |
and there nighed them |
a mickle noise |
as of many that sat there |
at meat.
And there came against the place |
as they stood |
a young learning·knight yclept Dixon.
And the traveller Leopold |
was couth to him |
sithen it had happed |
that they had had ado each with other |
in the house of misericord |
where this learning·knight lay |
by cause the traveller Leopold came there |
to be healed |
for he was sore wounded in his breast by a spear |
wherewith a horrible and dreadful dragon was smitten him |
\gab{130}
for which he did do |
make a salve of volatile salt and chrism |
as much as he might suffice.
And he said now |
that he should go in to that castle |
for to make merry with them |
that were there.

\N2:
And the traveller Leopold said |
that he should go other·whither |
for he was a man of cautels and a subtile.
Also the lady was of his avis |
and repreved the learning·knight |
though she trowed well |
that the traveller had said thing |
that was false for his subtility.
But the learning·knight would not hear say nay |
nor do her mandement |
ne have him in aught contrarious to his list |
and he said how |
it was a marvellous castle.
And the traveller Leopold went into the castle |
for to rest him for a space |
being sore of limb |
after many marches environing |
\gab{140}
in divers lands and sometime venery.

\week{10}

[omitted in 2008]

And in the castle |
was set a board |
that was of the birchwood of Finlandy |
and it was upheld by four dwarfmen of that country |
but they durst not move more |
for enchantment.
And on this board |
were frightful swords and knives |
that are made in a great cavern by swinking demons |
out of white flames |
that they fix then in the horns of buffalos and stags |
that there abound marvellously.
And there were vessels |
that are wrought by magic of Mahound |
out of sea·sand and the air |
by a warlock with his breath |
that he blases in to them |
like to bubbles.
And full fair cheer and rich |
was on the board |
that no wight could devise a fuller |
ne richer.
And there was a vat of silver |
\gab{150}
that was moved by craft to open |
in the which lay strange fishes withouten heads |
though misbelieving men nie that this be possible thing |
without they see it |
natheless they are so.
And these fishes lie in an oily water |
brought there from Portugal land |
because of the fatness that therein |
is like to the juices of the olive·press.
And also |
it was a marvel to see in that castle |
how by magic |
they make a compost out of fecund wheat·kidneys out of Chaldee |
that by aid of certain angry spirits |
that they do in to it |
swells up wondrously |
like to a vast mountain.
And they teach the serpents there |
to entwine themselves up on long sticks out of the ground |
and of the scales of these serpents |
they brew out a brewage |
like to mead.

\week{11}

\N3:
And the learning·knight |
\gab{160}
let pour for childe Leopold a draught |
and halp thereto the while |
all they that were there |
drank every each.
And childe Leopold did up his beaver |
for to pleasure him |
and took apertly somewhat in amity |
for he never drank no manner of mead |
which he then put by |
and anon full privily |
he voided the more part in his neighbour glass |
and his neighbour nist not |
of this wile.
And he sat down |
in that castle with them |
for to rest him there awhile.
Thanked be Almighty God.

\week{12}

\histstyle{Malory}

This meanwhile |
this good sister stood by the door |
and begged them |
at the reverence of Jesu |
our alther liege Lord |
to leave their wassailing |
for there was above |
one quick with child,
a gentle dame,
whose time hied fast.

\N1:
This meanwhile this good sister stood by the door:

\nursecallan:
I begged ye |
at the reverence of Jesu |
our alther liege Lord |
to leave your wassailing |
for there is above |
one quick with child,
a gentle dame,
whose time hies fast.

\All:
\emph{shriek!}

\N2:
Sir Leopold heard on the upfloor |
\gab{170}
cry on high |
and he wondered what cry that it was |
whether of child or woman.

\Bloom:
I marvel that it be not come or now.
Meseems |
it dureth overlong.

\N3:
And he was ware |
and saw a franklin |
that hight Lenehan |
on that side the table |
that was older than any of the tother |
and for that |
they both were knights virtuous in the one emprise |
and eke by cause that he was elder |
he spoke to him full gently.

\Bloom:
But or it be long too |
she will bring forth |
by God His bounty |
and have joy of her childing |
for she hath waited |
marvellous long.

\N1:
And the franklin |
that had drunken |
said:

\lenehan:
Expecting each moment |
to be her next.

\N2:
Also he took the cup |
that stood tofore him |
for him needed never |
none asking nor desiring of him to drink |
and,
Now drink,
said he,
fully delectably,
\gab{180}
and he quaffed |
as far as he might |
to their both's health |
for he was a passing good man |
of his lustiness.
And sir Leopold |
that was the goodliest guest that ever sat in scholars' hall |
and that was the meekest man and the kindest |
that ever laid husbandly hand under hen |
and that was the very truest knight of the world |
one that ever did |
minion service to lady gentle |
pledged him courtly in the cup.
Woman's woe |
with wonder |
pondering.

\week{13}

\N3:
Now let us speak of that fellowship |
that was there |
to the intent to be drunken |
an they might.
There was a sort of scholars |
along either side the board,
that is to wit:

\N1:
Dixon |
yclept junior |
of saint Mary Merciable's.
\stage{Dixon bows}

\N2:
With other his fellows,
Lynch.
\stage{Lynch bows}

\N3:
And Madden,
\gab{190}
scholars of medicine.
\stage{Madden bows}

\N1:
And the franklin that hight Lenehan.
\stage{Lenehan bows}

\N2:
And one from Alba Longa,
one Crotthers.
\stage{Crotthers bows}

\N3:
And young Stephen
\stage{Stephen bows}
that had mien of a frere |
that was at head of the board.

\N1:
And Costello |
that men clepen Punch Costello.
\stage{Costello bows}

\N2:
All long of a mastery of him |
erewhile gested
(and |
of all them,
reserved young Stephen,
he was the most drunken |
that demanded still of more mead)
and beside the meek |
sir Leopold.
But on young Malachi they waited |
for that he promised to have come |
and such as |
intended to no goodness |
said how he had broke his avow.
And sir Leopold sat with them |
for he bore fast friendship to sir Simon |
and to this |
his son young Stephen |
and for that |
his languor becalmed him there |
after longest wanderings |
\gab{200}
in·so·much as they feasted him |
for that time in the honourablest manner.
Ruth red him,
love led on |
with will to wander,
loth to leave.

\week{14}

\N3:
For they were right witty scholars.
And he heard their aresouns |
each gen other |
as touching birth and righteousness.

\sout{young Madden maintaining that}\\
\madden:
I put such case |
it were hard the wife to die
(for so it fell out |
a matter of some year agone |
with a woman of Eblana in Horne's house that now was |
trespassed out of this world |
and the self |
night next before her death |
all leeches and pothecaries |
had taken counsel of her case).

\N1:
And they said farther |
she should live |
because in the beginning,
they said,
the woman should bring forth in pain |
and wherefore they that were of this imagination |
affirmed how young Madden had said truth |
\gab{210}
for he had conscience |
to let her die.

\sout{And not few |
and of these |
was young Lynch were}\\
\lynch:
I am in doubt |
that the world was now right evil |
governed as it was |
never other |
howbeit the mean people believed it otherwise |
but the law nor his judges |
did provide no remedy.
A redress |
God grant.

\N2:
This was scant said |
but all cried |
with one acclaim:

\All:
Nay,
by our Virgin Mother,
the wife should live |
and the babe to die.

\N3:
In colour whereof |
they waxed hot upon that head |
what with argument |
and what for their drinking |
but the franklin Lenehan |
was prompt each |
when to pour them ale |
so that at the least way |
mirth might not lack.
Then young Madden showed all |
the whole affair:

\madden:
She was dead |
and how |
for holy religion sake |
by rede of palmer and bedesman |
\gab{220}
and for a vow he had made to Saint Ultan of Arbraccan |
her goodman husband |
would not let her death.

\N1:
Whereby |
they were all wondrous grieved.
To whom |
young Stephen had these words following:

\stephen:
Murmur,
sirs,
is eke oft among lay folk.
Both babe and parent now glorify their Maker,
the one in limbo gloom,
the other in purgefire.
But,
gramercy,
what of those God·possibled souls that we nightly impossibilise,
which is the sin against the Holy Ghost,
Very God,
Lord and Giver of Life?
For,
sirs,
our lust is brief.
We are means |
to those small creatures within us |
and nature has other ends |
than we.

\N2:
Then said Dixon junior to Punch Costello |
wist he what ends.
But he had overmuch drunken |
\gab{230}
and the best word he could have of him |
was that |
he would ever dishonest a woman |
whoso she were or wife or maid or leman |
if it so fortuned him |
to be delivered of his spleen of lustihead.
Whereat Crotthers of Alba Longa |
sang young Malachi's praise of that beast the unicorn |
how once in the millennium |
he cometh by his horn,
the other all this while,
pricked forward |
with their jibes |
wherewith they did malice him,
witnessing all and several |
by saint Foutinus |
his engines |
that he was able to do |
any manner of thing |
that lay in man to do.

\N3:
Thereat |
laughed they all right jocundly |
only young Stephen |
and sir Leopold which never durst laugh too open |
by reason of a strange humour |
which he would not bewray |
and also |
for that he rued |
for her that bare |
\gab{240}
whoso she might be |
or wheresoever.

\N1:
Then spake young Stephen |
orgulous of mother Church |
that would cast him out of her bosom,
of law,
of canons,
of Lilith,
patron of abortions,
of bigness wrought by wind of seeds of brightness |
or by potency of vampires |
mouth to mouth |
or,
as Virgilius saith,
by the influence of the occident |
or by the reek of moonflower |
or an she lie with a woman |
which her man has but lain with,
\latin{effectu secuto},
or peradventure |
in her bath |
according to the opinions of Averroes and Moses Maimonides.

\stephen:
At the end of the second month |
a human soul is infused |
and how in all |
our holy mother foldeth ever souls |
for God's greater glory |
whereas that earthly mother |
which was but a dam to bear beastly |
should die by canon |
\gab{250}
for so saith he |
that holdeth the fisherman's seal,
even that blessed Peter |
on which rock |
was holy church |
for all ages |
founded.

\N2:
All they bachelors |
then asked of sir Leopold:

\All:
Would thee in like case |
so jeopard her person |
as risk life to save life?

\N3:
A wariness of mind |
he would answer |
as fitted all |
and,
laying hand to jaw,
he said |
dissembling,
as his wont was,
that |
as it was informed him,
who had ever loved the art of physic |
as might a layman,
and agreeing also |
with his experience of so seldom·seen an accident |
it was good |
for that mother Church |
belike at one blow |
had birth and death pence |
and in such sort deliverly |
he scaped their questions.

\dixon:
That is truth,
pardy,
and,
or I err,
a pregnant word.

\N1:
Which hearing |
\gab{260}
young Stephen was a marvellous glad man |
and he averred
“he who stealeth from the poor |
lendeth to the Lord”
for he was of a wild manner |
when he was drunken |
and that he was now in that taking |
it appeared eftsoons.

\week{15}

\N2:
But sir Leopold was passing grave maugre his word |
by cause he still had pity |
of the terror·causing shrieking of shrill women |
in their labour |
and as he was minded of his good lady Marion |
that had borne him an only manchild |
which on his eleventh day on live had died |
and no man of art could save |
so dark is destiny.
And she was wondrous stricken of heart |
for that evil hap |
and for his burial did him |
on a fair corselet of lamb's wool,
the flower of the flock,
\gab{270}
lest he might perish utterly |
and lie akeled
(for it was then about the midst of the winter)
and now Sir Leopold |
that had of his body no manchild for an heir |
looked upon him his friend's son |
and was shut up in sorrow for his forepassed happiness |
and as sad as he was |
that him failed a son of such gentle courage
(for all accounted him of real parts)
so grieved he also |
in no less measure |
for young Stephen |
for that he lived riotously with those wastrels |
and murdered his goods with whores.

\week{16}

\histstyle{Holinshead}

\N3:
About that present time |
young Stephen filled all cups that stood empty |
so as there remained but little mo |
if the prudenter had not shadowed their approach from him |
that still plied it very busily |
who, praying for the intentions
\gab{280}
of the sovereign pontiff:

\stephen:
I give ye for a pledge |
the vicar of Christ |
which also is vicar of Bray.
Now drink we of this mazer |
and quaff ye this mead |
which is not indeed |
parcel of my body |
but my soul's bodiment.
Leave ye fraction of bread to them |
that live by bread alone.
Be not afeard |
neither for any want |
for this will comfort |
more than the other will dismay.
See ye here.

\N1:
And he showed them |
glistering coins of the tribute |
and goldsmith notes the worth of two pound nineteen shilling |
that he had,
he said,
for a song which he writ.
They all admired to see the foresaid riches |
in such dearth of money |
as was herebefore.
His words were then these |
as followeth:

\stephen:
Know all men,
time's ruins build eternity's mansions.
\gab{290}
What means this?
Desire's wind blasts the thorntree |
but after |
it becomes from a bramblebush to be a rose |
upon the rood of time.
Mark me now.
In woman's womb |
word is made flesh |
but in the spirit of the maker |
all flesh that passes |
becomes the word that shall not pass away.
This is the post·creation.
\latin{Omnis caro ad te veniet}.
No question |
but her name is puissant |
who aventried the dear corse of our Agenbuyer,
Healer and Herd,
our mighty mother and mother most venerable |
and Bernardus saith aptly |
that She hath an \latin{omnipotentiam deiparae supplicem},
that is |
to wit,
an almightiness of petition |
because she is the second Eve |
and she won us,
saith Augustine too,
whereas that other,
our grandam,
which we are linked up with |
by successive anastomosis of navel·cords
\gab{300}
sold us all,
seed,
breed and generation,
for a penny pippin.
But here is the matter now.
Or she knew him,
that second I say,
and was |
but creature of her creature,
\italian{vergine madre,
figlia di tuo figlio},
or she knew him not |
and then stands she |
in the one denial or ignorancy |
with Peter Piscator |
who lives in the house that Jack built |
and with Joseph the joiner |
patron of the happy demise of all unhappy marriages,
\french{parceque M.~Leo Taxil nous a dit |
que qui l'avait mise |
dans cette fichue position |
c'etait le sacre pigeon,
ventre de Dieu!}
\german{Entweder} tran·substantiality \german{oder} con·substantiality
but in no case |
sub·substantiality.

\N2:
And all cried out upon it:

\All:
A very scurvy word!

\stephen:
A pregnancy without joy,
a birth without pangs,
\gab{310}
a body without blemish,
a belly without bigness.
Let the lewd |
with faith and fervour |
worship.
With will |
will we withstand,
withsay.

\week{17}

\N3:
Hereupon |
Punch Costello dinged with his fist upon the board |
and would sing a bawdy catch \song{Staboo Stabella} |
about a wench that was put in pod |
of a jolly swashbuckler in Almany |
which he did straight·ways now attack:

\punch:
\song{The first three months |
she was not well,
staboo.}

\N1:
When here |
nurse Quigley from the door |
angerly bid them:

\nursequigley:
Hist ye |
should shame you |
nor was it not meet |
as I remembered you being.

\N2:
Her mind was to have all orderly |
\gab{320}
against lord Andrew came for |
because she was jealous |
that no gasteful turmoil might shorten |
the honour of her guard.
It was an ancient and a sad matron |
of a sedate look and christian walking,
in habit dun |
beseeming her megrims and wrinkled visage.

\N3:
Nor did her hortative want of it effect |
for incontinently Punch Costello was of them all embraided |
and they reclaimed the churl |
with civil rudeness some |
and shaked him with menace of blandishments others |
whiles they all chode with him:

\lynch:
A murrain seize the dolt!

\madden:
What a devil |
thee would be at!

\crotthers:
Thou chuff!

\dixon:
Thou puny!

\lenehan:
Thou got in peasestraw!

\lynch:
Thou losel!

\madden:
Thou chitterling!

\crotthers:
Thou spawn of a rebel!

\lenehan:
Thou dykedropt!

\dixon:
Thou abortion thou!

\stephen:
Shut up thy drunken drool out of that |
like a curse of God ape!

\N1:
The good sir Leopold
\gab{330}
that had for his cognisance
the flower of quiet,
margerain gentle,
advising also |
the time's occasion as most sacred |
and most worthy to be most sacred.
In Horne's house |
rest should reign.

\week{18}

\histstyle{Milton, Hooker}

\N2:
To be short |
this passage was scarce by |
when Master Dixon of Mary in Eccles,
goodly grinning,
asked young Stephen:

\dixon:
What was the reason |
why thou hast not cided |
to take friar's vows?

\stephen:
Obedience in the womb,
chastity in the tomb |
but involuntary poverty all mine days.

\N3:
Master Lenehan at this |
made return |
that he had heard of those nefarious deeds.

\lenehan:
And how,
as I heard hereof counted,
thou hast besmirched the lily virtue of a confiding female |
\gab{340}
which was corruption of minors.

\N1:
And they all intershowed it too,
waxing merry and toasting:

\All:
To thy fathership!

\stephen:
But it was clean contrary to your suppose |
for I am the eternal son and ever virgin.

\N2:
Thereat |
mirth grew in them the more |
and they rehearsed to him |
his curious rite of wedlock |
for the disrobing and deflowering of spouses,
as the priests use in Madagascar island,
she to be in guise of white and saffron,
her groom in white and grain,
with burning of nard and tapers,
on a bridebed |
while clerks sung kyries |
and the anthem \latin{Ut novetur sexus omnis corporis mysterium} |
till she was there |
unmaided.

\N3:
He gave them then |
a much admirable hymen minim |
by those delicate poets Master John Fletcher and Master Francis Beaumont |
\gab{350}
that is in their \emph{Maid's Tragedy} |
that was writ for a like twining of lovers:
\emph{To bed, to bed}
was the burden of it |
to be played with accompanable concent |
upon the virginals.
An exquisite dulcet epithalame |
of most mollificative suadency |
for juveniles amatory |
whom the odoriferous flambeaus of the paranymphs |
have escorted to the quadrupedal proscenium of connubial communion.

\dixon:
Well met they were,
joyed,
but,
harkee,
young sir,
better were they named Beau Mount and Lecher |
for,
by my troth,
of such a mingling,
much might come.

\stephen:
To my best remembrance |
they had but the one doxy between them |
and she of the stews |
to make shift with in delights amorous |
for life ran very high in those days
\gab{360}
and the custom of the country |
approved with it.
Greater love than this |
no man hath |
that a man lay down his wife |
for his friend.
Go thou
and do likewise.
Thus,
or words to that effect,
saith Zarathustra,
sometime regius professor of French letters |
to the university of Oxtail |
nor breathed there ever |
that man to whom |
mankind was more beholden.
Bring a stranger within thy tower |
it will go hard |
but thou wilt have the second·best bed.
\latin{Orate,
fratres,
pro memetipso}.
And all the people shall say,
Amen.
Remember,
Erin,
thy generations and thy days of old,
how thou settedst |
little by me and by my word |
and broughtedst in a stranger to my gates |
to commit fornication in my sight |
and to wax fat and kick like Jeshurum.
\gab{370}
Therefore |
hast thou sinned against my light |
and hast made me,
thy lord,
to be the slave of servants.

[omitted in 2008]

\stephen:
Return,
return,
Clan Milly:
forget me not,
O Milesian.
Why hast thou done this abomination |
before me |
that thou didst spurn me |
for a merchant of jalaps |
and didst deny me to the Roman |
and to the Indian of dark speech |
with whom thy daughters |
did lie luxuriously?
Look forth now,
my people,
upon the land of behest,
even from Horeb |
and from Nebo |
and from Pisgah |
and from the Horns of Hatten |
unto a land flowing with milk and money.
But thou hast suckled me |
with a bitter milk:
my moon and my sun |
thou hast quenched for ever.
And thou hast left me alone |
for ever in the dark ways of my bitterness:
and with a kiss of ashes |
hast thou kissed my mouth.
\gab{380}

\N1:
This tenebrosity of the interior,
he proceeded to say,
hath not been illumined |
by the wit of the septuagint |
nor so much as mentioned |
for the Orient from on high |
Which brake hell's gates |
visited a darkness that was foraneous.
Assuefaction minorates atrocities
(as Tully saith of his darling Stoics)
and Hamlet his father |
showeth the prince |
no blister of combustion.
The adiaphane in the noon of life |
is an Egypt's plague |
which in the nights of prenativity and postmortemity |
is their most proper \latin{ubi} and \latin{quomodo}.
And as the ends and ultimates of all things |
accord in some mean and measure |
with their inceptions and originals,
that same multiplicit concordance |
which leads forth growth from birth
\gab{390}
accomplishing by a retrogressive metamorphosis |
that minishing and ablation towards the final |
which is agreeable unto nature |
so is it with our subsolar being.
The aged sisters draw us into life:
we wail,
batten,
sport,
clip,
clasp,
sunder,
dwindle,
die:
over us dead |
they bend.
First,
saved from waters of old Nile,
among bulrushes,
a bed of fasciated wattles:
at last |
the cavity of a mountain,
an occulted sepulchre |
amid the conclamation of the hillcat and the ossifrage.
And as no man knows |
the ubicity of his tumulus |
nor to what processes we shall thereby be ushered |
nor whether to Tophet or to Edenville |
in the like way |
is all hidden |
when we would backward see |
from what region of remoteness |
the what·ness of our who·ness |
\gab{400}
hath fetched his whence·ness.

\week{19}

\N1:
Thereto Punch Costello roared out mainly \french{Étienne chanson} |
but he loudly bid them.

\stephen:
Lo,
wisdom hath built herself a house,
this vast majestic longstablished vault,
the crystal palace of the Creator,
all in apple·pie order,
a penny for him |
who finds the pea.

\stephen:
\song{Behold the mansion reared by dedal Jack{}\\
See the malt stored in many a refluent sack,\\
In the proud cirque of Jackjohn's bivouac.}

\week{20}

\histstyle{Burton, Browne}

\N2:
A black crack of noise in the street here,
alack,
bawled back.
Loud on left |
Thor thundered:
in anger awful |
the hammerhurler.
Came now the storm |
that hist his heart.
\gab{410}

\lynch:
Have a care |
to flout and wit·wanton |
as the god self is angered |
for your hell·prate and paganry.

\N3:
And he |
that had erst challenged to be so doughty |
waxed wan |
as they might all mark |
and shrank together |
and his pitch that was before so haught uplift |
was now of a sudden |
quite plucked down |
and his heart shook within the cage of his breast |
as he tasted the rumour of that storm.
Then did some mock |
and some jeer |
and Punch Costello fell hard again to his yale |
which Master Lenehan vowed he would do after |
and he was indeed |
but a word and a blow |
on any the least colour.
But the braggart boaster cried:

\stephen:
An old Nobodaddy |
is in his cups |
it was much·what indifferent |
and I shall not lag behind his lead.
\gab{420}

\N1:
But this was only to dye his desperation |
as cowed |
he crouched in Horne's hall.
He drank indeed at one draught |
to pluck up a heart of any grace |
for it thundered long rumblingly |
over all the heavens |
so that Master Madden,
being godly certain whiles,
knocked him on his ribs |
upon that crack of doom |
and Master Bloom,
at the braggart's side,
spoke to him |
calming words |
to slumber his great fear

\Bloom:
It was no other thing |
but a hubbub noise |
that you heard,
the discharge of fluid from the thunderhead,
look you,
having taken place,
and all of the order |
of a natural phenomenon.

\week{21}

\histstyle{Bunyan}

\N2:
But was young Boasthard's fear |
vanquished by Calmer's words?
No,
for he had in his bosom |
a spike named Bitterness |
\gab{430}
which could not by words be done away.
And was he then |
neither calm like the one |
nor godly like the other?
He was neither as much |
as he would have liked to be either.
But could he not |
have endeavoured to have found again |
as in his youth |
the bottle Holiness |
that then he lived withal?

\N3:
Indeed no |
for Grace was not there |
to find that bottle.
Heard he then |
in that clap |
the voice of the god Bringforth |
or,
what Calmer said,
a hubbub of Phenomenon?
Heard?
Why,
he could not but hear |
unless he had plugged him up |
the tube Understanding
(which he had not done).
For through that tube |
he saw that he was in the land of Phenomenon |
where he must for a certain |
one day die |
as he was like the rest too |
a passing show.
\gab{440}
And would he not |
accept to die |
like the rest |
and pass away?

\N1:
By no means |
would he |
though he must |
nor would he |
make more shows |
according as men do with wives |
which Phenomenon has commanded them to do |
by the book Law.
Then wotted he |
nought of that other land |
which is called Believe-on-Me,
that is the land of promise |
which behoves to the king Delightful |
and shall be for ever |
where there is no death and no birth |
neither wiving nor mothering |
at which all shall come |
as many as believe on it?

\N2:
Yes,
Pious had told him |
of that land |
and Chaste had pointed him |
to the way |
but the reason was |
that in the way |
he fell in with a certain whore |
of an eyepleasing exterior |
whose name,
she said,
is Bird-in-the-Hand |
\gab{450}
and she beguiled him wrongways |
from the true path |
by her flatteries |
that she said to him |
as,
Ho,
you pretty man,
turn aside hither |
and I will show you a brave place,
and she lay at him so flatteringly |
that she had him in her grot |
which is named Two-in-the-Bush |
or,
by some learned,
Carnal Concupiscence.

\week{22}

\N3:
This was it |
what all that company that sat there |
at commons in Manse of Mothers |
the most lusted after |
and if they met with this whore Bird-in-the-Hand
(which was within all foul plagues,
monsters and a wicked devil)
they would strain the last |
but they would make at her |
and know her.
For regarding Believe-on-Me |
they said it was nought else but notion |
and they could conceive no thought of it.
\gab{460}

\N1:
For,
First,
Two-in-the-Bush |
whither she ticed them |
was the very goodliest grot |
and in it were four pillows |
on which were four tickets |
with these words printed on them,
Pickaback and Topsyturvy and Shameface and Cheek by Jowl.

\N2:
And,
second,
for that foul plague |
Allpox and the monsters |
they cared not for them |
for Preservative had given them |
a stout shield of oxengut |
and,
third,
that they might take no hurt |
neither from Offspring |
that was that wicked devil |
by virtue of this same shield |
which was named Killchild.
So were they all |
in their blind fancy,
Mr Cavil and Mr Sometimes Godly,
Mr Ape Swillale,
Mr False Franklin,
Mr Dainty Dixon,
Young Boasthard
and Mr Cautious Calmer.

\N3:
Wherein,
O wretched company,
\gab{470}
were ye all deceived |
for that was the voice of the god |
that was in a very grievous rage |
that he would presently lift his arm up |
and spill their souls |
for their abuses |
and their spillings done by them |
contrariwise to his word |
which forth |
to bring brenningly biddeth.

\week{23}

\histstyle{Evelyn, Pepys}

\N1:
So Thursday sixteenth June |
Patrick Dignam laid in clay |
of an apoplexy |
and after hard drought,
please God,
rained,
a bargeman coming in by water |
a fifty mile or thereabout |
with turf |
saying the seed won't sprout,
fields athirst,
very sadcoloured and stunk mightily,
the quags and tofts too.
Hard to breathe |
and all the young quicks |
clean consumed without sprinkle |
this long while back |
as no man remembered to be without.
The rosy buds all gone brown |
and spread out blobs |
\gab{480}
and on the hills |
nought but dry flag and faggots |
that would catch at first fire.
All the world saying,
for aught they knew,
the big wind of last February a year |
that did havoc the land so pitifully |
a small thing beside this barrenness.

\N2:
But by and by,
as said,
this evening after sundown,
the wind sitting in the west,
biggish swollen clouds to be seen |
as the night increased |
and the weatherwise poring up at them |
and some sheet lightnings at first |
and after,
past ten of the clock,
one great stroke with a long thunder |
and in a brace of shakes |
all scamper pellmell within door |
for the smoking shower,
the men making shelter for their straws |
with a clout or kerchief,
womenfolk skipping off with kirtles catched up |
soon as the pour came.

\N3:
In Ely place,
\gab{490}
Baggot street,
Duke's lawn,
thence through Merrion green |
up to Holles street |
a swash of water flowing |
that was before bonedry |
and not one chair or coach or fiacre seen about |
but no more crack after that first.
Over against the Right Honourable Mr Justice Fitzgibbon's door
(that is to sit with Mr Healy the lawyer |
upon the college lands),
Malachi Mulligan |
a gentleman's gentleman |
that had but come from Mr Moore's the writer's
(that was a papish |
but is now,
folk say,
a good Williamite)
chanced against Alec Bannon |
in a cut bob
(which are now in with dance cloaks of Kendal green)
that was new got to town from Mullingar |
with the stage |
where his coz and Malachi M's brother |
will stay a month yet |
till Saint Swithin |
\gab{500}
and asks |
what in the earth he does there,
he bound home |
and he to Andrew Horne's |
being stayed for |
to crush a cup of wine,
so he said,
but would tell him |
of a skittish heifer,
big of her age and beef to the heel,
and all this |
while poured with rain |
and so both together |
on to Horne's.

\N1:
There |
Leopold Bloom of Crawford's journal |
sitting snug with a covey of wags,
likely brangling fellows,
Dixon junior,
scholar of my lady of Mercy's;
Vincent Lynch;
a Scots fellow;
William Madden;
T.~Lenehan,
very sad about a racer he fancied;
and Stephen D.
Leopold Bloom |
there for a languor he had |
but was now better,
be having dreamed tonight |
a strange fancy of his dame Mrs Moll |
with red slippers on |
in a pair of Turkey trunks |
which is thought by those in ken |
to be for a change |
\gab{510}
and Mistress Purefoy there,
that got in |
through pleading her belly,
and now on the stools,
poor body,
two days past her term,
the midwives sore put to it |
and can't deliver,
she queasy for a bowl of riceslop |
that is a shrewd drier up of the insides |
and her breath very heavy |
more than good |
and should be a bullyboy |
from the knocks,
they say,
but God give her soon |
issue.

\N2:
'Tis her ninth chick to live,
I hear,
and Lady day |
bit off her last chick's nails |
that was then a twelvemonth |
and with other three all breastfed |
that died |
written out in a fair hand |
in the king's bible.
Her hub fifty odd |
and a methodist |
but takes the sacrament |
and is to be seen any fair sabbath |
with a pair of his boys |
off Bullock harbour |
dapping on the sound |
with a heavy·braked reel |
\gab{520}
or in a punt he has |
trailing for flounder and pollock |
and catches a fine bag,
I hear.

\N3:
In sum |
an infinite great fall of rain |
and all refreshed |
and will much increase the harvest yet |
those in ken say |
after wind and water |
fire shall come |
for a prognostication of Malachi's almanac
(and I hear |
that Mr Russell has done a prophetical charm of the same gist |
out of the Hindustanish |
for his farmer's gazette)
to have three things in all |
but this a mere fetch |
without bottom of reason |
for old crones and bairns |
yet sometimes they are found in the right guess |
with their queerities |
no telling how.

\week{24}

\histstyle{Defoe}

\N1:
With this |
came up Lenehan |
to the feet of the table |
\gab{530}
to say how the letter |
was in that night's gazette |
and he made a show |
to find it about him
(for he swore with an oath |
that he had been at pains about it)
but on Stephen's persuasion |
he gave over the search |
and was bidden to sit near by |
which he did |
mighty brisk.

\N2:
He was a kind of sport gentleman |
that went for a merryandrew or honest pickle |
and what belonged of women,
horseflesh or hot scandal |
he had it pat.
To tell the truth |
he was mean in fortunes |
and for the most part |
hankered about the coffeehouses and low taverns |
with crimps,
ostlers,
bookies,
Paul's men,
\dots

\N3:
runners,
flatcaps,
waist·coat·eers,
ladies of the bagnio |
and other rogues of the game |
or with a chanceable catchpole |
or a tipstaff |
often at nights till broad day |
of whom he picked up |
between his sack·possets |
\gab{540}
much loose gossip.

\N1:
He took his ordinary at a boilingcook's |
and if he had but gotten into him |
a mess of broken victuals |
or a platter of tripes |
with a bare tester in his purse |
he could always bring himself off |
with his tongue,
some randy quip he had |
from a punk or whatnot |
that every mother's son of them |
would burst their sides.

\N2:
The other,
Costello that is,
hearing this talk |
asked:

\punch:
Is it poetry or a tale?

\lenehan:
Faith,
no,
Frank,
'tis all about Kerry cows |
that are to be butchered |
along of the plague.
But they can go hang,
for me |
with their bully beef,
a pox on it.
There's as good fish in this tin |
as ever came out of it.

\N3:
And very friendly |
he offered to take |
of some salty sprats that stood by |
which he had eyed wishly |
in the meantime |
\gab{550}
and found the place |
which was indeed the chief design of his embassy |
as he was sharpset.

\punch:
\french{Mort aux vaches.}

\N1:
Says Frank then |
in the French language |
that had been indentured to a brandyshipper |
that has a winelodge in Bordeaux |
and he spoke French like a gentleman too.
From a child |
this Frank had been a do-nought |
that his father,
a headborough,
who could ill keep him to school |
to learn his letters |
and the use of the globes,
matriculated at the university |
to study the mechanics |
but he took the bit between his teeth |
like a raw colt |
and was more familiar |
with the justiciary |
and the parish beadle |
than with his volumes.

\N2:
One time |
he would be a playactor,
then a sutler |
or a welsher,
then nought would keep him from the bearpit |
and the cocking main,
\gab{560}
then he was for the ocean sea |
or to hoof it on the roads |
with the romany folk,
kidnapping a squire's heir |
by favour of moonlight |
or fecking maids' linen |
or choking chicken behind a hedge.
He had been off |
as many times as a cat has lives |
and back again with naked pockets as many more |
to his father the headborough |
who shed a pint of tears |
as often as he saw him.

\Bloom:
What,
will they slaughter all?
I protest |
I saw them but this day morning |
going to the Liverpool boats.
I can scarce believe 'tis so bad.

\N3:
And he had experience |
of the like brood beasts |
and of springers,
greasy hoggets |
and wether wool,
\gab{570}
having been some years before |
actuary for Mr Joseph Cuffe,
a worthy salesmaster |
that drove his trade for live stock and meadow auctions |
hard by Mr Gavin Low's yard |
in Prussia street.

\Bloom:
I question with you there.
More like |
'tis the hoose or the timber tongue.

\N1:
Mr Stephen,
a little moved |
but very handsomely told him |
no such matter |
and that he had dispatches |
from the emperor's chief tailtickler |
thanking him for the hospitality,
that was sending over Doctor Rinderpest,
the bestquoted cowcatcher in all Muscovy,
with a bolus or two of physic |
to take the bull by the horns.

\lynch:
Come,
come,
plain dealing.
He'll find himself |
on the horns of a dilemma |
if he meddles with a bull that's Irish.

\stephen:
Irish by name |
and Irish by nature.
An Irish bull |
in an English chinashop.
\gab{580}

\dixon:
I conceive you.
It is that same bull |
that was sent to our island |
by farmer Nicholas,
the bravest cattlebreeder of them all,
with an emerald ring in his nose.

\lynch:
True for you,
and a bullseye into the bargain,
and a plumper and a portlier bull,
never shit on shamrock.
He had horns galore,
a coat of cloth of gold |
and a sweet smoky breath coming out of his nostrils |
so that the women of our island,
leaving doughballs and rollingpins,
followed after him |
hanging his bulliness in daisychains.

\dixon:
What for that,
but before he came over |
farmer Nicholas that was a eunuch |
had him properly gelded |
\gab{590}
by a college of doctors |
who were no better off than himself.
So be off now,
says he,
and do all |
my cousin german the lord Harry |
tells you |
and take a farmer's blessing,
and with that |
he slapped his posteriors very soundly.

\lynch:
But the slap and the blessing |
stood him friend,
for to make up |
he taught him a trick |
worth two of the other |
so that maid,
wife,
abbess and widow |
to this day affirm |
that they would rather |
any time of the month |
whisper in his ear |
in the dark of a cowhouse |
or get a lick on the nape |
from his long holy tongue |
than lie with the finest strapping young ravisher |
in the four fields of all Ireland.

\madden:
And they dressed him,
in a point shift and petticoat |
with a tippet and girdle |
\gab{600}
and ruffles on his wrists |
and clipped his forelock |
and rubbed him all over |
with spermacetic oil |
and built stables for him |
at every turn of the road |
with a gold manger in each |
full of the best hay in the market |
so that he could doss and dung |
to his heart's content.
By this time |
the father of the faithful
(for so they called him)
was grown so heavy |
that he could scarce walk to pasture.
To remedy which |
our cozening dames and damsels |
brought him his fodder |
in their apronlaps |
and as soon as his belly was full |
he would rear up on his hindquarters |
to show their ladyships a mystery |
and roar and bellow out of him |
in bulls' language |
and they all after him.

\lenehan:
Ay,
and so pampered was he |
that he would suffer nought to grow |
in all the land |
\gab{610}
but green grass for himself
(for that was the only colour |
to his mind)
and there was a board put up |
on a hillock |
in the middle of the island |
with a printed notice,
saying:
By the Lord Harry,
Green is the grass |
that grows on the ground.

\dixon:
And,
if ever he got scent |
of a cattleraider in Roscommon |
or the wilds of Connemara |
or a husbandman in Sligo |
that was sowing |
as much as a handful of mustard |
or a bag of rapeseed |
out he'd run amok |
over half the countryside |
rooting up with his horns |
whatever was planted |
and all |
by lord Harry's orders.

\lynch:
There was bad blood between them |
at first,
and the lord Harry called |
farmer Nicholas all the old Nicks in the world |
\gab{620}
and an old whoremaster |
that kept seven trulls in his house |
and I'll meddle in his matters,
says he.
I'll make that animal smell hell,
says he,
with the help of that good pizzle |
my father left me.

\dixon:
But one evening,
when the lord Harry |
was cleaning his royal pelt |
to go to dinner |
after winning a boatrace
(he had spade oars for himself |
but the first rule of the course |
was that the others |
were to row with pitchforks)
he discovered in himself |
a wonderful likeness to a bull |
and on picking up a blackthumbed chapbook |
that he kept in the pantry |
he found |
sure enough |
that he was a lefthanded descendant |
of the famous champion bull of the Romans,
\latin{Bos Bovum},
which is good bog Latin |
for boss of the show.

\lynch:
After that,
the lord Harry put his head |
into a cow's drinking·trough |
\gab{630}
in the presence of all his courtiers |
and pulling it out again |
told them all |
his new name.
Then,
with the water running off him,
he got into an old smock and skirt |
that had belonged to his grandmother |
and bought a grammar of the bulls' language |
to study |
but he could never learn a word of it |
except the first personal pronoun |
which he copied out big |
and got off by heart |
and if ever he went out for a walk |
he filled his pockets with chalk |
to write it upon what took his fancy,
the side of a rock |
or a teahouse table |
or a bale of cotton |
or a corkfloat.
In short,
he and the bull of Ireland |
were soon as fast friends |
as an arse and a shirt.

\stephen:
They were,
and the end was |
that the men of the island |
seeing no help was toward,
\gab{640}
as the ungrate women |
were all of one mind,
made a wherry raft,
loaded themselves and their bundles of chattels on shipboard,
set all masts erect,
manned the yards,
sprang their luff,
heaved to,
spread three sheets in the wind,
put her head between wind and water,
weighed anchor,
ported her helm,
ran up the jolly Roger,
gave three times three,
let the bullgine run,
pushed off in their bumboat |
and put to sea |
to recover the main of America.

\lynch:
Which was the occasion,
of the composing |
by a boatswain |
of that rollicking chanty:

\lynch:
\song{Pope Peter's but a pissabed.\\
A man's a man for a' that.}%
\gab{650}

\week{25}

\histstyle{Addison, Steele}

\stage{Enter Mulligan and Bannon}

\N2:
Our worthy acquaintance |
Mr Malachi Mulligan \stage{bows} |
now appeared in the doorway |
as the students were finishing their apologue |
accompanied with a friend |
whom he had just rencountered,
a young gentleman,
his name |
Alec Bannon \stage{bows},
who had late come to town,
it being his intention |
to buy a colour |
or a cornetcy in the fencibles |
and list for the wars.

\N3:
Mr Mulligan was civil enough |
to express some relish of it |
all the more |
as it jumped with a project of his own |
for the cure of the very evil |
that had been touched on.
Whereat |
he handed round to the company |
a set of pasteboard cards |
which he had had printed |
that day at Mr Quinnell's |
bearing a legend |
printed in fair italics:
\emph{Mr Malachi Mulligan.
Fertiliser and Incubator.
Lambay Island}.
\gab{660}
His project,
as he went on to expound,
was to withdraw |
from the round of idle pleasures |
such as form |
the chief business of sir Fopling Popinjay and sir Milksop Quidnunc in town |
and to devote himself |
to the noblest task |
for which our bodily organism has been framed.

\dixon:
Well,
let us hear of it,
good my friend.
I make no doubt |
it smacks of wenching.
Come,
be seated,
both.
'Tis as cheap |
sitting as standing.

\N1:
Mr Mulligan accepted of the invitation |
and,
expatiating upon his design,
told his hearers |
that he had been led into this thought |
by a consideration of the causes of sterility.

\mulligan:
Both the inhibitory and the prohibitory,
whether the inhibition |
in its turn |
are due to conjugal vexations |
\gab{670}
or to a parsimony of the balance |
as well as |
whether the prohibition proceeds from defects congenital |
or from proclivities acquired.
It grieves me plaguily,
to see the nuptial couch |
defrauded of its dearest pledges:
and to reflect upon |
so many agreeable females with rich jointures,
a prey to the vilest bonzes,
who hide their flambeau under a bushel |
in an uncongenial cloister |
or lose their womanly bloom |
in the embraces of some unaccountable muskin |
when they might multiply the inlets of happiness,
sacrificing the inestimable jewel of their sex |
when a hundred pretty fellows were at hand to caress,
this,
I assure you,
makes my heart weep.
To curb this inconvenient,
\gab{680}
having advised with certain counsellors of worth |
and inspected into this matter,
I have resolved |
to purchase in fee simple for ever |
the freehold of Lambay island from its holder,
lord Talbot de Malahide,
a Tory gentleman of note |
much in favour with our ascendancy party.
I propose to set up there |
a national fertilising farm |
to be named \latin{Omphalos} |
with an obelisk hewn and erected |
after the fashion of Egypt |
and to offer my dutiful yeoman services |
for the fecundation of any female |
of what grade of life soever |
who should there direct to me |
with the desire of fulfilling the functions |
of her natural.
Money is no object,
nor will I take a penny for my pains.
The poorest kitchenwench |
no less than the opulent lady of fashion,
\gab{690}
if so be their constructions and their tempers |
were warm persuaders for their petitions,
would find |
in me |
their man.

\N2:
For his nutriment |
he showed how he would feed himself |
exclusively upon a diet |
of savoury tubercles and fish and coneys there,
the flesh of these latter prolific rodents |
being highly recommended for his purpose,
both broiled |
and stewed with a blade of mace |
and a pod or two of capsicum chillies.

\N3:
After this homily |
which he delivered with much warmth of asseveration |
Mr Mulligan in a trice |
put off from his hat |
a kerchief with which |
he had shielded it.
They both,
it seems,
had been overtaken by the rain |
and for all their mending |
their pace had taken water,
\gab{700}
as might be observed |
by Mr Mulligan's small·clothes of a hodden grey |
which was now somewhat piebald.
His project meanwhile |
was very favourably entertained by his auditors |
and won hearty eulogies from all |
though Mr Dixon of Mary's |
excepted to it,
asking with a finicking air |
did he purpose also |
to carry coals to Newcastle.

\N1:
Mr Mulligan however |
made court to the scholarly |
by an apt quotation from the classics |
which,
as it dwelt upon his memory,
seemed to him |
a sound and tasteful support of his contention:

\mulligan:
\latin{Talis ac tanta depravatio hujus seculi,
O quirites,
ut matresfamiliarum nostrae lascivas cujuslibet semiviri libici titillationes
testibus ponderosis atque excelsis erectionibus
centurionum Romanorum magnopere anteponunt.}

\N2:
While for those of ruder wit |
\gab{710}
he drove home his point |
by analogies of the animal kingdom |
more suitable to their stomach,
the buck and doe of the forest glade,
the farmyard drake and duck.

\week{26}

\N3:
Valuing himself not a little |
upon his elegance,
being indeed |
a proper man of person,
this talkative |
now applied himself to his dress |
with animadversions of some heat |
upon the sudden whimsy of the atmospherics |
while the company lavished their encomiums |
upon the project he had advanced.
The young gentleman,
his friend,
overjoyed as he was |
at a passage that had late befallen him,
could not forbear to tell it |
his nearest neighbour.

\N1:
Mr Mulligan,
now perceiving the table,
asked for whom |
were those loaves and fishes
and,
\gab{720}
seeing the stranger,
he made him a civil bow |
and said:

\mulligan:
\stage{to Bloom} Pray,
sir,
was you in need |
of any professional assistance |
we could give?

\N2:
Who,
upon his offer,
thanked him very heartily,
though preserving his proper distance,
and replied that:

\Bloom:
I am come here about a lady,
now an inmate of Horne's house,
that is in an interesting condition,
poor body,
from woman's woe \stage{sighs} |
to know if her happiness |
has yet taken place.

\N3:
Mr Dixon,
to turn the table,
took on |
to ask of Mr Mulligan himself:

\dixon:
Whether your incipient ventripotence |
betokens an ovoblastic gestation |
in the prostatic utricle |
or male womb,
or is due,
as with the noted physician,
Mr Austin Meldon,
to a wolf in the stomach?%
\gab{730}

\N1:
For answer |
Mr Mulligan,
in a gale of laughter at his smalls,
smote himself bravely below the diaphragm,
exclaiming |
with an admirable droll mimic of Mother Grogan
(the most excellent creature of her sex |
though 'tis pity she's a trollop):

\mulligan:
There's a belly |
that never bore a bastard.

\N2:
This was so happy a conceit |
that it renewed the storm of mirth |
and threw the whole room |
into the most violent agitations of delight.
The spry rattle had run on |
in the same vein of mimicry |
but for some larum in the antechamber.

\week{27}

\histstyle{Sterne}

\N3:
Here the listener |
who was none other than the Scotch student,
a little fume of a fellow,
blond as tow,
congratulated in the liveliest fashion |
with the young gentleman |
and,
\gab{740}
interrupting the narrative at a salient point,
having desired his visavis with a polite beck |
to have the obligingness |
to pass him a flagon of cordial waters |
at the same time |
by a questioning poise of the head
(a whole century of polite breeding |
had not achieved so nice a gesture)
to which |
was united an equivalent but contrary balance of the bottle |
asked the narrator |
as plainly as was ever done in words |
if he might treat him |
with a cup of it.

\bannon:
\french{Mais bien sûr},
noble stranger,
\french{et mille compliments}.
That you may |
and very opportunely.
There wanted nothing |
but this cup to crown my felicity.
But,
gracious heaven,
was I left |
with but a crust in my wallet |
and a cupful of water from the well,
my God,
I would accept of them |
\gab{750}
and find it in my heart |
to kneel down upon the ground |
and give thanks to the powers above |
for the happiness vouchsafed me |
by the Giver of good things.

\N1:
With these words |
he approached the goblet to his lips,
took a complacent draught of the cordial,
slicked his hair |
and,
opening his bosom,
out popped a locket |
that hung from a silk riband,
that very picture |
which he had cherished |
ever since her hand |
had wrote therein.
Gazing upon those features |
with a world of tenderness:

\bannon:
Ah,
Monsieur,
had you but beheld her |
as I did |
with these eyes |
at that affecting instant |
with her dainty tucker |
and her new coquette cap
(a gift for her feastday |
as she told me prettily)
in such an artless disorder,
of so melting a tenderness,
'pon my conscience,
even you,
Monsieur,
\gab{760}
had been impelled by generous nature |
to deliver yourself wholly into the hands of such an enemy |
or to quit the field for ever.
I declare,
I was never so touched in all my life.
God,
I thank thee,
as the Author of my days!
Thrice happy will he be |
whom so amiable a creature |
will bless with her favours.

\N2:
A sigh of affection |
gave eloquence to these words |
and,
having replaced the locket in his bosom,
he wiped his eye |
and sighed again.

\bannon:
Beneficent Disseminator of blessings |
to all Thy creatures,
how great and universal must be |
that sweetest of Thy tyrannies |
which can hold in thrall |
the free and the bond,
the simple swain |
and the polished coxcomb,
the lover in the heyday of reckless passion |
and the husband of maturer years.
But indeed,
sir,
I wander from the point.
\gab{770}
How mingled and imperfect |
are all our sublunary joys.
Maledicity!
\stage{exclaimed in anguish}
Would to God |
that foresight had but remembered me |
to take my cloak along!
I could weep |
to think of it.
Then,
though it had poured seven showers,
we were neither of us |
a penny the worse.
But beshrew me,
\stage{clapping hand to his forehead}
tomorrow will be a new day |
and,
thousand thunders,
I know of a \french{marchand de capotes},
Monsieur Poyntz,
from whom I can have |
for a \french{livre} |
as snug a cloak of the French fashion |
as ever kept a lady from wetting.

\N3:
Tut,
tut!
cries Le Fecondateur,
tripping in.

\mulligan:
My friend Monsieur Moore,
that most accomplished traveller
(I have just cracked a half bottle \french{avec lui} |
\gab{780}
in a circle of the best wits of the town),
is my authority |
that in Cape Horn,
\french{ventre biche},
they have a rain |
that will wet through any,
even the stoutest cloak.
A drenching of that violence,
he tells me,
\french{sans blague},
has sent more than one luckless fellow |
in good earnest |
posthaste to another world.

\lynch:
Pooh!
A \french{livre}!
The clumsy things are dear at a sou.
One umbrella,
were it no bigger than a fairy mushroom,
is worth ten such stopgaps.
No woman of any wit |
would wear one.
My dear Kitty told me today |
that she would dance in a deluge |
before ever she would starve |
in such an ark of salvation |
for,
as she reminded me
(blushing piquantly
and whispering in my ear |
though there was none to snap her words |
but giddy butterflies),
\gab{790}
dame Nature,
by the divine blessing,
has implanted it in our hearts |
and it has become a household word |
that \french{il y a deux choses}
for which |
the innocence of our original garb,
in other circumstances |
a breach of the proprieties,
is the fittest,
nay,
the only garment.
The first,
said she
(and here my pretty philosopher,
as I handed her to her tilbury,
to fix my attention,
gently tipped with her tongue |
the outer chamber of my ear),
the first |
is a bath\dots.

\N1:
But at this point |
a bell tinkling in the hall |
cut short a discourse |
which promised so bravely |
for the enrichment of our store of knowledge.

\week{28}

\histstyle{Goldsmith}

\N2:
Amid the general vacant hilarity of the assembly |
a bell rang |
and,
while all were conjecturing |
what might be the cause,
\gab{800}
Miss Callan entered |
and,
having spoken a few words in a low tone |
to young Mr Dixon,
retired with a profound bow to the company.

\N3:
The presence |
even for a moment |
among a party of debauchees |
of a woman endued with every quality of modesty |
and not less severe than beautiful |
refrained the humourous sallies |
even of the most licentious |
but her departure |
was the signal for an outbreak of ribaldry.

\N1:
Strike me silly,
said Costello,
a low fellow |
who was fuddled.

\punch:
A monstrous fine bit of cowflesh!
I'll be sworn |
she has rendezvoused you.
What,
you dog?
Have you a way with them?

\lynch:
Gad's bud,
immensely so.
The bedside manner it is |
that they use in the Mater hospice.
Demme,
\gab{810}
does not Doctor O'Gargle |
chuck the nuns there under the chin.
As I look to be saved |
I had it from my Kitty |
who has been wardmaid there |
any time these seven months.
Lawksamercy,
doctor,
\stage{feigning a womanish simper |
and with immodest squirmings of his body}
how you do tease a body!
Drat the man!
Bless me,
I'm all of a wibbly
wobbly.
Why,
you're as bad |
as dear little Father Cantekissem,
that you are!

\punch:
May this pot of four |
half choke me,
if she aint in the family way.
I knows a lady |
what's got a white swelling |
quick as I claps eyes on her.

\N2:
The young surgeon,
however,
rose and begged the company to excuse his retreat |
as the nurse had just then informed him |
that he was needed in the ward.
\gab{820}
Merciful providence had been pleased |
to put a period to the sufferings |
of the lady who was \french{enceinte} |
which she had borne with a laudable fortitude |
and she had given birth |
to a bouncing boy.

\dixon:
I want patience,
with those who,
without wit to enliven |
or learning to instruct,
revile an ennobling profession |
which,
saving the reverence due to the Deity,
is the greatest power for happiness |
upon the earth.
I am positive |
when I say that if need were |
I could produce a cloud of witnesses |
to the excellence of her noble exercitations |
which,
so far from being a byword,
should be a glorious incentive |
in the human breast.
I cannot away with them.
What?
Malign such an one,
the amiable Miss Callan,
\gab{830}
who is the lustre of her own sex |
and the astonishment of ours?
And at an instant |
the most momentous |
that can befall a puny child of clay?
Perish the thought!
I shudder to think |
of the future of a race |
where the seeds of such malice have been sown |
and where no right reverence is rendered |
to mother and maid in house of Horne.

\N3:
Having delivered himself of this rebuke |
he saluted those present on the by |
and repaired to the door.
A murmur of approval arose from all |
and some were for ejecting the low soaker |
without more ado,
a design which would have been effected |
nor would he have received more than his bare deserts |
had he not abridged his transgression |
by affirming with a horrid imprecation
(for he swore a round hand)
\gab{840}
that he was as good a son of the true fold |
as ever drew breath.

\punch:
Stap my vitals,
them was always |
the sentiments of honest Frank Costello |
which I was bred up most particular |
to honour thy father and thy mother |
that had the best hand to a rolypoly or a hasty pudding |
as you ever see |
what I always looks back on |
with a loving heart.

\week{29}

\histstyle{Burke}

\N1:
To revert to Mr Bloom |
who,
after his first entry,
had been conscious |
of some impudent mocks |
which he however had borne with |
as being the fruits of that age |
upon which it is commonly charged |
that it knows not pity.
The young sparks,
it is true,
were as full of extravagancies |
as overgrown children:
the words of their tumultuary discussions |
were difficultly understood |
and not often nice:
\gab{850}
their testiness and outrageous \french{mots} |
were such that his intellects resiled from:
nor were they scrupulously sensible |
of the proprieties |
though their fund of strong animal spirits |
spoke in their behalf.

\N2:
But the word of Mr Costello |
was an unwelcome language for him |
for he nauseated the wretch |
that seemed to him |
a cropeared creature |
of a misshapen gibbosity,
born out of wedlock |
and thrust like a crookback |
toothed and feet first |
into the world,
which the dint of the surgeon's pliers in his skull |
lent indeed a colour to,
so as to put him in thought |
of that missing link of creation's chain |
desiderated by the late ingenious Mr Darwin.

[omitted in 2008]

It was now |
for more than the middle span of our allotted years |
that he had passed |
through the thousand vicissitudes of existence |
\gab{860}
and,
being of a wary ascendancy |
and self a man of rare forecast,
he had enjoined his heart |
to repress all motions of a rising choler |
and,
by intercepting them with the readiest precaution,
foster within his breast |
that plenitude of sufferance |
which base minds jeer at,
rash judgers scorn |
and all find tolerable and but tolerable.
To those who create themselves wits |
at the cost of feminine delicacy
(a habit of mind |
which he never did hold with)
to them he would concede |
neither to bear the name |
nor to herit the tradition |
of a proper breeding:
while for such that,
having lost all forbearance,
can lose no more,
there remained |
the sharp antidote of experience |
to cause their insolency |
to beat a precipitate and inglorious retreat.
\gab{870}
Not but what he could feel |
with mettlesome youth |
which,
caring nought for the mows of dotards |
or the gruntlings of the severe,
is ever
(as the chaste fancy of the Holy Writer |
expresses it)
for eating of the tree forbid it |
yet not so far forth |
as to pretermit humanity |
upon any condition soever |
towards a gentlewoman |
when she was about her lawful occasions.

\N3:
To conclude,
while from the sister's words |
he had reckoned upon a speedy delivery |
he was,
however,
it must be owned,
not a little alleviated by the intelligence |
that the issue so auspicated |
after an ordeal of such duress |
now testified once more |
to the mercy |
as well as to the bounty |
of the Supreme Being.

\week{30}

\histstyle{Sheridan}

\N1:
Accordingly |
he broke his mind to his neighbour,
\gab{880}
saying that,
to express his notion of the thing,
his opinion
(who ought not perchance to express one)
is:

\Bloom:
That one must have a cold constitution |
and a frigid genius |
not to be rejoiced by this freshest news |
of the fruition of her confinement |
since she had been in such pain |
through no fault of hers.

\N2:
The dressy young blade |
said it was her husband's |
that put her in that expectation |
or at least it ought to be |
unless she were another Ephesian matron.

\N3:
I must acquaint you,
said Mr Crotthers,
clapping on the table |
so as to evoke |
a resonant comment of emphasis.

\crotthers:
Old Glory Allelujurum was round again today,
an elderly man with dundrearies,
preferring through his nose |
a request to have word of Wilhelmina,
my life,
as he calls her.
\gab{890}
I bade him hold himself in readiness |
for that the event would burst anon.
'Slife,
I'll be round with you.
I cannot but extol |
the virile potency of the old bucko |
that could still |
knock another child out of her.

\N1:
All fell to praising of it,
each after his own fashion,
though the same young blade |
held with his former view |
that another than her conjugial |
had been the man in the gap,
a clerk in orders,
a linkboy
(virtuous)
or an itinerant vendor of articles |
needed in every household.

\N2:
Singular,
communed the guest |
with himself.

\Bloom:
\stage{to himself}
The wonderfully unequal faculty of metempsychosis |
possessed by them,
that the puerperal dormitory |
and the dissecting theatre |
should be the seminaries |
of such frivolity,
that the mere acquisition of academic titles |
should suffice to transform |
\gab{900}
in a pinch of time |
these votaries of levity |
into exemplary practitioners of an art |
which most men anywise eminent |
have esteemed the noblest.
But,
it is mayhap to relieve the pentup feelings |
that in common oppress them |
for I have more than once |
observed that birds of a feather |
laugh together.

\week{31}

[omitted in 2008]

But with what fitness,
let it be asked of the noble lord,
his patron,
has this alien,
whom the concession of a gracious prince |
has admitted to civic rights,
constituted himself the lord paramount |
of our internal polity?
Where is now |
that gratitude which loyalty should have counselled?
During the recent war |
whenever the enemy |
had a temporary advantage with his granados |
did this traitor to his kind |
\gab{910}
not seize that moment |
to discharge his piece |
against the empire |
of which he is a tenant at will |
while he trembled for the security |
of his four per cents?
Has he forgotten this |
as he forgets all benefits received?
Or is it |
that from being a deluder of others |
he has become at last |
his own dupe |
as he is,
if report belie him not,
his own |
and his only enjoyer?
Far be it from candour |
to violate the bedchamber of a respectable lady,
the daughter of a gallant major,
or to cast the most distant reflections |
upon her virtue |
but if he challenges attention there
(as it was indeed |
highly his interest |
not to have done)
then be it so.
Unhappy woman,
she has been |
too long and too persistently denied |
her legitimate prerogative |
to listen to his objurgations |
with any other feeling |
\gab{920}
than the derision of the desperate.
He says this,
a censor of morals,
a very pelican in his piety,
who did not scruple,
oblivious of the ties of nature,
to attempt illicit intercourse |
with a female domestic |
drawn from the lowest strata of society!
Nay,
had the hussy's scouringbrush |
not been her tutelary angel,
it had gone with her |
as hard as with Hagar,
the Egyptian!
In the question of the grazing lands |
his peevish asperity is notorious |
and in Mr Cuffe's hearing |
brought upon him |
from an indignant rancher |
a scathing retort |
couched in terms as straight·forward |
as they were bucolic.
It ill becomes him |
to preach that gospel.
Has he not nearer home |
a seed·field that lies fallow |
for the want of the plough·share?
\gab{930}
A habit reprehensible at puberty |
is second nature |
and an opprobrium in middle life.
If he must dispense his balm of Gilead |
in nostrums and apothegms of dubious taste |
to restore to health |
a generation of unfledged profligates |
let his practice |
consist better with the doctrines |
that now engross him.
His marital breast |
is the repository of secrets |
which decorum is reluctant to adduce.
The lewd suggestions |
of some faded beauty |
may console him |
for a consort neglected and debauched |
but this new exponent of morals and healer of ills |
is at his best |
an exotic tree |
which,
when rooted in its native orient,
throve and flourished |
and was abundant in balm |
but,
transplanted to a clime more temperate,
its roots have lost their quondam vigour
\gab{940}
while the stuff that comes away from it |
is stagnant,
acid and inoperative.

\week{32}

\histstyle{Gibbon}

\N3:
The news was imparted with a circumspection recalling the ceremonial
usage of the Sublime Porte by the second female infirmarian to the junior
medical officer in residence,
who in his turn announced to the delegation
that an heir had been born.
When he had betaken himself to the women's apartment to assist at the prescribed ceremony of the afterbirth in the
presence of the secretary of state for domestic affairs and the members of the privy council,
silent in unanimous exhaustion and approbation the delegates,
chafing under the length and solemnity of their vigil and hoping that the joyful occurrence
\gab{950}
would palliate a licence which the
simultaneous absence of abigail and obstetrician rendered the easier,
broke out at once into a strife of tongues.

[omitted in 2008]

In vain the voice of Mr
Canvasser Bloom was heard endeavouring to urge,
to mollify,
to refrain.
The moment was too propitious for the display of that discursiveness
which seemed the only bond of union among tempers so divergent.
Every
phase of the situation was successively eviscerated:
the prenatal
repugnance of uterine brothers,
the Caesarean section,
posthumity with
respect to the father and,
that rarer form,
with respect to the mother,
the fratricidal case known as the Childs Murder and rendered memorable by
the impassioned plea of Mr Advocate Bushe which secured the acquittal of
the wrongfully accused,
\gab{960}
the rights of primogeniture and king's bounty
touching twins and triplets,
miscarriages and infanticides,
simulated or
dissimulated,
the acardiac \latin{foetus in foetu} and aprosopia due to a
congestion,
the agnathia of certain chinless Chinamen
(cited by Mr Candidate Mulligan)
in consequence of defective reunion of the maxillary
knobs along the medial line so that
(as he said)
one ear could hear what the other spoke,
the benefits of anesthesia or twilight sleep,
the prolongation of labour pains in advanced gravidancy by reason of pressure on the vein,
the premature relentment of the amniotic fluid
(as exemplified in the actual case)
with consequent peril of sepsis to the matrix,
artificial insemination by means of syringes,
\gab{970}
involution of the womb consequent upon the menopause,
the problem of the perpetration of
the species in the case of females impregnated by delinquent rape,
that distressing manner of delivery called
by the Brandenburghers \german{Sturzgeburt},
the recorded instances of multiseminal,
twikindled and monstrous births
conceived during the catamenic period or of consanguineous parents---in a
word all the cases of human nativity which Aristotle has classified in
his masterpiece with chromolithographic illustrations.
The gravest
problems of obstetrics and forensic medicine were examined with as much
animation as the most popular beliefs on the state of pregnancy such as
the forbidding to a gravid woman to step over a countrystile lest,
by her movement,
\gab{980}
the navelcord should strangle her creature and the injunction
upon her in the event of a yearning,
ardently and ineffectually
entertained,
to place her hand against that part of her person which long
usage has consecrated as the seat of castigation.
The abnormalities of harelip,
breastmole,
supernumerary digits,
negro's inkle,
strawberry mark
and portwine stain were alleged by one as a \latin{prima facie} and natural
hypothetical explanation of those swineheaded
(the case of Madame Grissel
Steevens was not forgotten)
or doghaired infants occasionally born.
The hypothesis of a plasmic memory,
advanced by the Caledonian envoy and
worthy of the metaphysical traditions of the land he stood for,
\gab{990}
envisaged in such cases an arrest of embryonic development at some stage antecedent
to the human.
An outlandish delegate sustained against both these views,
with such heat as almost carried conviction,
the theory of copulation
between women and the males of brutes,
his authority being his own
avouchment in support of fables such as that of the Minotaur which the
genius of the elegant Latin poet has handed down to us in the pages of
his Metamorphoses.
The impression made by his words was immediate but
shortlived.
It was effaced as easily as it had been evoked by an
allocution from Mr Candidate Mulligan in that vein of pleasantry which
none better than he knew how to affect,
postulating as the supremest object of desire a nice clean old man.
\gab{1000}
Contemporaneously,

\N1:
A heated argument having arisen between Mr Delegate Madden and Mr
Candidate Lynch regarding the juridical and theological dilemma created in
the event of one Siamese twin predeceasing the other,
the difficulty by
mutual consent was referred to Mr Canvasser Bloom for instant submittal to
Mr Coadjutor Deacon Dedalus.

\N2:
Hitherto silent,
whether the better to show by preternatural gravity
that curious dignity of the garb with which he was invested or in
obedience to an inward voice,
he delivered briefly and,
as some thought,
perfunctorily the ecclesiastical ordinance forbidding man to put asunder
what God has joined.

\week{33}

\histstyle{Walpole}

[omitted in 2008]

But Malachias' tale began to freeze them with horror.
\gab{1010}
He conjured up the
scene before them.
The secret panel beside the chimney slid back and in
the recess appeared \dots Haines!
Which of us did not feel his flesh creep!
He had a portfolio full of Celtic literature in one hand,
in the other a
phial marked \emph{Poison}.
Surprise,
horror,
loathing were depicted on all
faces while he eyed them with a ghostly grin.
I anticipated some such
reception,
he began with an eldritch laugh,
for which,
it seems,
history
is to blame.
Yes,
it is true.
I am the murderer of Samuel Childs.
And how
I am punished!
The inferno has no terrors for me.
This is the appearance
is on me.
Tare and ages,
what way would I be resting at all,
he muttered
thickly,
and I tramping Dublin this while back
\gab{1020}
with my share of songs and himself after me the like of a soulth or a bullawurrus?
My hell,
and
Ireland's,
is in this life.
It is what I tried to obliterate my crime.
Distractions,
rookshooting,
the Erse language
(he recited some),
laudanum
(he raised the phial to his lips),
camping out.
In vain!
His spectre
stalks me.
Dope is my only hope \dots Ah!
Destruction!
The black panther!
With a cry he suddenly vanished and the panel slid back.
An instant later
his head appeared in the door opposite and said:
Meet me at Westland Row
station at ten past eleven.
He was gone.
Tears gushed from the eyes of
the dissipated host.
The seer raised his hand to heaven,
murmuring:
The
vendetta of Mananaun!
The sage repeated:
\latin{Lex talionis}.
\gab{1030}
The sentimentalist is he who would enjoy without incurring the immense
debtorship for a thing done.
Malachias,
overcome by emotion,
ceased.
The
mystery was unveiled.
Haines was the third brother.
His real name was
Childs.
The black panther was himself the ghost of his own father.
He
drank drugs to obliterate.
For this relief much thanks.
The lonely house
by the graveyard is uninhabited.
No soul will live there.
The spider
pitches her web in the solitude.
The nocturnal rat peers from his hole.
A curse is on it.
It is haunted.
Murderer's ground.

\week{34}

\histstyle{Lamb}

\N3:
What is the age of the soul of man?
As she hath the virtue of the
chameleon to change her hue at every new approach,
to be gay with the
merry and mournful with the downcast,
\gab{1040}
so too is her age changeable as her mood.
No longer is Leopold,
as he sits there,
ruminating,
chewing the cud
of reminiscence,
that staid agent of publicity and holder of a modest
substance in the funds.

\N1:
A score of years are blown away.
He is young Leopold.
There,
as in a retrospective arrangement,
a mirror within a
mirror
(hey,
presto!),
he beholdeth himself.
That young figure of then is
seen,
precociously manly,
walking on a nipping morning from the old house
in Clanbrassil street to the high school,
his booksatchel on him
bandolierwise,
and in it a goodly hunk of wheaten loaf,
a mother's
thought.

[omitted in 2008]

Or it is the same figure,
a year or so gone over,
in his first
hard hat
(ah,
that was a day!),
already on the road,
\gab{1050}
a full·fledged traveller for the family firm,
equipped with an orderbook,
a scented
handkerchief
(not for show only),
his case of bright trinketware
(alas!
a thing now of the past!)
and a quiverful of compliant smiles for this or
that half·won housewife reckoning it out upon her fingertips or for a budding virgin,
shyly acknowledging
(but the heart?
tell me!)
his studied baisemoins.
The scent,
the smile,
but,
more than these,
the dark eyes and
oleaginous address,
brought home at duskfall many a commission to the
head of the firm,
seated with Jacob's pipe after like labours in the
paternal ingle
(a meal of noodles,
you may be sure,
is aheating),
reading through round horned spectacles some paper from the Europe of a month
before.

\N2:
But hey,
presto,
\gab{1060}
the mirror is breathed on and the young
knighterrant recedes,
shrivels,
dwindles to a tiny speck within the mist.
Now he is himself paternal and these about him might be his sons.
Who can
say?
The wise father knows his own child.

\N3:
He thinks of a drizzling night in Hatch street,
hard by the bonded
stores there,
the first.
Together
(she is a poor waif,
a child of shame,
yours and mine and of all for a bare shilling and her luckpenny),
together
they hear the heavy tread of the watch as two raincaped shadows pass the
new royal university.
Bridie!
Bridie Kelly!

\N1:
He will never forget the name,
ever remember the night:
first night,
the bridenight.
They are entwined in nethermost darkness,
the willer with the willed,
\gab{1070}
and in an instant
(\latin{fiat!})
light shall flood the world.

\N2:
Did heart leap to heart?
Nay,
fair reader.
In a breath 'twas
done but---hold!
Back!
It must not be!
In terror the poor girl flees away
through the murk.
She is the bride of darkness,
a daughter of night.
She
dare not bear the sunnygolden babe of day.

\N3:
No,
Leopold.
Name and memory solace thee not.
That youthful illusion of thy strength was taken from
thee---and in vain.
No son of thy loins is by thee.
There is none now to be for Leopold,
what Leopold was for Rudolph.

[omitted in 2008]

\week{35}

\histstyle{de Quincey}

The voices blend and fuse in clouded silence:
silence that is the
infinite of space:
and swiftly,
silently the soul is wafted over regions of cycles of generations that have lived.
\gab{1080}
A region where grey twilight ever descends,
never falls on wide sagegreen pasturefields,
shedding her
dusk,
scattering a perennial dew of stars.
She follows her mother with
ungainly steps,
a mare leading her fillyfoal.
Twilight phantoms are they,
yet moulded in prophetic grace of structure,
slim shapely haunches,
a
supple tendonous neck,
the meek apprehensive skull.
They fade,
sad
phantoms:
all is gone.
Agendath is a waste land,
a home of screechowls
and the sandblind upupa.
Netaim,
the golden,
is no more.
And on the
highway of the clouds they come,
muttering thunder of rebellion,
the
ghosts of beasts.
Huuh!
Hark!
Huuh!
Parallax stalks behind and goads
them,
the lancinating lightnings of whose brow are scorpions.
\gab{1090}
Elk and yak,
the bulls of Bashan and of Babylon,
mammoth and mastodon,
they come
trooping to the sunken sea,
\latin{Lacus Mortis}.
Ominous revengeful
zodiacal host!
They moan,
passing upon the clouds,
horned and capricorned,
the trumpeted with the tusked,
the lionmaned,
the giantantlered,
snouter
and crawler,
rodent,
ruminant and pachyderm,
all their moving moaning
multitude,
murderers of the sun.

\week{36}

[omitted in 2008]

Onward to the dead sea they tramp to drink,
unslaked and with horrible
gulpings,
the salt somnolent inexhaustible flood.
And the equine portent
grows again,
magnified in the deserted heavens,
nay to heaven's own
magnitude,
till it looms,
vast,
over the house of Virgo.
And lo,
wonder of metempsychosis,
it is she,
\gab{1100}
the everlasting bride,
harbinger of the
daystar,
the bride,
ever virgin.
It is she,
Martha,
thou lost one,
Millicent,
the young,
the dear,
the radiant.
How serene does she now
arise,
a queen among the Pleiades,
in the penultimate antelucan hour,
shod in sandals of bright gold,
coifed with a veil of what do you call it
gossamer.
It floats,
it flows about her starborn flesh and loose it
streams,
emerald,
sapphire,
mauve and heliotrope,
sustained on currents
of the cold interstellar wind,
winding,
coiling,
simply swirling,
writhing in the skies a mysterious writing till,
after a myriad
metamorphoses of symbol,
it blazes,
Alpha,
a ruby and triangled sign
upon the forehead of Taurus.

\week{37}

\histstyle{Landor}

\N1:
Francis was reminding Stephen of years before
\gab{1110}
when they had been at school together in Conmee's time.
He asked about Glaucon,
Alcibiades,
Pisistratus.
Where were they now?
Neither knew.

\stephen:
You have spoken of the past and its phantoms.
Why think of them?
If I
call them into life across the waters of Lethe will not the poor ghosts
troop to my call?
Who supposes it?
I,
Bous Stephanoumenos,
bullockbefriending bard,
am lord and giver of their life.

\N2:
He encircled his gadding hair with a
coronal of vineleaves,
smiling at Vincent.

\lynch:
That answer and those leaves,
will adorn you more fitly when
something more,
and greatly more,
than a capful of light odes can call
your genius father.
All who wish you well hope this for you.
\gab{1120}
All desire to see you bring forth the work you meditate,
to acclaim you Stephaneforos.
I heartily wish you may not fail them.

\lenehan:
O no,
Vincent.
Have no fear.
He could not leave his mother an orphan.

\N3:
The young man's face grew dark.
All could see how hard it was for
him to be reminded of his promise and of his recent loss.
He would have withdrawn from the feast had not the noise of voices allayed the smart.

\N1:
Madden had lost five drachmas on Sceptre for a whim of the rider's name:
Lenehan as much more.
He told them of the race.
The flag fell and,
huuh!
off,
scamper,
the mare ran out freshly with O. Madden up.
She was leading
the field.
All hearts were beating.%
\gab{1130}

\N2:
Even Phyllis could not contain herself.
She waved her scarf and cried:
Huzzah!
Sceptre wins!

\N3:
But in the straight on the run home when all were in close order the
dark horse Throwaway drew level,
reached,
outstripped her.

\N1:
All was lost now.

\N2:
Phyllis was silent:
her eyes were sad anemones.
Juno,
she cried,
I am
undone.
But her lover consoled her and brought her a bright casket of gold
in which lay some oval sugarplums which she partook.
A tear fell:
one
only.

\lenehan:
A whacking fine whip is W. Lane.
Four winners yesterday and
three today.
What rider is like him?
Mount him on the camel or the
boisterous buffalo the victory in a hack canter is still his.
But let us
bear it as was the ancient wont.
Mercy on the luckless!
Poor Sceptre!
\stage{sighing}
\gab{1140}
She is not the filly that she was.
Never,
by this
hand,
shall we behold such another.
By gad,
sir,
a queen of them.
Do you
remember her,
Vincent?

\lynch:
I wish you could have seen my queen today.
How young she was and radiant
(Lalage were scarce fair beside her)
in her yellow shoes and frock of muslin,
I do not know the right name of it.
The chestnuts that shaded us were in bloom:
the air drooped with their persuasive odour and with pollen floating by us.
In the sunny patches one might easily have cooked on a stone a batch of those buns
with Corinth fruit in them that Periplipomenes sells in his booth near the bridge.
But she had nought for her teeth but the arm with which I held her
\gab{1150}
and in that she nibbled mischievously when I pressed too close.
A week ago she lay ill,
four days on the couch,
but today she was free,
blithe,
mocked at peril.
She is more taking then.
Her posies too!
Mad romp that she is,
she had pulled her fill as we reclined together.
And in your ear,
my friend,
you will not think who met us as we left the field.
Conmee himself!
He was walking by the hedge,
reading,
I think a brevier book with,
I doubt not,
a witty letter in it from Glycera or Chloe to keep the page.
The sweet creature turned all colours in her confusion,
feigning to reprove a slight disorder in her dress:
a slip of underwood
clung there for the very trees adore her.
When Conmee had passed she
glanced at her lovely echo in that little mirror she carries.
\gab{1160}
But he had been kind.
In going by he had blessed us.

\lenehan:
The gods too are ever kind.
If I had poor luck with Bass's mare perhaps this draught of
his may serve me more propensely.

\N3:
He was laying his hand upon a winejar:
Malachi saw it and withheld his act,
pointing to the stranger and to the
scarlet label.

\mulligan:
\stage{whispering} Warily,
preserve a druid silence.
His soul is far away.
It is as painful perhaps to be awakened from a vision
as to be born.
Any object,
intensely regarded,
may be a gate of access to
the incorruptible eon of the gods.
Do you not think it,
Stephen?

\stephen:
Theosophos told me so,
\gab{1170}
whom in a previous existence
Egyptian priests initiated into the mysteries of karmic law.
The lords of the moon,
Theosophos told me,
an orangefiery shipload from planet Alpha
of the lunar chain would not assume the etheric doubles and these were
therefore incarnated by the rubycoloured egos from the second
constellation.

\week{38}

\histstyle{Macauley}

\N1:
However,
as a matter of fact though,
the preposterous surmise about him
being in some description of a doldrums or other or mesmerised which was
entirely due to a misconception of the shallowest character,
was not the
case at all.
The individual whose visual organs while the above was going
on were at this juncture commencing to exhibit symptoms of animation was
as astute if not astuter than any man living and anybody that conjectured the contrary
\gab{1180}
would have found themselves pretty speedily in the wrong shop.

\N2:
During the past four minutes or thereabouts he had been staring
hard at a certain amount of number one Bass bottled by Messrs Bass and Co
at Burton-on-Trent which happened to be situated amongst a lot of others
right opposite to where he was and which was certainly calculated to
attract anyone's remark on account of its scarlet appearance.
He was
simply and solely,
as it subsequently transpired for reasons best known
to himself,
which put quite an altogether different complexion on the
proceedings,
after the moment before's observations about boyhood days
and the turf,
recollecting two or three private transactions of his own
which the other two were as mutually innocent of
\gab{1190}
as the babe unborn.

\N3:
Eventually,
however,
both their eyes met and as soon as it began to dawn
on him that the other was endeavouring to help himself to the thing he
involuntarily determined to help him himself and so he accordingly took
hold of the neck of the mediumsized glass recipient which contained the
fluid sought after and made a capacious hole in it by pouring a lot of it
out with,
also at the same time,
however,
a considerable degree of
attentiveness in order not to upset any of the beer that was in it about
the place.

\week{39}

\N1:
The debate which ensued was in its scope and progress an epitome of the
course of life.
Neither place nor council was lacking in dignity.
The debaters were the keenest in the land,
\gab{1200}
the theme they were engaged on the
loftiest and most vital.
The high hall of Horne's house had never beheld
an assembly so representative and so varied nor had the old rafters of
that establishment ever listened to a language so encyclopaedic.
A
gallant scene in truth it made.

\N2:
Crotthers was there at the foot of the
table in his striking Highland garb,
his face glowing from the briny airs
of the Mull of Galloway.
There too,
opposite to him,
was Lynch whose
countenance bore already the stigmata of early depravity and premature
wisdom.
Next the Scotchman was the place assigned to Costello,
the
eccentric,
while at his side was seated in stolid repose the squat form
of Madden.
The chair of the resident indeed stood vacant before the hearth
\gab{1210}
but on either flank of it the figure of Bannon in explorer's kit
of tweed shorts and salted cowhide brogues contrasted sharply with the
primrose elegance and townbred manners of Malachi Roland St John
Mulligan.

\N3:
Lastly at the head of the board was the young poet who found a
refuge from his labours of pedagogy and metaphysical inquisition in the
convivial atmosphere of Socratic discussion,
while to right and left of
him were accommodated the flippant prognosticator,
fresh from the
hippodrome,
and that vigilant wanderer,
soiled by the dust of travel and
combat and stained by the mire of an indelible dishonour,
but from whose
steadfast and constant heart no lure or peril or threat or degradation
\gab{1220}
could ever efface the image of that voluptuous loveliness which the
inspired pencil of Lafayette has limned for ages yet to come.

\week{40}

\histstyle{Huxley}

\N1:
It had better be stated here and now at the outset that the perverted
transcendentalism to which Mr S. Dedalus' contentions would
appear to prove him pretty badly addicted runs directly counter to
accepted scientific methods.
Science,
it cannot be too often repeated,
deals with tangible phenomena.
The man of science like the man in the
street has to face hardheaded facts that cannot be blinked and explain
them as best he can.
There may be,
it is true,
some questions which
science cannot answer---at present---such as the first problem submitted by
Mr~L.~Bloom
\gab{1230}
regarding the future determination of sex.

\N2:
Must we accept the view of Empedocles of Trinacria
that the right ovary
(the postmenstrual period,
assert others)
is responsible for the birth of males or
are the too long neglected spermatozoa or nemasperms the
differentiating factors or is it,
as most embryologists incline to opine,
such as Culpepper,
Spallanzani,
Blumenbach,
Lusk,
Hertwig,
Leopold and
Valenti,
a mixture of both?
This would be tantamount to a cooperation
(one of nature's favourite devices) between
the \latin{nisus formativus} of the
nemasperm on the one hand and on the other a happily chosen position,
\latin{succubitus felix} of the passive element.

\N3:
The other problem raised by the same inquirer is scarcely less vital:
\gab{1240}
infant mortality.
It is interesting because,
as he pertinently remarks,
we
are all born in the same way but we all die in different ways.

\mulligan:
I blame the sanitary conditions in which our greylunged citizens
contract adenoids,
pulmonary complaints etc.
by inhaling the bacteria
which lurk in dust.
These factors,
and the revolting spectacles offered by
our streets,
hideous publicity posters,
religious ministers of all
denominations,
mutilated soldiers and sailors,
exposed scorbutic
cardrivers,
the suspended carcases of dead animals,
paranoic bachelors and
unfructified duennas---these are accountable
\gab{1250}
for any and every falling·off in the calibre of the race.
Kalipedia will soon be generally adopted and
all the graces of life,
genuinely good music,
agreeable literature,
light
philosophy,
instructive pictures,
plastercast reproductions of the
classical statues such as Venus and Apollo,
artistic coloured photographs
of prize babies,
all these little attentions would enable ladies who were
in a particular condition to pass the intervening months in a most
enjoyable manner.

\crotthers:
I attribute some of these demises to abdominal trauma in the case
of women workers subjected to heavy labours in the workshop and to
marital discipline in the home but by far the vast majority to neglect,
private or official,
\gab{1260}
culminating in the exposure of newborn infants,
the practice of criminal abortion or in the atrocious crime of infanticide.

\N1:
Although the former
(we are thinking of neglect)
is undoubtedly only too
true the case he cites of nurses forgetting to count the sponges in the
peritoneal cavity is too rare to be normative.
In fact when one comes to
look into it the wonder is that so many pregnancies and deliveries go off
so well as they do,
all things considered and in spite of our human
shortcomings which often baulk nature in her intentions.

\N2:
An ingenious suggestion is that thrown out by Mr V. Lynch:

\lynch:
Both natality and mortality,
as well as all other phenomena of evolution,
tidal movements,
lunar phases,
blood temperatures,
diseases in general,
\gab{1270}
everything,
in fine,
in nature's vast workshop from the extinction of
some remote sun to the blossoming of one of the countless flowers which
beautify our public parks is subject to a law of numeration as yet
unascertained.
Still the plain straightforward question why a child of
normally healthy parents and seemingly a healthy child and properly
looked after succumbs unaccountably in early childhood
(though other children of the same marriage do not)
must certainly,
in the poet's
words,
give us pause.
Nature,
we may rest assured,
has her own good and
cogent reasons for whatever she does and in all probability such deaths
are due to some law of anticipation by which organisms in which morbous
germs have taken up their residence
\gab{1280}
(modern science has conclusively shown that only the plasmic substance can be said to be immortal)
tend to disappear at an increasingly earlier stage of development,
an arrangement which,
though productive of pain to some of our feelings
(notably the maternal),
is nevertheless,
some of us think,
in the long run beneficial
to the race in general in securing thereby the survival of the fittest.

\stephen:
An omnivorous being which can masticate,
deglute,
digest and apparently pass through the ordinary channel with
pluterperfect imperturbability such multifarious aliments as cancrenous
females emaciated by parturition,
corpulent professional gentlemen,
\gab{1290}
not to speak of jaundiced politicians and chlorotic nuns,
might possibly find
gastric relief in an innocent collation of staggering bob.

\N3:
Mr S. Dedalus' remark reveals as
nought else could and in a very unsavoury light the tendency above
alluded to.
For the enlightenment of those who are not so intimately
acquainted with the minutiae of the municipal abattoir as this
morbidminded esthete and embryo philosopher who for all his overweening
bumptiousness in things scientific can scarcely distinguish an acid from
an alkali prides himself on being,
it should perhaps be stated that
staggering bob in the vile parlance of our lowerclass licensed
victuallers signifies the cookable and eatable flesh of a calf newly
dropped from its mother.

\N1:
In a recent public controversy
\gab{1300}
with Mr L. Bloom
which took place in the commons' hall of the National Maternity Hospital,
29, 30 and 31 Holles street,
of which,
as is well known,
Dr A. Horne
(Licence in Midwifery, F.K.Q.C.P.I.)
is the able and popular master,
he is reported by eyewitnesses as having stated that:

\stephen:
Once a woman has let the cat into the bag \dots

\N2:
(an esthete's allusion,
presumably,
to one of the most complicated and marvellous of all nature's processes---the
act of sexual congress)

\stephen:
She must let it out again or give it life to save her own.

\Bloom:
At the risk of her own.

\N3:
Was the telling rejoinder of his interlocutor,
none the less
effective for the moderate and measured tone in which it was delivered.

\histstyle{Dickens}

\N1:
Meanwhile the skill and patience of the physician
\gab{1310}
had brought about a happy \french{accouchement}.
It had been a weary weary while both for
patient and doctor.
All that surgical skill could do was done and the
brave woman had manfully helped.
She had.
She had fought the good fight
and now she was very very happy.
Those who have passed on,
who have gone
before,
are happy too as they gaze down and smile upon the touching scene.
Reverently look at her as she reclines there with the motherlight in her
eyes,
that longing hunger for baby fingers
(a pretty sight it is to see),
in the first bloom of her new motherhood,
breathing a silent prayer of
thanksgiving to One above,
the Universal Husband.

\N2:
And as her loving eyes behold her babe she wishes only one blessing more,
\gab{1320}
to have her dear Doady there with her to share her joy,
to lay in
his arms that mite of God's clay,
the fruit of their lawful embraces.
He
is older now
(you and I may whisper it)
and a trifle stooped in the shoulders yet in the whirligig of years a grave dignity has come to the
conscientious second accountant of the Ulster bank,
College Green branch.
O Doady,
loved one of old,
faithful lifemate now,
it may never be again,
that faroff time of the roses!

\N3:
With the old shake of her pretty head she recalls those days.
God!
How beautiful now across the mist of years!
But their children are
grouped in her imagination about the bedside,
hers and his,
Charley,
Mary
Alice,
Frederick Albert
(if he had lived),
Mamy,
Budgy
(Victoria Frances),
\gab{1330}
Tom,
Violet Constance Louisa,
darling little Bobsy
(called after our famous hero of the South African war,
lord Bobs of Waterford and Candahar)
and now this last pledge of their union,
a Purefoy if ever there was one,
with the true Purefoy nose.
Young hopeful will be christened Mortimer Edward after the influential third cousin of Mr
Purefoy in the Treasury Remembrancer's office,
Dublin Castle.
And so time
wags on:
but father Cronion has dealt lightly here.
No,
let no sigh break
from that bosom,
dear gentle Mina.

\N1:
And Doady,
knock the ashes from your pipe,
the seasoned briar you
still fancy when the curfew rings for you
(may it be the distant day!)
and dout the light whereby you read in the Sacred Book for the oil too has run low,
\gab{1340}
and so with a tranquil heart to bed,
to rest.
He knows and will call in His own good time.
You too have fought the good fight and played
loyally your man's part.
Sir,
to you my hand.
Well done,
thou good and faithful servant!

\histstyle{Newman}

\N2:
There are sins or
(let us call them as the world calls them)
evil memories which are hidden away by man in the darkest places of the heart
but they abide there and wait.
He may suffer their memory to grow dim,
let them be as though they had not been and all but persuade himself that
they were not or at least were otherwise.
Yet a chance word will call
them forth suddenly and they will rise up to confront him in the most
various circumstances,
a vision or a dream,
\gab{1350}
or while timbrel and harp
soothe his senses or amid the cool silver tranquility of the evening or
at the feast,
at midnight,
when he is now filled with wine.
Not to insult
over him will the vision come as over one that lies under her wrath,
not
for vengeance to cut him off from the living but shrouded in the piteous
vesture of the past,
silent,
remote,
reproachful.

\histstyle{Pater}

\N3:
The stranger still regarded on the face before him a slow recession of
that false calm there,
imposed,
as it seemed,
by habit or some studied
trick,
upon words so embittered as to accuse in their speaker an
unhealthiness,
a \emph{flair},
for the cruder things of life.

[omitted in 2008]

A scene disengages itself in the observer's memory,
evoked,
\gab{1360}
it would seem,
by a word of so natural a homeliness as if those days were really present there
(as some thought)
with their immediate pleasures.
A shaven space of lawn one soft
May evening,
the wellremembered grove of lilacs at Roundtown,
purple and
white,
fragrant slender spectators of the game but with much real
interest in the pellets as they run slowly forward over the sward or
collide and stop,
one by its fellow,
with a brief alert shock.
And yonder
about that grey urn where the water moves at times in thoughtful
irrigation you saw another as fragrant sisterhood,
Floey,
Atty,
Tiny and
their darker friend with I know not what of arresting in her pose then,
Our Lady of the Cherries,
a comely brace of them pendent from an ear,
\gab{1370}
bringing out the foreign warmth of the skin so daintily against the cool ardent fruit.
A lad of four or five in linseywoolsey
(blossom·time but there will be cheer in the kindly hearth when ere long the bowls are gathered and hutched)
is standing on the urn secured by that circle of girlish fond hands.
He frowns a little just as this young man does now
with a perhaps too conscious enjoyment of the danger but must needs
glance at whiles towards where his mother watches
from the \italian{piazzetta} giving upon the flowerclose
with a faint shadow of remoteness or of reproach
(\german{alles Vergängliche}) in her glad look.

\histstyle{Ruskin}

\N1:
Mark this farther and remember.
The end comes suddenly.
Enter that antechamber of birth where the studious are assembled
\gab{1380}
and note their faces.
Nothing,
as it seems,
there of rash or violent.
Quietude of
custody,
rather,
befitting their station in that house,
the vigilant
watch of shepherds and of angels about a crib in Bethlehem of Juda long
ago.

\N2:
But as before the lightning the serried stormclouds,
heavy with
preponderant excess of moisture,
in swollen masses turgidly distended,
compass earth and sky in one vast slumber,
impending above parched field
and drowsy oxen and blighted growth of shrub and verdure till in an
instant a flash rives their centres and with the reverberation of the
thunder the cloudburst pours its torrent,
so and not otherwise was the
transformation,
violent and instantaneous,
\gab{1390}
upon the utterance of the word.

\histstyle{Carlyle}

\stephen:
Burke's!

\N3:
Outflings my lord Stephen,
giving the cry,
and a tag and bobtail
of all them after,
cockerel,
jackanapes,
welsher,
pilldoctor,
punctual
Bloom at heels with a universal grabbing at headgear,
ashplants,
bilbos,
Panama hats and scabbards,
Zermatt alpenstocks and what not.
A dedale of
lusty youth,
noble every student there.

\N1:
Nurse Callan taken aback in the hallway cannot stay them nor smiling
surgeon coming downstairs with news of placentation ended,
a full pound if
a milligramme.
They hark him on.
The door!
It is open?
Ha!

\N2:
They are out,
tumultuously,
off for a minute's race,
all bravely
legging it,
Burke's of Denzille and Holles their ulterior goal.
\gab{1400}
Dixon follows giving them sharp language but raps out an oath,
he too,
and on.

\N3:
Bloom stays with nurse a thought to send a kind
word to happy mother and nurseling up there.
Doctor Diet and Doctor
Quiet.
Looks she too not other now?
Ward of watching in Horne's house has
told its tale in that washedout pallor.
Then all being gone,
a glance of
motherwit helping,
he whispers close in going:

\Bloom:
Madam,
when comes the storkbird for thee?

\N1:
The air without is impregnated with raindew moisture,
life essence
celestial,
glistening on Dublin stone there under starshiny \latin{coelum}.
God's air,
the Allfather's air,
scintillant circumambient cessile air.

\N2:
Breathe it deep into thee.
\gab{1410}
By heaven,
Theodore Purefoy,
thou hast done a doughty deed and no botch!
Thou art,
I vow,
the remarkablest progenitor barring none in this chaffering allincluding most farraginous chronicle.
Astounding!
In her lay a Godframed Godgiven preformed
possibility which thou hast fructified with thy modicum of man's work.
Cleave to her!
Serve!
Toil on,
labour like a very bandog and let
scholarment and all Malthusiasts go hang.

\N3:
Thou art all their daddies,
Theodore.
Art drooping
under thy load,
bemoiled with butcher's bills at home and ingots
(not thine!)
in the countinghouse?
Head up!
For every newbegotten thou shalt
gather thy homer of ripe wheat.
See,
thy fleece is drenched.
Dost envy
Darby Dullman there with his Joan?
A canting jay and a rheumeyed curdog
is all their progeny.
\gab{1420}
Pshaw,
I tell thee!
He is a mule,
a dead
gasteropod,
without vim or stamina,
not worth a cracked kreutzer.

\N1:
Copulation without population!
No,
say I!
Herod's slaughter of the
innocents were the truer name.
Vegetables,
forsooth,
and sterile
cohabitation!
Give her beefsteaks,
red,
raw,
bleeding!
She is a hoary
pandemonium of ills,
enlarged glands,
mumps,
quinsy,
bunions,
hayfever,
bedsores,
ringworm,
floating kidney,
Derbyshire neck,
warts,
bilious
attacks,
gallstones,
cold feet,
varicose veins.
A truce to threnes and
trentals and jeremies and all such congenital defunctive music!

\N2:
Twenty years of it,
regret them not.
With thee it was not as with
many that will and would and wait and never---do.
\gab{1430}
Thou sawest thy America,
thy lifetask,
and didst charge to cover like the transpontine bison.

\N3:
How saith Zarathustra?
\german{Deine Kuh Trűbsal melkest Du.
Nun trinkst Du die susse Milch des Euters}.

\N1:
See!
it displodes for thee in abundance.
Drink,
man,
an
udderful!
Mother's milk,
Purefoy,
the milk of human kin,
milk too of
those burgeoning stars overhead rutilant in thin rainvapour,
punch milk,
such as those rioters will quaff in their guzzling den,
milk of madness,
the honeymilk of Canaan's land.
Thy cow's dug was tough,
what?
Ay,
but
her milk is hot and sweet and fattening.
No dollop this but thick rich
bonnyclaber.
To her,
old patriarch!
Pap!
\latin{Per deam Partulam et Pertundam nunc est bibendum}!

\histstyle{Coda}

% \histstyle{Coda {\#}1}

\N2:
All off for a buster,
\gab{1440}
armstrong,
hollering down the street.

Bonafides.

Where you slep las nigh?

Timothy of the battered naggin.

Like ole Billyo.

Any brollies or gumboots in the fambly?

Where the Henry Nevil's sawbones and ole clo?

Sorra one o' me knows.

\madden:
Hurrah there,
Dix!

\N3:
Forward to the ribbon counter.

\dixon:
Where's Punch?

\N1:
All serene.

\bystander:
Jay,
look at the drunken minister coming out of the maternity hospal!

\mulligan:
\latin{Benedicat vos omnipotens Deus,
Pater et Filius}.

A make,
mister.

The Denzille lane boys.

Hell,
blast ye!

Scoot.

Righto,
Isaacs,
shove em out of the bleeding limelight.

\punch:
Yous join uz,
dear sir?

No hentrusion in life.

Lou heap good man.

Allee samee dis bunch.

\bannon:
\french{En avant,
mes enfants}!

\lenehan:
Fire away number one on the gun.

\All:
Burke's!
Burke's!

\N2:
Thence they advanced five parasangs.
\gab{1450}
Slattery's mounted foot.

Where's that bleeding awfur?

Parson Steve,
apostates' creed!

\lynch:
No,
no,
Mulligan!
Abaft there!
Shove ahead.

\lenehan:
Keep a watch on the clock.
Chuckingout time.

\madden:
Mullee!
What's on you?

\mulligan:
\french{Ma mère m'a mariée}.

British Beatitudes!

\latin{Retamplatan digidi boumboum}.

Ayes have it.

\stephen:
To be printed and bound at the Druiddrum press by two designing females.
Calf covers of pissedon green.
Last word in art shades.
Most beautiful book come out of Ireland my time.

\crotthers:
\latin{Silentium}!
Get a spurt on.

\lenehan:
Tention.
Proceed to nearest canteen and there annex liquor stores.

\punch:
March!
Tramp,
tramp,
tramp,
the boys are
(atitudes!)
parching.

\madden:
Beer,
beef,
business,
bibles,
bulldogs,
battleships,
buggery and bishops.%
\gab{1460}

\dixon:
Whether on the scaffold high.

\madden:
Beer,
beef,
trample the bibles.

\dixon:
When for Irelandear.

\crotthers:
Trample the trampellers.

\lenehan:
Thunderation!
Keep the durned millingtary step.

We fall.

Bishops boosebox.

\punch:
Halt!
Heave to.

\dixon:
Rugger.
Scrum in.
No touch kicking.

\lynch:
Wow,
my tootsies!

\dixon:
You hurt?
Most amazingly sorry!

% \histstyle{Coda {\#}2}

\lenehan:
Query.
Who's astanding this here do?

\madden:
Proud possessor of damnall.

\lynch:
Declare misery.
Bet to the ropes.

Me nantee saltee.

\dixon:
Not a red at me this week gone.

\barman:
Yours?

\stephen:
Mead of our fathers for the \german{Übermensch}.

\bannon:
Dittoh.

\stephen:
Five number ones.

\barman:
You,
sir?

\Bloom:
Ginger cordial.

Chase me,
the cabby's caudle.

Stimulate the caloric.

\N2:
Winding of his ticker.

Bl:
Stopped short never to go again when the old.

\stephen:
Absinthe for me,
savvy?
\gab{1470}

\mulligan:
\spanish{Caramba!} Have an eggnog or a prairie oyster.

\Bloom:
Enemy?

Avuncular's got my timepiece.

\madden:
Ten to.

\Bloom:
Obligated awful.

\madden:
Don't mention it.

\lynch:
Got a pectoral trauma,
eh,
Dix?

\dixon:
Pos fact.
Got bet be a boomblebee whenever he wus
settin sleepin in hes bit garten.
Digs up near the Mater.

Buckled he is.

\mulligan:
Know his dona?

\dixon:
Yup,
sartin I do.
Full of a dure.
See her in her dishybilly.
Peels off a credit.
Lovey lovekin.
None of your lean kine,
not much.

Pull down the blind,
love.

\madden:
Two Ardilauns.

\lynch:
Same here.

Look slippery.

If you fall don't wait to get up.

\barman:
Five,
seven,
nine.

Fine!

\dixon:
Got a prime pair of mincepies,
no kid.
And her take me to rests and her anker of rum.
Must be seen to be believed.
Your starving eyes and allbeplastered neck you stole my heart,
O gluepot.

\barman:
Sir?
\gab{1480}

Spud again the rheumatiz?

All poppycock,
you'll scuse me saying.

For the hoi polloi.

I vear thee beest a gert vool.

Well,
doc?

Back fro Lapland?

Your corporosity sagaciating O K?

How's the squaws and papooses?

Womanbody after going on the straw?

\lenehan:
Stand and deliver.

\crotthers:
Password.

\punch:
There's hair.

\Bloom:
Ours the white death and the ruddy birth.

Hi!

\bannon:
Spit in your own eye,
boss!

Mummer's wire.

Cribbed out of Meredith.

\mulligan:
Jesified,
orchidised,
polycimical jesuit!
Aunty mine's writing Pa Kinch.
Baddybad Stephen lead astray goodygood Malachi.

% \histstyle{Coda {\#}3}

Hurroo!

Collar the leather,
youngun.

Roun wi the nappy.

\lenehan:
Here,
Jock braw Hielentman's your barleybree.
\gab{1490}

\crotthers:
Lang may your lum reek and your kailpot boil!

My tipple.

\french{Merci}.

\dixon:
Here's to us.

\madden:
How's that?

\lynch:
Leg before wicket.

\mulligan:
Don't stain my brandnew sitinems.

Give's a shake of pepper,
you there.

Catch aholt.

Caraway seed to carry away.

Twig?

\N1:
Shrieks of silence.

Every cove to his gentry mort.

Venus Pandemos.

\french{Les petites femmes}.

\bannon:
Bold bad girl from the town of Mullingar.
Tell her I was axing at her.
Hauding Sara by the wame.

On the road to Malahide.

Me?

If she who seduced me had left but the name.

\lenehan:
What do you want for ninepence?

\stephen:
Machree,
macruiskeen.

\lynch:
Smutty Moll for a mattress jig.

\punch:
And a pull all together.

\All:
\latin{Ex!}

% \histstyle{Coda {\#}4}

Waiting,
guvnor?

Most deciduously.

Bet your boots on.

Stunned like,
seeing as how no shiners is acoming.
\gab{1500}

Underconstumble?

He've got the chink \latin{ad lib}.

Seed near free poun on un a spell ago a said war hisn.

Us come right in on your invite,
see?

Up to you,
matey.

Out with the oof.

Two bar and a wing.

You larn that go off of they there Frenchy bilks?

Won't wash here for nuts nohow.

Lil chile velly solly.

Ise de cutest colour coon down our side.

Gawds teruth,
Chawley.

We are nae fou.

We're nae tha fou.

Au reservoir,
mossoo.

Tanks you.

% \histstyle{Coda {\#}5}

'Tis,
sure.

What say?

In the speakeasy.

Tight.

I shee you,
shir.

Bantam,
two days teetee.

Bowsing nowt but claretwine.

Garn!

Have a glint,
do.

Gum,
I'm jiggered.

And been to barber he have.

Too full for words.

With a railway bloke.

How come you so?
\gab{1510}

Opera he'd like?

Rose of Castile.

Rows of cast.

Police!

Some H2O for a gent fainted.

Look at Bantam's flowers.

Gemini.

He's going to holler.

The colleen bawn.

My colleen bawn.

O,
cheese it!

Shut his blurry Dutch oven with a firm hand.

\lenehan:
Had the winner today till I tipped him a dead cert.
The ruffin cly the nab of Stephen Hand as give me the jady coppaleen.

He strike a telegramboy paddock wire big bug Bass to the depot.

Shove him a joey and grahamise.

\lenehan:
Mare on form hot order.

Guinea to a goosegog.

Tell a cram,
that.

\lenehan:
Gospeltrue.

Criminal diversion?

I think that yes.

Sure thing.

Land him in chokeechokee if the harman beck copped the game.

Madden back Madden's a maddening back.

\lynch:
O lust our refuge and our strength.

\madden:
Decamping.
\gab{1520}

\bannon:
Must you go?

\madden:
Off to mammy.

\dixon:
Stand by.

\madden:
Hide my blushes someone.

\stage{Exit Madden.
Enter Bantam Lyons}

\Bloom:
All in if he spots me.

\punch:
Come ahome,
our Bantam.

\dixon:
Horryvar,
mong vioo.

\crotthers:
Dinna forget the cowslips for hersel.

Cornfide.

Wha gev ye thon colt?

Pal to pal.

Jannock.

Of John Thomas,
her spouse.

No fake,
old man Leo.

S'elp me,
honest injun.

Shiver my timbers if I had.

There's a great big holy friar.

Vyfor you no me tell?

Vel,
I ses,
if that aint a sheeny nachez,
vel,
I vil get misha mishinnah.

Through yerd our lord,
Amen.

% \histstyle{Coda {\#}6}

\stephen:
You move a motion?

\lenehan:
Steve boy,
you're going it some.

\dixon:
More bluggy drunkables?

\lynch:
Will immensely splendiferous stander permit one stooder of most extreme poverty
\gab{1530}
and one largesize grandacious thirst to terminate one expensive inaugurated libation?

\crotthers:
Give's a breather.

\punch:
Landlord,
landlord,
have you good wine,
staboo?

\crotthers:
Hoots,
mon,
a wee drap to pree.

\lenehan:
Cut and come again.

\barman:
Right.

Boniface!

\stephen:
Absinthe the lot.
\latin{Nos omnes biberimus viridum toxicum diabolus capiat posterioria nostria}.

\barman:
Closingtime,
gents.

Eh?

\stephen:
Rome boose for the Bloom toff.

I hear you say onions?

\bannon:
Bloo?

\dixon:
Cadges ads.

\bannon:
Photo's papli,
by all that's gorgeous.
Play low,
pardner.
Slide.

\mulligan:
\french{Bonsoir la compagnie}.

\stage{Exit Mulligan and Bannon}

\stephen:
And snares of the poxfiend.

\bantam:
Where's the buck and Namby Amby?

\stephen:
Skunked?

\lenehan:
Leg bail.

\crotthers:
Aweel,
ye maun e'en gang yer gates.

\stephen:
Checkmate.
King to tower.

Kind Kristyann wil yu help yung man hoose frend
tuk bungellow kee tu find plais whear tu lay crown of his hed 2 night.
\gab{1540}

Crickey,
I'm about sprung.

Tarnally dog gone my shins if this beent the
bestest puttiest longbreak yet.

\punch:
Item,
curate,
couple of cookies for this child.

Cot's plood and prandypalls,
none!

Not a pite of sheeses?

\stephen:
Thrust syphilis down to hell and with him those other licensed spirits.

\barman:
Time,
gents!

\stephen:
Who wander through the world.

\punch:
Health all!

\dixon:
\french{À la vôtre}!

% \histstyle{Coda {\#}7}

\crotthers:
Golly,
whatten tunket's yon guy in the mackintosh?

\lenehan:
Dusty Rhodes.

\lynch:
Peep at his wearables.

\crotthers:
By mighty!

\stephen:
What's he got?

\dixon:
Jubilee mutton.
Bovril,
by James.

\lenehan:
Wants it real bad.

\crotthers:
D'ye ken bare socks?
Seedy cuss in the Richmond?

\dixon:
Rawthere!
Thought he had a deposit of lead in his penis.
Trumpery insanity.
\gab{1550}
Bartle the Bread we calls him.

\lynch:
That,
sir,
was once a prosperous cit.

\Bloom:
Man all tattered and torn that married a maiden all forlorn.
Slung her hook,
she did.
Here see lost love.
Walking Mackintosh of lonely canyon.

Tuck and turn in.

Schedule time.

Nix for the hornies.

Pardon?

Seen him today at a runefal?

Chum o' yourn passed in his checks?

Ludamassy!
Pore piccaninnies!

Thou'll no be telling me thot,
Pold veg.

Did ums blubble bigsplash crytears cos fren Padney
was took off in black bag?

Of all de darkies Massa Pat was verra best.

I never see the like since I was born.

\french{Tiens,
tiens},
but it is well sad,
that,
my faith,
yes.

O,
get,
rev on a gradient one in nine.

Live axle drives are souped.

Lay you two to one Jenatzy licks him ruddy well hollow.
\gab{1560}

Jappies?

High angle fire,
inyah!

Sunk by war specials.

Be worse for him,
says he,
nor any Rooshian.

\barman:
Time all.
There's eleven of them.
Get ye gone.

\lenehan:
Forward,
woozy wobblers!

\lynch:
Night.

\punch:
Night.

\dixon:
May Allah the Excellent One your soul this night ever
tremendously conserve.

% \histstyle{Coda {\#}8}

\lenehan:
Your attention!

\crotthers:
We're nae tha fou.

\lynch:
The Leith police dismisseth us.

\lenehan:
The least tholice.

\punch:
Ware hawks for the chap puking.
Unwell in his abominable regions.

\stephen:
Yooka.

Night.

\crotthers:
Mona,
my true love.

\lynch:
Yook.

\crotthers:
Mona,
my own love.

Ook.

% \histstyle{Coda {\#}9}

\lynch:
Hark!
Shut your obstropolos.

\N2:
Pflaap!

\N3:
Pflaap!

\lenehan:
Blaze on.
There she goes.
Brigade!

\punch:
Bout ship.
\gab{1570}
Mount street way.
Cut up!

\N1:
Pflaap!

\lynch:
Tally ho.

You not come?

Run,
skelter,
race.

\N2:
Pflaaaap!

% \histstyle{Coda {\#}10}

\stephen:
Lynch!
Hey?
Sign on long o' me.

\lynch:
Denzille lane this way.
Change here for Bawdyhouse.

\stephen:
We two,
she said,
will seek the kips where shady Mary is.

\lynch:
Righto,
any old time.

\stephen:
\latin{Laetabuntur in cubilibus suis}.

\lynch:
You coming long?

\stephen:
Who the sooty hell's the johnny in the black duds?

\lynch:
Hush!

\dowie:
Sinned against the light and even now that day is at hand
when he shall come to judge the world by fire.

\N3:
Pflaap!

\stephen:
\latin{Ut implerentur scripturae}.
Strike up a ballad.

\N1:
Then outspake medical Dick to his comrade medical Davy.

\bystander:
Christicle,
who's this excrement yellow gospeller on the Merrion hall?

\dowie:
Elijah is coming!
\gab{1580}
Washed in the blood of the Lamb.

\dowie:
Come on you winefizzling,
ginsizzling,
booseguzzling existences!
Come on,
you dog-gone,
bullnecked,
beetlebrowed,
hogjowled,
peanutbrained,
weaseleyed fourflushers,
false alarms and excess baggage!
Come on,
you triple extract of infamy!

\dowie:
Alexander J Christ Dowie,
that's my name,
that's
yanked to glory most half this planet from Frisco beach to Vladivostok.
The Deity aint no nickel dime bumshow.
I put it to you that He's on the
square and a corking fine business proposition.
He's the grandest thing
yet and don't you forget it.
Shout salvation in King Jesus.
You'll need
to rise precious early you sinner there,
if you want to diddle the
Almighty God.

\N2:
Pflaaaap!

\stephen:
Not half.

\dowie:
He's got a cough·mixture with a punch in it for you,
\gab{1590}
my friend,
in his back pocket.
Just you try it on.
