\N1:
\gab{534}
Behind him he heard Buck Mulligan club with his heavy bathtowel
the leader shoots of ferns or grasses.

\Mulligan:
Down, sir! How dare you, sir!

\N1:
Haines asked:

\Haines:
Do you pay rent for this tower?

\Mulligan:
Twelve quid,
\sout{Buck Mulligan said}.

\Stephen:
To the secretary of state for war,
\gab{540}

\N2:
Stephen added over his shoulder.

\N1:
They halted while Haines surveyed the tower and said at last:

\Haines:
Rather bleak in wintertime,
I should say.
Martello you call it?

\Mulligan:
Billy Pitt had them built,
\sout{Buck Mulligan said},
when the French were on the sea.
But ours is the \greek{omphalos}.

\Haines:
What is your idea of Hamlet?

\N1:
Haines asked Stephen.

\Mulligan:
No, no,

\N1:
Buck Mulligan shouted in pain.

\Mulligan:
I'm not equal to Thomas Aquinas
and the fifty-five reasons he has made out to prop it up.
Wait till I have a few pints in me first.

\N1:
He turned to Stephen,
saying,
as he pulled down neatly the peaks of his primrose waistcoat:
\gab{550}

\Mulligan:
You couldn't manage it under three pints,
Kinch,
could you?

\Stephen:
It has waited so long,
\sout{Stephen said listlessly},
it can wait longer.

\Haines:
You pique my curiosity,
\sout{Haines said amiably}.
Is it some paradox?

\Mulligan:
Pooh!
\sout{Buck Mulligan said}.
We have grown out of Wilde and paradoxes.
It's quite simple.
He proves by algebra that Hamlet's grandson is Shakespeare's grandfather
and that he himself is the ghost of his own father.

\Haines:
What?

\N1:
Haines said,
beginning to point at Stephen.

\Haines:
He himself?


\N1:
Buck Mulligan slung his towel stolewise round his neck
and, bending in loose laughter,
\gab{560}
said to Stephen's ear:

\Mulligan:
O, shade of Kinch the elder!
Japhet in search of a father!

\Stephen:
We're always tired in the morning,
\sout{Stephen said to Haines}.
And it is rather long to tell.

\N1:
Buck Mulligan,
walking forward again,
raised his hands.

\Mulligan:
The sacred pint alone can unbind the tongue of Dedalus,
\sout{he said}.

\Haines:
I mean to say,

\N1:
Haines explained to Stephen as they followed,

\Haines:
this tower and these cliffs here
remind me somehow of Elsinore.
\emph{That beetles o'er his base into the sea}, isn't it?


\N1:
Buck Mulligan turned suddenly,
for an instant
towards Stephen but did not speak.
\gab{570}

\N2:
In the bright silent instant
Stephen saw his own image
in cheap dusty mourning
between their gay attires.

\Haines:
It's a wonderful tale,

\N1:
Haines said,
bringing them to halt again.

\N2:
Eyes,
pale as the sea |
the wind had freshened,
paler, firm and prudent.
The seas' ruler,
he gazed southward over the bay,
empty save for the smokeplume of the mailboat
vague on the bright skyline
and a sail tacking by the Muglins.

\Haines:
I read a theological interpretation of it somewhere,
\sout{he said bemused}.
The Father and the Son idea.
The Son striving to be atoned with the Father.

\N1:
Buck Mulligan at once put on a blithe broadly smiling face.
He looked at them,
\gab{580}
his wellshaped mouth open happily,
his eyes,
from which he had suddenly withdrawn all shrewd sense,
blinking with mad gaiety.
He moved a doll's head to and fro,
the brims of his Panama hat quivering,
and began to chant in a quiet happy foolish voice:

\Mulligan:
\begin{verse}
\song{%
    I'm the queerest young fellow that ever you heard. \\
    My mother's a jew, my father's a bird. \\
    With Joseph the joiner i cannot agree. \\
    So here's to disciples and Calvary.}
\end{verse}

\N1:
He held up a forefinger of warning.

\begin{verse}
\Mulligan:
\song{%
    If anyone thinks that I amn't divine \\
    He'll get no free drinks when I'm making the wine
\gab{590} \\
    But have to drink water and wish it were plain \\
    That I make when the wine becomes water again.}
\end{verse}

\N1:
He tugged swiftly at Stephen's ashplant in farewell
and, running forward to a brow of the cliff,
fluttered his hands at his sides
like fins or wings of one about to rise in the air,
and chanted:

\Mulligan:
\begin{verse}
\song{%
    Goodbye, now, goodbye! Write down all I said \\
    And tell Tom, Dick and Harry I rose from the dead. \\
    What's bred in the bone cannot fail me to fly \\
    And Olivet's breezy … Goodbye, now, goodbye!}
\end{verse}

\N1:
He capered before them down towards the forty-foot hole,
\gab{600}
fluttering his winglike hands,
leaping nimbly,
Mercury's hat quivering in the fresh wind
that bore back to them his brief birdsweet cries.


\N1:
Haines,
who had been laughing guardedly,
walked on beside Stephen
and said:

\Haines:
We oughtn't to laugh,
I suppose.
He's rather blasphemous.
I'm not a believer myself, that is to say.
Still his gaiety takes the harm out of it somehow, doesn't it?
What did he call it?
Joseph the Joiner?

\Stephen:
The ballad of joking Jesus,
\sout{Stephen answered}.

\Haines:
O,
\sout{Haines said},
you have heard it before?

\Stephen:
Three times a day, after meals,
\gab{610}
\sout{Stephen said drily}.

\Haines:
You're not a believer, are you?
\sout{Haines asked}.
I mean, a believer in the narrow sense of the word.
Creation from nothing and miracles and a personal God.

\Stephen:
There's only one sense of the word,
it seems to me,
\sout{Stephen said}.

\N1:
Haines stopped to take out a smooth silver case
in which twinkled a green stone.
He sprang it open with his thumb and offered it.

\Stephen:
Thank you,

\N2:
Stephen said,
taking a cigarette.

\N1:
Haines helped himself and snapped the case to.
He put it back in his sidepocket
and took from his waistcoatpocket
a nickel tinderbox,
sprang it open too,
and,
having lit his cigarette,
\gab{620}
held the flaming spunk towards Stephen
in the shell of his hands.

\Haines:
Yes, of course,

\N1:
he said, as they went on again.

\Haines:
Either you believe or you don't, isn't it?
Personally I couldn't stomach that idea of a personal God.
You don't stand for that, I suppose?

\Stephen:
You behold in me,

\N2:
Stephen said with grim displeasure,

\Stephen:
a horrible example of free thought.

\N2:
He walked on,
waiting to be spoken to,
trailing his ashplant by his side.
Its ferrule followed lightly on the path,
squealing at his heels.

\StephenInt:
My familiar, after me,
calling, Steeeeeeeeeeeephen!
A wavering line along the path.
They will walk on it tonight,
coming here in the dark.
\gab{630}
He wants that key.
It is mine.
I paid the rent.
Now I eat his salt bread.
Give him the key too.
All.
He will ask for it.
That was in his eyes.

\Haines:
After all,
\sout{Haines began} …

\N2:
Stephen turned
and saw that the cold gaze which had measured him
was not all unkind.

\Haines:
After all,
I should think you are able to free yourself.
You are your own master,
it seems to me.

\Stephen:
I am a servant of two masters,
\sout{Stephen said},
an English and an Italian.

\Haines:
Italian?
\sout{Haines said}.

\StephenInt:
A crazy queen, old and jealous.
\gab{640}
Kneel down before me.

\Stephen:
And a third,
\sout{Stephen said},
there is who wants me for odd jobs.

\Haines:
Italian?
\sout{Haines said again}.
What do you mean?

\Stephen:
The imperial British state,

\N2:
Stephen answered, his colour rising,

\Stephen:
and the holy Roman catholic and apostolic church.

\N1:
Haines detached from his underlip some fibres of tobacco
before he spoke.

\Haines:
I can quite understand that,
\sout{he said calmly}.
An Irishman must think like that, I daresay.
We feel in England that we have treated you rather unfairly.
It seems history is to blame.

\N2:
The proud potent titles clanged over Stephen's memory
\gab{650}
the triumph of their brazen bells:
\latin{et unam sanctam catholicam et apostolicam ecclesiam}\/:
the slow growth and change of rite and dogma
like his own rare thoughts, a chemistry of stars.
Symbol of the apostles in the mass for pope Marcellus,
the voices blended, singing alone loud in affirmation:
and behind their chant |
the vigilant angel of the church militant
disarmed and menaced her heresiarchs.
A horde of heresies fleeing with mitres awry:
Photius and the brood of mockers of whom Mulligan was one,
and Arius, warring his life long
upon the consubstantiality of the Son with the Father,
and Valentine, spurning Christ's terrene body,
and the subtle African heresiarch Sabellius
who held that the Father was Himself His own Son.
\gab{660}
Words Mulligan had spoken a moment since in mockery to the stranger.
Idle mockery.
The void awaits surely all them that weave the wind:
a menace, a disarming
and a worsting from those embattled angels of the church,
Michael's host, who defend her ever
in the hour of conflict with their lances and their shields.

\StephenInt:
Hear, hear!
Prolonged applause.
\french{Zut!
Nom de Dieu!}

\Haines:
Of course I'm a Britisher,

\N1:
Haines's voice said,

\Haines:
and I feel as one.
I don't want to see my country fall into the hands of German jews either.
That's our national problem, I'm afraid, just now.

\N1:
Two men stood at the verge of the cliff, watching:
businessman, boatman.
\gab{670}

\Businessman:
She's making for Bullock harbour.

\N1:
The boatman nodded towards the north of the bay with some disdain.

\Boatman:
There's five fathoms out there,
\sout{he said}.
It'll be swept up that way when the tide comes in about one.
It's nine days today.

\StephenInt:
The man that was drowned.
A sail veering about the blank bay |
waiting for a swollen bundle to bob up,
roll over to the sun a puffy face, saltwhite.
Here I am.

\N1:
They followed the winding path down to the creek.
Buck Mulligan stood on a stone,
in shirtsleeves, his unclipped tie rippling over his shoulder.
A young man clinging to a spur of rock near him,
\gab{680}
moved slowly frogwise his green legs in the deep jelly of the water.

\YoungMan:
Is the brother with you, Malachi?

\Mulligan:
Down in Westmeath.
With the Bannons.

\YoungMan:
Still there?
I got a card from Bannon.
Says he found a sweet young thing down there.
Photo girl he calls her.

\Mulligan:
Snapshot, eh?
Brief exposure.

\N1:
Buck Mulligan sat down to unlace his boots.
An elderly man shot up near the spur of rock |
a blowing red face.
He scrambled up by the stones,
water glistening on his pate and on its garland of grey hair,
water rilling over his chest and paunch
\gab{690}
and spilling jets out of his black sagging loincloth.

\N1:
Buck Mulligan made way for him to scramble past
and, glancing at Haines and Stephen,
crossed himself piously with his thumbnail at brow and lips and breastbone.

\YoungMan:
Seymour's back in town,

\N1:
the young man said,
grasping again his spur of rock.

\YoungMan:
Chucked medicine and going in for the army.

\Mulligan:
Ah, go to God!
\sout{Buck Mulligan said}.

\YoungMan:
Going over next week to stew.
You know that red Carlisle girl, Lily?

\Mulligan:
Yes.

\YoungMan:
Spooning with him last night on the pier.
\gab{700}
The father is rotto with money.

\Mulligan:
Is she up the pole?

\YoungMan:
Better ask Seymour that.

\Mulligan:
Seymour a bleeding officer!
\sout{Buck Mulligan said}.

\N1:
He nodded to himself as he drew off his trousers
and stood up, saying tritely:

\Mulligan:
Redheaded women buck like goats.

\N1:
He broke off in alarm,
feeling his side under his flapping shirt.

\Mulligan:
My twelfth rib is gone,
\sout{he cried}.
I'm the \german{Űbermensch}.
Toothless Kinch and I, the supermen.

\N1:
He struggled out of his shirt
\gab{710}
and flung it behind him to where his clothes lay.

\YoungMan:
Are you going in here, Malachi?

\Mulligan:
Yes. Make room in the bed.

\N1:
The young man shoved himself backward through the water
and reached the middle of the creek in two long clean strokes.
Haines sat down on a stone, smoking.

\Mulligan:
Are you not coming in?
\sout{Buck Mulligan asked}.

\Haines:
Later on,
\sout{Haines said}.
Not on my breakfast.

\N2:
Stephen turned away.

\Stephen:
I'm going, Mulligan,
\gab{720}
\sout{he said}.

\Mulligan:
Give us that key, Kinch,
\sout{Buck Mulligan said},
to keep my chemise flat.

\N2:
Stephen handed him the key.

\N1:
Buck Mulligan laid it across his heaped clothes.

\Mulligan:
And twopence,
\sout{he said},
for a pint.
Throw it there.

\N2:
Stephen threw two pennies on the soft heap.

\N1:
Dressing, undressing.
Buck Mulligan erect,
with joined hands before him,
said solemnly:

\Mulligan:
He who stealeth from the poor lendeth to the Lord.
Thus spake Zarathustra.

\N1:
His plump body plunged.

\Haines:
We'll see you again,

\N1:
Haines said,
\gab{730}
turning as Stephen walked up the path and smiling at wild Irish.

\StephenInt:
Horn of a bull,
hoof of a horse,
smile of a Saxon.

\Mulligan:
The Ship,
\sout{Buck Mulligan cried}.
Half twelve.

\Stephen:
Good,
\sout{Stephen said}.

\N2:
He walked along the upwardcurving path.

\StephenInt:
\begin{verse}
\song{%
    Liliata rutilantium. \\
    Turma circumdet. \\
    Iubilantium te virginum.}
\end{verse}

\N2:
The priest's grey nimbus in a niche where he dressed discreetly.

\StephenInt:
I will not sleep here tonight.
\gab{740}
Home also |
I cannot go.

\N2:
A voice, sweettoned and sustained,
called to him from the sea.
Turning the curve he waved his hand.
It called again.
A sleek brown head, a seal's,
far out on the water, round.

\StephenInt:
Usurper.
