\N1:
\gab{313}
In the gloomy domed livingroom of the tower |
Buck Mulligan's gowned form moved briskly to and fro
about the hearth,
hiding and revealing its yellow glow.
Two shafts of soft daylight
fell across the flagged floor from the high barbacans:
and at the meeting of their rays |
a cloud of coalsmoke and fumes of fried grease floated,
turning.

\Mulligan:
We'll be choked,
\sout{Buck Mulligan said}.
Haines, open that door, will you?

\N2:
Stephen laid the shavingbowl on the locker.

\N1:
A tall figure rose from the hammock where it had been sitting,
\gab{320}
went to the doorway
and pulled open the inner doors.

\Haines:
Have you the key?

\N1:
a voice asked.

\Mulligan:
Dedalus has it,
\sout{Buck Mulligan said}.
Janey Mack, I'm choked!

\N1:
He howled, without looking up from the fire:

\Mulligan:
Kinch!

\Stephen:
It's in the lock,

\N2:
Stephen said,
coming forward.

\N1:
The key scraped round harshly twice
and, when the heavy door had been set ajar,
welcome light and bright air entered.
Haines stood at the doorway, looking out.

\N2:
Stephen haled his upended valise to the table and sat down to wait.
\gab{330}

\N1:
Buck Mulligan tossed the fry on to the dish beside him.
Then he carried the dish and a large teapot over to the table,
set them down heavily
and sighed with relief.

\Mulligan:
I'm melting,
\sout{he said},
as the candle remarked when …
But, hush!
Not a word more on that subject!
Kinch, wake up!
Bread, butter, honey.
Haines, come in.
The grub is ready.
Bless us, O Lord, and these thy gifts.
Where's the sugar?
O, jay, there's no milk.

\N2:
Stephen fetched the loaf and the pot of honey and the buttercooler
from the locker.

\N1:
Buck Mulligan sat down in a sudden pet.

\Mulligan:
What sort of a kip is this?
\sout{he said}.
I told her to come after eight.

\Stephen:
We can drink it black,

\N2:
Stephen said thirstily.
\gab{340}

\Stephen:
There's a lemon in the locker.

\Mulligan:
O, damn you and your Paris fads!
\sout{Buck Mulligan said}.
I want Sandycove milk.

\N1:
Haines came in from the doorway and said quietly:

\Haines:
That woman is coming up with the milk.

\Mulligan:
The blessings of God on you!

\N1:
Buck Mulligan cried,
jumping up from his chair.

\Mulligan:
Sit down.
Pour out the tea there.
The sugar is in the bag.
Here, I can't go fumbling at the damned eggs.

\N1:
He hacked through the fry on the dish
and slapped it out on three plates,
\gab{350}
saying:

\Mulligan:
\latin{In nomine Patris et Filii et Spiritus Sancti.}

\N1:
Haines sat down to pour out the tea.

\Haines:
I'm giving you two lumps each,
\sout{he said}.
But, I say, Mulligan, you do make strong tea, don't you?

\N1:
Buck Mulligan,
hewing thick slices from the loaf,
said in an old woman's wheedling voice:

\Mulligan:
When I makes tea I makes tea, as old mother Grogan said.
And when I makes water I makes water.

\Haines:
By Jove, it is tea,
\sout{Haines said}.

\N1:
Buck Mulligan went on hewing and wheedling:
\gab{360}

\Mulligan:
\emph{So I do, Mrs Cahill}\/,
says she.
\emph{Begob, ma'am}\/,
says Mrs Cahill,
\emph{God send you don't make them in the one pot.}

\N1:
He lunged towards his messmates in turn |
a thick slice of bread, impaled on his knife.

\Mulligan:
That's folk,
\sout{he said very earnestly,}
for your book, Haines.
Five lines of text and ten pages of notes
about the folk and the fishgods of Dundrum.
Printed by the weird sisters in the year of the big wind.

\N1:
He turned to Stephen and asked in a fine puzzled voice,
lifting his brows:

\Mulligan:
Can you recall, brother,
\gab{370}
is mother Grogan's tea and water pot spoken of in the Mabinogion
or is it in the Upanishads?

\Stephen:
I doubt it,

\N2:
said Stephen gravely.

\Mulligan:
Do you now?

\N1:
Buck Mulligan said in the same tone.

\Mulligan:
Your reasons, pray?

\Stephen:
I fancy,

\N2:
Stephen said as he ate,

\Stephen:
it did not exist in or out of the Mabinogion.
Mother Grogan was, one imagines, a kinswoman of Mary Ann.

\N1:
Buck Mulligan's face smiled with delight.

\Mulligan:
Charming!

\N1:
he said in a finical sweet voice,
showing his white teeth and blinking his eyes pleasantly.

\Mulligan:
Do you think she was?
Quite charming!

\N1:
Then, suddenly overclouding all his features,
\gab{380}
he growled in a hoarsened rasping voice
as he hewed again vigorously at the loaf:

\Mulligan:
\begin{verse}
\song{%
    For old Mary Ann \\
    She doesn't care a damn. \\
    But, hising up her petticoats …}
\end{verse}

\N1:
He crammed his mouth with fry and munched and droned.

\N2:
The doorway was darkened by an entering form.

\OldWoman:
The milk, sir!

\Mulligan:
Come in, ma'am,
\sout{Mulligan said}.
Kinch, get the jug.

\N2:
An old woman came forward and stood by Stephen's elbow.

\OldWoman:
That's a lovely morning, sir,
\sout{she said}.
\gab{390}
Glory be to God.

\Mulligan:
To whom?

\N1:
Mulligan said, glancing at her.

\Mulligan:
Ah, to be sure!

\N2:
Stephen reached back and took the milkjug from the locker.

\Mulligan:
The islanders,

\N1:
Mulligan said to Haines casually,

\Mulligan:
speak frequently of the collector of prepuces.

\OldWoman:
How much, sir?
\sout{asked the old woman}.

\Stephen:
A quart,
\sout{Stephen said}.

\N2:
He watched her pour into the measure
and thence into the jug

\StephenInt:
rich white milk,
not hers.
Old shrunken paps.

\N2:
She poured again a measureful and a tilly.

\StephenInt:
Old and secret she had entered from a morning world,
maybe a messenger.
\gab{400}

\N2:
She praised the goodness of the milk, pouring it out.

\StephenInt:
Crouching by a patient cow at daybreak in the lush field,
a witch on her toadstool,
her wrinkled fingers quick at the squirting dugs.
They lowed about her whom they knew,
dewsilky cattle.
Silk of the kine and poor old woman,
names given her in old times.
A wandering crone,
lowly form of an immortal
serving her conqueror and her gay betrayer,
their common cuckquean,
a messenger from the secret morning.

\N2:
To serve or to upbraid,
whether he could not tell:
but scorned to beg her favour.

\Mulligan:
It is indeed, ma'am,

\N1:
Buck Mulligan said,
pouring milk into their cups.

\OldWoman:
Taste it, sir,
\sout{she said}.

\N1:
He drank at her bidding.
\gab{410}

\Mulligan:
If we could live on good food like that,

\N1:
he said to her somewhat loudly,

\Mulligan:
we wouldn't have the country full of rotten teeth and rotten guts.
Living in a bogswamp, eating cheap food
and the streets paved with dust, horsedung and consumptives' spits.

\OldWoman:
Are you a medical student, sir?
\sout{the old woman asked}.

\Mulligan:
I am, ma'am,
\sout{Buck Mulligan answered}.

\OldWoman:
Look at that now,
\sout{she said}.


\N2:
Stephen listened in scornful silence.

\StephenInt:
She bows her old head to a voice that speaks to her loudly,
her bonesetter, her medicineman:
me |
she slights.
To the voice that will shrive and oil for the grave |
\gab{420}
all there is of her
but her woman's unclean loins,
of man's flesh made not in God's likeness,
the serpent's prey.
And to the loud voice that now bids her be silent |
with wondering unsteady eyes.

\Stephen:
Do you understand what he says?
\sout{Stephen asked her}.

\OldWoman:
Is it French you are talking, sir?

\N2:
the old woman said to Haines.

\N1:
Haines spoke to her again a longer speech, confidently.

\Mulligan:
Irish,
\sout{Buck Mulligan said}.
Is there Gaelic on you?

\OldWoman:
I thought it was Irish,
\sout{she said},
by the sound of it.
Are you from the west, sir?

\Haines:
I am an Englishman,
\gab{430}
\sout{Haines answered}.

\Mulligan:
He's English,
\sout{Buck Mulligan said},
and he thinks we ought to speak Irish in Ireland.

\OldWoman:
Sure we ought to,
\sout{the old woman said},
and I'm ashamed I don't speak the language myself.
I'm told it's a grand language by them that knows.

\Mulligan:
Grand is no name for it,
\sout{said Buck Mulligan}.
Wonderful entirely.
Fill us out some more tea, Kinch.
Would you like a cup, ma'am?

\OldWoman:
No, thank you, sir,

\N2:
the old woman said,
slipping the ring of the milkcan on her forearm
and about to go.

\N1:
Haines said to her:

\Haines:
Have you your bill?
\gab{440}
We had better pay her, Mulligan, hadn't we?

\N2:
Stephen filled again the three cups.

\OldWoman:
Bill, sir?

\N2:
she said, halting.

\OldWoman:
Well, it's seven mornings a pint at twopence |
is seven twos |
is a shilling and twopence over |
and these three mornings a quart at fourpence |
is three quarts |
is a shilling.
That's a shilling and one and two |
is two and two, sir.

\N1:
Buck Mulligan sighed
and, having filled his mouth with a crust thickly buttered on both sides,
stretched forth his legs and began to search his trouser pockets.

\Haines:
Pay up and look pleasant,

\N1:
Haines said to him, smiling.

\N2:
Stephen filled a third cup,
\gab{450}
a spoonful of tea colouring faintly the thick rich milk.

\N1:
Buck Mulligan brought up a florin,
twisted it round in his fingers
and cried:

\Mulligan:
A miracle!

\N1:
He passed it along the table towards the old woman, saying:

\Mulligan:
Ask nothing more of me, sweet.
All I can give you I give.

\N2:
Stephen laid the coin in her uneager hand.

\Stephen:
We'll owe twopence,
\sout{he said}.

\OldWoman:
Time enough, sir,

\N2:
she said, taking the coin.

\OldWoman:
Time enough.
Good morning, sir.
\gab{460}

\N2:
She curtseyed and went out,
followed by Buck Mulligan's tender chant:

\begin{verse}
\Mulligan:
\song{%
    Heart of my heart, were it more, \\
    More would be laid at your feet.}
\end{verse}

\N1:
He turned to Stephen and said:

\Mulligan:
Seriously, Dedalus.
I'm stony.
Hurry out to your school kip
and bring us back some money.
Today the bards must drink and junket.
Ireland expects that every man
this day will do his duty.

\Haines:
That reminds me,

\N1:
Haines said, rising,

\Haines:
that I have to visit your national library today.
\gab{470}

\Mulligan:
Our swim first,
\sout{Buck Mulligan said}.

\N1:
He turned to Stephen and asked blandly:

\Mulligan:
Is this the day for your monthly wash, Kinch?

\N1:
Then he said to Haines:

\Mulligan:
The unclean bard makes a point of washing once a month.

\Stephen:
All Ireland is washed by the gulfstream,

\N2:
Stephen said as he let honey trickle over a slice of the loaf.

\N1:
Haines from the corner |
where he was knotting easily a scarf about the loose collar of his tennis shirt |
spoke:

\Haines:
I intend to make a collection of your sayings |
\gab{480}
if you will let me.

\StephenInt:
Speaking to me.
They wash and tub and scrub.
Agenbite of inwit.
Conscience.
Yet here's a spot.

\Haines:
That one about the cracked lookingglass of a servant
being the symbol of Irish art |
is deuced good.

\N1:
Buck Mulligan kicked Stephen's foot under the table
and said with warmth of tone:

\Mulligan:
Wait till you hear him on Hamlet, Haines.

\Haines:
Well, I mean it,

\N1:
Haines said, still speaking to Stephen.

\Haines:
I was just thinking of it when that poor old creature came in.

\Stephen:
Would I make any money by it?
\gab{490}
\sout{Stephen asked}.

\N1:
Haines laughed
and, as he took his soft grey hat from the holdfast of the hammock,
said:

\Haines:
I don't know, I'm sure.

\N1:
He strolled out to the doorway.
Buck Mulligan bent across to Stephen
and said with coarse vigour:

\Mulligan:
You put your hoof in it now.
What did you say that for?

\Stephen:
Well?
\sout{Stephen said}.
The problem is to get money.
From whom?
From the milkwoman or from him.
It's a toss up, I think.

\Mulligan:
I blow him out about you,
\sout{Buck Mulligan said},
and then you come along
with your lousy leer and your gloomy jesuit jibes.
\gab{500}

\Stephen:
I see little hope,
\sout{Stephen said},
from her or from him.

\N1:
Buck Mulligan sighed tragically and laid his hand on Stephen's arm.

\Mulligan:
From me, Kinch,
\sout{he said}.

\N1:
In a suddenly changed tone he added:

\Mulligan:
To tell you the God's truth |
I think you're right.
Damn all else they are good for.
Why don't you play them as I do?
To hell with them all.
Let us get out of the kip.

\N1:
He stood up,
gravely ungirdled and disrobed himself of his gown,
saying resignedly:

\Mulligan:
Mulligan is stripped of his garments.
\gab{510}

\N1:
He emptied his pockets on to the table.

\Mulligan:
There's your snotrag,
\sout{he said}.

\N1:
And putting on his stiff collar and rebellious tie
he spoke to them,
chiding them,
and to his dangling watchchain.
His hands plunged and rummaged in his trunk
while he called for a clean handkerchief.

\Mulligan:
God, we'll simply have to dress the character.
I want puce gloves and green boots.
Contradiction.
Do I contradict myself?
Very well then, I contradict myself.
Mercurial Malachi.

\N1:
A limp black missile flew out of his talking hands.

\Mulligan:
And there's your Latin quarter hat,
\sout{he said}.

\N2:
Stephen picked it up and put it on.
\gab{520}

\N1:
Haines called to them from the doorway:

\Haines:
Are you coming, you fellows?

\Mulligan:
I'm ready,

\N1:
Buck Mulligan answered,
going towards the door.

\Mulligan:
Come out, Kinch.
You have eaten all we left, I suppose.

\N1:
Resigned |
he passed out with grave words and gait,
saying,
wellnigh with sorrow:

\Mulligan:
And going forth |
he met Butterly.

\N2:
Stephen,
taking his ashplant from its leaningplace,
followed them out
and, as they went down the ladder,
pulled to the slow iron door and locked it.
\gab{530}
He put the huge key in his inner pocket.

\N1:
At the foot of the ladder
Buck Mulligan asked:

\Mulligan:
Did you bring the key?

\Stephen:
I have it,

\N2:
Stephen said,
preceding them.

\N2:
He walked on.
