\N1:
Stately, plump Buck Mulligan came from the stairhead,
bearing a bowl of lather
on which a mirror and a razor lay crossed.
A yellow dressinggown, ungirdled,
was sustained gently behind him by the mild morning air.
He held the bowl aloft and intoned:

\Mulligan:
\latin{Introibo ad altare Dei}.

\N1:
Halted, he peered down the dark winding stairs
and called out coarsely:

\Mulligan:
Come up, Kinch!
Come up, you fearful jesuit!

\N1:
Solemnly he came forward and mounted the round gunrest.
He faced about and blessed gravely thrice the tower,
\gab{10}
the surrounding land and the awaking mountains.
Then, catching sight of Stephen Dedalus,
he bent towards him and made rapid crosses in the air,
gurgling in his throat and shaking his head.

\N2:
Stephen Dedalus, displeased and sleepy,
leaned his arms on the top of the staircase
and looked coldly at the shaking gurgling face that blessed him,
equine in its length,
and at the light untonsured hair,
grained and hued like pale oak.

\N1:
Buck Mulligan peeped an instant under the mirror
and then covered the bowl smartly.

\Mulligan:
Back to barracks!
\sout{he said sternly}.

\N1:
He added in a preacher's tone:
\gab{20}

\Mulligan:
For this, O dearly beloved,
is the genuine Christine:
body and soul and blood and ouns.
Slow music, please.
Shut your eyes, gents.
One moment.
A little trouble about those white corpuscles.
Silence, all.

\N1:
He peered sideways up and gave a long slow whistle of call,
then paused awhile in rapt attention,
his even white teeth glistening here and there with gold points.

\StephenInt:
Chrysostomos.

\N1:
Two strong shrill whistles answered through the calm.

\Mulligan:
Thanks, old chap,
\sout{he cried briskly}.
That will do nicely.
Switch off the current, will you?

\N1:
He skipped off the gunrest
\gab{30}
and looked gravely at his watcher,
gathering about his legs the loose folds of his gown.
The plump shadowed face and sullen oval jowl
recalled a prelate,
patron of arts in the middle ages.
A pleasant smile broke quietly over his lips.

\Mulligan:
The mockery of it!
\sout{he said gaily}.
Your absurd name, an ancient Greek!

\N1:
He pointed his finger in friendly jest
and went over to the parapet,
laughing to himself.

\N2:
Stephen Dedalus stepped up,
followed him wearily halfway
and sat down on the edge of the gunrest,
watching him still
as he propped his mirror on the parapet,
dipped the brush in the bowl
and lathered cheeks and neck.

\N1:
Buck Mulligan's gay voice went on.
\gab{40}

\Mulligan:
My name is absurd too:
Malachi Mulligan, two dactyls.
But it has a Hellenic ring, hasn't it?
Tripping and sunny like the buck himself.
We must go to Athens.
Will you come
if I can get the aunt to fork out twenty quid?

\N1:
He laid the brush aside
and, laughing with delight, cried:

\Mulligan:
Will he come?
The jejune jesuit!

\N1:
Ceasing, he began to shave with care.

\Stephen:
Tell me, Mulligan,
\sout{Stephen said quietly}.

\Mulligan:
Yes, my love?

\Stephen:
How long is Haines going to stay in this tower?

\N1:
Buck Mulligan showed a shaven cheek over his right shoulder.
\gab{50}

\Mulligan:
God, isn't he dreadful?
\sout{he said frankly}.
A ponderous Saxon.
He thinks you're not a gentleman.
God, these bloody English!
Bursting with money and indigestion.
Because he comes from Oxford.
You know, Dedalus, you have the real Oxford manner.
He can't make you out.
O, my name for you is the best:
Kinch, the knife-blade.

\N1:
He shaved warily over his chin.

\Stephen:
He was raving all night about a black panther,
\sout{Stephen said}.
Where is his guncase?

\Mulligan:
A woful lunatic!
\sout{Mulligan said}.
Were you in a funk?

\Stephen:
I was,

\N2:
Stephen said with energy and growing fear.
\gab{60}

\Stephen:
Out here in the dark
with a man I don't know |
raving and moaning to himself
about shooting a black panther.
You saved men from drowning.
I'm not a hero, however.
If he stays on here
I am off.

\N1:
Buck Mulligan frowned at the lather on his razorblade.
He hopped down from his perch
and began to search his trouser pockets hastily.

\Mulligan:
Scutter!
\sout{he cried thickly}.

\N1:
He came over to the gunrest
and, thrusting a hand into Stephen's upper pocket,
said:

\Mulligan:
Lend us a loan of your noserag to wipe my razor.

\N2:
Stephen suffered him to pull out
\gab{70}
and hold up on show by its corner
a dirty crumpled handkerchief.

\N1:
Buck Mulligan wiped the razorblade neatly.
Then, gazing over the handkerchief, he said:

\Mulligan:
The bard's noserag!
A new art colour for our Irish poets:
snotgreen.
You can almost taste it, can't you?

\N1:
He mounted to the parapet again
and gazed out over Dublin bay,
his fair oakpale hair stirring slightly.

\Mulligan:
God!
\sout{he said quietly}.
Isn't the sea what Algy calls it:
a great sweet mother?
The snotgreen sea.
The scrotumtightening sea.
\greek{Epi oinopa ponton}.
Ah, Dedalus, the Greeks!
I must teach you.
You must read them in the original.
\greek{Thalatta! Thalatta!}
\gab{80}
She is our great sweet mother.
Come and look.

\N2:
Stephen stood up and went over to the parapet.
Leaning on it
he looked down on the water
and on the mailboat clearing the harbourmouth of Kingstown.

\Mulligan:
Our mighty mother!
\sout{Buck Mulligan said}.

\N1:
He turned abruptly his grey searching eyes
from the sea to Stephen's face.

\Mulligan:
The aunt thinks you killed your mother,
\sout{he said}.
That's why she won't let me have anything to do with you.

\Stephen:
Someone killed her,
\gab{90}
\sout{Stephen said gloomily}.

\Mulligan:
You could have knelt down,
damn it, Kinch,
when your dying mother asked you,
\sout{Buck Mulligan said}.
I'm hyperborean as much as you.
But to think of your mother
begging you with her last breath
to kneel down and pray for her.
And you refused.
There is something sinister in you …

\N1:
He broke off and lathered again lightly his farther cheek.
A tolerant smile curled his lips.

\Mulligan:
But a lovely mummer!

\N1:
he murmured to himself.

\Mulligan:
Kinch, the loveliest mummer of them all!

\N1:
He shaved evenly and with care,
in silence,
seriously.

\N2:
Stephen, an elbow rested on the jagged granite
\gab{100}
leaned his palm against his brow
and gazed at the fraying edge of his shiny black coat-sleeve.
Pain, that was not yet the pain of love,
fretted his heart.
Silently, in a dream
she had come to him after her death,
her wasted body within its loose brown graveclothes
giving off an odour of wax and rosewood,
her breath, that had bent upon him,
mute, reproachful,
a faint odour of wetted ashes.
Across the threadbare cuffedge
he saw the sea hailed as a great sweet mother
by the wellfed voice beside him.
The ring of bay and skyline held a dull green mass of liquid.
A bowl of white china had stood beside her deathbed
holding the green sluggish bile
which she had torn up from her rotting liver
\gab{110}
by fits of loud groaning vomiting.

\N1:
Buck Mulligan wiped again his razorblade.

\Mulligan:
Ah, poor dogsbody!
\sout{he said in a kind voice}.
I must give you a shirt and a few noserags.
How are the secondhand breeks?

\Stephen:
They fit well enough,
\sout{Stephen answered}.

\N1:
Buck Mulligan attacked the hollow beneath his underlip.

\Mulligan:
The mockery of it,
\sout{he said contentedly}.
Secondleg they should be.
God knows what poxy bowsy left them off.
I have a lovely pair with a hair stripe, grey.
You'll look spiffing in them.
I'm not joking, Kinch.
You look damn well when you're dressed.

\Stephen:
Thanks,
\sout{Stephen said}.
\gab{120}
I can't wear them if they are grey.

\Mulligan:
He can't wear them,

\N1:
Buck Mulligan told his face in the mirror.

\Mulligan:
Etiquette is etiquette.
He kills his mother but he can't wear grey trousers.

\N1:
He folded his razor neatly
and with stroking palps of fingers
felt the smooth skin.

\N2:
Stephen turned his gaze from the sea
and to the plump face
with its smokeblue mobile eyes.

\Mulligan:
That fellow I was with in the Ship last night,
\sout{said Buck Mulligan},
says you have g.p.i.
He's up in Dottyville with Connolly Norman.
General paralysis of the insane!

\N1:
He swept the mirror a half circle in the air
\gab{130}
to flash the tidings abroad
in sunlight now radiant on the sea.
His curling shaven lips laughed
and the edges of his white glittering teeth.
Laughter seized all his strong wellknit trunk.

\Mulligan:
Look at yourself,
\sout{he said},
you dreadful bard!

\N2:
Stephen bent forward
and peered at the mirror held out to him,
cleft by a crooked crack.

\StephenInt:
Hair on end.
As he and others see me.
Who chose this face for me?
This dogsbody to rid of vermin.
It asks me too.

\Mulligan:
I pinched it out of the skivvy's room,
\sout{Buck Mulligan said}.
It does her all right.
The aunt always keeps plainlooking servants for Malachi.
Lead him not into temptation.
\gab{140}
And her name is Ursula.

\N1:
Laughing again,
he brought the mirror away from Stephen's peering eyes.

\Mulligan:
The rage of Caliban
at not seeing his face in a mirror,
\sout{he said}.
If Wilde were only alive to see you!

\N2:
Drawing back and pointing,
Stephen said with bitterness:

\Stephen:
It is a symbol of Irish art.
The cracked looking-glass of a servant.

\N1:
Buck Mulligan suddenly linked his arm in Stephen's
and walked with him round the tower,
his razor and mirror clacking
in the pocket where he had thrust them.

\Mulligan:
It's not fair to tease you like that,
\gab{150}
Kinch, is it?
\sout{he said kindly}.
God knows you have more spirit than any of them.

\StephenInt:
Parried again.
He fears the lancet of my art as I fear that of his.
The cold steelpen.

\Mulligan:
Cracked lookingglass of a servant!
Tell that to the oxy chap downstairs and touch him for a guinea.
He's stinking with money and thinks you're not a gentleman.
His old fellow made his tin
by selling jalap to Zulus or some bloody swindle or other.
God, Kinch, if you and I could only work together
we might do something for the island.
Hellenise it.

\StephenInt:
Cranly's arm.
His arm.

\Mulligan:
And to think of your having to beg from these swine.
\gab{160}
I'm the only one that knows what you are.
Why don't you trust me more?
What have you up your nose against me?
Is it Haines?
If he makes any noise here
I'll bring down Seymour
and we'll give him a ragging worse than they gave Clive Kempthorpe.

\StephenInt:
Young shouts of moneyed voices in Clive Kempthorpe's rooms.
Palefaces:
they hold their ribs with laughter,
one clasping another.
O, I shall expire!
Break the news to her gently, Aubrey!
I shall die!
With slit ribbons of his shirt whipping the air
he hops and hobbles round the table,
with trousers down at heels,
chased by Ades of Magdalen with the tailor's shears.
A scared calf's face gilded with marmalade.
\gab{170}
I don't want to be debagged!
Don't you play the giddy ox with me!

\StephenInt:
Shouts from the open window
startling evening in the quadrangle.
A deaf gardener,
aproned, masked with Matthew Arnold's face,
pushes his mower on the sombre lawn
watching narrowly the dancing motes of grasshalms.

\StephenInt:
To ourselves … new paganism … \greek{omphalos}.

\Stephen:
Let him stay,
\sout{Stephen said}.
There's nothing wrong with him except at night.

\Mulligan:
Then what is it?
\sout{Buck Mulligan asked impatiently}.
Cough it up.
I'm quite frank with you.
\gab{180}
What have you against me now?

\N2:
They halted,
looking towards the blunt cape of Bray Head
that lay on the water like the snout of a sleeping whale.
Stephen freed his arm quietly.

\Stephen:
Do you wish me to tell you?
\sout{he asked}.

\Mulligan:
Yes, what is it?
\sout{Buck Mulligan answered}.
I don't remember anything.

\N1:
He looked in Stephen's face as he spoke.
A light wind passed his brow,
fanning softly his fair uncombed hair
and stirring silver points of anxiety in his eyes.

\N2:
Stephen,
depressed by his own voice,
said:

\Stephen:
Do you remember the first day I went to your house
after my mother's death?
\gab{190}

\N1:
Buck Mulligan frowned quickly and said:

\Mulligan:
What?
Where?
I can't remember anything.
I remember only ideas and sensations.
Why?
What happened in the name of God?

\Stephen:
You were making tea,
\sout{Stephen said},
and went across the landing to get more hot water.
Your mother and some visitor came out of the drawingroom.
She asked you who was in your room.

\Mulligan:
Yes?
\sout{Buck Mulligan said}.
What did I say?
I forget.

\Stephen:
You said,
\sout{Stephen answered},
\emph{O, it's only Dedalus whose mother is beastly dead.}

\N1:
A flush
which made him seem younger and more engaging
\gab{200}
rose to Buck Mulligan's cheek.

\Mulligan:
Did I say that?
\sout{he asked}.
Well?
What harm is that?

\N1:
He shook his constraint from him nervously.

\Mulligan:
And what is death,
\sout{he asked},
your mother's or yours or my own?
You saw only your mother die.
I see them pop off every day in the Mater and Richmond
and cut up into tripes in the dissectingroom.
It's a beastly thing and nothing else.
It simply doesn't matter.
You wouldn't kneel down to pray for your mother on her deathbed
when she asked you.
Why?
Because you have the cursed jesuit strain in you,
only it's injected the wrong way.
To me it's all a mockery and beastly.
\gab{210}
Her cerebral lobes are not functioning.
She calls the doctor sir Peter Teazle
and picks buttercups off the quilt.
Humour her till it's over.
You crossed her last wish in death
and yet you sulk with me
because I don't whinge like some hired mute from Lalouette's.
Absurd!
I suppose I did say it.
I didn't mean to offend the memory of your mother.

\N1:
He had spoken himself into boldness.

\N2:
Stephen,
shielding the gaping wounds
which the words had left in his heart,
said very coldly:

\Stephen:
I am not thinking of the offence to my mother.

\Mulligan:
Of what then?
\sout{Buck Mulligan asked}.

\Stephen:
Of the offence to me,
\gab{220}
\sout{Stephen answered}.

\N1:
Buck Mulligan swung round on his heel.

\Mulligan:
O, an impossible person!
\sout{he exclaimed}.

\N1:
He walked off quickly round the parapet.
Stephen stood at his post,
gazing over the calm sea towards the headland.
Sea and headland now grew dim.
Pulses were beating in his eyes,
veiling their sight,
and he felt the fever of his cheeks.

\N1:
A voice within the tower called loudly:

\Haines:
Are you up there, Mulligan?

\Mulligan:
I'm coming,
\sout{Buck Mulligan answered}.

\N1:
He turned towards Stephen and said:
\gab{230}

\Mulligan:
Look at the sea.
What does it care about offences?
Chuck Loyola, Kinch,
and come on down.
The Sassenach wants his morning rashers.

\N1:
His head halted again for a moment at the top of the staircase,
level with the roof:

\Mulligan:
Don't mope over it all day,
\sout{he said}.
I'm inconsequent.
Give up the moody brooding.

\N1:
His head vanished
but the drone of his descending voice boomed out of the stairhead:

\Mulligan:
\begin{verse}
\song{%
   And no more turn aside and brood \\
   Upon love's bitter mystery
\gab{240} \\
   For Fergus rules the brazen cars.}
\end{verse}

\N2:
Woodshadows floated silently by
through the morning peace
from the stairhead seaward where he gazed.
Inshore and farther out
the mirror of water whitened,
spurned by lightshod hurrying feet.
White breast of the dim sea.
The twining stresses, two by two.
A hand plucking the harpstrings,
merging their twining chords.
Wavewhite wedded words
shimmering on the dim tide.

\N2:
A cloud began to cover the sun
slowly, wholly,
shadowing the bay in deeper green.
It lay beneath him, a bowl of bitter waters.

\StephenInt:
Fergus' song:
I sang it alone in the house,
\gab{250}
holding down the long dark chords.
Her door was open:
she wanted to hear my music.
Silent with awe and pity
I went to her bedside.
She was crying in her wretched bed.
For those words,
Stephen:
love's bitter mystery.

\StephenInt:
Where now?

\StephenInt:
Her secrets:
old featherfans,
tasselled dancecards,
powdered with musk,
a gaud of amber beads in her locked drawer.
A birdcage hung in the sunny
window of her house when she was a girl.
She heard old Royce sing in the pantomime of \Title{Turko the Terrible}
and laughed with others when he sang:

\begin{verse}
\song{%
    I am the boy
\gab{260} \\
    That can enjoy \\
    Invisibility.}
\end{verse}

\StephenInt:
Phantasmal mirth,
folded away:
muskperfumed.

\begin{verse}
\song{%
    And no more turn aside and brood.}
\end{verse}

\N2:
Folded away in the memory of nature with her toys.
Memories beset his brooding brain.

\StephenInt:
Her glass of water from the kitchen tap
when she had approached the sacrament.
A cored apple, filled with brown sugar,
roasting for her at the hob on a dark autumn evening.
Her shapely fingernails
reddened by the blood of squashed lice from the children's shirts.

\N2:
In a dream,
silently,
she had come to him,
\gab{270}
her wasted body within its loose graveclothes
giving off an odour of wax and rosewood,
her breath,
bent over him with mute secret words,
a faint odour of wetted ashes.

\StephenInt:
Her glazing eyes,
staring out of death,
to shake and bend my soul.
On me alone.
The ghostcandle to light her agony.
Ghostly light on the tortured face.
Her hoarse loud breath rattling in horror,
while all prayed on their knees.
Her eyes on me to strike me down.
\latin{Liliata rutilantium te confessorum turma circumdet:
iubilantium te virginum chorus excipiat.}

\StephenInt:
Ghoul! Chewer of corpses!

\StephenInt:
No, mother! Let me be and let me live.

\Mulligan:
Kinch ahoy!
\gab{280}

\N1:
Buck Mulligan's voice sang from within the tower.
It came nearer up the staircase,
calling again.

\N2:
Stephen, still trembling at his soul's cry,
heard warm running sunlight
and in the air behind him
friendly words.

\Mulligan:
Dedalus,
come down, like a good mosey.
Breakfast is ready.
Haines is apologising for waking us last night.
It's all right.

\Stephen:
I'm coming,

\N2:
Stephen said, turning.

\Mulligan:
Do, for Jesus' sake,
\sout{Buck Mulligan said}.
For my sake and for all our sakes.

\N1:
His head disappeared and reappeared.

\Mulligan:
I told him your symbol of Irish art.
\gab{290}
He says it's very clever.
Touch him for a quid, will you?
A guinea, I mean.

\Stephen:
I get paid this morning,
\sout{Stephen said}.

\Mulligan:
The school kip?
\sout{Buck Mulligan said}.
How much?
Four quid?
Lend us one.

\Stephen:
If you want it,
\sout{Stephen said}.

\Mulligan:
Four shining sovereigns,
\sout{Buck Mulligan cried with delight}.
We'll have a glorious drunk to astonish the druidy druids.
Four omnipotent sovereigns.

\N1:
He flung up his hands
and tramped down the stone stairs,
singing out of tune
with a Cockney accent:

\Mulligan:
\begin{verse}
\song{%
    O, won't we have a merry time,
\gab{300} \\
    Drinking whisky, beer and wine! \\
    On coronation, \\
    Coronation day! \\
    O, won't we have a merry time \\
    On coronation day!}
\end{verse}

\N2:
Warm sunshine merrying over the sea.
The nickel shavingbowl shone,
forgotten,
on the parapet.

\StephenInt:
Why should I bring it down?
Or leave it there all day,
forgotten friendship?

\N2:
He went over to it,
held it in his hands awhile,
feeling its coolness,
smelling the clammy slaver of the lather
\gab{310}
in which the brush was stuck.

\StephenInt:
So I carried the boat of incense then at Clongowes.
I am another now and yet the same.
A servant too.
A server of a servant.
