%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

\input{defs}
% \input{largeprint} \renewcommand{\gab}[1]{\marginpar[\tiny\raggedleft #1]{\tiny #1}}


\title{\Huge Eumaeus%\\
%\small Second Draft%
}
\author{George V. Reilly\\
{\small for the}\\
Wild Geese Players of Seattle\\
{\emph{www.WildGeeseSeattle.org}}\\
\\
{\small Chapter~16 of \emph{Ulysses} by James Joyce}\\
{\small Adapted from the 1922~edition at Project Gutenberg}}
\date{May \nth{8}, 2011}
\raggedbottom

\begin{document}

% Page i, Title Page
\maketitle
\thispagestyle{empty}
\pagebreak

% Page ii, Cast
\begin{tabular}{ll}
\multicolumn{2}{c}{\large Cast in Order of Appearance} \\
\\
\N1 & First Narrator. An unreliable sort. \\
\N2 & Narrator the Second. No better. \\
\N3 & Supplier of Tertiary Narrative. \\
\N4 & Narrator of the Fourth Part. \\
\N5 & Narrator, Fifth of his kind. \\
\N6 & Narrator the Sixth, last, perhaps least. \\
\Bloom & A latter-day Odysseus. \\
\Stephen & Telemachus; intelligent, but not wise. \\
\LordJohnCorley & A young sot. \\
\Italian{1} & An icecream seller. \\
\Italian{2} & A screamer. \\
\Italian{3} & Irascible also. \\
\Murphy & An old salt; less than he seems. \\
\SkinTheGoat & An erstwhile assassin. \\
\Someone & Who shall not be named. \\
\Campbell & Doppelganger of the townclerk. \\
\OldSpecimen & The old specimen in the corner. \\
\end{tabular}

\newpage


% Page 1, first page of script proper
\setcounter{page}{1}

\N1:
Preparatory to anything else |
Mr~Bloom brushed off the greater bulk of the shavings |
and handed Stephen the hat and ashplant |
and bucked him up generally in orthodox Samaritan fashion

\N2:
which he very badly needed.

\N3:
His (Stephen's) mind was not exactly what you would call wandering but a bit unsteady |
and on his expressed desire for some beverage to drink

\N4:
Mr~Bloom in view of the hour it was

\N5:
    --- and there being no pump of Vartry water available for their ablutions |
        let alone drinking purposes ---

\N4:
hit upon an expedient by suggesting, off the reel, |
the propriety of the cabman's shelter, as it was called,

\N6:
    --- hardly a stonesthrow away near Butt bridge ---

\N4:
where they might hit upon some drinkables |
\gab{10}
in the shape of a milk and soda or a mineral.

\N2:
But how to get there was the rub.

\N1:
For the nonce he was rather nonplussed |
but inasmuch as the duty plainly devolved upon him |
to take some measures on the subject |
he pondered suitable ways and means

\N3:
during which Stephen repeatedly yawned.

\N5:
So far as he could see |
he was rather pale in the face |
so that it occurred to him as highly advisable |
to get a conveyance of some description |
which would answer in their then condition, |
both of them being E.D.'ed, particularly Stephen,

\N2:
always assuming that there was such a thing to be found.

\N6:
Accordingly after a few such preliminaries as brushing, |
in spite of his having forgotten to take up his rather soapsuddy handkerchief |
after it had done yeoman service in the shaving line,
\gab{20}

\N2:
they both walked together along Beaver street

\N3:
    --- or, more properly, lane ---

\N2:
as far as the farrier's |
and the distinctly fetid atmosphere of the livery stables |
at the corner of Montgomery street |
where they made tracks to the left from thence |
debouching into Amiens street |
round by the corner of Dan Bergin's.

\N1:
But as he confidently anticipated |
there was not a sign of a Jehu plying for hire anywhere to be seen |
except a fourwheeler,

\N5:
    --- probably engaged by some fellows inside on the spree ---

\N1:
outside the North Star hotel |
and there was no symptom of its budging a quarter of an inch

\N4:
when Mr~Bloom, who was anything but a professional whistler, |
endeavoured to hail it by emitting a kind of a whistle, |
holding his arms arched over his head, twice.
\gab{30}

\N2:
This was a quandary but, bringing common sense to bear on it, |
evidently there was nothing for it |
but put a good face on the matter and foot it |
which they accordingly did.

\N6:
So, bevelling around by Mullett's and the Signal House which they shortly reached, |
they proceeded perforce in the direction of Amiens street railway terminus,

\N1:
Mr~Bloom being handicapped by the circumstance |
that one of the back buttons of his trousers had,

\N2:
    --- to vary the timehonoured adage ---

\N1:
gone the way of all buttons |
though, entering thoroughly into the spirit of the thing, |
he heroically made light of the mischance.

\N4:
So as neither of them were particularly pressed for time, |
as it happened, and the temperature refreshing
\gab{40}

\N5:
    --- since it cleared up after the recent visitation of Jupiter Pluvius ---

\N4:
they dandered along past by where the empty vehicle was waiting without a fare or a jarvey.

\N3:
As it so happened a Dublin United Tramways Company's sandstrewer happened to be returning |
and the elder man recounted to his companion |
\french{à~propos} of the incident his own truly miraculous escape |
of some little while back.

\N1:
They passed the main entrance of the Great Northern railway station, |
the starting point for Belfast, where of course |
all traffic was suspended at that late hour

\N3:
and passing the backdoor of the morgue

\N2:
    --- (a not very enticing locality, not to say gruesome to a degree,
        more especially at night) ---

\N3:
ultimately gained the Dock Tavern |
\gab{50}
and in due course turned into Store street, |
famous for its C~division police station.

\N1:
Between this point and the high at-present unlit warehouses of Beresford place |
Stephen thought to think of Ibsen, |
associated with Baird's the stonecutter's in his mind somehow

\N4:
    --- in Talbot place, first turning on the right ---

\N2:
while the other who was acting as his \latin{fidus Achates}
inhaled with internal satisfaction |
the smell of James Rourke's city bakery,

\N4:
    --- situated quite close to where they were ---

\N2:
the very palatable odour indeed of our daily bread, |
of all commodities of the public the primary and most indispensable.

\N6:
Bread, the staff of life, earn your bread, |
O tell me where is fancy bread, |
at Rourke's the baker's it is said.

\N1:
\french{En route} to his taciturn and,
\gab{60}

\N5:
    --- not to put too fine a point on it ---

\N1:
not yet perfectly sober companion |
Mr~Bloom who at all events was in complete possession of his faculties,

\N5:
    --- never more so, |
        in fact disgustingly sober ---

\N1:
spoke a word of caution re the dangers of nighttown, |
women of ill fame and swell mobsmen, |
which,

\N5:
    --- barely permissible once in a while though not as a habitual practice ---

\N1:
was of the nature of a regular deathtrap for young fellows of his age |
particularly if they had acquired drinking habits |
under the influence of liquor

\N3:
unless you knew a little jiujitsu for every contingency |
as even a fellow on the broad of his back could administer a nasty kick |
if you didn't look out.

\N2:
Highly providential was the appearance on the scene of Corny Kelleher |
\gab{70}
when Stephen was blissfully unconscious |
but for that man in the gap turning up at the eleventh hour |
the \french{finis} might have been |
that he might have been a candidate for the accident ward |
or, failing that, the bridewell |
and an appearance in the court next day before Mr~Tobias

\N4:
    --- or, he being the solicitor rather ---

\N2:
old Wall, he meant to say, |
or Mahony |
which simply spelt ruin for a chap when it got bruited about.

\N6:
The reason he mentioned the fact was that a lot of those policemen, |
whom he cordially disliked, |
were admittedly unscrupulous in the service of the Crown |
and, as Mr~Bloom put it,

\N1:
    --- recalling a case or two in the A~division in Clanbrassil street ---

\N6:
prepared to swear a hole through a ten-gallon pot.

\N5:
Never on the spot when wanted |
but in quiet parts of the city, Pembroke road for example, |
\gab{80}
the guardians of the law were well in evidence, |
the obvious reason being they were paid to protect the upper classes.

\N1:
Another thing he commented on was equipping soldiers with firearms |
or sidearms of any description liable to go off at any time |
which was tantamount to inciting them against civilians |
should by any chance they fall out over anything.

\N3:
You frittered away your time, he very sensibly maintained, |
and health and also character

\N4:
    --- besides which, the squandermania of the thing ---

\N3:
fast women of the \french{demimonde} ran away with
a lot of pounds, shillings, and pence into the bargain |
and the greatest danger of all was who you got drunk with.

\N2:
Though, touching the much vexed question of stimulants, |
\gab{90}
he relished a glass of choice old wine in season |
as both nourishing and bloodmaking |
and possessing aperient virtues

\N4:
    --- notably a good burgundy which he was a staunch believer in ---

\N2:
still never beyond a certain point |
where he invariably drew the line |
as it simply led to trouble all round |
to say nothing of your being at the tender mercy of others practically.

\N6:
Most of all he commented adversely on the desertion of Stephen |
by all his pubhunting \french{confrères} but one, |
a most glaring piece of ratting |
on the part of his brother medicos under all the circs.

\Stephen:
And that \emph{one} was Judas.

\N5:
Stephen said, |
who up to then had said nothing whatsoever of any kind.

\N1:
Discussing these and kindred topics |
\gab{100}
they made a beeline across the back of the Customhouse |
and passed under the Loop Line bridge |
where a brazier of coke burning in front of a sentrybox or something like one |
attracted their rather lagging footsteps.

\N4:
Stephen of his own accord stopped for no special reason |
to look at the heap of barren cobblestones |
and by the light emanating from the brazier |
he could just make out |
the darker figure of the corporation watchman |
inside the gloom of the sentrybox.

\N3:
He began to remember that this had happened |
or had been mentioned as having happened before |
but it cost him no small effort before he remembered |
that he recognised in the sentry |
a \latin{quondam} friend of his father's, Gumley.

\N2:
To avoid a meeting he drew nearer to the pillars of the railway bridge.
\gab{110}

\Bloom:
Someone saluted you

\N1:
Mr~Bloom said.

\N6:
A figure of middle height on the prowl evidently |
under the arches saluted again, calling:

\LordJohnCorley:
\emph{Night!}

\N3:
Stephen of course started rather dizzily |
and stopped to return the compliment.

\N4:
Mr~Bloom actuated by motives of inherent delicacy |
inasmuch as he always believed in minding his own business |
moved off but nevertheless remained on the \latin{qui vive} |
with just a shade of anxiety though not funkyish in the least.

\N1:
Though unusual in the Dublin area |
he knew that it was not by any means unknown for desperadoes
\gab{120}

\N5:
--- who had next to nothing to live on ---

\N1:
to be abroad waylaying and generally terrorising peaceable pedestrians |
by placing a pistol at their head |
in some secluded spot outside the city proper,

\N2:
famished loiterers of the Thames embankment category |
they might be hanging about there

\N3:
or simply marauders ready to decamp with whatever boodle they could |
in one fell swoop at a moment's notice,

\N1:
your money or your life, |
leaving you there to point a moral, |
gagged and garrotted.

\N2:
Stephen, that is when the accosting figure came to close quarters, |
though he was not in an over sober state himself |
recognised Corley's breath |
redolent of rotten cornjuice.
\gab{130}

\N3:
Lord John Corley some called him |
and his genealogy came about in this wise.

\N6:
He was the eldest son of inspector Corley of the G division, lately deceased, |
who had married a certain Katherine Brophy, |
the daughter of a Louth farmer.

\N4:
His grandfather Patrick Michael Corley of New Ross |
had married the widow of a publican there |
whose maiden name had been Katherine (also) Talbot.

\N1:
Rumour had it (though not proved) |
that she descended from the house of the lords Talbot de Malahide |
in whose mansion

\N5:
    --- really an unquestionably fine residence of its kind and well worth seeing ---

\N1:
her mother or aunt or some relative, |
a woman, as the tale went, of extreme beauty, |
had enjoyed the distinction of being in service in the washkitchen.
\gab{140}

\N2:
This therefore was the reason why |
the still comparatively young though dissolute man |
who now addressed Stephen was spoken of |
by some with facetious proclivities |
as Lord John Corley.

\N4:
Taking Stephen on one side |
he had the customary doleful ditty to tell.
Not as much as a farthing to purchase a night's lodgings.
His friends had all deserted him.
Furthermore he had a row with Lenehan and called him to Stephen |
a mean bloody swab |
with a sprinkling of a number of other uncalled-for expressions.

\N3:
He was out of a job and implored of Stephen to tell him |
where on God's earth he could get something, anything at all, to do.

\N1:
    --- No, it was the daughter of the mother in the washkitchen |
        \gab{150}
        that was fostersister to the heir of the house |
        or else they were connected through the mother in some way, |
        both occurrences happening at the same time |
        if the whole thing wasn't a complete fabrication from start to finish. ---

\N3:
Anyhow he was all in.

\LordJohnCorley:
I wouldn't ask you only, |
% pursued he, |
on my solemn oath |
and God knows I'm on the rocks.

\Stephen:
There'll be a job tomorrow or next day, |
% Stephen told him, |
in a boys' school at Dalkey for a gentleman usher.
Mr~Garrett Deasy.
Try it.
You may mention my name.

\LordJohnCorley:
Ah, God, |
% Corley replied, |
sure I couldn't teach in a school, man.
\gab{160}
I was never one of your bright ones, |
\stage{he added with a half laugh.}
I got stuck twice in the junior at the Christian Brothers.

\Stephen:
I have no place to sleep myself.
% Stephen informed him.

\N5:
Corley at the first go-off was inclined to suspect |
it was something to do with Stephen being fired out of his digs |
for bringing in a bloody tart off the street.
There was a dosshouse in Marlborough street, Mrs~Maloney's, |
but it was only a tanner touch and full of undesirables

\N6:
but M'Conachie told him you got a decent enough do |
in the Brazen Head over in Winetavern street

\N1:
    --- which was distantly suggestive to the person addressed |
        of friar Bacon ---

\N6:
for a bob.
\gab{170}

\N2:
He was starving too though he hadn't said a word about it.

\N4:
Though this sort of thing went on every other night or very near it |
still Stephen's feelings got the better of him in a sense |
though he knew that Corley's brandnew rigmarole |
on a par with the others was hardly deserving of much credence.

\N2:
However \latin{haud ignarus malorum miseris succurrere disco} etcetera |
as the Latin poet remarks |
especially as luck would have it |
he got paid his screw after every middle of the month on the sixteenth

\N1:
    --- which was the date of the month as a matter of fact ---

\N2:
though a good bit of the wherewithal was demolished.

\N1:
But the cream of the joke was |
nothing would get it out of Corley's head |
that Stephen was living in affluence |
and hadn't a thing to do but hand out the needful.
Whereas.
\gab{180}

\N2:
Stephen put his hand in a pocket anyhow |
not with the idea of finding any food there |
but thinking he might lend him |
anything up to a bob or so in lieu |
so that he might endeavour at all events |
and get sufficient to eat |
but the result was in the negative |
for, to his chagrin, he found his cash missing.

\N3:
A few broken biscuits were all the result of his investigation.

\N4:
He tried his hardest to recollect for the moment |
whether he had lost |
as well he might have |
or left

\N5:
because in that contingency |
it was not a pleasant lookout, very much the reverse in fact.
He was altogether too fagged out |
to institute a thorough search |
though he tried to recollect.

\N3:
About biscuits he dimly remembered.
\gab{190}
Who now exactly gave them he wondered or where was or did he buy.

\N6:
However in another pocket |
he came across what he surmised in the dark |
were pennies, erroneously however, as it turned out.

\LordJohnCorley:
Those are halfcrowns, man.
% Corley corrected him.

\N1:
And so in point of fact they turned out to be.

\N2:
Stephen anyhow lent him one of them.

\LordJohnCorley:
Thanks.
% Corley answered, |
You're a gentleman.
I'll pay you back one time.
Who's that with you?
I saw him a few times in the Bleeding Horse in Camden street |
with Boylan, the billsticker.
You might put in a good word for us to get me taken on there.
\gab{200}
I'd carry a sandwichboard |
only the girl in the office told me |
they're full up for the next three weeks, man.
God, you've to book ahead, man, |
you'd think it was for the Carl Rosa.
I don't give a shite anyway |
so long as I get a job, even as a crossing sweeper.

\N2:
Subsequently being not quite so down in the mouth |
after the two-and-six he got |
he informed Stephen about a fellow by the name of Bags Comisky |
that he said Stephen knew well out of Fullam's, the shipchandler's,

\N6:
    --- bookkeeper there that used to be often round in Nagle's back with O'Mara ---

\N2:
and a little chap with a stutter the name of Tighe.

\N4:
Anyhow he was lagged the night before last |
and fined ten bob for a drunk and disorderly |
and refusing to go with the constable.
\gab{210}

\N3:
Mr~Bloom in the meanwhile |
kept dodging about in the vicinity of the cobblestones |
near the brazier of coke |
in front of the corporation watchman's sentrybox

\N1:
who evidently a glutton for work, |
it struck him, |
was having a quiet forty winks for all intents and purposes |
on his own private account while Dublin slept.

\N5:
He threw an odd eye at the same time now and then |
at Stephen's anything but immaculately attired interlocutor |
as if he had seen that nobleman somewhere or other |
though where he was not in a position to truthfully state |
nor had he the remotest idea when.

\N2:
Being a levelheaded individual who could give points to not a few |
in point of shrewd observation |
he also remarked
\gab{220}
on the other's very dilapidated hat and slouchy wearing apparel |
generally testifying to a chronic impecuniosity.

\N6:
Palpably he was one of his hangers-on |
but for the matter of that |
it was merely a question of one preying on his nextdoor neighbour all round,

\N3:
    --- in every deep, so to put it, a deeper depth ---

\N6:
and for the matter of that |
if the man in the street chanced to be in the dock himself |
penal servitude with or without the option of a fine |
would be a very rara avis altogether.

\N4:
In any case he had a consummate amount of cool assurance |
intercepting people at that hour of the night or morning.
Pretty thick that was certainly.

\N3:
The pair parted company and Stephen rejoined Mr~Bloom |
who, with his practised eye, was not without perceiving |
\gab{230}
that he had succumbed to the blandiloquence of the other parasite.

\N5:
Alluding to the encounter he said, laughingly,

\N1:
Stephen, that is:

\Stephen:
He is down on his luck.
He asked me to ask you to ask somebody named Boylan, a billsticker, |
to give him a job as a sandwichman.

\N1:
At this intelligence, in which he seemingly evinced little interest, |
Mr~Bloom gazed abstractedly for the space of a half a second or so

\N2:
    --- in the direction of a bucket-dredger, rejoicing in the farfamed name of \Title{Eblana}, |
        moored alongside Customhouse quay and quite possibly out of repair ---

\N1:
whereupon he observed evasively:

\Bloom:
Everybody gets their own ration of luck, they say.
\gab{240}
Now you mention it his face was familiar to me.
But, leaving that for the moment, |
how much did you part with, |
% he queried, |
if I am not too inquisitive?

\Stephen:
Half a crown.
% Stephen responded.
I daresay he needs it to sleep somewhere.

\Bloom:
Needs!

\N1:
    --- Mr~Bloom ejaculated, |
        professing not the least surprise at the intelligence ---

\Bloom:
I can quite credit the assertion and I guarantee he invariably does.
Everyone according to his needs or everyone according to his deeds.
But, talking about things in general, |
where,
% added he with a smile, |
will you sleep yourself?
Walking to Sandycove is out of the question.
And even supposing you did |
\gab{250}
you won't get in after what occurred at Westland Row station.
Simply fag out there for nothing.
I don't mean to presume to dictate to you in the slightest degree |
but why did you leave your father's house?

\Stephen:
To seek misfortune.
% was Stephen's answer.

\Bloom:
I met your respected father on a recent occasion, |
% Mr~Bloom diplomatically returned, |
today in fact, |
or to be strictly accurate, on yesterday.
Where does he live at present?
I gathered in the course of conversation that he had moved.

\Stephen:
\stage{unconcernedly}
I believe he is in Dublin somewhere.
% Stephen answered unconcernedly.
Why?

\Bloom:
A gifted man,
% Mr~Bloom said of Mr~Dedalus senior, |
\gab{260}
in more respects than one |
and a born \french{raconteur} if ever there was one.
He takes great pride, quite legitimate, out of you.
You could go back perhaps.

\N2:
Bloom hasarded, |
still thinking of the very unpleasant scene at Westland Row terminus |
when it was perfectly evident that the other two,

\N4:
    --- Mulligan, that is, and that English tourist friend of his, |
        who eventually euchred their third companion ---

\N2:
were patently trying as if the whole bally station belonged to them |
to give Stephen the slip in the confusion, which they did.

\N5:
There was no response forthcoming to the suggestion however, |
such as it was, |
Stephen's mind's eye being too busily engaged
in repicturing his family hearth |
the last time he saw it |
\gab{270}
with his sister Dilly sitting by the ingle, |
her hair hanging down, |
waiting for some weak Trinidad shell cocoa |
that was in the soot-coated kettle to be done |
so that she and he could drink it |
with the oatmealwater for milk |
after the Friday herrings they had eaten at two a penny |
with an egg apiece for Maggy, Boody and Katey,

\N6:
    --- the cat meanwhile under the mangle |
        devouring a mess of eggshells and charred fish heads and bones |
        on a square of brown paper ---

\N5:
in accordance with the third precept of the church |
to fast and abstain on the days commanded, |
it being quarter tense |
or if not, ember days or something like that.

\Bloom:
No, |
% Mr~Bloom repeated again, |
I wouldn't personally repose much trust in that boon companion of yours |
\gab{280}
who contributes the humorous element, |
Dr~Mulligan, |
as a guide, philosopher and friend if I were in your shoes.
He knows which side his bread is buttered on |
though in all probability he never realised |
what it is to be without regular meals.
Of course you didn't notice as much as I did.
But it wouldn't occasion me the least surprise |
to learn that a pinch of tobacco or some narcotic |
was put in your drink for some ulterior object.

\N1:
Mr~Bloom understood however from all he heard |
that Dr~Mulligan was a versatile all-round man, |
by no means confined to medicine only, |
who was rapidly coming to the fore in his line |
and, if the report was verified, |
bade fair to enjoy a flourishing practice |
\gab{290}
in the not too distant future as a tony medical practitioner |
drawing a handsome fee for his services

\N2:
    --- in addition to which professional status |
        his rescue of that man from certain drowning |
        by artificial respiration and what they call first aid |
        at Skerries, or Malahide was it?, |
        was, he was bound to admit, |
        an exceedingly plucky deed |
        which he could not too highly praise ---

\N1:
so that frankly he was utterly at a loss |
to fathom what earthly reason could be at the back of it |
except he put it down to sheer cussedness or jealousy, pure and simple.

\Bloom:
Except it simply amounts to one thing |
and he is what they call picking your brains.
% he ventured to throw out.

\N3:
The guarded glance of half solicitude half curiosity |
\gab{300}
augmented by friendliness which he gave |
at Stephen's at-present morose expression of features |
did not throw a flood of light, none at all in fact |
on the problem |
as to whether he had let himself be badly bamboozled

\N4:
    --- to judge by two or three lowspirited remarks he let drop ---

\N3:
or the other way about, saw through the affair |
and for some reason or other best known to himself |
allowed matters to more or less.

\N3:
Grinding poverty did have that effect |
and he more than conjectured that, |
high educational abilities though he possessed, |
he experienced no little difficulty in making both ends meet.

\N5:
Adjacent to the men's public urinal |
they perceived an icecream car |
round which a group of presumably Italians in heated altercation |
\gab{310}
were getting rid of voluble expressions |
in their vivacious language |
in a particularly animated way, |
there being some little differences between the parties.

\Italian{1}:
\italian{Puttana madonna, che ci dia i quattrini! Ho ragione? Culo rotto!}

\Italian{2}:
\italian{Intendiamoci. Mezzo sovrano più\ldots.}

\Italian{3}:
\italian{Dice lui, però!}

\Italian{2}:
\italian{Mezzo.}

\Italian{1}:
\italian{Farabutto! Mortacci sui!}

\Italian{3}:
\italian{Ma ascolta! Cinque la testa più\ldots.}

\N6:
Mr~Bloom and Stephen entered the cabman's shelter, |
\gab{320}
an unpretentious wooden structure, |
where, prior to then, he had rarely if ever been before,

\N2:
the former having previously whispered to the latter |
a few hints anent the keeper of it |
said to be the once famous Skin-the-Goat Fitzharris, the invincible, |
though he could not vouch for the actual facts |
which quite possibly there was not one vestige of truth in.

\N1:
A few moments later |
saw our two noctambules safely seated in a discreet corner |
only to be greeted by stares |
from the decidedly miscellaneous collection
of waifs and strays and other nondescript specimens |
of the genus \latin{homo} already there |
engaged in eating and drinking diversified by conversation |
for whom they seemingly formed an object of marked curiosity.
\gab{330}

\Bloom:
Now touching a cup of coffee.

\N2:
    --- Mr~Bloom ventured to plausibly suggest |
        to break the ice ---

\Bloom:
It occurs to me you ought to sample something |
in the shape of solid food, say, a roll of some description.

\N3:
Accordingly his first act was |
with characteristic \french{sangfroid} |
to order these commodities quietly.

\N4:
The \latin{hoi polloi} of jarvies or stevedores or whatever they were |
after a cursory examination |
turned their eyes apparently dissatisfied, away |
though one redbearded bibulous individual, |
portion of whose hair was greyish, |
a sailor probably, |
still stared for some appreciable time |
before transferring his rapt attention to the floor.

\N5:
Mr~Bloom, availing himself of the right of free speech, |
\gab{340}
he having just a bowing acquaintance with the language in dispute,

\N6:
    --- though, to be sure, rather in a quandary over \italian{voglio} ---

\N5:
remarked to his \french{protégé} in an audible tone of voice |
\french{à~propos} of the battle royal in the street |
which was still raging fast and furious:

\Bloom:
A beautiful language.
I mean for singing purposes.
Why do you not write your poetry in that language?
\italian{Bella Poetria!}
It is so melodious and full.
\italian{Belladonna. Voglio.}

\N1:
Stephen, who was trying his dead best to yawn if he could, |
suffering from lassitude generally, |
replied:

\Stephen:
To fill the ear of a cow elephant.
They were haggling over money.
\gab{350}

\Bloom:
Is that so?
% Mr~Bloom asked.
Of course\ldots

\N6:
    --- He subjoined pensively, |
        at the inward reflection of there being more languages to start with
        than were absolutely necessary ---

\Bloom:
It may be only the southern glamour that surrounds it.

\N1:
The keeper of the shelter |
in the middle of this \french{tête-à-tête} |
put a boiling swimming cup of a choice concoction labelled coffee |
on the table and a rather antediluvian specimen of a bun, or so it seemed.
After which he beat a retreat to his counter.

\N2:
Mr~Bloom determining to have a good square look at him later on |
so as not to appear to.

\N3:
For which reason |
he encouraged Stephen to proceed with his eyes |
while he did the honours |
by surreptitiously pushing the cup of |
\gab{360}
what was temporarily supposed to be called coffee |
gradually nearer him.

\N4:
After a pause of some little time, Stephen said:

\Stephen:
Sounds are impostures, like names.
Cicero, Podmore. Napoleon, Mr~Goodbody. Jesus, Mr~Doyle.
Shakespeares were as common as Murphies.
What's in a name?

\Bloom:
Yes, to be sure.
% Mr~Bloom unaffectedly concurred.
Of course.
Our name was changed too.
% he added, pushing the so-called roll across.

\N5:
The redbearded sailor who had his weather eye on the newcomers |
boarded Stephen, |
whom he had singled out for attention in particular, |
squarely by asking:

\Murphy:
And what might your name be?
\gab{370}

\N6:
Just in the nick of time |
Mr~Bloom touched his companion's boot |
but Stephen, apparently disregarding the warm pressure from an unexpected quarter, |
answered:

\Stephen:
Dedalus.

\N5:
The sailor stared at him heavily from a pair of drowsy baggy eyes, |
rather bunged up from excessive use of boose, |
preferably good old Hollands and water.
At length, he asked:

\Murphy:
You know Simon Dedalus?

\Stephen:
I've heard of him.
% Stephen said.

\N6:
Mr~Bloom was all at sea for a moment, |
\gab{380}
seeing the others evidently eavesdropping too.

\Murphy:
He's Irish.

\N5:
The seaman bold affirmed, |
staring still in much the same way and nodding.

\Murphy:
All Irish.

\Stephen:
All too Irish.
% Stephen rejoined.

\N1:
As for Mr~Bloom |
he could neither make head or tail of the whole business |
and he was just asking himself what possible connection

\N2:
when the sailor of his own accord
turned to the other occupants of the shelter |
with the remark:

\Murphy:
I seen him shoot two eggs off two bottles |
at fifty yards over his shoulder.
The lefthand dead shot.
\gab{390}

\N2:
Though he was slightly hampered by an occasional stammer |
and his gestures being also clumsy as it was |
still he did his best to explain.

\Murphy:
Bottles out there, say.
Fifty yards measured.
Eggs on the bottles.
Cocks his gun over his shoulder.
Aims.

\N2:
He turned his body half round, |
shut up his right eye completely.
Then he screwed his features up someway sideways |
and glared out into the night |
with an unprepossessing cast of countenance.

\Murphy:
Pom!
% he then shouted once.

\N3:
The entire audience waited, anticipating an additional detonation, |
there being still a further egg.
\gab{400}

\stage{pause}

\Murphy:
Pom!
% he shouted twice.

\N3:
Egg two evidently demolished, he nodded and winked, |
adding bloodthirstily:

\Murphy:
\begin{verse}
    Buffalo Bill shoots to kill,\\
    Never missed nor he never will.
\end{verse}

\N4:
A silence ensued |
till Mr~Bloom for agreeableness' sake |
just felt like asking him |
whether it was for a marksmanship competition like the Bisley.

\Murphy:
Beg pardon.
% the sailor said.

\Bloom:
Long ago?

\N4:
Mr~Bloom pursued |
without flinching a hairsbreadth.

\N5:
The sailor replied, |
relaxing to a certain extent |
\gab{410}
under the magic influence of diamond cut diamond.

\Murphy:
Why, it might be a matter of ten years.
He toured the wide world with Hengler's Royal Circus.
I seen him do that in Stockholm.

\Bloom:
Curious coincidence.

\N4:
    --- Mr~Bloom confided to Stephen unobtrusively ---

\Murphy:
Murphy's my name, |
% the sailor continued.
D.B.~Murphy of Carrigaloe.
Know where that is?

\Stephen:
Queenstown harbour.
% Stephen replied.

\Murphy:
That's right,
% the sailor said.
Fort Camden and Fort Carlisle.
That's where I hails from.
I belongs there.
That's where I hails from.
My little woman's down there.
She's waiting for me, I know.
\gab{420}
\emph{For England, home and beauty.}
She's my own true wife I haven't seen for seven years now, sailing about.

\N4:
Mr~Bloom could easily picture his advent on this scene,

\N5:
the homecoming to the mariner's roadside shieling |
after having diddled Davy Jones, |
a rainy night with a blind moon.
Across the world for a wife.

\N6:
Quite a number of stories there were |
on that particular Alice Ben Bolt topic, |
Enoch Arden and Rip van Winkle

\N1:
and does anybody hereabouts remember Caoc O'Leary, |
a favourite and most trying declamation piece |
by the way of poor John Casey |
and a bit of perfect poetry in its own small way.

\N5:
Never about the runaway wife coming back, |
however much devoted to the absentee.

\N2:
The face at the window!
\gab{430}

\N3:
Judge of his astonishment when he finally did breast the tape |
and the awful truth dawned upon him anent his better half, |
wrecked in his affections.

\N2:
You little expected me |
but I've come to stay and make a fresh start.

\N3:
There she sits, a grasswidow, at the selfsame fireside.

\N2:
Believes me dead, rocked in the cradle of the deep.

\N3:
And there sits uncle Chubb or Tomkin, as the case might be, |
the publican of the Crown and Anchor, in shirtsleeves, |
eating rumpsteak and onions.

\N2:
No chair for father.

\N1:
Broo! The wind!

\N3:
Her brandnew arrival is on her knee, \latin{postmortem} child.

\N6:
With a high ro! and a randy ro! and my galloping tearing tandy, O!

\N5:
Bow to the inevitable.
Grin and bear it.

\N2:
I remain with much love |
your brokenhearted husband D.B.~Murphy.
\gab{440}

\N4:
The sailor, who scarcely seemed to be a Dublin resident, |
turned to one of the jarvies with the request:

\Murphy:
You don't happen to have such a thing as a spare chaw about you?

\N1:
The jarvey addressed as it happened had not |
but the keeper took a die of plug from his good jacket hanging on a nail |
and the desired object was passed from hand to hand.

\Murphy:
Thank you.
% the sailor said.

\N4:
He deposited the quid in his gob |
and, chewing and with some slow stammers, |
proceeded:

\Murphy:
We come up this morning eleven o'clock.
\gab{450}
The threemaster \Title{Rosevean} from Bridgwater with bricks.
I shipped to get over.
Paid off this afternoon.
There's my discharge.
See?
D.B.~Murphy. A.B.S.

\N4:
In confirmation of which statement |
he extricated from an inside pocket |
and handed to his neighbour |
a not very cleanlooking folded document.

\SkinTheGoat:
You must have seen a fair share of the world.

\N2:
The keeper remarked, leaning on the counter.

\Murphy:
Why, |
% the sailor answered upon reflection upon it, |
I've circumnavigated a bit since I first joined on.
I was in the Red Sea.
I was in China and North America and South America.
We was chased by pirates one voyage.
\gab{460}
I seen icebergs plenty, growlers.
I was in Stockholm and the Black Sea, |
the Dardanelles under Captain Dalton, |
the best bloody man that ever scuttled a ship.
I seen Russia.
\russian{Gospodi pomilyou.}
That's how the Russians prays.

\Campbell:
You seen queer sights, don't be talking.

\N3:
Put in a jarvey.

\Murphy:
Why---

\N4:
The sailor said, shifting his partially chewed plug.

\Murphy:
Why, I seen queer things too, ups and downs.
I seen a crocodile bite the fluke of an anchor |
same as I chew that quid.

\N4:
He took out of his mouth the pulpy quid |
and, lodging it between his teeth, |
bit ferociously:

\Murphy:
Khaan! Like that.
\gab{470}
And I seen maneaters in Peru that eats corpses and the livers of horses.
Look here.
Here they are. A friend of mine sent me.

\N5:
He fumbled out a picture postcard from his inside pocket |
which seemed to be in its way |
a species of repository and pushed it along the table.
The printed matter on it stated: |
\spanish{Choza de Indios. Beni, Bolivia.}

\N6:
All focussed their attention at the scene exhibited, |
a group of savage women in striped loincloths, |
squatted, blinking, suckling, frowning, |
sleeping amid a swarm of infants

\N1:
    --- there must have been quite a score of them ---

\N6:
outside some primitive shanties of osier.

\Murphy:
Chews coca all day.
% the communicative tarpaulin added.
Stomachs like breadgraters.
Cuts off their diddies when they can't bear no more children.
\gab{480}
See them sitting there stark ballocknaked |
eating a dead horse's liver raw.

\N2:
His postcard proved a centre of attraction for Messrs the greenhorns |
for several minutes |
if not more.

\Murphy:
Know how to keep them off?
% he inquired generally.

\N2:
Nobody volunteering a statement |
he winked, saying:

\Murphy:
Glass.
That boggles 'em.
Glass.

\N3:
Mr~Bloom, without evincing surprise, |
unostentatiously turned over the card |
to peruse the partially obliterated address and postmark.

\N4:
It ran as follows: |
\spanish{Tarjeta Postal, Señor A.~Boudin, Galeria Becche, Santiago, Chile.}

\N3:
There was no message evidently, as he took particular notice.
\gab{490}
Though not an implicit believer in the lurid story narrated

\N4:
    --- or the eggsniping transaction for that matter |
        despite William Tell and the Lazarillo--Don Cesar de Bazan incident
        depicted in \Title{Maritana} |
        on which occasion the former's ball passed through the latter's hat ---

\N3:
having detected a discrepancy between his name

\N4:
    --- assuming he was the person he represented himself to be |
        and not sailing under false colours |
        after having boxed the compass on the strict~q.t.\ somewhere ---

\N3:
and the fictitious addressee of the missive |
which made him nourish some suspicions of our friend's \latin{bona fides}.

\N5:
Nevertheless it reminded him in a way |
of a longcherished plan he meant to one day realise |
some Wednesday or Saturday |
\gab{500}
of travelling to London via long sea |
not to say that he had ever travelled extensively to any great extent |
but he was at heart a born adventurer |
though by a trick of fate |
he had consistently remained a landlubber |
except you call going to Holyhead which was his longest.

\N6:
Martin Cunningham frequently said he would work a pass through Egan |
but some deuced hitch or other eternally cropped up |
with the net result that the scheme fell through.
But even suppose it did come to planking down the needful |
and breaking Boyd's heart |
it was not so dear, purse permitting, |
a few guineas at the outside |
considering the fare to Mullingar |
where he figured on going was five and six, there and back.

\N1:
The trip would benefit health |
on account of the bracing ozone |
\gab{510}
and be in every way thoroughly pleasurable, |
especially for a chap whose liver was out of order, |
seeing the different places along the route, |
Plymouth, Falmouth, Southampton and so on |
culminating in an instructive tour |
of the sights of the great metropolis, |
the spectacle of our modern Babylon |
where doubtless he would see the greatest improvement, |
tower, abbey, wealth of Park lane to renew acquaintance with.

\N2:
Another thing just struck him |
as a by no means bad notion |
was he might have a gaze around on the spot |
to see about trying to make arrangements |
about a concert tour of summer music |
embracing the most prominent pleasure resorts, |
Margate with mixed bathing and first-rate hydros and spas, |
\gab{520}
Eastbourne, Scarborough, Margate and so on, |
beautiful Bournemouth, the Channel islands and similar bijou spots, |
which might prove highly remunerative.

\N3:
Not, of course, with a hole and corner scratch company |
or local ladies on the job,

\N4:
    --- witness Mrs~C.P. M'Coy type |
        lend me your valise and I'll post you the ticket ---

\N3:
No, something top notch, an all star Irish caste, |
the Tweedy--Flower grand opera company |
with his own legal consort as leading lady |
as a sort of counterblast to the Elster Grimes and Moody--Manners.
Perfectly simple matter |
and he was quite sanguine of success, |
providing puffs in the local papers |
could be managed by some fellow with a bit of bounce |
who could pull the indispensable wires |
and thus combine business with pleasure.

\N4:
But who?
That was the rub.
\gab{530}

\N5:
Also, without being actually positive, |
it struck him a great field was to be opened up |
in the line of opening up new routes |
to keep pace with the times |
\french{apropos} of the Fishguard--Rosslare route |
which, it was mooted, |
was once more on the \french{tapis} in the circumlocution departments |
with the usual quantity of red tape |
and dillydallying of effete fogeydom and dunderheads generally.

\N6:
A great opportunity there certainly was for push and enterprise |
to meet the travelling needs of the public at large, |
the average man, i.e.\ Brown, Robinson and Co.

\N1:
It was a subject of regret |
and absurd as well on the face of it |
and no small blame to our vaunted society |
that the man in the street, |
\gab{540}
when the system really needed toning up, |
for the matter of a couple of paltry pounds |
was debarred from seeing more of the world they lived in |
instead of being always and ever cooped up |
since my old stick-in-the-mud took me for a wife.

\N2:
After all, hang it, |
they had their eleven and more humdrum months of it |
and merited a radical change of \french{venue} after the grind of city life |
in the summertime for choice |
when dame Nature is at her spectacular best |
constituting nothing short of a new lease of life.

\N3:
There were equally excellent opportunities for vacationists in the home island, |
delightful sylvan spots for rejuvenation, |
offering a plethora of attractions |
as well as a bracing tonic for the system in and around Dublin |
\gab{550}
and its picturesque environs |
even Poulaphouca to which there was a steamtram, |
but also farther away from the madding crowd |
in Wicklow, rightly termed the garden of Ireland, |
an ideal neighbourhood for elderly wheelmen |
so long as it didn't come down.

\N4:
And in the wilds of Donegal |
where if report spoke true |
the \french{coup d'œil} was exceedingly grand |
though the lastnamed locality was not easily get-at-able |
so that the influx of visitors |
was not as yet all that it might be |
considering the signal benefits to be derived from it.

\N5:
While Howth with its historic associations and otherwise,

\N6:
    --- Silken Thomas, Grace O'Malley, George~IV, |
        rhododendrons several hundred feet above sealevel ---

\N5:
was a favourite haunt with all sorts and conditions of men |
\gab{560}
especially in the spring |
when young men's fancy, |
though it had its own toll of deaths |
by falling off the cliffs by design or accidentally, |
usually, by the way, on their left leg, |
it being only about three quarters of an hour's run from the pillar.

\N6:
Because of course up-to-date tourist travelling was |
as yet merely in its infancy, so to speak, |
and the accommodation left much to be desired.
Interesting to fathom |
it seemed to him |
from a motive of curiosity, pure and simple, |
was whether it was the traffic that created the route |
or vice-versa |
or the two sides in fact.

\N1:
He turned back the other side of the card, picture, |
and passed it along to Stephen.

\Murphy:
I seen a Chinese one time, |
% related the doughty narrator, |
\gab{570}
that had little pills like putty and he put them in the water |
and they opened and every pill was something different.
One was a ship, another was a house, another was a flower.
Cooks rats in your soup, |
% he appetisingly added, |
the chinks does.

\N2:
Possibly perceiving an expression of dubiosity on their faces |
the globetrotter went on, |
adhering to his adventures.

\Murphy:
And I seen a man killed in Trieste by an Italian chap.
Knife in his back.
Knife like that.

\N2:
Whilst speaking he produced a dangerous-looking claspknife |
quite in keeping with his character |
and held it in the striking position.

\Murphy:
In a knockingshop it was |
\gab{580}
count of a try-on between two smugglers.
Fellow hid behind a door, come up behind him.
Like that.
\emph{Prepare to meet your God}, says he.
Chuk!
It went into his back up to the butt.

\N2:
His heavy glance drowsily roaming about |
kind of defied their further questions |
even should they |
by any chance |
want to.

\Murphy:
That's a good bit of steel.

\N2:
Repeated he, |
examining his formidable \italian{stiletto}.
After which harrowing \french{dénouement} |
sufficient to appal the stoutest |
he snapped the blade to |
and stowed the weapon in question away as before |
in his chamber of horrors |
otherwise pocket.

\Someone:
They're great for the cold steel.

\N3:
Somebody who was evidently quite in the dark |
said for the benefit of them all.
\gab{590}

\Someone:
That was why they thought the park murders of the invincibles |
was done by foreigners on account of them using knives.

\N3:
At this remark passed |
obviously in the spirit of \emph{where ignorance is bliss}

\N4:
Mr~B. and Stephen, each in his own particular way, |
both instinctively exchanged meaning glances,

\N1:
    --- in a religious silence of the strictly \french{entre nous} variety however ---

\N4:
towards where Skin-the-Goat, \latin{alias} the keeper, |
not turning a hair, |
was drawing spurts of liquid from his boiler affair.

\N5:
His inscrutable face which was really a work of art, |
a perfect study in itself, beggaring description, |
conveyed the impression that he didn't understand one jot of what was going on.
Funny, very!
\gab{600}

\N6:
There ensued a somewhat lengthy pause.
One man was reading in fits and starts |
a stained-by-coffee evening journal, |
another the card with the natives \spanish{choza de}, |
another the seaman's discharge.

\N1:
Mr~Bloom, |
so far as he was personally concerned, |
was just pondering in pensive mood.
He vividly recollected when the occurrence alluded to took place |
as well as yesterday, |
roughly some score of years previously |
in the days of the land troubles, |
when it took the civilised world by storm, figuratively speaking, |
early in the eighties, |
eighty-one to be correct, |
when he was just turned fifteen.

\Murphy:
Ay, boss.
% the sailor broke in.
Give us back them papers.

\N2:
The request being complied with |
\gab{610}
he clawed them up with a scrape.

\Bloom:
Have you seen the rock of Gibraltar?
% Mr~Bloom inquired.

\N2:
The sailor grimaced, |
chewing, |
in a way that might be read as yes, ay, or no.

\Bloom:
Ah, you've touched there too, |
Europa point.

\N3:
Mr~Bloom said, |
thinking he had, |
in the hope that the rover might possibly by some reminiscences

\N2:
but he failed to do so, |
simply letting spirt a jet of spew into the sawdust, |
and shook his head with a sort of lazy scorn.

\Bloom:
What year would that be about?
% Mr~B. interrogated.
Can you recall the boats?

\N2:
Our \french{soi-disant} sailor munched heavily awhile hungrily |
\gab{620}
before answering:

\Murphy:
I'm tired of all them rocks in the sea, |
% he said, |
and boats and ships.
Salt junk all the time.

\N2:
Tired seemingly, |
he ceased.

\N3:
His questioner perceiving that he was not likely |
to get a great deal of change out of such a wily old customer,

\N4:
fell to woolgathering on the enormous dimensions of the water about the globe, |
suffice it to say that, |
as a casual glance at the map revealed, |
it covered fully three-fourths of it |
and he fully realised accordingly what it meant to rule the waves.

\N5:
On more than one occasion, a dozen at the lowest, |
near the North Bull at Dollymount |
\gab{630}
he had remarked a superannuated old salt, evidently derelict, |
seated habitually near the not particularly redolent sea on the wall, |
staring quite obliviously at it and it at him, |
dreaming of fresh woods and pastures new |
as someone somewhere sings.

\N5:
And it left him wondering why.
Possibly he had tried to find out the secret for himself, |
floundering up and down the antipodes and all that sort of thing |
and over and under, |
well, not exactly under, |
tempting the fates.
And the odds were twenty to nil there was really no secret about it at all.

\N6:
Nevertheless, without going into the \latin{minutiae} of the business, |
the eloquent fact remained |
that the sea was there in all its glory |
and in the natural course of things |
somebody or other had to sail on it |
and fly in the face of providence.
\gab{640}

\N1:
Though it merely went to show how people usually contrived |
to load that sort of onus on to the other fellow |
like the hell idea and the lottery and insurance |
which were run on identically the same lines

\N2:
so that for that very reason if no other |
lifeboat Sunday was a highly laudable institution |
to which the public at large, |
no matter where living inland or seaside, as the case might be, |
having it brought home to them like that |
should extend its gratitude also |
to the harbourmasters and coastguard service |
who had to man the rigging and push off and out |
amid the elements |
whatever the season |
when duty called

\N3:
\emph{Ireland expects that every man} and so on |
and sometimes had a terrible time of it in the wintertime |
not forgetting the Irish lights, Kish and others, |
\gab{650}
liable to capsize at any moment,

\N4:
rounding which he once with his daughter |
had experienced some remarkably choppy, not to say stormy, weather.

\Murphy:
There was a fellow sailed with me in the \Title{Rover}, |
% the old seadog, himself a rover, proceeded, |
went ashore and took up a soft job as gentleman's valet at six quid a month.
Them are his trousers I've on me and he gave me an oilskin and that jackknife.
I'm game for that job, shaving and brushup.
I hate roaming about.
There's my son now, Danny, run off to sea |
and his mother got him took in a draper's in Cork |
where he could be drawing easy money.

\Campbell:
What age is he?

\N4:
Queried one hearer |
\gab{660}
who, by the way, seen from the side, |
bore a distant resemblance to Henry Campbell, the townclerk, |
away from the carking cares of office, |
unwashed of course and in a seedy getup |
and a strong suspicion of nosepaint about the nasal appendage.

\Murphy:
Why, |
\stage{the sailor answered with a slow puzzled utterance}
my son, Danny?
He'd be about eighteen now, way I figure it.

\N5:
The Skibbereen father
hereupon tore open his grey or unclean anyhow shirt with his two hands |
and scratched away at his chest |
on which was to be seen an image |
tattooed in blue Chinese ink intended to represent an anchor.

\Murphy:
There was lice in that bunk in Bridgwater,
\gab{670}
% he remarked, |
sure as nuts.
I must get a wash tomorrow or next day.
It's them black lads I objects to.
I hate those buggers.
Suck your blood dry, they does.

\N6:
Seeing they were all looking at his chest |
he accommodatingly dragged his shirt more open |
so that on top of the timehonoured symbol of the mariner's hope and rest |
they had a full view of the figure~16 |
and a young man's sideface looking frowningly rather.

\Murphy:
Tattoo.
% the exhibitor explained.
That was done when we were lying becalmed |
off Odessa in the Black Sea under Captain Dalton.
Fellow, the name of Antonio, done that.
There he is himself, a Greek.

\Someone:
Did it hurt much doing it?
\gab{680}
% one asked the sailor.

\N1:
That worthy, however, was busily engaged in collecting round the.
Someway in his.
Squeezing or.

\Murphy:
See here.

\N1:
The sailor said, |
showing Antonio.

\Murphy:
There he is cursing the mate.
And there he is now, |
the same fellow.

\N1:
He added, |
pulling the skin with his fingers, |
some special knack evidently, and he laughing at a yarn.

\N2:
And in point of fact the young man named Antonio's livid face |
did actually look like forced smiling |
and the curious effect excited the unreserved admiration of everybody |
including Skin-the-Goat, who this time stretched over.

\Murphy:
Ay, ay.

\N2:
Sighed the sailor, |
\gab{690}
looking down on his manly chest.

\Murphy:
He's gone too.
Ate by sharks after.
Ay, ay.

\N2:
He let go of the skin so
that the profile resumed the normal expression of before.

\N3:
``Neat bit of work,''
one longshoreman said.

\N4:
``And what's the number for?''
loafer number two queried.

\N5:
``Eaten alive?''
a third asked the sailor.

\Murphy:
Ay, ay.

\N2:
Sighed again the latter personage, |
more cheerily this time |
with some sort of a half smile for a brief duration only |
in the direction of the questioner about the number.

\Murphy:
Ate.
A Greek he was.

\N2:
And then he added with rather gallowsbird humour |
\gab{700}
considering his alleged end:

\Murphy:
\begin{verse}
    As bad as old Antonio,\\
    For he left me on my ownio.
\end{verse}

\N6:
The face of a streetwalker glazed and haggard under a black straw hat |
peered askew round the door of the shelter |
palpably reconnoitring on her own |
with the object of bringing more grist to her mill.

\N1:
Mr~Bloom, scarcely knowing which way to look, |
turned away on the moment flusterfied but outwardly calm, |
and, picking up from the table the pink sheet of the Abbey street organ |
which the jarvey, if such he was, had laid aside, |
he picked it up and looked at the pink of the paper |
though why pink.
\gab{710}

\N6:
His reason for so doing |
was he recognised on the moment round the door the same face |
he had caught a fleeting glimpse of that afternoon on Ormond quay, |
the partially idiotic female, namely, of the lane |
who knew the lady in the brown costume does be with you |
(Mrs~B.) |
and begged the chance of his washing.

\N1:
Also why washing |
which seemed rather vague than not, your washing.
Still candour compelled him to admit |
he had washed his wife's undergarments when soiled in Holles street |
and women would and did too |
a man's similar garments

\N3:
    --- initialled with Bewley and Draper's marking ink |
        (hers were, that is) ---

\N1:
if they really loved him, |
that is to say, love me, love my dirty shirt.

\N6:
Still just then, being on tenterhooks, |
\gab{720}
he desired the female's room more than her company |
so it came as a genuine relief |
when the keeper made her a rude sign to take herself off.
Round the side of the \Title{Evening Telegraph} |
he just caught a fleeting glimpse of her face |
round the side of the door with a kind of demented glassy grin |
showing that she was not exactly all there, |
viewing with evident amusement
the group of gazers round skipper Murphy's nautical chest |
and then there was no more of her.

\SkinTheGoat:
The gunboat.

\N2:
The keeper said.

\Bloom:
It beats me.

\N3:
    --- Mr~Bloom confided to Stephen ---

\Bloom:
Medically I am speaking, |
how a wretched creature like that from the Lock hospital |
reeking with disease |
can be barefaced enough to solicit |
\gab{730}
or how any man in his sober senses, |
if he values his health in the least.
Unfortunate creature!
Of course I suppose some man is ultimately responsible for her condition.
Still no matter what the cause is from\ldots.

\N4:
Stephen had not noticed her and shrugged his shoulders, |
merely remarking:

\Stephen:
In this country people sell much more than she ever had |
and do a roaring trade.
Fear not them that sell the body but have not power to buy the soul.
She is a bad merchant.
She buys dear and sells cheap.

\N5:
The elder man,

\N6:
    --- though not by any manner of means an old maid or a prude ---

\N5:
said it was nothing short of a crying scandal |
\gab{740}
that ought to be put a stop to \latin{instanter} |
to say that women of that stamp

\N6:
    --- quite apart from any oldmaidish squeamishness on the subject, |
        a necessary evil ---

\N5:
were not licensed and medically inspected by the proper authorities, |
a thing, he could truthfully state, |
he, as a \latin{paterfamilias}, |
was a stalwart advocate of from the very first start.
Whoever embarked on a policy of the sort, |
he said, |
and ventilated the matter thoroughly |
would confer a lasting boon on everybody concerned.

\Bloom:
You as a good catholic, |
% he observed, |
talking of body and soul, |
believe in the soul.
Or do you mean the intelligence, the brainpower as such, |
as distinct from any outside object, |
the table, let us say, that cup.
\gab{750}
I believe in that myself because it has been explained by competent men |
as the convolutions of the grey matter.
Otherwise we would never have such inventions as X~rays, for instance.
Do you?

\N1:
Thus cornered, |
Stephen had to make a superhuman effort of memory |
to try and concentrate and remember before he could say:

\Stephen:
They tell me on the best authority |
it is a simple substance and therefore incorruptible.
It would be immortal, I understand, |
but for the possibility of its annihilation |
by its First Cause |
Who, from all I can hear, |
is quite capable of adding that to the number of His other practical jokes, |
\latin{corruptio per se} and \latin{corruptio per accidens} |
both being excluded by court etiquette.
\gab{760}

\N2:
Mr~Bloom thoroughly acquiesced in the general gist of this |
though the mystical finesse involved |
was a bit out of his sublunary depth |
still he felt bound to enter a demurrer on the head of simple, |
promptly rejoining:

\Bloom:
Simple? I shouldn't think that is the proper word.
Of course, I grant you, to concede a point, |
you do knock across a simple soul once in a blue moon.
But what I am anxious to arrive at is |
it is one thing for instance |
to invent those rays Röntgen did |
or the telescope like Edison, |
though I believe it was before his time |
Galileo was the man, I mean, |
and the same applies to the laws, for example, |
of a farreaching natural phenomenon such as electricity |
but it's a horse of quite another colour |
\gab{770}
to say you believe in the existence of a supernatural God.

\Stephen:
O that, |
\stage{Stephen expostulated}
has been proved conclusively |
by several of the bestknown passages in Holy Writ, |
apart from circumstantial evidence.

\N3:
On this knotty point however |
the views of the pair,

\N4:
    --- poles apart as they were both in schooling and everything else |
        with the marked difference in their respective ages ---

\N3:
clashed.

\Bloom:
Has been?

\N1:
The more experienced of the two objected, |
sticking to his original point with a smile of unbelief.

\Bloom:
I'm not so sure about that.
That's a matter for everyman's opinion |
and, without dragging in the sectarian side of the business, |
I beg to differ with you \latin{in toto} there.
\gab{780}
My belief is, |
to tell you the candid truth, |
that those bits were genuine forgeries |
all of them put in by monks most probably |
or it's the big question of our national poet over again, |
who precisely wrote them |
like \Title{Hamlet} and Bacon, |
as, you who know your Shakespeare infinitely better than I, |
of course I needn't tell you.
Can't you drink that coffee, by the way? Let me stir it.
And take a piece of that bun.
It's like one of our skipper's bricks disguised.
Still no-one can give what he hasn't got.
Try a bit.

\Stephen:
Couldn't.

\N2:
Stephen contrived to get out, |
his mental organs for the moment refusing to dictate further.

\N3:
Faultfinding being a proverbially bad hat |
\gab{790}
Mr~Bloom thought well to stir |
or try to |
the clotted sugar from the bottom |
and reflected with something approaching acrimony |
on the Coffee Palace and its temperance (and lucrative) work.
To be sure it was a legitimate object |
and beyond yea or nay did a world of good, |
shelters such as the present one they were in |
run on teetotal lines for vagrants at night, |
concerts, dramatic evenings and useful lectures (admittance free) |
by qualified men for the lower orders.

\N4:
On the other hand |
he had a distinct and painful recollection |
they paid his wife, Madam Marion Tweedy |
who had been prominently associated with it at one time, |
a very modest remuneration indeed for her pianoplaying.
The idea, |
he was strongly inclined to believe, |
\gab{800}
was to do good and net a profit, |
there being no competition to speak of.

\N5:
Sulphate of copper poison SO$_4$ or something in some dried peas |
he remembered reading of |
in a cheap eatinghouse somewhere |
but he couldn't remember when it was or where.
Anyhow inspection, medical inspection, of all eatables |
seemed to him more than ever necessary |
which possibly accounted for the vogue of Dr~Tibble's Vi--Cocoa |
on account of the medical analysis involved.

\Bloom:
Have a shot at it now.

\N6:
He ventured to say of the coffee |
after being stirred.

\N1:
Thus prevailed on to at any rate taste it |
Stephen lifted the heavy mug from the brown puddle it clopped out of |
when taken up by the handle |
and took a sip of the offending beverage.
\gab{810}

\Bloom:
Still, it's solid food.
I'm a stickler for solid food.

\N2:
His good genius urged,
his one and only reason being |
not gormandising in the least |
but regular meals as the \latin{sine qua non} for any kind of proper work, |
mental or manual.

\Bloom:
You ought to eat more solid food.
You would feel a different man.

\Stephen:
Liquids I can eat.
% Stephen said.
But O, oblige me by taking away that knife.
I can't look at the point of it.
It reminds me of Roman history.

\N3:
Mr~Bloom promptly did as suggested and removed the incriminated article, |
a blunt hornhandled ordinary knife |
with nothing particularly Roman or antique about it to the lay eye, |
observing that the point was the least conspicuous point about it.
\gab{820}

\Bloom:
Our mutual friend's stories are like himself.

\N4:
    --- Mr~Bloom \french{apropos} of knives
        remarked to his \italian{confidante} \italian{sotto voce} ---

\Bloom:
Do you think they are genuine?
He could spin those yarns for hours on end all night long |
and lie like old boots.
Look at him.

\N5:
Yet still though his eyes were thick with sleep and sea air |
life was full of a host of things and coincidences of a terrible nature |
and it was quite within the bounds of possibility |
that it was not an entire fabrication.
Though at first blush |
there was not much inherent probability |
in all the spoof he got off his chest |
being strictly accurate gospel.

\N6:
Mr~Bloom had been meantime taking stock |
\gab{830}
of the individual in front of him |
and Sherlock-Holmes-ing him up |
ever since he clapped eyes on him.

\N5:
Though a well-preserved man of no little stamina, |
if a trifle prone to baldness, |
there was something spurious in the cut of his jib |
that suggested a jail delivery |
and it required no violent stretch of imagination |
to associate such a weird-looking specimen |
with the oakum and treadmill fraternity.

\N1:
He might even have done for his man |
supposing it was his own case he told, |
as people often did about others, |
namely, that he killed him himself |
and had served his four or five good-looking years in durance vile |
to say nothing of the Antonio personage

\N2:
    --- no relation to the dramatic personage of identical name |
        who sprang from the pen of our national poet ---
        \gab{840}

\N1:
who expiated his crimes in the melodramatic manner above described.

\N3:
On the other hand he might be only bluffing, |
a pardonable weakness |
because meeting unmistakable mugs

\N2:
    --- Dublin residents, like those jarvies waiting news from abroad ---

\N3:
would tempt any ancient mariner who sailed the ocean seas |
to draw the long bow about the schooner \Title{Hesperus} and etcetera.
And when all was said and done |
the lies a fellow told about himself |
couldn't probably hold a proverbial candle |
to the wholesale whoppers other fellows coined about him.

\Bloom:
Mind you, I'm not saying that it's all a pure invention.
% he resumed.
Analogous scenes are occasionally, if not often, met with.
Giants, though that is rather a far cry, |
you see once in a way, |
\gab{850}
Marcella the midget queen.
In those waxworks in Henry street |
I myself saw some Aztecs, as they are called, sitting bowlegged, |
they couldn't straighten their legs if you paid them |
because the muscles here, you see---

\N4:
    --- He proceeded, |
        indicating on his companion |
        the brief outline of the sinews ---

\Bloom:
Or whatever you like to call them behind the right knee, |
were utterly powerless from sitting that way so long cramped up, |
being adored as gods.
There's an example again of simple souls.

\N5:
However reverting to friend Sinbad and his horrifying adventures

\N6:
    --- who reminded him a bit of Ludwig, \latin{alias} Ledwidge, |
        when he occupied the boards of the Gaiety
        \gab{860}

        \N1:
            (when Michael Gunn was identified with the management)

\N6:
        in the \Title{Flying Dutchman}, a stupendous success, |
        and his host of admirers came in large numbers, |
        everyone simply flocking to hear him

        \N1:
            (though ships of any sort, phantom or the reverse, on the stage |
            usually fell a bit flat |
            as also did trains)

\N5:
there was nothing intrinsically incompatible about it, |
he conceded.

\N2:
On the contrary |
that stab in the back touch |
was quite in keeping with those italianos |
though candidly he was none the less free to admit |
those icecreamers and friers in the fish way

\N3:
    --- not to mention the chip potato variety and so forth |
        over in little Italy there near the Coombe ---

\N2:
were sober thrifty hardworking fellows |
except perhaps a bit too given
\gab{870}
to pothunting the harmless necessary animal of the feline persuasion |
of others at night |
so as to have a good old succulent tuck~in

\N3:
    --- with garlic \french{de rigueur} ---

\N2:
off him or her next day on the quiet and, he added, on the cheap.

\Bloom:
Spaniards, for instance, |
% he continued, |
passionate temperaments like that, |
impetuous as Old Nick, |
are given to taking the law into their own hands |
and give you your quietus doublequick |
with those poignards they carry |
in the abdomen.
It comes from the great heat, climate generally.
My wife is, so to speak, Spanish, half that is.
Point of fact she could actually claim Spanish nationality if she wanted, |
having been born in (technically) Spain, i.e. Gibraltar.
She has the Spanish type.
Quite dark, regular brunette, black.
I for one certainly believe climate accounts for character.
\gab{880}
That's why I asked you if you wrote your poetry in Italian.

\Stephen:
The temperaments at the door, |
% Stephen interposed with, |
were very passionate about ten shillings.
\latin{Roberto ruba roba sua.}

\Bloom:
Quite so.
% Mr~Bloom dittoed.

\Stephen:
Then---

\N4:
    --- Stephen said staring and rambling on
        to himself or some unknown listener somewhere ---

\Stephen:
Then we have the impetuosity of Dante and the isosceles triangle |
miss Portinari he fell in love with |
and Leonardo and san Tommaso Mastino.

\Bloom:
It's in the blood.
% Mr~Bloom acceded at once.
All are washed in the blood of the sun.
Coincidence I just happened to be
\gab{890}
in the Kildare street museum today, |
shortly prior to our meeting |
if I can so call it, |
and I was just looking at those antique statues there.
The splendid proportions of hips, bosom.
You simply don't knock against those kind of women here.
An exception here and there.
Handsome yes, pretty in a way you find |
but what I'm talking about is the female form.
Besides |
they have so little taste in dress, most of them, |
which greatly enhances a woman's natural beauty, |
no matter what you say.
Rumpled stockings, it may be, possibly is, a foible of mine |
but still it's a thing I simply hate to see.

\N5:
Interest, however, was starting to flag somewhat all round |
and then the others got on to talking about accidents at sea, |
\gab{900}
ships lost in a fog, |
collisions with icebergs, |
all that sort of thing.

\N6:
Shipahoy of course had his own say to say.
He had doubled the cape a few odd times |
and weathered a monsoon, a kind of wind, in the China seas |
and through all those perils of the deep |
there was one thing, he declared, stood to him |
or words to that effect, |
a pious medal he had |
that saved him.

\N1:
So then after that |
they drifted on to the wreck off Daunt's rock, |
wreck of that illfated Norwegian barque |
nobody could think of her name for the moment |
till the jarvey who had really quite a look of Henry Campbell |
remembered it

\Campbell:
\Title{Palme} on Booterstown strand.

\N2:
That was the talk of the town that year
\gab{910}

\N3:
    --- Albert William Quill wrote a fine piece of original verse of distinctive merit |
        on the topic for the \Title{Irish Times} ---

\N2:
breakers running over her |
and crowds and crowds on the shore in commotion |
petrified with horror.

\N3:
Then someone said something |
about the case of the \Title{S.S.~Lady Cairns} of Swansea |
run into by the \Title{Mona} |
which was on an opposite tack in rather muggyish weather |
and lost with all hands on deck.
No aid was given.
Her master, the \Title{Mona's}, |
said he was afraid his collision bulkhead would give way.
She had no water, it appears, in her hold.

\N4:
At this stage an incident happened.
It having become necessary for him to unfurl a reef |
the sailor vacated his seat.

\Murphy:
Let me cross your bows mate.
\gab{920}

\N4:
He said to his neighbour |
who was just gently dropping off into a peaceful doze.
He made tracks heavily, slowly |
with a dumpy sort of a gait to the door, |
stepped heavily down the one step there was |
out of the shelter and bore due left.

\N5:
While he was in the act of getting his bearings |
Mr~Bloom |
who noticed when he stood up |
that he had two flasks of presumably ship's rum |
sticking one out of each pocket |
for the private consumption of his burning interior, |
saw him produce a bottle and uncork it or unscrew |
and, applying its nozzle to his lips, |
take a good old delectable swig
out of it with a gurgling noise.

\N6:
The irrepressible Bloom,

\N1:
    --- who also had a shrewd suspicion
        that the old stager went out on a manoeuvre |
        \gab{930}
        after the counterattraction in the shape of a female |
        who however had disappeared to all intents and purposes ---

\N6:
could by straining just perceive him, |
when duly refreshed by his rum puncheon exploit, |
gaping up at the piers and girders of the Loop line |
rather out of his depth |
as of course it was all radically altered
since his last visit and greatly improved.

\N4:
Some person or persons invisible
directed him to the male urinal |
erected by the cleansing committee

\N1:
    --- all over the place ---

\N4:
for the purpose

\N1:
    --- but after a brief space of time |
        during which silence reigned supreme |
        the sailor, evidently giving it a wide berth, |
        eased himself closer at hand |
        the noise of his bilgewater |
        some little time subsequently |
        splashing on the ground ---

\N2:
where it apparently awoke a horse of the cab rank.
\gab{940}
A hoof scooped anyway |
for new foothold |
after sleep and harness jingled.

\N3:
Slightly disturbed in his sentrybox
by the brazier of live coke |
the watcher of the corporation stones

\N5:
    --- who, though now broken down and fast breaking up, |
        was none other in stern reality than the Gumley aforesaid, |
        now practically on the parish rates, |
        given the temporary job by Pat Tobin |
        in all human probability from dictates of humanity |
        knowing him before  ---

\N3:
shifted about and shuffled in his box |
before composing his limbs again |
into the arms of Morpheus,

\N5:
    --- a truly amazing piece of hard lines |
        in its most virulent form |
        on a fellow most respectably connected |
        and familiarised with decent home comforts all his life |
        \gab{950}
        who came in for a~cool £100 a~year at one time |
        which of course |
        the doublebarrelled ass proceeded to make general ducks and drakes of. ---

\N6:
And there he was at the end of his tether |
after having often painted the town tolerably pink |
without a beggarly stiver.
He drank needless to be told |
and it pointed only once more a moral |
when he might quite easily be in a large way of business |
if---a big if, however---he had contrived to cure himself
of his particular partiality.

--- \stage{Halfway point} ---

\N4:
All meantime were loudly lamenting |
the falling off in Irish shipping, |
coastwise and foreign as well, |
which was all part and parcel of the same thing.
A Palgrave Murphy boat was put off the ways at Alexandra basin, |
the only launch that year.
\gab{960}
Right enough the harbours were there |
only no ships ever called.

\N1:
There were wrecks and wreckers, |
the keeper said, |
who was evidently \french{au fait}.
What he wanted to ascertain |
was why that ship ran bang against the only rock in Galway bay |
when the Galway harbour scheme
was mooted by a Mr~Worthington or some name like that, eh?
Ask the then-captain, he advised them, |
how much palmoil the British government gave him
for that day's work, |
Captain John Lever of the Lever Line.

\SkinTheGoat:
Am I right, skipper?

\N2:
The keeper queried of the sailor, |
now returning after his private potation and the rest of his exertions.
\gab{970}

\N3:
That worthy |
picking up the scent of the fag end of the song or words |
growled in would-be music |
but with great vim |
some kind of chanty or other in seconds or thirds.

\N4:
Mr~Bloom's sharp ears heard him |
then expectorate the plug probably (which it was), |
so that he must have lodged it for the time being in his fist |
while he did the drinking and making water jobs |
and found it a bit sour after the liquid fire in question.

\N5:
Anyhow in he rolled after his successful libation-\latin{cum}-potation, |
introducing an atmosphere of drink into the \french{soirée}, |
boisterously trolling, |
like a veritable son of a seacook:

\Murphy:
    \begin{verse}
      The biscuits was as hard as brass\\
      And the beef as salt as Lot's wife's arse.%
\gab{980}\\
      O, Johnny Lever!\\
      Johnny Lever, O!
    \end{verse}

\N6:
After which effusion
the redoubtable specimen duly arrived on the scene |
and regaining his seat |
he sank rather than sat heavily on the form provided.

\N1:
Skin-the-Goat, assuming he was he, |
evidently with an axe to grind, |
was airing his grievances in a forcible-feeble philippic |
anent the natural resources of Ireland or something of that sort |
which he described in his lengthy dissertation |
as the richest country bar none on the face of God's earth, |
far and away superior to England, |
with coal in large quantities, |
six million pounds worth of pork exported every year, |
\gab{990}
ten millions between butter and eggs |
and all the riches drained out of it |
by England levying taxes on the poor people |
that paid through the nose always |
and gobbling up the best meat in the market |
and a lot more surplus steam in the same vein.

\N2:
Their conversation accordingly became general |
and all agreed that that was a fact.

\SkinTheGoat:
You could grow any mortal thing in Irish soil |
%, he stated, |
and there is that colonel Everard down there in Navan growing tobacco.
Where would you find anywhere the like of Irish bacon?

\N3:
But a day of reckoning, |
he stated \italian{crescendo} with no uncertain voice, |
thoroughly monopolising all the conversation, |
was in store for mighty England, |
despite her power of pelf |
on account of her crimes.
There would be a fall and the greatest fall in history.
\gab{1000}

\SkinTheGoat:
The Germans and the Japs are going to have their little look-in.
% he affirmed.
The Boers were the beginning of the end.
Brummagem England is toppling already |
and her downfall will be Ireland, |
her Achilles heel.

\N4:
He explained to them
about the vulnerable point of Achilles, the Greek hero, |
a point his auditors at once seized |
as he completely gripped their attention
by showing the tendon referred to on his boot.

\N5:
His advice to every Irishman:

\SkinTheGoat:
Stay in the land of your birth and work for Ireland and live for Ireland.
Ireland, Parnell said, could not spare a single one of her sons.

\N6:
Silence all round marked the termination of his \italian{finale}.
\gab{1010}
The impervious navigator heard these lurid tidings, undismayed.

\Murphy:
Take a bit of doing, boss.

\N1:
    --- Retaliated that rough diamond |
        palpably a bit peeved in response to the foregoing truism. ---

\N2:
To which cold douche |
referring to downfall and so on |
the keeper concurred |
but nevertheless held to his main view.

\SkinTheGoat:
Who's the best troops in the army?
% the grizzled old veteran irately interrogated.
And the best jumpers and racers?
And the best admirals and generals we've got?
Tell me that.

\Campbell:
The Irish, for choice.
\gab{1020}

\N3:
    --- Retorted the cabby like Campbell, facial blemishes apart. ---

\Murphy:
That's right.
% the old tarpaulin corroborated.
The Irish catholic peasant.
He's the backbone of our empire.
You know Jem Mullins?

\N4:
While allowing him his individual opinions as everyman |
the keeper added he cared nothing for any empire, ours or his, |
and considered no Irishman worthy of his salt that served it.

\N5:
Then they began to have a few irascible words |
when it waxed hotter, |
both, needless to say, |
appealing to the listeners |
who followed the passage of arms with interest |
so long as they didn't indulge in recriminations and come to blows.

\N6:
From inside information extending over a series of years |
\gab{1030}
Mr~Bloom was rather inclined to poohpooh the suggestion as egregious balderdash |
for, pending that consummation devoutly to be or not to be wished for, |
he was fully cognisant of the fact that their neighbours across the channel, |
unless they were much bigger fools than he took them for, |
rather concealed their strength than the opposite.

\N1:
It was quite on a par |
with the quixotic idea in certain quarters |
that in a hundred million years |
the coal seam of the sister island would be played out |
and if, as time went on, |
that turned out to be how the cat jumped |
all he could personally say on the matter |
was that as a host of contingencies, |
equally relevant to the issue, |
might occur ere then |
it was highly advisable in the interim |
to try to make the most of both countries |
even though poles apart.
\gab{1040}

\N2:
Another little interesting point,

\N5:
    --- the amours of whores and chummies, |
        to put it in common parlance ---

\N2:
reminded him Irish soldiers had as often fought for England as against her, |
more so, in fact.
And now, why?

\N3:
So the scene between the pair of them, |
the licensee of the place rumoured to be |
or have been |
Fitzharris, the famous invincible, |
and the other, obviously bogus, |
reminded him forcibly |
as being on all fours with the confidence trick, |
supposing, that is, |
it was prearranged as the looker-on, |
a student of the human soul if anything, |
the others seeing least of the game.

\N4:
And as for the lessee or keeper, |
who probably wasn't the other person at all, |
he~(B.) couldn't help feeling |
and most properly |
it was better to give people like that the go-by |
\gab{1050}
unless you were a blithering idiot altogether |
and refuse to have anything to do with them |
% as a golden rule in private life |
and their felonsetting, |
there always being the offchance of |
a Dannyman coming forward and turning queen's evidence or king's now |
like Denis or Peter Carey, |
an idea he utterly repudiated.

\N5:
Quite apart from that |
he disliked those careers of wrongdoing and crime on principle.
Yet,

\N6:
    --- though such criminal propensities
        had never been an inmate of his bosom in any shape or form ---

\N5:
he certainly did feel

\N6:
    --- and no denying it |
        (while inwardly remaining what he was) ---

\N5:
a certain kind of admiration
for a man who had actually brandished a knife, cold steel, |
with the courage of his political convictions

\N6:
    --- (though, personally, he would never be a party to any such thing) ---
    \gab{1060}

\N5:
off the same bat as those love vendettas of the south,

\N6:
    --- have her or swing for her ---

\N1:
when the husband |
frequently, after some words passed between the two |
concerning her relations with the other lucky mortal

\N6:
    --- (he having had the pair watched) ---

\N1:
inflicted fatal injuries on his adored one |
as a result of an alternative postnuptial \french{liaison} |
by plunging his knife into her.

\N2:
Until it just struck him |
that Fitz, nicknamed Skin-the-Goat, merely drove the car |
for the actual perpetrators of the outrage |
and so was not, |
if he was reliably informed, |
actually party to the ambush |
which, in point of fact, |
was the plea some legal luminary saved his skin on.

\N2:
In any case that was very ancient history by now |
and as for our friend, the pseudo Skin-the-etcetera, |
\gab{1070}
he had transparently outlived his welcome.
He ought to have either died naturally or on the scaffold high.

\N3:
Like actresses, always farewell |
positively last performance |
then come up smiling again.
Generous to a fault of course, |
temperamental, |
no economising or any idea of the sort, |
always snapping at the bone for the shadow.

\N4:
So |
similarly |
he had a very shrewd suspicion |
that Mr~Johnny Lever got rid of some pounds, shillings, and pence
in the course of his perambulations round the docks |
in the congenial atmosphere of the \Title{Old Ireland} tavern, |
come back to Erin and so on.

\N5:
Then |
as for the other |
he had heard not so long before |
the same identical lingo |
as he told Stephen how he simply but effectually silenced the offender.
\gab{1080}

\Bloom:
He took umbrage at something or other.

\N5:
    --- That much injured but on the whole even-tempered person declared ---

\Bloom:
I let slip.
He called me a jew |
and in a heated fashion |
offensively.
So I without deviating from plain facts in the least |
told him his God, |
I mean Christ, |
was a jew too and all his family like me |
though in reality I'm not.
That was one for him.
A soft answer turns away wrath.
He hadn't a word to say for himself as everyone saw.
Am I not right?

\N6:
He turned a long |
you-are-wrong gaze on Stephen |
of timorous dark pride |
at the soft impeachment with a glance also of entreaty |
for he seemed to glean in a kind of a way |
that it wasn't all |
exactly.
\gab{1090}

\Stephen:
\latin{Ex quibus}.

\N1:
Stephen mumbled in a noncommittal accent, |
their two or four eyes conversing.

\Stephen:
\latin{Christus} or Bloom his name is |
or after all any other, |
\latin{secundum carnem}.

\Bloom:
Of course, |
% Mr~B. proceeded to stipulate, |
you must look at both sides of the question.
It is hard to lay down any hard and fast rules as to right and wrong |
but room for improvement all round there certainly is |
though every country, they say, |
our own distressful included, |
has the government it deserves.
But with a little goodwill all round.
It's all very fine |
to boast of mutual superiority |
but what about mutual equality.
I resent violence and intolerance in any shape or form.
\gab{1100}
It never reaches anything or stops anything.
A revolution must come on the due instalments plan.
It's a patent absurdity on the face of it |
to hate people because they live round the corner
and speak another vernacular, |
in the next house so to speak.

\Stephen:
Memorable bloody bridge battle and seven minutes' war, |
% Stephen assented, |
between Skinner's alley and Ormond market.

\N2:
Yes, |
Mr~Bloom thoroughly agreed, entirely endorsing the remark, |
that was overwhelmingly right.
And the whole world was full of that sort of thing.

\Bloom:
You just took the words out of my mouth.
% he said.
A hocuspocus of conflicting evidence |
that candidly you couldn't remotely\ldots.
\gab{1110}

\N3:
All those wretched quarrels, |
in his humble opinion, |
stirring up bad blood, |
from some bump of combativeness or gland of some kind, |
erroneously supposed to be
about a punctilio of honour and a flag, |
were very largely a question of the money |
question which was at the back of everything |
greed and jealousy, |
people never knowing when to stop.

\Bloom:
They accuse.

\N4:
Remarked he audibly.

\N5:
He turned away from the others |
who probably |
and spoke nearer to, |
so as the others |
in case they.

\Bloom:
Jews,

\N6:
    --- he softly imparted in an aside in Stephen's ear ---

\Bloom:
are accused of ruining.
\gab{1120}
Not a vestige of truth in it, |
I can safely say.
History, would you be surprised to learn, |
proves up to the hilt |
Spain decayed when the inquisition hounded the jews out |
and England prospered when Cromwell, |
an uncommonly able ruffian |
who in other respects has much to answer for, |
imported them.
Why?
Because they are imbued with the proper spirit.
They are practical and are proved to be so.
I don't want to indulge in any |
because you know the standard works on the subject and then |
orthodox as you are.
But in the economic, not touching religion, domain |
the priest spells poverty.
Spain again, you saw in the war, compared with go-ahead America.
Turks.
It's in the dogma.
Because if they didn't believe they'd go straight to heaven when they die |
\gab{1130}
they'd try to live better, |
at least so I think.
That's the juggle on which the parish priests raise the wind on false pretences.
I'm
\stage{he resumed with dramatic force}
as good an Irishman as that rude person I told you about at the outset |
and I want to see everyone, |
% concluded he, |
all creeds and classes \latin{pro rata} |
having a comfortable tidysized income, |
in no niggard fashion either, |
something in the neighbourhood of £300 per annum.
That's the vital issue at stake |
and it's feasible |
and would be provocative of friendlier intercourse between man and man.
At least that's my idea for what it's worth.
I call that patriotism.
\latin{Ubi patria}, |
as we learned a smattering of in our classical days in \latin{Alma mater}, |
\latin{vita bene}.
Where you can live well, |
the sense is, |
if you work.
\gab{1140}

\N1:
Over his untastable apology for a cup of coffee, |
listening to this synopsis of things in general, |
Stephen stared at nothing in particular.
He could hear, |
of course, |
all kinds of words changing colour |
like those crabs about Ringsend in the morning |
burrowing quickly into all colours of different sorts of the same sand |
where they had a home somewhere beneath |
or seemed to.

\N2:
Then he looked up and saw the eyes that said or didn't say |
the words the voice he heard |
said, |
``if you work''.

\Stephen:
Count me out.

\N2:
He managed to remark, |
meaning ``work''.

\N3:
The eyes were surprised at this observation |
because |
as he, the person who owned them pro tem.\ observed |
\gab{1150}
or rather his voice speaking did, |
all must work, |
have to, |
together.

\Bloom:
I mean, of course,

\N3:
    --- the other hastened to affirm ---

\Bloom:
work in the widest possible sense.
Also literary labour not merely for the kudos of the thing.
Writing for the newspapers which is the readiest channel nowadays.
That's work too.
Important work.
After all, from the little I know of you, |
after all the money expended on your education |
you are entitled to recoup yourself and command your price.
You have every bit as much right |
to live by your pen in pursuit of your philosophy |
as the peasant has.
What?
You both belong to Ireland, |
the brain and the brawn.
Each is equally important.

\Stephen:
You suspect, |
\gab{1160}
\stage{Stephen retorted with a sort of a half laugh}
that I may be important |
because I belong to the \french{faubourg Saint Patrice} |
called Ireland for short.

\Bloom:
I would go a step farther.
\stage{Mr~Bloom insinuated}

\Stephen:
But I suspect, |
\stage{Stephen interrupted}
that Ireland must be important because it \emph{belongs to me}.

\Bloom:
What belongs?

\N3:
Queried Mr~Bloom bending, |
fancying he was perhaps under some misapprehension.

\Bloom:
Excuse me.
Unfortunately, I didn't catch the latter portion.
What was it you\ldots?

\N4:
Stephen, |
patently crosstempered, |
repeated |
and shoved aside his mug of coffee or whatever you like to call it |
\gab{1170}
none too politely, |
adding:

\Stephen:
We can't change the country.
Let us change the subject.

\N5:
At this pertinent suggestion |
Mr~Bloom, to change the subject, looked down |
but in a quandary, |
as he couldn't tell exactly what construction to put on |
``belongs to'' |
which sounded rather a far cry.
The rebuke of some kind was clearer than the other part.
Needless to say the fumes of his recent orgy spoke then |
with some asperity |
in a curious bitter way foreign to his sober state.

\N6:
Probably the homelife |
to which Mr~B. attached the utmost importance |
had not been all that was needful |
or he hadn't been familiarised with the right sort of people.
With a touch of fear for the young man beside him |
whom he furtively scrutinised |
\gab{1180}
with an air of some consternation |
remembering he had just come back from Paris,

\N1:
    --- the eyes more especially reminding him forcibly of father and sister, |
        failing to throw much light on the subject, however ---

\N6:
he brought to mind instances of cultured fellows |
that promised so brilliantly |
nipped in the bud of premature decay |
and nobody to blame but themselves.

\N2:
For instance |
there was the case of O'Callaghan, for one, |
the halfcrazy faddist, |
respectably connected though of inadequate means, |
with his mad vagaries |
among whose other gay doings |
when rotto and making himself a nuisance to everybody all round |
he was in the habit of ostentatiously sporting in public |
a suit of brown paper (a fact).
And then the usual \french{dénouement} |
after the fun had gone on fast and furious |
\gab{1190}
he got landed into hot water |
and had to be spirited away by a few friends, |
after a strong hint to a blind horse from John Mallon of Lower Castle Yard, |
so as not to be made amenable |
under section two of the criminal law amendment act, |
certain names of those subpœnaed being handed in but not divulged |
for reasons which will occur to anyone with a pick of brains.

\N3:
Briefly, putting two and two together,

\N4:
    --- six sixteen which he pointedly turned a deaf ear to, |
        Antonio and so forth, |
        jockeys and esthetes and the tattoo |
        which was all the go in the seventies or thereabouts |
        even in the house of lords |
        because early in life the occupant of the throne, then heir apparent, |
        the other members of the upper ten and other high personages |
        \gab{1200}
        simply following in the footsteps of the head of the state ---

\N3:
he reflected about the errors of notorieties and crowned heads |
running counter to morality |
such as the Cornwall case a number of years before |
under their veneer in a way scarcely intended by nature, |
a thing good Mrs~Grundy, as the law stands, was terribly down on |
though not for the reason they thought they were |
probably whatever it was |
except women chiefly |
who were always fiddling more or less at one another |
it being largely a matter of dress and all the rest of it.

\N5:
Ladies who like distinctive underclothing |
should, and every welltailored man must, |
trying to make the gap wider between them by innuendo |
and give more of a genuine filip to acts of impropriety between the two,
\gab{1210}

\N6:
    --- she unbuttoned his and then he untied her, mind the pin ---

\N5:
whereas savages in the cannibal islands, say, |
at ninety degrees in the shade not caring a continental.

\N2:
However, reverting to the original, |
there were on the other hand |
others who had forced their way to the top |
from the lowest rung by the aid of their bootstraps.
Sheer force of natural genius, that.
With brains, sir.

\N3:
For which and further reasons |
he felt it was his interest and duty even |
to wait on and profit by |
the unlookedfor occasion |
though why he could not exactly tell |
being as it was already several shillings to the bad |
having in fact let himself in for it.

\N4:
Still to cultivate the acquaintance of someone of no uncommon calibre |
\gab{1220}
who could provide food for reflection |
would amply repay any small.
Intellectual stimulation, as such, was, |
he felt, |
from time to time |
a firstrate tonic for the mind.

\N5:
Added to which |
was the coincidence of meeting, discussion, dance, row, |
old salt of the here today and gone tomorrow type, |
night loafers, the whole galaxy of events, |
all went to make up a miniature cameo of the world we live in |
especially as the lives of the submerged tenth, |
namely, coalminers, divers, scavengers etc., |
were very much under the microscope lately.

\N6:
To improve the shining hour |
he wondered whether he might meet with anything
approaching the same luck as Mr~Philip Beaufoy |
if taken down in writing |
suppose he were to pen something out of the common groove |
\gab{1230}
(as he fully intended doing) |
at the rate of one guinea per column.
\Title{My Experiences}, let us say, \Title{in a Cabman's Shelter}.

\N1:
The pink edition extra sporting of the \Title{Telegraph}

\N2:
    --- tell a graphic lie ---

\N1:
lay, as luck would have it, beside his elbow

\N2:
    --- and as he was just puzzling again, |
        far from satisfied, |
        over a country belonging to him and the preceding rebus |
        the vessel came from Bridgwater |
        and the postcard was addressed A.~Boudin |
        find the captain's age ---

\N1:
his eyes went aimlessly over the respective captions |
which came under his special province |
the all-embracing give us this day our daily press.

\N3:
First he got a bit of a start |
but it turned out to be only something about somebody named H.~du~Boyes, |
agent for typewriters or something like that.
\gab{1240}

\N4:
Great battle, Tokio.

\N3:
Lovemaking in Irish, £200~damages.

\N4:
Gordon Bennett.

\N3:
Emigration Swindle.

\N4:
Letter from His Grace.

\N3:
William.

\N4:
Ascot meeting, the Gold Cup.

\N3:
Victory of outsider \Title{Throwaway} recalls Derby of~'92 |
when Capt.\ Marshall's dark horse \Title{Sir Hugo} captured the blue ribband at long odds.

\N4:
New York disaster.

\N3:
Thousand lives lost.

\N4:
Foot and Mouth.

\N3:
Funeral of the late Mr~Patrick Dignam.

\N5:
So to change the subject he read about Dignam~R.I.P. |
which, he reflected, was anything but a gay sendoff.
Or a change of address anyway.

\Bloom:
\stage{to himself}
\emph{This morning |
\emph{(Hynes put it in of course)} |
the remains of the late Mr~Patrick Dignam were removed from his residence, |
number~9 Newbridge Avenue, Sandymount, for interment in Glasnevin.
\gab{1250}
The deceased gentleman was a most popular and genial personality in city life |
and his demise after a brief illness |
came as a great shock to citizens of all classes |
by whom he is deeply regretted.
The obsequies, at which many friends of the deceased were present, |
were carried out |
\emph{(certainly Hynes wrote it with a nudge from Corny)} |
by Messrs H.J.~O'Neill and Son, 164~North Strand Road.
The mourners included: Patrick Dignam (son), Bernard Corrigan (brother-in-law), |
John Henry Menton, solicitor, Martin Cunningham, John Power, |
% .)eatondph 1/8 ador dorador douradora |
% \emph{(must be where he called Monks the dayfather about Keyes's ad)} |
Thomas Kernan, Simon Dedalus, Stephen Dedalus~B.A., Edward~J. Lambert, |
Cornelius~T. Kelleher, Joseph M'C Hynes, L.~Boom, |
\gab{1260}
C.P.~M'Coy, --- M'Intosh and several others.}

\N6:
Nettled not a little by \emph{L. Boom}

\N1:
    --- (as it incorrectly stated) ---
        % and the line of bitched type

\N6:
but tickled to death simultaneously by C.P.~M'Coy and Stephen Dedalus~B.A.
who were conspicuous, needless to say, by their total absence

\N1:
    --- (to say nothing of M'Intosh) ---

\N6:
L.~Boom pointed it out to his companion~B.A. |
engaged in stifling another yawn, half nervousness, |
not forgetting the usual crop of nonsensical howlers of misprints.

\Stephen:
Is that first epistle to the Hebrews in?

\N2:
Stephen asked as soon as his bottom jaw would let him.

\Stephen:
Text: open thy mouth and put thy foot in it.

\Bloom:
It is.

\N3:
Mr~Bloom said
\gab{1270}

\N4:
    --- (though first he fancied he alluded to the archbishop |
        till he added about foot and mouth |
        with which there could be no possible connection) ---

\N3:
Really overjoyed to set his mind at rest |
and a bit flabbergasted at Myles Crawford's after all managing to.

\Bloom:
There.

\N5:
While the other was reading it on page two |
Boom (to give him for the nonce his new misnomer) |
whiled away a few odd leisure moments in fits and starts |
with the account of the third event at Ascot on page three, his side.

\N6:
Value 1000~sovs with 3000~sovs in specie added.
For entire colts and fillies.
Mr~F. Alexander's \Title{Throwaway}, |
b.h.~by \Title{Rightaway}, 5~yrs, 9~st 4~lbs (W.~Lane)~1, |
lord Howard de Walden's \Title{Zinfandel} (M.~Cannon)~2, |
Mr~W. Bass's \Title{Sceptre}~3.
\gab{1280}

\N1:
Betting 5 to~4 on \Title{Zinfandel}, 20 to~1 \Title{Throwaway} (off).
\Title{Sceptre} a shade heavier, |
5 to~4 on \Title{Zinfandel}, |
20 to~1 \Title{Throwaway} (off).
\Title{Throwaway} and \Title{Zinfandel} stood close order.

\N2:
It was anybody's race |
then the rank outsider drew to the fore, got long lead, |
beating lord Howard de Walden's chestnut colt
and Mr~W. Bass's bay filly \Title{Sceptre} |
on a 2\nicefrac{1}{2}-mile course.

\N3:
Winner trained by Braime
so that Lenehan's version of the business was all pure buncombe.

\N4:
Secured the verdict cleverly by a length.

\N3:
1000~sovs with 3000 in specie.

\N4:
Also ran: J~de Bremond's \Title{Maximum~II}.

\N3:
(French horse |
Bantam Lyons was anxiously inquiring after |
not in yet but expected any minute.)

\N4:
Different ways of bringing off a coup.
Lovemaking damages.

\N3:
Though that halfbaked Lyons ran off at a tangent |
\gab{1290}
in his impetuosity to get left.
Of course gambling eminently lent itself to that sort of thing |
though as the event turned out |
the poor fool hadn't much reason |
to congratulate himself on his pick, the forlorn hope.
Guesswork it reduced itself to eventually.

\Bloom:
There was every indication they would arrive at that.
% he, Bloom, said.

\Stephen:
Who?

\N5:
    --- The other, whose hand by the way was hurt, said ---

\N6:
One morning you would open the paper, |
the cabman affirmed, |
and read: \Title{Return of Parnell}.
He bet them what they liked.

\N1:
A Dublin fusilier was in that shelter one night and said he saw him in South Africa.

\N6:
Pride it was killed him.
\gab{1300}
He ought to have done away with himself or lain low for a time after committee room no~15 |
until he was his old self again |
with no-one to point a finger at him.
Then they would all to a man have gone down on their marrowbones to him |
to come back when he had recovered his senses.

\N1:
Dead he wasn't.
Simply absconded somewhere.

\N2:
The coffin they brought over was full of stones.

\N1:
He changed his name to De Wet, the Boer general.

\N2:
He made a mistake to fight the priests.

\N6:
And so forth and so on.

\N3:
All the same Bloom (properly so dubbed) |
was rather surprised at their memories |
for in nine cases out of ten |
it was a case of tarbarrels |
and not singly but in their thousands |
and then complete oblivion because it was twenty-odd years.
\gab{1310}
Highly unlikely of course there was even a shadow of truth in the stones |
and, even supposing, |
he thought a return highly inadvisable, all things considered.

\N4:
Something evidently riled them in his death.
Either he petered out too tamely of acute pneumonia

\N5:
    --- just when his various different political arrangements were nearing completion ---

\N4:
or whether it transpired he owed his death |
to his having neglected to change his boots and clothes after a wetting

\N5:
    --- when a cold resulted and failing to consult a specialist |
        he being confined to his room till he eventually died of it |
        amid widespread regret before a fortnight was at an end ---

\N4:
or quite possibly they were distressed to find the job was taken out of their hands.

\N6:
Of course nobody being acquainted with his movements even before |
\gab{1320}
there was absolutely no clue as to his whereabouts |
which were decidedly of the \Title{Alice, where art thou} order |
even prior to his starting to go under several aliases such as Fox and Stewart |
so the remark which emanated from friend cabby might be within the bounds of possibility.
Naturally then it would prey on his mind as a born leader of men

\N1:
    --- which undoubtedly he was |
        and a commanding figure, |
        a six-footer or at any rate five feet ten or eleven in his stockinged feet ---

\N6:
whereas Messrs So and So |
who, though they weren't even a patch on the former man, |
ruled the roost after |
their redeeming features were very few and far between.

\N4:
It certainly pointed a moral, the idol with feet of clay, |
\gab{1330}
and then seventy-two of his trusty henchmen rounding on him with mutual mudslinging.
And the identical same with murderers.
You had to come back.
That haunting sense kind of drew you.
To show the understudy in the title role |
how to.

\N5:
He saw him once

\N6:
    --- on the auspicious occasion |
        when they broke up the type in the \Title{Insuppressible} |
        or was it \Title{United Ireland} ---

\N1:
    --- a privilege he keenly appreciated ---

\N5:
and, in point of fact, handed him his silk hat when it was knocked off |
and he said ``Thank you,'' |
excited as he undoubtedly was under his frigid exterior |
notwithstanding the little misadventure
mentioned between the cup and the lip:
what's bred in the bone.

\N1:
Still as regards return.
You were a lucky dog if they didn't set the terrier at you directly you got back.
\gab{1340}
Then a lot of shillyshally usually followed, |
Tom for and Dick and Harry against.

\N2:
And then, number one, |
you came up against the man in possession |
and had to produce your credentials |
like the claimant in the Tichborne case, Roger Charles Tichborne,

\N3:
\Title{Bella} was the boat's name to the best of his recollection |
he, the heir, went down in |
as the evidence went to show |
and there was a tattoo mark too in Indian ink, |
lord Bellew was it? |
as he might very easily have picked up the details from some pal on board ship |
and then, when got up to tally with the description given, |
introduce himself with: |
``Excuse me, my name is So and So'' |
or some such commonplace remark.

\N4:
A more prudent course, |
as Bloom said to the not over effusive, |
\gab{1350}
in fact |
like the distinguished personage under discussion beside him, |
would have been to sound the lie of the land first.

\SkinTheGoat:
That bitch, that English whore, did for him.
% the shebeen proprietor commented.
She put the first nail in his coffin.

\Campbell:
Fine lump of a woman all the same, |
% the \french{soi-disant} townclerk Henry Campbell remarked, |
and plenty of her.
She loosened many a man's thighs.
I seen her picture in a barber's.
The husband was a captain or an officer.

\SkinTheGoat:
Ay,
% Skin-the-Goat amusingly added, |
he was |
and a cottonball one.

\N5:
This gratuitous contribution of a humorous character |
occasioned a fair amount of laughter among his \french{entourage}.

\N6:
As regards Bloom |
he, without the faintest suspicion of a smile, |
\gab{1360}
merely gazed in the direction of the door |
and reflected upon the historic story |
which had aroused extraordinary interest at the time |
when the facts, to make matters worse, were made public |
with the usual affectionate letters that passed between them full of sweet nothings.

\N1:
First it was strictly Platonic |
till nature intervened and an attachment sprang up between them |
till bit by bit matters came to a climax |
and the matter became the talk of the town

\N2:
till the staggering blow came as a welcome intelligence
to not a few evildisposed, however, |
who were resolved upon encompassing his downfall |
though the thing was public property all along |
though not to anything like the sensational extent that it subsequently blossomed into.
\gab{1370}

\N3:
Since their names were coupled, though, |
since he was her declared favourite, |
where was the particular necessity to proclaim it
to the rank and file from the housetops, |
the fact, namely, that he had shared her bedroom |
which came out in the witnessbox on oath

\N4:
when a thrill went through the packed court |
literally electrifying everybody |
in the shape of witnesses swearing
to having witnessed him on such and such a particular date |
in the act of scrambling out of an upstairs apartment |
with the assistance of a ladder |
in night apparel, |
having gained admittance in the same fashion,

\N5:
a fact the weeklies, addicted to the lubric a little, |
simply coined shoals of money out of.

\N6:
Whereas the simple fact of the case was |
it was simply a case of the husband not being up to the scratch, |
\gab{1380}
with nothing in common between them beyond the name, |
and then a real man arriving on the scene, |
strong to the verge of weakness, falling a victim to her siren charms |
and forgetting home ties, |
the usual sequel, |
to bask in the loved one's smiles.

\N1:
The eternal question of the life connubial, needless to say, cropped up.
Can real love, |
supposing there happens to be another chap in the case, |
exist between married folk?
Poser.

\N2:
Though it was no concern of theirs absolutely |
if he regarded her with affection, |
carried away by a wave of folly.
A magnificent specimen of manhood he was truly |
augmented obviously by gifts of a high order, |
as compared with the other military supernumerary that is 
\gab{1390}

\N3:
    --- (who was just the usual everyday |
        \emph{farewell, my gallant captain} kind of an individual |
        in the light dragoons, the \nth{18}~hussars to be accurate) ---

\N2:
and inflammable doubtless

\N3:
    --- (the fallen leader, that is, not the other) ---

\N2:
in his own peculiar way |
which she of course, woman, quickly perceived |
as highly likely to carve his way to fame |
which he almost bid fair to do.

\N4:
Till the priests and ministers of the gospel

\N5:
    --- as a whole, his erstwhile staunch adherents, ---

\N4:
and his beloved evicted tenants

\N5:
    --- for whom he had done yeoman service in the rural parts of the country |
        by taking up the cudgels on their behalf
        in a way that exceeded their most sanguine expectations, ---

\N4:
very effectually cooked his matrimonial goose, |
thereby heaping coals of fire on his head |
much in the same way as the fabled ass's kick.
\gab{1400}

\N6:
Looking back now in a retrospective kind of arrangement |
all seemed a kind of dream.
And then coming back was the worst thing you ever did |
because it went without saying |
you would feel out of place as things always moved with the times.

\N1:
Why, as he reflected, |
Irishtown strand, |
a locality he had not been in for quite a number of years |
looked different somehow |
since, as it happened, he went to reside on the north side.

\N2:
North or south, however, |
it was just the wellknown case of hot passion, pure and simple, |
upsetting the applecart with a vengeance |
and just bore out the very thing he was saying |
as she also was Spanish or half so, |
types that wouldn't do things by halves, |
passionate abandon of the south, |
casting every shred of decency to the winds.
\gab{1410}

\Bloom:
Just bears out what I was saying, |
% he, with glowing bosom said to Stephen, |
about blood and the sun.
And, if I don't greatly mistake |
she was Spanish too.

\Stephen:
The king of Spain's daughter.

\N3:
Stephen answered, |
adding something or other rather muddled about farewell |
and adieu to you Spanish onions and |
the first land called the Deadman |
and from Ramhead to Scilly was so and so many.

\Bloom:
Was she?

\N4:
    --- Bloom ejaculated, |
        surprised though not astonished by any means ---

\Bloom:
I never heard that rumour before.
Possible, especially there, it was as she lived there.
So, Spain.
\gab{1420}

\N4:
Carefully avoiding a book in his pocket \Title{Sweets of}, |
which reminded him by the by of that Capel street library book out of date, |
he took out his pocketbook |
and, turning over the various contents it contained rapidly |
finally he.

\Bloom:
Do you consider, by the by, that a Spanish type?

\N4:
He said, |
thoughtfully selecting a faded photo which he laid on the table.

\N3:
Stephen, obviously addressed, |
looked down on the photo showing a large sized lady |
with her fleshy charms on evidence in an open fashion |
as she was in the full bloom of womanhood |
in evening dress cut ostentatiously low for the occasion |
to give a liberal display of bosom, |
\gab{1430}
with more than vision of breasts, |
her full lips parted and some perfect teeth, |
standing near, ostensibly with gravity, a piano |
on the rest of which was \Title{In Old Madrid}, |
a ballad, pretty in its way, which was then all the vogue.
Her (the lady's) eyes, dark, large, |
looked at Stephen, |
about to smile about something to be admired, |
Lafayette of Westmoreland street, Dublin's premier photographic artist, |
being responsible for the æsthetic execution.

\Bloom:
Mrs~Bloom, my wife the \italian{prima donna} Madam Marion Tweedy.
% Bloom indicated.
Taken a few years since.
In or about ninety-six.
Very like her then.

\N5:
Beside the young man |
he looked also at the photo of the lady |
\gab{1440}
now his legal wife |
who, he intimated, was the accomplished daughter of Major Brian Tweedy |
and displayed at an early age remarkable proficiency as a singer |
having even made her bow to the public |
when her years numbered barely sweet sixteen.

\N6:
As for the face |
it was a speaking likeness in expression |
but it did not do justice to her figure |
which came in for a lot of notice usually |
and which did not come out to the best advantage in that getup.
She could without difficulty, he said, have posed for the ensemble, |
not to dwell on certain opulent curves of the. |
He dwelt, being a bit of an artist in his spare time, |
on the female form in general developmentally |
because, as it so happened, no later than that afternoon |
\gab{1450}
he had seen those Grecian statues, perfectly developed as works of art, |
in the National Museum.

\N1:
Marble could give the original, |
shoulders, back, all the symmetry, all the rest.

\N2:
    --- Yes, puritanisme, it does |
        though Saint Joseph's sovereign thievery alors |
        \french{(Bandez!) |
        Figne toi trop.} ---

\N1:
Whereas no photo could |
because it simply wasn't art |
in a word.

\N3:
The spirit moving him |
he would much have liked to follow Jack Tar's good example |
and leave the likeness there for a very few minutes to speak for itself |
on the plea he |
so that the other could drink in the beauty for himself, |
her stage presence being, frankly, a treat in itself |
which the camera could not at all do justice to.
\gab{1460}
But it was scarcely professional etiquette so.

\N4:
Though it was a warm pleasant sort of a night now |
yet wonderfully cool for the season considering, |
for sunshine after storm.
And he did feel a kind of need there and then |
to follow suit like a kind of inward voice |
and satisfy a possible need by moving a motion.

\N5:
Nevertheless he sat tight |
just viewing the slightly soiled photo creased by opulent curves, |
none the worse for wear however, |
and looked away thoughtfully |
with the intention of not further increasing the other's possible embarrassment |
while gauging her symmetry of heaving \french{embonpoint}.

\N6:
In fact the slight soiling was only an added charm |
like the case of linen slightly soiled, good as new, |
much better in fact with the starch out.
\gab{1470}

\N1:
Suppose she was gone when he?
``I looked for the lamp which she told me'' |
came into his mind but merely as a passing fancy of his |
because he then recollected the morning-littered bed etcetera |
and the book about Ruby with ``met him pike hoses''
%(\latin{sic})
in it |
which must have fell down sufficiently appropriately beside the domestic chamberpot |
with apologies to Lindley Murray.

\N2:
The vicinity of the young man
\stage{emphasize}
he certainly relished, |
educated, \french{distingué} and impulsive into the bargain, |
far and away the pick of the bunch |
though you wouldn't think he had it in him |
yet you would.
Besides |
he said the picture was handsome |
which, say what you like, it was |
though at the moment she was distinctly stouter.
\gab{1480}

\N3:
And why not?
An awful lot of makebelieve went on about that sort of thing |
involving a lifelong slur |
with the usual splash page of gutterpress |
about the same old matrimonial tangle |
alleging misconduct with professional golfer or the newest stage favourite |
instead of being honest and aboveboard about the whole business.

\N4:
How they were fated to meet and an attachment sprang up between the two |
so that their names were coupled in the public eye |
was told in court |
with letters containing the habitual mushy and compromising expressions |
leaving no loophole |
to show that they openly cohabited two or three times a week |
at some wellknown seaside hotel |
and relations, |
when the thing ran its normal course, |
became in due course intimate.
\gab{1490}

\N5:
Then the decree \latin{nisi} |
and the King's proctor tries to show cause why |
and, he failing to quash it, |
\latin{nisi} was made absolute.
But as for that |
the two misdemeanants, |
wrapped up as they largely were in one another, |
could safely afford to ignore it |
as they very largely did |
till the matter was put in the hands of a solicitor |
who filed a petition for the party wronged in due course.

\N6:
He,~B., enjoyed the distinction of being close
to Erin's uncrowned king in the flesh |
when the thing occurred on the historic \french{fracas} |
when the fallen leader's,

\N1:
    --- who notoriously stuck to his guns to the last drop |
        even when clothed in the mantle of adultery ---

\N6:
(leader's) trusty henchmen

\N1:
    --- to the number of ten or a dozen or possibly even more than that ---

\N6:
penetrated into the printing works of the \Title{Insuppressible}
\gab{1500}

\N1:
    --- or no it was \Title{United Ireland}
        (a by no means by-the-by appropriate appellative) ---

\N6:
and broke up the typecases with hammers or something like that

\N1:
    --- all on account of some scurrilous effusions |
        from the facile pens of the O'Brienite scribes |
        at the usual mudslinging occupation |
        reflecting on the erstwhile tribune's private morals. ---

\N2:
Though palpably a radically altered man |
he was still a commanding figure |
though carelessly garbed as usual |
with that look of settled purpose |
which went a long way with the shillyshallyers |
till they discovered to their vast discomfiture |
that their idol had feet of clay |
after placing him upon a pedestal |
which she, however, was the first to perceive.

\N3:
As those were particularly hot times in the general hullaballoo |
\gab{1510}
Bloom sustained a minor injury from |
a nasty prod of some chap's elbow |
in the crowd that of course congregated |
lodging some place about the pit of the stomach, |
fortunately not of a grave character.

\N4:
His hat (Parnell's) a silk one |
was inadvertently knocked off |
and, as a matter of strict history, |
Bloom was the man who picked it up in the crush |
after witnessing the occurrence |
meaning to return it to him

\N3:
    --- (and return it to him he did with the utmost celerity) ---

\N4:
who panting and hatless
and whose thoughts were miles away from his hat at the time 

\N5:
all the same |
being a gentleman born with a stake in the country 

\N6:
    --- he, as a matter of fact, having gone into it |
        more for the kudos of the thing than anything else ---

\N5:
what's bred in the bone
\gab{1520}

\N6:
    --- instilled into him in infancy at his mother's knee |
        in the shape of knowing what good form was ---

\N5:
came out at once

\N4:
because he turned round to the donor
and thanked him with perfect \french{aplomb}, saying: |
\emph{Thank you, sir},

\N5:
though in a very different tone of voice |
from the ornament of the legal profession |
% Menton
whose headgear Bloom also set to rights earlier in the course of the day,

\N6:
history repeating itself with a difference, |
after the burial of a mutual friend |
when they had left him alone in his glory |
after the grim task of having committed his remains to the grave.

\N1:
On the other hand what incensed him more inwardly |
was the blatant jokes of the cabman and so on |
\gab{1530}
who passed it all off as a jest, laughing immoderately, |
pretending to understand everything, the why and the wherefore, |
and in reality not knowing their own minds,

\N2:
it being a case for the two parties themselves |
unless it ensued that the legitimate husband happened to be a party to it |
owing to some anonymous letter from the usual boy Jones, |
who happened to come across them at the crucial moment |
in a loving position locked in one another's arms, |
drawing attention to their illicit proceedings |
and leading up to a domestic rumpus

\N3:
and the erring fair one begging forgiveness of her lord and master upon her knees |
and promising to sever the connection and not receive his visits any more |
if only the aggrieved husband would overlook the matter |
\gab{1540}
and let bygones be bygones |
with tears in her eyes |
though possibly with her tongue in her fair cheek at the same time |
as quite possibly there were several others.

\N4:
He personally, being of a sceptical bias, believed |
and didn't make the smallest bones about saying so either |
that man or men in the plural were always hanging around on the waiting list about a lady,

\N5:
    --- even supposing she was the best wife in the world |
        and they got on fairly well together for the sake of argument ---

\N6:
    --- when, neglecting her duties, |
        she chose to be tired of wedded life |
        and was on for a little flutter in polite debauchery ---

\N4:
to press their attentions on her with improper intent, |
the upshot being that her affections centred on another, |
the cause of many \french{liaisons} |
\gab{1550}
between still attractive married women |
getting on for fair and forty |
and younger men, |
no doubt as several famous cases of feminine infatuation proved up to the hilt.

\N4:
It was a thousand pities a young fellow, |
blessed with an allowance of brains as his neighbour obviously was, |
should waste his valuable time with profligate women |
who might present him with a nice dose to last him his lifetime.
In the nature of single blessedness |
he would one day take unto himself a wife when Miss Right came on the scene |
but in the interim |
ladies' society was a \latin{conditio sine qua non}

\N5:
    --- though he had the gravest possible doubts, |
        not that he wanted in the smallest to pump Stephen about Miss Ferguson ---
        \gab{1560}

\N6:
    --- (who was very possibly the particular lodestar |
        who brought him down to Irishtown so early in the morning) ---

\N5:
as to whether he would find much satisfaction basking in the boy and girl courtship idea |
and the company of smirking misses without a penny to their names bi or triweekly |
with the orthodox preliminary canter of compliment playing and walking out |
leading up to fond lovers' ways and flowers and chocs.

\N4:
To think of him house and homeless, |
rooked by some landlady worse than any stepmother, |
was really too bad at his age.

\N1:
The queer suddenly things he popped out with |
attracted the elder man who was several years the other's senior |
or like his father |
but something substantial he certainly ought to eat |
even were it only an eggflip made on unadulterated maternal nutriment |
\gab{1570}
or, failing that, the homely Humpty Dumpty boiled.

\Bloom:
At what o'clock did you dine?

\N1:
He questioned of the slim form and tired though unwrinkled face.

\Stephen:
Some time yesterday.
% Stephen said.

\Bloom:
Yesterday!

\N1:
Exclaimed Bloom till he remembered it was already tomorrow Friday.

\Bloom:
Ah, you mean it's after twelve!

\Stephen:
The day before yesterday.

\N2:
Stephen said, improving on himself.

\N1:
Literally astounded at this piece of intelligence |
Bloom reflected.
Though they didn't see eye to eye in everything |
a certain analogy there somehow was |
as if both their minds were travelling, |
so to speak, |
\gab{1580}
in the one train of thought.

\N3:
At his age when dabbling in politics |
roughly some score of years previously |
when he had been a \latin{quasi} aspirant to parliamentary honours in the Buckshot Foster days |
he too recollected in retrospect |
(which was a source of keen satisfaction in itself) |
he had a sneaking regard for those same ultra ideas.

\N4:
For instance when the evicted tenants question, |
then at its first inception, |
bulked largely in people's mind

\N5:
    --- though, it goes without saying, not contributing a copper |
        or pinning his faith absolutely to its dictums, |
        some of which wouldn't exactly hold water ---

\N4:
he at the outset in principle at all events |
was in thorough sympathy with peasant possession |
as voicing the trend of modern opinion
\gab{1590}

\N5:
    --- (a partiality, however, |
        which, realising his mistake, |
        he was subsequently partially cured of) ---

\N4:
and even was twitted with going a step farther than Michael Davitt |
in the striking views he at one time inculcated as a back-to-the-lander
which was one reason he strongly resented the innuendo |
put upon him in so barefaced a fashion by our friend |
at the gathering of the clans in Barney Kiernan's

\N6:
so that he, |
though often considerably misunderstood |
and the least pugnacious of mortals, be it repeated, |
departed from his customary habit to give him (metaphorically) one in the gizzard |
though, so far as politics themselves were concerned, |
he was only too conscious of the casualties |
invariably resulting from propaganda and displays of mutual animosity |
\gab{1600}
and the misery and suffering it entailed |
as a foregone conclusion |
on fine young fellows, |
chiefly, destruction of the fittest, in a word.

\N1:
Anyhow upon weighing up the pros and cons, |
getting on for one, as it was, |
it was high time to be retiring for the night.
The crux was |
it was a bit risky to bring him home |
as eventualities might possibly ensue

\N2:
    --- (somebody having a temper of her own sometimes) ---

\N1:
and spoil the hash altogether |
as on the night he misguidedly brought home a dog (breed unknown) with a lame paw

\N2:
    --- (not that the cases were either identical or the reverse |
        though he had hurt his hand too) ---

\N1:
to Ontario Terrace |
as he very distinctly remembered, |
having been there, so to speak.
\gab{1610}

\N3:
On the other hand |
it was altogether far and away too late |
for the Sandymount or Sandycove suggestion |
so that he was in some perplexity as to which of the two alternatives.

\N4:
Everything pointed to the fact that it behoved him |
to avail himself to the full of the opportunity, all things considered.
His initial impression was |
he was a shade standoffish or not over effusive |
but it grew on him someway.

\N5:
For one thing he mightn't what you call jump at the idea, if approached, |
and what mostly worried him |
was he didn't know how to lead up to it or word it exactly, |
supposing he did entertain the proposal, |
as it would afford him very great personal pleasure |
if he would allow him to help to put coin in his way |
or some wardrobe, if found suitable.
\gab{1620}

\N6:
At all events he wound up by concluding, |
eschewing for the nonce |
hidebound precedent, |
a cup of Epps's cocoa and a shakedown for the night |
plus the use of a rug or two and overcoat doubled into a pillow |
at least he would be in safe hands and as warm as a toast on a trivet

\N1:
he failed to perceive any very vast amount of harm in that |
always with the proviso no rumpus of any sort was kicked up.

\N2:
A move had to be made |
because that merry old soul, the grasswidower in question |
who appeared to be glued to the spot, |
didn't appear in any particular hurry |
to wend his way home to his dearly beloved Queenstown |
and it was highly likely some sponger's bawdyhouse of retired beauties |
where age was no bar |
off Sheriff street lower |
would be the best clue |
\gab{1630}
to that equivocal character's whereabouts for a few days to come, |
alternately racking their feelings (the mermaids') |
with sixchamber revolver anecdotes verging on the tropical |
calculated to freeze the marrow of anybody's bones |
and mauling their largesized charms betweenwhiles with rough and tumble gusto |
to the accompaniment of large potations of potheen |
and the usual blarney about himself |
for as to who he in reality was |
let~$x$ equal my right name and address, |
as Mr~Algebra remarks \latin{passim}.

\N3:
At the same time he inwardly chuckled |
over his gentle repartee to the blood and ouns champion |
about his god being a jew.
People could put up with being bitten by a wolf |
but what properly riled them |
was a bite from a sheep.

\N4:
    --- The most vulnerable point too of tender Achilles. ---
    \gab{1640}

\N3:
``Your god was a jew.''
Because mostly they appeared to imagine he came from Carrick-on-Shannon |
or somewhereabouts in the county Sligo.

\Bloom:
I propose---

\N5:
    --- Our hero eventually suggested after mature reflection |
    while prudently pocketing her photo ---

\Bloom:
As it's rather stuffy here |
you just come home with me and talk things over.
My diggings are quite close in the vicinity.
You can't drink that stuff.
Do you like cocoa?
Wait.
I'll just pay this lot.

\N6:
The best plan clearly being to clear out, |
the remainder being plain sailing, he beckoned, |
while prudently pocketing the photo, |
to the keeper of the shanty |
who didn't seem to.

\Bloom:
Yes, that's the best.
\gab{1650}

\N1:
    --- He assured Stephen |
        to whom for the matter of that |
        Brazen Head or him or anywhere else |
        was all more or less ---

\N2:
All kinds of Utopian plans were flashing through his (B.'s) busy brain, |
education (the genuine article), literature, journalism, |
prize titbits, up to date billing, |
concert tours in English watering resorts |
packed with hydros and seaside theatres, |
turning money away, |
duets in Italian with the accent perfectly true to nature |
and a quantity of other things, |
no necessity, of course, |
to tell the world and his wife from the housetops about it, |
and a slice of luck.

\N3:
An opening was all was wanted.
Because he more than suspected |
he had his father's voice to bank his hopes on |
which it was quite on the cards he had |
so it would be just as well, |
\gab{1660}
by the way no harm, |
to trail the conversation in the direction of that particular red herring |
just to.

\N4:
The cabby read out of the paper he had got hold of |
that the former viceroy, earl Cadogan, |
had presided at the cabdrivers' association dinner in London somewhere.
Silence with a yawn or two accompanied this thrilling announcement.

\N5:
Then the old specimen in the corner |
who appeared to have some spark of vitality left |
read out that sir Anthony MacDonnell
had left Euston for the chief secretary's lodge |
or words to that effect.
To which absorbing piece of intelligence |
echo answered why.

\Murphy:
Give us a squint at that literature, grandfather.

\N6:
The ancient mariner put in, |
manifesting some natural impatience.
\gab{1670}

\OldSpecimen:
And welcome.

\N5:
Answered the elderly party thus addressed.

\N6:
The sailor lugged out from a case he had |
a pair of greenish goggles |
which he very slowly hooked over his nose and both ears.

\Campbell:
Are you bad in the eyes?

\N1:
The sympathetic personage like the townclerk queried.

\Murphy:
Why---

\N6:
Answered the seafarer with the tartan beard, |
who seemingly was a bit of a literary cove in his own small way, |
staring out of seagreen portholes as you might well describe them as.

\Murphy:
Why, I uses goggles reading.
Sand in the Red Sea done that.
One time I could read a book in the dark, manner of speaking.
\gab{1680}
\Title{The Arabian Nights Entertainment} was my favourite |
and \Title{Red as a Rose is She}.

\N6:
Hereupon he pawed the journal open |
and pored upon Lord only knows what, found drowned |
or the exploits of King Willow, |
Iremonger having made a hundred and something second wicket not out for Notts,

\N2:
during which time (completely regardless of Ire) |
the keeper was intensely occupied loosening an apparently new or secondhand boot |
which manifestly pinched him as he muttered against whoever it was sold it, |
all of them who were sufficiently awake enough |
to be picked out by their facial expressions, |
that is to say, either simply looking on glumly or passing a trivial remark.
\gab{1690}

\N3:
To cut a long story short |
Bloom, grasping the situation, was the first to rise from his seat |
so as not to outstay their welcome |
having first and foremost, |
being as good as his word that he would foot the bill for the occasion, |
taken the wise precaution to unobtrusively motion to mine host |
as a parting shot a scarcely perceptible sign when the others were not looking |
to the effect that the amount due was forthcoming, |
making a grand total of fourpence |
(the amount he deposited unobtrusively in four coppers, literally the last of the Mohicans), |
he having previously spotted on the printed pricelist |
for all who ran to read |
opposite him in unmistakable figures, coffee two pence, confectionery ditto, |
\gab{1700}
and honestly well worth twice the money once in a way |
as Wetherup used to remark.

\Bloom:
Come.

\N4:
He counselled to close the \french{séance}.
Seeing that the ruse worked and the coast was clear |
they left the shelter or shanty together |
and the \french{élite} society of oilskin and company |
whom nothing short of an earthquake would move out of their \italian{dolce far niente}.

\N5:
Stephen, who confessed to still feeling poorly and fagged out, |
paused at the, |
for a moment, |
the door.

\Stephen:
One thing I never understood.

\N5:
He said to be original on the spur of the moment.

\Stephen:
Why they put tables upside down at night.
I mean chairs upside down, |
on the tables in cafés.
\gab{1710}

\N6:
To which impromptu |
the neverfailing Bloom replied |
without a moment's hesitation, |
saying straight off:

\Bloom:
To sweep the floor in the morning.

\N6:
So saying he skipped around, nimbly considering, |
frankly at the same time apologetic to get on his companion's right, |
a habit of his, by the bye, |
his right side being, in classical idiom, his tender Achilles.
The night air was certainly now a treat to breathe |
though Stephen was a bit weak on his pins.

\Bloom:
It will (the air) do you good,

\N6:
    --- Bloom said, meaning also the walk ---

\Bloom:
in a moment.
The only thing is to walk |
then you'll feel a different man.
Come.
It's not far.
Lean on me.
\gab{1720}

\N1:
Accordingly he passed his left arm in Stephen's right and led him on accordingly.

\Stephen:
Yes.

\N4:
Stephen said uncertainly |
because he thought he felt a strange kind of flesh of a different man approach him, |
sinewless and wobbly and all that.

\N2:
Anyhow they passed the sentrybox with stones, brazier etc.\
where the municipal supernumerary, ex Gumley, was still |
to all intents and purposes |
wrapped in the arms of Murphy, as the adage has it, |
dreaming of fresh fields and pastures new.

\N3:
And \french{apropos} of coffin of stones |
the analogy was not at all bad |
as it was in fact a stoning to death |
on the part of seventy-two out of eighty-odd constituencies |
\gab{1730}
that ratted at the time of the split |
and chiefly the belauded peasant class, |
probably the selfsame evicted tenants he had put in their holdings.

\N4:
So they turned on to chatting about music, |
a form of art for which |
Bloom, as a pure amateur, possessed the greatest love, |
as they made tracks arm in arm across Beresford place.
Wagnerian music, though confessedly grand in its way, |
was a bit too heavy for Bloom and hard to follow at the first go-off |
but the music of Mercadante's \Title{Huguenots},
Meyerbeer's \Title{Seven Last Words on the Cross}
and Mozart's \Title{Twelfth Mass} |
he simply revelled in, |
the \Title{Gloria} in that |
being, to his mind, |
the acme of first class music as such, |
literally knocking everything else into a cocked hat.
\gab{1740}

\N5:
He infinitely preferred the sacred music of the catholic church |
to anything the opposite shop could offer in that line |
such as those Moody and Sankey hymns |
or \Title{Bid me to live and I will live thy protestant to be}.
He also yielded to none in his admiration of Rossini's \Title{Stabat Mater}, |
a work simply abounding in immortal numbers, |
in which his wife, Madam Marion Tweedy, made a hit, |
a veritable sensation, he might safely say, |
greatly adding to her other laurels |
and putting the others totally in the shade, |
in the jesuit fathers' church in upper Gardiner street, |
the sacred edifice being thronged to the doors to hear her with virtuosos, |
or \italian{virtuosi} rather.
There was the unanimous opinion that there was none to come up to her |
\gab{1750}
and suffice it to say in a place of worship for music of a sacred character |
there was a generally voiced desire for an encore.

\N6:
On the whole though |
favouring preferably light opera of the \Title{Don Giovanni} description |
and \Title{Martha}, a gem in its line, |
he had a \french{penchant}, |
though with only a surface knowledge, |
for the severe classical school such as Mendelssohn.

\N1:
And talking of that, |
taking it for granted he knew all about the old favourites, |
he mentioned \french{par excellence} |
Lionel's air in \Title{Martha}, |
\Title{M'appari}, |
which, curiously enough, he had heard or overheard, |
to be more accurate, on yesterday, a privilege he keenly appreciated, |
from the lips of Stephen's respected father, sung to perfection, |
a study of the number, in fact, which made all the others take a back seat.
\gab{1760}

\N2:
Stephen, in reply to a politely put query, said he didn't sing it |
but launched out into praises of Shakespeare's songs, |
at least of in or about that period, |
the lutenist Dowland who lived in Fetter lane near Gerard the herbalist, |
who \latin{annos ludendo hausi, Doulandus}, |
an instrument he was contemplating purchasing from Mr~Arnold Dolmetsch,

\N3:
    --- whom B.~did not quite recall though the name certainly sounded familiar ---

\N2:
for sixty-five guineas |
and Farnaby and son with their \latin{dux} and \latin{comes} conceits |
and Byrd (William) who played the virginals, he said, |
in the Queen's chapel or anywhere else he found them |
and one Tomkins who made toys or airs |
and John Bull.

\N3:
On the roadway which they were approaching whilst still speaking |
\gab{1770}
beyond the swingchains a horse, |
dragging a sweeper, |
paced on the paven ground, |
brushing a long swathe of mire up

\N4:
so that with the noise |
Bloom was not perfectly certain whether he had caught aright |
the allusion to sixty-five guineas and John Bull.
He inquired if it was John Bull the political celebrity of that ilk, |
as it struck him, |
the two identical names, as a striking coincidence.

\N5:
By the chains the horse slowly swerved to turn, |
which perceiving, |
Bloom, who was keeping a sharp lookout as usual, |
plucked the other's sleeve gently, jocosely remarking:

\Bloom:
Our lives are in peril tonight.
\gab{1780}
Beware of the steamroller.

\N6:
They thereupon stopped.
Bloom looked at the head of a horse not worth anything like sixty-five guineas, |
suddenly in evidence in the dark quite near |
so that it seemed new, |
a different grouping of bones and even flesh |
because palpably it was |
a four-walker, a hip-shaker, a black-buttocker,
a tail-dangler, a head-hanger putting his hind foot foremost |
the while the lord of his creation sat on the perch, busy with his thoughts.

\N1:
But such a good poor brute |
he was sorry he hadn't a lump of sugar |
but, as he wisely reflected, |
you could scarcely be prepared for every emergency that might crop up.
He was just a big nervous foolish noodly kind of a horse, |
without a second care in the world.
\gab{1790}

\N2:
But even a dog, he reflected, |
take that mongrel in Barney Kiernan's, of the same size, |
would be a holy horror to face.
But it was no animal's fault in particular |
if he was built that way |
like the camel, ship of the desert, |
distilling grapes into potheen in his hump.
Nine tenths of them all could be caged or trained, |
nothing beyond the art of man |
barring the bees.
Whale with a harpoon hairpin, |
alligator tickle the small of his back and he sees the joke, |
chalk a circle for a rooster, |
tiger my eagle eye.

\N3:
These timely reflections anent the brutes of the field occupied his mind |
somewhat distracted from Stephen's words |
while the ship of the street was manoeuvring |
and Stephen went on about the highly interesting old.

\Bloom:
What's this I was saying?
\gab{1800}
Ah, yes!
My wife, |
% he intimated, plunging \latin{in medias res}, |
would have the greatest of pleasure in making your acquaintance |
as she is passionately attached to music of any kind.

\N4:
He looked sideways in a friendly fashion at the sideface of Stephen, |
image of his mother, |
which was not quite the same as the usual handsome blackguard type |
they unquestionably had an insatiable hankering after |
as he was perhaps not that way built.
Still, |
supposing he had his father's gift as he more than suspected, |
it opened up new vistas in his mind |
such as Lady Fingall's Irish industries, |
concert on the preceding Monday, |
and aristocracy in general.

\N5:
Exquisite variations he was now describing |
\gab{1810}
on an air \Title{Youth here has End} by Jans Pieter Sweelinck, |
a Dutchman of Amsterdam where the frows come from.
Even more he liked an old German song of Johannes Jeep |
about the clear sea and the voices of sirens, sweet murderers of men, |
which boggled Bloom a bit:

\Stephen:
\begin{verse}
  \german{% |
    Von der Sirenen Listigkeit\\
    Tun die Poeten dichten.}
\end{verse}

\N6:
These opening bars he sang and translated \latin{extempore}.

\N1:
Bloom, nodding, said he perfectly understood |
and begged him to go on by all means |
which he did.
A phenomenally beautiful tenor voice like that,
\gab{1820}

\N2:
    --- the rarest of boons, |
        which Bloom appreciated at the very first note he got out  ---

\N1:
could easily,

\N2:
    --- if properly handled by some recognised authority on voice production |
        such as Barraclough |
        and being able to read music into the bargain ---

\N1:
command its own price |
where baritones were ten a penny |
and procure for its fortunate possessor in the near future |
an \french{entrée} into fashionable houses in the best residential quarters |
of financial magnates in a large way of business |
and titled people.

\N3:
With his university degree of~B.A. |
(a huge ad in its way) |
and gentlemanly bearing |
to all the more influence the good impression |
he would infallibly score a distinct success, |
being blessed with brains which also could be utilised for the purpose |
\gab{1830}
and other requisites, |
if his clothes were properly attended to |
so as to the better worm his way into their good graces |
as he, a youthful tyro in society's sartorial niceties, |
hardly understood how a little thing like that could militate against you.

\N4:
It was in fact only a matter of months |
and he could easily foresee him participating
in their musical and artistic \italian{conversaziones} |
during the festivities of the Christmas season, for choice, |
causing a slight flutter in the dovecotes of the fair sex |
and being made a lot of |
by ladies out for sensation, |
cases of which, as he happened to know, were on record |
--- in fact, without giving the show away, |
he himself once upon a time, if he cared to, could easily have.

\N5:
Added to which of course |
would be the pecuniary emolument |
\gab{1840}
by no means to be sneezed at, |
going hand in hand with his tuition fees.
Not, he parenthesised, that for the sake of filthy lucre |
he need necessarily embrace the lyric platform as a walk in life |
for any lengthy space of time.

\N6:
But a step in the required direction it was |
beyond yea or nay |
and both monetarily and mentally |
it contained no reflection on his dignity in the smallest |
and it often turned in uncommonly handy |
to be handed a cheque at a muchneeded moment when every little helped.

\N1:
Besides, though taste latterly had deteriorated to a degree, |
original music like that, |
different from the conventional rut, |
would rapidly have a great vogue |
as it would be a decided novelty for Dublin's musical world |
\gab{1850}
after the usual hackneyed run of catchy tenor solos |
foisted on a confiding public |
by Ivan St Austell and Hilton St Just and their \latin{genus omne}.

\N2:
Yes, beyond a shadow of a doubt |
he could with all the cards in his hand |
and he had a capital opening to make a name for himself
and win a high place in the city's esteem |
where he could command a stiff figure |
and, booking ahead, |
give a grand concert for the patrons of the King street house, |
given a backer-up, if one were forthcoming |
to kick him upstairs, so to speak, |
a big \emph{if}, however, |
with some impetus of the go-ahead sort |
to obviate the inevitable procrastination |
which often tripped up a too-much fêted prince of good fellows.

\N3:
And it need not detract from the other by one iota |
\gab{1860}
as, being his own master, |
he would have heaps of time to practise literature in his spare moments when desirous of so doing |
without its clashing with his vocal career |
or containing anything derogatory whatsoever |
as it was a matter for himself alone.
In fact, he had the ball at his feet |
and that was the very reason why the other, |
possessed of a remarkably sharp nose for smelling a rat of any sort, |
hung on to him at all.

\N5:
The horse was just then.

\N4:
And later on |
at a propitious opportunity |
he purposed (Bloom did), |
without anyway prying into his private affairs |
on the \emph{fools step in where angels} principle, |
advising him to sever his connection with a certain budding practitioner |
who, he noticed, |
was prone to disparage and even to a slight extent |
\gab{1870}
with some hilarious pretext |
when not present, |
deprecate him, |
or whatever you like to call it |
which in Bloom's humble opinion |
threw a nasty sidelight on that side of a person's character, |
no pun intended.

\N5:
The horse having reached the end of his tether, so to speak, halted |
and, rearing high a proud feathering tail, |
added his quota by letting fall on the floor |
which the brush would soon brush up and polish, |
three smoking globes of turds.
Slowly three times, one after another, from a full crupper |
he mired.
And humanely his driver waited till he (or she) had ended, |
patient in his scythed car.

\N6:
Side by side Bloom, |
\gab{1880}
profiting by the \french{contretemps}, |
with Stephen passed through the gap of the chains, divided by the upright, |
and, stepping over a strand of mire, |
went across towards Gardiner street lower,

\N1:
Stephen singing more boldly, but not loudly, |
the end of the ballad.

\Stephen:
\begin{verse}
    \german{Und alle Schiffe brücken.}
\end{verse}

\N2:
The driver never said a word, good, bad or indifferent, |
but merely watched the two figures, |
as he sat on his lowbacked car, |
both black, one full, one lean, |
walk towards the railway bridge, |
\emph{to be married by Father Maher.}

\N3:
As they walked |
they at times stopped and walked again |
continuing their \french{tête-à-tête} |
(which, of course, he was utterly out of)

\N3:
about sirens |
enemies of man's reason, |
\gab{1890}
mingled with a number of other topics of the same category, |
usurpers, historical cases of the kind

\N2:
while the man in the sweeper car |
or you might as well call it in the sleeper car |
who in any case couldn't possibly hear |
because they were too far |
simply sat in his seat near the end of lower Gardiner street |
\emph{and looked after their lowbacked car.}

\end{document}
